---
title: "PANTHER PFAS Analysis"
author: "Hailey Hampson"
date: "2025-12-19"
output: html_document
---

#1. Set Up 
##a. Load Libraries & Directores
```{r}
#Libraries
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

#Directories
# computer <- "mac studio"
computer <- "mac laptop"
if (computer == "mac studio") {
  user <- Sys.info()[["user"]]
  if (user == "hhampson") {
    dir.dat <- c("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive")
    dir.results <- c("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results")
    git_path <- "Users/hhampson/CHCO-Code/Petter Bjornstad"
  } else if (user == "shivaniramesh") {
    dir.dat <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive")
    dir.results <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results")
  } else if (user == "choiyej") {
    dir.dat <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  } else if (user == "pylell") {
    dir.dat <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  }
} else {
  user <- Sys.info()[["user"]]
  if (user == "hhampson") {
    dir.dat <- c("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive")
    dir.results <- c("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/PANTHER/Results")
    git_path <- "Users/hhampson/CHCO-Code/Petter Bjornstad"
  } else {
    dir.dat <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive")
    dir.results <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results")
  }
}

# source(file.path(git_path, "PANTHER PFAS/Libraries.R"))
source("Libraries.R")
source("Functions.R")
```

##b. Load Data
```{r}
#Load Metadata
dat <- read.csv(fs::path(dir.dat,"Data Harmonization","Data Clean","harmonized_dataset.csv"),na.strings = "")
lod <- read.csv(fs::path(dir.dat,"/PANTHER/Data_Raw/PFAS Emory 2025/8- PFAS Quantification/PFAS_LODs.csv"))

##c. Clean & Format Data----
# dat <- dat %>% 
#   mutate(group=ifelse(record_id=="PAN-110-O","Obese Control",
#                       ifelse(record_id=="PAN-61-C","Lean Control",
#                              ifelse(record_id=="PAN-71-C","Lean Control",group))))

pfas <- c("N.EtFOSAA","N.MeFOSAA","PFBA","PFDA","PFDoA","PFHpA","PFHxA","PFNA","PFOA","PFBS","PFPeA","PFTeDA","PFTrDA","PFUnA","PFBS","PFDoS","PFHps","PFHxS","PFNS","PFOS","PFOSA","PFPeAS")
dat_collapsed1 <- dat %>%
  filter(study=="PANTHER") %>% 
  filter(visit=="baseline" | visit=="screening") %>% 
  # dplyr::select(record_id, date, rh2_id, visit,mrn, group, age, sex, bmi, hba1c,albuminuria_cat,microalbumin_u,starts_with("egfr"),cystatin_c_s,starts_with("creat"),all_of(pfas)) %>% 
  dplyr::select(record_id, dob, date, rh2_id, visit, mrn, group, age, sex, bmi,race,ethnicity,race_ethnicity, hba1c,albuminuria_cat,acr_u,microalbumin_u,eGFR_CKiD_U25_avg,cystatin_c_s,creatinine_s,creatinine_u,erpf_bsa_plasma,erpf_bsa_plasma_urine,eGFR_CKiD_U25_avg,mm_si,pah_clear_bsa,avg_k_r2,avg_c_r2) %>% 
  # group_by(record_id, visit) %>%
  # tidyr::fill(date, .direction = "updown") %>% ungroup() %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id)) %>% 
  mutate(visit="baseline_screening") %>% 
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                                     race == "Black or African American" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                                     ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                                     T ~ "Not Hispanic or Latino Other"))
dat_collapsed2 <- dat %>%
  filter(study=="PANTHER") %>% 
  filter(visit=="year_1" | visit=="year_2") %>% 
  # dplyr::select(record_id, date, rh2_id, visit,mrn, group, age, sex, bmi, hba1c,albuminuria_cat,microalbumin_u,starts_with("egfr"),cystatin_c_s,starts_with("creat"),all_of(pfas)) %>% 
  dplyr::select(record_id, dob, date, rh2_id, visit,mrn, group, age, sex, bmi, race,ethnicity, race_ethnicity, hba1c,acr_u,albuminuria_cat,microalbumin_u,eGFR_CKiD_U25_avg,cystatin_c_s,creatinine_s,creatinine_u,erpf_bsa_plasma,erpf_bsa_plasma_urine,eGFR_CKiD_U25_avg,mm_si,pah_clear_bsa,avg_k_r2,avg_c_r2) %>% 
  # group_by(record_id, visit) %>%
  # tidyr::fill(date, .direction = "updown") %>% ungroup() %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id, visit)) %>% 
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                                     race == "Black or African American" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                                     ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                                     T ~ "Not Hispanic or Latino Other"))
meta <- rbind(dat_collapsed1,dat_collapsed2)

pfas_dat <- dat %>% 
  filter(study=="PANTHER") %>% 
  filter(visit=="baseline") %>% 
  dplyr::select("record_id",all_of(pfas)) %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id)) %>% 
  filter(!is.na(N.EtFOSAA)) #91 participants with baseline pfas data
rm(dat_collapsed1,dat_collapsed2,dat_followup1,dat_followup2)

#Formatted imputed pfas data
pfas_imputed <- readRDS("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/PANTHER/Data_Cleaned/PFAS_Data_Imputed_11_03.rds")

#Format metadata
# meta <- dat_collapsed
dat <- tidylog::left_join(pfas_imputed,meta,by="record_id")

#Particpants 
dat_baseline <- dat %>% 
  filter(visit=="baseline_screening")
table1(~age+sex+bmi+race_ethnicity_condensed+hba1c+microalbumin_u+acr_u+creatinine_s+creatinine_u+eGFR_CKiD_U25_avg|group,data=dat_baseline)
# saveRDS(dat_baseline,fs::path(dir.dat,"panther_pfas_baseline_12_17.rds"))

#Folowoup Year 1
dat_followup1 <- dat %>% 
  filter(visit=="year_1") %>% 
  dplyr::select(c("record_id","microalbumin_u","acr_u","eGFR_CKiD_U25_avg")) %>% 
  dplyr::rename(microalbumin_u_yr1=microalbumin_u,
                acr_u_yr1=acr_u,
                eGFR_CKiD_U25_avg_yr1=eGFR_CKiD_U25_avg)

#Followup Year 2
dat_followup2 <- dat %>% 
  filter(visit=="year_2") %>% 
  dplyr::select(c("record_id","microalbumin_u","acr_u","eGFR_CKiD_U25_avg")) %>% 
  dplyr::rename(microalbumin_u_yr2=microalbumin_u,
                acr_u_yr2=acr_u,
                eGFR_CKiD_U25_avg_yr2=eGFR_CKiD_U25_avg)

dat_all <- tidylog::left_join(dat_baseline,dat_followup1,by="record_id")
dat_all <- tidylog::left_join(dat_all,dat_followup2,by="record_id")
```

##c. Impute PFAS Data
```{r}
# t_lod <- t(lod)
# colnames(t_lod) <- t_lod[1,]
# t_lod <- as.data.frame(t_lod)
# t_lod <- t_lod[-1,]
# lod <- t_lod
# rm(t_lod)
# lod <- lod %>% 
#   mutate(across(everything(), ~as.numeric(.)))
# colnames(lod) <- str_replace_all(colnames(lod),"-",".")
# 
# # pfas
# N.EtFOSAA <- length(which(pfas_dat$N.EtFOSAA < lod$`N-EtFOSAA`))/length(unique(pfas_dat$record_id))
# N.EtFOSAA < 2/3 #False, exclude
# N.MeFOSAA <- length(which(pfas_dat$N.MeFOSAA < lod$`N-MeFOSAA`))/length(unique(pfas_dat$record_id))
# N.MeFOSAA < 2/3 #True, impute 
# PFBA <- length(which(pfas_dat$PFBA < lod$`PFBA`))/length(unique(pfas_dat$record_id))
# PFBA < 2/3 #False, exclude
# PFDA <- length(which(pfas_dat$PFDA < lod$`PFDA`))/length(unique(pfas_dat$record_id))
# PFDA < 2/3 #True, impute
# PFDoA <- length(which(pfas_dat$PFDoA < lod$`PFDoA`))/length(unique(pfas_dat$record_id))
# PFDoA < 2/3 #False, exclude
# PFHpA <- length(which(pfas_dat$PFHpA < lod$`PFHpA`))/length(unique(pfas_dat$record_id))
# PFHpA < 2/3 #True, impute
# PFHxA <- length(which(pfas_dat$PFHxA < lod$`PFHxA`))/length(unique(pfas_dat$record_id))
# PFHxA < 2/3 #False, exclude
# PFNA <- length(which(pfas_dat$PFNA < lod$`PFNA`))/length(unique(pfas_dat$record_id))
# PFNA < 2/3 #True, impute
# PFOA <- length(which(pfas_dat$PFOA < lod$`PFOA`))/length(unique(pfas_dat$record_id))
# PFOA < 2/3 #True, impute
# PFBS <- length(which(pfas_dat$PFBS < lod$`PFBS`))/length(unique(pfas_dat$record_id))
# PFBS < 2/3 #True, impute
# PFPeA <- length(which(pfas_dat$PFPeA < lod$`PFPeA`))/length(unique(pfas_dat$record_id))
# PFPeA < 2/3 #False, exclude
# PFTeDA <- length(which(pfas_dat$PFTeDA < lod$`PFTeDA`))/length(unique(pfas_dat$record_id))
# PFTeDA < 2/3 #False, exclude
# PFTrDA <- length(which(pfas_dat$PFTrDA < lod$`PFTrDA`))/length(unique(pfas_dat$record_id))
# PFTrDA < 2/3 #False, exclude
# PFUnA <- length(which(pfas_dat$PFUnA < lod$`PFUnA`))/length(unique(pfas_dat$record_id))
# PFUnA < 2/3 #False, exclude
# PFBS <- length(which(pfas_dat$PFBS < lod$`PFBS`))/length(unique(pfas_dat$record_id))
# PFBS < 2/3 #True, impute
# PFDoS <- length(which(pfas_dat$PFDoS < lod$`PFDoS`))/length(unique(pfas_dat$record_id))
# PFDoS < 2/3 #False, exclude
# PFHps <- length(which(pfas_dat$PFHps < lod$`PFHps`))/length(unique(pfas_dat$record_id))
# PFHps < 2/3 #True, impute
# PFHxS <- length(which(pfas_dat$PFHxS < lod$`PFHxS`))/length(unique(pfas_dat$record_id))
# PFHxS < 2/3 #True, impute
# PFNS <- length(which(pfas_dat$PFNS < lod$`PFNS`))/length(unique(pfas_dat$record_id))
# PFNS < 2/3 #FALSE, exclude
# PFOS <- length(which(pfas_dat$PFOS < lod$`PFOS`))/length(unique(pfas_dat$record_id))
# PFOS < 2/3 #True, impute
# PFOSA <- length(which(pfas_dat$PFOSA < lod$`PFOSA`))/length(unique(pfas_dat$record_id))
# PFOSA < 2/3 #False, exclude
# PFPeAS <- length(which(pfas_dat$PFPeAS < lod$`PFPeAS`))/length(unique(pfas_dat$record_id))
# PFPeAS < 2/3 #True, impute
# 
# pfas_to_impute <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA","PFOA" ,"PFBS", 
#                     "PFHps", "PFHxS", "PFOS", "PFPeAS")
# pfas_imputed <- pfas_dat %>% 
#   dplyr::select(c("record_id",all_of(pfas_to_impute)))
# for (pfas in pfas_to_impute){
#   pfas_imputed[[pfas]] <- ifelse(pfas_imputed[[pfas]]<lod[[pfas]],(lod[[pfas]]/sqrt(2)),pfas_imputed[[pfas]])
# }

#Save formatted and imputed pfas data
# saveRDS(pfas_imputed,fs::path(dir.dat,"PFAS_Data_Imputed_12_17.rds"))
# pfas_imputed <- readRDS("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/PANTHER/Data_Cleaned/PFAS_Data_Imputed_11_03.rds")
# pfas_to_impute
table1(~N.MeFOSAA+PFDA+PFHpA+PFNA+PFOA+PFBS+PFHps+PFHxS+PFOS+PFPeAS,data=pfas_imputed,render.continuous = render.geometric)

# #Format metadata
# # meta <- dat_collapsed
# dat <- tidylog::left_join(pfas_imputed,meta,by="record_id")
# 
# #Particpants 
# dat_baseline <- dat %>% 
#   filter(visit=="baseline_screening")
# table1(~age+sex+bmi+race_ethnicity_condensed+hba1c+microalbumin_u+acr_u+creatinine_s+creatinine_u+eGFR_CKiD_U25_avg|group,data=dat_baseline)
# # saveRDS(dat_baseline,fs::path(dir.dat,"panther_pfas_baseline_12_17.rds"))
# 
# #Folowoup Year 1
# dat_followup1 <- dat %>% 
#   filter(visit=="year_1") %>% 
#   dplyr::select(c("record_id","microalbumin_u","acr_u","eGFR_CKiD_U25_avg")) %>% 
#   dplyr::rename(microalbumin_u_yr1=microalbumin_u,
#                 acr_u_yr1=acr_u,
#                 eGFR_CKiD_U25_avg_yr1=eGFR_CKiD_U25_avg)
# 
# #Followup Year 2
# dat_followup2 <- dat %>% 
#   filter(visit=="year_2") %>% 
#   dplyr::select(c("record_id","microalbumin_u","acr_u","eGFR_CKiD_U25_avg")) %>% 
#   dplyr::rename(microalbumin_u_yr2=microalbumin_u,
#                 acr_u_yr2=acr_u,
#                 eGFR_CKiD_U25_avg_yr2=eGFR_CKiD_U25_avg)
# 
# dat_all <- tidylog::left_join(dat_baseline,dat_followup1,by="record_id")
# dat_all <- tidylog::left_join(dat_all,dat_followup2,by="record_id")

table1(~N.MeFOSAA+PFDA+PFHpA+PFNA+PFBS+PFHps+PFHxS+PFOS+PFPeAS,data=pfas_imputed,render.continuous = render.geometric)

table1(~age+sex+bmi+race_ethnicity_condensed+hba1c+microalbumin_u+microalbumin_u_yr1+microalbumin_u_yr2+acr_u+
         acr_u_yr1+acr_u_yr2+creatinine_s+creatinine_u+eGFR_CKiD_U25_avg+eGFR_CKiD_U25_avg_yr1+eGFR_CKiD_U25_avg_yr2|group,data=dat_all)
table1(~N.MeFOSAA+PFDA+PFHpA+PFNA+PFBS+PFHps+PFHxS+PFOA+PFOS+PFPeAS|group,data=dat_all,render.continuous = render.geometric)

# #Filter out albumin and acru outlier
# dat_baseline2 <- dat_baseline %>%
#   mutate(microalbumin_u=ifelse(microalbumin_u>3000,NA,microalbumin_u)) %>%
#   mutate(acr_u=ifelse(acr_u>2367,NA,acr_u))
# #Filter out albumin outlier and acru outlier
# dat2 <- dat %>%
#   mutate(microalbumin_u=ifelse(microalbumin_u>3000,NA,microalbumin_u)) %>%
#   mutate(acr_u=ifelse(acr_u>2367,NA,acr_u))
# dat_all2 <- dat_all %>% 
#   mutate(microalbumin_u=ifelse(microalbumin_u>3000,NA,microalbumin_u)) %>% 
#   mutate(acr_u=ifelse(acr_u>2367,NA,acr_u))
# 
# label(dat_all2$age) <- "Age (y)"
# label(dat_all2$sex) <- "Sex"
# label(dat_all2$bmi) <- "BMI (kg/m2)"
# label(dat_all2$race_ethnicity_condensed) <- "Race/Ethnicity"
# label(dat_all2$hba1c) <- "HbA1c (%)"
# label(dat_all2$microalbumin_u) <- "Urine Microalbumin (mcg/mL) - Baseline"
# label(dat_all2$microalbumin_u_yr1) <- "Urine Microalbumin (mcg/mL) - Year 1"
# label(dat_all2$microalbumin_u_yr2) <- "Urine Microalbumin (mcg/mL) - Year 2"
# label(dat_all2$acr_u) <- "Urine Albumin-Creatinine Ratio (mg/g) - Baseline"
# label(dat_all2$acr_u_yr1) <- "Urine Albumin-Creatinine Ratio	(mg/g) - Year 1"
# label(dat_all2$acr_u_yr2) <- "Urine Albumin-Creatinine Ratio	(mg/g) - Year 2"
# label(dat_all2$creatinine_s) <- "Serum Creatinine (mg/dL)"
# label(dat_all2$creatinine_u) <- "Urine Creatinine (mg/dL)"
# label(dat_all2$eGFR_CKiD_U25_avg) <- "eGFR CKiD U25  (mL/min/1.73 m²) - Baseline"
# label(dat_all2$eGFR_CKiD_U25_avg_yr1) <- "eGFR CKiD U25 (mL/min/1.73 m²) - Year 1"
# label(dat_all2$eGFR_CKiD_U25_avg_yr2) <- "eGFR CKiD U25 (mL/min/1.73 m²) - Year 2"
# 
# table1(~age+sex+bmi+race_ethnicity_condensed+hba1c+microalbumin_u+microalbumin_u_yr1+microalbumin_u_yr2+acr_u+acr_u_yr1+acr_u_yr2+creatinine_s+creatinine_u+eGFR_CKiD_U25_avg+eGFR_CKiD_U25_avg_yr1+eGFR_CKiD_U25_avg_yr2|group,data=dat_all2)
# table1(~ age + sex + bmi + race_ethnicity_condensed + hba1c + 
#          microalbumin_u + microalbumin_u_yr1 + microalbumin_u_yr2 + 
#          acr_u + acr_u_yr1 + acr_u_yr2 + creatinine_s + creatinine_u + 
#          eGFR_CKiD_U25_avg + eGFR_CKiD_U25_avg_yr1 + eGFR_CKiD_U25_avg_yr2 | group, 
#        data = dat_all2,
#        render.continuous = c("Mean (SD)" = "Mean (SD)"))
# 
# table1(~N.MeFOSAA+PFDA+PFHpA+PFNA+PFBS+PFHps+PFHxS+PFOA+PFOS+PFPeAS|group,data=dat_all2,render.continuous = render.geometric)
```


#2. Exposure assessment analysis (X = PFAS, Y = ACRu)
```{r}
pdf("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results/PFAS Results/pfas_acru_plots_no_outlier_withPFOA.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOA","PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("acr_u~",pfas,"+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=.data[[pfas]],y=acr_u))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between PFAS and ACRu, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,3), ", pval = ", round(pval,3)),
         x = pfas,  # This will use the variable name from pfas
         y = "ACRu")
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results/PFAS Results/pfas_microalbumin_plots_no_outlier.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("microalbumin_u~",pfas,"+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=.data[[pfas]],y=microalbumin_u))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between PFAS and Microalbuminuria, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,3), ", pval = ", round(pval,3)),
         x = pfas,  # This will use the variable name from pfas
         y = "Microalbumin (U)")
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results/PFAS Results/pfas_eGFR_plots.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("eGFR_CKiD_U25_avg~",pfas,"+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=.data[[pfas]],y=eGFR_CKiD_U25_avg))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between PFAS and eGFR, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,3), ", pval = ", round(pval,3)),
         x = pfas,  # This will use the variable name from pfas
         y = "eGFR")
  print(p)
  
}
dev.off()

#Mixed Model Analysis with Time
dat2$time_numeric <- as.numeric(factor(dat2$visit)) - 1  # Start at 0 for baseline
m0 <- as.formula(paste0("acr_u ~ time + (1|record_id)"))
model_simple <- lmer(m0, data = dat2)
summary(model_simple)$coef

# Scatter plot with individual trajectories
p3 <- ggplot(dat2, aes(x = time, y = acr_u)) +
  geom_point(aes(color = factor(visit)), alpha = 0.6) +
  geom_path(aes(group = record_id), alpha = 0.2) +
  geom_smooth(method = "lm", color = "red4", linewidth= 0.5,se=F) +
  scale_color_viridis_d(name = "Time") +
  theme_minimal() +
  labs(title = "ACRu Levels over Time",
       subtitle = "Beta = -3.15, P-value = 0.418",
       x = "Time",
       y = "acr_u Level") +
  theme(plot.title = element_text(face = "bold"))
# legend.position="none")

print(p3)

#Baseline PFAS with Follow-Up ACRu
#Exposure assessment: 
pdf("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results/PFAS Results/baseline_pfas_followup_acru_yr1.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOA","PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("acr_u_yr1~",pfas,"+age+sex+acr_u"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_all2,aes(x=.data[[pfas]],y=acr_u_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between Baseline PFAS and ACRu 1 Year Follow-Up, adj. Age, Sex and Baseline ACRu"),
         subtitle = paste0("Beta = ", round(beta,3), ", pval = ", round(pval,3)),
         x = pfas,  # This will use the variable name from pfas
         y = "ACRu (1 Yr Follow-Up)")
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results/PFAS Results/baseline_pfas_followup_acru_yr2.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOA","PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("acr_u_yr1~",pfas,"+age+sex+acr_u"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_all2,aes(x=.data[[pfas]],y=acr_u_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between Baseline PFAS and ACRu 2 Year Follow-Up, adj. Age, Sex and Baseline ACRu"),
         subtitle = paste0("Beta = ", round(beta,3), ", pval = ", round(pval,3)),
         x = pfas,  # This will use the variable name from pfas
         y = "ACRu (2 Yr Follow-Up)")
  print(p)
  
}
dev.off()

```

#3. Exposure assessment analysis (X = ACRu, Y = PFAS)
```{r}
pdf("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/Hailey Hampson/1_Ongoing Projects/PFAS & IR in Panther/Results/Baseline_ACRu_PFAS.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOA","PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0(pfas,"~acr_u","+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=acr_u,y=.data[[pfas]]))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between ACRu and PFAS, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,5), ", pval = ", round(pval,3)),
         x = "ACRu",  # This will use the variable name from pfas
         y = pfas)
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/Hailey Hampson/1_Ongoing Projects/PFAS & IR in Panther/Results/Baseline_Microalbuminuria_PFAS.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0(pfas,"~microalbumin_u","+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=microalbumin_u,y=.data[[pfas]]))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between Microalbuminuria and PFAS, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,5), ", pval = ", round(pval,3)),
         x = "Microalbumin (U)",  # This will use the variable name from pfas
         y = pfas)
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/Hailey Hampson/1_Ongoing Projects/PFAS & IR in Panther/Results/Baseline_eGFR_PFAS.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0(pfas,"~eGFR_CKiD_U25_avg+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=eGFR_CKiD_U25_avg,y=.data[[pfas]]))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between eGFR and PFAS, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,5), ", pval = ", round(pval,3)),
         x = "eGFR",  # This will use the variable name from pfas
         y = pfas)
  print(p)
  
}
dev.off()

#Mixed Model Analysis with Time
dat2$time_numeric <- as.numeric(factor(dat2$visit)) - 1  # Start at 0 for baseline
m0 <- as.formula(paste0("acr_u ~ time + (1|record_id)"))
model_simple <- lmer(m0, data = dat2)
summary(model_simple)$coef

# Scatter plot with individual trajectories
p3 <- ggplot(dat2, aes(x = time, y = acr_u)) +
  geom_point(aes(color = factor(visit)), alpha = 0.6) +
  geom_path(aes(group = record_id), alpha = 0.2) +
  geom_smooth(method = "lm", color = "red4", linewidth= 0.5,se=F) +
  scale_color_viridis_d(name = "Time") +
  theme_minimal() +
  labs(title = "ACRu Levels over Time",
       subtitle = "Beta = -3.15, P-value = 0.418",
       x = "Time",
       y = "acr_u Level") +
  theme(plot.title = element_text(face = "bold"))
# legend.position="none")

print(p3)

#PFAS
dat2$time_numeric <- as.numeric(factor(dat2$visit)) - 1  # Start at 0 for baseline
m0 <- as.formula(paste0("acr_u ~ time + (1|record_id)"))
model_simple <- lmer(m0, data = dat2)
summary(model_simple)$coef

# Scatter plot with individual trajectories
p3 <- ggplot(dat2, aes(x = time, y = acr_u)) +
  geom_point(aes(color = factor(visit)), alpha = 0.6) +
  geom_path(aes(group = record_id), alpha = 0.2) +
  geom_smooth(method = "lm", color = "red4", linewidth= 0.5,se=F) +
  scale_color_viridis_d(name = "Time") +
  theme_minimal() +
  labs(title = "ACRu Levels over Time",
       subtitle = "Beta = -3.15, P-value = 0.418",
       x = "Time",
       y = "acr_u Level") +
  theme(plot.title = element_text(face = "bold"))
# legend.position="none")

print(p3)

#Baseline PFAS with Follow-Up ACRu
#Exposure assessment: 
pdf("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results/PFAS Results/baseline_pfas_followup_acru_yr1.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOA","PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0(pfas,"~acr_u_yr1+age+sex+acr_u"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_all2,aes(x=.data[[pfas]],y=acr_u_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between Baseline PFAS and ACRu 1 Year Follow-Up, adj. Age, Sex and Baseline ACRu"),
         subtitle = paste0("Beta = ", round(beta,3), ", pval = ", round(pval,3)),
         x = pfas,  # This will use the variable name from pfas
         y = "ACRu (1 Yr Follow-Up)")
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results/PFAS Results/baseline_pfas_followup_acru_yr2.pdf", width = 8, height = 6)
#Baseline Analysis
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
               "PFHps", "PFHxS", "PFOA","PFOS", "PFPeAS")
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("acr_u_yr1~",pfas,"+age+sex+acr_u"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_all2,aes(x=.data[[pfas]],y=acr_u_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between Baseline PFAS and ACRu 2 Year Follow-Up, adj. Age, Sex and Baseline ACRu"),
         subtitle = paste0("Beta = ", round(beta,3), ", pval = ", round(pval,3)),
         x = pfas,  # This will use the variable name from pfas
         y = "ACRu (2 Yr Follow-Up)")
  print(p)
  
}
dev.off()

#Log and Scale Variables
```

##a.Log and Scale Variables
```{r}
# Natural log transform and scale the PFAS variables
dat_baseline2 <- dat_baseline2 %>%
  mutate(across(all_of(pfas_vars), ~ scale(log(.))[,1], .names = "{.col}_ln_scaled"))
pfas_vars_log_scale <- paste0(pfas_vars,"_ln_scaled")
pdf("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/Hailey Hampson/1_Ongoing Projects/PFAS & IR in Panther/Results/Baseline_ACRu_PFAS_Log_Scaled.pdf", width = 8, height = 6)
#Baseline Analysis
# pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
#                "PFHps", "PFHxS", "PFOA","PFOS", "PFPeAS")
for (pfas in pfas_vars_log_scale) {
  m0 <- as.formula(paste0(pfas,"~acr_u","+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=acr_u,y=.data[[pfas]]))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between ACRu and PFAS, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,5), ", pval = ", round(pval,3)),
         x = "ACRu",  # This will use the variable name from pfas
         y = pfas)
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/Hailey Hampson/1_Ongoing Projects/PFAS & IR in Panther/Results/Baseline_Microalbuminuria_PFAS_Log_Scaled.pdf", width = 8, height = 6)
#Baseline Analysis
# pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
#                     "PFHps", "PFHxS", "PFOS", "PFPeAS")
for (pfas in pfas_vars_log_scale) {
  m0 <- as.formula(paste0(pfas,"~microalbumin_u","+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=microalbumin_u,y=.data[[pfas]]))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between Microalbuminuria and PFAS, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,5), ", pval = ", round(pval,3)),
         x = "Microalbumin (U)",  # This will use the variable name from pfas
         y = pfas)
  print(p)
  
}
dev.off()

pdf("/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/Hailey Hampson/1_Ongoing Projects/PFAS & IR in Panther/Results/Baseline_eGFR_PFAS_Log_Scaled.pdf", width = 8, height = 6)
#Baseline Analysis
# pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS", 
#                "PFHps", "PFHxS", "PFOS", "PFPeAS")
for (pfas in pfas_vars_log_scale) {
  m0 <- as.formula(paste0(pfas,"~eGFR_CKiD_U25_avg+age+sex"))
  m1 <- lm(m0,dat=dat_baseline2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  
  p <- ggplot(data=dat_baseline2,aes(x=eGFR_CKiD_U25_avg,y=.data[[pfas]]))+
    geom_point()+
    geom_smooth(method="lm",se=F)+
    theme_bw()+
    labs(title = c("Association between eGFR and PFAS, adj. Age & Sex"),
         subtitle = paste0("Beta = ", round(beta,5), ", pval = ", round(pval,3)),
         x = "eGFR",  # This will use the variable name from pfas
         y = pfas)
  print(p)
  
}
dev.off()
```

#4.Changes over time in the outcomes by group
```{r}
outcome_vars <- c("erpf_bsa_plasma","erpf_bsa_plasma_urine","eGFR_CKiD_U25_avg",
                  "mm_si","pah_clear_bsa","acr_u","avg_c_r2","avg_k_r2")

label(dat$avg_c_r2) <- "Average Cortical R2*"
label(dat$avg_k_r2) <- "Average Kidney R2* "
label(dat$erpf_bsa_plasma)      <- "BSA indexed ERPF from plasma clearance (mL/min/1.73m\u00b2)"
label(dat$erpf_bsa_plasma_urine)<- "BSA indexed ERPF from plasma and urine clearance (mL/min/1.73m\u00b2)"
label(dat$eGFR_CKiD_U25_avg)   <- "eGFR from CKiD U25 equation (mL/min/1.73m\u00b2)"
label(dat$mm_si)                <- "Insulin sensitivity (\u03bcU/mL)\u207b\u00b9\u00b7min\u207b\u00b9"
label(dat$pah_clear_bsa)        <- "BSA Normalized PAH Clearance (Plasma) (mL/min/1.73m\u00b2)"
label(dat$acr_u)                <- "Urine Albumin-Creatinine Ratio (mg/g)"

group_var    <- "group"
group_colors <- c("#C2185B", "#E8634A", "#F5A623")

# Output directory for figures
fig_dir <- fs::path(dir.dat, "PANTHER/Results/PFAS Results/figures")
fs::dir_create(fig_dir)

outlier_log <- list()
plot_list   <- list()
fig_paths   <- list()  # track saved PNG paths

# ── Identify baseline-only variables ─────────────────────────────────────────
baseline_only_vars <- outcome_vars[sapply(outcome_vars, function(var) {
  dat %>%
    filter(!is.na(.data[[var]])) %>%
    pull(visit) %>%
    unique() %>%
    setdiff("baseline_screening") %>%
    length() == 0
})]

# ── Main loop ─────────────────────────────────────────────────────────────────
for (var in outcome_vars) {
  
  var_label        <- if (!is.null(label(dat[[var]]))) label(dat[[var]]) else var
  is_baseline_only <- var %in% baseline_only_vars
  
  # ── Outlier removal ──────────────────────────────────────────────────────────
  dat_var  <- dat %>%
    filter(!is.na(.data[[var]]), !is.na(.data[[group_var]]))
  
  mean_val <- mean(dat_var[[var]], na.rm = TRUE)
  sd_val   <- sd(dat_var[[var]],   na.rm = TRUE)
  
  outliers <- dat_var %>%
    filter(abs(.data[[var]] - mean_val) > 3 * sd_val) %>%
    select(record_id, visit, all_of(group_var), all_of(var)) %>%
    mutate(variable = var, var_label = var_label)
  
  if (nrow(outliers) > 0) outlier_log[[var]] <- outliers
  
  dat_clean <- dat_var %>%
    filter(abs(.data[[var]] - mean_val) <= 3 * sd_val) %>%
    dplyr::rename(group = all_of(group_var))
  
  # ── Baseline N per group (for facet strip labels only) ───────────────────────
  n_df <- dat_clean %>%
    filter(visit == "baseline_screening") %>%
    dplyr::count(group) %>%
    mutate(n_label = paste0(group, "\nn = ", n))
  
  n_lookup <- setNames(n_df$n_label, n_df$group)
  
  # ── Footnote ─────────────────────────────────────────────────────────────────
  footnote <- if (var %in% names(outlier_log)) {
    ol <- outlier_log[[var]]
    paste0("Outlier(s) removed: ",
           paste(paste0(ol$record_id, " (", ol[[group_var]],
                        ", value = ", round(ol[[var]], 2), ")"),
                 collapse = "; "))
  } else NULL
  
  # ── Build plot ────────────────────────────────────────────────────────────────
  if (is_baseline_only) {
    
    dat_base <- dat_clean %>% filter(visit == "baseline_screening")
    
    p <- ggplot(dat_base, aes(x = group, y = .data[[var]], color = group)) +
      geom_boxplot(outlier.shape = NA, width = 0.5, linewidth = 0.8) +
      geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
      scale_color_manual(values = group_colors, name = "Group") +
      scale_x_discrete(labels = n_lookup) +
      labs(x = NULL, y = var,
           title   = paste("Baseline distribution:", var_label),
           caption = footnote) +
      theme_bw(base_size = 12) +
      theme(legend.position = "none",
            plot.title   = element_text(face = "bold"),
            plot.caption = element_text(color = "firebrick", size = 9))
    
  } else {
    
    summary_df <- dat_clean %>%
      group_by(visit, group) %>%
      summarise(
        mean_val = mean(.data[[var]], na.rm = TRUE),
        se_val   = sd(.data[[var]], na.rm = TRUE) / sqrt(n()),
        n        = n(),
        t_crit   = qt(0.975, df = n - 1),
        ci_lo    = mean_val - t_crit * se_val,
        ci_hi    = mean_val + t_crit * se_val,
        .groups  = "drop"
      )
    
    p <- ggplot() +
      geom_line(data = dat_clean,
                aes(x = visit, y = .data[[var]], group = record_id, color = group),
                alpha = 0.25, linewidth = 0.35) +
      geom_point(data = dat_clean,
                 aes(x = visit, y = .data[[var]], color = group),
                 alpha = 0.25, size = 0.7) +
      geom_line(data = summary_df,
                aes(x = visit, y = mean_val, group = group, color = group),
                linewidth = 1.4) +
      geom_point(data = summary_df,
                 aes(x = visit, y = mean_val, color = group),
                 size = 3) +
      geom_errorbar(data = summary_df,
                    aes(x = visit, ymin = ci_lo, ymax = ci_hi, color = group),
                    width = 0.15, linewidth = 0.9) +
      scale_color_manual(values = group_colors, name = "Group") +
      facet_wrap(~ group, nrow = 1,
                 labeller = labeller(group = n_lookup)) +
      labs(x = "Visit", y = var,
           # title   = paste("Longitudinal trajectories:", var_label),
           caption = paste0(
             "Faint lines: individual trajectories  |  Bold: mean \u00b1 95% CI",
             if (!is.null(footnote)) paste0("\n", footnote) else ""
           )) +
      theme_bw(base_size = 12) +
      theme(legend.position  = "none",
            plot.title       = element_text(face = "bold"),
            strip.text       = element_text(face = "bold"),
            strip.background = element_rect(fill = "grey92"),
            plot.caption     = element_text(
              color = ifelse(!is.null(footnote), "firebrick", "grey40"),
              size  = 9))
  }
  
  plot_list[[var]] <- p
  
  # ── Save PNG ──────────────────────────────────────────────────────────────────
  fig_path <- fs::path(fig_dir, paste0(var, ".png"))
  ggsave(fig_path, plot = p, width = 10, height = 5, dpi = 300)
  fig_paths[[var]] <- fig_path
}

# ── Save outlier log and fig_paths for Quarto ─────────────────────────────────
saveRDS(outlier_log, fs::path(fig_dir, "outlier_log.rds"))
saveRDS(fig_paths,   fs::path(fig_dir, "fig_paths.rds"))


```

#5. Associate PFAS with ERPF, GFR, Insulin Sensitivity, Renal Perfusion, renal oxygenation, adj. for baseline ACRu
##a. Cross sectional analysis baseline, adj. baseline ACRu, age, sex & group - scatterplots & tables
```{r}

##a. Cross sectional analysis baseline - scatterplots & tables
# ── Baseline PFAS Association Plots ──────────────────────────────────────────

dat_baseline <- dat %>% filter(visit == "baseline_screening")

# ── Log transform and scale PFAS ─────────────────────────────────────────────
pfas_vars <- c("N.MeFOSAA", "PFDA", "PFHpA", "PFNA", "PFBS",
               "PFHps", "PFHxS", "PFOA", "PFOS", "PFPeAS")

dat_baseline <- dat_baseline %>%
  mutate(across(all_of(pfas_vars), ~ scale(log(.))[,1], .names = "{.col}_ln_scaled"))

pfas_vars_log_scale <- paste0(pfas_vars, "_ln_scaled")

outcome_vars <- c("erpf_bsa_plasma", "erpf_bsa_plasma_urine", "eGFR_CKiD_U25_avg",
                  "mm_si", "pah_clear_bsa", "acr_u", "avg_c_r2", "avg_k_r2")

# ── Output directory ──────────────────────────────────────────────────────────
out_dir <- fs::path(dir.results, "PANTHER/Results/PFAS Results")
fs::dir_create(out_dir)

# ── Helpers ───────────────────────────────────────────────────────────────────
remove_outliers_baseline <- function(df, outcome_var) {
  m  <- mean(df[[outcome_var]], na.rm = TRUE)
  s  <- sd(df[[outcome_var]],   na.rm = TRUE)
  removed <- df %>%
    filter(abs(.data[[outcome_var]] - m) > 3 * s) %>%
    pull(record_id)
  clean <- df %>% filter(abs(.data[[outcome_var]] - m) <= 3 * s)
  list(data = clean, removed = removed)
}

get_label <- function(df, var) {
  lbl <- attr(df[[var]], "label")
  if (!is.null(lbl) && nchar(lbl) > 0) lbl else var
}

# ── Model specifications ──────────────────────────────────────────────────────
models <- list(
  crude       = list(covars = NULL,            suffix = "crude"),
  adj_age_sex = list(covars = c("age", "sex"), suffix = "adj_age_sex"),
  adj_group   = list(covars = "group",         suffix = "adj_group")
)

# ── Step 1: Collect all raw p-values first (needed for FDR) ──────────────────
cs_results_raw <- list()

for (model_name in names(models)) {
  covars <- models[[model_name]]$covars
  suffix <- models[[model_name]]$suffix
  
  for (outcome in outcome_vars) {
    outcome_label <- get_label(dat_baseline, outcome)
    
    ol_result <- remove_outliers_baseline(
      dat_baseline %>% filter(!is.na(.data[[outcome]])), outcome
    )
    dat_clean   <- ol_result$data
    removed_ids <- ol_result$removed
    
    for (pfas in pfas_vars_log_scale) {
      pfas_raw  <- gsub("_ln_scaled", "", pfas)
      dat_plot  <- dat_clean %>% filter(!is.na(.data[[pfas]]))
      
      formula_str <- if (!is.null(covars)) {
        paste0(outcome, " ~ ", pfas, " + ", paste(covars, collapse = " + "))
      } else {
        paste0(outcome, " ~ ", pfas)
      }
      
      m1   <- lm(as.formula(formula_str), data = dat_plot)
      beta <- round(coef(summary(m1))[pfas, "Estimate"], 3)
      pval <- coef(summary(m1))[pfas, "Pr(>|t|)"]
      
      cs_results_raw[[paste(suffix, outcome, pfas_raw, sep = "_")]] <- data.frame(
        model         = suffix,
        outcome       = outcome,
        outcome_label = outcome_label,
        pfas          = pfas_raw,
        beta          = beta,
        pval          = pval,
        n             = nrow(dat_plot),
        removed       = ifelse(length(removed_ids) > 0,
                               paste(removed_ids, collapse = ", "), "None"),
        stringsAsFactors = FALSE
      )
    }
  }
}

# ── Step 2: Apply FDR correction within each model ───────────────────────────
cs_results_df <- bind_rows(cs_results_raw) %>%
  arrange(model, outcome, pfas) %>%
  group_by(model) %>%
  mutate(
    fdr      = p.adjust(pval, method = "fdr"),
    pval_fmt = ifelse(pval < 0.001, "<0.001", as.character(round(pval, 3))),
    fdr_fmt  = ifelse(fdr  < 0.001, "<0.001", as.character(round(fdr,  3))),
    sig      = fdr < 0.05
  ) %>%
  ungroup()

# Save
saveRDS(cs_results_df, fs::path(out_dir, "cs_results.rds"))
write.csv(cs_results_df, fs::path(out_dir, "cs_results.csv"), row.names = FALSE)
message("Cross-sectional results: ", nrow(cs_results_df), " rows, ",
        sum(cs_results_df$sig, na.rm = TRUE), " FDR-significant")

# ── Step 3: Regenerate plots with FDR-adjusted p-values on subtitle ───────────
for (model_name in names(models)) {
  covars <- models[[model_name]]$covars
  suffix <- models[[model_name]]$suffix
  
  model_dir <- fs::path(out_dir, suffix)
  fs::dir_create(model_dir)
  
  for (outcome in outcome_vars) {
    outcome_label <- get_label(dat_baseline, outcome)
    
    ol_result   <- remove_outliers_baseline(
      dat_baseline %>% filter(!is.na(.data[[outcome]])), outcome
    )
    dat_clean   <- ol_result$data
    removed_ids <- ol_result$removed
    
    n_total  <- nrow(dat_clean %>% filter(!is.na(.data[[outcome]])))
    footnote <- if (length(removed_ids) > 0) {
      paste0("n = ", n_total, "  |  Outlier(s) removed: ",
             paste(removed_ids, collapse = ", "))
    } else {
      paste0("n = ", n_total)
    }
    
    for (pfas in pfas_vars_log_scale) {
      pfas_raw   <- gsub("_ln_scaled", "", pfas)
      pfas_label <- paste0(pfas_raw, " (log-scaled)")
      
      dat_plot <- dat_clean %>% filter(!is.na(.data[[pfas]]))
      
      # Pull pre-computed beta, pval, fdr for this combination
      # res <- cs_results_df %>%
      #   filter(model == suffix, outcome == outcome, pfas == pfas_raw)
      res <- cs_results_df %>%
        filter(model == .env$suffix, 
               outcome == .env$outcome, 
               pfas == .env$pfas_raw)
      
      beta     <- res$beta
      pval_fmt <- res$pval_fmt
      fdr_fmt  <- res$fdr_fmt
      is_sig   <- res$sig
      
      p <- ggplot(dat_plot, aes(x = .data[[pfas]], y = .data[[outcome]])) +
        geom_point(aes(color = group), alpha = 0.6, size = 1.2) +
        geom_smooth(method = "lm", se = TRUE,
                    color     = "black",
                    fill      = "grey80",
                    alpha     = 0.2,
                    linewidth = 0.9) +
        scale_color_manual(values = group_colors, name = "Group") +
        labs(
          title    = outcome_label,
          subtitle = paste0(
            "\u03b2 = ", beta,
            "  |  p = ", pval_fmt,
            "  |  FDR = ", fdr_fmt,
            if (is_sig) "  \u2605" else "",   # star if FDR significant
            if (!is.null(covars))
              paste0("\nAdjusted for: ", paste(covars, collapse = ", "))
            else ""
          ),
          x       = pfas_label,
          y       = outcome,
          caption = footnote
        ) +
        theme_bw(base_size = 12) +
        theme(
          plot.title    = element_text(face = "bold", size = 11),
          plot.subtitle = element_text(
            size  = 9.5,
            color = ifelse(is_sig, "#C2185B", "grey30")   # pink if significant
          ),
          plot.caption     = element_text(
            size  = 8,
            color = ifelse(length(removed_ids) > 0, "firebrick", "grey50")
          ),
          legend.position  = "bottom",
          panel.grid.minor = element_blank()
        )
      
      fname <- paste0(pfas_raw, "_", outcome, "_", suffix, ".png")
      ggsave(fs::path(model_dir, fname),
             plot = p, width = 6, height = 5, dpi = 300)
    }
  }
  message("Done plots: ", suffix)
}
```

##b. Linear mixed model for baseline pfas and longitudinal primary outcomes, adj. baseline acru, age, sex & group - scatterplots & tables
###i. Models
```{r}
# Longitudinal outcome vars (exclude baseline-only)
outcome_vars_long <- setdiff(outcome_vars, baseline_only_vars)
outcome_vars_long <- setdiff(outcome_vars_long, "acr_u")

# ── Add log-scaled AND standardized PFAS to dat ───────────────────────────────
# (standardizing already happens via scale(), so ln_scaled = log + standardized)
dat <- dat %>%
  mutate(across(all_of(pfas_vars), ~ scale(log(.))[,1], .names = "{.col}_ln_scaled"))

pfas_vars_log_scale <- paste0(pfas_vars, "_ln_scaled")

# ── Build dat_long ────────────────────────────────────────────────────────────
dat_long <- dat %>%
  mutate(visit_num = case_when(
    visit == "baseline_screening" ~ 0,
    visit == "year_1"             ~ 1,
    visit == "year_2"             ~ 2
  ))

# Baseline ACR (time-invariant)
acr_baseline <- dat %>%
  filter(visit == "baseline_screening") %>%
  dplyr::select(record_id, acr_u_baseline = acr_u)

dat_long <- dat_long %>%
  left_join(acr_baseline, by = "record_id")

# ── Standardize outcomes AFTER outlier removal would be ideal but we do it  ───
# globally here for simplicity — betas will be in SD units
dat_long <- dat_long %>%
  mutate(across(all_of(outcome_vars_long), ~ scale(.)[,1],
                .names = "{.col}_z"))

outcome_vars_long_z <- paste0(outcome_vars_long, "_z")

# ── Model specifications ──────────────────────────────────────────────────────
models_lme <- list(
  crude           = list(covars = NULL,                                suffix = "crude"),
  adj_acr         = list(covars = "acr_u_baseline",                    suffix = "adj_acr_u"),
  adj_age_sex_acr = list(covars = c("age", "sex", "acr_u_baseline"),   suffix = "adj_age_sex_acr"),
  adj_group_acr   = list(covars = c("group", "acr_u_baseline"),        suffix = "adj_group_acr")
)

# ── Output directory ──────────────────────────────────────────────────────────
lme_dir <- fs::path(dir.results, "PANTHER/Results/PFAS Results/longitudinal_lme")
fs::dir_create(lme_dir)

lme_results <- list()

# ── Main loop ─────────────────────────────────────────────────────────────────
for (model_name in names(models_lme)) {
  
  covars <- models_lme[[model_name]]$covars
  suffix <- models_lme[[model_name]]$suffix
  
  for (outcome in outcome_vars_long) {
    
    outcome_z     <- paste0(outcome, "_z")       # standardized version for model
    outcome_label <- get_label(dat, outcome)     # label from original variable
    
    # ── Outlier removal on ORIGINAL scale before standardization ─────────────
    # (outliers already baked into _z since scale() was applied to full data)
    # Flag outliers using original variable
    dat_out <- dat_long %>%
      filter(!is.na(.data[[outcome]]), !is.na(visit_num))
    
    m_out <- mean(dat_out[[outcome]], na.rm = TRUE)
    s_out <- sd(dat_out[[outcome]],   na.rm = TRUE)
    
    removed_ids <- dat_out %>%
      filter(abs(.data[[outcome]] - m_out) > 3 * s_out) %>%
      pull(record_id) %>% unique()
    
    dat_out <- dat_out %>%
      filter(abs(.data[[outcome]] - m_out) <= 3 * s_out)
    
    for (pfas in pfas_vars_log_scale) {
      
      pfas_raw <- gsub("_ln_scaled", "", pfas)
      
      dat_model <- dat_out %>%
        filter(!is.na(.data[[pfas]]), !is.na(.data[[outcome_z]]))
      
      if (!is.null(covars)) {
        dat_model <- dat_model %>%
          filter(if_all(all_of(covars), ~ !is.na(.)))
      }
      
      n_subj <- dat_model %>% pull(record_id) %>% n_distinct()
      n_obs  <- nrow(dat_model)
      
      # ── Formula uses outcome_z (standardized) ─────────────────────────────
      fixed_covars <- if (!is.null(covars)) paste("+", paste(covars, collapse = " + ")) else ""
      formula_str  <- paste0(
        outcome_z, " ~ ", pfas, " + visit_num + ", pfas, ":visit_num",
        fixed_covars,
        " + (1 + visit_num | record_id)"
      )
      
      m1 <- tryCatch(
        lmer(as.formula(formula_str), data = dat_model, REML = TRUE),
        error = function(e) {
          lmer(as.formula(gsub("1 \\+ visit_num", "1", formula_str)),
               data = dat_model, REML = TRUE)
        }
      )
      
      coef_tbl         <- as.data.frame(coef(summary(m1)))
      interaction_term <- paste0(pfas, ":visit_num")
      
      beta_main <- round(coef_tbl[pfas, "Estimate"], 3)
      pval_main <- coef_tbl[pfas, "Pr(>|t|)"]
      beta_int  <- round(coef_tbl[interaction_term, "Estimate"], 3)
      pval_int  <- coef_tbl[interaction_term, "Pr(>|t|)"]
      
      lme_results[[paste(model_name, outcome, pfas, sep = "_")]] <- data.frame(
        model         = model_name,
        outcome       = outcome,
        outcome_label = outcome_label,
        pfas          = pfas_raw,
        beta_main     = beta_main,
        pval_main     = pval_main,
        beta_int      = beta_int,
        pval_int      = pval_int,
        n_subj        = n_subj,
        n_obs         = n_obs
      )
    }
  }
  message("Done LME: ", suffix)
}

# ── Save results ──────────────────────────────────────────────────────────────
lme_results_df <- bind_rows(lme_results) %>%
  arrange(model, outcome, pfas) %>%
  mutate(
    pval_main_fmt = ifelse(pval_main < 0.001, "<0.001", as.character(round(pval_main, 3))),
    pval_int_fmt  = ifelse(pval_int  < 0.001, "<0.001", as.character(round(pval_int,  3)))
  )

saveRDS(lme_results_df, fs::path(lme_dir, "lme_results.rds"))
write.csv(lme_results_df, fs::path(lme_dir, "lme_results.csv"), row.names = FALSE)

message("Results saved to: ", lme_dir)

```

```{r}
library(lme4)
library(lmerTest)

# Longitudinal outcome vars (exclude baseline-only)
outcome_vars_long <- setdiff(outcome_vars, baseline_only_vars)
outcome_vars_long <- setdiff(outcome_vars_long, "acr_u") 

# ── Add log-scaled PFAS to dat (needed for dat_long) ─────────────────────────
dat <- dat %>%
  mutate(across(all_of(pfas_vars), ~ scale(log(.))[,1], .names = "{.col}_ln_scaled"))

# visit_num for LME
dat_long <- dat %>%
  mutate(visit_num = case_when(
    visit == "baseline_screening" ~ 0,
    visit == "year_1"             ~ 1,
    visit == "year_2"             ~ 2
  ))

# Baseline ACR (time-invariant) — pull from baseline rows
acr_baseline <- dat %>%
  filter(visit == "baseline_screening") %>%
  dplyr::select(record_id, acr_u_baseline = acr_u)

dat_long <- dat_long %>%
  left_join(acr_baseline, by = "record_id")

# ── Model specifications ──────────────────────────────────────────────────────
models_lme <- list(
  crude           = list(covars = NULL,                                suffix = "crude"),
  adj_acr         = list(covars = "acr_u_baseline",                    suffix = "adj_acr_u"),
  adj_age_sex_acr = list(covars = c("age", "sex", "acr_u_baseline"),   suffix = "adj_age_sex_acr"),
  adj_group_acr   = list(covars = c("group", "acr_u_baseline"),        suffix = "adj_group_acr")
)

# ── Output directory ──────────────────────────────────────────────────────────
lme_dir <- fs::path(dir.results, "PANTHER/Results/PFAS Results/longitudinal_lme")
fs::dir_create(lme_dir)

# Standardize outcome variables so betas are comparable across outcomes
dat_long <- dat_long %>%
  mutate(across(all_of(outcome_vars_long), ~ scale(.)[,1],
                .names = "{.col}_z"))

# Update outcome_vars_long to use standardized versions
outcome_vars_long_z <- paste0(outcome_vars_long, "_z")

lme_results <- list()

# ── Main loop ─────────────────────────────────────────────────────────────────
for (model_name in names(models_lme)) {
  
  covars <- models_lme[[model_name]]$covars
  suffix <- models_lme[[model_name]]$suffix
  
  model_dir <- fs::path(lme_dir, suffix)
  fs::dir_create(model_dir)
  
  for (outcome in outcome_vars_long) {
    
    outcome_label <- get_label(dat, outcome)
    
    # ── Outlier removal (per outcome across all visits) ───────────────────────
    dat_out   <- dat_long %>% filter(!is.na(.data[[outcome]]), !is.na(visit_num))
    m_out     <- mean(dat_out[[outcome]], na.rm = TRUE)
    s_out     <- sd(dat_out[[outcome]],   na.rm = TRUE)
    removed_ids <- dat_out %>%
      filter(abs(.data[[outcome]] - m_out) > 3 * s_out) %>%
      pull(record_id) %>% unique()
    dat_out <- dat_out %>%
      filter(abs(.data[[outcome]] - m_out) <= 3 * s_out)
    
    for (pfas in pfas_vars_log_scale) {
      
      pfas_raw   <- gsub("_ln_scaled", "", pfas)
      
      dat_model <- dat_out %>%
        filter(!is.na(.data[[pfas]]), !is.na(.data[[outcome]]))
      
      if (!is.null(covars)) {
        dat_model <- dat_model %>%
          filter(if_all(all_of(covars), ~ !is.na(.)))
      }
      
      n_subj <- dat_model %>% pull(record_id) %>% n_distinct()
      n_obs  <- nrow(dat_model)
      
      footnote <- paste0(
        "n = ", n_subj, " subjects, ", n_obs, " observations",
        if (length(removed_ids) > 0)
          paste0("  |  Outlier subject(s) removed: ", paste(removed_ids, collapse = ", "))
        else ""
      )
      
      # ── Formula ───────────────────────────────────────────────────────────
      fixed_covars <- if (!is.null(covars)) paste("+", paste(covars, collapse = " + ")) else ""
      formula_str  <- paste0(
        outcome, " ~ ", pfas, " + visit_num + ", pfas, ":visit_num",
        fixed_covars,
        " + (1 + visit_num | record_id)"
      )
      
      # Fit — fall back to random intercept only if singular
      m1 <- tryCatch(
        lmer(as.formula(formula_str), data = dat_model, REML = TRUE),
        error = function(e) {
          lmer(as.formula(gsub("1 \\+ visit_num", "1", formula_str)),
               data = dat_model, REML = TRUE)
        }
      )
      
      coef_tbl         <- as.data.frame(coef(summary(m1)))
      interaction_term <- paste0(pfas, ":visit_num")
      
      beta_main <- round(coef_tbl[pfas, "Estimate"], 3)
      pval_main <- coef_tbl[pfas, "Pr(>|t|)"]
      plab_main <- ifelse(pval_main < 0.001, "p < 0.001", paste0("p = ", round(pval_main, 3)))
      
      beta_int  <- round(coef_tbl[interaction_term, "Estimate"], 3)
      pval_int  <- coef_tbl[interaction_term, "Pr(>|t|)"]
      plab_int  <- ifelse(pval_int < 0.001, "p < 0.001", paste0("p = ", round(pval_int, 3)))
      
      # Store
      lme_results[[paste(model_name, outcome, pfas, sep = "_")]] <- data.frame(
        model         = model_name,
        outcome       = outcome,
        outcome_label = outcome_label,
        pfas          = pfas_raw,
        beta_main     = beta_main,
        pval_main     = pval_main,
        beta_int      = beta_int,
        pval_int      = pval_int,
        n_subj        = n_subj,
        n_obs         = n_obs
      )
    }
  }
  message("Done LME: ", suffix)
}

# ── Save results ──────────────────────────────────────────────────────────────
lme_results_df <- bind_rows(lme_results) %>%
  arrange(model, outcome, pfas) %>%
  lme_results_df <- bind_rows(lme_results) %>%
  arrange(model, outcome, pfas) %>%
  # FDR correction within each model, adjusting for n_pfas x n_outcomes tests
  group_by(model) %>%
  mutate(
    fdr_main = p.adjust(pval_main, method = "fdr"),
    fdr_int  = p.adjust(pval_int,  method = "fdr")
  ) %>%
  ungroup() %>%
  mutate(
    pval_main_fmt = ifelse(pval_main < 0.001, "<0.001", as.character(round(pval_main, 3))),
    pval_int_fmt  = ifelse(pval_int  < 0.001, "<0.001", as.character(round(pval_int,  3))),
    fdr_main_fmt  = ifelse(fdr_main  < 0.001, "<0.001", as.character(round(fdr_main,  3))),
    fdr_int_fmt   = ifelse(fdr_int   < 0.001, "<0.001", as.character(round(fdr_int,   3))),
    sig_main      = fdr_main < 0.05,
    sig_int       = fdr_int  < 0.05
  )

saveRDS(lme_results_df, fs::path(lme_dir, "lme_results.rds"))
write.csv(lme_results_df, fs::path(lme_dir, "lme_results.csv"), row.names = FALSE)

message("Results saved to: ", lme_dir)
```
###ii. Display results
```{r}
# ── Heatmap of PFAS-outcome associations across visits ────────────────────────
library(ggplot2)
library(tidyr)

# Choose which model to visualize — change as needed
model_to_plot <- "crude"  # or "adj_age_sex_acr", "adj_group_acr" etc.

heatmap_df <- lme_results_df %>%
  filter(model == model_to_plot) %>%
  mutate(
    beta_visit0 = beta_main,
    beta_visit1 = beta_main + beta_int * 1,
    beta_visit2 = beta_main + beta_int * 2
  ) %>%
  select(pfas, outcome, beta_visit0, beta_visit1, beta_visit2,
         pval_main, pval_int) %>%
  pivot_longer(
    cols      = c(beta_visit0, beta_visit1, beta_visit2),
    names_to  = "visit",
    values_to = "beta"
  ) %>%
  mutate(
    visit = factor(visit,
                   levels = c("beta_visit0", "beta_visit1", "beta_visit2"),
                   labels = c("Baseline", "Year 1", "Year 2")),
    # significance markers based on most relevant p-value per visit
    sig = case_when(
      visit == "Baseline" & pval_main < 0.001 ~ "***",
      visit == "Baseline" & pval_main < 0.01  ~ "**",
      visit == "Baseline" & pval_main < 0.05  ~ "*",
      visit != "Baseline" & pval_int  < 0.001 ~ "***",
      visit != "Baseline" & pval_int  < 0.01  ~ "**",
      visit != "Baseline" & pval_int  < 0.05  ~ "*",
      TRUE ~ ""
    ),
    # Clean up labels
    pfas    = gsub("_ln_scaled", "", pfas),
    outcome = factor(outcome,
                     levels = outcome_vars_long,
                     labels = sapply(outcome_vars_long, get_label, df = dat))
  )

# Symmetric color scale centered at 0
beta_limit <- max(abs(heatmap_df$beta), na.rm = TRUE)

p_heatmap <- ggplot(heatmap_df,
                    aes(x = pfas, y = outcome, fill = beta)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sig), size = 3.5, vjust = 0.75, color = "black") +
  scale_fill_gradient2(
    low      = "#2166AC",   # blue = negative association
    mid      = "white",
    high     = "#B2182B",   # red = positive association
    midpoint = 0,
    limits   = c(-beta_limit, beta_limit),
    name     = "\u03b2 coefficient"
  ) +
  facet_wrap(~ visit, nrow = 1) +
  labs(
    title    = paste0("PFAS-Outcome Associations Across Visits (", model_to_plot, ")"),
    subtitle = "Baseline: main effect \u03b2  |  Year 1 & 2: \u03b2 main + \u03b2 interaction \u00d7 visit\n* p<0.05  ** p<0.01  *** p<0.001 (main effect at baseline; interaction term at Year 1/2)",
    x        = "PFAS",
    y        = NULL,
    caption  = "Blue = negative association, Red = positive association"
  ) +
  theme_bw(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 12),
    plot.subtitle    = element_text(size = 8.5, color = "grey30"),
    axis.text.x      = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y      = element_text(size = 9),
    strip.text       = element_text(face = "bold"),
    strip.background = element_rect(fill = "grey92"),
    legend.position  = "right",
    panel.grid       = element_blank()
  )

# Save
ggsave(fs::path(lme_dir, paste0("heatmap_", model_to_plot, ".png")),
       plot   = p_heatmap,
       width  = 14,
       height = 6,
       dpi    = 300)

# ── Save one heatmap per model ────────────────────────────────────────────────
for (mod in unique(lme_results_df$model)) {
  
  heatmap_df_mod <- lme_results_df %>%
    filter(model == mod) %>%
    mutate(
      beta_visit0 = beta_main,
      beta_visit1 = beta_main + beta_int * 1,
      beta_visit2 = beta_main + beta_int * 2
    ) %>%
    select(pfas, outcome, beta_visit0, beta_visit1, beta_visit2,
           pval_main, pval_int) %>%
    pivot_longer(
      cols      = c(beta_visit0, beta_visit1, beta_visit2),
      names_to  = "visit",
      values_to = "beta"
    ) %>%
    mutate(
      visit = factor(visit,
                     levels = c("beta_visit0", "beta_visit1", "beta_visit2"),
                     labels = c("Baseline", "Year 1", "Year 2")),
      sig = case_when(
        visit == "Baseline" & pval_main < 0.001 ~ "***",
        visit == "Baseline" & pval_main < 0.01  ~ "**",
        visit == "Baseline" & pval_main < 0.05  ~ "*",
        visit != "Baseline" & pval_int  < 0.001 ~ "***",
        visit != "Baseline" & pval_int  < 0.01  ~ "**",
        visit != "Baseline" & pval_int  < 0.05  ~ "*",
        TRUE ~ ""
      ),
      pfas    = gsub("_ln_scaled", "", pfas),
      outcome = factor(outcome,
                       levels = outcome_vars_long,
                       labels = sapply(outcome_vars_long, get_label, df = dat))
    )
  
  beta_limit_mod <- max(abs(heatmap_df_mod$beta), na.rm = TRUE)
  
  p_mod <- ggplot(heatmap_df_mod,
                  aes(x = pfas, y = outcome, fill = beta)) +
    geom_tile(color = "white", linewidth = 0.4) +
    geom_text(aes(label = sig), size = 3.5, vjust = 0.75, color = "black") +
    scale_fill_gradient2(
      low      = "#2166AC",
      mid      = "white",
      high     = "#B2182B",
      midpoint = 0,
      limits   = c(-beta_limit_mod, beta_limit_mod),
      name     = "\u03b2 coefficient"
    ) +
    facet_wrap(~ visit, nrow = 1) +
    labs(
      title    = paste0("PFAS-Outcome Associations Across Visits (", mod, ")"),
      subtitle = "Baseline: main effect \u03b2  |  Year 1 & 2: \u03b2 main + \u03b2 interaction \u00d7 visit\n* p<0.05  ** p<0.01  *** p<0.001",
      x        = "PFAS",
      y        = NULL,
      caption  = "Blue = negative association, Red = positive association"
    ) +
    theme_bw(base_size = 11) +
    theme(
      plot.title       = element_text(face = "bold", size = 12),
      plot.subtitle    = element_text(size = 8.5, color = "grey30"),
      axis.text.x      = element_text(angle = 45, hjust = 1, size = 9),
      axis.text.y      = element_text(size = 9),
      strip.text       = element_text(face = "bold"),
      strip.background = element_rect(fill = "grey92"),
      legend.position  = "right",
      panel.grid       = element_blank()
    )
  
  ggsave(fs::path(lme_dir, paste0("heatmap_", mod, ".png")),
         plot   = p_mod,
         width  = 14,
         height = 6,
         dpi    = 300)
  
  message("Heatmap saved: ", mod)
}

```

