---
title: "PFAS_data cleaning"
author: "Savanah leidholt & Hailey Hampson"
date: "2026-03-12"
output: html_document
  theme: cerulean
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)
```

```{r}
#load libraries

library(ggplot2)
library(tidyverse)
library(janitor)
library(lme4)
library(lmerTest)
library(broom)
library(broom.mixed)
library(emmeans)
library(ggeffects)
library(patchwork)
library(gWQS)
library(corrplot)
library(gt)
library(table1)
#install.packages("lme4")
library(lme4)
#install.packages("ggpubr")
library(ggpubr)
```
#---------------------------------------------------
# 1. inputting all directories and storage locations
#---------------------------------------------------
```{r}
#Directories
computer <- "mac studio"
#computer <- "mac laptop"
if (computer == "mac studio") {
  user <- Sys.info()[["user"]]
  if (user == "sleidholt") {
    dir.dat <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  } else if (user == "shivaniramesh") {
    dir.dat <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive")
    dir.results <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results")
  } else if (user == "choiyej") {
    dir.dat <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  } else if (user == "pylell") {
    dir.dat <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  }
} else {
  user <- Sys.info()[["user"]]
  if (user == "savanahleidholt") {
    dir.dat <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
  } else {
    dir.dat <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive")
    dir.results <- c("/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results")
  }
}
```

```{r}
source(file.path(git_path, "Panther PFAS/Functions.R"))

dat <- read.csv(fs::path(dir.dat,"Data Harmonization","Data Clean","harmonized_dataset.csv"),na.strings = "")
lod <- read.csv(fs::path(dir.dat,"/PANTHER/Data_Raw/PFAS Emory 2025/8- PFAS Quantification/PFAS_LODs.csv"))
```
#-------------------------------------
#formatting Data identical to Hailey's
#-------------------------------------
```{r}
pfas <- c("N.EtFOSAA","N.MeFOSAA","PFBA","PFDA","PFDoA","PFHpA","PFHxA","PFNA","PFOA","PFBS","PFPeA","PFTeDA","PFTrDA","PFUnA","PFBS","PFDoS","PFHps","PFHxS","PFNS","PFOS","PFOSA","PFPeAS")
dat_collapsed1 <- dat %>%
  filter(study=="PANTHER") %>% 
  filter(visit=="baseline" | visit=="screening") %>% 
  # dplyr::select(record_id, date, rh2_id, visit,mrn, group, age, sex, bmi, hba1c,albuminuria_cat,microalbumin_u,starts_with("egfr"),cystatin_c_s,starts_with("creat"),all_of(pfas)) %>% 
  # dplyr::select(record_id, dob, date, rh2_id, visit, mrn, group, age, sex, bmi,race,ethnicity,race_ethnicity, hba1c,albuminuria_cat,acr_u,microalbumin_u,eGFR_CKiD_U25_avg,cystatin_c_s,creatinine_s,creatinine_u) %>% 
  # group_by(record_id, visit) %>%
  # tidyr::fill(date, .direction = "updown") %>% ungroup() %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id)) %>% 
  mutate(visit="baseline_screening") %>% 
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                                     race == "Black or African American" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                                     ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                                     T ~ "Not Hispanic or Latino Other"))
dat_collapsed2 <- dat %>%
  filter(study=="PANTHER") %>% 
  filter(visit=="year_1" | visit=="year_2") %>% 
  # dplyr::select(record_id, date, rh2_id, visit,mrn, group, age, sex, bmi, hba1c,albuminuria_cat,microalbumin_u,starts_with("egfr"),cystatin_c_s,starts_with("creat"),all_of(pfas)) %>% 
  # dplyr::select(record_id, dob, date, rh2_id, visit,mrn, group, age, sex, bmi, race,ethnicity, race_ethnicity, hba1c,acr_u,albuminuria_cat,microalbumin_u,eGFR_CKiD_U25_avg,cystatin_c_s,creatinine_s,creatinine_u) %>% 
  # group_by(record_id, visit) %>%
  # tidyr::fill(date, .direction = "updown") %>% ungroup() %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id, visit)) %>% 
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                                     race == "Black or African American" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                                     ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                                     T ~ "Not Hispanic or Latino Other"))
```
#----------------------------------------
#combine metadata & extract baseline pfas
#----------------------------------------
```{r}
meta <- rbind(dat_collapsed1,dat_collapsed2)
meta <- meta %>% 
  dplyr::select(-all_of(pfas))

pfas_dat <- dat %>% 
  filter(study=="PANTHER") %>% 
  filter(visit=="baseline") %>% 
  dplyr::select("record_id",all_of(pfas)) %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id)) %>% 
  filter(!is.na(N.EtFOSAA)) #91 participants with baseline pfas data
```
#-------------------------------------
#Create limit of detection table (lod)
#-------------------------------------
```{r}
#Impute PFAS Data
t_lod <- t(lod)
colnames(t_lod) <- t_lod[1,]
t_lod <- as.data.frame(t_lod)
t_lod <- t_lod[-1,]
lod <- t_lod
rm(t_lod)
lod <- lod %>% 
  mutate(across(everything(), ~as.numeric(.)))
colnames(lod) <- str_replace_all(colnames(lod),"-",".")
```
#-------------------------------------
#imputation function below LOD
#-------------------------------------
```{r}

impute_pfas_lod <- function(pfas_dat, lod, pfas_vars, threshold = 2/3) {
  n_participants <- length(unique(pfas_dat$record_id))
  pfas_to_impute <- c()
  for (pf in pfas_vars) {
    if (!(pf %in% names(lod))) next
    prop_below <- sum(pfas_dat[[pf]] < lod[[pf]], na.rm = TRUE) / n_participants
    if (prop_below < threshold) {
      pfas_to_impute <- c(pfas_to_impute, pf)
    }
  }
  pfas_imputed <- pfas_dat %>%
    dplyr::select(record_id, all_of(pfas_to_impute))
  for (pf in pfas_to_impute) {
    pfas_imputed[[pf]] <- ifelse(
      pfas_imputed[[pf]] < lod[[pf]],
      lod[[pf]] / sqrt(2),
      pfas_imputed[[pf]]
    )
  }
  return(list(
    imputed_data = pfas_imputed,
    pfas_used = pfas_to_impute
  ))
}

pfas_results <- impute_pfas_lod(
  pfas_dat = pfas_dat,
  lod = lod,
  pfas_vars = pfas
)

pfas_imputed <- pfas_results$imputed_data
pfas_to_impute <- pfas_results$pfas_used
```
#-------------------------------
#merge PFAS & baseline metadata
#-------------------------------
```{r}
meta_baseline <- meta %>% 
  filter(visit=="baseline_screening")

dat <- tidylog::left_join(
  pfas_imputed,
  meta_baseline,
  by="record_id"
)

dat_long <- meta %>%
  left_join(pfas_imputed, by="record_id")

dat_long <- dat_long %>%
  mutate(
    time_numeric = case_when(
      visit=="baseline_screening" ~ 0,
      visit=="year_1" ~ 1,
      visit=="year_2" ~ 2
    )
  )
```
```{r}
dat <- dat %>%
  mutate(
    microalbumin_u=ifelse(microalbumin_u>3000,NA,microalbumin_u),
    acr_u=ifelse(acr_u>2367,NA,acr_u)
  )

dat_long <- dat_long %>%
  mutate(
    microalbumin_u=ifelse(microalbumin_u>3000,NA,microalbumin_u),
    acr_u=ifelse(acr_u>2367,NA,acr_u)
  )
```
#-------------------------------
#Baseline participante metadata
#-------------------------------
```{r}
#Particpants 
table1(~age+sex+bmi+hba1c+hct+race_ethnicity_condensed+hba1c+microalbumin_u+acr_u+eGFR_CKiD_U25_avg+dexa_lean_kg+dexa_fat_kg+igf_1+bl_dheas|group,data=dat)
```

```{r}
table1(~N.MeFOSAA+PFDA+PFHpA+PFNA+PFBS+PFHps+
       PFHxS+PFOA+PFOS+PFPeAS | group,
       data=dat,
       render.continuous = render.geometric)
```

```{r}
#setting pfas variable to call back to

pfas_vars <- c(
"N.MeFOSAA","PFDA","PFHpA","PFNA","PFBS",
"PFHps","PFHxS","PFOA","PFOS","PFPeAS"
)

```

#------------------------------------------------------
#Performing ANOVA of Groups/PFAS with Tukey Corrections
#------------------------------------------------------
```{r}
pfas_group_dat <- dat %>%
  dplyr::select(record_id, group, sex, all_of(pfas_vars))

table(pfas_group_dat$group)
table(pfas_group_dat$sex)

# log transform PFAS
pfas_group_dat <- pfas_group_dat %>%
  mutate(across(all_of(pfas_vars), log))
```

```{r}
#pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah #Leidholt/PFAS/Results/PFAS_group_concentrations.pdf",width = 8, height = 6)

for (pfas in pfas_vars) {
  # ANOVA model
  formula0 <- as.formula(paste0(pfas, " ~ group + sex + group:sex"))
  model <- aov(formula0, data = pfas_group_dat)
  cat("\n----------------------\n")
  cat("PFAS:", pfas, "\n")
  print(summary(model))
  # Tukey post-hoc test
  tk <- TukeyHSD(model, "group")
  tk_df <- as.data.frame(tk$group)
  tk_df$comparison <- rownames(tk_df)
  tk_df <- tk_df %>%
    mutate(
      group1 = sub("-.*", "", comparison),
      group2 = sub(".*-", "", comparison),
      p.adj = `p adj`
    ) %>%
    filter(p.adj < 0.05)
  # format p-values (scientific notation, ~3 sig figs)
  if (nrow(tk_df) > 0) {
    tk_df$p_label <- formatC(tk_df$p.adj, format = "e", digits = 2)
    # set annotation heights
    tk_df$y.position <- max(pfas_group_dat[[pfas]], na.rm = TRUE) *
      seq(1.05, 1.2, length.out = nrow(tk_df))
  }
  # create plot
  p <- ggplot(pfas_group_dat,
              aes(x = factor(group), y = .data[[pfas]], fill = sex)) +
    geom_boxplot() +
    theme_bw() +
    labs(
      title = paste("PFAS Distribution:", pfas),
      x = "Group",
      y = "Log PFAS concentration"
    )
  # add significant comparisons
  if (nrow(tk_df) > 0) {
    p <- p +
      stat_pvalue_manual(
        tk_df,
        label = "p_label",
        xmin = "group1",
        xmax = "group2",
        y.position = "y.position"
      )
  }
  print(p)
}
dev.off()
```

# =========================================================
# PFAS + kidney/metabolic/hormonal analysis workflow
# Using pre-cleaned longitudinal dataset: dat_lmm.rds
# =========================================================
```{r}
# -------------------------
# 2. Define variables
# -------------------------

# Baseline PFAS exposures
pfas_vars <- c("N.MeFOSAA","PFDA","PFHpA","PFNA","PFBS","PFHps","PFHxS","PFOA","PFOS","PFPeAS")


# Outcomes with longitudinal data
outcome_vars <- c(
  "acr_u",
  "microalbumin_u",
  "eGFR_CKiD_U25_avg",
  "dexa_lean_kg",
  "dexa_fat_kg",
  "igf_1",
  "bl_dheas"
)

# Baseline-only outcomes
baseline_only_outcomes <- c(
  "mm_si",
  "mm_ir",
  "mm_di"
)

# All baseline outcomes
baseline_outcomes <- c(outcome_vars, baseline_only_outcomes)

```



```{r}
model_pfas <- lmer(
  acr_u ~ time_numeric + PFOS + age + sex + group +
  (1|record_id),
  data = dat_long
)

summary(model_pfas)
```


##---------------------------------------
## PFAS and Baseline Kidney Associations
##---------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Kidney Associations/uACR_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("acr_u ~ ",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y=acr_u))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and ACRu",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="ACRu"
    )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Kidney Associations/microalbumin_urine_PFAS_baseline.pdf", width = 8, height = 6)
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("microalbumin_u~ ",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y=microalbumin_u))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Microalbuminuria adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="microalbumin_u"
    )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Kidney Associations/eGFR_CKiD_U25_avg_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("eGFR_CKiD_U25_avg~ ",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= eGFR_CKiD_U25_avg))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and eGFR_CKiD_U25_avg adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="eGFR_CKiD_U25_avg"
      )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Kidney Associations/eGFR_fas_cr_PFAS_baseline.pdf", width = 8, height = 6)
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("eGFR_fas_cr~ ",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= eGFR_fas_cr))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and eGFR_fas_cr adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="eGFR_fas_cr"
      )
  
  print(p)
}
dev.off()
```
##--------------------------------------------
## PFAS & Baseline Metabolic Associations
##--------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Metabolic Associations/Fat_mass_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("dexa_fat_kg~",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= dexa_fat_kg))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Fat Mass adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="dexa_fat_kg"
      )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Metabolic Associations/Lean_mass_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("dexa_lean_kg~",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= dexa_lean_kg))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Lean Mass adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="dexa_lean_kg"
      )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Metabolic Associations/Insulin_Sensitivity_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("mm_si~",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= mm_si))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Insulin Sensitivity adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="mm_si"
      )
  
  print(p)
}

dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Metabolic Associations/Insulin_Resistance_PFAS_baseline.pdf", width = 8, height = 6)
for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("mm_ir~",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= mm_ir))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Insulin Resistance adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="mm_ir"
      )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Metabolic Associations/Insulin_Secretion_Capacity_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("mm_di~",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= mm_di))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Insulin Secretion Capacity adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="mm_di"
      )
  
  print(p)
}
dev.off()
```
##--------------------------------------------
## PFAS & Baseline Hormonal Associations
##--------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Hormonal Associations/IGF1_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("igf_1~",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= igf_1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and igf_1 adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="igf_1"
      )
  
  print(p)
}
dev.off()
```

```{r}
##"bl_estrad","bl_tot_test" --not in dataset 

pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Baseline Associations/Hormonal Associations/DHEAS_PFAS_baseline.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("bl_dheas~",pfas," + age + sex"))
  m1 <- lm(m0,dat=dat)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat,aes(x=.data[[pfas]],y= bl_dheas))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and DHEA-S adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="bl_dheas"
      )
  
  print(p)
}
dev.off()
```

```{r}
ggplot(dat_long,aes(time_numeric,acr_u))+

  geom_point(aes(color=factor(visit)),alpha=0.6)+
  geom_path(aes(group=record_id),alpha=0.2)+
  geom_smooth(method="lm",color="red4",se=FALSE)+

  theme_minimal()+
  labs(
    title="ACRu Levels Over Time",
    x="Time",
    y="ACRu"
  )

ggsave("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/uACR_over_time.pdf", width = 8, height = 6)
dev.off()
```
#---------------------------------
#Creating Follow up Data-sets
#---------------------------------
```{r}
#Folowoup Year 1

dat_followup1 <- dat_long %>% 
  filter(visit=="year_1") %>% 
  dplyr::select(c("record_id","microalbumin_u","acr_u","eGFR_CKiD_U25_avg","eGFR_fas_cr", "mm_si","mm_ir","mm_di","dexa_lean_kg","dexa_fat_kg", "igf_1", "bl_dheas")) %>% 
  dplyr::rename(microalbumin_u_yr1=microalbumin_u,
                acr_u_yr1=acr_u,
                eGFR_CKiD_U25_avg_yr1=eGFR_CKiD_U25_avg,
                eGRF_fas_cr_y1=eGFR_fas_cr,
                mm_si_yr1 = mm_si,
                mm_ir_yr1 = mm_ir,
                mm_di_yr1 = mm_di,
                dexa_lean_kg_yr1 = dexa_lean_kg, 
                dexa_fat_kg_yr1 = dexa_fat_kg, 
                igf_1_yr1 = igf_1, 
                bl_dheas_yr1 = bl_dheas)
dat_all <- left_join(dat,dat_followup1,by="record_id")
```
##----------------------------------------------------
## PFAS & Year 1 Followup Renal Clearance Associations
##----------------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Kidney Associations/uACR_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("acr_u_yr1 ~ ",pfas," + age + sex + acr_u")
  )
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,
              aes(x=.data[[pfas]],y=acr_u_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Associations Between PFAS vs ACRu at 1 Year Adj. for Age and Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="ACRu Year 1"
    )

  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Kidney Associations/Microalbumin_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("microalbumin_u_yr1 ~ ",pfas," + age + sex + microalbumin_u")
  )
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,
              aes(x=.data[[pfas]],y=microalbumin_u_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Assocations Between PFAS vs Microalbumin at 1 Year Adj. for Age and Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="Microalbuminuria Year 1"
    )

  print(p)
}

dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Kidney Associations/eGFR_CKiD_U25_avg_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("eGFR_CKiD_U25_avg_yr1 ~ ",pfas," + age + sex + eGFR_CKiD_U25_avg")
  )
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,
              aes(x=.data[[pfas]],y=eGFR_CKiD_U25_avg_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Associations Between PFAS vs eGFR_CKiD_U25_avg at 1 Year",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="eGFR_CKiD_U25_avg Year 1"
    )

  print(p)
}

dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Kidney Associations/eGFR_fas_cr_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("eGRF_fas_cr_y1 ~ ",pfas," + age + sex + eGFR_fas_cr")
  )
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,
              aes(x=.data[[pfas]],y=eGRF_fas_cr_y1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Associations between PFAS vs eGFR_fas_cr at 1 Year adj. for Age and Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="eGFR_fas_cr_y1 Year 1"
    )

  print(p)
}
dev.off()
```

##--------------------------------------------------
## PFAS & Year 1 Metabolic Associations
## no data for mm_si, mm_di, or mm_ir at year 1 mark
##--------------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Metabolic Associations/Fat_Mass_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("dexa_fat_kg_yr1~",pfas," + age + sex + dexa_fat_kg"))
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,aes(x=.data[[pfas]],y= dexa_fat_kg_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Fat Mass at Year 1 Adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="dexa_fat_kg_yr1"
      )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Metabolic Associations/Lean_Mass_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("dexa_lean_kg_yr1~",pfas," + age + sex+dexa_lean_kg"))
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,aes(x=.data[[pfas]],y= dexa_lean_kg_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Lean Mass at Year 1 adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="dexa_lean_kg_yr1"
      )
  
  print(p)
}

dev.off()
```
##--------------------------------------------
## PFAS & Year 1 Hormonal Associations
##--------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Hormonal Associations/IGF1_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("igf_1_yr1~",pfas," + age + sex + igf_1"))
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,aes(x=.data[[pfas]],y= igf_1_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and igf_1_yr1 at Year 1 adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="igf_1_yr1"
      )
  
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 1 Associations/Hormonal Associations/DHEAS_PFAS_yr1.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("bl_dheas_yr1~",pfas," + age + sex + bl_dheas"))
  m1 <- lm(m0,dat=dat_all)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all,aes(x=.data[[pfas]],y= bl_dheas_yr1))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and DHEA-S adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="bl_dheas_yr1"
      )
  
  print(p)
}
dev.off()
```

```{r}
#--------------------------------------
#Create Year 2 follow-up dataset
#--------------------------------------
dat_followup2 <- dat_long %>% 
  filter(visit=="year_2") %>% 
  dplyr::select(c("record_id","microalbumin_u","acr_u","eGFR_CKiD_U25_avg","eGFR_fas_cr", "mm_si","mm_ir","mm_di","dexa_lean_kg","dexa_fat_kg", "igf_1", "bl_dheas")) %>% 
  dplyr::rename(microalbumin_u_yr2=microalbumin_u,
                acr_u_yr2 = acr_u,
                eGFR_CKiD_U25_avg_yr2=eGFR_CKiD_U25_avg,
                eGRF_fas_cr_yr2=eGFR_fas_cr,
                mm_si_yr2 = mm_si,
                mm_ir_yr2 = mm_ir,
                mm_di_yr2 = mm_di,
                dexa_lean_kg_yr2 = dexa_lean_kg, 
                dexa_fat_kg_yr2 = dexa_fat_kg, 
                igf_1_yr2 = igf_1, 
                bl_dheas_yr2 = bl_dheas)

# join to baseline dataset
dat_all2 <- left_join(dat_all, dat_followup2, by="record_id")

dat_all2 <- dat_all2 %>%
  mutate(
    acr_u_yr2 = ifelse(acr_u_yr2 > 2367, NA, acr_u_yr2)
  )
```
##----------------------------------------------------
## PFAS & Year 1 Followup Renal Clearance Associations
##----------------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Kidney Associations/uACR_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("acr_u_yr2 ~ ",pfas," + age + sex + acr_u")
  )
  m1 <- lm(m0, dat = dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,
              aes(x=.data[[pfas]], y=acr_u_yr2)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    theme_bw() +
    labs(
      title="Baseline PFAS vs ACRu at 2 Years",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="ACRu Year 2"
    )

  print(p)
}

dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Kidney Associations/Microalbumin_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("microalbumin_u_yr2 ~ ",pfas," + age + sex + microalbumin_u")
  )
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,
              aes(x=.data[[pfas]],y=microalbumin_u_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Baseline PFAS vs Microalbumin at Year 2",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="Microalbuminuria Year 2"
    )

  print(p)
}

dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Kidney Associations/eGFR_CKiD_U25_avg_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("eGFR_CKiD_U25_avg_yr2 ~ ",pfas," + age + sex + eGFR_CKiD_U25_avg")
  )
  m1 <- lm(m0, dat = dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,
              aes(x=.data[[pfas]], y=eGFR_CKiD_U25_avg_yr2)) +
    geom_point() +
    geom_smooth(method="lm", se=FALSE) +
    theme_bw() +
    labs(
      title="Baseline PFAS vs eGFR_CKiD_U25_avg at 2 Years",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="eGFR_CKiD_U25_avg_yr2"
    )

  print(p)
}

dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Kidney Associations/eGFR_fas_cr_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(
    paste0("eGRF_fas_cr_yr2 ~ ",pfas," + age + sex + eGFR_fas_cr")
  )
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,
              aes(x=.data[[pfas]],y=eGRF_fas_cr_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Baseline PFAS vs eGFR_fas_cr at Year 2",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="eGFR_fas_cr Year 2"
    )

  print(p)
}

dev.off()
```

##--------------------------------------------------
## PFAS & Year 2 Metabolic Associations
## no data for mm_si, mm_di, or mm_ir at year 1 mark
##--------------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Metabolic Associations/Fat_Mass_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("dexa_fat_kg_yr2~",pfas," + age + sex + dexa_fat_kg"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,aes(x=.data[[pfas]],y= dexa_fat_kg_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Fat Mass Adj. for Age & Sex & Baseline DXA",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="dexa_fat_kg_yr2"
      )
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Metabolic Associations/Lean_Mass_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("dexa_lean_kg_yr2~",pfas," + age + sex+dexa_lean_kg"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,aes(x=.data[[pfas]],y= dexa_lean_kg_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and Lean Mass adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="dexa_lean_kg_yr2"
      )
  
  print(p)
}
dev.off()
```
##--------------------------------------------
## PFAS & Year 2 Hormonal Associations
##--------------------------------------------
```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Hormonal Associations/IGF1_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("igf_1_yr2~",pfas," + age + sex + igf_1"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,aes(x=.data[[pfas]],y= igf_1_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and igf_1 at Year 2 adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="igf_1_yr2"
      )
  print(p)
}
dev.off()
```

```{r}
pdf("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/Year 2 Associations/Hormonal Associations/DHEAS_PFAS_yr2.pdf", width = 8, height = 6)

for (pfas in pfas_vars) {
  m0 <- as.formula(paste0("bl_dheas_yr2~",pfas," + age + sex + bl_dheas"))
  m1 <- lm(m0,dat=dat_all2)
  beta <- summary(m1)$coef[2,1]
  pval <- summary(m1)$coef[2,4]
  p <- ggplot(dat_all2,aes(x=.data[[pfas]],y= bl_dheas_yr2))+
    geom_point()+
    geom_smooth(method="lm",se=FALSE)+
    theme_bw()+
    labs(
      title="Association between PFAS and DHEA-S at Year 2 adj. for Age & Sex",
      subtitle=paste0("Beta=",round(beta,3),
                      " p=",round(pval,3)),
      x=pfas,
      y="bl_dheas_yr2"
      )
  
  print(p)
    
}

dev.off()
```




```{r}
# =========================================================
# PFAS + kidney/metabolic/hormonal analysis workflow
# Using pre-cleaned longitudinal dataset: dat_lmm.rds
# =========================================================

# -------------------------
# 1. Read cleaned data
# -------------------------
dat_lmm <- readRDS(file.path(dir.dat2, "dat_lmm.rds"))

# -------------------------
# 2. Define variables
# -------------------------

# Baseline PFAS exposures
pfas_vars <- c(
  "N.MeFOSAA_total_ng", "PFDA_total_ng", "PFHpA_total_ng", "PFNA_total_ng", "PFBS_total_ng",
  "PFHps_total_ng", "PFHxS_total_ng", "PFOA_total_ng", "PFOS_total_ng", "PFPeAS_total_ng"
)

# Outcomes with longitudinal data
outcome_vars <- c(
  "log_acr_u",
  "log_microalbumin_u",
  "eGFR_CKiD_U25_avg",
  "dexa_lean_kg",
  "dexa_fat_kg",
  "igf_1",
  "bl_dheas"
)

# Baseline-only outcomes
baseline_only_outcomes <- c(
  "mm_si",
  "mm_ir",
  "mm_di"
)

# All baseline outcomes
baseline_outcomes <- c(outcome_vars, baseline_only_outcomes)

# -------------------------
# 3. Prepare longitudinal dataset
# -------------------------
dat_lmm2 <- dat_lmm %>%
  mutate(
    record_id = as.character(record_id),
    sex = factor(sex),
    group = factor(group),
    visit_f = factor(visit, levels = c("Baseline", "Year 1", "Year 2")),
    )

# -------------------------
# 4. Create baseline PFAS and baseline ACRu variables
# -------------------------
# Rename baseline PFAS so they are clearly "baseline exposure" variables
baseline_pfas <- dat_lmm2 %>%
  filter(visit == "Baseline") %>%
  select(record_id, all_of(pfas_vars)) %>%
  rename_with(~ paste0("bl_", .x), all_of(pfas_vars))

# Rename baseline log ACRu for use as an adjustment covariate
baseline_acru <- dat_lmm2 %>%
  filter(visit == "Baseline") %>%
  select(record_id, log_acr_u) %>%
  rename(bl_log_acr_u = log_acr_u)

# Join baseline exposure/covariate values onto all visits
dat_lmm2 <- dat_lmm2 %>%
  left_join(baseline_pfas, by = "record_id") %>%
  left_join(baseline_acru, by = "record_id")

# Define renamed baseline PFAS variables
bl_pfas_vars <- paste0("bl_", pfas_vars)

# Baseline-only dataset for cross-sectional analyses
dat_bl <- dat_lmm2 %>%
  filter(visit == "Baseline")

# Quick checks
table(dat_lmm2$visit_f, useNA = "ifany")
table(dat_lmm2$group, useNA = "ifany")
table(dat_lmm2$visit_f, dat_lmm2$group, useNA = "ifany")
```



```{r}
model_pfas <- lmer(
  acr_u ~ time_numeric + PFOS + age + sex + group +
  (1|record_id),
  data = dat_long
)

summary(model_pfas)
```

```{r}
wqs_model <- gwqs(
  acr_u ~ wqs + age + sex,
  mix_name = pfas_vars,
  data = dat,
  q = 4,
  validation = 0.6,
  b = 1000,
  family = "gaussian"
)

summary(wqs_model)
```