---
title: "PFAS Analysis — PANTHER Study"
author: "Savanah Leidholt & Hailey Hampson"
date: last-modified
format:
  html:
    theme: cerulean
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# libraries, directories, data loading, variable definitions,
# helper functions, and all data prep go here

#load libraries

library(ggplot2)
library(tidyverse)
library(janitor)
library(lme4)
library(lmerTest)
library(broom)
library(broom.mixed)
library(emmeans)
library(ggeffects)
library(patchwork)
library(corrplot)
library(gt)
library(table1)
library(stringr)
library(ggpubr)
library(patchwork)

#Directories
#computer <- "mac studio"
computer <- "mac laptop"
if (computer == "mac studio") {
  user <- Sys.info()[["user"]]
  if (user == "sleidholt") {
    dir.dat <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
  }
} else {
  user <- Sys.info()[["user"]]
  if (user == "savanahleidholt") {
    dir.dat <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/unadjusted_pfas_results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad/Panther PFAS"
  } 
}


source(file.path(git_path, "PFAS_Analysis_SLL.R"))


```

------------------------------------------------------------------------

# Participant Characteristics

## Table 1: Demographics & Clinical Outcomes by Group

```{r}
#| label: table1-demographics
table1(
  ~ age + sex + bmi + race_ethnicity_condensed + hba1c + acr_u + 
    microalbumin_u + eGFR_CKiD_U25_avg + avg_c_r2| group,
  data = dat_baseline
)
```

## Table 2: Baseline PFAS Levels by Group

```{r}
#| label: table2-pfas

```

------------------------------------------------------------------------

# Baseline Group Comparisons

## Clinical Outcomes by Group

```{r}
#| label: baseline-outcomes-plots
#| fig-width: 20
#| fig-height: 18

plot_list <- lapply(baseline_model_outcomes, function(outcome) {
  plot_baseline_group_boxplot(
    data = dat_baseline,
    outcome = outcome,
    results_list = baseline_group_outcome_results
  )
})

wrap_plots(plot_list, ncol = 3)

```

## PFAS Levels by Group

```{r}
#| label: baseline-pfas-plots
#| fig-width: 20
#| fig-height: 18

plots_by_group <- lapply(
  pfas_group_vars,
  function(pfas) {
    plot_baseline_pfas_group_boxplot(
      data = dat_baseline,
      exposure = pfas,
      results_list = baseline_group_pfas_results
    )
  }
)

wrap_plots(plots_by_group, ncol = 3)
```

------------------------------------------------------------------------

# Model 1: Cross-Sectional PFAS–Outcome Associations

```{r}
#| label: Forest Plots
#| fig-width: 16
#| fig-height: 14

forest_df <- baseline_table %>%
  dplyr::mutate(
    conf.low = estimate - 1.96 * std.error,
    conf.high = estimate + 1.96 * std.error
  )

plot_list <- list()


for (outcome in unique(forest_df$outcome)) {
  df <- forest_df %>%
    dplyr::filter(outcome == !!outcome)
  p <- ggplot(df, aes(x = estimate, y = exposure)) +
  geom_point(aes(shape = sig_fdr)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    title = paste("PFAS associations with", get_outcome_label(outcome)),
    x = "Effect size (β)",
    y = "PFAS",
    shape = "FDR <0.05"
  )
plot_list[[outcome]] <- p
}


wrap_plots(plot_list, ncol = 3)
```


------------------------------------------------------------------------

# Model 2: Outcome Trajectories Over Time by Group
##Model 2A: Trajectory Plots Base Model
```{r}
#| label: model2a-trajectory-plots
#| fig-width: 20
#| fig-height: 18
plot_group_time_set_unadjusted <- function(data,
                                           outcomes = group_time_outcomes,
                                           save_pdf = FALSE,
                                           out_dir = dir.results) {
  
  plot_list <- vector("list", length(outcomes))
  names(plot_list) <- outcomes
  
  for (i in seq_along(outcomes)) {
    outcome <- outcomes[i]
    
    rhs <- c("visit_f", "group", "visit_f:group", "age", "sex")
    
    m0 <- as.formula(
      paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
    )
    
    plot_data <- data %>%
      dplyr::select(record_id, all.vars(m0)) %>%
      tidyr::drop_na()
    
    if (nrow(plot_data) == 0) {
      plot_list[[i]] <- NULL
      next
    }
    
    m1 <- lmerTest::lmer(m0, data = plot_data, REML = FALSE)
    
    pred_df <- as.data.frame(
      ggeffects::ggpredict(m1, terms = c("visit_f", "group"))
    )
    
    int_rows <- broom.mixed::tidy(m1, effects = "fixed") %>%
      dplyr::filter(stringr::str_detect(term, "visit_f.*:group|group.*:visit_f"))
    
    subtitle_text <- if (nrow(int_rows) > 0) {
      paste0(
        "Visit × Group: ",
        paste0(int_rows$term, " p = ", fmt_p(int_rows$p.value), collapse = "; ")
      )
    } else {
      NULL
    }
    
    p <- ggplot(pred_df, aes(x = x, y = predicted, color = group, group = group)) +
      geom_line(linewidth = 1.1) +
      geom_point(size = 2.5) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.08) +
      scale_color_viridis_d(option = "C") +
      theme_bw() +
      labs(
        title = paste("Unadjusted", get_outcome_label(outcome), "over time by group"),
        subtitle = subtitle_text,
        x = "Visit",
        y = get_outcome_label(outcome),
        color = "Group"
      )
    
    plot_list[[i]] <- p
  }
  
  plot_list <- plot_list[!vapply(plot_list, is.null, logical(1))]
  
  if (save_pdf && length(plot_list) > 0) {
    pdf(
      file.path(out_dir, "group_time_all_outcomes.pdf"),
      width = 8,
      height = 6
    )
    for (nm in names(plot_list)) {
      print(plot_list[[nm]])
    }
    dev.off()
  }
  
  return(plot_list)
}

wrap_plots(plot_list, ncol = 3)

```
##Model 2B: Trajectory Plots Base Model Adjusted for Baseline ACR
```{r}
#| label: model2b-trajectory-plots
#| fig-width: 20
#| fig-height: 18
plot_group_time_set <- function(data,
                                outcomes = group_time_outcomes,
                                adjust_acr = TRUE,
                                save_pdf = FALSE,
                                out_dir = dir.results) {
  
  plot_list <- vector("list", length(outcomes))
  names(plot_list) <- outcomes
  
  for (i in seq_along(outcomes)) {
    outcome <- outcomes[i]
    rhs <- c("visit_f", "group", "visit_f:group", "age", "sex") 
    
    if (adjust_acr && outcome != "acr_u" && outcome != "log_acr_u") {
      rhs <- c(rhs, "log_acr_u_bl")
    }
    m0 <- as.formula(
      paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
    )
    
    plot_data <- data %>%
      dplyr::select(record_id, all.vars(m0)) %>%
      drop_na()
    
    if (nrow(plot_data) == 0) {
      plot_list[[i]] <- NULL
      next
    }
    m1 <- lmerTest::lmer(m0, data = plot_data, REML = FALSE)
    
    pred_df <- as.data.frame(
      ggeffects::ggpredict(m1, terms = c("visit_f", "group"))
    )
    
    int_rows <- broom.mixed::tidy(m1, effects = "fixed") %>%
      filter(str_detect(term, "visit_f.*:group|group.*:visit_f"))
    
    subtitle_text <- if (nrow(int_rows) > 0) {
      paste0(
        "Visit × Group: ",
        paste0(int_rows$term, " p = ", fmt_p(int_rows$p.value), collapse = "; ")
      )
    } else {
      NULL
    }
    
    p <- ggplot(pred_df, aes(x = x, y = predicted, color = group, group = group)) +
      geom_line(linewidth = 1.1) +
      geom_point(size = 2.5) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.08) +
      scale_color_viridis_d(option = "C")+
      theme_bw() +
      labs(
        title = paste("PFAS predicted", get_outcome_label(outcome), "over time by group"),
        subtitle = subtitle_text,
        x = "Visit",
        y = get_outcome_label(outcome),
        color = "Group"
      )
    
    plot_list[[i]] <- p
  }
  
  plot_list <- plot_list[!vapply(plot_list, is.null, logical(1))]
  
  if (save_pdf && length(plot_list) > 0) {
    pdf(
      file.path(out_dir, "group_time_all_outcomes.pdf"),
      width = 8,
      height = 6
    )
    for (nm in names(plot_list)) {
      print(plot_list[[nm]])
    }
    dev.off()
  }
  
  return(plot_list)
}


wrap_plots(plot_list, ncol = 3)
```
##Model 2C: Trajectory Plots Base Model Adjusted for Tanner Stage
```{r}
#| label: model2c-trajectory-plots
#| fig-width: 20
#| fig-height: 18
plot_group_time_set_tanner <- function(data,
                                       outcomes = group_time_outcomes,
                                       save_pdf = FALSE,
                                       out_dir = dir.results) {
  
  plot_list <- vector("list", length(outcomes))
  names(plot_list) <- outcomes
  
  for (i in seq_along(outcomes)) {
    outcome <- outcomes[i]
    
    rhs <- c("visit_f", "group", "visit_f:group", "age", "sex", "tan_stage")
    
    m0 <- as.formula(
      paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
    )
    
    plot_data <- data %>%
      dplyr::select(record_id, all.vars(m0)) %>%
      tidyr::drop_na()
    
    if (nrow(plot_data) == 0) {
      plot_list[[i]] <- NULL
      next
    }
    
    m1 <- lmerTest::lmer(m0, data = plot_data, REML = FALSE)
    
    pred_df <- as.data.frame(
      ggeffects::ggpredict(m1, terms = c("visit_f", "group"))
    )
    
    int_rows <- broom.mixed::tidy(m1, effects = "fixed") %>%
      dplyr::filter(stringr::str_detect(term, "visit_f.*:group|group.*:visit_f"))
    
    subtitle_text <- if (nrow(int_rows) > 0) {
      paste0(
        "Visit × Group: ",
        paste0(int_rows$term, " p = ", fmt_p(int_rows$p.value), collapse = "; ")
      )
    } else {
      NULL
    }
    
    p <- ggplot(pred_df, aes(x = x, y = predicted, color = group, group = group)) +
      geom_line(linewidth = 1.1) +
      geom_point(size = 2.5) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.08) +
      scale_color_viridis_d(option = "C") +
      theme_bw() +
      labs(
        title = paste("Tanner-adjusted", get_outcome_label(outcome), "over time by group"),
        subtitle = subtitle_text,
        x = "Visit",
        y = get_outcome_label(outcome),
        color = "Group"
      )
    
    plot_list[[i]] <- p
  }
  
  plot_list <- plot_list[!vapply(plot_list, is.null, logical(1))]
  
  if (save_pdf && length(plot_list) > 0) {
    pdf(
      file.path(out_dir, "group_time_all_outcomes_tanner.pdf"),
      width = 8,
      height = 6
    )
    for (nm in names(plot_list)) {
      print(plot_list[[nm]])
    }
    dev.off()
  }
  
  return(plot_list)
}

wrap_plots(plot_list, ncol = 3)
```
##Model 2D: Trajectory Plots Base Model Adjusted for Baseline uACR & Tanner Stage
```{r}
#| label: model2d-trajectory-plots
#| fig-width: 20
#| fig-height: 18
plot_group_time_acr_tan_set <- function(data,
                                outcomes = group_time_outcomes,
                                adjust_acr = TRUE,
                                save_pdf = FALSE,
                                out_dir = dir.results) {
  
  plot_list <- vector("list", length(outcomes))
  names(plot_list) <- outcomes
  
  for (i in seq_along(outcomes)) {
    outcome <- outcomes[i]
    rhs <- c("visit_f", "group", "visit_f:group", "age", "sex", "tan_stage") 
    
    if (adjust_acr && outcome != "acr_u" && outcome != "log_acr_u") {
      rhs <- c(rhs, "log_acr_u_bl")
    }
    m0 <- as.formula(
      paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
    )
    
    plot_data <- data %>%
      dplyr::select(record_id, all.vars(m0)) %>%
      drop_na()
    
    if (nrow(plot_data) == 0) {
      plot_list[[i]] <- NULL
      next
    }
    m1 <- lmerTest::lmer(m0, data = plot_data, REML = FALSE)
    
    pred_df <- as.data.frame(
      ggeffects::ggpredict(m1, terms = c("visit_f", "group"))
    )
    
    int_rows <- broom.mixed::tidy(m1, effects = "fixed") %>%
      filter(str_detect(term, "visit_f.*:group|group.*:visit_f"))
    
    subtitle_text <- if (nrow(int_rows) > 0) {
      paste0(
        "Visit × Group: ",
        paste0(int_rows$term, " p = ", fmt_p(int_rows$p.value), collapse = "; ")
      )
    } else {
      NULL
    }
    
    p <- ggplot(pred_df, aes(x = x, y = predicted, color = group, group = group)) +
      geom_line(linewidth = 1.1) +
      geom_point(size = 2.5) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.08) +
      scale_color_viridis_d(option = "C")+
      theme_bw() +
      labs(
        title = paste("PFAS predicted", get_outcome_label(outcome), "over time by group"),
        subtitle = subtitle_text,
        x = "Visit",
        y = get_outcome_label(outcome),
        color = "Group"
      )
    
    plot_list[[i]] <- p
  }
  
  plot_list <- plot_list[!vapply(plot_list, is.null, logical(1))]
  
  if (save_pdf && length(plot_list) > 0) {
    pdf(
      file.path(out_dir, "group_time_all_outcomes.pdf"),
      width = 8,
      height = 6
    )
    for (nm in names(plot_list)) {
      print(plot_list[[nm]])
    }
    dev.off()
  }
  
  return(plot_list)
}

wrap_plots(plot_list, ncol = 3)
```

------------------------------------------------------------------------

# Model 3: Baseline PFAS Predicting Longitudinal Outcomes

## Model 3A: Longitudinal Heatmap of Base Model
```{r}
#| label: model3a-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_base
```
## Model 3B: Longitudinal Heatmap Adjusted for Baseline uACR
```{r}
#| label: model3b-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_acr
```
## Model 3C: Longitudinal Heatmap Adjusted for Tanner Stage
```{r}
#| label: model3c-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_tanner
```
## Model 3D: Longitudinal Heatmap Adjusted for Baseline uACR & Tanner Stage
```{r}
#| label: model3d-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_acr_tanner
```
------------------------------------------------------------------------

# Appendix

```{r}
#| label: session-info
sessionInfo()
```
