---
title: "PFAS_Analysis"
author: "Savanah leidholt & Hailey Hampson"
date: "2026-03-12"
output: html_document
  theme: cerulean
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)
```

```{r}
#load libraries

library(ggplot2)
library(tidyverse)
library(janitor)
library(lme4)
library(lmerTest)
library(broom)
library(broom.mixed)
library(emmeans)
library(ggeffects)
library(patchwork)
library(gWQS)
library(corrplot)
library(gt)
library(table1)
library(stringr)
library(ggpubr)
```
#---------------------------------------------------
# inputting all directories and storage locations
#---------------------------------------------------
```{r}
#Directories
#computer <- "mac studio"
computer <- "mac laptop"
if (computer == "mac studio") {
  user <- Sys.info()[["user"]]
  if (user == "sleidholt") {
    dir.dat <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  }
} else {
  user <- Sys.info()[["user"]]
  if (user == "savanahleidholt") {
    dir.dat <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
  } 
}
```

```{r}
source(file.path(git_path, "Panther PFAS/Functions.R"))

dat_raw <- read.csv(fs::path(dir.dat,"Data Harmonization","Data Clean","harmonized_dataset.csv"),na.strings = "")
lod_raw <- read.csv(fs::path(dir.dat,"/PANTHER/Data_Raw/PFAS Emory 2025/8- PFAS Quantification/PFAS_LODs.csv"))
```

```{r}

pfas_all <- c(
  "N.EtFOSAA","N.MeFOSAA","PFBA","PFDA","PFDoA","PFHpA","PFHxA","PFNA","PFOA",
  "PFBS","PFPeA","PFTeDA","PFTrDA","PFUnA","PFDoS","PFHps","PFHxS","PFNS",
  "PFOS","PFOSA","PFPeAS"
)

pfas_vars <- c(
  "N.MeFOSAA","PFDA","PFHpA","PFNA","PFBS",
  "PFHps","PFHxS","PFOA","PFOS","PFPeAS"
)

baseline_model_outcomes <- c(
  "log_acr_u",
  "log_microalbumin_u",
  "eGFR_CKiD_U25_avg",
  "erpf_bsa_plasma",
  "avg_c_r2",
  "avg_k_r2",
  "dexa_lean_kg",
  "dexa_fat_kg",
  "igf_1",
  "bl_dheas",
  "mm_si",
  "mm_ir",
  "mm_di"
)

group_time_outcomes <- c(
  "log_acr_u",
  "log_microalbumin_u",
  "eGFR_CKiD_U25_avg",
  "erpf_bsa_plasma",
  "avg_c_r2",
  "avg_k_r2",
  "dexa_lean_kg",
  "dexa_fat_kg",
  "igf_1",
  "bl_dheas"
)


outcome_label_map <- c(
  #kidney function
  acr_u = "uACR",
  log_acr_u = "log1p(uACR)",
  microalbumin_u = "Microalbuminuria",
  log_microalbumin_u = "log1p(Microalbuminuria)",
  eGFR_CKiD_U25_avg = "eGFR",
  erpf_bsa_plasma = "ERPF",
  pah_clear_bsa = "PAH clearance",
  #kidney oxygenation
  avg_c_r2 = "Cortical R2",
  avg_k_r2= "Kidney R2",
  #metabolism
  dexa_lean_kg = "DEXA Lean Mass (kg)",
  dexa_fat_kg = "DEXA Fat Mass (kg)",
  mm_si = "Insulin Sensitivity",
  mm_ir = "Insulin Resistance",
  mm_di = "Disposition Index",
  #hormones
  igf_1 = "IGF-1",
  bl_dheas = "DHEA-S"
)

pfas_label_map <- c(
  log2_N.MeFOSAA_bl = "log2(N-MeFOSAA)",
  log2_PFDA_bl = "log2(PFDA)",
  log2_PFHpA_bl = "log2(PFHpA)",
  log2_PFNA_bl = "log2(PFNA)",
  log2_PFBS_bl = "log2(PFBS)",
  log2_PFHps_bl = "log2(PFHpS)",
  log2_PFHxS_bl = "log2(PFHxS)",
  log2_PFOA_bl = "log2(PFOA)",
  log2_PFOS_bl = "log2(PFOS)",
  log2_PFPeAS_bl = "log2(PFPeAS)"
)

```

```{r}
collapse_visit <- function(data, by_vars) {
  data %>%
    summarise(
      across(
        where(is.character),
        ~ if (all(is.na(.x))) NA_character_ else dplyr::last(na.omit(.x))
      ),
      across(
        where(is.factor),
        ~ if (all(is.na(.x))) NA else dplyr::last(na.omit(.x))
      ),
      across(
        where(is.numeric),
        ~ if (all(is.na(.x))) NA_real_ else mean(.x, na.rm = TRUE)
      ),
      .by = all_of(by_vars)
    )
}

make_race_ethnicity <- function(data) {
  data %>%
    mutate(
      race_ethnicity_condensed = case_when(
        race == "White" & ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
        race == "Black or African American" & ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
        ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
        TRUE ~ "Not Hispanic or Latino Other"
      )
    )
}
```

#---------------------------------------
#Helper fucntions for pvalues and labels
#---------------------------------------
```{r}
fmt_p <- function(p) {
  case_when(
    is.na(p) ~ NA_character_,
    p < 0.001 ~ "<0.001",
    TRUE ~ formatC(p, format = "f", digits = 3)
  )
}

get_pfas_label <- function(x) {
  dplyr::recode(x, !!!pfas_label_map, .default = x)
}

get_outcome_label <- function(outcome) {
  dplyr::recode(outcome, !!!outcome_label_map, .default = outcome)
}
```
#--------------------------------
#building analysis dataset
#--------------------------------
```{r}
#baseline only
dat_base <- dat_raw %>%
  filter(study == "PANTHER", visit %in% c("baseline", "screening")) %>%
  collapse_visit(by_vars = "record_id") %>%
  mutate(visit = "baseline_screening") %>%
  make_race_ethnicity()
#all data
dat_full <- dat_raw %>%
  filter(study == "PANTHER", visit %in% c("year_1", "year_2")) %>%
  collapse_visit(by_vars = c("record_id", "visit")) %>%
  make_race_ethnicity()

meta <- bind_rows(dat_base, dat_full)
```
#---------------------------------
#Baseline PFAS VALUES & Dataframe
#---------------------------------
```{r}
pfas_raw <- dat_raw %>%
  filter(study == "PANTHER", visit == "baseline") %>%
  select(record_id, all_of(pfas_all)) %>%
  collapse_visit(by_vars = "record_id")

lod <- lod_raw %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column("row_id")

colnames(lod) <- lod[1, ]
lod <- lod[-1, , drop = FALSE] %>%
  mutate(across(everything(), as.numeric))

colnames(lod) <- str_replace_all(colnames(lod), "-", ".")
```
#-------------------------------
#PFAS level of Detection (LOD) 
#-------------------------------
```{r}
impute_pfas_lod <- function(pfas_dat, lod, pfas_vars, threshold = 2/3) {
  n_participants <- n_distinct(pfas_dat$record_id)

  pfas_to_keep <- purrr::keep(pfas_vars, function(pf) {
    pf %in% names(lod) &&
      (sum(pfas_dat[[pf]] < lod[[pf]], na.rm = TRUE) / n_participants) < threshold
  })

  pfas_imputed <- pfas_dat %>%
    select(record_id, all_of(pfas_to_keep))

  for (pf in pfas_to_keep) {
    pfas_imputed[[pf]] <- ifelse(
      pfas_imputed[[pf]] < lod[[pf]],
      lod[[pf]] / sqrt(2),
      pfas_imputed[[pf]]
    )
  }

  list(
    imputed_data = pfas_imputed,
    pfas_used = pfas_to_keep
  )
}

pfas_results <- impute_pfas_lod(
  pfas_dat = pfas_raw,
  lod = lod,
  pfas_vars = pfas_all
)

pfas_imputed <- pfas_results$imputed_data %>%
  select(record_id, all_of(pfas_vars)) %>%
  rename_with(~ paste0(.x, "_bl"), -record_id)
```
#----------------------------------------
#Joining PFAS data to metadata
#----------------------------------------
```{r}
baseline_acru <- meta %>%
  filter(visit == "baseline_screening") %>%
  select(record_id, acr_u) %>%
  rename(acr_u_bl = acr_u)

meta_nopfas <- meta %>%
  select(-any_of(pfas_all))

dat_baseline <- meta_nopfas %>%
  filter(visit == "baseline_screening") %>%
  left_join(pfas_imputed, by = "record_id") %>%
  left_join(baseline_acru, by = "record_id")

dat_long <- meta_nopfas %>%
  left_join(pfas_imputed, by = "record_id") %>%
  left_join(baseline_acru, by = "record_id") %>%
  mutate(
    visit_f = factor(
      visit,
      levels = c("baseline_screening", "year_1", "year_2"),
      labels = c("Baseline", "Year 1", "Year 2")
    ),
    visit_num = case_when(
      visit == "baseline_screening" ~ 0,
      visit == "year_1" ~ 1,
      visit == "year_2" ~ 2,
      TRUE ~ NA_real_
    )
  )


```

#--------------------------------------------------
# Removing outliers and log transforming covariates
#--------------------------------------------------
```{r}
dat_baseline <- dat_baseline %>%
  mutate(
    microalbumin_u = ifelse(microalbumin_u > 3000, NA, microalbumin_u),
    acr_u = ifelse(acr_u > 2367, NA, acr_u),
    log_acr_u = log1p(acr_u),
    log_microalbumin_u = log1p(microalbumin_u),
    log_acr_u_bl = log1p(acr_u_bl)
  )

dat_long <- dat_long %>%
  mutate(
    microalbumin_u = ifelse(microalbumin_u > 3000, NA, microalbumin_u),
    acr_u = ifelse(acr_u > 2367, NA, acr_u),
    log_acr_u = log1p(acr_u),
    log_microalbumin_u = log1p(microalbumin_u),
    log_acr_u_bl = log1p(acr_u_bl)
  )

#Creating column of total pfas burden and then log transforming all PFAS data
pfas_bl_vars <- paste0(pfas_vars, "_bl")

dat_baseline <- dat_baseline %>%
  mutate(
    n_pfas_nonmiss = rowSums(!is.na(across(all_of(pfas_bl_vars)))),
    total_pfas_bl = ifelse(
      n_pfas_nonmiss == 0,
      NA_real_,
      rowSums(across(all_of(pfas_bl_vars)), na.rm = TRUE)
    ),
    log2_total_pfas_bl = ifelse(total_pfas_bl > 0, log2(total_pfas_bl), NA_real_)
  ) %>%
  select(-n_pfas_nonmiss)

dat_long <- dat_long %>%
  mutate(
    n_pfas_nonmiss = rowSums(!is.na(across(all_of(pfas_bl_vars)))),
    total_pfas_bl = ifelse(
      n_pfas_nonmiss == 0,
      NA_real_,
      rowSums(across(all_of(pfas_bl_vars)), na.rm = TRUE)
    ),
    log2_total_pfas_bl = ifelse(total_pfas_bl > 0, log2(total_pfas_bl), NA_real_)
  ) %>%
  select(-n_pfas_nonmiss)

dat_baseline <- dat_baseline %>%
  mutate(across(all_of(pfas_bl_vars), log2, .names = "log2_{.col}"))

dat_long <- dat_long %>%
  mutate(across(all_of(pfas_bl_vars), log2, .names = "log2_{.col}"))

log2_pfas_bl_vars <- paste0("log2_", pfas_bl_vars)
```


#-------------------------------
#Baseline participante metadata
#-------------------------------
```{r}
dat_baseline$dexa_fat_kg
table1(
  ~ age + sex + bmi + race_ethnicity_condensed + hba1c + acr_u + 
    microalbumin_u + eGFR_CKiD_U25_avg + avg_c_r2| group,
  data = dat_baseline
)

table1(
  ~ N.MeFOSAA_bl + PFDA_bl + PFHpA_bl + PFNA_bl +
    PFBS_bl + PFHps_bl + PFHxS_bl + PFOA_bl +
    PFOS_bl + PFPeAS_bl | group,
  data = dat_baseline,
  render.continuous = render.geometric
)
```
#---------------------------------------------------
# Baseline group comparison function outside of PFAS
#---------------------------------------------------
```{r}
run_baseline_group_model <- function(data, outcome, covars = c("age", "sex")) {
  
  rhs <- c("group", covars)
  form <- as.formula(paste0(outcome, " ~ ", paste(rhs, collapse = " + ")))
  
  model_data <- data %>%
    dplyr::select(all.vars(form)) %>%
    tidyr::drop_na()
  
  fit <- lm(form, data = model_data)
  
  # omnibus test for group
  group_rows <- broom::tidy(stats::anova(fit)) %>%
    dplyr::filter(term == "group") %>%
    dplyr::mutate(
      outcome = outcome,
      n = nrow(model_data)
    )
  
  # pairwise Tukey-adjusted comparisons
  emm <- emmeans::emmeans(fit, ~ group)
  pairwise_tbl <- summary(
    pairs(emm, adjust = "tukey"),
    infer = TRUE
  ) %>%
    as.data.frame() %>%
    dplyr::mutate(
      outcome = outcome,
      n = nrow(model_data)
    )
  
  list(
    model = fit,
    omnibus = group_rows,
    emmeans = as.data.frame(summary(emm)),
    pairwise = pairwise_tbl
  )
}
```

#-----------------------------------
# baseline outcome differences by group
#-----------------------------------
```{r}
baseline_group_outcome_results <- lapply(
  baseline_model_outcomes,
  function(outcome) {
    run_baseline_group_model(
      data = dat_baseline,
      outcome = outcome,
      covars = c("age", "sex")
    )
  }
)
names(baseline_group_outcome_results) <- baseline_model_outcomes

baseline_group_outcome_omnibus <- bind_rows(
  lapply(baseline_group_outcome_results, `[[`, "omnibus")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_outcome_label),
    p_fdr = p.adjust(p.value, method = "fdr"),
    p_fmt = fmt_p(p.value),
    p_fdr_fmt = fmt_p(p_fdr)
  )

baseline_group_outcome_pairwise <- bind_rows(
  lapply(baseline_group_outcome_results, `[[`, "pairwise")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_outcome_label),
    p_fmt = fmt_p(p.value)
  )

baseline_group_outcome_omnibus
baseline_group_outcome_pairwise
```
#-----------------------------------
# Plot baseline outcome by group
#-----------------------------------
```{r}
plot_baseline_group_boxplot <- function(data,
                                        outcome,
                                        results_list,
                                        y_label = NULL) {
  
  plot_data <- data %>%
    dplyr::select(group, sex, age, all_of(outcome)) %>%
    tidyr::drop_na()
  
  pair_tbl <- results_list[[outcome]]$pairwise %>%
    dplyr::mutate(
      group1 = stringr::str_trim(stringr::word(contrast, 1, sep = " - ")),
      group2 = stringr::str_trim(stringr::word(contrast, 2, sep = " - "))
    )
  
  # y positions for significance bars
  y_max <- max(plot_data[[outcome]], na.rm = TRUE)
  y_min <- min(plot_data[[outcome]], na.rm = TRUE)
  y_step <- 0.08 * (y_max - y_min)
  
  pair_tbl <- pair_tbl %>%
    dplyr::mutate(
      y.position = y_max + seq_len(n()) * y_step,
      p_label = dplyr::case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE ~ "ns"
      )
    )
  
  omnibus_p <- results_list[[outcome]]$omnibus$p.value[1]
  
  ggplot(plot_data, aes(x = group, y = .data[[outcome]], fill = group)) + 
    scale_fill_viridis_d(option = "C") +
    geom_boxplot(outlier.shape = NA, alpha = 0.7) +
    geom_jitter(width = 0.12, alpha = 0.6, size = 2) +
    ggpubr::stat_pvalue_manual(
      pair_tbl,
      label = "p_label",
      xmin = "group1",
      xmax = "group2",
      y.position = "y.position",
      tip.length = 0.01,
      hide.ns = TRUE
    ) +
    theme_bw() +
    labs(
      title = paste("Baseline", get_outcome_label(outcome), "by group"),
      subtitle = paste0("Omnibus group p = ", fmt_p(omnibus_p), "; pairwise p adjusted by Tukey"),
      x = "Group",
      y = ifelse(is.null(y_label), get_outcome_label(outcome), y_label)
    ) +
    theme(
      legend.position = "none",
      plot.title = element_text(face = "bold")
    )
}
```

```{r}
for (outcome in baseline_model_outcomes) {
  print(
    plot_baseline_group_boxplot(
      data = dat_baseline,
      outcome = outcome,
      results_list = baseline_group_outcome_results
    )
  )
}
```
#-----------------------------------
# Question 5: PFAS differences by group
#-----------------------------------
```{r}
pfas_group_vars <- c(log2_pfas_bl_vars, "log2_total_pfas_bl")

baseline_group_pfas_results <- lapply(
  pfas_group_vars,
  function(outcome) {
    run_baseline_group_model(
      data = dat_baseline,
      outcome = outcome,
      covars = c("age", "sex")
    )
  }
)
names(baseline_group_pfas_results) <- pfas_group_vars

baseline_group_pfas_omnibus <- bind_rows(
  lapply(baseline_group_pfas_results, `[[`, "omnibus")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_pfas_label),
    p_fdr = p.adjust(p.value, method = "fdr"),
    p_fmt = fmt_p(p.value),
    p_fdr_fmt = fmt_p(p_fdr)
  )

baseline_group_pfas_pairwise <- bind_rows(
  lapply(baseline_group_pfas_results, `[[`, "pairwise")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_pfas_label),
    p_fmt = fmt_p(p.value)
  )

baseline_group_pfas_omnibus
baseline_group_pfas_pairwise
```

```{r}
plot_baseline_pfas_group_boxplot <- function(data,
                                             exposure,
                                             results_list) {
  
  plot_data <- data %>%
    dplyr::select(group, sex, age, all_of(exposure)) %>%
    tidyr::drop_na()
  
  pair_tbl <- results_list[[exposure]]$pairwise %>%
    dplyr::mutate(
      group1 = stringr::str_trim(stringr::word(contrast, 1, sep = " - ")),
      group2 = stringr::str_trim(stringr::word(contrast, 2, sep = " - "))
    )
  
  y_max <- max(plot_data[[exposure]], na.rm = TRUE)
  y_min <- min(plot_data[[exposure]], na.rm = TRUE)
  y_step <- 0.08 * (y_max - y_min)
  
  pair_tbl <- pair_tbl %>%
    dplyr::mutate(
      y.position = y_max + seq_len(n()) * y_step,
      p_label = dplyr::case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE ~ "ns"
      )
    )
  
  omnibus_p <- results_list[[exposure]]$omnibus$p.value[1]
  
  ggplot(plot_data, aes(x = group, y = .data[[exposure]], fill = group)) +
    scale_fill_viridis_d(option = "C") +
    geom_boxplot(outlier.shape = NA, alpha = 0.7) +
    geom_jitter(width = 0.12, alpha = 0.6, size = 2) +
    ggpubr::stat_pvalue_manual(
      pair_tbl,
      label = "p_label",
      xmin = "group1",
      xmax = "group2",
      y.position = "y.position",
      tip.length = 0.01,
      hide.ns = TRUE
    ) +
    theme_bw() +
    labs(
      title = paste("Baseline", get_pfas_label(exposure), "by group"),
      subtitle = paste0("Omnibus group p = ", fmt_p(omnibus_p), "; pairwise p adjusted by Tukey"),
      x = "Group",
      y = get_pfas_label(exposure)
    ) +
    theme(
      legend.position = "none",
      plot.title = element_text(face = "bold")
    )
}
```

```{r}
for (pfas in pfas_group_vars) {
  print(
    plot_baseline_pfas_group_boxplot(
      data = dat_baseline,
      exposure = pfas,
      results_list = baseline_group_pfas_results
    )
  )
}
```

#------------------------------------------------
# Model 1: Baseline PFAS Cross-Sectional Analysis
#------------------------------------------------
```{r}
run_baseline_pfas_lm <- function(data, outcome, exposure, adjust_acr = TRUE) {
  
  rhs <- c(exposure, "age", "sex", "group")
  if (adjust_acr && !outcome %in% c("acr_u", "log_acr_u")) {
   rhs <- c(rhs, "log_acr_u_bl")
  }
  
  m0 <- as.formula(paste0(outcome, " ~ ", paste(rhs, collapse = " + ")))
  
  model_data <- data %>%
    dplyr::select(all.vars(m0)) %>%
    tidyr::drop_na()
  
  m1 <- lm(m0, data = model_data)
  
  list(
    model = m1,
    tidy = broom::tidy(m1) %>%
      dplyr::mutate(
        analysis = "baseline_cross_sectional",
        outcome = outcome,
        exposure = exposure,
        model_formula = paste(deparse(m0), collapse = " ")
      )
  )
}
```

```{r}
baseline_results_list <- lapply(baseline_model_outcomes, function(outcome) {
  lapply(log2_pfas_bl_vars, function(exposure) {
    run_baseline_pfas_lm(dat_baseline, outcome, exposure)
  })
})

baseline_results <- dplyr::bind_rows(
  lapply(baseline_results_list, function(x) dplyr::bind_rows(lapply(x, `[[`, "tidy")))
)
```

##-----------------------------------
##Baseline Results Tables
##-----------------------------------
```{r}
baseline_pfas_only <- baseline_results %>%
  dplyr::mutate(
    exposure_clean = stringr::str_replace_all(exposure, "`", "")
  ) %>%
  dplyr::filter(term %in% c(exposure, exposure_clean)) %>%
  dplyr::group_by(analysis, outcome) %>%
  dplyr::mutate(
    p_fdr = p.adjust(p.value, method = "fdr"),
    sig_fdr = p_fdr < 0.05,
    p_fmt = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value)),
    p_fdr_fmt = ifelse(p_fdr < 0.001, "<0.001", sprintf("%.3f", p_fdr))
  ) %>%
  dplyr::ungroup()

baseline_results_final <- baseline_results %>%
  dplyr::left_join(
    baseline_pfas_only %>%
      dplyr::select(analysis, outcome, exposure, term, p_fdr, sig_fdr, p_fdr_fmt),
    by = c("analysis", "outcome", "exposure", "term")
  )

baseline_table <- baseline_results_final %>%
  dplyr::filter(term %in% c(exposure, stringr::str_replace_all(exposure, "`", ""))) %>%
  dplyr::mutate(
    estimate_ci = paste0(
      round(estimate, 3), " (",
      round(estimate - 1.96 * std.error, 3), ", ",
      round(estimate + 1.96 * std.error, 3), ")"
    ),
    p_fmt = fmt_p(p.value)
  ) %>%
  dplyr::select(
    analysis, outcome, exposure, estimate, std.error, statistic, p.value,
    p_fdr, sig_fdr, estimate_ci, p_fmt, p_fdr_fmt
  )

baseline_table
```

```{r}
plot_baseline_pfas_set <- function(data,
                                   outcome,
                                   exposures = log2_pfas_bl_vars,
                                   adjust_acr = TRUE,
                                   results_tbl = baseline_results_final,
                                   save_pdf = FALSE,
                                   out_dir = dir.results) {
  
  plot_list <- vector("list", length(exposures))
  names(plot_list) <- exposures
  
  for (i in seq_along(exposures)) {
    exposure <- exposures[i]
    
    rhs <- c(exposure, "age", "sex", "group")
    if (adjust_acr && outcome != "log_acr_u") {
      rhs <- c(rhs, "log_acr_u_bl")
    }
    
    m0 <- as.formula(
      paste0(outcome, " ~ ", paste(rhs, collapse = " + "))
    )
    
    plot_data <- data %>%
      dplyr::select(all.vars(m0), group) %>%
      tidyr::drop_na()
    
    if (nrow(plot_data) == 0) {
      plot_list[[i]] <- NULL
      next
    }
    
    m1 <- lm(m0, data = plot_data)
    
    est_row <- broom::tidy(m1) %>%
      dplyr::filter(term == exposure)
    
    fdr_row <- results_tbl %>%
      dplyr::mutate(exposure_clean = stringr::str_replace_all(exposure, "`", "")) %>%
      dplyr::filter(
        outcome == !!outcome,
        exposure %in% c(!!exposure, stringr::str_replace_all(!!exposure, "`", "")) |
          exposure_clean %in% c(!!exposure, stringr::str_replace_all(!!exposure, "`", "")),
        term %in% c(!!exposure, stringr::str_replace_all(!!exposure, "`", ""))
      ) %>%
      dplyr::slice(1)
    
    pred_df <- as.data.frame(ggeffects::ggpredict(m1, terms = exposure))
    
    subtitle_txt <- paste0(
      "Beta = ", round(est_row$estimate, 3),
      "; p = ", fmt_p(est_row$p.value),
      "; FDR p = ", fmt_p(fdr_row$p_fdr)
    )
    
    p <- ggplot(plot_data, aes(x = .data[[exposure]], y = .data[[outcome]], color = group)) +
      geom_point(alpha = 0.7, size = 2) +
      geom_line(
        data = pred_df,
        aes(x = x, y = predicted),
        inherit.aes = FALSE,
        linewidth = 1.1,
        color = "black"
      ) +
      geom_ribbon(
        data = pred_df,
        aes(x = x, ymin = conf.low, ymax = conf.high),
        inherit.aes = FALSE,
        alpha = 0.2
      ) +
      theme_bw() +
      labs(
        title = paste(get_outcome_label(outcome), "vs", get_pfas_label(exposure)),
        subtitle = subtitle_txt,
        x = get_pfas_label(exposure),
        y = get_outcome_label(outcome),
        color = "Group"
      )
    
    plot_list[[i]] <- p
  }
  
  plot_list <- plot_list[!vapply(plot_list, is.null, logical(1))]
  
  if (save_pdf && length(plot_list) > 0) {
    pdf(
      file.path(out_dir, paste0("baseline_", outcome, "_all_pfas.pdf")),
      width = 8,
      height = 6
    )
    for (nm in names(plot_list)) {
      print(plot_list[[nm]])
    }
    dev.off()
  }
  
  return(plot_list)
}
```

```{r}
plot_baseline_pfas_set(data = dat_baseline,outcome = "log_acr_u")
plot_baseline_pfas_set(data = dat_baseline,outcome = "log_microalbumin_u")
plot_baseline_pfas_set(data = dat_baseline,outcome = "erpf_bsa_plasma")
plot_baseline_pfas_set(data = dat_baseline,outcome = "eGFR_CKiD_U25_avg")
plot_baseline_pfas_set(data = dat_baseline,outcome = "avg_c_r2")
plot_baseline_pfas_set(data = dat_baseline,outcome = "avg_k_r2")
plot_baseline_pfas_set(data = dat_baseline,outcome = "dexa_fat_kg")
plot_baseline_pfas_set(data = dat_baseline,outcome = "dexa_lean_kg")
plot_baseline_pfas_set(data = dat_baseline,outcome = "igf_1")
```


#--------------------------------------------
# Model 2: changes over time by group
#--------------------------------------------
```{r}
run_group_time_lmm <- function(data, outcome, adjust_acr = TRUE) {
  rhs <- c("visit_f:group", "age", "sex")
  
  if (adjust_acr && outcome != "acr_u" && outcome != "log_acr_u") {
    rhs <- c(rhs, "log_acr_u_bl")
  }
  
  m0 <- as.formula(
    paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
  )
  
  model_data <- data %>%
    select(record_id, all.vars(m0)) %>%
    drop_na()
  
  m1 <- lmerTest::lmer(m0, data = model_data, REML = FALSE)
  
  list(
    model = m1,
    tidy = broom.mixed::tidy(m1, effects = "fixed") %>%
      mutate(
        analysis = "time_by_group",
        outcome = outcome,
        model_formula = paste(deparse(m0), collapse = " ")
      )
  )
}

```

```{r}
group_time_results_list <- lapply(group_time_outcomes, function(outcome) {
  run_group_time_lmm(dat_long, outcome)
})
names(group_time_results_list) <- group_time_outcomes

group_time_results <- bind_rows(
  lapply(group_time_results_list, `[[`, "tidy")
)

group_time_results
```

```{r}
plot_group_time_set <- function(data,
                                outcomes = group_time_outcomes,
                                adjust_acr = TRUE,
                                save_pdf = FALSE,
                                out_dir = dir.results) {
  
  plot_list <- vector("list", length(outcomes))
  names(plot_list) <- outcomes
  
  for (i in seq_along(outcomes)) {
    outcome <- outcomes[i]
    rhs <- c("visit_f", "group", "visit_f:group", "age", "sex")
    
    if (adjust_acr && outcome != "acr_u" && outcome != "log_acr_u") {
      rhs <- c(rhs, "log_acr_u_bl")
    }
    m0 <- as.formula(
      paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
    )
    
    plot_data <- data %>%
      select(record_id, all.vars(m0)) %>%
      drop_na()
    
    if (nrow(plot_data) == 0) {
      plot_list[[i]] <- NULL
      next
    }
    m1 <- lmerTest::lmer(m0, data = plot_data, REML = FALSE)
    
    pred_df <- as.data.frame(
      ggeffects::ggpredict(m1, terms = c("visit_f", "group"))
    )
    
    int_rows <- broom.mixed::tidy(m1, effects = "fixed") %>%
      filter(str_detect(term, "visit_f.*:group|group.*:visit_f"))
    
    subtitle_text <- if (nrow(int_rows) > 0) {
      paste0(
        "Visit × Group: ",
        paste0(int_rows$term, " p = ", fmt_p(int_rows$p.value), collapse = "; ")
      )
    } else {
      NULL
    }
    
    p <- ggplot(pred_df, aes(x = x, y = predicted, color = group, group = group)) +
      geom_line(linewidth = 1.1) +
      geom_point(size = 2.5) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.08) +
      theme_bw() +
      labs(
        title = paste("PFAS predicted", get_outcome_label(outcome), "over time by group"),
        subtitle = subtitle_text,
        x = "Visit",
        y = get_outcome_label(outcome),
        color = "Group"
      )
    
    plot_list[[i]] <- p
  }
  
  plot_list <- plot_list[!vapply(plot_list, is.null, logical(1))]
  
  if (save_pdf && length(plot_list) > 0) {
    pdf(
      file.path(out_dir, "group_time_all_outcomes.pdf"),
      width = 8,
      height = 6
    )
    for (nm in names(plot_list)) {
      print(plot_list[[nm]])
    }
    dev.off()
  }
  
  return(plot_list)
}
```

```{r}
plot_group_time_set(data = dat_long,outcome = group_time_outcomes)
```

#-------------------------------------------------------------------------------
# Model 3: baseline PFAS predicting longitudinal outcomes adjusted for base uACR
#-------------------------------------------------------------------------------
```{r}
run_baseline_pfas_lmm <- function(data, outcome, exposure, adjust_acr = TRUE) {
  
  rhs <- c(exposure, "visit_f", paste0(exposure, ":visit_f"), "age", "sex", "group")
  
  if (adjust_acr && outcome != "acr_u" && outcome != "log_acr_u") {
    rhs <- c(rhs, "log_acr_u_bl")
  }
  
  m0 <- as.formula(
    paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
  )
  
  model_data <- data %>%
    select(record_id, all.vars(m0)) %>%
    drop_na()
  
  m1 <- lmerTest::lmer(m0, data = model_data, REML = FALSE)
  
  list(
    model = m1,
    tidy = broom.mixed::tidy(m1, effects = "fixed") %>%
      mutate(
        analysis = "baseline_pfas_longitudinal",
        outcome = outcome,
        exposure = exposure,
        model_formula = paste(deparse(m0), collapse = " ")
      )
  )
}
```

```{r}
longitudinal_results_list <- lapply(group_time_outcomes, function(outcome) {
  lapply(log2_pfas_bl_vars, function(exposure) {
    run_baseline_pfas_lmm(dat_long, outcome, exposure)
  })
})
names(longitudinal_results_list) <- group_time_outcomes

longitudinal_results <- bind_rows(
  lapply(longitudinal_results_list, function(x) bind_rows(lapply(x, `[[`, "tidy")))
)
```

```{r}
plot_longitudinal_pfas_set <- function(data,
                                       outcome,
                                       exposures = log2_pfas_bl_vars,
                                       adjust_acr = TRUE,
                                       save_pdf = FALSE,
                                       out_dir = dir.results) {
  
  plot_list <- vector("list", length(exposures))
  names(plot_list) <- exposures
  
  for (i in seq_along(exposures)) {
    exposure <- exposures[i]
    
    rhs <- c(exposure, "visit_f", paste0(exposure, ":visit_f"), "age", "sex", "group")
    
    if (adjust_acr && outcome != "acr_u" && outcome != "log_acr_u") {
      rhs <- c(rhs, "log_acr_u_bl")
    }
    
    m0 <- as.formula(
      paste0(outcome, " ~ ", paste(rhs, collapse = " + "), " + (1 | record_id)")
    )
    
    plot_data <- data %>%
      select(record_id, all.vars(m0)) %>%
      drop_na()
    
    if (nrow(plot_data) == 0) {
      plot_list[[i]] <- NULL
      next
    }
    
    m1 <- lmerTest::lmer(m0, data = plot_data, REML = FALSE)
    
    est_rows <- broom.mixed::tidy(m1, effects = "fixed") %>%
      filter(
        str_detect(term, fixed(paste0(exposure, ":visit_f"))) |
          str_detect(term, fixed(paste0("visit_f:", exposure)))
      )
    
    qs <- quantile(plot_data[[exposure]], probs = c(0.25, 0.50, 0.75), na.rm = TRUE)
    qs <- unique(as.numeric(qs))
    qs_round <- signif(qs, 3)
    
    pred_df <- as.data.frame(
      ggeffects::ggpredict(
        m1,
        terms = c(
          "visit_f",
          paste0(exposure, " [", paste(qs, collapse = ","), "]")
        )
      )
    )
    
    group_levels <- unique(pred_df$group)
    
    label_pool <- c(
      paste0("Low (25th: ", qs_round[1], ")"),
      if (length(qs_round) >= 2) paste0("Mid (50th: ", qs_round[min(2, length(qs_round))], ")"),
      if (length(qs_round) >= 3) paste0("High (75th: ", qs_round[3], ")")
    )
    
    label_pool <- label_pool[!is.na(label_pool)]
    label_pool <- label_pool[seq_along(group_levels)]
    
    pred_df$exposure_level <- factor(
      pred_df$group,
      levels = group_levels,
      labels = label_pool
    )
    
    subtitle_text <- if (nrow(est_rows) > 0) {
      paste0(
        "PFAS × Visit: ",
        paste0(est_rows$term, " p = ", fmt_p(est_rows$p.value), collapse = "; ")
      )
    } else {
      NULL
    }
    
    p <- ggplot(pred_df, aes(x = x, y = predicted, color = exposure_level, group = exposure_level)) +
      geom_line(linewidth = 1.1) +
      geom_point(size = 2.5) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.08) +
      theme_bw() +
      labs(
        title = paste(get_outcome_label(outcome), "trajectory by", get_pfas_label(exposure)),
        subtitle = subtitle_text,
        x = "Visit",
        y = get_outcome_label(outcome),
        color = "Baseline PFAS"
      )
    
    plot_list[[i]] <- p
  }
  
  plot_list <- plot_list[!vapply(plot_list, is.null, logical(1))]
  
  if (save_pdf && length(plot_list) > 0) {
    pdf(
      file.path(out_dir, paste0("longitudinal_", outcome, "_all_pfas.pdf")),
      width = 9,
      height = 6
    )
    for (nm in names(plot_list)) {
      print(plot_list[[nm]])
    }
    dev.off()
  }
  
  return(plot_list)
}
```

```{r}
plot_longitudinal_pfas_set(data = dat_long, outcome = "log_acr_u")
plot_longitudinal_pfas_set(data = dat_long, outcome = "log_microalbumin_u")
plot_longitudinal_pfas_set(data = dat_long, outcome = "eGFR_CKiD_U25_avg")
plot_longitudinal_pfas_set(data = dat_long, outcome = "dexa_lean_kg")
plot_longitudinal_pfas_set(data = dat_long, outcome = "dexa_fat_kg")
plot_longitudinal_pfas_set(data = dat_long, outcome = "igf_1")
plot_longitudinal_pfas_set(data = dat_long, outcome = "bl_dheas")
```

#-----------------------------------
#Heatmap of longitudinal outcomes
#-----------------------------------
```{r}
heatmap_longitudinal_df <- longitudinal_results %>%
  group_by(outcome, exposure) %>%
  summarise(
    beta_main = estimate[term == exposure][1],
    pval_main = p.value[term == exposure][1],
    
    beta_int_y1 = estimate[
      str_detect(term, fixed(paste0(exposure, ":visit_fYear 1"))) |
        str_detect(term, fixed(paste0("visit_fYear 1:", exposure)))
    ][1],
    pval_int_y1 = p.value[
      str_detect(term, fixed(paste0(exposure, ":visit_fYear 1"))) |
        str_detect(term, fixed(paste0("visit_fYear 1:", exposure)))
    ][1],
    
    beta_int_y2 = estimate[
      str_detect(term, fixed(paste0(exposure, ":visit_fYear 2"))) |
        str_detect(term, fixed(paste0("visit_fYear 2:", exposure)))
    ][1],
    pval_int_y2 = p.value[
      str_detect(term, fixed(paste0(exposure, ":visit_fYear 2"))) |
        str_detect(term, fixed(paste0("visit_fYear 2:", exposure)))
    ][1],
    .groups = "drop"
  ) %>%
  mutate(
    beta_visit0 = beta_main,
    beta_visit1 = beta_main + beta_int_y1,
    beta_visit2 = beta_main + beta_int_y2
  ) %>%
  select(
    outcome, exposure,
    beta_visit0, beta_visit1, beta_visit2,
    pval_main, pval_int_y1, pval_int_y2
  ) %>%
  pivot_longer(
    cols = c(beta_visit0, beta_visit1, beta_visit2),
    names_to = "visit",
    values_to = "beta"
  ) %>%
  mutate(
    visit = factor(
      visit,
      levels = c("beta_visit0", "beta_visit1", "beta_visit2"),
      labels = c("Baseline", "Year 1", "Year 2")
    ),
    sig = case_when(
      visit == "Baseline" & !is.na(pval_main)   & pval_main   < 0.001 ~ "***",
      visit == "Baseline" & !is.na(pval_main)   & pval_main   < 0.01  ~ "**",
      visit == "Baseline" & !is.na(pval_main)   & pval_main   < 0.05  ~ "*",
      visit == "Year 1"   & !is.na(pval_int_y1) & pval_int_y1 < 0.001 ~ "***",
      visit == "Year 1"   & !is.na(pval_int_y1) & pval_int_y1 < 0.01  ~ "**",
      visit == "Year 1"   & !is.na(pval_int_y1) & pval_int_y1 < 0.05  ~ "*",
      visit == "Year 2"   & !is.na(pval_int_y2) & pval_int_y2 < 0.001 ~ "***",
      visit == "Year 2"   & !is.na(pval_int_y2) & pval_int_y2 < 0.01  ~ "**",
      visit == "Year 2"   & !is.na(pval_int_y2) & pval_int_y2 < 0.05  ~ "*",
      TRUE ~ ""
    ),
    exposure = factor(
      exposure,
      levels = log2_pfas_bl_vars,
      labels = sapply(log2_pfas_bl_vars, get_pfas_label)
    ),
    outcome = factor(
      outcome,
      levels = group_time_outcomes,
      labels = sapply(group_time_outcomes, get_outcome_label)
    )
  )
```

```{r}
heatmap_longitudinal_df <- heatmap_longitudinal_df %>%
  group_by(visit) %>%
  mutate(beta_z = as.numeric(scale(beta))) %>%
  ungroup()
```

```{r}
z_limit <- max(abs(heatmap_longitudinal_df$beta_z), na.rm = TRUE)

p_longitudinal_heatmap <- ggplot(
  heatmap_longitudinal_df,
  aes(x = exposure, y = outcome, fill = beta_z)
) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sig), size = 3.5, vjust = 0.75, color = "black") +
  scale_fill_gradient2(
    low = "#2166AC",
    mid = "white",
    high = "#B2182B",
    midpoint = 0,
    limits = c(-z_limit, z_limit),
    name = "Z-score"
  ) +
  facet_wrap(~ visit, nrow = 1) +
  labs(
    title = "Baseline PFAS associations with longitudinal outcomes",
    subtitle = paste0(
      "Tiles show standardized model effect sizes within visit panels.",
      "\nBaseline: PFAS main effect; Year 1 and Year 2: PFAS main effect + PFAS×visit interaction.",
      "\n* p<0.05  ** p<0.01  *** p<0.001"
    ),
    x = "Baseline PFAS",
    y = NULL,
    caption = "Blue = negative association; Red = positive association"
  ) +
  theme_bw(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 8.5, color = "grey30"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y = element_text(size = 9),
    strip.text = element_text(face = "bold"),
    strip.background = element_rect(fill = "grey92"),
    legend.position = "right",
    panel.grid = element_blank()
  )

p_longitudinal_heatmap
```
