---
title: "PFAS_Analysis"
author: "Savanah leidholt"
date: "2026-03-12"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load libraries
library(ggplot2)
library(tidyverse)
library(janitor)
library(lme4)
library(lmerTest)
library(broom)
library(broom.mixed)
library(emmeans)
library(ggeffects)
library(patchwork)
library(gWQS)
library(corrplot)
library(gt)
library(table1)
library(stringr)
library(ggpubr)
library(jsonlite)
library(influence.ME)
library(aws.s3)
library(readr)
```

```{r}
#directories
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
  
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/UW/bjornstad_pyle_tommerdahl_lab - Data Science/Projects/P015_PANTHER_PFAS_dw_sl/results/")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
    
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/UW/bjornstad_pyle_tommerdahl_lab - Documents/Data Science/Projects/P015_PANTHER_PFAS_dw_sl/results/")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
    
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r}
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```

```{r}
obj <- aws.s3::get_object(
  object = "harmonized dataset/harmonized_dataset.csv",
  bucket = "core.data",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE,
  url_style = "path"
)

dat_raw <- readr::read_csv(
  rawConnection(obj)
)

obj2 <- aws.s3::get_object(
  object = "P015_PANTHER_PFAS_dw_sl/clean_data/PFAS_LODs.csv",
  bucket = "togo.projects",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE,
  url_style = "path"
)

lod_raw <- readr::read_csv(
  rawConnection(obj2)
)
```

```{r}
#setting up PFAS that passed LOD, outcomes, labels, and covariates
pfas_all <- c(
  "N.EtFOSAA","N.MeFOSAA","PFBA","PFDA","PFDoA","PFHpA","PFHxA","PFNA","PFOA",
  "PFBS","PFPeA","PFTeDA","PFTrDA","PFUnA","PFDoS","PFHps","PFHxS","PFNS",
  "PFOS","PFOSA","PFPeAS"
)

pfas_vars <- c(
  "N.MeFOSAA","PFDA","PFHpA","PFNA","PFBS",
  "PFHps","PFHxS","PFOA","PFOS","PFPeAS"
)


baseline_model_outcomes <- c(
  "log_uACR",
  #"log_microalbumin_u",
  "eGFR_CKiD_U25_avg",
  "erpf_bsa_plasma",
  "avg_c_r2",
  "avg_k_r2",
  "bmi",
  #"pah_clear_bsa",
  "dexa_lean_kg",
  "dexa_fat_kg",
  "igf_1",
  "bl_dheas",
  "mm_si",
  "mm_ir",
  "mm_di"
)


group_time_outcomes <- c(
  "log_uACR",
 # "log_microalbumin_u",
  "eGFR_CKiD_U25_avg",
  "erpf_bsa_plasma",
  #"pah_clear_bsa",
  "avg_c_r2",
  "avg_k_r2",
  "bmi",
  "dexa_lean_kg",
  "dexa_fat_kg",
  "igf_1",
  "bl_dheas"
)



outcome_label_map <- c(
  #kidney function
  acr_u = "uACR",
  log_uACR = "log1p(uACR)",
 # microalbumin_u = "Microalbuminuria",
#  log_microalbumin_u = "log1p(Microalbuminuria)",
  eGFR_CKiD_U25_avg = "eGFR",
  erpf_bsa_plasma = "ERPF",
  #pah_clear_bsa = "PAH clearance", #functionally the same as erpf
  #kidney oxygenation
  avg_c_r2 = "Cortical Oxygenation",
  avg_k_r2= "Medullary Oxygenation",
  #metabolism
  bmi = "Body Mass Index",
  dexa_lean_kg = "DEXA Lean Mass (kg)",
  dexa_fat_kg = "DEXA Fat Mass (kg)",
  mm_si = "Insulin Sensitivity",
  mm_ir = "Insulin Resistance",
  mm_di = "Disposition Index",
  #hormones
  igf_1 = "IGF-1",
  bl_dheas = "DHEA-S",
  tan_mgd = "Tanner Stage",
  tan_fgd = "Tanner Stage"
)


pfas_label_map <- c(
  log2_N.MeFOSAA_bl = "log2(N-MeFOSAA)",
  log2_PFDA_bl = "log2(PFDA)",
  log2_PFHpA_bl = "log2(PFHpA)",
  log2_PFNA_bl = "log2(PFNA)",
  log2_PFBS_bl = "log2(PFBS)",
  log2_PFHps_bl = "log2(PFHpS)",
  log2_PFHxS_bl = "log2(PFHxS)",
  log2_PFOA_bl = "log2(PFOA)",
  log2_PFOS_bl = "log2(PFOS)",
  log2_PFPeAS_bl = "log2(PFPeAS)"
)


source(file.path(git_path, "Functions.R")) 
```

#------------------------------------------------------
#building analysis dataset based on how Hailey filtered
#built an R ojbect so I don't need to run all over again
#------------------------------------------------------
```{r}
#baseline only
dat_base <- dat_raw %>%
  filter(study == "PANTHER", visit %in% c("baseline", "screening")) %>%
  collapse_visit(by_vars = "record_id") %>%
  mutate(visit = "baseline_screening") %>%
  make_race_ethnicity()


#all data
dat_full <- dat_raw %>%
  filter(study == "PANTHER", visit %in% c("year_1", "year_2")) %>%
  collapse_visit(by_vars = c("record_id", "visit")) %>%
  make_race_ethnicity()

meta <- bind_rows(dat_base, dat_full) %>%
  mutate(
    sex_lower = stringr::str_to_lower(as.character(sex)),
    tan_stage = case_when(
      sex_lower == "male" ~ tan_mgd,
      sex_lower == "female" ~ tan_fgd,
      TRUE ~ dplyr::coalesce(tan_mgd, tan_fgd)
    ),
    tan_stage = as.numeric(as.character(tan_stage))
  )
```
#---------------------------------
#Baseline PFAS VALUES & Dataframe
#---------------------------------
```{r}
lod_raw <- lod_raw %>%
  mutate(
    Compound = dplyr::recode(
      Compound,
      "N.MeFOSAA" = "N-MeFOSAA",
      "N.EtFOSAA" = "N-EtFOSAA"
    )
  )
unique(lod_raw$Compound)

dat_raw <- dat_raw %>%
  dplyr::rename(
    N.EtFOSAA = `N-EtFOSAA`,
    N.MeFOSAA = `N-MeFOSAA`
  )

pfas_raw <- dat_raw %>%
  dplyr::filter(study == "PANTHER", visit == "baseline") %>%
  dplyr::select(record_id, all_of(pfas_all)) %>%
  collapse_visit(by_vars = "record_id") %>%
  dplyr::summarise(
    across(
      where(negate(is.numeric)),
      ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))
    ),
    across(
      where(is.numeric),
      ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))
    ),
    .by = record_id
  ) %>%
  dplyr::filter(
    rowSums(!is.na(dplyr::across(all_of(pfas_all)))) > 0
  )

pfas_raw

lod <- lod_raw %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column("row_id")

colnames(lod) <- lod[1, ]
lod <- lod[-1, , drop = FALSE] %>%
  dplyr::mutate(across(everything(), as.numeric))

colnames(lod) <- str_replace_all(colnames(lod), "-", ".")
lod
```
#---------------------------------------------------------
#PFAS level of Detection (LOD) based off Hailey's LOD Step
#---------------------------------------------------------
```{r}
#pfas_results <- impute_pfas_lod(
#  pfas_dat = pfas_raw,
#  lod = lod,
#  pfas_vars = pfas_all
#)

#pfas_imputed <- pfas_results$imputed_data %>%
#  dplyr::select(record_id, all_of(pfas_vars)) %>%
#  rename_with(~ paste0(.x, "_baseline"), -record_id)

#tmp <- tempfile(fileext = "csv")

#write.csv(
#  pfas_imputed,
#  tmp,
#  row.names = FALSE
#)

#httr::set_config(httr::config(ssl_verifypeer = FALSE))
#put_object(
#  file = tmp,
#  object = "P015_PANTHER_PFAS_dw_sl/data_freeze/Plasma PFAS/pfas_imputed.csv",
#  bucket = "togo.projects",
#  base_url = "s3.kopah.uw.edu",
#  region = "",
#  use_https = TRUE
#)

obj3 <- aws.s3::get_object(
  object = "P015_PANTHER_PFAS_dw_sl/data_freeze/Plasma PFAS/pfas_imputed_06242026.csv",
  bucket = "togo.projects",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE,
  url_style = "path"
)

pfas_imputed <- readr::read_csv(
  rawConnection(obj3)
)
```

#-------------------------------------------------------------
#Joining PFAS data to metadata--Should have someone check this
#-------------------------------------------------------------
```{r}
baseline_acru <- meta %>%
  dplyr::filter(visit == "baseline_screening") %>%
  dplyr::select(record_id, acr_u) %>%
  dplyr::rename(uACR_baseline = acr_u)

meta_nopfas <- meta %>%
  dplyr::select(-any_of(pfas_all))

dat_baseline <- meta_nopfas %>%
  filter(visit == "baseline_screening") %>%
  left_join(pfas_imputed, by = "record_id") %>%
  left_join(baseline_acru, by = "record_id") # %>% left_join(baseline_tanner, by = "record_id")

dat_long <- meta_nopfas %>%
  left_join(pfas_imputed, by = "record_id") %>%
  left_join(baseline_acru, by = "record_id") %>%
  mutate(
    visit_f = factor(
      visit,
      levels = c("baseline_screening", "year_1", "year_2"),
      labels = c("Baseline", "Year 1", "Year 2")
    ),
    visit_num = case_when(
      visit == "baseline_screening" ~ 0,
      visit == "year_1" ~ 1,
      visit == "year_2" ~ 2,
      TRUE ~ NA_real_
    )
  )

#write_rds(baseline_acru, file = file.path(dir.dat, "Savanah Leidholt", "PFAS", "Data", "baseline_acru.rds"))
#write_rds(meta_nopfas,   file = file.path(dir.dat, "Savanah Leidholt", "PFAS", "Data", "meta_nopfas.rds"))
#write_rds(dat_baseline,  file = file.path(dir.dat, "Savanah Leidholt", "PFAS", "Data", "dat_baseline.rds"))
#write_rds(dat_long,      file = file.path(dir.dat, "Savanah Leidholt", "PFAS", "Data", "dat_long.rds"))

#baseline_acru <- readRDS(file.path(root_path, "Savanah Leidholt", "PFAS", "Data", "baseline_acru.rds"))
#meta_nopfas <- readRDS(file.path(root_path, "Savanah Leidholt", "PFAS", "Data", "meta_nopfas.rds"))
#dat_baseline <- readRDS(file.path(root_path, "Savanah Leidholt", "PFAS", "Data", "dat_baseline.rds"))
#dat_long <- readRDS(file.path(root_path, "Savanah Leidholt", "PFAS", "Data", "dat_long.rds"))

```
#---------------------------------------------------------------------------------
# Removing outliers and log transforming covariates
# not doing microalbumin in this analysis anymore --redundant since uACR is better
# check for other outliers
#---------------------------------------------------------------------------------
```{r}
dat_baseline <- dat_baseline %>%
  mutate(
    #microalbumin_u = ifelse(microalbumin_u > 3000, NA, microalbumin_u),
 #   acr_u = ifelse(acr_u > 2367, NA, acr_u),
    log_uACR = log1p(acr_u),
    #log_microalbumin_u = log1p(microalbumin_u),
    log_uACR_base = log1p(log_uACR)
  )

dat_long <- dat_long %>%
  mutate(
    #microalbumin_u = ifelse(microalbumin_u > 3000, NA, microalbumin_u),
#    acr_u = ifelse(acr_u > 2367, NA, acr_u),
    log_uACR = log1p(acr_u),
    #log_microalbumin_u = log1p(microalbumin_u),
    log_uACR_base = log1p(log_uACR)
  )

# Creating column of total PFAS burden and ln transforming all PFAS data
# ln because that is what most epidiomology studies use (rather than log2)
pfas_base_vars <- paste0(pfas_vars, "_baseline")

dat_baseline <- dat_baseline %>%
  mutate(
    n_pfas_nonmiss = rowSums(!is.na(across(all_of(pfas_base_vars)))),
    total_pfas_baseline = ifelse(
      n_pfas_nonmiss == 0,
      NA_real_,
      rowSums(across(all_of(pfas_base_vars)), na.rm = TRUE)
    ),
    ln_total_pfas_baseline = ifelse(total_pfas_baseline > 0, log(total_pfas_baseline), NA_real_)
  ) %>%
  dplyr::select(-n_pfas_nonmiss)

dat_long <- dat_long %>%
  mutate(
    n_pfas_nonmiss = rowSums(!is.na(across(all_of(pfas_base_vars)))),
    total_pfas_baseline = ifelse(
      n_pfas_nonmiss == 0,
      NA_real_,
      rowSums(across(all_of(pfas_base_vars)), na.rm = TRUE)
    ),
    ln_total_pfas_baseline = ifelse(total_pfas_baseline > 0, log(total_pfas_baseline), NA_real_)
  ) %>%
  dplyr::select(-n_pfas_nonmiss)

dat_baseline <- dat_baseline %>%
  mutate(across(all_of(pfas_base_vars), ~ log(.x), .names = "ln_{.col}"))

dat_long <- dat_long %>%
  mutate(across(all_of(pfas_base_vars), ~ log(.x), .names = "ln_{.col}"))

ln_pfas_base_vars <- paste0("ln_", pfas_base_vars)
```
#---------------------------------------------------------------------
#Correlation matrix of PFAS
#PFAS tend to be correlated not sure if I should do anything with this
#---------------------------------------------------------------------
```{r}
# data for correlation matrix
pfas_corr_df <- dat_baseline %>%
  dplyr::select(all_of(ln_pfas_base_vars)) %>%
  dplyr::rename_with(~ sapply(.x, get_pfas_label))

# correlation matrix using pairwise complete observations
pfas_corr_mat <- cor(
  pfas_corr_df,
  use = "pairwise.complete.obs",
  method = "spearman"   
)

pfas_corr_mat
```

```{r}
pfas_cor_tests <- expand.grid(
  var1 = ln_pfas_base_vars,
  var2 = ln_pfas_base_vars,
  stringsAsFactors = FALSE
) %>%
  dplyr::filter(var1 < var2) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    test = list(
      cor.test(
        dat_baseline[[var1]],
        dat_baseline[[var2]],
        method = "spearman",
        use = "pairwise.complete.obs",
        exact = FALSE
      )
    ),
    rho = test$estimate,
    p_value = test$p.value
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    var1_label = sapply(var1, get_pfas_label),
    var2_label = sapply(var2, get_pfas_label),
    p_fdr = p.adjust(p_value, method = "fdr")
  ) %>%
  dplyr::select(var1_label, var2_label, rho, p_value, p_fdr)

pfas_cor_tests
```

```{r}
pfas_corr_df <- dat_baseline %>%
  dplyr::select(all_of(ln_pfas_base_vars)) %>%
  dplyr::rename_with(~ sapply(.x, get_pfas_label))

pfas_corr_mat <- cor(
  pfas_corr_df,
  use = "pairwise.complete.obs",
  method = "spearman"
)

pfas_p_mat <- cor_mtest(pfas_corr_df, method = "spearman")
```

```{r fig.width=16, fig.height=12}
#pdf(file = file.path(dir.results,"figures", "plasma_pfas_unadjusted", "pfas_correlation_plot.pdf"))

corrplot::corrplot(
  pfas_corr_mat,
  method = "color",
  type = "upper",
  order = "hclust",
  tl.col = "black",
  tl.srt = 45,
  col = colorRampPalette(c("#2166AC", "white", "#B2182B"))(200),
  p.mat = pfas_p_mat,
  sig.level = c(0.001, 0.01, 0.05),
  insig = "label_sig",
  pch.col = "black",
  pch.cex = 1.2,
  mar = c(0, 0, 1, 0),
  title = "Spearman correlation matrix of baseline ln-transformed PFAS"
)

```

#-------------------------------
#Baseline participante metadata
#Baseline PFAS data at baseline
#-------------------------------
```{r}
dat_baseline_pfas <- dat_baseline %>%
  filter(if_any(all_of(pfas_base_vars), ~ !is.na(.)))

participant_data <- table1(
  ~ age + sex + bmi + race_ethnicity_condensed + hba1c + acr_u + erpf_bsa_plasma +
    eGFR_CKiD_U25_avg + avg_c_r2 + avg_k_r2 + mm_ir + bl_dheas | group,
  data = dat_baseline_pfas
)
participant_data
#participant_data_save <- as.data.frame(participant_data)

pfas_data <- table1(
  ~ N.MeFOSAA_baseline + PFDA_baseline + PFHpA_baseline + PFNA_baseline +
    PFBS_baseline + PFHps_baseline + PFHxS_baseline + PFOA_baseline +
    PFOS_baseline + PFPeAS_baseline | group,
  data = dat_baseline_pfas,
  render.continuous = render.geometric
)
pfas_data
#pfas_data_save <- as.data.frame(pfas_data)


#write_csv(participant_data_save, 
  #        file = file.path(dir.results, "csv", "plasma_pfas_unadjusted","participant_clinical_data.csv"))

#write_csv(pfas_data_save, 
 #         file = file.path(dir.results, "csv", "plasma_pfas_unadjusted","pfas_data.csv"))

```
#---------------------------------------------------
# Baseline group comparison function outside of PFAS
#---------------------------------------------------
```{r}
baseline_group_outcome_results <- lapply(
  baseline_model_outcomes,
  function(outcome) {
    run_baseline_group_model(
      data = dat_baseline,
      outcome = outcome,
      covars = c("age", "sex")
    )
  }
)

names(baseline_group_outcome_results) <- baseline_model_outcomes

baseline_group_outcome_omnibus <- bind_rows(
  lapply(baseline_group_outcome_results, `[[`, "omnibus")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_outcome_label),
    p_fdr = p.adjust(p.value, method = "fdr"),
    p_fmt = fmt_p(p.value),
    p_fdr_fmt = fmt_p(p_fdr)
  )

baseline_group_outcome_pairwise <- bind_rows(
  lapply(baseline_group_outcome_results, `[[`, "pairwise")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_outcome_label),
    p_fmt = fmt_p(p.value)
  )


baseline_group_outcome_omnibus
#write_csv(baseline_group_outcome_omnibus, file = file.path(dir.results, "csv", "plasma_pfas_unadjusted", "baseline_group_outcomes.csv"))

baseline_group_outcome_pairwise
#write_csv(baseline_group_outcome_pairwise, file = file.path(dir.results, "csv", "plasma_pfas_unadjusted", "baseline_group_outcomes_pairwise.csv"))
```
##-----------------------------------
## Plot baseline outcome by group
##-----------------------------------

```{r fig.height=18}
#pdf(file = file.path(dir.results, "figures", "plasma_pfas_unadjusted", "group_outcomes.pdf"), width = 16, height = 12)

plot_list <- lapply(baseline_model_outcomes, function(outcome) {
  plot_baseline_group_boxplot(
    data = dat_baseline,
    outcome = outcome,
    results_list = baseline_group_outcome_results
  )
})

# Combine the plots
combined_plot <- wrap_plots(plot_list, ncol = 3)
combined_plot

#dev.off()

```
#---------------------------------------------------------
# Baseline PFAS differences by group
# Hypotheizing equal PFAS exposures so no group differences
#----------------------------------------------------------
```{r}
pfas_group_vars <- c(ln_pfas_base_vars, "ln_total_pfas_baseline")


baseline_group_pfas_results <- lapply(
  pfas_group_vars,
  function(outcome) {
    run_baseline_group_model(
      data = dat_baseline,
      outcome = outcome,
      covars = c("age", "sex")
    )
  }
)

names(baseline_group_pfas_results) <- pfas_group_vars

baseline_group_pfas_omnibus <- bind_rows(
  lapply(baseline_group_pfas_results, `[[`, "omnibus")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_pfas_label),
    p_fdr = p.adjust(p.value, method = "fdr"),
    p_fmt = fmt_p(p.value),
    p_fdr_fmt = fmt_p(p_fdr)
  )

baseline_group_pfas_pairwise <- bind_rows(
  lapply(baseline_group_pfas_results, `[[`, "pairwise")
) %>%
  dplyr::mutate(
    outcome_label = sapply(outcome, get_pfas_label),
    p_fmt = fmt_p(p.value)
  )

baseline_group_pfas_omnibus
#write_csv(baseline_group_pfas_omnibus, file = file.path(dir.results,"csv", "plasma_pfas_unadjusted", "baseline_pfas_omnibus.csv"))

baseline_group_pfas_pairwise
#write_csv(baseline_group_pfas_omnibus, file = file.path(dir.results,"csv", "plasma_pfas_unadjusted", "baseline_pfas_pairwise.csv"))
```

```{r fig.width=18}
plots_by_group <- lapply(
  pfas_group_vars,
  function(pfas) {
    plot_baseline_pfas_group_boxplot(
      data = dat_baseline,
      exposure = pfas,
      results_list = baseline_group_pfas_results
    )
  }
)

combined_plot2 <- wrap_plots(plots_by_group, ncol = 3)
combined_plot2

#pdf(file.path(dir.results, "figures", "pfas_by_group.pdf"), width = 18, height = 16)
#print(combined_plot2)
#dev.off()

```
#------------------------------------------------
# Model 1: Baseline PFAS Cross-Sectional Analysis
# no uACR adjustment at baseline
#------------------------------------------------

```{r}
baseline_results_list <- lapply(baseline_model_outcomes, function(outcome) {
  lapply(pfas_group_vars, function(exposure) {
    run_baseline_pfas_lm(dat_baseline, outcome, exposure)
  })
})

baseline_results <- dplyr::bind_rows(
  lapply(baseline_results_list, function(x) dplyr::bind_rows(lapply(x, `[[`, "tidy")))
)
```

#--------------------------
#Baseline Results Tables
#--------------------------
```{r}
baseline_pfas_only <- baseline_results %>%
  dplyr::mutate(
    exposure_clean = stringr::str_replace_all(exposure, "`", "")
  ) %>%
  dplyr::filter(term %in% c(exposure, exposure_clean)) %>%
  dplyr::group_by(analysis, outcome) %>%
  dplyr::mutate(
    p_fdr = p.adjust(p.value, method = "fdr"),
    sig_fdr = p_fdr < 0.05,
    p_fmt = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value)),
    p_fdr_fmt = ifelse(p_fdr < 0.001, "<0.001", sprintf("%.3f", p_fdr))
  ) %>%
  dplyr::ungroup()

baseline_results_final <- baseline_results %>%
  dplyr::left_join(
    baseline_pfas_only %>%
      dplyr::select(analysis, outcome, exposure, term, p_fdr, sig_fdr, p_fdr_fmt),
    by = c("analysis", "outcome", "exposure", "term")
  )

baseline_table <- baseline_results_final %>%
  dplyr::filter(term %in% c(exposure, stringr::str_replace_all(exposure, "`", ""))) %>%
  dplyr::mutate(
    estimate_ci = paste0(
      round(estimate, 3), " (",
      round(estimate - 1.96 * std.error, 3), ", ",
      round(estimate + 1.96 * std.error, 3), ")"
    ),
    p_fmt = fmt_p(p.value)
  ) %>%
  dplyr::select(
    analysis, outcome, exposure, estimate, std.error, statistic, p.value,
    p_fdr, sig_fdr, estimate_ci, p_fmt, p_fdr_fmt
  )

baseline_table
#write_csv(baseline_table, file=file.path(dir.results, "csv", "plasma_pfas_unadjusted", "baseline_cross_sectional_results.csv"))
```
#-------------------------
#individual baseline plots
#-------------------------
```{r}
#plot_baseline_pfas_set(data = dat_baseline,outcome = "log_acr_u", save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "log_microalbumin_u",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "erpf_bsa_plasma",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "eGFR_CKiD_U25_avg",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "pah_clear_bsa",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "avg_c_r2",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "avg_k_r2",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "dexa_fat_kg",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "dexa_lean_kg",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "igf_1",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "bl_dheas",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "mm_si",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "mm_ir",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
#plot_baseline_pfas_set(data = dat_baseline,outcome = "mm_di",save_pdf = T, out_dir = file.path(dir.results, "figures", "plasma_pfas_unadjusted"))
```

```{r}
forest_df <- baseline_table %>%
  dplyr::mutate(
    conf.low = estimate - 1.96 * std.error,
    conf.high = estimate + 1.96 * std.error
  )

forest_df
```

```{r}
plot_list <- list()


for (outcome in unique(forest_df$outcome)) {
  df <- forest_df %>%
    dplyr::filter(outcome == !!outcome)
  p <- ggplot(df, aes(x = estimate, y = exposure)) +
  geom_point(aes(colour = sig_fdr), size = 2.5) +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "darkred")
) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    title = paste("PFAS associations with", get_outcome_label(outcome)),
    x = "Effect size (β)",
    y = "PFAS",
    shape = "FDR <0.05"
  )
plot_list[[outcome]] <- p
}
```

```{r fig.width=14, fig.height=18}
#pdf(file.path(dir.results, "figures", "plasma_pfas_unadjusted", "cross_sectional_analysis", "cross-sectional_forest_plots.pdf"), width = 22, height = 20)
wrap_plots(plot_list, ncol=3)
```


#--------------------------------------------------------------------
# Model 2: changes Over time by group
# a) base model of age + sex + group
# b) baseline uACR adjustment 
# c) tanner stage adjustment (laura suggested)
# d) baseline uACR + tanner stage adjustment (might be too strict)
#--------------------------------------------------------------------

##model 2a: results list + plots
```{r}
library(WriteXLS)
mod2a_results_list <- lapply(group_time_outcomes, function(outcome) {
  model_2a(dat_long, outcome)
})

names(mod2a_results_list) <- group_time_outcomes

mod2a_results <- dplyr::bind_rows(
  lapply(mod2a_results_list, `[[`, "tidy"))

#mod2a_results 
WriteXLS(mod2a_results,file.path(dir.results, "csv", "plasma_pfas_unadjusted","longitudinal_group_outcomes", "model_2a_results.2.xlsx"))



mod2a_plot <- plot_model_2a(
  data = dat_long,
  outcomes = group_time_outcomes ,
  save_pdf = TRUE,
  out_dir = dir.results
)

#mod2a_plot
pdf(file.path(dir.results, "figures", "plasma_pfas_unadjusted", "longitudinal_group_outcomes", "model_2a_plots.2.pdf"), width = 22, height = 20)
wrap_plots(mod2a_plot, ncol=3)

```

#model 2b results list + plots
```{r}
mod2b_list <- lapply(group_time_outcomes, function(outcome) {
  model_2b(dat_long, outcome)
})
names(mod2b_list) <- group_time_outcomes

mod2b_results <- bind_rows(
  lapply(mod2b_list, `[[`, "tidy")
)

#mod2b_results 
WriteXLS(mod2b_results,file.path(dir.results, "csv", "plasma_pfas_unadjusted","longitudinal_group_outcomes", "model_2b_results.2.xlsx"))

mod2b_plot <- plot_model_2b(data = dat_long)
#mod2b_plot

pdf(file.path(dir.results, "figures", "plasma_pfas_unadjusted", "longitudinal_group_outcomes", "model_2b_plots.2.pdf"), width = 22, height = 20)
wrap_plots(mod2b_plot, ncol=3)
```

#model 2c results + plots
```{r fig.height=16}
mod2c_list <- lapply(group_time_outcomes, function(outcome) {
  model_2c(dat_long, outcome)
})
names(mod2c_list) <- group_time_outcomes

mod2c_results <- dplyr::bind_rows(
  lapply(mod2c_list, `[[`, "tidy")
)
#mod2c_results
WriteXLS(mod2c_results,file.path(dir.results, "csv", "plasma_pfas_unadjusted","longitudinal_group_outcomes", "model_2c_results.2.xlsx"))


mod2c_plots <- plot_model_2c(data = dat_long)

pdf(file.path(dir.results, "figures", "plasma_pfas_unadjusted", "longitudinal_group_outcomes", "model_2c_plots.2.pdf"), width = 22, height = 20)
wrap_plots(mod2c_plots, ncol=3)
```

#model 2d results +plots
```{r}
mod2d_list <- lapply(group_time_outcomes, function(outcome) {
  model_2d(dat_long, outcome)
})
names(mod2d_list) <- group_time_outcomes

mod2d_results<- bind_rows(
  lapply(mod2d_list, `[[`, "tidy")
)
mod2d_results
WriteXLS(mod2d_results,file.path(dir.results, "csv", "plasma_pfas_unadjusted","longitudinal_group_outcomes", "model_2d_results.2.xlsx"))

mod2d_plots <- plot_model_2d(data = dat_long)

pdf(file.path(dir.results, "figures", "plasma_pfas_unadjusted", "longitudinal_group_outcomes", "model_2d_plots.2.pdf"), width = 22, height = 20)
wrap_plots(mod2d_plots, ncol=3)
```
#--------------------------------------------
# Outlier diagnostics for longitudinal models
# Models 2a-2d: group outcomes only
#--------------------------------------------
```{r}
#model adjustment
make_model2_formula <- function(outcome, model = "2d") {
  rhs <- c("visit_f", "group", "visit_f:group", "age", "sex")
  
  if (model %in% c("2c", "2d")) {
    rhs <- c(rhs, "tan_stage")
  }
  
  if (model %in% c("2b", "2d") &&
      !outcome %in% c("acr_u", "log_acr_u")) {
    rhs <- c(rhs, "log_uACR_base")
  }
  
  as.formula(
    paste0(outcome, " ~ ", paste(rhs, collapse = " + "),
           " + (1 | record_id)")
  )
}

```

```{r}
run_longitudinal_outlier_check <- function(data,
                                           outcome,
                                           model = "2a",
                                           resid_cutoff = 3) {
  
  form <- make_model2_formula(outcome, model = model)
  
  needed_vars <- unique(
    c("record_id", "group", "visit_f", "visit_num", all.vars(form))
  )
  
  model_data <- data %>%
    dplyr::select(dplyr::all_of(needed_vars)) %>%
    tidyr::drop_na()
  
  mod <- lmerTest::lmer(
    formula = form,
    data = model_data,
    REML = FALSE
  )
  
  #Raw outcome outliers by group x visit
  outcome_outliers <- model_data %>%
    dplyr::group_by(group, visit_f) %>%
    dplyr::mutate(
      outcome_z = as.numeric(scale(.data[[outcome]])),
      outcome_outlier = abs(outcome_z) > resid_cutoff
    ) %>%
    dplyr::ungroup() %>%
    dplyr::filter(outcome_outlier)
  
  #Model residual outliers
  resid_tbl <- model_data %>%
    dplyr::mutate(
      outcome = outcome,
      model = model,
      fitted = fitted(mod),
      resid = resid(mod),
      std_resid = as.numeric(scale(resid)),
      resid_outlier = abs(std_resid) > resid_cutoff
    )
  
  residual_outliers <- resid_tbl %>%
    dplyr::filter(resid_outlier)
  
  #Random intercept outliers
  random_effects <- ranef(mod)$record_id %>%
    tibble::rownames_to_column("record_id") %>%
    dplyr::rename(random_intercept = `(Intercept)`) %>%
    dplyr::mutate(
      outcome = outcome,
      model = model,
      random_intercept_z = as.numeric(scale(random_intercept)),
      random_intercept_outlier =
        abs(random_intercept_z) > resid_cutoff
    ) %>%
    dplyr::arrange(dplyr::desc(abs(random_intercept_z)))
  
  random_effect_outliers <- random_effects %>%
    dplyr::filter(random_intercept_outlier)
  
  # 4. Leave-one-subject-out sensitivity
  ids <- unique(model_data$record_id)
  
  loo_results <- purrr::map_dfr(ids, function(id) {
    
    dat_sub <- model_data %>%
      dplyr::filter(record_id != id)
    
    mod_sub <- tryCatch(
      lmerTest::lmer(
        formula = form,
        data = dat_sub,
        REML = FALSE
      ),
      error = function(e) NULL
    )
    
    if (is.null(mod_sub)) {
      return(NULL)
    }
    
    broom.mixed::tidy(
      mod_sub,
      effects = "fixed"
    ) %>%
      dplyr::mutate(
        outcome = outcome,
        model = model,
        removed_record_id = id
      )
  })
  
  primary_tidy <- broom.mixed::tidy(
    mod,
    effects = "fixed"
  ) %>%
    dplyr::mutate(
      outcome = outcome,
      model = model,
      removed_record_id = NA_character_
    )
  
  list(
    model = mod,
    model_data = model_data,
    primary_tidy = primary_tidy,
    outcome_outliers = outcome_outliers,
    residuals = resid_tbl,
    residual_outliers = residual_outliers,
    random_effects = random_effects,
    random_effect_outliers = random_effect_outliers,
    loo_results = loo_results
  )
}
```

```{r}
outlier_list <- lapply(group_time_outcomes, function(outcome) {
  
  message("Running outlier diagnostics for: ", outcome)
  
  tryCatch(
    run_longitudinal_outlier_check(
      data = dat_long,
      outcome = outcome,
      model = "2a"
    ),
    error = function(e) {
      message("FAILED: ", outcome)
      message(conditionMessage(e))
      NULL
    }
  )
})

names(outlier_list) <- group_time_outcomes

failed_outcomes <- names(outlier_list)[
  sapply(outlier_list, is.null)
]

failed_outcomes

outlier_list <- outlier_list[
  !sapply(outlier_list, is.null)
]

```
```{r}
outcome_outliers <- bind_rows(
  lapply(outlier_list, `[[`, "outcome_outliers") #is the value unusual for others in the same group and visit?
)

residual_outliers <- bind_rows(
  lapply(outlier_list, `[[`, "residual_outliers") #is the value unusual after accounting for the model
)

random_effects<- bind_rows(
  lapply(outlier_list, `[[`, "random_effects") #is the participent consistently different across visits? are they always different from everyone else
)

loo_results <- bind_rows(
  lapply(outlier_list, `[[`, "loo_results") #leaving one participant out sensitivity analysis
)

primary_results <- bind_rows(
  lapply(outlier_list, `[[`, "primary_tidy")
)
```

```{r}
loo_key_terms <- loo_results %>%
  filter(grepl(c("group",":"), term))
  

loo_summary<- loo_key_terms %>%
  left_join(
    primary_results %>%
      select(outcome, term, primary_estimate = estimate),
    by = c("outcome", "term")
  ) %>%
  mutate(
    estimate_change = estimate - primary_estimate,
    abs_change = abs(estimate_change),
    percent_change = 100 * abs_change / abs(primary_estimate)
  ) %>%
  group_by(outcome, term) %>%
  summarise(
    primary_estimate = first(primary_estimate),
    min_loo_estimate = min(estimate, na.rm = TRUE),
    max_loo_estimate = max(estimate, na.rm = TRUE),
    max_abs_change = max(abs_change, na.rm = TRUE),
    max_percent_change = max(percent_change, na.rm = TRUE),
    most_influential_id = removed_record_id[which.max(abs_change)],
    .groups = "drop"
  ) %>%
  arrange(desc(max_abs_change))

loo_summary <- loo_summary2 %>% 
  filter(max_percent_change >= 50)
```

```{r}
plot_loo_summary <- function(loo_summary_df,
                             outcome_keep = NULL,
                             term_keep = "visit_fYear 2:groupObese Control") {
  
  plot_df <- loo_summary_df %>%
    dplyr::filter(term == term_keep)
  
  if (!is.null(outcome_keep)) {
    plot_df <- plot_df %>%
      dplyr::filter(outcome %in% outcome_keep)
  }
  
  plot_df <- plot_df %>%
    dplyr::mutate(
      outcome = factor(outcome, levels = outcome[order(primary_estimate)])
    )
  
  ggplot(
    plot_df,
    aes(y = outcome)
  ) +
    geom_segment(
      aes(
        x = min_loo_estimate,
        xend = max_loo_estimate,
        yend = outcome
      ),
      linewidth = 1
    ) +
    geom_point(
      aes(x = primary_estimate),
      size = 3
    ) +
    geom_vline(
      xintercept = 0,
      linetype = "dashed"
    ) +
    theme_bw() +
    labs(
      title = paste0("Leave-one-out sensitivity: ", term_keep),
      x = "Coefficient estimate",
      y = "Outcome"
    )
}
```

```{r}
plot_loo_summary(
  loo_summary2,
  term_keep = "visit_fYear 2:groupObese Control"
)
```
```{r}
plot_loo_percent_change <- function(loo_summary_df,
                                    term_keep = "visit_fYear 2:groupType 2 Diabetes") {
  
  loo_summary_df %>%
    dplyr::filter(term == term_keep) %>%
    dplyr::mutate(
      outcome = factor(
        outcome,
        levels = outcome[order(max_percent_change)]
      )
    ) %>%
    ggplot(
      aes(
        x = max_percent_change,
        y = outcome
      )
    ) +
    geom_col() +
    geom_vline(
      xintercept = c(25, 50, 100),
      linetype = "dashed"
    ) +
    theme_bw() +
    labs(
      title = paste0("Maximum leave-one-out percent change: ", term_keep),
      x = "Maximum percent change from primary estimate",
      y = "Outcome"
    )
}
```

```{r}
plot_loo_percent_change(
  loo_summary2,
  term_keep = "visit_fYear 2:groupType 2 Diabetes"
)
```
```{r}
plot_influential_trajectories <- function(data,
                                          outcome,
                                          ids,
                                          y_label = outcome) {
  
  data %>%
    dplyr::filter(!is.na(.data[[outcome]])) %>%
    dplyr::mutate(
      highlight = record_id %in% ids
    ) %>%
    ggplot(
      aes(
        x = visit_num,
        y = .data[[outcome]],
        group = record_id
      )
    ) +
    geom_line(
      data = . %>% dplyr::filter(!highlight),
      alpha = 0.2
    ) +
    geom_point(
      data = . %>% dplyr::filter(!highlight),
      alpha = 0.2
    ) +
    geom_line(
      data = . %>% dplyr::filter(highlight),
      aes(color = record_id),
      linewidth = 1.2
    ) +
    geom_point(
      data = . %>% dplyr::filter(highlight),
      aes(color = record_id),
      size = 2.5
    ) +
    facet_wrap(~ group) +
    theme_bw() +
    labs(
      title = paste("Influential participant trajectories:", y_label),
      x = "Visit",
      y = y_label,
      color = "Participant"
    )
}
```

```{r}
plot_influential_trajectories(
  data = dat_long,
  outcome = "bl_dheas",
  ids = c("PAN-08-T", "PAN-01-T", "PAN-75-O", "PAN-98-C")
)

plot_influential_trajectories(
  data = dat_long,
  outcome = "igf_1",
  ids = c("PAN-08-T", "PAN-36-O", "PAN-21-C")
)

plot_influential_trajectories(
  data = dat_long,
  outcome = "bmi",
  ids = c("PAN-08-T", "PAN-38-O", "PAN-5-T"))


plot_influential_trajectories(
  data = dat_long,
  outcome = "dexa_fat_kg",
  ids = c("PAN-08-T", "PAN-94-O")
)

plot_influential_trajectories(
  data = dat_long,
  outcome = "dexa_lean_kg",
  ids = c("PAN-02-T", "PAN-5-T", "PAN-90-O", "PAN-60-O")
)

plot_influential_trajectories(
  data = dat_long,
  outcome = "eGFR_CKiD_U25_avg",
  ids = c("PAN-05-T", "PAN-24-T", "PAN-28-C")
)

plot_influential_trajectories(
  data = dat_long,
  outcome = "erpf_bsa_plasma",
  ids = c("PAN-40-C")
)

plot_influential_trajectories(
  data = dat_long,
  outcome = "avg_k_r2",
  ids = c("PAN-94-O")
)


dat_long %>% select(record_id, visit, group, avg_k_r2, bmi, eGFR_CKiD_U25_avg) %>%
  filter(group == "Obese Control")
```


```{r}
PAN05 <- dat_long %>% 
  filter(record_id == "PAN-05-T")
PAN05$dexa_lean_kg


PAN40 <- dat_long %>% 
  filter(record_id == "PAN-40-C")

PAN28 <- dat_long %>% 
  filter(record_id == "PAN-28-C")

PAN28 %>% select(visit, creatinine_s, cystatin_c_s, sex, age, height)
PAN40 %>% select(visit, creatinine_s, cystatin_c_s, sex, age, height)
```


#-------------------------------------------------------------------------------
# Model 3: baseline PFAS predicting longitudinal outcomes 
# a) age and sex adjusted
# b) adjusted for basline acr
# c) adjusted for tanner stage
# d) adjusted for baseline acr and tanner stage
#-------------------------------------------------------------------------------
```{r}
mod3a_list <- lapply(group_time_outcomes, function(outcome) {
  lapply(pfas_group_vars, function(exposure) {
    model_3a(dat_long, outcome, exposure)
  })
})
names(mod3a_list) <- group_time_outcomes

mod3a_results <- dplyr::bind_rows(
  lapply(mod3a_list, function(x) {
    dplyr::bind_rows(lapply(x, `[[`, "tidy"))
  })
)

mod3a_results

mod3a_results_fdr <- mod3a_results %>%
  dplyr::mutate(
    term_type = dplyr::case_when(
      term == exposure ~ "main",
      term == paste0(exposure, ":visit_num") ~ "interaction",
      term == paste0("visit_num:", exposure) ~ "interaction",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::group_by(term_type) %>%
  dplyr::mutate(
    p_fdr = ifelse(!is.na(term_type), p.adjust(p.value, method = "fdr"), NA_real_),
    sig_fdr = dplyr::case_when(
      !is.na(p_fdr) & p_fdr < 0.001 ~ "***",
      !is.na(p_fdr) & p_fdr < 0.01  ~ "**",
      !is.na(p_fdr) & p_fdr < 0.05  ~ "*",
      TRUE ~ ""
    )
  ) %>%
  dplyr::ungroup()


#WriteXLS(mod3a_results, file.path(dir.results, "csv", "plasma_pfas_unadjusted", "pfas_predicting_longitudinal_outcomes", "model_3a_results.xlsx"))
```
#-----------------
#results vis heatmaps
#-----------------
```{r}
#------------------------------
#plotting model 3a: heatmaps
#------------------------------
hm_mod3a <- mod3a_results_fdr %>%
  dplyr::group_by(outcome, exposure) %>%
  dplyr::summarise(
    beta_main = estimate[term == exposure][1],
    pval_main = p.value[term == exposure][1],
    pval_main_fdr = p_fdr[term == exposure][1],
    beta_int = estimate[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    pval_int = p.value[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    pval_int_fdr = p_fdr[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    beta_visit0 = beta_main,
    beta_visit1 = beta_main + beta_int,
    beta_visit2 = beta_main + 2 * beta_int
  ) %>%
  dplyr::select(
    outcome, exposure,
    beta_visit0, beta_visit1, beta_visit2,
    pval_main, pval_main_fdr,
    pval_int, pval_int_fdr
  ) %>%
  tidyr::pivot_longer(
    cols = c(beta_visit0, beta_visit1, beta_visit2),
    names_to = "visit",
    values_to = "beta"
  ) %>%
  dplyr::mutate(
    visit = factor(
      visit,
      levels = c("beta_visit0", "beta_visit1", "beta_visit2"),
      labels = c("Baseline", "Year 1", "Year 2")
    ),
    sig = dplyr::case_when(
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.001 ~ "***",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.01  ~ "**",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.05  ~ "*",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.001 ~ "***",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.01  ~ "**",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.05  ~ "*",
      TRUE ~ ""
    ),
    exposure = factor(
      exposure,
      levels = pfas_group_vars,
      labels = sapply(pfas_group_vars, get_pfas_label)
    ),
    outcome = factor(
      outcome,
      levels = group_time_outcomes,
      labels = sapply(group_time_outcomes, get_outcome_label)
    )
  )
```

```{r}
beta_mod3a <- max(abs(hm_mod3a$beta), na.rm = TRUE)

plot_model_3a<- ggplot(
  hm_mod3a,
  aes(x = exposure, y = outcome, fill = beta)
) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sig), size = 3.5, color = "black") +
  scale_fill_gradient2(
    low = "#2166AC",
    mid = "white",
    high = "#B2182B",
    midpoint = 0,
    limits = c(-beta_mod3a, beta_mod3a),
    name = "Beta"
  ) +
  facet_wrap(~ visit, nrow = 1) +
  labs(
    title = "Baseline PFAS associations with longitudinal outcomes",
    subtitle = paste0(
      "Outcome ~ PFAS + PFAS*visit + Age + Sex)\n",
      "* FDR<0.05  ** FDR<0.01  *** FDR<0.001"
    ),
    x = "Baseline PFAS",
    y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )



plot_model_3a
```

#---------------------------
#model 3b results + heatmaps
#---------------------------
```{r}
mod3b_list <- lapply(group_time_outcomes, function(outcome) {
  lapply(pfas_group_vars, function(exposure) {
    model_3b(dat_long, outcome, exposure)
  })
})
names(mod3b_list) <- group_time_outcomes

mod_3b_results <- dplyr::bind_rows(
  lapply(mod3b_list, function(x) {
    dplyr::bind_rows(lapply(x, `[[`, "tidy"))
  })
)

mod_3b_results
#WriteXLS(mod_3b_results, file.path(dir.results, "csv", "plasma_pfas_unadjusted", "pfas_predicting_longitudinal_outcomes", "model_3b_results.xlsx"))
```

```{r}
mod3b_results_fdr <- mod_3b_results %>%
  dplyr::mutate(
    term_type = dplyr::case_when(
      term == exposure ~ "main",
      term == paste0(exposure, ":visit_num") ~ "interaction",
      term == paste0("visit_num:", exposure) ~ "interaction",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::group_by(term_type) %>%
  dplyr::mutate(
    p_fdr = ifelse(!is.na(term_type), p.adjust(p.value, method = "fdr"), NA_real_),
    sig_fdr = dplyr::case_when(
      !is.na(p_fdr) & p_fdr < 0.001 ~ "***",
      !is.na(p_fdr) & p_fdr < 0.01  ~ "**",
      !is.na(p_fdr) & p_fdr < 0.05  ~ "*",
      TRUE ~ ""
    )
  ) %>%
  dplyr::ungroup()
```

```{r}
#-----------
#plotting 3b
#-----------
hm_mod3b <- mod3b_results_fdr %>%
  dplyr::group_by(outcome, exposure) %>%
  dplyr::summarise(
    beta_main = estimate[term == exposure][1],
    pval_main = p.value[term == exposure][1],
    pval_main_fdr = p_fdr[term == exposure][1],
    beta_int = estimate[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    pval_int = p.value[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    pval_int_fdr = p_fdr[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    beta_visit0 = beta_main,
    beta_visit1 = beta_main + beta_int,
    beta_visit2 = beta_main + 2 * beta_int
  ) %>%
  dplyr::select(
    outcome, exposure,
    beta_visit0, beta_visit1, beta_visit2,
    pval_main, pval_main_fdr,
    pval_int, pval_int_fdr
  ) %>%
  tidyr::pivot_longer(
    cols = c(beta_visit0, beta_visit1, beta_visit2),
    names_to = "visit",
    values_to = "beta"
  ) %>%
  dplyr::mutate(
    visit = factor(
      visit,
      levels = c("beta_visit0", "beta_visit1", "beta_visit2"),
      labels = c("Baseline", "Year 1", "Year 2")
    ),
    sig = dplyr::case_when(
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.001 ~ "***",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.01  ~ "**",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.05  ~ "*",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.001 ~ "***",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.01  ~ "**",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.05  ~ "*",
      TRUE ~ ""
    ),
    exposure = factor(
      exposure,
      levels = pfas_group_vars,
      labels = sapply(pfas_group_vars, get_pfas_label)
    ),
    outcome = factor(
      outcome,
      levels = group_time_outcomes,
      labels = sapply(group_time_outcomes, get_outcome_label)
    )
  )
```

```{r}
beta_mod3b <- max(abs(hm_mod3b$beta), na.rm = TRUE)

plot_mod3b <- ggplot(
  hm_mod3b,
  aes(x = exposure, y = outcome, fill = beta)
) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sig), size = 3.5, color = "black") +
  scale_fill_gradient2(
    low = "#2166AC",
    mid = "white",
    high = "#B2182B",
    midpoint = 0,
    limits = c(-beta_mod3b, beta_mod3b),
    name = "Beta"
  ) +
  facet_wrap(~ visit, nrow = 1) +
  labs(
    title = "Baseline PFAS associations with longitudinal outcomes",
    subtitle = paste0(
      "Baseline uACR-adjusted mixed models.\n",
      "* FDR<0.05  ** FDR<0.01  *** FDR<0.001"
    ),
    x = "Baseline PFAS",
    y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

plot_mod3b
```

#------------------------
#model 3c results + plots
#------------------------
```{r}
mod3c_list<- lapply(group_time_outcomes, function(outcome) {
  lapply(pfas_group_vars, function(exposure) {
    model_3c(dat_long, outcome, exposure)
  })
})
names(mod3c_list) <- group_time_outcomes

mod3c_results <- dplyr::bind_rows(
  lapply(mod3c_list, function(x) {
    dplyr::bind_rows(lapply(x, `[[`, "tidy"))
  })
)

mod3c_results
#WriteXLS(mod3c_results, file.path(dir.results, "csv", "plasma_pfas_unadjusted", "pfas_predicting_longitudinal_outcomes", "model_3c_results.xlsx"))
```

```{r}
mod3c_results_fdr <- mod3c_results %>%
  dplyr::mutate(
    term_type = dplyr::case_when(
      term == exposure ~ "main",
      term == paste0(exposure, ":visit_num") ~ "interaction",
      term == paste0("visit_num:", exposure) ~ "interaction",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::group_by(term_type) %>%
  dplyr::mutate(
    p_fdr = ifelse(!is.na(term_type), p.adjust(p.value, method = "fdr"), NA_real_),
    sig_fdr = dplyr::case_when(
      !is.na(p_fdr) & p_fdr < 0.001 ~ "***",
      !is.na(p_fdr) & p_fdr < 0.01  ~ "**",
      !is.na(p_fdr) & p_fdr < 0.05  ~ "*",
      TRUE ~ ""
    )
  ) %>%
  dplyr::ungroup()
```

```{r}
hm_mod3c <- mod3c_results_fdr %>%
  dplyr::group_by(outcome, exposure) %>%
  dplyr::summarise(
    beta_main = estimate[term == exposure][1],
    pval_main = p.value[term == exposure][1],
    pval_main_fdr = p_fdr[term == exposure][1],
    beta_int = estimate[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    pval_int = p.value[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    pval_int_fdr = p_fdr[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    beta_visit0 = beta_main,
    beta_visit1 = beta_main + beta_int,
    beta_visit2 = beta_main + 2 * beta_int
  ) %>%
  dplyr::select(
    outcome, exposure,
    beta_visit0, beta_visit1, beta_visit2,
    pval_main, pval_main_fdr,
    pval_int, pval_int_fdr
  ) %>%
  tidyr::pivot_longer(
    cols = c(beta_visit0, beta_visit1, beta_visit2),
    names_to = "visit",
    values_to = "beta"
  ) %>%
  dplyr::mutate(
    visit = factor(
      visit,
      levels = c("beta_visit0", "beta_visit1", "beta_visit2"),
      labels = c("Baseline", "Year 1", "Year 2")
    ),
    sig = dplyr::case_when(
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.001 ~ "***",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.01  ~ "**",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.05  ~ "*",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.001 ~ "***",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.01  ~ "**",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.05  ~ "*",
      TRUE ~ ""
    ),
    exposure = factor(
      exposure,
      levels = pfas_group_vars,
      labels = sapply(pfas_group_vars, get_pfas_label)
    ),
    outcome = factor(
      outcome,
      levels = group_time_outcomes,
      labels = sapply(group_time_outcomes, get_outcome_label)
    )
  )

```

```{r}
beta_mod3c <- max(abs(hm_mod3c$beta), na.rm = TRUE)

plot_mod3c <- ggplot(
  hm_mod3c,
  aes(x = exposure, y = outcome, fill = beta)
) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sig), size = 3.5, color = "black") +
  scale_fill_gradient2(
    low = "#2166AC",
    mid = "white",
    high = "#B2182B",
    midpoint = 0,
    limits = c(-beta_mod3c, beta_mod3c),
    name = "Beta"
  ) +
  facet_wrap(~ visit, nrow = 1) +
  labs(
    title = "Baseline PFAS associations with longitudinal outcomes",
    subtitle = paste0(
      "Tanner-adjusted mixed models.\n",
      "* FDR<0.05  ** FDR<0.01  *** FDR<0.001"
    ),
    x = "Baseline PFAS",
    y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

plot_mod3c
```


#model 3d results + plots
```{r}
mod3d_list<- lapply(group_time_outcomes, function(outcome) {
  lapply(pfas_group_vars, function(exposure) {
    model_3d(dat_long, outcome, exposure)
  })
})
names(mod3d_list) <- group_time_outcomes

mod3d_results<- dplyr::bind_rows(
  lapply(mod3d_list, function(x) {
    dplyr::bind_rows(lapply(x, `[[`, "tidy"))
  })
)

mod3d_results
#WriteXLS(mod3d_results, file.path(dir.results, "csv", "plasma_pfas_unadjusted", "pfas_predicting_longitudinal_outcomes", "model_3d_results.xlsx"))
```




```{r}
mod3d_results_fdr<- mod3d_results %>%
  dplyr::mutate(
    term_type = dplyr::case_when(
      term == exposure ~ "main",
      term == paste0(exposure, ":visit_num") ~ "interaction",
      term == paste0("visit_num:", exposure) ~ "interaction",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::group_by(term_type) %>%
  dplyr::mutate(
    p_fdr = ifelse(!is.na(term_type), p.adjust(p.value, method = "fdr"), NA_real_),
    sig_fdr = dplyr::case_when(
      !is.na(p_fdr) & p_fdr < 0.001 ~ "***",
      !is.na(p_fdr) & p_fdr < 0.01  ~ "**",
      !is.na(p_fdr) & p_fdr < 0.05  ~ "*",
      TRUE ~ ""
    )
  ) %>%
  dplyr::ungroup()
```


```{r}
hm_mod3d <- mod3d_results_fdr %>%
  dplyr::group_by(outcome, exposure) %>%
  dplyr::summarise(
    
    # Main PFAS effect (baseline)
    beta_main = estimate[term == exposure][1],
    pval_main = p.value[term == exposure][1],
    pval_main_fdr = p_fdr[term == exposure][1],
    
    # Interaction (change over time)
    beta_int = estimate[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    
    pval_int = p.value[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    
    pval_int_fdr = p_fdr[
      term == paste0(exposure, ":visit_num") |
        term == paste0("visit_num:", exposure)
    ][1],
    
    .groups = "drop"
  ) %>%
  
  # Convert to visit-specific effects
  dplyr::mutate(
    beta_visit0 = beta_main,
    beta_visit1 = beta_main + beta_int,
    beta_visit2 = beta_main + 2 * beta_int
  ) %>%
  
  dplyr::select(
    outcome, exposure,
    beta_visit0, beta_visit1, beta_visit2,
    pval_main, pval_main_fdr,
    pval_int, pval_int_fdr
  ) %>%
  
  # Long format for plotting
  tidyr::pivot_longer(
    cols = c(beta_visit0, beta_visit1, beta_visit2),
    names_to = "visit",
    values_to = "beta"
  ) %>%
  
  dplyr::mutate(
    
    visit = factor(
      visit,
      levels = c("beta_visit0", "beta_visit1", "beta_visit2"),
      labels = c("Baseline", "Year 1", "Year 2")
    ),
    
    # Significance stars
    sig = dplyr::case_when(
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.001 ~ "***",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.01  ~ "**",
      visit == "Baseline" & !is.na(pval_main_fdr) & pval_main_fdr < 0.05  ~ "*",
      
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.001 ~ "***",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.01  ~ "**",
      visit %in% c("Year 1", "Year 2") & !is.na(pval_int_fdr) & pval_int_fdr < 0.05  ~ "*",
      
      TRUE ~ ""
    ),
    
    exposure = factor(
      exposure,
      levels = pfas_group_vars,
      labels = sapply(pfas_group_vars, get_pfas_label)
    ),
    
    outcome = factor(
      outcome,
      levels = group_time_outcomes,
      labels = sapply(group_time_outcomes, get_outcome_label)
    )
  )
```

```{r}
beta_mod3d <- max(abs(hm_mod3d$beta), na.rm = TRUE)

plot_mod3d <- ggplot(
  hm_mod3d,
  aes(x = exposure, y = outcome, fill = beta)
) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sig), size = 3.5, color = "black") +
  
  scale_fill_gradient2(
    low = "#2166AC",
    mid = "white",
    high = "#B2182B",
    midpoint = 0,
    limits = c(-beta_mod3d, beta_mod3d),
    name = "Beta"
  ) +
  
  facet_wrap(~ visit, nrow = 1) +
  
  labs(
    title = "Baseline PFAS associations with longitudinal outcomes",
    subtitle = paste0(
      "uACR + Tanner-adjusted mixed models\n",
      "* FDR<0.05  ** FDR<0.01  *** FDR<0.001"
    ),
    x = "Baseline PFAS",
    y = NULL
  ) +
  
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

plot_mod3d
```