---
title: "PFAS Analysis — PANTHER Study"
author: "Savanah Leidholt & Hailey Hampson"
date: last-modified
format:
  html:
    theme: cerulean
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# libraries, directories, data loading, variable definitions,
# helper functions, and all data prep go here

#load libraries

library(ggplot2)
library(tidyverse)
library(janitor)
library(lme4)
library(lmerTest)
library(broom)
library(broom.mixed)
library(emmeans)
library(ggeffects)
library(patchwork)
library(corrplot)
library(gt)
library(table1)
library(stringr)
library(ggpubr)
library(patchwork)

#Directories
#computer <- "mac studio"
computer <- "mac laptop"
if (computer == "mac studio") {
  user <- Sys.info()[["user"]]
  if (user == "sleidholt") {
    dir.dat <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
  }
} else {
  user <- Sys.info()[["user"]]
  if (user == "savanahleidholt") {
    dir.dat <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/PFAS/Results/unadjusted_pfas_results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad/Panther PFAS"
  } 
}


source(file.path(git_path, "PFAS_Analysis_SLL.R"))


```

------------------------------------------------------------------------

# Participant Characteristics

## Table 1: Demographics & Clinical Outcomes by Group

```{r}
#| label: table1-demographics
table1(
  ~ age + sex + bmi + race_ethnicity_condensed + hba1c + acr_u + 
    microalbumin_u + eGFR_CKiD_U25_avg + avg_c_r2| group,
  data = dat_baseline
)
```

## Table 2: Baseline PFAS Levels by Group

```{r}
#| label: table2-pfas

```

------------------------------------------------------------------------

# Baseline Group Comparisons

## Clinical Outcomes by Group

```{r}
#| label: baseline-outcomes-plots
#| fig-width: 20
#| fig-height: 18

plot_list <- lapply(baseline_model_outcomes, function(outcome) {
  plot_baseline_group_boxplot(
    data = dat_baseline,
    outcome = outcome,
    results_list = baseline_group_outcome_results
  )
})

wrap_plots(plot_list, ncol = 3)

```

## PFAS Levels by Group

```{r}
#| label: baseline-pfas-plots
#| fig-width: 20
#| fig-height: 18

plots_by_group <- lapply(
  pfas_group_vars,
  function(pfas) {
    plot_baseline_pfas_group_boxplot(
      data = dat_baseline,
      exposure = pfas,
      results_list = baseline_group_pfas_results
    )
  }
)

wrap_plots(plots_by_group, ncol = 3)
```

------------------------------------------------------------------------

# Model 1: Cross-Sectional PFAS–Outcome Associations

```{r}
#| label: Forest Plots
#| fig-width: 16
#| fig-height: 14

forest_df <- baseline_table %>%
  dplyr::mutate(
    conf.low = estimate - 1.96 * std.error,
    conf.high = estimate + 1.96 * std.error
  )

plot_list <- list()


for (outcome in unique(forest_df$outcome)) {
  df <- forest_df %>%
    dplyr::filter(outcome == !!outcome)
  p <- ggplot(df, aes(x = estimate, y = exposure)) +
  geom_point(aes(shape = sig_fdr)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(
    title = paste("PFAS associations with", get_outcome_label(outcome)),
    x = "Effect size (β)",
    y = "PFAS",
    shape = "FDR <0.05"
  )
plot_list[[outcome]] <- p
}


wrap_plots(plot_list, ncol = 3)
```


------------------------------------------------------------------------

# Model 2: Outcome Trajectories Over Time by Group

## Model 2A: Trajectory Plots Base Model
```{r}
#| label: model2a-trajectory-plots
#| fig-width: 20
#| fig-height: 18

plots <- plot_group_time_set_unadjusted(
  data = dat_long
)
wrap_plots(plots, ncol = 3)

```
## Model 2B: Trajectory Plots Base Model Adjusted for Baseline ACR
```{r}
#| label: model2b-trajectory-plots
#| fig-width: 20
#| fig-height: 18


plots <- plot_group_time_set(
  data = dat_long,
  adjust_acr = TRUE
)

wrap_plots(plots, ncol = 3)
```
## Model 2C: Trajectory Plots Base Model Adjusted for Tanner Stage
```{r}
#| label: model2c-trajectory-plots
#| fig-width: 20
#| fig-height: 18


plots <- plot_group_time_set_tanner(
  data = dat_long
)

wrap_plots(plots, ncol = 3)

```
## Model 2D: Trajectory Plots Base Model Adjusted for Baseline uACR & Tanner Stage
```{r}
#| label: model2d-trajectory-plots
#| fig-width: 20
#| fig-height: 18

plots <- plot_group_time_acr_tan_set(
  data = dat_long,
  adjust_acr = TRUE
)

wrap_plots(plots, ncol = 3)
```

------------------------------------------------------------------------

# Model 3: Baseline PFAS Predicting Longitudinal Outcomes

## Model 3A: Longitudinal Heatmap of Base Model
```{r}
#| label: model3a-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_base
```
## Model 3B: Longitudinal Heatmap Adjusted for Baseline uACR
```{r}
#| label: model3b-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_acr
```
## Model 3C: Longitudinal Heatmap Adjusted for Tanner Stage
```{r}
#| label: model3c-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_tanner
```
## Model 3D: Longitudinal Heatmap Adjusted for Baseline uACR & Tanner Stage
```{r}
#| label: model3d-heatmap
#| fig-width: 12
#| fig-height: 7
p_longitudinal_heatmap_acr_tanner
```
------------------------------------------------------------------------

# Appendix

```{r}
#| label: session-info
sessionInfo()
```
