---
title: "PANTHER PFAS Analysis — Handoff README"
author: "Hailey Hampson"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

---

# Project Overview

This project examines associations between per- and polyfluoroalkyl substance (PFAS) exposures and kidney function outcomes in the **PANTHER** study (a longitudinal pediatric cohort of youth-onset Type 2 Diabetes and controls). PFAS concentrations were measured from plasma samples collected by Emory University.

Key scientific questions:

1. Are baseline PFAS levels associated with albuminuria (ACR-urine) cross-sectionally and longitudinally?
2. Are baseline PFAS levels associated with kidney function outcomes (GFR, ERPF, insulin sensitivity, renal perfusion/oxygenation) at baseline and over follow-up?
3. How do kidney outcomes change over time by diabetes group (T2D, Obese Control, Lean Control)?

**PI:** Petter Bjornstad
**Biostatisticians:** Hailey Hampson; Savanah Leidholt (plasma volume adjustment)
**Core:** Laura Pyle — Biostatistics Core, University of Washington

---

# Directory Structure

## Code Repository

```
PANTHER PFAS/
├── README.Rmd                          # This file
├── PANTHER PFAS Analysis.Rmd           # PRIMARY analysis script
├── PANTHER_PFAS_Cleaning.R             # Data cleaning script
├── PFAS_Adj_Plasma_Volume.Rmd          # Plasma volume adjustment analysis (Leidholt & Hampson)
├── PFAS_Analysis_SLL.qmd/.R/.Rmd       # Savanah Leidholt analysis files
├── Albuminuria TK model Simplified_Panther_heh.R  # Toxicokinetic model (HEH version)
├── Albuminuria TK model Simplified.R   # Original TK model
├── Functions.R                         # Custom R functions
├── Libraries.R                         # All package imports
├── PANTHER_PFAS_Analysis.R             # R script version of the analysis
├── PANTHER PFAS.qmd                    # Presentation slides (Quarto)
└── PANTHER PFAS.Rproj                  # RStudio project file
```

## Results Directory (OneDrive)

```
PANTHER/Results/PFAS Results/
├── figures/                    # Summary outcome figures (used in Quarto presentation)
├── crude/                      # Unadjusted scatter plots (per PFAS × outcome)
├── adj_age_sex/                # Age & sex adjusted scatter plots
├── adj_group/                  # Diabetes group adjusted scatter plots
├── longitudinal_lme/           # Longitudinal LME results
│   ├── crude/                  # Unadjusted LME plots
│   ├── adj_acr_u/              # Adj. for baseline ACR-urine
│   ├── adj_age_sex_acr/        # Adj. for age, sex, ACR-urine
│   ├── adj_group_acr/          # Adj. for group + ACR-urine
│   ├── heatmap_*.png           # Result heatmaps (4 adjustment models)
│   ├── lme_results.csv/.rds    # LME result tables
│   └── lme_results2.csv/.rds   # LME result tables (secondary)
├── cs_results.csv/.rds         # Cross-sectional result tables
├── PANTHER Tables.xlsx         # Summary tables for manuscript/report
├── PANTHER_PFAS_Table1_Baseline.png  # Table 1 (baseline characteristics)
├── PFAS_PANTHER_03_17.html     # Rendered HTML report (March 17 version)
├── pfas_acru_plots*.pdf        # ACR-urine trajectory plots (with/without outlier, with/without PFOA)
├── pfas_eGFR_plots.pdf         # eGFR trajectory plots
├── pfas_microalbumin_plots*.pdf # Microalbumin plots
├── baseline_pfas_followup_acru_yr*.pdf # Baseline PFAS × follow-up ACRu (yr 1 & 2)
└── longitudinal_outcomes.pdf   # Longitudinal outcome summary plots
```

---

# Data Sources

## Primary Data

All data are sourced from the harmonized PANTHER dataset:

```{r}
# Update dir.dat for your machine (see path setup in §1 of the analysis script)
dir.dat <- "/Users/[your_username]/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive"

dat <- read.csv(fs::path(dir.dat, "Data Harmonization", "Data Clean", "harmonized_dataset.csv"), na.strings = "")
lod <- read.csv(fs::path(dir.dat, "PANTHER", "Data_Raw", "PFAS Emory 2025", "8- PFAS Quantification", "PFAS_LODs.csv"))
```

## Study Visits

| Visit Label | Description |
|-------------|-------------|
| `baseline` / `screening` | Collapsed to a single baseline timepoint (averaged) |
| `year_1` | Year 1 follow-up |
| `year_2` | Year 2 follow-up |

## Participant Groups

| Group | Description |
|-------|-------------|
| T2D | Youth-onset Type 2 Diabetes |
| Obese Control | Obese youth without diabetes |
| Lean Control | Lean youth without diabetes |

---

# PFAS Compounds

## LOD Filtering

PFAS compounds were included only if **< 2/3 of values were below the limit of detection (LOD)**. Values below the LOD were imputed with **LOD / √2**.

The following compounds were assessed (from Emory PFAS quantification panel):

**Included (< 2/3 below LOD — retained and imputed):**

| Compound | Class |
|----------|-------|
| N.MeFOSAA | Sulfonamide |
| PFDA | Carboxylic acid |
| PFHpA | Carboxylic acid |
| PFNA | Carboxylic acid |
| PFOA | Carboxylic acid |
| PFBS | Sulfonate |
| PFHps | Sulfonate |
| PFHxS | Sulfonate |
| PFOS | Sulfonate |
| PFPeAS | Sulfonate |

**Excluded (≥ 2/3 below LOD):** N.EtFOSAA, PFBA, PFDoA, PFHxA, PFPeA, PFTeDA, PFTrDA, PFUnA, PFDoS, PFNS, PFOSA

All retained PFAS were **log-transformed and scaled** prior to modeling.

---

# Outcomes

| Variable | Description |
|----------|-------------|
| `acr_u` | Albumin-to-creatinine ratio (urine) — primary albuminuria outcome |
| `eGFR_CKiD_U25_avg` | Estimated GFR (CKiD U25 equation, averaged) |
| `erpf_bsa_plasma` | Effective renal plasma flow (plasma-based, BSA-adjusted) |
| `erpf_bsa_plasma_urine` | ERPF (plasma + urine-based, BSA-adjusted) |
| `mm_si` | Insulin sensitivity (mixed-meal clamp) |
| `pah_clear_bsa` | PAH clearance (BSA-adjusted) |
| `avg_k_r2` | Renal cortex R2* (MRI oxygenation marker) |
| `avg_c_r2` | Cortical R2* (MRI oxygenation marker) |
| `pcasl3d_left` | Left kidney perfusion (3D pCASL MRI) |
| `pcasl3d_right` | Right kidney perfusion (3D pCASL MRI) |

---

# Analysis Workflow

## 1. Data Cleaning (`PANTHER_PFAS_Cleaning.R`)

- Filters harmonized dataset to PANTHER participants only
- Collapses baseline/screening visits (numeric variables averaged, character variables take last non-NA)
- Constructs `race_ethnicity_condensed` variable
- Loads PFAS LOD data and applies LOD/√2 imputation for retained compounds

## 2. PFAS → ACRu (`§2` in main Rmd)

Cross-sectional and longitudinal associations between baseline PFAS and albumin-to-creatinine ratio.

- **Baseline:** Linear regression of ACRu ~ PFAS (crude, adj. age/sex, adj. group)
- **Longitudinal:** LME of ACRu trajectory ~ baseline PFAS × time

## 3. ACRu → PFAS (`§3` in main Rmd)

Reverse-direction analysis: ACRu as predictor of PFAS levels (exploratory, cross-sectional and longitudinal).

## 4. Longitudinal Outcome Trends by Group (`§4` in main Rmd)

Characterizes how each kidney outcome changes over time (baseline → year 1 → year 2) stratified by diabetes group. Generates trajectory plots saved to `figures/` and `longitudinal_outcomes.pdf`. Outlier flagging is logged to `figures/outlier_log.rds`.

## 5. PFAS → Kidney Outcomes (`§5` in main Rmd)

Associates baseline PFAS with the full set of kidney outcomes (GFR, ERPF, insulin sensitivity, renal perfusion, oxygenation), adjusting for baseline ACRu, age, sex, and group.

| Sub-analysis | Method | Output |
|-------------|--------|--------|
| §5a — Cross-sectional | Linear regression at baseline | Scatter plots in `crude/`, `adj_age_sex/`, `adj_group/`; `cs_results.csv` |
| §5b — Longitudinal | LME (baseline PFAS × time) | Plots in `longitudinal_lme/`; `lme_results.csv` |

Result heatmaps (`heatmap_*.png`) summarize significance across all PFAS × outcome combinations for each adjustment model.

---

# Path Setup

The analysis script supports multiple users via a `computer` toggle. Update this at the top of `PANTHER PFAS Analysis.Rmd`:

```{r}
computer <- "mac laptop"  # or "mac studio"
```

Then add your username/paths in the appropriate `if (user == ...)` block, or simply hardcode:

```{r}
dir.dat     <- "/Users/[your_username]/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive"
dir.results <- file.path(dir.dat, "PANTHER/Results")
git_path    <- "/Users/[your_username]/CHCO-Code/Petter Bjornstad"
```

Then source helpers:

```{r}
source("Libraries.R")
source("Functions.R")
```

---

# Key R Packages

| Package | Purpose |
|---------|---------|
| `lme4` / `lmerTest` | Linear mixed effects models |
| `tidyverse` | Data wrangling and plotting |
| `ggplot2` | Scatter plots, heatmaps, trajectory plots |
| `table1` | Baseline characteristics (Table 1) |
| `janitor` | Data cleaning utilities |
| `broom` | Tidy model output |
| `fs` | File path management |
| `openxlsx` | Excel output |

---

# Results Files Summary

| File | Description |
|------|-------------|
| `cs_results.csv/.rds` | Cross-sectional linear regression results (all PFAS × outcomes) |
| `lme_results.csv/.rds` | LME results (baseline PFAS × longitudinal outcomes) |
| `lme_results2.csv/.rds` | LME results (secondary model specification) |
| `PANTHER Tables.xlsx` | Formatted tables for manuscript/collaborator review |
| `PANTHER_PFAS_Table1_Baseline.png` | Table 1 — baseline characteristics by group |
| `PFAS_PANTHER_03_17.html` | Full rendered HTML report (March 17, 2026 version) |

---

# Notes

- PFOA is included/excluded across different sensitivity analyses — some plot files are labeled `_withPFOA` vs. without; check `pfas_acru_plots_no_outlier_withPFOA.pdf` vs. `pfas_acru_plots_no_outlier.pdf`.
- Outlier sensitivity analyses are built into §4; the outlier log is saved to `figures/outlier_log.rds` for reference.
- The `PFAS_Adj_Plasma_Volume.Rmd` script (Leidholt & Hampson) adjusts raw PFAS concentrations for plasma volume — this was an exploratory sensitivity analysis and results are separate from the main analysis.
- `PFAS_Analysis_SLL.*` files were developed by Savanah Leidholt and may overlap with the main analysis — check for any non-redundant analyses before archiving.

---

# Contact

- **Hailey Hampson** — hehampson@uw.edu (primary biostatistician)
- **Savanah Leidholt** — sleidholdt@uw.edu (plasma volume adjustment)
- **Laura Pyle** — lpyle@uw.edu (Biostatistics Core lead)
- **Petter Bjornstad** — petter.bjornstad@cuanschutz.edu (PI)
