---
title: "RPC2 Endothelial Cell CMR Report"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

# Setup

```{r setup, echo = F, message = F, warning = F}
library(dplyr)
library(aws.s3)
library(jsonlite)
library(table1)
library(knitr)
library(kableExtra)
library(tidyr)
library(stringr)
library(lubridate)
```

```{r paths, echo = F}
user <- Sys.info()[["user"]]

if (user == "choiyej") {
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") {
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") {
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else {
  stop("Unknown user: please specify root path for this user.")
}

Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```

```{r helpers, echo = F}
# Read an object from S3 into R (e.g. a CSV via read.csv)
s3read_using_region <- function(FUN, ..., object, bucket, region = NULL, opts = NULL, filename = NULL) {
  if (missing(bucket)) bucket <- get_bucketname(object)
  object <- get_objectkey(object)
  tmp <- if (is.character(filename)) file.path(tempdir(TRUE), filename) else tempfile(fileext = paste0(".", tools::file_ext(object)))
  on.exit(unlink(tmp))
  if (!is.null(region)) {
    if (is.null(opts)) opts <- list(region = region) else opts$region <- region
  }
  if (is.null(opts)) {
    save_object(bucket = bucket, object = object, file = tmp)
  } else {
    do.call("save_object", c(list(bucket = bucket, object = object, file = tmp), opts))
  }
  FUN(tmp, ...)
}

# Download a saved figure into a document-relative folder for embedding.
# A document-relative path (not the R tempdir) is required so pandoc can
# resolve and embed the image at render time.
# Returns the local relative path, or NULL if the object is not found.
fig_dir <- "ec_cmr_figs"
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)

get_s3_fig <- function(object, bucket = "scrna") {
  dest <- file.path(fig_dir, gsub("[/ ]", "_", object))
  ok <- tryCatch({
    save_object(object = object, bucket = bucket, file = dest, region = "")
    TRUE
  }, error = function(e) FALSE)
  if (ok && file.exists(dest) && file.info(dest)$size > 0) dest else NULL
}

# Imaging variable labels (carried over from the analysis script)
im_labels <- c(
  lv_hr                  = "LV Heart rate (1/min)",
  rvedv                  = "RV EDV (mL)",
  lvedv                  = "LV EDV (mL)",
  rvesv                  = "RV ESV (mL)",
  lvesv                  = "LV ESV (mL)",
  rv_stroke_volume       = "RV stroke volume (mL)",
  lv_stroke_volume       = "LV stroke volume (mL)",
  rvedv_bsa              = "RVEDV BSA (mL/m²)",
  lvedv_bsa              = "LVEDV BSA (mL/m²)",
  rvesv_bsa              = "RVESV BSA (mL/m²)",
  lvesv_bsa              = "LVESV BSA (mL/m²)",
  rv_stroke_volume_bsa   = "RV stroke volume BSA (mL/m²)",
  lv_stroke_volume_bsa   = "LV stroke volume BSA (mL/m²)",
  rv_cardiac_output      = "RV cardiac output",
  lv_cardiac_output      = "LV cardiac output",
  lves_mass              = "LV myocardial mass (systole)",
  lved_mass              = "LV mass (g)",
  lves_mass_bsa          = "LV myocardial mass BSA (systole)",
  lved_mass_bsa          = "LV mass BSA (g/m²)",
  lv_myo_thickness_total = "LV thickness total (mm)",
  lv_myo_thickness_dias  = "LV thickness (mm)",
  rvef                   = "RV ejection fraction (%)",
  lvef                   = "LV ejection fraction (%)",
  gls                    = "Global longitudinal strain (%)",
  gcs                    = "Global circumferential strain (%)",
  grs                    = "Global radial strain (%)",
  af_high5pctwss         = "Aortic wall shear stress (top 5%)",
  af_medianwss           = "Aortic wall shear stress (median)",
  af_meanwss             = "Mean wall shear stress (Pa)",
  af_pwv                 = "Pulse wave velocity (m/s)",
  tt1                    = "CMR T1 (ms)",
  tt2                    = "CMR T2* (ms)",
  tr2                    = "R2* (1/s)",
  mv_flow_e              = "E wave (cm/s)",
  mv_flow_a              = "A wave (cm/s)",
  mv_flow_ea             = "E/A ratio"
)
```

# Included participants


```{r descriptive, echo = F, warning = F, message = F, results='asis'}
meta <- s3read_using_region(FUN = read.csv,
                            bucket = "scrna",
                            object = "Projects/EC CMR/Data Clean/pb90_ec_cmr_meta.csv",
                            region = "")

# Drop the barcode row-name column written by write.csv
meta <- meta[, setdiff(names(meta), c("X", "X.1")), drop = FALSE]

# Collapse repeated per-cell rows to one row per participant
meta_subj <- meta %>% dplyr::distinct(kit_id, .keep_all = T)

cat("Participants included:", nrow(meta_subj), "  |  Endothelial cells included:", nrow(meta), "\n")

summary(arsenal::tableby(group ~ age + sex + hba1c + kwt(acr_u, "medianq1q3", "range") +  study, data = meta_subj), digits = 1)
```

# EC association with CMR parameters (NEBULA)

Gene level associations between endothelial cell expression and each CMR parameter.

```{r nebula-read, echo = F}
nebula_results <- s3read_using_region(FUN = read.csv,
                                      bucket = "scrna",
                                      object = "Projects/EC CMR/Results/cmr_ec_nebula_compiled.csv",
                                      region = "")
vars_nebula <- unique(nebula_results$var)
```

## Significant genes per parameter

Count of significant genes per CMR parameter, at both raw p < 0.05 and FDR < 0.05.

```{r nebula-sig, echo = F}
n_p   <- nebula_results %>% filter(p   < 0.05) %>% count(var, name = "n_p_sig")
n_fdr <- nebula_results %>% filter(fdr < 0.05) %>% count(var, name = "n_fdr_sig")

sig_tab <- full_join(n_p, n_fdr, by = "var") %>%
  mutate(across(c(n_p_sig, n_fdr_sig), ~ tidyr::replace_na(., 0L)),
         label = ifelse(var %in% names(im_labels), im_labels[var], var)) %>%
  arrange(desc(n_p_sig)) %>%
  select(Parameter = label, `Variable` = var,
         `p < 0.05 genes` = n_p_sig, `FDR significant genes` = n_fdr_sig)

sig_tab %>% kable() %>% kable_styling(full_width = FALSE)
```

## Combined volcano plots

```{r facet-volcano, echo = F, out.width = "100%"}
fp <- get_s3_fig("Projects/EC CMR/Results/Figures/cmr_faceted_volcano.png")
if (!is.null(fp)) knitr::include_graphics(fp)
```

```{r facet-volcano-fdr, echo = F, out.width = "100%"}
fp <- get_s3_fig("Projects/EC CMR/Results/Figures/cmr_faceted_volcano_fdr.png")
if (!is.null(fp)) knitr::include_graphics(fp)
```

## Per parameter volcano plots

For each CMR parameter, the raw p-value volcano is shown alongside the FDR volcano.

```{r per-var-volcano, echo = F, results = "asis", out.width = "48%"}
for (v in vars_nebula) {
  ttl <- if (v %in% names(im_labels)) im_labels[[v]] else v
  cat("\n\n### ", ttl, "\n\n")
  paths <- c(
    get_s3_fig(paste0("Projects/EC CMR/Results/Figures/", v, ".png")),
    get_s3_fig(paste0("Projects/EC CMR/Results/Figures/", v, "_fdr.png"))
  )
  paths <- paths[!sapply(paths, is.null)]
  if (length(paths)) {
    cat(paste0("![](", paths, ")", collapse = " "))
    cat("\n\n")
  } else {
    cat("_No saved figure found._\n\n")
  }
}
```

# Gene set enrichment (Hallmark)

Hallmark GSEA results per CMR parameter.

```{r gsea-read, echo = F}
gsea_results <- s3read_using_region(FUN = read.csv,
                                    bucket = "scrna",
                                    object = "Projects/EC CMR/Results/cmr_ec_hallmark_compiled.csv",
                                    region = "")
```

## Significant pathways

Hallmark pathways with padj < 0.05.

```{r gsea-sig, echo = F}
gsea_sig <- gsea_results %>%
  filter(padj < 0.05) %>%
  mutate(label = ifelse(var %in% names(im_labels), im_labels[var], var),
         pathway = str_to_title(str_replace_all(str_remove(pathway, "^HALLMARK_"), "_", " "))) %>%
  arrange(var, padj) %>%
  select(Parameter = label, Pathway = pathway, NES, pval, padj, size)

if (nrow(gsea_sig)) {
  gsea_sig %>% kable(digits = 3) %>% kable_styling(full_width = FALSE)
} else {
  cat("No pathways reached padj < 0.05.")
}
```

## NES heatmap

```{r gsea-heat, echo = F, out.width = "100%"}
fp <- get_s3_fig("Projects/EC CMR/Results/Figures/gsea/ec_cmr_nes_heatmap.png")
if (!is.null(fp)) knitr::include_graphics(fp)
```

## FDR significant genes per parameter

```{r gsea-bar, echo = F, out.width = "80%"}
fp <- get_s3_fig("Projects/EC CMR/Results/Figures/gsea/ec_cmr_siggene_bar.png")
if (!is.null(fp)) knitr::include_graphics(fp)
```

## Per parameter pathway lollipops

Lollipop plots were saved only for parameters with significant Hallmark pathways; others are skipped.

```{r gsea-lollipop, echo = F, results = "asis", out.width = "80%"}
for (v in vars_nebula) {
  fp <- get_s3_fig(paste0("Projects/EC CMR/Results/Figures/gsea/hallmark_", v, ".png"))
  if (!is.null(fp)) {
    ttl <- if (v %in% names(im_labels)) im_labels[[v]] else v
    cat("\n\n### ", ttl, "\n\n")
    cat(paste0("![](", fp, ")"))
    cat("\n\n")
  }
}
```
