---
title: "PANTHER - IVGTT mixed models"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r libraries, echo=F, include = F}
library(SomaDataIO)
library(tidyverse)
library(growthcleanr)
library(arsenal)
library(dplyr)
library(naniar)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(pedbp)
library(psych)
library(EnhancedVolcano)
library(gt)
library(SomaPlotr)
library(readxl)
library(limma)
library(glue)
library(purrr)
library(stringr)
library(nlme)
library(emmeans)

options(scipen=999)

user <- Sys.info()[["user"]]
if (user == "laurapyle") {
  home_dir <- "/Users/laurapyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Data Harmonization/Data Clean"
  r_dir <- '/Users/laurapyle/Documents/GitHub/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions'
  panther_dir <- "/Users/laurapyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/PANTHER/Data_Cleaned"
} else if (user == "lpyle") {
  home_dir <- '/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Data Harmonization/Data Clean'
  r_dir <- '/Users/lpyle/Documents/GitHub/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions'
  panther_dir <- "/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/PANTHER/Data_Cleaned"
} else if (user == "pylell") {
  home_dir <- '/Users/pylell/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Data Harmonization/Data Clean'
  r_dir <- '/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions'
  panther_dir <- "/Users/pylell/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/PANTHER/Data_Cleaned"
} else {
  home_dir <- "/Users/tim/Library/CloudStorage/OneDrive-UW/UWMDI/Andrea Steck/Advanced CGM and OGTT"
  r_dir <- '/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions'
}
```

```{r load-data, echo = F, include = F}
# Load dictionary function and file
filename <- "label_harmonized_function.R"
full_path <- file.path(r_dir, filename)
source(full_path)
filename <- "correlation_function.R"
full_path <- file.path(r_dir, filename)
source(full_path)
```

```{r prep-data, echo = F, include = F}
# Load clinical data
filename <- "soma_harmonized_dataset.csv"
full_path <- file.path(home_dir, filename)
harm_dat <- read.csv(full_path)

filename <- "panther_withdrew_ltfu_list.csv"
full_path <- file.path(panther_dir, filename)
exclude <- read.csv(full_path)

dat <- harm_dat %>% filter(study == "PANTHER" & procedure == "ivgtt" & visit %in% c("baseline", "screening")) %>%
  mutate(visit = case_when(visit == "screening" ~ "baseline", T ~ visit),
         mrn = case_when(study != "PANTHER" ~ paste0(as.character(mrn), as.character(screen_date)), T~ as.character(mrn))) %>%
  filter(record_id %nin% exclude) 

# PANTHER protocol does not include 0 timepoint - will not include any data prior to this timepoint

# keep only the variables we need
matching_cols_cpep <- names(dat)[startsWith(names(dat), "cpep_")]
matching_cols_cpep <- matching_cols_cpep[1:29]
matching_cols_ins <- names(dat)[startsWith(names(dat), "insulin")]
matching_cols_ins <- matching_cols_ins[1:33]
matching_cols_ins <- matching_cols_ins[-1]
matching_cols_ins <- matching_cols_ins[-12]
matching_cols_ins <- matching_cols_ins[-20]
matching_cols_ins <- matching_cols_ins[-28]
matching_cols_glu <- names(dat)[startsWith(names(dat), "gluc_")]
matching_cols_glu <- matching_cols_glu[1:29]
dat_keep <- dat %>% select(mrn, visit, record_id, matching_cols_cpep, matching_cols_ins, matching_cols_glu)

# need to compare by risk groups and Tanner stage so need to pull from harmonized data
temp <- harm_dat %>% filter(study == "PANTHER" & visit %in% c("baseline", "screening")) %>%
   mutate(visit = case_when(visit == "screening" ~ "baseline", T ~ visit),
          mrn = case_when(study != "PANTHER" ~ paste0(as.character(mrn), as.character(screen_date)), T~ as.character(mrn))) %>%
   filter(record_id %nin% exclude) 
preds <- temp %>% 
   dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                    across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                    .by = c(mrn, visit)) %>%
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" &
                                                ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                              race == "Black or African American" &
                                                ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                              ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                              T ~ "Not Hispanic (or Latino Other"),
         tanner_stage_comp = coalesce(tan_fgd, tan_fph, tan_tveq, tan_mgd, tan_mph, breast_tanner),
         tanner_stage_comp_panther = case_when(tanner_stage_comp > 3 ~ 4, T~ tanner_stage_comp),
         tanner_stage_cat = case_when(tanner_stage_comp <3 ~ "I_II",
                                      tanner_stage_comp >=3 ~ "III_V"),
         tanner_stage_cat_2 = case_when(tanner_stage_comp == 1 ~ "I",
                                        tanner_stage_comp <4 ~ "II_III",
                                        tanner_stage_comp >=4 ~ "IV_V"),
         tanner_stage_cat_3 = case_when(tanner_stage_comp < 2 ~ "I_II",
                                        tanner_stage_comp == 3 ~ "III",
                                        tanner_stage_comp >= 4 ~ "IV_V"),
         tanner_stage_comp_panther_cat =
                  case_when(tanner_stage_comp_panther == 4 ~ "4 & 5",
                            T ~ as.character(tanner_stage_comp_panther)),
         age_mo = (age * 12),
         sex = case_when(sex == "Male" ~ "male",
                         sex == "Female" ~ "female"),
         male_ind = case_when(sex == "male" ~ 1, sex == "female" ~ 0),
         combined_tkv = coalesce(total_kidney_volume_ml, total_kidney_volume_ml_manual),
         combined_ht_tkv = coalesce(ht_adj_tkv, ht_adj_tkv_manual),
         log_tot_test = log(tot_test),
         log_free_test = log(free_test),
         log_lh = log(lh),
         log_estrad = log(estrad),
         log_acr_u = log(acr_u),
         avg_c_adc_tkv = avg_c_adc/combined_tkv,
         avg_c_r2_tkv = avg_c_r2/combined_tkv,
         avg_pcascl_tkv = avg_pcascl/combined_tkv,
         avg_c_t1_tkv = avg_c_t1/combined_tkv,
         mm_airg = case_when(mm_airg > 0 ~ mm_airg),
         mm_bcell = case_when(mm_bcell > 0 ~ mm_bcell),
         mm_di = case_when(mm_di > 0 & mm_di < 10000 ~ mm_di),
         mm_ir = case_when(mm_ir > 0 ~ mm_ir),
         mm_si = case_when(mm_si > 0 & mm_si < 500 ~ mm_si),
         birthweight = gsub(" kg", "", birthweight),
         birthweight = as.numeric(birthweight),
         glp1_agonist_timepoint = case_when(is.na(glp1_agonist_timepoint) ~ "No", T ~ glp1_agonist_timepoint),
         uacr_cat = case_when(acr_u <= 10 ~ "UACR [0, 10]",
                              acr_u <= 20 ~ "UACR (10, 20]",
                              acr_u <= 30 ~ "UACR (20, 30]",
                              acr_u > 30 ~ "UACR > 30")) %>%
  arrange(record_id) %>%
  # filter(!is.na(dexa_body_fat)) %>% # filter based on DEXA (to filter those who may just be screen fails)
  filter(!is.na(male_ind)) %>%
  filter(record_id %nin% c("PAN_205_O", "PAN_209_O", "PAN_210_O")) # awaiting study visit
preds$tanner_stage_comp <- as.factor(preds$tanner_stage_comp)
bp_percentile = p_bp(q_sbp = preds$sbp, q_dbp = preds$dbp, age = preds$age_mo, male = preds$male_ind)
preds$sbp_p <- bp_percentile$sbp_p
preds$dbp_p <- bp_percentile$dbp_p
preds$bmip <- NULL
preds$bmiz <- NULL
bmi_percentile = ext_bmiz(data = subset(preds,
                                        select = c("record_id", "sex", "age_mo", "weight", "height", "bmi")),
                          age = "age_mo",
                          wt = "weight",
                          ht = "height",
                          bmi = "bmi",
                          adjust.integer.age = F) %>%
  dplyr:: select(record_id, bmip, bmiz) %>%
  filter(!is.na(bmip))
preds <- left_join(preds, bmi_percentile, by = "record_id") 
preds <- preds %>% dplyr::mutate(group_risk = case_when(bmip >= 95 | hba1c >=6 | group == "Type 2 Diabetes" ~ "High",
                                       T ~ "Low"),
                sex_group_risk = case_when(sex == "male" & group_risk == "High" ~ "M, High",
                                           sex == "male" & group_risk == "Low" ~ "M, Low",
                                           sex == "female" & group_risk == "High" ~ "F, High",
                                           sex == "female" & group_risk == "Low" ~ "F, Low"))
preds_keep <- preds[,c("record_id", "visit", "tanner_stage_comp", "sex", "sex_group_risk", "group_risk")]

final_dat <- left_join(dat_keep, preds_keep, by = c("record_id", "visit"))
final_dat$tanner_combined <- as.factor(ifelse(final_dat$tanner_stage_comp %in% c(4,5), "4/5", final_dat$tanner_stage_comp))

# reshape
ins_temp <- final_dat[,c("mrn", "group_risk", "tanner_combined", matching_cols_ins)]
ins_temp <- ins_temp %>% filter(!is.na(tanner_combined))
ins_temp <- unique(ins_temp)
insulin_long <- reshape(ins_temp, direction = "long", varying = matching_cols_ins, 
                          v.names = "insulin",times = c(2,3,4,5,6,8,10,12,14,16,19,22,23,24,25,27,30,35,40,50,60,70,80,90,100,120,140,160,180),
                          idvar = c("mrn","group_risk","tanner_combined"))
insulin_long <- insulin_long %>% arrange(mrn,time)

glu_temp <- final_dat[,c("mrn", "group_risk", "tanner_combined", matching_cols_glu)]
glu_temp <- glu_temp %>% filter(!is.na(tanner_combined))
glu_temp <- unique(glu_temp)
glucose_long <- reshape(glu_temp, direction = "long", varying = matching_cols_glu, 
                          v.names = "glucose",times = c(2,3,4,5,6,8,10,12,14,16,19,22,23,24,25,27,30,35,40,50,60,70,80,90,100,120,140,160,180),
                          idvar = c("mrn","group_risk","tanner_combined"))
glucose_long <- glucose_long %>% arrange(mrn,time)

cpep_temp <- final_dat[,c("mrn", "group_risk", "tanner_combined", matching_cols_cpep)]
cpep_temp <- cpep_temp %>% filter(!is.na(tanner_combined))
cpep_temp <- unique(cpep_temp)
cpep_long <- reshape(cpep_temp, direction = "long", varying = matching_cols_cpep, 
                          v.names = "Cpeptide",times = c(2,3,4,5,6,8,10,12,14,16,19,22,23,24,25,27,30,35,40,50,60,70,80,90,100,120,140,160,180),
                          idvar = c("mrn","group_risk","tanner_combined"))
cpep_long <- cpep_long %>% arrange(mrn,time)
```


```{r echo=FALSE, include=FALSE}
mixed <- function(data, outcome, grouping_var){
  data <- as.data.frame(data)
  data$time <- as.factor(data$time)
  form = as.formula(paste0(outcome,"~ ", grouping_var, "*time"))
  mod <- lme(as.formula(form),random=~1|mrn,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,c(grouping_var,"time"))
  prs <-  pairs(m,adjust="tukey",by="time")
  return(list(anova,m,prs))
}

mixed_adj <- function(data, outcome, grouping_var){
  data <- as.data.frame(data)
  data$time <- as.factor(data$time)
  form = as.formula(paste0(outcome,"~ ", grouping_var, "*time + group_risk"))
  mod <- lme(as.formula(form),random=~1|mrn,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,c(grouping_var,"time"))
  prs <-  pairs(m,adjust="tukey",by="time")
  return(list(anova,m,prs))
}

plot_mixed_group_risk <- function(mixed_output, data, ylab) {
  emmeans_result <- mixed_output[[2]]
  emmeans_result <- as.data.frame(emmeans_result)
  emmeans_result$group <- ifelse(emmeans_result$group_risk == "High", "High", "Low")
  # Create line plot
  p <-  ggplot(emmeans_result, aes(x = time, y = emmean, color = group, group = group, 
                                  line = group)) +
    geom_line() +
    geom_point() +
    labs(x = "Time", y = ylab, color = "group") +
    theme_bw() + 
    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), linetype = 1, alpha = 0.1) +
    theme(legend.title=element_blank()) + geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL))
}

plot_mixed_tanner <- function(mixed_output, data, ylab) {
  emmeans_result <- mixed_output[[2]]
  emmeans_result <- as.data.frame(emmeans_result)
  emmeans_result$group <- emmeans_result$tanner_combined
  # Create line plot
  p <-  ggplot(emmeans_result, aes(x = time, y = emmean, color = group, group = group, 
                                  line = group)) +
    geom_line() +
    geom_point() +
    labs(x = "Time", y = ylab, color = "group") +
    theme_bw() + 
    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), linetype = 1, alpha = 0.1) +
    theme(legend.title=element_blank()) + geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL))
}

```

# Questions/comments

1) Original IVGTT code read in two data files - the harmonized data and a lab data csv file. I checked MRN and visit dates and it seemed like all the data in the csv were contained in the harmonized data, so I didn't merge it in.
2) PANTHER protocol says that first time point for IVGTT is 2 minutes. There are data in the insulin_minus_15 and gluc_minus_15 variables, but I did not include them in analysis because there were no corresponding data for C-peptide.

# Results

## Insulin

### By risk group

```{r echo=FALSE}
mod <- mixed(data = insulin_long, outcome = "insulin", grouping_var = "group_risk")
```

```{r echo = FALSE, fig.width=8}
p <- plot_mixed_group_risk(mixed_output = mod, data = insulin_long, ylab = "Insulin")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

### By Tanner stage

```{r echo=FALSE}
mod <- mixed(data = insulin_long, outcome = "insulin", grouping_var = "tanner_combined")
```

```{r echo = FALSE, fig.width=8}
p <- plot_mixed_tanner(mixed_output = mod, data = insulin_long, ylab = "Insulin")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

### By Tanner stage - adjusted for risk group

```{r echo=FALSE}
mod <- mixed_adj(data = insulin_long, outcome = "insulin", grouping_var = "tanner_combined")
```

```{r echo = FALSE, fig.width=8}
p <- plot_mixed_tanner(mixed_output = mod, data = insulin_long, ylab = "Insulin")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

## Glucose

### By risk group

```{r echo=FALSE}
mod <- mixed(data = glucose_long, outcome = "glucose", grouping_var = "group_risk")
```

```{r echo=FALSE, fig.width = 8}
p <- plot_mixed_group_risk(mixed_output = mod, data = glucose_long, ylab = "Glucose")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

### By Tanner stage

```{r echo=FALSE}
mod <- mixed(data = glucose_long, outcome = "glucose", grouping_var = "tanner_combined")
```

```{r echo = FALSE, fig.width=8}
p <- plot_mixed_tanner(mixed_output = mod, data = glucose_long, ylab = "Glucose")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

### By Tanner stage - adjusted for risk group

```{r echo=FALSE}
mod <- mixed_adj(data = glucose_long, outcome = "glucose", grouping_var = "tanner_combined")
```

```{r echo = FALSE, fig.width=8}
p <- plot_mixed_tanner(mixed_output = mod, data = glucose_long, ylab = "Glucose")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

## C-peptide

### By risk group

```{r echo=FALSE}
mod <- mixed(data = cpep_long, outcome = "Cpeptide", grouping_var = "group_risk")
```

```{r echo=FALSE, fig.width = 8}
p <- plot_mixed_group_risk(mixed_output = mod, data = cpep_long, ylab = "C-peptide")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

### By Tanner stage

```{r echo=FALSE}
mod <- mixed(data = cpep_long, outcome = "Cpeptide", grouping_var = "tanner_combined")
```

```{r echo = FALSE, fig.width=8}
p <- plot_mixed_tanner(mixed_output = mod, data = cpep_long, ylab = "C-peptide")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```

### By Tanner stage - adjusted for risk group

```{r echo=FALSE}
mod <- mixed_adj(data = cpep_long, outcome = "Cpeptide", grouping_var = "tanner_combined")
```

```{r echo = FALSE, fig.width=8}
p <- plot_mixed_tanner(mixed_output = mod, data = cpep_long, ylab = "C-peptide")
p
```

#### ANOVA table

```{r echo=FALSE}
kable(mod[[1]])
```

#### Estimated marginal means

```{r echo=FALSE}
kable(mod[[2]])
```

#### Pairwise comparisons

```{r echo=FALSE}
kable(mod[[3]])
```
