---
title: "PANTHER Baseline Manuscript"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format: 
  docx:
    fig-cap-location: top
    tbl-cap-location: top
    title-block-banner: false
    title-block-style: none
execute:
  echo: false
  warning: false
  message: false
---

Data for: Pubertal Effects on Kidney Health Across the Spectrum of Metabolic Risk: Results from the PANTHER Study

```{r libraries, echo=F, include = F}
library(SomaDataIO)
library(tidyverse)
library(growthcleanr)
library(arsenal)
library(dplyr)
library(naniar)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(pedbp)
library(psych)
library(EnhancedVolcano)
library(gt)
library(SomaPlotr)
library(readxl)
library(limma)
library(glue)
library(purrr)
library(aws.s3)
library(jsonlite)
library(gtsummary)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/pylell/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r load-data, echo = F, include = F}
# Load SOMAScan analytes data
load(file.path(root_path, "Data Harmonization/Combined SomaScan/analytes.Rdata"))
# write.csv(analytes, file.path(root_path, "Data Harmonization/Combined SomaScan/analytes.csv"), row.names = F)

# Load dictionary function and file
source("/Users/choiyej/GitHub/YC_CHCO/R Functions/label_harmonized_function.R")
source("/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad/Resources/YC/R Functions/correlation_function.R")
```

```{r}
# IGF-Z score calculation
# 1) Reference table from your screenshots (children)
igf_ref_child <- tribble(
  ~sex,     ~tanner_stage_comp_panther, ~age_years, ~lower_ref, ~upper_ref, ~mean_ref,
  "male",   1,           3,          20,         141,        63,
  "male",   1,           4,          25,         157,        73,
  "male",   1,           5,          30,         174,        84,
  "male",   1,           6,          37,         192,        96,
  "male",   1,           7,          44,         211,        109,
  "male",   1,           8,          52,         231,        123,
  "male",   1,           9,          61,         252,        137,
  "male",   1,           10,         71,         275,        153,
  "male",   1,           11,         82,         299,        170,
  "male",   1,           12,         93,         324,        188,
  "male",   1,           13,         106,        350,        207,
  "male",   1,           14,         120,        377,        227,
  "male",   1,           15,         127,        391,        237,
  
  "male",   2,           8,          39,         264,        121,
  "male",   2,           9,          52,         304,        146,
  "male",   2,           10,         67,         347,        174,
  "male",   2,           11,         86,         393,        206,
  "male",   2,           12,         106,        443,        240,
  "male",   2,           13,         130,        497,        277,
  "male",   2,           14,         156,        554,        318,
  "male",   2,           15,         185,        616,        362,
  "male",   2,           16,         201,        648,        386,
  
  "male",   3,           8,          39,         264,        121,
  "male",   3,           9,          52,         304,        146,
  "male",   3,           10,         67,         347,        174,
  "male",   3,           11,         86,         393,        206,
  "male",   3,           12,         106,        443,        240,
  "male",   3,           13,         130,        497,        277,
  "male",   3,           14,         156,        554,        318,
  "male",   3,           15,         185,        616,        362,
  "male",   3,           16,         201,        648,        386,
  
  "male",   4,           11,         277,        673,        449,
  "male",   4,           12,         265,        652,        433,
  "male",   4,           13,         241,        612,        401,
  "male",   4,           14,         220,        574,        372,
  "male",   4,           15,         209,        537,        343,
  "male",   4,           16,         180,        501,        316,
  "male",   4,           17,         161,        467,        290,
  "male",   4,           18,         144,        434,        266,
  
  "female", 1,           3,          36,         162,        77,
  "female", 1,           4,          26,         179,        88,
  "female", 1,           5,          39,         198,        100,
  "female", 1,           6,          47,         217,        113,
  "female", 1,           7,          55,         238,        127,
  "female", 1,           8,          64,         259,        142,
  "female", 1,           9,          74,         282,        158,
  "female", 1,           10,         85,         306,        175,
  "female", 1,           11,         97,         332,        194,
  "female", 1,           12,         110,        358,        213,
  
  "female", 2,           8,          89,         369,        181,
  "female", 2,           9,          96,         399,        196,
  "female", 2,           10,         104,        431,        212,
  "female", 2,           11,         112,        466,        229,
  "female", 2,           12,         121,        504,        247,
  "female", 2,           13,         131,        545,        267,
  "female", 2,           14,         136,        566,        278,
  
  "female", 3,           9,          192,        568,        350,
  "female", 3,           10,         192,        568,        350,
  "female", 3,           11,         192,        568,        350,
  "female", 3,           12,         192,        568,        350,
  "female", 3,           13,         192,        568,        350,
  "female", 3,           14,         192,        568,        350,
  "female", 3,           15,         192,        568,        350,
  
  "female", 4,           10,         279,        664,        446,
  "female", 4,           11,         268,        646,        433,
  "female", 4,           12,         248,        612,        406,
  "female", 4,           13,         229,        579,        380,
  "female", 4,           14,         211,        547,        356,
  "female", 4,           15,         194,        516,        332,
  "female", 4,           16,         177,        487,        309,
  "female", 4,           17,         162,        458,        287,
  "female", 4,           18,         147,        430,        266
)
```

```{r prep-data, echo = F, include = F, eval = F}
# Load clinical data
harm_dat <- read.csv(file.path(root_path, "Data Harmonization/Data Clean/soma_harmonized_dataset.csv"), na.strings = "")
exclude <- read.csv(file.path(root_path, "PANTHER/Data_Cleaned/panther_withdrew_ltfu_list.csv"))$record_id
dict <- read.csv(file.path(root_path, "Data Harmonization/data_dictionary_master.csv"))

dat <- harm_dat %>% filter((study == "PANTHER" & visit %in% c("baseline", "screening"))|study == "RENAL-HEIR"|
                             study == "RENAL-HEIRitage" | (study == "IMPROVE" & visit == "baseline")) %>%
  mutate(visit = case_when(visit == "screening" ~ "baseline", T ~ visit),
         mrn = case_when(study != "PANTHER" ~ paste0(as.character(mrn), as.character(screen_date)), T~ as.character(mrn))) %>%
  filter(record_id %nin% exclude) %>% # Not removing anyone with proteomics data
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(mrn, visit)) %>%
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" & 
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                                     race == "Black or African American" & 
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                                     ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                                     T ~ "Not Hispanic or Latino Other"),
                tanner_stage_comp = coalesce(tan_fgd, tan_fph, tan_tveq, tan_mgd, tan_mph, breast_tanner),
                tanner_stage_comp_panther = case_when(tanner_stage_comp > 3 ~ 4, T~ tanner_stage_comp),
                tanner_stage_cat = case_when(tanner_stage_comp <3 ~ "I_II",
                                             tanner_stage_comp >=3 ~ "III_V"),
                tanner_stage_cat_2 = case_when(tanner_stage_comp == 1 ~ "I",
                                               tanner_stage_comp <4 ~ "II_III",
                                               tanner_stage_comp >=4 ~ "IV_V"),
                tanner_stage_cat_3 = case_when(tanner_stage_comp < 2 ~ "I_II",
                                               tanner_stage_comp == 3 ~ "III",
                                               tanner_stage_comp >= 4 ~ "IV_V"),
                tanner_stage_comp_panther_cat = 
                  case_when(tanner_stage_comp_panther == 4 ~ "4 & 5",
                            T ~ as.character(tanner_stage_comp_panther)),
                age_mo = (age * 12),
                sex = case_when(sex == "Male" ~ "male",
                                sex == "Female" ~ "female"),
                male_ind = case_when(sex == "male" ~ 1, sex == "female" ~ 0),
                combined_tkv = coalesce(total_kidney_volume_ml, total_kidney_volume_ml_manual),
                combined_ht_tkv = coalesce(ht_adj_tkv, ht_adj_tkv_manual),
                log_tot_test = log(tot_test),
                log_free_test = log(free_test),
                log_lh = log(lh),
                log_estrad = log(estrad),
                log_acr_u = log(acr_u),
                avg_c_adc_tkv = avg_c_adc/combined_tkv,
                avg_c_r2_tkv = avg_c_r2/combined_tkv,
                avg_pcascl_tkv = avg_pcascl/combined_tkv,
                avg_c_t1_tkv = avg_c_t1/combined_tkv,
                mm_airg = case_when(mm_airg > 0 ~ mm_airg),
                mm_bcell = case_when(mm_bcell > 0 ~ mm_bcell),
                mm_di = case_when(mm_di > 0 & mm_di < 10000 ~ mm_di),
                mm_ir = case_when(mm_ir > 0 ~ mm_ir),
                mm_si = case_when(mm_si > 0 & mm_si < 500 ~ mm_si),
                birthweight = gsub(" kg", "", birthweight),
                birthweight = as.numeric(birthweight),
                glp1_agonist_timepoint = case_when(is.na(glp1_agonist_timepoint) ~ "No", T ~ glp1_agonist_timepoint),
                uacr_cat = case_when(acr_u <= 10 ~ "UACR [0, 10]",
                                     acr_u <= 20 ~ "UACR (10, 20]",
                                     acr_u <= 30 ~ "UACR (20, 30]",
                                     acr_u > 30 ~ "UACR > 30"),
                tanner_stage_comp = as.factor(tanner_stage_comp)) %>%
  arrange(record_id) %>%
  filter(!is.na(male_ind)) %>%
  mutate(
    age_years = floor(age)
  ) %>%
  left_join(igf_ref_child, by = c("sex", "tanner_stage_comp_panther", "age_years")) %>%
  mutate(sd_ref = case_when(igf_1 < mean_ref ~ (mean_ref - lower_ref) / 2,
                            igf_1 >= mean_ref ~ (upper_ref - mean_ref) / 2), 
         igf1_z_score_calc = (igf_1 - mean_ref) / sd_ref
  ) %>%
  select(-ends_with("_adult"))


dat$uacr_cat <- factor(dat$uacr_cat, levels = c("UACR [0, 10]",
                                                "UACR (10, 20]",
                                                "UACR (20, 30]",
                                                "UACR > 30"))

bp_percentile = p_bp(q_sbp = dat$sbp, q_dbp = dat$dbp, age = dat$age_mo, male = dat$male_ind)
dat$sbp_p <- bp_percentile$sbp_p
dat$dbp_p <- bp_percentile$dbp_p

bmi_percentile = ext_bmiz(data = subset(dat, 
                                        select = c("record_id", "sex", "age_mo", "weight", "height", "bmi")), 
                          age = "age_mo", 
                          wt = "weight", 
                          ht = "height", 
                          bmi = "bmi", 
                          adjust.integer.age = F) %>% 
  dplyr:: select(record_id, bmip, bmiz) %>%
  filter(!is.na(bmip))

dat <- dat %>%
  dplyr::mutate(group_risk = case_when(bmip >= 95 | hba1c >= 6 | group == "Type 2 Diabetes" ~ "High",
                                       bmip < 85 & hba1c <= 5.6 ~ "Low"),
                group_2 = case_when(group == "Lean Control" & group_risk == "High" ~ "Obesity/Prediabetes",
                                    group == "Lean Control" & group_risk == "Low" ~ "Normal weight & Normoglycemic",
                                    group == "Obese Control" ~ "Obesity/Prediabetes",
                                    T ~ group),
                sex_group_risk = case_when(sex == "male" & group_risk == "High" ~ "M, High",
                                           sex == "male" & group_risk == "Low" ~ "M, Low",
                                           sex == "female" & group_risk == "High" ~ "F, High",
                                           sex == "female" & group_risk == "Low" ~ "F, Low"))

panther <- dat %>% filter(study == "PANTHER")
sens_dat <- dat %>% filter(study == "PANTHER" | study == "RENAL-HEIR") %>%
  filter(!is.na(tanner_stage_comp))

# Prepare SOMA data
# prepare data set for analysis
cleanData_comb <- dat |>
  mutate_at(getAnalytes(dat), ~ as.numeric(scale(log(.))))

cleanData_panther <- cleanData_comb %>%
  filter(study == "PANTHER") %>%
  filter(!is.na(tanner_stage_comp))
cleanData_sens <- cleanData_comb %>%
  filter(study == "PANTHER" | study == "RENAL-HEIR") %>%
  filter(!is.na(tanner_stage_comp))

analytes$Target_nodup <- paste0(analytes$Target, " (", analytes$AptName, ")")

# Load and modify dictionary file
dict <- label_harmonized_dict(dat, dict) %>%
  dplyr::mutate(age = "Age",
                tanner_stage_comp = "Tanner stage",
                tanner_stage_comp_panther = "Tanner stage",
                tanner_stage_comp_ord = "Tanner stage",
                tanner_stage_cat = "Tanner stage categorized (1/2, 3/4/5)",
                tanner_stage_cat2 = "Tanner stage categorized (1, 2/3, 4/5)",
                mm_airg = "AIRg",
                mm_bcell = "Beta cell function",
                mm_di = "DI",
                mm_ir = "Insulin resistance",
                mm_si = "Insulin sensitivity",
                bmip = "BMI percentile",
                bmiz = "BMI Z score",
                total_kidney_volume_ml = "Total kidney volume (TKV)",
                ht_adj_tkv = "Height-adjusted TKV",
                total_kidney_volume_ml_manual = "TKV (mL; manual)",
                ht_adj_tkv_manual = "Height Adjusted TKV (mL/m; manual)",
                combined_tkv = "Collapsed Total Kidney Volume (mL)", 
                combined_ht_tkv = "Collapsed Height Adjusted Total Kidney Volume (mL/m)",
                fsh = "FSH", 
                lh = "LH", 
                estrad = "Estradiol",
                tot_test = "Total testosterone",
                free_test = "Free testosterone",
                igf_1 = "IGF-1",
                igf1_z_score = "IGF-1 Z-score",
                igf1_z_score_calc = "IGF-1 Z-score",
                avg_c_t1 = "Avg cortical T1",
                avg_pcascl = "Avg pCASCL",
                avg_c_adc_tkv = "Avg cortical ADC/TKV",
                avg_c_r2_tkv = "Avg cortical R2*/TKV",
                avg_pcascl_tkv = "Avg cortical ASL/TKV",
                avg_c_t1_tkv = "Avg cortical T1/TKV", 
                sbp_p = "SBP percentile",
                dbp_p = "DBP percentile",
                birthweight = "Birthweight",
                bl_dheas = "DHEA-S",
                log_acr_u = "ln(UACR)",
                log_tot_test = "ln(Total testosterone)",
                log_free_test = "ln(Free testosterone)",
                log_lh = "ln(LH)",
                log_estrad = "ln(Estradiol)",
                race_ethnicity_condensed = "Race/Ethnicity", 
                eGFR_CKiD_U25_CystatinC = "eGFR CKiD U25 CystC",
                eGFR_CKiD_U25_Creat ="eGFR CKiD U25 Creat",
                eGFR_CKiD_U25_avg ="eGFR CKiD U25 CystC & Creat",
                bl_dheas = "DHEA-S",
                gfr_raw_plasma = "mGFR",
                gfr_bsa_plasma = "mGFR (BSA)",
                erpf_raw_plasma = "ERPF",
                erpf_bsa_plasma = "ERPF (BSA)",
                eGFR_fas_cr = "eGFR - EKFC",
                glomerular_pressure = "Glomerular pressure",
                dexa_bone_mineral_density = "BMD",
                mm_bcell = "Insulin secretion capacity",
                u24_mab = "Urine AER 24 hours",
                group_2 = "Group",
                rbf = "Renal blood flow")%>%
  dplyr::select(-starts_with("seq."))

analytes_dict <- analytes %>%
  dplyr::mutate(target_mod = Target_nodup) %>%
  column_to_rownames("AptName") %>%
  dplyr::select(target_mod) %>%
  t()

dict <- cbind(dict[1,], analytes_dict)

# Label data
dat <- label_harmonized(dat, dict)
panther <- label_harmonized(panther, dict)
sens_dat <- label_harmonized(sens_dat, dict)
```

```{r save-data, include = F, eval = F}
save(analytes,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther_analytes_022326.RData"))
save(dict,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther_dict_022326.RData"))
save(dat,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther+rh+rh2+imp_data_022326.RData"))
save(panther,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther_data_022326.RData"))
save(sens_dat,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther_rh_rh2_imp_data_022326.RData"))
save(cleanData_comb,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther+rh+rh2+imp_data_soma_std_022326.RData"))
save(cleanData_panther,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther_data_soma_std_022326.RData"))
save(cleanData_sens,
     file = file.path(root_path, "PANTHER/Data_Cleaned/panther_data_rh_rh2_imp_std_022326.RData"))
```


```{r, include = F, eval = T}
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther_analytes_022326.RData"))
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther_dict_022326.RData"))
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther+rh+rh2+imp_data_022326.RData"))
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther_data_022326.RData"))
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther_rh_rh2_imp_data_022326.RData"))
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther+rh+rh2+imp_data_soma_std_022326.RData"))
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther_data_soma_std_022326.RData"))
load(file = file.path(root_path, "PANTHER/Data_Cleaned/panther_data_rh_rh2_imp_std_022326.RData"))
```

```{r echo = F}
# missing avg_c_t1
# missing_t1 <- harm_dat %>%
#   filter(study == "PANTHER") %>%
#   filter(!is.na(visit)) %>%
#   distinct(record_id, visit, .keep_all = T) %>%
#   select(record_id, visit, avg_c_t1, avg_c_r2)
# write.csv(missing_t1, file.path(root_path, "PANTHER/Data_Cleaned/panther_T1_ids.csv", row.names = F, na = ""))
```

```{r include = F}
table_controls <- tableby.control(
  total = T,
  numeric.stats = c("meansd"),
  cat.stats = c("countpct"),
  cat.test = "chisq",
  digits = 1, 
  numeric.simplify = TRUE,
  test.always = T,
  cat.droplevels = F
)

```

# Descriptive tables

## By risk

```{r include = F}
# # identify UACR outliers/high UACRs
# panther_uacr <- subset(panther, 
#                        select = c(record_id, mrn, group_2, aer_24, u24_mab, u24_hrs, 
#                                   acr_u, creatinine_u, microalbumin_u)) %>%
#   arrange(desc(u24_mab))
# 
# panther_uacr$flag <- ifelse(panther_uacr$u24_mab > 20 & panther_uacr$acr_u > 30, "flag", "")
# 
# panther_uacr_2 <- harm_dat %>% filter(study == "PANTHER") %>%
#   mutate(mrn = case_when(study != "PANTHER" ~ paste0(as.character(mrn), as.character(screen_date)), T~ as.character(mrn))) %>%
#   filter(record_id %nin% exclude) %>% # Not removing anyone with proteomics data
#   dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
#                    across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
#                    .by = c(mrn, visit)) %>%
#   dplyr::select(record_id, visit, group, u24_mab, acr_u, creatinine_u, microalbumin_u, creatinine_s)
# 
# panther_uacr_wide <- panther_uacr_2 %>%
#   pivot_wider(
#     id_cols = c(record_id, group),  # person ID and group
#     names_from = 2,     # visit variable (baseline, year_1, year_2)
#     values_from = 4:8   # the numeric columns
#   ) %>%
#   dplyr::select(where(~ !all(is.na(.x))))


```

```{r desc-risk, results = 'asis', echo = F}
# Table 1
kable(summary(tableby(group_risk ~ age + sex + group_2 + race_ethnicity_condensed + height + weight + 
                        bmi +  bmiz + bmip + sbp + dbp +  hba1c + 
                        eGFR_fas_cr + eGFR_fas_cr_cysc + eGFR_CKiD_U25_avg + 
                        kwt(acr_u, "medianq1q3") + fe(uacr_cat) + kwt(u24_mab, "medianq1q3") + 
                        gotast_base + gptalt_base +  
                        fe(metformin_timepoint) + fe(insulin_med_timepoint) + fe(glp1_agonist_timepoint) + fe(sglti_timepoint), 
                      data = panther, 
                      # control = table_controls,
                      test = T), digits = 1))

# chisq.test(table(panther$group_risk, panther$glp1_agonist_timepoint))
# chisq.test(table(panther$group_risk))
# gtsummary::tbl_summary(panther, by = group_risk,
#                        include = c(metformin_timepoint, insulin_med_timepoint, glp1_agonist_timepoint, sglti_timepoint)) %>%
#   add_p(test = everything() ~ "chisq.test.no.correct")

# people missing cystatin C and not creatinine
subset(panther, is.na(panther$cystatin_c) & !is.na(panther$creatinine_s))
subset(panther, is.na(eGFR_fas_cr_cysc))$cystatin_c
subset(panther, is.na(eGFR_fas_cr_cysc))$creatinine_s
subset(panther, is.na(eGFR_fas_cr_cysc))$age
subset(panther, is.na(eGFR_fas_cr_cysc))$sex


test <- panther %>%
  mutate(qcr = floor(age),
         qcr = case_when(qcr == 8 ~ 0.46,
                         qcr == 9 ~ 0.49,
                         qcr == 10 ~ 0.51,
                         qcr == 11 ~ 0.53,
                         qcr == 12 ~ 0.57,
                         qcr == 13 ~ 0.59,
                         qcr == 14 ~ 0.61,
                         sex == "female" & qcr == 15 ~ 0.64,
                         sex == "female" & qcr == 15 ~ 0.64,
                         sex == "female" & qcr == 16 ~ 0.67,
                         sex == "female" & qcr == 17 ~ 0.69,
                         sex == "female" & qcr == 18 ~ 0.69,
                         sex == "female" & qcr >= 19 ~ 0.70,
                         
                         sex == "male" & qcr == 15 ~ 0.72,
                         sex == "male" & qcr == 16 ~ 0.78,
                         sex == "male" & qcr == 17 ~ 0.82,
                         sex == "male" & qcr == 18 ~ 0.85,
                         sex == "male" & qcr == 19 ~ 0.88,
                         sex == "male" & qcr > 19 ~ 0.90),
         f1 = creatinine_s / qcr,
         f3 = cystatin_c_s / 0.82,
         egfr_calc = 107.3 / ((0.5 * f1) + (0.5 * f3))) %>%
  dplyr::select(record_id, visit, age, sex, qcr, cystatin_c_s, creatinine_s, f1, f3, eGFR_fas_cr_cysc, egfr_calc)

```

```{r results = 'asis', echo = F}
# Table 2
kable(summary(tableby(group_risk ~  total_kidney_volume_ml + ht_adj_tkv + avg_c_r2 + avg_k_r2 + avg_c_t1 + avg_pcascl + avg_c_adc + 
                        mm_si + mm_ir + mm_bcell + mm_di + gfr_raw_plasma + gfr_bsa_plasma + erpf_raw_plasma + erpf_bsa_plasma +
                        ra + re + anova(rvr, digits = 3) + glomerular_pressure, 
                      data = panther, 
                      control = table_controls,
                      test = T), digits = 1))
```

```{r results = 'asis', echo = F}
# Table 3
kable(summary(tableby(group_risk ~  kwt(lh, "medianq1q3") + kwt(estrad, "medianq1q3") + kwt(tot_test, "medianq1q3") + 
                        kwt(free_test, "medianq1q3") + bl_dheas + igf_1 + igf1_z_score_calc + cpeptide + 
                        dexa_lean_mass + dexa_body_fat + dexa_fat_kg + dexa_trunk_mass, 
                      data = panther,                      
                      control = table_controls,
                      test = T), digits = 1))
```

```{r include = F}
subset(panther, is.na(acr_u))
subset(panther, is.na(dexa_lean_mass))
subset(panther, is.na(bl_dheas))
subset(panther, is.na(igf_1))
subset(panther, is.na(cpeptide), select = c(mrn, visit, record_id, date, cpep_minus_15))
```

## By tanner stage

### By tanner stage (4 & 5 combined)

```{r desc-tanner-stage-45, results = 'asis', echo = F}
# By tanner stage
summary(tableby(tanner_stage_comp_panther_cat ~ age + sex + group_2 + race_ethnicity_condensed + height + weight + 
                  bmi +  bmiz + bmip + sbp + dbp +  hba1c + 
                  eGFR_CKD_epi + eGFR_fas_cr+ eGFR_fas_cr_cysc + eGFR_CKiD_U25_CystatinC + eGFR_CKiD_U25_Creat + eGFR_CKiD_U25_avg + 
                  kwt(acr_u, "medianq1q3") + uacr_cat + kwt(u24_mab, "medianq1q3") + 
                  gotast_base + gptalt_base + total_kidney_volume_ml + ht_adj_tkv + avg_c_r2 + avg_k_r2 + avg_c_t1 + avg_pcascl + avg_c_adc + 
                  mm_si + mm_ir + mm_bcell + mm_di + gfr_raw_plasma + gfr_bsa_plasma + erpf_raw_plasma + erpf_bsa_plasma +
                  ra + re + anova(rvr, digits = 3) + glomerular_pressure + kwt(lh, "medianq1q3") + kwt(estrad, "medianq1q3") + kwt(tot_test, "medianq1q3") + 
                  kwt(free_test, "medianq1q3") + bl_dheas + igf_1 + igf1_z_score_calc + cpeptide + 
                  dexa_lean_mass + dexa_body_fat + dexa_fat_kg + dexa_trunk_mass,
                data = subset(panther, !is.na(group_risk)), 
                # control = table_controls, 
                total = T, test = F, digits = 1))
```

## Figures

```{r echo = F}
panther <- panther %>%
  subset(!is.na(tanner_stage_comp) & !is.na(group_risk))
```

### IQR plots
```{r echo = F, warning = F}
iqr_plot <- function(data, x_var, y_var, x_var_lab, y_var_lab) {
  # Perform ANOVA test
  formula <- as.formula(paste(y_var, "~", x_var))
  anova_res <- aov(formula, data = data)
  p_value <- summary(anova_res)[[1]][["Pr(>F)"]][1]
  
  # Format p-value for display
  p_text <- ifelse(p_value < 0.001, "ANOVA p < 0.001", paste0("ANOVA p = ", round(p_value, 3)))
  
  ggplot(data = data, aes_string(x = x_var, y = y_var)) +
    geom_jitter(width = 0.2, alpha = 0.5) +
    stat_summary(
      fun = median,
      geom = "crossbar",
      width = 0.1) +
    stat_summary(
      fun = median,
      geom = "line",
      group_2 = 1) +
    stat_summary(
      fun.data = function(x) {
        data.frame(
          y = median(x),
          ymin = quantile(x, 0.25),
          ymax = quantile(x, 0.75))
      },
      geom = "errorbar",
      width = 0.2) +
    labs(x = x_var_lab,
         y = y_var_lab) +
    theme(panel.background = element_blank(),
          text = element_text(size = 18),
          axis.title.y = element_text(size = 18)) +
    annotate("text", x = 1, y = max(data[[y_var]], na.rm = TRUE),
             label = p_text, hjust = 0, size = 6)
}
```

```{r echo = F, warning = F}
vars <- c("bmi", "birthweight", "eGFR_CKD_epi", "eGFR_fas_cr", "sbp", "dbp", "hba1c", "gotast_base", "gptalt_base", "log_acr_u", "log_lh", "log_estrad", "log_tot_test", "log_free_test", "bl_dheas", "igf_1", "igf1_z_score_calc", "dexa_lean_mass", "dexa_body_fat", "dexa_fat_kg", "dexa_trunk_mass", "avg_c_r2", "avg_k_r2", "avg_pcascl", "avg_c_adc", "avg_c_adc_tkv", "avg_c_r2_tkv", "avg_pcascl_tkv", "total_kidney_volume_ml", "ht_adj_tkv", "mm_si", "mm_ir", "mm_bcell", "mm_di", "gfr_raw_plasma", "erpf_raw_plasma", "erpf_bsa_plasma", "ra", "re", "rvr", "glomerular_pressure")
```

```{r echo = F, warning = F, eval = F}
plots <- map(vars, ~ iqr_plot(
  data = panther,
  x_var = "tanner_stage_comp_panther_cat",
  y_var = .x,
  x_var_lab = "Tanner stage",
  y_var_lab = label(panther[[.x]])))
walk(plots, print)
out_dir <- file.path(root_path, "PANTHER/Results/Figures/022326/IQR Pooled")
walk2(
  .x = plots,
  .y = vars,
  .f = ~ ggsave(
    filename = glue("{out_dir}/plot_{.y}.jpeg"),
    plot = .x,
    units = "in"
  )
)
```


```{r echo = F, warning = F}
iqr_plot_grp <- function(data, x_var, y_var, group_var = NULL,
                         x_var_lab, y_var_lab, group_lab = NULL,
                         colors = NULL) {
  p <- ggplot(data = data, aes_string(x = x_var, y = y_var,
                                      color = group_var)) +
    geom_jitter(width = 0.2, alpha = 0.5) +
    stat_summary(
      fun = median,
      geom = "crossbar",
      width = 0.2) +
    stat_summary(
      fun.data = function(x) {
        data.frame(
          y = median(x),
          ymin = quantile(x, 0.25),
          ymax = quantile(x, 0.75))
      },
      geom = "errorbar",
      width = 0.2,
      size = 2,
      aes_string(linetype = group_var)) +
    stat_summary(
      fun = median,
      geom = "line",
      aes_string(group_2 = group_var,
                 linetype = group_var),
      size = 2) +
    labs(x = x_var_lab,
         y = y_var_lab,
         color = group_lab,
         linetype = group_lab) +
    theme(panel.background = element_blank(),
          text = element_text(size = 18),
          axis.title.y = element_text(size = 18))
  
  # Apply custom colors if provided
  if (!is.null(colors)) {
    p <- p + scale_color_manual(values = colors)
  }
  
  # Compute ANOVA p-values for each group_2
  p_values <- data %>%
    group_by_at(group_var) %>%
    dplyr::summarise(p_value = summary(aov(get(y_var) ~ get(x_var), data = cur_data()))[[1]][["Pr(>F)"]][1]) %>%
    ungroup()
  
  # Format p-values for display
  p_values <- p_values %>%
    mutate(p_text = ifelse(p_value < 0.001, "p < 0.001", paste0("p = ", round(p_value, 3))))
  
  # Add p-value annotations with matching colors
  for (i in seq_len(nrow(p_values))) {
    grp <- p_values[[group_var]][i]
    p_val <- p_values$p_text[i]
    color <- if (!is.null(colors)) colors[grp] else "black"
    
    p <- p + annotate("text", x = 1, y = max(data[[y_var]], na.rm = TRUE) - i * 0.1 * diff(range(data[[y_var]], na.rm = TRUE)),
                      label = paste0("ANOVA (",grp,")", ": ", p_val), hjust = 0, size = 6, color = color)
  }
  
  return(p)
}
```

#### Risk
```{r echo = F, warning = F, eval = F}
plots <- map(vars, ~ iqr_plot_grp(
  data = panther,
  x_var = "tanner_stage_comp_panther_cat",
  y_var = .x,
  group_var = "group_risk",
  x_var_lab = "Tanner stage",
  y_var_lab = label(panther[[.x]]),
  group_lab = "Risk group_2"))
walk(plots, print)

out_dir <- file.path(root_path, "PANTHER/Results/Figures/022326/IQR Risk")
walk2(
  .x = plots,
  .y = vars,
  .f = ~ ggsave(
    filename = glue("{out_dir}/plot_{.y}.jpeg"),
    plot = .x,
    units = "in"
  )
)
```

#### Sex
```{r echo = F, warning = F, eval = F}
plots <- map(vars, ~ iqr_plot_grp(
  data = panther,
  x_var = "tanner_stage_comp_panther_cat",
  y_var = .x,
  group_var = "sex",
  x_var_lab = "Tanner stage",
  y_var_lab = label(panther[[.x]]),
  group_lab = "Sex"))
walk(plots, print)

out_dir <- file.path(root_path, "PANTHER/Results/Figures/022326/IQR Sex")
walk2(
  .x = plots,
  .y = vars,
  .f = ~ ggsave(
    filename = glue("{out_dir}/plot_{.y}.jpeg"),
    plot = .x,
    units = "in"
  )
)
```

## Heatmap

```{r}
# corr_plot_modified <- function(
#     data, X, Y, cor_method = "pearson", adj_var = NULL, dict = dict,
#     method = "color", insig = "pch", coef_col = NULL,
#     pch = 4, pch.col = "black",
#     pch.cex = 0, tl.cex = .8, cl.cex = 0.8, number.cex = 1,
#     coef_text_cex = 0.9,
#     star_text_cex = 0.9,
#     col = colorRampPalette(c("#0096c7", "white", "#d62828"))(200),
#     x_groups = NULL,        # NEW: named list for column grouping
#     y_groups = NULL,        # NEW: named list for row grouping
#     group_line_col = "#495057",
#     group_label_cex = 0.9,
#     group_offset = 1.2      # how far outside the plot the bracket sits
# ) {
#   n_cols <- length(Y)
#   M <- cor(y = subset(data, select = Y),
#            x = subset(data, select = X),
#            use = "pairwise.complete.obs",
#            method = cor_method)
#   
#   colnames(M) <- as.list(dict[match(colnames(M), names(dict))])
#   rownames(M) <- as.list(dict[match(rownames(M), names(dict))])
#   
#   # --- adjusted-variable block (unchanged) ---
#   if (!is.null(adj_var)) {
#     x_vars <- rep(X, times = length(Y))
#     x_coord <- rep(seq(1, length(Y)), each = length(X))
#     y_vars <- rep(Y, each = length(X))
#     y_coord <- rep(seq(length(X), 1), times = length(Y))
#     lm_extracted <- data.frame(yName = character(0),
#                                xName = character(0),
#                                x = numeric(0),
#                                y = numeric(0),
#                                adj_var = character(0),
#                                adj_x_coef = numeric(0),
#                                adj_x_pval = numeric(0))
#     for (i in seq_along(x_vars)) {
#       if (x_vars[i] != y_vars[i]) {
#         lm_formula <- as.formula(paste0(y_vars[i], "~", x_vars[i], "+", adj_var))
#         lm_model <- lm(lm_formula, data = data)
#         lm_summary <- summary(lm_model)
#         lm_coef <- lm_summary$coefficient[x_vars[i], "Estimate"]
#         lm_p_val <- lm_summary$coefficient[x_vars[i], "Pr(>|t|)"]
#         lm_extracted <- rbind(lm_extracted,
#                               data.frame(
#                                 yName = y_vars[i],
#                                 xName = x_vars[i],
#                                 x = x_coord[i],
#                                 y = y_coord[i],
#                                 adj_var = adj_var,
#                                 adj_x_coef = lm_coef,
#                                 adj_x_pval = lm_p_val))
#       }
#     }
#   }
#   
#   correlation_p_value <- correlation_p_value_matrix(
#     data, relevant_vars = c(Y, X), n_cols = n_cols, cor_method = cor_method
#   )
#   
#   
#   corrplot(M,
#            p.mat = correlation_p_value,
#            method = method,
#            tl.col = "black",
#            tl.srt = 45,
#            tl.cex = tl.cex,
#            insig = insig,
#            addCoef.col = coef_col,
#            addgrid.col = 'white',
#            number.digits = 2,
#            number.cex = number.cex,
#            pch.col = pch.col,
#            pch = pch,
#            pch.cex = pch.cex,
#            col = col,
#            cl.cex = cl.cex, 
#            mar = c(0, group_offset + 1, 0,0) # bottom, left, top, right
#   )$corrPos -> p1
#   
#   p1_sub  <- subset(p1, p.value <= 0.05)
#   p1_sub2 <- subset(p1, p.value <= 0.05 & abs(corr) >= 0.7)
#   
#   if (nrow(p1) > 0) {
#     graphics::text(p1$x, p1$y, sprintf("%.2f", p1$corr),
#                    adj = c(0.5, 0), cex = coef_text_cex)
#   }
#   if (nrow(p1_sub) > 0) {
#     graphics::text(p1_sub$x, p1_sub$y, stars.pval(p1_sub$p.value),
#                    adj = c(0.5, 2), cex = star_text_cex)
#   }
#   if (nrow(p1_sub2) > 0) {
#     graphics::text(p1_sub2$x, p1_sub2$y, sprintf("%.2f", p1_sub2$corr),
#                    col = "white", adj = c(0.5, 0), cex = coef_text_cex)
#     graphics::text(p1_sub2$x, p1_sub2$y, stars.pval(p1_sub2$p.value),
#                    col = "white", adj = c(0.5, 2), cex = star_text_cex)
#   }
#   if (!is.null(adj_var)) {
#     lm_extracted <- subset(lm_extracted, adj_x_pval <= 0.05)
#     if (nrow(lm_extracted) > 0) {
#       graphics::rect(xleft = lm_extracted$x - 0.45,
#                      ybottom = lm_extracted$y - 0.45,
#                      xright = lm_extracted$x + 0.45,
#                      ytop = lm_extracted$y + 0.45)
#     }
#   }
#   
#  # ──────────────────────────────────────────────
#   #   NEW: draw grouping brackets
#   # ──────────────────────────────────────────────
#   
#   par(xpd = TRUE)
#   
#   # Helper: translate original variable names -> dict labels
#   to_label <- function(v) {
#     lbl <- dict[match(v, names(dict))]
#     ifelse(is.na(lbl), v, lbl)
#   }
#   
#   # X-groups: X variables are plotted on the ROW axis (left side of plot)
#   # Rows of M map to X variables; corrplot y-coord: row 1 (top) = y nrow(M), last row = y 1
#   if (!is.null(x_groups)) {
#     row_labels <- rownames(M)
#     n_rows <- nrow(M)
#     x_base <- 0.5 - group_offset
#     
#     for (grp_name in names(x_groups)) {
#       grp_labels <- to_label(x_groups[[grp_name]])
#       idx <- which(row_labels %in% grp_labels)
#       if (length(idx) == 0) next
#       
#       y_positions <- n_rows - idx + 1
#       y_lo <- min(y_positions) - 0.3
#       y_hi <- max(y_positions) + 0.3
#       y_mid <- (y_lo + y_hi) / 2
#       
#       # vertical span line
#       graphics::segments(x_base, y_lo, x_base, y_hi,
#                          col = group_line_col, xpd = TRUE)
#       # centered rotated label
#       graphics::text(x_base - 0.35, y_mid, grp_name,
#                      col = "#495057",
#                      cex = group_label_cex, srt = 90, xpd = TRUE)
#     }
#   }
#   
#   # Y-groups: Y variables are plotted on the COLUMN axis (bottom/top of plot)
#   # Columns of M map to Y variables; position 1..length(Y) left-to-right
#   if (!is.null(y_groups)) {
#     col_labels <- colnames(M)
#     y_base <- 0.5 - group_offset
#     
#     for (grp_name in names(y_groups)) {
#       grp_labels <- to_label(y_groups[[grp_name]])
#       idx <- which(col_labels %in% grp_labels)
#       if (length(idx) == 0) next
#       
#       x_lo <- min(idx) - 0.3
#       x_hi <- max(idx) + 0.3
#       x_mid <- (x_lo + x_hi) / 2
#       
#       # horizontal span line
#       graphics::segments(x_lo, y_base, x_hi, y_base,
#                          col = group_line_col, xpd = TRUE)
#       # centered label
#       graphics::text(x_mid, y_base - 0.35, grp_name,
#                      cex = group_label_cex, xpd = TRUE)
#     }
#   }
# }  
```

```{r}
corr_plot_modified_gg <- function(
    data, X, Y, cor_method = "pearson", adj_var = NULL, dict = dict,
    coef_text_size = 5,
    star_text_size = 6,
    axis_text_size = 11,
    legend_text_size = 8,
    col_low = "#0096c7", col_mid = "white", col_high = "#d62828",
    x_groups = NULL,
    y_groups = NULL,
    group_line_col = "#495057",
    group_label_size = 3.5,
    bracket_line_pos = -1.8,    # x-coordinate for the bracket line
    bracket_label_pos = -2.5    # x-coordinate for the bracket label
) {
  
  library(ggplot2)
  library(dplyr)
  library(tidyr)
  
  # ── 1. Correlation matrix ──────────────────────────────────
  M <- cor(y = subset(data, select = Y),
           x = subset(data, select = X),
           use = "pairwise.complete.obs",
           method = cor_method)
  
  # ── 2. P-value matrix ─────────────────────────────────────
  p_mat <- matrix(NA, nrow = length(X), ncol = length(Y),
                  dimnames = list(X, Y))
  for (xi in X) {
    for (yi in Y) {
      complete <- complete.cases(data[[xi]], data[[yi]])
      if (sum(complete) > 3) {
        test <- cor.test(data[[xi]][complete], data[[yi]][complete], method = cor_method)
        p_mat[xi, yi] <- test$p.value
      }
    }
  }
  
  # ── 3. Adjusted variable analysis ─────────────────────────
  adj_df <- NULL
  if (!is.null(adj_var)) {
    adj_list <- list()
    for (xi in X) {
      for (yi in Y) {
        if (xi != yi) {
          lm_formula <- as.formula(paste0(yi, "~", xi, "+", adj_var))
          lm_model <- lm(lm_formula, data = data)
          lm_summary <- summary(lm_model)
          adj_list[[length(adj_list) + 1]] <- data.frame(
            xvar = xi, yvar = yi,
            adj_coef = lm_summary$coefficients[xi, "Estimate"],
            adj_pval = lm_summary$coefficients[xi, "Pr(>|t|)"],
            stringsAsFactors = FALSE
          )
        }
      }
    }
    adj_df <- bind_rows(adj_list) %>% filter(adj_pval <= 0.05)
  }
  
  # ── 4. Build long-form data frame ─────────────────────────
  to_label <- function(v) {
    lbl <- dict[match(v, names(dict))]
    ifelse(is.na(lbl), v, as.character(lbl))
  }
  
  cor_df <- reshape2::melt(M, varnames = c("xvar", "yvar"), value.name = "corr")
  p_df   <- reshape2::melt(p_mat, varnames = c("xvar", "yvar"), value.name = "pval")
  plot_df <- left_join(cor_df, p_df, by = c("xvar", "yvar"))
  
  plot_df$x_label <- factor(to_label(plot_df$xvar), levels = rev(to_label(X)))
  plot_df$y_label <- factor(to_label(plot_df$yvar), levels = to_label(Y))
  
  plot_df$stars <- ifelse(is.na(plot_df$pval), "",
                          ifelse(plot_df$pval <= 0.001, "***",
                                 ifelse(plot_df$pval <= 0.01, "**",
                                        ifelse(plot_df$pval <= 0.05, "*", ""))))
  
  plot_df$text_col <- ifelse(plot_df$pval <= 0.05 & abs(plot_df$corr) >= 0.7, "white", "black")
  plot_df$star_col <- ifelse(plot_df$pval <= 0.05 & abs(plot_df$corr) >= 0.7, "white", "black")
  
  # ── 5. Adjusted variable boxes ────────────────────────────
  if (!is.null(adj_df) && nrow(adj_df) > 0) {
    adj_df$x_label <- factor(to_label(adj_df$xvar), levels = rev(to_label(X)))
    adj_df$y_label <- factor(to_label(adj_df$yvar), levels = to_label(Y))
  }
  
  # ── 6. Build the plot ─────────────────────────────────────
  sig_df <- filter(plot_df, stars != "")
  
  p <- ggplot(plot_df, aes(x = y_label, y = x_label)) +
    geom_tile(aes(fill = corr), color = "white", linewidth = 1) +
    scale_fill_gradient2(low = col_low, mid = col_mid, high = col_high,
                         midpoint = 0, limits = c(-1, 1), name = NULL) +
    geom_text(aes(label = sprintf("%.2f", corr), color = text_col),
              size = coef_text_size, vjust = -0.3, show.legend = FALSE) +
    scale_color_identity() +
    geom_text(data = sig_df,
              aes(label = stars),
              color = sig_df$star_col,
              size = star_text_size, vjust = 1.8) +
    {if (!is.null(adj_df) && nrow(adj_df) > 0)
      geom_tile(data = adj_df, aes(x = y_label, y = x_label),
                fill = NA, color = "black", linewidth = 0.8)
    } +
    scale_x_discrete(position = "top") +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x.top = element_text(angle = 45, hjust = 0, vjust = 0,
                                     size = axis_text_size, color = "black"),
      axis.text.y = element_text(size = axis_text_size, color = "black"),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      legend.text = element_text(size = legend_text_size),
      plot.margin = margin(
        t = 10,
        r = if (!is.null(x_groups)) 60 else 10,
        b = if (!is.null(y_groups)) 60 else 10,
        l = 10
      )
    ) +
    coord_fixed(clip = "off")
  
# ── 7. Grouping brackets ──────────────────────────────────
  y_levels <- rev(to_label(X))
  x_levels <- to_label(Y)
  
  n_x <- length(x_levels)  # number of columns
  n_y <- length(y_levels)  # number of rows
  
  # X-groups: bracket on the RIGHT side
  if (!is.null(x_groups)) {
    for (grp_name in names(x_groups)) {
      grp_labels <- to_label(x_groups[[grp_name]])
      idx <- which(y_levels %in% grp_labels)
      if (length(idx) == 0) next
      
      y_lo <- min(idx) - 0.4
      y_hi <- max(idx) + 0.4
      y_mid <- (y_lo + y_hi) / 2
      line_x <- n_x + bracket_line_pos
      label_x <- n_x + bracket_label_pos
      
      p <- p + annotate("segment",
                        x = line_x, xend = line_x,
                        y = y_lo, yend = y_hi,
                        color = group_line_col, linewidth = 0.6)
      p <- p + annotate("text",
                        x = label_x, y = y_mid,
                        label = grp_name,
                        color = group_line_col,
                        size = group_label_size, angle = 270)
    }
  }
  
  # Y-groups: bracket on the BOTTOM
  if (!is.null(y_groups)) {
    for (grp_name in names(y_groups)) {
      grp_labels <- to_label(y_groups[[grp_name]])
      idx <- which(x_levels %in% grp_labels)
      if (length(idx) == 0) next
      
      x_lo <- min(idx) - 0.4
      x_hi <- max(idx) + 0.4
      x_mid <- (x_lo + x_hi) / 2
      line_y <- 1 - bracket_line_pos
      label_y <- 1 - bracket_label_pos
      
      p <- p + annotate("segment",
                        x = x_lo, xend = x_hi,
                        y = line_y, yend = line_y,
                        color = group_line_col, linewidth = 0.6)
      p <- p + annotate("text",
                        x = x_mid, y = label_y,
                        label = grp_name,
                        color = group_line_col,
                        size = group_label_size)
    }
  }
  
  return(p)
}
```

```{r echo = F}
xvars <- c("age", "tanner_stage_comp_panther", "log_lh", "fsh", "log_tot_test", "log_free_test", "log_estrad", "bl_dheas", "igf_1", "mm_si", "mm_bcell")

dict_heat <- dict %>%
  mutate(
    # Demographics
    age = "Age",
    tanner_stage_comp_panther = "Tanner stage",
    
    # Sex hormones
    log_lh = "ln(LH)",
    fsh = "FSH",
    log_tot_test = "ln(Total\ntestosterone)",
    log_free_test = "ln(Free\ntestosterone)",
    log_estrad = "ln(Estradiol)",
    
    # Metabolic parameters
    bl_dheas = "DHEA-S",
    igf_1 = "IGF-1",
    mm_si = "Insulin\nsensitivity",
    mm_bcell = "Beta-cell\nfunction",
    
    # Y variables (renal outcomes)
    gfr_raw_plasma = "mGFR",
    erpf_raw_plasma = "ERPF",
    erpf_bsa_plasma = "ERPF (BSA)",
    eGFR_fas_cr = "eGFR (EKFC)",
    log_acr_u = "ln(UACR)",
    glomerular_pressure = "Glomerular\npressure",
    ra = "Afferent\nresistance",
    re = "Efferent\nresistance",
    rvr = "Renal vascular\nresistance",
    
    total_kidney_volume_ml = "TKV",
    ht_adj_tkv = "Ht-adjusted\nTKV",
    avg_c_r2 = "Cortical R2*",
    avg_k_r2 = "Kidney R2*",
    avg_pcascl = "pCASL",
    avg_c_adc = "Cortical ADC",
    
    dexa_lean_mass = "Lean mass\npercentage",
    dexa_lean_kg = "Lean mass",
    dexa_body_fat = "Body fat mass\npercentage",
    dexa_fat_kg = "Body fat mass",
    dexa_trunk_mass = "Trunk mass\npercentage",
    dexa_trunk_kg = "Trunk mass",
    bmi = "BMI"
  )

corr_plot_modified_gg(
  panther,
  X = xvars,
  Y = c("gfr_raw_plasma", "erpf_raw_plasma", "erpf_bsa_plasma", "eGFR_fas_cr",
        "log_acr_u", "glomerular_pressure", "ra", "re", "rvr"),
  dict = dict_heat,
  x_groups = list(
    "Demographics" = c("age", "tanner_stage_comp_panther"),
    "Sex hormones" = c("log_lh", "fsh", "log_tot_test", "log_free_test", "log_estrad"),
    "Metabolic parameters" = c("bl_dheas", "igf_1", "mm_si", "mm_bcell")
  ),
  bracket_line_pos = 0.7,
  bracket_label_pos = 1,
  group_label_size = 3.5
)

ggsave(file.path(root_path, "PANTHER/Results/Figures/panther_cor1_brackets_gg.png"),
       width = 10, height = 10, dpi = 170)


corr_plot_modified_gg(
  panther,
  X = xvars,
  Y = c("total_kidney_volume_ml", "ht_adj_tkv","avg_c_r2",
        "avg_k_r2","avg_c_t1", "avg_pcascl", "avg_c_adc"),
  dict = dict_heat,
  x_groups = list(
    "Demographics" = c("age", "tanner_stage_comp_panther"),
    "Sex hormones" = c("log_lh", "fsh", "log_tot_test", "log_free_test", "log_estrad"),
    "Metabolic parameters" = c("bl_dheas", "igf_1", "mm_si", "mm_bcell")
  ),
  bracket_line_pos = 0.7,
  bracket_label_pos = 1,
  group_label_size = 3.5
)

ggsave(file.path(root_path, "PANTHER/Results/Figures/panther_cor2_brackets_gg.png"),
       width = 10, height = 10, dpi = 170)

xvars2 <- c("dexa_lean_kg","dexa_lean_mass",  "dexa_fat_kg", "dexa_body_fat", "dexa_trunk_kg", "dexa_trunk_mass", 
            "bmi")


corr_plot_modified_gg(
  panther,
  X = xvars2,
  Y = c("total_kidney_volume_ml", "ht_adj_tkv", "rvr", "rbf", "glomerular_pressure", "ra", "re"),
  dict = dict_heat,
  x_groups = list(
    "Lean" = c("dexa_lean_mass", "dexa_lean_kg"),
    "Fat" = c("dexa_body_fat", "dexa_fat_kg"),
    "Trunk" = c("dexa_trunk_mass", "dexa_trunk_kg")
  ),
  y_groups = list(
    "Kidney\nstructure" = c("total_kidney_volume_ml", "ht_adj_tkv"),
    "Renal\nhemodynamics" = c("rvr", "rbf"),
    "Glomerular\nhemodynamics" = c("glomerular_pressure", "ra", "re")
  ),
  bracket_line_pos = 0.7,
  bracket_label_pos = 1,
  group_label_size = 3.5
)

ggsave(file.path(root_path, "PANTHER/Results/Figures/panther_cor5_brackets_gg.png"),
       width = 10, height = 10, dpi = 170)
```

#### Sex hormones adjusting for age

```{r echo = F}
summary(lm(log_lh ~ tanner_stage_comp_panther + age, panther))
summary(lm(fsh ~ tanner_stage_comp_panther + age, panther))
summary(lm(log_tot_test ~ tanner_stage_comp_panther + age, panther))
summary(lm(log_free_test ~ tanner_stage_comp_panther + age, panther))
summary(lm(log_estrad ~ tanner_stage_comp_panther + age, panther))
```

Partial correlations adjusting for age (tiles with black rectangles are significant at p < 0.05 in partial correlation adjusting for age).

```{r echo = F}
# partial correlations adjusting for age
# png(height = 500, width = 1500, file = file.path(root_path, "PANTHER/Results/Figures/panther_cor5.png"), res = 170)
corr_plot_modified(panther,
                   X = "tanner_stage_comp_panther",
                   Y = c("log_lh", "fsh", "log_tot_test", "log_free_test",
                         "log_estrad"),
                   dict = dict,
                   adj_var = "age")
# dev.off()
```

```{r echo = F}
# correlations, regression adjusting for age
# png(height = 500, width = 1500, file = file.path(root_path, "PANTHER/Results/Figures/panther_cor6.png"), res = 170)
corr_plot_ver2(panther,
               X = c("tanner_stage_comp_panther"),
               Y = c("log_lh", "fsh", "log_tot_test", "log_free_test",
                     "log_estrad"),
               dict = dict,
               adj_var = "age")
# dev.off()
```

```{r echo = F}
# correlations, regression adjusting for sex and age
# png(height = 800, width = 1000, file = file.path(root_path, "PANTHER/Results/Figures/panther_cor7.png"), res = 170)
corr_plot_ver2(panther,
               X = c("tanner_stage_comp_panther", "log_lh", "fsh", "log_tot_test", "log_free_test",
                     "log_estrad"),
               Y = c("total_kidney_volume_ml", "ht_adj_tkv", "avg_c_r2", "avg_k_r2", "avg_c_t1", "avg_pcascl", "avg_c_adc"),
               dict = dict,
               adj_var = "age+sex+bmi")
# dev.off()

# png(height = 800, width = 1400, file = file.path(root_path, "PANTHER/Results/Figures/panther_cor8.png"), res = 170)
corr_plot_ver2(panther,
               X = c("tanner_stage_comp_panther", "log_lh", "fsh", "log_tot_test", "log_free_test",
                     "log_estrad"),
               Y = c("gfr_raw_plasma", "erpf_raw_plasma", "erpf_bsa_plasma","eGFR_fas_cr",
                     "log_acr_u", "glomerular_pressure", "ra", "re", "rvr"),
               dict = dict,
               adj_var = "age+sex+bmi")
# dev.off()
```


## Volcano plots for proteomics
### Risk

```{r echo = F}
panther <- panther %>% arrange(record_id)
# Look at differential expressed proteins (DEPs) between low and high risk in PANTHER
# filter soma scan data to baseline data only and log transform
soma_risk <- panther %>%
  remove_rownames() %>% column_to_rownames(var = "record_id") %>%
  dplyr::select(starts_with("seq"))
y <- t(soma_risk)
y <- log2(y)

# Design matrix
cond = as.factor(panther$group_risk)
design_mat <- model.matrix(~0 + group_risk, data = panther)
colnames(design_mat) = gsub("group_risk","",colnames(design_mat))
# Fit
contrast = makeContrasts(contrasts = c("High-Low"), levels = design_mat) # positive = higher expression in high risk grp
fit <- lmFit(y, design_mat)
fit <- contrasts.fit(fit, contrasts = contrast)
fit <- eBayes(fit)
res_risk <- topTable(fit, coef = 1, number = dim(y)[1], sort.by = "p")
res_risk$Target <- analytes$Target[match(rownames(res_risk), analytes$AptName)]
res_risk$TargetFullName <- analytes$TargetFullName[match(rownames(res_risk), analytes$AptName)]
kbl(head(res_risk, 50)) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "500px")
res_risk_save <- res_risk %>%
  rownames_to_column("AptName") %>%
  dplyr::select(-c(Target, TargetFullName)) %>%
  left_join(analytes, by = "AptName")

EnhancedVolcano(res_risk_save,
                lab = res_risk_save$Target,
                x = 'logFC',
                y = 'adj.P.Val',
                title = paste0("High vs. Low Risk"),
                subtitle = "",
                pCutoff = 0.05,
                FCcutoff = 0.75,
                labFace = 'bold',
                pointSize = 4,
                labSize = 5,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                legendPosition=c(0.9,0.5),
                boxedLabels = TRUE,
                max.overlaps=30,
                caption = "")
# ggsave(file.path(root_path, "PANTHER/Results/Figures/volcano_risk_dep.jpeg"),
#       width = 12, height = 10)

```

### Across Tanner stage 1-2 vs. 3-5

```{r echo = F}
soma <- panther %>%
  filter(!is.na(tanner_stage_cat)) %>%
  remove_rownames() %>% column_to_rownames(var = "record_id") %>%
  dplyr::select(starts_with("seq"))
y <- t(soma)
y <- log2(y)
# Design matrix
cond = as.factor(panther$tanner_stage_cat)
panther <- panther %>% arrange(record_id)
design_mat <- model.matrix(~0 + tanner_stage_cat, data = panther)
colnames(design_mat) = gsub("tanner_stage_cat","",colnames(design_mat))
# Fit
contrast = makeContrasts(contrasts = c("III_V-I_II"), levels = design_mat) # positive = higher expression in TS 3-5
fit <- lmFit(y, design_mat)
fit <- contrasts.fit(fit, contrasts = contrast)
fit <- eBayes(fit)
res_ts <- topTable(fit, coef = 1, number = dim(y)[1], sort.by = "p")
res_ts$Target <- analytes$Target[match(rownames(res_ts), analytes$AptName)]
res_ts$TargetFullName <- analytes$TargetFullName[match(rownames(res_ts), analytes$AptName)]
kbl(head(res_ts, 50)) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "500px")
res_ts_save <- res_ts %>%
  rownames_to_column("AptName") %>%
  dplyr::select(-c(Target, TargetFullName)) %>%
  left_join(analytes, by = "AptName")

EnhancedVolcano(res_ts_save,
                lab = res_ts_save$Target,
                x = 'logFC',
                y = 'P.Value',
                title = paste0("TS>=3 vs. TS<3 (with unadjusted p-values)"),
                subtitle = "",
                pCutoff = 0.05,
                FCcutoff = 0.5,
                labFace = 'bold',
                pointSize = 4,
                labSize = 5,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                legendPosition=c(0.9,0.5),
                boxedLabels = TRUE,
                max.overlaps=30,
                caption = "")
# ggsave(file.path(root_path, "PANTHER/Results/Figures/volcano_ts_dep.jpeg"),
#       width = 12, height = 10)
```

### Across Tanner stage 1-2 vs. 3-5 (High risk only)

```{r echo = F}
panther_high <- subset(panther, group_risk == "High")
soma <- panther_high %>%
  filter(!is.na(tanner_stage_cat)) %>%
  remove_rownames() %>% column_to_rownames(var = "record_id") %>%
  dplyr::select(starts_with("seq"))
y <- t(soma)
y <- log2(y)
# Design matrix
cond = as.factor(panther_high$tanner_stage_cat)
panther_high <- panther_high %>% arrange(record_id)
design_mat <- model.matrix(~0 + tanner_stage_cat, data = panther_high)
colnames(design_mat) = gsub("tanner_stage_cat","",colnames(design_mat))
# Fit
contrast = makeContrasts(contrasts = c("III_V-I_II"), levels = design_mat) # positive = higher expression in TS 3-5
fit <- lmFit(y, design_mat)
fit <- contrasts.fit(fit, contrasts = contrast)
fit <- eBayes(fit)
res_ts_high <- topTable(fit, coef = 1, number = dim(y)[1], sort.by = "p")
res_ts_high$Target <- analytes$Target[match(rownames(res_ts_high), analytes$AptName)]
res_ts_high$TargetFullName <- analytes$TargetFullName[match(rownames(res_ts_high), analytes$AptName)]
kbl(head(res_ts_high, 50)) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "500px")
res_ts_high_save <- res_ts_high %>%
  rownames_to_column("AptName") %>%
  dplyr::select(-c(Target, TargetFullName)) %>%
  left_join(analytes, by = "AptName")

EnhancedVolcano(res_ts_high_save,
                lab = res_ts_high_save$Target,
                x = 'logFC',
                y = 'P.Value',
                title = paste0("TS>=3 vs. TS<3 (High risk only, with unadjusted p-values)"),
                subtitle = "",
                pCutoff = 0.05,
                FCcutoff = 0.5,
                labFace = 'bold',
                pointSize = 4,
                labSize = 5,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                legendPosition=c(0.9,0.5),
                boxedLabels = TRUE,
                max.overlaps=30,
                caption = "")
# ggsave(file.path(root_path, "PANTHER/Results/Figures/volcano_ts_high_highrisk_dep.jpeg"),
#       width = 12, height = 10)
```
### Across Tanner stage 1-2 vs. 3-5 (Low risk only)

```{r echo = F}
panther_low <- subset(panther, group_risk == "Low")
soma <- panther_low %>%
  filter(!is.na(tanner_stage_cat)) %>%
  remove_rownames() %>% column_to_rownames(var = "record_id") %>%
  dplyr::select(starts_with("seq"))
y <- t(soma)
y <- log2(y)
# Design matrix
cond = as.factor(panther_low$tanner_stage_cat)
panther_low <- panther_low %>% arrange(record_id)
design_mat <- model.matrix(~0 + tanner_stage_cat, data = panther_low)
colnames(design_mat) = gsub("tanner_stage_cat","",colnames(design_mat))
# Fit
contrast = makeContrasts(contrasts = c("III_V-I_II"), levels = design_mat) # positive = higher expression in TS 3-5
fit <- lmFit(y, design_mat)
fit <- contrasts.fit(fit, contrasts = contrast)
fit <- eBayes(fit)
res_ts_low <- topTable(fit, coef = 1, number = dim(y)[1], sort.by = "p")
res_ts_low$Target <- analytes$Target[match(rownames(res_ts_low), analytes$AptName)]
res_ts_low$TargetFullName <- analytes$TargetFullName[match(rownames(res_ts_low), analytes$AptName)]
kbl(head(res_ts_low, 50)) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "500px")
res_ts_low_save <- res_ts_low %>%
  rownames_to_column("AptName") %>%
  dplyr::select(-c(Target, TargetFullName)) %>%
  left_join(analytes, by = "AptName")

EnhancedVolcano(res_ts_low_save,
                lab = res_ts_low_save$Target,
                x = 'logFC',
                y = 'P.Value',
                title = paste0("TS>=3 vs. TS<3 (Low risk only, with unadjusted p-values)"),
                subtitle = "",
                pCutoff = 0.05,
                FCcutoff = 0.5,
                labFace = 'bold',
                pointSize = 4,
                labSize = 5,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                legendPosition=c(0.9,0.5),
                boxedLabels = TRUE,
                max.overlaps=30,
                caption = "")
# ggsave(file.path(root_path, "PANTHER/Results/Figures/volcano_ts_low_lowrisk_dep.jpeg"),
#       width = 12, height = 10)
```

### Sex

```{r echo = F}
soma <- panther %>%
  remove_rownames() %>% column_to_rownames(var = "record_id") %>%
  dplyr::select(starts_with("seq"))
y <- t(soma)
y <- log2(y)
# Design matrix
panther <- panther %>% arrange(record_id)
design_mat <- model.matrix(~0 + sex, data = panther)
colnames(design_mat) = gsub("sex","",colnames(design_mat))
# Fit
contrast = makeContrasts(contrasts = c("male-female"), levels = design_mat) # positive = higher expression in males
fit <- lmFit(y, design_mat)
fit <- contrasts.fit(fit, contrasts = contrast)
fit <- eBayes(fit)
res_sex <- topTable(fit, coef = 1, number = dim(y)[1], sort.by = "p")
res_sex$Target <- analytes$Target[match(rownames(res_sex), analytes$AptName)]
res_sex$TargetFullName <- analytes$TargetFullName[match(rownames(res_sex), analytes$AptName)]

kbl(head(res_sex, 50)) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "500px")
res_sex_save <- res_sex %>%
  rownames_to_column("AptName") %>%
  dplyr::select(-c(Target, TargetFullName)) %>%
  left_join(analytes, by = "AptName")

EnhancedVolcano(res_sex_save,
                lab = res_sex_save$Target,
                x = 'logFC',
                y = 'P.Value',
                title = paste0("Male vs. Female (with unadjusted p-values)"),
                subtitle = "",
                pCutoff = 0.05,
                FCcutoff = 0.5,
                labFace = 'bold',
                pointSize = 4,
                labSize = 5,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                legendPosition=c(0.9,0.5),
                boxedLabels = TRUE,
                max.overlaps=30,
                caption = "")
# ggsave(file.path(root_path, "PANTHER/Results/Figures/volcano_sex_dep.jpeg"),
#       width = 12, height = 10)
```

### Correlation w/ key parameters shown to change with puberty above (in heatmaps)

```{r echo = F}
plotVolcano_mod <- function (data, FC, p.value, labels, identify = F,
                             identify_manual = NULL,
                             fc.cutoff = 1,
                             label_apt = F, pt.size = 8, text.size = 20,
                             cutoff = 0.05/nrow(data),
                             sig_fc_pos_lab = "Significant & Fold-Change (+)",
                             sig_fc_neg_lab = "Significant & Fold-Change (-)",
                             sig_lab = "Significant",
                             fc_lab = "Fold-Change",
                             ns_lab = "Non-Significant",
                             sig_fc_pos_lab_col = "#f28482",
                             sig_fc_neg_lab_col = "#8ecae6",
                             sig_lab_col = "#f6bd60",
                             fc_lab_col = "#84a59d",
                             ns_lab_col = "#dad7cd",
                             main = NULL, x.lab = NULL, ...)
{
  .fc <- enquo(FC)
  .p <- enquo(p.value)
  .label <- enquo(labels)
  
  if (all(pull(data, !!.fc) >= 0)) {
    warning("It appears you are not passing log2-transformed ",
            "fold-change values. Please check.", call. = FALSE)
  }
  if (is.null(main)) {
    main <- "Volcano Plot of Significant Fold Changes"
  }
  if (is.null(x.lab)) {
    x.lab <- bquote(italic(log)[2] ~ (Fold ~ Change))
  }
  y.lab <- bquote(-italic(log)[10] ~ (p - value))
  plot_df <- dplyr::mutate(data,
                           group_2 = case_when((-log10(!!.p) >= -log10(cutoff)) &
                                                 (abs(!!.fc) >= fc.cutoff & !!.fc > 0) ~
                                                 sig_fc_pos_lab,
                                               (-log10(!!.p) >= -log10(cutoff)) &
                                                 (abs(!!.fc) >= fc.cutoff & !!.fc < 0) ~
                                                 sig_fc_neg_lab,
                                               -log10(!!.p) >= -log10(cutoff) ~
                                                 sig_lab, abs(!!.fc) >= fc.cutoff ~
                                                 fc_lab,
                                               TRUE ~ ns_lab),
                           type = grepl(paste0("^", sig_lab, "|Significant"), group_2))
  
  create_col_vector <- function(label_values, colors) {
    cols <- setNames(colors, label_values)
    return(cols)
  }
  
  label_values = c(ns_lab, fc_lab, sig_lab, sig_fc_pos_lab, sig_fc_neg_lab)
  label_colors = c(ns_lab_col, fc_lab_col, sig_lab_col, sig_fc_pos_lab_col, sig_fc_neg_lab_col)
  cols <- create_col_vector(label_values, label_colors)
  
  p <- ggplot(plot_df, aes(x = !!.fc, y = -log10(!!.p), color = group_2)) +
    geom_point(alpha = 0.5, size = pt.size) +
    scale_color_manual(values = cols, name = "") +
    labs(x = x.lab, y = y.lab, title = main) +
    geom_vline(xintercept = 0,
               color = "grey50", linetype = "dashed", alpha = 0.75) +
    geom_hline(yintercept = -log10(0.05),
               color = "grey50", linetype = "dashed", alpha = 0.75) +
    theme_soma() +
    theme(text = element_text(size = text.size),
          panel.grid = element_blank(),
          plot.background = element_rect(color = "transparent")) +
    NULL
  
  if (identify) { # Label ALL significant points
    p <- p + geom_text_repel(data = dplyr::filter(plot_df, type),
                             aes(label = !!.label),
                             hjust = "inward",
                             size = text.size,
                             max.overlaps = Inf,
                             color = "black",
                             face = "bold")
  }
  if (!is.null(identify_manual) & label_apt) { # Label manually specified points only (with apt)
    plot_df <- plot_df %>%
      dplyr::mutate(label_col = rownames(plot_df))
    p <- p + geom_point(data = dplyr::filter(plot_df, !!.label %in% identify_manual),
                        size = pt.size + 3, alpha = 0.3) +
      geom_text_repel(data = dplyr::filter(plot_df, !!.label %in% identify_manual),
                      aes(label = label_col, color = group_2),
                      hjust = "inward",
                      size = text.size,
                      max.overlaps = 20,
                      color = "grey20",
                      force_pull = -.002,
                      force = 8)
    
  }
  if (!is.null(identify_manual) & !label_apt) { # Label manually specified points only (without apt)
    p <- p +
      geom_point(data = dplyr::filter(plot_df, !!.label %in% identify_manual),
                 size = pt.size + 3, alpha = 0.3) +
      geom_text_repel(data = dplyr::filter(plot_df, !!.label %in% identify_manual),
                      aes(label = !!.label, color = group_2),
                      hjust = "inward",
                      size = text.size,
                      max.overlaps = 20,
                      color = "grey20",
                      force_pull = -.002,
                      force = 8)
  }
  p
}
# Heatmap tables (combined)
ts_tbl <- function(x, columns = NULL){
  gt(x)  %>%
    data_color(columns = columns,
               method = "numeric",
               palette = c("#023e8a","#ade8f4",
                           "#ffffff",
                           "#ffe5ec","#d62828"),
               domain = c(-0.8, -0.2,
                          0,
                          0.2, 0.8)) %>%
    tab_options(column_labels.hidden = T) %>%
    as_raw_html()
}

# Function to run linear models & extract significant proteins
run_lm <- function(var) {
  analytes %>%
    select(AptName, SeqId, Target_nodup, EntrezGeneSymbol, UniProt) %>%
    mutate(
      formula   = map(AptName, ~ as.formula(paste(.x, "~", var))),  # Create dynamic formula
      lm_model  = map(formula, ~ lm(.x, data = panther)),  # Fit models
      p.value   = map_dbl(lm_model, ~ summary(.x)$coefficients[var, "Pr(>|t|)"]),  # Extract p-values
      fdr       = p.adjust(p.value, method = "BH")  # FDR correction
    ) %>%
    arrange(p.value) %>%
    filter(p.value < 0.05) %>%  # Select significant proteins
    pull(AptName) %>%
    unique()
}

run_corr <- function(var, sig_vars) {
  pearson_cor <- corr.test(x = panther[[var]],
                           y = panther[sig_vars],
                           method = "pearson",
                           adjust = "none", alpha = 0.05)
  pearson_cor_r <- data.frame(t(pearson_cor$r))
  colnames(pearson_cor_r) <- var
  pearson_cor_p <- data.frame(p.value = (pearson_cor$p %>% t() ))
  pearson_cor_mat <- cbind(pearson_cor_r, pearson_cor_p) %>%
    rownames_to_column(var = "AptName") %>%
    left_join(analytes, by = "AptName") %>%
    column_to_rownames("Target_nodup")
  return(pearson_cor_mat)
}
```

#### mGFR (raw plasma)

```{r echo = F}
# Find significant lm results w/ gfr_raw_plasma
# tbl_lm_top <- run_lm("gfr_raw_plasma")

# Pearson correlations
pearson_cor_mat <- run_corr("gfr_raw_plasma", analytes$AptName) %>%
  dplyr::select(gfr_raw_plasma, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(gfr_raw_plasma < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  # filter(p.value < 0.05) %>%
  arrange(desc(abs(gfr_raw_plasma)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(gfr_raw_plasma < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(gfr_raw_plasma > 0)  %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

mgfr_plot <- plotVolcano_mod(data = pearson_cor_mat,
                             FC = gfr_raw_plasma,
                             p.value = p.value,
                             labels = Target,
                             identify = F,
                             identify_manual = top_10_correlations,
                             fc.cutoff = 0,
                             cutoff = 0.05,
                             x.lab = "Correlation",
                             main = "mGFR",
                             pt.size = 1.5,
                             text.size = 5,
                             fc_lab = "p > 0.05",
                             fc_lab_col = "#dad7cd",
                             sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                             sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/gfr_cor_plot.jpeg"), mgfr_plot,
      width = 12, height = 10, bg = "transparent")
```

#### ERPF (raw plasma)

```{r echo = F}
# Find significant lm results w/ erpf_raw_plasma
# tbl_lm_top <- run_lm("erpf_raw_plasma")

# Pearson correlations
pearson_cor_mat <- run_corr("erpf_raw_plasma", analytes$AptName) %>%
  dplyr::select(erpf_raw_plasma, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(erpf_raw_plasma < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  arrange(desc(abs(erpf_raw_plasma)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(erpf_raw_plasma < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(erpf_raw_plasma > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = erpf_raw_plasma,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$erpf_raw_plasma)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
# ggsave(file.path(root_path, "PANTHER/Results/Figures/erpf_cor_plot.jpeg"),
#       width = 12, height = 10)
```

#### eGFR (CKD EPI)

```{r echo = F}
# Find significant lm results w/ eGFR_CKD_epi
# tbl_lm_top <- run_lm("eGFR_CKD_epi")

# Pearson correlations
pearson_cor_mat <- run_corr("eGFR_CKD_epi", analytes$AptName) %>%
  dplyr::select(eGFR_CKD_epi, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(eGFR_CKD_epi < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(eGFR_CKD_epi)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(eGFR_CKD_epi < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(eGFR_CKD_epi > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = eGFR_CKD_epi,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$eGFR_CKD_epi)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/egfr_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### PGLO

```{r echo = F}
# Find significant lm results w/ glomerular_pressure
# tbl_lm_top <- run_lm("glomerular_pressure")

# Pearson correlations
pearson_cor_mat <- run_corr("glomerular_pressure", analytes$AptName) %>%
  dplyr::select(glomerular_pressure, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(glomerular_pressure < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(glomerular_pressure)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(glomerular_pressure < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(glomerular_pressure > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = glomerular_pressure,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$glomerular_pressure)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/pglo_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### RA

```{r echo = F}
# Find significant lm results w/ ra
# tbl_lm_top <- run_lm("ra")

# Pearson correlations
pearson_cor_mat <- run_corr("ra", analytes$AptName) %>%
  dplyr::select(ra, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(ra < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(ra)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(ra < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(ra > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = ra,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$ra)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/ra_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### RVR

```{r echo = F}
# Find significant lm results w/ rvr
# tbl_lm_top <- run_lm("rvr")

# Pearson correlations
pearson_cor_mat <- run_corr("rvr", analytes$AptName) %>%
  dplyr::select(rvr, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(rvr < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(rvr)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(rvr < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(rvr > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = rvr,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$rvr)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/rvr_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### TKV

```{r echo = F}
# Find significant lm results w/ total_kidney_volume_ml
# tbl_lm_top <- run_lm("total_kidney_volume_ml")

# Pearson correlations
pearson_cor_mat <- run_corr("total_kidney_volume_ml", analytes$AptName) %>%
  dplyr::select(total_kidney_volume_ml, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(total_kidney_volume_ml < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(total_kidney_volume_ml)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(total_kidney_volume_ml < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(total_kidney_volume_ml > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = total_kidney_volume_ml,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$total_kidney_volume_ml)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/tkv_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### Ht adjusted TKV

```{r echo = F}
# Find significant lm results w/ ht_adj_tkv
# tbl_lm_top <- run_lm("ht_adj_tkv")

# Pearson correlations
pearson_cor_mat <- run_corr("ht_adj_tkv", analytes$AptName) %>%
  dplyr::select(ht_adj_tkv, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(ht_adj_tkv < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(ht_adj_tkv)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(ht_adj_tkv < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(ht_adj_tkv > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = ht_adj_tkv,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$ht_adj_tkv)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/httkv_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### Avg ASL

```{r echo = F}
# Find significant lm results w/ avg_pcascl
# tbl_lm_top <- run_lm("avg_pcascl")

# Pearson correlations
pearson_cor_mat <- run_corr("avg_pcascl", analytes$AptName) %>%
  dplyr::select(avg_pcascl, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(avg_pcascl < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  # filter(p.value < 0.05) %>%
  arrange(desc(abs(avg_pcascl)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(avg_pcascl < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(avg_pcascl > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = avg_pcascl,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$avg_pcascl)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/asl_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### Avg cortical ADC

```{r echo = F}
# Find significant lm results w/ avg_c_adc
# tbl_lm_top <- run_lm("avg_c_adc")

# Pearson correlations
pearson_cor_mat <- run_corr("avg_c_adc", analytes$AptName) %>%
  dplyr::select(avg_c_adc, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(avg_c_adc < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(avg_c_adc)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(avg_c_adc < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(avg_c_adc > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = avg_c_adc,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$avg_c_adc)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/adc_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### Minmod SI

```{r echo = F}
# Find significant lm results w/ mm_si
# tbl_lm_top <- run_lm("mm_si")

# Pearson correlations
pearson_cor_mat <- run_corr("mm_si", analytes$AptName) %>%
  dplyr::select(mm_si, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(mm_si < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(mm_si)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(mm_si < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(mm_si > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = mm_si,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$mm_si)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/si_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### DEXA fat (kg)

```{r echo = F}
# Find significant lm results w/ dexa_fat_kg
# tbl_lm_top <- run_lm("dexa_fat_kg")

# Pearson correlations
pearson_cor_mat <- run_corr("dexa_fat_kg", analytes$AptName) %>%
  dplyr::select(dexa_fat_kg, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(dexa_fat_kg < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(dexa_fat_kg)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(dexa_fat_kg < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(dexa_fat_kg > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = dexa_fat_kg,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$dexa_fat_kg)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/fatkg_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### DEXA trunk mass (%)

```{r echo = F}
# Find significant lm results w/ dexa_trunk_mass
# tbl_lm_top <- run_lm("dexa_trunk_mass")

# Pearson correlations
pearson_cor_mat <- run_corr("dexa_trunk_mass", analytes$AptName) %>%
  dplyr::select(dexa_trunk_mass, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(dexa_trunk_mass < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(dexa_trunk_mass)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(dexa_trunk_mass < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(dexa_trunk_mass > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = dexa_trunk_mass,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$dexa_trunk_mass)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/trunkmass_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### DEXA BMD

```{r echo = F}
# Find significant lm results w/ dexa_bone_mineral_density
# tbl_lm_top <- run_lm("dexa_bone_mineral_density")

# Pearson correlations
pearson_cor_mat <- run_corr("dexa_bone_mineral_density", analytes$AptName) %>%
  dplyr::select(dexa_bone_mineral_density, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(dexa_bone_mineral_density < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(dexa_bone_mineral_density)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(dexa_bone_mineral_density < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(dexa_bone_mineral_density > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = dexa_bone_mineral_density,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$dexa_bone_mineral_density)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/bmd_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### Weight

```{r echo = F}
# Find significant lm results w/ weight
# tbl_lm_top <- run_lm("weight")

# Pearson correlations
pearson_cor_mat <- run_corr("weight", analytes$AptName) %>%
  dplyr::select(weight, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(weight < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  #filter(p.value < 0.05) %>%
  arrange(desc(abs(weight)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(weight < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(weight > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = weight,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$weight)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/weight_cor_plot.jpeg"),
       width = 12, height = 10)
```

#### Height

```{r echo = F}
# Find significant lm results w/ height
# tbl_lm_top <- run_lm("height")

# Pearson correlations
pearson_cor_mat <- run_corr("height", analytes$AptName) %>%
  dplyr::select(height, p.value, Target) %>%
  dplyr::mutate(direction = ifelse(height < 0, "Negative", "Positive"),
                Target_nodup = row.names(.))  %>%
  # filter(p.value < 0.05) %>%
  arrange(desc(abs(height)))

negative_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(height < 0) %>%
  distinct(Target, .keep_all = T)
positive_correlations <- arrange(pearson_cor_mat, p.value) %>%
  filter(p.value < 0.05) %>%
  filter(height > 0) %>%
  distinct(Target, .keep_all = T)
top_10_correlations <- c(negative_correlations$Target[1:10],
                         positive_correlations$Target[1:10])

plotVolcano_mod(data = pearson_cor_mat,
                FC = height,
                p.value = p.value,
                labels = Target,
                identify = F,
                identify_manual = top_10_correlations,
                fc.cutoff = 0,
                cutoff = 0.05,
                x.lab = "Correlation",
                main = paste0("Pearson correlations for ", label(panther$height)),
                pt.size = 1.5,
                text.size = 5,
                fc_lab = "p > 0.05",
                fc_lab_col = "#dad7cd",
                sig_fc_pos_lab = "p < 0.05 (Positive correlation)",
                sig_fc_neg_lab = "p < 0.05 (Negative correlation)")
ggsave(file.path(root_path, "PANTHER/Results/Figures/height_cor_plot.jpeg"),
       width = 12, height = 10)
```

## Pathway analysis
```{r echo = F}
# write.csv(res_risk_save, file.path(root_path, "PANTHER/Data_Cleaned/ipa/panther_risk_dep.csv"), row.names = F)
# write.csv(res_ts_save, file.path(root_path, "PANTHER/Data_Cleaned/ipa/panther_ts_dep.csv"), row.names = F)
# write.csv(res_sex_save, file.path(root_path, "PANTHER/Data_Cleaned/ipa/panther_sex_dep.csv"), row.names = F)

risk_pathway <- read_excel(file.path(root_path, "PANTHER/results/IPA/panther_risk_dep_pathway.xls"), skip = 1)
sex_pathway <- read_excel(file.path(root_path, "PANTHER/results/IPA/panther_sex_dep_pathway.xls"), skip = 1)
ts_pathway <- read_excel(file.path(root_path, "PANTHER/results/IPA/panther_ts_dep_pathway.xls"), skip = 1)

colnames(risk_pathway) <- c("pathway", "neg_log_p", "ratio", "z", "genes")
colnames(sex_pathway) <- c("pathway", "neg_log_p", "ratio", "z", "genes")
colnames(ts_pathway) <- c("pathway", "neg_log_p", "ratio", "z", "genes")

text1 = 6.5
text2 = 18
text3 = 20
```

### Risk DEP pathways
#### Negative Pathways
```{r echo = F}
pathways_negative <- risk_pathway %>%
  filter(z < 0) %>%
  arrange(desc(neg_log_p))

pathways_negative$pathway <- reorder(pathways_negative$pathway, pathways_negative$neg_log_p)

top50_negative <- pathways_negative[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = abs(z)), alpha = 0.8, color = "#2c7da0") +
  scale_size_continuous(range = c(0,10),labels = function(x) paste0("-", x)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#2c7da0", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Inhibited Pathways")   +
  scale_x_continuous(limits = c(0, 10), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))
n_pathways_negative <- pathways_negative %>%
  filter(neg_log_p > -log10(0.05)) %>%
  nrow()

top50_negative

ggsave(file.path(root_path, "PANTHER/results/Figures/top50_risk_negative_pathways.jpeg"),
       width = 27.5, height = 14, scale = 1)

########## Limited to z < -2
pathways_negative_sig <- pathways_negative %>%
  filter(z < -2)
top50_negative <- pathways_negative_sig[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = abs(z)), alpha = 0.8, color = "#2c7da0") +
  scale_size_continuous(range = c(0,10),labels = function(x) paste0("-", x)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#2c7da0", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Inhibited Pathways")   +
  scale_x_continuous(limits = c(0, 9), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 0.2))

top50_negative
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_risk_negative_pathways_2.jpeg"),
       width = 27.5, height = 14, scale = 1)
```

#### Positive Pathways
```{r echo = F}
pathways_positive <- risk_pathway %>%
  filter(z > 0) %>%
  arrange(desc(neg_log_p))

pathways_positive$pathway <- reorder(pathways_positive$pathway, pathways_positive$neg_log_p)

top50_positive <- pathways_positive[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = z), color = "#c75146", alpha = 0.8) +
  scale_size_continuous(range = c(0,10)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#c75146", hjust = 0, size = text1, nudge_x = 0.05) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       color = "Scenario",
       size = "Z-score",
       title = "Top 50 Activated Pathways")   +
  scale_x_continuous(limits = c(0, 6), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

n_pathways_positive <- pathways_positive %>%
  filter(neg_log_p > -log10(0.05)) %>%
  nrow()

top50_positive
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_risk_positive_pathways.jpeg"),
       width = 27.5, height = 14, scale = 1)

########## Limited to z > 2
pathways_positive_sig <- pathways_positive %>%
  filter(z > 2)
top50_positive <- pathways_positive_sig[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = z), color = "#c75146", alpha = 0.8) +
  scale_size(range = c(0,10)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#c75146", hjust = 0, size = text1, nudge_x = 0.05) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Activated Pathways") +
  scale_x_continuous(limits = c(0, 4), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

top50_positive
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_risk_positive_pathways_2.jpeg"),
       width = 27.5, height = 14, scale = 1)
```
### Sex DEP pathways
#### Negative Pathways
```{r echo = F}
pathways_negative <- sex_pathway %>%
  filter(z < 0) %>%
  arrange(desc(neg_log_p))

pathways_negative$pathway <- reorder(pathways_negative$pathway, pathways_negative$neg_log_p)

top50_negative <- pathways_negative[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = abs(z)), alpha = 0.8, color = "#2c7da0") +
  scale_size_continuous(range = c(0,10),labels = function(x) paste0("-", x)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#2c7da0", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Inhibited Pathways")   +
  scale_x_continuous(limits = c(0, 10), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))
n_pathways_negative <- pathways_negative %>%
  filter(neg_log_p > -log10(0.05)) %>%
  nrow()

top50_negative

ggsave(file.path(root_path, "PANTHER/results/Figures/top50_sex_negative_pathways.jpeg"),
       width = 27.5, height = 14, scale = 1)

########## Limited to z < -2
pathways_negative_sig <- pathways_negative %>%
  filter(z < -2)
top50_negative <- pathways_negative_sig[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = abs(z)), alpha = 0.8, color = "#2c7da0") +
  scale_size_continuous(range = c(0,10),labels = function(x) paste0("-", x)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#2c7da0", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Inhibited Pathways")   +
  scale_x_continuous(limits = c(0, 9), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

top50_negative
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_sex_negative_pathways_2.jpeg"),
       width = 27.5, height = 14, scale = 1)
```

#### Positive Pathways
```{r echo = F}
pathways_positive <- sex_pathway %>%
  filter(z > 0) %>%
  arrange(desc(neg_log_p))

pathways_positive$pathway <- reorder(pathways_positive$pathway, pathways_positive$neg_log_p)

top50_positive <- pathways_positive[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = z), color = "#c75146", alpha = 0.8) +
  scale_size_continuous(range = c(0,10)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#c75146", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       color = "Scenario",
       size = "Z-score",
       title = "Top 50 Activated Pathways")   +
  scale_x_continuous(limits = c(0, 7), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

n_pathways_positive <- pathways_positive %>%
  filter(neg_log_p > -log10(0.05)) %>%
  nrow()

top50_positive
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_sex_positive_pathways.jpeg"),
       width = 27.5, height = 14, scale = 1)

########## Limited to z > 2
pathways_positive_sig <- pathways_positive %>%
  filter(z > 2)
top50_positive <- pathways_positive_sig[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = z), color = "#c75146", alpha = 0.8) +
  scale_size(range = c(0,10)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#c75146", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Activated Pathways") +
  scale_x_continuous(limits = c(0, 5), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

top50_positive
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_sex_negative_pathways_2.jpeg"),
       width = 27.5, height = 14, scale = 1)
```

### Tanner stage DEP pathways
#### Negative Pathways
```{r echo = F}
pathways_negative <- ts_pathway %>%
  filter(z < 0) %>%
  arrange(desc(neg_log_p))

pathways_negative$pathway <- reorder(pathways_negative$pathway, pathways_negative$neg_log_p)

top50_negative <- pathways_negative[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = abs(z)), alpha = 0.8, color = "#2c7da0") +
  scale_size_continuous(range = c(0,10),labels = function(x) paste0("-", x)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#2c7da0", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Inhibited Pathways")   +
  scale_x_continuous(limits = c(0, 6), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))
n_pathways_negative <- pathways_negative %>%
  filter(neg_log_p > -log10(0.05)) %>%
  nrow()

top50_negative

ggsave(file.path(root_path, "PANTHER/results/Figures/top50_ts_negative_pathways.jpeg"),
       width = 27.5, height = 14, scale = 1)

########## Limited to z < -2
pathways_negative_sig <- pathways_negative %>%
  filter(z < -2)
top50_negative <- pathways_negative_sig[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = abs(z)), alpha = 0.8, color = "#2c7da0") +
  scale_size_continuous(range = c(0,10),labels = function(x) paste0("-", x)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#2c7da0", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Inhibited Pathways")   +
  scale_x_continuous(limits = c(0, 6), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

top50_negative
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_ts_negative_pathways_2.jpeg"),
       width = 27.5, height = 14, scale = 1)
```

#### Positive Pathways
```{r echo = F}
pathways_positive <- ts_pathway %>%
  filter(z > 0) %>%
  arrange(desc(neg_log_p))

pathways_positive$pathway <- reorder(pathways_positive$pathway, pathways_positive$neg_log_p)

top50_positive <- pathways_positive[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = z), color = "#c75146", alpha = 0.8) +
  scale_size_continuous(range = c(0,10)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#c75146", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       color = "Scenario",
       size = "Z-score",
       title = "Top 50 Activated Pathways")   +
  scale_x_continuous(limits = c(0, 7), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

n_pathways_positive <- pathways_positive %>%
  filter(neg_log_p > -log10(0.05)) %>%
  nrow()

top50_positive
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_ts_positive_pathways.jpeg"),
       width = 27.5, height = 14, scale = 1)

########## Limited to z > 2
pathways_positive_sig <- pathways_positive %>%
  filter(z > 2)
top50_positive <- pathways_positive_sig[1:50,] %>%
  ggplot(aes(x = neg_log_p, y = pathway, label = pathway)) +
  geom_point(aes(size = z), color = "#c75146", alpha = 0.8) +
  scale_size(range = c(0,10)) +
  geom_vline(xintercept = -log10(0.05), linetype = "dashed") +
  geom_text(aes(group_2 = pathway), color = "#c75146", hjust = 0, size = text1, nudge_x = 0.1) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = text3),
        axis.title = element_text(size = text3),
        axis.ticks.y = element_blank(),
        legend.position = c(0.9,0.2),
        legend.background = element_blank(),
        legend.box.background = element_rect(color = "black"),
        legend.title = element_text(size = text2),
        legend.text = element_text(size = text2),
        title = element_text (size = text3)) +
  labs(x = "-log(p-value)",
       y = "Pathways",
       size = "Z-score",
       title = "Top 50 Activated Pathways") +
  scale_x_continuous(limits = c(0, 4), expand = c(0, 0)) +
  scale_y_discrete(expand = expansion(add = 1))

top50_positive
ggsave(file.path(root_path, "PANTHER/results/Figures/top50_ts_positive_pathways_2.jpeg"),
       width = 27.5, height = 14, scale = 1)
```
