---
title: "PANTHER - preliminary data for renewal"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r libraries, echo=F, include = F}
library(SomaDataIO)
library(tidyverse)
library(growthcleanr)
library(arsenal)
library(dplyr)
library(naniar)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(pedbp)
library(psych)
library(EnhancedVolcano)
library(gt)
library(SomaPlotr)
library(readxl)
library(limma)
library(glue)
library(purrr)
library(stringr)
library(nlme)
library(emmeans)
library(simr)
library(effectsize)
library(lme4)

options(scipen=999)

# specify user for paths
user <- Sys.info()[["user"]]
if (user == "laurapyle") {
  data_path <- "/Users/laurapyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Data Harmonization/Data Clean"
  git_path <- "/Users/laurapyle/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  panther_path <- "/Users/laurapyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/PANTHER/Data_Cleaned"
} else if (user == "lpyle") {
  data_path <- "/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Data Harmonization/Data Clean"
  git_path <- "/Users/lpyle/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  panther_path <- "/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/PANTHER/Data_Cleaned"
} else if (user == "pylell") {
  data_path <- "/Users/pylell/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Data Harmonization/Data Clean"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  panther_path <- "/Users/pylell/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/PANTHER/Data_Cleaned"
} else {
  stop("Unknown user: please specify root path for this user.")
}


```

```{r load-data, echo = F, include = F}
# Load dictionary function and file
filename <- "/Resources/YC/R Functions/label_harmonized_function.R"
full_path <- file.path(git_path, filename)
source(full_path)
filename <- "/Resources/YC/R Functions/correlation_function.R"
full_path <- file.path(git_path, filename)
source(full_path)
```

```{r prep-data, echo = F, include = F}
# Load clinical data
filename <- "/harmonized_dataset.csv"
full_path <- file.path(data_path, filename)
harm_dat <- read.csv(full_path, na.strings = "")
filename <- "/panther_withdrew_ltfu_list.csv"
full_path <- file.path(panther_path, filename)
exclude <- read.csv(full_path) %>%
  filter(category == "SF") %>%
  pull(record_id) # Remove screen fails only (not lost to follow up)

dat <- harm_dat %>% filter(study == "PANTHER") %>%
  mutate(visit = case_when(visit == "screening" ~ "baseline", T ~ visit),
         mrn = case_when(study != "PANTHER" ~ paste0(as.character(mrn), as.character(date_of_screen)), T~ as.character(mrn))) %>%
  filter(record_id %nin% exclude) %>% 
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                    across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                    .by = c(mrn, visit)) %>%
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" &
                                                ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                              race == "Black or African American" &
                                                ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                              ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                              T ~ "Not Hispanic (or Latino Other"),
         tanner_stage_comp = coalesce(tan_fgd, tan_fph, tan_tveq, tan_mgd, tan_mph, breast_tanner),
         tanner_stage_comp_panther = case_when(tanner_stage_comp > 3 ~ 4, T~ tanner_stage_comp),
         tanner_stage_cat = case_when(tanner_stage_comp <3 ~ "I_II",
                                      tanner_stage_comp >=3 ~ "III_V"),
         tanner_stage_cat_2 = case_when(tanner_stage_comp == 1 ~ "I",
                                        tanner_stage_comp <4 ~ "II_III",
                                        tanner_stage_comp >=4 ~ "IV_V"),
         tanner_stage_cat_3 = case_when(tanner_stage_comp < 2 ~ "I_II",
                                        tanner_stage_comp == 3 ~ "III",
                                        tanner_stage_comp >= 4 ~ "IV_V"),
         tanner_stage_comp_panther_cat =
                  case_when(tanner_stage_comp_panther == 4 ~ "4 & 5",
                            T ~ as.character(tanner_stage_comp_panther)),
         age_mo = (age * 12),
         sex = case_when(sex == "Male" ~ "male",
                         sex == "Female" ~ "female"),
         male_ind = case_when(sex == "male" ~ 1, sex == "female" ~ 0),
         combined_tkv = coalesce(total_kidney_volume_ml, total_kidney_volume_ml_manual),
         combined_ht_tkv = coalesce(ht_adj_tkv, ht_adj_tkv_manual),
         log_tot_test = log(tot_test),
         log_free_test = log(free_test),
         log_lh = log(lh),
         log_estrad = log(estrad),
         log_acr_u = log(acr_u),
         avg_c_adc_tkv = avg_c_adc/combined_tkv,
         avg_c_r2_tkv = avg_c_r2/combined_tkv,
         avg_pcascl_tkv = avg_pcascl/combined_tkv,
         avg_c_t1_tkv = avg_c_t1/combined_tkv,
         mm_airg = case_when(mm_airg > 0 ~ mm_airg),
         mm_bcell = case_when(mm_bcell > 0 ~ mm_bcell),
         mm_di = case_when(mm_di > 0 & mm_di < 10000 ~ mm_di),
         mm_ir = case_when(mm_ir > 0 ~ mm_ir),
         mm_si = case_when(mm_si > 0 & mm_si < 500 ~ mm_si),
         birthweight = gsub(" kg", "", birthweight),
         birthweight = as.numeric(birthweight),
         glp1_agonist_timepoint = case_when(is.na(glp1_agonist_timepoint) ~ "No", T ~ glp1_agonist_timepoint),
         uacr_cat = case_when(acr_u <= 10 ~ "UACR [0, 10]",
                              acr_u <= 20 ~ "UACR (10, 20]",
                              acr_u <= 30 ~ "UACR (20, 30]",
                              acr_u > 30 ~ "UACR > 30")) %>%
  arrange(record_id) %>%
  # filter(!is.na(dexa_body_fat)) %>% # filter based on DEXA (to filter those who may just be screen fails)
  filter(!is.na(male_ind)) 
dat$tanner_stage_comp <- as.factor(dat$tanner_stage_comp)
bp_percentile = p_bp(q_sbp = dat$sbp, q_dbp = dat$dbp, age = dat$age_mo, male = dat$male_ind)
dat$sbp_p <- bp_percentile$sbp_p
dat$dbp_p <- bp_percentile$dbp_p
bmi_percentile = ext_bmiz(data = subset(dat,
                                        select = c("mrn", "visit", "sex", "age_mo", "weight", "height", "bmi")),
                          age = "age_mo",
                          wt = "weight",
                          ht = "height",
                          bmi = "bmi",
                          adjust.integer.age = F) %>%
  dplyr:: select(mrn, visit, bmip, bmiz) 
dat <- left_join(dat, bmi_percentile, by = c("mrn", "visit")) %>%
  dplyr::mutate(group_risk = case_when(bmip >= 95 | hba1c >=6 | group == "Type 2 Diabetes" ~ "High",
                                       T ~ "Low"),
                sex_group_risk = case_when(sex == "male" & group_risk == "High" ~ "M, High",
                                           sex == "male" & group_risk == "Low" ~ "M, Low",
                                           sex == "female" & group_risk == "High" ~ "F, High",
                                           sex == "female" & group_risk == "Low" ~ "F, Low"))
dat$tanner_combined <- as.factor(ifelse(dat$tanner_stage_comp %in% c(4,5), "4/5", dat$tanner_stage_comp))

# separate datasets by visit for descriptive statistics
base <- dat %>% filter(visit == "baseline")
yr1 <- dat %>% filter(visit == "year_1")
yr2 <- dat %>% filter(visit == "year_2")
  
# create a variable to indicate cohort (i.e., the TS at which they were enrolled)
cohortdf <- base %>% select(mrn, tanner_stage_comp)
cohortdf$cohort <- paste0("Cohort", cohortdf$tanner_stage_comp)
cohortdf <- cohortdf %>% select(mrn, cohort)
dat <- merge(dat, cohortdf, by = "mrn")

# Remove people TS5 at baseline
dat <- dat %>% filter(!tanner_stage_comp == "5")
```


```{r echo=FALSE, include=FALSE}
mixed <- function(data, outcome, grouping_var){
  data <- as.data.frame(data)
  data$time <- as.factor(data$time)
  form = as.formula(paste0(outcome,"~ ", grouping_var, "*time"))
  mod <- lme(as.formula(form),random=~1|mrn,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,c(grouping_var,"time"))
  prs <-  pairs(m,adjust="tukey",by="time")
  return(list(anova,m,prs))
}

mixed_adj <- function(data, outcome, grouping_var){
  data <- as.data.frame(data)
  data$time <- as.factor(data$time)
  form = as.formula(paste0(outcome,"~ ", grouping_var, "*time + group_risk"))
  mod <- lme(as.formula(form),random=~1|mrn,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,c(grouping_var,"time"))
  prs <-  pairs(m,adjust="tukey",by="time")
  return(list(anova,m,prs))
}

plot_mixed_group_risk <- function(mixed_output, data, ylab) {
  emmeans_result <- mixed_output[[2]]
  emmeans_result <- as.data.frame(emmeans_result)
  emmeans_result$group <- ifelse(emmeans_result$group_risk == "High", "High", "Low")
  # Create line plot
  p <-  ggplot(emmeans_result, aes(x = time, y = emmean, color = group, group = group, 
                                  line = group)) +
    geom_line() +
    geom_point() +
    labs(x = "Time", y = ylab, color = "group") +
    theme_bw() + 
    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), linetype = 1, alpha = 0.1) +
    theme(legend.title=element_blank()) + geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL))
}

plot_mixed_tanner <- function(mixed_output, data, ylab) {
  emmeans_result <- mixed_output[[2]]
  emmeans_result <- as.data.frame(emmeans_result)
  emmeans_result$group <- emmeans_result$tanner_combined
  # Create line plot
  p <-  ggplot(emmeans_result, aes(x = time, y = emmean, color = group, group = group, 
                                  line = group)) +
    geom_line() +
    geom_point() +
    labs(x = "Time", y = ylab, color = "group") +
    theme_bw() + 
    #geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), linetype = 1, alpha = 0.1) +
    theme(legend.title=element_blank()) + geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL))
}

```

# Descriptive statistics

## Baseline

```{r results = 'asis'}
summary(arsenal::tableby(tanner_stage_comp ~ age + bmiz + sex + group_risk + albuminuria_cat + 
                           kwt(acr_u, "medianq1q3") + total_kidney_volume_ml + mm_si + mm_bcell + 
                           gfr_bsa_plasma + glomerular_pressure + hba1c, data = base),
        digits = 1)
```

## Year 1

```{r results = 'asis'}
summary(arsenal::tableby(tanner_stage_comp ~ age + bmiz + sex + group_risk + albuminuria_cat + 
                           kwt(acr_u, "medianq1q3") + total_kidney_volume_ml + 
                           gfr_bsa_plasma + glomerular_pressure , data = yr1),
        digits = 1)
```

## Year 2

```{r results = 'asis'}
summary(arsenal::tableby(tanner_stage_comp ~ age + bmiz + sex + group_risk + albuminuria_cat + 
                           kwt(acr_u, "medianq1q3") + total_kidney_volume_ml + 
                           gfr_bsa_plasma + glomerular_pressure , data = yr2),
        digits = 1)
```

# Mixed-effects models

## Glomerular pressure

```{r}
mod <- lme(glomerular_pressure ~ tanner_stage_comp + cohort,random=~1|mrn,data = dat,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
anova <- anova.lme(mod, type="marginal")
m <- emmeans(mod,"tanner_stage_comp")
prs <-  pairs(m,adjust="tukey")
```

```{r, fig.width=8}
 p <-  ggplot(dat, aes(x = as.numeric(tanner_stage_comp), y = glomerular_pressure, color = cohort, group = cohort)) +
    geom_point() +
    labs(x = "TS", color = "cohort") +
    theme_bw() 
p
```

### ANOVA table

```{r}
anova
```

### Estimated marginal means

```{r}
m
```

### Pairwise comparisons

```{r echo=FALSE}
prs
```

## GFR

### With cohort effect

```{r}
mod <- lme(gfr_bsa_plasma ~ tanner_stage_comp + cohort,random=~1|mrn,data = dat,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
anova <- anova.lme(mod, type="marginal")
m <- emmeans(mod,"tanner_stage_comp")
prs <-  pairs(m,adjust="tukey")
```

```{r, fig.width=8}
 p <-  ggplot(dat, aes(x = as.numeric(tanner_stage_comp), y = gfr_bsa_plasma, color = cohort, group = cohort)) +
    geom_point() +
    labs(x = "TS", color = "cohort") +
    theme_bw() 
p
```

#### ANOVA table

```{r}
anova
```

#### Estimated marginal means

```{r}
m
```

#### Pairwise comparisons

```{r}
prs
```

### Without cohort effect

```{r}
mod <- lme(gfr_bsa_plasma ~ tanner_stage_comp ,random=~1|mrn,data = dat,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
anova <- anova.lme(mod, type="marginal")
m <- emmeans(mod,"tanner_stage_comp")
prs <-  pairs(m,adjust="tukey")
```

```{r, fig.width=8}
 p <-  ggplot(dat, aes(x = as.numeric(tanner_stage_comp), y = gfr_bsa_plasma, color = cohort, group = cohort)) +
    geom_point() +
    labs(x = "TS", color = "cohort") +
    theme_bw() 
p
```

#### ANOVA table

```{r}
anova
```

#### Estimated marginal means

```{r}
m
```

#### Pairwise comparisons

```{r}
prs
```

## Left T1 Kidney

### With cohort effect

```{r}
mod <- lme(bold_l_t1_kidney ~ tanner_stage_comp + cohort,random=~1|mrn,data = dat,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
anova <- anova.lme(mod, type="marginal")
m <- emmeans(mod,"tanner_stage_comp")
prs <-  pairs(m,adjust="tukey")
```

```{r, fig.width=8}
 p <-  ggplot(dat, aes(x = as.numeric(tanner_stage_comp), y = bold_l_t1_kidney, color = cohort, group = cohort)) +
    geom_point() +
    labs(x = "TS", color = "cohort") +
    theme_bw() 
p
```

#### ANOVA table

```{r}
anova
```

#### Estimated marginal means

```{r}
m
```

#### Pairwise comparisons

```{r}
prs
```

### Without cohort effect

```{r}
mod <- lme(bold_l_t1_kidney ~ tanner_stage_comp ,random=~1|mrn,data = dat,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
anova <- anova.lme(mod, type="marginal")
m <- emmeans(mod,"tanner_stage_comp")
prs <-  pairs(m,adjust="tukey")
```

```{r, fig.width=8}
 p <-  ggplot(dat, aes(x = as.numeric(tanner_stage_comp), y = bold_l_t1_kidney, color = cohort, group = cohort)) +
    geom_point() +
    labs(x = "TS", color = "cohort") +
    theme_bw() 
p
```

#### ANOVA table

```{r}
anova
```

#### Estimated marginal means

```{r}
m
```

#### Pairwise comparisons

```{r}
prs
```

## Right T1 Kidney

### With cohort effect

```{r}
mod <- lme(bold_r_t1_kidney ~ tanner_stage_comp + cohort,random=~1|mrn,data = dat,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
anova <- anova.lme(mod, type="marginal")
m <- emmeans(mod,"tanner_stage_comp")
prs <-  pairs(m,adjust="tukey")
```

```{r, fig.width=8}
 p <-  ggplot(dat, aes(x = as.numeric(tanner_stage_comp), y = bold_r_t1_kidney, color = cohort, group = cohort)) +
    geom_point() +
    labs(x = "TS", color = "cohort") +
    theme_bw() 
p
```

#### ANOVA table

```{r}
anova
```

#### Estimated marginal means

```{r}
m
```

#### Pairwise comparisons

```{r }
prs
```

### Without cohort effect

```{r}
mod <- lme(bold_r_t1_kidney ~ tanner_stage_comp ,random=~1|mrn,data = dat,na.action = na.omit, correlation = corCompSymm(form = ~1|mrn))
anova <- anova.lme(mod, type="marginal")
m <- emmeans(mod,"tanner_stage_comp")
prs <-  pairs(m,adjust="tukey")
```

```{r, fig.width=8}
p <-  ggplot(dat, aes(x = as.numeric(tanner_stage_comp), y = bold_r_t1_kidney, color = cohort, group = cohort)) +
    geom_point() +
    labs(x = "TS", color = "cohort") +
    theme_bw() 
p
```

#### ANOVA table

```{r echo=FALSE}
anova
```

#### Estimated marginal means

```{r}
m
```

#### Pairwise comparisons

```{r}
prs
```

# Power calculations

## GFR

### Using simr on pilot data - this is not working

```{r}
mod <- lmer(gfr_bsa_plasma ~ as.numeric(tanner_stage_comp) + (1|mrn), 
            data = dat, 
            na.action = na.omit,
            REML = FALSE)
doTest(mod, fixed("as.numeric(tanner_stage_comp)"))
fixef(mod)["as.numeric(tanner_stage_comp)"] <- -10
mod1 <- extend(mod, along = "mrn", n = 90, 
               values = list(tanner_stage_comp = 1:5))
doTest(mod1, fixed("as.numeric(tanner_stage_comp)"))
powerSim(mod1, nsim=50, test = "as.numeric(tanner_stage_comp)")
powerCurve(mod1, within = "mrn", breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130), 
           test = fixed("as.numeric(tanner_stage_comp)"))
```

### Using simr with makeLmer() to specify data

```{r}

# time, MRN, and variable being tested from PANTHER (Z)
years <- 1:5
set.seed(123)  # for reproducibility
mrn <- replicate(90, paste0(sample(c(letters, LETTERS, 0:9), 8, replace = TRUE), collapse = ""))
while(length(unique(mrn)) < 90) {
  mrn <- replicate(90, paste0(sample(c(letters, LETTERS, 0:9), 8, replace = TRUE), collapse = ""))
}
mrn <- unique(mrn)[1:90]
X <- expand.grid(years, mrn)
temp <- rnorm(90)
Z <- NULL
for (i in 1:90) {
  c <- rep(temp[i],5)
  Z <- c(Z, c)
}
X <- cbind(X, Z)
colnames(X) <- c("years", "mrn", "Z")

# specify parameters
# fixed intercept, slope over time, association with Z
b <- c(135, -2, 0.5)
# random intercept variance
V1 <- 2
# residual standard deviation
s <- 5

model1 <- makeLmer(y ~ years + Z + (1|mrn), data = X, fixef = b, VarCorr = V1, sigma = s)
dat1 <- getData(model1)

p <-  ggplot(dat1, aes(x = years, y = y, color = Z)) +
    geom_point() +
    labs(x = "year", color = "Z") +
    theme_bw() 
p

powerSim(model1, nsim = 50)

# function
sim_power <- function(n = 90, intercept = 135, beta_time = -2, beta_pred = 0.5, var_int = 2, resid_sd = 5) {
  # time, MRN, and variable being tested from PANTHER (Z)
  years <- 1:5
  set.seed(123)  # for reproducibility
  mrn <- replicate(n, paste0(sample(c(letters, LETTERS, 0:9), 8, replace = TRUE), collapse = ""))
  while(length(unique(mrn)) < n) {
    mrn <- replicate(n, paste0(sample(c(letters, LETTERS, 0:9), 8, replace = TRUE), collapse = ""))
  }
  mrn <- unique(mrn)[1:n]
  X <- expand.grid(years, mrn)
  temp <- rnorm(n)
  Z <- NULL
  for (i in 1:n) {
    c <- rep(temp[i],5)
    Z <- c(Z, c)
  }
  X <- cbind(X, Z)
  colnames(X) <- c("years", "mrn", "Z")
  
  # specify parameters
  # fixed intercept, slope over time, association with Z
  b <- c(intercept, beta_time, beta_pred)
  # random intercept variance
  V1 <- var_int
  # residual standard deviation
  s <- resid_sd
  
  model1 <- makeLmer(y ~ years + Z + (1|mrn), data = X, fixef = b, VarCorr = V1, sigma = s)
  dat1 <- getData(model1)
  
  p <-  ggplot(dat1, aes(x = years, y = y, color = Z)) +
    geom_point() +
    labs(x = "year", color = "Z") +
    theme_bw() 
  return(p)
  
  powerSim(model1, nsim = 50)
}

sim_power(n = 90, intercept = 135, beta_time = -2, beta_pred = 0.5, var_int = 2, resid_sd = 5)
p
```