---
title: "CKM in TODAY integrated Analysis LP/JAPC.qmd"
author: "Laura Pyle / Jairo A Pinzon"
format: html
editor: visual
echo: FALSE
warning: FALSE
toc: TRUE
---

```{r include = F}
# =============================================================================
# LIBRARIES
# =============================================================================
library(dplyr)
library(arsenal)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(ggalluvial)
library(FSA)
library(fuzzyjoin)
library(purrr)
library(stringr)
library(arsenal)
library(Hmisc)
library(SomaDataIO)
library(caret)
library(survival)
library(broom)
library(openxlsx)
library(readxl)
library(knitr)

# Working directory setup
getwd()
#usethis::use_git()

# specify user for paths
user <- Sys.info()[["user"]]
if (user == "laurapyle") {
  data_path <- "/Users/laurapyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  github_path <- "/Users/laurapyle/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "lpyle") {
  data_path <- "/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  github_path <- "/Users/lpyle/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else {
  stop("Unknown user: please specify root path for this user.")
}

filename <- "/TODAY proteomics metabolomics/CKM/CKM functions.R"
full_path <- file.path(github_path, filename)
source(full_path)
```

```{r include = F}
# =============================================================================
# DATA LOADING AND BASIC VARIABLES
# =============================================================================

# Core datasets
#comorb <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/COMORB.csv")
filename <- "/Clinical data/comorb.csv"
full_path <- file.path(data_path, filename)
comorb <- read.csv(full_path)

## Fix that path because the OneDrive is not allowing proper access to it.
#load("/Users/jpcortes/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
#load("/Users/jpcortes/Documents/Temp_files_TODAY/clinical_data_long.Rdata")
filename <- "/Clinical data/TODAY/clinical_data_long.Rdata"
full_path <- file.path(data_path, filename)
load(full_path)

#load("/Users/jpcortes/Documents/Temp_files_TODAY/baserisk.Rdata")
filename <- "/Clinical data/TODAY/baserisk.Rdata"
full_path <- file.path(data_path, filename)
load(file = full_path)

#ame <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/AME.csv")
filename <- "/Clinical data/TODAY2/AME.csv"
full_path <- file.path(data_path, filename)
ame <- read.csv(full_path)

ame_tab <- table(ame$AMENAME, ame$AMETYPE)
ame_pt <- ame %>% select(releaseid, AMENAME, AMETYPE) %>% unique()
ame_pt_tab <- table(ame_pt$AMENAME, ame_pt$AMETYPE)
ametype <- c("Heart", "Vascular", "Cerebrovascular", "Liver", "Nerve", "Renal", "Eye", "Death")
amename <- c("Arrhythmia", "CAD", "CHD", "LVSD", "MI", "PAD", "DVT", "Stroke", "TIA",
             "Pancreatitis", "Gallbladder disease", "Peripheral neuropathy", "Autonomic neuropathy", "Diabetic mononeuropathy", "CKD", "ESKD", "NPDR", 
             "PDR", "Macular edema", "Vitreous hemorrhage", "Blindness", "Cataracts", "Glaucoma", "Death")
ame_tab <- as.data.frame.matrix(ame_tab)
ame_pt_tab <- as.data.frame.matrix(ame_pt_tab)
colnames(ame_tab) <- ametype
rownames(ame_tab) <- amename
colnames(ame_pt_tab) <- ametype
rownames(ame_pt_tab) <- amename
ame_tab <- as.data.frame(ame_tab)
ame_pt_tab <- as.data.frame(ame_pt_tab)
ame_tab <- ame_tab %>%  mutate(Total = rowSums(across(where(is.numeric))))
ame_pt_tab <- ame_pt_tab %>%  mutate(Total = rowSums(across(where(is.numeric))))
ame_tab_col_tot <- colSums(ame_tab)
ame_tab <- rbind(ame_tab, ame_tab_col_tot)
rownames(ame_tab)[nrow(ame_tab)] <- "Total"
ame_pt_tab_col_tot <- colSums(ame_pt_tab)
ame_pt_tab <- rbind(ame_pt_tab, ame_pt_tab_col_tot)
rownames(ame_pt_tab)[nrow(ame_pt_tab)] <- "Total"



#baseline <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/BASELINE.csv")
filename <- "/Clinical data/TODAY/BASELINE.csv"
full_path <- file.path(data_path, filename)
baseline <- read.csv(full_path)

# Echo data
#today_echo <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/ECHO_today.csv")
filename <- "/Clinical data/TODAY/ECHO.csv"
full_path <- file.path(data_path, filename)
today_echo <- read.csv(full_path)
today_echo$TIMEPOINT <- "1"

#today2_echo <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/ECHO_today2.csv")
filename <- "/Clinical data/TODAY2/ECHO.csv"
full_path <- file.path(data_path, filename)
today2_echo <- read.csv(full_path)
today2_echo$TIMEPOINT <- "2"

# Lab data
#cbl_1 <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/CBL_TODAY1.csv")
filename <- "/Clinical data/TODAY/CBL.csv"
full_path <- file.path(data_path, filename)
cbl_1 <- read.csv(full_path)
cbl_1$TIMEPOINT <- "1"
# remove R visits
cbl_1 <- cbl_1 %>% filter(!mvisit == "R")

#cbl_2 <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/CBL_TODAY2.csv")
filename <- "/Clinical data/TODAY2/CBL.csv"
full_path <- file.path(data_path, filename)
cbl_2 <- read.csv(full_path)
cbl_2$TIMEPOINT <- "2"
# rename columns to match TODAY
cbl_2 <- cbl_2 %>% rename(HbA1c = hba1c, LDL = ldl, Chol = chol, hsCRP = hscrp, ALT = alt, AST = ast,
                          FFA = ffa, FIB = fib, Glucose = glucose, HOM = hom, LDLB = ldlb, LDLC = ldlc,
                          PIN = pin, Rf = rf, Trig = trig, ApoB = apob, EstCreatClear = estcreatclear,
                          HDL = hdl, LDLCB = ldlcb, SerumCreat = serumcreat, UAlb = ualb,
                          UAlbCreat = ualbcreat, UCreat = ucreat, VB12 = vb12, VLDL = vldl, 
                          IL6 = il6, PAI1 = pai1)

#addcbl_1 <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/ADDCBL_TODAY1.csv")
filename <- "/Clinical data/TODAY/ADDCBL.csv"
full_path <- file.path(data_path, filename)
addcbl_1 <- read.csv(full_path)
addcbl_1$TIMEPOINT <- "1"
# LP: remove R visits
addcbl_1 <- addcbl_1 %>% filter(!mvisit == "R")

#addcbl_2 <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/ADDCBL_TODAY2.csv")
filename <- "/Clinical data/TODAY2/ADDCBL.csv"
full_path <- file.path(data_path, filename)
addcbl_2 <- read.csv(full_path)
addcbl_2$TIMEPOINT <- "2"
# ONCE WE HAVE CORRECTED DATA FROM DCC, MAY NEED TO RENAME COLUMNS TO MATCH TODAY

# Specialized datasets
#speckle <- read.csv('/Users/jpcortes/Documents/Temp_files_TODAY/SPECKLE.csv')
filename <- "/Clinical data/TODAY2/SPECKLE.csv"
full_path <- file.path(data_path, filename)
speckle <- read.csv(full_path)

#pwv <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/PWV.csv")
filename <- "/Clinical data/TODAY2/PWV.csv"
full_path <- file.path(data_path, filename)
pwv <- read.csv(full_path)

# Quality of life questionnaires
# LP: why are TODAY data not used and merged to echo timepoint?
# TODAY
filename <- "/Clinical data/TODAY/PEDSQLGC.csv"
full_path <- file.path(data_path, filename)
pedsQL_C_today <- read.csv(full_path)
names(pedsQL_C_today) <- toupper(names(pedsQL_C_today))

filename <- "/Clinical data/TODAY/PEDSQLGT.csv"
full_path <- file.path(data_path, filename)
pedsQL_T_today <- read.csv(full_path)
names(pedsQL_T_today) <- toupper(names(pedsQL_T_today))

# TODAY2
# LP: filenames don't match TODAY2 documentation - emailed Laure
#pedsQL_A <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/PEDSQLGA.csv")
filename <- "/Clinical data/TODAY2/PEDSQLGA.csv"
full_path <- file.path(data_path, filename)
pedsQL_A <- read.csv(full_path)

#pedsQL_Y <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/PEDSQLGY.csv")
filename <- "/Clinical data/TODAY2/PEDSQLGY.csv"
full_path <- file.path(data_path, filename)
pedsQL_Y <- read.csv(full_path)

#pedsQL_T <- read.csv("/Users/jpcortes/Documents/Temp_files_TODAY/PEDSQLGT.csv")
filename <- "/Clinical data/TODAY2/PEDSQLGT.csv"
full_path <- file.path(data_path, filename)
pedsQL_T <- read.csv(full_path)

filename <- "/Clinical data/TODAY2/TME.csv"
full_path <- file.path(data_path, filename)
tme <- read.csv(full_path)

# Alternative paths (backup)
#comorb <- read.csv("/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad#/Biostatistics Core Shared Drive/TODAY subaward/Clinical data/COMORB.csv")
#ame <- read.csv('/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad#/Biostatistics Core Shared Drive/TODAY subaward/Clinical data/TODAY2/AME.csv')
#today_echo <- read.csv("/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward/Clinical data/TODAY/echo.csv")
#today2_echo <- read.csv("/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward/Clinical data/TODAY2/echo.csv")
#speckle <- read.csv('/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward/Clinical data/TODAY2/SPECKLE.csv')
#pwv <- read.csv("/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward//Clinical data/TODAY2/pwv.csv")
```

```{r include = F}
# =============================================================================
# DATA PREPARATION AND PROCESSING
# =============================================================================

# AME data processing - create individual event datasets
arrhythmia <- ame %>% filter(AMENAME == 1) %>% select(releaseid, DAYSTOAME)
colnames(arrhythmia) <- c("RELEASEID", "DAYSTOARRHYTHMIA")
arrhythmia$ARRHYTHMIA <- 1
arrhythmia <- arrhythmia %>% arrange(RELEASEID, DAYSTOARRHYTHMIA) %>% group_by(RELEASEID) %>% filter(row_number()==1)

cad <- ame %>% filter(AMENAME == 2) %>% select(releaseid, DAYSTOAME)
colnames(cad) <- c("RELEASEID", "DAYSTOCAD")
cad$CAD <- 1
cad <- cad %>% arrange(RELEASEID, DAYSTOCAD) %>% group_by(RELEASEID) %>% filter(row_number()==1)

chf <- ame %>% filter(AMENAME == 3) %>% select(releaseid, DAYSTOAME)
colnames(chf) <- c("RELEASEID", "DAYSTOCHF")
chf$CHF <- 1
chf <- chf %>% arrange(RELEASEID, DAYSTOCHF) %>% group_by(RELEASEID) %>% filter(row_number()==1)

LVSD <- ame %>% filter(AMENAME == 4) %>% select(releaseid, DAYSTOAME)
colnames(LVSD) <- c("RELEASEID", "DAYSTOLVSD")
LVSD$LVSD <- 1
LVSD <- LVSD %>% arrange(RELEASEID, DAYSTOLVSD) %>% group_by(RELEASEID) %>% filter(row_number()==1)

mi <- ame %>% filter(AMENAME == 5) %>% select(releaseid, DAYSTOAME)
colnames(mi) <- c("RELEASEID", "DAYSTOMI")
mi$MI <- 1
mi <- mi %>% arrange(RELEASEID, DAYSTOMI) %>% group_by(RELEASEID) %>% filter(row_number()==1)

pad <- ame %>% filter(AMENAME == 6) %>% select(releaseid, DAYSTOAME)
colnames(pad) <- c("RELEASEID", "DAYSTOPAD")
pad$PAD <- 1
pad <- pad %>% arrange(RELEASEID, DAYSTOPAD) %>% group_by(RELEASEID) %>% filter(row_number()==1)

dvt <- ame %>% filter(AMENAME == 8) %>% select(releaseid, DAYSTOAME)
colnames(dvt) <- c("RELEASEID", "DAYSTODVT")
dvt$DVT <- 1
dvt <- dvt %>% arrange(RELEASEID, DAYSTODVT) %>% group_by(RELEASEID) %>% filter(row_number()==1)

stroke <- ame %>% filter(AMENAME == 9) %>% select(releaseid, DAYSTOAME)
colnames(stroke) <- c("RELEASEID", "DAYSTOSTROKE")
stroke$STROKE <- 1
stroke <- stroke %>% arrange(RELEASEID, DAYSTOSTROKE) %>% group_by(RELEASEID) %>% filter(row_number()==1)

tia <- ame %>% filter(AMENAME == 11) %>% select(releaseid, DAYSTOAME)
colnames(tia) <- c("RELEASEID", "DAYSTOTIA")
tia$TIA <- 1
tia <- tia %>% arrange(RELEASEID, DAYSTOTIA) %>% group_by(RELEASEID) %>% filter(row_number()==1)

ckd <- ame %>% filter(AMENAME == 17) %>% select(releaseid, DAYSTOAME)
colnames(ckd) <- c("RELEASEID", "DAYSTOCKD")
ckd$CKD <- 1
ckd <- ckd %>% arrange(RELEASEID, DAYSTOCKD) %>% group_by(RELEASEID) %>% filter(row_number()==1)

eskd <- ame %>% filter(AMENAME == 18) %>% select(releaseid, DAYSTOAME)
colnames(eskd) <- c("RELEASEID", "DAYSTOESKD")
eskd$ESKD <- 1
eskd <- eskd %>% arrange(RELEASEID, DAYSTOESKD) %>% group_by(RELEASEID) %>% filter(row_number()==1)

death <- ame %>% filter(AMENAME == 26) %>% select(releaseid, DAYSTOAME)
colnames(death) <- c("RELEASEID", "DAYSTODEATH")
death$DEATH <- 1
death <- death %>% arrange(RELEASEID, DAYSTODEATH) %>% group_by(RELEASEID) %>% filter(row_number()==1)

ulcer <- tme %>% filter(TMETYPE == 5) %>% select(RELEASEID, DAYSTOTME)
colnames(ulcer) <- c("RELEASEID", "DAYSTOULCER")
ulcer$ULCER <- 1
ulcer <- ulcer %>% arrange(RELEASEID, DAYSTOULCER) %>% group_by(RELEASEID) %>% filter(row_number()==1)

# Merge all AME events
ame_summary <- full_join(arrhythmia, cad, by = "RELEASEID")
ame_summary <- full_join(ame_summary, chf, by = "RELEASEID")
ame_summary <- full_join(ame_summary, LVSD, by = "RELEASEID")
ame_summary <- full_join(ame_summary, mi, by = "RELEASEID")
ame_summary <- full_join(ame_summary, pad, by = "RELEASEID")
ame_summary <- full_join(ame_summary, dvt, by = "RELEASEID")
ame_summary <- full_join(ame_summary, stroke, by = "RELEASEID")
ame_summary <- full_join(ame_summary, tia, by = "RELEASEID")
ame_summary <- full_join(ame_summary, ckd, by = "RELEASEID")
ame_summary <- full_join(ame_summary, eskd, by = "RELEASEID")
ame_summary <- full_join(ame_summary, death, by = "RELEASEID")
ame_summary <- full_join(ame_summary, ulcer, by = "RELEASEID")

# Convert NAs to 0 and create factors
ame_summary <- ame_summary %>% 
    mutate_at(c("ARRHYTHMIA", "CAD", "CHF", "LVSD", "MI", "PAD", "DVT", "STROKE", "TIA", "CKD", "ESKD", "DEATH", "ULCER"), ~replace_na(.,0))
ame_summary <- ame_summary %>% 
    mutate_at(c("ARRHYTHMIA", "CAD", "CHF", "LVSD", "MI", "PAD", "DVT", "STROKE", "TIA", "CKD", "ESKD", "DEATH", "ULCER"), ~as.factor(.))

# Clinical CVD composite
ame_summary$clin_cvd <- ifelse(
  rowSums(ame_summary[, c("ARRHYTHMIA", "CAD", "CHF", "MI", "PAD", "DVT", "STROKE", "TIA", "ULCER", "LVSD")] == 1, na.rm = TRUE) > 0,
  1,
  0)

# Merge with comorb data
comorb_ckm <- merge(comorb, ame_summary, by = "RELEASEID", all.x = TRUE)
# LP NOTE: comorb_ckm has NAs, need to set events to 0 and days to event to total follow up time

# Follow-up time calculation
# CHECK THIS
# comorb$fup_time <- ifelse(comorb$HTN == 0, comorb$DAYSTOHTN, 
#                           ifelse(comorb$LDLDLP == 0, comorb$DAYSTOLDL,
#                                  ifelse(comorb$NEURO == 0, comorb$DAYSTONEURO, 
#                                         ifelse(comorb$DNE == 0, comorb$DAYSTODNE, 
#                                                ifelse(comorb$FILAM == 0, comorb$DAYSTOFILAM, 
#                                                      ifelse(comorb$RETINO == 0, comorb$DAYSTORETINO, 
#                                                             ifelse(comorb$TGDLP == 0, comorb$DAYSTOTG, 
#                                                                    ifelse(comorb$NEPHRO == 0, comorb$DAYSTONEPHRO, 
#                                                                           ifelse(comorb$HYP ==0, comorb$DAYSTOHYP, NA)))))))))

for (i in 1:nrow(comorb_ckm)) {
  comorb_ckm$fup_time[i] <- max(comorb_ckm$DAYSTOHTN[i], comorb_ckm$DAYSTOLDL[i], comorb_ckm$DAYSTOLDL[i],
                                comorb_ckm$daystotg[i], comorb_ckm$DAYSTOANYDLP[i], comorb_ckm$DAYSTOMIC[i],
                                comorb_ckm$DAYSTOMAC[i], comorb_ckm$DAYSTONEPHRO[i], comorb_ckm$DAYSTOHYP[i],
                                comorb_ckm$DAYSTORAPID[i], comorb_ckm$DAYSTODNE[i], comorb_ckm$DAYSTOFILAM[i],
                                comorb_ckm$DAYSTONEURO[i], comorb_ckm$DAYSTORETINO[i], comorb_ckm$DAYSTOMVD[i],
                                comorb_ckm$DAYSTOGLYC[i], comorb_ckm$DAYSTOARRHYTHMIA[i], comorb_ckm$DAYSTOCAD[i],
                                comorb_ckm$DAYSTOCHF[i], comorb_ckm$DAYSTOLVSD[i], comorb_ckm$DAYSTOMI[i],
                                comorb_ckm$DAYSTOPAD[i], comorb_ckm$DAYSTODVT[i], comorb_ckm$DAYSTOSTROKE[i],
                                comorb_ckm$DAYSTOTIA[i], comorb_ckm$DAYSTOCKD[i], comorb_ckm$DAYSTOESKD[i],
                                comorb_ckm$DAYSTODEATH[i], comorb_ckm$DAYSTOULCER[i], na.rm = T)
}

# need to update this to process all events
comorb_ckm <- replace_missing_events(comorb_ckm)

# now make a long dataset with number of days and flag 0/1 for event having occurred
# we only have the first event of each so can't look at repeat events
# first comorb events which have a baseline and follow up variable
HTN0 <- comorb_ckm %>% select(RELEASEID, HTN0)
HTN0 <- HTN0 %>% rename(HTN = HTN0)
HTN0$days <- 0
HTN <- comorb_ckm %>% select(RELEASEID, DAYSTOHTN, HTN)
colnames(HTN) <- c("RELEASEID", "days", "HTN")
HTN <- rbind(HTN0, HTN)
HTN <- HTN %>% filter(HTN == 1)
HTN <- unique(HTN)
ANYDLP0 <- comorb_ckm %>% select(RELEASEID, ANYDLP0)
ANYDLP0 <- ANYDLP0 %>% rename(ANYDLP = ANYDLP0)
ANYDLP0$days <- 0
ANYDLP <- comorb_ckm %>% select(RELEASEID, DAYSTOANYDLP, ANYDLP)
colnames(ANYDLP) <- c("RELEASEID", "days", "ANYDLP")
ANYDLP <- rbind(ANYDLP0, ANYDLP)
ANYDLP <- ANYDLP %>% filter(ANYDLP == 1)
ANYDLP <- unique(ANYDLP)
temp <- full_join(HTN, ANYDLP, by = c("RELEASEID", "days"))
arrhythmia <- arrhythmia %>% rename("days" = "DAYSTOARRHYTHMIA")
comorb_long <- full_join(temp, arrhythmia, by = c("RELEASEID", "days"))
cad <- cad %>% rename("days" = "DAYSTOCAD")
comorb_long <- full_join(comorb_long, cad, by = c("RELEASEID", "days"))
chf <- chf %>% rename("days" = "DAYSTOCHF")
comorb_long <- full_join(comorb_long, chf, by = c("RELEASEID", "days"))
LVSD <- LVSD %>% rename("days" = "DAYSTOLVSD")
comorb_long <- full_join(comorb_long, LVSD, by = c("RELEASEID", "days"))
mi <- mi %>% rename("days" = "DAYSTOMI")
comorb_long <- full_join(comorb_long, mi, by = c("RELEASEID", "days"))
dvt <- dvt %>% rename("days" = "DAYSTODVT")
comorb_long <- full_join(comorb_long, dvt, by = c("RELEASEID", "days"))
pad <- pad %>% rename("days" = "DAYSTOPAD")
comorb_long <- full_join(comorb_long, pad, by = c("RELEASEID", "days"))
stroke <- stroke %>% rename("days" = "DAYSTOSTROKE")
comorb_long <- full_join(comorb_long, stroke, by = c("RELEASEID", "days"))
tia <- tia %>% rename("days" = "DAYSTOTIA")
comorb_long <- full_join(comorb_long, tia, by = c("RELEASEID", "days"))
ulcer <- ulcer %>% rename("days" = "DAYSTOULCER")
comorb_long <- full_join(comorb_long, ulcer, by = c("RELEASEID", "days"))
fup <- comorb_ckm %>% select(RELEASEID, fup_time)

# Biomarker extraction for various markers
bnp_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "bnp")
troponin_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "troponin")
adiponectin_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "adiponectin")
hmwa_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "hmwa")
edol_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "edol")
shbg_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "shbg")
test_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "testosterone")
gfr_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "ckd_gfr")
fas_vals <- extract_biomarker_values_long(list(addcbl_1, addcbl_2), "eGFR_FAS")

# More biomarker extractions
uacr_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "UAlbCreat")
il6_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "IL6")
il1_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "il1")
hscrp_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "hsCRP")
apob_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "ApoB")
mcp1_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "mcp1")
tnfa_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "tnfa")
tnfr1_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "tnfr1")
tnfr2_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "tnfr2")
icam1_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "icam1")
vcam1_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "vcam1")
eselectin_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "eselectin")
vegf_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "vegf")
fgf23_vals <- extract_biomarker_values_long(list(cbl_1, cbl_2), "fgf23")

# Clean infinite values
clean_infinite <- function(df) {
  df %>% mutate(across(everything(), ~ifelse(. == -Inf, NA_real_, .)))
}

# Apply cleaning to all biomarker datasets
biomarker_list <- list(bnp_vals, troponin_vals, adiponectin_vals, hmwa_vals, edol_vals, 
                      shbg_vals, test_vals, gfr_vals, fas_vals, uacr_vals, 
                      il6_vals, il1_vals, hscrp_vals, apob_vals, mcp1_vals, tnfa_vals,
                      tnfr1_vals, tnfr2_vals, icam1_vals, vcam1_vals, eselectin_vals,
                      vegf_vals, fgf23_vals)

biomarker_list <- lapply(biomarker_list, clean_infinite)

# Standardize RELEASEID to uppercase for all biomarker datasets
biomarker_list <- lapply(biomarker_list, function(df) {
  df %>% rename(RELEASEID = releaseid) 
})

# Create consolidated biomarker dataset
biom_ckm <- list(biomarker_list[[1]], biomarker_list[[2]], biomarker_list[[3]],
                 biomarker_list[[8]], biomarker_list[[9]], biomarker_list[[10]],
                 biomarker_list[[11]], biomarker_list[[12]], biomarker_list[[13]], biomarker_list[[14]],
                 biomarker_list[[16]]) %>%
  reduce(full_join, by = c("RELEASEID", "days")) %>%
  distinct(RELEASEID, days, .keep_all = TRUE)

# Filter out rows where all biomarker columns are NA
cols_to_check <- setdiff(names(biom_ckm), c("RELEASEID"))
biom_ckm <- biom_ckm %>%
  filter(!if_all(all_of(cols_to_check), is.na))

# Create high troponin indicators
biom_ckm <- biom_ckm %>%
  mutate(
    high_trop = ifelse(is.na(troponin), NA, 
                       ifelse(troponin > 0, 1, 0))
  )

# Create high BNP indicators
biom_ckm <- biom_ckm %>%
  mutate(
    high_bnp = ifelse(is.na(bnp), NA, 
                       ifelse(bnp > 80, 1, 0))
  )

# convert UAlbCreat
biom_ckm$UAlbCreat <- biom_ckm$UAlbCreat*1000

# create dataframe of UACR and eGFR values for classifying KDIGO criteria
gfr1 <- addcbl_1 %>% select(releaseid, days, ckd_gfr) 
#gfr2 <- addcbl_2 %>% select(releaseid, days, ckd_gfr)
#all_gfr <- rbind(gfr1, gfr2)
all_gfr <- gfr1
uacr1 <- cbl_1 %>% select(releaseid, days, UAlbCreat)
uacr2 <- cbl_2 %>% select(releaseid, days, UAlbCreat)
all_uacr <- rbind(uacr1, uacr2)
kdigo <- full_join(all_gfr, all_uacr, by = c("releaseid", "days"))
kdigo <- kdigo %>% arrange(releaseid, days)
kdigo$UAlbCreat_mg_g <- kdigo$UAlbCreat*1000
kdigo <- classify_ckd_batch(kdigo, gfr_col = "ckd_gfr", albumin_mg_g_col = "UAlbCreat_mg_g")
kdigo <- kdigo %>% rename(RELEASEID = releaseid)
kdigo <- kdigo %>% group_by(RELEASEID) %>% fill(risk_level, .direction = "down") %>%
  fill(recommendation, .direction = "down") %>% ungroup() %>% select(RELEASEID, days, gfr_category, gfr_description, albumin_category, albumin_description, risk_level, recommendation)
```

```{r include = F}
# =============================================================================
# ECHO AND PWV DATA PROCESSING
# =============================================================================

# Process echo data
drop_cols <- c("dopplerqc","mmodeqc","overallqc","plaxqc","saxqc","apicalqc","mmodeqc")
today2_echo <- today2_echo %>% select(-one_of(drop_cols))

# Combine echo datasets
echo <- bind_rows(today2_echo, today_echo)
echo$TIMEPOINT <- as.factor(echo$TIMEPOINT)
echo$RELEASEID <- echo$releaseid
echo$releaseid <- NULL

# Keep only needed echo variables
echo_keep <- echo %>% select(RELEASEID, TIMEPOINT, lvsepem, lvem, lvsepratio, lvratio, 
                            peakvelo, laarea2d, lvmass, walthick, ivsdias, ivssyst, 
                            walldias, wallsyst)
echo_keep$average_E_Em <- (echo_keep$lvsepratio + echo_keep$lvratio) / 2

# Process speckle tracking
speckle_keep <- speckle %>% select(RELEASEID, TIMEPOINT, GLS_4CH, GLS_2CH, GLS_3CH)
speckle_keep$TIMEPOINT <- gsub("SPECKLE", "", speckle_keep$TIMEPOINT)

# Merge echo and speckle
echo_keep <- merge(echo_keep, speckle_keep, by = c("RELEASEID", "TIMEPOINT"), all.x = TRUE)

# Process PWV data
pwv_keep <- pwv %>% select(RELEASEID, DAYS, PWVF)
pwv_keep$high_pwv <- ifelse(is.na(pwv_keep$PWVF), NA, 
                               ifelse(pwv_keep$PWVF > 10, 1, 0))
pwv_keep <- pwv_keep %>% rename(days = DAYS)

# Add height and weight data for BSA calculation
long_clean <- long %>%
  mutate(
    RELEASEID = toupper(releaseid),
    visit_type = substr(visit, 1, 1),
    visit_num = as.numeric(str_extract(visit, "\\d+"))
  )

# Get heights by visit type
# LP: This is taking the last height - is that what we want?
height_M <- long_clean %>%
  filter(visit_type == "M", !is.na(height)) %>%
  group_by(RELEASEID) %>%
  slice_max(order_by = visit_num, n = 1, with_ties = FALSE) %>%
  select(RELEASEID, height) %>%
  rename(height_M = height)

# LP: This is taking the last height - is that what we want?
height_P <- long_clean %>%
  filter(visit_type == "P", !is.na(height)) %>%
  group_by(RELEASEID) %>%
  slice_max(order_by = visit_num, n = 1, with_ties = FALSE) %>%
  select(RELEASEID, height) %>%
  rename(height_P = height)

# Add heights to echo data
echo_keep <- echo_keep %>%
  mutate(RELEASEID = toupper(RELEASEID)) %>%
  left_join(height_M, by = "RELEASEID") %>%
  left_join(height_P, by = "RELEASEID") %>%
  mutate(
    height_final = case_when(
      TIMEPOINT == 1 ~ height_M,
      TIMEPOINT == 2 ~ height_P,
      TRUE ~ NA_real_
    )
  ) %>%
  select(-height_M, -height_P)

# Add weight data through visit matching
long_clean <- long %>%
  mutate(RELEASEID = toupper(releaseid)) %>%
  select(RELEASEID, visit, visit_days, weight, height) %>%
  filter(!is.na(visit))

# Match closest visits for missing weight data
echo_keep <- echo_keep %>%
  left_join(echo %>% select(RELEASEID, TIMEPOINT, days), by = c("RELEASEID", "TIMEPOINT"))

# weight addition based on closest day before echo
echo_keep <- echo_keep %>%
  rowwise() %>%
  mutate(
    weight = {
      current_id <- RELEASEID
      weights <- long_clean %>% 
        filter(RELEASEID == current_id, 
               !is.na(weight),
               visit_days < days)  # Filter here
      
      if(nrow(weights) > 0) {
        # Since all weights are before 'days', get the one with max visit_days (closest)
        weights$weight[which.max(weights$visit_days)]
      } else {
        NA_real_
      }
    }
  ) %>%
  ungroup()

# Calculate BSA using Du Bois and Mosteller formulas
echo_keep <- echo_keep %>%
  mutate(
    BSA_dubois = ifelse(!is.na(height_final) & !is.na(weight),
                        0.007184 * (height_final ^ 0.725) * (weight ^ 0.425),
                        NA_real_),
    BSA_mosteller = ifelse(!is.na(height_final) & !is.na(weight),
                           sqrt((height_final * weight) / 3600),
                           NA_real_)
  )

# Add sex information for LVMI calculation
echo_keep <- echo_keep %>%
  mutate(releaseid = tolower(RELEASEID)) %>%
  left_join(baserisk %>% select(releaseid, sex), by = "releaseid") %>%
  select(-releaseid)

# Calculate LVMI and related scores
# LP: I think this logic is incorrect
# echo_keep <- echo_keep %>%
#   mutate(
#     lvmi = ifelse(!is.na(lvmass) & !is.na(BSA_dubois), lvmass / BSA_dubois, NA_real_),
#     lvmi_score = case_when(
#       sex == 2 & lvmi >= 149 ~ 2,
#       sex == 2 & lvmi >= 122 ~ 1,
#       sex == 2 ~ 0,
#       sex == 1 & lvmi >= 115 ~ 2,
#       sex == 1 & lvmi >= 95  ~ 1,
#       sex == 1 ~ 0,
#       TRUE ~ NA_real_
#     )
#   )
echo_keep <- echo_keep %>%
  mutate(
    lvmi = ifelse(!is.na(lvmass) & !is.na(BSA_dubois), lvmass / BSA_dubois, NA_real_),
    lvmi_score = case_when(
      sex == 2 & lvmi >= 149 ~ 2,  # Male severe
      sex == 2 & lvmi >= 116 ~ 1,  # Male mild (using 116 as threshold for abnormal)
      sex == 2 ~ 0,                 # Male normal
      sex == 1 & lvmi >= 122 ~ 2,  # Female severe  
      sex == 1 & lvmi >= 96  ~ 1,  # Female mild (using 96 as threshold for abnormal)
      sex == 1 ~ 0,                 # Female normal
      TRUE ~ NA_real_
    )
  )

# LP: doesn't this need to take timepoint into account?
# echo_keep <- echo_keep %>%
#   left_join(
#     echo %>% select(RELEASEID, ladimen),
#     by = "RELEASEID"
#   )
echo_keep <- echo_keep %>%
  left_join(
    echo %>% select(RELEASEID, TIMEPOINT, ladimen),
    by = c("RELEASEID", "TIMEPOINT")
  )

# Calculate LAVI estimation
echo_keep <- echo_keep %>%
  mutate(
    LA_volume_est = ifelse(!is.na(laarea2d) & !is.na(ladimen),
                           (8 / (3 * pi)) * (laarea2d^2 / ladimen),
                           NA_real_),
    LAVI_est = ifelse(!is.na(LA_volume_est) & !is.na(BSA_dubois),
                      LA_volume_est / BSA_dubois,
                      NA_real_)
  )
```

```{r include = F}
# =============================================================================
# HFpEF SCORING SYSTEM
# =============================================================================

# Initialize HFpEF points
echo_keep$HFpEF_points <- 0

# Get maximum BNP per participant
# LP: this gets the max BNP and merges it to both echo time points 
# would we want to use only BNP data up to the point of each echo?
# max_bnp_by_id <- aggregate(addcbl_1$bnp ~ addcbl_1$releaseid, data = addcbl_1, FUN = function(x) {
#   if(all(is.na(x))) return(NA)
#   return(max(x, na.rm = TRUE))
# })
# colnames(max_bnp_by_id) <- c("RELEASEID", "bnp")

# Get maximum BNP up to each echo time point
# This ensures we're not using future BNP data
echo_keep_with_bnp <- echo_keep  # Create a copy to preserve original
# Initialize BNP column
echo_keep_with_bnp$bnp <- NA
# Loop through each row in echo_keep to find max BNP up to that time point
for(i in 1:nrow(echo_keep_with_bnp)) {
  current_id <- echo_keep_with_bnp$RELEASEID[i]
  current_days <- echo_keep_with_bnp$days[i]
  # Get all BNP values for this participant up to current echo time
  bnp_subset <- addcbl_1$bnp[addcbl_1$releaseid == current_id & 
                              addcbl_1$days <= current_days]
  # Calculate max BNP if there are any non-NA values
  if(length(bnp_subset) > 0 && !all(is.na(bnp_subset))) {
    echo_keep_with_bnp$bnp[i] <- max(bnp_subset, na.rm = TRUE)
  }
}
# Replace original echo_keep with the updated version
echo_keep <- echo_keep_with_bnp

# HFpEF scoring algorithm
for (i in 1:nrow(echo_keep)) {
  scored_2pts <- FALSE
  
  # 2-point functional scoring
  if (!is.na(echo_keep$lvsepem[i]) && echo_keep$lvsepem[i] < 7) {
    echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 2
    scored_2pts <- TRUE
  } else if (!is.na(echo_keep$lvem[i]) && echo_keep$lvem[i] < 10) {
    echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 2
    scored_2pts <- TRUE
  } else if (!is.na(echo_keep$average_E_Em[i]) && echo_keep$average_E_Em[i] >= 15) {
    echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 2
    scored_2pts <- TRUE
  } else if (!is.na(echo_keep$peakvelo[i]) && echo_keep$peakvelo[i] > 280) {
    echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 2
    scored_2pts <- TRUE
  }
  
  # 1-point functional scoring (only if no 2-point condition was met)
  if (!scored_2pts) {
    if (
      (!is.na(echo_keep$GLS_4CH[i]) && echo_keep$GLS_4CH[i] > -16) ||
      (!is.na(echo_keep$GLS_3CH[i]) && echo_keep$GLS_3CH[i] > -16) ||
      (!is.na(echo_keep$GLS_2CH[i]) && echo_keep$GLS_2CH[i] > -16) ||
      (!is.na(echo_keep$average_E_Em[i]) && echo_keep$average_E_Em[i] >= 9 && echo_keep$average_E_Em[i] < 15)
    ) {
      echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 1
    }
  }
  
  # BNP scoring
  if(!is.na(echo_keep$bnp[i])) {
    if(echo_keep$bnp[i] >= 80) {
      echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 2
    } else if(echo_keep$bnp[i] >= 35 && echo_keep$bnp[i] < 80) {
      echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 1
    }
  }
  
  # Morphological scoring
  if(!is.na(echo_keep$lvmi_score[i])) {
    echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + echo_keep$lvmi_score[i]
  }
  
  # Wall thickness scoring
  if (
    (!is.na(echo_keep$ivsdias[i]) && echo_keep$ivsdias[i] > 1.2) ||
    (!is.na(echo_keep$walldias[i]) && echo_keep$walldias[i] > 1.2) ||
    (!is.na(echo_keep$walthick[i]) && echo_keep$walthick[i] > 0.42)
  ) {
    echo_keep$HFpEF_points[i] <- echo_keep$HFpEF_points[i] + 1
  }
}

# Create HFpEF diagnosis variable
echo_keep$hfpef <- ifelse(echo_keep$HFpEF_points >= 5, 1, 0)
echo_keep <- echo_keep %>% select(-bnp)

```

```{r include = F}
# =============================================================================
# QUALITY OF LIFE (PedsQL) PROCESSING
# =============================================================================

# Define columns to keep from PedsQL
cols_to_keep <- c("RELEASEID", "DAYS", "G01WALK", "G02RUN", "G03SPORT",
                  "G04LIFT", "G05BATH", "G06CHORE", "G07HURT", "G08ENERG")

# Combine all PedsQL datasets
pedsQL_all <- bind_rows(
  pedsQL_Y %>% select(all_of(cols_to_keep)),
  pedsQL_T %>% select(all_of(cols_to_keep)),
  pedsQL_A %>% select(all_of(cols_to_keep)),
  pedsQL_C_today %>% select(all_of(cols_to_keep)),
  pedsQL_T_today %>% select(all_of(cols_to_keep))
)

# Define PedsQL items
pedsQL_items <- c("G01WALK", "G02RUN", "G03SPORT", "G04LIFT", 
                  "G05BATH", "G06CHORE", "G07HURT", "G08ENERG")

# Calculate physical activity total
pedsQL_all <- pedsQL_all %>%
  rowwise() %>%
  mutate(physical_total = sum(c_across(all_of(pedsQL_items)), na.rm = TRUE)) %>%
  ungroup()

# Ensure DAYS is numeric
pedsQL_all <- pedsQL_all %>%
  mutate(days = as.numeric(DAYS)) %>% select(-DAYS)

# Get baseline, latest, lowest and highest scores per participant
pedsQL_summary <- pedsQL_all %>%
  group_by(RELEASEID) %>%
  summarise(
    pedsql_baseline = physical_total[which.min(days)],
    days_baseline = days[which.min(days)],
    pedsql_last = physical_total[which.max(days)],
    days_last = days[which.max(days)],
    pedsql_highest = max(physical_total, na.rm = TRUE),
    days_highest = days[which.max(physical_total)],
    pedsql_lowest = min(physical_total, na.rm = TRUE),
    days_lowest = days[which.min(physical_total)],
    .groups = "drop"
  )
```

```{r include = F}
# =============================================================================
# DATA INTEGRATION AND CKM STAGING
# =============================================================================

# set any days <0 to zero
baseline$days <- 0
baserisk$days <- 0
biom_ckm$days <- ifelse(biom_ckm$days <0, 0, biom_ckm$days)

# Fix the column name first
baseline <- baseline %>% 
 rename(RELEASEID = releaseid)  # or whatever the actual column name is

baserisk <- baserisk %>% 
 rename(RELEASEID = releaseid)  # or whatever the actual column name is

# Create main analysis dataset
comorb_ckm2 <- comorb_long %>%
  full_join(pwv_keep, by = c("RELEASEID", "days")) %>%
  full_join(echo_keep, by = c("RELEASEID", "days")) %>%
  full_join(pedsQL_all, by = c("RELEASEID", "days")) %>%
  full_join(baseline %>% select(RELEASEID, bmi, sbp, dbp, wastcirc, days), by = c("RELEASEID", "days")) %>%
  full_join(baserisk %>% select(RELEASEID, AGEBASE, days), by = c("RELEASEID", "days")) %>% 
  full_join(kdigo, , by = c("RELEASEID", "days"))
comorb_ckm2 <- comorb_ckm2 %>% arrange("RELEASEID", "days")
comorb_ckm2 <- comorb_ckm2 %>% group_by(RELEASEID) %>% fill(gfr_category, .direction = "down") %>% fill(gfr_description, .direction = "down") %>%
  fill(albumin_category, .direction = "down") %>% fill(albumin_description, .direction = "down") %>% fill(risk_level, .direction = "down") %>%
  fill(recommendation, .direction = "down") %>% ungroup()

# Merge all biomarker datasets at once
# all_biomarkers <- biomarker_list %>%
#   reduce(full_join, by = c("RELEASEID", "days"))

# convert UAlbCreat
# all_biomarkers$UAlbCreat_baseline <- all_biomarkers$UAlbCreat_baseline*1000
# all_biomarkers$UAlbCreat_highest <- all_biomarkers$UAlbCreat_highest*1000

# Add to main dataset
comorb_ckm2 <- comorb_ckm2 %>%
  full_join(biom_ckm, by = c("RELEASEID", "days"))

# need to fill down, and make earlier NA's into zero
comorb_ckm2 <- comorb_ckm2 %>% arrange(RELEASEID, days)
comorb_ckm2 <- comorb_ckm2 %>% group_by(RELEASEID) %>% fill(HTN, .direction = "down") %>% fill(ANYDLP, .direction = "down") %>% fill(ARRHYTHMIA, .direction = "down") %>%
  fill(CAD, .direction = "down") %>% fill(CHF, .direction = "down") %>% fill(LVSD, .direction = "down") %>% fill(MI, .direction = "down") %>% fill(DVT, .direction = "down") %>%
  fill(PAD, .direction = "down") %>% fill(STROKE, .direction = "down") %>% fill(TIA, .direction = "down") %>% fill(ULCER, .direction = "down") %>% fill(high_pwv, .direction = "down") %>%
  fill(hfpef, .direction = "down") %>% fill(high_bnp, .direction = "down") %>% fill(high_trop, .direction = "down")

comorb_ckm2$clin_cvd <- ifelse(
  rowSums(comorb_ckm2[, c("ARRHYTHMIA", "CAD", "CHF", "MI", "PAD", "DVT", "STROKE", "TIA", "ULCER", "LVSD")] == 1, na.rm = TRUE) > 0,
  1,
  0)
# fill down clincvd
comorb_ckm2 <- comorb_ckm2 %>% group_by(RELEASEID) %>% fill(clin_cvd, .direction = "down") %>% ungroup()

# fill in zeroes
comorb_ckm2 <- replace_na_tidyverse(comorb_ckm2, c("HTN", "ANYDLP", "ARRHYTHMIA", "CAD", "CHF", "LVSD", "MI", "DVT", "PAD", "STROKE", "TIA", "ULCER", "high_pwv", "hfpef",
                                            "high_trop", "high_bnp"))

# drop unneeded variables
comorb_ckm2 <- comorb_ckm2 %>% select(-TIMEPOINT, -lvsepem, -lvem, -lvsepratio, -lvratio, -peakvelo, -laarea2d, -lvmass, -walthick, -ivsdias, -ivssyst, -walldias, -wallsyst,
                                      -average_E_Em, -GLS_4CH, -GLS_2CH, -GLS_3CH, -height_final, -weight, -BSA_mosteller, -BSA_dubois, -lvmi, -lvmi_score, -ladimen, 
                                      -LA_volume_est, -LAVI_est, -HFpEF_points, -G01WALK, -G02RUN, -G03SPORT, -G04LIFT, -G05BATH, -G06CHORE, -G07HURT, -G08ENERG)

# CKM Staging - Main classification
comorb_ckm2 <- comorb_ckm2 %>%
  mutate(
    CKM_syn = case_when(
      # Stage 4: most severe
      clin_cvd == 1 |
      hfpef == 1 ~ "Stage4",
      
      # Stage 3: elevated biomarkers
       (high_bnp == 1 |
    #   high_trop == 1 |
      high_pwv == 1 |
       risk_level == "Very high") ~ "Stage3",
      
      # Stage 2_plus: risk factors
      # need to add KDIGO criteria
      (HTN == 1 |
       ANYDLP == 1 |
       risk_level %in% c("Moderate", "High")) ~ "Stage2_plus",
      
      TRUE ~ "Stage2"
    )
  )

# CKM Staging with BMI inclusion
comorb_ckm2 <- comorb_ckm2 %>%
  mutate(
    CKM_syn2 = case_when(
       # Stage 4: most severe
      clin_cvd == 1 |
      hfpef == 1 ~ "Stage4",
      
      # Stage 3: elevated biomarkers
       (high_bnp == 1 |
#       high_trop == 1 |
       high_pwv == 1 |
       risk_level == "Very high") ~ "Stage3",
      
      # Stage 2_plus: risk factors
      # need to add KDIGO criteria
      (HTN == 1 |
       ANYDLP == 1 |
       risk_level %in% c("Moderate", "High") |
       bmi >= 30) ~ "Stage2_plus",
      
      TRUE ~ "Stage2"
    )
  )

# Set factor levels for proper ordering
comorb_ckm2$CKM_syn <- factor(comorb_ckm2$CKM_syn, levels = c("Stage2", "Stage2_plus", "Stage3", "Stage4"))
comorb_ckm2$CKM_syn2 <- factor(comorb_ckm2$CKM_syn2, levels = c("Stage2", "Stage2_plus", "Stage3", "Stage4"))

# need to create the following variables
# CKM_syn_base, CKM_syn_base2
ckm_base <- comorb_ckm2 %>% filter(days == 0) 
ckm_base <- ckm_base %>% select(RELEASEID, CKM_syn, CKM_syn2)
colnames(ckm_base) <- c("RELEASEID", "CKM_syn_base", "CKM_syn_base2")
comorb_ckm2 <- full_join(comorb_ckm2, ckm_base, by = "RELEASEID")

# create a numeric variable for stages
comorb_ckm2 <- comorb_ckm2 %>%
  mutate(
    # Convert stages to numeric (without BMI)
    CKM_syn_num = case_when(
      CKM_syn == "Stage2" ~ 1,
      CKM_syn == "Stage2_plus" ~ 2,
      CKM_syn == "Stage3" ~ 3,
      CKM_syn == "Stage4" ~ 4
    ),
    # Convert stages to numeric (with BMI)
    CKM_syn_num2 = case_when(
      CKM_syn2 == "Stage2" ~ 1,
      CKM_syn2 == "Stage2_plus" ~ 2,
      CKM_syn2 == "Stage3" ~ 3,
      CKM_syn2 == "Stage4" ~ 4
    ))

# create indicator for progression from previous visit and number of stages
comorb_ckm2 <- comorb_ckm2 %>% arrange(RELEASEID, days)
comorb_ckm2 <- comorb_ckm2 %>% group_by(RELEASEID) %>%
  mutate(
    CKM_syn_lag = ifelse(days == 0, NA, lag(CKM_syn_num)),
    CKM_syn2_lag = ifelse(days == 0, NA, lag(CKM_syn_num2))
  )
comorb_ckm2$progress_CKM <- ifelse(comorb_ckm2$CKM_syn_num > comorb_ckm2$CKM_syn_lag, 1, 0)
comorb_ckm2$progress_CKM_stages <- pmax(0, comorb_ckm2$CKM_syn_num - comorb_ckm2$CKM_syn_lag)
comorb_ckm2$regress_CKM <- ifelse(comorb_ckm2$CKM_syn_num < comorb_ckm2$CKM_syn_lag, 1, 0)
comorb_ckm2$regress_CKM_stages <- pmax(0, comorb_ckm2$CKM_syn_lag - comorb_ckm2$CKM_syn_num)
comorb_ckm2$progress_CKM2 <- ifelse(comorb_ckm2$CKM_syn_num2 > comorb_ckm2$CKM_syn2_lag, 1, 0)
comorb_ckm2$progress_CKM2_stages <- pmax(0, comorb_ckm2$CKM_syn_num2 - comorb_ckm2$CKM_syn2_lag)
comorb_ckm2$regress_CKM2 <- ifelse(comorb_ckm2$CKM_syn_num2 < comorb_ckm2$CKM_syn2_lag, 1, 0)
comorb_ckm2$regress_CKM2_stages <- pmax(0, comorb_ckm2$CKM_syn2_lag - comorb_ckm2$CKM_syn_num2)

# max CKM_syn, max CKM_syn2
ckm_max <- comorb_ckm2 %>% group_by(RELEASEID) %>% arrange(RELEASEID, desc(CKM_syn_num)) %>% filter(row_number() == 1) %>% select(RELEASEID, CKM_syn_num)
colnames(ckm_max) <- c("RELEASEID", "CKM_max")
ckm_max2 <- comorb_ckm2 %>% group_by(RELEASEID) %>% arrange(RELEASEID, desc(CKM_syn_num2)) %>% filter(row_number() == 1) %>% select(RELEASEID, CKM_syn_num2)
colnames(ckm_max2) <- c("RELEASEID", "CKM_max2")
ckm_max <- full_join(ckm_max, ckm_max2, by = "RELEASEID")
comorb_ckm2 <- full_join(comorb_ckm2, ckm_max, by = "RELEASEID")

# time to progression (first? all?)
# first get follow-up time
fup <- comorb_ckm %>% select(RELEASEID, fup_time)
comorb_ckm2 <- full_join(comorb_ckm2, fup)
# get max days for each person
max_days <- comorb_ckm2 %>% select(RELEASEID, days) %>% arrange(RELEASEID, desc(days)) %>% filter(row_number() == 1)
colnames(max_days) <- c("RELEASEID", "maxdays")
comorb_ckm2 <- left_join(comorb_ckm2, max_days, by = "RELEASEID")
comorb_ckm2$fup_time <- ifelse(is.na(comorb_ckm2$fup_time), comorb_ckm2$maxdays, comorb_ckm2$fup_time)
ckm_prog_time <- comorb_ckm2 %>% select(RELEASEID, days, progress_CKM, progress_CKM2, fup_time)
ckm_prog_time1 <- ckm_prog_time %>% arrange(RELEASEID, desc(progress_CKM)) %>% filter(row_number() == 1) %>% select(-progress_CKM2)
ckm_prog_time1$days_to_ckm_prog <- ifelse(ckm_prog_time1$progress_CKM == 1, ckm_prog_time1$days, ckm_prog_time1$fup_time)
ckm_prog_time1 <- ckm_prog_time1 %>% select(RELEASEID, days_to_ckm_prog)
ckm_prog_time2 <- ckm_prog_time %>% arrange(RELEASEID, desc(progress_CKM2)) %>% filter(row_number() == 1) %>% select(-progress_CKM)
ckm_prog_time2$days_to_ckm2_prog <- ifelse(ckm_prog_time2$progress_CKM2 == 1, ckm_prog_time2$days, ckm_prog_time2$fup_time)
ckm_prog_time2 <- ckm_prog_time2 %>% select(RELEASEID, days_to_ckm2_prog)
ckm_prog_time_final <- full_join(ckm_prog_time1, ckm_prog_time2, by = "RELEASEID")
comorb_ckm2 <- full_join(comorb_ckm2, ckm_prog_time_final, by = "RELEASEID")

# number of progressions
a <- comorb_ckm2 %>% filter(progress_CKM == 1) %>% select(RELEASEID, progress_CKM)
b <- comorb_ckm2 %>% filter(progress_CKM2 == 1) %>% select(RELEASEID, progress_CKM2)
tab_a <- as.data.frame(table(a$RELEASEID))
colnames(tab_a) <- c("RELEASEID", "num_prog")
tab_b <- as.data.frame(table(b$RELEASEID))
colnames(tab_b) <- c("RELEASEID", "num_prog2")
prog_table <- full_join(tab_a, tab_b, by = "RELEASEID")
comorb_ckm2 <- full_join(comorb_ckm2, prog_table, by = "RELEASEID")
comorb_ckm2$num_prog <- ifelse(is.na(comorb_ckm2$progress_CKM), NA, 
                                     ifelse(is.na(comorb_ckm2$num_prog), 0, comorb_ckm2$num_prog))
comorb_ckm2$num_prog2 <- ifelse(is.na(comorb_ckm2$progress_CKM2), NA, 
                                ifelse(is.na(comorb_ckm2$num_prog2), 0, comorb_ckm2$num_prog2))

# merge in death
death_keep <- death %>% select(RELEASEID, DEATH)
comorb_ckm2 <- left_join(comorb_ckm2, death_keep, by = "RELEASEID")
comorb_ckm2$DEATH <- ifelse(is.na(comorb_ckm2$DEATH), 0, comorb_ckm2$DEATH)

# convert to factors
cols_to_convert <- c("HTN", "ANYDLP", "high_pwv", "hfpef",
                     "high_trop", "CKM_syn_base", "CKM_max", "CKM_syn_base2", "CKM_max2",
                     "progress_CKM", "progress_CKM2", "regress_CKM", "regress_CKM_stages", "regress_CKM2", "regress_CKM2_stages", "DEATH", 
                     "num_prog", "num_prog2")
comorb_ckm2[cols_to_convert] <- lapply(comorb_ckm2[cols_to_convert], as.factor)

# collapse to a dataset with 1 row per person with baseline CKM, max CKM, time to progression, number of progressions?
# this is an oversimplification because people progress, regress, progress
prog_summary <- comorb_ckm2 %>% select(RELEASEID, fup_time, CKM_syn_base, CKM_max, progress_CKM,  days_to_ckm_prog, num_prog) %>%
  arrange(RELEASEID, desc(progress_CKM)) %>% filter(row_number() == 1)
prog_summary2 <-  comorb_ckm2 %>% select(RELEASEID, fup_time, CKM_syn_base2, CKM_max2, progress_CKM2, days_to_ckm2_prog, num_prog2, -fup_time) %>%
  arrange(RELEASEID, desc(progress_CKM2)) %>% filter(row_number() == 1)
prog_summary_final <- full_join(prog_summary, prog_summary2, by = "RELEASEID")    
# create character variable corresponding to max stages
prog_summary_final$CKM_max_char <- case_when(
  prog_summary_final$CKM_max == 1 ~ "Stage 2",
  prog_summary_final$CKM_max == 2 ~ "Stage 2+",
  prog_summary_final$CKM_max == 3 ~ "Stage 3",
  prog_summary_final$CKM_max == 4 ~ "Stage 4"
)
prog_summary_final$CKM_max_char2 <- case_when(
  prog_summary_final$CKM_max2 == 1 ~ "Stage 2",
  prog_summary_final$CKM_max2 == 2 ~ "Stage 2+",
  prog_summary_final$CKM_max2 == 3 ~ "Stage 3",
  prog_summary_final$CKM_max2 == 4 ~ "Stage 4"
)

# Labels
label(prog_summary_final$progress_CKM) <- "CKM syndrome progression (binary: 0=no progression, 1=progressed, without BMI)"
label(prog_summary_final$progress_CKM2) <- "CKM syndrome progression (binary: 0=no progression, 1=progressed, with BMI)"
label(prog_summary_final$CKM_syn_base) <- "CKM syndrome stage at baseline, without BMI"
label(prog_summary_final$CKM_syn_base2) <- "CKM syndrome stage at baseline, with BMI"
label(prog_summary_final$CKM_max) <- "Maximum CKM stage, without BMI"
label(prog_summary_final$CKM_max_char) <- "Maximum CKM stage, without BMI"
label(prog_summary_final$CKM_max2) <- "Maximum CKM stage, with BMI"
label(prog_summary_final$CKM_max_char2) <- "Maximum CKM stage, without BMI"
label(prog_summary_final$num_prog) <- "Number of progression events, without BMI"
label(prog_summary_final$num_prog2) <- "Number of progression events, with BMI"
label(prog_summary_final$days_to_ckm_prog) <- "Days to first progression event, without BMI"
label(prog_summary_final$days_to_ckm2_prog) <- "Days to first progression event, with BMI"

# make a descriptive dataset using progression summary data and baseline/comorb info
prog_summary_final <- left_join(prog_summary_final, comorb, by = "RELEASEID")
prog_summary_final <- left_join(prog_summary_final, baserisk, by = "RELEASEID") 
prog_summary_final <- prog_summary_final %>% select(-days)
prog_summary_final <- left_join(prog_summary_final, death, by = "RELEASEID")
prog_summary_final$DEATH <- ifelse(is.na(prog_summary_final$DEATH), 0, prog_summary_final$DEATH)

cols_to_convert <- c("HTN", "ANYDLP", "MIC", "MAC", "GLYC", "tx", 
                     "CKM_syn_base", "CKM_max", "CKM_syn_base2", "CKM_max2",
                     "progress_CKM", "progress_CKM2", "DEATH", 
                     "num_prog", "num_prog2")
prog_summary_final[cols_to_convert] <- lapply(prog_summary_final[cols_to_convert], as.factor)

# write file for other analyses
filename <- "/CKM/Results/CKM stage summary.csv"
full_path <- file.path(data_path, filename)
write.csv(prog_summary_final, full_path, row.names = FALSE)

# merge in proteins
# load somalogic data, with QC samples already excluded
filename <- "/Somalogic data raw/soma.Rdata"
full_path <- file.path(data_path, filename)
load(full_path)
# load analyte info
filename <- "/Somalogic data raw/analytes.Rdata"
full_path <- file.path(data_path, filename)
load(full_path)
# take only the baseline soma samples
base <- soma %>% arrange(releaseid,Date.Drawn) %>% dplyr::group_by(releaseid) %>% dplyr::filter(row_number()==1)
# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))
# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))
# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)
# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)
# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[,paste0(predictors[i])] <- base_log_scale[,paste0(predictors[i])]/sd(unlist(base_log[,paste0(predictors[i])]))
}
base_log_scale <- base_log_scale %>% rename(RELEASEID = releaseid)

prog_summary_final_soma <- left_join(base_log_scale, prog_summary_final, by = "RELEASEID")

# # create another data frame with change in proteins
# # find follow-up samples
# fup <- soma %>% anti_join(base, soma, by=c("releaseid","Date.Drawn"))
# # identify columns corresponding to proteins and convert to numeric
# fup <- fup %>%
#        mutate(across(starts_with("seq"),
#               ~ as.numeric(as.character(.))))
# # calculate deltas
# base_for_deltas <- base
# base_for_deltas <- base_for_deltas %>% filter(releaseid %in% fup$releaseid)
# fup <- fup
# %>% filter(releaseid %in% base_for_deltas$releaseid)
# fup <- fup %>% arrange(releaseid)
# base_for_deltas <- base_for_deltas %>% arrange(releaseid)
# # log transform
# # scale by SD
```

# Results and Statistical Analysis

## Descriptive statistics - adjudicated medical events

### Events

```{r}
kable(ame_tab)
```

### Patients

```{r}
kable(ame_pt_tab)
```

## CKM stage summary

```{r include=F}
t1 <- tableby( ~ CKM_syn_base + CKM_syn_base2 + CKM_max_char + CKM_max_char2 + progress_CKM + progress_CKM2 +
                 num_prog + num_prog2 + notest(days_to_ckm_prog, "Nmiss", "median", "q1q3") + notest(days_to_ckm2_prog, "Nmiss", "median", "q1q3"), 
              data = prog_summary_final, total = F)
```

```{r results='asis'}
summary(t1,  pfootnote = F)
```

## Descriptive statistics by maximum CKM stage, without BMI

```{r include=F}
t2 <- tableby(CKM_max_char ~  days_to_ckm_prog + HTN + ANYDLP + MIC + MAC + GLYC + HbA1c +
                log_trig + sbp + dbp + uacid + si_1_ins0 + UAlbCreat + bmi + HDL + codi + ALT + AST + sex_char + racedesc + AGEBASE +
                txdesc + DEATH, 
              data = prog_summary_final, total = F)
```

```{r results='asis'}
summary(t2,  pfootnote = F)
```

## Descriptive statistics by maximum CKM stage, with BMI

```{r include=F}
t3 <- tableby(CKM_max_char2 ~  days_to_ckm2_prog + HTN + ANYDLP + MIC + MAC + GLYC + HbA1c +
                log_trig + sbp + dbp + uacid + si_1_ins0 + UAlbCreat + bmi + HDL + codi + ALT + AST + sex_char + racedesc + AGEBASE +
                txdesc + DEATH, 
              data = prog_summary_final, total = F)
```

```{r results='asis'}
summary(t3,  pfootnote = F)
```

```{r, include=FALSE}
# Run univariable Cox models for time to first progression without BMI
result_noBMI <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(days_to_ckm_prog, as.numeric(progress_CKM))~",predictors[i])
  mod <- coxph(data = prog_summary_final_soma, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_noBMI <- rbind(result_noBMI,tidymod)
}
result_noBMI$adj.p.value <- p.adjust(result_noBMI$p.value, "fdr")
result_noBMI$AptName <- result_noBMI$term
result_noBMI$term <- NULL
result_noBMI <- merge(result_noBMI,analytes,by="AptName",all.x = T, all.y = F)
result_noBMI <- result_noBMI[order(result_noBMI$p.value),] 
```

```{r, include=FALSE}
# Run univariable Cox models for time to first progression with BMI
result_BMI <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(days_to_ckm_prog, as.numeric(progress_CKM2))~",predictors[i])
  mod <- coxph(data = prog_summary_final_soma, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_BMI <- rbind(result_BMI,tidymod)
}
result_BMI$adj.p.value <- p.adjust(result_BMI$p.value, "fdr")
result_BMI$AptName <- result_BMI$term
result_BMI$term <- NULL
result_BMI <- merge(result_BMI,analytes,by="AptName",all.x = T, all.y = F)
result_BMI <- result_BMI[order(result_BMI$p.value),] 
```

```{r, include=FALSE}
# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"CKM without BMI CPH")
writeData(wb,"CKM without BMI CPH",result_noBMI,rowNames = F)
addWorksheet(wb,"CKM with BMI CPH")
writeData(wb,"CKM with BMI CPH",result_BMI,rowNames = F)
full_path <- file.path(data_path, "CKM/Results/TODAY CKM Cox models scaled baseline adjusted.xlsx")
saveWorkbook(wb, full_path, overwrite = TRUE)

# separate files for IPA
wb <- createWorkbook()
addWorksheet(wb,"CKM without BMI CPH")
writeData(wb,"CKM without BMI CPH",result_noBMI,rowNames = F)
full_path <- file.path(data_path, "CKM/Results/IPA/TODAY somalogic Cox models scaled baseline adjusted without BMI.xlsx")
saveWorkbook(wb, full_path, overwrite = TRUE)

wb <- createWorkbook()
addWorksheet(wb,"CKM with BMI CPH")
writeData(wb,"CKM with BMI CPH",result_BMI,rowNames = F)
full_path <- file.path(data_path, "CKM/Results/IPA/TODAY somalogic Cox models scaled baseline adjusted with BMI.xlsx")
saveWorkbook(wb, full_path, overwrite = TRUE)
```

```{r, include=FALSE}
# Elastic net Cox models

# filter to top 100 proteins by effect size
f <- file.path(data_path, "CKM/Results/TODAY CKM Cox models scaled baseline adjusted.xlsx")

# CKM without BMI
ckm_without_bmi_keep = read_excel(f,sheet = "CKM without BMI CPH")
ckm_without_bmi_keep <- arrange(ckm_without_bmi_keep,adj.p.value)
ckm_without_bmi_drop <- ckm_without_bmi_keep[101:nrow(ckm_without_bmi_keep),"AptName"]
rownames(ckm_without_bmi_drop) <- ckm_without_bmi_drop$AptName
ckm_without_bmi_drop <- as.data.frame(t(ckm_without_bmi_drop))
ckm_without_bmi_endata <- prog_summary_final_soma %>% select(-colnames(ckm_without_bmi_drop))

# CKM with BMI
ckm_with_bmi_keep = read_excel(f,sheet = "CKM with BMI CPH")
ckm_with_bmi_keep <- arrange(ckm_with_bmi_keep,adj.p.value)
ckm_with_bmi_drop <- ckm_with_bmi_keep[101:nrow(ckm_with_bmi_keep),"AptName"]
rownames(ckm_with_bmi_drop) <- ckm_with_bmi_drop$AptName
ckm_with_bmi_drop <- as.data.frame(t(ckm_with_bmi_drop))
ckm_with_bmi_endata <- prog_summary_final_soma %>% select(-colnames(ckm_with_bmi_drop))

set.seed(3654)
filename <- "TODAY proteomics metabolomics/Proteomics analyses/Elastic net/easy_elasticnet.R"
full_path <- file.path(github_path, filename)
source(full_path)
```

# Elastic net models

## CKM without BMI

```{r, include=FALSE, comment=""}
ckm_without_bmi_endata_nozero <- ckm_without_bmi_endata[!ckm_without_bmi_endata$days_to_ckm_prog == 0,]
seq <- is_seq(names(ckm_without_bmi_endata_nozero))

ckm_without_bmi_select = easy_elasticnet(data = ckm_without_bmi_endata_nozero,outcome = "progress_CKM",
                           predictors = ckm_without_bmi_endata_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "days_to_ckm_prog", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% ckm_without_bmi_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(days_to_ckm_prog, as.numeric(progress_CKM))~",paste0(ckm_without_bmi_select,collapse = "+")))
mod = coxph(formula = form,data = ckm_without_bmi_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")

#concordance_df <- as.data.frame(cbind("Severe albuminuria","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

## CKM with BMI

```{r, include=FALSE, comment=""}
ckm_with_bmi_endata_nozero <- ckm_with_bmi_endata[!ckm_with_bmi_endata$days_to_ckm2_prog == 0,]
seq <- is_seq(names(ckm_with_bmi_endata_nozero))

ckm_with_bmi_select = easy_elasticnet(data = ckm_with_bmi_endata_nozero,outcome = "progress_CKM2",
                           predictors = ckm_with_bmi_endata_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "days_to_ckm2_prog", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% ckm_without_bmi_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(days_to_ckm2_prog, as.numeric(progress_CKM2))~",paste0(ckm_with_bmi_select,collapse = "+")))
mod = coxph(formula = form,data = ckm_with_bmi_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")

#concordance_df <- as.data.frame(cbind("Severe albuminuria","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```
