---
title: "CKM in the TODAY study"
author: "Laura Pyle / Jairo A Pinzon"
date: "`r lubridate::today()`"
format: html
editor: visual
echo: FALSE
warning: FALSE
toc: TRUE
embed-resources: TRUE
---

```{r include = F}
# =============================================================================
# LIBRARIES
# =============================================================================
library(dplyr)
library(arsenal)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(ggalluvial)
library(FSA)
library(fuzzyjoin)
library(purrr)
library(stringr)
library(arsenal)
library(Hmisc)
library(SomaDataIO)
library(caret)
library(survival)
library(broom)
library(openxlsx)
library(readxl)
library(knitr)
library(limma)
library(gtsummary)
library(cardx)
library(forestploter)
library(forcats)
library(fgsea)
# Install from CRAN
#install.packages("ensr")
library(ensr)

# Working directory setup
getwd()
#usethis::use_git()

# specify user for paths
user <- Sys.info()[["user"]]
if (user == "laurapyle") {
  data_path <- "/Users/laurapyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  github_path <- "/Users/laurapyle/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  root_path <- "/Users/laurapyle/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
} else if (user == "lpyle") {
  data_path <- "/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  github_path <- "/Users/lpyle/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  root_path <- "/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  teen_labs_path <- "/Users/lpyle/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Teen Labs/"
} else if (user == "pylell") {
  data_path <- "/Users/pylell/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  github_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  teen_labs_path <- "/Users/pylell/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/Teen Labs/"
} else if (user == "jpcortes") {
  # Update the data_path to your OneDrive or cloud storage location
  data_path <- "/Users/jpcortes/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive/TODAY subaward"
  github_path <- "/Users/jpcortes/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  root_path <- "/Users/jpcortes/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive"
  } else {
  stop("Unknown user: please specify root path for this user.")
}

filename <- "/TODAY proteomics metabolomics/CKM/CKM functions.R"
full_path <- file.path(github_path, filename)
source(full_path)
```

```{r include = F}
# load CKM long file
filename <- "/Clinical data/ckm_long.Rdata"
full_path <- file.path(data_path, filename)
load(full_path)

# load CKM summary file
filename <- "/Clinical data/ckm_summary.Rdata"
full_path <- file.path(data_path, filename)
load(full_path)

# merge in proteins
# load somalogic data, with QC samples already excluded
filename <- "/Somalogic data raw/soma.Rdata"
full_path <- file.path(data_path, filename)
load(full_path)
# load analyte info
filename <- "/Somalogic data raw/analytes.Rdata"
full_path <- file.path(data_path, filename)
load(full_path)
# take only the baseline soma samples
base <- soma %>% arrange(releaseid,Date.Drawn) %>% dplyr::group_by(releaseid) %>% dplyr::filter(row_number()==1)
# these 3 release IDs have first sample after end of recruitment, so they must be missing the baseline visit
base <- base %>% filter(!releaseid %in% c("65-85903","65-47984","65-25901"))
# identify columns corresponding to proteins
#is_seq <- function(.x) grepl("^seq\\.[0-9]{4}", .x) # regex for analytes
is_seq <- function(.x) grepl("seq", .x)
seq <- is_seq(names(base))
# convert to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)
# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log2)
# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[,paste0(predictors[i])] <- base_log_scale[,paste0(predictors[i])]/sd(unlist(base_log[,paste0(predictors[i])]))
}
base_log_scale <- base_log_scale %>% rename(RELEASEID = releaseid)

prog_summary_final_soma <- left_join(base_log_scale, prog_summary_final, by = "RELEASEID")

# create another data frame with change in proteins
# find follow-up samples
fup <- soma %>% anti_join(base, soma, by=c("releaseid","Date.Drawn"))
# identify columns corresponding to proteins and convert to numeric
fup <- fup %>%
       mutate(across(starts_with("seq"),
              ~ as.numeric(as.character(.))))

# calculate deltas
base_for_deltas <- base
base_for_deltas <- base_for_deltas %>% filter(releaseid %in% fup$releaseid)
fup <- fup %>% filter(releaseid %in% base_for_deltas$releaseid)
fup <- fup %>% arrange(releaseid)
base_for_deltas <- base_for_deltas %>% arrange(releaseid)
a <- fup %>% select(starts_with("seq"))
b <- base_for_deltas %>% ungroup() %>% select(starts_with("seq"))
deltas <- as.data.frame(a) - as.data.frame(b)
# can't log transform because many delta values are negative
# scale by SD
deltas_scale <- deltas
for (i in 1:length(predictors)) {
  deltas_scale[,paste0(predictors[i])] <- deltas_scale[,paste0(predictors[i])]/sd(unlist(deltas_scale[,paste0(predictors[i])]))
}
# add back in releaseid and progression information
deltas_scale <- cbind(base_for_deltas$releaseid, deltas_scale)
colnames(deltas_scale)[1] <- "RELEASEID"
deltas_scale <- left_join(deltas_scale, prog_summary_final, by = "RELEASEID")
# create dataframe that is not scaled by SD
deltas_no_scale <- cbind(base_for_deltas$releaseid, deltas)
colnames(deltas_no_scale)[1] <- "RELEASEID"
deltas_no_scale <- left_join(deltas_no_scale, prog_summary_final, by = "RELEASEID")

# analyzing the deltas doesn't work well because they can't be log transformed
# instead, create a long dataset with log transformed proteins and use paired limma
temp1 <- base_for_deltas %>% rename("RELEASEID" = "releaseid")
temp2 <- fup %>% rename("RELEASEID" = "releaseid")
soma_long <- rbind(temp1, temp2) 
soma_long <- left_join(soma_long, prog_summary_final, by = "RELEASEID")
soma_long_log <- soma_long %>% modify_if(is_seq(names(.)), log)

# adjudicated medical events
filename <- "/Clinical data/TODAY2/AME.csv"
full_path <- file.path(data_path, filename)
ame <- read.csv(full_path)

ame_tab <- table(ame$AMENAME, ame$AMETYPE)
ame_pt <- ame %>% select(releaseid, AMENAME, AMETYPE) %>% unique()
ame_pt_tab <- table(ame_pt$AMENAME, ame_pt$AMETYPE)
ametype <- c("Heart", "Vascular", "Cerebrovascular", "GI", "Nerve", "Renal", "Eye", "Death")
amename <- c("Arrhythmia", "CAD", "CHD", "LVSD", "MI", "PAD", "DVT", "Stroke", "TIA",
             "Pancreatitis", "Gallbladder disease", "Peripheral neuropathy", "Autonomic neuropathy", "Diabetic mononeuropathy", "CKD", "ESKD", "NPDR", 
             "PDR", "Macular edema", "Vitreous hemorrhage", "Blindness", "Cataracts", "Glaucoma", "Death")
ame_tab <- as.data.frame.matrix(ame_tab)
ame_pt_tab <- as.data.frame.matrix(ame_pt_tab)
colnames(ame_tab) <- ametype
rownames(ame_tab) <- amename
colnames(ame_pt_tab) <- ametype
rownames(ame_pt_tab) <- amename
ame_tab <- as.data.frame(ame_tab)
ame_pt_tab <- as.data.frame(ame_pt_tab)
ame_tab <- ame_tab %>%  mutate(Total = rowSums(across(where(is.numeric))))
ame_pt_tab <- ame_pt_tab %>%  mutate(Total = rowSums(across(where(is.numeric))))
ame_tab_col_tot <- colSums(ame_tab)
ame_tab <- rbind(ame_tab, ame_tab_col_tot)
rownames(ame_tab)[nrow(ame_tab)] <- "Total"
ame_pt_tab_col_tot <- colSums(ame_pt_tab)
ame_pt_tab <- rbind(ame_pt_tab, ame_pt_tab_col_tot)
rownames(ame_pt_tab)[nrow(ame_pt_tab)] <- "Total"
```

# Results and Statistical Analysis

## Descriptive statistics - adjudicated medical events

### Events

```{r}
kable(ame_tab)
```

### Patients

```{r}
kable(ame_pt_tab)
```

## Criteria for CKM stages

```{r include=F}
binary_vars <- c("max_clin_cvd", "max_hfpef", "max_ARRHYTHMIA", "max_CAD",
                 "max_CHF", "max_LVSD", "max_MI", "max_DVT", "max_PAD", "max_STROKE",
                 "max_TIA", "max_ULCER", "max_DEATH", "max_KDIGO_char", "max_HTN", "max_TGDLP",
                 "max_high_pwv", "max_high_bnp")

my_controls <- tableby.control(
  cat.simplify = TRUE
)

#simplify_list <- setNames(rep("1", length(binary_vars)), binary_vars)
#my_controls <- tableby.control(cat.simplify = simplify_list)
t2a <- tableby(CKM_max_char ~ max_clin_cvd + max_hfpef + max_ARRHYTHMIA + max_CAD + max_CHF + max_LVSD + max_MI + max_DVT + max_PAD + max_STROKE + 
                 max_TIA + max_ULCER + max_DEATH + max_KDIGO_char + max_HTN + max_TGDLP + max_high_pwv + max_high_bnp, 
              data = comorb_ckm2_criteria, total = FALSE, control = my_controls)
```

```{r results='asis'}
summary(t2a,  pfootnote = F)
```

## Descriptive statistics by maximum CKM stage

### Full table

```{r include=F}
# Complete table with both continuous and binary variables
t2data <- prog_summary_final %>% dplyr::mutate(age_num = as.numeric(AGEBASE),
                                               yrs_to_first_ckm_prog = days_to_first_ckm_prog/365.25,
                                               progress_CKM_char = case_when(progress_CKM == 0 ~ "Did not progress",
                                                    progress_CKM == 1 ~ "Progressed"))
label(t2data$yrs_to_first_ckm_prog) <- "Years to first progression event"
label(t2data$progress_CKM_char) <- "CKM progression"
t2data$UAlbCreat_mg_g <- t2data$UAlbCreat*1000

t2a <- t2data %>%
  gtsummary::tbl_summary(
    by = CKM_max_char,
    missing = "no",
    include = c(
      AGEBASE, sex_char, racedesc, bmi, txdesc, HbA1c, log_trig, sbp, dbp, 
      uacid, si_1_ins0, UAlbCreat_mg_g, HDL, codi, ALT, AST,
      HTN, TGDLP, MIC, MAC, GLYC, DEATH, CKM_syn_base, progress_CKM_char, years_to_first_ckm_prog_nocensor, num_prog, LIPO_Ala, LIPO_ApoA1, LIPO_ApoB, LIPO_BCAA, LIPO_Ctr, LIPO_DRF5, LIPO_Glu, LIPO_GlycA, LIPO_GxH13_I, LIPO_GxH5_I, LIPO_GxH6_I, LIPO_H1P, LIPO_H2P, LIPO_H3P, LIPO_H4P, LIPO_H5P, LIPO_H6P, LIPO_H7P, LIPO_HDLZ, LIPO_IRDRF, LIPO_Ileu, LIPO_LDLP, LIPO_LDLZ, LIPO_LGVX, LIPO_LPIR, LIPO_L_TRLP, LIPO_L_cHDLP, LIPO_L_cLDLP, LIPO_Leu, LIPO_M_TRLP, LIPO_M_cHDLP, LIPO_M_cLDLP, LIPO_NHDLC, LIPO_NLDLC, LIPO_NTC, LIPO_NTG, LIPO_NTRLC, LIPO_NTRLTG, LIPO_SDRF_I, LIPO_SDRF_M, LIPO_S_TRLP, LIPO_S_cHDLP, LIPO_S_cLDLP, LIPO_TRLP, LIPO_TRLZ, LIPO_VL_TRLP, LIPO_VS_TRLP, LIPO_Val, LIPO_cHDLP, homa_ir, wastcirc, waist_height, log_trig
    ),
    type = list(
      c(AGEBASE, years_to_first_ckm_prog_nocensor, HbA1c, log_trig, sbp, dbp, 
        uacid, si_1_ins0, UAlbCreat_mg_g, bmi, HDL, codi, ALT, AST, LIPO_Ala, LIPO_ApoA1, LIPO_ApoB, LIPO_BCAA, LIPO_Ctr, LIPO_DRF5, LIPO_Glu, LIPO_GlycA, LIPO_GxH13_I, LIPO_GxH5_I, LIPO_GxH6_I, LIPO_H1P, LIPO_H2P, LIPO_H3P, LIPO_H4P, LIPO_H5P, LIPO_H6P, LIPO_H7P, LIPO_HDLZ, LIPO_IRDRF, LIPO_Ileu, LIPO_LDLP, LIPO_LDLZ, LIPO_LGVX, LIPO_LPIR, LIPO_L_TRLP, LIPO_L_cHDLP, LIPO_L_cLDLP, LIPO_Leu, LIPO_M_TRLP, LIPO_M_cHDLP, LIPO_M_cLDLP, LIPO_NHDLC, LIPO_NLDLC, LIPO_NTC, LIPO_NTG, LIPO_NTRLC, LIPO_NTRLTG, LIPO_SDRF_I, LIPO_SDRF_M, LIPO_S_TRLP, LIPO_S_cHDLP, LIPO_S_cLDLP, LIPO_TRLP, LIPO_TRLZ, LIPO_VL_TRLP, LIPO_VS_TRLP, LIPO_Val, LIPO_cHDLP, homa_ir, wastcirc, waist_height, log_trig) ~ "continuous",
      c(HTN, TGDLP, MIC, MAC, GLYC, DEATH, progress_CKM_char) ~ "dichotomous"
    ),
    statistic = list(
      gtsummary::all_continuous() ~ "{mean} ({sd})",
      c(UAlbCreat_mg_g) ~ "{median} ({p25}, {p75})",
      gtsummary::all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      # Continuous variables
      AGEBASE ~ 1,
      years_to_first_ckm_prog_nocensor ~ 1,
      c(codi, si_1_ins0) ~ 3,
      c(HbA1c, log_trig, sbp, dbp, uacid, UAlbCreat_mg_g, bmi, HDL) ~ 1,
      c(ALT, AST) ~ 0,
      # Binary variables
      gtsummary::all_categorical() ~ c(0, 1)
    ),
    value = list(  # Show only "1" level for binary variables
      HTN ~ "1",
      TGDLP ~ "1",
      MIC ~ "1",
      MAC ~ "1",
      GLYC ~ "1",
      DEATH ~ "1",
      progress_CKM_char ~ "Progressed"
    )
  ) %>%
  gtsummary::add_p(test = list(gtsummary::all_continuous() ~ "aov",
      c(UAlbCreat_mg_g) ~ "wilcox.test",                         
      c(sex_char, HTN, TGDLP, MIC, MAC, GLYC, DEATH) ~ "fisher.test",
      c(racedesc, txdesc, CKM_syn_base ) ~ "chisq.test"  # Use chi-square instead of Fisher's exact
  )) %>%
  gtsummary::modify_header(label ~ "**Characteristic**") %>%
  gtsummary::bold_labels()

```

```{r}
t2a
```

### Brief table

```{r include=F, results='hide'}
t2b <- t2data %>%
  gtsummary::tbl_summary(
    by = CKM_max_char,
    missing = "no",
    include = c(
      AGEBASE, sex_char, racedesc, bmi, txdesc, HbA1c, 
      si_1_ins0, UAlbCreat_mg_g, 
      GLYC, DEATH, num_prog
    ),
    type = list(
      c(AGEBASE, HbA1c, 
        si_1_ins0, UAlbCreat_mg_g, bmi) ~ "continuous",
      c(GLYC, DEATH) ~ "dichotomous"
    ),
    statistic = list(
      # Specify median and IQR for UAlbCreat_mg_g
      UAlbCreat_mg_g ~ "{median} ({p25}, {p75})",
      # Mean and SD for other continuous variables
      c(AGEBASE, HbA1c, si_1_ins0, bmi) ~ "{mean} ({sd})",
      # Standard format for categorical variables
      gtsummary::all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      # Continuous variables
      AGEBASE ~ 1,
      c(HbA1c, UAlbCreat_mg_g, bmi) ~ 1,
      c(si_1_ins0) ~ 3,
      # Binary variables
      gtsummary::all_categorical() ~ c(0, 1)
    ),
    value = list(  # Show only "1" level for binary variables
      GLYC ~ "1",
      DEATH ~ "1"
    )
  ) %>%
  gtsummary::add_p(test = list(gtsummary::all_continuous() ~ "aov",  # or "aov" for ANOVA
      c(sex_char, GLYC, DEATH, num_prog) ~ "fisher.test",
      c(racedesc, txdesc) ~ "chisq.test"  # Use chi-square instead of Fisher's exact
  ))   %>%
  gtsummary::modify_header(label ~ "**Characteristic**") %>%
  gtsummary::bold_labels() %>%
  as_gt() %>%
  gt::cols_width(
    label ~ px(350),
    stat_1 ~ px(150),  # First group column
    stat_2 ~ px(150),  # Second group column
    stat_3 ~ px(150)
    #,  # Third group column (adjust based on your groups)
    #p.value ~ px(100)
  ) 
```

```{r include=F}
# Sankey diagram for CKM transitions
# save and incorporate into figures markdown
df_alluvial <- prog_summary_final %>%
  select(RELEASEID, CKM_syn_base, CKM_max_char) %>%
  mutate(CKM_syn_base = as.factor(CKM_syn_base), CKM_max = as.factor(CKM_max_char))
```

```{r}
p_sankey <- ggplot(df_alluvial, aes(axis1 = CKM_syn_base, axis2 = CKM_max_char)) +
  geom_alluvium(aes(fill = CKM_syn_base), width = 1/12, alpha = 0.8) +
  geom_stratum(width = 1/12, fill = "grey80", color = "black") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Baseline CKM Stage", "Follow-up CKM Stage"), expand = c(.05, .05)) +
  labs(title = "", y = "Number of Participants") +
  theme_minimal() + theme(legend.position = "none")
ggsave(
  filename = 'CKM_sankey.jpg', plot = p_sankey,
  width = 1600, height = 900, units = "px", scale = 2
)

# First, calculate transitions
transitions <- df_alluvial %>%
  group_by(CKM_syn_base, CKM_max_char) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(
    total = sum(n),
    percent = round(100 * n / total, 1),
    label = paste0(n, "\n(", percent, "%)")
  )

# Create the plot with flow labels
p_sankey <- ggplot(df_alluvial, aes(axis1 = CKM_syn_base, axis2 = CKM_max_char)) +
  geom_alluvium(aes(fill = CKM_syn_base), width = 1/12, alpha = 0.8) +
  geom_stratum(width = 1/12, fill = "grey80", color = "black") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  # Add flow labels
  geom_text(stat = "flow", aes(label = after_stat(n)),
            nudge_x = 0.0, size = 3) +
  scale_x_discrete(limits = c("Baseline CKM Stage", "Follow-up CKM Stage"), 
                   expand = c(.05, .05)) +
  labs(title = "", y = "Number of Participants") +
  theme_minimal() + 
  theme(legend.position = "none")

# Calculate flow percentages
df_flow_labels <- df_alluvial %>%
  count(CKM_syn_base, CKM_max_char) %>%
  group_by(CKM_syn_base) %>%
  mutate(
    pct = round(100 * n / sum(n), 1),
    label = paste0(n, "\n(", pct, "%)")
  )

# Create plot with centered flow labels
p_sankey <- ggplot(df_alluvial, 
                   aes(axis1 = CKM_syn_base, axis2 = CKM_max_char)) +
  geom_alluvium(aes(fill = CKM_syn_base), 
                width = 1/12, alpha = 0.8) +
  geom_stratum(width = 1/12, fill = "grey80", color = "black") +
  geom_text(stat = "stratum", 
            aes(label = after_stat(stratum)),
            size = 4) +
  # Add flow labels in the middle
  geom_text(stat = "flow",
            aes(label = paste0(after_stat(n), "(",
                              round(after_stat(prop) * 100, 1), "%)")),
            nudge_x = 0.2,  # This centers labels between stages
            size = 3,
            color = "black",
            fontface = "bold") +
  scale_x_discrete(limits = c("Baseline CKM Stage", "Follow-up CKM Stage"), 
                   expand = c(.05, .05)) +
  labs(title = "CKM Stage Transitions", 
       y = "Number of Participants") +
  theme_minimal() + 
  theme(legend.position = "none")

ggsave(filename = 'CKM_sankey.jpg', plot = p_sankey,
       width = 600, height = 900, units = "px", scale = 2)

## Trying a minimal version 

p_sankey <- ggplot(df_alluvial, 
                   aes(axis1 = CKM_syn_base, axis2 = CKM_max_char)) +
  geom_alluvium(aes(fill = CKM_syn_base), 
                width = 1/12, alpha = 0.8) +
  geom_stratum(width = 1/12, fill = "grey80", color = "black") +
  # Only label the baseline (first) stages
  geom_text(stat = "stratum", 
            aes(label = ifelse(after_stat(x) == "CKM_syn_base", 
                              after_stat(stratum), 
                              "")),
            size = 4) +
  # Flow labels in middle
  #geom_text(stat = "flow",
          #  aes(label = paste0(after_stat(n), " (", 
          #                    round(after_stat(prop) * 100, 1), "%)")),
           # nudge_x = 0.1,
            #size = 3,
           # color = "black",
        #    fontface = "bold") +
  #scale_x_discrete(limits = c("Baseline CKM Stage", "Follow-up CKM Stage"), 
   #                expand = c(.05, .05)) +
  labs(title = "", 
       y = "") +  # Remove y-axis title
  theme_minimal() + 
  theme(legend.position = "none",
        axis.text.y = element_blank(),  # Remove y-axis labels
        axis.ticks.y = element_blank(), # Remove y-axis ticks
        panel.grid = element_blank())   # Optional: remove gridlines

ggsave(filename = 'CKM_sankey.jpg', plot = p_sankey,
       width = 300, height = 900, units = "px", scale = 2)

```

```{r}
t2b
```

```{r, include=FALSE}
# Run univariable Cox models for time to first progression without BMI
result_noBMI <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(days_to_first_ckm_prog, as.numeric(progress_CKM))~",predictors[i])
  mod <- coxph(data = prog_summary_final_soma, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_noBMI <- rbind(result_noBMI,tidymod)
}
result_noBMI$adj.p.value <- p.adjust(result_noBMI$p.value, "fdr")
result_noBMI$AptName <- result_noBMI$term
result_noBMI$term <- NULL
result_noBMI <- merge(result_noBMI,analytes,by="AptName",all.x = T, all.y = F)
result_noBMI <- result_noBMI[order(result_noBMI$p.value),] 
```

```{r, include=FALSE}
# Run univariable Cox models for time to progression to stage 3
result_noBMI_3 <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(days_to_first_ckm_stage3_prog, as.numeric(progress_CKM_stage3))~",predictors[i])
  mod <- coxph(data = prog_summary_final_soma, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_noBMI_3 <- rbind(result_noBMI_3,tidymod)
}
result_noBMI_3$adj.p.value <- p.adjust(result_noBMI_3$p.value, "fdr")
result_noBMI_3$AptName <- result_noBMI_3$term
result_noBMI_3$term <- NULL
result_noBMI_3 <- merge(result_noBMI_3,analytes,by="AptName",all.x = T, all.y = F)
result_noBMI_3 <- result_noBMI_3[order(result_noBMI_3$p.value),] 
```

```{r, include=FALSE}
# Run univariable Cox models for time to progression to stage 4
result_noBMI_4 <- NULL
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(days_to_first_ckm_stage4_prog, as.numeric(progress_CKM_stage3))~",predictors[i])
  mod <- coxph(data = prog_summary_final_soma, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_noBMI_4 <- rbind(result_noBMI_4,tidymod)
}
result_noBMI_4$adj.p.value <- p.adjust(result_noBMI_4$p.value, "fdr")
result_noBMI_4$AptName <- result_noBMI_4$term
result_noBMI_4$term <- NULL
result_noBMI_4 <- merge(result_noBMI_4,analytes,by="AptName",all.x = T, all.y = F)
result_noBMI_4 <- result_noBMI_4[order(result_noBMI_4$p.value),] 
```

```{r, include=FALSE}
# Run univariable Cox models for clinical variables
result_clinical <- NULL
predictors <- c("AGEBASE", "sex_char", "bmi", "HbA1c", "sbp", "codi", "si_1_ins0", "GLYC", "mean_a1c", "mean_adipo", "mean_il6", "mean_il1",
                "mean_hscrp", "mean_apob", "mean_tnfa", "mean_insinv", "mean_codi", "tx")
for (i in 1:length(predictors)) {
  form <- paste0("Surv(days_to_first_ckm_prog, as.numeric(progress_CKM))~",predictors[i])
  mod <- coxph(data = prog_summary_final_soma, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  #tidymod <- tidymod[1,]
  tidymod <- cbind(tidymod, mod$concordance[[6]])
  result_clinical <- rbind(result_clinical,tidymod)
}
result_clinical$adj.p.value <- p.adjust(result_clinical$p.value, "fdr")
result_clinical$AptName <- result_clinical$term
result_clinical$term <- NULL
result_clinical$concordance <- result_clinical$`mod$concordance[[6]]`
result_clinical$`mod$concordance[[6]]` <- NULL
result_clinical <- merge(result_clinical,analytes,by="AptName",all.x = T, all.y = F)
#result_clinical <- result_clinical[order(result_clinical$p.value),] 
result_clinical$VarName <- result_clinical$AptName
result_clinical <- result_clinical %>% select(VarName, estimate, std.error,	statistic,	p.value,	conf.low,	conf.high, concordance)
```

```{r, include=FALSE}
# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"Proteins all stages")
writeData(wb,"Proteins all stages",result_noBMI,rowNames = F)
addWorksheet(wb,"Proteins CKM 3")
writeData(wb,"Proteins CKM 3",result_noBMI_3,rowNames = F)
addWorksheet(wb,"Proteins CKM 4")
writeData(wb,"Proteins CKM 4",result_noBMI_4,rowNames = F)
addWorksheet(wb,"Clinical all stages")
writeData(wb,"Clinical all stages",result_clinical,rowNames = F)
full_path <- file.path(data_path, "CKM/Results/TODAY CKM Cox models scaled baseline unadjusted.xlsx")
saveWorkbook(wb, full_path, overwrite = TRUE)
```

```{r, include=FALSE}
# compare change in proteins in those who did and did not progress
# using limma mixed effects model
soma_long_log$sample_id <- paste(soma_long_log$RELEASEID, "v", soma_long_log$visit, sep = "")
protein_cols <- grep("^seq\\.", names(soma_long_log), value = TRUE)
protein_data <- soma_long_log[, c("sample_id", protein_cols)]
expression_matrix <- t(protein_data[, -1]) 
colnames(expression_matrix) <- protein_data$sample_id
metadata <- soma_long_log[, c("sample_id", "RELEASEID", "visit", "progress_CKM")]
metadata <- metadata[!duplicated(metadata$sample_id), ]
rownames(metadata) <- metadata$sample_id
metadata <- metadata[colnames(expression_matrix), ]
metadata$visit_factor <- factor(metadata$visit)
metadata$progress_CKM <- factor(metadata$progress_CKM, levels = c(0, 1))
design <- model.matrix(~ progress_CKM * visit_factor, data = metadata)
colnames(design) <- c("Intercept", "ProgressCKM", "Visit2", "ProgressCKM_Visit2")
block <- factor(metadata$RELEASEID)
correlation <- duplicateCorrelation(expression_matrix, design, block = block)
fit <- lmFit(expression_matrix, design, block = block, 
             correlation = correlation$consensus.correlation)
fit <- eBayes(fit)
results_interaction <- topTable(fit, coef = "ProgressCKM_Visit2", adjust.method = "BH", number = Inf)
results_interaction$AptName <- row.names(results_interaction)
results_interaction <- merge(results_interaction,analytes,by="AptName",all.x = T, all.y = F)
results_interaction <- results_interaction[order(results_interaction$P.Value),] 
```

```{r, include=FALSE}
# write DiD results to file
wb <- createWorkbook()
addWorksheet(wb,"DiD proteins vs CKM progression")
writeData(wb,"DiD proteins vs CKM progression",results_interaction,rowNames = F)
full_path <- file.path(data_path, "CKM/Results/TODAY change in proteins CKM progression.xlsx")
saveWorkbook(wb, full_path, overwrite = TRUE)
```

```{r, include=FALSE}
# Elastic net Cox models

# filter to top 100 proteins by effect size
f <- file.path(data_path, "CKM/Results/TODAY CKM Cox models scaled baseline unadjusted.xlsx")

# CKM first progression to any stage
ckm_without_bmi_keep = read_excel(f,sheet = "Proteins all stages")
ckm_without_bmi_keep <- arrange(ckm_without_bmi_keep,adj.p.value)
ckm_without_bmi_drop <- ckm_without_bmi_keep[101:nrow(ckm_without_bmi_keep),"AptName"]
rownames(ckm_without_bmi_drop) <- ckm_without_bmi_drop$AptName
ckm_without_bmi_drop <- as.data.frame(t(ckm_without_bmi_drop))
ckm_without_bmi_endata <- prog_summary_final_soma %>% select(-colnames(ckm_without_bmi_drop))

# CKM time to progression for those who progressed to stage 3
ckm_stage3_keep = read_excel(f,sheet = "Proteins CKM 3")
ckm_stage3_keep <- arrange(ckm_stage3_keep,adj.p.value)
ckm_stage3_drop <- ckm_stage3_keep[101:nrow(ckm_stage3_keep),"AptName"]
rownames(ckm_stage3_drop) <- ckm_stage3_drop$AptName
ckm_stage3_drop <- as.data.frame(t(ckm_stage3_drop))
ckm_stage3_endata <- prog_summary_final_soma %>% select(-colnames(ckm_stage3_drop))

# CKM time to progression for those who progressed to stage 4
ckm_stage4_keep = read_excel(f,sheet = "Proteins CKM 4")
ckm_stage4_keep <- arrange(ckm_stage4_keep,adj.p.value)
ckm_stage4_drop <- ckm_stage4_keep[101:nrow(ckm_stage4_keep),"AptName"]
rownames(ckm_stage4_drop) <- ckm_stage4_drop$AptName
ckm_stage4_drop <- as.data.frame(t(ckm_stage4_drop))
ckm_stage4_endata <- prog_summary_final_soma %>% select(-colnames(ckm_stage4_drop))

set.seed(3654)
filename <- "TODAY proteomics metabolomics/Proteomics analyses/Elastic net/easy_elasticnet.R"
full_path <- file.path(github_path, filename)
source(full_path)
```

#Searching for target genes

```{r}
#Change gene name to check if exists in EntrexGeneSymbol column
analytes$EntrezGeneSymbol[which(grepl("RIPK2",analytes$EntrezGeneSymbol))]

#Change gene name
"RIPK2" %in% analytes$EntrezGeneSymbol

```

# Elastic net models

## Time to CKM progression, any stage

### Proteins

```{r, include=FALSE, comment=""}
ckm_without_bmi_endata_nozero <- ckm_without_bmi_endata[!ckm_without_bmi_endata$days_to_first_ckm_prog == 0,]
seq <- is_seq(names(ckm_without_bmi_endata_nozero))

ckm_without_bmi_select = easy_elasticnet(data = ckm_without_bmi_endata_nozero,outcome = "progress_CKM",
                           predictors = ckm_without_bmi_endata_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "days_to_first_ckm_prog", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% ckm_without_bmi_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(days_to_first_ckm_prog, as.numeric(progress_CKM))~",paste0(ckm_without_bmi_select,collapse = "+")))
mod = coxph(formula = form,data = ckm_without_bmi_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df <- as.data.frame(cbind("Time to CKM progression","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with only age and sex

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_prog, as.numeric(progress_CKM))~sex_char + AGEBASE"))
mod = coxph(formula = form,data = ckm_without_bmi_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression","Age, sex",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with age, sex, BMI, HbA1c, SBP

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_prog, as.numeric(progress_CKM))~sex_char + AGEBASE + bmi + HbA1c + sbp"))
mod = coxph(formula = form,data = ckm_without_bmi_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression","Age, sex, BMI, HbA1c, SBP",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with proteins, age, and sex

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_prog, as.numeric(progress_CKM))~",paste0(ckm_without_bmi_select,collapse = "+"),paste0("+sex_char+AGEBASE")))
mod = coxph(formula = form,data = ckm_without_bmi_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf[1:(nrow(moddf)-2),]$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression","Proteins, age, sex",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with proteins, age, sex, BMI, HbA1c, SBP

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_prog, as.numeric(progress_CKM))~",paste0(ckm_without_bmi_select,collapse = "+"),paste0("+HbA1c+log_trig+sbp+si_1_ins0+sex_char+bmi+AGEBASE+racedesc")))
mod = coxph(formula = form,data = ckm_without_bmi_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf[1:(nrow(moddf)-10),]$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression","Proteins + all covariates",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

```{r, include=TRUE, comment=""}
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE")
concordance_df$Concordance <- as.numeric(concordance_df$Concordance)
concordance_df$SE <- as.numeric(concordance_df$SE)
concordance_df$ci_low <- concordance_df$Concordance - (1.96*concordance_df$SE)
concordance_df$ci_high <- concordance_df$Concordance + (1.96*concordance_df$SE)
concordance_df[,c("Concordance","SE","ci_low", "ci_high")] <- round(concordance_df[,c("Concordance","SE","ci_low", "ci_high")],2)
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")
#concordance_df <- rbind(concordance_df[5,],concordance_df[7,],
#                            concordance_df[1,],
#                            concordance_df[4,],concordance_df[6,])
concordance_df <- concordance_df[,-1]
colnames(concordance_df) <- c("Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")

# forest plot of concordance
concordance_df_for_plot <- concordance_df
concordance_df_for_plot$` ` <- paste(rep(" ", 40), collapse = " ")
concordance_df_for_plot$`Concordance (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     concordance_df_for_plot$Concordance, concordance_df_for_plot$`Lower bound 95% CI`, concordance_df_for_plot$`Upper bound 95% CI`)
p <- forest(concordance_df_for_plot[, c(1, 6:7)], est = concordance_df_for_plot$Concordance, lower = concordance_df_for_plot$`Lower bound 95% CI`,
       upper = concordance_df_for_plot$`Upper bound 95% CI`, ci_column = 2, xlim = c(0,1))
#p$widths <- unit(c(1,100,50,50), "mm")
png('forest_plot_elastic_net_concordance_any_max_stage.png', res = 300, width = 9, height = 3, units = "in")
p
dev.off()

concordance_df <- NULL
concordance_df_temp <- NULL
concordance_df_for_plot <- NULL
```

## Time to CKM progression, for those who reached stage 3

### Proteins

```{r, include=FALSE, comment=""}
ckm_stage3_endata_nozero <- ckm_stage3_endata[!ckm_stage3_endata$days_to_first_ckm_stage3_prog == 0,]
seq <- is_seq(names(ckm_stage3_endata_nozero))

ckm_stage3_select = easy_elasticnet(data = ckm_stage3_endata_nozero,outcome = "progress_CKM_stage3",
                           predictors = ckm_stage3_endata_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "days_to_first_ckm_stage3_prog", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% ckm_stage3_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(days_to_first_ckm_stage3_prog, as.numeric(progress_CKM_stage3))~",paste0(ckm_stage3_select,collapse = "+")))
mod = coxph(formula = form,data = ckm_stage3_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df <- as.data.frame(cbind("Time to CKM progression, stage 3","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with only age and sex

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage3_prog, as.numeric(progress_CKM_stage3))~sex_char + AGEBASE"))
mod = coxph(formula = form,data = ckm_stage3_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 3","Age, sex",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with age, sex, BMI, HbA1c, SBP

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage3_prog, as.numeric(progress_CKM_stage3))~sex_char + AGEBASE + bmi + HbA1c + sbp"))
mod = coxph(formula = form,data = ckm_stage3_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 3","Age, sex, BMI, HbA1c, SBP",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with proteins, age, and sex

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage3_prog, as.numeric(progress_CKM_stage3))~",paste0(ckm_stage3_select,collapse = "+"),paste0("+sex_char+AGEBASE")))
mod = coxph(formula = form,data = ckm_stage3_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf[1:(nrow(moddf)-2),]$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 3","Proteins, age, sex",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with proteins, age, sex, BMI, HbA1c, SBP

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage3_prog, as.numeric(progress_CKM_stage3))~",paste0(ckm_stage3_select,collapse = "+"),paste0("+HbA1c+log_trig+sbp+si_1_ins0+sex_char+bmi+AGEBASE+racedesc")))
mod = coxph(formula = form,data = ckm_stage3_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf[1:(nrow(moddf)-10),]$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 3","Proteins + all covariates",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

```{r, include=TRUE, comment=""}
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE")
concordance_df$Concordance <- as.numeric(concordance_df$Concordance)
concordance_df$SE <- as.numeric(concordance_df$SE)
concordance_df$ci_low <- concordance_df$Concordance - (1.96*concordance_df$SE)
concordance_df$ci_high <- concordance_df$Concordance + (1.96*concordance_df$SE)
concordance_df[,c("Concordance","SE","ci_low", "ci_high")] <- round(concordance_df[,c("Concordance","SE","ci_low", "ci_high")],2)
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")
#concordance_df <- rbind(concordance_df[5,],concordance_df[7,],
#                            concordance_df[1,],
#                            concordance_df[4,],concordance_df[6,])
concordance_df <- concordance_df[,-1]
colnames(concordance_df) <- c("Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")

# forest plot of concordance
concordance_df_for_plot <- concordance_df
concordance_df_for_plot$` ` <- paste(rep(" ", 40), collapse = " ")
concordance_df_for_plot$`Concordance (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     concordance_df_for_plot$Concordance, concordance_df_for_plot$`Lower bound 95% CI`, concordance_df_for_plot$`Upper bound 95% CI`)
p <- forest(concordance_df_for_plot[, c(1, 6:7)], est = concordance_df_for_plot$Concordance, lower = concordance_df_for_plot$`Lower bound 95% CI`,
       upper = concordance_df_for_plot$`Upper bound 95% CI`, ci_column = 2, xlim = c(0,1))
#p$widths <- unit(c(1,100,50,50), "mm")
png('forest_plot_elastic_net_concordance_stage3.png', res = 300, width = 9, height = 3, units = "in")
p
dev.off()

concordance_df <- NULL
concordance_df_temp <- NULL
concordance_df_for_plot <- NULL
```

## Time to CKM progression, for those who reached stage 4

### Proteins

```{r, include=FALSE, comment=""}
ckm_stage4_endata_nozero <- ckm_stage4_endata[!ckm_stage4_endata$days_to_first_ckm_stage4_prog == 0,]
seq <- is_seq(names(ckm_stage4_endata_nozero))

ckm_stage4_select = easy_elasticnet(data = ckm_stage4_endata_nozero,outcome = "progress_CKM_stage4",
                           predictors = ckm_stage4_endata_nozero[,seq], out = "min.error",
                           model_type = "cox",time = "days_to_first_ckm_stage4_prog", cv_method="kfold", folds = 10)
retained_vars = analytes[analytes$AptName %in% ckm_stage4_select, c("Target","TargetFullName")]

form = as.formula(paste0("Surv(days_to_first_ckm_stage4_prog, as.numeric(progress_CKM_stage4))~",paste0(ckm_stage4_select,collapse = "+")))
mod = coxph(formula = form,data = ckm_stage4_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df <- as.data.frame(cbind("Time to CKM progression, stage 4","Proteins",mod$concordance[[6]], mod$concordance[[7]]))
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with only age and sex

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage4_prog, as.numeric(progress_CKM_stage4))~sex_char + AGEBASE"))
mod = coxph(formula = form,data = ckm_stage4_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 4","Age, sex",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with age, sex, BMI, HbA1c, SBP

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage4_prog, as.numeric(progress_CKM_stage4))~sex_char + AGEBASE + bmi + HbA1c + sbp"))
mod = coxph(formula = form,data = ckm_stage4_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 4","Age, sex, BMI, HbA1c, SBP",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with proteins, age, and sex

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage4_prog, as.numeric(progress_CKM_stage4))~",paste0(ckm_stage4_select,collapse = "+"),paste0("+sex_char+AGEBASE")))
mod = coxph(formula = form,data = ckm_stage4_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf[1:(nrow(moddf)-2),]$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 4","Proteins, age, sex",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

### Model with proteins, age, sex, BMI, HbA1c, SBP

```{r, include=FALSE, comment=""}
form = as.formula(paste0("Surv(days_to_first_ckm_stage4_prog, as.numeric(progress_CKM_stage4))~",paste0(ckm_stage4_select,collapse = "+"),paste0("+HbA1c+log_trig+sbp+si_1_ins0+sex_char+bmi+AGEBASE+racedesc")))
mod = coxph(formula = form,data = ckm_stage4_endata_nozero)
kable(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)

# make a summary table
moddf <- as.data.frame(tidy(mod, exponentiate = TRUE, conf.int = TRUE),digits = 3)
moddf[1:(nrow(moddf)-10),]$term <- analytes[analytes$AptName %in% moddf$term, ]$Target
moddf <- cbind(moddf$term,round(moddf$estimate, digits=3),round(moddf$std.error, digits=3),round(moddf$p.value, digits=3))
colnames(moddf) <- c("Variable","Hazard Ratio","SE","P-value")
concordance_df_temp <- as.data.frame(cbind("Time to CKM progression, stage 4","Proteins + all covariates",mod$concordance[[6]], mod$concordance[[7]]))
concordance_df <- rbind(concordance_df, concordance_df_temp)
```

```{r, include=TRUE, comment=""}
kable(moddf, row.names=F)
```

There are `r nrow(moddf)` variables in the model.

```{r, include=TRUE, comment=""}
kable(round(mod$concordance,digits=3))
```

```{r, include=TRUE, comment=""}
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE")
concordance_df$Concordance <- as.numeric(concordance_df$Concordance)
concordance_df$SE <- as.numeric(concordance_df$SE)
concordance_df$ci_low <- concordance_df$Concordance - (1.96*concordance_df$SE)
concordance_df$ci_high <- concordance_df$Concordance + (1.96*concordance_df$SE)
concordance_df[,c("Concordance","SE","ci_low", "ci_high")] <- round(concordance_df[,c("Concordance","SE","ci_low", "ci_high")],2)
colnames(concordance_df) <- c("Outcome","Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")
#concordance_df <- rbind(concordance_df[5,],concordance_df[7,],
#                            concordance_df[1,],
#                            concordance_df[4,],concordance_df[6,])
concordance_df <- concordance_df[,-1]
colnames(concordance_df) <- c("Model","Concordance","SE","Lower bound 95% CI","Upper bound 95% CI")

# forest plot of concordance
concordance_df_for_plot <- concordance_df
concordance_df_for_plot$` ` <- paste(rep(" ", 40), collapse = " ")
concordance_df_for_plot$`Concordance (95% CI)` <- sprintf("%.2f (%.2f to %.2f)",
                                     concordance_df_for_plot$Concordance, concordance_df_for_plot$`Lower bound 95% CI`, concordance_df_for_plot$`Upper bound 95% CI`)
p <- forest(concordance_df_for_plot[, c(1, 6:7)], est = concordance_df_for_plot$Concordance, lower = concordance_df_for_plot$`Lower bound 95% CI`,
       upper = concordance_df_for_plot$`Upper bound 95% CI`, ci_column = 2, xlim = c(0,1))
#p$widths <- unit(c(1,100,50,50), "mm")
png('forest_plot_elastic_net_concordance_stage4.png', res = 300, width = 9, height = 3, units = "in")
p
dev.off()

concordance_df <- NULL
concordance_df_temp <- NULL
concordance_df_for_plot <- NULL
```

```{r include=FALSE}
# GSEA for any progression
# for aptamers, latest version of fgsea throws an error
# keep the one with the largest summary statistic
ckm_without_bmi_keep_unique <- ckm_without_bmi_keep %>%
  group_by(EntrezGeneSymbol) %>%
  slice_max(abs(statistic), n = 1, with_ties = FALSE) %>%
  ungroup()
bg_path <- file.path(root_path, "GSEA/")
soma_fgsea <- run_fgsea_analysis(bg_path = bg_path,
                                     results_annotated = ckm_without_bmi_keep_unique,
                                     stat_col = "statistic")
soma_fgsea$summary
```

```{r echo = F}
# KEGG
plot_fgsea_transpose(soma_fgsea$kegg, 
                     title = "SOMA Top 30 KEGG",
                     xmax = 30)
ggsave("plasma_SOMA_KEGG.png",
       width = 27.5, height = 14, scale = 1)
# REACTOME
plot_fgsea_transpose(soma_fgsea$reactome, 
                     title = "SOMA Top 30 REACTOME", 
                     xmax = 30)
ggsave("plasma_SOMA_REACTOME.png",
       width = 27.5, height = 14, scale = 1)
# GO
plot_fgsea_transpose(soma_fgsea$go, 
                     title = "SOMA Top 30 GO", 
                     xmax = 30)
ggsave("plasma_SOMA_GO.png",
       width = 27.5, height = 14, scale = 1)
# HALLMARK
plot_fgsea_transpose(soma_fgsea$hallmark, 
                     title = "SOMA Top 30 Hallmark", 
                     xmax = 30)
ggsave("plasma_SOMA_Hallmark.png",
       width = 27.5, height = 14, scale = 1)
```

```{r include=FALSE}
# GSEA for progression to stage 3
ckm_stage3_keep_unique <- ckm_stage3_keep %>%
  group_by(EntrezGeneSymbol) %>%
  slice_max(abs(statistic), n = 1, with_ties = FALSE) %>%
  ungroup()
bg_path <- file.path(root_path, "GSEA/")
soma_fgsea_stage3 <- run_fgsea_analysis(bg_path = bg_path,
                                     results_annotated = ckm_stage3_keep_unique,
                                     stat_col = "statistic")
soma_fgsea_stage3$summary
```

```{r echo = F}
# KEGG
plot_fgsea_transpose(soma_fgsea_stage3$kegg, 
                     title = "SOMA Top 30 KEGG, Stage 3",
                     xmax = 30)
ggsave("plasma_SOMA_KEGG_stage3.png",
       width = 27.5, height = 14, scale = 1)
# REACTOME
plot_fgsea_transpose(soma_fgsea_stage3$reactome, 
                     title = "SOMA Top 30 REACTOME, Stage 3", 
                     xmax = 30)
ggsave("plasma_SOMA_REACTOME_stage3.png",
       width = 27.5, height = 14, scale = 1)
# GO
plot_fgsea_transpose(soma_fgsea_stage3$go, 
                     title = "SOMA Top 30 GO, Stage 3", 
                     xmax = 30)
ggsave("plasma_SOMA_GO_stage3.png",
       width = 27.5, height = 14, scale = 1)
# HALLMARK
plot_fgsea_transpose(soma_fgsea_stage3$hallmark, 
                     title = "SOMA Top 30 Hallmark, stage 3", 
                     xmax = 30)
ggsave("plasma_SOMA_Hallmark_stage3.png",
       width = 27.5, height = 14, scale = 1)
```

```{r include=FALSE}
# GSEA for progression to stage 4
ckm_stage4_keep_unique <- ckm_stage4_keep %>%
  group_by(EntrezGeneSymbol) %>%
  slice_max(abs(statistic), n = 1, with_ties = FALSE) %>%
  ungroup()
bg_path <- file.path(root_path, "GSEA/")
soma_fgsea_stage4 <- run_fgsea_analysis(bg_path = bg_path,
                                     results_annotated = ckm_stage4_keep_unique,
                                     stat_col = "statistic")
soma_fgsea_stage4$summary
```

```{r echo = F}
# KEGG
plot_fgsea_transpose(soma_fgsea_stage4$kegg, 
                     title = "SOMA Top 30 KEGG, Stage 4",
                     xmax = 30)
ggsave("plasma_SOMA_KEGG_stage4.png",
       width = 27.5, height = 14, scale = 1)
# REACTOME
plot_fgsea_transpose(soma_fgsea_stage4$reactome, 
                     title = "SOMA Top 30 REACTOME, Stage 4", 
                     xmax = 30)
ggsave("plasma_SOMA_REACTOME_stage4.png",
       width = 27.5, height = 14, scale = 1)
# GO
plot_fgsea_transpose(soma_fgsea_stage4$go, 
                     title = "SOMA Top 30 GO, Stage 4", 
                     xmax = 30)
ggsave("plasma_SOMA_GO_stage4.png",
       width = 27.5, height = 14, scale = 1)
# HALLMARK
plot_fgsea_transpose(soma_fgsea_stage4$hallmark, 
                     title = "SOMA Top 30 Hallmark, stage 4", 
                     xmax = 30)
ggsave("plasma_SOMA_Hallmark_stage4.png",
       width = 27.5, height = 14, scale = 1)
```

```{r clean data,echo=FALSE}
# Read in Teen-LABS data
# source("/Users/timvigers/GitHub/CHCO-Code/Petter Bjornstad/Teen-LABS/create_teen_labs_analysis_dataset.R")
filename <- "Data_Cleaned/analysis_dataset.RData"
full_path <- file.path(teen_labs_path, filename)
load(full_path)
tldata <- df

# keep top proteins
#load("/Volumes/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Petter Bjornstad/Proteomics and DKD/Data_Cleaned/analysis_dataset.RData")
#df <- df %>% filter(visit == "baseline")
# select top 20 aptamers
#top_glyc_df <- top_glyc_df %>% arrange(adj.p.value) %>% slice_head(n = 10)
#seq_include <- top_glyc_df$AptName 

seq_include <- analyte_info %>% filter(AptName %in% c("seq.9116.28","seq.18386.36","seq.4135.84","seq.6918.183", "seq.24975.3"))
apt_keep <- seq_include$AptName

# read in estimated marginal means
filename <- "Results/Teen LABS emmeans.xlsx"
full_path <- file.path(teen_labs_path, filename)
emmdf <- read_xlsx(full_path)
emmdf_keep <- emmdf %>% filter(AptName %in% apt_keep)
emmdf_keep$uniqname <- paste0(emmdf_keep$Target,"_",emmdf_keep$AptName)  
```

```{r,echo=FALSE}
p <- ggplot(emmdf_keep, aes(x=visit, y=emmean, colour=Target, group=uniqname)) +
    geom_line() + 
    geom_errorbar(aes(ymin=emmean-SE, ymax=emmean+SE), width=.1) +  
    xlab("") +
    ylab("Estimated marginal mean") + theme_bw() 
png('Teen-LABS_changes_top_TODAY_CKM_proteins.png', res = 300, width = 10, height = 10, units = "in")
p
dev.off()

p <- NULL
# loop through top 10 proteins and make separate plot for each
for (i in apt_keep) {
  p[[i]] <- ggplot(emmdf_keep[emmdf_keep$AptName == i,], aes(x=visit, y=emmean, colour=Target, group=uniqname)) +
    geom_line(size = 1.5) + 
    geom_errorbar(aes(ymin=emmean-SE, ymax=emmean+SE), width=.2) +  
    xlab("") +
    ylab("Mean") + theme_bw() + theme(legend.position = "top") + scale_color_discrete(name="")
}

panel_p <- ggarrange(p[["seq.9116.28"]],p[["seq.18386.36"]],p[["seq.4135.84"]],p[["seq.6918.183"]],p[["seq.24975.3"]],
                     ncol = 2, nrow = 5)
png('Teen-LABS_changes_top_TODAY_CKM_proteins.png', res = 300, width = 8, height = 10, units = "in")
panel_p
dev.off()
```