---
title: "Overall Data Summary"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

```{r include = F}
library(jsonlite)
library(dplyr)
library(kableExtra)
library(knitr)
library(purrr)
library(tidyr)
library(ggplot2)
library(GGally)
library(gtExtras)
library(gt)
library(purrr)
library(Hmisc)
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/pylell/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}
harm_dat_raw <- read.csv(file.path(root_path, "Data Harmonization/Data Clean/soma_harmonized_dataset.csv"), na = "")
```

```{r, include = F}
harm_dat <- harm_dat_raw %>%
  dplyr::mutate(sex = case_when(study == "ULTRA" ~ "Male", T ~ sex)) %>%
  dplyr::select(-starts_with("seq"), seq.10000.28, seq.10000.28_urine)

harm_dat_recid <- harm_dat %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id, visit)) %>%
  dplyr::select(-mrn, -dob)  %>%
  filter(visit %nin% c("treatment_period_1", "treatment_period_2",
                       "treatment_period_3", "treatment_period_4", 
                       "Early Withdrawal", "Unscheduled 1", "Unscheduled 2")) %>%
  bind_rows(data.frame(study = c(rep("ADAPT-CKD", 2),
                                 rep("MANATEE", 1),
                                 rep("INSPIRE", 1),
                                 rep("SUGAR-N-SALT", 1),
                                 rep("T1-DISCO", 1),
                                 rep("MetMod", 1),
                                 rep("OTTER", 1)),
                       visit = c("MO6", "YR1", 
                                 "MO4",
                                 "MO3", 
                                 "YR3",
                                 "MO6",
                                 "MO6",
                                 "MO6"),
                       group = c("Type 2 Diabetes", "Obese Control",
                                 "Type 1 Diabetes",
                                 "Type 2 Diabetes",
                                 "Type 1 Diabetes",
                                 "Type 1 Diabetes",
                                 "Type 1 Diabetes",
                                 "Type 1 Diabetes"))) %>%
  dplyr::mutate(study_design = case_when(study %in% c("RENAL-HEIR", "RENAL-HEIRitage", "PENGUIN",
                                                      "CROCODILE", "SWEETHEART", "CASPER") ~ "Observational: Cross-sectional",
                                         study %in% c("ULTRA", "COFFEE") ~ "Interventional: Longitudinal",
                                         study %in% c("PANTHER") ~ "Observational: Accelerated Longitudinal",
                                         study %in% c("ADAPT-CKD", "PANDA", "IMPROVE") ~ "Observational: Longitudinal",
                                         study %in% c("ATTEMPT", "RPC2", "MANATEE", "T1-DISCO", "MetMod", 
                                                      "OTTER", "SUGAR-N-SALT", "INSPIRE") ~ "Interventional: RCT"),
                exposure_design = case_when(study %in% c("COFFEE", "ULTRA", "ATTEMPT", "INSPIRE", "MANATEE", "MetMod",
                                                         "OTTER", "RPC2", "SUGAR-N-SALT", "T1-DISCO") ~ "Interventional",
                                            T ~ "Observational"),
                design_struc = case_when(study %in% c("COFFEE", "ULTRA", "ATTEMPT", "INSPIRE", "MANATEE", "MetMod",
                                                      "OTTER", "RPC2", "SUGAR-N-SALT", "T1-DISCO", "ADAPT-CKD",
                                                      "IMPROVE", "PANDA") ~ "Longitudinal",
                                         study == "PANTHER" ~ "Longitudinal (Accelerated)",
                                         T ~ "Cross-Sectional"),
                study_design_sub = case_when(study %in% c("ATTEMPT", "INSPIRE", "MANATEE", "MetMod", "OTTER", "RPC2", 
                                                          "SUGAR-N-SALT", "T1-DISCO") ~ "RCT",
                                             study %in% c("COFFEE", "ULTRA") ~ "Single-arm trial",
                                             study %in% c("IMPROVE", "ADAPT-CKD") ~ "Registry cohort",
                                             study %in% c("PANTHER", "PANDA") ~ "Prospective cohort", 
                                             T ~ "Cohort study"),
                visit = case_when(visit == "screening" ~ "baseline",
                                  study == "ATTEMPT" & visit == "follow_up" ~ "4_months_post",
                                  is.na(visit) & study_design == "Cross-sectional" ~ "baseline",
                                  is.na(visit) ~ "follow up",
                                  T ~ visit),
                intervention = case_when(study == "COFFEE" ~ "Caffeine (205mg/day)",
                                         study == "IMPROVE" ~ "Bariatric Surgery (VSG)",
                                         study == "ULTRA" ~ "Pegloticase (Krystexxa 8mg/once)",
                                         study == "ATTEMPT" ~ "SGLT2i (Dapagliflozin 5mg/day)",
                                         study == "RPC2" ~ "SGLT2i (Canagliflozin 10mg/day)",
                                         study == "MANATEE" ~ "AID System + Metformin (2000mg/day)",
                                         study == "INSPIRE" ~ "IMST (75% PImax, 30 breaths/day, 6 days/wk)",
                                         study == "SUGAR-N-SALT" ~ "SGLT2i (Sotagliflozin 200mg/day)",
                                         study == "T1-DISCO" ~ "GLP-1RA (Semaglutide 1mg/week)",
                                         study == "MetMod" ~ "AMX0035 (Phenylbutyrate + TUDCA)",
                                         study == "ADAPT-CKD" ~ "SGLT2i, GLP-1RA, OR ns-MRA",
                                         study == "OTTER" ~ "ns-MRA (Finerenone 20mg/day)")) %>%
  filter(!is.na(study_design)) %>%
  filter(study != "RPC2" | visit != "follow up") 


# Deidentify data for LLM
# harm_dat_recid <- harm_dat_recid[, !grepl("date", names(harm_dat_recid), ignore.case = TRUE)]
# write.csv(harm_dat_recid, file.path(root_path, "Ye Ji Choi/harm_data_deidentified_022326.csv"), row.names = F, na = "")
```

# Data availability by study

```{r}
# ── 1. Variable lists per procedure domain ────────────────────────────────────
cols <- names(harm_dat_recid)

# ── IVGTT (Intravenous Glucose Tolerance Test — MinMod) ─────────────────────
ivgtt <- c(
  # MinMod-derived outcomes
  "mm_si",       # Insulin Sensitivity (Si)
  "mm_airg",     # AIRg (Acute Insulin Response to Glucose)
  "mm_di",       # Disposition Index
  "mm_ir",       # Insulin Resistance
  "mm_bcell",    # Beta Cell Function
  "mm_sg",       # Glucose Effectiveness (Sg, min⁻¹)
  "mm_gb",       # Basal Glucose (mg/dL)
  "mm_ib",       # Basal Insulin (µU/mL)
  "mm_r2",       # MinMod R-squared (goodness of fit)
  # Endocrine labs drawn during IVGTT
  "lh",          # LH (mIU/mL)
  "fsh",         # FSH (mIU/mL)
  "igf1_z_score",# IGF-1 Z-score
  "bl_androst",  # Androstenedione (ng/dL)
  "bl_dheas"     # DHEA-S (ng/dL)
)

kidney_pet <- c(
  "avg_c_k1",
  "avg_k_k1",
  "avg_c_k2",
  "avg_k_k2",
  "avg_c_f",
  "avg_k_f"
)

blood_labs <- intersect(c(
  "glucose","hba1c","fasting_insulin","cpeptide","homa_ir","creatinine_s",
  "cystatin_c_s","bun_s","na_s","cholesterol","hdl","ldl","triglycerides",
  "hemoglobin","hematocrit","gotast_base","gptalt_base","alkphos_base",
  "bili_base","alb_base","tot_protein","bicarb_base","ca_base","cl_base",
  "k_base","sodium_base","lh","fsh","bl_androst","bl_dheas","copeptin",
  "igf1_z_score","alb","alk_phos","bili","bun","bg"
), cols)

urine_labs <- intersect(c(
  "creatinine_u","microalbumin_u","acr_u","elevated_albuminuria",
  "albuminuria_cat","na_u","urine_osm","u24_vl","u24_mab","u24_na",
  "u24_gluc","acr"
), cols)

clamp_outcomes <- intersect(c(
  "m_value","gir","m_ffm","hepatic_glucose_output","steady_state_glucose",
  "steady_state_insulin","glucose_infusion_rate_ss","endogenous_glucose_production",
  "hgc_bg_ss","hgc_insulin_ss","hgc_gir_ss","hgc_cpeptide_ss"
), cols)
clamp_ts_vars <- grep("^(glucose|insulin|cpeptide|ffa|d20_infusion_rate)_\\d",
                      cols, value = TRUE)
clamp_all <- unique(c(clamp_outcomes, clamp_ts_vars))

renal_clearance <- intersect(c(
  "gfr_bsa_plasma","gfr_bsa_plasma_urine","gfr_raw_plasma","gfr_raw_plasma_urine",
  "gfr_ecv_std","erpf_bsa_plasma","erpf_bsa_plasma_urine","erpf_raw_plasma",
  "erpf_raw_plasma_urine","rbf","ff","ra","re","rvr","pg","deltapf",
  "glomerular_pressure","kfg","gfr_p1","gfr_p2","pah_bsa","pah_raw",
  "plasma_ioh_1","plasma_ioh_2"
), cols)

bold_mri <- intersect(c(
  "bold_l_bl_cortex","bold_l_bl_medulla","bold_r_bl_cortex","bold_r_bl_medulla",
  "bold_l_pf_cortex","bold_l_pf_medulla","bold_r_pf_cortex","bold_r_pf_medulla",
  "avg_k_r2","avg_c_r2","avg_m_r2","fsoc_l_cortex","fsoc_r_cortex",
  "left_cortex_ox_metab_mean","right_cortex_ox_metab_mean"
), cols)

dxa <- intersect(c(
  "dexa_body_fat","dexa_lean_mass","dexa_lean_kg","dexa_fat_kg","dexa_trunk_mass",
  "dexa_trunk_kg","dexa_ag_ratio","dexa_est_vat","dexa_bone_mineral_density"
), cols)

kidney_bx <- intersect(c(
  "gloms","gloms_gs","glom_volume_con","glom_volume_weibel","glom_volume_wiggins",
  "glom_tuft_area","glom_nuc_count","fia","acan"
), cols)

cardiac_mri <- intersect(c(
  "lvedv","lvedvi","lvesv","lvesvi","lvef","lvsv","lv_myo_mass_dias",
  "lv_myo_mass_syst","lv_cardiac_output","lv_cardiac_index","lv_stroke_volume",
  "rvedv","rvef","rvsv","cardiac_output","ef_percent","edv_ml","esv_ml",
  "sv_ml","lved_mass","lved_massi"
), cols)

kidney_size_vars <- intersect(c(
  "left_kidney_volume_ml","right_kidney_volume_ml","volume_left_manual",
  "volume_right_manual","length_left","length_right","httkv",
  "left_kidney_length","right_kidney_length","total_kidney_volume"
), cols)
kidney_perf_vars <- grep("^(pcasl|pasl|asl).*_(left|right)", cols, value = TRUE)
kidney_mri_all   <- unique(c(kidney_size_vars, kidney_perf_vars, bold_mri))

# lipidomics: lipid species columns
lipid_vars <- grep("^(lpc|pc|pe|sm|cer|dg|tg_|mg|lpe|fa_|hex|lac)_",
                   cols, value = TRUE)

# proteomics (not in current extract — patterns retained for completeness)
prot_blood_vars <- intersect(c("seq.10000.28"), cols)
prot_urine_vars <-intersect(c("seq.10000.28_urine"), cols)
```

```{r}
# ── 2. Helper: does a study have ≥1 non-missing value in a domain? ─────────
has_any <- function(study_name, varlist) {
  present <- intersect(varlist, cols)
  if (length(present) == 0) return(FALSE)
  sub <- harm_dat_recid[!is.na(harm_dat_recid$study) & harm_dat_recid$study == study_name, present, drop = FALSE]
  any(rowSums(!is.na(sub)) > 0)
}

# ── 3. Build procedure matrix for studies present in the data ─────────────
studies_in_data <- sort(unique(na.omit(harm_dat_recid$study)))

proc_matrix <- lapply(studies_in_data, function(s) {
  sub   <- harm_dat_recid[!is.na(harm_dat_recid$study) & harm_dat_recid$study == s, ]
  n_pts <- length(unique(subset(sub, visit == "baseline")$record_id))
  groups <- sort(unique(na.omit(sub$group)))
  
  list(
    study           = s,
    n               = n_pts,
    groups          = groups,
    exposure_design    = unique(sub$exposure_design),
    design_struc = unique(sub$design_struc),
    study_design_sub = unique(sub$study_design_sub),
    intervention    = unique(sub$intervention),
    visits          = unique(subset(sub, visit != "baseline")$visit),
    blood_labs      = has_any(s, blood_labs),
    urine_labs      = has_any(s, urine_labs),
    clamp           = has_any(s, clamp_all),
    renal_clear     = has_any(s, renal_clearance),
    dxa             = has_any(s, dxa),
    kidney_bx       = has_any(s, kidney_bx),
    cardiac_mri     = has_any(s, cardiac_mri),
    kidney_mri      = has_any(s, kidney_mri_all),
    lipidomics      = has_any(s, lipid_vars),
    blood_proteomics      = has_any(s, prot_blood_vars),
    urine_proteomics      = has_any(s, prot_urine_vars),
    kidney_pet = has_any(s, kidney_pet),
    ivgtt = has_any(s, ivgtt)
  )
})

proc_harm_dat_recid <- bind_rows(lapply(proc_matrix, function(x) {
  as.data.frame(x[names(x) %nin% c("groups", "visits")], stringsAsFactors = FALSE)
})) %>%
  mutate(kidney_bx = case_when(study %in% c("ATTEMPT", "RPC2") ~ T, T ~ kidney_bx),
         intervention = case_when(is.na(intervention) ~ "", T ~ intervention),
         n = as.character(n),
         n = case_when(study == "ADAPT-CKD" ~ "≤100",
                       study == "INSPIRE" ~ "≤90",
                       study == "MANATEE" ~ "≤60",
                       study == "MetMod" ~ "≤60",
                       study == "OTTER" ~ "≤60",
                       study == "SUGAR-N-SALT" ~ "≤150",
                       study == "SWEETHEART" ~ "≤30",
                       study == "T1-DISCO" ~ "≤60",
                       T ~ n),
         vitals = T,
         blood_labs = T,
         urine_labs = T,
         sphyg = case_when(study %in% c("INSPIRE", "MANATEE", "OTTER", 
                                        "T1-DISCO", "RENAL-HEIRitage") ~ T,
                           T ~ F),
         cgm = case_when(study %in% c("MANATEE", "MetMod", "PANDA", 
                                      "PANTHER", "T1-DISCO") ~ T,
                         T ~ F),
         jwire = case_when(study %in% c("INSPIRE", "MANATEE", "OTTER", 
                                        "T1-DISCO", "RENAL-HEIRitage") ~ T,
                           T ~ F),
         muscle_bx = case_when(study %in% c("MetMod") ~ T,
                               T ~ F),
         liver_bx = case_when(study %in% c("IMPROVE") ~ T,
                              T ~ F),
         oroboros = case_when(study %in% c("IMPROVE", "MetMod") ~ T,
                              T ~ F),
         blood_metabolomics = case_when(study %in% c("RENAL-HEIR", "IMPROVE", 
                                                     "RENAL-HEIRitage") ~ T,
                                        T ~ F),
         urine_metabolomics = case_when(study %in% c("RENAL-HEIR", "IMPROVE", 
                                                     "PANDA", "RENAL-HEIRitage") ~ T,
                                        T ~ F),
         clamp = case_when(study == "ADAPT-CKD" ~ F,
                           study == "INSPIRE" ~ F,
                           study == "MANATEE" ~ T,
                           study == "MetMod" ~ T,
                           study == "OTTER" ~ F,
                           study == "SUGAR-N-SALT" ~ F,
                           study == "SWEETHEART" ~ F,
                           study == "T1-DISCO" ~ T,
                           study == "PANTHER" ~ F,
                           T ~ clamp),
         renal_clear = case_when(study == "ATTEMPT" ~ T,
                                 study == "ADAPT-CKD" ~ F,
                                 study == "INSPIRE" ~ F,
                                 study == "MANATEE" ~ T,
                                 study == "MetMod" ~ F,
                                 study == "OTTER" ~ T,
                                 study == "SUGAR-N-SALT" ~ F,
                                 study == "SWEETHEART" ~ F,
                                 study == "T1-DISCO" ~ F,
                                 T ~ renal_clear),
         dxa = case_when(study == "ADAPT-CKD" ~ F,
                         study == "INSPIRE" ~ F,
                         study == "MANATEE" ~ T, # BIS
                         study == "MetMod" ~ T,
                         study == "OTTER" ~ T,
                         study == "SUGAR-N-SALT" ~ F,
                         study == "SWEETHEART" ~ T,
                         study == "T1-DISCO" ~ T,
                         T ~ dxa),
         kidney_bx = case_when(study == "ADAPT-CKD" ~ T,
                               study == "INSPIRE" ~ F,
                               study == "MANATEE" ~ F,
                               study == "MetMod" ~ F,
                               study == "OTTER" ~ F,
                               study == "SUGAR-N-SALT" ~ F,
                               study == "SWEETHEART" ~ F,
                               study == "T1-DISCO" ~ F,
                               T ~ kidney_bx),
         cardiac_mri = case_when(study == "ADAPT-CKD" ~ F,
                                 study == "INSPIRE" ~ F,
                                 study == "MANATEE" ~ F,
                                 study == "MetMod" ~ F,
                                 study == "OTTER" ~ T,
                                 study == "SUGAR-N-SALT" ~ F,
                                 study == "SWEETHEART" ~ T,
                                 study == "T1-DISCO" ~ F,
                                 T ~ cardiac_mri),
         kidney_mri = case_when(study == "ADAPT-CKD" ~ F,
                                study == "INSPIRE" ~ F,
                                study == "MANATEE" ~ F,
                                study == "MetMod" ~ T,
                                study == "OTTER" ~ T,
                                study == "SUGAR-N-SALT" ~ F,
                                study == "SWEETHEART" ~ F,
                                study == "T1-DISCO" ~ F,
                                T ~ kidney_mri),
         lipidomics = case_when(study == "ADAPT-CKD" ~ F,
                                study == "INSPIRE" ~ F,
                                study == "MANATEE" ~ F,
                                study == "MetMod" ~ F,
                                study == "OTTER" ~ F,
                                study == "SUGAR-N-SALT" ~ F,
                                study == "SWEETHEART" ~ F,
                                study == "T1-DISCO" ~ F,
                                T ~ lipidomics),
         blood_proteomics = case_when(study == "ADAPT-CKD" ~ T,
                                      study == "INSPIRE" ~ F,
                                      study == "MANATEE" ~ F,
                                      study == "MetMod" ~ F,
                                      study == "OTTER" ~ T,
                                      study == "SUGAR-N-SALT" ~ F,
                                      study == "SWEETHEART" ~ F,
                                      study == "T1-DISCO" ~ F,
                                      T ~ blood_proteomics),
         urine_proteomics = case_when(study == "ADAPT-CKD" ~ T,
                                      study == "INSPIRE" ~ F,
                                      study == "MANATEE" ~ F,
                                      study == "MetMod" ~ F,
                                      study == "OTTER" ~ F,
                                      study == "SUGAR-N-SALT" ~ F,
                                      study == "SWEETHEART" ~ F,
                                      study == "T1-DISCO" ~ F,
                                      T ~ urine_proteomics),
         decode = case_when(study %in% c("CROCODILE", "RENAL-HEIR", 
                                         "RENAL-HEIRitage", "RPC2",
                                         "PANDA", "CASPER", "COFFEE",
                                         "PENGUIN", "T1-DISCO", "REMODEL-T1D",
                                         "ATTEMPT", "PANTHER") ~ "DECODE",
                            T ~ "NOT PART OF DECODE")
  )

# Define all procedures
flagged_procs <- c("vitals", "blood_labs", "urine_labs", "dxa",
                   "clamp", "ivgtt", "oroboros", 
                   "renal_clear", "kidney_mri", "kidney_pet", 
                   "sphyg",  "cardiac_mri", 
                   "kidney_bx", "muscle_bx", "liver_bx", "jwire",    
                   "lipidomics", "blood_proteomics", 
                   "urine_proteomics", "blood_metabolomics", "urine_metabolomics",
                   "cgm")
# Add cohort string
groups_list <- lapply(proc_matrix, `[[`, "groups")
names(groups_list) <- proc_harm_dat_recid$study

visits_list <- lapply(proc_matrix, `[[`, "visits")
names(visits_list) <- proc_harm_dat_recid$study
visits_list$COFFEE <- "WK1"

decode_list <- as.list(proc_harm_dat_recid$decode)
names(decode_list) <- proc_harm_dat_recid$study


# Map group labels to abbreviations
abbrev_group <- function(g) {
  dplyr::case_when(
    grepl("Type 1",  g, ignore.case = TRUE) ~ "T1D",
    grepl("Type 2",  g, ignore.case = TRUE) ~ "T2D",
    grepl("^Lean",   g, ignore.case = TRUE) ~ "HC",
    grepl("^Obese",  g, ignore.case = TRUE) ~ "OB",
    grepl("PKD|ADPKD", g, ignore.case = TRUE) ~ "ADPKD",
    TRUE ~ g
  )
}

# Map visit labels to abbreviations
abbrev_visit <- function(v) {
  dplyr::case_when(
    grepl("baseline",   v, ignore.case = TRUE) ~ "0",
    grepl("4_weeks_post",  v, ignore.case = TRUE) ~ "MO1",
    grepl("3_months_post_surgery",  v, ignore.case = TRUE) ~ "MO3",
    grepl("4_months_post",  v, ignore.case = TRUE) ~ "MO4",
    grepl("12_months_post_surgery", v, ignore.case = TRUE) ~ "YR1",
    grepl("year_1",  v, ignore.case = TRUE) ~ "YR1",
    grepl("year_2", v, ignore.case = TRUE) ~ "YR2",
    grepl("year_3",  v, ignore.case = TRUE) ~ "YR3",
    grepl("year_4", v, ignore.case = TRUE) ~ "YR4",
    grepl("follow up", v, ignore.case = TRUE) ~ "WK1",
    grepl("post_treatment", v, ignore.case = TRUE) ~ "MO6",
    TRUE ~ v
  )
}
```

```{r}
# ── 4. HTML helpers ───────────────────────────────────────────────────────────
cohort_colors <- c(
  T1D   = "#e77c76",
  T2D   = "#e37902",
  HC    = "#8cb369",
  OB    = "#5b8e7d",
  ADPKD = "#bc4b51"
)

visit_colors <- c(
  `0`   = "#d6ccc2",
  `WK1` = "#f8ad9d", 
  `MO1`   = "#ff9e00",
  `MO3`    = "#ff9100",
  `MO4`    = "#ff8500",
  `MO6` = "#ff6d00",
  `YR1`   = "#64a4ce",
  `YR2`    = "#4893c6",
  `YR3`    = "#2d82bd",
  `YR4` = "#2476b1"
)

decode_colors <- c(
  "DECODE"   = "#00314D",
  "NOT PART OF DECODE" = "#c1121f"
)

badge_html <- function(label) {
  bg <- cohort_colors[label]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.8em;font-weight:600;',
           'padding:2px 7px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, label
  )
}

badge_visit_html <- function(label) {
  bg <- visit_colors[label]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.8em;font-weight:600;',
           'padding:2px 7px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, label
  )
}

badge_decode_html <- function(label) {
  bg <- decode_colors[label]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.6em;font-weight:600;',
           'padding:2px 7px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, label
  )
}

make_cohort_html <- function(study_name) {
  gs   <- groups_list[[study_name]]
  abbr <- unique(abbrev_group(gs))
  paste(vapply(abbr, badge_html, character(1)), collapse = "")
}

make_visit_html <- function(study_name) {
  vs   <- visits_list[[study_name]]
  abbr <- sort(unique(abbrev_visit(vs)))
  paste(vapply(abbr, badge_visit_html, character(1)), collapse = "")
}

make_decode_html <- function(study_name) {
  dc   <- decode_list[[study_name]]
  paste(vapply(sort(unique(dc)), badge_decode_html, character(1)), collapse = "")
}

dot_yes <- '<span style="color:#cb1b16;font-size:1.5em;">&#9679;</span>'
dot_no  <- '<span style="color:#d1d5db;font-size:1.5em;">&#9675;</span>'
dot     <- function(x) ifelse(x, dot_yes, dot_no)

# ── 5. Assemble display data frame ────────────────────────────────────────────
display_harm_dat_recid <- proc_harm_dat_recid %>%
  mutate(
    cohorts_html = sapply(study, make_cohort_html),
    visits_html = sapply(study, make_visit_html),
    decode_html = sapply(study, make_decode_html),
    across(flagged_procs,
           ~ sapply(.x, dot))
  ) %>%
  select(decode_html, study, exposure_design, design_struc, study_design_sub, 
         intervention, cohorts_html, visits_html, n,
         flagged_procs) %>%
  arrange(decode_html, exposure_design, desc(design_struc), study_design_sub)
```

```{r}
# ── 6. Build gt table ─────────────────────────────────────────────────────────
study_gt_table <- display_harm_dat_recid %>%
  gt(id = "harm-study-table") %>%
  
  cols_label(
    decode_html = "PART OF DECODE",
    study        = "STUDY",
    exposure_design = "STUDY DESIGN",
    design_struc = "TEMPORAL DESIGN",
    study_design_sub = "STUDY SUB-DESIGN",
    intervention = "EXPOSURE",
    visits_html = "FOLLOW-UP VISITS",
    cohorts_html = "COHORTS",
    n            = "N AT BASELINE",
    vitals = "VITALS",
    blood_labs   = "BLOOD",
    urine_labs   = "URINE",
    clamp        = "CLAMP",
    renal_clear  = "RENAL CLEARANCE TEST",
    dxa          = "DXA",
    kidney_bx    = "KIDNEY BX",
    cardiac_mri  = "CARDIAC MRI",
    kidney_mri   = "KIDNEY MRI",
    lipidomics   = "LIPID",
    blood_proteomics   = "PROTEOMICS (PLASMA)",
    urine_proteomics = "PROTEOMICS (URINE)",
    blood_metabolomics = "METABO (PLASMA)",
    urine_metabolomics = "METABO (URINE)",
    muscle_bx = "MUSLCE BX",
    liver_bx = "LIVER BX",
    jwire = "J-WIRE",
    sphyg = "SPHYGMOCOR",
    kidney_pet = "KIDNEY PET"
  ) %>%
  
  fmt_markdown(columns = everything()) %>%
  
  # Spanners
  tab_spanner(
    label = "Labs",
    columns = c(blood_labs, urine_labs)
  ) %>%
  tab_spanner(
    label = "Metabolic Testing",
    columns = c(clamp, ivgtt, oroboros)
  ) %>%
  tab_spanner(
    label = "Kidney Function",
    columns = c(renal_clear, kidney_mri, kidney_pet)
  ) %>%
  tab_spanner(
    label = "Cardiac",
    columns = c(sphyg, cardiac_mri)
  ) %>%
  tab_spanner(
    label = "Biopsies/Procedures",
    columns = c(kidney_bx, muscle_bx, liver_bx, jwire)
  ) %>%
  tab_spanner(
    label = "Omics",
    columns = c(lipidomics, blood_proteomics, urine_proteomics,
                blood_metabolomics, urine_metabolomics)
  ) %>%
  # Column headers: uppercase, small, bold
  tab_style(
    style = cell_text(
      weight = "bold", size = "12px",
      transform = "uppercase", color = "#374151"
    ),
    locations = cells_column_spanners()
  ) %>%
  
  # Study name: bold
  tab_style(
    style = cell_text(weight = "bold", size = "15px"),
    locations = cells_body(columns = study)
  ) %>%
  
  # Column headers: uppercase, small, bold
  tab_style(
    style = cell_text(
      weight = "bold", size = "12px",
      transform = "uppercase", color = "#374151"
    ),
    locations = cells_column_labels()
  ) %>%
  
  # Body text style
  tab_style(
    style = cell_text(
      size = "15px"
    ),
    locations = cells_body()
  ) %>%
  
  
  # Procedure columns: center
  cols_align(align = "center",
             columns = c(
               decode_html, study, exposure_design, design_struc, study_design_sub,
               visits_html, intervention, cohorts_html, n,
               flagged_procs)) %>%
  
  # Column widths
  cols_width(
    study        ~ px(145),
    exposure_design ~ px(150),
    design_struc ~ px(150),
    study_design_sub ~ px(150),
    visits_html ~ px(120),
    intervention ~ px(150),
    cohorts_html ~ px(120),
    n            ~ px(70),
    flagged_procs   ~ px(70)) %>%
  opt_row_striping(row_striping = TRUE)

# ── 7. View / save ────────────────────────────────────────────────────────────
study_gt_table

study_gt_table %>% 
  gtsave(file.path(root_path, "Data Harmonization/Report_files/Summary Report/summary_table.png"),
         vwidth = 3000, vheight = 1000)

study_gt_table %>%
  gt::opt_interactive()
```

## Version with N's

```{r}
# ── N-version table: replace dots with per-procedure observed N ───────────────
# For studies present in the dataset: count unique baseline participants with
# ≥1 non-NA value in each procedure's variable set.
# For hard-coded / still-enrolling studies: show ≤[hard-coded N].

hardcoded_studies_n <- c("ADAPT-CKD", "INSPIRE", "MANATEE", "MetMod",
                         "OTTER", "SUGAR-N-SALT", "SWEETHEART", "T1-DISCO")

# Variable sets per procedure flag (for N-counting)
proc_n_varlist <- list(
  vitals           = c("height", "weight", "bmi", "sbp", "dbp"),
  blood_labs       = blood_labs,
  urine_labs       = urine_labs,
  clamp            = clamp_all,
  ivgtt            = ivgtt,
  oroboros         = character(0),
  renal_clear      = renal_clearance,
  kidney_mri       = kidney_mri_all,
  kidney_pet       = kidney_pet,
  sphyg            = character(0),
  cardiac_mri      = cardiac_mri,
  kidney_bx        = kidney_bx,
  muscle_bx        = character(0),
  liver_bx         = character(0),
  jwire            = intersect(c("jwire_side", "jwire_num"), cols),
  lipidomics       = lipid_vars,
  blood_proteomics = prot_blood_vars,
  urine_proteomics = prot_urine_vars,
  blood_metabolomics = character(0),
  urine_metabolomics = character(0),
  cgm              = intersect(c("cgm_v_high", "cgm_high", "cgm_low", "cgm_v_low"), cols)
)

count_proc_n <- function(study_name, proc_name) {
  proc_on <- proc_harm_dat_recid[proc_harm_dat_recid$study == study_name,
                                 proc_name, drop = TRUE]
  if (!isTRUE(proc_on)) return("")                       # procedure not done

  hard_n <- proc_harm_dat_recid$n[proc_harm_dat_recid$study == study_name]

  if (study_name %in% hardcoded_studies_n) return(hard_n) # show ≤N for hard-coded

  vlist <- intersect(proc_n_varlist[[proc_name]], cols)
  if (length(vlist) == 0) return(hard_n)                 # no vars mapped → ≤N

  sub <- harm_dat_recid[
    !is.na(harm_dat_recid$study) &
    harm_dat_recid$study == study_name &
    harm_dat_recid$visit == "baseline",
    vlist, drop = FALSE
  ]
  n_obs <- sum(rowSums(!is.na(sub)) > 0)
  if (n_obs == 0) hard_n else as.character(n_obs)
}

# Build N display data frame (same structure as the dot table)
display_n <- proc_harm_dat_recid %>%
  mutate(
    cohorts_html = sapply(study, make_cohort_html),
    visits_html  = sapply(study, make_visit_html),
    decode_html  = sapply(study, make_decode_html)
  )

for (proc in flagged_procs) {
  display_n[[proc]] <- vapply(display_n$study,
                              function(s) count_proc_n(s, proc),
                              character(1))
}

display_n <- display_n %>%
  select(decode_html, study, exposure_design, design_struc, study_design_sub,
         intervention, cohorts_html, visits_html, n, all_of(flagged_procs)) %>%
  arrange(decode_html, exposure_design, desc(design_struc), study_design_sub)

study_gt_n_table <- display_n %>%
  gt(id = "harm-study-table-n") %>%
  cols_label(
    decode_html = "PART OF DECODE",
    study        = "STUDY",
    exposure_design = "STUDY DESIGN",
    design_struc = "TEMPORAL DESIGN",
    study_design_sub = "STUDY SUB-DESIGN",
    intervention = "EXPOSURE",
    visits_html = "FOLLOW-UP VISITS",
    cohorts_html = "COHORTS",
    n            = "N AT BASELINE",
    vitals = "VITALS",
    blood_labs   = "BLOOD",
    urine_labs   = "URINE",
    clamp        = "CLAMP",
    renal_clear  = "RENAL CLEARANCE TEST",
    dxa          = "DXA",
    kidney_bx    = "KIDNEY BX",
    cardiac_mri  = "CARDIAC MRI",
    kidney_mri   = "KIDNEY MRI",
    lipidomics   = "LIPID",
    blood_proteomics   = "PROTEOMICS (PLASMA)",
    urine_proteomics = "PROTEOMICS (URINE)",
    blood_metabolomics = "METABO (PLASMA)",
    urine_metabolomics = "METABO (URINE)",
    muscle_bx = "MUSCLE BX",
    liver_bx = "LIVER BX",
    jwire = "J-WIRE",
    sphyg = "SPHYGMOCOR",
    kidney_pet = "KIDNEY PET",
    cgm = "CGM"
  ) %>%
  fmt_markdown(columns = everything()) %>%
  tab_spanner(label = "Labs",              columns = c(blood_labs, urine_labs)) %>%
  tab_spanner(label = "Metabolic Testing", columns = c(clamp, ivgtt, oroboros)) %>%
  tab_spanner(label = "Kidney Function",   columns = c(renal_clear, kidney_mri, kidney_pet)) %>%
  tab_spanner(label = "Cardiac",           columns = c(sphyg, cardiac_mri)) %>%
  tab_spanner(label = "Biopsies/Procedures",
              columns = c(kidney_bx, muscle_bx, liver_bx, jwire)) %>%
  tab_spanner(label = "Omics",
              columns = c(lipidomics, blood_proteomics, urine_proteomics,
                          blood_metabolomics, urine_metabolomics)) %>%
  tab_style(style = cell_text(weight = "bold", size = "12px",
                              transform = "uppercase", color = "#374151"),
            locations = cells_column_spanners()) %>%
  tab_style(style = cell_text(weight = "bold", size = "15px"),
            locations = cells_body(columns = study)) %>%
  tab_style(style = cell_text(weight = "bold", size = "12px",
                              transform = "uppercase", color = "#374151"),
            locations = cells_column_labels()) %>%
  tab_style(style = cell_text(size = "15px"),
            locations = cells_body()) %>%
  cols_align(align = "center",
             columns = c(decode_html, study, exposure_design, design_struc,
                         study_design_sub, visits_html, intervention,
                         cohorts_html, n, all_of(flagged_procs))) %>%
  cols_width(study        ~ px(145),
             exposure_design ~ px(150),
             design_struc ~ px(150),
             study_design_sub ~ px(150),
             visits_html ~ px(120),
             intervention ~ px(150),
             cohorts_html ~ px(120),
             n            ~ px(70),
             flagged_procs   ~ px(70)) %>%
  opt_row_striping(row_striping = TRUE)

study_gt_n_table

study_gt_n_table %>%
  gtsave(file.path(root_path, "Data Harmonization/Report_files/Summary Report/summary_table_n.png"),
         vwidth = 3000, vheight = 1000)

study_gt_n_table %>% gt::opt_interactive()
```


# Data availability by cohort

```{r}
# ── Cohort table: new helpers (study_palette, design_colors, tag_html,
#    study_badge_html, cohort_label_html, visit_badge_cohort_html) ─────────────
# NOTE: cohort_colors, visit_colors, badge_html, badge_visit_html,
#       abbrev_group, abbrev_visit, dot_yes, dot_no, dot, groups_list,
#       visits_list, proc_harm_dat_recid, and flagged_procs are already
#       defined above and remain in scope here.

# Neutral colored bubble per study name (QMD study names: "IMPROVE", "ULTRA")
study_palette <- c(
  # RCT
  "ATTEMPT"         = "#606c38",
  "INSPIRE"         = "#606c38",
  "MANATEE"         = "#606c38",
  "MetMod"          = "#606c38",
  "OTTER"           = "#606c38",
  "RPC2"            = "#606c38",
  "SUGAR-N-SALT"    = "#606c38",
  "T1-DISCO"        = "#606c38",
  # Single-arm trial
  "COFFEE"          = "#f48c06",
  "ULTRA"           = "#f48c06",
  # Registry cohort
  "IMPROVE"         = "#f28482",
  "ADAPT-CKD"       = "#f28482",
  # Prospective cohort
  "PANTHER"         = "#0891b2",
  "PANDA"           = "#0891b2",
  # Cohort study
  "CASPER"          = "#14213d",
  "CROCODILE"       = "#14213d",
  "PENGUIN"         = "#14213d",
  "RENAL-HEIR"      = "#14213d",
  "RENAL-HEIRitage" = "#14213d",
  "SWEETHEART"      = "#14213d"
)

# Colors for pooled design-type tags
design_colors <- c(
  "RCT"                        = "#606c38",
  "Single-arm trial"           = "#f48c06",
  "Registry cohort"            = "#f28482",
  "Prospective cohort"         = "#0891b2",
  "Cohort study"               = "#14213d"
)

# Outlined design tag badge
tag_html <- function(label, color) {
  sprintf(
    paste0('<span style="background:%s20;color:%s;border:1px solid %s40;',
           'font-size:0.7em;font-weight:600;padding:1px 5px;border-radius:4px;',
           'margin:1px 2px;display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    color, color, color, label
  )
}

# Colored solid bubble for a study name
study_badge_html <- function(study_name) {
  bg <- study_palette[study_name]
  if (is.na(bg)) bg <- "#6b7280"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.75em;font-weight:700;',
           'padding:2px 8px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, study_name
  )
}

# Larger cohort-type pill (row label)
cohort_label_html <- function(cohort) {
  bg <- cohort_colors[cohort]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.9em;font-weight:700;',
           'padding:3px 10px;border-radius:999px;display:inline-block;',
           'white-space:nowrap;line-height:1.7">%s</span>'),
    bg, cohort
  )
}

# Smaller visit badge for cohort table (reuses visit_colors already in scope)
visit_badge_cohort_html <- function(v) {
  col <- visit_colors[v]
  if (is.na(col)) col <- "#9ca3af"
  sprintf(
    paste0('<span style="background:white;color:%s;border:1px solid %s;',
           'font-size:0.7em;font-weight:600;padding:1px 5px;border-radius:999px;',
           'margin:1px 2px;display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    col, col, v
  )
}
```

```{r}
# ── Cohort table: data preparation ────────────────────────────────────────────

# 1. Derive per-study cohort abbreviations from groups_list (already in scope)
cohort_abbrev_per_study <- lapply(names(groups_list), function(s) {
  gs <- groups_list[[s]]
  sort(unique(abbrev_group(gs)))
})
names(cohort_abbrev_per_study) <- names(groups_list)

# 2. Derive per-study follow-up visit string from visits_list (already in scope)
#    Exclude baseline (abbrev "0")
followup_str_per_study <- lapply(names(visits_list), function(s) {
  vs   <- visits_list[[s]]
  vs   <- vs[!is.na(vs)]
  abbr <- sort(unique(abbrev_visit(vs)))
  abbr <- abbr[abbr != "0"]
  if (length(abbr) == 0) return(NA_character_)
  paste(abbr, collapse = ",")
})
names(followup_str_per_study) <- names(visits_list)

# 3. Build a slim metadata frame: study | cohorts_str | followup_visits | n_str
#    n_str comes from proc_harm_dat_recid$n (already character with <= overrides)
study_meta_cohort <- data.frame(
  study           = proc_harm_dat_recid$study,
  cohorts_str     = sapply(proc_harm_dat_recid$study, function(s) {
    v <- cohort_abbrev_per_study[[s]]
    if (is.null(v) || length(v) == 0) NA_character_ else paste(v, collapse = ",")
  }),
  followup_visits = sapply(proc_harm_dat_recid$study, function(s) {
    v <- followup_str_per_study[[s]]
    if (is.null(v) || length(v) == 0) NA_character_ else v
  }),
  n_str           = proc_harm_dat_recid$n,
  stringsAsFactors = FALSE
)

# 4. Procedure columns for cohort table


# 5. Join metadata with procedure flags from proc_harm_dat_recid
study_full_cohort <- study_meta_cohort %>%
  left_join(
    proc_harm_dat_recid %>%
      select(study, decode, exposure_design, design_struc, study_design_sub,
             all_of( flagged_procs)),
    by = "study"
  )

# 6. Expand to one row per study x cohort
study_cohort_long <- study_full_cohort %>%
  mutate(cohort_list = strsplit(cohorts_str, ",")) %>%
  tidyr::unnest(cohort_list) %>%
  mutate(cohort = trimws(cohort_list))

cohort_order <- c("HC", "OB", "T1D", "T2D", "ADPKD")

# ── build_cohort_display(): aggregate to one gt-ready row per cohort ──────────
build_cohort_display <- function(subset_decode = c("ALL", "DECODE", "NOT PART OF DECODE")) {
  subset_decode <- match.arg(subset_decode)

  dat <- if (subset_decode == "ALL") {
    study_cohort_long
  } else {
    study_cohort_long %>% filter(decode == subset_decode)
  }

  cohort_rows <- lapply(cohort_order, function(ch) {
    ch_dat <- dat %>% filter(cohort == ch)
    if (nrow(ch_dat) == 0) return(NULL)

    studies_in_ch <- ch_dat$study

    # Cohort label pill
    cohort_html <- cohort_label_html(ch)

    # N: sum numeric Ns; mark capped (<=xx) studies with +
    ns         <- ch_dat$n_str
    numeric_ns <- suppressWarnings(as.numeric(ns))
    capped_ns  <- as.numeric(gsub("≤", "", ns))
    sum_num    <- sum(numeric_ns, capped_ns, na.rm = TRUE)
    n_display  <- if (length(capped_ns) == 0) {
      as.character(sum_num)
    } else if (sum_num == 0) {
      paste(collapse = "≤", unique(capped_ns))
    } else {
      paste0("≤", sum_num)
    }

    # Study bubbles + follow-up visit badges (one line per study)
    study_visit_parts <- vapply(studies_in_ch, function(s) {
      sb  <- study_badge_html(s)
      row <- ch_dat[ch_dat$study == s, ]
      vstr <- row$followup_visits
      if (!is.na(vstr) && nchar(vstr) > 0) {
        vv  <- trimws(unlist(strsplit(vstr, ",")))
        vbs <- paste(vapply(vv, visit_badge_cohort_html, character(1)), collapse = "")
        paste0(sb, vbs)
      } else {
        sb
      }
    }, character(1))
    studies_html <- paste(study_visit_parts, collapse = "<br>")

    # Pooled design tags — unique values across all three design columns
    designs     <- unique(c(ch_dat$study_design_sub))
    designs     <- designs[!is.na(designs) & nchar(trimws(designs)) > 0]
    design_html <- paste(vapply(designs, function(d) {
      col <- design_colors[d]
      if (is.na(col)) col <- "#6b7280"
      tag_html(d, col)
    }, character(1)), collapse = " ")

    # Procedure dots: ANY study in this cohort = filled
    proc_vals <- vapply( flagged_procs, function(p) {
      vals <- ch_dat[[p]]
      if (is.logical(vals)) any(vals, na.rm = TRUE)
      else any(as.logical(vals), na.rm = TRUE)
    }, logical(1))

    row_data <- c(
      list(
        cohort_html  = cohort_html,
        studies_html = studies_html,
        design_html  = design_html,
        n            = n_display
      ),
      as.list(proc_vals)
    )
    as.data.frame(row_data, stringsAsFactors = FALSE, check.names = FALSE)
  })

  cohort_df <- bind_rows(Filter(Negate(is.null), cohort_rows))

  # Convert logical procedure values to HTML dots
  cohort_df <- cohort_df %>%
    mutate(across(all_of( flagged_procs),
                  ~ sapply(.x, function(v) dot(as.logical(v)))))
  cohort_df
}

# Build the three display data frames
df_cohort_all    <- build_cohort_display("ALL")
df_cohort_decode <- build_cohort_display("DECODE")
df_cohort_nondec <- build_cohort_display("NOT PART OF DECODE")

# ── make_cohort_gt(): build a gt table from a cohort display data frame ────────
make_cohort_gt <- function(display_df, title_suffix = "") {
  display_df %>%
    gt(id = paste0("harm-cohort-table", gsub("[^A-Za-z0-9]", "_", title_suffix))) %>%

    cols_label(
      cohort_html      = "COHORT",
      studies_html     = "STUDIES & FOLLOW-UP",
      design_html      = "STUDY DESIGNS",
      n                = "N",
      vitals = "VITALS",
      blood_labs   = "BLOOD",
      urine_labs   = "URINE",
      clamp        = "CLAMP",
      renal_clear  = "RENAL CLEARANCE TEST",
      dxa          = "DXA",
      kidney_bx    = "KIDNEY BX",
      cardiac_mri  = "CARDIAC MRI",
      kidney_mri   = "KIDNEY MRI",
      lipidomics   = "LIPID",
      blood_proteomics   = "PROTEOMICS (PLASMA)",
      urine_proteomics = "PROTEOMICS (URINE)",
    blood_metabolomics = "METABO (PLASMA)",
    urine_metabolomics = "METABO (URINE)",
      muscle_bx = "MUSLCE BX",
      liver_bx = "LIVER BX",
      jwire = "J-WIRE",
      sphyg = "SPHYG MOCOR",
      kidney_pet = "KIDNEY PET"
    ) %>%
    
    # Spanners
    tab_spanner(
      label = "Labs",
      columns = c(blood_labs, urine_labs)
    ) %>%
    tab_spanner(
      label = "Metabolic Testing",
      columns = c(clamp, ivgtt, oroboros)
    ) %>%
    tab_spanner(
      label = "Kidney Function",
      columns = c(renal_clear, kidney_mri, kidney_pet)
    ) %>%
    tab_spanner(
      label = "Cardiac",
      columns = c(sphyg, cardiac_mri)
    ) %>%
    tab_spanner(
      label = "Biopsies/Procedures",
      columns = c(kidney_bx, muscle_bx, liver_bx, jwire)
    ) %>%
    tab_spanner(
      label = "Omics",
      columns = c(lipidomics, blood_proteomics, urine_proteomics,
                  blood_metabolomics, urine_metabolomics)
    ) %>%
    
    
    fmt_markdown(columns = everything()) %>%

    # Styling
    # Column headers: uppercase, small, bold
    tab_style(
      style = cell_text(
        weight = "bold", size = "12px",
        transform = "uppercase", color = "#374151"
      ),
      locations = cells_column_spanners()
    ) %>%
    
    # Column headers: uppercase, small, bold
    tab_style(
      style = cell_text(
        weight = "bold", size = "12px",
        transform = "uppercase", color = "#374151"
      ),
      locations = cells_column_labels()
    ) %>%
    
    # Body text style
    tab_style(
      style = cell_text(
        size = "15px"
      ),
      locations = cells_body()
    ) %>%
    # Alignment
    cols_align(align = "center") %>%

    # Column widths
    cols_width(
      cohort_html      ~ px(80),
      studies_html     ~ px(200),
      design_html      ~ px(150),
      n                ~ px(50),
      blood_proteomics ~ px(90),
      urine_proteomics ~ px(90),
       flagged_procs ~ px(80)

    ) %>%

    opt_row_striping(row_striping = TRUE)

}
```

## All studies

```{r}
all_cohort_table <- make_cohort_gt(df_cohort_all, "")
all_cohort_table

all_cohort_table %>% 
  gtsave(file.path(root_path, "Data Harmonization/Report_files/Summary Report/summary_table_by_cohort.png"),
         vwidth = 4500, vheight = 1000)

all_cohort_table %>%
  gt::opt_interactive()
```

## DECODE studies only

```{r}
decode_cohort_table <- make_cohort_gt(df_cohort_decode, " · DECODE Studies")
decode_cohort_table
decode_cohort_table %>% 
  gtsave(file.path(root_path, "Data Harmonization/Report_files/Summary Report/summary_table_by_cohort_decode_only.png"),
         vwidth = 4500, vheight = 1000)

decode_cohort_table %>%
  gt::opt_interactive()
```

## Non-DECODE studies only

```{r}
non_decode_cohort_table <- make_cohort_gt(df_cohort_nondec, " · Non-DECODE Studies")
non_decode_cohort_table
non_decode_cohort_table %>% 
  gtsave(file.path(root_path, "Data Harmonization/Report_files/Summary Report/summary_table_by_cohort_nondecode_only.png"),
         vwidth = 4500, vheight = 1000)

non_decode_cohort_table %>%
  gt::opt_interactive()
```
---

# Summary visualizations

```{r}
# ── Shared theme and palette ──────────────────────────────────────────────────
theme_harm <- theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor  = element_blank(),
    strip.text        = element_text(face = "bold"),
    legend.position   = "bottom",
    axis.text.x       = element_text(angle = 45, hjust = 1)
  )

cohort_fill <- c(T1D = "#e77c76", T2D = "#e37902",
                 HC  = "#8cb369", OB  = "#5b8e7d", ADPKD = "#bc4b51")

# Output dir for saved PNGs
fig_dir <- file.path(root_path,
  "Data Harmonization/Report_files/Summary Report/figures")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)

save_fig <- function(p, filename, width = 10, height = 5) {
  ggsave(file.path(fig_dir, filename), plot = p,
         width = width, height = height, dpi = 300, bg = "white")
  p
}
```

```{r}
# ── Hard-coded study metadata (group composition + ≤N) ───────────────────────
# Used to supplement observed data with upper-bound estimates for studies that
# are still enrolling / not yet fully entered in the dataset.
hardcoded_meta <- list(
  "ADAPT-CKD"    = list(groups = c("T2D"), n_total = 100),
  "INSPIRE"      = list(groups = c("T1D"), n_total = 90),
  "MANATEE"      = list(groups = c("T1D"), n_total = 60),
  "MetMod"       = list(groups = c("T1D"), n_total = 60),
  "OTTER"        = list(groups = c("T2D"), n_total = 60),
  "SUGAR-N-SALT" = list(groups = c("T2D"), n_total = 150),
  "SWEETHEART"   = list(groups = c("T1D"), n_total = 30),
  "T1-DISCO"     = list(groups = c("T1D"), n_total = 60)
)

# Procedure order matches flagged_procs / main summary table column order.
# Only procedures with a mappable variable set are included for N-counting.
proc_varlist_plot <- list(
  "Blood Labs"       = blood_labs,
  "Urine Labs"       = urine_labs,
  "DXA"              = dxa,
  "Clamp"            = clamp_all,
  "IVGTT"            = ivgtt,
  "Renal Clearance"  = renal_clearance,
  "Kidney MRI"       = kidney_mri_all,
  "Kidney PET"       = kidney_pet,
  "Cardiac MRI"      = cardiac_mri,
  "Kidney Biopsy"    = kidney_bx,
  "Lipidomics"       = lipid_vars,
  "Blood Proteomics" = prot_blood_vars,
  "Urine Proteomics" = prot_urine_vars,
  "CGM"              = intersect(c("cgm_v_high","cgm_high","cgm_low","cgm_v_low"), cols)
)
proc_order <- names(proc_varlist_plot)

# proc_flag_to_label: maps proc_harm_dat_recid column name → plot label
proc_flag_to_label <- c(
  blood_labs       = "Blood Labs",
  urine_labs       = "Urine Labs",
  dxa              = "DXA",
  clamp            = "Clamp",
  ivgtt            = "IVGTT",
  renal_clear      = "Renal Clearance",
  kidney_mri       = "Kidney MRI",
  kidney_pet       = "Kidney PET",
  cardiac_mri      = "Cardiac MRI",
  kidney_bx        = "Kidney Biopsy",
  lipidomics       = "Lipidomics",
  blood_proteomics = "Blood Proteomics",
  urine_proteomics = "Urine Proteomics",
  cgm              = "CGM"
)
```

```{r}
# ── Data-building helpers ─────────────────────────────────────────────────────
# decode_map: study → DECODE / NOT PART OF DECODE
decode_map <- setNames(proc_harm_dat_recid$decode, proc_harm_dat_recid$study)

# ── make_pie_dat() ─────────────────────────────────────────────────────────────
# Builds cohort composition counts.
# include_hardcoded = TRUE adds ≤N estimates from hard-coded studies.
# ≤ prefix only applied to groups that receive an estimated contribution.
make_pie_dat <- function(include_hardcoded = FALSE,
                         decode_filter     = c("ALL","DECODE","NOT PART OF DECODE")) {
  decode_filter <- match.arg(decode_filter)

  real_dat <- harm_dat_recid %>%
    filter(visit == "baseline", !is.na(group), !is.na(record_id)) %>%
    distinct(record_id, .keep_all = TRUE) %>%
    mutate(group_abbr   = abbrev_group(group),
           study_decode = decode_map[study]) %>%
    { if (decode_filter == "DECODE")
        filter(., study_decode == "DECODE")
      else if (decode_filter == "NOT PART OF DECODE")
        filter(., study_decode == "NOT PART OF DECODE")
      else . } %>%
    count(group_abbr) %>%
    mutate(has_estimated = FALSE)

  if (!include_hardcoded) {
    return(real_dat %>%
      mutate(pct   = n / sum(n),
             label = paste0(group_abbr, "\n(n=", n, ")")) %>%
      select(-has_estimated))
  }

  hc_rows <- bind_rows(lapply(names(hardcoded_meta), function(s) {
    meta <- hardcoded_meta[[s]]
    if (decode_filter == "DECODE"             && decode_map[s] != "DECODE")             return(NULL)
    if (decode_filter == "NOT PART OF DECODE" && decode_map[s] != "NOT PART OF DECODE") return(NULL)
    data.frame(group_abbr    = meta$groups,
               n             = floor(meta$n_total / length(meta$groups)),
               has_estimated = TRUE, stringsAsFactors = FALSE)
  }))

  bind_rows(real_dat, hc_rows) %>%
    group_by(group_abbr) %>%
    summarise(n = sum(n), has_estimated = any(has_estimated), .groups = "drop") %>%
    mutate(pct   = n / sum(n),
           label = ifelse(has_estimated,
                          paste0(group_abbr, "\n(≤", n, ")"),
                          paste0(group_abbr, "\n(n=",  n, ")")))
}

# ── make_bar_dat() ─────────────────────────────────────────────────────────────
# Builds per-procedure N by cohort group.
# ≤ prefix only applied to group×procedure cells that include an estimated study.
make_bar_dat <- function(include_hardcoded = FALSE,
                         decode_filter     = c("ALL","DECODE","NOT PART OF DECODE")) {
  decode_filter <- match.arg(decode_filter)

  real_rows <- bind_rows(lapply(names(proc_varlist_plot), function(proc_name) {
    vlist <- intersect(proc_varlist_plot[[proc_name]], cols)
    if (length(vlist) == 0) return(NULL)
    harm_dat_recid %>%
      filter(visit == "baseline", !is.na(group), !is.na(record_id)) %>%
      distinct(record_id, .keep_all = TRUE) %>%
      mutate(has_proc     = rowSums(!is.na(select(., all_of(vlist)))) > 0,
             study_decode = decode_map[study]) %>%
      { if (decode_filter == "DECODE")
          filter(., study_decode == "DECODE")
        else if (decode_filter == "NOT PART OF DECODE")
          filter(., study_decode == "NOT PART OF DECODE")
        else . } %>%
      filter(has_proc) %>%
      mutate(group_abbr = abbrev_group(group)) %>%
      count(group_abbr) %>%
      mutate(procedure = proc_name, has_estimated = FALSE)
  }))

  if (!include_hardcoded) {
    return(real_rows %>%
      mutate(procedure = factor(procedure, levels = proc_order)))
  }

  hc_rows <- bind_rows(lapply(names(hardcoded_meta), function(s) {
    meta <- hardcoded_meta[[s]]
    if (decode_filter == "DECODE"             && decode_map[s] != "DECODE")             return(NULL)
    if (decode_filter == "NOT PART OF DECODE" && decode_map[s] != "NOT PART OF DECODE") return(NULL)
    study_flags <- proc_harm_dat_recid[proc_harm_dat_recid$study == s, ]
    if (nrow(study_flags) == 0) return(NULL)
    bind_rows(lapply(names(proc_flag_to_label), function(flag) {
      if (!isTRUE(study_flags[[flag]])) return(NULL)
      data.frame(group_abbr    = meta$groups,
                 n             = floor(meta$n_total / length(meta$groups)),
                 procedure     = proc_flag_to_label[[flag]],
                 has_estimated = TRUE, stringsAsFactors = FALSE)
    }))
  }))

  bind_rows(real_rows, hc_rows) %>%
    group_by(procedure, group_abbr) %>%
    summarise(n = sum(n), has_estimated = any(has_estimated), .groups = "drop") %>%
    mutate(procedure = factor(procedure, levels = proc_order))
}

# ── make_proc_summary_tbl() ────────────────────────────────────────────────────
# Builds a numeric summary table: rows = procedures, cols = cohort groups,
# cells = N (observed) or ≤N (estimated). Used for the compiled N table.
make_proc_summary_tbl <- function(include_hardcoded = TRUE,
                                  decode_filter     = c("ALL","DECODE","NOT PART OF DECODE")) {
  decode_filter <- match.arg(decode_filter)

  bar_dat <- make_bar_dat(include_hardcoded = include_hardcoded,
                          decode_filter     = decode_filter)

  bar_dat %>%
    mutate(n_label = ifelse(has_estimated, paste0("≤", n), as.character(n))) %>%
    select(procedure, group_abbr, n_label) %>%
    pivot_wider(names_from = group_abbr, values_from = n_label, values_fill = "—") %>%
    arrange(factor(procedure, levels = proc_order)) %>%
    rename(Procedure = procedure)
}
```

```{r}
# ── Plot helpers ──────────────────────────────────────────────────────────────

make_pie_plot <- function(pie_dat, filename = NULL) {
  p <- ggplot(pie_dat, aes(x = "", y = pct, fill = group_abbr)) +
    geom_col(width = 1, color = "white") +
    coord_polar(theta = "y") +
    scale_fill_manual(values = cohort_fill, name = "Cohort") +
    geom_text(aes(label = label),
              position = position_stack(vjust = 0.5),
              size = 3.5, color = "white", fontface = "bold") +
    labs(x = NULL, y = NULL) +
    theme_void(base_size = 12) +
    theme(legend.position = "none")
  if (!is.null(filename)) save_fig(p, filename, width = 6, height = 6) else p
}

make_bar_plot <- function(bar_dat, filename = NULL) {
  p <- ggplot(bar_dat, aes(x = procedure, y = n, fill = group_abbr)) +
    geom_col(position = "dodge", width = 0.7) +
    scale_fill_manual(values = cohort_fill, name = "Cohort") +
    labs(x = NULL, y = "N") +
    theme_harm
  if (!is.null(filename)) save_fig(p, filename) else p
}

bp_base <- harm_dat_recid %>%
  filter(visit == "baseline", !is.na(group), !is.na(record_id)) %>%
  distinct(record_id, .keep_all = TRUE) %>%
  mutate(group_abbr = factor(abbrev_group(group),
                             levels = c("HC", "OB", "T1D", "T2D", "ADPKD")))

make_boxplot <- function(data, var, ylabel, filename = NULL, log_transform = FALSE) {
  d <- data %>% filter(!is.na(.data[[var]]))
  p <- ggplot(d, aes(x = group_abbr, y = !!sym(var), fill = group_abbr)) +
    { if (log_transform) scale_y_log10() } +
    geom_boxplot(outlier.shape = 21, outlier.size = 1.5, alpha = 0.85) +
    geom_jitter(width = 0.15, size = 0.8, alpha = 0.35, color = "grey30") +
    scale_fill_manual(values = cohort_fill, guide = "none") +
    labs(x = NULL, y = ylabel) +
    theme_harm +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
  if (!is.null(filename)) save_fig(p, filename, width = 6, height = 5) else p
}
```

---

## Data-only (observed data in dataset)

> Baseline participants with ≥1 non-missing value per procedure variable set.
> Hard-coded / still-enrolling studies are **excluded**.

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(FALSE, "ALL"), "pie_data_only_all.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(FALSE, "ALL"), "bar_data_only_all.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = FALSE, decode_filter = "ALL") %>%
  knitr::kable(caption = "N per procedure × cohort — data only")
```

### Key outcome box plots

```{r}
make_boxplot(subset(bp_base, is.na(acr_u) | acr_u > 0), "acr_u",
             "UACR (mg/g) [log scale]", "box_data_only_uacr.png", log_transform = TRUE)
```

```{r}
make_boxplot(bp_base, "p2_raw_m", "M-value (mg/kg/min)", "box_data_only_mvalue.png")
```

```{r}
make_boxplot(bp_base, "gfr_bsa_plasma", "GFR (mL/min/1.73 m²)", "box_data_only_gfr.png")
```

```{r}
make_boxplot(bp_base, "avg_c_r2", "Avg Cortical R2* (s⁻¹)", "box_data_only_r2star.png")
```

---

## Full (observed + hard-coded estimates, all studies)

> Combines observed N with ≤N estimates for hard-coded / still-enrolling studies.
> ≤ appears only on cohort groups that include at least one estimated study.

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(TRUE, "ALL"), "pie_full_all.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(TRUE, "ALL"), "bar_full_all.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = TRUE, decode_filter = "ALL") %>%
  knitr::kable(caption = "N per procedure × cohort — full (observed + ≤ estimated)")
```

---

## DECODE studies only (observed + hard-coded estimates)

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(TRUE, "DECODE"), "pie_full_decode.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(TRUE, "DECODE"), "bar_full_decode.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = TRUE, decode_filter = "DECODE") %>%
  knitr::kable(caption = "N per procedure × cohort — DECODE studies (observed + ≤ estimated)")
```

---

## Non-DECODE studies only (observed + hard-coded estimates)

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(TRUE, "NOT PART OF DECODE"), "pie_full_nondecode.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(TRUE, "NOT PART OF DECODE"), "bar_full_nondecode.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = TRUE, decode_filter = "NOT PART OF DECODE") %>%
  knitr::kable(caption = "N per procedure × cohort — non-DECODE studies (observed + ≤ estimated)")
```

---

## Box plots — key outcomes by cohort

> Observed data only; hard-coded studies have no measured values.

```{r}
make_boxplot(subset(bp_base, is.na(acr_u) | acr_u > 0), "acr_u",
             "UACR (mg/g) [log scale]", "box_uacr.png", log_transform = TRUE)
```

```{r}
make_boxplot(bp_base, "p2_raw_m", "M-value (mg/kg/min)", "box_mvalue.png")
```

```{r}
make_boxplot(bp_base, "gfr_bsa_plasma", "GFR (mL/min/1.73 m²)", "box_gfr.png")
```

```{r}
make_boxplot(bp_base, "avg_c_r2", "Avg Cortical R2* (s⁻¹)", "box_r2star.png")
```
