---
title: "Overall Data Summary"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

```{r include = F}
# ── Libraries ─────────────────────────────────────────────────────────────────
library(jsonlite)
library(dplyr)
library(kableExtra)
library(knitr)
library(purrr)
library(tidyr)
library(ggplot2)
library(GGally)
library(gtExtras)
library(gt)
library(patchwork)
library(Hmisc)

# ── User-specific paths ───────────────────────────────────────────────────────
user <- Sys.info()[["user"]]

if (user == "choiyej") {
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") {
  root_path <- ""
  git_path  <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") {
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path  <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/mmfs1/home/pylell/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}

# ── Load raw harmonized dataset ───────────────────────────────────────────────
harm_dat_raw <- read.csv(
  file.path(root_path, "Data Harmonization/Data Clean/soma_harmonized_dataset.csv"),
  na = ""
)
```


```{r}
# ── TODAY study data ─────────────────────────────────────────────────────────

# Load TODAY clinical data (long format; object name is "long")
load(file.path(root_path, "TODAY subaward/Clinical data/TODAY/clinical_data_long.Rdata"))

# TODAY ECHO
today_echo <- read.csv(file.path(root_path, "TODAY subaward/Clinical data/TODAY/echo.csv"))
today_echo$TIMEPOINT <- "1"
# TODAY2 ECHO
today2_echo <- read.csv(file.path(root_path, "TODAY subaward/Clinical data/TODAY2/echo.csv"))
today2_echo$TIMEPOINT <- "2"
drop_cols <- c("dopplerqc","mmodeqc","overallqc","plaxqc","saxqc","apicalqc","mmodeqc")
today2_echo <- today2_echo %>% dplyr::select(-dplyr::one_of(drop_cols))
# combine echo datasets
echo <- dplyr::bind_rows(today2_echo, today_echo)
echo$TIMEPOINT <- as.factor(echo$TIMEPOINT)
echo$RELEASEID <- echo$releaseid
echo$releaseid <- NULL

# TODAY2 PWV
pwv <- read.csv(file.path(root_path, "TODAY subaward//Clinical data/TODAY2/pwv.csv"))
pwv_keep <- pwv %>% dplyr::select(RELEASEID, TIMEPOINT, PWVF)
pwv_keep$pwv_cf_gt10 <- ifelse(is.na(pwv_keep$PWVF), NA,
                               ifelse(pwv_keep$PWVF >10, 1, 0))
pwv_keep <- pwv_keep %>% dplyr::arrange(RELEASEID, dplyr::desc(pwv_cf_gt10)) %>%
  dplyr::group_by(RELEASEID) %>% dplyr::filter(dplyr::row_number()==1)

# TODAY DEXA
today_dexa <- read.csv(file.path(root_path, "TODAY subaward/Clinical data/TODAY/DEXA.csv"))

# ── Compute TODAY baseline Ns ────────────────────────────────────────────────
# Baseline = M00 (or visit_num == 0) from the clinical long data
today_baseline <- long %>% dplyr::filter(visit == "M00")
today_n_total     <- length(unique(today_baseline$releaseid))

# Blood labs: hba1c, trig, ins0min, alt, ast
today_lab_cols    <- intersect(c("hba1c", "trig", "ins0min", "alt", "ast"), names(today_baseline))
today_n_blood     <- sum(rowSums(!is.na(today_baseline[, today_lab_cols, drop = FALSE])) > 0)

# Urine labs: check for urine-related columns
today_urine_cols  <- intersect(c("codi"), names(today_baseline))
today_n_urine     <- if (length(today_urine_cols) > 0) {
  sum(rowSums(!is.na(today_baseline[, today_urine_cols, drop = FALSE])) > 0)
} else 0L

# Vitals: bmi, height, weight, wastcirc
today_vitals_cols <- intersect(c("bmi", "height", "weight", "wastcirc"), names(today_baseline))
today_n_vitals    <- sum(rowSums(!is.na(today_baseline[, today_vitals_cols, drop = FALSE])) > 0)

# DEXA at M00
today_n_dxa       <- nrow(today_dexa %>% dplyr::filter(mvisit == "M00") %>%
                            dplyr::filter(dplyr::if_any(tidyselect::starts_with("WB_TOT"), ~ !is.na(.x))))

# ECHO (follow-up only — no baseline ECHO in TODAY)
today_n_echo      <- length(unique(echo$RELEASEID))

# PWV / sphygmocor (follow-up only — PWV1 timepoint)
today_n_pwv       <- length(unique(pwv_keep$RELEASEID))

# TODAY metabolomics: plasma and urine
load(file.path(root_path, "TODAY subaward/Metabolomic data/plasma.Rdata"))
load(file.path(root_path, "TODAY subaward/Metabolomic data/urine.Rdata"))
today_plasma_bl    <- plasma %>% dplyr::filter(visnum == 0)
today_n_blood_met  <- length(unique(today_plasma_bl$releaseid[
  rowSums(!is.na(today_plasma_bl %>% dplyr::select(-dplyr::any_of(c(
    "current_label","Freezerworks.ID","Sample.Name","site","releaseid",
    "material_type","bsi_id","Date.Drawn","visnum","location","MASK.ID","SAMPLE_ID"
  ))))) > 0]))
today_urine_met_bl <- urine %>% dplyr::filter(visnum == 0)
today_n_urine_met  <- length(unique(today_urine_met_bl$releaseid[
  rowSums(!is.na(today_urine_met_bl %>% dplyr::select(-dplyr::any_of(c(
    "current_label","Freezerworks.ID","Sample.Name","site","releaseid",
    "material_type","bsi_id","Date.Drawn","visnum","location","MASK.ID","SAMPLE_ID"
  ))))) > 0]))

# TODAY blood proteomics: SomaLogic
load(file.path(root_path, "TODAY subaward/Somalogic data raw/soma.Rdata"))
today_soma_bl      <- soma %>% dplyr::filter(visnum == 0)
today_n_blood_prot <- length(unique(today_soma_bl$releaseid[
  rowSums(!is.na(today_soma_bl %>% dplyr::select(tidyselect::starts_with("seq.")))) > 0]))
```

```{r}
# ── TEEN-LABS study data ──────────────────────────────────────────────────────

# Load TEEN-LABS cleaned data frame (longitudinal: 64 subjects, 7 visits)
teenlabs_df <- readRDS(file.path(root_path, "Teen Labs/Data_Cleaned/df.rds"))

# Baseline = "Month 1" (first post-surgery assessment)
teenlabs_bl <- teenlabs_df %>% dplyr::filter(visit == "Month 1")

# ── Compute TEEN-LABS baseline Ns ─────────────────────────────────────────────
teenlabs_n_total <- length(unique(teenlabs_bl$ID))

# Blood labs
teenlabs_lab_cols <- intersect(
  c("HBA1C","GLU","CHOL","HDL","LDL","TGNET","CREAS","CYSC","ALBS",
    "CRPM","FER","FOLAT","INSTOSOH","PTH","VB12","VD","VITA","VITB1G"),
  toupper(names(teenlabs_bl))
)
teenlabs_n_blood <- sum(rowSums(!is.na(
  teenlabs_bl[, names(teenlabs_bl)[toupper(names(teenlabs_bl)) %in% teenlabs_lab_cols], drop = FALSE]
)) > 0)

# Urine labs
teenlabs_urine_cols <- intersect(
  c("ALBU","CREAU","UACRATIO"),
  toupper(names(teenlabs_bl))
)
teenlabs_n_urine <- sum(rowSums(!is.na(
  teenlabs_bl[, names(teenlabs_bl)[toupper(names(teenlabs_bl)) %in% teenlabs_urine_cols], drop = FALSE]
)) > 0)

# Vitals
teenlabs_vitals_cols <- intersect(
  c("bmi","weight","height","sbp","dbp","iwaist","uwaist","neck","sad","bfat"),
  names(teenlabs_bl)
)
teenlabs_n_vitals <- sum(rowSums(!is.na(
  teenlabs_bl[, teenlabs_vitals_cols, drop = FALSE]
)) > 0)

# SomaLogic blood proteomics (seq.* columns)
teenlabs_soma_cols <- grep("^seq\\.", names(teenlabs_bl), value = TRUE)
teenlabs_n_soma <- sum(rowSums(!is.na(
  teenlabs_bl[, teenlabs_soma_cols, drop = FALSE]
)) > 0)

# Olink blood proteomics (columns after seq.* that are protein symbols)
# Identify: everything after the last seq.* column, excluding derived clinical vars
teenlabs_derived <- c("albuminuria","eGFR_FAS_cr","eGFR_FAS_cr_cysc",
  "eGFR_FAS_cysc","eGFR_Schwartz","eGFR_CKD_EPI_cr","eGFR_CKD_EPI_cr_cysc",
  "eGFR_CKD_EPI_cysc","nafld_fs","fli","hsi","usfli","diab","diab_resolved",
  "diab_baseline","diabetes_duration","SURG","anthdat","surgdat","potentialpreg",
  "ethn","race","sex","age","labid","fasting","date_drawn","volume_unit",
  "sample_id","hrate","neck","sad","uwaist","iwaist","bfat","dbp_percentile",
  "sbp_percentile","lmcbcplt","lmcbcwbc")
last_seq_idx <- max(grep("^seq\\.", names(teenlabs_bl)))
olink_candidates <- names(teenlabs_bl)[(last_seq_idx + 1):ncol(teenlabs_bl)]
teenlabs_olink_cols <- setdiff(olink_candidates, c(teenlabs_derived, "ID", "visit"))
teenlabs_n_olink <- sum(rowSums(!is.na(
  teenlabs_bl[, teenlabs_olink_cols, drop = FALSE]
)) > 0)

# DXA: TEEN-LABS has bfat (body fat %) — count as DXA equivalent
teenlabs_n_dxa <- sum(!is.na(teenlabs_bl$bfat))
```

```{r, include = F}
# ── Data cleaning & variable engineering ──────────────────────────────────────

# Drop SomaScan seq columns (keep only the two legacy seq variables)
harm_dat <- harm_dat_raw %>%
  dplyr::mutate(sex = dplyr::case_when(study == "ULTRA" ~ "Male", T ~ sex)) %>%
  dplyr::select(-tidyselect::starts_with("seq"), seq.10000.28, seq.10000.28_urine)

# Collapse duplicate record_id × visit rows:
#   - character columns: last non-NA value
#   - numeric columns:   mean of non-NA values
# Then filter out treatment-period and unscheduled visits.
harm_dat_recid <- harm_dat %>%
  dplyr::summarise(
    dplyr::across(tidyselect::where(purrr::negate(is.numeric)),
           ~ ifelse(all(is.na(.x)), NA_character_, dplyr::last(na.omit(.x)))),
    dplyr::across(tidyselect::where(is.numeric),
           ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
    .by = c(record_id, visit)
  ) %>%
  dplyr::select(-mrn, -dob) %>%
  dplyr::filter(visit %nin% c(
    "treatment_period_1", "treatment_period_2",
    "treatment_period_3", "treatment_period_4",
    "Early Withdrawal", "Unscheduled 1", "Unscheduled 2"
  )) %>%

  # ── Hard-coded placeholder rows for studies still enrolling / not yet in data ──
  # These studies are either still collecting data or have not yet been entered;
  # N values below are upper-bound estimates (≤).
  dplyr::bind_rows(data.frame(
    study = c(
      rep("ADAPT-CKD",    2),
      rep("MANATEE",      1),
      rep("INSPIRE",      1),
      rep("T1-DISCO",     1),
      rep("MetMod",       1),
      rep("OTTER",        1),
      rep("TODAY",         1),
      rep("TEEN-LABS",     2)
    ),
    visit = c("MO6", "YR1",
              "MO4",
              "MO3",
              "MO6",
              "MO6",
              "MO6",
              "baseline",
              "baseline", "baseline"),
    group = c(
      "Type 2 Diabetes", "Obese Control",
      "Type 1 Diabetes",
      "Type 2 Diabetes",
      "Type 1 Diabetes",
      "Type 1 Diabetes",
      "Type 1 Diabetes",
      "Type 2 Diabetes",
      "Type 2 Diabetes", "Obese Control"
    )
  )) %>%

  # ── Study-level metadata ───────────────────────────────────────────────────
  dplyr::mutate(
    study_design = dplyr::case_when(
      study %in% c("RENAL-HEIR", "RENAL-HEIRitage", "PENGUIN",
                   "CROCODILE", "SWEETHEART", "CASPER")            ~ "Observational: Cross-sectional",
      study %in% c("ULTRA", "COFFEE")                              ~ "Interventional: Longitudinal",
      study %in% c("PANTHER")                                      ~ "Observational: Accelerated Longitudinal",
      study %in% c("ADAPT-CKD", "PANDA", "IMPROVE", "TEEN-LABS")    ~ "Observational: Longitudinal",
      study %in% c("ATTEMPT", "RPC2", "MANATEE", "T1-DISCO",
                   "MetMod", "OTTER", "INSPIRE", "TODAY")          ~ "Interventional: RCT"
    ),
    exposure_design = dplyr::case_when(
      study %in% c("COFFEE", "ULTRA", "ATTEMPT", "INSPIRE", "MANATEE",
                   "MetMod", "OTTER", "RPC2", "T1-DISCO", "TODAY") ~ "Interventional",
      T                                                                     ~ "Observational"
    ),
    design_struc = dplyr::case_when(
      study %in% c("COFFEE", "ULTRA", "ATTEMPT", "INSPIRE", "MANATEE",
                   "MetMod", "OTTER", "RPC2", "T1-DISCO", "TODAY",
                   "ADAPT-CKD", "IMPROVE", "PANDA", "TEEN-LABS")   ~ "Longitudinal",
      study == "PANTHER"                                            ~ "Longitudinal (Accelerated)",
      T                                                             ~ "Cross-Sectional"
    ),
    study_design_sub = dplyr::case_when(
      study %in% c("ATTEMPT", "INSPIRE", "MANATEE", "MetMod",
                   "OTTER", "RPC2", "T1-DISCO", "TODAY")          ~ "RCT",
      study %in% c("COFFEE", "ULTRA")                              ~ "Single-arm trial",
      study %in% c("IMPROVE", "ADAPT-CKD", "TEEN-LABS")              ~ "Registry cohort",
      study %in% c("PANTHER", "PANDA")                             ~ "Prospective cohort",
      T                                                             ~ "Cohort study"
    ),
    # Standardise visit labels
    visit = dplyr::case_when(
      visit == "screening"                                         ~ "baseline",
      study == "ATTEMPT" & visit == "follow_up"                   ~ "4_months_post",
      is.na(visit) & study_design == "Cross-sectional"            ~ "baseline",
      is.na(visit)                                                ~ "follow up",
      T                                                           ~ visit
    ),
    intervention = dplyr::case_when(
      study == "COFFEE"       ~ "Caffeine (205mg/day)",
      study == "IMPROVE"      ~ "Bariatric Surgery (VSG)",
      study == "ULTRA"        ~ "Pegloticase (Krystexxa 8mg/once)",
      study == "ATTEMPT"      ~ "SGLT2i (Dapagliflozin 5mg/day)",
      study == "RPC2"         ~ "SGLT2i (Canagliflozin 10mg/day)",
      study == "MANATEE"      ~ "AID System + Metformin (2000mg/day)",
      study == "INSPIRE"      ~ "IMST (75% PImax, 30 breaths/day, 6 days/wk)",
      study == "T1-DISCO"     ~ "GLP-1RA (Semaglutide 1mg/week)",
      study == "MetMod"       ~ "AMX0035 (Phenylbutyrate + TUDCA)",
      study == "ADAPT-CKD"    ~ "SGLT2i, GLP-1RA, OR ns-MRA",
      study == "OTTER"        ~ "ns-MRA (Finerenone 20mg/day)",
      study == "TODAY"        ~ "Metformin vs Rosiglitazone vs Lifestyle",
      study == "TEEN-LABS"    ~ "Bariatric Surgery (RYGB, SG, LAGB)"
    )
  ) %>%
  dplyr::filter(!is.na(study_design)) %>%
  # Drop spurious RPC2 "follow up" rows introduced by the bind_rows
  dplyr::filter(study != "RPC2" | visit != "follow up")

# ── (Optional) De-identify for sharing with LLM ───────────────────────────────
# harm_dat_recid <- harm_dat_recid[, !grepl("date", names(harm_dat_recid), ignore.case = TRUE)]
# write.csv(harm_dat_recid,
#           file.path(root_path, "Ye Ji Choi/harm_data_deidentified_022326.csv"),
#           row.names = FALSE, na = "")
```

# Data availability by study

```{r}
# ── 1. Variable lists per procedure domain ─────────────────────────────────────
# Each vector defines the outcome/measurement variables that constitute "having"
# that procedure. intersect() restricts to columns actually present in the data.
cols <- names(harm_dat_recid)

# IVGTT — MinMod-derived insulin kinetics + endocrine labs
ivgtt <- c(
  "mm_si",        # Insulin Sensitivity (Si)
  "mm_airg",      # AIRg (Acute Insulin Response to Glucose)
  "mm_di",        # Disposition Index
  "mm_ir",        # Insulin Resistance
  "mm_bcell",     # Beta Cell Function
  "mm_sg",        # Glucose Effectiveness (Sg, min⁻¹)
  "mm_gb",        # Basal Glucose (mg/dL)
  "mm_ib",        # Basal Insulin (µU/mL)
  "mm_r2"        # MinMod R-squared (goodness of fit)
)

# Kidney PET — perfusion kinetics
kidney_pet <- c(
  "avg_c_k1", "avg_k_k1",
  "avg_c_k2", "avg_k_k2",
  "avg_c_f",  "avg_k_f"
)

# Blood labs (standard clinical + research panel)
blood_labs <- intersect(c(
  "glucose", "hba1c", "fasting_insulin", "cpeptide", "homa_ir",
  "creatinine_s", "cystatin_c_s", "bun_s", "na_s",
  "cholesterol", "hdl", "ldl", "triglycerides",
  "hemoglobin", "hematocrit",
  "gotast_base", "gptalt_base", "alkphos_base",
  "bili_base", "alb_base", "tot_protein",
  "bicarb_base", "ca_base", "cl_base", "k_base", "sodium_base",
  "lh", "fsh", "bl_androst", "bl_dheas",
  "copeptin", "igf1_z_score",
  "alb", "alk_phos", "bili", "bun", "bg"
), cols)

# Urine labs (spot + 24hr collections)
urine_labs <- intersect(c(
  "creatinine_u", "microalbumin_u", "acr_u",
  "elevated_albuminuria", "albuminuria_cat",
  "na_u", "urine_osm",
  "u24_vl", "u24_mab", "u24_na", "u24_gluc", "acr"
), cols)

# Hyperinsulinemic-euglycemic clamp — steady-state outcomes + time-series
clamp_outcomes <- intersect(c(
  "m_value", "gir", "m_ffm",
  "hepatic_glucose_output", "steady_state_glucose", "steady_state_insulin",
  "glucose_infusion_rate_ss", "endogenous_glucose_production",
  "hgc_bg_ss", "hgc_insulin_ss", "hgc_gir_ss", "hgc_cpeptide_ss"
), cols)
clamp_ts_vars <- grep("^(glucose|insulin|cpeptide|ffa|d20_infusion_rate)_\\d", cols, value = TRUE)
clamp_all     <- unique(c(clamp_outcomes, clamp_ts_vars))

# Renal clearance test — iohexol/PAH-based GFR & ERPF
renal_clearance <- intersect(c(
  "gfr_bsa_plasma", "gfr_bsa_plasma_urine", "gfr_raw_plasma", "gfr_raw_plasma_urine",
  "gfr_ecv_std",
  "erpf_bsa_plasma", "erpf_bsa_plasma_urine", "erpf_raw_plasma", "erpf_raw_plasma_urine",
  "rbf", "ff", "ra", "re", "rvr", "pg", "deltapf",
  "glomerular_pressure", "kfg",
  "gfr_p1", "gfr_p2", "pah_bsa", "pah_raw",
  "plasma_ioh_1", "plasma_ioh_2"
), cols)

# BOLD MRI — R2* cortex/medulla + fat signal fraction / FSOC
bold_mri <- intersect(c(
  "bold_l_bl_cortex", "bold_l_bl_medulla", "bold_r_bl_cortex", "bold_r_bl_medulla",
  "bold_l_pf_cortex", "bold_l_pf_medulla", "bold_r_pf_cortex", "bold_r_pf_medulla",
  "avg_k_r2", "avg_c_r2", "avg_m_r2",
  "fsoc_l_cortex", "fsoc_r_cortex",
  "left_cortex_ox_metab_mean", "right_cortex_ox_metab_mean"
), cols)

# DXA — body composition
dxa <- intersect(c(
  "dexa_body_fat", "dexa_lean_mass", "dexa_lean_kg", "dexa_fat_kg",
  "dexa_trunk_mass", "dexa_trunk_kg", "dexa_ag_ratio",
  "dexa_est_vat", "dexa_bone_mineral_density"
), cols)

# Kidney biopsy — morphometric outcomes
kidney_bx <- intersect(c(
  "gloms", "gloms_gs",
  "glom_volume_con", "glom_volume_weibel", "glom_volume_wiggins",
  "glom_tuft_area", "glom_nuc_count", "fia", "acan"
), cols)

# Cardiac MRI — ventricular volumes, EF, output
cardiac_mri <- intersect(c(
  "lvedv", "lvedvi", "lvesv", "lvesvi", "lvef",
  "lvsv", "lv_myo_mass_dias", "lv_myo_mass_syst",
  "lv_cardiac_output", "lv_cardiac_index", "lv_stroke_volume",
  "rvedv", "rvef", "rvsv",
  "cardiac_output", "ef_percent", "edv_ml", "esv_ml", "sv_ml",
  "lved_mass", "lved_massi"
), cols)

# Kidney MRI — size (volume/length), perfusion (ASL/pCASL), BOLD
kidney_size_vars <- intersect(c(
  "left_kidney_volume_ml", "right_kidney_volume_ml",
  "volume_left_manual", "volume_right_manual",
  "length_left", "length_right", "httkv",
  "left_kidney_length", "right_kidney_length", "total_kidney_volume"
), cols)
kidney_perf_vars <- grep("^(pcasl|pasl|asl).*_(left|right)", cols, value = TRUE)
kidney_mri_all   <- unique(c(kidney_size_vars, kidney_perf_vars, bold_mri))

# CGM — time-in-range categories (from Dexcom/CGM download)
cgm_vars <- intersect(c("cgm_v_high", "cgm_high", "cgm_low", "cgm_v_low"), cols)

# J-Wire — renal vein catheter hemodynamics
jwire_vars <- intersect(c("jwire_side", "jwire_num"), cols)

# Lipidomics — lipid species
lipid_vars <- grep("^(lpc|pc|pe|sm|cer|dg|tg_|mg|lpe|fa_|hex|lac)_", cols, value = TRUE)

# Proteomics (SomaScan; retained for completeness even if dropped from main extract)
prot_blood_vars <- intersect(c("seq.10000.28"),        cols)
prot_urine_vars <- intersect(c("seq.10000.28_urine"),  cols)
```

```{r}
# ── 2. Helper: does a study have ≥1 non-missing value across a variable set? ──
has_any <- function(study_name, varlist) {
  present <- intersect(varlist, cols)
  if (length(present) == 0) return(FALSE)
  sub <- harm_dat_recid[
    !is.na(harm_dat_recid$study) & harm_dat_recid$study == study_name,
    present, drop = FALSE
  ]
  any(rowSums(!is.na(sub)) > 0)
}

# ── 3. Build per-study procedure matrix from studies present in the data ───────
studies_in_data <- sort(unique(na.omit(harm_dat_recid$study)))

proc_matrix <- lapply(studies_in_data, function(s) {
  sub    <- harm_dat_recid[!is.na(harm_dat_recid$study) & harm_dat_recid$study == s, ]
  n_pts  <- length(unique(subset(sub, visit == "baseline")$record_id))
  groups <- sort(unique(na.omit(sub$group)))

  list(
    study            = s,
    n                = n_pts,
    groups           = groups,
    exposure_design  = unique(sub$exposure_design),
    design_struc     = unique(sub$design_struc),
    study_design_sub = unique(sub$study_design_sub),
    intervention     = unique(sub$intervention),
    visits           = unique(subset(sub, visit != "baseline")$visit),
    blood_labs       = has_any(s, blood_labs),
    urine_labs       = has_any(s, urine_labs),
    clamp            = has_any(s, clamp_all),
    renal_clear      = has_any(s, renal_clearance),
    dxa              = has_any(s, dxa),
    kidney_bx        = has_any(s, kidney_bx),
    cardiac_mri      = has_any(s, cardiac_mri),
    kidney_mri       = has_any(s, kidney_mri_all),
    lipidomics       = has_any(s, lipid_vars),
    blood_proteomics = has_any(s, prot_blood_vars),
    urine_proteomics = has_any(s, prot_urine_vars),
    kidney_pet       = has_any(s, kidney_pet),
    ivgtt            = has_any(s, ivgtt),
    cgm              = has_any(s, cgm_vars),
    jwire            = has_any(s, jwire_vars)
  )
})

# ── 4. Convert to data frame and apply hard-coded overrides ───────────────────
# Studies still enrolling or actively collecting data are not fully represented
# in the dataset. Procedure flags and N values for those studies are manually
# overridden below. N values represent enrollment targets / upper-bound estimates.
proc_harm_dat_recid <- dplyr::bind_rows(lapply(proc_matrix, function(x) {
  as.data.frame(x[names(x) %nin% c("groups", "visits")], stringsAsFactors = FALSE)
})) %>%
  # ── Procedure flag overrides (hard-coded / still enrolling) ─────────────────
  dplyr::mutate(
    kidney_bx = dplyr::case_when(
      study %in% c("ATTEMPT", "RPC2") ~ TRUE,
      study == "ADAPT-CKD"            ~ TRUE,
      study %in% c("INSPIRE", "MANATEE", "MetMod", "OTTER",
                   "SWEETHEART", "T1-DISCO", "TODAY", "TEEN-LABS") ~ FALSE,
      T ~ kidney_bx
    ),
    vitals = TRUE,
    blood_labs = TRUE,
    urine_labs = TRUE,
    sphyg = study %in% c("INSPIRE", "MANATEE", "OTTER", "T1-DISCO", "RENAL-HEIRitage", "TODAY"),
    # CGM: detected from dataset for PANTHER; hard-coded for others
    cgm = dplyr::case_when(
      study %in% c("MANATEE", "MetMod", "T1-DISCO")  ~ TRUE,  # hard-coded (≤N)
      study == "PANDA"                                ~ TRUE,  # hard-coded (≤N) — PANDA CGM being added
      T ~ cgm                                                  # from data
    ),
    # J-Wire: no vars in dataset yet; hard-coded
    jwire = study %in% c("INSPIRE", "MANATEE", "OTTER", "T1-DISCO", "RENAL-HEIRitage"),
    muscle_bx = study %in% c("MetMod"),
    liver_bx  = study %in% c("IMPROVE"),
    oroboros  = study %in% c("IMPROVE", "MetMod"),
    blood_metabolomics = study %in% c("RENAL-HEIR", "IMPROVE", "RENAL-HEIRitage", "TODAY", "TEEN-LABS"),
    urine_metabolomics = study %in% c("RENAL-HEIR", "IMPROVE", "PANDA", "RENAL-HEIRitage", "TODAY", "TEEN-LABS"),
    clamp = dplyr::case_when(
      study %in% c("ADAPT-CKD", "INSPIRE", "OTTER",
                   "SWEETHEART", "PANTHER", "TODAY", "TEEN-LABS") ~ FALSE,
      study %in% c("MANATEE", "MetMod", "T1-DISCO")         ~ TRUE,
      T ~ clamp
    ),
    renal_clear = dplyr::case_when(
      study %in% c("ATTEMPT", "MANATEE", "OTTER")                         ~ TRUE,
      study %in% c("ADAPT-CKD", "INSPIRE", "MetMod",
                   "SWEETHEART", "T1-DISCO", "TODAY", "TEEN-LABS") ~ FALSE,
      T ~ renal_clear
    ),
    dxa = dplyr::case_when(
      study %in% c("ADAPT-CKD", "INSPIRE") ~ FALSE,
      study %in% c("MANATEE", "MetMod", "OTTER", "SWEETHEART", "T1-DISCO",
                   "TODAY", "TEEN-LABS") ~ TRUE,
      T ~ dxa
    ),
    cardiac_mri = dplyr::case_when(
      study %in% c("ADAPT-CKD", "INSPIRE", "MANATEE",
                   "MetMod", "TODAY", "TEEN-LABS") ~ FALSE,
      study %in% c("OTTER", "SWEETHEART", "T1-DISCO")    ~ TRUE,
      T ~ cardiac_mri
    ),
    echo = study %in% c("T1-DISCO", "OTTER", "TODAY"),
    fmd  = study %in% c("T1-DISCO", "OTTER"),
    kidney_mri = dplyr::case_when(
      study %in% c("ADAPT-CKD", "INSPIRE", "MANATEE",
                   "SWEETHEART", "T1-DISCO", "TODAY", "TEEN-LABS") ~ FALSE,
      study %in% c("MetMod", "OTTER")                        ~ TRUE,
      T ~ kidney_mri
    ),
    lipidomics = dplyr::case_when(
      study %in% c("ADAPT-CKD", "INSPIRE", "MANATEE", "MetMod",
                   "OTTER", "SWEETHEART", "T1-DISCO", "TODAY", "TEEN-LABS") ~ FALSE,
      T ~ lipidomics
    ),
    blood_proteomics = dplyr::case_when(
      study %in% c("ADAPT-CKD", "OTTER", "TEEN-LABS", "TODAY") ~ TRUE,
      study %in% c("INSPIRE", "MANATEE", "MetMod",
                   "SWEETHEART", "T1-DISCO") ~ FALSE,
      T ~ blood_proteomics
    ),
    urine_proteomics = dplyr::case_when(
      study == "ADAPT-CKD"                             ~ TRUE,
      study %in% c("INSPIRE", "MANATEE", "MetMod",
                   "OTTER", "SWEETHEART", "T1-DISCO", "TODAY", "TEEN-LABS") ~ FALSE,
      T ~ urine_proteomics
    ),
    # DECODE membership
    decode = dplyr::case_when(
      study %in% c("CROCODILE", "RENAL-HEIR", "RENAL-HEIRitage", "RPC2",
                   "PANDA", "CASPER", "COFFEE", "PENGUIN", "T1-DISCO",
                   "REMODEL-T1D", "ATTEMPT", "PANTHER", "IMPROVE",
                   "TODAY", "TEEN-LABS")  ~ "DECODE",
      T                                                  ~ "NOT PART OF DECODE"
    ),
    # Intervention label: blank for observational
    intervention = dplyr::case_when(is.na(intervention) ~ "", T ~ intervention),
    # N: replace with upper-bound estimates for studies still enrolling
    n = as.character(n),
    n = dplyr::case_when(
      study == "ADAPT-CKD"    ~ "≤100",
      study == "INSPIRE"      ~ "≤90",
      study == "MANATEE"      ~ "≤60",
      study == "MetMod"       ~ "≤60",
      study == "OTTER"        ~ "≤60",
      study == "SWEETHEART"   ~ "≤30",
      study == "T1-DISCO"     ~ "≤60",
      study == "TODAY"        ~ as.character(today_n_total),
      study == "TEEN-LABS"    ~ "≤242",
      T ~ n
    )
  )

# ── 5. Define ordered procedure columns ───────────────────────────────────────
flagged_procs <- c(
  "vitals", "blood_labs", "urine_labs", "dxa",
  "clamp", "ivgtt", "oroboros",
  "renal_clear", "kidney_mri", "kidney_pet",
  "sphyg", "cardiac_mri", "echo", "fmd",
  "kidney_bx", "muscle_bx", "liver_bx", "jwire",
  "lipidomics", "blood_proteomics", "urine_proteomics",
  "blood_metabolomics", "urine_metabolomics",
  "cgm"
)

# ── 6. Build lookup lists (groups, visits, decode) for HTML badge helpers ──────
groups_list <- lapply(proc_matrix, `[[`, "groups")
names(groups_list) <- proc_harm_dat_recid$study

visits_list <- lapply(proc_matrix, `[[`, "visits")
names(visits_list) <- proc_harm_dat_recid$study
visits_list$COFFEE    <- "WK1"               # override: COFFEE follow-up is week 1
visits_list$TODAY     <- "ANNUAL UP TO 15 YRS"  # single special label
visits_list$`TEEN-LABS` <- c("MO6", "YR1", "YR2", "YR3", "YR5")

decode_list <- as.list(proc_harm_dat_recid$decode)
names(decode_list) <- proc_harm_dat_recid$study
```

```{r}
# ── 7. Abbreviation helpers ────────────────────────────────────────────────────

# Map group labels to display abbreviations
abbrev_group <- function(g) {
  dplyr::case_when(
    grepl("Type 1",       g, ignore.case = TRUE)  ~ "T1D",
    grepl("Type 2",       g, ignore.case = TRUE)  ~ "T2D",
    grepl("^Lean",        g, ignore.case = TRUE)  ~ "HC",
    grepl("^Obese",       g, ignore.case = TRUE)  ~ "OB",
    grepl("PKD|ADPKD",    g, ignore.case = TRUE)  ~ "ADPKD",
    TRUE ~ g
  )
}

# Map visit labels to short codes
abbrev_visit <- function(v) {
  dplyr::case_when(
    grepl("baseline",                v, ignore.case = TRUE)  ~ "0",
    grepl("4_weeks_post",            v, ignore.case = TRUE)  ~ "MO1",
    grepl("3_months_post_surgery",   v, ignore.case = TRUE)  ~ "MO3",
    grepl("4_months_post",           v, ignore.case = TRUE)  ~ "MO4",
    grepl("12_months_post_surgery",  v, ignore.case = TRUE)  ~ "YR1",
    grepl("year_1",                  v, ignore.case = TRUE)  ~ "YR1",
    grepl("year_2",                  v, ignore.case = TRUE)  ~ "YR2",
    grepl("year_3",                  v, ignore.case = TRUE)  ~ "YR3",
    grepl("year_4",                  v, ignore.case = TRUE)  ~ "YR4",
    grepl("follow up",               v, ignore.case = TRUE)  ~ "WK1",
    grepl("post_treatment",          v, ignore.case = TRUE)  ~ "MO6",
    TRUE ~ v
  )
}
```

```{r}
# ── 8. HTML badge / dot helpers ────────────────────────────────────────────────

# Color palettes
cohort_colors <- c(
  T1D   = "#e77c76",
  T2D   = "#e37902",
  HC    = "#8cb369",
  OB    = "#5b8e7d",
  ADPKD = "#bc4b51"
)

visit_colors <- c(
  `0`   = "#d6ccc2",
  `WK1` = "#f8ad9d",
  `MO1` = "#ff9e00",
  `MO3` = "#ff9100",
  `MO4` = "#ff8500",
  `MO6` = "#ff6d00",
  `YR1` = "#64a4ce",
  `YR2` = "#4893c6",
  `YR3` = "#2d82bd",
  `YR4` = "#2476b1",
  `YR5` = "#1a6aa5",
  `ANNUAL UP TO 15 YRS` = "#00314D"
)

decode_colors <- c(
  "DECODE"            = "#00314D",
  "NOT PART OF DECODE" = "#c1121f"
)

# Solid pill badge (cohort)
badge_html <- function(label) {
  bg <- cohort_colors[label]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.8em;font-weight:600;',
           'padding:2px 7px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, label
  )
}

# Solid pill badge (visit)
badge_visit_html <- function(label) {
  bg <- visit_colors[label]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.8em;font-weight:600;',
           'padding:2px 7px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, label
  )
}

# Small pill badge (DECODE membership)
badge_decode_html <- function(label) {
  bg <- decode_colors[label]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.6em;font-weight:600;',
           'padding:2px 7px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, label
  )
}

# Compose cohort, visit, and DECODE badge strings per study
make_cohort_html <- function(study_name) {
  abbr <- unique(abbrev_group(groups_list[[study_name]]))
  paste(vapply(abbr, badge_html, character(1)), collapse = "")
}

make_visit_html <- function(study_name) {
  abbr <- sort(unique(abbrev_visit(visits_list[[study_name]])))
  paste(vapply(abbr, badge_visit_html, character(1)), collapse = "")
}

make_decode_html <- function(study_name) {
  paste(vapply(sort(unique(decode_list[[study_name]])),
               badge_decode_html, character(1)), collapse = "")
}

# Red filled dot = procedure present; grey hollow dot = absent
dot_yes <- '<span style="color:#cb1b16;font-size:1.5em;">&#9679;</span>'
dot_no  <- '<span style="color:#d1d5db;font-size:1.5em;">&#9675;</span>'
dot     <- function(x) ifelse(x, dot_yes, dot_no)
```

```{r}
# ── 9. Assemble display data frame ─────────────────────────────────────────────
display_harm_dat_recid <- proc_harm_dat_recid %>%
  dplyr::mutate(
    cohorts_html = sapply(study, make_cohort_html),
    visits_html  = sapply(study, make_visit_html),
    decode_html  = sapply(study, make_decode_html),
    dplyr::across(dplyr::all_of(flagged_procs), ~ sapply(.x, dot))
  ) %>%
  dplyr::select(
    decode_html, study, exposure_design, design_struc, study_design_sub,
    intervention, cohorts_html, visits_html, n,
    dplyr::all_of(flagged_procs)
  ) %>%
  dplyr::arrange(decode_html, exposure_design, dplyr::desc(design_struc), study_design_sub)
```

```{r}
# ── 10. Shared gt styling helper ──────────────────────────────────────────────
# Applies consistent column labels, spanners, and styles to any procedure matrix
# gt table. Used by both the dot version and the N version.
apply_proc_gt_style <- function(tbl) {
  tbl %>%
    cols_label(
      decode_html      = "PART OF DECODE",
      study            = "STUDY",
      exposure_design  = "STUDY DESIGN",
      design_struc     = "TEMPORAL DESIGN",
      study_design_sub = "STUDY SUB-DESIGN",
      intervention     = "EXPOSURE",
      visits_html      = "FOLLOW-UP VISITS",
      cohorts_html     = "COHORTS",
      n                = "N AT BASELINE",
      vitals           = "VITALS",
      blood_labs       = "BLOOD",
      urine_labs       = "URINE",
      clamp            = "CLAMP",
      renal_clear      = "RENAL CLEARANCE TEST",
      dxa              = "DXA",
      kidney_bx        = "KIDNEY BX",
      cardiac_mri      = "CARDIAC MRI",
      echo             = "ECHO",
      fmd              = "FMD",
      kidney_mri       = "KIDNEY MRI",
      lipidomics       = "LIPID",
      blood_proteomics = "PROTEOMICS (PLASMA)",
      urine_proteomics = "PROTEOMICS (URINE)",
      blood_metabolomics = "METABO (PLASMA)",
      urine_metabolomics = "METABO (URINE)",
      muscle_bx        = "MUSCLE BX",
      liver_bx         = "LIVER BX",
      jwire            = "J-WIRE",
      sphyg            = "SPHYGMOCOR",
      kidney_pet       = "KIDNEY PET",
      cgm              = "CGM"
    ) %>%
    fmt_markdown(columns = everything()) %>%
    tab_spanner(label = "Labs",             columns = c(blood_labs, urine_labs)) %>%
    tab_spanner(label = "Metabolic Testing",columns = c(clamp, ivgtt, oroboros)) %>%
    tab_spanner(label = "Kidney Function",  columns = c(renal_clear, kidney_mri, kidney_pet)) %>%
    tab_spanner(label = "Cardiac",          columns = c(sphyg, cardiac_mri, echo, fmd)) %>%
    tab_spanner(label = "Biopsies/Procedures",
                columns = c(kidney_bx, muscle_bx, liver_bx, jwire)) %>%
    tab_spanner(label = "Omics",
                columns = c(lipidomics, blood_proteomics, urine_proteomics,
                            blood_metabolomics, urine_metabolomics)) %>%
    tab_style(
      style     = cell_text(weight = "bold", size = "12px",
                            transform = "uppercase", color = "#374151"),
      locations = cells_column_spanners()
    ) %>%
    tab_style(
      style     = cell_text(weight = "bold", size = "15px"),
      locations = cells_body(columns = study)
    ) %>%
    tab_style(
      style     = cell_text(weight = "bold", size = "12px",
                            transform = "uppercase", color = "#374151"),
      locations = cells_column_labels()
    ) %>%
    tab_style(
      style     = cell_text(size = "15px"),
      locations = cells_body()
    ) %>%
    cols_align(
      align   = "center",
      columns = c(decode_html, study, exposure_design, design_struc, study_design_sub,
                  visits_html, intervention, cohorts_html, n, all_of(flagged_procs))
    ) %>%
    cols_width(
      study            ~ px(145),
      exposure_design  ~ px(150),
      design_struc     ~ px(150),
      study_design_sub ~ px(150),
      visits_html      ~ px(120),
      intervention     ~ px(150),
      cohorts_html     ~ px(120),
      n                ~ px(70),
      all_of(flagged_procs) ~ px(70)
    ) %>%
    opt_row_striping(row_striping = TRUE)
}
```

## Dot version (presence / absence)

```{r}
# ── 11. Build and render the dot-style procedure table ─────────────────────────
study_gt_table <- display_harm_dat_recid %>%
  gt(id = "harm-study-table") %>%
  apply_proc_gt_style()

study_gt_table

study_gt_table %>%
  gtsave(
    file.path(root_path,
              "Data Harmonization/Report_files/Summary Report/summary_table.png"),
    vwidth = 3000, vheight = 1000
  )

study_gt_table %>% gt::opt_interactive()
```

## N version (per-procedure sample sizes)

```{r}
# ── 12. Per-procedure N computation ───────────────────────────────────────────
# For studies present in the dataset: count unique participants with ≥1 non-NA
# value in the relevant variable set (at baseline visit).
# For hard-coded / still-enrolling studies: show ≤[hard-coded N].

# Map each procedure flag to its variable set
proc_varlist <- list(
  vitals           = c("height", "weight", "bmi", "sbp", "dbp"),   # common vitals
  blood_labs       = blood_labs,
  urine_labs       = urine_labs,
  clamp            = clamp_all,
  ivgtt            = ivgtt,
  oroboros         = character(0),   # no harmonized vars yet
  renal_clear      = renal_clearance,
  kidney_mri       = kidney_mri_all,
  kidney_pet       = kidney_pet,
  sphyg            = character(0),   # no harmonized vars yet
  cardiac_mri      = cardiac_mri,
  echo             = character(0),   # no harmonized vars yet
  fmd              = character(0),   # no harmonized vars yet
  kidney_bx        = kidney_bx,
  muscle_bx        = character(0),   # no harmonized vars yet
  liver_bx         = character(0),   # no harmonized vars yet
  jwire            = jwire_vars,
  lipidomics       = lipid_vars,
  blood_proteomics = prot_blood_vars,
  urine_proteomics = prot_urine_vars,
  blood_metabolomics = character(0), # no harmonized vars yet
  urine_metabolomics = character(0), # no harmonized vars yet
  cgm              = cgm_vars
)

# Studies that are hard-coded (not yet in dataset) — show ≤N
hardcoded_studies <- c("ADAPT-CKD", "INSPIRE", "MANATEE", "MetMod",
                       "OTTER", "SWEETHEART", "T1-DISCO")

# Count baseline non-missing participants per study × procedure
count_proc_n <- function(study_name, proc_name) {
  vlist   <- intersect(proc_varlist[[proc_name]], cols)
  proc_on <- proc_harm_dat_recid[proc_harm_dat_recid$study == study_name,
                                 proc_name, drop = TRUE]
  # Procedure not done for this study
  if (!isTRUE(proc_on)) return("")

  # Hard-coded study: return ≤N
  hard_n <- proc_harm_dat_recid$n[proc_harm_dat_recid$study == study_name]
  if (study_name %in% hardcoded_studies) return(paste0(hard_n))

  # TODAY: Ns computed from separate data files (not in harm_dat_recid)
  if (study_name == "TODAY") {
    today_proc_ns <- c(
      vitals = today_n_vitals, blood_labs = today_n_blood,
      urine_labs = today_n_urine, dxa = today_n_dxa,
      echo = today_n_echo, sphyg = today_n_pwv,
      blood_proteomics = today_n_blood_prot,
      blood_metabolomics = today_n_blood_met,
      urine_metabolomics = today_n_urine_met
    )
    if (proc_name %in% names(today_proc_ns)) return(as.character(today_proc_ns[[proc_name]]))
    return(as.character(hard_n))
  }

  # TEEN-LABS: hardcoded ≤242 for most procedures; proteomics uses exact count
  if (study_name == "TEEN-LABS") {
    teenlabs_hardcoded_ns <- c(
      vitals = 242L, blood_labs = 242L,
      urine_labs = 242L, dxa = 242L,
      blood_metabolomics = 242L, urine_metabolomics = 242L
    )
    if (proc_name %in% names(teenlabs_hardcoded_ns)) return(paste0("≤", teenlabs_hardcoded_ns[[proc_name]]))
    if (proc_name == "blood_proteomics") return(as.character(teenlabs_n_soma))
    return(paste0("≤", hard_n))
  }

  # No vars mapped: return ≤N as fallback
  if (length(vlist) == 0) return(paste0(hard_n))

  # Compute from data (baseline rows only)
  sub <- harm_dat_recid[
    !is.na(harm_dat_recid$study) &
      harm_dat_recid$study == study_name &
      harm_dat_recid$visit == "baseline",
    vlist, drop = FALSE
  ]
  n <- sum(rowSums(!is.na(sub)) > 0)
  if (n == 0) as.character(hard_n) else as.character(n)
}

# Build N display data frame — same structure as display_harm_dat_recid
display_n_table <- proc_harm_dat_recid %>%
  dplyr::mutate(
    cohorts_html = sapply(study, make_cohort_html),
    visits_html  = sapply(study, make_visit_html),
    decode_html  = sapply(study, make_decode_html)
  )

# Replace dot columns with N strings
for (proc in flagged_procs) {
  display_n_table[[proc]] <- vapply(
    display_n_table$study,
    function(s) count_proc_n(s, proc),
    character(1)
  )
}

display_n_table <- display_n_table %>%
  dplyr::select(
    decode_html, study, exposure_design, design_struc, study_design_sub,
    intervention, cohorts_html, visits_html, n,
    dplyr::all_of(flagged_procs)
  ) %>%
  dplyr::arrange(decode_html, exposure_design, dplyr::desc(design_struc), study_design_sub)
```

```{r}
# ── 13. Render N-version table ─────────────────────────────────────────────────
study_gt_n_table <- display_n_table %>%
  gt(id = "harm-study-table-n") %>%
  apply_proc_gt_style()

study_gt_n_table

study_gt_n_table %>%
  gtsave(
    file.path(root_path,
              "Data Harmonization/Report_files/Summary Report/summary_table_n.png"),
    vwidth = 3000, vheight = 1000
  )

study_gt_n_table %>% gt::opt_interactive()
```

---

# Data availability by cohort

```{r}
# ── 14. Cohort-table HTML helpers ──────────────────────────────────────────────
# NOTE: cohort_colors, visit_colors, badge_html, badge_visit_html,
# abbrev_group, abbrev_visit, dot_yes/dot_no/dot, groups_list, visits_list,
# proc_harm_dat_recid, and flagged_procs are all defined in Section 1 above.

# Color map: study name → design-type color (for solid study bubbles)
study_palette <- c(
  # RCT
  "ATTEMPT"         = "#606c38",
  "INSPIRE"         = "#606c38",
  "MANATEE"         = "#606c38",
  "MetMod"          = "#606c38",
  "OTTER"           = "#606c38",
  "RPC2"            = "#606c38",
  "T1-DISCO"        = "#606c38",
  "TODAY"           = "#606c38",
  # Single-arm trial
  "COFFEE"          = "#f48c06",
  "ULTRA"           = "#f48c06",
  # Registry cohort
  "IMPROVE"         = "#f28482",
  "ADAPT-CKD"       = "#f28482",
  "TEEN-LABS"       = "#f28482",
  # Prospective cohort
  "PANTHER"         = "#0891b2",
  "PANDA"           = "#0891b2",
  # Cohort study
  "CASPER"          = "#14213d",
  "CROCODILE"       = "#14213d",
  "PENGUIN"         = "#14213d",
  "RENAL-HEIR"      = "#14213d",
  "RENAL-HEIRitage" = "#14213d",
  "SWEETHEART"      = "#14213d"
)

# Colors for pooled design-type outline tags
design_colors <- c(
  "RCT"              = "#606c38",
  "Single-arm trial" = "#f48c06",
  "Registry cohort"  = "#f28482",
  "Prospective cohort" = "#0891b2",
  "Cohort study"     = "#14213d"
)

# Outlined design-type tag (used in cohort table)
tag_html <- function(label, color) {
  sprintf(
    paste0('<span style="background:%s20;color:%s;border:1px solid %s40;',
           'font-size:0.7em;font-weight:600;padding:1px 5px;border-radius:4px;',
           'margin:1px 2px;display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    color, color, color, label
  )
}

# Solid study-name bubble
study_badge_html <- function(study_name) {
  bg <- study_palette[study_name]
  if (is.na(bg)) bg <- "#6b7280"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.75em;font-weight:700;',
           'padding:2px 8px;border-radius:999px;margin:1px 2px;',
           'display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    bg, study_name
  )
}

# Large cohort-type row-label pill
cohort_label_html <- function(cohort) {
  bg <- cohort_colors[cohort]
  if (is.na(bg)) bg <- "#9ca3af"
  sprintf(
    paste0('<span style="background:%s;color:white;font-size:0.9em;font-weight:700;',
           'padding:3px 10px;border-radius:999px;display:inline-block;',
           'white-space:nowrap;line-height:1.7">%s</span>'),
    bg, cohort
  )
}

# Outlined visit badge (cohort table uses outlined style)
visit_badge_cohort_html <- function(v) {
  col <- visit_colors[v]
  if (is.na(col)) col <- "#9ca3af"
  sprintf(
    paste0('<span style="background:white;color:%s;border:1px solid %s;',
           'font-size:0.7em;font-weight:600;padding:1px 5px;border-radius:999px;',
           'margin:1px 2px;display:inline-block;white-space:nowrap;line-height:1.7">%s</span>'),
    col, col, v
  )
}
```

```{r}
# ── 15. Cohort table: data preparation ────────────────────────────────────────

# Derive per-study cohort abbreviations
cohort_abbrev_per_study <- lapply(names(groups_list), function(s) {
  sort(unique(abbrev_group(groups_list[[s]])))
})
names(cohort_abbrev_per_study) <- names(groups_list)

# Derive per-study follow-up visit string (exclude baseline)
followup_str_per_study <- lapply(names(visits_list), function(s) {
  vs   <- visits_list[[s]]
  vs   <- vs[!is.na(vs)]
  abbr <- sort(unique(abbrev_visit(vs)))
  abbr <- abbr[abbr != "0"]
  if (length(abbr) == 0) return(NA_character_)
  paste(abbr, collapse = ",")
})
names(followup_str_per_study) <- names(visits_list)

# Slim metadata frame: study | cohorts_str | followup_visits | n_str
study_meta_cohort <- data.frame(
  study = proc_harm_dat_recid$study,
  cohorts_str = sapply(proc_harm_dat_recid$study, function(s) {
    v <- cohort_abbrev_per_study[[s]]
    if (is.null(v) || length(v) == 0) NA_character_ else paste(v, collapse = ",")
  }),
  followup_visits = sapply(proc_harm_dat_recid$study, function(s) {
    v <- followup_str_per_study[[s]]
    if (is.null(v) || length(v) == 0) NA_character_ else v
  }),
  n_str = proc_harm_dat_recid$n,
  stringsAsFactors = FALSE
)

# Join with procedure flags
study_full_cohort <- study_meta_cohort %>%
  dplyr::left_join(
    proc_harm_dat_recid %>%
      dplyr::select(study, decode, exposure_design, design_struc, study_design_sub,
             dplyr::all_of(flagged_procs)),
    by = "study"
  )

# Expand to one row per study × cohort
study_cohort_long <- study_full_cohort %>%
  dplyr::mutate(cohort_list = strsplit(cohorts_str, ",")) %>%
  tidyr::unnest(cohort_list) %>%
  dplyr::mutate(cohort = trimws(cohort_list))

cohort_order <- c("HC", "OB", "T1D", "T2D", "ADPKD")

# ── build_cohort_display(): aggregate to one gt-ready row per cohort ───────────
build_cohort_display <- function(subset_decode = c("ALL", "DECODE", "NOT PART OF DECODE")) {
  subset_decode <- match.arg(subset_decode)

  dat <- if (subset_decode == "ALL") {
    study_cohort_long
  } else {
    study_cohort_long %>% dplyr::filter(decode == subset_decode)
  }

  cohort_rows <- lapply(cohort_order, function(ch) {
    ch_dat <- dat %>% dplyr::filter(cohort == ch)
    if (nrow(ch_dat) == 0) return(NULL)

    studies_in_ch <- ch_dat$study

    # Cohort row label
    cohort_html <- cohort_label_html(ch)

    # Aggregate N: sum numeric Ns, keep ≤ prefix if any study is capped
    ns         <- ch_dat$n_str
    numeric_ns <- suppressWarnings(as.numeric(ns))
    capped_ns  <- as.numeric(gsub("≤", "", ns))
    sum_num    <- sum(numeric_ns, capped_ns, na.rm = TRUE)
    n_display  <- if (length(capped_ns) == 0) {
      as.character(sum_num)
    } else if (sum_num == 0) {
      paste(collapse = "≤", unique(capped_ns))
    } else {
      paste0("≤", sum_num)
    }

    # Study bubbles + follow-up visit badges (one line per study)
    study_visit_parts <- vapply(studies_in_ch, function(s) {
      sb  <- study_badge_html(s)
      row <- ch_dat[ch_dat$study == s, ]
      vstr <- row$followup_visits
      if (!is.na(vstr) && nchar(vstr) > 0) {
        vv  <- trimws(unlist(strsplit(vstr, ",")))
        vbs <- paste(vapply(vv, visit_badge_cohort_html, character(1)), collapse = "")
        paste0(sb, vbs)
      } else {
        sb
      }
    }, character(1))
    studies_html <- paste(study_visit_parts, collapse = "<br>")

    # Design-type tags (unique across studies in this cohort)
    designs     <- unique(ch_dat$study_design_sub)
    designs     <- designs[!is.na(designs) & nchar(trimws(designs)) > 0]
    design_html <- paste(vapply(designs, function(d) {
      col <- design_colors[d]; if (is.na(col)) col <- "#6b7280"
      tag_html(d, col)
    }, character(1)), collapse = " ")

    # Procedure dots: TRUE if ANY study in this cohort has the procedure
    proc_vals <- vapply(flagged_procs, function(p) {
      vals <- ch_dat[[p]]
      if (is.logical(vals)) any(vals, na.rm = TRUE)
      else any(as.logical(vals), na.rm = TRUE)
    }, logical(1))

    row_data <- c(
      list(cohort_html  = cohort_html,
           studies_html = studies_html,
           design_html  = design_html,
           n            = n_display),
      as.list(proc_vals)
    )
    as.data.frame(row_data, stringsAsFactors = FALSE, check.names = FALSE)
  })

  cohort_df <- dplyr::bind_rows(Filter(Negate(is.null), cohort_rows)) %>%
    dplyr::mutate(dplyr::across(dplyr::all_of(flagged_procs),
                  ~ sapply(.x, function(v) dot(as.logical(v)))))
  cohort_df
}

# Build display data frames for all / DECODE-only / non-DECODE subsets
df_cohort_all    <- build_cohort_display("ALL")
df_cohort_decode <- build_cohort_display("DECODE")
df_cohort_nondec <- build_cohort_display("NOT PART OF DECODE")
```

```{r}
# ── 16. Cohort gt table builder ────────────────────────────────────────────────
make_cohort_gt <- function(display_df, title_suffix = "") {
  display_df %>%
    gt(id = paste0("harm-cohort-table", gsub("[^A-Za-z0-9]", "_", title_suffix))) %>%
    cols_label(
      cohort_html      = "COHORT",
      studies_html     = "STUDIES & FOLLOW-UP",
      design_html      = "STUDY DESIGNS",
      n                = "N",
      vitals           = "VITALS",
      blood_labs       = "BLOOD",
      urine_labs       = "URINE",
      clamp            = "CLAMP",
      renal_clear      = "RENAL CLEARANCE TEST",
      dxa              = "DXA",
      kidney_bx        = "KIDNEY BX",
      cardiac_mri      = "CARDIAC MRI",
      echo             = "ECHO",
      fmd              = "FMD",
      kidney_mri       = "KIDNEY MRI",
      lipidomics       = "LIPID",
      blood_proteomics = "PROTEOMICS (PLASMA)",
      urine_proteomics = "PROTEOMICS (URINE)",
      blood_metabolomics = "METABO (PLASMA)",
      urine_metabolomics = "METABO (URINE)",
      muscle_bx        = "MUSCLE BX",
      liver_bx         = "LIVER BX",
      jwire            = "J-WIRE",
      sphyg            = "SPHYGMOCOR",
      kidney_pet       = "KIDNEY PET",
      cgm              = "CGM"
    ) %>%
    fmt_markdown(columns = everything()) %>%
    tab_spanner(label = "Labs",              columns = c(blood_labs, urine_labs)) %>%
    tab_spanner(label = "Metabolic Testing", columns = c(clamp, ivgtt, oroboros)) %>%
    tab_spanner(label = "Kidney Function",   columns = c(renal_clear, kidney_mri, kidney_pet)) %>%
    tab_spanner(label = "Cardiac",           columns = c(sphyg, cardiac_mri, echo, fmd)) %>%
    tab_spanner(label = "Biopsies/Procedures",
                columns = c(kidney_bx, muscle_bx, liver_bx, jwire)) %>%
    tab_spanner(label = "Omics",
                columns = c(lipidomics, blood_proteomics, urine_proteomics,
                            blood_metabolomics, urine_metabolomics)) %>%
    tab_style(
      style     = cell_text(weight = "bold", size = "12px",
                            transform = "uppercase", color = "#374151"),
      locations = cells_column_spanners()
    ) %>%
    tab_style(
      style     = cell_text(weight = "bold", size = "12px",
                            transform = "uppercase", color = "#374151"),
      locations = cells_column_labels()
    ) %>%
    tab_style(
      style     = cell_text(size = "15px"),
      locations = cells_body()
    ) %>%
    cols_align(align = "center") %>%
    cols_width(
      cohort_html      ~ px(80),
      studies_html     ~ px(200),
      design_html      ~ px(150),
      n                ~ px(50),
      blood_proteomics ~ px(90),
      urine_proteomics ~ px(90),
      all_of(flagged_procs) ~ px(80)
    ) %>%
    opt_row_striping(row_striping = TRUE)
}
```

## All studies

```{r}
all_cohort_table <- make_cohort_gt(df_cohort_all, "")
all_cohort_table

all_cohort_table %>%
  gtsave(
    file.path(root_path,
              "Data Harmonization/Report_files/Summary Report/summary_table_by_cohort.png"),
    vwidth = 4500, vheight = 1000
  )

all_cohort_table %>% gt::opt_interactive()
```

## DECODE studies only

```{r}
decode_cohort_table <- make_cohort_gt(df_cohort_decode, " · DECODE Studies")
decode_cohort_table

decode_cohort_table %>%
  gtsave(
    file.path(root_path,
              "Data Harmonization/Report_files/Summary Report/summary_table_by_cohort_decode_only.png"),
    vwidth = 4500, vheight = 1000
  )

decode_cohort_table %>% gt::opt_interactive()
```

## Non-DECODE studies only

```{r}
non_decode_cohort_table <- make_cohort_gt(df_cohort_nondec, " · Non-DECODE Studies")
non_decode_cohort_table

non_decode_cohort_table %>%
  gtsave(
    file.path(root_path,
              "Data Harmonization/Report_files/Summary Report/summary_table_by_cohort_nondecode_only.png"),
    vwidth = 4500, vheight = 1000
  )

non_decode_cohort_table %>% gt::opt_interactive()
```

---

# Summary visualizations

```{r}
# ── 17. Shared ggplot theme and palette ───────────────────────────────────────
theme_harm <- theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor  = element_blank(),
    strip.text        = element_text(face = "bold"),
    legend.position   = "bottom",
    axis.text.x       = element_text(angle = 45, hjust = 1)
  )

# Cohort fill colors matching badge palette
cohort_fill <- c(T1D = "#e77c76", T2D = "#e37902",
                 HC  = "#8cb369", OB  = "#5b8e7d", ADPKD = "#bc4b51")

# Procedure order matches flagged_procs / main summary table column order.
# Only procedures that have a variable set (can be counted from data or estimated)
# are included here; vitals, oroboros, sphyg, jwire, muscle_bx, liver_bx,
# metabolomics are excluded because they have no mapped variable sets for counting.
proc_varlist_plot <- list(
  "Blood Labs"        = blood_labs,
  "Urine Labs"        = urine_labs,
  "DXA"               = dxa,
  "Clamp"             = clamp_all,
  "IVGTT"             = ivgtt,
  "Renal Clearance"   = renal_clearance,
  "Kidney MRI"        = kidney_mri_all,
  "Kidney PET"        = kidney_pet,
  "Cardiac MRI"       = cardiac_mri,
  "ECHO"              = character(0),
  "FMD"               = character(0),
  "Kidney Biopsy"     = kidney_bx,
  "Lipidomics"        = lipid_vars,
  "Blood Proteomics"  = prot_blood_vars,
  "Urine Proteomics"  = prot_urine_vars,
  "Blood Metabolomics" = character(0),  # no harmonized vars; Ns from separate data
  "Urine Metabolomics" = character(0),  # no harmonized vars; Ns from separate data
  "CGM"               = cgm_vars
)

# Ordered factor levels for the x-axis (used in bar plots)
proc_order <- names(proc_varlist_plot)
```

```{r}
# ── 18. Hard-coded study metadata for "full" visualizations ───────────────────
# Studies still enrolling / data not yet entered — procedure flags and
# estimated N by group from proc_harm_dat_recid (≤ upper-bound values).
# Group compositions are known from study protocols.

# Map each hard-coded study to: group abbreviation, and estimated ≤N per group.
# These Ns are the hard-coded totals; per-procedure N ≤ total N.
hardcoded_meta <- list(
  # group_n: named vector of group_abbr → estimated N (≤N upper bound)
  "ADAPT-CKD"    = list(group_n = c(OB = 50, T2D = 50)),
  "INSPIRE"      = list(group_n = c(T2D = 90)),
  "MANATEE"      = list(group_n = c(T1D = 60)),
  "MetMod"       = list(group_n = c(T1D = 60)),
  "OTTER"        = list(group_n = c(T1D = 60)),
  "SWEETHEART"   = list(group_n = c(T2D = 18, OB = 9, HC = 4)),
  "T1-DISCO"     = list(group_n = c(T1D = 60)),
  "TODAY"        = list(group_n = c(T2D = today_n_total)),
  "TEEN-LABS"    = list(group_n = c(T2D = 34, OB = 208))
)

# Set of hard-coded study names (used for ≤ labelling logic)
# TODAY is in hardcoded_meta but uses exact (not estimated) Ns,
# so it is excluded from this vector to avoid the ≤ prefix in labels.
# TEEN-LABS uses hardcoded ≤242 until the full dataset is available.
hardcoded_studies <- setdiff(names(hardcoded_meta), c("TODAY"))

# Map proc flag column name → plot label (must match proc_varlist_plot keys)
proc_flag_to_label <- c(
  blood_labs       = "Blood Labs",
  urine_labs       = "Urine Labs",
  dxa              = "DXA",
  clamp            = "Clamp",
  ivgtt            = "IVGTT",
  renal_clear      = "Renal Clearance",
  kidney_mri       = "Kidney MRI",
  kidney_pet       = "Kidney PET",
  cardiac_mri      = "Cardiac MRI",
  echo             = "ECHO",
  fmd              = "FMD",
  kidney_bx        = "Kidney Biopsy",
  lipidomics       = "Lipidomics",
  blood_proteomics = "Blood Proteomics",
  urine_proteomics = "Urine Proteomics",
  blood_metabolomics = "Blood Metabolomics",
  urine_metabolomics = "Urine Metabolomics",
  cgm              = "CGM"
)
```


```{r}
# ── 18b. Pre-compute deduplicated baseline frame ────────────────────────────
# This expensive summarise (collapse duplicate record_id rows) is needed by
# make_pie_dat, make_bar_dat, and bp_base.  Computing it once here avoids
# re-running it inside every function call / lapply iteration.
# Deduplicate baseline rows — collapse duplicate record_id rows by taking
# last non-NA for character columns and mean for numeric columns.
# Uses a single across() with a type-aware lambda to avoid double-across issues.
baseline_dedup <- harm_dat_recid %>%
  dplyr::filter(visit == "baseline", !is.na(group), !is.na(record_id)) %>%
  dplyr::group_by(record_id) %>%
  dplyr::summarise(
    dplyr::across(dplyr::everything(), function(.x) {
      if (is.numeric(.x)) {
        if (all(is.na(.x))) NA_real_ else mean(.x, na.rm = TRUE)
      } else {
        .x <- as.character(.x)
        if (all(is.na(.x))) NA_character_ else dplyr::last(na.omit(.x))
      }
    }),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    group_abbr   = abbrev_group(group),
    study_decode = setNames(proc_harm_dat_recid$decode,
                            proc_harm_dat_recid$study)[study]
  )
```

```{r}
# ── 19. Data-building helpers ─────────────────────────────────────────────────

# Build pie data — cohort composition by participant counts.
#   include_hardcoded : if TRUE adds ≤N estimates from hard-coded studies
#   decode_filter     : "ALL", "DECODE", or "NOT PART OF DECODE"
#
# ≤ prefix logic: a cohort group only gets a ≤ prefix if at least one of the
# studies contributing to its count is a hard-coded (estimated) study.
make_pie_dat <- function(include_hardcoded = FALSE,
                         decode_filter     = "ALL") {

  decode_map <- setNames(proc_harm_dat_recid$decode, proc_harm_dat_recid$study)

  # Observed participants — use precomputed baseline_dedup
  bd <- if (decode_filter == "DECODE") {
    baseline_dedup %>% dplyr::filter(study_decode == "DECODE")
  } else if (decode_filter == "NOT PART OF DECODE") {
    baseline_dedup %>% dplyr::filter(study_decode == "NOT PART OF DECODE")
  } else {
    baseline_dedup
  }
  real_dat <- bd %>%
    dplyr::count(group_abbr) %>%
    dplyr::mutate(has_estimated = FALSE)  # all observed rows are real

  if (!include_hardcoded) {
    real_dat %>%
      dplyr::mutate(pct   = n / sum(n),
                    label = paste0(group_abbr, "\n(n=", n, ")")) %>%
      dplyr::select(-has_estimated)
  } else {
    # Hard-coded rows: track which groups receive an estimate contribution
    hc_rows <- dplyr::bind_rows(lapply(names(hardcoded_meta), function(s) {
      meta         <- hardcoded_meta[[s]]
      study_decode <- decode_map[s]
      if (decode_filter == "DECODE"             && study_decode != "DECODE")             return(NULL)
      if (decode_filter == "NOT PART OF DECODE" && study_decode != "NOT PART OF DECODE") return(NULL)
      data.frame(group_abbr    = names(meta$group_n),
                 n             = unname(meta$group_n),
                 has_estimated = s %in% hardcoded_studies,
                 stringsAsFactors = FALSE)
    }))

    combined <- dplyr::bind_rows(real_dat, hc_rows) %>%
      dplyr::group_by(group_abbr) %>%
      dplyr::summarise(n             = sum(n),
                       has_estimated = any(has_estimated),
                       .groups = "drop") %>%
      dplyr::mutate(pct   = n / sum(n),
                    label = ifelse(has_estimated,
                                   paste0(group_abbr, "\n(≤", n, ")"),
                                   paste0(group_abbr, "\n(n=", n, ")")))
    combined
  }
}

# Build bar data — per-procedure N by cohort group.
#   include_hardcoded : if TRUE adds ≤N estimates from hard-coded studies
#   decode_filter     : "ALL", "DECODE", or "NOT PART OF DECODE"
#
# Returns a data frame with columns: procedure (ordered factor), group_abbr, n,
# has_estimated (TRUE if any hard-coded study contributed to that cell).
make_bar_dat <- function(include_hardcoded = FALSE,
                         decode_filter     = "ALL") {

  decode_map <- setNames(proc_harm_dat_recid$decode, proc_harm_dat_recid$study)

  # Pre-filter baseline_dedup by decode once (instead of inside each iteration)
  bd <- if (decode_filter == "DECODE") {
    baseline_dedup %>% dplyr::filter(study_decode == "DECODE")
  } else if (decode_filter == "NOT PART OF DECODE") {
    baseline_dedup %>% dplyr::filter(study_decode == "NOT PART OF DECODE")
  } else {
    baseline_dedup
  }

  # Observed rows — use precomputed baseline_dedup
  real_rows <- dplyr::bind_rows(lapply(names(proc_varlist_plot), function(proc_name) {
    vlist <- intersect(proc_varlist_plot[[proc_name]], cols)
    if (length(vlist) == 0) return(NULL)

    bd %>%
      dplyr::filter(rowSums(!is.na(dplyr::select(., dplyr::all_of(vlist)))) > 0) %>%
      dplyr::count(group_abbr) %>%
      dplyr::mutate(procedure     = proc_name,
                    has_estimated = FALSE)
  }))

  if (!include_hardcoded) {
    real_rows %>%
      dplyr::mutate(procedure = factor(procedure, levels = proc_order))
  } else {
    # Hard-coded rows
    hc_rows <- dplyr::bind_rows(lapply(names(hardcoded_meta), function(s) {
      meta         <- hardcoded_meta[[s]]
      study_decode <- decode_map[s]
      if (decode_filter == "DECODE"             && study_decode != "DECODE")             return(NULL)
      if (decode_filter == "NOT PART OF DECODE" && study_decode != "NOT PART OF DECODE") return(NULL)

      study_flags <- proc_harm_dat_recid[proc_harm_dat_recid$study == s, ]
      if (nrow(study_flags) == 0) return(NULL)

      dplyr::bind_rows(lapply(names(proc_flag_to_label), function(flag) {
        proc_label <- proc_flag_to_label[[flag]]
        if (!isTRUE(study_flags[[flag]])) return(NULL)
        data.frame(group_abbr    = names(meta$group_n),
                   n             = unname(meta$group_n),
                   procedure     = proc_label,
                   has_estimated = s %in% hardcoded_studies,
                   stringsAsFactors = FALSE)
      }))
    }))

    dplyr::bind_rows(real_rows, hc_rows) %>%
      dplyr::group_by(procedure, group_abbr) %>%
      dplyr::summarise(n             = sum(n),
                       has_estimated = any(has_estimated),
                       .groups = "drop") %>%
      dplyr::mutate(procedure = factor(procedure, levels = proc_order))
  }
}
```

```{r}
# ── 20a. Compiled procedure N table helper ────────────────────────────────────
# Pivots bar_dat wide: rows = procedures, columns = cohort groups, cells = n / ≤n
make_proc_summary_tbl <- function(include_hardcoded = TRUE,
                                  decode_filter = c("ALL", "DECODE", "NOT PART OF DECODE")) {
  decode_filter <- match.arg(decode_filter)
  bar_dat <- make_bar_dat(include_hardcoded = include_hardcoded,
                          decode_filter     = decode_filter)
  bar_dat %>%
    dplyr::mutate(n_label = ifelse(has_estimated, paste0("≤", n), as.character(n))) %>%
    dplyr::select(procedure, group_abbr, n_label) %>%
    tidyr::pivot_wider(names_from = group_abbr, values_from = n_label,
                       values_fill = "—") %>%
    dplyr::arrange(factor(procedure, levels = proc_order)) %>%
    dplyr::rename(Procedure = procedure)
}
```

```{r}
# ── 20. Plot helper functions ─────────────────────────────────────────────────

# Output path for saved PNGs
fig_dir <- file.path(root_path, "Data Harmonization/Report_files/Summary Report/figures")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)

# Utility: save a ggplot as PNG then return it (so it also renders inline)
save_fig <- function(p, filename, width = 10, height = 5) {
  ggsave(file.path(fig_dir, filename), plot = p,
         width = width, height = height, dpi = 300, bg = "transparent")
  p
}

# ── Pie chart ─────────────────────────────────────────────────────────────────
# No title; slice labels carry group name + n.
make_pie_plot <- function(pie_dat, filename = NULL) {
  p <- ggplot(pie_dat, aes(x = "", y = pct, fill = group_abbr)) +
    geom_col(width = 1, color = "white") +
    coord_polar(theta = "y") +
    scale_fill_manual(values = cohort_fill, name = "Cohort") +
    geom_text(aes(label = label),
              position = position_stack(vjust = 0.5),
              size = 3.5, color = "white", fontface = "bold") +
    labs(x = NULL, y = NULL) +
    theme_void(base_size = 12) +
    theme(legend.position = "none")
  if (!is.null(filename)) save_fig(p, filename, width = 6, height = 6) else p
}

# ── Bar chart ─────────────────────────────────────────────────────────────────
# No title; no x-axis label. ≤ annotation on bars that include estimated data.
make_bar_plot <- function(bar_dat, filename = NULL) {
  p <- ggplot(bar_dat, aes(x = procedure, y = n, fill = group_abbr)) +
    geom_col(position = "dodge", width = 0.7) +
    scale_fill_manual(values = cohort_fill, name = "Cohort") +
    labs(x = NULL, y = "N") +
    theme_harm +
    theme(text = element_text(color = "white"),
          axis.text = element_text(color = "white"),
          axis.title = element_text(color = "white"))
  if (!is.null(filename)) save_fig(p, filename) else p
}

# ── Box plot ──────────────────────────────────────────────────────────────────
# Baseline observed data only (hard-coded studies have no measured values).
# Reuses precomputed baseline_dedup; just coerce group_abbr to ordered factor.
bp_base <- baseline_dedup %>%
  dplyr::mutate(group_abbr = factor(group_abbr,
                                    levels = c("HC", "OB", "T1D", "T2D", "ADPKD")))

make_boxplot <- function(data, var, ylabel, filename = NULL, log_transform = FALSE) {
  d <- data %>% dplyr::filter(!is.na(.data[[var]]))
  p <- ggplot(d, aes(x = group_abbr, y = !!sym(var), fill = group_abbr)) +
    { if (log_transform) scale_y_log10() } +
    geom_boxplot(outlier.shape = 21, outlier.size = 1.5, alpha = 0.85) +
    geom_jitter(width = 0.15, size = 0.8, alpha = 0.35, color = "grey30") +
    scale_fill_manual(values = cohort_fill, guide = "none") +
    labs(x = NULL, y = ylabel) +
    theme_harm +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
  if (!is.null(filename)) save_fig(p, filename, width = 6, height = 5) else p
}

# make_proc_pie_plot: one pie per procedure, each saved as its own PNG.
# bar_dat    → output of make_bar_dat()
# file_prefix → string prepended to each file, e.g. "pie_proc_full_all"
#               produces "pie_proc_full_all_Blood_Labs.png" etc.
# Returns a named list of ggplot objects (one per procedure) invisibly,
# and prints them inline via patchwork::wrap_plots for the rendered doc.
make_proc_pie_plot <- function(bar_dat, file_prefix = NULL) {
  procs <- intersect(proc_order, unique(bar_dat$procedure))

  plots <- lapply(procs, function(proc) {
    d <- bar_dat %>%
      dplyr::filter(procedure == proc) %>%
      dplyr::mutate(
        pct   = n / sum(n),
        label = ifelse(has_estimated,
                       paste0(group_abbr, "\n(≤", n, ")"),
                       paste0(group_abbr, "\n(n=",  n, ")"))
      )

    p <- ggplot(d, aes(x = "", y = pct, fill = group_abbr)) +
      geom_col(width = 1, color = "white") +
      coord_polar(theta = "y") +
      scale_fill_manual(values = cohort_fill, name = "Cohort") +
      geom_text(aes(label = label),
                position = position_stack(vjust = 0.5),
                size = 3, color = "white", fontface = "bold") +
      labs(x = NULL, y = NULL,
           title = proc) +
      theme_void(base_size = 11) +
      theme(
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 18,
                                  color = "white",
                                  hjust = 0.5, vjust = -4)
      )

    if (!is.null(file_prefix)) {
      safe_name <- gsub("[^A-Za-z0-9]+", "_", proc)
      save_fig(p, paste0(file_prefix, "_", safe_name, ".png"),
               width = 5, height = 5)
    }
    p
  })

  names(plots) <- procs
  # Display as a wrapped grid in the rendered document
  patchwork::wrap_plots(plots, ncol = 4) +
    patchwork::plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
}
```

---

## Data-only (observed data in dataset)
## Data-only (observed data in dataset)

> Counts participants with ≥1 non-missing value in each procedure's variable set,
> at baseline. Hard-coded / still-enrolling studies are **excluded**.

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(include_hardcoded = FALSE, decode_filter = "ALL"),
              filename = "pie_data_only_all.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(include_hardcoded = FALSE, decode_filter = "ALL"),
              filename = "bar_data_only_all.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = FALSE, decode_filter = "ALL") %>%
  knitr::kable(caption = "N per procedure × cohort — data only")
```

### Per-procedure cohort breakdown (pie charts)

```{r fig.width=14, fig.height=10}
make_proc_pie_plot(make_bar_dat(FALSE, "ALL"), "pie_proc_data_only_all")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = FALSE, decode_filter = "ALL") %>%
  knitr::kable(caption = "N per procedure x cohort --- data only")
```

### Per-procedure cohort breakdown (pie charts)

```{r fig.width=14, fig.height=10}
make_proc_pie_plot(make_bar_dat(FALSE, "ALL"), "pie_proc_data_only_all")
```

### Key outcome box plots

```{r}
# log(UACR)
make_boxplot(bp_base, "acr_u", "UACR (mg/g) [log scale]",
             filename = "box_data_only_uacr.png", log_transform = TRUE)
```

```{r}
# Clamp M-value
make_boxplot(bp_base, "p2_raw_m", "M-value (mg/kg/min)",
             filename = "box_data_only_mvalue.png")
```

```{r}
# GFR (BSA-corrected plasma)
make_boxplot(bp_base, "gfr_bsa_plasma", "GFR (mL/min/1.73 m²)",
             filename = "box_data_only_gfr.png")
```

```{r}
# Avg cortical R2*
make_boxplot(bp_base, "avg_c_r2", "Avg Cortical R2* (s⁻¹)",
             filename = "box_data_only_r2star.png")
```

---

## Full (observed + hard-coded estimates, all studies)

> Combines observed N from the dataset with ≤N estimates for hard-coded /
> still-enrolling studies. ≤ prefix appears only for cohort groups that include
> at least one hard-coded study contribution.

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(include_hardcoded = TRUE, decode_filter = "ALL"),
              filename = "pie_full_all.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(include_hardcoded = TRUE, decode_filter = "ALL"),
              filename = "bar_full_all.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = TRUE, decode_filter = "ALL") %>%
  knitr::kable(caption = "N per procedure × cohort — full (observed + ≤ estimated)")
```

### Per-procedure cohort breakdown (pie charts)

```{r fig.width=14, fig.height=10}
make_proc_pie_plot(make_bar_dat(TRUE, "ALL"), "pie_proc_full_all")
```

---

## DECODE studies only (observed + hard-coded estimates)

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(include_hardcoded = TRUE, decode_filter = "DECODE"),
              filename = "pie_full_decode.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(include_hardcoded = TRUE, decode_filter = "DECODE"),
              filename = "bar_full_decode.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = TRUE, decode_filter = "DECODE") %>%
  knitr::kable(caption = "N per procedure × cohort — DECODE studies (observed + ≤ estimated)")
```

### Per-procedure cohort breakdown (pie charts)

```{r fig.width=14, fig.height=10}
make_proc_pie_plot(make_bar_dat(TRUE, "DECODE"), "pie_proc_full_decode")
```

---

## Non-DECODE studies only (observed + hard-coded estimates)

### Cohort composition

```{r}
make_pie_plot(make_pie_dat(include_hardcoded = TRUE, decode_filter = "NOT PART OF DECODE"),
              filename = "pie_full_nondecode.png")
```

### N per procedure × cohort

```{r}
make_bar_plot(make_bar_dat(include_hardcoded = TRUE, decode_filter = "NOT PART OF DECODE"),
              filename = "bar_full_nondecode.png")
```

### Compiled procedure N table

```{r}
make_proc_summary_tbl(include_hardcoded = TRUE, decode_filter = "NOT PART OF DECODE") %>%
  knitr::kable(caption = "N per procedure × cohort — non-DECODE studies (observed + ≤ estimated)")
```

### Per-procedure cohort breakdown (pie charts)

```{r fig.width=14, fig.height=10}
make_proc_pie_plot(make_bar_dat(TRUE, "NOT PART OF DECODE"), "pie_proc_full_nondecode")
```

---

## Box plots — key outcomes by cohort

> Box plots reflect observed data only (hard-coded studies have no measured values).

```{r}
# log(UACR)
make_boxplot(bp_base, "acr_u", "UACR (mg/g) [log scale]",
             filename = "box_uacr.png", log_transform = TRUE)
```

```{r}
# Clamp M-value
make_boxplot(bp_base, "p2_raw_m", "M-value (mg/kg/min)",
             filename = "box_mvalue.png")
```

```{r}
# GFR (BSA-corrected plasma)
make_boxplot(bp_base, "gfr_bsa_plasma", "GFR (mL/min/1.73 m²)",
             filename = "box_gfr.png")
```

```{r}
# Avg cortical R2*
make_boxplot(bp_base, "avg_c_r2", "Avg Cortical R2* (s⁻¹)",
             filename = "box_r2star.png")
```