---
title: "Plasma_DEP_anlaysis"
author: "Savanah Leidholt"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ---- packages ----
library(dplyr)
library(tidyr)
library(limma)
library(EnhancedVolcano)
library(ggplot2)
library(patchwork)
library(ggrepel)
library(vegan)
library(tidyverse)
library(data.table)
library(purrr)
library(haven)
library(ggVennDiagram)
library(caret)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "sleidholt") {
    dir.dat <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Savanah Leidholt/"
} else if (user == "savanahleidholt") {
    dir.dat <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Savanah Leidholt/"
} else if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
    
  } 
```
#-----------------------------------------------------------------------------
#Clean and setup data
# a) all proteomic data DPE analysis
# b) sensitivty analysis where I remove participants who have T1D and are obese
# c) sensitivty analysis (rm attempt)
# d) DPE analysis with samples that contain both proteomic and metabolomic data
#------------------------------------------------------------------------------
```{r}
obj <- readRDS(file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "omics_split_objects_clean.rds"))

meta   <- obj$meta_prot
prot_plasma <- obj$prot_plasma
n_table     <- obj$n_table

# quick check
n_table
dim(prot_plasma)
dim(meta)


```
##---------------------------------------------
##open panther_data.RData to add in hba1c data
##---------------------------------------------
```{r}
load(file.path(dir.dat, "PANTHER", "Data_cleaned", "panther_data.RData"))

#trasnforming data to match the harmonized data
panther_hba1c <- panther %>%
  dplyr::select(record_id, hba1c) %>%
  transmute(
    record_id = gsub("_", "-", as.character(record_id)),
    hba1c_panther = as.numeric(hba1c)
  )

panther_hba1c$record_id <- as.character(panther_hba1c$record_id)

#Make sure harm$hba1c is numeric too (if it exists)
meta <- meta %>%
  mutate(hba1c = as.numeric(zap_labels(hba1c))) %>%
  left_join(panther_hba1c, by = "record_id") %>%
  mutate(
    hba1c = if_else(
      study == "PANTHER" & is.na(hba1c),
      hba1c_panther,
      hba1c
    )
  ) %>%
  dplyr::select(-hba1c_panther)
```
##---------------------
##cleaning metadata
##---------------------
```{r}
##--------------------------------------------------------------
## Clean metadata and match to plasma proteomics
##--------------------------------------------------------------

meta2 <- meta %>%
  mutate(
    record_id = trimws(as.character(record_id)),
    study     = trimws(as.character(study)),
    group     = trimws(as.character(group)),
    sex       = trimws(as.character(sex)),
    ethnicity = trimws(as.character(race_ethnicity_condensed)),
    age       = suppressWarnings(as.numeric(age)),
    uacr      = suppressWarnings(as.numeric(acr_u)),
    log_uacr  = log(uacr + 1)
  ) %>%
  filter(
    !is.na(age),
    !is.na(sex),
    sex != ""
  ) %>%
  mutate(
    group4 = factor(
      group,
      levels = c("Lean Control", "Obese Control", "Type 1 Diabetes", "Type 2 Diabetes"),
      labels = c("LC", "OC", "T1D", "T2D")
    ),
    ethnicity = factor(ethnicity),
    study6 = dplyr::recode(
      study,
      "ATTEMPT"          = "ATT",
      "CROCODILE"        = "CRC",
      "PANTHER"          = "PAN",
      "RENAL-HEIR"       = "RH",
      "IMPROVE"          = "IT",
      "RENAL-HEIRitage"  = "RH2",
      .default = study
    ),
    study6 = factor(study6)
  )

#removing rows with no proteomics data 
rownames(prot_plasma) <- trimws(rownames(prot_plasma))

has_prot <- rowSums(!is.na(prot_plasma)) > 0
prot_plasma <- prot_plasma[has_prot, , drop = FALSE]

meta2 <- meta2 %>%
  filter(record_id %in% rownames(prot_plasma))

olap <- intersect(meta2$record_id, rownames(prot_plasma))
cat("meta n:", nrow(meta), "\n")
cat("meta2 n:", nrow(meta2), "\n")

cat("prot_plasma dim:", dim(prot_plasma), "\n")
cat("overlap n:", length(olap), "\n")
cat("meta not in prot:", sum(!meta2$record_id %in% rownames(prot_plasma)), "\n")
cat("prot not in meta:", sum(!rownames(prot_plasma) %in% meta2$record_id), "\n")

stopifnot(length(olap) > 0)

meta3 <- meta2[match(olap, meta2$record_id), , drop = FALSE]
prot2 <- prot_plasma[olap, , drop = FALSE]

stopifnot(identical(rownames(prot2), meta3$record_id))

## Check near-zero variance proteins
nzv_idx <- caret::nearZeroVar(prot2) #zero
```

##--------------------------------------------------------------
## Protein Filter and Log Transformation
##filter to proteins that are only in the analyte panel (~7000)
##---------------------------------------------------------------
```{r}
analytes <- read.csv(
  file.path(dir.dat, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

prot_panel <- unique(trimws(as.character(analytes$AptName)))
aptamers_keep <- intersect(prot_panel, colnames(prot2))
prot3 <- prot2[, aptamers_keep, drop = FALSE]

dim(prot3)   # should be ~7000
dim(prot2)   # should be ~10000

length(aptamers_keep)
length(setdiff(prot_panel, colnames(prot2)))

## Log transform protein data
prot_log <- log2(as.matrix(prot3) + 1)
prot_log[!is.finite(prot_log)] <- NA_real_

## Check near-zero variance after filtering
nzv_idx <- caret::nearZeroVar(prot3) #zero

```

```{r}
#not really necessary to do as all protiens pass filter once I remove the extra 3000 from the ATTEMPT Panel
##filtering step where protein must exist in 60% of samples in a study

#study_idx_all <- split(seq_len(nrow(prot_log)), meta_common$study)
#present_by_study_all <- sapply(study_idx_all, function(idx) {
#  colMeans(!is.na(prot_log[idx, ])) >= 0.6
#})

#present_by_study_all <- as.matrix(present_by_study_all)
#n_studies_present_all <- colSums(present_by_study_all)
#keep_study_all <- n_studies_present_all >= 2

#removed_proteins <- colnames(prot_log)[!keep_study_all]
#kept_proteins    <- colnames(prot_log)[keep_study_all]

#cat("All proteomic features before study filter:", ncol(prot_log), "\n")
#cat("Removed (<2 studies meeting 60% presence threshold):", length(removed_proteins), "\n")
#cat("Kept (>=2 studies meeting 60% presence threshold):", length(kept_proteins), "\n")
```

##--------------------------------
## Find and Remove Outliers in PCA
##---------------------------------
```{r}
pca0 <- prcomp(prot_log, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2])
scores$record_id <- rownames(scores)
scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

threshold <- mean(scores$dist) + 3 * sd(scores$dist)
outliers <- scores$record_id[scores$dist > threshold]
outliers #"10424"   "RH-84-T"

meta4 <- meta3 %>%
  filter(!record_id %in% outliers)

prot_log2 <- prot_log[
  !rownames(prot_log) %in% outliers,
  ,
  drop = FALSE
]

cat("Still present in metadata? ", any(outliers %in% meta4$record_id), "\n")
cat("Still present in matrix? ", any(outliers %in% rownames(prot_log2)), "\n")

stopifnot(!any(outliers %in% meta4$record_id))
stopifnot(!any(outliers %in% rownames(prot_log2)))
stopifnot(identical(rownames(prot_log2), meta4$record_id))

```

#---------------
# Plotting Funciton
#---------------
```{r}
make_pca_plot <- function(prot_mat, meta_df, title_str) {
  pca  <- prcomp(prot_mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1    = pca$x[, 1],
    PC2    = pca$x[, 2],
    group4 = meta_df$group4,
    study  = meta_df$study6
  ) %>%
    ggplot(aes(PC1, PC2, color = group4, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        LC  = "Cyan4",
        OC  = "Gold",
        T1D = "DarkRed",
        T2D = "Purple"
      ),
      name = "Group"
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
##-----------
## PCAs
##-----------
```{r}
design <- model.matrix(~ group4, data = meta4)

## no adjustment
p0 <- make_pca_plot(prot_log2, meta4, "No adjustment")

## age + sex adjustment
cov1 <- model.matrix(~ age + sex, data = meta4)[, -1, drop = FALSE]
prot_adj1 <- t(limma::removeBatchEffect(t(prot_log2), covariates = cov1, design = design))
p1 <- make_pca_plot(prot_adj1, meta4, "Age + Sex")

## age + sex + study adjustment
cov2 <- model.matrix(~ age + sex + study6, data = meta4)[, -1, drop = FALSE]
prot_adj2 <- t(limma::removeBatchEffect(t(prot_log2), covariates = cov2, design = design))
p2 <- make_pca_plot(prot_adj2, meta4, "Age + Sex + Study")

## combine plots
pca_combined <- p0 | p1 | p2
pca_combined

ggsave(filename = file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Results", "pca_plots.pdf"),
       plot = pca_combined,
       width = 16,
       height = 14)
```

##------------------------------
##limma testing all prots data
## model1: age + sex
## model2: age + sex + study
##------------------------------
```{r}
# helper function-- run limma + return list of topTables
run_limma_contrasts <- function(Y, design, contr) {
  fit <- lmFit(Y, design)
  fit2 <- contrasts.fit(fit, contr)
  fit2 <- eBayes(fit2)

  out <- setNames(vector("list", ncol(contr)), colnames(contr))
  for (cc in colnames(contr)) {
    tt <- topTable(fit2, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
    tt$feature_id <- rownames(tt)
    tt$contrast <- cc
    out[[cc]] <- tt
  }
  list(fit = fit2, out = out)
}

# ---- helper: summarise sig hits per contrast ----
summarise_sig <- function(out_list, model_label, n_samp) {
  tibble(
    contrast = names(out_list),
    n_FDR05  = sapply(out_list, \(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
    n_FDR01  = sapply(out_list, \(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
    model    = model_label,
    n_samples = n_samp
  )
}

## response matrices: proteins x samples
Y0 <- t(prot_log)
Y1 <- t(prot_log2)

## check sample alignment
stopifnot(identical(colnames(Y0), meta3$record_id))
stopifnot(identical(colnames(Y1), meta4$record_id))
```

##--------------------------------------------------------------
## Model 1: age + sex, before outlier removal
##--------------------------------------------------------------
```{r}
meta_m0 <- meta3
Y_m0 <- Y0[, meta_m0$record_id, drop = FALSE]

design_m0 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m0)
colnames(design_m0) <- sub("^group4", "", colnames(design_m0))
colnames(design_m0) <- make.names(colnames(design_m0))

contr_m0 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, 
                                 levels = design_m0)

res_m0 <- run_limma_contrasts(Y_m0, design_m0, contr_m0)
fit_m0 <- res_m0$fit
out_m0 <- res_m0$out

sig_m0 <- summarise_sig(out_m0, "age + sex, before outlier removal", nrow(meta_m0))

sig_m0
```
##--------------------------------------------------------------
## Model 2: age + sex, after outlier removal
##--------------------------------------------------------------
```{r}
meta_m1 <- meta4
Y_m1 <- Y1[, meta_m1$record_id, drop = FALSE]

design_m1 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m1)
colnames(design_m1) <- sub("^group4", "", colnames(design_m1))
colnames(design_m1) <- make.names(colnames(design_m1))

contr_m1 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, 
                                 levels = design_m1)

res_m1 <- run_limma_contrasts(Y_m1, design_m1, contr_m1)
fit_m1 <- res_m1$fit
out_m1 <- res_m1$out

sig_m1 <- summarise_sig(out_m1, "age + sex, after outlier removal", nrow(meta_m1))

sig_m1
```
##--------------------------------------------------------------
## Model 3: age + sex + study, before outlier removal
##--------------------------------------------------------------
```{r}
meta_m2 <- meta3 %>%
  filter(!is.na(study6))

Y_m2 <- Y0[, meta_m2$record_id, drop = FALSE]

design_m2 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m2)
colnames(design_m2) <- sub("^group4", "", colnames(design_m2))
colnames(design_m2) <- make.names(colnames(design_m2))

contr_m2 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, levels = design_m2)

res_m2 <- run_limma_contrasts(Y_m2, design_m2, contr_m2)
fit_m2 <- res_m2$fit
out_m2 <- res_m2$out

sig_m2 <- summarise_sig(out_m2, "age + sex + study, before outlier removal", nrow(meta_m2))

sig_m2
```
##--------------------------------------------------------------
## Model 4: age + sex + study, after outlier removal
##--------------------------------------------------------------
```{r}
meta_m3 <- meta4 %>%
  filter(!is.na(study6))

Y_m3 <- Y1[, meta_m3$record_id, drop = FALSE]

design_m3 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m3)
colnames(design_m3) <- sub("^group4", "", colnames(design_m3))
colnames(design_m3) <- make.names(colnames(design_m3))

contr_m3 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, levels = design_m3)

res_m3 <- run_limma_contrasts(Y_m3, design_m3, contr_m3)
fit_m3 <- res_m3$fit
out_m3 <- res_m3$out

sig_m3 <- summarise_sig(out_m3, "age + sex + study, after outlier removal", nrow(meta_m3))

sig_m3
```
```{r}
sig_overlap_m1 <- bind_rows(
  lapply(names(out_m0), function(cc) {

    pre_ids <- out_m0[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    post_ids <- out_m1[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_pre <- length(pre_ids)
    n_overlap <- length(intersect(pre_ids, post_ids))

    tibble(
      model = "age + sex",
      contrast = cc,
      n_pre = n_pre,
      n_post = length(post_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_pre, 1)
    )
  })
)

sig_overlap_m2 <- bind_rows(
  lapply(names(out_m2), function(cc) {

    pre_ids <- out_m2[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    post_ids <- out_m3[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_pre <- length(pre_ids)
    n_overlap <- length(intersect(pre_ids, post_ids))

    tibble(
      model = "age + sex + study",
      contrast = cc,
      n_pre = n_pre,
      n_post = length(post_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_pre, 1)
    )
  })
)

sig_overlap_all <- bind_rows(sig_overlap_m1, sig_overlap_m2)

sig_overlap_all
```
#-------------------------------------
# Sensitivity analysis:
# Remove T1D participants who are obese
#--------------------------------------
```{r}
#obesity status--check if this is ok. patients range in age from 13-25 so I may need to do percentiles for youth
meta4 <- meta4 %>%
  dplyr::mutate(
    bmi = suppressWarnings(as.numeric(bmi)),
    bmip = suppressWarnings(as.numeric(bmip)),
    bmi_obesity = dplyr::case_when(
      age > 19 & bmi < 25 ~ "Normal",
      age > 19 & bmi >= 25 & bmi < 30 ~ "Overweight",
      age > 19 & bmi >= 30 ~ "Obese",
      age <= 19 & bmip < 85 ~ "Normal",
      age <= 19 & bmip >= 85 & bmip < 95 ~ "Overweight",
      age <= 19 & bmip >= 95 ~ "Obese",
      TRUE ~ NA_character_
    ),
    bmi_obesity = factor(bmi_obesity, levels = c("Normal", "Overweight", "Obese"))
  )

table(meta4$bmi_obesity, useNA = "ifany")
meta4 %>%
  count(group4, bmi_obesity) %>%
  tidyr::pivot_wider(
    names_from = bmi_obesity,
    values_from = n,
    values_fill = 0
  )
```

```{r}
meta5 <- meta4 %>%
  mutate(remove_t1d_obese = group4 == "T1D" & bmi_obesity == "Obese") %>%
  filter(!remove_t1d_obese) %>%
  droplevels()

prot_log3 <- prot_log2[meta5$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_log3), meta5$record_id))

cat("Primary analysis N:", nrow(meta4), "\n")
cat("Sensitivity analysis N:", nrow(meta5), "\n")
cat("Removed obese T1D N:", nrow(meta4) - nrow(meta5), "\n")

meta5 %>%
  count(group4, bmi_obesity)

meta5 %>%
  count(group4, study6) %>%
  arrange(group4, study6)

table(meta5$group4)
```

# -----------------------------
# PCA after removing obese T1D
# -----------------------------
```{r}
design_sens <- model.matrix(~ group4, data = meta5)

p_sens0 <- make_pca_plot(prot_log3, meta5, "Sensitivity: No adjustment")

cov_sens1 <- model.matrix(~ age + sex, data = meta5)[, -1, drop = FALSE]
prot_sens1 <- t(limma::removeBatchEffect(t(prot_log3), covariates = cov_sens1, design = design_sens))
p_sens1 <- make_pca_plot(prot_sens1, meta5, "Sensitivity: Age + Sex")

cov_sens2 <- model.matrix(~ age + sex + study6, data = meta5)[, -1, drop = FALSE]
prot_sens2 <- t(limma::removeBatchEffect(t(prot_log3), covariates = cov_sens2, design = design_sens))
p_sens2 <- make_pca_plot(prot_sens2, meta5, "Sensitivity: Age + Sex + Study")

pca_sens_combined <- p_sens0 | p_sens1 | p_sens2

pca_sens_combined

ggsave(filename = file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Results", "pca_plots_sensitivity_no_obese_T1D.pdf"),
       plot = pca_sens_combined,
       width = 16,
       height = 14)
```

# -----------------------------
#limma adiposity sensitivity data
# -----------------------------
```{r}
Y_sens <- t(prot_log3)
stopifnot(identical(colnames(Y_sens), meta5$record_id))
```

##---------------------------------
## adiposity model 1: age + sex
##----------------------------------
```{r}
meta_m4 <- meta5
Y_m4 <- Y_sens[, meta_m4$record_id, drop = FALSE]

design_m4 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m4)
colnames(design_m4) <- sub("^group4", "", colnames(design_m4))
colnames(design_m4) <- make.names(colnames(design_m4))

contr_m4 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m4)

res_m4 <- run_limma_contrasts(Y_m4, design_m4, contr_m4)
fit_m4 <- res_m4$fit
out_m4 <- res_m4$out

sig_m4 <- summarise_sig(out_m4, "sensitivity: age + sex, no obese T1D", nrow(meta_m4))

sig_m4


```
##----------------------------------------
## adiposity model 2: age + sex + study
##-----------------------------------------
```{r}
meta_m5 <- meta5 %>%
  filter(!is.na(study6)) %>%
  droplevels()

Y_m5 <- Y_sens[, meta_m5$record_id, drop = FALSE]

design_m5 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m5)
colnames(design_m5) <- sub("^group4", "", colnames(design_m5))
colnames(design_m5) <- make.names(colnames(design_m5))

contr_m5 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m5)

res_m5 <- run_limma_contrasts(Y_m5, design_m5, contr_m5)
fit_m5 <- res_m5$fit
out_m5 <- res_m5$out

sig_m5 <- summarise_sig(out_m5, "sensitivity: age + sex + study, no obese T1D", nrow(meta_m5))

sig_m5
```
##-------------------------------------------------
## Compare primary and adipoisty sensitivity limma summaries
##-------------------------------------------------
```{r}
summary_sensitivity <- bind_rows(
  sig_m1 %>% mutate(model = "primary: age + sex"),
  sig_m3 %>% mutate(model = "primary: age + sex + study"),
  sig_m4,
  sig_m5
) %>%
  select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

summary_sensitivity

summary_sensitivity_wide <- summary_sensitivity %>%
  select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

summary_sensitivity_wide


sig_overlap_study <- bind_rows(
  lapply(names(out_m3), function(cc) {

    primary_ids <- out_m3[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    sens_ids <- out_m5[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_primary <- length(primary_ids)
    n_overlap <- length(intersect(primary_ids, sens_ids))

    tibble(
      contrast = cc,
      n_primary = n_primary,
      n_sensitivity = length(sens_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_primary, 1)
    )
  })
)

sig_overlap_study
```
#--------------------------------------------------------
# SENSITIVITY ANALYSIS:
# Exclude ATTEMPT study (different sequencing platform)
#--------------------------------------------------------
```{r}
meta6 <- meta4 %>%
  filter(study != "ATTEMPT", study6 != "ATT") %>%
  droplevels()

prot_log4 <- prot_log2[meta6$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_log4), meta6$record_id))

cat("Primary analysis N:", nrow(meta4), "\n")
cat("No-ATTEMPT sensitivity N:", nrow(meta6), "\n")
cat("Removed ATTEMPT N:", nrow(meta4) - nrow(meta6), "\n")

meta6 %>%
  count(group4, study6) %>%
  arrange(group4, study6)

table(meta6$group4)
```
##---------------------------
##outlier check and removal
##---------------------------
```{r}
pca_noATT <- prcomp(prot_log4, center = TRUE, scale. = TRUE)

scores_noATT <- as.data.frame(pca_noATT$x[, 1:2])
scores_noATT$record_id <- rownames(scores_noATT)
scores_noATT$dist <- sqrt(scores_noATT$PC1^2 + scores_noATT$PC2^2)

threshold_noATT <- mean(scores_noATT$dist) + 3 * sd(scores_noATT$dist)

outliers_noATT <- scores_noATT$record_id[scores_noATT$dist > threshold_noATT]

outliers_noATT

#meta7 <- meta6 %>%
 # filter(!record_id %in% outliers_noATT) %>%
  #droplevels()

#prot_log5 <- prot_log4[!rownames(prot_log4) %in% outliers_noATT, , drop = FALSE]

#cat("Still present in metadata? ", any(outliers_noATT %in% meta7$record_id), "\n")
#cat("Still present in matrix? ", any(outliers_noATT %in% rownames(prot_log5)), "\n")

#stopifnot(!any(outliers_noATT %in% meta7$record_id))
#stopifnot(!any(outliers_noATT %in% rownames(prot_log5)))
#stopifnot(identical(rownames(prot_log5), meta7$record_id))
```
##--------------
##PCA no ATTEMPt
##--------------
```{r}
design_noATT <- model.matrix(~ group4, data = meta6)

p_noATT0 <- make_pca_plot(prot_log4, meta6, "Sensitivity: No adjustment, excluding ATTEMPT")

cov_noATT1 <- model.matrix(~ age + sex, data = meta6)[, -1, drop = FALSE]
prot_noATT1 <- t(limma::removeBatchEffect(t(prot_log4), covariates = cov_noATT1, design = design_noATT))
p_noATT1 <- make_pca_plot(prot_noATT1, meta6, "Sensitivity: Age + Sex, excluding ATTEMPT")

cov_noATT2 <- model.matrix(~ age + sex + study6, data = meta6)[, -1, drop = FALSE]
prot_noATT2 <- t(limma::removeBatchEffect(t(prot_log4), covariates = cov_noATT2, design = design_noATT))
p_noATT2 <- make_pca_plot(prot_noATT2, meta6, "Sensitivity: Age + Sex + Study, excluding ATTEMPT")

pca_noATT_combined <- (p_noATT0 | p_noATT1 | p_noATT2) +
  patchwork::plot_annotation(
    title = "PCA: Sensitivity Analysis Excluding ATTEMPT",
    subtitle = "Color = group | Shape = study | ATTEMPT removed",
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 9, hjust = 0.5, color = "grey40")
    )
  )

pca_noATT_combined

ggsave(filename = file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Results", "pca_noATTEMPT.pdf"),
       plot = pca_noATT_combined,
       width = 16,
       height = 14)
```
##----------------------------------
## limma using no-ATTEMPT dataset
## model 1: age + sex
## model 2: age + sex + study
## ---------------------------------
```{r}
Y_noATT <- t(prot_log4)
stopifnot(identical(colnames(Y_noATT), meta6$record_id))

# Model 1: age + sex
meta_m6 <- meta6
Y_m6 <- Y_noATT[, meta_m6$record_id, drop = FALSE]

design_m6 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m6)
colnames(design_m6) <- sub("^group4", "", colnames(design_m6))
colnames(design_m6) <- make.names(colnames(design_m6))

contr_m6 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m6)

res_m6 <- run_limma_contrasts(Y_m6, design_m6, contr_m6)
fit_m6 <- res_m6$fit
out_m6 <- res_m6$out

sig_m6 <- summarise_sig(out_m6, "sensitivity: age + sex, no ATTEMPT", nrow(meta_m6))

sig_m6


# Model 2: age + sex + study
meta_m7 <- meta6 %>%
  filter(!is.na(study6)) %>%
  droplevels()

Y_m7 <- Y_noATT[, meta_m7$record_id, drop = FALSE]

design_m7 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m7)
colnames(design_m7) <- sub("^group4", "", colnames(design_m7))
colnames(design_m7) <- make.names(colnames(design_m7))

contr_m7 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m7)

res_m7 <- run_limma_contrasts(Y_m7, design_m7, contr_m7)
fit_m7 <- res_m7$fit
out_m7 <- res_m7$out

sig_m7 <- summarise_sig(out_m7, "sensitivity: age + sex + study, no ATTEMPT", nrow(meta_m7))

sig_m7

```
## -----------------------------
## Summarize no-ATTEMPT results
## -----------------------------
```{r}
summary_noATT <- bind_rows(
  sig_m1 %>% mutate(model = "primary: age + sex"),
  sig_m3 %>% mutate(model = "primary: age + sex + study"),
  sig_m6,
  sig_m7
) %>%
  select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

summary_noATT

summary_noATT_wide <- summary_noATT %>%
  select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

summary_noATT_wide

sig_overlap_noATT_study <- bind_rows(
  lapply(names(out_m3), function(cc) {

    primary_ids <- out_m3[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    noATT_ids <- out_m7[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_primary <- length(primary_ids)
    n_overlap <- length(intersect(primary_ids, noATT_ids))

    tibble(
      contrast = cc,
      n_primary = n_primary,
      n_noATT = length(noATT_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_primary, 1)
    )
  })
)

sig_overlap_noATT_study
```

#------------------------------------------------------------------------
# Sensitivity Analysis #3:
# Restrict proteomics workflow to participants with all 3 omics types
#fix to have scRNA and not lipidomic data
#------------------------------------------------------------------------
```{r}
# Identify participants with all three omics types
omics_objects <- readRDS(file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "omics_split_objects3.rds"))


# HELPER: count non-NA samples in a matrix for a given metadata
count_samples_with_data <- function(meta_df, mat_df) {
  # A sample "has data" if it has at least one non-NA value in the omics matrix
  has_data <- rowSums(!is.na(mat_df)) > 0
  meta_df %>%
    mutate(has_omics = has_data) %>%
    filter(has_omics)
}
```

##---------------------------------------
## samples counts Study x Group
##---------------------------------------
```{r}

make_count_table <- function(meta_df, mat_df, omics_label) {
  count_samples_with_data(meta_df, mat_df) %>%
    count(study, group) %>%
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>%
    arrange(study) %>%
    mutate(Omics = omics_label, .before = 1)
}

tbl_metabolites <- make_count_table(omics_objects$meta_metab, omics_objects$metabolites,  "Metabolomics")
tbl_lipids      <- make_count_table(omics_objects$meta_lips,  omics_objects$lipids,        "Lipidomics")
tbl_proteins    <- make_count_table(omics_objects$meta_prot,  omics_objects$prot_plasma,   "Proteomics (Plasma)")

cat("\n===== METABOLOMICS: Samples per Study x Group =====\n")
print(tbl_metabolites)

cat("\n===== LIPIDOMICS: Samples per Study x Group =====\n")
print(tbl_lipids)

cat("\n===== PROTEOMICS (Plasma): Samples per Study x Group =====\n")
print(tbl_proteins)
```

```{r}
#A sample "has data" if it has at least one non-missing value in that matrix
ids_metab <- rownames(omics_objects$metabolites)[rowSums(!is.na(omics_objects$metabolites)) > 0]
ids_lipids <- rownames(omics_objects$lipids)[rowSums(!is.na(omics_objects$lipids)) > 0]
ids_prot <- rownames(omics_objects$prot_plasma)[rowSums(!is.na(omics_objects$prot_plasma)) > 0]

ids_metab <- trimws(as.character(ids_metab))
ids_lipids <- trimws(as.character(ids_lipids))
ids_prot <- trimws(as.character(ids_prot))

ids_all_three <- Reduce(intersect, list(ids_metab, ids_lipids, ids_prot))

cat("Total participants with all 3 omics:", length(ids_all_three), "\n")
```

##--------------------------------------------------
## Restrict the proteomics analysis-ready dataset
##--------------------------------------------------
```{r}
meta_all3 <- meta_common %>%
  mutate(record_id = trimws(as.character(record_id))) %>%
  filter(record_id %in% ids_all_three)

prot_all3 <- prot_plasma_common[meta_all3$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_all3), meta_all3$record_id))

cat("Original proteomics analysis N:", nrow(meta_common), "\n")
cat("Proteomics analysis N with all 3 omics:", nrow(meta_all3), "\n")
cat("Excluded because not all 3 omics:",
    nrow(meta_common) - nrow(meta_all3), "\n")

cat("Group counts in all-3-omics subset:\n")
print(table(meta_all3$group4, useNA = "ifany"))

cat("Study counts in all-3-omics subset:\n")
print(table(meta_all3$study, useNA = "ifany"))
```
##--------------------------------------------------
##data cleaning for data subset with all 3 Omics
##--------------------------------------------------
```{r}
meta_pca_all3 <- meta_all3
prot_pca_all3 <- prot_all3[meta_pca_all3$record_id, , drop = FALSE]
stopifnot(identical(rownames(prot_pca_all3), meta_pca_all3$record_id))

prot_pca_all3 <- as.matrix(prot_pca_all3)

# ---- log2 transform + median impute ----
prot_log_all3 <- log2(prot_pca_all3 + 1)
prot_log_all3[!is.finite(prot_log_all3)] <- NA_real_

prot_imp_all3 <- prot_log_all3
meds_all3 <- apply(prot_imp_all3, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(prot_imp_all3))) {
  if (!is.na(meds_all3[j])) prot_imp_all3[is.na(prot_imp_all3[, j]), j] <- meds_all3[j]
}

keep_cols_all3 <- colSums(is.na(prot_imp_all3)) == 0 &
  apply(prot_imp_all3, 2, var) > 0 &
  is.finite(apply(prot_imp_all3, 2, var))

prot_imp_all3 <- prot_imp_all3[, keep_cols_all3, drop = FALSE]

cat("Proteins kept for all-3-omics PCA:", ncol(prot_imp_all3), "\n")
```
##---------------------------
##outlier check and removal
##---------------------------
```{r}
pca0_all3 <- prcomp(prot_imp_all3, center = TRUE, scale. = TRUE)
scores3 <- as.data.frame(pca0_all3$x[, 1:2])  # PC1 + PC2 (or use more PCs)
scores3$dist <- sqrt(scores3$PC1^2 + scores3$PC2^2)

# identify outliers > 3 SD
threshold3 <- mean(scores3$dist) + 3 * sd(scores3$dist)
outliers_all3 <- rownames(scores3)[scores3$dist > threshold3]
outliers_all3

meta_pca_all3b<- meta_pca_all3 %>%
  filter(!record_id %in% outliers_all3)

prot_imp_all3b <- prot_imp_all3[
  !rownames(prot_imp_all3) %in% outliers_all3, , drop = FALSE
]

cat("Still present in metadata? ", any(outliers_all3 %in% meta_pca_all3b$record_id), "\n")
cat("Still present in matrix? ", any(outliers_all3 %in% rownames(prot_imp_all3b)), "\n")

stopifnot(!any(outliers_all3 %in% meta_pca_all3b$record_id))
stopifnot(!any(outliers_all3 %in% rownames(prot_imp_all3b)))
stopifnot(identical(rownames(prot_imp_all3b), meta_pca_all3b$record_id))

##checking for low variance prots--none exist
nzv_idx2 <- nearZeroVar(prot_imp_all3b)
length(nzv_idx2)   
colnames(prot_imp_all3b)[nzv_idx2]
```

```{r}
# ---- PCA panels ----
p_all3 <- make_pca_plot(prot_imp_all3b, meta_pca_all3b, "No adjustment (all 3 omics subset)")

cov_m1_all3 <- model.matrix(~ age + sex, data = meta_pca_all3b)[, -1, drop = FALSE]
prot_m1_all3 <- t(removeBatchEffect(t(prot_imp_all3b), covariates = cov_m1_all3))
p1_all3 <- make_pca_plot(prot_m1_all3, meta_pca_all3b, "Age + Sex (all 3 omics subset)")

keep_m2_all3 <- !is.na(meta_pca_all3b$bmi)
meta_m2p_all3 <- meta_pca_all3b[keep_m2_all3, ]
prot_m2p_all3 <- prot_imp_all3b[keep_m2_all3, , drop = FALSE]
cov_m2_all3 <- model.matrix(~ age + sex + study6, data = meta_m2p_all3)[, -1, drop = FALSE]
prot_m2_adj_all3 <- t(removeBatchEffect(t(prot_m2p_all3), covariates = cov_m2_all3))
p2_all3 <- make_pca_plot(prot_m2_adj_all3, meta_m2p_all3, "Age + Sex + Study (all 3 omics subset)")

pca_plot3 <- (p_all3 | p1_all3 | p2_all3)+
  plot_annotation(
    title = "PCA: Proteomics Sensitivity Analysis in Participants with All 3 Omics Types",
    subtitle = "Color = group | Shape = study | Restricted to shared metabolomics + proteomics participants",
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 9, hjust = 0.5, color = "grey40")
    )
  )

ggsave(filename = file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Results","pca_allOmics.pdf"),
                            plot = pca_plot3,
                            width = 16, height = 14)
```
##--------------------------------------------------
##limma using all-3-omics subset
## model1: age + sex
## model2: age + sex + study
##--------------------------------------------------
```{r}
Y_all_all3 <- t(prot_log_all3)


# Model 1: age + sex
meta_m1_all3 <- meta_pca_all3b
Y_m1_all3 <- Y_all_all3[, meta_m1_all3$record_id, drop = FALSE]

design_m1_all3 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m1_all3)
colnames(design_m1_all3) <- sub("^group4", "", colnames(design_m1_all3))
colnames(design_m1_all3) <- make.names(colnames(design_m1_all3))

contr_m1_all3 <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_m1_all3
)

res_m1_all3 <- run_limma_contrasts(Y_m1_all3, design_m1_all3, contr_m1_all3)
fit_m1_all3 <- res_m1_all3$fit
out_m1_all3 <- res_m1_all3$out

sig_m1_all3 <- tibble(
  contrast = colnames(contr_m1_all3),
  n_FDR05 = sapply(out_m1_all3, \(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_m1_all3, \(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE))
) %>% mutate(model = "age_sex_all3", n_samples = nrow(meta_m1_all3))



# Model 2: age + sex + study
Y_m2_all3 <- Y_all_all3[, meta_m1_all3$record_id, drop = FALSE]

design_m2_all3 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m1_all3)
colnames(design_m2_all3) <- sub("^group4", "", colnames(design_m2_all3))
colnames(design_m2_all3) <- make.names(colnames(design_m2_all3))

contr_m2_all3 <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_m2_all3
)

res_m2_all3 <- run_limma_contrasts(Y_m2_all3, design_m2_all3, contr_m2_all3)
fit_m2_all3 <- res_m2_all3$fit
out_m2_all3 <- res_m2_all3$out
sig_m2_all3 <- summarise_sig(out_m2_all3, "age_sex_study_all3", nrow(meta_m1_all3))
sig_m2_all3
```
## --------------------------------------------------
## Summarize all-3-omics results
##--------------------------------------------------
```{r}
summary_all_all3 <- bind_rows(
  sig_m1_all3,
  sig_m2_all3
) %>%
  dplyr::select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

print(summary_all_all3)

summary_wide_all3 <- summary_all_all3 %>%
  dplyr::select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

print(summary_wide_all3)

summary_compare_all3 <- bind_rows(
  summary_all %>% mutate(dataset = "original"),
  summary_all_all3 %>% mutate(dataset = "all_3_omics_subset")
) %>%
  arrange(contrast, dataset, model)

print(summary_compare_all3)

```

```{r}
FDR_THRESH <- 0.05

contrasts <- c("T1D_vs_LC", "T2D_vs_LC", "T1D_vs_T2D",
               "OC_vs_LC",  "T1D_vs_OC", "T2D_vs_OC")

get_sig <- function(out_list, contrast, fdr = FDR_THRESH) {
  tt <- out_list[[contrast]]
  if (is.null(tt)) return(character(0))
  tt %>%
    dplyr::filter(!is.na(adj.P.Val), adj.P.Val < fdr) %>%
    dplyr::pull(feature_id) %>%
    unique()
}

overlap_results <- lapply(contrasts, function(cc) {
  
  all_studies <- get_sig(out_m2, cc)
  noATT       <- get_sig(out_m2_noATT, cc)
  all3        <- get_sig(out_m2_all3, cc)
  
  tibble::tibble(
    contrast = cc,
    
    n_all_studies = length(all_studies),
    n_noATT       = length(noATT),
    n_all3        = length(all3),
    
    overlap_all_vs_noATT = length(intersect(all_studies, noATT)),
    overlap_all_vs_all3  = length(intersect(all_studies, all3)),
    overlap_noATT_vs_all3 = length(intersect(noATT, all3)),
    
    overlap_all_three = length(Reduce(intersect, list(all_studies, noATT, all3))),
    
    union_all_vs_noATT = length(union(all_studies, noATT)),
    union_all_vs_all3  = length(union(all_studies, all3)),
    union_noATT_vs_all3 = length(union(noATT, all3)),
    union_all_three    = length(Reduce(union, list(all_studies, noATT, all3))),
    
    jaccard_all_vs_noATT = ifelse(length(union(all_studies, noATT)) > 0,
                                  length(intersect(all_studies, noATT)) / length(union(all_studies, noATT)),
                                  NA_real_),
    
    jaccard_all_vs_all3 = ifelse(length(union(all_studies, all3)) > 0,
                                 length(intersect(all_studies, all3)) / length(union(all_studies, all3)),
                                 NA_real_),
    
    jaccard_noATT_vs_all3 = ifelse(length(union(noATT, all3)) > 0,
                                   length(intersect(noATT, all3)) / length(union(noATT, all3)),
                                   NA_real_),
    
    jaccard_all_three = ifelse(length(Reduce(union, list(all_studies, noATT, all3))) > 0,
                               length(Reduce(intersect, list(all_studies, noATT, all3))) /
                                 length(Reduce(union, list(all_studies, noATT, all3))),
                               NA_real_)
  )
})

overlap_summary <- dplyr::bind_rows(overlap_results)
print(overlap_summary)
```

#---------------------------------------------------
# saving limma outputs
# Three analyses, all using age + sex + study model
#---------------------------------------------------
```{r}
prep_gsea_tables <- function(out_list) {
  lapply(out_list, function(tt) {
    tt %>%
      dplyr::select(feature_id, logFC, AveExpr, t, P.Value, adj.P.Val, B, contrast) %>%
      dplyr::arrange(desc(t)) 
  })
}

out_m2 <- prep_gsea_tables(out_m2)
out_m2_noATT <- prep_gsea_tables(out_m2_noATT)
out_m2_all3 <- prep_gsea_tables(out_m2_all3)


contrast_order <- c(
  "T1D_vs_LC", "T2D_vs_LC", "T1D_vs_T2D",
  "OC_vs_LC", "T1D_vs_OC", "T2D_vs_OC"
)

limma_gsea_inputs <- list(
  
  metadata = list(
    created = Sys.time(),
    response = "plasma proteomics",
    transform = "log2(x + 1)",
    model_formula = "0 + group4 + age + sex + study6",
    contrasts = contrast_order
  ),
  
  all_studies = list(
    model_name = "age_sex_study6",
    samples = meta_m2,
    design = design_m2,
    fit = fit_m2,
    out = out_m2,
    gsea_tables = out_m2
  ),
  
  no_attempt = list(
    model_name = "age_sex_study_noATT",
    samples = meta_m2_noATT,
    design = design_m2_noATT,
    fit = fit_m2_noATT,
    out = out_m2_noATT,
    gsea_tables = out_m2_noATT
  ),
  
  all3_omics = list(
    model_name = "age_sex_study4",
    samples = meta_all3,
    design = design_m2_all3,
    fit = fit_m2_all3,
    out = out_m2_all3,
    gsea_tables = out_m2_all3
  )
)

saveRDS(
  limma_gsea_inputs,
  file = file.path(
    dir.dat, "Savanah Leidholt", "T1D-T2D", "Data",
    "limma_gsea_inputs_age_sex_study.rds"
  )
)

cat("Saved: limma_gsea_inputs_age_sex_study.rds\n")
```

#---------------------------------------------------
# Three analyses with reduced model
#---------------------------------------------------
```{r}

limma_gsea_inputs <- list(
  
  metadata = list(
    created = Sys.time(),
    response = "plasma proteomics",
    transform = "log2(x + 1)",
    model_formula = "0 + group4 + age + sex",
    contrasts = contrast_order
  ),
  
  no_attempt_m1 = list(
    model_name = "age_sex_noATT",
    samples = meta_m1_noATT,
    design = design_m1_noATT,
    fit = fit_m1_noATT,
    out = out_m1_noATT,
    gsea_tables = prep_gsea_tables(out_m1_noATT)
  ),
  
  all3_omics_m1 = list(
    model_name = "age_sex_multiOmics",
    samples = meta_all3,
    design = design_m1_all3,
    fit = fit_m1_all3,
    out = out_m1_all3,
    gsea_tables = prep_gsea_tables(out_m1_all3)
  )
)

saveRDS(
  limma_gsea_inputs,
  file = file.path(
    dir.dat, "Savanah Leidholt", "T1D-T2D", "Data",
    "limma_gsea_inputs_age_sex.rds"
  )
)

cat("Saved: limma_gsea_inputs_age_sex.rds\n")

```

