---
title: "Plasma_DEP_anlaysis"
author: "Savanah Leidholt"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ---- packages ----
library(dplyr)
library(tidyr)
library(limma)
library(EnhancedVolcano)
library(ggplot2)
library(patchwork)
library(ggrepel)
library(vegan)
library(tidyverse)
library(data.table)
library(purrr)
library(haven)
library(ggVennDiagram)
library(caret)
library(jsonlite)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- "/mmfs1/gscratch/togo/leidholt/"
  git_path  <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys      <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
  root_path <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
  dir.results <- file.path(root_path, "Savanah Leidholt", "TRIAD", "Results")
  git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "savanahleidholt") {
  root_path <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
  dir.results <- file.path(root_path, "Savanah Leidholt", "TRIAD", "Results")
  git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}
```
#-----------------------------------------------------------------------------
#Clean and setup data
# a) all proteomic data DPE analysis
# b) sensitivty analysis where I remove participants who have T1D and are obese
# c) sensitivty analysis (rm attempt)
# d) DPE analysis with samples that contain both proteomic and metabolomic data
#------------------------------------------------------------------------------
```{r}
obj <- readRDS(file.path(root_path, "Savanah Leidholt", "TRIAD", "Data", "omics_split_objects_clean.rds"))

meta   <- obj$meta_prot
prot_plasma <- obj$prot_plasma
n_table     <- obj$n_table

# quick check
n_table
dim(prot_plasma)
dim(meta)


```
##---------------------------------------------
##open panther_data.RData to add in hba1c data
##---------------------------------------------
```{r}
load(file.path(root_path, "PANTHER", "Data_cleaned", "panther_data.RData"))

#trasnforming data to match the harmonized data
panther_hba1c <- panther %>%
  dplyr::select(record_id, hba1c) %>%
  transmute(
    record_id = gsub("_", "-", as.character(record_id)),
    hba1c_panther = as.numeric(hba1c)
  )

panther_hba1c$record_id <- as.character(panther_hba1c$record_id)

#Make sure harm$hba1c is numeric too (if it exists)
meta <- meta %>%
  mutate(hba1c = as.numeric(zap_labels(hba1c))) %>%
  left_join(panther_hba1c, by = "record_id") %>%
  mutate(
    hba1c = if_else(
      study == "PANTHER" & is.na(hba1c),
      hba1c_panther,
      hba1c
    )
  ) %>%
  dplyr::select(-hba1c_panther)
```
##---------------------
##cleaning metadata
##---------------------
```{r}
meta2 <- meta %>%
  mutate(
    record_id = trimws(as.character(record_id)),
    study     = trimws(as.character(study)),
    group     = trimws(as.character(group)),
    sex       = trimws(as.character(sex)),
    ethnicity = trimws(as.character(race_ethnicity_condensed)),
    age       = suppressWarnings(as.numeric(age)),
    uacr      = suppressWarnings(as.numeric(acr_u)),
    log_uacr  = log(uacr + 1)
  ) %>%
  filter(
    !is.na(age),
    !is.na(sex),
    sex != ""
  ) %>%
  mutate(
    group4 = factor(
      group,
      levels = c("Lean Control", "Obese Control", "Type 1 Diabetes", "Type 2 Diabetes"),
      labels = c("LC", "OC", "T1D", "T2D")
    ),
    ethnicity = factor(ethnicity),
    study6 = dplyr::recode(
      study,
      "ATTEMPT"          = "ATT",
      "CROCODILE"        = "CRC",
      "PANTHER"          = "PAN",
      "RENAL-HEIR"       = "RH",
      "IMPROVE"          = "IT",
      "RENAL-HEIRitage"  = "RH2",
      .default = study
    ),
    study6 = factor(study6)
  )

#removing rows with no proteomics data 
rownames(prot_plasma) <- trimws(rownames(prot_plasma))

has_prot <- rowSums(!is.na(prot_plasma)) > 0
prot_plasma <- prot_plasma[has_prot, , drop = FALSE]

meta2 <- meta2 %>%
  filter(record_id %in% rownames(prot_plasma))

olap <- intersect(meta2$record_id, rownames(prot_plasma))
cat("meta n:", nrow(meta), "\n")
cat("meta2 n:", nrow(meta2), "\n")

cat("prot_plasma dim:", dim(prot_plasma), "\n")
cat("overlap n:", length(olap), "\n")
cat("meta not in prot:", sum(!meta2$record_id %in% rownames(prot_plasma)), "\n")
cat("prot not in meta:", sum(!rownames(prot_plasma) %in% meta2$record_id), "\n")

stopifnot(length(olap) > 0)

meta3 <- meta2[match(olap, meta2$record_id), , drop = FALSE]
prot2 <- prot_plasma[olap, , drop = FALSE]

stopifnot(identical(rownames(prot2), meta3$record_id))

## Check near-zero variance proteins
nzv_idx <- caret::nearZeroVar(prot2) #zero
```

##--------------------------------------------------------------
## Protein Filter and Log Transformation
##filter to proteins that are only in the analyte panel (~7000)
##---------------------------------------------------------------
```{r}
analytes <- read.csv(
  file.path(root_path, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

prot_panel <- unique(trimws(as.character(analytes$AptName)))
aptamers_keep <- intersect(prot_panel, colnames(prot2))
prot3 <- prot2[, aptamers_keep, drop = FALSE]

dim(prot3)   # should be ~7000
dim(prot2)   # should be ~10000

length(aptamers_keep)
length(setdiff(prot_panel, colnames(prot2)))

## Log transform protein data
prot_log <- log2(as.matrix(prot3) + 1)
prot_log[!is.finite(prot_log)] <- NA_real_

## Check near-zero variance after filtering
nzv_idx <- caret::nearZeroVar(prot3) #zero

```

```{r}
#not really necessary to do as all protiens pass filter once I remove the extra 3000 from the ATTEMPT Panel
##filtering step where protein must exist in 60% of samples in a study

#study_idx_all <- split(seq_len(nrow(prot_log)), meta_common$study)
#present_by_study_all <- sapply(study_idx_all, function(idx) {
#  colMeans(!is.na(prot_log[idx, ])) >= 0.6
#})

#present_by_study_all <- as.matrix(present_by_study_all)
#n_studies_present_all <- colSums(present_by_study_all)
#keep_study_all <- n_studies_present_all >= 2

#removed_proteins <- colnames(prot_log)[!keep_study_all]
#kept_proteins    <- colnames(prot_log)[keep_study_all]

#cat("All proteomic features before study filter:", ncol(prot_log), "\n")
#cat("Removed (<2 studies meeting 60% presence threshold):", length(removed_proteins), "\n")
#cat("Kept (>=2 studies meeting 60% presence threshold):", length(kept_proteins), "\n")
```

##--------------------------------
## Find and Remove Outliers in PCA
##---------------------------------
```{r}
pca0 <- prcomp(prot_log, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2])
scores$record_id <- rownames(scores)
scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

threshold <- mean(scores$dist) + 3 * sd(scores$dist)
outliers <- scores$record_id[scores$dist > threshold]
outliers #"10424"   "RH-84-T"

meta4 <- meta3 %>%
  filter(!record_id %in% outliers)

prot_log2 <- prot_log[
  !rownames(prot_log) %in% outliers,
  ,
  drop = FALSE
]

cat("Still present in metadata? ", any(outliers %in% meta4$record_id), "\n")
cat("Still present in matrix? ", any(outliers %in% rownames(prot_log2)), "\n")

stopifnot(!any(outliers %in% meta4$record_id))
stopifnot(!any(outliers %in% rownames(prot_log2)))
stopifnot(identical(rownames(prot_log2), meta4$record_id))

saveRDS(prot_log2,
  file = file.path(root_path,"Savanah Leidholt","TRIAD","Data","prot_log2.rds"
  )
)
```

#---------------
# Plotting Funciton
#---------------
```{r}
make_pca_plot <- function(prot_mat, meta_df, title_str) {
  pca  <- prcomp(prot_mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1    = pca$x[, 1],
    PC2    = pca$x[, 2],
    group4 = meta_df$group4,
    study  = meta_df$study6
  ) %>%
    ggplot(aes(PC1, PC2, color = group4, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        LC  = "Cyan4",
        OC  = "Gold",
        T1D = "DarkRed",
        T2D = "Purple"
      ),
      name = "Group"
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
##-----------
## PCAs
##-----------
```{r}
design <- model.matrix(~ group4, data = meta4)

## no adjustment
p0 <- make_pca_plot(prot_log2, meta4, "No adjustment")

## age + sex adjustment
cov1 <- model.matrix(~ age + sex, data = meta4)[, -1, drop = FALSE]
prot_adj1 <- t(limma::removeBatchEffect(t(prot_log2), covariates = cov1, design = design))
p1 <- make_pca_plot(prot_adj1, meta4, "Age + Sex")

## age + sex + study adjustment
cov2 <- model.matrix(~ age + sex + study6, data = meta4)[, -1, drop = FALSE]
prot_adj2 <- t(limma::removeBatchEffect(t(prot_log2), covariates = cov2, design = design))
p2 <- make_pca_plot(prot_adj2, meta4, "Age + Sex + Study")

## combine plots
pca_combined <- p0 | p1 | p2
pca_combined

#ggsave(filename = file.path(root_path, "Proteomics Figures", "limma", "pca_plots.pdf"),
#       plot = pca_combined,
#       width = 16,
#       height = 14)
```

##------------------------------
##limma testing all prots data
## model1: age + sex
## model2: age + sex + study
##------------------------------
```{r}
# helper function-- run limma + return list of topTables
run_limma_contrasts <- function(Y, design, contr) {
  fit <- lmFit(Y, design)
  fit2 <- contrasts.fit(fit, contr)
  fit2 <- eBayes(fit2)

  out <- setNames(vector("list", ncol(contr)), colnames(contr))
  for (cc in colnames(contr)) {
    tt <- topTable(fit2, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
    tt$feature_id <- rownames(tt)
    tt$contrast <- cc
    out[[cc]] <- tt
  }
  list(fit = fit2, out = out)
}

# ---- helper: summarise sig hits per contrast ----
summarise_sig <- function(out_list, model_label, n_samp) {
  tibble(
    contrast = names(out_list),
    n_FDR05  = sapply(out_list, \(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
    n_FDR01  = sapply(out_list, \(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
    model    = model_label,
    n_samples = n_samp
  )
}

## response matrices: proteins x samples
Y0 <- t(prot_log)
Y1 <- t(prot_log2)

## check sample alignment
stopifnot(identical(colnames(Y0), meta3$record_id))
stopifnot(identical(colnames(Y1), meta4$record_id))
```

##--------------------------------------------------------------
## Model 1: age + sex, before outlier removal
##--------------------------------------------------------------
```{r}
meta_m0 <- meta3
Y_m0 <- Y0[, meta_m0$record_id, drop = FALSE]

design_m0 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m0)
colnames(design_m0) <- sub("^group4", "", colnames(design_m0))
colnames(design_m0) <- make.names(colnames(design_m0))

contr_m0 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, 
                                 levels = design_m0)

res_m0 <- run_limma_contrasts(Y_m0, design_m0, contr_m0)
fit_m0 <- res_m0$fit
out_m0 <- res_m0$out

sig_m0 <- summarise_sig(out_m0, "age + sex, before outlier removal", nrow(meta_m0))

sig_m0
```
##--------------------------------------------------------------
## Model 2: age + sex, after outlier removal
##--------------------------------------------------------------
```{r}
meta_m1 <- meta4
Y_m1 <- Y1[, meta_m1$record_id, drop = FALSE]

design_m1 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m1)
colnames(design_m1) <- sub("^group4", "", colnames(design_m1))
colnames(design_m1) <- make.names(colnames(design_m1))

contr_m1 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, 
                                 levels = design_m1)

res_m1 <- run_limma_contrasts(Y_m1, design_m1, contr_m1)
fit_m1 <- res_m1$fit
out_m1 <- res_m1$out

sig_m1 <- summarise_sig(out_m1, "age + sex, after outlier removal", nrow(meta_m1))

sig_m1
```
##--------------------------------------------------------------
## Model 3: age + sex + study, before outlier removal
##--------------------------------------------------------------
```{r}
meta_m2 <- meta3 %>%
  filter(!is.na(study6))

Y_m2 <- Y0[, meta_m2$record_id, drop = FALSE]

design_m2 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m2)
colnames(design_m2) <- sub("^group4", "", colnames(design_m2))
colnames(design_m2) <- make.names(colnames(design_m2))

contr_m2 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, levels = design_m2)

res_m2 <- run_limma_contrasts(Y_m2, design_m2, contr_m2)
fit_m2 <- res_m2$fit
out_m2 <- res_m2$out

sig_m2 <- summarise_sig(out_m2, "age + sex + study, before outlier removal", nrow(meta_m2))

sig_m2
```
##--------------------------------------------------------------
## Model 4: age + sex + study, after outlier removal
##--------------------------------------------------------------
```{r}
meta_m3 <- meta4 %>%
  filter(!is.na(study6))

Y_m3 <- Y1[, meta_m3$record_id, drop = FALSE]

design_m3 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m3)
colnames(design_m3) <- sub("^group4", "", colnames(design_m3))
colnames(design_m3) <- make.names(colnames(design_m3))

contr_m3 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, 
                                 T2D_vs_LC = T2D - LC, 
                                 T1D_vs_T2D = T1D - T2D, 
                                 OC_vs_LC = OC - LC, 
                                 T1D_vs_OC = T1D - OC, 
                                 T2D_vs_OC = T2D - OC, levels = design_m3)

res_m3 <- run_limma_contrasts(Y_m3, design_m3, contr_m3)
fit_m3 <- res_m3$fit
out_m3 <- res_m3$out

sig_m3 <- summarise_sig(out_m3, "age + sex + study, after outlier removal", nrow(meta_m3))

sig_m3
```
```{r}
sig_overlap_m1 <- bind_rows(
  lapply(names(out_m0), function(cc) {

    pre_ids <- out_m0[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    post_ids <- out_m1[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_pre <- length(pre_ids)
    n_overlap <- length(intersect(pre_ids, post_ids))

    tibble(
      model = "age + sex",
      contrast = cc,
      n_pre = n_pre,
      n_post = length(post_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_pre, 1)
    )
  })
)

sig_overlap_m2 <- bind_rows(
  lapply(names(out_m2), function(cc) {

    pre_ids <- out_m2[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    post_ids <- out_m3[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_pre <- length(pre_ids)
    n_overlap <- length(intersect(pre_ids, post_ids))

    tibble(
      model = "age + sex + study",
      contrast = cc,
      n_pre = n_pre,
      n_post = length(post_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_pre, 1)
    )
  })
)

sig_overlap_all <- bind_rows(sig_overlap_m1, sig_overlap_m2)

sig_overlap_all
```
#-------------------------------------
# Sensitivity analysis:
# Remove T1D participants who are obese
#--------------------------------------
```{r}
#obesity status--check if this is ok. patients range in age from 13-25 so I may need to do percentiles for youth
meta4 <- meta4 %>%
  dplyr::mutate(
    bmi = suppressWarnings(as.numeric(bmi)),
    bmip = suppressWarnings(as.numeric(bmip)),
    bmi_obesity = dplyr::case_when(
      age > 19 & bmi < 25 ~ "Normal",
      age > 19 & bmi >= 25 & bmi < 30 ~ "Overweight",
      age > 19 & bmi >= 30 ~ "Obese",
      age <= 19 & bmip < 85 ~ "Normal",
      age <= 19 & bmip >= 85 & bmip < 95 ~ "Overweight",
      age <= 19 & bmip >= 95 ~ "Obese",
      TRUE ~ NA_character_
    ),
    bmi_obesity = factor(bmi_obesity, levels = c("Normal", "Overweight", "Obese"))
  )

table(meta4$bmi_obesity, useNA = "ifany")
meta4 %>%
  count(group4, bmi_obesity) %>%
  tidyr::pivot_wider(
    names_from = bmi_obesity,
    values_from = n,
    values_fill = 0
  )
```

```{r}
meta5 <- meta4 %>%
  mutate(remove_t1d_obese = group4 == "T1D" & bmi_obesity == "Obese") %>%
  filter(!remove_t1d_obese) %>%
  droplevels()

prot_log3 <- prot_log2[meta5$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_log3), meta5$record_id))

cat("Primary analysis N:", nrow(meta4), "\n")
cat("Sensitivity analysis N:", nrow(meta5), "\n")
cat("Removed obese T1D N:", nrow(meta4) - nrow(meta5), "\n")

meta5 %>%
  count(group4, bmi_obesity)

meta5 %>%
  count(group4, study6) %>%
  arrange(group4, study6)

table(meta5$group4)
```

# -----------------------------
# PCA after removing obese T1D
# -----------------------------
```{r}
design_sens <- model.matrix(~ group4, data = meta5)

p_sens0 <- make_pca_plot(prot_log3, meta5, "Sensitivity: No adjustment")

cov_sens1 <- model.matrix(~ age + sex, data = meta5)[, -1, drop = FALSE]
prot_sens1 <- t(limma::removeBatchEffect(t(prot_log3), covariates = cov_sens1, design = design_sens))
p_sens1 <- make_pca_plot(prot_sens1, meta5, "Sensitivity: Age + Sex")

cov_sens2 <- model.matrix(~ age + sex + study6, data = meta5)[, -1, drop = FALSE]
prot_sens2 <- t(limma::removeBatchEffect(t(prot_log3), covariates = cov_sens2, design = design_sens))
p_sens2 <- make_pca_plot(prot_sens2, meta5, "Sensitivity: Age + Sex + Study")

pca_sens_combined <- p_sens0 | p_sens1 | p_sens2

pca_sens_combined

#ggsave(filename = file.path(root_path, "Proteomics Figures", "limma","pca_plots_sensitivity_no_obese_T1D.pdf"),
#       plot = pca_sens_combined,
#       width = 16,
#       height = 14)
```

# -----------------------------
#limma adiposity sensitivity data
# -----------------------------
```{r}
Y_sens <- t(prot_log3)
stopifnot(identical(colnames(Y_sens), meta5$record_id))
```

##---------------------------------
## adiposity model 1: age + sex
##----------------------------------
```{r}
meta_m4 <- meta5
Y_m4 <- Y_sens[, meta_m4$record_id, drop = FALSE]

design_m4 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m4)
colnames(design_m4) <- sub("^group4", "", colnames(design_m4))
colnames(design_m4) <- make.names(colnames(design_m4))

contr_m4 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m4)

res_m4 <- run_limma_contrasts(Y_m4, design_m4, contr_m4)
fit_m4 <- res_m4$fit
out_m4 <- res_m4$out

sig_m4 <- summarise_sig(out_m4, "sensitivity: age + sex, no obese T1D", nrow(meta_m4))

sig_m4


```
##----------------------------------------
## adiposity model 2: age + sex + study
##-----------------------------------------
```{r}
meta_m5 <- meta5 %>%
  filter(!is.na(study6)) %>%
  droplevels()

Y_m5 <- Y_sens[, meta_m5$record_id, drop = FALSE]

design_m5 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m5)
colnames(design_m5) <- sub("^group4", "", colnames(design_m5))
colnames(design_m5) <- make.names(colnames(design_m5))

contr_m5 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m5)

res_m5 <- run_limma_contrasts(Y_m5, design_m5, contr_m5)
fit_m5 <- res_m5$fit
out_m5 <- res_m5$out

sig_m5 <- summarise_sig(out_m5, "sensitivity: age + sex + study, no obese T1D", nrow(meta_m5))

sig_m5
```
##-------------------------------------------------
## Compare primary and adipoisty sensitivity limma summaries
##-------------------------------------------------
```{r}
summary_sensitivity <- bind_rows(
  sig_m1 %>% mutate(model = "primary: age + sex"),
  sig_m3 %>% mutate(model = "primary: age + sex + study"),
  sig_m4,
  sig_m5
) %>%
  select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

summary_sensitivity

summary_sensitivity_wide <- summary_sensitivity %>%
  select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

summary_sensitivity_wide


sig_overlap_study <- bind_rows(
  lapply(names(out_m3), function(cc) {

    primary_ids <- out_m3[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    sens_ids <- out_m5[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_primary <- length(primary_ids)
    n_overlap <- length(intersect(primary_ids, sens_ids))

    tibble(
      contrast = cc,
      n_primary = n_primary,
      n_sensitivity = length(sens_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_primary, 1)
    )
  })
)

sig_overlap_study
```
#--------------------------------------------------------
# SENSITIVITY ANALYSIS:
# Exclude ATTEMPT study (different sequencing platform)
#--------------------------------------------------------
```{r}
meta6 <- meta4 %>%
  filter(study != "ATTEMPT", study6 != "ATT") %>%
  droplevels()

prot_log4 <- prot_log2[meta6$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_log4), meta6$record_id))

cat("Primary analysis N:", nrow(meta4), "\n")
cat("No-ATTEMPT sensitivity N:", nrow(meta6), "\n")
cat("Removed ATTEMPT N:", nrow(meta4) - nrow(meta6), "\n")

meta6 %>%
  count(group4, study6) %>%
  arrange(group4, study6)

table(meta6$group4)
```
##---------------------------
##outlier check and removal
##---------------------------
```{r}
pca_noATT <- prcomp(prot_log4, center = TRUE, scale. = TRUE)

scores_noATT <- as.data.frame(pca_noATT$x[, 1:2])
scores_noATT$record_id <- rownames(scores_noATT)
scores_noATT$dist <- sqrt(scores_noATT$PC1^2 + scores_noATT$PC2^2)

threshold_noATT <- mean(scores_noATT$dist) + 3 * sd(scores_noATT$dist)

outliers_noATT <- scores_noATT$record_id[scores_noATT$dist > threshold_noATT]

outliers_noATT

#meta7 <- meta6 %>%
 # filter(!record_id %in% outliers_noATT) %>%
  #droplevels()

#prot_log5 <- prot_log4[!rownames(prot_log4) %in% outliers_noATT, , drop = FALSE]

#cat("Still present in metadata? ", any(outliers_noATT %in% meta7$record_id), "\n")
#cat("Still present in matrix? ", any(outliers_noATT %in% rownames(prot_log5)), "\n")

#stopifnot(!any(outliers_noATT %in% meta7$record_id))
#stopifnot(!any(outliers_noATT %in% rownames(prot_log5)))
#stopifnot(identical(rownames(prot_log5), meta7$record_id))
```
##--------------
##PCA no ATTEMPt
##--------------
```{r}
design_noATT <- model.matrix(~ group4, data = meta6)

p_noATT0 <- make_pca_plot(prot_log4, meta6, "Sensitivity: No adjustment, excluding ATTEMPT")

cov_noATT1 <- model.matrix(~ age + sex, data = meta6)[, -1, drop = FALSE]
prot_noATT1 <- t(limma::removeBatchEffect(t(prot_log4), covariates = cov_noATT1, design = design_noATT))
p_noATT1 <- make_pca_plot(prot_noATT1, meta6, "Sensitivity: Age + Sex, excluding ATTEMPT")

cov_noATT2 <- model.matrix(~ age + sex + study6, data = meta6)[, -1, drop = FALSE]
prot_noATT2 <- t(limma::removeBatchEffect(t(prot_log4), covariates = cov_noATT2, design = design_noATT))
p_noATT2 <- make_pca_plot(prot_noATT2, meta6, "Sensitivity: Age + Sex + Study, excluding ATTEMPT")

pca_noATT_combined <- (p_noATT0 | p_noATT1 | p_noATT2) +
  patchwork::plot_annotation(
    title = "PCA: Sensitivity Analysis Excluding ATTEMPT",
    subtitle = "Color = group | Shape = study | ATTEMPT removed",
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 9, hjust = 0.5, color = "grey40")
    )
  )

pca_noATT_combined

#ggsave(filename = file.path(root_path, "Proteomics Figures", "limma", "pca_noATTEMPT.pdf"),
#       plot = pca_noATT_combined,
#       width = 16,
#       height = 14)
```
##----------------------------------
## limma using no-ATTEMPT dataset
## model 1: age + sex
## model 2: age + sex + study
## ---------------------------------
```{r}
Y_noATT <- t(prot_log4)
stopifnot(identical(colnames(Y_noATT), meta6$record_id))

# Model 1: age + sex
meta_m6 <- meta6
Y_m6 <- Y_noATT[, meta_m6$record_id, drop = FALSE]

design_m6 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m6)
colnames(design_m6) <- sub("^group4", "", colnames(design_m6))
colnames(design_m6) <- make.names(colnames(design_m6))

contr_m6 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m6)

res_m6 <- run_limma_contrasts(Y_m6, design_m6, contr_m6)
fit_m6 <- res_m6$fit
out_m6 <- res_m6$out

sig_m6 <- summarise_sig(out_m6, "sensitivity: age + sex, no ATTEMPT", nrow(meta_m6))

sig_m6


# Model 2: age + sex + study
meta_m7 <- meta6 %>%
  filter(!is.na(study6)) %>%
  droplevels()

Y_m7 <- Y_noATT[, meta_m7$record_id, drop = FALSE]

design_m7 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m7)
colnames(design_m7) <- sub("^group4", "", colnames(design_m7))
colnames(design_m7) <- make.names(colnames(design_m7))

contr_m7 <- limma::makeContrasts(T1D_vs_LC = T1D - LC, T2D_vs_LC = T2D - LC, T1D_vs_T2D = T1D - T2D, OC_vs_LC = OC - LC, T1D_vs_OC = T1D - OC, T2D_vs_OC = T2D - OC, levels = design_m7)

res_m7 <- run_limma_contrasts(Y_m7, design_m7, contr_m7)
fit_m7 <- res_m7$fit
out_m7 <- res_m7$out

sig_m7 <- summarise_sig(out_m7, "sensitivity: age + sex + study, no ATTEMPT", nrow(meta_m7))

sig_m7

```
## -----------------------------
## Summarize no-ATTEMPT results
## -----------------------------
```{r}
summary_noATT <- bind_rows(
  sig_m1 %>% mutate(model = "primary: age + sex"),
  sig_m3 %>% mutate(model = "primary: age + sex + study"),
  sig_m6,
  sig_m7
) %>%
  select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

summary_noATT

summary_noATT_wide <- summary_noATT %>%
  select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

summary_noATT_wide

sig_overlap_noATT_study <- bind_rows(
  lapply(names(out_m3), function(cc) {

    primary_ids <- out_m3[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    noATT_ids <- out_m7[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_primary <- length(primary_ids)
    n_overlap <- length(intersect(primary_ids, noATT_ids))

    tibble(
      contrast = cc,
      n_primary = n_primary,
      n_noATT = length(noATT_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_primary, 1)
    )
  })
)

sig_overlap_noATT_study
```

#------------------------------------------------------------------------
# Sensitivity Analysis #3:
# Restrict proteomics workflow to participants with all 3 omics types
#fix to have scRNA and not lipidomic data
#------------------------------------------------------------------------
```{r}
obj <- readRDS(
  file.path(root_path, "Savanah Leidholt", "TRIAD", "Data", "omics_split_objects_clean.rds")
)

prot_log <- readRDS(
  file.path(root_path, "Savanah Leidholt", "TRIAD", "Data", "prot_log2.rds")
)

prot_plasma <- obj$prot_plasma
metabolites <- obj$metabolites
meta_prot <- obj$meta_prot
meta_metab <- obj$meta_metab

# clean rownames
rownames(prot_plasma) <- trimws(as.character(rownames(prot_plasma)))
rownames(prot_log)    <- trimws(as.character(rownames(prot_log)))
rownames(metabolites) <- trimws(as.character(rownames(metabolites)))

# IDs with any measured data in each omics matrix
ids_metab <- rownames(metabolites)[rowSums(!is.na(metabolites)) > 0]
ids_prot  <- rownames(prot_plasma)[rowSums(!is.na(prot_plasma)) > 0]
ids_log   <- rownames(prot_log)[rowSums(!is.na(prot_log)) > 0]

ids_metab <- trimws(as.character(ids_metab))
ids_prot  <- trimws(as.character(ids_prot))
ids_log   <- trimws(as.character(ids_log))

# paired IDs present in metabolomics, raw proteomics, and logged proteomics
paired_ids <- Reduce(intersect, list(ids_metab, ids_prot, ids_log))
```

```{r}
# HELPER: Study x Group count table for one omics type
make_count_table <- function(meta_df, mat_df, omics_label) {
  
  count_samples_with_data(meta_df, mat_df) %>%
    count(study, group) %>%
    pivot_wider(
      names_from = group,
      values_from = n,
      values_fill = 0
    ) %>%
    arrange(study) %>%
    mutate(Omics = omics_label, .before = 1)
}

make_count_table_group4 <- function(meta_df, omics_label = "Paired metabolomics + proteomics") {
  
  meta_df %>%
    count(study6, group4) %>%
    pivot_wider(
      names_from = group4,
      values_from = n,
      values_fill = 0
    ) %>%
    mutate(
      Total = rowSums(across(any_of(c("LC", "OC", "T1D", "T2D")))),
      Omics = omics_label,
      .before = 1
    ) %>%
    arrange(study6)
}
```

##---------------------------------------
## samples counts Study x Group
##---------------------------------------
```{r}
meta_mp <- meta_prot %>%
  mutate(
    record_id = trimws(as.character(record_id)),
    study     = trimws(as.character(study)),
    group     = trimws(as.character(group)),
    sex       = trimws(as.character(sex)),
    ethnicity = trimws(as.character(race_ethnicity_condensed)),
    age       = suppressWarnings(as.numeric(age)),
    uacr      = suppressWarnings(as.numeric(acr_u)),
    log_uacr  = log(uacr + 1)
  ) %>%
  filter(
    record_id %in% paired_ids,
    !is.na(age),
    !is.na(sex),
    sex != ""
  ) %>%
  mutate(
    group4 = factor(
      group,
      levels = c("Lean Control", "Obese Control", "Type 1 Diabetes", "Type 2 Diabetes"),
      labels = c("LC", "OC", "T1D", "T2D")
    ),
    ethnicity = factor(ethnicity),
    study6 = dplyr::recode(
      study,
      "ATTEMPT"          = "ATT",
      "CROCODILE"        = "CRC",
      "PANTHER"          = "PAN",
      "RENAL-HEIR"       = "RH",
      "IMPROVE"          = "IT",
      "RENAL-HEIRitage"  = "RH2",
      .default = study
    )
  ) %>%
  filter(study6 != "ATT") %>%
  mutate(study6 = factor(study6)) %>%
  arrange(record_id) %>%
  droplevels()

saveRDS(
  meta_mp,
  file = file.path(
    root_path,
    "Savanah Leidholt",
    "TRIAD",
    "Data",
    "paired_metadata.rds"
  )
)
```


```{r}
paired_ids_final <- meta_mp$record_id

prot_log6 <- prot_log[paired_ids_final, , drop = FALSE]
saveRDS(
  prot_log6,
  file = file.path(
    root_path,
    "Savanah Leidholt",
    "TRIAD",
    "Data",
    "proteomics_paired_matrix.rds"
  )
)

metab_mp  <- metabolites[paired_ids_final, , drop = FALSE]

stopifnot(identical(rownames(prot_log6), meta_mp$record_id))
stopifnot(identical(rownames(metab_mp), meta_mp$record_id))
```

```{r}
cat("Paired metabolomics + proteomics N:", nrow(meta_mp), "\n")
cat("Proteins:", nrow(prot_log6), "\n")
cat("Metabolites:", nrow(metab_mp), "\n")

meta_mp %>%
  count(study6, group4) %>%
  pivot_wider(
    names_from = group4,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(study6)
```
```{r}
tbl_mp <- meta_mp %>%
  count(study6, group4) %>%
  pivot_wider(
    names_from = group4,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(Total = rowSums(across(any_of(c("LC", "OC", "T1D", "T2D"))))) %>%
  arrange(study6)

tbl_mp
```


#-------------------
#sensitivty 3: PCAs 
#-------------------
```{r}
design_allomics <- model.matrix(~ group4, data = meta_mp)

p_all0 <- make_pca_plot(prot_log6,meta_mp,"Sensitivity: No adjustment, all omics")
cov_all1 <- model.matrix(~ age + sex,data = meta_mp)[, -1, drop = FALSE]
prot_all1 <- t(limma::removeBatchEffect(t(prot_log6),covariates = cov_all1, design = design_allomics))

p_all1 <- make_pca_plot(prot_all1,meta_mp,"Sensitivity: Age + Sex, all omics")
cov_all2 <- model.matrix(~ age + sex + study6,data = meta_mp)[, -1, drop = FALSE]
prot_all2 <- t(limma::removeBatchEffect( t(prot_log6),covariates = cov_all2,design = design_allomics))

p_all2 <- make_pca_plot(prot_all2,meta_mp,"Sensitivity: Age + Sex + Study, all omics")
pca_allomics <- p_all0 | p_all1 | p_all2
pca_allomics

ggsave(
  filename = file.path(root_path, "Savanah Leidholt", "TRIAD", "figures","Proteomics Figures","Limma","pca_allomics_sensitivity.pdf"),
  plot = pca_allomics,
  width = 16,
  height = 14
)
```
##--------------------------------------------------
##limma using all omics subset
## model1: age + sex
## model2: age + sex + study
##--------------------------------------------------
```{r}
Y_allomics <- t(prot_log6)
stopifnot(identical(colnames(Y_allomics), meta_mp$record_id))

# Model 1: age + sex
meta_m8 <- meta_mp
Y_m8 <- Y_allomics[, meta_m8$record_id, drop = FALSE]

design_m8 <- model.matrix(~ 0 + group4 + age + sex,data = meta_m8)
colnames(design_m8) <- sub("^group4", "", colnames(design_m8))
colnames(design_m8) <- make.names(colnames(design_m8))

contr_m8 <- limma::makeContrasts(
  T1D_vs_LC = T1D - LC,
  T2D_vs_LC = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC = OC - LC,
  T1D_vs_OC = T1D - OC,
  T2D_vs_OC = T2D - OC,
  levels = design_m8
)

res_m8 <- run_limma_contrasts(Y_m8, design_m8, contr_m8)
fit_m8 <- res_m8$fit
out_m8 <- res_m8$out

sig_m8 <- summarise_sig(out_m8,"sensitivity: age + sex, all omics",nrow(meta_m8))
sig_m8


# Model 2: age + sex + study
Y_m9 <- Y_allomics[, meta_m8$record_id, drop = FALSE]

design_m9 <- model.matrix(~ 0 + group4 + age + sex + study6,data = meta_m8)
colnames(design_m9) <- sub("^group4", "", colnames(design_m9))
colnames(design_m9) <- make.names(colnames(design_m9))

contr_m9 <- limma::makeContrasts(
  T1D_vs_LC = T1D - LC,
  T2D_vs_LC = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC = OC - LC,
  T1D_vs_OC = T1D - OC,
  T2D_vs_OC = T2D - OC,
  levels = design_m9
)

res_m9 <- run_limma_contrasts(Y_m9, design_m9, contr_m9)
fit_m9 <- res_m9$fit
out_m9 <- res_m9$out

sig_m9 <- summarise_sig( out_m9,"sensitivity: age + sex + study, all omics",nrow(meta_m8))
sig_m9
```
## --------------------------------------------------
## Summarize all-omics results compared to total
##--------------------------------------------------
```{r}
summary_allomics <- bind_rows(
  sig_m1 %>% mutate(model = "primary: age + sex"),
  sig_m3 %>% mutate(model = "primary: age + sex + study"),
  sig_m8,
  sig_m9
) %>%
  select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

summary_allomics

summary_allomics_wide <- summary_allomics %>%
  select(model, contrast, n_FDR05) %>%
  pivot_wider(
    names_from = model,
    values_from = n_FDR05
  )

summary_allomics_wide

sig_overlap_allomics_study <- bind_rows(
  lapply(names(out_m3), function(cc) {

    primary_ids <- out_m3[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    allomics_ids <- out_m9[[cc]] %>%
      filter(adj.P.Val < 0.05) %>%
      pull(feature_id)

    n_primary <- length(primary_ids)
    n_overlap <- length(intersect(primary_ids, allomics_ids))

    tibble(
      contrast = cc,
      n_primary = n_primary,
      n_allomics = length(allomics_ids),
      n_overlap = n_overlap,
      pct_overlap = round(100 * n_overlap / n_primary, 1)
    )
  })
)

sig_overlap_allomics_study
```
#----------------------------
#sensitivity analysis tables
#----------------------------
```{r}
compare_to_primary <- function(primary_out, sens_out, analysis_name){

  bind_rows(
    lapply(names(primary_out), function(cc){

      primary_ids <- primary_out[[cc]] %>%
        filter(adj.P.Val < 0.05) %>%
        pull(feature_id)

      sens_ids <- sens_out[[cc]] %>%
        filter(adj.P.Val < 0.05) %>%
        pull(feature_id)

      n_overlap <- length(intersect(primary_ids, sens_ids))

      tibble(
        Contrast = cc,
        `Primary\n(FDR<0.05)` = length(primary_ids),
        `Sensitivity\n(FDR<0.05)` = length(sens_ids),
        Overlap = n_overlap,
        `% overlap` = round(
          100 * n_overlap / length(primary_ids),
          1
        ),
        Analysis = analysis_name
      )
    })
  )
}
```


```{r}
tab_obesity <- compare_to_primary(
  primary_out = out_m3,
  sens_out = out_m5,
  analysis_name = "Remove obese T1D"
)

tab_noATT <- compare_to_primary(
  primary_out = out_m3,
  sens_out = out_m7,
  analysis_name = "Exclude ATTEMPT"
)

tab_allomics <- compare_to_primary(
  primary_out = out_m3,
  sens_out = out_m9,
  analysis_name = "Paired multi-omics subset"
)

sens_table <- bind_rows(
  tab_obesity,
  tab_noATT,
  tab_allomics
)
```


```{r}
pub_table <- full_join(
  tab_obesity %>%
    select(
      Contrast,
      Primary = `Primary\n(FDR<0.05)`,
      ObeseT1D = `Sensitivity\n(FDR<0.05)`,
      ObeseT1D_overlap = `% overlap`
    ),

  tab_noATT %>%
    select(
      Contrast,
      NoATT = `Sensitivity\n(FDR<0.05)`,
      NoATT_overlap = `% overlap`
    ),

  by = "Contrast"

) %>%
  full_join(

    tab_allomics %>%
      select(
        Contrast,
        AllOmics = `Sensitivity\n(FDR<0.05)`,
        AllOmics_overlap = `% overlap`
      ),

    by = "Contrast"
  )
```

```{r}
library(gt)

pub_table %>%
  gt() %>%
  tab_header(
    title = md("**Sensitivity Analyses Compared with the Primary Proteomic Analysis**")
  ) %>%
  cols_label(
    Contrast = "Contrast",
    Primary = "Primary",
    ObeseT1D = "Remove obese T1D",
    ObeseT1D_overlap = "% DEP overlap",
    NoATT = "Exclude ATTEMPT",
    NoATT_overlap = "% DEP overlap",
    AllOmics = "Paired multi-omics",
    AllOmics_overlap = "% DEP overlap"
  ) %>%
  tab_spanner(
    label = "Remove obese T1D",
    columns = c(ObeseT1D, ObeseT1D_overlap)
  ) %>%
  tab_spanner(
    label = "Exclude ATTEMPT",
    columns = c(NoATT, NoATT_overlap)
  ) %>%
  tab_spanner(
    label = "Paired multi-omics subset",
    columns = c(AllOmics, AllOmics_overlap)
  )
```
```{r}
gt_tab <- pub_table %>%
  gt() %>%
  tab_header(
    title = md("**Sensitivity Analyses Compared with the Primary Proteomic Analysis**")
  )

gtsave(
  gt_tab,
  filename = file.path(
    root_path,
    "Savanah Leidholt",
    "TRIAD",
    "figures",
    "Proteomics Figures",
    "Limma",
    "sensitivity_analysis_table.png"
  )
)
```
#--------------------
#T-stat correlations
#--------------------
```{r}
# Primary + sensitivity objects
primary_out <- out_m3

sensitivity_outs <- list(
  no_obese_T1D = out_m5,
  no_ATTEMPT   = out_m7,
  all_omics    = out_m9
)


#Function to compare t-statistics for one contrast
compare_t_stat <- function(primary_out, sens_out, sens_name, contrast_name) {
  
  df <- primary_out[[contrast_name]] %>%
    dplyr::select(feature_id, t_primary = t) %>%
    inner_join(
      sens_out[[contrast_name]] %>%
        dplyr::select(feature_id, t_sensitivity = t),
      by = "feature_id"
    ) %>%
    filter(
      is.finite(t_primary),
      is.finite(t_sensitivity)
    )
  
  rho <- cor(
    df$t_primary,
    df$t_sensitivity,
    method = "spearman",
    use = "complete.obs"
  )
  
  p <- ggplot(df, aes(x = t_primary, y = t_sensitivity)) +
    geom_point(alpha = 0.25, size = 1) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    labs(
      title = paste0(contrast_name, ": ", sens_name),
      subtitle = paste0("Spearman rho = ", round(rho, 3)),
      x = "Primary limma t-statistic",
      y = "Sensitivity limma t-statistic"
    ) +
    theme_bw()
  
  list(
    plot = p,
    summary = tibble(
      sensitivity = sens_name,
      contrast = contrast_name,
      n_features = nrow(df),
      spearman_t = rho
    )
  )
}
```

```{r}
contrast_order <- c(
  "T1D_vs_LC",
  "T2D_vs_LC",
  "T1D_vs_T2D",
  "OC_vs_LC",
  "T1D_vs_OC",
  "T2D_vs_OC"
)

tstat_results <- purrr::imap(
  sensitivity_outs,
  function(sens_out, sens_name) {
    purrr::map(
      contrast_order,
      ~ compare_t_stat(
        primary_out = primary_out,
        sens_out = sens_out,
        sens_name = sens_name,
        contrast_name = .x
      )
    )
  }
)
```

```{r}
tstat_summary <- purrr::map_dfr(
  tstat_results,
  ~ purrr::map_dfr(.x, "summary")
)

tstat_summary
```
```{r}
plots_no_obese_T1D <- purrr::map(tstat_results$no_obese_T1D, "plot")
plots_no_ATTEMPT   <- purrr::map(tstat_results$no_ATTEMPT, "plot")
plots_all_omics    <- purrr::map(tstat_results$all_omics, "plot")

panel_no_obese_T1D <- wrap_plots(plots_no_obese_T1D, ncol = 3) +
  plot_annotation(title = "Primary vs sensitivity: remove obese T1D")

panel_no_ATTEMPT <- wrap_plots(plots_no_ATTEMPT, ncol = 3) +
  plot_annotation(title = "Primary vs sensitivity: exclude ATTEMPT")

panel_all_omics <- wrap_plots(plots_all_omics, ncol = 3) +
  plot_annotation(title = "Primary vs sensitivity: paired multi-omics subset")

out_dir <- file.path(
  root_path,
  "Savanah Leidholt",
  "TRIAD",
  "figures",
  "Proteomics Figures",
  "Limma"
)
ggsave(
  file.path(out_dir, "tstat_correlation_no_obese_T1D.pdf"),
  panel_no_obese_T1D,
  width = 14,
  height = 8
)

ggsave(
  file.path(out_dir, "tstat_correlation_no_ATTEMPT.pdf"),
  panel_no_ATTEMPT,
  width = 14,
  height = 8
)

ggsave(
  file.path(out_dir, "tstat_correlation_all_omics.pdf"),
  panel_all_omics,
  width = 14,
  height = 8
)

write.csv(
  tstat_summary,
  file.path(dir.results, "Proteomics Results", "Limma", "tstat_correlation_summary.csv"),
  row.names = FALSE
)

panel_no_obese_T1D
panel_no_ATTEMPT
panel_all_omics
```
#---------------------------------------------------
#correlations with only top 100 overlapping proteins
#---------------------------------------------------
```{r}
compare_top100 <- function(primary_out,
                           sens_out,
                           sens_name,
                           contrast_name,
                           top_n = 500){

  primary_df <- primary_out[[contrast_name]]
  sens_df    <- sens_out[[contrast_name]]

  top_primary <- primary_df %>%
    arrange(desc(abs(t))) %>%
    slice_head(n = top_n) %>%
    select(feature_id, t_primary = t)

  top_sens <- sens_df %>%
    arrange(desc(abs(t))) %>%
    slice_head(n = top_n) %>%
    select(feature_id, t_sens = t)

  overlap_df <- inner_join(
    top_primary,
    top_sens,
    by = "feature_id"
  )

  n_overlap <- nrow(overlap_df)

  rho <- if(n_overlap > 1){
    cor(
      overlap_df$t_primary,
      overlap_df$t_sens,
      method = "spearman"
    )
  } else {
    NA
  }

  tibble(
    sensitivity = sens_name,
    contrast = contrast_name,
    top_n = top_n,
    n_overlap = n_overlap,
    pct_overlap = round(100*n_overlap/top_n,1),
    spearman_overlap_t = round(rho,3)
  )
}
```

```{r}
top100_summary <- imap_dfr(
  sensitivity_outs,
  function(sens_out, sens_name){

    map_dfr(
      contrast_order,
      ~compare_top100(
        primary_out = out_m3,
        sens_out = sens_out,
        sens_name = sens_name,
        contrast_name = .x,
        top_n = 500
      )
    )

  }
)

top100_summary
write.csv(
  top100_summary,
  file.path(dir.results, "Proteomics Results", "Limma", "top_overlap_summary.csv"),
  row.names = FALSE
)
```

```{r}
top100_summary %>%
  ggplot(
    aes(
      x = sensitivity,
      y = contrast,
      fill = pct_overlap
    )
  ) +
  geom_tile() +
  geom_text(
    aes(label = paste0(n_overlap, "/500")),
    size = 4
  ) +
  scale_fill_gradient(
    low = "grey",
    high = "purple4"
  ) +
  theme_bw()
```



#---------------------------------------------------
# saving limma outputs
# 4 analyses, all using age + sex + study model
#---------------------------------------------------
```{r}
prep_gsea_tables <- function(out_list) {
  lapply(out_list, function(tt) {
    tt %>%
      dplyr::select(feature_id,logFC,AveExpr,t,P.Value,adj.P.Val,B,contrast) %>%
      dplyr::arrange(desc(t))
  })
}

## prepare ranked tables
out_m3_gsea <- prep_gsea_tables(out_m3)
out_m7_gsea <- prep_gsea_tables(out_m7)
out_m9_gsea <- prep_gsea_tables(out_m9)
out_m5_gsea <- prep_gsea_tables(out_m5)

contrast_order <- c(
  "T1D_vs_LC",
  "T2D_vs_LC",
  "T1D_vs_T2D",
  "OC_vs_LC",
  "T1D_vs_OC",
  "T2D_vs_OC"
)

limma_gsea_inputs <- list(

  metadata = list(
    created = Sys.time(),
    response = "plasma proteomics",
    transform = "log2(x + 1)",
    model_formula = "0 + group4 + age + sex + study6",
    contrasts = contrast_order
  ),

  primary = list(
    model_name = "primary_age_sex_study",
    samples = meta_m3,
    design = design_m3,
    fit = fit_m3,
    out = out_m3,
    gsea_tables = out_m3_gsea
  ),
  
  noObeseT1D = list(
    model_name = "sensitivty_age_sex_study_noObeseT1D",
    samples = meta_m5,
    design = design_m5,
    fit = fit_m5,
    out = out_m5,
    gsea_tables = out_m5_gsea
  ),

  no_ATTEMPT = list(
    model_name = "sensitivity_age_sex_study_noATT",
    samples = meta_m7,
    design = design_m7,
    fit = fit_m7,
    out = out_m7,
    gsea_tables = out_m7_gsea
  ),

  all_omics = list(
    model_name = "sensitivity_age_sex_study_allomics",
    samples = meta_m8, # this might seem wrong but it's not I want to use meta8 not meta9
    design = design_m9,
    fit = fit_m9,
    out = out_m9,
    gsea_tables = out_m9_gsea
  )
)

saveRDS(
  limma_gsea_inputs,
  file = file.path(
    root_path,
    "Savanah Leidholt",
    "TRIAD",
    "Data",
    "limma_gsea_inputs_age_sex_study.rds"
  )
)

cat("Saved: limma_gsea_inputs_age_sex_study.rds\n")
```

#---------------------------------------------------
# Three analyses with reduced model
#---------------------------------------------------
```{r}
out_m1_gsea <- prep_gsea_tables(out_m1)
out_m6_gsea <- prep_gsea_tables(out_m6)
out_m8_gsea <- prep_gsea_tables(out_m8)


limma_gsea_inputs_m1 <- list(

  metadata = list(
    created = Sys.time(),
    response = "plasma proteomics",
    transform = "log2(x + 1)",
    model_formula = "0 + group4 + age + sex",
    contrasts = contrast_order
  ),

  primary = list(
    model_name = "primary_age_sex",
    samples = meta_m1,
    design = design_m1,
    fit = fit_m1,
    out = out_m1,
    gsea_tables = out_m1_gsea
  ),

  no_ATTEMPT = list(
    model_name = "sensitivity_age_sex_noATT",
    samples = meta_m6,
    design = design_m6,
    fit = fit_m6,
    out = out_m6,
    gsea_tables = out_m6_gsea
  ),

  all_omics = list(
    model_name = "sensitivity_age_sex_allomics",
    samples = meta_m8,
    design = design_m8,
    fit = fit_m8,
    out = out_m8,
    gsea_tables = out_m8_gsea
  )
)

saveRDS(
  limma_gsea_inputs_m1,
  file = file.path(
    root_path,
    "Savanah Leidholt",
    "TRIAD",
    "Data",
    "limma_gsea_inputs_age_sex.rds"
  )
)

cat("Saved: limma_gsea_inputs_age_sex.rds\n")

```

