---
title: "Plasma_DEP_anlaysis"
author: "Savanah Leidholt"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ---- packages ----
library(dplyr)
library(tidyr)
library(limma)
library(EnhancedVolcano)
library(ggplot2)
library(patchwork)
library(ggrepel)
library(vegan)
library(tidyverse)
library(data.table)
library(purrr)
library(haven)
library(ggVennDiagram)
library(caret)
```

```{r}
#Directories
#computer <- "mac studio"
computer <- "mac laptop"

user <- Sys.info()[["user"]]

if (computer == "mac studio") {
  
  if (user == "sleidholt") {
    dir.dat <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
    
  } else if (user == "hhampson") {
    dir.dat <- "/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results"
    git_path <- "/Users/hhampson/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
    
  } else if (user == "choiyej") {
    dir.dat <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "pylell") {
    dir.dat <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  }

} else if (computer == "mac laptop") {

  if (user == "savanahleidholt") {
    dir.dat <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "hhampson") {
    dir.dat <- "/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive"
    dir.results <- "/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/PANTHER/Results"
    git_path <- "/Users/hhampson/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
  }
}
```
#-----------------------------------------------------------------------------
#Clean and setup data
# a) all proteomic data DPE analysis
# b) sensitivty analysis (rm attempt)
# c) DPE analysis with samples that contain both proteomic and metabolomic data
#------------------------------------------------------------------------------
```{r}
obj <- readRDS(file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "omics_split_objects3.rds"))

meta   <- obj$meta_prot
prot_plasma <- obj$prot_plasma
n_table     <- obj$n_table

# quick check
n_table
dim(prot_plasma)
dim(meta)


```
##---------------------------------------------
##open panther_data.RData to add in hba1c data
##---------------------------------------------
```{r}
load(file.path(dir.dat, "PANTHER", "Data_cleaned", "panther_data.RData"))


#trasnforming data to match the harmonized data
panther_hba1c <- panther %>%
  dplyr::select(record_id, hba1c) %>%
  transmute(
    record_id = gsub("_", "-", as.character(record_id)),
    hba1c_panther = as.numeric(hba1c)
  )

panther_hba1c$record_id <- as.character(panther_hba1c$record_id)

#Make sure harm$hba1c is numeric too (if it exists)
meta <- meta %>%
  mutate(hba1c = as.numeric(zap_labels(hba1c))) %>%
  left_join(panther_hba1c, by = "record_id") %>%
  mutate(
    hba1c = if_else(
      study == "PANTHER" & is.na(hba1c),
      hba1c_panther,
      hba1c
    )
  ) %>%
  dplyr::select(-hba1c_panther)
```
##---------------------
##cleaning metadata
##---------------------
```{r}
meta$record_id <- trimws(as.character(meta$record_id))
meta$study     <- trimws(as.character(meta$study))
meta$group     <- trimws(as.character(meta$group))
meta$sex       <- trimws(as.character(meta$sex))
meta$ethnicity <- trimws(as.character(meta$race_ethnicity_condensed))

meta$age <- suppressWarnings(as.numeric(meta$age))

# drop missing covariates 
keep <- !is.na(meta$age) & !is.na(meta$sex) & meta$sex != "" 
meta <- meta[keep, , drop = FALSE]

meta$group4 <- factor(
  meta$group,
  levels = c("Lean Control","Obese Control","Type 1 Diabetes","Type 2 Diabetes"),
  labels = c("LC","OC","T1D","T2D")
)

meta$ethnicity <- factor(meta$ethnicity)
meta$study6 <- dplyr::recode(
  meta$study,
  "ATTEMPT"        = "ATT",
  "CROCODILE"      = "CRC",
  "PANTHER"        = "PAN",
  "RENAL-HEIR"     = "RH",
  "IMPROVE"        = "IT",
  "RENAL-HEIRitage"= "RH2",
  .default = meta$study
)
meta$study6 <- factor(meta$study6)

meta_common <- meta
meta_common <- meta_common %>%
  mutate(
    uacr = suppressWarnings(as.numeric(acr_u)),
    log_uacr = if_else(is.na(uacr), NA_real_, log(uacr + 1))
  )

meta_common$record_id <- trimws(as.character(meta_common$record_id))
prot_ids <- trimws(rownames(prot_plasma))

cat("meta_common n:", nrow(meta_common), "\n")
cat("prot_plasma dim:", dim(prot_plasma), "\n")

cat("example meta IDs:", paste(head(meta_common$record_id), collapse=", "), "\n")
cat("example prot rownames:", paste(head(prot_ids), collapse=", "), "\n")

# overlap
olap <- intersect(meta_common$record_id, prot_ids)
cat("overlap n:", length(olap), "\n")
cat("meta not in prot:", sum(!meta_common$record_id %in% prot_ids), "\n")
cat("prot not in meta:", sum(!prot_ids %in% meta_common$record_id), "\n")
meta_common$record_id <- trimws(as.character(meta_common$record_id))
rownames(prot_plasma) <- trimws(rownames(prot_plasma))

olap <- intersect(meta_common$record_id, rownames(prot_plasma))
cat("Overlap samples (meta vs plasma):", length(olap), "\n")
stopifnot(length(olap) > 0)

meta_common <- meta_common[match(olap, meta_common$record_id), , drop = FALSE]
prot_plasma_common <- prot_plasma[olap, , drop = FALSE]
prot_plasma_common
stopifnot(identical(rownames(prot_plasma_common), meta_common$record_id))


##checking for low variance prots--none exist
nzv_idx <- nearZeroVar(prot_plasma_common)

length(nzv_idx)   
colnames(prot_plasma_common)[nzv_idx]  
```

##--------------------------------------------------------------
## Protein Filter and Log Transformation
##---------------------------------------------------------------
```{r}
##filter to proteins that are only in the analyte panel (~7000)
##check if this is correct for the newer somascan prot pannel
##should filter ~3000 all from ATTEMPT
##then filter to keep proteins that are in at least 2 studies
analytes <- read.csv(
  file.path(dir.dat, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

prot_panel <- unique(trimws(as.character(analytes$AptName)))

##keep only aptamers that are present in the protein matrix
aptamers_keep <- intersect(prot_panel, colnames(prot_plasma_common))

##subset protein matrix to those aptamers
prot_plasma_common2 <- prot_plasma_common[, aptamers_keep, drop = FALSE]

dim(prot_plasma_common2) #should be ~7000
dim(prot_plasma_common) #should be ~10000
length(aptamers_keep)
length(setdiff(prot_panel, colnames(prot_plasma_common)))

##log transform protein data
prot_pca <- as.matrix(prot_plasma_common2)
prot_log <- log2(prot_pca + 1)
prot_log[!is.finite(prot_log)] <- NA_real_

##checking to see if there are any proteins with low to near zero variance
nzv_idx <- caret::nearZeroVar(prot_plasma_common2)
length(nzv_idx)   #comes out as zero
```

```{r}
#not really necessary to do as all protiens pass filter once I remove the extra 3000 from the ATTEMPT Panel
##filtering step where protein must exist in 60% of samples in a study
#study_idx_all <- split(seq_len(nrow(prot_log)), meta_common$study)
#present_by_study_all <- sapply(study_idx_all, function(idx) {
#  colMeans(!is.na(prot_log[idx, ])) >= 0.6
#})

#present_by_study_all <- as.matrix(present_by_study_all)
#n_studies_present_all <- colSums(present_by_study_all)
#keep_study_all <- n_studies_present_all >= 2

#removed_proteins <- colnames(prot_log)[!keep_study_all]
#kept_proteins    <- colnames(prot_log)[keep_study_all]

#cat("All proteomic features before study filter:", ncol(prot_log), "\n")
#cat("Removed (<2 studies meeting 60% presence threshold):", length(removed_proteins), "\n")
#cat("Kept (>=2 studies meeting 60% presence threshold):", length(kept_proteins), "\n")
```

```{r}
set.seed(18)
prot_imp <- prot_log
meds <- apply(prot_imp, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(prot_imp))) {
  if (!is.na(meds[j])) {
    prot_imp[is.na(prot_imp[, j]), j] <- meds[j]
  }
}

keep_cols <- colSums(is.na(prot_imp)) == 0 &
  apply(prot_imp, 2, var) > 0 &
  is.finite(apply(prot_imp, 2, var))

prot_imp <- prot_imp[, keep_cols, drop = FALSE]

```

##--------------------------------
## Find and Remove Outliers in PCA
##---------------------------------
```{r}
pca0 <- prcomp(prot_imp, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2])  # PC1 + PC2 (or use more PCs)

scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

# identify outliers > 3 SD
threshold <- mean(scores$dist) + 3 * sd(scores$dist)
outliers <- rownames(scores)[scores$dist > threshold]
outliers

meta_common2<- meta_common %>%
  filter(!record_id %in% outliers)

prot_imp2 <- prot_imp[
  !rownames(prot_imp) %in% outliers, , drop = FALSE
]

cat("Still present in metadata? ", any(outliers %in% meta_common2$record_id), "\n")
cat("Still present in matrix? ", any(outliers %in% rownames(prot_imp2)), "\n")

stopifnot(!any(outliers %in% meta_common2$record_id))
stopifnot(!any(outliers %in% rownames(prot_imp2)))
stopifnot(identical(rownames(prot_imp2), meta_common2$record_id))


##checking for low variance prots--none exist
nzv_idx2 <- nearZeroVar(prot_imp2)
length(nzv_idx2)   
colnames(prot_imp2)[nzv_idx2]


write.csv(
  prot_imp2,
  file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "prot_imp")
)
```

#---------------
# Plotting Funciton
#---------------
```{r}
make_pca_plot <- function(prot_mat, meta_df, title_str) {
  pca  <- prcomp(prot_mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1    = pca$x[, 1],
    PC2    = pca$x[, 2],
    group4 = meta_df$group4,
    study  = meta_df$study6
  ) %>%
    ggplot(aes(PC1, PC2, color = group4, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        LC  = "Cyan4",
        OC  = "Gold",
        T1D = "DarkRed",
        T2D = "Purple"
      ),
      name = "Group"
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
##-----------
## PCAs
##-----------
```{r}
p  <- make_pca_plot(prot_imp2, meta_common2, "No adjustment")

cov_m1 <- model.matrix(~ age + sex, data = meta_common2)[, -1, drop = FALSE]
prot_m1 <- t(removeBatchEffect(t(prot_imp2), covariates = cov_m1))
p1 <- make_pca_plot(prot_m1, meta_common2, "Age + Sex")


cov_m2 <- model.matrix(~ age + sex + study6, data = meta_common2)[, -1, drop = FALSE]
prot_m2 <- t(removeBatchEffect(t(prot_imp2), covariates = cov_m2))
p2 <- make_pca_plot(prot_m2, meta_common2, "Age + Sex + Study")

(p | p1 | p2)
```



##------------------------------
##limma testing all prots data
## model1: age + sex
## model2: age + sex + study
##------------------------------
```{r}
meta_common2$sex    <- factor(meta_common2$sex)
meta_common2$group4 <- factor(meta_common2$group4)

# helper function-- run limma + return list of topTables
run_limma_contrasts <- function(Y, design, contr) {
  fit <- lmFit(Y, design)
  fit2 <- contrasts.fit(fit, contr)
  fit2 <- eBayes(fit2)

  out <- setNames(vector("list", ncol(contr)), colnames(contr))
  for (cc in colnames(contr)) {
    tt <- topTable(fit2, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
    tt$feature_id <- rownames(tt)
    tt$contrast <- cc
    out[[cc]] <- tt
  }
  list(fit = fit2, out = out)
}

# ---- helper: summarise sig hits per contrast ----
summarise_sig <- function(out_list, model_label, n_samp) {
  tibble(
    contrast = names(out_list),
    n_FDR05  = sapply(out_list, \(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
    n_FDR01  = sapply(out_list, \(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
    model    = model_label,
    n_samples = n_samp
  )
}

# response matrix (proteins x samples) base
Y_m0 <- t(prot_imp2)
```

```{r}
# MODEL 1: age + sex 
meta_m1 <- meta_common2
Y_m1 <- Y_m0[, meta_m1$record_id, drop = FALSE]  

design <- model.matrix(~ 0 + group4 + age + sex, data = meta_m1)
colnames(design) <- sub("^group4", "", colnames(design))
colnames(design) <- make.names(colnames(design))

contr_m1 <- makeContrasts(
  T1D_vs_LC   = T1D - LC,
  T2D_vs_LC   = T2D - LC,
  T1D_vs_T2D  = T1D - T2D,
  OC_vs_LC    = OC  - LC,
  T1D_vs_OC   = T1D - OC,
  T2D_vs_OC   = T2D - OC,
  levels = design
)

res_m1 <- run_limma_contrasts(Y_m1, design, contr_m1)
fit_m1 <- res_m1$fit
out_m1  <- res_m1$out

sig_m1 <- tibble(
  contrast = colnames(contr_m1),
  n_FDR05 = sapply(out_m1, \(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_m1, \(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE))
) %>% mutate(model = "age + sex", n_samples = nrow(meta_m1))

sig_m1


# MODEL 2: age + sex + study
meta_m2 <- meta_common2 %>%
  filter(!is.na(study6))       

Y_m2 <- Y_m0[, meta_m2$record_id, drop = FALSE]

design_m2 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m2)
colnames(design_m2) <- sub("^group4", "", colnames(design_m2))
colnames(design_m2) <- make.names(colnames(design_m2))

contr_m2 <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC  - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_m2
)

res_m2  <- run_limma_contrasts(Y_m2, design_m2, contr_m2)
out_m2  <- res_m2$out
fit_m2  <- res_m2$fit
sig_m2  <- summarise_sig(out_m2, "age_sex_study", nrow(meta_m2))

cat("Model 2 (age+sex+study) complete. N =", nrow(meta_m2), "\n")
sig_m2
```
##-------------------------------------
## combined summary tables (model 1 & 2)
##-------------------------------------
```{r}
summary_all <- bind_rows(
  sig_m1  %>% mutate(model = "age_sex"),
  sig_m2
) %>%
  dplyr::select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

print(summary_all)

summary_wide <- summary_all %>%
  dplyr::select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

print(summary_wide)
```

#--------------------------------------------------------
# SENSITIVITY ANALYSIS:
# Exclude ATTEMPT study (different sequencing platform)
#--------------------------------------------------------
```{r}
# Subset metadata + protein matrix
meta_noATT <- meta_common %>%
  filter(study != "ATTEMPT")

prot_noATT <- prot_plasma_common[meta_noATT$record_id, , drop = FALSE]
stopifnot(identical(rownames(prot_noATT), meta_noATT$record_id))

cat("Original sample size:", nrow(meta_common), "\n")
cat("Sample size excluding ATTEMPT:", nrow(meta_noATT), "\n")
cat("ATTEMPT samples removed:", nrow(meta_common) - nrow(meta_noATT), "\n")

cat("Studies remaining:\n")
print(table(meta_noATT$study, useNA = "ifany"))
```
##--------------------------------------------------
## PCA using no-ATTEMPT dataset + transform + imputing
##--------------------------------------------------
```{r}
meta_pca_noATT <- meta_noATT
prot_pca_noATT <- prot_noATT[meta_pca_noATT$record_id, , drop = FALSE]
stopifnot(identical(rownames(prot_pca_noATT), meta_pca_noATT$record_id))

prot_pca_noATT <- as.matrix(prot_pca_noATT)

# log2 transform + median impute
prot_log_noATT <- log2(prot_pca_noATT + 1)
prot_log_noATT[!is.finite(prot_log_noATT)] <- NA_real_

prot_imp_noATT <- prot_log_noATT
meds_noATT <- apply(prot_imp_noATT, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(prot_imp_noATT))) {
  if (!is.na(meds_noATT[j])) prot_imp_noATT[is.na(prot_imp_noATT[, j]), j] <- meds_noATT[j]
}

keep_cols_noATT <- colSums(is.na(prot_imp_noATT)) == 0 &
  apply(prot_imp_noATT, 2, var) > 0 &
  is.finite(apply(prot_imp_noATT, 2, var))

prot_imp_noATT <- prot_imp_noATT[, keep_cols_noATT, drop = FALSE]

cat("Proteins kept for no-ATTEMPT PCA:", ncol(prot_imp_noATT), "\n")
```
##---------------------------
##outlier check and removal
##---------------------------
```{r}
pca0_noATT <- prcomp(prot_imp_noATT, center = TRUE, scale. = TRUE)
scores2 <- as.data.frame(pca0_noATT$x[, 1:2])  # PC1 + PC2 (or use more PCs)
scores2$dist <- sqrt(scores2$PC1^2 + scores2$PC2^2)

# identify outliers > 3 SD
threshold2 <- mean(scores2$dist) + 3 * sd(scores2$dist)
outliers_noATT <- rownames(scores2)[scores2$dist > threshold2]
outliers_noATT

meta_pca_noATT2<- meta_pca_noATT %>%
  filter(!record_id %in% outliers_noATT)

prot_imp_noATT2 <- prot_imp_noATT[
  !rownames(prot_imp_noATT) %in% outliers_noATT, , drop = FALSE
]

cat("Still present in metadata? ", any(outliers_noATT %in% meta_pca_noATT2$record_id), "\n")
cat("Still present in matrix? ", any(outliers_noATT %in% rownames(prot_imp_noATT2)), "\n")

stopifnot(!any(outliers_noATT %in% meta_pca_noATT2$record_id))
stopifnot(!any(outliers_noATT %in% rownames(prot_imp_noATT2)))
stopifnot(identical(rownames(prot_imp_noATT2), meta_pca_noATT2$record_id))

##checking for low variance prots--none exist
nzv_idx2 <- nearZeroVar(prot_imp_noATT)
length(nzv_idx2)   
colnames(prot_imp_noATT)[nzv_idx2]
```
##--------------
##PCA no ATTEMPt
##--------------
```{r}
p_noATT <- make_pca_plot(prot_imp_noATT2, meta_pca_noATT2, "No adjustment (excluding ATTEMPT)")

cov_m1_noATT <- model.matrix(~ age + sex, data = meta_pca_noATT2)[, -1, drop = FALSE]
prot_m1_noATT <- t(removeBatchEffect(t(prot_imp_noATT2), covariates = cov_m1_noATT))
p1_noATT <- make_pca_plot(prot_m1_noATT, meta_pca_noATT2, "Age + Sex (excluding ATTEMPT)")

keep_m2_noATT <- !is.na(meta_pca_noATT2$study6)
meta_m2p_noATT <- meta_pca_noATT2[keep_m2_noATT, ]
prot_m2p_noATT <- prot_imp_noATT2[keep_m2_noATT, , drop = FALSE]
cov_m2_noATT <- model.matrix(~ age + sex + study6, data = meta_m2p_noATT)[, -1, drop = FALSE]
prot_m2_adj_noATT <- t(removeBatchEffect(t(prot_m2p_noATT), covariates = cov_m2_noATT))
p2_noATT <- make_pca_plot(prot_m2_adj_noATT, meta_m2p_noATT, "Age + Sex + Study (excluding ATTEMPT)")

# Render no-ATTEMPT PCA panel
(p_noATT | p1_noATT | p2_noATT) 
  plot_annotation(
    title = "PCA: Sensitivity Analysis Excluding ATTEMPT",
    subtitle = "Color = group | Shape = study | ATTEMPT removed due to platform differences",
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 9, hjust = 0.5, color = "grey40")
    )
  )
```
##----------------------------------
## limma using no-ATTEMPT dataset
## model 1: age + sex
## model 2: age + sex + study
## ---------------------------------
```{r}
Y_all_noATT <- t(prot_imp_noATT2)


# Model 1: age + sex

meta_m1_noATT <- meta_pca_noATT2
Y_m1_noATT <- Y_all_noATT[, meta_m1_noATT$record_id, drop = FALSE]

design_m1_noATT <- model.matrix(~ 0 + group4 + age + sex, data = meta_m1_noATT)
colnames(design_m1_noATT) <- sub("^group4", "", colnames(design_m1_noATT))
colnames(design_m1_noATT) <- make.names(colnames(design_m1_noATT))

contr_m1_noATT <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_m1_noATT
)

res_m1_noATT <- run_limma_contrasts(Y_m1_noATT, design_m1_noATT, contr_m1_noATT)
fit_m1_noATT <- res_m1_noATT$fit
out_m1_noATT <- res_m1_noATT$out

sig_m1_noATT <- tibble(
  contrast = colnames(contr_m1_noATT),
  n_FDR05 = sapply(out_m1_noATT, \(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_m1_noATT, \(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE))
) %>% mutate(model = "age_sex_noATT", n_samples = nrow(meta_m1_noATT))


# Model 2: age + sex + study

meta_m2_noATT <- meta_pca_noATT2 %>% filter(!is.na(study6))
Y_m2_noATT <- Y_all_noATT[, meta_m2_noATT$record_id, drop = FALSE]

design_m2_noATT <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m2_noATT)
colnames(design_m2_noATT) <- sub("^group4", "", colnames(design_m2_noATT))
colnames(design_m2_noATT) <- make.names(colnames(design_m2_noATT))

contr_m2_noATT <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_m2_noATT
)

res_m2_noATT <- run_limma_contrasts(Y_m2_noATT, design_m2_noATT, contr_m2_noATT)
fit_m2_noATT <- res_m2_noATT$fit
out_m2_noATT <- res_m2_noATT$out
sig_m2_noATT <- summarise_sig(out_m2_noATT, "age_sex_study_noATT", nrow(meta_m2_noATT))

```
## -----------------------------
## Summarize no-ATTEMPT results
## -----------------------------
```{r}
summary_all_noATT <- bind_rows(
  sig_m1_noATT,
  sig_m2_noATT
) %>%
  dplyr::select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

print(summary_all_noATT)

summary_wide_noATT <- summary_all_noATT %>%
  dplyr::select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

print(summary_wide_noATT)


# Compare original vs no-ATTEMPT summary tables
summary_compare_noATT <- bind_rows(
  summary_all %>% mutate(dataset = "original"),
  summary_all_noATT %>% mutate(dataset = "excluding_ATTEMPT")
) %>%
  arrange(contrast, dataset, model)

print(summary_compare_noATT)
```

#------------------------------------------------------------------------
# Sensitivity Analysis #2:
# Restrict proteomics workflow to participants with all 3 omics types
# (metabolomics + lipidomics + proteomics) = (metabolomics + proteomics)
#------------------------------------------------------------------------
```{r}
# Identify participants with all three omics types
omics_objects <- readRDS(file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "omics_split_objects3.rds"))


# HELPER: count non-NA samples in a matrix for a given metadata
count_samples_with_data <- function(meta_df, mat_df) {
  # A sample "has data" if it has at least one non-NA value in the omics matrix
  has_data <- rowSums(!is.na(mat_df)) > 0
  meta_df %>%
    mutate(has_omics = has_data) %>%
    filter(has_omics)
}
```

##---------------------------------------
## samples counts Study x Group
##---------------------------------------
```{r}

make_count_table <- function(meta_df, mat_df, omics_label) {
  count_samples_with_data(meta_df, mat_df) %>%
    count(study, group) %>%
    pivot_wider(names_from = group, values_from = n, values_fill = 0) %>%
    arrange(study) %>%
    mutate(Omics = omics_label, .before = 1)
}

tbl_metabolites <- make_count_table(omics_objects$meta_metab, omics_objects$metabolites,  "Metabolomics")
tbl_lipids      <- make_count_table(omics_objects$meta_lips,  omics_objects$lipids,        "Lipidomics")
tbl_proteins    <- make_count_table(omics_objects$meta_prot,  omics_objects$prot_plasma,   "Proteomics (Plasma)")

cat("\n===== METABOLOMICS: Samples per Study x Group =====\n")
print(tbl_metabolites)

cat("\n===== LIPIDOMICS: Samples per Study x Group =====\n")
print(tbl_lipids)

cat("\n===== PROTEOMICS (Plasma): Samples per Study x Group =====\n")
print(tbl_proteins)
```

```{r}
#A sample "has data" if it has at least one non-missing value in that matrix
ids_metab <- rownames(omics_objects$metabolites)[rowSums(!is.na(omics_objects$metabolites)) > 0]
ids_lipids <- rownames(omics_objects$lipids)[rowSums(!is.na(omics_objects$lipids)) > 0]
ids_prot <- rownames(omics_objects$prot_plasma)[rowSums(!is.na(omics_objects$prot_plasma)) > 0]

ids_metab <- trimws(as.character(ids_metab))
ids_lipids <- trimws(as.character(ids_lipids))
ids_prot <- trimws(as.character(ids_prot))

ids_all_three <- Reduce(intersect, list(ids_metab, ids_lipids, ids_prot))

cat("Total participants with all 3 omics:", length(ids_all_three), "\n")
```

##--------------------------------------------------
## Restrict the proteomics analysis-ready dataset
##--------------------------------------------------
```{r}
meta_all3 <- meta_common %>%
  mutate(record_id = trimws(as.character(record_id))) %>%
  filter(record_id %in% ids_all_three)

prot_all3 <- prot_plasma_common[meta_all3$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_all3), meta_all3$record_id))

cat("Original proteomics analysis N:", nrow(meta_common), "\n")
cat("Proteomics analysis N with all 3 omics:", nrow(meta_all3), "\n")
cat("Excluded because not all 3 omics:",
    nrow(meta_common) - nrow(meta_all3), "\n")

cat("Group counts in all-3-omics subset:\n")
print(table(meta_all3$group4, useNA = "ifany"))

cat("Study counts in all-3-omics subset:\n")
print(table(meta_all3$study, useNA = "ifany"))
```
##--------------------------------------------------
##data cleaning for data subset with all 3 Omics
##--------------------------------------------------
```{r}
meta_pca_all3 <- meta_all3
prot_pca_all3 <- prot_all3[meta_pca_all3$record_id, , drop = FALSE]
stopifnot(identical(rownames(prot_pca_all3), meta_pca_all3$record_id))

prot_pca_all3 <- as.matrix(prot_pca_all3)

# ---- log2 transform + median impute ----
prot_log_all3 <- log2(prot_pca_all3 + 1)
prot_log_all3[!is.finite(prot_log_all3)] <- NA_real_

prot_imp_all3 <- prot_log_all3
meds_all3 <- apply(prot_imp_all3, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(prot_imp_all3))) {
  if (!is.na(meds_all3[j])) prot_imp_all3[is.na(prot_imp_all3[, j]), j] <- meds_all3[j]
}

keep_cols_all3 <- colSums(is.na(prot_imp_all3)) == 0 &
  apply(prot_imp_all3, 2, var) > 0 &
  is.finite(apply(prot_imp_all3, 2, var))

prot_imp_all3 <- prot_imp_all3[, keep_cols_all3, drop = FALSE]

cat("Proteins kept for all-3-omics PCA:", ncol(prot_imp_all3), "\n")
```
##---------------------------
##outlier check and removal
##---------------------------
```{r}
pca0_all3 <- prcomp(prot_imp_all3, center = TRUE, scale. = TRUE)
scores3 <- as.data.frame(pca0_all3$x[, 1:2])  # PC1 + PC2 (or use more PCs)
scores3$dist <- sqrt(scores3$PC1^2 + scores3$PC2^2)

# identify outliers > 3 SD
threshold3 <- mean(scores3$dist) + 3 * sd(scores3$dist)
outliers_all3 <- rownames(scores3)[scores3$dist > threshold3]
outliers_all3

meta_pca_all3b<- meta_pca_all3 %>%
  filter(!record_id %in% outliers_all3)

prot_imp_all3b <- prot_imp_all3[
  !rownames(prot_imp_all3) %in% outliers_all3, , drop = FALSE
]

cat("Still present in metadata? ", any(outliers_all3 %in% meta_pca_all3b$record_id), "\n")
cat("Still present in matrix? ", any(outliers_all3 %in% rownames(prot_imp_all3b)), "\n")

stopifnot(!any(outliers_all3 %in% meta_pca_all3b$record_id))
stopifnot(!any(outliers_all3 %in% rownames(prot_imp_all3b)))
stopifnot(identical(rownames(prot_imp_all3b), meta_pca_all3b$record_id))

##checking for low variance prots--none exist
nzv_idx2 <- nearZeroVar(prot_imp_all3b)
length(nzv_idx2)   
colnames(prot_imp_all3b)[nzv_idx2]
```

```{r}
# ---- PCA panels ----
p_all3 <- make_pca_plot(prot_imp_all3b, meta_pca_all3b, "No adjustment (all 3 omics subset)")

cov_m1_all3 <- model.matrix(~ age + sex, data = meta_pca_all3b)[, -1, drop = FALSE]
prot_m1_all3 <- t(removeBatchEffect(t(prot_imp_all3b), covariates = cov_m1_all3))
p1_all3 <- make_pca_plot(prot_m1_all3, meta_pca_all3b, "Age + Sex (all 3 omics subset)")

keep_m2_all3 <- !is.na(meta_pca_all3b$bmi)
meta_m2p_all3 <- meta_pca_all3b[keep_m2_all3, ]
prot_m2p_all3 <- prot_imp_all3b[keep_m2_all3, , drop = FALSE]
cov_m2_all3 <- model.matrix(~ age + sex + study6, data = meta_m2p_all3)[, -1, drop = FALSE]
prot_m2_adj_all3 <- t(removeBatchEffect(t(prot_m2p_all3), covariates = cov_m2_all3))
p2_all3 <- make_pca_plot(prot_m2_adj_all3, meta_m2p_all3, "Age + Sex + Study (all 3 omics subset)")

(p_all3 | p1_all3 | p2_all3)+
  plot_annotation(
    title = "PCA: Proteomics Sensitivity Analysis in Participants with All 3 Omics Types",
    subtitle = "Color = group | Shape = study | Restricted to shared metabolomics + lipidomics + proteomics participants",
    theme = theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 9, hjust = 0.5, color = "grey40")
    )
  )
```
##--------------------------------------------------
##limma using all-3-omics subset
## model1: age + sex
## model2: age + sex + study
##--------------------------------------------------
```{r}
Y_all_all3 <- t(prot_imp_all3b)


# Model 1: age + sex
meta_m1_all3 <- meta_pca_all3b
Y_m1_all3 <- Y_all_all3[, meta_m1_all3$record_id, drop = FALSE]

design_m1_all3 <- model.matrix(~ 0 + group4 + age + sex, data = meta_m1_all3)
colnames(design_m1_all3) <- sub("^group4", "", colnames(design_m1_all3))
colnames(design_m1_all3) <- make.names(colnames(design_m1_all3))

contr_m1_all3 <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_m1_all3
)

res_m1_all3 <- run_limma_contrasts(Y_m1_all3, design_m1_all3, contr_m1_all3)
fit_m1_all3 <- res_m1_all3$fit
out_m1_all3 <- res_m1_all3$out

sig_m1_all3 <- tibble(
  contrast = colnames(contr_m1_all3),
  n_FDR05 = sapply(out_m1_all3, \(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_m1_all3, \(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE))
) %>% mutate(model = "age_sex_all3", n_samples = nrow(meta_m1_all3))



# Model 2: age + sex + study
Y_m2_all3 <- Y_all_all3[, meta_m1_all3$record_id, drop = FALSE]

design_m2_all3 <- model.matrix(~ 0 + group4 + age + sex + study6, data = meta_m1_all3)
colnames(design_m2_all3) <- sub("^group4", "", colnames(design_m2_all3))
colnames(design_m2_all3) <- make.names(colnames(design_m2_all3))

contr_m2_all3 <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_m2_all3
)

res_m2_all3 <- run_limma_contrasts(Y_m2_all3, design_m2_all3, contr_m2_all3)
fit_m2_all3 <- res_m2_all3$fit
out_m2_all3 <- res_m2_all3$out
sig_m2_all3 <- summarise_sig(out_m2_all3, "age_sex_study_all3", nrow(meta_m1_all3))
sig_m2_all3
```
## --------------------------------------------------
## Summarize all-3-omics results
##--------------------------------------------------
```{r}
summary_all_all3 <- bind_rows(
  sig_m1_all3,
  sig_m2_all3
) %>%
  dplyr::select(model, n_samples, contrast, n_FDR05, n_FDR01) %>%
  arrange(contrast, model)

print(summary_all_all3)

summary_wide_all3 <- summary_all_all3 %>%
  dplyr::select(model, contrast, n_FDR05) %>%
  pivot_wider(names_from = model, values_from = n_FDR05)

print(summary_wide_all3)

summary_compare_all3 <- bind_rows(
  summary_all %>% mutate(dataset = "original"),
  summary_all_all3 %>% mutate(dataset = "all_3_omics_subset")
) %>%
  arrange(contrast, dataset, model)

print(summary_compare_all3)

```

```{r}
FDR_THRESH <- 0.05

contrasts <- c("T1D_vs_LC", "T2D_vs_LC", "T1D_vs_T2D",
               "OC_vs_LC",  "T1D_vs_OC", "T2D_vs_OC")

get_sig <- function(out_list, contrast, fdr = FDR_THRESH) {
  tt <- out_list[[contrast]]
  if (is.null(tt)) return(character(0))
  tt %>%
    dplyr::filter(!is.na(adj.P.Val), adj.P.Val < fdr) %>%
    dplyr::pull(feature_id) %>%
    unique()
}

overlap_results <- lapply(contrasts, function(cc) {
  
  all_studies <- get_sig(out_m2, cc)
  noATT       <- get_sig(out_m2_noATT, cc)
  all3        <- get_sig(out_m2_all3, cc)
  
  tibble::tibble(
    contrast = cc,
    
    n_all_studies = length(all_studies),
    n_noATT       = length(noATT),
    n_all3        = length(all3),
    
    overlap_all_vs_noATT = length(intersect(all_studies, noATT)),
    overlap_all_vs_all3  = length(intersect(all_studies, all3)),
    overlap_noATT_vs_all3 = length(intersect(noATT, all3)),
    
    overlap_all_three = length(Reduce(intersect, list(all_studies, noATT, all3))),
    
    union_all_vs_noATT = length(union(all_studies, noATT)),
    union_all_vs_all3  = length(union(all_studies, all3)),
    union_noATT_vs_all3 = length(union(noATT, all3)),
    union_all_three    = length(Reduce(union, list(all_studies, noATT, all3))),
    
    jaccard_all_vs_noATT = ifelse(length(union(all_studies, noATT)) > 0,
                                  length(intersect(all_studies, noATT)) / length(union(all_studies, noATT)),
                                  NA_real_),
    
    jaccard_all_vs_all3 = ifelse(length(union(all_studies, all3)) > 0,
                                 length(intersect(all_studies, all3)) / length(union(all_studies, all3)),
                                 NA_real_),
    
    jaccard_noATT_vs_all3 = ifelse(length(union(noATT, all3)) > 0,
                                   length(intersect(noATT, all3)) / length(union(noATT, all3)),
                                   NA_real_),
    
    jaccard_all_three = ifelse(length(Reduce(union, list(all_studies, noATT, all3))) > 0,
                               length(Reduce(intersect, list(all_studies, noATT, all3))) /
                                 length(Reduce(union, list(all_studies, noATT, all3))),
                               NA_real_)
  )
})

overlap_summary <- dplyr::bind_rows(overlap_results)
print(overlap_summary)
```

#---------------------------------------------------
# saving limma outputs
# Three analyses, all using age + sex + study model
#---------------------------------------------------
```{r}
prep_gsea_tables <- function(out_list) {
  lapply(out_list, function(tt) {
    tt %>%
      dplyr::select(feature_id, logFC, AveExpr, t, P.Value, adj.P.Val, B, contrast) %>%
      dplyr::arrange(desc(t)) 
  })
}

out_m2 <- prep_gsea_tables(out_m2)
out_m2_noATT <- prep_gsea_tables(out_m2_noATT)
out_m2_all3 <- prep_gsea_tables(out_m2_all3)


contrast_order <- c(
  "T1D_vs_LC", "T2D_vs_LC", "T1D_vs_T2D",
  "OC_vs_LC", "T1D_vs_OC", "T2D_vs_OC"
)

limma_gsea_inputs <- list(
  
  metadata = list(
    created = Sys.time(),
    response = "plasma proteomics",
    transform = "log2(x + 1)",
    model_formula = "0 + group4 + age + sex + study6",
    contrasts = contrast_order
  ),
  
  all_studies = list(
    model_name = "age_sex_study6",
    samples = meta_m2,
    design = design_m2,
    fit = fit_m2,
    out = out_m2,
    gsea_tables = out_m2
  ),
  
  no_attempt = list(
    model_name = "age_sex_study_noATT",
    samples = meta_m2_noATT,
    design = design_m2_noATT,
    fit = fit_m2_noATT,
    out = out_m2_noATT,
    gsea_tables = out_m2_noATT
  ),
  
  all3_omics = list(
    model_name = "age_sex_study4",
    samples = meta_all3,
    design = design_m2_all3,
    fit = fit_m2_all3,
    out = out_m2_all3,
    gsea_tables = out_m2_all3
  )
)

saveRDS(
  limma_gsea_inputs,
  file = file.path(
    dir.dat, "Savanah Leidholt", "T1D-T2D", "Data",
    "limma_gsea_inputs_age_sex_study.rds"
  )
)

cat("Saved: limma_gsea_inputs_age_sex_study.rds\n")
```

#---------------------------------------------------
# Three analyses with reduced model
#---------------------------------------------------
```{r}

limma_gsea_inputs <- list(
  
  metadata = list(
    created = Sys.time(),
    response = "plasma proteomics",
    transform = "log2(x + 1)",
    model_formula = "0 + group4 + age + sex",
    contrasts = contrast_order
  ),
  
  no_attempt_m1 = list(
    model_name = "age_sex_noATT",
    samples = meta_m1_noATT,
    design = design_m1_noATT,
    fit = fit_m1_noATT,
    out = out_m1_noATT,
    gsea_tables = prep_gsea_tables(out_m1_noATT)
  ),
  
  all3_omics_m1 = list(
    model_name = "age_sex_multiOmics",
    samples = meta_all3,
    design = design_m1_all3,
    fit = fit_m1_all3,
    out = out_m1_all3,
    gsea_tables = prep_gsea_tables(out_m1_all3)
  )
)

saveRDS(
  limma_gsea_inputs,
  file = file.path(
    dir.dat, "Savanah Leidholt", "T1D-T2D", "Data",
    "limma_gsea_inputs_age_sex.rds"
  )
)

cat("Saved: limma_gsea_inputs_age_sex.rds\n")

```

