---
title: "Elastic Net Analysis"
output: pdf_document
author: "Savanah Leidholt"
date: "02/17/2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths("~/leidholt/R")
```

```{r}
#load packages
library(dplyr)
library(tidyr)
library(glmnet)
library(pROC)
library(ggplot2)
library(tibble)
```

```{r}
#Directories
#computer <- "mac studio"
computer <- "mac laptop"

user <- Sys.info()[["user"]]

if (computer == "mac studio") {
  
  if (user == "sleidholt") {
    dir.dat <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
    
  } else if (user == "choiyej") {
    dir.dat <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "pylell") {
    dir.dat <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  }

} else if (computer == "mac laptop") {

  if (user == "savanahleidholt") {
    dir.dat <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/T1D-T2D/Results/ElasticNet"
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
  }
}
```

#-----------------------------
#Data cleaning and naming setup
#---------------------------
```{r}
prot_mat_out1 <- read.csv(file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "prot_imp"), row.names = "X")

obj2 <- readRDS(file.path(
  dir.dat, "Savanah Leidholt", "T1D-T2D", "Data",
  "limma_gsea_inputs_age_sex_study.rds"
))

dat1 <- obj2$all_studies
meta1 <- dat1$samples
meta1
#dat2 <- obj2$no_attempt
dat3 <- obj2$all3_omics

out1 <- dat1$out
#out2 <- dat2$out
out3 <- dat3$out


dim(prot_mat_out1)
```

```{r}
# clean IDs
meta1$record_id <- trimws(as.character(meta1$record_id))
rownames(prot_mat_out1) <- trimws(rownames(prot_mat_out1))

# keep only overlapping samples
olap <- intersect(meta1$record_id, rownames(prot_mat_out1))

meta1 <- meta1 %>%
  dplyr::filter(record_id %in% olap) %>%
  dplyr::mutate(
    age = suppressWarnings(as.numeric(age)),
    sex = as.character(sex),
    study6 = as.character(study6)
  )

# drop rows missing anything needed for removeBatchEffect
meta1 <- meta1 %>%
  dplyr::filter(!is.na(age), !is.na(sex), sex != "", !is.na(study6), study6 != "")


prot_mat <- prot_mat_out1[meta1$record_id, , drop = FALSE]

#final checks
stopifnot(identical(rownames(prot_mat), meta1$record_id))
stopifnot(nrow(prot_mat) == nrow(meta1))

meta1 <- meta1 %>%
  dplyr::mutate(
    sex = factor(sex),
    study6 = factor(study6),
    group4 = factor(group4, levels = c("LC", "OC", "T1D", "T2D"))
  )

# check row counts
nrow(prot_mat)
nrow(meta1)
length(meta1$study6)

# prot_mat currently: samples x proteins
dim(prot_mat)

# transpose to proteins x samples for limma
prot_mat_t <- t(prot_mat)

stopifnot(ncol(prot_mat_t) == nrow(meta1))
stopifnot(colnames(prot_mat_t) == meta1$record_id)

design <- model.matrix(~ group4, data = meta1)

sum(is.na(prot_mat))
table(rowSums(is.na(prot_mat)) > 0)
table(colSums(is.na(prot_mat)) > 0)


prot_resid_t <- limma::removeBatchEffect(
  prot_mat_t,
  batch = meta1$study6,
  design = design
)

# transpose back to samples x proteins for glmnet
prot_resid <- t(prot_resid_t)

# restore rownames check
stopifnot(rownames(prot_resid) == meta1$record_id)
dim(prot_resid)

```
#----------------------------------------------
#reading in analyte file for easier annotation
#---------------------------------------------
```{r}
analytes <- read.csv(
  file.path(dir.dat, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

prot_panel <- unique(trimws(as.character(analytes$AptName)))

aptamers_keep <- intersect(prot_panel, colnames(prot_mat))
protein2symbol <- analytes %>%
  transmute(
    feature_id  = trimws(as.character(AptName)),
    gene_symbol = trimws(as.character(EntrezGeneSymbol))
  ) %>%
  filter(feature_id != "", !is.na(feature_id)) %>%
  distinct(feature_id, .keep_all = TRUE)
```

#------------------------
#elasticNet function
#------------------------
```{r}
run_elastic_net <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1,
  min_per_class = 5,
  maxit = 1e6
) {
  
  cat("\n============================\n")
  cat("Running contrast:", cc, "\n")
  
  if (!cc %in% names(out_list)) {
    stop("Contrast not found in out_list: ", cc)
  }
  
  if (!grepl("_vs_", cc)) {
    stop("Contrast name must contain '_vs_'")
  }
  
  parts <- strsplit(cc, "_vs_")[[1]]
  if (length(parts) != 2) {
    stop("Contrast could not be split into two groups")
  }
  
  g_pos <- parts[1]
  g_neg <- parts[2]
  
 idx <- which(meta_df$group4 %in% c(g_pos, g_neg))
meta_sub <- meta_df[idx, , drop = FALSE]
meta_sub$record_id <- as.character(meta_sub$record_id)

meta_sub$group4 <- droplevels(meta_sub$group4)

y <- ifelse(meta_sub$group4 == g_pos, 1, 0)
y <- as.numeric(y)

tab <- table(meta_sub$group4)
if (any(tab < min_per_class)) {
  stop(
    "Too few samples in at least one class for contrast: ",
    cc,
    " | counts = ",
    paste(names(tab), tab, collapse = ", ")
  )
}
  
  t <- out_list[[cc]]
  if (is.null(t)) {
    stop("Contrast table is NULL for: ", cc)
  }
  t <- as.data.frame(t)
  
  cat("Columns in contrast table:\n")
  print(names(t))
  
  if (!"feature_id" %in% names(t)) {
    fid <- rownames(t)
    if (is.null(fid)) stop("No feature_id column or rownames found")
    t$feature_id <- fid
  }
  
  if (!all(c("adj.P.Val", "P.Value") %in% names(t))) {
    stop("Contrast table is missing adj.P.Val and/or P.Value")
  }
  
  cat("Rows in limma table:", nrow(t), "\n")
  cat("Overlap with protein matrix:", sum(t$feature_id %in% colnames(X_mat)), "\n")
  
  top_feats <- t %>%
  dplyr::filter(!is.na(adj.P.Val)) %>%
  dplyr::left_join(protein2symbol, by = "feature_id") %>%
  dplyr::mutate(
    gene_symbol = dplyr::if_else(
      is.na(gene_symbol) | gene_symbol == "",
      feature_id,
      gene_symbol
    )
  ) %>%
  dplyr::filter(feature_id %in% colnames(X_mat)) %>%
  dplyr::arrange(adj.P.Val, P.Value) %>%
  dplyr::group_by(gene_symbol) %>%
  dplyr::slice_head(n = 1) %>%   # keeps best aptamer per gene
  dplyr::ungroup() %>%
  dplyr::slice_head(n = top_n) %>%
  dplyr::pull(feature_id)
  
  cat("Top overlapping features:", length(top_feats), "\n")
  cat("First few top features:\n")
  print(head(top_feats, 10))
  
  if (length(top_feats) < 5) {
    stop("Fewer than 5 overlapping top features for contrast: ", cc)
  }
  
  meta_sub <- meta_sub %>%
    dplyr::mutate(
      sex = factor(sex),
      study6 = factor(study6)
    )
  
  covars_present <- covars[covars %in% colnames(meta_sub)]
  cat("Covariates used:", paste(covars_present, collapse = ", "), "\n")
  
  if (length(covars_present) > 0) {
    form_cov <- as.formula(paste("~", paste(covars_present, collapse = " + ")))
    Z <- model.matrix(form_cov, data = meta_sub)
  } else {
    Z <- matrix(1, nrow = nrow(meta_sub), ncol = 1)
    colnames(Z) <- "(Intercept)"
  }
  
  Xp <- X_mat[meta_sub$record_id, top_feats, drop = FALSE]
  
  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"
  
  pf <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))
  
  sds <- apply(X, 2, sd)
  keep <- which(is.finite(sds) & sds > 0)
  
  X <- X[, keep, drop = FALSE]
  pf <- pf[keep]
  
  cat("Final design matrix dim:", dim(X), "\n")
  
  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))
  cat("nfolds used:", nfolds_use, "\n")
  
  if (nfolds_use < 2) {
    stop("Not enough samples per class for CV")
  }
  
  set.seed(seed)
  
  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      cv.glmnet(
        x = X,
        y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = pf,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) {
        message("glmnet failed at alpha = ", a, ": ", conditionMessage(e))
        NULL
      }
    )
    
    if (is.null(cv)) return(NULL)
    list(alpha = a, cvfit = cv)
  })
  
  fit_list <- Filter(Negate(is.null), fit_list)
  
  if (length(fit_list) == 0) {
    stop("All glmnet fits failed")
  }
  
  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]
  
  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))
  coef_df <- tibble::tibble(
    term = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>%
    dplyr::filter(beta != 0)
  
  protein_terms <- intersect(colnames(Xp), coef_df$term)
  
  p_hat <- as.numeric(
    predict(best$cvfit, newx = X, s = "lambda.1se", type = "response")
  )
  
  auc_val <- as.numeric(
    pROC::auc(pROC::roc(y, p_hat, quiet = TRUE))
  )
  
  cat("Best alpha:", best$alpha, "\n")
  cat("AUC:", auc_val, "\n")
  cat("Nonzero coefficients:", nrow(coef_df), "\n")
  cat("Selected protein terms:", length(protein_terms), "\n")
  
  list(
    contrast         = cc,
    positive         = g_pos,
    negative         = g_neg,
    best_alpha       = best$alpha,
    lambda_1se       = best$cvfit$lambda.1se,
    lambda_min       = best$cvfit$lambda.min,
    auc_train        = auc_val,
    n_input_proteins = length(top_feats),
    coef_df          = coef_df,
    protein_terms    = protein_terms,
    cvfit            = best$cvfit,
    top_feats        = top_feats
  )
}
```

#-----------------------------------
#Contrast specific feature selection
#------------------------------------
```{r}
res_T1D_LC <- run_elastic_net(
  cc = "T1D_vs_LC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex")
)
```

```{r}
res_T1D_OC <- run_elastic_net(
  cc = "T1D_vs_OC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex")
)
```

```{r}
res_T1D_T2D <- run_elastic_net(
  cc = "T1D_vs_T2D",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_mat,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex")
)
```

```{r}
res_T2D_LC<- run_elastic_net(
  cc = "T2D_vs_LC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex")
)
```

```{r}
res_T2D_OC<- run_elastic_net(
  cc = "T2D_vs_OC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex")
)
```

```{r}
res_OC_LC<- run_elastic_net(
  cc = "OC_vs_LC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex")
)
```

#------------------------------
#reuslts tables per contrast
#------------------------------
```{r}
#sig T1D vs LC
tibble::tibble(
  contrast = res_T1D_LC$contrast,
  alpha = res_T1D_LC$best_alpha,
  lambda_1se = res_T1D_LC$lambda_1se,
  lambda_min = res_T1D_LC$lambda_min,
  auc = res_T1D_LC$auc_train,
  n_input_proteins = res_T1D_LC$n_input_proteins,
  n_selected_proteins = length(res_T1D_LC$protein_terms)
)

sig_tbl_T1D_LC <- res_T1D_LC$coef_df %>%
  dplyr::filter(term %in% res_T1D_LC$protein_terms) %>%
  dplyr::rename(feature_id = term) %>%
  dplyr::left_join(protein2symbol, by = "feature_id") %>%
  dplyr::mutate(abs_beta = abs(beta)) %>%
  dplyr::arrange(desc(abs_beta))

head(sig_tbl_T1D_LC)
```

```{r}
#sig T1D vs OC
tibble::tibble(
  contrast = res_T1D_OC$contrast,
  alpha = res_T1D_OC$best_alpha,
  lambda_1se = res_T1D_OC$lambda_1se,
  lambda_min = res_T1D_OC$lambda_min,
  auc = res_T1D_OC$auc_train,
  n_input_proteins = res_T1D_OC$n_input_proteins,
  n_selected_proteins = length(res_T1D_OC$protein_terms)
)

sig_tbl_T1D_OC <- res_T1D_OC$coef_df %>%
  dplyr::filter(term %in% res_T1D_OC$protein_terms) %>%
  dplyr::rename(feature_id = term) %>%
  dplyr::left_join(protein2symbol, by = "feature_id") %>%
  dplyr::mutate(abs_beta = abs(beta)) %>%
  dplyr::arrange(desc(abs_beta))

head(sig_tbl_T1D_OC)
```

```{r}
#sig T1D vs T2D
tibble::tibble(
  contrast = res_T1D_T2D$contrast,
  alpha = res_T1D_T2D$best_alpha,
  lambda_1se = res_T1D_T2D$lambda_1se,
  lambda_min = res_T1D_T2D$lambda_min,
  auc = res_T1D_T2D$auc_train,
  n_input_proteins = res_T1D_T2D$n_input_proteins,
  n_selected_proteins = length(res_T1D_T2D$protein_terms)
)

sig_tbl_T1D_T2D<- res_T1D_T2D$coef_df %>%
  dplyr::filter(term %in% res_T1D_T2D$protein_terms) %>%
  dplyr::rename(feature_id = term) %>%
  dplyr::left_join(protein2symbol, by = "feature_id") %>%
  dplyr::mutate(abs_beta = abs(beta)) %>%
  dplyr::arrange(desc(abs_beta))

head(sig_tbl_T1D_T2D)
```

```{r}
#sig T2D vs LC
tibble::tibble(
  contrast = res_T2D_LC$contrast,
  alpha = res_T2D_LC$best_alpha,
  lambda_1se = res_T2D_LC$lambda_1se,
  lambda_min = res_T2D_LC$lambda_min,
  auc = res_T2D_LC$auc_train,
  n_input_proteins = res_T2D_LC$n_input_proteins,
  n_selected_proteins = length(res_T2D_LC$protein_terms)
)

sig_tbl_T2D_LC <- res_T2D_LC$coef_df %>%
  dplyr::filter(term %in% res_T2D_LC$protein_terms) %>%
  dplyr::rename(feature_id = term) %>%
  dplyr::left_join(protein2symbol, by = "feature_id") %>%
  dplyr::mutate(abs_beta = abs(beta)) %>%
  dplyr::arrange(desc(abs_beta))

head(sig_tbl_T2D_LC)
```
#-----------------------------------------------
#bootstrapping function (100x)
#-----------------------------------------------
```{r}
run_elastic_net_bootstrap <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  n_boot = 100,
  seed = 1835,
  min_per_class = 5, 
  verbose = F) {
  set.seed(seed)
  
  boot_res <- vector("list", n_boot)
  
  for (b in seq_len(n_boot)) {
    
    cat("Bootstrap", b, "of", n_boot, "for", cc, "\n")
    
    parts <- strsplit(cc, "_vs_")[[1]]
    g_pos <- parts[1]
    g_neg <- parts[2]
    
    idx0 <- which(meta_df$group4 %in% c(g_pos, g_neg))
    meta_sub0 <- meta_df[idx0, , drop = FALSE]
    
    # stratified bootstrap: sample within each group
    idx_boot <- unlist(lapply(split(seq_len(nrow(meta_sub0)), meta_sub0$group4), function(ii) {
      sample(ii, size = length(ii), replace = TRUE)
    }))
    
    meta_boot <- meta_sub0[idx_boot, , drop = FALSE]
    meta_boot$record_id_boot <- paste0(meta_boot$record_id, "_boot", seq_len(nrow(meta_boot)))
    
    X_boot <- X_mat[meta_boot$record_id, , drop = FALSE]
    rownames(X_boot) <- meta_boot$record_id_boot
    
    meta_boot$record_id <- meta_boot$record_id_boot
    
    fit_b <- tryCatch(
      run_elastic_net(
        cc = cc,
        out_list = out_list,
        meta_df = meta_boot,
        X_mat = X_boot,
        protein2symbol = protein2symbol,
        top_n = top_n,
        covars = covars,
        alpha_grid = alpha_grid,
        seed = seed + b,
        min_per_class = min_per_class
      ),
      error = function(e) {
        message("Bootstrap failed at b = ", b, ": ", conditionMessage(e))
        NULL
      }
    )
    
    boot_res[[b]] <- fit_b
  }
  
  boot_res <- Filter(Negate(is.null), boot_res)
  
 selected_tbl <- lapply(seq_along(boot_res), function(i) {
  tibble(
    boot = i,
    contrast = cc,
    best_alpha = boot_res[[i]]$best_alpha,
    auc_train = boot_res[[i]]$auc_train,
    feature_id = unique(boot_res[[i]]$protein_terms)
  )
}) %>%
  bind_rows() %>%
  distinct(contrast, boot, feature_id, .keep_all = TRUE)
  
  stability_tbl <- selected_tbl %>%
  distinct(contrast, boot, feature_id) %>%
  count(contrast, feature_id, name = "n_selected") %>%
  mutate(
    n_boot_success = length(boot_res),
    selection_frequency = n_selected / n_boot_success
  ) %>%
  left_join(protein2symbol, by = "feature_id") %>%
  arrange(desc(selection_frequency))
  
  stability_tbl <- stability_tbl %>%
  mutate(selection_percent = 100 * selection_frequency)
  
  model_summary <- tibble(
    contrast = cc,
    n_boot_requested = n_boot,
    n_boot_success = length(boot_res),
    mean_auc = mean(sapply(boot_res, function(x) x$auc_train), na.rm = TRUE),
    median_auc = median(sapply(boot_res, function(x) x$auc_train), na.rm = TRUE),
    mean_alpha = mean(sapply(boot_res, function(x) x$best_alpha), na.rm = TRUE),
    median_alpha = median(sapply(boot_res, function(x) x$best_alpha), na.rm = TRUE)
  )
  
  list(
    contrast = cc,
    boot_res = boot_res,
    selected_tbl = selected_tbl,
    stability_tbl = stability_tbl,
    model_summary = model_summary
  )
}
```
#----------------------
#boostrapping T1D vs LC
#----------------------
```{r}
boot_T1D_LC <- run_elastic_net_bootstrap(
  cc = "T1D_vs_LC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1835,
  verbose = FALSE
)
```

```{r}
boot_T1D_LC$stability_tbl %>%
  filter(selection_frequency >= 0.7)
```
#--------------------------
#boostrapping T1D vs OC
#--------------------------
```{r}
boot_T1D_OC <- run_elastic_net_bootstrap(
  cc = "T1D_vs_OC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1
)
```

```{r}
boot_T1D_OC$stability_tbl %>%
  filter(selection_frequency >= 0.7)
```
#--------------------------
#boostrapping T1D vs T2D
#--------------------------
```{r, message=FALSE, warning=FALSE, results='hide'}
boot_T1D_T2D <- run_elastic_net_bootstrap(
  cc = "T1D_vs_T2D",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1
)
```

```{r}
boot_T1D_T2D$stability_tbl %>%
  filter(selection_frequency >= 0.7)
```

#--------------------------
#boostrapping T2D vs LC
#--------------------------
```{r, message=FALSE, warning=FALSE, results='hide'}
boot_T2D_LC <- run_elastic_net_bootstrap(
  cc = "T2D_vs_LC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1
)
```

```{r}
boot_T2D_LC$stability_tbl %>%
  filter(selection_frequency >= 0.7)
```

#--------------------------
#boostrapping T2D vs OC
#--------------------------
```{r}
boot_T2D_OC <- run_elastic_net_bootstrap(
  cc = "T2D_vs_OC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1
)
```

```{r}
boot_T2D_OC$stability_tbl %>%
  filter(selection_frequency >= 0.7)
```

#--------------------------
#boostrapping OC vs LC
#--------------------------
```{r}
boot_OC_LC <- run_elastic_net_bootstrap(
  cc = "OC_vs_LC",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1
)
```

```{r}
boot_OC_LC$stability_tbl %>%
  filter(selection_frequency >= 0.7)
```
#---------------------------------------
#feature selection frequency plotting funciton
#--------------------------------------
```{r}
plot_stability <- function(boot_obj, min_freq = 0.7, top_n = 25) {
  
  plot_df <- boot_obj$stability_tbl %>%
    dplyr::mutate(
      selection_percent = 100 * selection_frequency,
      label = dplyr::if_else(
        is.na(gene_symbol) | gene_symbol == "",
        feature_id,
        gene_symbol
      )
    ) %>%
    dplyr::filter(selection_frequency >= min_freq) %>%
    dplyr::slice_max(selection_frequency, n = top_n, with_ties = FALSE)
  
  ggplot(plot_df, aes(
    x = reorder(label, selection_percent),
    y = selection_percent
  )) +
    geom_col() +
    coord_flip() +
    labs(
      title = paste0("Elastic net feature stability: ", boot_obj$contrast),
      x = NULL,
      y = "Bootstrap selection frequency (%)"
    ) +
    theme_bw()
}
```

```{r}

p_stability_T1D_LC <- plot_stability(boot_T1D_LC)
p_stability_T1D_LC
ggsave(filename = file.path(dir.results, "p_stability_T1D_LC.png"),
       plot = p_stability_T1D_LC,
       width = 10,
       height = 6)
```
```{r}
p_stability_T1D_OC <- plot_stability(boot_T1D_OC)
p_stability_T1D_OC
ggsave(filename = file.path(dir.results, "p_stability_T1D_OC.png"),
       plot = p_stability_T1D_OC,
       width = 10,
       height = 6)
```

```{r}
p_stability_T1D_T2D <- plot_stability(boot_T1D_T2D)
p_stability_T1D_T2D
ggsave(filename = file.path(dir.results, "p_stability_T1D_T2D.png"),
       plot = p_stability_T1D_T2D,
       width = 10,
       height = 6)
```
```{r}
p_stability_T2D_LC <- plot_stability(boot_T2D_LC)
p_stability_T2D_LC
ggsave(filename = file.path(dir.results, "p_stability_T2D_LC.png"),
       plot = p_stability_T2D_LC,
       width = 10,
       height = 6)
```

```{r}
p_stability_T2D_OC <- plot_stability(boot_T2D_OC)
p_stability_T2D_OC
ggsave(filename = file.path(dir.results, "p_stability_T2D_OC.png"),
       plot = p_stability_T2D_OC,
       width = 10,
       height = 6)
```

```{r}
p_stability_OC_LC <- plot_stability(boot_OC_LC)
p_stability_OC_LC
ggsave(filename = file.path(dir.results, "p_stability_OC_LC.png"),
       plot = p_stability_OC_LC,
       width = 10,
       height = 6)
```

```{r}
plot_boot_auc <- function(boot_obj) {
  
  auc_df <- tibble(
    contrast = boot_obj$contrast,
    auc = sapply(boot_obj$boot_res, function(x) x$auc_train)
  )
  
  ggplot(auc_df, aes(x = auc)) +
    geom_histogram(bins = 20) +
    geom_vline(xintercept = median(auc_df$auc, na.rm = TRUE), linetype = "dashed") +
    labs(
      title = paste0("Bootstrap AUC distribution: ", boot_obj$contrast),
      x = "Training AUC across bootstrap models",
      y = "Number of bootstrap models"
    ) +
    theme_bw()
}

```

```{r}
p_auc_T1D_LC <- plot_boot_auc(boot_T1D_LC)
p_auc_T1D_LC
```

```{r}
p_auc_T1D_OC <- plot_boot_auc(boot_T1D_OC)
p_auc_T1D_OC
```

```{r}
p_auc_T1D_T2D <- plot_boot_auc(boot_T1D_T2D)
p_auc_T1D_T2D
```
```{r}
p_auc_T2D_LC <- plot_boot_auc(boot_T2D_LC)
p_auc_T2D_LC
```

```{r}
p_auc_T2D_OC <- plot_boot_auc(boot_T2D_OC)
p_auc_T2D_OC
```

```{r}
p_auc_OC_LC <- plot_boot_auc(boot_OC_LC)
p_auc_OC_LC
```

```{r}
plot_alpha_distribution <- function(boot_obj) {
  
  alpha_df <- tibble(
    contrast = boot_obj$contrast,
    alpha = sapply(boot_obj$boot_res, function(x) x$best_alpha)
  )
  
  ggplot(alpha_df, aes(x = factor(alpha))) +
    geom_bar() +
    labs(
      title = paste0("Selected alpha values: ", boot_obj$contrast),
      x = "Best alpha",
      y = "Number of bootstrap models"
    ) +
    theme_bw()
}

```

```{r}
p_alpha_T1D_LC <- plot_alpha_distribution(boot_T1D_LC)
p_alpha_T1D_LC
```

```{r}
p_alpha_T1D_OC <- plot_alpha_distribution(boot_T1D_OC)
p_alpha_T1D_OC
```

```{r}
p_alpha_T1D_T2D <- plot_alpha_distribution(boot_T1D_T2D)
p_alpha_T1D_T2D
```

```{r}
p_alpha_T2D_LC <- plot_alpha_distribution(boot_T2D_LC)
p_alpha_T2D_LC
```

```{r}
p_alpha_T2D_OC <- plot_alpha_distribution(boot_T2D_OC)
p_alpha_T2D_OC
```

```{r}
p_alpha_OC_LC <- plot_alpha_distribution(boot_OC_LC)
p_alpha_OC_LC
```