---
title: "Metabolomics Data Analysis"
author: "Savanah leidholt"
date: "2026-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#load libraries
library(dplyr)
library(tidyr)
library(tibble)
library(limma)
library(ggplot2)
library(EnhancedVolcano)
library(patchwork)
library(stringr)
library(purrr)
library(KEGGREST)
library(caret)
library(glmnet)
library(jsonlite)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- "/mmfs1/gscratch/togo/leidholt/"
  git_path  <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys      <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
  root_path <- "/Users/Shared/UW - bjornstad_pyle_tommerdahl_lab - Data Science/Projects/"
  dir.results <- "/Users/Shared/UW - bjornstad_pyle_tommerdahl_lab - Data Science/Projects/P010_endothelial_multiomics_sl/results/"
  git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "savanahleidholt") {
  root_path <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
  dir.results <- file.path(root_path, "Savanah Leidholt", "TRIAD", "Results")
  git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

#------------------------------------
# raw metabolomics QC/summary plots
#------------------------------------
```{r}
omics_objects <- readRDS(file.path(root_path, "Savanah Leidholt", "TRIAD", "Data", "omics_split_objects_clean.rds"))
meta_qc <- omics_objects$meta_metab
mat_qc  <- omics_objects$metabolites

meta_qc$record_id <- trimws(as.character(meta_qc$record_id))
meta_qc$study     <- trimws(as.character(meta_qc$study))
meta_qc$group     <- trimws(as.character(meta_qc$group))

meta_qc$study6 <- dplyr::recode(
  meta_qc$study,
  "ATTEMPT"         = "ATT",
  "CROCODILE"       = "CRC",
  "RENAL-HEIR"      = "RH",
  "IMPROVE"         = "IT",
  "PANTHER" = "PANTH",
  "RENAL-HEIRitage" = "RH2",
  .default = meta_qc$study
)

rownames(mat_qc) <- trimws(as.character(rownames(mat_qc)))

# keep only overlapping samples
ids_qc <- intersect(meta_qc$record_id, rownames(mat_qc))
meta_qc <- meta_qc[match(ids_qc, meta_qc$record_id), , drop = FALSE]
mat_qc  <- mat_qc[ids_qc, , drop = FALSE]

stopifnot(identical(meta_qc$record_id, rownames(mat_qc)))

cat("QC samples:", nrow(mat_qc), "\n")
cat("QC metabolites:", ncol(mat_qc), "\n")

```
#--------------------
#PCA plotting Function
#--------------------
```{r}
make_pca_plot <- function(mat_in, meta_df, title_str) {
  pca <- prcomp(mat_in, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1 = pca$x[, 1],
    PC2 = pca$x[, 2],
    group4 = meta_df$group4,
    study  = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = group4, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        LC  = "#2166AC",
        OC  = "#F4A582",
        T1D = "#D6604D",
        T2D = "#4DAC26"
      )
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
## ----------------------
## sample level coverage
## ----------------------
```{r}
sample_qc <- meta_qc %>%
  mutate(
    n_metabolites_detected = rowSums(!is.na(mat_qc)),
    n_metabolites_missing  = rowSums(is.na(mat_qc)),
    pct_detected           = 100 * n_metabolites_detected / ncol(mat_qc),
    pct_missing            = 100 * n_metabolites_missing / ncol(mat_qc)
  )

# boxplot: number of detected metabolites per sample by study
p_cov_n <- ggplot(sample_qc, aes(x = study6, y = n_metabolites_detected, fill = study6)) +
  scale_fill_brewer(palette = "Dark2") + 
  geom_boxplot(alpha = 0.75, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
  theme_bw() +
  labs(
    title = "Raw metabolite coverage per sample",
    x = "Study",
    y = "Number of detected metabolites"
  ) +
  theme(legend.position = "none")

# boxplot: percent missing per sample by study
p_miss_sample <- ggplot(sample_qc, aes(x = study6, y = pct_missing, fill = study6)) +
  scale_fill_brewer(palette = "Dark2") +
  geom_boxplot(alpha = 0.75, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
  theme_bw() +
  labs(
    title = "Percent missingness per sample",
    x = "Study",
    y = "% missing metabolites"
  ) +
  theme(legend.position = "none")

p_cov_n | p_miss_sample
```
##---------------------------------
## study level coverage
##---------------------------------
```{r}
study_idx <- split(seq_len(nrow(mat_qc)), meta_qc$study6)

present_by_study <- sapply(study_idx, function(idx) {
  colSums(!is.na(mat_qc[idx, , drop = FALSE])) > 0
})

present_by_study <- as.matrix(present_by_study)

study_coverage <- tibble(
  study6 = colnames(present_by_study),
  n_metabolites_present = colSums(present_by_study),
  pct_metabolites_present = 100 * n_metabolites_present / nrow(present_by_study)
)

p_study_cov <- ggplot(study_coverage, aes(x = study6, y = pct_metabolites_present, fill = study6)) +
  scale_fill_brewer(palette = "Dark2")+
  geom_col(alpha = 0.85) +
  geom_text(aes(label = round(pct_metabolites_present, 1)), vjust = -0.3, size = 3.5) +
  theme_bw() +
  labs(
    title = "Raw metabolite coverage in each study",
    x = "Study",
    y = "% of metabolites observed in at least one sample"
  ) +
  ylim(0, 105) +
  theme(legend.position = "none")

p_study_cov

```
# ----------------------------------------------------
# METABOLOMICS-ONLY ANALYSIS WORKFLOW
# a) all metabolomics saples
# b) metabolomics samples that also have proteomics
#-----------------------------------------------------
```{r}
contrast_names <- c(
  "T1D_vs_LC", "T2D_vs_LC", "T1D_vs_T2D",
  "OC_vs_LC", "T1D_vs_OC", "T2D_vs_OC"
)
```
#------------------
# a. metabolomics only
#------------------
```{r}
meta_metab <- omics_objects$meta_metab
mat_metab  <- omics_objects$metabolites


rownames(mat_metab) <- trimws(rownames(mat_metab))
meta_metab <- meta_metab %>%
  mutate(
    record_id = trimws(as.character(record_id)),
    study     = trimws(as.character(study)),
    group     = trimws(as.character(group)),
    sex       = trimws(as.character(sex)),
    age       = suppressWarnings(as.numeric(age)),
    bmi       = suppressWarnings(as.numeric(bmi)),

    group4 = factor(
      group,
      levels = c(
        "Lean Control",
        "Obese Control",
        "Type 1 Diabetes",
        "Type 2 Diabetes"
      ),
      labels = c("LC", "OC", "T1D", "T2D")
    ),

    study = dplyr::recode(
      study,
      "ATTEMPT"         = "ATT",
      "CROCODILE"       = "CRC",
      "RENAL-HEIR"      = "RH",
      "IMPROVE"         = "IT",
      "PANTHER"         = "PANTH",
      "RENAL-HEIRitage" = "RH2",
      .default = study
    ),

    study = factor(study),
    sex   = factor(sex)
  ) %>%
  filter(
    !is.na(age),
    !is.na(sex),
    sex != "",
    study != "ATT" #attempt only contains one metabolite
  )

#remove samples with no metabolite data
mat_metab <- mat_metab[rowSums(!is.na(mat_metab)) > 0, , drop = FALSE]

# keep and align shared samples
ids_use <- intersect(meta_metab$record_id, rownames(mat_metab))
meta_all <- meta_metab %>%
  filter(record_id %in% ids_use) %>%
  arrange(match(record_id, ids_use))

mat_all <- mat_metab[meta_all$record_id, , drop = FALSE]

stopifnot(identical(meta_all$record_id, rownames(mat_all)))

cat("All metabolomics participant before filtering:", nrow(meta_metab), "\n")
cat("All metabolomics participant after filtering:", nrow(meta_all), "\n")
cat("All metabolomics features before filtiner:", ncol(mat_metab), "\n")
cat("All metabolomics features after filtering:", ncol(mat_all), "\n")
print(table(meta_all$study, meta_all$group4, useNA = "ifany"))
```
#----------------------------
#filtering step where protein 
#must exist in 60% of samples 
#in a study &b be in 2 studies
#----------------------------
```{r}
study_idx <- split(seq_len(nrow(mat_all)), meta_all$study)

present_by_study <- vapply(
  study_idx,
  FUN = function(idx) {
    colMeans(!is.na(mat_all[idx, , drop = FALSE])) >= 0.6
  },
  FUN.VALUE = setNames(logical(ncol(mat_all)), colnames(mat_all))
)

present_by_study[is.na(present_by_study)] <- FALSE

n_studies_present <- rowSums(present_by_study)

keep_metabolites <- n_studies_present >= 2

kept_metabolites <- rownames(present_by_study)[keep_metabolites]
removed_metabolites <- rownames(present_by_study)[!keep_metabolites]

cat("Removed metabolites:", length(removed_metabolites), "\n")
cat("Kept metabolites:", length(kept_metabolites), "\n")

mat_filt <- mat_all[, kept_metabolites, drop = FALSE]

```

##----------------------------------------------------------------------------
## log transform & removing participants without any data in indivudal studies
##----------------------------------------------------------------------------
```{r}
mat_log_all <- as.matrix(log2(mat_filt + 1))
mat_log_all[!is.finite(mat_log_all)] <- NA_real_


# remove near-zero variance features
nzv_idx <- nearZeroVar(mat_log_all)
nzv_idx

mat_log_filt <- if (length(nzv_idx) > 0) {
  mat_log_all[, -nzv_idx, drop = FALSE]
} else {
  mat_log_all
}

cat("Features before filtering:", ncol(mat_filt), "\n") #165
cat("Features after filtering:", ncol(mat_log_filt), "\n") #165
cat("Participants:", nrow(mat_log_filt), "\n") #132

stopifnot(identical(rownames(mat_log_filt), meta_all$record_id))
```
#------------------------------------
# Impute missing values for PCA only
#------------------------------------
```{r}
mat_pca <- mat_log_filt

# convert Inf / -Inf / NaN to NA
mat_pca[!is.finite(mat_pca)] <- NA_real_

# median impute each metabolite
impute_col_median <- function(x) {
  med <- median(x, na.rm = TRUE)
  x[is.na(x)] <- med
  x
}

mat_pca <- apply(mat_pca, 2, impute_col_median)
mat_pca <- as.matrix(mat_pca)

# checks
stopifnot(identical(rownames(mat_pca), meta_all$record_id))
stopifnot(!anyNA(mat_pca))
stopifnot(all(is.finite(mat_pca)))

cat("Missing values before PCA imputation:", sum(is.na(mat_log_filt)), "\n")
cat("Missing values after PCA imputation:", sum(is.na(mat_pca)), "\n")
```


##---------------------------
## PCA-based outlier removal
##---------------------------
```{r}
pca_all <- prcomp(mat_pca, center = TRUE, scale. = TRUE)

pca_dist_all <- sqrt(rowSums(pca_all$x[, 1:5, drop = FALSE]^2))
cutoff_all <- median(pca_dist_all) + 3 * mad(pca_dist_all, constant = 1)

outlier_df_all <- meta_all %>%
  mutate(
    pca_dist = pca_dist_all,
    pca_outlier = pca_dist > cutoff_all
  ) %>%
  arrange(desc(pca_dist))

outlier_df_all %>%
  filter(pca_outlier) %>%
  select(record_id, study, group4, pca_dist) #RH-48-T, RH-49-T, CRC-11, CRC-50, CRC-42, CRC-43

keep_ids_all <- outlier_df_all %>%
  filter(!pca_outlier) %>%
  pull(record_id)

mat_log_filt2 <- mat_log_filt[keep_ids_all, , drop = FALSE]

meta_all2 <- meta_all %>%
  filter(record_id %in% keep_ids_all) %>%
  arrange(match(record_id, keep_ids_all))

stopifnot(identical(meta_all2$record_id, rownames(mat_log_filt2)))

cat("All metabolomics features kept for PCA/LIMMA:", ncol(mat_log_filt2), "\n")
cat("Number of participants kept:", nrow(mat_log_filt2), "\n")
```

##-----------------------
## PCA: all metabolomics
##-----------------------
```{r}
p0_all <- make_pca_plot(
  mat_pca,
  meta_all,
  "Metabolomics - No adjustment"
)

cov1_all <- model.matrix(~ age + sex, data = meta_all)[, -1, drop = FALSE]

mat_adj1_all <- t(limma::removeBatchEffect(
  t(mat_pca),
  covariates = cov1_all
))

p1_all <- make_pca_plot(
  mat_adj1_all,
  meta_all,
  "Metabolomics - Age + Sex"
)

cov2_all <- model.matrix(~ age + sex + study, data = meta_all)[, -1, drop = FALSE]

mat_adj2_all <- t(limma::removeBatchEffect(
  t(mat_pca),
  covariates = cov2_all
))

p2_all <- make_pca_plot(
  mat_adj2_all,
  meta_all,
  "Metabolomics - Age + Sex + Study"
)

(p0_all | p1_all | p2_all) +
  plot_annotation(
    title = "PCA: All metabolomics samples",
    subtitle = "Color = group | Shape = study | removeBatchEffect used for visualization only"
  )
```
##------------------------
## LIMMA: all metabolomics
## ------------------------
```{r}
Y_all <- t(as.matrix(mat_log_filt))
meta_all <- meta_all %>%
  mutate(
    group4 = droplevels(group4),
    sex    = droplevels(sex),
    study  = droplevels(study)
  )

design_all <- model.matrix(~ 0 + group4 + age + sex + study, data = meta_all)
colnames(design_all) <- sub("^group4", "", colnames(design_all))
colnames(design_all) <- make.names(colnames(design_all))

contr_all <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_all
)

fit_all <- lmFit(Y_all, design_all)
fit_all <- contrasts.fit(fit_all, contr_all)
fit_all <- eBayes(fit_all)

out_all <- list()
for (cc in colnames(contr_all)) {
  tt <- topTable(fit_all, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
  tt$feature_id <- rownames(tt)
  tt$contrast <- cc
  out_all[[cc]] <- tt
}

sig_all <- tibble(
  contrast = names(out_all),
  n_FDR05 = sapply(out_all, function(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_all, function(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
  model = "metabolomics_age_sex_study_all",
  n_samples = nrow(meta_all)
)

print(sig_all)

table(meta_all$study, useNA = "ifany")
colnames(design_all)
```

##--------------------------------
## VOLCANO PLOTS--all metabolomics
##---------------------------------
```{r}
volcano_list_all <- lapply(contrast_names, function(cc) {
  tt <- out_all[[cc]]

  EnhancedVolcano(
    tt,
    lab = tt$feature_id,
    x = "logFC",
    y = "adj.P.Val",
    pCutoff = 0.05,
    FCcutoff = 0.5,
    title = gsub("_", " ", cc),
    subtitle = "Age + sex + study",
    pointSize = 2.0,
    labSize = 3.0
  )
})

wrap_plots(volcano_list_all, ncol = 3)
```


```{r}

dict <- read.csv2(file.path(root_path, "Data Harmonization", "data_dictionary_master.csv"), sep = ",") 

dict <- dict %>% dplyr::select(variable_name, label, form_name)

dict_metab <- dict %>%
  mutate(
    variable_name = trimws(as.character(variable_name)),
    label = trimws(as.character(label)),
    form_name = trimws(as.character(form_name))
  ) %>%
  filter(stringr::str_detect(form_name, regex("^metabolomics", ignore_case = TRUE)))

dict_lookup <- dict_metab %>%
  distinct(variable_name, label) %>%
  filter(!is.na(variable_name), variable_name != "") %>%
  tibble::deframe()

out_all <- lapply(out_all, function(x) {
  if (is.matrix(x) || is.data.frame(x)) {
    old_names <- rownames(x)
    if (!is.null(old_names)) {
      new_names <- dict_lookup[old_names]
      new_names[is.na(new_names) | new_names == ""] <- old_names[is.na(new_names) | new_names == ""]
      rownames(x) <- make.unique(new_names)
    }
  }
  x
})

out_all
```

#---------------------------------
# KEGG compound -> pathway mapping
#---------------------------------
```{r}
# all human KEGG pathways
kegg_pathways_all <- KEGGREST::keggList("pathway", "hsa")

pathway_df <- tibble(
  pathway_id = sub("^path:", "", names(kegg_pathways_all)),
  pathway_name = unname(kegg_pathways_all)
)

pathway_df <- pathway_df %>%
  filter(str_detect(pathway_name, regex("metabolism", ignore_case = TRUE)) |
           pathway_id %in% c(
             "hsa00010","hsa00020","hsa00030","hsa00040","hsa00051","hsa00052",
             "hsa00053","hsa00500","hsa00520","hsa00620","hsa00630","hsa00640",
             "hsa00650","hsa00670","hsa00730","hsa00830","hsa00230","hsa00240",
             "hsa00250","hsa00260","hsa00270","hsa00280","hsa00300","hsa00310",
             "hsa00330","hsa00340","hsa00350","hsa00360","hsa00400","hsa00410",
             "hsa00430","hsa00450","hsa00460","hsa00471","hsa00472","hsa00480"
           ))

# get compounds in each pathway
pathway2compound <- map_dfr(pathway_df$pathway_id, function(pid) {
  x <- KEGGREST::keggGet(pid)[[1]]

  cpds <- x$COMPOUND
  if (is.null(cpds)) return(NULL)

  tibble(
    pathway_id = pid,
    compound_id = names(cpds)
  )
}) %>%
  distinct() %>%
  left_join(pathway_df, by = "pathway_id")

head(pathway2compound)
```
```{r}
normalize_kegg_compound <- function(x) {
  x <- trimws(as.character(x))
  x <- toupper(x)
  x
}
```

#----------------------------------------------------------------
#Performing Over-representation analysis
#using metabolites that remained after filtering as background
#check to see if I should include all metabolites in assay instead
#-----------------------------------------------------------------
```{r}
run_kegg_metab_ora <- function(
  tt,
  pathway_map,
  p_cutoff = 0.05,
  fc_cutoff = 0,
  feature_col = "feature_id",
  min_overlap = 3
) {
  tt <- tt %>%
    mutate(
      compound_id = normalize_kegg_compound(.data[[feature_col]])
    )

  # background = all tested metabolites that survived filtering
  bg <- intersect(tt$compound_id, unique(pathway_map$compound_id))

  # significant metabolites
  sig <- tt %>%
    filter(adj.P.Val < p_cutoff, abs(logFC) >= fc_cutoff) %>%
    pull(compound_id) %>%
    intersect(unique(pathway_map$compound_id))

  if (length(sig) == 0 || length(bg) == 0) {
    return(tibble(
      pathway_id = character(),
      pathway_name = character(),
      n_pathway = integer(),
      n_overlap = integer(),
      n_sig = integer(),
      n_bg = integer(),
      p_value = numeric(),
      padj = numeric(),
      overlap_compounds = character()
    ))
  }

  res <- pathway_map %>%
    group_by(pathway_id, pathway_name) %>%
    summarise(
      pathway_compounds = list(unique(compound_id)),
      n_pathway = length(unique(compound_id)),
      .groups = "drop"
    ) %>%
    rowwise() %>%
    mutate(
      pathway_bg = list(intersect(bg, pathway_compounds)),
      overlap = list(intersect(sig, pathway_bg)),
      n_overlap = length(overlap),
      n_in_path_bg = length(pathway_bg),
      n_sig = length(sig),
      n_bg = length(bg),

      # Fisher table components
      n_sig_in_path = n_overlap, # sig metabolite & in this pathway
      n_sig_not_in_path = n_sig - n_overlap, # sig metabolite but not in this pathway
      n_nonsig_in_path = n_in_path_bg - n_overlap, #not sig metab but in this pathway
      n_nonsig_not_path = n_bg - n_sig - (n_in_path_bg - n_overlap), #not sig, not in pathway

      p_value = if (
        all(c(
          n_sig_in_path,
          n_sig_not_in_path,
          n_nonsig_in_path,
          n_nonsig_not_path
        ) >= 0)
      ) {
        fisher.test(
          matrix(
            c(
              n_sig_in_path,
              n_sig_not_in_path,
              n_nonsig_in_path,
              n_nonsig_not_path
            ),
            nrow = 2
          ),
          alternative = "greater"
        )$p.value
      } else {
        NA_real_
      }
    ) %>%
    ungroup() %>%
    mutate(
      padj = p.adjust(p_value, method = "BH"),
      overlap_compounds = purrr::map_chr(overlap, ~ paste(.x, collapse = ", "))
    ) %>%
    filter(n_overlap >= min_overlap) %>%
    arrange(padj, p_value)

  res
}

```

```{r}
#plotting function
plot_kegg_ora <- function(res_df, title_text, n = 15) {
  if (nrow(res_df) == 0) {
    return(ggplot() + theme_void() + labs(title = paste(title_text, "- no enriched pathways")))
  }

  df <- res_df %>%
    filter(!is.na(padj)) %>%
    slice_min(order_by = padj, n = n, with_ties = FALSE) %>%
    mutate(
      pathway_name = str_replace_all(pathway_name, " - Homo sapiens \\(human\\)", ""),
      pathway_name = factor(pathway_name, levels = rev(pathway_name)),
      neglog10_padj = -log10(padj)
    )

  ggplot(df, aes(x = neglog10_padj, y = pathway_name, size = n_overlap)) +
    geom_point(alpha = 0.85) +
    theme_bw() +
    labs(
      title = title_text,
      x = expression(-log[10]("BH-adjusted p-value")),
      y = NULL,
      size = "Overlap"
    )
}
```
##-----------------------------
##a. ORA on all metabolite data
##-----------------------------
```{r}
kegg_res <- lapply(out_all, function(tt) {
  run_kegg_metab_ora(
    tt = tt,
    pathway_map = pathway2compound,
    p_cutoff = 0.05,
    fc_cutoff = 0
  )
})

names(kegg_res) <- names(out_all)

kegg_summary<- bind_rows(lapply(names(kegg_res), function(cc) {
  x <- kegg_res[[cc]]
  tibble(
    contrast = cc,
    n_pathways_FDR05 = sum(x$padj < 0.05, na.rm = TRUE),
    n_pathways_p05   = sum(x$p_value < 0.05, na.rm = TRUE)
  )
}))

kegg_summary
```

```{r}
kegg_res[["T1D_vs_OC"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(20)
```

```{r}
plot_kegg_ora(kegg_res[["T1D_vs_OC"]], "KEGG ORA: T1D vs OC")
kegg_plots <- lapply(names(kegg_res), function(cc) {
  plot_kegg_ora(kegg_res[[cc]], gsub("_", " ", cc), n = 12)
})

patchwork::wrap_plots(kegg_plots, ncol = 2)
```

#--------------------------------------------------
#ElasticNet on top 50 metabolites per contrast
#a) all metabolite data
#--------------------------------------------------
```{r}
run_elastic_net <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  top_n = 50,
  covars = c("age", "sex", "study"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1,
  min_per_class = 5,
  maxit = 1e6
) {
  
  cat("\n============================\n")
  cat("Running contrast:", cc, "\n")
  
  if (!cc %in% names(out_list)) {
    stop("Contrast not found in out_list: ", cc)
  }
  
  if (!grepl("_vs_", cc)) {
    stop("Contrast name must contain '_vs_'")
  }
  
  parts <- strsplit(cc, "_vs_")[[1]]
  if (length(parts) != 2) {
    stop("Contrast could not be split into two groups")
  }
  
  g_pos <- parts[1]
  g_neg <- parts[2]
  
 idx <- which(meta_df$group4 %in% c(g_pos, g_neg))
meta_sub <- meta_df[idx, , drop = FALSE]
meta_sub$record_id <- as.character(meta_sub$record_id)

meta_sub$group4 <- droplevels(meta_sub$group4)

y <- ifelse(meta_sub$group4 == g_pos, 1, 0)
y <- as.numeric(y)

tab <- table(meta_sub$group4)
if (any(tab < min_per_class)) {
  stop(
    "Too few samples in at least one class for contrast: ",
    cc,
    " | counts = ",
    paste(names(tab), tab, collapse = ", ")
  )
}
  
  t <- out_list[[cc]]
  if (is.null(t)) {
    stop("Contrast table is NULL for: ", cc)
  }
  t <- as.data.frame(t)
  
  cat("Columns in contrast table:\n")
  print(names(t))
  
  if (!"feature_id" %in% names(t)) {
    fid <- rownames(t)
    if (is.null(fid)) stop("No feature_id column or rownames found")
    t$feature_id <- fid
  }
  
  if (!all(c("adj.P.Val", "P.Value") %in% names(t))) {
    stop("Contrast table is missing adj.P.Val and/or P.Value")
  }
  
  cat("Rows in limma table:", nrow(t), "\n")
  cat("Overlap with metabolite matrix:", sum(t$feature_id %in% colnames(X_mat)), "\n")
  
  if (!"t" %in% names(t)) {
  stop("Contrast table is missing limma test statistic column: t")
}

top_feats <- t %>%
  dplyr::filter(
    !is.na(t),
    feature_id %in% colnames(X_mat)
  ) %>%
  dplyr::mutate(abs_t = abs(t)) %>%
  dplyr::arrange(dplyr::desc(abs_t)) %>%
  dplyr::slice_head(n = top_n) %>%
  dplyr::pull(feature_id)
  
  cat("Top overlapping features:", length(top_feats), "\n")
  cat("First few top features:\n")
  print(head(top_feats, 10))
  
  if (length(top_feats) < 5) {
    stop("Fewer than 5 overlapping top features for contrast: ", cc)
  }
  
  meta_sub <- meta_sub %>%
    dplyr::mutate(
      sex = factor(sex),
      study = factor(study)
    )
  
  covars_present <- covars[covars %in% colnames(meta_sub)]
  cat("Covariates used:", paste(covars_present, collapse = ", "), "\n")
  
  if (length(covars_present) > 0) {
    form_cov <- as.formula(paste("~", paste(covars_present, collapse = " + ")))
    Z <- model.matrix(form_cov, data = meta_sub)
  } else {
    Z <- matrix(1, nrow = nrow(meta_sub), ncol = 1)
    colnames(Z) <- "(Intercept)"
  }
  
  Xp <- X_mat[meta_sub$record_id, top_feats, drop = FALSE]
  
  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"
  
  pf <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))
  
  sds <- apply(X, 2, sd)
  keep <- which(is.finite(sds) & sds > 0)
  
  X <- X[, keep, drop = FALSE]
  pf <- pf[keep]
  
  cat("Final design matrix dim:", dim(X), "\n")
  
  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))
  cat("nfolds used:", nfolds_use, "\n")
  
  if (nfolds_use < 2) {
    stop("Not enough samples per class for CV")
  }
  
  set.seed(seed)
  
  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      cv.glmnet(
        x = X,
        y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = pf,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) {
        message("glmnet failed at alpha = ", a, ": ", conditionMessage(e))
        NULL
      }
    )
    
    if (is.null(cv)) return(NULL)
    list(alpha = a, cvfit = cv)
  })
  
  fit_list <- Filter(Negate(is.null), fit_list)
  
  if (length(fit_list) == 0) {
    stop("All glmnet fits failed")
  }
  
  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]
  
  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))
  coef_df <- tibble::tibble(
    term = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>%
    dplyr::filter(beta != 0)
  
  protein_terms <- intersect(colnames(Xp), coef_df$term)
  
  p_hat <- as.numeric(
    predict(best$cvfit, newx = X, s = "lambda.1se", type = "response")
  )
  
  auc_val <- as.numeric(
    pROC::auc(pROC::roc(y, p_hat, quiet = TRUE))
  )
  
  cat("Best alpha:", best$alpha, "\n")
  cat("AUC:", auc_val, "\n")
  cat("Nonzero coefficients:", nrow(coef_df), "\n")
  cat("Selected metabolites:", length(protein_terms), "\n")
  
  list(
    contrast         = cc,
    positive         = g_pos,
    negative         = g_neg,
    best_alpha       = best$alpha,
    lambda_1se       = best$cvfit$lambda.1se,
    lambda_min       = best$cvfit$lambda.min,
    auc_train        = auc_val,
    n_input_proteins = length(top_feats),
    coef_df          = coef_df,
    protein_terms    = protein_terms,
    cvfit            = best$cvfit,
    top_feats        = top_feats
  )
}
```

```{r}
res_T1D_LC <- run_elastic_net(
  cc = "T1D_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T1D_OC <- run_elastic_net(
  cc = "T1D_vs_OC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T1D_T2D <- run_elastic_net(
  cc = "T1D_vs_T2D",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T2D_LC <- run_elastic_net(
  cc = "T2D_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T2D_OC <- run_elastic_net(
  cc = "T2D_vs_OC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_OC_LC <- run_elastic_net(
  cc = "OC_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50,
  covars = c("age", "sex")
)
```
#------------------------------------
#boostrapping function for elasticnet
#-------------------------------------
```{r}
run_elastic_net_bootstrap_metab <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  top_n = 50,
  covars = c("age", "sex"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  n_boot = 100,
  seed = 1835,
  min_per_class = 5
) {
  
  set.seed(seed)
  boot_res <- vector("list", n_boot)
  
  for (b in seq_len(n_boot)) {
    
    idx0 <- which(meta_df$group4 %in% strsplit(cc, "_vs_")[[1]])
    meta_sub0 <- meta_df[idx0, , drop = FALSE]
    
    idx_boot <- unlist(lapply(split(seq_len(nrow(meta_sub0)), meta_sub0$group4), function(ii) {
      sample(ii, size = length(ii), replace = TRUE)
    }))
    
    meta_boot <- meta_sub0[idx_boot, , drop = FALSE]
    meta_boot$record_id_boot <- paste0(meta_boot$record_id, "_boot", seq_len(nrow(meta_boot)))
    
    X_boot <- X_mat[meta_boot$record_id, , drop = FALSE]
    rownames(X_boot) <- meta_boot$record_id_boot
    meta_boot$record_id <- meta_boot$record_id_boot
    
    fit_b <- tryCatch(
      run_elastic_net(
        cc = cc,
        out_list = out_list,
        meta_df = meta_boot,
        X_mat = X_boot,
        top_n = top_n,
        covars = covars,
        seed = seed + b
      ),
      error = function(e) NULL
    )
    
    boot_res[[b]] <- fit_b
  }
  
  boot_res <- Filter(Negate(is.null), boot_res)
  
  selected_tbl <- lapply(seq_along(boot_res), function(i) {
    tibble(
      boot = i,
      contrast = cc,
      feature_id = unique(boot_res[[i]]$protein_terms)  # still named this internally
    )
  }) %>%
    bind_rows() %>%
    distinct(contrast, boot, feature_id)
  
  stability_tbl <- selected_tbl %>%
    count(contrast, feature_id, name = "n_selected") %>%
    mutate(
      n_boot_success = length(boot_res),
      selection_frequency = n_selected / n_boot_success,
      selection_percent = 100 * selection_frequency
    ) %>%
    arrange(desc(selection_frequency))
  
  list(
    contrast = cc,
    stability_tbl = stability_tbl,
    boot_res = boot_res
  )
}
```

#t1d vs lc
```{r}
boot_metab_T1D_LC <- run_elastic_net_bootstrap_metab(
  cc = "T1D_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50
)
```

```{r}
core_metab_T1D_LC <- boot_metab_T1D_LC$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_T1D_LC
```

```{r}
boot_metab_T1D_T2D <- run_elastic_net_bootstrap_metab(
  cc = "T1D_vs_T2D",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50
)
```
```{r}
core_metab_T1D_T2D <- boot_metab_T1D_T2D$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_T1D_T2D
```

```{r}
plot_stability_metab <- function(boot_obj, min_freq = 0.25, top_n = 25) {
  
  plot_df <- boot_obj$stability_tbl %>%
    filter(selection_frequency >= min_freq) %>%
    slice_max(selection_frequency, n = top_n, with_ties = FALSE)
  
  ggplot(plot_df, aes(
    x = reorder(feature_id, selection_percent),
    y = selection_percent
  )) +
    geom_col() +
    coord_flip() +
    labs(
      title = paste0("Metabolite stability: ", boot_obj$contrast),
      x = NULL,
      y = "Bootstrap selection frequency (%)"
    ) +
    theme_bw()
}
```

```{r}
plot_stability_metab(boot_metab_T1D_LC)
plot_stability_metab(boot_metab_T1D_T2D)
```
#--------------------------------------------------
# b: Metabolomic + Proteomic Samples 
#--------------------------------------------------
```{r}
obj <- readRDS(
  file.path(root_path, "Savanah Leidholt", "TRIAD", "Data", "omics_split_objects_clean.rds")
)

prot_log <- readRDS(
  file.path(root_path, "Savanah Leidholt", "TRIAD", "Data", "prot_log2.rds")
)

prot_plasma <- obj$prot_plasma
metabolites <- obj$metabolites
meta_prot <- obj$meta_prot
meta_metab <- obj$meta_metab

# clean rownames
rownames(prot_plasma) <- trimws(as.character(rownames(prot_plasma)))
rownames(prot_log)    <- trimws(as.character(rownames(prot_log)))
rownames(metabolites) <- trimws(as.character(rownames(metabolites)))

# IDs with any measured data in each omics matrix
ids_metab <- rownames(metabolites)[rowSums(!is.na(metabolites)) > 0]
ids_prot  <- rownames(prot_plasma)[rowSums(!is.na(prot_plasma)) > 0]
ids_log   <- rownames(prot_log)[rowSums(!is.na(prot_log)) > 0]

ids_metab <- trimws(as.character(ids_metab))
ids_prot  <- trimws(as.character(ids_prot))
ids_log   <- trimws(as.character(ids_log))

# paired IDs present in metabolomics, raw proteomics, and logged proteomics
paired_ids <- Reduce(intersect, list(ids_metab, ids_prot, ids_log))
```

```{r}
# HELPER: Study x Group count table for one omics type
make_count_table <- function(meta_df, mat_df, omics_label) {
  
  count_samples_with_data(meta_df, mat_df) %>%
    count(study, group) %>%
    pivot_wider(
      names_from = group,
      values_from = n,
      values_fill = 0
    ) %>%
    arrange(study) %>%
    mutate(Omics = omics_label, .before = 1)
}

make_count_table_group4 <- function(meta_df, omics_label = "Paired metabolomics + proteomics") {
  
  meta_df %>%
    count(study6, group4) %>%
    pivot_wider(
      names_from = group4,
      values_from = n,
      values_fill = 0
    ) %>%
    mutate(
      Total = rowSums(across(any_of(c("LC", "OC", "T1D", "T2D")))),
      Omics = omics_label,
      .before = 1
    ) %>%
    arrange(study6)
}
```

##---------------------------------------
## samples counts Study x Group
##---------------------------------------
```{r}
meta_mp <- meta_metab %>%
  mutate(
    record_id = trimws(as.character(record_id)),
    study     = trimws(as.character(study)),
    group     = trimws(as.character(group)),
    sex       = trimws(as.character(sex)),
    ethnicity = trimws(as.character(race_ethnicity_condensed)),
    age       = suppressWarnings(as.numeric(age)),
    uacr      = suppressWarnings(as.numeric(acr_u)),
    log_uacr  = log(uacr + 1)
  ) %>%
  filter(
    record_id %in% paired_ids,
    !is.na(age),
    !is.na(sex),
    sex != ""
  ) %>%
  mutate(
    group4 = factor(
      group,
      levels = c("Lean Control", "Obese Control", "Type 1 Diabetes", "Type 2 Diabetes"),
      labels = c("LC", "OC", "T1D", "T2D")
    ),
    ethnicity = factor(ethnicity),
    study6 = dplyr::recode(
      study,
      "ATTEMPT"          = "ATT",
      "CROCODILE"        = "CRC",
      "PANTHER"          = "PAN",
      "RENAL-HEIR"       = "RH",
      "IMPROVE"          = "IT",
      "RENAL-HEIRitage"  = "RH2",
      .default = study
    )
  ) %>%
  filter(study6 != "ATT") %>%
  mutate(study6 = factor(study6)) %>%
  arrange(record_id) %>%
  droplevels()
```

```{r}
metab_log <- as.matrix(log2(metabolites + 1))
metab_log[!is.finite(metab_log)] <- NA_real_
saveRDS(
  metab_log,
  file = file.path(
    root_path,
    "Savanah Leidholt",
    "TRIAD",
    "Data",
    "metabolomics_paired_matrix.rds"
  )
)
```

```{r}
paired_ids_final <- meta_mp$record_id

prot_log6 <- prot_log[paired_ids_final, , drop = FALSE]
metab_mp  <- metab_log[paired_ids_final, , drop = FALSE]

stopifnot(identical(rownames(prot_log6), meta_mp$record_id))
stopifnot(identical(rownames(metab_mp), meta_mp$record_id))
```

```{r}
cat("Paired metabolomics + proteomics N:", nrow(meta_mp), "\n")
cat("Proteins:", nrow(prot_log6), "\n")
cat("Metabolites:", nrow(metab_mp), "\n")

meta_mp %>%
  count(study6, group4) %>%
  pivot_wider(
    names_from = group4,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(study6)
```
```{r}
tbl_mp <- meta_mp %>%
  count(study6, group4) %>%
  pivot_wider(
    names_from = group4,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(Total = rowSums(across(any_of(c("LC", "OC", "T1D", "T2D"))))) %>%
  arrange(study6)

tbl_mp
```
#------------------------------------------------------
#study level filter
#just a sanity check shouldn't filter anything else out
#-------------------------------------------------------
```{r}
study_idx <- split(seq_len(nrow(metab_mp)), meta_mp$study)
present_by_study <- vapply(
  study_idx,
  FUN = function(idx) {
    colMeans(!is.na(metab_mp[idx, , drop = FALSE])) >= 0.6
  },
  FUN.VALUE = setNames(logical(ncol(metab_mp)), colnames(metab_mp))
)

present_by_study[is.na(present_by_study)] <- FALSE

n_studies_present <- rowSums(present_by_study)

keep_metabolites <- n_studies_present >= 2

kept_metabolites <- rownames(present_by_study)[keep_metabolites]
removed_metabolites <- rownames(present_by_study)[!keep_metabolites]

cat("Removed metabolites:", length(removed_metabolites), "\n") #92
cat("Kept metabolites:", length(kept_metabolites), "\n") #165

mat_filt <- metab_mp[, kept_metabolites, drop = FALSE]
cat("particpant n:", nrow(mat_filt))
```

#-------------------------
#checking for low variance
#-------------------------
```{r}
nzv_idx <- nearZeroVar(mat_filt)

mat_filt2 <- if (length(nzv_idx) > 0) {
 mat_filt[, -nzv_idx, drop = FALSE]
} else {
  mat_filt
}

cat("Features before filtering:", ncol(mat_filt), "\n") #165
cat("Features after filtering:", ncol(mat_filt2), "\n") #165
cat("Participants:", nrow(mat_filt2), "\n") #122

stopifnot(identical(rownames(mat_filt2), meta_mp$record_id))
```
#-----------------
#imputing for pcas
#-----------------
```{r}
mat_pca <- mat_filt2

# convert Inf / -Inf / NaN to NA
mat_pca[!is.finite(mat_pca)] <- NA_real_

# median impute each metabolite
impute_col_median <- function(x) {
  med <- median(x, na.rm = TRUE)
  x[is.na(x)] <- med
  x
}

mat_pca <- apply(mat_pca, 2, impute_col_median)
mat_pca <- as.matrix(mat_pca)

# checks
stopifnot(identical(rownames(mat_pca), meta_mp$record_id))
stopifnot(!anyNA(mat_pca))
stopifnot(all(is.finite(mat_pca)))

cat("Missing values before PCA imputation:", sum(is.na(mat_log_filt)), "\n") #1050
cat("Missing values after PCA imputation:", sum(is.na(mat_pca)), "\n") #0
```

##---------------------------
## PCA-based OUTLIER REMOVAL
##---------------------------
```{r}
pca_all <- prcomp(mat_pca, center = TRUE, scale. = TRUE)

pca_dist_all <- sqrt(rowSums(pca_all$x[, 1:5, drop = FALSE]^2))
cutoff_all <- median(pca_dist_all) + 3 * mad(pca_dist_all, constant = 1)

outlier_df_all <- meta_mp %>%
  mutate(
    pca_dist = pca_dist_all,
    pca_outlier = pca_dist > cutoff_all
  ) %>%
  arrange(desc(pca_dist))

outlier_df_all %>%
  filter(pca_outlier) %>%
  select(record_id, study, group4, pca_dist) #RH-48-T, RH-49-T, CRC-11, CRC-50, CRC-42, CRC-38

keep_ids_all <- outlier_df_all %>%
  filter(!pca_outlier) %>%
  pull(record_id)

mat_no_outliers <- mat_pca[keep_ids_all, , drop = FALSE] #no outlier matrix for pca
mat_filt3 <- mat_filt2[keep_ids_all, , drop = FALSE] # no outlier matrix for limma

meta_mp2 <- meta_mp %>%
  filter(record_id %in% keep_ids_all) %>%
  arrange(match(record_id, keep_ids_all))

stopifnot(identical(meta_mp2$record_id, rownames(mat_no_outliers)))

cat("All metabolomics features kept for PCA/LIMMA:", ncol(mat_no_outliers), "\n")
cat("Number of participants kept:", nrow(mat_no_outliers), "\n")
```

##---------------------------------------
## PCA: METABOLOMICS + PROTEOMICS SUBSET
##---------------------------------------
```{r}
p0_mp <- make_pca_plot(
  mat_pca,
  meta_mp,
  "Metabolomics - No adjustment"
)

cov1_all <- model.matrix(~ age + sex, data = meta_mp)[, -1, drop = FALSE]

mat_adj1_all <- t(limma::removeBatchEffect(
  t(mat_pca),
  covariates = cov1_all
))

p1_mp <- make_pca_plot(
  mat_adj1_all,
  meta_mp,
  "Metabolomics - Age + Sex"
)

cov2_all <- model.matrix(~ age + sex + study, data = meta_mp)[, -1, drop = FALSE]

mat_adj2_all <- t(limma::removeBatchEffect(
  t(mat_pca),
  covariates = cov2_all
))

p2_mp <- make_pca_plot(
  mat_adj2_all,
  meta_mp,
  "Metabolomics - Age + Sex + Study"
)

(p0_mp | p1_mp | p2_mp) +
  plot_annotation(
    title = "PCA: metabolomics/proteomics subset",
    subtitle = "Color = group | Shape = study "
  )
```
```{r}
p0_No <- make_pca_plot(
  mat_no_outliers,
  meta_mp2,
  "Metabolomics - No adjustment"
)

cov1_all <- model.matrix(~ age + sex, data = meta_mp2)[, -1, drop = FALSE]

mat_adj1_all <- t(limma::removeBatchEffect(
  t(mat_no_outliers),
  covariates = cov1_all
))

p1_No <- make_pca_plot(
  mat_adj1_all,
  meta_mp2,
  "Metabolomics - Age + Sex"
)

cov2_all <- model.matrix(~ age + sex + study, data = meta_mp2)[, -1, drop = FALSE]

mat_adj2_all <- t(limma::removeBatchEffect(
  t(mat_no_outliers),
  covariates = cov2_all
))

p2_No <- make_pca_plot(
  mat_adj2_all,
  meta_mp2,
  "Metabolomics - Age + Sex + Study"
)

(p0_No | p1_No | p2_No) +
  plot_annotation(
    title = "PCA: metabolomics/proteomics subset samples--no outliers",
    subtitle = "Color = group | Shape = study "
  )

(p2_mp |p2_No) +
  plot_annotation(
    title = "PCA: outlier comparison",
    subtitle = "Color = group | Shape = study "
  )
```

##---------------------------------------
##LIMMA: METABOLOMICS + PROTEOMICS SUBSET
##---------------------------------------
```{r}
Y_mp <- t(as.matrix(mat_filt2))

design_mp <- model.matrix(~ 0 + group4 + age + sex + study, data = meta_mp)
colnames(design_mp) <- sub("^group4", "", colnames(design_mp))
colnames(design_mp) <- make.names(colnames(design_mp))

contr_mp <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_mp
)

fit_mp <- lmFit(Y_mp, design_mp)
fit_mp <- contrasts.fit(fit_mp, contr_mp)
fit_mp <- eBayes(fit_mp)

out_mp <- list()
for (cc in colnames(contr_mp)) {
  tt <- topTable(fit_mp, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
  tt$feature_id <- rownames(tt)
  tt$contrast <- cc
  out_mp[[cc]] <- tt
}

sig_mp <- tibble(
  contrast = names(out_mp),
  n_FDR05 = sapply(out_mp, function(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_mp, function(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
  model = "metabolomics_age_sex_study_metab_prot_subset",
  n_samples = nrow(meta_mp)
)

print(sig_mp)
```
```{r}
prep_tables <- function(out_list) {
  lapply(out_list, function(tt) {
    tt %>%
      dplyr::select(
        feature_id,
        logFC,
        AveExpr,
        t,
        P.Value,
        adj.P.Val,
        B,
        contrast
      ) %>%
      dplyr::arrange(desc(t))
  })
}

out_mp_table <- prep_tables(out_mp)

contrast_order <- c(
  "T1D_vs_LC",
  "T2D_vs_LC",
  "T1D_vs_T2D",
  "OC_vs_LC",
  "T1D_vs_OC",
  "T2D_vs_OC"
)


limma_metab_inputs <- list(
  metadata = list(
    created = Sys.time(),
    response = "plasma metabolomics",
    transform = "log2(x + 1)",
    model_formula = "0 + group4 + age + sex + study",
    contrasts = contrast_order
  ),

  all_omics = list(
    model_name = "metabolomics_age_sex_study_mp_subset",
    samples = meta_mp,
    matrix = mat_filt2,
    design = design_mp,
    fit = fit_mp,
    out = out_mp,
    sig_summary = sig_mp
  )
)


saveRDS(
  limma_metab_inputs,
  file = file.path(
    root_path,
    "Savanah Leidholt",
    "TRIAD",
    "Data",
    "limma_metabolomics_inputs_age_sex_study.rds"
  )
)

cat("Saved: limma_metabolomics_inputs_age_sex_study.rds\n")
```



##-----------------------------------------------
##VOLCANO PLOTS: METABOLOMICS + PROTEOMICS SUBSET
##------------------------------------------------
```{r}
volcano_list_mp <- lapply(contrast_names, function(cc) {
  tt <- out_mp[[cc]]

  EnhancedVolcano(
    tt,
    lab = tt$feature_id,
    x = "logFC",
    y = "adj.P.Val",
    pCutoff = 0.05,
    FCcutoff = 0.5,
    title = gsub("_", " ", cc),
    subtitle = "Age + sex + study (metabolomics samples with proteomics overlap)",
    pointSize = 2.0,
    labSize = 3.0
  )
})

wrap_plots(volcano_list_mp, ncol = 3)
```


##---------------------------------------
##LIMMA: METABOLOMICS + PROTEOMICS SUBSET
## without outliers sensitivity check
##---------------------------------------
```{r}
Y_mp2 <- t(as.matrix(mat_filt3))

design_mp2 <- model.matrix(~ 0 + group4 + age + sex + study, data = meta_mp2)
colnames(design_mp2) <- sub("^group4", "", colnames(design_mp2))
colnames(design_mp2) <- make.names(colnames(design_mp2))

contr_mp <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_mp2
)

fit_mp2 <- lmFit(Y_mp2, design_mp2)
fit_mp2 <- contrasts.fit(fit_mp2, contr_mp)
fit_mp2 <- eBayes(fit_mp2)

out_mp2 <- list()
for (cc in colnames(contr_mp)) {
  tt <- topTable(fit_mp2, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
  tt$feature_id <- rownames(tt)
  tt$contrast <- cc
  out_mp2[[cc]] <- tt
}

sig_mp2 <- tibble(
  contrast = names(out_mp2),
  n_FDR05 = sapply(out_mp2, function(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_mp2, function(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
  model = "metabolomics_age_sex_study_metab_prot_subset",
  n_samples = nrow(meta_mp2)
)

print(sig_mp2)
```
##-----------------------------------------------
##VOLCANO PLOTS: METABOLOMICS + PROTEOMICS SUBSET
##------------------------------------------------
```{r}
volcano_list_mp2 <- lapply(contrast_names, function(cc) {
  tt <- out_mp2[[cc]]

  EnhancedVolcano(
    tt,
    lab = tt$feature_id,
    x = "logFC",
    y = "adj.P.Val",
    pCutoff = 0.05,
    FCcutoff = 0.5,
    title = gsub("_", " ", cc),
    subtitle = "Age + sex + study (metabolomics samples with proteomics overlap)",
    pointSize = 2.0,
    labSize = 3.0
  )
})

wrap_plots(volcano_list_mp2, ncol = 3)
```

```{r}
out_mp_df <- bind_rows(out_mp, .id = "contrast_from_list") %>%
  select(-contrast_from_list)

out_mp2_df <- bind_rows(out_mp2, .id = "contrast_from_list") %>%
  select(-contrast_from_list)

sig_features_mp <- out_mp_df %>%
  filter(adj.P.Val < 0.05) %>%
  select(
    contrast,
    feature_id,
    logFC,
    P.Value,
    adj.P.Val
  )

sig_features_mp2 <- out_mp2_df %>%
  filter(adj.P.Val < 0.05) %>%
  select(
    contrast,
    feature_id,
    logFC,
    P.Value,
    adj.P.Val
  )

overlap_mp <- sig_features_mp %>%
  inner_join(
    sig_features_mp2,
    by = c("contrast", "feature_id"),
    suffix = c("_with_outliers", "_without_outliers")
  ) %>%
  mutate(
    same_direction =
      sign(logFC_with_outliers) == sign(logFC_without_outliers),
    delta_logFC =
      logFC_without_outliers - logFC_with_outliers,
    delta_adjP =
      adj.P.Val_without_outliers - adj.P.Val_with_outliers
  )

overlap_summary_mp <- overlap_mp %>%
  group_by(contrast) %>%
  summarise(
    n_overlap = n(),
    n_same_direction = sum(same_direction, na.rm = TRUE),
    n_opposite_direction = sum(!same_direction, na.rm = TRUE),
    .groups = "drop"
  )

overlap_summary_mp

lost_after_outlier_removal <- sig_features_mp %>%
  anti_join(
    sig_features_mp2,
    by = c("contrast", "feature_id")
  ) %>%
  mutate(status = "Lost after outlier removal")

gained_after_outlier_removal <- sig_features_mp2 %>%
  anti_join(
    sig_features_mp,
    by = c("contrast", "feature_id")
  ) %>%
  mutate(status = "Gained after outlier removal")

overlap_mp %>%
  arrange(contrast, adj.P.Val_with_outliers) %>%
  select(
    contrast,
    feature_id,
    logFC_with_outliers,
    adj.P.Val_with_outliers,
    logFC_without_outliers,
    adj.P.Val_without_outliers,
    same_direction,
    delta_logFC,
    delta_adjP
  )

overlap_mp
lost_after_outlier_removal
gained_after_outlier_removal
```

##---------------------------------------------
## b. ORA on metabolite data with proteomics dat
##----------------------------------------------
```{r}
kegg_res_mp <- lapply(out_mp, function(tt) {
  run_kegg_metab_ora(
    tt = tt,
    pathway_map = pathway2compound,
    p_cutoff = 0.05,
    fc_cutoff = 0
  )
})

names(kegg_res_mp) <- names(out_mp)

kegg_summary_mp <- bind_rows(lapply(names(kegg_res_mp), function(cc) {
  x <- kegg_res_mp[[cc]]
  tibble(
    contrast = cc,
    n_pathways_FDR05 = sum(x$padj < 0.05, na.rm = TRUE),
    n_pathways_p05   = sum(x$p_value < 0.05, na.rm = TRUE)
  )
}))

kegg_summary_mp
```

```{r}
kegg_res_mp[["T1D_vs_T2D"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(5)
kegg_res_mp[["T2D_vs_LC"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(5)
kegg_res_mp[["OC_vs_LC"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(5)
kegg_res_mp[["T1D_vs_OC"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(20)
```

```{r}
plot_kegg_ora(kegg_res_mp[["OC_vs_LC"]], "KEGG ORA: OC vs LC")
```

```{r}
kegg_plots_mp <- lapply(names(kegg_res_mp), function(cc) {
  plot_kegg_ora(kegg_res_mp[[cc]], gsub("_", " ", cc), n = 12)
})

patchwork::wrap_plots(kegg_plots_mp, ncol = 2)
```
#-------------------------------------------
#ElasticNet for metabolomic/proteomic subset
#t1d vs lc
#-------------------------------------------
```{r}
boot_metab_T1D_LC <- run_elastic_net_bootstrap_metab(
  cc = "T1D_vs_LC",
  out_list = out_mp,
  meta_df = meta_mp,
  X_mat = mat_filt2,
  top_n = 50
)
```

```{r}
core_metab_T1D_LC <- boot_metab_T1D_LC$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_T1D_LC
```
```{r}
boot_metab_T1D_OC <- run_elastic_net_bootstrap_metab(
  cc = "T1D_vs_OC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50
)
```

```{r}
core_metab_T1D_OC <- boot_metab_T1D_OC$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_T1D_OC
```

```{r}
boot_metab_T1D_T2D <- run_elastic_net_bootstrap_metab(
  cc = "T1D_vs_T2D",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_log_filt,
  top_n = 50
)
```

```{r}
core_metab_T1D_T2D <- boot_metab_T1D_T2D$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_T1D_T2D
```

```{r}
boot_metab_T2D_LC <- run_elastic_net_bootstrap_metab(
  cc = "T2D_vs_LC",
  out_list = out_mp,
  meta_df = meta_mp,
  X_mat = mat_filt2,
  top_n = 50
)
```

```{r}
core_metab_T2D_LC <- boot_metab_T2D_LC$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_T2D_LC
```

```{r}
boot_metab_T2D_OC <- run_elastic_net_bootstrap_metab(
  cc = "T2D_vs_OC",
  out_list = out_mp,
  meta_df = meta_mp,
  X_mat = mat_filt2,
  top_n = 50
)
```

```{r}
core_metab_T2D_OC <- boot_metab_T2D_OC$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_T2D_OC
```


```{r}
boot_metab_OC_LC <- run_elastic_net_bootstrap_metab(
  cc = "OC_vs_LC",
  out_list = out_mp,
  meta_df = meta_mp,
  X_mat = mat_filt2,
  top_n = 50
)
```

```{r}
core_metab_OC_LC <- boot_metab_OC_LC$stability_tbl %>%
  filter(selection_frequency >= 0.6)
core_metab_OC_LC
```

```{r}
plot_stability_metab <- function(boot_obj, min_freq = 0.25, top_n = 25) {
  
  plot_df <- boot_obj$stability_tbl %>%
    filter(selection_frequency >= min_freq) %>%
    slice_max(selection_frequency, n = top_n, with_ties = FALSE)
  
  ggplot(plot_df, aes(
    x = reorder(feature_id, selection_percent),
    y = selection_percent
  )) +
    geom_col() +
    coord_flip() +
    labs(
      title = paste0("Metabolite stability: ", boot_obj$contrast),
      x = NULL,
      y = "Bootstrap selection frequency (%)"
    ) +
    theme_bw()
}
```

```{r}
plot_stability_metab(boot_metab_T1D_LC)
plot_stability_metab(boot_metab_T1D_T2D)
```
```

