---
title: "Metabolomics Data Analysis"
author: "Savanah leidholt"
date: "2026-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Directories
#computer <- "mac studio"
computer <- "mac laptop"

user <- Sys.info()[["user"]]

if (computer == "mac studio") {
  
  if (user == "sleidholt") {
    dir.dat <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
    
  } else if (user == "hhampson") {
    dir.dat <- "/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/hhampson/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/PANTHER/Results"
    git_path <- "/Users/hhampson/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
    
  } else if (user == "choiyej") {
    dir.dat <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "pylell") {
    dir.dat <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  }

} else if (computer == "mac laptop") {

  if (user == "savanahleidholt") {
    dir.dat <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/T1D-T2D/Results/Metabolomics"
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "hhampson") {
    dir.dat <- "/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive"
    dir.results <- "/Users/hhampson/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/PANTHER/Results"
    git_path <- "/Users/hhampson/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
  }
}
```

```{r}
#load libraries
library(dplyr)
library(tidyr)
library(tibble)
library(limma)
library(ggplot2)
library(EnhancedVolcano)
library(patchwork)
library(stringr)
library(purrr)
library(KEGGREST)
library(caret)
library(glmnet)
```
#------------------------------------
# raw metabolomics QC/summary plots
#------------------------------------
```{r}
omics_objects <- readRDS(file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "omics_split_objects3.rds"))
meta_qc <- omics_objects$meta_metab
mat_qc  <- omics_objects$metabolites

meta_qc$record_id <- trimws(as.character(meta_qc$record_id))
meta_qc$study     <- trimws(as.character(meta_qc$study))
meta_qc$group     <- trimws(as.character(meta_qc$group))

meta_qc$study6 <- dplyr::recode(
  meta_qc$study,
  "ATTEMPT"         = "ATT",
  "CROCODILE"       = "CRC",
  "RENAL-HEIR"      = "RH",
  "IMPROVE"         = "IT",
  "PANTHER" = "PANTH",
  "RENAL-HEIRitage" = "RH2",
  .default = meta_qc$study
)

rownames(mat_qc) <- trimws(as.character(rownames(mat_qc)))

# keep only overlapping samples
ids_qc <- intersect(meta_qc$record_id, rownames(mat_qc))
meta_qc <- meta_qc[match(ids_qc, meta_qc$record_id), , drop = FALSE]
mat_qc  <- mat_qc[ids_qc, , drop = FALSE]

stopifnot(identical(meta_qc$record_id, rownames(mat_qc)))

cat("QC samples:", nrow(mat_qc), "\n")
cat("QC metabolites:", ncol(mat_qc), "\n")

```
#--------------------
#PCA plotting Function
#--------------------
```{r}
make_pca_plot <- function(mat_in, meta_df, title_str) {
  pca <- prcomp(mat_in, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1 = pca$x[, 1],
    PC2 = pca$x[, 2],
    group4 = meta_df$group4,
    study  = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = group4, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        LC  = "#2166AC",
        OC  = "#F4A582",
        T1D = "#D6604D",
        T2D = "#4DAC26"
      )
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
## ----------------------
## sample level coverage
## ----------------------
```{r}
sample_qc <- meta_qc %>%
  mutate(
    n_metabolites_detected = rowSums(!is.na(mat_qc)),
    n_metabolites_missing  = rowSums(is.na(mat_qc)),
    pct_detected           = 100 * n_metabolites_detected / ncol(mat_qc),
    pct_missing            = 100 * n_metabolites_missing / ncol(mat_qc)
  )

# boxplot: number of detected metabolites per sample by study
p_cov_n <- ggplot(sample_qc, aes(x = study6, y = n_metabolites_detected, fill = study6)) +
  scale_fill_brewer(palette = "Dark2") + 
  geom_boxplot(alpha = 0.75, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
  theme_bw() +
  labs(
    title = "Raw metabolite coverage per sample",
    x = "Study",
    y = "Number of detected metabolites"
  ) +
  theme(legend.position = "none")

# boxplot: percent missing per sample by study
p_miss_sample <- ggplot(sample_qc, aes(x = study6, y = pct_missing, fill = study6)) +
  scale_fill_brewer(palette = "Dark2") +
  geom_boxplot(alpha = 0.75, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
  theme_bw() +
  labs(
    title = "Percent missingness per sample",
    x = "Study",
    y = "% missing metabolites"
  ) +
  theme(legend.position = "none")

p_cov_n | p_miss_sample
```
##---------------------------------
## study level coverage
##---------------------------------
```{r}
study_idx <- split(seq_len(nrow(mat_qc)), meta_qc$study6)

present_by_study <- sapply(study_idx, function(idx) {
  colSums(!is.na(mat_qc[idx, , drop = FALSE])) > 0
})

present_by_study <- as.matrix(present_by_study)

study_coverage <- tibble(
  study6 = colnames(present_by_study),
  n_metabolites_present = colSums(present_by_study),
  pct_metabolites_present = 100 * n_metabolites_present / nrow(present_by_study)
)

p_study_cov <- ggplot(study_coverage, aes(x = study6, y = pct_metabolites_present, fill = study6)) +
  scale_fill_brewer(palette = "Dark2")+
  geom_col(alpha = 0.85) +
  geom_text(aes(label = round(pct_metabolites_present, 1)), vjust = -0.3, size = 3.5) +
  theme_bw() +
  labs(
    title = "Raw metabolite coverage in each study",
    x = "Study",
    y = "% of metabolites observed in at least one sample"
  ) +
  ylim(0, 105) +
  theme(legend.position = "none")

p_study_cov

```
# ----------------------------------------------------
# METABOLOMICS-ONLY ANALYSIS WORKFLOW
# a) all metabolomics saples
# b) metabolomics samples that also have proteomics
#-----------------------------------------------------
```{r}
contrast_names <- c(
  "T1D_vs_LC", "T2D_vs_LC", "T1D_vs_T2D",
  "OC_vs_LC", "T1D_vs_OC", "T2D_vs_OC"
)
```
#------------------
# a. metabolomics only
#------------------
```{r}
meta_metab <- omics_objects$meta_metab
mat_metab  <- omics_objects$metabolites

meta_metab$record_id <- trimws(as.character(meta_metab$record_id))
meta_metab$study     <- trimws(as.character(meta_metab$study))
meta_metab$group     <- trimws(as.character(meta_metab$group))
meta_metab$sex       <- trimws(as.character(meta_metab$sex))

meta_metab$age <- suppressWarnings(as.numeric(meta_metab$age))
meta_metab$bmi <- suppressWarnings(as.numeric(meta_metab$bmi))

meta_metab <- meta_metab[
  !is.na(meta_metab$age) & !is.na(meta_metab$sex) & meta_metab$sex != "",
  ,
  drop = FALSE
]

meta_metab$group4 <- factor(
  meta_metab$group,
  levels = c("Lean Control", "Obese Control", "Type 1 Diabetes", "Type 2 Diabetes"),
  labels = c("LC", "OC", "T1D", "T2D")
)

# recode study names directly into study
meta_metab$study <- dplyr::recode(
  meta_metab$study,
  "ATTEMPT"         = "ATT",
  "CROCODILE"       = "CRC",
  "RENAL-HEIR"      = "RH",
  "IMPROVE"         = "IT",
  "PANTHER" = "PANTH",
  "RENAL-HEIRitage" = "RH2",
  .default = meta_metab$study
)

# remove ATTEMPT from all downstream analyses
meta_metab <- meta_metab %>%
  filter(!study %in% c("ATT", "PANTH"))

meta_metab$study <- factor(meta_metab$study)
meta_metab$sex   <- factor(meta_metab$sex)

rownames(mat_metab) <- trimws(as.character(rownames(mat_metab)))

# keep only IDs remaining after ATT removal
keep_ids <- intersect(meta_metab$record_id, rownames(mat_metab))

meta_metab <- meta_metab[match(keep_ids, meta_metab$record_id), , drop = FALSE]
mat_metab  <- mat_metab[keep_ids, , drop = FALSE]

stopifnot(identical(meta_metab$record_id, rownames(mat_metab)))

cat("ATT & PANTH removed from all analyses\n")
cat("Remaining samples:", nrow(meta_metab), "\n")
cat("Remaining metabolites:", ncol(mat_metab), "\n")
print(table(meta_metab$study, useNA = "ifany"))
```
#--------------
#IDS WITH DATA
#--------------
```{r}
ids_metab <- rownames(mat_metab)[rowSums(!is.na(mat_metab)) > 0]

prot_tmp <- omics_objects$prot_plasma
rownames(prot_tmp) <- trimws(as.character(rownames(prot_tmp)))
ids_prot <- rownames(prot_tmp)[rowSums(!is.na(prot_tmp)) > 0]

ids_metab <- trimws(as.character(ids_metab))
ids_prot  <- trimws(as.character(ids_prot))

ids_metab_prot <- intersect(ids_metab, ids_prot)

cat("Total metabolomics IDs with data after ATT & PANTH removal:", length(ids_metab), "\n")
cat("Total metabolomics + proteomics overlap IDs:", length(ids_metab_prot), "\n")

```
#----------------------------
# a) ALL METABOLOMICS SAMPLES
#----------------------------
```{r}
ids_use_all <- intersect(meta_metab$record_id, rownames(mat_metab))

meta_all <- meta_metab[match(ids_use_all, meta_metab$record_id), , drop = FALSE]
mat_all  <- mat_metab[ids_use_all, , drop = FALSE]

stopifnot(identical(meta_all$record_id, rownames(mat_all)))

cat("All metabolomics analysis-ready N:", nrow(meta_all), "\n")
cat("All metabolomics features before filtering:", ncol(mat_all), "\n")
print(table(meta_all$study, meta_all$group4, useNA = "ifany"))
```
```{r}
##filtering step where protein must exist in 60% of samples in a study &b be in 2 studies
study_idx_all <- split(seq_len(nrow(mat_all)), meta_metab$study)
present_by_study_all <- sapply(study_idx_all, function(idx) {
  colMeans(!is.na(mat_all[idx, ])) >= 0.6
})

present_by_study_all <- as.matrix(present_by_study_all)
n_studies_present_all <- colSums(present_by_study_all)
keep_study_all <- n_studies_present_all >= 2

removed_metabolites <- colnames(mat_all)[!keep_study_all]
kept_metabolites    <- colnames(mat_all)[keep_study_all]

cat("All metabolomic features before study filter:", ncol(mat_all), "\n")
cat("Removed (<2 studies meeting 60% presence threshold):", length(removed_metabolites), "\n")
cat("Kept (>=2 studies meeting 60% presence threshold):", length(kept_metabolites), "\n")

mat_filt <- mat_all[, kept_metabolites, drop = FALSE]
```

##--------------------------
## log transform & imputation
##-------------------
```{r}
mat_filt <- as.matrix(mat_filt)
rownames(mat_filt) <- trimws(as.character(rownames(mat_filt)))

mat_log_all <- log2(mat_filt + 1)

mat_log_all[!is.finite(mat_log_all)] <- NA_real_

meds_all <- apply(mat_filt, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

mat_imp_all <- mat_filt

for (j in seq_len(ncol(mat_imp_all))) {
  if (!is.na(meds_all[j])) {
    mat_imp_all[is.na(mat_imp_all[, j]), j] <- meds_all[j]
  }
}

vars_all <- apply(mat_imp_all, 2, var, na.rm = TRUE)

keep_cols_all <- colSums(is.na(mat_imp_all)) == 0 &
  vars_all > 0 &
  is.finite(vars_all)

mat_imp_all <- mat_imp_all[, keep_cols_all, drop = FALSE]

cat("All metabolomics features kept for PCA/LIMMA:", ncol(mat_imp_all), "\n")
cat("number of participants kept", nrow(mat_imp_all), "\n")
stopifnot(identical(rownames(mat_imp_all), meta_metab$record_id))
```

```{r}
#removing zero variance features in the imputed matrix
nzv_idx <- nearZeroVar(mat_imp_all)

length(nzv_idx)   # how many features will be removed
colnames(mat_imp_all)[nzv_idx]  #view removed features

mat_imp_filt <- mat_imp_all[, -nzv_idx, drop = FALSE]
```

##---------------------------
## PCA-based outlier removal
##---------------------------
```{r}
pca_all <- prcomp(mat_imp_filt, center = TRUE, scale. = TRUE)
pca_dist_all <- sqrt(rowSums(pca_all$x[, 1:5]^2))

cutoff_all <- median(pca_dist_all) + 3 * mad(pca_dist_all, constant = 1)

outlier_df_all <- meta_all %>%
  dplyr::mutate(
    pca_dist = pca_dist_all,
    pca_outlier = pca_dist > cutoff_all
  ) %>%
  dplyr::arrange(dplyr::desc(pca_dist))

# inspect outliers
outlier_df_all %>%
  dplyr::filter(pca_outlier) %>%
  dplyr::select(record_id, study, group4, pca_dist)

# remove outliers
keep_ids_all <- outlier_df_all %>%
  dplyr::filter(!pca_outlier) %>%
  dplyr::pull(record_id)

# subset matrix
mat_imp_filt <- mat_imp_filt[keep_ids_all, , drop = FALSE]

# reorder metadata to match matrix
meta_all <- meta_all[match(keep_ids_all, meta_all$record_id), , drop = FALSE]
cat("All metabolomics features kept for PCA/LIMMA:", ncol(mat_imp_all), "\n")
cat("number of participants kept", nrow(mat_imp_all), "\n")

stopifnot(identical(meta_all$record_id, rownames(mat_imp_filt)))
```

##-----------------------
## PCA: all metabolomics
##-----------------------
```{r}
p0_all <- make_pca_plot(mat_imp_filt, meta_all, "Metabolomics - No adjustment")

cov1_all <- model.matrix(~ age + sex, data = meta_all)[, -1, drop = FALSE]
mat_adj1_all <- t(removeBatchEffect(t(mat_imp_filt), covariates = cov1_all))
p1_all <- make_pca_plot(mat_adj1_all, meta_all, "Metabolomics - Age + Sex")

cov2_all <- model.matrix(~ age + sex + study, data = meta_all)[, -1, drop = FALSE]
mat_adj2_all <- t(removeBatchEffect(t(mat_imp_filt), covariates = cov2_all))
p2_all <- make_pca_plot(mat_adj2_all, meta_all, "Metabolomics - Age + Sex + Study")

(p0_all | p1_all | p2_all) +
  plot_annotation(
    title = "PCA: All metabolomics samples",
    subtitle = "Color = group | Shape = study | removeBatchEffect used for visualization only"
  )
```
##------------------------
## LIMMA: all metabolomics
## ------------------------
```{r}
Y_all <- t(as.matrix(mat_imp_filt))

design_all <- model.matrix(~ 0 + group4 + age + sex + study, data = meta_all)
colnames(design_all) <- sub("^group4", "", colnames(design_all))
colnames(design_all) <- make.names(colnames(design_all))

contr_all <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_all
)

fit_all <- lmFit(Y_all, design_all)
fit_all <- contrasts.fit(fit_all, contr_all)
fit_all <- eBayes(fit_all)

out_all <- list()
for (cc in colnames(contr_all)) {
  tt <- topTable(fit_all, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
  tt$feature_id <- rownames(tt)
  tt$contrast <- cc
  out_all[[cc]] <- tt
}

sig_all <- tibble(
  contrast = names(out_all),
  n_FDR05 = sapply(out_all, function(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_all, function(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
  model = "metabolomics_age_sex_study_all",
  n_samples = nrow(meta_all)
)

print(sig_all)
```
##--------------------------------
## VOLCANO PLOTS--all metabolomics
##---------------------------------
```{r}
volcano_list_all <- lapply(contrast_names, function(cc) {
  tt <- out_all[[cc]]

  EnhancedVolcano(
    tt,
    lab = tt$feature_id,
    x = "logFC",
    y = "adj.P.Val",
    pCutoff = 0.05,
    FCcutoff = 0.5,
    title = gsub("_", " ", cc),
    subtitle = "Age + sex + study",
    pointSize = 2.0,
    labSize = 3.0
  )
})

wrap_plots(volcano_list_all, ncol = 3)
```

#--------------------------------------------------
# b: Metabolomic + Proteomic Samples 
#--------------------------------------------------
```{r}
#identify participants that have both proteomic and metabolomic data
ids_use_mp <- intersect(ids_metab_prot, intersect(meta_metab$record_id, rownames(mat_metab)))


meta_mp <- meta_metab[match(ids_use_mp, meta_metab$record_id), , drop = FALSE]
mat_mp  <- mat_metab[ids_use_mp, , drop = FALSE]

stopifnot(identical(meta_mp$record_id, rownames(mat_mp)))

cat("Metabolomics + proteomics subset N:", nrow(meta_mp), "\n")
cat("Metabolomics + proteomics subset features before filtering:", ncol(mat_mp), "\n")

tbl_mp <- meta_mp %>%
  count(study, group4) %>%
  pivot_wider(names_from = group4, values_from = n, values_fill = 0) %>%
  mutate(Total = rowSums(across(c("LC", "OC", "T1D", "T2D")))) %>%
  arrange(study)

tbl_mp
```

##-------------------------------------------------
##PREP MATRIX: Metabolomic + Proteomics data subset
##-------------------------------------------------
```{r}
mat_mp <- as.matrix(mat_mp)
rownames(mat_mp) <- trimws(as.character(rownames(mat_mp)))

mat_log_mp <- log2(mat_mp + 1)
mat_log_mp[!is.finite(mat_log_mp)] <- NA_real_

study_idx_mp <- split(seq_len(nrow(mat_log_mp)), meta_mp$study)

present_by_study_mp <- sapply(study_idx_mp, function(idx) {
  colSums(!is.na(mat_log_mp[idx, , drop = FALSE])) > 0
})

present_by_study_mp <- as.matrix(present_by_study_mp)

n_studies_present_mp <- colSums(present_by_study_mp)
keep_study_mp <- n_studies_present_mp >= 2

cat("Subset features before study filter:", ncol(mat_log_mp), "\n")
cat("Removed (<2 studies):", sum(!keep_study_mp), "\n")
cat("Kept (>=2 studies):", sum(keep_study_mp), "\n")

mat_filt_mp <- mat_log_mp[, keep_study_mp, drop = FALSE]

meds_mp <- apply(mat_filt_mp, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

mat_imp_mp <- mat_filt_mp
for (j in seq_len(ncol(mat_imp_mp))) {
  if (!is.na(meds_mp[j])) {
    mat_imp_mp[is.na(mat_imp_mp[, j]), j] <- meds_mp[j]
  }
}

vars_mp <- apply(mat_imp_mp, 2, var, na.rm = TRUE)

keep_cols_mp <- colSums(is.na(mat_imp_mp)) == 0 &
  vars_mp > 0 &
  is.finite(vars_mp)

mat_imp_mp <- mat_imp_mp[, keep_cols_mp, drop = FALSE]

cat("Subset features kept for PCA/LIMMA:", ncol(mat_imp_mp), "\n")
cat("number of participants kept", nrow(mat_imp_mp), "\n")
stopifnot(identical(rownames(mat_imp_mp), meta_mp$record_id))
```


```{r}
#removing near zero variance metabolites
nzv_idx <- nearZeroVar(mat_imp_mp)

length(nzv_idx)   
colnames(mat_imp_mp)[nzv_idx]  #check whic metabolites are removed

mat_imp_mp <- mat_imp_mp[, -nzv_idx, drop = FALSE]
```
##---------------------------
## PCA-based OUTLIER REMOVAL
##---------------------------
```{r}
pca_mp <- prcomp(mat_imp_mp, center = TRUE, scale. = TRUE)
pca_dist_mp <- sqrt(rowSums(pca_mp$x[, 1:5]^2))

cutoff_mp <- median(pca_dist_mp) + 3 * mad(pca_dist_mp, constant = 1)

outlier_df_mp <- meta_mp %>%
  dplyr::mutate(
    pca_dist = pca_dist_mp,
    pca_outlier = pca_dist > cutoff_mp
  ) %>%
  dplyr::arrange(dplyr::desc(pca_dist))

# inspect outliers
outlier_df_mp %>%
  dplyr::filter(pca_outlier) %>%
  dplyr::select(record_id, study, group4, pca_dist)

# remove outliers
keep_ids_mp <- outlier_df_mp %>%
  dplyr::filter(!pca_outlier) %>%
  dplyr::pull(record_id)
keep_ids_mp
# subset matrix
mat_imp_mp <- mat_imp_mp[keep_ids_mp, , drop = FALSE]

# reorder metadata to match matrix
meta_mp <- meta_mp[match(keep_ids_mp, meta_mp$record_id), , drop = FALSE]
cat("All metabolomics features kept for PCA/LIMMA:", ncol(mat_imp_mp), "\n")
cat("number of participants kept", nrow(mat_imp_mp), "\n")

stopifnot(identical(meta_mp$record_id, rownames(mat_imp_mp)))
```

##---------------------------------------
## PCA: METABOLOMICS + PROTEOMICS SUBSET
##---------------------------------------
```{r}
p0_mp <- make_pca_plot(mat_imp_mp, meta_mp, "Metabolomics - No adjustment (metab+prot subset)")

cov1_mp <- model.matrix(~ age + sex, data = meta_mp)[, -1, drop = FALSE]
mat_adj1_mp <- t(removeBatchEffect(t(mat_imp_mp), covariates = cov1_mp))
p1_mp <- make_pca_plot(mat_adj1_mp, meta_mp, "Metabolomics - Age + Sex (metab+prot subset)")

cov2_mp <- model.matrix(~ age + sex + study, data = meta_mp)[, -1, drop = FALSE]
mat_adj2_mp <- t(removeBatchEffect(t(mat_imp_mp), covariates = cov2_mp))
p2_mp <- make_pca_plot(mat_adj2_mp, meta_mp, "Metabolomics - Age + Sex + Study (metab+prot subset)")

(p0_mp | p1_mp | p2_mp) +
  plot_annotation(
    title = "PCA: Metabolomics samples with proteomics overlap",
    subtitle = "Color = group | Shape = study | removeBatchEffect used for visualization only"
  )
```
##---------------------------------------
##LIMMA: METABOLOMICS + PROTEOMICS SUBSET
##---------------------------------------
```{r}
Y_mp <- t(as.matrix(mat_imp_mp))

design_mp <- model.matrix(~ 0 + group4 + age + sex + study, data = meta_mp)
colnames(design_mp) <- sub("^group4", "", colnames(design_mp))
colnames(design_mp) <- make.names(colnames(design_mp))

contr_mp <- makeContrasts(
  T1D_vs_LC  = T1D - LC,
  T2D_vs_LC  = T2D - LC,
  T1D_vs_T2D = T1D - T2D,
  OC_vs_LC   = OC - LC,
  T1D_vs_OC  = T1D - OC,
  T2D_vs_OC  = T2D - OC,
  levels = design_mp
)

fit_mp <- lmFit(Y_mp, design_mp)
fit_mp <- contrasts.fit(fit_mp, contr_mp)
fit_mp <- eBayes(fit_mp)

out_mp <- list()
for (cc in colnames(contr_mp)) {
  tt <- topTable(fit_mp, coef = cc, number = Inf, adjust.method = "BH", sort.by = "P")
  tt$feature_id <- rownames(tt)
  tt$contrast <- cc
  out_mp[[cc]] <- tt
}

sig_mp <- tibble(
  contrast = names(out_mp),
  n_FDR05 = sapply(out_mp, function(tt) sum(tt$adj.P.Val < 0.05, na.rm = TRUE)),
  n_FDR01 = sapply(out_mp, function(tt) sum(tt$adj.P.Val < 0.01, na.rm = TRUE)),
  model = "metabolomics_age_sex_study_metab_prot_subset",
  n_samples = nrow(meta_mp)
)

print(sig_mp)
```
##-----------------------------------------------
##VOLCANO PLOTS: METABOLOMICS + PROTEOMICS SUBSET
##------------------------------------------------
```{r}
volcano_list_mp <- lapply(contrast_names, function(cc) {
  tt <- out_mp[[cc]]

  EnhancedVolcano(
    tt,
    lab = tt$feature_id,
    x = "logFC",
    y = "adj.P.Val",
    pCutoff = 0.05,
    FCcutoff = 0.5,
    title = gsub("_", " ", cc),
    subtitle = "Age + sex + study (metabolomics samples with proteomics overlap)",
    pointSize = 2.0,
    labSize = 3.0
  )
})

wrap_plots(volcano_list_mp, ncol = 3)
```
```{r}

dict <- read.csv2(file.path(dir.dat, "Data Harmonization", "data_dictionary_master.csv"), sep = ",") 

dict <- dict %>% dplyr::select(variable_name, label, form_name)

dict_metab <- dict %>%
  mutate(
    variable_name = trimws(as.character(variable_name)),
    label = trimws(as.character(label)),
    form_name = trimws(as.character(form_name))
  ) %>%
  filter(stringr::str_detect(form_name, regex("^metabolomics", ignore_case = TRUE)))

dict_lookup <- dict_metab %>%
  distinct(variable_name, label) %>%
  filter(!is.na(variable_name), variable_name != "") %>%
  tibble::deframe()

out_mp <- lapply(out_mp, function(x) {
  if (is.matrix(x) || is.data.frame(x)) {
    old_names <- rownames(x)
    if (!is.null(old_names)) {
      new_names <- dict_lookup[old_names]
      new_names[is.na(new_names) | new_names == ""] <- old_names[is.na(new_names) | new_names == ""]
      rownames(x) <- make.unique(new_names)
    }
  }
  x
})

out_mp
```

#---------------------------------
# KEGG compound -> pathway mapping
#---------------------------------
```{r}
# all human KEGG pathways
kegg_pathways_all <- KEGGREST::keggList("pathway", "hsa")

pathway_df <- tibble(
  pathway_id = sub("^path:", "", names(kegg_pathways_all)),
  pathway_name = unname(kegg_pathways_all)
)

pathway_df <- pathway_df %>%
  filter(str_detect(pathway_name, regex("metabolism", ignore_case = TRUE)) |
           pathway_id %in% c(
             "hsa00010","hsa00020","hsa00030","hsa00040","hsa00051","hsa00052",
             "hsa00053","hsa00500","hsa00520","hsa00620","hsa00630","hsa00640",
             "hsa00650","hsa00670","hsa00730","hsa00830","hsa00230","hsa00240",
             "hsa00250","hsa00260","hsa00270","hsa00280","hsa00300","hsa00310",
             "hsa00330","hsa00340","hsa00350","hsa00360","hsa00400","hsa00410",
             "hsa00430","hsa00450","hsa00460","hsa00471","hsa00472","hsa00480"
           ))

# get compounds in each pathway
pathway2compound <- map_dfr(pathway_df$pathway_id, function(pid) {
  x <- KEGGREST::keggGet(pid)[[1]]

  cpds <- x$COMPOUND
  if (is.null(cpds)) return(NULL)

  tibble(
    pathway_id = pid,
    compound_id = names(cpds)
  )
}) %>%
  distinct() %>%
  left_join(pathway_df, by = "pathway_id")

head(pathway2compound)
```
```{r}
normalize_kegg_compound <- function(x) {
  x <- trimws(as.character(x))
  x <- toupper(x)
  x
}
```

#----------------------------------------------------------------
#Performing Over-representation analysis
#using metabolites that remained after filtering as background
#check to see if I should include all metabolites in assay instead
#-----------------------------------------------------------------
```{r}
run_kegg_metab_ora <- function(
  tt,
  pathway_map,
  p_cutoff = 0.05,
  fc_cutoff = 0,
  feature_col = "feature_id",
  min_overlap = 3
) {
  tt <- tt %>%
    mutate(
      compound_id = normalize_kegg_compound(.data[[feature_col]])
    )

  # background = all tested metabolites that survived filtering
  bg <- intersect(tt$compound_id, unique(pathway_map$compound_id))

  # significant metabolites
  sig <- tt %>%
    filter(adj.P.Val < p_cutoff, abs(logFC) >= fc_cutoff) %>%
    pull(compound_id) %>%
    intersect(unique(pathway_map$compound_id))

  if (length(sig) == 0 || length(bg) == 0) {
    return(tibble(
      pathway_id = character(),
      pathway_name = character(),
      n_pathway = integer(),
      n_overlap = integer(),
      n_sig = integer(),
      n_bg = integer(),
      p_value = numeric(),
      padj = numeric(),
      overlap_compounds = character()
    ))
  }

  res <- pathway_map %>%
    group_by(pathway_id, pathway_name) %>%
    summarise(
      pathway_compounds = list(unique(compound_id)),
      n_pathway = length(unique(compound_id)),
      .groups = "drop"
    ) %>%
    rowwise() %>%
    mutate(
      pathway_bg = list(intersect(bg, pathway_compounds)),
      overlap = list(intersect(sig, pathway_bg)),
      n_overlap = length(overlap),
      n_in_path_bg = length(pathway_bg),
      n_sig = length(sig),
      n_bg = length(bg),

      # Fisher table components
      n_sig_in_path = n_overlap, # sig metabolite & in this pathway
      n_sig_not_in_path = n_sig - n_overlap, # sig metabolite but not in this pathway
      n_nonsig_in_path = n_in_path_bg - n_overlap, #not sig metab but in this pathway
      n_nonsig_not_path = n_bg - n_sig - (n_in_path_bg - n_overlap), #not sig, not in pathway

      p_value = if (
        all(c(
          n_sig_in_path,
          n_sig_not_in_path,
          n_nonsig_in_path,
          n_nonsig_not_path
        ) >= 0)
      ) {
        fisher.test(
          matrix(
            c(
              n_sig_in_path,
              n_sig_not_in_path,
              n_nonsig_in_path,
              n_nonsig_not_path
            ),
            nrow = 2
          ),
          alternative = "greater"
        )$p.value
      } else {
        NA_real_
      }
    ) %>%
    ungroup() %>%
    mutate(
      padj = p.adjust(p_value, method = "BH"),
      overlap_compounds = purrr::map_chr(overlap, ~ paste(.x, collapse = ", "))
    ) %>%
    filter(n_overlap >= min_overlap) %>%
    arrange(padj, p_value)

  res
}

```

```{r}
#plotting function
plot_kegg_ora <- function(res_df, title_text, n = 15) {
  if (nrow(res_df) == 0) {
    return(ggplot() + theme_void() + labs(title = paste(title_text, "- no enriched pathways")))
  }

  df <- res_df %>%
    filter(!is.na(padj)) %>%
    slice_min(order_by = padj, n = n, with_ties = FALSE) %>%
    mutate(
      pathway_name = str_replace_all(pathway_name, " - Homo sapiens \\(human\\)", ""),
      pathway_name = factor(pathway_name, levels = rev(pathway_name)),
      neglog10_padj = -log10(padj)
    )

  ggplot(df, aes(x = neglog10_padj, y = pathway_name, size = n_overlap)) +
    geom_point(alpha = 0.85) +
    theme_bw() +
    labs(
      title = title_text,
      x = expression(-log[10]("BH-adjusted p-value")),
      y = NULL,
      size = "Overlap"
    )
}
```
##-----------------------------
##a. ORA on all metabolite data
##-----------------------------
```{r}
kegg_res <- lapply(out_all, function(tt) {
  run_kegg_metab_ora(
    tt = tt,
    pathway_map = pathway2compound,
    p_cutoff = 0.05,
    fc_cutoff = 0
  )
})

names(kegg_res) <- names(out_all)

kegg_summary<- bind_rows(lapply(names(kegg_res), function(cc) {
  x <- kegg_res[[cc]]
  tibble(
    contrast = cc,
    n_pathways_FDR05 = sum(x$padj < 0.05, na.rm = TRUE),
    n_pathways_p05   = sum(x$p_value < 0.05, na.rm = TRUE)
  )
}))

kegg_summary
```

```{r}
kegg_res[["T1D_vs_T2D"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(20)
```

```{r}
plot_kegg_ora(kegg_res[["T1D_vs_T2D"]], "KEGG ORA: T1D vs T2D")
kegg_plots <- lapply(names(kegg_res), function(cc) {
  plot_kegg_ora(kegg_res[[cc]], gsub("_", " ", cc), n = 12)
})

patchwork::wrap_plots(kegg_plots, ncol = 2)
```

##---------------------------------------------
## b. ORA on metabolite data with proteomics dat
##----------------------------------------------
```{r}
kegg_res_mp <- lapply(out_mp, function(tt) {
  run_kegg_metab_ora(
    tt = tt,
    pathway_map = pathway2compound,
    p_cutoff = 0.05,
    fc_cutoff = 0
  )
})

names(kegg_res_mp) <- names(out_mp)

kegg_summary_mp <- bind_rows(lapply(names(kegg_res_mp), function(cc) {
  x <- kegg_res_mp[[cc]]
  tibble(
    contrast = cc,
    n_pathways_FDR05 = sum(x$padj < 0.05, na.rm = TRUE),
    n_pathways_p05   = sum(x$p_value < 0.05, na.rm = TRUE)
  )
}))

kegg_summary_mp
```

```{r}
kegg_res_mp[["T1D_vs_T2D"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(5)
kegg_res_mp[["T2D_vs_LC"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(5)
kegg_res_mp[["OC_vs_LC"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(5)
kegg_res_mp[["T1D_vs_OC"]] %>%
  select(pathway_name, n_overlap, p_value, padj, overlap_compounds) %>%
  head(20)
```

```{r}
plot_kegg_ora(kegg_res_mp[["T1D_vs_T2D"]], "KEGG ORA: T1D vs T2D")
```

```{r}
kegg_plots_mp <- lapply(names(kegg_res_mp), function(cc) {
  plot_kegg_ora(kegg_res_mp[[cc]], gsub("_", " ", cc), n = 12)
})

patchwork::wrap_plots(kegg_plots_mp, ncol = 2)
```

#--------------------------------------------------
#ElasticNet on top 50 metabolites per contrast
#a) all metabolite data
#b) data with metabolite and proteomics data
#--------------------------------------------------
```{r}
#change for metabolite data

```

```{r}
run_elastic_net <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  top_n = 50,
  covars = c("age", "sex"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1,
  min_per_class = 5,
  maxit = 1e6
) {
  
  cat("\n============================\n")
  cat("Running contrast:", cc, "\n")
  
  if (!cc %in% names(out_list)) {
    stop("Contrast not found in out_list: ", cc)
  }
  
  if (!grepl("_vs_", cc)) {
    stop("Contrast name must contain '_vs_'")
  }
  
  parts <- strsplit(cc, "_vs_")[[1]]
  if (length(parts) != 2) {
    stop("Contrast could not be split into two groups")
  }
  
  g_pos <- parts[1]
  g_neg <- parts[2]
  
 idx <- which(meta_df$group4 %in% c(g_pos, g_neg))
meta_sub <- meta_df[idx, , drop = FALSE]
meta_sub$record_id <- as.character(meta_sub$record_id)

meta_sub$group4 <- droplevels(meta_sub$group4)

y <- ifelse(meta_sub$group4 == g_pos, 1, 0)
y <- as.numeric(y)

tab <- table(meta_sub$group4)
if (any(tab < min_per_class)) {
  stop(
    "Too few samples in at least one class for contrast: ",
    cc,
    " | counts = ",
    paste(names(tab), tab, collapse = ", ")
  )
}
  
  t <- out_list[[cc]]
  if (is.null(t)) {
    stop("Contrast table is NULL for: ", cc)
  }
  t <- as.data.frame(t)
  
  cat("Columns in contrast table:\n")
  print(names(t))
  
  if (!"feature_id" %in% names(t)) {
    fid <- rownames(t)
    if (is.null(fid)) stop("No feature_id column or rownames found")
    t$feature_id <- fid
  }
  
  if (!all(c("adj.P.Val", "P.Value") %in% names(t))) {
    stop("Contrast table is missing adj.P.Val and/or P.Value")
  }
  
  cat("Rows in limma table:", nrow(t), "\n")
  cat("Overlap with protein matrix:", sum(t$feature_id %in% colnames(X_mat)), "\n")
  
  top_feats <- t %>%
    dplyr::filter(!is.na(adj.P.Val)) %>%
    dplyr::arrange(adj.P.Val, P.Value) %>%
    dplyr::slice_head(n = top_n) %>%
    dplyr::pull(feature_id) %>%
    intersect(colnames(X_mat))
  
  cat("Top overlapping features:", length(top_feats), "\n")
  cat("First few top features:\n")
  print(head(top_feats, 10))
  
  if (length(top_feats) < 5) {
    stop("Fewer than 5 overlapping top features for contrast: ", cc)
  }
  
  meta_sub <- meta_sub %>%
    dplyr::mutate(
      sex = factor(sex),
      study = factor(study)
    )
  
  covars_present <- covars[covars %in% colnames(meta_sub)]
  cat("Covariates used:", paste(covars_present, collapse = ", "), "\n")
  
  if (length(covars_present) > 0) {
    form_cov <- as.formula(paste("~", paste(covars_present, collapse = " + ")))
    Z <- model.matrix(form_cov, data = meta_sub)
  } else {
    Z <- matrix(1, nrow = nrow(meta_sub), ncol = 1)
    colnames(Z) <- "(Intercept)"
  }
  
  Xp <- X_mat[meta_sub$record_id, top_feats, drop = FALSE]
  
  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"
  
  pf <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))
  
  sds <- apply(X, 2, sd)
  keep <- which(is.finite(sds) & sds > 0)
  
  X <- X[, keep, drop = FALSE]
  pf <- pf[keep]
  
  cat("Final design matrix dim:", dim(X), "\n")
  
  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))
  cat("nfolds used:", nfolds_use, "\n")
  
  if (nfolds_use < 2) {
    stop("Not enough samples per class for CV")
  }
  
  set.seed(seed)
  
  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      cv.glmnet(
        x = X,
        y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = pf,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) {
        message("glmnet failed at alpha = ", a, ": ", conditionMessage(e))
        NULL
      }
    )
    
    if (is.null(cv)) return(NULL)
    list(alpha = a, cvfit = cv)
  })
  
  fit_list <- Filter(Negate(is.null), fit_list)
  
  if (length(fit_list) == 0) {
    stop("All glmnet fits failed")
  }
  
  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]
  
  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))
  coef_df <- tibble::tibble(
    term = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>%
    dplyr::filter(beta != 0)
  
  protein_terms <- intersect(colnames(Xp), coef_df$term)
  
  p_hat <- as.numeric(
    predict(best$cvfit, newx = X, s = "lambda.1se", type = "response")
  )
  
  auc_val <- as.numeric(
    pROC::auc(pROC::roc(y, p_hat, quiet = TRUE))
  )
  
  cat("Best alpha:", best$alpha, "\n")
  cat("AUC:", auc_val, "\n")
  cat("Nonzero coefficients:", nrow(coef_df), "\n")
  cat("Selected metabolites:", length(protein_terms), "\n")
  
  list(
    contrast         = cc,
    positive         = g_pos,
    negative         = g_neg,
    best_alpha       = best$alpha,
    lambda_1se       = best$cvfit$lambda.1se,
    lambda_min       = best$cvfit$lambda.min,
    auc_train        = auc_val,
    n_input_proteins = length(top_feats),
    coef_df          = coef_df,
    protein_terms    = protein_terms,
    cvfit            = best$cvfit,
    top_feats        = top_feats
  )
}
```

```{r}
res_T1D_LC <- run_elastic_net(
  cc = "T1D_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_filt,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T1D_OC <- run_elastic_net(
  cc = "T1D_vs_OC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_all,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T1D_T2D <- run_elastic_net(
  cc = "T1D_vs_T2D",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_all,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T2D_LC <- run_elastic_net(
  cc = "T2D_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_all,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_T2D_OC <- run_elastic_net(
  cc = "T2D_vs_OC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_all,
  top_n = 50,
  covars = c("age", "sex")
)
```

```{r}
res_OC_LC <- run_elastic_net(
  cc = "OC_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_all,
  top_n = 50,
  covars = c("age", "sex")
)
```
#------------------------------------
#boostrapping function for elasticnet
#-------------------------------------
```{r}
run_elastic_net_bootstrap_metab <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  top_n = 50,
  covars = c("age", "sex"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  n_boot = 100,
  seed = 1835,
  min_per_class = 5
) {
  
  set.seed(seed)
  boot_res <- vector("list", n_boot)
  
  for (b in seq_len(n_boot)) {
    
    idx0 <- which(meta_df$group4 %in% strsplit(cc, "_vs_")[[1]])
    meta_sub0 <- meta_df[idx0, , drop = FALSE]
    
    idx_boot <- unlist(lapply(split(seq_len(nrow(meta_sub0)), meta_sub0$group4), function(ii) {
      sample(ii, size = length(ii), replace = TRUE)
    }))
    
    meta_boot <- meta_sub0[idx_boot, , drop = FALSE]
    meta_boot$record_id_boot <- paste0(meta_boot$record_id, "_boot", seq_len(nrow(meta_boot)))
    
    X_boot <- X_mat[meta_boot$record_id, , drop = FALSE]
    rownames(X_boot) <- meta_boot$record_id_boot
    meta_boot$record_id <- meta_boot$record_id_boot
    
    fit_b <- tryCatch(
      run_elastic_net(
        cc = cc,
        out_list = out_list,
        meta_df = meta_boot,
        X_mat = X_boot,
        top_n = top_n,
        covars = covars,
        seed = seed + b
      ),
      error = function(e) NULL
    )
    
    boot_res[[b]] <- fit_b
  }
  
  boot_res <- Filter(Negate(is.null), boot_res)
  
  selected_tbl <- lapply(seq_along(boot_res), function(i) {
    tibble(
      boot = i,
      contrast = cc,
      feature_id = unique(boot_res[[i]]$protein_terms)  # still named this internally
    )
  }) %>%
    bind_rows() %>%
    distinct(contrast, boot, feature_id)
  
  stability_tbl <- selected_tbl %>%
    count(contrast, feature_id, name = "n_selected") %>%
    mutate(
      n_boot_success = length(boot_res),
      selection_frequency = n_selected / n_boot_success,
      selection_percent = 100 * selection_frequency
    ) %>%
    arrange(desc(selection_frequency))
  
  list(
    contrast = cc,
    stability_tbl = stability_tbl,
    boot_res = boot_res
  )
}
```

#t1d vs lc
```{r}
boot_metab_T1D_LC <- run_elastic_net_bootstrap_metab(
  cc = "T1D_vs_LC",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_all,
  top_n = 50
)
```
```{r}
core_metab_T1D_LC <- boot_metab_T1D_LC$stability_tbl %>%
  filter(selection_frequency >= 0.7)
core_metab_T1D_LC
```

```{r}
boot_metab_T1D_T2D <- run_elastic_net_bootstrap_metab(
  cc = "T1D_vs_T2D",
  out_list = out_all,
  meta_df = meta_all,
  X_mat = mat_imp_all,
  top_n = 50
)
```
```{r}
core_metab_T1D_T2D <- boot_metab_T1D_T2D$stability_tbl %>%
  filter(selection_frequency >= 0.7)
core_metab_T1D_T2D
```

```{r}
plot_stability_metab <- function(boot_obj, min_freq = 0.7, top_n = 25) {
  
  plot_df <- boot_obj$stability_tbl %>%
    filter(selection_frequency >= min_freq) %>%
    slice_max(selection_frequency, n = top_n, with_ties = FALSE)
  
  ggplot(plot_df, aes(
    x = reorder(feature_id, selection_percent),
    y = selection_percent
  )) +
    geom_col() +
    coord_flip() +
    labs(
      title = paste0("Metabolite stability: ", boot_obj$contrast),
      x = NULL,
      y = "Bootstrap selection frequency (%)"
    ) +
    theme_bw()
}
```

```{r}
plot_stability_metab(boot_metab_T1D_LC)
plot_stability_metab(boot_metab_T1D_T2D)
```

```{r}
run_kegg_metab_ora(
  tt = out_all[["T1D_vs_LC"]],
  pathway_map = pathway2compound,
  p_cutoff = 0.05,
  fc_cutoff = 0,
  min_overlap = 3
)
```

```{r}
sig_metabs <- core_metab_T1D_LC$feature_id

ora_input <- out_all[["T1D_vs_LC"]] %>%
  mutate(is_sig = feature_id %in% sig_metabs)

run_kegg_metab_ora(ora_input, pathway2compound)
```

