---
title: "scRNA_data_analysis"
author: "Savanah Leidholt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths("/mmfs1/gscratch/togo/leidholt/R_SLL_Seurat/")
```

```{r}
install_if_missing_cran <- function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(
      pkg,
      lib = ("/mmfs1/gscratch/togo/leidholt/R_SLL_Seurat/"),
      dependencies = TRUE
    )
  }
}

install_if_missing_bioc <- function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    BiocManager::install(
      pkg,
      lib = ("/mmfs1/gscratch/togo/leidholt/R_SLL_Seurat/"),
      update = FALSE,
      ask = FALSE
    )
  }
}
```

```{r}
library(Seurat)
library(tidyverse)
library(arsenal)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(Hmisc)
library(readxl)
library(purrr)
library(aws.s3)
library(jsonlite)
library(RColorBrewer)
library(ggtrace)
library(kableExtra)
library(data.table)
library(table1)
library(BiocManager)
library(GSA)
library(emmeans)
library(pheatmap)
library(enrichplot)
library(enrichR)
library(nebula)
library(doParallel)
library(Matrix)
library(edgeR)
library(stringr)
library(patchwork)
library(msigdbr)
library(fgsea)


#install_if_missing_bioc("speckle")
library(speckle)
#install_if_missing_bioc("DESeq2")
library(DESeq2)
```

```{r}
#directories
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
  
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/TRIAD/Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
    
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/TRIAD/Results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
    
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r}
## Create an S3 client
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```




```{r}
pb90_multiomics_subset <- s3readRDS(
          object = "data/pb90_multiomics_SLL_subset_20260527.rds", 
          bucket = "triad",
          region = "")


```

```{r}
my_colors <- colorRampPalette(brewer.pal(12, "Set3"))(43)
# Read Seurat object and run UMAP, obtain cell numbers, etc
pb90_multiomics_subset_meta <- pb90_multiomics_subset@meta.data %>%
  cbind(pb90_multiomics_subset@reductions$umap.TRIAD@cell.embeddings)

```

```{r}
centers <- pb90_multiomics_subset_meta %>%
  group_by(KPMP_celltype) %>%
  summarise(
    x = median(TRIAD_UMAP1),
    y = median(TRIAD_UMAP2),
    .groups = "drop"
  )

umap_p <- ggplot(pb90_multiomics_subset_meta,
                 aes(x = TRIAD_UMAP1, y = TRIAD_UMAP2, color = KPMP_celltype)) +
  geom_point(alpha = 0.2, shape = 16) +
  geom_text_repel(
    data = centers,
    aes(x = x, y = y, label = KPMP_celltype),
    inherit.aes = FALSE,
    size = 5,                
    fontface = "bold",
    color = "black"
  ) +
  scale_color_manual(values = my_colors) +
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),
    
    # axis titles + tick labels (match previous plots)
    axis.title.x = element_text(size = 16, hjust = 0),
    axis.title.y = element_text(size = 16, hjust = 0),
    axis.text = element_text(size = 14),
    
    # no legend
    legend.position = "none"
  ) +
  labs(x = "UMAP1", y = "UMAP2")

umap_p
```

```{r}
pb90_multiomics_subset_meta%>%
  distinct(record_id, group)

umap_p_study<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = TRIAD_UMAP1, y = TRIAD_UMAP2, color = study)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("RENAL-HEIR" = "darkcyan",
                                "RENAL-HEIRitage"= "gold4",
                                "IMPROVE" = "orange",
                                "CROCODILE" = "black",
                                "ATTEMPT" = "lightpink"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(size = 16, hjust = 0),   
        axis.title.y = element_text(size = 16, hjust = 0), 
        axis.text = element_text(size = 14),

 # legend
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)) + 
  labs(x = "UMAP1", y = "UMAP2",
       color = "study") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_study
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)
# UMAP by source
umap_p_group<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = TRIAD_UMAP1, y = TRIAD_UMAP2, color = group)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("Obese Control" = "darkcyan",
                                "Lean Control"= "plum4",
                                "Type 1 Diabetes" = "orange",
                                "Type 2 Diabetes" = "black"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(size = 16, hjust = 0),   
        axis.title.y = element_text(size = 16, hjust = 0),
        axis.title = element_text(size = 14),

        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14))+
  labs(x = "UMAP1", y = "UMAP2",
       color = "group") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_group
```
```{r}
aws.s3::s3HTTP(
  verb = "GET",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  path = "/"
)
```

```{r}
umap_cells_study <- umap_p + umap_p_study
tmp <- tempfile(fileext = ".png")

ggsave(
  tmp,
  plot = umap_cells_study,
  width = 20,
  height = 16,
  dpi = 300
)

httr::set_config(httr::config(ssl_verifypeer = FALSE))
put_object(
  file = tmp,
  object = "figures/metadata/Umap_cells_study.png",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)


umap_cells_group <- umap_p + umap_p_group
tmp <- tempfile(fileext = ".png")

ggsave(tmp,
       plot = umap_cells_group,
       width = 20,
       height =16,
       dpi = 300
       )

put_object(
  file = tmp,
  object = "figures/metadata/Umap_cells_group.png",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)
```

## b. Descriptive Statistics
```{r}
meta_table1 <- pb90_multiomics_subset_meta %>%
  distinct(record_id, .keep_all = TRUE) %>%
  mutate(
    group = factor(group)
  ) %>%
  filter(!is.na(group)) %>%
  droplevels()

table(meta_table1$group, useNA = "ifany")
```

```{r}
#Load Metadata cleaned
#Table 1
label(meta_table1$age) <- "Age (years)"
label(meta_table1$sex) <- "Sex"
label(meta_table1$race_ethnicity_condensed) <- "Ethnicity"
label(meta_table1$bmi) <- "BMI (kg/m^2)"
label(meta_table1$hba1c) <- "HbA1c (%)"
label(meta_table1$acr_u) <- "Urine Albumin-Creatinine Ratio (mg/g)"
label(meta_table1$eGFR_CKiD_U25_avg) <- "eGFR by Full Age Spectrum (mL/min/1.73m^2)"
label(meta_table1$group) <- "Diabetes Status"


pvalue_table1 <- function(x, ...) {
  
  y <- unlist(x)
  g <- factor(rep(names(x), times = sapply(x, length)))
  
  keep <- !is.na(y) & !is.na(g)
  y <- y[keep]
  g <- g[keep]
  
  if (is.numeric(y)) {
    if (length(unique(g)) < 2) return(c("", NA))
    p <- kruskal.test(y ~ g)$p.value
    test <- "Kruskal-Wallis"
  } else {
    tab <- table(y, g)
    if (nrow(tab) < 2 || ncol(tab) < 2) return(c("", NA))
    chi <- suppressWarnings(chisq.test(tab))
    if (any(chi$expected < 5)) {
      chi <- chisq.test(tab, simulate.p.value = TRUE, B = 10000)
      test <- "Chi-square simulated"
    } else {
      test <- "Chi-square"
    }
    p <- chi$p.value
  }
  c("", format.pval(p, digits = 3, eps = 0.001))
}

tble1 <- table1(
  ~ age + sex + race_ethnicity_condensed + hba1c + bmi + acr_u + eGFR_CKiD_U25_avg | group,
  data = meta_table1,
  extra.col = list(`Overall p-value` = pvalue_table1)
)

tmp <- tempfile(fileext = ".html")

htmltools::save_html(
  htmltools::HTML(as.character(tble1)),
  file = tmp
)

put_object(
  file = tmp,
  object = "tables/metadata_table.html",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)
```

```{r}
contrasts <- list(
  LC_vs_OC    = c("Lean Control", "Obese Control"),
  T1D_vs_T2D  = c("Type 1 Diabetes", "Type 2 Diabetes"),
  T1D_vs_LC   = c("Type 1 Diabetes", "Lean Control"),
  T2D_vs_LC   = c("Type 2 Diabetes", "Lean Control"),
  T1D_vs_OC   = c("Type 1 Diabetes", "Obese Control"),
  T2D_vs_OC   = c("Type 2 Diabetes", "Obese Control")
)

continuous_vars <- c("age", "bmi", "hba1c", "acr_u", "eGFR_CKiD_U25_avg") 
pairwise_results <- map_dfr(names(contrasts), function(comp) {
  
  groups_keep <- contrasts[[comp]]
  
  map_dfr(continuous_vars, function(var) {
    
    dat_test <- meta_table1 %>%
      filter(group %in% groups_keep) %>%
      select(group, value = all_of(var)) %>%
      filter(!is.na(group), !is.na(value)) %>%
      mutate(group = droplevels(factor(group, levels = groups_keep)))
    
    n_group1 <- sum(dat_test$group == groups_keep[1], na.rm = TRUE)
    n_group2 <- sum(dat_test$group == groups_keep[2], na.rm = TRUE)
    
    if (nlevels(dat_test$group) < 2 || n_group1 == 0 || n_group2 == 0) {
      return(tibble(
        contrast = comp,
        variable = var,
        group1 = groups_keep[1],
        group2 = groups_keep[2],
        n_group1 = n_group1,
        n_group2 = n_group2,
        p_value = NA_real_,
        note = "Skipped: fewer than 2 groups with non-missing data"
      ))
    }
    
    test <- wilcox.test(value ~ group, data = dat_test, exact = FALSE)
    
    tibble(
      contrast = comp,
      variable = var,
      group1 = groups_keep[1],
      group2 = groups_keep[2],
      n_group1 = n_group1,
      n_group2 = n_group2,
      p_value = test$p.value,
      note = NA_character_
    )
  })
})

pairwise_results <- pairwise_results %>%
  group_by(variable) %>%
  mutate(
    p_adj = p.adjust(p_value, method = "BH"),
    p_value_fmt = format.pval(p_value, digits = 3, eps = 0.001),
    p_adj_fmt = format.pval(p_adj, digits = 3, eps = 0.001)
  ) %>%
  ungroup()

pairwise_results
tmp <- tempfile(fileext = ".csv")

write.csv(
  pairwise_results,
  tmp,
  row.names = FALSE
)

httr::set_config(httr::config(ssl_verifypeer = FALSE))
put_object(
  file = tmp,
  object = "tables/pairwise_results.csv",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)
```

```{r}
cell_counts <- pb90_multiomics_subset_meta %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n))

max_n <- max(cell_counts$n, na.rm = TRUE)

label_pos <- -max_n * 0.05  
xmin_lim  <- -max_n * 0.35  
xmax_lim  <-  max_n * 1.15   

cell_count_bar <- ggplot(cell_counts, 
                         aes(x = reorder(KPMP_celltype_general, n), y = n)) +
  geom_col(fill = "#97a97c") +
  

  geom_text(aes(y = n, label = n),
            hjust = -0.1, size = 4) +
  
  geom_text(aes(y = label_pos, label = KPMP_celltype_general),
            hjust = 1, size = 4) +
  
  coord_flip(clip = "off") +
  
  scale_y_continuous(
    limits = c(xmin_lim, xmax_lim),
    expand = expansion(mult = c(0, 0))
  ) +
  
  labs(x = NULL, y = "Count") +
  
  theme(
    text = element_text(size = 15),
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.margin = margin(10, 30, 10, 30)
  ) +
  theme_transparent()

cell_count_bar

#ggsave("cell_counts.pdf",
 #      plot = cell_count_bar,
  #     path = file.path(dir.results, "T1D-T2D", "Results"),
   #    width = 10,
    #   height = 6)
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)

# cell counts per group
cell_counts_group <- pb90_multiomics_subset_meta %>%
  group_by(group) %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n))

cell_group_bar <- ggplot(cell_counts_group, aes(x = reorder(KPMP_celltype_general, n), y = n, fill = group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Lean Control" = "plum4",
                               "Obese Control" = "darkcyan",
                               "Type 1 Diabetes" = "orange2",
                               "Type 2 Diabetes"= "black")) +
  coord_flip() +
  labs(x = NULL, 
       y = "Count",
       fill = "Group") +
  theme_transparent() +
  theme(
    text = element_text(size = 15),
    panel.grid = element_blank(),
    plot.margin = margin(1, 0.5, 1, 0.5),
    
    legend.position = c(0.7, 0.5),
    legend.justification = c(0, 0.5),
    legend.background = element_blank(),
    legend.margin = margin(0, 0, 0, 0)
  ) +
  
  scale_y_continuous(
    limits = c(xmin_lim, xmax_lim),
    expand = expansion(mult = c(0, 0))
  ) 

cell_group_bar
#ggsave("cell_gounts_group.pdf",
 #      plot = cell_group_bar,
  #     path = file.path(dir.results, "T1D-T2D", "Results"),
   #    width = 10,
    #   height = 6)
```

```{r}
cells <- cell_count_bar + cell_group_bar 
tmp <- tempfile(fileext = ".png")
ggsave(tmp,
       plot = cells,
       width = 16,
       height = 10,
       dpi = 600)

put_object(
  file = tmp,
  object = "figures/cell_counts.png",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)

```


```{r}
# Barplot sample proportion by celltype 
ggplot(pb90_multiomics_subset@meta.data) +
    geom_bar(aes(x=study, fill=KPMP_celltype_general), 
                 position=position_fill(), color="black") +
    theme_classic() +
    labs(x="Study", y="Proportion of Cells")

ggplot(pb90_multiomics_subset@meta.data) +
    geom_bar(aes(x=group, fill=KPMP_celltype_general), 
                 position=position_fill(), color="black") +
    theme_classic() +
    labs(x="group", y="Proportion of Cells")
```


```{r}
pb90_multiomics_subset@meta.data$cell_type_legend <- paste0(
  pb90_multiomics_subset@meta.data$KPMP_celltype_general, 
  " = ", 
  pb90_multiomics_subset@meta.data$cell_type
)
nrow(pb90_multiomics_subset) # genes 31974
ncol(pb90_multiomics_subset) # cells 307386
```

```{r}
# Check your metadata
table(pb90_multiomics_subset$group)
table(pb90_multiomics_subset$kit_id)
table(pb90_multiomics_subset$source)
table(pb90_multiomics_subset$study)
table(pb90_multiomics_subset$KPMP_celltype_general)
head(pb90_multiomics_subset@meta.data[, c("group", "kit_id", "record_id", "KPMP_celltype_general")])
```
#------------------
#propeller analysis
#------------------
```{r}
groups <- unique(pb90_multiomics_subset_meta$group)
groups <- groups[!is.na(groups)]

group_pairs <- combn(groups, 2, simplify = FALSE)

propeller_results <- list()

for (pair in group_pairs) {
  
  pair_name <- paste(pair, collapse = "_vs_")
  pair_name_clean <- gsub("[^A-Za-z0-9_]+", "_", pair_name)
  
  cat("\n=== Propeller:", pair_name, "===\n")
  
  meta_sub <- pb90_multiomics_subset_meta[
    pb90_multiomics_subset_meta$group %in% pair,
  ]
  
  meta_sub$group <- droplevels(factor(meta_sub$group))
  
  res <- propeller(
    clusters  = meta_sub$KPMP_celltype_general,
    sample    = meta_sub$record_id,
    group     = meta_sub$group,
    transform = "logit"
  )
  
  res$comparison <- pair_name
  propeller_results[[pair_name_clean]] <- res
  
  tmp <- tempfile(fileext = ".csv")
  
  write.csv(
    res,
    tmp,
    row.names = TRUE
  )
  
  put_object(
    file = tmp,
    object = paste0("results/propeller/propeller_", pair_name_clean,".csv"),
    bucket = "triad",
    base_url = "s3.kopah.uw.edu",
    region = "",
    use_https = TRUE
  )
  
  unlink(tmp)
}

  #write.csv(
  #  x    = res,
  #  file = dir.results, paste0("propeller_", pair_name, ".csv"),
  #  row.names = TRUE
  #)

```
#-----------------------------------------------------
#psuedobulk analysis with edgeR & Deseq2
#-----------------------------------------------------
```{r}
#setting up save function to kopah
bucket <- "triad"

upload_to_kopah <- function(file, object, bucket = "triad") {
  put_object(
    file = file,
    object = object,
    bucket = bucket,
    base_url = "s3.kopah.uw.edu",
    region = "",
    use_https = TRUE
  )
}

min_cells_per_sample <- 20
min_donors_per_group <- 3

```

#-------------------------------------------------------
# metadata and building pseudobulk counts
# one pseudobulk sample = one donor within one cell type
#-------------------------------------------------------
```{r}
meta <- pb90_multiomics_subset@meta.data %>%
  mutate(
    cell_barcode = rownames(pb90_multiomics_subset@meta.data),
    record_id = as.character(record_id),
    group = as.character(group),
    sex = as.character(sex),
    study = as.character(study),
    celltype = as.character(KPMP_celltype_general)
  ) %>%
  filter(
    !is.na(record_id),
    !is.na(group),
    !is.na(age),
    !is.na(sex),
    !is.na(study),
    !is.na(celltype)
  ) %>%
  mutate(
    sample_celltype_id = paste(record_id, celltype, sep = "__")
  )

counts <- GetAssayData(pb90_multiomics_subset, assay = "RNA", layer = "counts")
counts <- counts[, meta$cell_barcode, drop = FALSE]

pb_ids <- unique(meta$sample_celltype_id)

pb_count_list <- list()
pb_meta_list <- list()

for (sid in pb_ids) {
  
  idx <- which(meta$sample_celltype_id == sid)
  
  if (length(idx) < min_cells_per_sample) next
  
  meta_sub <- meta[idx, ]
  
  pb_count_list[[sid]] <- Matrix::rowSums(counts[, idx, drop = FALSE])
  
  pb_meta_list[[sid]] <- data.frame(
    pseudobulk_id = sid,
    record_id = meta_sub$record_id[1],
    group = meta_sub$group[1],
    age = meta_sub$age[1],
    sex = meta_sub$sex[1],
    study = meta_sub$study[1],
    celltype = meta_sub$celltype[1],
    n_cells = nrow(meta_sub),
    stringsAsFactors = FALSE
  )
}

pb_counts <- do.call(cbind, pb_count_list)
pb_meta <- bind_rows(pb_meta_list)

pb_meta <- pb_meta %>%
  dplyr::slice(match(colnames(pb_counts), pseudobulk_id))

stopifnot(identical(colnames(pb_counts), pb_meta$pseudobulk_id))
```

#-----------------------
#volcano plot function
#-----------------------
```{r}
plot_volcano <- function(tt, title = "", fdr_col = "FDR",
                         lfc_col = "logFC",
                         stat_col = NULL,
                         fdr_cutoff = 0.05,
                         lfc_cutoff = 0.5,
                         n_labels = 10) {
  
  tt <- tt %>%
    dplyr::mutate(
      neg_log10_fdr = -log10(.data[[fdr_col]]),
      sig = dplyr::case_when(
        .data[[fdr_col]] < fdr_cutoff & .data[[lfc_col]] > lfc_cutoff  ~ "Up",
        .data[[fdr_col]] < fdr_cutoff & .data[[lfc_col]] < -lfc_cutoff ~ "Down",
        TRUE ~ "NS"
      )
    )
  
  if (!is.null(stat_col) && stat_col %in% colnames(tt)) {
    
    label_df <- tt %>%
      dplyr::filter(.data[[fdr_col]] < fdr_cutoff) %>%
      dplyr::arrange(dplyr::desc(abs(.data[[stat_col]]))) %>%
      dplyr::slice_head(n = n_labels)
    
  } else {
    
    label_df <- tt %>%
      dplyr::filter(.data[[fdr_col]] < fdr_cutoff) %>%
      dplyr::arrange(.data[[fdr_col]]) %>%
      dplyr::slice_head(n = n_labels)
  }
  
  ggplot(tt, aes(x = .data[[lfc_col]], y = neg_log10_fdr, color = sig)) +
    geom_point(alpha = 0.6, size = 1.5) +
    geom_text_repel(
      data = label_df,
      aes(label = gene),
      size = 3,
      max.overlaps = Inf
    ) +
    scale_color_manual(values = c("Up" = "orange3", "Down" = "purple4", "NS" = "grey")) +
    geom_vline(xintercept = c(-lfc_cutoff, lfc_cutoff), linetype = "dashed") +
    geom_hline(yintercept = -log10(fdr_cutoff), linetype = "dashed") +
    labs(title = title, x = "log2 Fold Change", y = "-log10(FDR)") +
    theme_classic()
}
```

#------------------------
#running edgeR pseudobulk
#-----------------------
```{r}
groups <- sort(unique(pb_meta$group))
group_pairs <- combn(groups, 2, simplify = FALSE)

celltypes <- sort(unique(pb_meta$celltype))

edgeR_results <- list()

for (cell_name in celltypes) {
  
  cell_idx <- which(pb_meta$celltype == cell_name)
  counts_ct <- pb_counts[, cell_idx, drop = FALSE]
  meta_ct <- pb_meta[cell_idx, , drop = FALSE]
  
  for (pair in group_pairs) {
    
    pair_name <- paste(pair, collapse = "_vs_")
    pair_name_clean <- gsub("[^A-Za-z0-9_]+", "_", pair_name)
    cell_name_clean <- gsub("[^A-Za-z0-9_]+", "_", cell_name)
    
    cat("\n=== edgeR:", cell_name, pair_name, "===\n")
    
    keep_pair <- meta_ct$group %in% pair
    counts_sub <- counts_ct[, keep_pair, drop = FALSE]
    meta_sub <- meta_ct[keep_pair, , drop = FALSE]
    
    meta_sub$group <- factor(meta_sub$group, levels = pair)
    meta_sub$sex <- factor(meta_sub$sex)
    
    donor_tab <- table(meta_sub$group)
    
    if (any(donor_tab < min_donors_per_group)) {
      cat("Skipping: too few donors\n")
      next
    }
    
    design <- model.matrix(~ group + age + sex, data = meta_sub)
    
    y <- DGEList(counts = counts_sub)
    keep_genes <- filterByExpr(y, design = design)
    y <- y[keep_genes, , keep.lib.sizes = FALSE]
    
    if (nrow(y) == 0) next
    
    y <- calcNormFactors(y)
    y <- estimateDisp(y, design)
    
    fit <- glmQLFit(y, design, robust = TRUE)
    
    coef_name <- grep("^group", colnames(design), value = TRUE)
    
    qlf <- glmQLFTest(fit, coef = coef_name)
    tt <- topTags(qlf, n = Inf)$table
    
    tt$gene <- rownames(tt)
    tt$comparison <- pair_name
    tt$celltype <- cell_name
    tt$n_donors_group1 <- as.numeric(donor_tab[1])
    tt$n_donors_group2 <- as.numeric(donor_tab[2])
    
    edgeR_results[[paste(cell_name_clean, pair_name_clean, sep = "__")]] <- tt
    
    tmp_csv <- tempfile(fileext = ".csv")
    write.csv(tt, tmp_csv, row.names = FALSE)
    
    upload_to_kopah(
      file = tmp_csv,
      object = paste0(
        "results/pseudobulk/edgeR/all_cells/pseudo_all_edgeR_",
        pair_name_clean, "_", cell_name_clean, ".csv"
      )
    )
    
    volcano_plot <- plot_volcano(
      tt,
      title = paste("edgeR", cell_name, pair_name),
      fdr_col = "FDR",
      lfc_col = "logFC"
    )
    
    tmp_png <- tempfile(fileext = ".png")
    
    ggsave(
      tmp_png,
      plot = volcano_plot,
      width = 6,
      height = 5,
      dpi = 300
    )
    
    upload_to_kopah(
      file = tmp_png,
      object = paste0(
        "figures/pseudobulk/edgeR/all_cells/volcano_all_edgeR_",
        pair_name_clean, "_", cell_name_clean, ".png"
      )
    )
    
    unlink(c(tmp_csv, tmp_png))
  }
}

```
#-------------------
#DESeq2 pseudobulk
#-------------------
```{r}
deseq2_results <- list()

for (cell_name in celltypes) {
  
  cell_idx <- which(pb_meta$celltype == cell_name)
  counts_ct <- pb_counts[, cell_idx, drop = FALSE]
  meta_ct <- pb_meta[cell_idx, , drop = FALSE]
  
  for (pair in group_pairs) {
    
    pair_name <- paste(pair, collapse = "_vs_")
    pair_name_clean <- gsub("[^A-Za-z0-9_]+", "_", pair_name)
    cell_name_clean <- gsub("[^A-Za-z0-9_]+", "_", cell_name)
    
    cat("\n=== DESeq2:", cell_name, pair_name, "===\n")
    
    keep_pair <- meta_ct$group %in% pair
    counts_sub <- round(counts_ct[, keep_pair, drop = FALSE])
    meta_sub <- meta_ct[keep_pair, , drop = FALSE]
    
    meta_sub$group <- factor(meta_sub$group, levels = pair)
    meta_sub$sex <- factor(meta_sub$sex)
    rownames(meta_sub) <- colnames(counts_sub)
    
    donor_tab <- table(meta_sub$group)
    
    if (any(donor_tab < min_donors_per_group)) {
      cat("Skipping: too few donors\n")
      next
    }
    
    # Use the same expression filtering strategy as edgeR
    filter_design <- model.matrix(~ group + age + sex, data = meta_sub)
    
    y_tmp <- DGEList(counts = counts_sub)
    
    keep_genes <- filterByExpr(
      y_tmp,
      design = filter_design
    )
    
    counts_sub <- counts_sub[keep_genes, , drop = FALSE]
    
    if (nrow(counts_sub) == 0) next
    
    dds <- DESeqDataSetFromMatrix(
      countData = counts_sub,
      colData = meta_sub,
      design = ~ age + sex + group
    )
    
    dds <- DESeq(dds)
    
    res <- results(
      dds,
      contrast = c("group", pair[2], pair[1])
    )
    
    tt <- as.data.frame(res)
    
    cat("\nColumns returned:\n")
    print(colnames(tt))
    
    tt$gene <- rownames(tt)
    tt$comparison <- pair_name
    tt$celltype <- cell_name
    tt$n_donors_group1 <- as.numeric(donor_tab[1])
    tt$n_donors_group2 <- as.numeric(donor_tab[2])
    
    tt <- tt %>%
      dplyr::rename(
        logFC = log2FoldChange,
        FDR = padj,
        PValue = pvalue
      )
    
    deseq2_results[[paste(cell_name_clean, pair_name_clean, sep = "__")]] <- tt
    
    tmp_csv <- tempfile(fileext = ".csv")
    write.csv(tt, tmp_csv, row.names = FALSE)
    
    upload_to_kopah(
      file = tmp_csv,
      object = paste0(
        "results/pseudobulk/DESeq2/all_cells/pseudo_all_DESeq2_",
        pair_name_clean, "_", cell_name_clean, ".csv"
      )
    )
    
    volcano_plot <- plot_volcano(
      tt,
      title = paste("DESeq2", cell_name, pair_name),
      fdr_col = "FDR",
      lfc_col = "logFC"
    )
    
    tmp_png <- tempfile(fileext = ".png")
    
    ggsave(
      tmp_png,
      plot = volcano_plot,
      width = 6,
      height = 5,
      dpi = 300
    )
    
    upload_to_kopah(
      file = tmp_png,
      object = paste0(
        "figures/pseudobulk/DESeq2/all_cells/volcano_all_DESeq2_",
        pair_name_clean, "_", cell_name_clean, ".png"
      )
    )
    
    unlink(c(tmp_csv, tmp_png))
  }
}
```


```{r}
method_compare_summary <- purrr::map_dfr(
  intersect(names(edgeR_results), names(deseq2_results)),
  function(nm) {
    
    edgeR_tt <- edgeR_results[[nm]]
    deseq2_tt <- deseq2_results[[nm]]
    
    merged_de <- dplyr::inner_join(
      edgeR_tt %>%
        dplyr::select(gene, edgeR_logFC = logFC, edgeR_FDR = FDR),
      deseq2_tt %>%
        dplyr::select(gene, DESeq2_logFC = logFC, DESeq2_FDR = FDR),
      by = "gene"
    )
    
    tibble::tibble(
      result = nm,
      n_genes = nrow(merged_de),
      edgeR_sig = sum(merged_de$edgeR_FDR < 0.05, na.rm = TRUE),
      DESeq2_sig = sum(merged_de$DESeq2_FDR < 0.05, na.rm = TRUE),
      both_sig = sum(
        merged_de$edgeR_FDR < 0.05 &
          merged_de$DESeq2_FDR < 0.05,
        na.rm = TRUE
      ),
      logFC_cor = cor(
        merged_de$edgeR_logFC,
        merged_de$DESeq2_logFC,
        use = "complete.obs"
      )
    )
  }
)

method_compare_summary
method_compare_summary <- method_compare_summary %>%
  dplyr::rename(
    `Correlation of LFC effect size` = logFC_cor
  )

method_compare_summary
tmp <- tempfile(fileext = ".csv")
write.csv(method_compare_summary, tmp, row.names = FALSE)

upload_to_kopah(
  file = tmp,
  object = "results/pseudobulk/method_compare_edgeR_vs_DESeq2.csv"
)
```
#------------------------------------------------------------
#plotting cell type specific expression patterns per contrast
#------------------------------------------------------------
```{r}
pb_files <- list.files(
  file.path(dir.results,"T1D-T2D", "Results", "pseudobulk_1"),
  pattern = "^pseudobulk_edgeR_.*\\.csv$",
  full.names = TRUE
)

de_all <- map_dfr(pb_files, read.csv)
de_all

de_counts <- de_all %>%
  mutate(
    direction = dplyr::case_when(
      FDR < 0.05 & logFC > 0  ~ "Positive",
      FDR < 0.05 & logFC < 0  ~ "Negative",
      TRUE ~ "NS"
    )
  ) %>%
  filter(direction != "NS") %>%
  group_by(comparison, celltype, direction) %>%
  summarise(n_de = n(), .groups = "drop") %>%
  mutate(
    n_plot = ifelse(direction == "Negative", -n_de, n_de)
  )
```

```{r}
ggplot(de_counts, aes(x = n_plot, y = celltype, fill = direction)) +
  geom_col(width = 0.75) +
  geom_vline(xintercept = 0, color = "gray40") +
  facet_wrap(~ comparison, scales = "free_x") +
  scale_x_continuous(labels = abs) +
  scale_fill_manual(
    values = c("Negative" = "purple4", "Positive" = "orange2")
  ) +
  labs(
    x = "Number of DE genes (FDR < 0.05)",
    y = NULL,
    fill = NULL
  ) +
  theme_classic() +
  theme(
    strip.text = element_text(face = "bold", size = 10),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 9),
    legend.position = "top"
  )

ggsave("cell_type_exrpession.png",
       path = file.path(dir.results, "T1D-T2D", "Results"),
       width = 10,
       height = 6)
```

#------------------------------------------------------------------
#plot to create panel volcanos for a specific cell type of interest
#-------------------------------------------------------------------
```{r}
celltype_of_interest <- "PT"  # <-- change this

de_ct <- de_all %>%
  filter(celltype == celltype_of_interest) %>%
  mutate(
    neg_log10_fdr = -log10(FDR),
    sig = case_when(
      FDR < 0.05 & logFC > 0.5  ~ "Up",
      FDR < 0.05 & logFC < -0.5 ~ "Down",
      TRUE ~ "NS"
    )
  )

label_df <- de_ct %>%
  filter(FDR < 0.05) %>%
  group_by(comparison) %>%
  arrange(FDR, .by_group = TRUE) %>%
  slice_head(n = 5) %>%
  ungroup()

volcano_celltype_panel <- ggplot(de_ct, aes(x = logFC, y = neg_log10_fdr, color = sig)) +
  geom_point(alpha = 0.6, size = 1.3) +
  geom_text_repel(
    data = label_df,
    aes(label = gene),
    size = 5,
    max.overlaps = Inf
  ) +
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ comparison) +
  scale_color_manual(values = c("Up" = "orange2", "Down" = "purple4", "NS" = "grey")) +
  labs(
    title = paste("Volcano plots for", celltype_of_interest),
    x = "log2 Fold Change",
    y = "-log10(FDR)",
    color = NULL
  ) +
  theme_classic() +
  theme(
    strip.text = element_text(face = "bold", size = 14),
    legend.position = "top" + 
  theme(axis.title = element_text(size = 12))
  )

volcano_celltype_panel
  
#ggsave(filename = "TAL_volcano_plots.png",
 #       path = file.path(dir.results, "T1D-T2D", "RESULTS"),
  #      width = 12,
   #     height = 10)

#ggsave(filename = "EC_volcano_plots.png",
 #       path = file.path(dir.results, "T1D-T2D", "RESULTS"),
  #      width = 12,
   #     height = 10)

#ggsave(filename = "Immune_volcano_plots.png",
 #       path = file.path(dir.results, "T1D-T2D", "RESULTS"),
  #      width = 12,
   #     height = 10)

ggsave(filename = "PT_volcano_plots.png",
        path = file.path(dir.results, "T1D-T2D", "RESULTS"),
        width = 12,
        height = 10)
```

#GSEA anlaysis
```{r}
# choose cell type
celltype_of_interest <- "TAL"  # <-- change this

# choose MSigDB gene sets
msig_reactome <- msigdbr(species = "Homo sapiens", collection = "C2", subcollection = "CP:REACTOME")
msig_go_bp <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:BP")
msig_go_mf <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:MF")
msig_go_cc <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:CC")

msig_all <- bind_rows(
  msig_reactome,
  msig_go_bp,
  msig_go_mf,
  msig_go_cc
)

pathways <- msig_all %>%
  split(x = .$gene_symbol, f = .$gs_name)
```

# -----------------------------
# fgsea function for one celltype
# -----------------------------
```{r}
run_fgsea_one <- function(tt, pathways, minSize = 15, maxSize = 500) {
  
  ranks <- tt %>%
    filter(!is.na(gene), !is.na(logFC), !is.na(PValue)) %>%
    mutate(rank_metric = sign(logFC) * -log10(PValue)) %>%
    group_by(gene) %>%
    slice_max(order_by = abs(rank_metric), n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    arrange(desc(rank_metric)) %>%
    select(gene, rank_metric)
  
  stats <- ranks$rank_metric
  names(stats) <- ranks$gene
  
  fgsea(
    pathways = pathways,
    stats = stats,
    minSize = minSize,
    maxSize = maxSize
  ) %>%
    as_tibble() %>%
    arrange(padj)
}
```

```{r}
de_ct <- de_all %>%
  filter(celltype == celltype_of_interest)

gsea_results <- list()

for (comp in unique(de_ct$comparison)) {
  
  message("Running GSEA: ", celltype_of_interest, " | ", comp)
  
  tt_sub <- de_ct %>%
    filter(comparison == comp)
  
  gsea_results[[comp]] <- run_fgsea_one(
    tt = tt_sub,
    pathways = pathways
  ) %>%
    mutate(
      comparison = comp,
      celltype = celltype_of_interest
    )
}
```

```{r}
gsea_table <- bind_rows(gsea_results)

gsea_table %>%
  select(celltype, comparison, pathway, NES, pval, padj, size) %>%
  arrange(padj)
```
```{r}
#a lot of ribosomal & mRNA processing check for contamination 
gsea_plot_sig <- gsea_table %>%
  filter(comparison == "Type 1 Diabetes_vs_Type 2 Diabetes", padj < 0.05) %>% 
  filter(celltype == celltype_of_interest)
gsea_plot_sig
```

```{r}
# choose one contrast
contrast_of_interest <- unique(gsea_table$comparison)[1]
```

```{r}
library(forcats)

celltype_of_interest <- "TAL"  # change this
gene_set_to_plot <- "REACTOME"     # options: , "REACTOME", "GO:..."

gsea_plot_df <- gsea_table %>%
  filter(celltype == celltype_of_interest) %>%
  mutate(
    gene_set = case_when(
      str_detect(pathway, "REACTOME") ~ "REACTOME",
      str_detect(pathway, "GOBP|GOMF|GOCC|GO_") ~ "GO",
      TRUE ~ "Other"
    )
  ) %>%
  filter(gene_set == gene_set_to_plot) %>%
  filter(padj < 0.05) %>%
  mutate(
    Direction = ifelse(NES > 0, "Up", "Down"),
    neg_log10_fdr = -log10(padj),
    pathway_clean = pathway %>%
      str_replace_all("_", " ") %>%
      str_replace("^REACTOME ", "Reactome ") %>%
      str_replace("^GOBP ", "GO BP ") %>%
      str_to_sentence()
  ) %>%
  group_by(comparison) %>%
  arrange(desc(abs(NES)), .by_group = TRUE) %>%
  slice_head(n = 10) %>%
  ungroup()

gsea_lollipop <- ggplot(
  gsea_plot_df,
  aes(x = NES, y = fct_reorder(pathway_clean, NES))
) +
  geom_segment(
    aes(x = 0, xend = NES, yend = pathway_clean),
    color = "grey70",
    linewidth = 0.5
  ) +
  geom_point(
    aes(color = Direction, size = neg_log10_fdr),
    alpha = 0.95
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~ comparison, scales = "free_y", ncol = 3) +
  scale_color_manual(
    values = c("Up" = "orange2", "Down" = "purple4")
  ) +
  scale_size_continuous(name = "-log10(FDR)") +
  labs(
    title = paste0(celltype_of_interest, "\n", gene_set_to_plot),
    x = "Normalized enrichment score",
    y = NULL,
    color = "Direction"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    strip.text = element_text(face = "bold", size = 10),
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(size = 9),
    legend.position = "right"
  )

gsea_lollipop


ggsave(filename = "GSEA_TAL.png",
       path = file.path(dir.results, "T1D-T2D", "Results"),
       width = 26,
       height = 18)
```

#do endothelial cell subpopulation DEPs
