---
title: "scRNA_data_analysis"
author: "Savanah Leidholt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths("~/leidholt/R")
```

```{r}
library(Seurat)
library(tidyverse)
library(arsenal)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(Hmisc)
library(readxl)
library(purrr)
library(aws.s3)
library(jsonlite)
library(RColorBrewer)
library(ggtrace)
library(kableExtra)
library(data.table)
library(table1)
library(speckle)
library(GSA)
library(emmeans)
library(pheatmap)
library(enrichplot)
library(enrichR)
library(nebula)
library(doParallel)
library(Matrix)
library(edgeR)
library(stringr)
library(patchwork)
library(msigdbr)
library(fgsea)
```

```{r}
#directories
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "choiyej") {
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad/"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "yejichoi") {
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "hhampson") {
  root_path <- "/Volumes/Peds Endo"
} else if (user == "sleidholt") {
  root_path <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
  git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/T1D Adipostiy/"
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r}
## Create an S3 client
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)


pb90_multiomics_subset <- s3readRDS(
          object = "data_clean/pb90_multiomics_subset_savanah_20260420.rds", 
          bucket = "scrna",
          region = "")
```

```{r}
my_colors <- colorRampPalette(brewer.pal(12, "Set3"))(43)
# Read Seurat object and run UMAP, obtain cell numbers, etc
#pb90_multiomics_subset_meta <- pb90_multiomics_subset@meta.data %>%
 # cbind(pb90_multiomics_subset@reductions$umap.harmony@cell.embeddings)

#s3saveRDS(pb90_multiomics_subset_meta, 
# object = "data_clean/pb90_multiomics_subset_meta_savanah_20260420.rds", 
# bucket = "scrna", 
# region = "", 
# multipart = T)
```

```{r}
# start from here if file already saved
pb90_multiomics_subset_meta <- s3readRDS(
  object = "data_clean/pb90_multiomics_subset_meta_savanah_20260420.rds", 
  bucket = "scrna", 
  region = "")

centers <- pb90_multiomics_subset_meta %>%
  group_by(KPMP_celltype) %>%
  summarise(
    x = median(umapharmony_1),
    y = median(umapharmony_2),
    .groups = "drop"
  )

umap_p <- ggplot(pb90_multiomics_subset_meta,
                 aes(x = umapharmony_1, y = umapharmony_2, color = KPMP_celltype)) +
  geom_point(alpha = 0.2, shape = 16) +
  geom_text_repel(
    data = centers,
    aes(x = x, y = y, label = KPMP_celltype),
    inherit.aes = FALSE,
    size = 5,                 # slightly larger for readability
    fontface = "bold",
    color = "black"
  ) +
  scale_color_manual(values = my_colors) +
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),
    
    # axis titles + tick labels (match previous plots)
    axis.title.x = element_text(size = 16, hjust = 0),
    axis.title.y = element_text(size = 16, hjust = 0),
    axis.text = element_text(size = 14),
    
    # no legend
    legend.position = "none"
  ) +
  labs(x = "UMAP1", y = "UMAP2")

umap_p
```

```{r}
pb90_multiomics_subset_meta%>%
  distinct(record_id, group)
# UMAP by source
umap_p_study<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = umapharmony_1, y = umapharmony_2, color = study)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("RENAL-HEIR" = "darkcyan",
                                "RENAL-HEIRitage"= "gold4",
                                "IMPROVE" = "orange",
                                "CROCODILE" = "black",
                                "ATTEMPT" = "lightpink"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(size = 16, hjust = 0),   
        axis.title.y = element_text(size = 16, hjust = 0), 
        axis.text = element_text(size = 14),

 # legend
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)) + 
  labs(x = "UMAP1", y = "UMAP2",
       color = "study") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_study
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)
# UMAP by source
umap_p_group<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = umapharmony_1, y = umapharmony_2, color = group)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("Obese Control" = "darkcyan",
                                "Lean Control"= "plum4",
                                "Type 1 Diabetes" = "orange",
                                "Type 2 Diabetes" = "black"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(size = 16, hjust = 0),   
        axis.title.y = element_text(size = 16, hjust = 0),
        axis.title = element_text(size = 14),

        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14))+
  labs(x = "UMAP1", y = "UMAP2",
       color = "group") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_group
```
```{r}
umap_cells_study <- umap_p + umap_p_study 
umap_cells_group <- umap_p + umap_p_group

ggsave("Umap_cells_study.png",
       plot = umap_cells_study,
       path = file.path(dir.results, "T1D-T2D", "Results"),
       width = 20,
       height =16,
       dpi = 300
       )

ggsave("Umap_cells_group.png",
       plot = umap_cells_group,
       path = file.path(dir.results, "T1D-T2D", "Results"),
       width = 20,
       height =16,
       dpi = 300
       )

```

## b. Descriptive Statistics
```{r}
#Load Metadata cleaned
#Table 1
label(pb90_multiomics_subset_meta$age) <- "Age (years)"
label(pb90_multiomics_subset_meta$sex) <- "Sex"
label(pb90_multiomics_subset_meta$race_ethnicity_condensed) <- "Ethnicity"
label(pb90_multiomics_subset_meta$bmi) <- "BMI (kg/m^2)"
label(pb90_multiomics_subset_meta$hba1c) <- "HbA1c (%)"
label(pb90_multiomics_subset_meta$acr_u) <- "Urine Albumin-Creatinine Ratio (mg/g)"
label(pb90_multiomics_subset_meta$eGFR_fas_cr) <- "eGFR by Full Age Spectrum (mL/min/1.73m^2)"
label(pb90_multiomics_subset_meta$group) <- "Diabetes Status (T1D, T2D, OC, LC)"

table1(~ age + sex + race_ethnicity_condensed + bmi  + hba1c + acr_u + eGFR_fas_cr + group| group, data= pb90_multiomics_subset_meta)

continuous_vars <- c("age", "bmi", "hba1c", "acr_u", "eGFR_fas_cr")

#uneven groups so using kruskal wallis
for (var in continuous_vars) {
  formula <- as.formula(paste(var, "~ group"))
  kw <- kruskal.test(formula, data = pb90_multiomics_subset_meta)
  cat(var, "Kruskal-Wallis p-value:", round(kw$p.value, 4), "\n")
}

# Categorical variables - chi-square tests
categorical_vars <- c("sex", "race_ethnicity_condensed")

for (var in categorical_vars) {
  chi_test <- chisq.test(table(pb90_multiomics_subset_meta[[var]], pb90_multiomics_subset_meta$group))
  cat(var, "p-value:", round(chi_test$p.value, 4), "\n")
}

```

```{r}
cell_counts <- pb90_multiomics_subset_meta %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n))

max_n <- max(cell_counts$n, na.rm = TRUE)

label_pos <- -max_n * 0.05  
xmin_lim  <- -max_n * 0.35  
xmax_lim  <-  max_n * 1.15   

cell_count_bar <- ggplot(cell_counts, 
                         aes(x = reorder(KPMP_celltype_general, n), y = n)) +
  geom_col(fill = "#97a97c") +
  

  geom_text(aes(y = n, label = n),
            hjust = -0.1, size = 4) +
  
  geom_text(aes(y = label_pos, label = KPMP_celltype_general),
            hjust = 1, size = 4) +
  
  coord_flip(clip = "off") +
  
  scale_y_continuous(
    limits = c(xmin_lim, xmax_lim),
    expand = expansion(mult = c(0, 0))
  ) +
  
  labs(x = NULL, y = "Count") +
  
  theme(
    text = element_text(size = 15),
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.margin = margin(10, 30, 10, 30)
  ) +
  theme_transparent()

cell_count_bar

#ggsave("cell_counts.pdf",
 #      plot = cell_count_bar,
  #     path = file.path(dir.results, "T1D-T2D", "Results"),
   #    width = 10,
    #   height = 6)
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)

# cell counts per group
cell_counts_group <- pb90_multiomics_subset_meta %>%
  group_by(group) %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n))

cell_group_bar <- ggplot(cell_counts_group, aes(x = reorder(KPMP_celltype_general, n), y = n, fill = group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Lean Control" = "plum4",
                               "Obese Control" = "darkcyan",
                               "Type 1 Diabetes" = "orange2",
                               "Type 2 Diabetes"= "black")) +
  coord_flip() +
  labs(x = NULL, 
       y = "Count",
       fill = "Group") +
  theme_transparent() +
  theme(
    text = element_text(size = 15),
    panel.grid = element_blank(),
    plot.margin = margin(1, 0.5, 1, 0.5),
    
    legend.position = c(0.7, 0.5),
    legend.justification = c(0, 0.5),
    legend.background = element_blank(),
    legend.margin = margin(0, 0, 0, 0)
  ) +
  
  scale_y_continuous(
    limits = c(xmin_lim, xmax_lim),
    expand = expansion(mult = c(0, 0))
  ) 

cell_group_bar
#ggsave("cell_gounts_group.pdf",
 #      plot = cell_group_bar,
  #     path = file.path(dir.results, "T1D-T2D", "Results"),
   #    width = 10,
    #   height = 6)
```

```{r}
cells <- cell_count_bar + cell_group_bar 
cells
ggsave("cell_counts.pdf",
       plot = cells,
       path = file.path(dir.results, "T1D-T2D", "Results"),
       width = 16,
       height = 10,
       dpi = 600)

```


```{r}
# Barplot sample proportion by celltype 
ggplot(pb90_multiomics_subset@meta.data) +
    geom_bar(aes(x=study, fill=KPMP_celltype_general), 
                 position=position_fill(), color="black") +
    theme_classic() +
    labs(x="Study", y="Proportion of Cells")

ggplot(pb90_multiomics_subset@meta.data) +
    geom_bar(aes(x=group, fill=KPMP_celltype_general), 
                 position=position_fill(), color="black") +
    theme_classic() +
    labs(x="group", y="Proportion of Cells")
```


```{r}
pb90_multiomics_subset@meta.data$cell_type_legend <- paste0(
  pb90_multiomics_subset@meta.data$KPMP_celltype_general, 
  " = ", 
  pb90_multiomics_subset@meta.data$cell_type
)
nrow(pb90_multiomics_subset) # genes 31974
ncol(pb90_multiomics_subset) # cells 307386
```
```{r}
# Check your metadata
table(pb90_multiomics_subset$group)
table(pb90_multiomics_subset$kit_id)
table(pb90_multiomics_subset$source)
table(pb90_multiomics_subset$study)
table(pb90_multiomics_subset$KPMP_celltype_general)
head(pb90_multiomics_subset@meta.data[, c("group", "kit_id", "record_id", "KPMP_celltype_general")])
```
#------------------
#propeller analysis
#------------------
```{r}
groups      <- unique(pb90_multiomics_subset_meta$group)
group_pairs <- combn(groups, 2, simplify = FALSE)  # all pairwise combos

propeller_results <- list()

for (pair in group_pairs) {
  pair_name <- paste(pair, collapse = "_vs_")
  cat("\n=== Propeller:", pair_name, "===\n")
  
  # Subset metadata to just these two groups
  meta_sub <- pb90_multiomics_subset_meta[pb90_multiomics_subset_meta$group %in% pair, ]
  meta_sub$group <- droplevels(factor(meta_sub$group))
  
  res <- propeller(
    clusters  = meta_sub$KPMP_celltype_general,
    sample    = meta_sub$record_id,
    group     = meta_sub$group,
    transform = "logit"
  )
  
  res$comparison <- pair_name
  propeller_results[[pair_name]] <- res
  
 # write.csv(
  #  x    = res,
   # file = file.path(dir.results, "T1D-T2D", "Results",
    #                 paste0("propeller_", pair_name, ".csv")),
    #row.names = TRUE
  #)
}
```
#-----------------------------------------------------
#psuedobulk analysis since OC only have 3 particpants
#-----------------------------------------------------
```{r}
min_cells_per_sample <- 20     # minimum cells contributing to a pseudobulk sample
min_donors_per_group <- 3      # minimum donors per group required to run DE
min_total_count <- 10          # edgeR gene filter helper
min_samples_expr <- 2          # edgeR gene filter helper

# output directory
pb_out_dir <- file.path(dir.results, "T1D-T2D", "Results", "pseudobulk_1")
dir.create(pb_out_dir, recursive = TRUE, showWarnings = FALSE)
```

#----------------------------
# prepping metadata for edgeR
#----------------------------
```{r}
meta <- pb90_multiomics_subset@meta.data %>%
  mutate(
    cell_barcode = rownames(pb90_multiomics_subset@meta.data),
    record_id = as.character(record_id),
    group = as.character(group),
    sex = as.character(sex),
    study = as.character(study),
    celltype = as.character(KPMP_celltype_general)
  )

#keeping only complete metadata for modeling
meta <- meta %>%
  filter(
    !is.na(record_id),
    !is.na(group),
    !is.na(age),
    !is.na(sex),
    !is.na(study),
    !is.na(celltype)
  )

counts <- GetAssayData(pb90_multiomics_subset, layer = "counts")
counts <- counts[, meta$cell_barcode, drop = FALSE]

# Number of cells by sample and celltype
bulk <- AggregateExpression(
            pb90_multiomics_subset,
            return.seurat = TRUE,
            assays = "RNA",
            group.by = c("KPMP_celltype_general", "record_id", "group"))

n_cells <- pb90_multiomics_subset@meta.data %>% 
  dplyr::count(record_id, KPMP_celltype_general) %>% 
  dplyr::rename(n_cells = n)

n_cells$record_id <- str_replace(n_cells$record_id, "_", "-")

meta_bulk <- left_join(bulk@meta.data, n_cells)
rownames(meta_bulk) <- meta_bulk$orig.ident
bulk@meta.data <- meta_bulk
```

#----------------------------------------------------------
# building pseudobulk cell counts
# one pseudobulk sample = one donor within one cell type
#---------------------------------------------------------
```{r}
meta <- meta %>%
  mutate(sample_celltype_id = paste(record_id, celltype, sep = "__"))

pb_ids <- unique(meta$sample_celltype_id)

pb_count_list <- vector("list", length(pb_ids))
pb_meta_list  <- vector("list", length(pb_ids))

for (i in seq_along(pb_ids)) {
  sid <- pb_ids[i]
  idx <- which(meta$sample_celltype_id == sid)
  
  # skip tiny pseudobulks
  if (length(idx) < min_cells_per_sample) next
  
  sub_counts <- counts[, idx, drop = FALSE]
  agg_counts <- Matrix::rowSums(sub_counts)
  
  meta_sub <- meta[idx, ]
  
  pb_count_list[[i]] <- agg_counts
  pb_meta_list[[i]] <- data.frame(
    pseudobulk_id = sid,
    record_id = meta_sub$record_id[1],
    group = meta_sub$group[1],
    age = meta_sub$age[1],
    sex = meta_sub$sex[1],
    study = meta_sub$study[1],
    celltype = meta_sub$celltype[1],
    n_cells = nrow(meta_sub),
    stringsAsFactors = FALSE
  )
}

# remove NULL entries
keep_idx <- which(!sapply(pb_count_list, is.null))
pb_count_list <- pb_count_list[keep_idx]
pb_meta_list  <- pb_meta_list[keep_idx]

#combine into matrix + metadata
pb_counts <- do.call(cbind, pb_count_list)
colnames(pb_counts) <- sapply(pb_meta_list, function(x) x$pseudobulk_id)
pb_meta <- bind_rows(pb_meta_list)


pb_meta <- pb_meta %>% slice(match(colnames(pb_counts), pseudobulk_id))
stopifnot(identical(colnames(pb_counts), pb_meta$pseudobulk_id))
```

#------------------------
#running edgeR/cell type
#-----------------------
```{r}
#be careful how I compute the mean as this can be sensitive to outliers (less so with normalized data)
#not sure if I should aggregate by mean or by the sum of the cells. 
#setting up pairwise comparisons
groups <- sort(unique(pb_meta$group))
group_pairs <- combn(groups, 2, simplify = FALSE)

celltypes <- sort(unique(pb_meta$celltype))

for (cell_name in celltypes) {
  
  cat("\n============================\n")
  cat("Cell type:", cell_name, "\n")
  cat("============================\n")
  
  cell_idx <- which(pb_meta$celltype == cell_name)
  
  if (length(cell_idx) == 0) next
  
  counts_ct <- pb_counts[, cell_idx, drop = FALSE]
  meta_ct   <- pb_meta[cell_idx, , drop = FALSE]
  
  for (pair in group_pairs) {
    
    pair_name <- paste(pair, collapse = "_vs_")
    cat("\n--- Comparison:", pair_name, "---\n")
    
    keep_pair <- meta_ct$group %in% pair
    counts_sub <- counts_ct[, keep_pair, drop = FALSE]
    meta_sub   <- meta_ct[keep_pair, , drop = FALSE]
    
    meta_sub$group <- factor(meta_sub$group, levels = pair)
    meta_sub$sex   <- factor(meta_sub$sex)
    meta_sub$study <- factor(meta_sub$study)
    
    # donor counts per group
    donor_tab <- table(meta_sub$group)
    print(donor_tab)
    
    if (any(donor_tab < min_donors_per_group)) {
      cat("Skipping", pair_name, "for", cell_name, "- too few donors in at least one group\n")
      next
    }
    
    #design matrix
    design <- model.matrix(~ group + age + sex, data = meta_sub)
    
    #edgeR object
    y <- DGEList(counts = counts_sub)
    
    #filter low-expression genes
    keep_genes <- filterByExpr(y, design = design)
    y <- y[keep_genes, , keep.lib.sizes = FALSE]
    
    if (nrow(y) == 0) {
      cat("Skipping", pair_name, "for", cell_name, "- no genes passed filtering\n")
      next
    }
    
    #normalization
    y <- calcNormFactors(y)
    
    #estimate dispersion
    y <- estimateDisp(y, design)
    
    # fit the model
    fit <- glmQLFit(y, design, robust = TRUE)
    
    coef_name <- grep("^group", colnames(design), value = TRUE)
    
    if (length(coef_name) != 1) {
      cat("Skipping", pair_name, "for", cell_name, "- could not identify unique group coefficient\n")
      next
    }
    
    qlf <- glmQLFTest(fit, coef = which(colnames(design) == coef_name))
    tt <- topTags(qlf, n = Inf)$table
    tt$gene <- rownames(tt)
    tt$comparison <- pair_name
    tt$celltype <- cell_name
    tt$n_donors_group1 <- as.numeric(donor_tab[1])
    tt$n_donors_group2 <- as.numeric(donor_tab[2])
    tt$coef_tested <- coef_name
    
    #add average CPM
    logcpm <- cpm(y, log = TRUE)
    tt$mean_logCPM_check <- rowMeans(logcpm[tt$gene, , drop = FALSE])
    
    # save
    cell_name_file <- gsub("[[:space:]/-]+", "_", cell_name)
    out_file <- file.path(
      pb_out_dir,
      paste0("pseudobulk_edgeR_", pair_name, "_", cell_name_file, ".csv")
    )
    
    write.csv(tt, out_file, row.names = FALSE)
    cat("Saved:", out_file, "\n")
  
    
  plot_volcano <- function(tt, title = "", fdr_cutoff = 0.05, lfc_cutoff = 0.5, n_labels = 10) {
  
  tt <- tt %>%
    mutate(
      neg_log10_fdr = -log10(FDR),
      sig = case_when(
        FDR < fdr_cutoff & logFC > lfc_cutoff  ~ "Up",
        FDR < fdr_cutoff & logFC < -lfc_cutoff ~ "Down",
        TRUE ~ "NS"
      )
    )
 
  label_df <- tt %>%
    filter(FDR < fdr_cutoff) %>%
    arrange(FDR) %>%
    slice_head(n = n_labels)
  
  ggplot(tt, aes(x = logFC, y = neg_log10_fdr, color = sig)) +
    geom_point(alpha = 0.6, size = 1.5) +
    geom_text_repel(
      data = label_df,
      aes(label = gene),
      size = 3,
      max.overlaps = Inf,
      box.padding = 0.4,
      point.padding = 0.3
    ) +
    scale_color_manual(values = c("Up" = "red", "Down" = "blue", "NS" = "grey")) +
    geom_vline(xintercept = c(-lfc_cutoff, lfc_cutoff), linetype = "dashed") +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
    labs(
      title = title,
      x = "log2 Fold Change",
      y = "-log10(FDR)",
      color = ""
    ) +
    theme_classic()
}

    volcano_plot <- plot_volcano(
      tt,
      title = paste(cell_name, pair_name),
      fdr_cutoff = 0.05,
      lfc_cutoff = 0.5
    )
    
    ggsave(
      filename = file.path(pb_out_dir, paste0("volcano_", pair_name, "_", cell_name_file, ".png")),
      plot = volcano_plot,
      width = 6,
      height = 5,
      dpi = 300
    )
    
    cat("Saved volcano:", pair_name, cell_name, "\n")
  }
}

```

#------------------------------------------------------------
#plotting cell type specific expression patterns per contrast
#------------------------------------------------------------
```{r}
pb_files <- list.files(
  file.path(dir.results,"T1D-T2D", "Results", "pseudobulk_1"),
  pattern = "^pseudobulk_edgeR_.*\\.csv$",
  full.names = TRUE
)

de_all <- map_dfr(pb_files, read.csv)
de_all

de_counts <- de_all %>%
  mutate(
    direction = dplyr::case_when(
      FDR < 0.05 & logFC > 0  ~ "Positive",
      FDR < 0.05 & logFC < 0  ~ "Negative",
      TRUE ~ "NS"
    )
  ) %>%
  filter(direction != "NS") %>%
  group_by(comparison, celltype, direction) %>%
  summarise(n_de = n(), .groups = "drop") %>%
  mutate(
    n_plot = ifelse(direction == "Negative", -n_de, n_de)
  )
```

```{r}
ggplot(de_counts, aes(x = n_plot, y = celltype, fill = direction)) +
  geom_col(width = 0.75) +
  geom_vline(xintercept = 0, color = "gray40") +
  facet_wrap(~ comparison, scales = "free_x") +
  scale_x_continuous(labels = abs) +
  scale_fill_manual(
    values = c("Negative" = "purple4", "Positive" = "orange2")
  ) +
  labs(
    x = "Number of DE genes (FDR < 0.05)",
    y = NULL,
    fill = NULL
  ) +
  theme_classic() +
  theme(
    strip.text = element_text(face = "bold", size = 10),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 9),
    legend.position = "top"
  )

ggsave("cell_type_exrpession.png",
       path = file.path(dir.results, "T1D-T2D", "Results"),
       width = 10,
       height = 6)
```

#------------------------------------------------------------------
#plot to create panel volcanos for a specific cell type of interest
#-------------------------------------------------------------------
```{r}
celltype_of_interest <- "PT"  # <-- change this

de_ct <- de_all %>%
  filter(celltype == celltype_of_interest) %>%
  mutate(
    neg_log10_fdr = -log10(FDR),
    sig = case_when(
      FDR < 0.05 & logFC > 0.5  ~ "Up",
      FDR < 0.05 & logFC < -0.5 ~ "Down",
      TRUE ~ "NS"
    )
  )

label_df <- de_ct %>%
  filter(FDR < 0.05) %>%
  group_by(comparison) %>%
  arrange(FDR, .by_group = TRUE) %>%
  slice_head(n = 5) %>%
  ungroup()

volcano_celltype_panel <- ggplot(de_ct, aes(x = logFC, y = neg_log10_fdr, color = sig)) +
  geom_point(alpha = 0.6, size = 1.3) +
  geom_text_repel(
    data = label_df,
    aes(label = gene),
    size = 5,
    max.overlaps = Inf
  ) +
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  facet_wrap(~ comparison) +
  scale_color_manual(values = c("Up" = "orange2", "Down" = "purple4", "NS" = "grey")) +
  labs(
    title = paste("Volcano plots for", celltype_of_interest),
    x = "log2 Fold Change",
    y = "-log10(FDR)",
    color = NULL
  ) +
  theme_classic() +
  theme(
    strip.text = element_text(face = "bold", size = 14),
    legend.position = "top" + 
  theme(axis.title = element_text(size = 12))
  )

volcano_celltype_panel
  
#ggsave(filename = "TAL_volcano_plots.png",
 #       path = file.path(dir.results, "T1D-T2D", "RESULTS"),
  #      width = 12,
   #     height = 10)

#ggsave(filename = "EC_volcano_plots.png",
 #       path = file.path(dir.results, "T1D-T2D", "RESULTS"),
  #      width = 12,
   #     height = 10)

#ggsave(filename = "Immune_volcano_plots.png",
 #       path = file.path(dir.results, "T1D-T2D", "RESULTS"),
  #      width = 12,
   #     height = 10)

ggsave(filename = "PT_volcano_plots.png",
        path = file.path(dir.results, "T1D-T2D", "RESULTS"),
        width = 12,
        height = 10)
```

#GSEA anlaysis
```{r}
# choose cell type
celltype_of_interest <- "TAL"  # <-- change this

# choose MSigDB gene sets
msig_reactome <- msigdbr(species = "Homo sapiens", collection = "C2", subcollection = "CP:REACTOME")
msig_go_bp <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:BP")
msig_go_mf <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:MF")
msig_go_cc <- msigdbr(species = "Homo sapiens", collection = "C5", subcollection = "GO:CC")

msig_all <- bind_rows(
  msig_reactome,
  msig_go_bp,
  msig_go_mf,
  msig_go_cc
)

pathways <- msig_all %>%
  split(x = .$gene_symbol, f = .$gs_name)
```

# -----------------------------
# fgsea function for one celltype
# -----------------------------
```{r}
run_fgsea_one <- function(tt, pathways, minSize = 15, maxSize = 500) {
  
  ranks <- tt %>%
    filter(!is.na(gene), !is.na(logFC), !is.na(PValue)) %>%
    mutate(rank_metric = sign(logFC) * -log10(PValue)) %>%
    group_by(gene) %>%
    slice_max(order_by = abs(rank_metric), n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    arrange(desc(rank_metric)) %>%
    select(gene, rank_metric)
  
  stats <- ranks$rank_metric
  names(stats) <- ranks$gene
  
  fgsea(
    pathways = pathways,
    stats = stats,
    minSize = minSize,
    maxSize = maxSize
  ) %>%
    as_tibble() %>%
    arrange(padj)
}
```

```{r}
de_ct <- de_all %>%
  filter(celltype == celltype_of_interest)

gsea_results <- list()

for (comp in unique(de_ct$comparison)) {
  
  message("Running GSEA: ", celltype_of_interest, " | ", comp)
  
  tt_sub <- de_ct %>%
    filter(comparison == comp)
  
  gsea_results[[comp]] <- run_fgsea_one(
    tt = tt_sub,
    pathways = pathways
  ) %>%
    mutate(
      comparison = comp,
      celltype = celltype_of_interest
    )
}
```

```{r}
gsea_table <- bind_rows(gsea_results)

gsea_table %>%
  select(celltype, comparison, pathway, NES, pval, padj, size) %>%
  arrange(padj)
```
```{r}
#a lot of ribosomal & mRNA processing check for contamination 
gsea_plot_sig <- gsea_table %>%
  filter(comparison == "Type 1 Diabetes_vs_Type 2 Diabetes", padj < 0.05) %>% 
  filter(celltype == celltype_of_interest)
gsea_plot_sig
```

```{r}
# choose one contrast
contrast_of_interest <- unique(gsea_table$comparison)[1]
```

```{r}
library(forcats)

celltype_of_interest <- "TAL"  # change this
gene_set_to_plot <- "REACTOME"     # options: , "REACTOME", "GO:..."

gsea_plot_df <- gsea_table %>%
  filter(celltype == celltype_of_interest) %>%
  mutate(
    gene_set = case_when(
      str_detect(pathway, "REACTOME") ~ "REACTOME",
      str_detect(pathway, "GOBP|GOMF|GOCC|GO_") ~ "GO",
      TRUE ~ "Other"
    )
  ) %>%
  filter(gene_set == gene_set_to_plot) %>%
  filter(padj < 0.05) %>%
  mutate(
    Direction = ifelse(NES > 0, "Up", "Down"),
    neg_log10_fdr = -log10(padj),
    pathway_clean = pathway %>%
      str_replace_all("_", " ") %>%
      str_replace("^REACTOME ", "Reactome ") %>%
      str_replace("^GOBP ", "GO BP ") %>%
      str_to_sentence()
  ) %>%
  group_by(comparison) %>%
  arrange(desc(abs(NES)), .by_group = TRUE) %>%
  slice_head(n = 10) %>%
  ungroup()

gsea_lollipop <- ggplot(
  gsea_plot_df,
  aes(x = NES, y = fct_reorder(pathway_clean, NES))
) +
  geom_segment(
    aes(x = 0, xend = NES, yend = pathway_clean),
    color = "grey70",
    linewidth = 0.5
  ) +
  geom_point(
    aes(color = Direction, size = neg_log10_fdr),
    alpha = 0.95
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~ comparison, scales = "free_y", ncol = 3) +
  scale_color_manual(
    values = c("Up" = "orange2", "Down" = "purple4")
  ) +
  scale_size_continuous(name = "-log10(FDR)") +
  labs(
    title = paste0(celltype_of_interest, "\n", gene_set_to_plot),
    x = "Normalized enrichment score",
    y = NULL,
    color = "Direction"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    strip.text = element_text(face = "bold", size = 10),
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(size = 9),
    legend.position = "right"
  )

gsea_lollipop


ggsave(filename = "GSEA_TAL.png",
       path = file.path(dir.results, "T1D-T2D", "Results"),
       width = 26,
       height = 18)
```

#do endothelial cell subpopulation DEPs
