---
title: "scRNA_data_analysis"
author: "Savanah Leidholt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths("~/leidholt/R")
```

```{r}
#BiocManager::install("Seurat")
#BiocManager::install("SeuratObject")
library(Seurat)
library(tidyverse)
library(arsenal)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(Hmisc)
library(readxl)
library(purrr)
library(aws.s3)
library(jsonlite)
library(RColorBrewer)
#install.packages("ggtrace")
library(ggtrace)
library(kableExtra)
library(data.table)
library(table1)
#BiocManager::install("speckle")
library(speckle)
#BiocManager::install("GSA")
library(GSA)
library(emmeans)
library(pheatmap)
#BiocManager::install("enrichplot")
library(enrichplot)
library(enrichR)
#BiocManager::install("nebula")
library(nebula)
#BiocManager::install("doParallel")
library(doParallel)
library(Matrix)
library(edgeR)
library(stringr)
```

```{r}
#directores
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "choiyej") {
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad/"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "yejichoi") {
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "hhampson") {
  root_path <- "/Volumes/Peds Endo"
} else if (user == "sleidholt") {
  root_path <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
  git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/T1D Adipostiy/"
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r}
## Create an S3 client
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)


pb90_multiomics_subset <- s3readRDS(
          object = "data_clean/pb90_multiomics_subset_savanah_20260420.rds", 
          bucket = "scrna",
          region = "")
```

```{r}
my_colors <- colorRampPalette(brewer.pal(12, "Set3"))(43)
# Read Seurat object and run UMAP, obtain cell numbers, etc
#pb90_multiomics_subset_meta <- pb90_multiomics_subset@meta.data %>%
 # cbind(pb90_multiomics_subset@reductions$umap.harmony@cell.embeddings)

#s3saveRDS(pb90_multiomics_subset_meta, 
# object = "data_clean/pb90_multiomics_subset_meta_savanah_20260420.rds", 
# bucket = "scrna", 
# region = "", 
# multipart = T)
```


```{r}
# start from here if file already saved
pb90_multiomics_subset_meta <- s3readRDS(
  object = "data_clean/pb90_multiomics_subset_meta_savanah_20260420.rds", 
  bucket = "scrna", 
  region = "")

centers <- pb90_multiomics_subset_meta %>%
  group_by(KPMP_celltype) %>%
  summarise(x = median(umapharmony_1),
            y = median(umapharmony_2), .groups = "drop")

umap_p <- ggplot(pb90_multiomics_subset_meta,
                 aes(x = umapharmony_1, y = umapharmony_2, color = KPMP_celltype)) +
  geom_point(alpha = 0.2, shape = 16) +
  geom_text_repel(data = centers,
                  aes(x = x, y = y, label = KPMP_celltype),
                  inherit.aes = FALSE,
                  size = 4, fontface = "bold", color = "black") +
  scale_color_manual(values = my_colors) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(hjust = 0),   # move x label to left
        axis.title.y = element_text(hjust = 0))  + # push y label down to bottom) 
  labs(x = "UMAP1", y = "UMAP2")

umap_p
```

```{r}
# Save to a temporary connection
#tmp <- tempfile(fileext = ".png")
#ggsave(tmp, plot = umap_p, width = 10, height = 8, dpi = 300)

# Upload directly
#put_object(
 # file = tmp,
#  object = "results/figures/t1d_hc_umap_kpmpcelltype.png",
#  bucket = "scrna",
#  region = ""
#)
```

```{r}
pb90_multiomics_subset_meta%>%
  distinct(record_id, group)
# UMAP by source
umap_p_study<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = umapharmony_1, y = umapharmony_2, color = study)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("RENAL-HEIR" = "darkcyan",
                                "RENAL-HEIRitage"= "gold4",
                                "IMPROVE" = "orange",
                                "CROCODILE" = "black",
                                "ATTEMPT" = "lightpink"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(hjust = 0),   # move x label to left
        axis.title.y = element_text(hjust = 0))  + # push y label down to bottom) 
  labs(x = "UMAP1", y = "UMAP2",
       color = "study") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_study
```

```{r}
# Save to a temporary connection
#tmp <- tempfile(fileext = ".png")
#ggsave(tmp, plot = umap_p_source, width = 10, height = 8, dpi = 300)

# Upload directly
#put_object(
#  file = tmp,
#  object = "results/figures/t1d_hc_umap_source.png",
#  bucket = "t1d.adiposity",
#  region = ""
#)
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)
# UMAP by source
umap_p_study<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = umapharmony_1, y = umapharmony_2, color = group)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("Obese Control" = "darkcyan",
                                "Lean Control"= "plum4",
                                "Type 1 Diabetes" = "orange",
                                "Type 2 Diabetes" = "black"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(hjust = 0),   # move x label to left
        axis.title.y = element_text(hjust = 0))  + # push y label down to bottom) 
  labs(x = "UMAP1", y = "UMAP2",
       color = "group") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_study
```
```{r}
cell_counts <- pb90_multiomics_subset_meta %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n))

# Create bar plot
cell_count_bar <- ggplot(cell_counts, aes(x = reorder(KPMP_celltype_general, n), y = n)) +
  geom_bar(stat = "identity", fill = "#97a97c") +
  geom_text(aes(label = n), hjust = -0.1, size = 4) +
  coord_flip() +
  labs(x = NULL, 
       y = "Count") +
  theme(text = element_text(size = 15),
        legend.position = "none",
        panel.grid = element_blank()) +
  theme_transparent() +
  scale_y_continuous(expand = expansion(add = 5000)) 

cell_count_bar
```
```{r}
# Save to a temporary connection
#tmp <- tempfile(fileext = ".png")
#ggsave(tmp, plot = cell_count_bar, width = 10, height = 8, dpi = 300)

# Upload directly
#put_object(
#  file = tmp,
#  object = "results/figures/t1d_hc_kpmp_cell_count_bar.png",
#  bucket = "t1d.adiposity",
#  region = ""
#)
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)

# cell counts per group
cell_counts_group <- pb90_multiomics_subset_meta %>%
  group_by(group) %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n)) %>%
  dplyr::mutate(group = case_when(group == "Lean Control" ~ "LC",
                                  group == "Obese Control" ~ "OC",
                                  group == "Type 1 Diabetes" ~ "T1D",
                                  group == "Type 2 Diabetes" ~ "T2D"))

ggplot(cell_counts_group, aes(x = reorder(KPMP_celltype_general, n), y = n, fill = group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("T1D" = "orange2",
                               "LC" = "plum4",
                               "OC" = "darkcyan",
                               "T2D"= "black")) +
  coord_flip() +
  labs(x = NULL, 
       y = "Count",
       fill = "Group") +
  theme(text = element_text(size = 15),
        panel.grid = element_blank()) +
  theme_transparent() +
  scale_y_continuous(expand = expansion(add = 5000)) 
```

```{r}
# Save to a temporary connection
#tmp <- tempfile(fileext = ".png")
#ggsave(tmp, width = 10, height = 8, dpi = 300)

# Upload directly
#put_object(
#  file = tmp,
#  object = "results/figures/t1d_hc_grp_kpmp_cell_count_bar.png",
#  bucket = "t1d.adiposity",
#  region = ""
#)
```

## b. Descriptive Statistics
```{r}
#Load Metadata cleaned
#Table 1
label(pb90_multiomics_subset_meta$age) <- "Age (years)"
label(pb90_multiomics_subset_meta$sex) <- "Sex"
label(pb90_multiomics_subset_meta$race_ethnicity_condensed) <- "Ethnicity"
label(pb90_multiomics_subset_meta$bmi) <- "BMI (kg/m^2)"
label(pb90_multiomics_subset_meta$hba1c) <- "HbA1c (%)"
label(pb90_multiomics_subset_meta$acr_u) <- "Urine Albumin-Creatinine Ratio (mg/g)"
label(pb90_multiomics_subset_meta$eGFR_fas_cr) <- "eGFR by Full Age Spectrum (mL/min/1.73m^2)"
label(pb90_multiomics_subset_meta$group) <- "Diabetes Status (T1D, T2D, OC, LC)"

table1(~ age + sex + race_ethnicity_condensed + bmi  + hba1c + acr_u + eGFR_fas_cr| group, data= pb90_multiomics_subset_meta)

continuous_vars <- c("age", "bmi", "hba1c", "acr_u", "eGFR_fas_cr")

#uneven groups so using kruskal wallis
for (var in continuous_vars) {
  formula <- as.formula(paste(var, "~ group"))
  kw <- kruskal.test(formula, data = pb90_multiomics_subset_meta)
  cat(var, "Kruskal-Wallis p-value:", round(kw$p.value, 4), "\n")
}

# Categorical variables - chi-square tests
categorical_vars <- c("sex", "race_ethnicity_condensed")

for (var in categorical_vars) {
  chi_test <- chisq.test(table(pb90_multiomics_subset_meta[[var]], pb90_multiomics_subset_meta$group))
  cat(var, "p-value:", round(chi_test$p.value, 4), "\n")
}

```

```{r}
## a. Aim 1.1: Perform differential expression between all group comparisons (T1D, T2D, LC, OC)
### using nebula
# Verify dimensions
pb90_multiomics_subset@meta.data$cell_type_legend <- paste0(
  pb90_multiomics_subset@meta.data$cell_type_abbrev, 
  " = ", 
  pb90_multiomics_subset@meta.data$cell_type
)
nrow(pb90_multiomics_subset) # genes 31974
ncol(pb90_multiomics_subset) # cells 307386
```
```{r}
# Check your metadata
table(pb90_multiomics_subset$group)
table(pb90_multiomics_subset$kit_id)
table(pb90_multiomics_subset$source)
table(pb90_multiomics_subset$study)
table(pb90_multiomics_subset$KPMP_celltype_general)
head(pb90_multiomics_subset@meta.data[, c("group", "kit_id", "record_id", "KPMP_celltype_general")])
```

```{r}
groups      <- unique(pb90_multiomics_subset_meta$group)
group_pairs <- combn(groups, 2, simplify = FALSE)  # all pairwise combos

propeller_results <- list()

for (pair in group_pairs) {
  pair_name <- paste(pair, collapse = "_vs_")
  cat("\n=== Propeller:", pair_name, "===\n")
  
  # Subset metadata to just these two groups
  meta_sub <- pb90_multiomics_subset_meta[pb90_multiomics_subset_meta$group %in% pair, ]
  meta_sub$group <- droplevels(factor(meta_sub$group))  # drop unused levels
  
  res <- propeller(
    clusters  = meta_sub$KPMP_celltype_general,
    sample    = meta_sub$record_id,
    group     = meta_sub$group,
    transform = "logit"
  )
  
  res$comparison <- pair_name
  propeller_results[[pair_name]] <- res
  
 # write.csv(
  #  x    = res,
   # file = file.path(dir.results, "T1D-T2D", "Results",
    #                 paste0("propeller_", pair_name, ".csv")),
    #row.names = TRUE
  #)
}
```

# =========================================================
# CLEANER NEBULA PIPELINE
# - pairwise group comparisons
# - loop over cell types
# - run one gene at a time
# - append results directly to disk
# - minimal memory growth
# =========================================================

```{r}
n_hvg <- 500 #change to to 1000 then 2000 to see if it can still run
min_cells_per_celltype <- 50   # skip tiny cell types
overwrite_existing <- TRUE     # FALSE = skip if output file exists

# -----------------------------
# group pairs
# -----------------------------
groups <- unique(pb90_multiomics_subset_meta$group)
group_pairs <- combn(groups, 2, simplify = FALSE)

celltypes_list <- unique(pb90_multiomics_subset_meta$KPMP_celltype_general)

# -----------------------------
# main loop
# -----------------------------
for (pair in group_pairs) {
  
  pair_name <- paste(pair, collapse = "_vs_")
  cat("\n============================\n")
  cat("Comparison:", pair_name, "\n")
  cat("============================\n")
  
  # subset once per comparison
  seurat_pair <- subset(pb90_multiomics_subset, group %in% pair)
  seurat_pair$group <- droplevels(factor(seurat_pair$group, levels = pair))
  
  for (cell_name in celltypes_list) {
    
    cat("\n--- Cell type:", cell_name, "---\n")
    
    # subset once per cell type
    seurat_sub <- subset(seurat_pair, KPMP_celltype_general == cell_name)
    
    n_cells <- ncol(seurat_sub)
    if (n_cells < min_cells_per_celltype) {
      cat("Skipping", cell_name, "- too few cells:", n_cells, "\n")
      next
    }
    
    DefaultAssay(seurat_sub) <- "RNA"
    meta_full <- seurat_sub@meta.data
    
    # make sure factors are clean
    meta_full$group <- droplevels(factor(meta_full$group, levels = pair))
    meta_full$sex   <- factor(meta_full$sex)
    meta_full$study <- factor(meta_full$study)
    
    # model matrix ONCE per cell type
    pred_gene <- model.matrix(~ group + age + sex + study, data = meta_full)
    library_offset <- meta_full$pooled_offset
    subject_id <- meta_full$record_id
    
    # choose HVGs once per cell type
    seurat_sub <- FindVariableFeatures(
      seurat_sub,
      selection.method = "vst",
      nfeatures = n_hvg
    )
    
    genes_list <- VariableFeatures(seurat_sub)
    cat("Cells:", n_cells, "| HVGs:", length(genes_list), "\n")
    
    # counts matrix once per cell type
    counts_hvg <- GetAssayData(seurat_sub, layer = "counts")[genes_list, , drop = FALSE]
    counts_hvg <- round(counts_hvg)
    
    # output file
    cell_name_file <- gsub("[[:space:]/-]+", "_", cell_name)
    out_path <- file.path(
      dir.results, "T1D-T2D", "Results",
      paste0("nebula_", pair_name, "_", cell_name_file, ".csv")
    )
    
    if (file.exists(out_path) && !overwrite_existing) {
      cat("Skipping existing file:", out_path, "\n")
      next
    }
    
    if (file.exists(out_path) && overwrite_existing) {
      file.remove(out_path)
    }
    
    first_write <- TRUE
    error_log <- character(0)
    converged_vec <- c()
    
    start_time <- Sys.time()
    
    for (g in genes_list) {
      
      tryCatch({
        count_gene <- counts_hvg[g, , drop = FALSE]
        
        data_g_gene <- group_cell(
          count  = count_gene,
          id     = subject_id,
          pred   = pred_gene,
          offset = library_offset
        )
        
        if (is.null(data_g_gene)) {
          data_g_gene <- list(
            count  = count_gene,
            id     = subject_id,
            pred   = pred_gene,
            offset = library_offset
          )
        }
        
        fit <- nebula(
          count      = data_g_gene$count,
          id         = data_g_gene$id,
          pred       = data_g_gene$pred,
          offset     = data_g_gene$offset,
          ncore      = 1,
          reml       = TRUE,
          model      = "NBLMM",
          output_re  = FALSE,
          covariance = FALSE
        )
        
        # store convergence code separately
        conv_code <- fit$convergence
        converged_vec[g] <- conv_code
        
        # extract result summary
        res <- as.data.frame(fit$summary)
        res$gene <- g
        res$comparison <- pair_name
        res$celltype <- cell_name
        res$cells <- n_cells
        
        # append directly to disk
        write.table(
          res,
          file = out_path,
          sep = ",",
          row.names = FALSE,
          col.names = first_write,
          append = !first_write
        )
        
        first_write <- FALSE
        
        rm(count_gene, data_g_gene, fit, res)
        gc(verbose = FALSE)
        
      }, error = function(e) {
        msg <- paste("Gene:", g, "| Error:", conditionMessage(e))
        cat(msg, "\n")
        error_log <<- c(error_log, msg)
      })
    }
    
    elapsed <- difftime(Sys.time(), start_time, units = "mins")
    cat("Finished", cell_name, "in", round(elapsed, 2), "minutes\n")
    
    # save convergence table
    conv_df <- data.frame(
      gene = names(converged_vec),
      convergence_code = as.numeric(converged_vec),
      comparison = pair_name,
      celltype = cell_name
    )
    
    conv_path <- file.path(
      dir.results, "T1D-T2D", "Results",
      paste0("nebula_", pair_name, "_", cell_name_file, "_convergence.csv")
    )
    write.csv(conv_df, conv_path, row.names = FALSE)
    
    # save errors if any
    if (length(error_log) > 0) {
      err_path <- file.path(
        dir.results, "T1D-T2D", "Results",
        paste0("nebula_", pair_name, "_", cell_name_file, "_errors.txt")
      )
      writeLines(error_log, err_path)
    }
    
    # clean memory before next cell type
    rm(seurat_sub, meta_full, pred_gene, library_offset, subject_id,
       genes_list, counts_hvg, conv_df, converged_vec, error_log)
    gc(verbose = FALSE)
  }
  
  rm(seurat_pair)
  gc(verbose = FALSE)
}
```
#-----------------------------------------------------
#psuedobulk analysis since OC only have 3 particpants
#-----------------------------------------------------
```{r}
#settings should already be min cell counts but just in case
min_cells_per_sample <- 20     # minimum cells contributing to a pseudobulk sample
min_donors_per_group <- 3      # minimum donors per group required to run DE
min_total_count <- 10          # edgeR gene filter helper
min_samples_expr <- 2          # edgeR gene filter helper

# output directory
pb_out_dir <- file.path(dir.results, "T1D-T2D", "Results", "pseudobulk")
dir.create(pb_out_dir, recursive = TRUE, showWarnings = FALSE)
```

#----------------------------
# prepping metadata for edgeR
#----------------------------
```{r}
meta <- pb90_multiomics_subset@meta.data %>%
  mutate(
    cell_barcode = rownames(pb90_multiomics_subset@meta.data),
    record_id = as.character(record_id),
    group = as.character(group),
    sex = as.character(sex),
    study = as.character(study),
    celltype = as.character(KPMP_celltype_general)
  )

#keeping only complete metadata for modeling
meta <- meta %>%
  filter(
    !is.na(record_id),
    !is.na(group),
    !is.na(age),
    !is.na(sex),
    !is.na(study),
    !is.na(celltype)
  )

# make sure cell order matches metadata row order
counts <- GetAssayData(pb90_multiomics_subset, layer = "counts")
counts <- counts[, meta$cell_barcode, drop = FALSE]
```
#----------------------------------------------------------
# building pseudobulk cell counts
# one pseudobulk sample = one donor within one cell type
#---------------------------------------------------------
```{r}
meta <- meta %>%
  mutate(sample_celltype_id = paste(record_id, celltype, sep = "__"))

pb_ids <- unique(meta$sample_celltype_id)

pb_count_list <- vector("list", length(pb_ids))
pb_meta_list  <- vector("list", length(pb_ids))

for (i in seq_along(pb_ids)) {
  sid <- pb_ids[i]
  idx <- which(meta$sample_celltype_id == sid)
  
  # skip tiny pseudobulks
  if (length(idx) < min_cells_per_sample) next
  
  sub_counts <- counts[, idx, drop = FALSE]
  agg_counts <- Matrix::rowSums(sub_counts)
  
  meta_sub <- meta[idx, ]
  
  pb_count_list[[i]] <- agg_counts
  pb_meta_list[[i]] <- data.frame(
    pseudobulk_id = sid,
    record_id = meta_sub$record_id[1],
    group = meta_sub$group[1],
    age = meta_sub$age[1],
    sex = meta_sub$sex[1],
    study = meta_sub$study[1],
    celltype = meta_sub$celltype[1],
    n_cells = nrow(meta_sub),
    stringsAsFactors = FALSE
  )
}

# remove NULL entries
keep_idx <- which(!sapply(pb_count_list, is.null))
pb_count_list <- pb_count_list[keep_idx]
pb_meta_list  <- pb_meta_list[keep_idx]

# combine into matrix + metadata
pb_counts <- do.call(cbind, pb_count_list)
colnames(pb_counts) <- sapply(pb_meta_list, function(x) x$pseudobulk_id)
pb_meta <- bind_rows(pb_meta_list)

# make sure order matches
pb_meta <- pb_meta %>% slice(match(colnames(pb_counts), pseudobulk_id))
stopifnot(identical(colnames(pb_counts), pb_meta$pseudobulk_id))
```

#------------------------
#running edgeR/cell type
#-----------------------
```{r}
#setting up pairwise comparisons
groups <- sort(unique(pb_meta$group))
group_pairs <- combn(groups, 2, simplify = FALSE)

celltypes <- sort(unique(pb_meta$celltype))

for (cell_name in celltypes) {
  
  cat("\n============================\n")
  cat("Cell type:", cell_name, "\n")
  cat("============================\n")
  
  cell_idx <- which(pb_meta$celltype == cell_name)
  
  if (length(cell_idx) == 0) next
  
  counts_ct <- pb_counts[, cell_idx, drop = FALSE]
  meta_ct   <- pb_meta[cell_idx, , drop = FALSE]
  
  for (pair in group_pairs) {
    
    pair_name <- paste(pair, collapse = "_vs_")
    cat("\n--- Comparison:", pair_name, "---\n")
    
    keep_pair <- meta_ct$group %in% pair
    counts_sub <- counts_ct[, keep_pair, drop = FALSE]
    meta_sub   <- meta_ct[keep_pair, , drop = FALSE]
    
    meta_sub$group <- factor(meta_sub$group, levels = pair)
    meta_sub$sex   <- factor(meta_sub$sex)
    meta_sub$study <- factor(meta_sub$study)
    
    # donor counts per group
    donor_tab <- table(meta_sub$group)
    print(donor_tab)
    
    if (any(donor_tab < min_donors_per_group)) {
      cat("Skipping", pair_name, "for", cell_name, "- too few donors in at least one group\n")
      next
    }
    
    #design matrix
    design <- model.matrix(~ group + age + sex + study, data = meta_sub)
    
    # edgeR object
    y <- DGEList(counts = counts_sub)
    
    # filter low-expression genes
    keep_genes <- filterByExpr(y, design = design)
    y <- y[keep_genes, , keep.lib.sizes = FALSE]
    
    if (nrow(y) == 0) {
      cat("Skipping", pair_name, "for", cell_name, "- no genes passed filtering\n")
      next
    }
    
    # normalization
    y <- calcNormFactors(y)
    
    # estimate dispersion
    y <- estimateDisp(y, design)
    
    # fit the model
    fit <- glmQLFit(y, design, robust = TRUE)
    
    coef_name <- grep("^group", colnames(design), value = TRUE)
    
    if (length(coef_name) != 1) {
      cat("Skipping", pair_name, "for", cell_name, "- could not identify unique group coefficient\n")
      next
    }
    
    qlf <- glmQLFTest(fit, coef = which(colnames(design) == coef_name))
    tt <- topTags(qlf, n = Inf)$table
    tt$gene <- rownames(tt)
    tt$comparison <- pair_name
    tt$celltype <- cell_name
    tt$n_donors_group1 <- as.numeric(donor_tab[1])
    tt$n_donors_group2 <- as.numeric(donor_tab[2])
    tt$coef_tested <- coef_name
    
    #add average CPM
    logcpm <- cpm(y, log = TRUE)
    tt$mean_logCPM_check <- rowMeans(logcpm[tt$gene, , drop = FALSE])
    
    # save
    cell_name_file <- gsub("[[:space:]/-]+", "_", cell_name)
    out_file <- file.path(
      pb_out_dir,
      paste0("pseudobulk_edgeR_", pair_name, "_", cell_name_file, ".csv")
    )
    
    write.csv(tt, out_file, row.names = FALSE)
    cat("Saved:", out_file, "\n")
  }
}
```






#----------------------
#archived code
----------------------
```{r}
groups      <- unique(pb90_multiomics_subset_meta$group)
group_pairs <- combn(groups, 2, simplify = FALSE)

celltypes_list <- unique(pb90_multiomics_subset_meta$KPMP_celltype_general)

for (pair in group_pairs) {
  pair_name <- paste(pair, collapse = "_vs_")
  cat("\n\n============================")
  cat("\nComparison:", pair_name)
  cat("\n============================\n")
  
  # Subset to this pair of groups
  seurat_pair <- subset(pb90_multiomics_subset,
                        group %in% pair)
  seurat_pair$group <- droplevels(factor(seurat_pair$group, levels = pair))  # pair[1] = reference
  
  for (cell_name in celltypes_list) {
    cat("\n=== Processing:", cell_name, "===\n")
    
    seurat_sub <- subset(seurat_pair, KPMP_celltype_general == cell_name)
    DefaultAssay(seurat_sub) <- "RNA"

    
    cat("Genes:", nrow(seurat_sub), "| Cells:", ncol(seurat_sub), "\n")
    
    seurat_sub <- FindVariableFeatures(seurat_sub,
                                       selection.method = "vst",
                                       nfeatures = 500)
    
    hvgs       <- VariableFeatures(seurat_sub)
    genes_list <- hvgs
    meta_full  <- seurat_sub@meta.data
    cat("HVGs:", length(hvgs), "\n")
    
    counts_hvg <- round(GetAssayData(seurat_sub, layer = "counts")[hvgs, , drop = FALSE])
    
    #setting up parallel processing
    cl <- makeCluster(4)
    registerDoParallel(cl)
    start_time <- Sys.time()
    
    nebula_results_list <- foreach(g = genes_list,
                                   .packages = c("nebula", "Matrix")) %dopar% {
      tryCatch({
        count_gene <- counts_hvg[g, , drop = FALSE]
        pred_gene  <- model.matrix(~ group + age + sex + study, data = meta_full)
        library    <- meta_full$pooled_offset
        
        data_g_gene <- group_cell(count  = count_gene,
                                  id     = meta_full$record_id,
                                  pred   = pred_gene,
                                  offset = library)
        
        if (is.null(data_g_gene)) {
          data_g_gene <- list(count  = count_gene,
                              id     = meta_full$record_id,
                              pred   = pred_gene,
                              offset = library)
        }
        
        result <- nebula(count      = data_g_gene$count,
                         id         = data_g_gene$id,
                         pred       = data_g_gene$pred,
                         ncore      = 1,
                         reml       = TRUE,
                         model      = "NBLMM",
                         output_re  = TRUE,
                         covariance = TRUE,
                         offset     = data_g_gene$offset)
        
        list(gene = g, result = result)
        
      }, error = function(e) {
        cat("Error on gene", g, ":", conditionMessage(e), "\n")
        NULL
      })
    }
    
    stopCluster(cl)
    cat("Time elapsed:", round(difftime(Sys.time(), start_time, units = "mins"), 2), "mins\n")
    
    out_path <- file.path(dir.results, "T1D-T2D", "Results",
                          paste0("nebula_", pair_name, "_",
                                 gsub(" ", "_", cell_name), ".rds"))
    saveRDS(nebula_results_list, file = out_path)
    cat("Saved:", out_path, "\n")
  }
}
```
