---
title: "scRNA_data_analysis"
author: "Savanah Leidholt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths("/mmfs1/gscratch/togo/leidholt/R_SLL_Seurat/")
```

```{r}
install_if_missing_cran <- function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(
      pkg,
      lib = ("/mmfs1/gscratch/togo/leidholt/R_SLL_Seurat/"),
      dependencies = TRUE
    )
  }
}

install_if_missing_bioc <- function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    BiocManager::install(
      pkg,
      lib = ("/mmfs1/gscratch/togo/leidholt/R_SLL_Seurat/"),
      update = FALSE,
      ask = FALSE
    )
  }
}
```

```{r}
library(Seurat)
library(tidyverse)
library(arsenal)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(Hmisc)
library(readxl)
library(purrr)
library(aws.s3)
library(jsonlite)
library(RColorBrewer)
library(ggtrace)
library(kableExtra)
library(data.table)
library(table1)
library(BiocManager)
library(GSA)
library(emmeans)
library(pheatmap)
library(enrichplot)
library(enrichR)
library(nebula)
library(doParallel)
library(Matrix)
library(edgeR)
library(stringr)
library(patchwork)
library(msigdbr)
library(fgsea)


#install_if_missing_bioc("speckle")
library(speckle)
#install_if_missing_bioc("DESeq2")
```
#directories
```{r}
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
  
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/TRIAD/Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
    
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/TRIAD/Results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
    
} else {
  stop("Unknown user: please specify root path for this user.")
}
```
#-------------------
#Create an S3 client
#-------------------
```{r}
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```

#--------------------------
#reading in cleaned subset
#--------------------------
```{r}
pb90_multiomics_subset <- s3readRDS(
          object = "data/pb90_multiomics_SLL_subset_20260527.rds", 
          bucket = "triad",
          region = "")


```

```{r}
my_colors <- colorRampPalette(brewer.pal(12, "Set3"))(43)
# Read Seurat object and run UMAP, obtain cell numbers, etc
pb90_multiomics_subset_meta <- pb90_multiomics_subset@meta.data %>%
  cbind(pb90_multiomics_subset@reductions$umap.TRIAD@cell.embeddings)

```
#------------
#making umaps
#------------
```{r}
centers <- pb90_multiomics_subset_meta %>%
  group_by(KPMP_celltype) %>%
  summarise(
    x = median(TRIAD_UMAP1),
    y = median(TRIAD_UMAP2),
    .groups = "drop"
  )

umap_p <- ggplot(pb90_multiomics_subset_meta,
                 aes(x = TRIAD_UMAP1, y = TRIAD_UMAP2, color = KPMP_celltype)) +
  geom_point(alpha = 0.2, shape = 16) +
  geom_text_repel(
    data = centers,
    aes(x = x, y = y, label = KPMP_celltype),
    inherit.aes = FALSE,
    size = 5,                
    fontface = "bold",
    color = "black"
  ) +
  scale_color_manual(values = my_colors) +
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),
    
    # axis titles + tick labels (match previous plots)
    axis.title.x = element_text(size = 16, hjust = 0),
    axis.title.y = element_text(size = 16, hjust = 0),
    axis.text = element_text(size = 14),
    
    # no legend
    legend.position = "none"
  ) +
  labs(x = "UMAP1", y = "UMAP2")

umap_p
```

```{r}
pb90_multiomics_subset_meta%>%
  distinct(record_id, group)

umap_p_study<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = TRIAD_UMAP1, y = TRIAD_UMAP2, color = study)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("RENAL-HEIR" = "darkcyan",
                                "RENAL-HEIRitage"= "gold4",
                                "IMPROVE" = "orange",
                                "CROCODILE" = "black",
                                "ATTEMPT" = "lightpink"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(size = 16, hjust = 0),   
        axis.title.y = element_text(size = 16, hjust = 0), 
        axis.text = element_text(size = 14),

 # legend
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)) + 
  labs(x = "UMAP1", y = "UMAP2",
       color = "study") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_study
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)
# UMAP by source
umap_p_group<- ggplot(pb90_multiomics_subset_meta,
                 aes(x = TRIAD_UMAP1, y = TRIAD_UMAP2, color = group)) +
  geom_point(alpha = 0.05, shape = 16) +
  scale_color_manual(values = c("Obese Control" = "darkcyan",
                                "Lean Control"= "plum4",
                                "Type 1 Diabetes" = "orange",
                                "Type 2 Diabetes" = "black"
                                )) +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        axis.title.x = element_text(size = 16, hjust = 0),   
        axis.title.y = element_text(size = 16, hjust = 0),
        axis.title = element_text(size = 14),

        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14))+
  labs(x = "UMAP1", y = "UMAP2",
       color = "group") +
  guides(color = guide_legend(override.aes = list(alpha = 1)))

umap_p_group
```
```{r}
aws.s3::s3HTTP(
  verb = "GET",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  path = "/"
)
```

```{r}
umap_cells_study <- umap_p + umap_p_study
tmp <- tempfile(fileext = ".png")

ggsave(
  tmp,
  plot = umap_cells_study,
  width = 20,
  height = 16,
  dpi = 300
)

httr::set_config(httr::config(ssl_verifypeer = FALSE))
put_object(
  file = tmp,
  object = "figures/metadata/Umap_cells_study.png",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)


umap_cells_group <- umap_p + umap_p_group
tmp <- tempfile(fileext = ".png")

ggsave(tmp,
       plot = umap_cells_group,
       width = 20,
       height =16,
       dpi = 300
       )

put_object(
  file = tmp,
  object = "figures/metadata/Umap_cells_group.png",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)
```
#----------------------
#Descriptive Statistics
#----------------------
```{r}
meta_table1 <- pb90_multiomics_subset_meta %>%
  distinct(record_id, .keep_all = TRUE) %>%
  mutate(
    group = factor(group)
  ) %>%
  filter(!is.na(group)) %>%
  droplevels()

table(meta_table1$group, useNA = "ifany")
```

```{r}
#Load Metadata cleaned
#Table 1
label(meta_table1$age) <- "Age (years)"
label(meta_table1$sex) <- "Sex"
label(meta_table1$race_ethnicity_condensed) <- "Ethnicity"
label(meta_table1$bmi) <- "BMI (kg/m^2)"
label(meta_table1$hba1c) <- "HbA1c (%)"
label(meta_table1$acr_u) <- "Urine Albumin-Creatinine Ratio (mg/g)"
label(meta_table1$eGFR_CKiD_U25_avg) <- "eGFR by Full Age Spectrum (mL/min/1.73m^2)"
label(meta_table1$group) <- "Diabetes Status"


pvalue_table1 <- function(x, ...) {
  
  y <- unlist(x)
  g <- factor(rep(names(x), times = sapply(x, length)))
  
  keep <- !is.na(y) & !is.na(g)
  y <- y[keep]
  g <- g[keep]
  
  if (is.numeric(y)) {
    if (length(unique(g)) < 2) return(c("", NA))
    p <- kruskal.test(y ~ g)$p.value
    test <- "Kruskal-Wallis"
  } else {
    tab <- table(y, g)
    if (nrow(tab) < 2 || ncol(tab) < 2) return(c("", NA))
    chi <- suppressWarnings(chisq.test(tab))
    if (any(chi$expected < 5)) {
      chi <- chisq.test(tab, simulate.p.value = TRUE, B = 10000)
      test <- "Chi-square simulated"
    } else {
      test <- "Chi-square"
    }
    p <- chi$p.value
  }
  c("", format.pval(p, digits = 3, eps = 0.001))
}

tble1 <- table1(
  ~ age + sex + race_ethnicity_condensed + hba1c + bmi + acr_u + eGFR_CKiD_U25_avg | group,
  data = meta_table1,
  extra.col = list(`Overall p-value` = pvalue_table1)
)

tmp <- tempfile(fileext = ".html")

htmltools::save_html(
  htmltools::HTML(as.character(tble1)),
  file = tmp
)

put_object(
  file = tmp,
  object = "tables/metadata_table.html",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)
```

```{r}
contrasts <- list(
  LC_vs_OC    = c("Lean Control", "Obese Control"),
  T1D_vs_T2D  = c("Type 1 Diabetes", "Type 2 Diabetes"),
  T1D_vs_LC   = c("Type 1 Diabetes", "Lean Control"),
  T2D_vs_LC   = c("Type 2 Diabetes", "Lean Control"),
  T1D_vs_OC   = c("Type 1 Diabetes", "Obese Control"),
  T2D_vs_OC   = c("Type 2 Diabetes", "Obese Control")
)

continuous_vars <- c("age", "bmi", "hba1c", "acr_u", "eGFR_CKiD_U25_avg") 
pairwise_results <- map_dfr(names(contrasts), function(comp) {
  
  groups_keep <- contrasts[[comp]]
  
  map_dfr(continuous_vars, function(var) {
    
    dat_test <- meta_table1 %>%
      filter(group %in% groups_keep) %>%
      select(group, value = all_of(var)) %>%
      filter(!is.na(group), !is.na(value)) %>%
      mutate(group = droplevels(factor(group, levels = groups_keep)))
    
    n_group1 <- sum(dat_test$group == groups_keep[1], na.rm = TRUE)
    n_group2 <- sum(dat_test$group == groups_keep[2], na.rm = TRUE)
    
    if (nlevels(dat_test$group) < 2 || n_group1 == 0 || n_group2 == 0) {
      return(tibble(
        contrast = comp,
        variable = var,
        group1 = groups_keep[1],
        group2 = groups_keep[2],
        n_group1 = n_group1,
        n_group2 = n_group2,
        p_value = NA_real_,
        note = "Skipped: fewer than 2 groups with non-missing data"
      ))
    }
    
    test <- wilcox.test(value ~ group, data = dat_test, exact = FALSE)
    
    tibble(
      contrast = comp,
      variable = var,
      group1 = groups_keep[1],
      group2 = groups_keep[2],
      n_group1 = n_group1,
      n_group2 = n_group2,
      p_value = test$p.value,
      note = NA_character_
    )
  })
})

pairwise_results <- pairwise_results %>%
  group_by(variable) %>%
  mutate(
    p_adj = p.adjust(p_value, method = "BH"),
    p_value_fmt = format.pval(p_value, digits = 3, eps = 0.001),
    p_adj_fmt = format.pval(p_adj, digits = 3, eps = 0.001)
  ) %>%
  ungroup()

pairwise_results
tmp <- tempfile(fileext = ".csv")

write.csv(
  pairwise_results,
  tmp,
  row.names = FALSE
)

httr::set_config(httr::config(ssl_verifypeer = FALSE))
put_object(
  file = tmp,
  object = "tables/pairwise_results.csv",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)
```

```{r}
cell_counts <- pb90_multiomics_subset_meta %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n))

max_n <- max(cell_counts$n, na.rm = TRUE)

label_pos <- -max_n * 0.05  
xmin_lim  <- -max_n * 0.35  
xmax_lim  <-  max_n * 1.15   

cell_count_bar <- ggplot(cell_counts, 
                         aes(x = reorder(KPMP_celltype_general, n), y = n)) +
  geom_col(fill = "#97a97c") +
  

  geom_text(aes(y = n, label = n),
            hjust = -0.1, size = 4) +
  
  geom_text(aes(y = label_pos, label = KPMP_celltype_general),
            hjust = 1, size = 4) +
  
  coord_flip(clip = "off") +
  
  scale_y_continuous(
    limits = c(xmin_lim, xmax_lim),
    expand = expansion(mult = c(0, 0))
  ) +
  
  labs(x = NULL, y = "Count") +
  
  theme(
    text = element_text(size = 15),
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.margin = margin(10, 30, 10, 30)
  ) +
  theme_transparent()

cell_count_bar

#ggsave("cell_counts.pdf",
 #      plot = cell_count_bar,
  #     path = file.path(dir.results, "T1D-T2D", "Results"),
   #    width = 10,
    #   height = 6)
```

```{r}
table((pb90_multiomics_subset_meta%>%
  distinct(record_id, visit, .keep_all = T))$group)

# cell counts per group
cell_counts_group <- pb90_multiomics_subset_meta %>%
  group_by(group) %>%
  dplyr::count(KPMP_celltype_general) %>%
  arrange(desc(n))

cell_group_bar <- ggplot(cell_counts_group, aes(x = reorder(KPMP_celltype_general, n), y = n, fill = group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Lean Control" = "plum4",
                               "Obese Control" = "darkcyan",
                               "Type 1 Diabetes" = "orange2",
                               "Type 2 Diabetes"= "black")) +
  coord_flip() +
  labs(x = NULL, 
       y = "Count",
       fill = "Group") +
  theme_transparent() +
  theme(
    text = element_text(size = 15),
    panel.grid = element_blank(),
    plot.margin = margin(1, 0.5, 1, 0.5),
    
    legend.position = c(0.7, 0.5),
    legend.justification = c(0, 0.5),
    legend.background = element_blank(),
    legend.margin = margin(0, 0, 0, 0)
  ) +
  
  scale_y_continuous(
    limits = c(xmin_lim, xmax_lim),
    expand = expansion(mult = c(0, 0))
  ) 

cell_group_bar
#ggsave("cell_gounts_group.pdf",
 #      plot = cell_group_bar,
  #     path = file.path(dir.results, "T1D-T2D", "Results"),
   #    width = 10,
    #   height = 6)
```

```{r}
cells <- cell_count_bar + cell_group_bar 
tmp <- tempfile(fileext = ".png")
ggsave(tmp,
       plot = cells,
       width = 16,
       height = 10,
       dpi = 600)

put_object(
  file = tmp,
  object = "figures/cell_counts.png",
  bucket = "triad",
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)

```


```{r}
# Barplot sample proportion by celltype 
ggplot(pb90_multiomics_subset@meta.data) +
    geom_bar(aes(x=study, fill=KPMP_celltype_general), 
                 position=position_fill(), color="black") +
    theme_classic() +
    labs(x="Study", y="Proportion of Cells")

ggplot(pb90_multiomics_subset@meta.data) +
    geom_bar(aes(x=group, fill=KPMP_celltype_general), 
                 position=position_fill(), color="black") +
    theme_classic() +
    labs(x="group", y="Proportion of Cells")
```


```{r}
pb90_multiomics_subset@meta.data$cell_type_legend <- paste0(
  pb90_multiomics_subset@meta.data$KPMP_celltype_general, 
  " = ", 
  pb90_multiomics_subset@meta.data$cell_type
)
nrow(pb90_multiomics_subset) # genes 31974
ncol(pb90_multiomics_subset) # cells 307386
```

```{r}
# Check your metadata
table(pb90_multiomics_subset$group)
table(pb90_multiomics_subset$kit_id)
table(pb90_multiomics_subset$source)
table(pb90_multiomics_subset$study)
table(pb90_multiomics_subset$KPMP_celltype_general)
head(pb90_multiomics_subset@meta.data[, c("group", "kit_id", "record_id", "KPMP_celltype_general")])
```
#------------------
#propeller analysis
#------------------
```{r}
groups <- unique(pb90_multiomics_subset_meta$group)
groups <- groups[!is.na(groups)]

group_pairs <- combn(groups, 2, simplify = FALSE)

propeller_results <- list()

for (pair in group_pairs) {
  
  pair_name <- paste(pair, collapse = "_vs_")
  pair_name_clean <- gsub("[^A-Za-z0-9_]+", "_", pair_name)
  
  cat("\n=== Propeller:", pair_name, "===\n")
  
  meta_sub <- pb90_multiomics_subset_meta[
    pb90_multiomics_subset_meta$group %in% pair,
  ]
  
  meta_sub$group <- droplevels(factor(meta_sub$group))
  
  res <- propeller(
    clusters  = meta_sub$KPMP_celltype_general,
    sample    = meta_sub$record_id,
    group     = meta_sub$group,
    transform = "logit"
  )
  
  res$comparison <- pair_name
  propeller_results[[pair_name_clean]] <- res
  
  tmp <- tempfile(fileext = ".csv")
  
  write.csv(
    res,
    tmp,
    row.names = TRUE
  )
  
  put_object(
    file = tmp,
    object = paste0("results/propeller/propeller_", pair_name_clean,".csv"),
    bucket = "triad",
    base_url = "s3.kopah.uw.edu",
    region = "",
    use_https = TRUE
  )
  
  unlink(tmp)
}

  #write.csv(
  #  x    = res,
  #  file = dir.results, paste0("propeller_", pair_name, ".csv"),
  #  row.names = TRUE
  #)

```
#-----------------------------------------------------
#psuedobulk analysis with edgeR & Deseq2
#-----------------------------------------------------
```{r}
#setting up save function to kopah
bucket <- "triad"

upload_to_kopah <- function(file, object, bucket = "triad") {
  put_object(
    file = file,
    object = object,
    bucket = bucket,
    base_url = "s3.kopah.uw.edu",
    region = "",
    use_https = TRUE
  )
}

min_cells_per_sample <- 20
min_donors_per_group <- 3

```

#-------------------------------------------------------
# metadata and building pseudobulk counts
# one pseudobulk sample = one donor within one cell type
#-------------------------------------------------------
```{r}
meta <- pb90_multiomics_subset@meta.data %>%
  mutate(
    cell_barcode = rownames(pb90_multiomics_subset@meta.data),
    record_id = as.character(record_id),
    group = as.character(group),
    sex = as.character(sex),
    study = as.character(study),
    celltype = as.character(KPMP_celltype_general)
  ) %>%
  filter(
    !is.na(record_id),
    !is.na(group),
    !is.na(age),
    !is.na(sex),
    !is.na(study),
    !is.na(celltype)
  ) %>%
  mutate(
    sample_celltype_id = paste(record_id, celltype, sep = "__")
  )

counts <- GetAssayData(pb90_multiomics_subset, assay = "RNA", layer = "counts")
counts <- counts[, meta$cell_barcode, drop = FALSE]

pb_ids <- unique(meta$sample_celltype_id)

pb_count_list <- list()
pb_meta_list <- list()

for (sid in pb_ids) {
  
  idx <- which(meta$sample_celltype_id == sid)
  
  if (length(idx) < min_cells_per_sample) next
  
  meta_sub <- meta[idx, ]
  
  pb_count_list[[sid]] <- Matrix::rowSums(counts[, idx, drop = FALSE])
  
  pb_meta_list[[sid]] <- data.frame(
    pseudobulk_id = sid,
    record_id = meta_sub$record_id[1],
    group = meta_sub$group[1],
    age = meta_sub$age[1],
    sex = meta_sub$sex[1],
    study = meta_sub$study[1],
    celltype = meta_sub$celltype[1],
    n_cells = nrow(meta_sub),
    stringsAsFactors = FALSE
  )
}

pb_counts <- do.call(cbind, pb_count_list)
pb_meta <- bind_rows(pb_meta_list)

pb_meta <- pb_meta %>%
  dplyr::slice(match(colnames(pb_counts), pseudobulk_id))

stopifnot(identical(colnames(pb_counts), pb_meta$pseudobulk_id))
```

#-----------------------
#volcano plot function
#-----------------------
```{r}
plot_volcano <- function(tt, title = "", fdr_col = "FDR",
                         lfc_col = "logFC",
                         stat_col = NULL,
                         fdr_cutoff = 0.05,
                         lfc_cutoff = 0.5,
                         n_labels = 10) {
  
  tt <- tt %>%
    dplyr::mutate(
      neg_log10_fdr = -log10(.data[[fdr_col]]),
      sig = dplyr::case_when(
        .data[[fdr_col]] < fdr_cutoff & .data[[lfc_col]] > lfc_cutoff  ~ "Up",
        .data[[fdr_col]] < fdr_cutoff & .data[[lfc_col]] < -lfc_cutoff ~ "Down",
        TRUE ~ "NS"
      )
    )
  
  if (!is.null(stat_col) && stat_col %in% colnames(tt)) {
    
    label_df <- tt %>%
      dplyr::filter(.data[[fdr_col]] < fdr_cutoff) %>%
      dplyr::arrange(dplyr::desc(abs(.data[[stat_col]]))) %>%
      dplyr::slice_head(n = n_labels)
    
  } else {
    
    label_df <- tt %>%
      dplyr::filter(.data[[fdr_col]] < fdr_cutoff) %>%
      dplyr::arrange(.data[[fdr_col]]) %>%
      dplyr::slice_head(n = n_labels)
  }
  
  ggplot(tt, aes(x = .data[[lfc_col]], y = neg_log10_fdr, color = sig)) +
    geom_point(alpha = 0.6, size = 1.5) +
    geom_text_repel(
      data = label_df,
      aes(label = gene),
      size = 3,
      max.overlaps = Inf
    ) +
    scale_color_manual(values = c("Up" = "orange3", "Down" = "purple4", "NS" = "grey")) +
    geom_vline(xintercept = c(-lfc_cutoff, lfc_cutoff), linetype = "dashed") +
    geom_hline(yintercept = -log10(fdr_cutoff), linetype = "dashed") +
    labs(title = title, x = "log2 Fold Change", y = "-log10(FDR)") +
    theme_classic()
}
```

#------------------------
#running edgeR pseudobulk
#-----------------------
```{r}
groups <- sort(unique(pb_meta$group))
group_pairs <- combn(groups, 2, simplify = FALSE)

celltypes <- sort(unique(pb_meta$celltype))

edgeR_results <- list()

for (cell_name in celltypes) {
  
  cell_idx <- which(pb_meta$celltype == cell_name)
  counts_ct <- pb_counts[, cell_idx, drop = FALSE]
  meta_ct <- pb_meta[cell_idx, , drop = FALSE]
  
  for (pair in group_pairs) {
    
    pair_name <- paste(pair, collapse = "_vs_")
    pair_name_clean <- gsub("[^A-Za-z0-9_]+", "_", pair_name)
    cell_name_clean <- gsub("[^A-Za-z0-9_]+", "_", cell_name)
    
    cat("\n=== edgeR:", cell_name, pair_name, "===\n")
    
    keep_pair <- meta_ct$group %in% pair
    counts_sub <- counts_ct[, keep_pair, drop = FALSE]
    meta_sub <- meta_ct[keep_pair, , drop = FALSE]
    
    meta_sub$group <- factor(meta_sub$group, levels = pair)
    meta_sub$sex <- factor(meta_sub$sex)
    
    donor_tab <- table(meta_sub$group)
    
    if (any(donor_tab < min_donors_per_group)) {
      cat("Skipping: too few donors\n")
      next
    }
    
    design <- model.matrix(~ group + age + sex, data = meta_sub)
    
    y <- DGEList(counts = counts_sub)
    keep_genes <- filterByExpr(y, design = design)
    y <- y[keep_genes, , keep.lib.sizes = FALSE]
    
    if (nrow(y) == 0) next
    
    y <- calcNormFactors(y)
    y <- estimateDisp(y, design)
    
    fit <- glmQLFit(y, design, robust = TRUE)
    
    coef_name <- grep("^group", colnames(design), value = TRUE)
    
    qlf <- glmQLFTest(fit, coef = coef_name)
    tt <- topTags(qlf, n = Inf)$table
    
    tt$gene <- rownames(tt)
    tt$comparison <- pair_name
    tt$celltype <- cell_name
    tt$n_donors_group1 <- as.numeric(donor_tab[1])
    tt$n_donors_group2 <- as.numeric(donor_tab[2])
    
    edgeR_results[[paste(cell_name_clean, pair_name_clean, sep = "__")]] <- tt
    
    tmp_csv <- tempfile(fileext = ".csv")
    write.csv(tt, tmp_csv, row.names = FALSE)
    
    upload_to_kopah(
      file = tmp_csv,
      object = paste0(
        "results/pseudobulk/edgeR/all_cells/pseudo_all_edgeR_",
        pair_name_clean, "_", cell_name_clean, ".csv"
      )
    )
    
    volcano_plot <- plot_volcano(
      tt,
      title = paste("edgeR", cell_name, pair_name),
      fdr_col = "FDR",
      lfc_col = "logFC"
    )
    
    tmp_png <- tempfile(fileext = ".png")
    
    ggsave(
      tmp_png,
      plot = volcano_plot,
      width = 6,
      height = 5,
      dpi = 300
    )
    
    upload_to_kopah(
      file = tmp_png,
      object = paste0(
        "figures/pseudobulk/edgeR/all_cells/volcano_all_edgeR_",
        pair_name_clean, "_", cell_name_clean, ".png"
      )
    )
    
    unlink(c(tmp_csv, tmp_png))
  }
}

```
#-------------------
#DESeq2 pseudobulk
#-------------------
```{r}
deseq2_results <- list()

for (cell_name in celltypes) {
  
  cell_idx <- which(pb_meta$celltype == cell_name)
  counts_ct <- pb_counts[, cell_idx, drop = FALSE]
  meta_ct <- pb_meta[cell_idx, , drop = FALSE]
  
  for (pair in group_pairs) {
    
    pair_name <- paste(pair, collapse = "_vs_")
    pair_name_clean <- gsub("[^A-Za-z0-9_]+", "_", pair_name)
    cell_name_clean <- gsub("[^A-Za-z0-9_]+", "_", cell_name)
    
    cat("\n=== DESeq2:", cell_name, pair_name, "===\n")
    
    keep_pair <- meta_ct$group %in% pair
    counts_sub <- round(counts_ct[, keep_pair, drop = FALSE])
    meta_sub <- meta_ct[keep_pair, , drop = FALSE]
    
    meta_sub$group <- factor(meta_sub$group, levels = pair)
    meta_sub$sex <- factor(meta_sub$sex)
    rownames(meta_sub) <- colnames(counts_sub)
    
    donor_tab <- table(meta_sub$group)
    
    if (any(donor_tab < min_donors_per_group)) {
      cat("Skipping: too few donors\n")
      next
    }
    
    # Use the same expression filtering strategy as edgeR
    filter_design <- model.matrix(~ group + age + sex, data = meta_sub)
    
    y_tmp <- DGEList(counts = counts_sub)
    
    keep_genes <- filterByExpr(
      y_tmp,
      design = filter_design
    )
    
    counts_sub <- counts_sub[keep_genes, , drop = FALSE]
    
    if (nrow(counts_sub) == 0) next
    
    dds <- DESeqDataSetFromMatrix(
      countData = counts_sub,
      colData = meta_sub,
      design = ~ age + sex + group
    )
    
    dds <- DESeq(dds)
    
    res <- results(
      dds,
      contrast = c("group", pair[2], pair[1])
    )
    
    tt <- as.data.frame(res)
    
    cat("\nColumns returned:\n")
    print(colnames(tt))
    
    tt$gene <- rownames(tt)
    tt$comparison <- pair_name
    tt$celltype <- cell_name
    tt$n_donors_group1 <- as.numeric(donor_tab[1])
    tt$n_donors_group2 <- as.numeric(donor_tab[2])
    
    tt <- tt %>%
      dplyr::rename(
        logFC = log2FoldChange,
        FDR = padj,
        PValue = pvalue
      )
    
    deseq2_results[[paste(cell_name_clean, pair_name_clean, sep = "__")]] <- tt
    
    tmp_csv <- tempfile(fileext = ".csv")
    write.csv(tt, tmp_csv, row.names = FALSE)
    
    upload_to_kopah(
      file = tmp_csv,
      object = paste0(
        "results/pseudobulk/DESeq2/all_cells/pseudo_all_DESeq2_",
        pair_name_clean, "_", cell_name_clean, ".csv"
      )
    )
    
    volcano_plot <- plot_volcano(
      tt,
      title = paste("DESeq2", cell_name, pair_name),
      fdr_col = "FDR",
      lfc_col = "logFC"
    )
    
    tmp_png <- tempfile(fileext = ".png")
    
    ggsave(
      tmp_png,
      plot = volcano_plot,
      width = 6,
      height = 5,
      dpi = 300
    )
    
    upload_to_kopah(
      file = tmp_png,
      object = paste0(
        "figures/pseudobulk/DESeq2/all_cells/volcano_all_DESeq2_",
        pair_name_clean, "_", cell_name_clean, ".png"
      )
    )
    
    unlink(c(tmp_csv, tmp_png))
  }
}
```


```{r}
method_compare_summary <- purrr::map_dfr(
  intersect(names(edgeR_results), names(deseq2_results)),
  function(nm) {
    
    edgeR_tt <- edgeR_results[[nm]]
    deseq2_tt <- deseq2_results[[nm]]
    
    merged_de <- dplyr::inner_join(
      edgeR_tt %>%
        dplyr::select(gene, edgeR_logFC = logFC, edgeR_FDR = FDR),
      deseq2_tt %>%
        dplyr::select(gene, DESeq2_logFC = logFC, DESeq2_FDR = FDR),
      by = "gene"
    )
    
    tibble::tibble(
      result = nm,
      n_genes = nrow(merged_de),
      edgeR_sig = sum(merged_de$edgeR_FDR < 0.05, na.rm = TRUE),
      DESeq2_sig = sum(merged_de$DESeq2_FDR < 0.05, na.rm = TRUE),
      both_sig = sum(
        merged_de$edgeR_FDR < 0.05 &
          merged_de$DESeq2_FDR < 0.05,
        na.rm = TRUE
      ),
      logFC_cor = cor(
        merged_de$edgeR_logFC,
        merged_de$DESeq2_logFC,
        use = "complete.obs"
      )
    )
  }
)

method_compare_summary
method_compare_summary <- method_compare_summary %>%
  dplyr::rename(
    `Correlation of LFC effect size` = logFC_cor
  )

method_compare_summary
tmp <- tempfile(fileext = ".csv")
write.csv(method_compare_summary, tmp, row.names = FALSE)

upload_to_kopah(
  file = tmp,
  object = "results/pseudobulk/method_compare_edgeR_vs_DESeq2.csv"
)
```
#------------------------------------------------------------
#START HERE IF YOU'VE ALREADY DONE DE ANLAYSIS
#plotting cell type specific expression patterns per contrast
#------------------------------------------------------------
```{r}
#setting up save function to kopah
bucket <- "triad"

upload_to_kopah <- function(file, object, bucket = "triad") {
  put_object(
    file = file,
    object = object,
    bucket = bucket,
    base_url = "s3.kopah.uw.edu",
    region = "",
    use_https = TRUE
  )
}


kopah_opts <- list(
  base_url = "s3.kopah.uw.edu",
  region = "",
  use_https = TRUE
)

read_kopah_csvs <- function(prefix, method_name) {
  
  objs <- get_bucket(
    bucket = "triad",
    prefix = prefix,
    base_url = kopah_opts$base_url,
    region = kopah_opts$region,
    use_https = kopah_opts$use_https
  )
  
  csv_keys <- purrr::map_chr(objs, "Key") %>%
    stringr::str_subset("\\.csv$")
  
  if (length(csv_keys) == 0) {
    stop("No CSV files found for prefix: ", prefix)
  }
  
  purrr::map_dfr(csv_keys, function(key) {
    
    message("Reading: ", key)
    
    dat <- aws.s3::s3read_using(
      FUN = function(x) readr::read_csv(x, show_col_types = FALSE),
      object = key,
      bucket = "triad",
      opts = kopah_opts
    )
    
    dat %>%
      dplyr::mutate(
        method = method_name,
        source_file = key
      )
  })
}
```

```{r}
edgeR_all <- read_kopah_csvs(
  prefix = "results/pseudobulk/edgeR/all_cells/",
  method_name = "edgeR"
)

deseq2_all <- read_kopah_csvs(
  prefix = "results/pseudobulk/DESeq2/all_cells/",
  method_name = "DESeq2"
)

de_all <- bind_rows(edgeR_all, deseq2_all)

tmp_de_all <- tempfile(fileext = ".csv")
write.csv(de_all, tmp_de_all, row.names = F)
upload_to_kopah(
  file = tmp_de_all,
  object = "results/pseudobulk/DE_all.csv"
)

unlink(tmp_de_all)
```
#---------------------------------
#count significant genes/cell type
#--------------------------------
```{r}
de_counts <- de_all %>%
  mutate(
    direction = case_when(
      FDR < 0.05 & logFC > 0 ~ "Positive",
      FDR < 0.05 & logFC < 0 ~ "Negative",
      TRUE ~ "NS"
    )
  ) %>%
  filter(direction != "NS") %>%
  group_by(method, comparison, celltype, direction) %>%
  summarise(n_de = n(), .groups = "drop") %>%
  mutate(
    n_plot = ifelse(direction == "Negative", -n_de, n_de)
  )

#save back to kopha
#tmp_counts <- tempfile(fileext = ".csv")

#write.csv(de_counts, tmp_counts, row.names = FALSE)

#upload_to_kopah(
#  file = tmp_counts,
#  object = "results/pseudobulk/celltype_DE_counts_edgeR_DESeq2.csv"
#)

#unlink(tmp_counts)
```

```{r}
plot_celltype_panel <- function(dat, method_name) {
  
  dat %>%
    dplyr::filter(method == method_name) %>%
    ggplot(
      aes(
        x = n_plot,
        y = reorder(celltype, n_plot),
        fill = direction
      )
    ) +
    geom_col(width = 0.75) +
    geom_vline(xintercept = 0, color = "gray40") +
    facet_wrap(~ comparison, scales = "free_x") +
    scale_x_continuous(labels = abs) +
    scale_fill_manual(
      values = c(
        "Negative" = "purple4",
        "Positive" = "orange3"
      )
    ) +
    labs(
      title = paste0(method_name, ": cell type-specific DE genes by contrast"),
      x = "Number of DE genes (FDR < 0.05)",
      y = NULL,
      fill = NULL
    ) +
    theme_classic() +
    theme(
      strip.text = element_text(face = "bold", size = 10),
      axis.text.y = element_text(size = 9),
      axis.text.x = element_text(size = 9),
      legend.position = "top"
    )
}
```

```{r fig.width=14, fig.height=8}
# edgeR panel plot
p_edgeR_panel <- plot_celltype_panel(
  dat = de_counts,
  method_name = "edgeR"
)

p_edgeR_panel
```
```{r fig.width=14, fig.height=8}
# DESeq2 panel plot
p_DESeq2_panel <- plot_celltype_panel(
  dat = de_counts,
  method_name = "DESeq2"
)

p_DESeq2_panel
```

```{r}
#tmp_edgeR <- tempfile(fileext = ".png")
#tmp_DESeq2 <- tempfile(fileext = ".png")

#ggsave(
#  filename = tmp_edgeR,
#  plot = p_edgeR_panel,
#  width = 14,
#  height = 8,
#  dpi = 300
#)

#ggsave(
#  filename = tmp_DESeq2,
#  plot = p_DESeq2_panel,
#  width = 14,
#  height = 8,
#  dpi = 300
#)

#upload_to_kopah(
#  file = tmp_edgeR,
#  object = "figures/pseudobulk/edgeR/all_cells/celltype_expression_patterns.png"
#)

#upload_to_kopah(
#  file = tmp_DESeq2,
#  object = "figures/pseudobulk/DESeq2/all_cells/celltype_expression_patterns.png"
#)

#unlink(c(tmp_edgeR, tmp_DESeq2))
```

#-------------
#GSEA anlaysis
#-------------
```{r}
hallmark <- msigdbr(
    species = "Homo sapiens",
    collection = "H")

hallmark <- hallmark %>%
  mutate(gs_name = str_remove(gs_name, "^HALLMARK_"),
  gs_name = str_replace_all(gs_name, "_", " "),
  gs_name = str_to_title(gs_name)
)


reactome<-  msigdbr(
    species = "Homo sapiens",
    collection = "C2",
    subcollection = "CP:REACTOME"
  )

reactome <- reactome %>%
  mutate(gs_name = str_remove(gs_name, "^REACTOME_"),
  gs_name = str_replace_all(gs_name, "_", " "),
  gs_name = str_to_title(gs_name)
)

go <- bind_rows(
   msigdbr(
    species = "Homo sapiens",
    collection = "C5",
    subcollection = "GO:BP"
    ),
   msigdbr(
    species = "Homo sapiens",
    collection = "C5",
    subcollection = "GO:BP"
    ),
   msigdbr(
    species = "Homo sapiens",
    collection = "C5",
    subcollection = "GO:BP")
)

go <- go %>%
  mutate(gs_name = str_remove(gs_name, "^GO"),
  gs_name = str_to_title(gs_name)
)

hallmark_list <- split(
  hallmark$gene_symbol,
  hallmark$gs_name)

reactome_list <- split(
  reactome$gene_symbol,
  reactome$gs_name)

go_list <- split(
  go$gene_symbol,
  go$gs_name)
```

```{r}
make_rank_vector <- function(dat){

  ranks <- dat %>%
    filter(
      !is.na(gene),
      !is.na(F),
      !is.na(logFC)
    ) %>%
    mutate(
      rank_metric = sign(logFC) * sqrt(F)
    ) %>%
    group_by(gene) %>%
    slice_max(
      abs(rank_metric),
      n = 1,
      with_ties = FALSE
    ) %>%
    ungroup()

  stats <- ranks$rank_metric
  names(stats) <- ranks$gene

  sort(stats, decreasing = TRUE)
}
```
#--------------
#reactome plots
#--------------
```{r}
gsea_results_reactome <- edgeR_all %>%
  group_by(celltype, comparison) %>%
  nest() %>%
  dplyr::mutate(
    stats = map(data, make_rank_vector),

    fgsea = map(
      stats,
      ~ fgsea(
        pathways = reactome_list,
        stats = .x,
        minSize = 15,
        maxSize = 500
      )
    )
  ) %>%
  dplyr::select(-data, -stats) %>%
  unnest(fgsea)

gsea_top_reactome <- gsea_results_reactome %>%
  filter(!is.na(padj)) %>%
  filter(padj < 0.05) %>%
  group_by(celltype, comparison) %>%
  slice_max(
    order_by = abs(NES),
    n = 10,
    with_ties = FALSE
  ) %>%
  ungroup()
```

```{r}
plot_gsea_lollipop <- function(dat, celltype_name, contrast_name = NULL) {
  
  plot_dat <- dat %>%
    filter(celltype == celltype_name) %>%
    { 
      if (!is.null(contrast_name)) {
        filter(., comparison == contrast_name)
      } else {
        .
      }
    } %>%
    mutate(
      Direction = ifelse(NES > 0, "Positive", "Negative"),
      pathway = stringr::str_wrap(pathway, 55),
      pathway = forcats::fct_reorder(pathway, NES)
    )
  
  p <- ggplot(plot_dat, aes(x = NES, y = pathway)) +
    geom_segment(
      aes(x = 0, xend = NES, yend = pathway),
      linewidth = 0.7,
      color = "grey70"
    ) +
    geom_point(aes(color = Direction), size = 8) +
    geom_text(
      data = ~ dplyr::filter(.x, NES > 0),
      aes(x = 0, label = pathway),
      hjust = 1.02,
      size = 6,
      fontface = "bold",
      color = "black"
    ) +
    geom_text(
      data = ~ dplyr::filter(.x, NES < 0),
      aes(x = 0, label = pathway),
      hjust = -0.02,
      size = 6,
      fontface = "bold",
      color = "black"
    ) +
    geom_vline(xintercept = 0, linewidth = 0.8, color = "black") +
    scale_color_manual(
      values = c(
        "Positive" = "orange3",
        "Negative" = "purple4"
      )
    ) +
    scale_x_continuous(
  limits = c(-4, 4),
  breaks = c(-4, -2, 0, 2, 4)
  )+
    labs(
      title = paste(
        celltype_name,
        ifelse(is.null(contrast_name), "all contrasts", contrast_name)
      ),
      x = "Normalized enrichment score",
      y = NULL,
      color = NULL
    ) +
    theme_classic() +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      legend.position = "top"
    )
  
  if (is.null(contrast_name)) {
    p <- p +
      facet_wrap(~ comparison, scales = "free_y")
  }
  
  p
}
```

```{r fig.width=14}
PT_panel <- plot_gsea_lollipop(
  dat = gsea_top_reactome,
  celltype_name = "PT"
)

#tmp <- tempfile(fileext = ".png")

#ggsave(
#  filename = tmp,
#  plot = PT_panel,
#  width = 24,
#  height = 12,
#  dpi = 300
#)

#upload_to_kopah(
#  file = tmp,
#  object = "figures/GSEA/Reactome/PT_panel_reactome.png"
#)

#unlink(tmp)

pt_individual_plots <- gsea_top_reactome %>%
  filter(celltype == "PT") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_reactome,
      celltype_name = "PT",
      contrast_name = .x
    )
  )

pt_individual_plots[[1]] #lean vs obese
pt_individual_plots[[2]] #lean vs T1d
pt_individual_plots[[3]] #lean vs T2d
pt_individual_plots[[4]] #OC vs t1d
pt_individual_plots[[5]] #oc vs t2d
pt_individual_plots[[6]] #t1d vs t2d
```

```{r}
PC_panel <- plot_gsea_lollipop(
  dat = gsea_top_reactome,
  celltype_name = "PC"
)

#tmp <- tempfile(fileext = ".png")

#ggsave(
#  filename = tmp,
#  plot = PC_panel,
#  width = 24,
#  height = 12,
#  dpi = 300
#)

#upload_to_kopah(
#  file = tmp,
#  object = "figures/GSEA/Reactome/PC_panel_reactome.png"
#)

#unlink(tmp)

pc_individual_plots <- gsea_top_reactome %>%
  filter(celltype == "PC") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_reactome,
      celltype_name = "PC",
      contrast_name = .x
    )
  )

pc_individual_plots[[1]] #lean vs obese
#pc_individual_plots[[2]] #lean vs T1d
pc_individual_plots[[2]] #lean vs T2d
pc_individual_plots[[3]] #OC vs t1d
pc_individual_plots[[4]] #oc vs t2d
pc_individual_plots[[5]] #t1d vs t2d
```

```{r fig.width=14}
EC_panel <- plot_gsea_lollipop(
  dat = gsea_top_reactome,
  celltype_name = "EC"
)

#tmp <- tempfile(fileext = ".png")

#ggsave(
#  filename = tmp,
#  plot = EC_panel,
#  width = 24,
#  height = 12,
#  dpi = 300
#)

#upload_to_kopah(
#  file = tmp,
#  object = "figures/GSEA/Reactome/EC_panel_reactome.png"
#)

#unlink(tmp)

ec_individual_plots <- gsea_top_reactome %>%
  filter(celltype == "EC") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_reactome,
      celltype_name = "EC",
      contrast_name = .x
    )
  )

ec_individual_plots[[1]] #lc vs oc
ec_individual_plots[[2]] #lc vs t2d
ec_individual_plots[[3]] #oc vs t1d
ec_individual_plots[[4]] #oc vs t2d
ec_individual_plots[[5]] #t1d vs t2d
#ec_individual_plots[[6]] #no t1d vs lc
```

```{r fig.width=14}
TAL_panel <- plot_gsea_lollipop(
  dat = gsea_top_reactome,
  celltype_name = "TAL"
)

#tmp <- tempfile(fileext = ".png")

#ggsave(
#  filename = tmp,
#  plot = TAL_panel,
#  width = 24,
#  height = 12,
#  dpi = 300
#)

#upload_to_kopah(
#  file = tmp,
#  object = "figures/GSEA/Reactome/TAL_panel_reactome.png"
#)

#unlink(tmp)

tal_individual_plots <- gsea_top_reactome %>%
  filter(celltype == "TAL") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_reactome,
      celltype_name = "TAL",
      contrast_name = .x
    )
  )

tal_individual_plots[[1]] #lc vs oc
tal_individual_plots[[2]] #lc vs t1d
tal_individual_plots[[3]] #lc vs t2d
tal_individual_plots[[4]] #oc vs t1d
tal_individual_plots[[5]] #oc vs t2d
tal_individual_plots[[6]] #t1d vs t2d
```

```{r}
Immune_panel <- plot_gsea_lollipop(
  dat = gsea_top_reactome,
  celltype_name = "Immune"
)

#tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = Immune_panel,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Reactome/Immune_panel_reactome.png"
)

unlink(tmp)

immune_individual_plots <- gsea_top_reactome %>%
  filter(celltype == "Immune") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_reactome,
      celltype_name = "Immune",
      contrast_name = .x
    )
  )

immune_individual_plots[[1]] #lc vs oc
immune_individual_plots[[2]] #lc vs t2d
immune_individual_plots[[3]] #oc vs t1d
immune_individual_plots[[4]] #oc vs t2d
immune_individual_plots[[5]] #t1d vs t2d
immune_individual_plots[[6]] #t1d vs lc
```

```{r}
POD_panel <-plot_gsea_lollipop(
  dat = gsea_top_reactome,
  celltype_name = "POD"
)

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = POD_panel,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Reactome/POD_panel_reactome.png"
)

unlink(tmp)

pod_individual_plots <- gsea_top_reactome %>%
  filter(celltype == "POD") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_reactome,
      celltype_name = "POD",
      contrast_name = .x
    )
  )

#pod_individual_plots[[1]] #lc vs oc not sig
#pod_individual_plots[[2]] #lc vs t1d not sig
#pod_individual_plots[[3]] #lc vs t2d not sig
#pod_individual_plots[[4]] #oc vs t1d not sig
#pod_individual_plots[[5]] #oc vs t2d not sig
pod_individual_plots[[1]] #t1d vs t2d
```
```{r}
DCT_CNT_panel <- plot_gsea_lollipop(
  dat = gsea_top_reactome,
  celltype_name = "DCT_CNT"
)

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = DCT_CNT_panel,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Reactome/DCT_CNT_panel_reactome.png"
)

unlink(tmp)

dctCNT_individual_plots <- gsea_top_reactome %>%
  filter(celltype == "DCT_CNT") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_reactome,
      celltype_name = "DCT_CNT",
      contrast_name = .x
    )
  )

dctCNT_individual_plots[[1]] #lc vs oc
dctCNT_individual_plots[[2]] #lc vs t1d
dctCNT_individual_plots[[3]] #lc vs t2d
dctCNT_individual_plots[[4]] #oc vs t1d
dctCNT_individual_plots[[5]] #oc vs t2d
dctCNT_individual_plots[[6]] #t1d vs t2d
```

#-------------------------------
#plotting top GO pathways
#-------------------------------
```{r}
gsea_results_GO <- edgeR_all %>%
  group_by(celltype, comparison) %>%
  nest() %>%
  dplyr::mutate(
    stats = map(data, make_rank_vector),

    fgsea = map(
      stats,
      ~ fgsea(
        pathways = go_list,
        stats = .x,
        minSize = 15,
        maxSize = 500
      )
    )
  ) %>%
  dplyr::select(-data, -stats) %>%
  unnest(fgsea)


gsea_top_GO<- gsea_results_GO%>%
  filter(!is.na(padj)) %>%
  filter(padj < 0.05) %>%
  group_by(celltype, comparison) %>%
  slice_max(
    order_by = abs(NES),
    n = 10,
    with_ties = FALSE
  ) %>%
  ungroup()
```

```{r}
PT_panel_GO <- plot_gsea_lollipop(
  dat = gsea_top_GO,
  celltype_name = "PT"
)

PT_panel_GO
tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = PT_panel_GO,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/GO/PT_panel_GO.png"
)

unlink(tmp)


pt_individual_plots <- gsea_top_GO %>%
  filter(celltype == "PT") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_GO,
      celltype_name = "PT",
      contrast_name = .x
    )
  )

pt_individual_plots[[1]] #lean vs obese
pt_individual_plots[[2]] #lean vs T1d
pt_individual_plots[[3]] #lean vs T2d
pt_individual_plots[[4]] #OC vs t1d
pt_individual_plots[[5]] #oc vs t2d
pt_individual_plots[[6]] #t1d vs t2d
```

```{r}
PC_panel_GO <- plot_gsea_lollipop(
  dat = gsea_top_GO,
  celltype_name = "PC"
)

PC_panel_GO

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = PC_panel_GO,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/GO/PC_panel_GO.png"
)

unlink(tmp)


pc_individual_plots <- gsea_top_GO %>%
  filter(celltype == "PC") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_GO,
      celltype_name = "PC",
      contrast_name = .x
    )
  )

pc_individual_plots[[1]] #lean vs obese
pc_individual_plots[[2]] #lean vs T1d
pc_individual_plots[[3]] #lean vs T2d
pc_individual_plots[[4]] #OC vs t1d
pc_individual_plots[[5]] #oc vs t2d
pc_individual_plots[[6]] #t1d vs t2d
```

```{r}
EC_panel_GO <- plot_gsea_lollipop(
  dat = gsea_top_GO,
  celltype_name = "EC"
)
EC_panel_GO
tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = EC_panel_GO,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/GO/EC_panel_GO.png"
)

unlink(tmp)

ec_individual_plots <- gsea_top_GO %>%
  filter(celltype == "EC") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_GO,
      celltype_name = "EC",
      contrast_name = .x
    )
  )

ec_individual_plots[[1]] #lc vs oc
ec_individual_plots[[2]] #lc vs t2d
ec_individual_plots[[3]] #oc vs t1d
ec_individual_plots[[4]] #oc vs t2d
ec_individual_plots[[5]] #t1d vs t2d
#ec_individual_plots[[6]] #no t1d vs lc
```

```{r}
TAL_panel_GO <- plot_gsea_lollipop(
  dat = gsea_top_GO,
  celltype_name = "TAL"
)

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = TAL_panel_GO,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/GO/TAL_panel_GO.png"
)

unlink(tmp)

tal_individual_plots <- gsea_top_GO %>%
  filter(celltype == "TAL") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_GO,
      celltype_name = "TAL",
      contrast_name = .x
    )
  )

tal_individual_plots[[1]] #lc vs oc
tal_individual_plots[[2]] #lc vs t1d
tal_individual_plots[[3]] #lc vs t2d
tal_individual_plots[[4]] #oc vs t1d
tal_individual_plots[[5]] #oc vs t2d
tal_individual_plots[[6]] #t1d vs t2d
```

```{r}
Immune_panel_GO <- plot_gsea_lollipop(
  dat = gsea_top_GO,
  celltype_name = "Immune")

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = Immune_panel_GO,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/GO/Immune_panel_GO.png"
)

unlink(tmp)

immune_individual_plots <- gsea_top_GO %>%
  filter(celltype == "Immune") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_GO,
      celltype_name = "Immune",
      contrast_name = .x
    )
  )

immune_individual_plots[[1]] #lc vs oc
immune_individual_plots[[2]] #lc vs t2d
immune_individual_plots[[3]] #oc vs t1d
immune_individual_plots[[4]] #oc vs t2d
immune_individual_plots[[5]] #t1d vs t2d
immune_individual_plots[[6]] #t1d vs lc
```

```{r}
POD_panel_GO <-plot_gsea_lollipop(
  dat = gsea_top_GO,
  celltype_name = "POD"
)
POD_panel_GO

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = POD_panel_GO,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/GO/POD_panel_GO.png"
)

unlink(tmp)

pod_individual_plots <- gsea_top_GO %>%
  filter(celltype == "POD") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_GO,
      celltype_name = "POD",
      contrast_name = .x
    )
  )

#pod_individual_plots[[1]] #lc vs oc not sig
#pod_individual_plots[[2]] #lc vs t1d not sig
#pod_individual_plots[[3]] #lc vs t2d not sig
#pod_individual_plots[[4]] #oc vs t1d not sig
#pod_individual_plots[[5]] #oc vs t2d not sig
pod_individual_plots[[1]] #t1d vs t2d
```

```{r}
DCT_CNT_panel_GO <-plot_gsea_lollipop(
  dat = gsea_top_GO,
  celltype_name = "DCT_CNT"
)

DCT_CNT_panel_GO

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = DCT_CNT_panel_GO,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/GO/DCT_CNT_panel_GO.png"
)

unlink(tmp)

dctCNT_individual_plots <- gsea_top_GO %>%
  filter(celltype == "DCT_CNT") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_GO,
      celltype_name = "DCT_CNT",
      contrast_name = .x
    )
  )

dctCNT_individual_plots[[1]] #lc vs oc
dctCNT_individual_plots[[2]] #lc vs t1d
dctCNT_individual_plots[[3]] #lc vs t2d
dctCNT_individual_plots[[4]] #oc vs t1d
dctCNT_individual_plots[[5]] #oc vs t2d
dctCNT_individual_plots[[6]] #t1d vs t2d
```

#--------------------
#plotting hallmark results
#--------------------
```{r}
gsea_results_hallmark <- edgeR_all %>%
  group_by(celltype, comparison) %>%
  nest() %>%
  dplyr::mutate(
    stats = map(data, make_rank_vector),

    fgsea = map(
      stats,
      ~ fgsea(
        pathways = hallmark_list,
        stats = .x,
        minSize = 15,
        maxSize = 500
      )
    )
  ) %>%
  dplyr::select(-data, -stats) %>%
  unnest(fgsea)

gsea_top_hallmark <- gsea_results_hallmark %>%
  filter(!is.na(padj)) %>%
  filter(padj < 0.05) %>%
  group_by(celltype, comparison) %>%
  slice_max(
    order_by = abs(NES),
    n = 10,
    with_ties = FALSE
  ) %>%
  ungroup()
```

```{r}
PT_panel_hallmark <- plot_gsea_lollipop(
  dat = gsea_top_hallmark,
  celltype_name = "PT"
)

PT_panel_hallmark

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = PT_panel_hallmark,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Hallmark/PT_panel_hallmark.png"
)

unlink(tmp)

pt_individual_plots <- gsea_top_hallmark %>%
  filter(celltype == "PT") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_hallmark,
      celltype_name = "PT",
      contrast_name = .x
    )
  )

pt_individual_plots[[1]] #lean vs obese
pt_individual_plots[[2]] #lean vs T1d
pt_individual_plots[[3]] #lean vs T2d
pt_individual_plots[[4]] #OC vs t1d
pt_individual_plots[[5]] #oc vs t2d
pt_individual_plots[[6]] #t1d vs t2d
```

```{r}
PC_panel_hallmark <-plot_gsea_lollipop(
  dat = gsea_top_hallmark,
  celltype_name = "PC"
)

PC_panel_hallmark

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = PC_panel_hallmark,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Hallmark/PC_panel_hallmark.png"
)

unlink(tmp)

pc_individual_plots <- gsea_top_hallmark %>%
  filter(celltype == "PC") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_hallmark,
      celltype_name = "PC",
      contrast_name = .x
    )
  )

pc_individual_plots[[1]] #lean vs obese
#pc_individual_plots[[2]] #lean vs T1d
pc_individual_plots[[2]] #lean vs T2d
pc_individual_plots[[3]] #OC vs t1d
pc_individual_plots[[4]] #oc vs t2d
pc_individual_plots[[5]] #t1d vs t2d
```

```{r}
EC_panel_hallmark <- plot_gsea_lollipop(
  dat = gsea_top_hallmark,
  celltype_name = "EC"
)

EC_panel_hallmark

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = EC_panel_hallmark,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Hallmark/EC_panel_hallmark.png"
)

unlink(tmp)

ec_individual_plots <- gsea_top_hallmark %>%
  filter(celltype == "EC") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_hallmark,
      celltype_name = "EC",
      contrast_name = .x
    )
  )

ec_individual_plots[[1]] #lc vs oc
ec_individual_plots[[2]] #lc vs t2d
ec_individual_plots[[3]] #oc vs t1d
ec_individual_plots[[4]] #oc vs t2d
ec_individual_plots[[5]] #t1d vs t2d
ec_individual_plots[[6]] #no t1d vs lc
```

```{r}
TAL_panel_hallmark <- plot_gsea_lollipop(
  dat = gsea_top_hallmark,
  celltype_name = "TAL"
)

TAL_panel_hallmark

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = TAL_panel_hallmark,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Hallmark/TAL_panel_hallmark.png"
)

unlink(tmp)

tal_individual_plots <- gsea_top_hallmark %>%
  filter(celltype == "TAL") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_hallmark,
      celltype_name = "TAL",
      contrast_name = .x
    )
  )

tal_individual_plots[[1]] #lc vs oc
tal_individual_plots[[2]] #lc vs t1d
tal_individual_plots[[3]] #lc vs t2d
tal_individual_plots[[4]] #oc vs t1d
tal_individual_plots[[5]] #oc vs t2d
tal_individual_plots[[6]] #t1d vs t2d
```

```{r}
Immune_panel_hallmark <-plot_gsea_lollipop(
  dat = gsea_top_hallmark,
  celltype_name = "Immune"
)

Immune_panel_hallmark

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = Immune_panel_hallmark,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Hallmark/Immune_panel_hallmark.png"
)

unlink(tmp)

immune_individual_plots <- gsea_top_hallmark %>%
  filter(celltype == "Immune") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_hallmark,
      celltype_name = "Immune",
      contrast_name = .x
    )
  )

immune_individual_plots[[1]] #lc vs oc
immune_individual_plots[[2]] #lc vs t2d
immune_individual_plots[[3]] #oc vs t1d
immune_individual_plots[[4]] #oc vs t2d
immune_individual_plots[[5]] #t1d vs t2d
immune_individual_plots[[6]] #t1d vs lc
```

```{r}
POD_panel_hallmark <-plot_gsea_lollipop(
  dat = gsea_top_hallmark,
  celltype_name = "POD"
)

POD_panel_hallmark

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = POD_panel_hallmark,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Hallmark/POD_panel_hallmark.png"
)

unlink(tmp)

pod_individual_plots <- gsea_top_hallmark %>%
  filter(celltype == "POD") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_hallmark,
      celltype_name = "POD",
      contrast_name = .x
    )
  )

#pod_individual_plots[[1]] #lc vs oc not sig
#pod_individual_plots[[2]] #lc vs t1d not sig
#pod_individual_plots[[3]] #lc vs t2d not sig
#pod_individual_plots[[4]] #oc vs t1d not sig
#pod_individual_plots[[5]] #oc vs t2d not sig
pod_individual_plots[[1]] #t1d vs t2d
```

```{r}
DCT_CNT_panel_hallmark <- plot_gsea_lollipop(
  dat = gsea_top_hallmark,
  celltype_name = "DCT_CNT"
)

DCT_CNT_panel_hallmark

tmp <- tempfile(fileext = ".png")

ggsave(
  filename = tmp,
  plot = DCT_CNT_panel_hallmark,
  width = 24,
  height = 12,
  dpi = 300
)

upload_to_kopah(
  file = tmp,
  object = "figures/GSEA/Hallmark/DCT_CNT_panel_hallmark.png"
)

unlink(tmp)

dctCNT_individual_plots <- gsea_top_hallmark %>%
  filter(celltype == "DCT_CNT") %>%
  distinct(comparison) %>%
  pull(comparison) %>%
  set_names() %>%
  purrr::map(
    ~ plot_gsea_lollipop(
      dat = gsea_top_hallmark,
      celltype_name = "DCT_CNT",
      contrast_name = .x
    )
  )

dctCNT_individual_plots[[1]] #lc vs oc
#dctCNT_individual_plots[[2]] #lc vs t1d
dctCNT_individual_plots[[2]] #lc vs t2d
dctCNT_individual_plots[[3]] #oc vs t1d
dctCNT_individual_plots[[4]] #oc vs t2d
dctCNT_individual_plots[[5]] #t1d vs t2d
```

