---
title: "11_disease_trajectory"
format: html
editor: visual
---

```{r}
library(Seurat)
library(dplyr)
library(harmony)
library(aws.s3)
```

```{r}
```{r}
#directores
user <- Sys.info()[["user"]]

if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/") #hyak
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/") #mac studio
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/") 
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/Savanah Leidholt"
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/") #laptop
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/") 
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad/Savanah Leidholt"
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r}
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```
```


```{r}
clean_trajectory_object <- function(obj,
                                    object_name,
                                    dims_use = 1:20,
                                    harmony_var = "study",
                                    resolution = 0.4) {
  
  message("Processing: ", object_name)
  
  print(table(obj$group, useNA = "ifany"))
  print(table(obj$study, useNA = "ifany"))
  
  DefaultAssay(obj) <- "RNA"
  
  obj@reductions <- list()
  obj@graphs <- list()
  
  obj$trajectory_object <- object_name
  
  obj <- FindVariableFeatures(
    obj,
    selection.method = "vst",
    nfeatures = 3000,
    verbose = FALSE
  )
  
  obj <- ScaleData(
    obj,
    features = VariableFeatures(obj),
    verbose = FALSE
  )
  
  obj <- RunPCA(
    obj,
    features = VariableFeatures(obj),
    npcs = max(dims_use),
    verbose = FALSE
  )
  
  obj <- RunHarmony(
    object = obj,
    group.by.vars = harmony_var,
    reduction = "pca",
    dims.use = dims_use,
    verbose = FALSE
  )
  
  obj <- RunUMAP(
    obj,
    reduction = "harmony",
    dims = dims_use,
    reduction.name = "umap_harmony",
    reduction.key = "UMAPH_",
    verbose = FALSE
  )
  
  obj <- FindNeighbors(
    obj,
    reduction = "harmony",
    dims = dims_use,
    verbose = FALSE
  )
  
  obj <- FindClusters(
    obj,
    resolution = resolution,
    verbose = FALSE
  )
  
  return(obj)
}
```

```{r}
pb90_PT <- s3readRDS(
  object = "data/pb90_multiomics_PT_trajectory_20260529.rds",
  bucket = bucket_name,
  region = "",
  endpoint = s3_endpoint
)
```

