---
title: "Elastic Net Analysis"
output: pdf_document
author: "Savanah Leidholt"
date: "02/17/2026"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths("~/leidholt/R")
```

```{r}
#load packages
library(dplyr)
library(tidyr)
library(glmnet)
library(pROC)
library(ggplot2)
```
#=====================================================
# inputting all directories and storage locations
#=====================================================
```{r}
#Directories
#computer <- "mac studio"
computer <- "mac laptop"

user <- Sys.info()[["user"]]

if (computer == "mac studio") {
  
  if (user == "sleidholt") {
    dir.dat <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/PANTHER PFAS/"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
    
  } else if (user == "choiyej") {
    dir.dat <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "pylell") {
    dir.dat <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
    dir.results <- paste0(dir.dat, "/PANTHER/Results")
    git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  }

} else if (computer == "mac laptop") {

  if (user == "savanahleidholt") {
    dir.dat <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/"
    dir.results <- "/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/"
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    
  } else if (user == "shivaniramesh") {
    dir.dat <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive"
    dir.results <- "/Users/shivaniramesh/Library/CloudStorage/OneDrive-UW/Biostatistics Core Shared Drive/Liver Pet IR/Results"
  }
}
```


#load in R objects and split them
```{r}


# ---- load split objects created in your first Rmd ----
obj <- readRDS(file.path(dir.dat, "Savanah Leidholt", "T1D-T2D", "Data", "omics_split_objects3.rds"))

meta  <- obj$meta_prot
prot_plasma <- obj$prot_plasma
n_table     <- obj$n_table


obj2 <- readRDS(file.path(
  dir.dat, "Savanah Leidholt", "T1D-T2D", "Data",
  "limma_gsea_inputs_age_sex_study.rds"
))

dat1 <- obj2$all_studies
dat2 <- obj2$no_attempt
dat3 <- obj2$all3_omics

out1 <- dat1$out
out2 <- dat2$out
out3 <- dat3$out


# quick checks
dim(prot_plasma)
dim(meta)



```

#reproducing the same alignment/record
```{r}
meta <- meta %>%
  mutate(
    record_id = trimws(as.character(record_id)),
    study     = trimws(as.character(study)),
    group     = trimws(as.character(group)),
    sex       = trimws(as.character(sex)),
    ethnicity = trimws(as.character(ethnicity))
  ) %>%
  filter(!is.na(age), !is.na(sex)) %>%
  mutate(
    group4 = case_when(
      group %in% c("Lean Control") ~ "LC",
      group %in% c("Obese Control") ~ "OC",
      group %in% c("Type 1 Diabetes") ~ "T1D",
      group %in% c("Type 2 Diabetes") ~ "T2D",
      TRUE ~ NA_character_
    ),
    group4 = factor(group4, levels = c("LC","OC","T1D","T2D"))
  ) %>%
  filter(!is.na(group4))

table(meta$group4, useNA="ifany")

meta <- meta %>% filter(record_id %in% rownames(prot_plasma))
prot_plasma <- prot_plasma[meta$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_plasma), meta$record_id))


```
#logfold and imputation
```{r}
# ----------------------------------------
# LOG2 + ROBUST IMPUTATION
# (median + tiny jitter to avoid ties)
# ----------------------------------------
X_log <- log2(as.matrix(prot_plasma) + 1)

# OPTIONAL: drop proteins that are basically all missing
# (good idea for glmnet stability)
min_non_missing <- 10  # adjust if needed
keep_feat <- colSums(!is.na(X_log)) >= min_non_missing
X_log <- X_log[, keep_feat, drop = FALSE]

# Median impute per protein (column-wise)
X_imp <- X_log
for (j in seq_len(ncol(X_imp))) {
  xj <- X_imp[, j]
  med <- median(xj, na.rm = TRUE)
  if (is.finite(med)) {
    xj[is.na(xj)] <- med
    X_imp[, j] <- xj
  }
}

# OPTIONAL: add tiny jitter so columns with many imputed values
# don't create perfect separation / numerical issues
set.seed(1)
X_imp <- X_imp + matrix(rnorm(length(X_imp), mean = 0, sd = 1e-6),
                        nrow = nrow(X_imp), ncol = ncol(X_imp))

# final alignment check
stopifnot(identical(rownames(X_imp), meta$record_id))

dim(X_imp)
summary(colSums(is.na(X_imp)))   # should be all 0



```

```{r}
# ---- protein -> gene symbol mapping (seq.* IDs) ----
analytes <- read.csv(file.path(dir.dat, "Data Harmonization", "Combined SomaScan", "analytes.csv"))

protein2symbol <- analytes %>%
  transmute(
    feature_id  = trimws(as.character(AptName)),
    gene_symbol = trimws(as.character(EntrezGeneSymbol))
  ) %>%
  distinct()

head(protein2symbol)


# fast lookup
apt2symbol <- setNames(protein2symbol$gene_symbol, protein2symbol$feature_id)
head(apt2symbol)


# QC: overlap + missing symbols
sum(colnames(prot_plasma) %in% names(apt2symbol))

table(is.na(apt2symbol[colnames(prot_plasma)]) | apt2symbol[colnames(prot_plasma)] == "")



```


```{r}
run_elasticnet_topN <- function(
  cc,
  out_list,
  meta,
  X_imp,
  top_n = 100,
  covars = c("age", "sex", "study"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1,
  min_per_class = 5,
  maxit = 1e6
) {
  if (!grepl("_vs_", cc)) return(NULL)

  parts <- strsplit(cc, "_vs_")[[1]]
  if (length(parts) != 2) return(NULL)

  g_pos <- parts[1]
  g_neg <- parts[2]

  idx <- which(meta$group4 %in% c(g_pos, g_neg))
  meta_sub <- meta[idx, , drop = FALSE]
  meta_sub$record_id <- as.character(meta_sub$record_id)

  y <- ifelse(meta_sub$group4 == g_pos, 1, 0)
  y <- as.numeric(y)

  tab <- table(meta_sub$group4)
  if (any(tab < min_per_class)) return(NULL)

  tt <- out_list[[cc]]
  if (is.null(tt)) return(NULL)
  tt <- as.data.frame(tt)

  if (!"feature_id" %in% names(tt)) {
    fid <- rownames(tt)
    if (is.null(fid)) return(NULL)
    tt$feature_id <- fid
  }

  if (!all(c("adj.P.Val", "P.Value") %in% names(tt))) return(NULL)

  top_feats <- tt %>%
    filter(!is.na(adj.P.Val)) %>%
    arrange(adj.P.Val, P.Value) %>%
    slice_head(n = top_n) %>%
    pull(feature_id) %>%
    intersect(colnames(X_imp))

  if (length(top_feats) < 5) return(NULL)

  meta_sub <- meta_sub %>%
    mutate(
      sex = factor(sex),
      study = factor(study)
    )

  covars_present <- covars[covars %in% colnames(meta_sub)]
  form_cov <- as.formula(paste("~", paste(covars_present, collapse = " + ")))
  Z <- model.matrix(form_cov, data = meta_sub)

  Xp <- X_imp[meta_sub$record_id, top_feats, drop = FALSE]

  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"

  pf <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))

  sds <- apply(X, 2, sd)
  keep_cols <- which(is.finite(sds) & sds > 0)
  X  <- X[, keep_cols, drop = FALSE]
  pf <- pf[keep_cols]

  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))
  if (nfolds_use < 2) return(NULL)

  set.seed(seed)

  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      cv.glmnet(
        x = X, y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = pf,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) NULL
    )
    if (is.null(cv)) return(NULL)
    list(alpha = a, cvfit = cv)
  })

  fit_list <- Filter(Negate(is.null), fit_list)
  if (length(fit_list) == 0) return(NULL)

  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]

  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))
  coef_df <- tibble(
    term = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>%
    filter(beta != 0)

  p_hat <- as.numeric(predict(best$cvfit, newx = X, s = "lambda.1se", type = "response"))
  auc_val <- as.numeric(pROC::auc(pROC::roc(y, p_hat, quiet = TRUE)))

  protein_terms <- intersect(colnames(Xp), coef_df$term)

  list(
    contrast      = cc,
    positive      = g_pos,
    negative      = g_neg,
    best_alpha    = best$alpha,
    lambda_1se    = best$cvfit$lambda.1se,
    lambda_min    = best$cvfit$lambda.min,
    auc_train     = auc_val,
    coef_df       = coef_df,
    protein_terms = protein_terms,
    cvfit         = best$cvfit
  )
}
```

```{r}
# ---------------------------
#run across contrasts top 100
#----------------------------

glmnet_res_all <- lapply(names(out3), function(cc) {
  run_elasticnet_topN(
    cc       = cc,
    out_list = out1,
    meta     = meta,
    X_imp    = X_imp,
    top_n    = 100,
    covars   = c("age", "sex", "study")
  )
})

names(glmnet_res_all) <- names(out1)
glmnet_res_all <- Filter(Negate(is.null), glmnet_res_all)
```



```{r}
# ---- summary table ----
summary_tbl <- bind_rows(lapply(glmnet_res_all, function(r) {
  tibble(
    contrast   = r$contrast,
    alpha      = r$best_alpha,
    lambda_1se = r$lambda_1se,
    lambda_min = r$lambda_min,
    auc_train  = r$auc_train,
    n_proteins = sum(r$coef_df$term %in% r$protein_terms)
  )
}))

summary_tbl
```


```{r}
r <- glmnet_res_all[["T1D_vs_T2D"]]

r$best_alpha
r$auc_train
head(r$coef_df, 30)

plot(r$cvfit)
title(paste0("CV: ", r$contrast, " (alpha=", r$best_alpha, ")"))

```

```{r}
sig_tbl_all <- bind_rows(lapply(glmnet_res, function(r) {
  r$coef_df %>%
    mutate(
      contrast   = r$contrast,
      alpha      = r$best_alpha,
      lambda_1se = r$lambda_1se,
      lambda_min = r$lambda_min,
      auc_train  = r$auc_train
    )
}))

# keep only protein terms (drop intercept + covariates)
sig_tbl_proteins <- bind_rows(lapply(glmnet_res, function(r) {
  r$coef_df %>%
    filter(term %in% r$protein_terms) %>%
    mutate(
      contrast   = r$contrast,
      alpha      = r$best_alpha,
      lambda_1se = r$lambda_1se,
      lambda_min = r$lambda_min,
      auc_train  = r$auc_train
    )
}))

head(sig_tbl_proteins)

```

```{r}
sig_proteins <- sig_tbl_proteins %>%
  rename(feature_id = term) %>%
  left_join(protein2symbol, by = "feature_id")

# optional: if you want to drop unmapped symbols
sig_proteins_mapped <- sig_proteins %>%
  filter(!is.na(gene_symbol), gene_symbol != "")

# write outputs
#write.csv(
 # sig_proteins,
#  "/mmfs1/gscratch/togo/leidholt/Results/enriched_proteins/elasticnet_plasma_top100_signatures_ALL_PROTEINS_WITH_NA_SYMBOLS.csv",
 # row.names = FALSE
#)

#write.csv(
 # sig_proteins_mapped,
  #"/mmfs1/gscratch/togo/leidholt/Results/enriched_proteins/elasticnet_plasma_top100_signatures_MAPPED_SYMBOLS_ONLY.csv",
  #row.names = FALSE
#)

```




#--------
#TOP 500
#--------
```{r}
# ---- elastic net runner ----
run_elasticnet_topN <- function(
  cc,
  out_list,
  meta,
  X_imp,
  top_n = 500,
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1,
  min_per_class = 5,
  maxit = 1e6
) {
  # parse contrast like "T1D_vs_T2D"
  if (!grepl("_vs_", cc)) return(NULL)
  parts <- unlist(strsplit(cc, "_vs_"))
  if (length(parts) != 2) return(NULL)
  g_pos <- parts[1]
  g_neg <- parts[2]

  # subset to the two groups
  idx <- which(meta$group4 %in% c(g_pos, g_neg))
  meta_sub <- meta[idx, , drop = FALSE]
  meta_sub$record_id <- as.character(meta_sub$record_id)

  y <- ifelse(meta_sub$group4 == g_pos, 1, 0)
  y <- as.numeric(y)

  tab <- table(as.character(meta_sub$group4))
  if (any(tab < min_per_class)) return(NULL)
  
  # get limma topTable for that contrast (already per-contrast)
  tt <- out_list[[cc]]
  if (is.null(tt)) return(NULL)
  tt <- as.data.frame(tt)

  # ensure feature_id
  if (!"feature_id" %in% names(tt)) {
    fid <- rownames(tt)
    if (is.null(fid)) return(NULL)
    tt$feature_id <- fid
  }

  # pick top N by adj.P.Val then P.Value
  if (!all(c("adj.P.Val", "P.Value") %in% names(tt))) return(NULL)

  top_feats <- tt %>%
    filter(!is.na(adj.P.Val)) %>%
    arrange(adj.P.Val, P.Value) %>%
    slice_head(n = top_n) %>%
    pull(feature_id) %>%
    intersect(colnames(X_imp))

  if (length(top_feats) < 5) return(NULL)

  # covariates (unpenalized)
  meta_sub <- meta_sub %>% mutate(sex = factor(sex))
  Z <- model.matrix(~ age + sex, data = meta_sub)

  # proteins (penalized)
  Xp <- X_imp[meta_sub$record_id, top_feats, drop = FALSE]

  # build X (covariates + proteins)
  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"

  # penalty factors: 0 for covariates, 1 for proteins
  pf <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))

  # drop zero-variance columns
  sds <- apply(X, 2, sd)
  keep_cols <- which(is.finite(sds) & sds > 0)
  X  <- X[, keep_cols, drop = FALSE]
  pf <- pf[keep_cols]

  # usable folds
  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))
  if (nfolds_use < 2) return(NULL)

  set.seed(seed)

  # fit across alpha grid
  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      cv.glmnet(
        x = X, y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = pf,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) NULL
    )
    if (is.null(cv)) return(NULL)
    list(alpha = a, cvfit = cv)
  })
  fit_list <- Filter(Negate(is.null), fit_list)
  if (length(fit_list) == 0) return(NULL)

  # choose alpha with minimum cvm
  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]

  # coefficients at lambda.1se
  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))
  coef_df <- tibble(
    term = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>% filter(beta != 0)

  # training AUC (predictions at lambda.1se)
  p_hat <- as.numeric(predict(best$cvfit, newx = X, s = "lambda.1se", type = "response"))
  auc_val <- as.numeric(pROC::auc(pROC::roc(y, p_hat, quiet = TRUE)))

  protein_terms <- intersect(colnames(Xp), coef_df$term)

  list(
    contrast      = cc,
    positive      = g_pos,
    negative      = g_neg,
    best_alpha    = best$alpha,
    lambda_1se    = best$cvfit$lambda.1se,
    lambda_min    = best$cvfit$lambda.min,
    auc_train     = auc_val,
    coef_df       = coef_df,
    protein_terms = protein_terms,
    cvfit         = best$cvfit
  )
}


```


```{r}
# ---------------------------
#run across contrasts top 500
#----------------------------

glmnet_500 <- lapply(names(out_no), function(cc) {
  run_elasticnet_topN(
    cc      = cc,
    out_list= out_no,
    meta    = meta,
    X_imp   = X_imp,
    top_n   = 500
  )
})
names(glmnet_500) <- names(out_no)
glmnet_500 <- Filter(Negate(is.null), glmnet_500)
```


```{r}
# ---- summary table ----
summary_tbl_500 <- bind_rows(lapply(glmnet_res, function(r) {
  tibble(
    contrast   = r$contrast,
    alpha      = r$best_alpha,
    lambda_1se = r$lambda_1se,
    lambda_min = r$lambda_min,
    auc_train  = r$auc_train,
    n_proteins = sum(r$coef_df$term %in% r$protein_terms)
  )
}))

summary_tbl_500


```

```{r}
r_500 <- glmnet_500[["T1D_vs_T2D"]]

r_500$best_alpha
r_500$auc_train
head(r_500$coef_df, 30)

plot(r_500$cvfit)
title(paste0("CV: ", r$contrast, " (alpha=", r_500$best_alpha, ")"))

```

```{r}
sig_tbl_500 <- bind_rows(lapply(glmnet_500, function(r_500) {
  r_500$coef_df %>%
    mutate(
      contrast   = r_500$contrast,
      alpha      = r_500$best_alpha,
      lambda_1se = r_500$lambda_1se,
      lambda_min = r_500$lambda_min,
      auc_train  = r_500$auc_train
    )
}))

# keep only protein terms (drop intercept + covariates)
sig_tbl_500 <- bind_rows(lapply(glmnet_500, function(r_500) {
  r_500$coef_df %>%
    filter(term %in% r_500$protein_terms) %>%
    mutate(
      contrast   = r_500$contrast,
      alpha      = r_500$best_alpha,
      lambda_1se = r_500$lambda_1se,
      lambda_min = r_500$lambda_min,
      auc_train  = r_500$auc_train
    )
}))

head(sig_tbl_500)

```

```{r}
sig_proteins_500 <- sig_tbl_500 %>%
  rename(feature_id = term) %>%
  left_join(protein2symbol, by = "feature_id")

# optional: if you want to drop unmapped symbols
sig_proteins_mapped_500 <- sig_proteins_500 %>%
  filter(!is.na(gene_symbol), gene_symbol != "")

# write outputs
write.csv(
  sig_proteins_500,
  "/mmfs1/gscratch/togo/leidholt/Data/enriched_proteins/elasticnet_plasma_top500_signatures_ALL_PROTEINS_WITH_NA_SYMBOLS.csv",
  row.names = FALSE
)

write.csv(
  sig_proteins_mapped_500,
  "/mmfs1/gscratch/togo/leidholt/Data/enriched_proteins/elasticnet_plasma_top500_signatures_MAPPED_SYMBOLS_ONLY.csv",
  row.names = FALSE
)

```








#----------------------------------------
#running with ethnicity in the model
#---------------------------------------

#load in R objects and split them
```{r}


# split objects (matrix + meta)
obj <- readRDS("/mmfs1/gscratch/togo/leidholt/Data/R_objects/omics_split_objects3.rds")
meta        <- obj$meta_prot
prot_plasma <- obj$prot_plasma

# limma outputs (out_plasma contains topTable per contrast)
limma_obj_eth <- readRDS("/mmfs1/gscratch/togo/leidholt/Data/R_objects/limma_outputs_eth.rds")
out_eth <- limma_obj_eth$out_eth

# quick checks
dim(prot_plasma)
dim(meta)
names(out_eth)


```

#reproducing the same alignment/record
```{r}
meta <- meta %>%
  mutate(
    record_id = trimws(as.character(record_id)),
    study     = trimws(as.character(study)),
    group     = trimws(as.character(group)),
    sex       = trimws(as.character(sex)),
    ethnicity = trimws(as.character(ethnicity))
  ) %>%
  filter(!is.na(age), !is.na(sex), !is.na(ethnicity)) %>%
  mutate(
    group4 = case_when(
      group %in% c("Lean Control") ~ "LC",
      group %in% c("Obese Control") ~ "OC",
      group %in% c("Type 1 Diabetes") ~ "T1D",
      group %in% c("Type 2 Diabetes") ~ "T2D",
      TRUE ~ NA_character_
    ),
    group4 = factor(group4, levels = c("LC","OC","T1D","T2D"))
  ) %>%
  filter(!is.na(group4))

table(meta$group4, useNA="ifany")

meta <- meta %>% filter(record_id %in% rownames(prot_plasma))
prot_plasma <- prot_plasma[meta$record_id, , drop = FALSE]

stopifnot(identical(rownames(prot_plasma), meta$record_id))


```
#logfold and imputation
```{r}
# ----------------------------------------
# LOG2 + ROBUST IMPUTATION
# (median + tiny jitter to avoid ties)
# ----------------------------------------
X_log <- log2(as.matrix(prot_plasma) + 1)

# OPTIONAL: drop proteins that are basically all missing
# (good idea for glmnet stability)
min_non_missing <- 10  # adjust if needed
keep_feat <- colSums(!is.na(X_log)) >= min_non_missing
X_log <- X_log[, keep_feat, drop = FALSE]

# Median impute per protein (column-wise)
X_imp <- X_log
for (j in seq_len(ncol(X_imp))) {
  xj <- X_imp[, j]
  med <- median(xj, na.rm = TRUE)
  if (is.finite(med)) {
    xj[is.na(xj)] <- med
    X_imp[, j] <- xj
  }
}

# OPTIONAL: add tiny jitter so columns with many imputed values
# don't create perfect separation / numerical issues
set.seed(1)
X_imp <- X_imp + matrix(rnorm(length(X_imp), mean = 0, sd = 1e-6),
                        nrow = nrow(X_imp), ncol = ncol(X_imp))

# final alignment check
stopifnot(identical(rownames(X_imp), meta$record_id))

dim(X_imp)
summary(colSums(is.na(X_imp)))   # should be all 0


# if feature_id column exists:
if ("feature_id" %in% colnames(tt1)) head(tt1$feature_id, 10)

```

```{r}
# ---- protein -> gene symbol mapping (seq.* IDs) ----
analytes <- read.csv("~/leidholt/Data/analyte_data/analytes.csv", stringsAsFactors = FALSE)



protein2symbol <- analytes %>%
  transmute(
    feature_id  = trimws(as.character(AptName)),
    gene_symbol = trimws(as.character(EntrezGeneSymbol))
  ) %>%
  distinct()




# fast lookup
apt2symbol <- setNames(protein2symbol$gene_symbol, protein2symbol$feature_id)
head(apt2symbol)


# QC: overlap + missing symbols
sum(colnames(prot_plasma) %in% names(apt2symbol))

table(is.na(apt2symbol[colnames(prot_plasma)]) | apt2symbol[colnames(prot_plasma)] == "")



```


```{r}
# ---- elastic net runner ----
run_elasticnet_topN <- function(
  cc,
  out_list,
  meta,
  X_imp,
  top_n = 100,
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1,
  min_per_class = 5,
  maxit = 1e6
) {
  # parse contrast like "T1D_vs_T2D"
  if (!grepl("_vs_", cc)) return(NULL)
  parts <- unlist(strsplit(cc, "_vs_"))
  if (length(parts) != 2) return(NULL)
  g_pos <- parts[1]
  g_neg <- parts[2]

  # subset to the two groups
  idx <- which(meta$group4 %in% c(g_pos, g_neg))
  meta_sub <- meta[idx, , drop = FALSE]
  meta_sub$record_id <- as.character(meta_sub$record_id)

  y <- ifelse(meta_sub$group4 == g_pos, 1, 0)
  y <- as.numeric(y)

  tab <- table(as.character(meta_sub$group4))
  if (any(tab < min_per_class)) return(NULL)
  
  # get limma topTable for that contrast (already per-contrast)
  tt <- out_list[[cc]]
  if (is.null(tt)) return(NULL)
  tt <- as.data.frame(tt)

  # ensure feature_id
  if (!"feature_id" %in% names(tt)) {
    fid <- rownames(tt)
    if (is.null(fid)) return(NULL)
    tt$feature_id <- fid
  }

  # pick top N by adj.P.Val then P.Value
  if (!all(c("adj.P.Val", "P.Value") %in% names(tt))) return(NULL)

  top_feats <- tt %>%
    filter(!is.na(adj.P.Val)) %>%
    arrange(adj.P.Val, P.Value) %>%
    slice_head(n = top_n) %>%
    pull(feature_id) %>%
    intersect(colnames(X_imp))

  if (length(top_feats) < 5) return(NULL)

  # covariates (unpenalized)
  meta_sub <- meta_sub %>% mutate(sex = factor(sex))
  Z <- model.matrix(~ age + sex, data = meta_sub)

  # proteins (penalized)
  Xp <- X_imp[meta_sub$record_id, top_feats, drop = FALSE]

  # build X (covariates + proteins)
  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"

  # penalty factors: 0 for covariates, 1 for proteins
  pf <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))

  # drop zero-variance columns
  sds <- apply(X, 2, sd)
  keep_cols <- which(is.finite(sds) & sds > 0)
  X  <- X[, keep_cols, drop = FALSE]
  pf <- pf[keep_cols]

  # usable folds
  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))
  if (nfolds_use < 2) return(NULL)

  set.seed(seed)

  # fit across alpha grid
  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      cv.glmnet(
        x = X, y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = pf,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) NULL
    )
    if (is.null(cv)) return(NULL)
    list(alpha = a, cvfit = cv)
  })
  fit_list <- Filter(Negate(is.null), fit_list)
  if (length(fit_list) == 0) return(NULL)

  # choose alpha with minimum cvm
  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]

  # coefficients at lambda.1se
  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))
  coef_df <- tibble(
    term = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>% filter(beta != 0)

  # training AUC (predictions at lambda.1se)
  p_hat <- as.numeric(predict(best$cvfit, newx = X, s = "lambda.1se", type = "response"))
  auc_val <- as.numeric(pROC::auc(pROC::roc(y, p_hat, quiet = TRUE)))

  protein_terms <- intersect(colnames(Xp), coef_df$term)

  list(
    contrast      = cc,
    positive      = g_pos,
    negative      = g_neg,
    best_alpha    = best$alpha,
    lambda_1se    = best$cvfit$lambda.1se,
    lambda_min    = best$cvfit$lambda.min,
    auc_train     = auc_val,
    coef_df       = coef_df,
    protein_terms = protein_terms,
    cvfit         = best$cvfit
  )
}


```

```{r}
# ---------------------------
#run across contrasts top 100
#----------------------------

glmnet_res_eth <- lapply(names(out_eth), function(cc) {
  run_elasticnet_topN(
    cc      = cc,
    out_list= out_eth,
    meta    = meta,
    X_imp   = X_imp,
    top_n   = 100
  )
})
names(glmnet_res_eth) <- names(out_eth)
glmnet_res_eth <- Filter(Negate(is.null), glmnet_res_eth)

names(glmnet_res_eth)
length(glmnet_res_eth)
```


```{r}
# ---- summary table ----
summary_tbl <- bind_rows(lapply(glmnet_res_eth, function(r) {
  tibble(
    contrast   = r$contrast,
    alpha      = r$best_alpha,
    lambda_1se = r$lambda_1se,
    lambda_min = r$lambda_min,
    auc_train  = r$auc_train,
    n_proteins = sum(r$coef_df$term %in% r$protein_terms)
  )
}))

summary_tbl


```


```{r}
r <- glmnet_res[["T1D_vs_T2D"]]

r$best_alpha
r$auc_train
head(r$coef_df, 30)

plot(r$cvfit)
title(paste0("CV: ", r$contrast, " (alpha=", r$best_alpha, ")"))

```

```{r}
sig_tbl_all <- bind_rows(lapply(glmnet_res_eth, function(r) {
  r$coef_df %>%
    mutate(
      contrast   = r$contrast,
      alpha      = r$best_alpha,
      lambda_1se = r$lambda_1se,
      lambda_min = r$lambda_min,
      auc_train  = r$auc_train
    )
}))

# keep only protein terms (drop intercept + covariates)
sig_tbl_proteins <- bind_rows(lapply(glmnet_res_eth, function(r) {
  r$coef_df %>%
    filter(term %in% r$protein_terms) %>%
    mutate(
      contrast   = r$contrast,
      alpha      = r$best_alpha,
      lambda_1se = r$lambda_1se,
      lambda_min = r$lambda_min,
      auc_train  = r$auc_train
    )
}))

head(sig_tbl_proteins)

```

```{r}
sig_proteins2 <- sig_tbl_proteins %>%
  rename(feature_id = term) %>%
  left_join(protein2symbol, by = "feature_id")

# optional: if you want to drop unmapped symbols
sig_proteins_mapped2 <- sig_proteins %>%
  filter(!is.na(gene_symbol), gene_symbol != "")

# write outputs
write.csv(
  sig_proteins2,
  "/mmfs1/gscratch/togo/leidholt/Data/enriched_proteins/top100_signatures_ALL_PROTS_ETHNICITY.csv",
  row.names = FALSE
)

write.csv(
  sig_proteins_mapped2,
  "/mmfs1/gscratch/togo/leidholt/Data/enriched_proteins/top100_signatures_MAPPED_PROTS_ETHNICITY.csv",
  row.names = FALSE
)

```


#-------------------------------------------
#checking overlap across the two main models
#-------------------------------------------

```{r}
extract_selected <- function(glmnet_list) {
  if (length(glmnet_list) == 0) return(tibble())

  bind_rows(lapply(glmnet_list, function(r) {
    if (is.null(r) || is.null(r$coef_df)) return(NULL)

    r$coef_df %>%
      filter(term %in% r$protein_terms) %>%   # proteins only
      transmute(
        contrast = r$contrast,
        feature_id = term,
        beta = beta,
        sign = if_else(beta > 0, "pos", "neg")
      )
  }))
}
```

```{r}
sel_no  <- extract_selected(glmnet_res)   # change to your object name
sel_eth <- extract_selected(glmnet_res_eth)  # change to your object name

dplyr::glimpse(sel_no)
dplyr::glimpse(sel_eth)

# optional: map to gene symbols for readability
sel_no  <- sel_no  %>% left_join(protein2symbol, by = "feature_id")
sel_eth <- sel_eth %>% left_join(protein2symbol, by = "feature_id")
```

```{r}
overlap_per_contrast <- function(sel_a, sel_b) {
  contrasts <- intersect(unique(sel_a$contrast), unique(sel_b$contrast))

  bind_rows(lapply(contrasts, function(cc) {
    a <- sel_a %>% filter(contrast == cc) %>% pull(feature_id) %>% unique()
    b <- sel_b %>% filter(contrast == cc) %>% pull(feature_id) %>% unique()

    inter <- intersect(a, b)
    uni   <- union(a, b)

    tibble(
      contrast = cc,
      n_a = length(a),
      n_b = length(b),
      n_intersect = length(inter),
      jaccard = ifelse(length(uni) == 0, NA_real_, length(inter) / length(uni))
    )
  })) %>% arrange(desc(jaccard), desc(n_intersect))
}

overlap_tbl <- overlap_per_contrast(sel_no, sel_eth)
overlap_tbl
```

```{r}
direction_overlap <- sel_no %>%
  select(contrast, feature_id, beta_no = beta, sign_no = sign) %>%
  inner_join(
    sel_eth %>% select(contrast, feature_id, beta_eth = beta, sign_eth = sign),
    by = c("contrast", "feature_id")
  ) %>%
  mutate(sign_match = (sign_no == sign_eth))

# summary per contrast
dir_summary <- direction_overlap %>%
  group_by(contrast) %>%
  summarise(
    n_overlap = n(),
    n_sign_match = sum(sign_match),
    prop_sign_match = mean(sign_match),
    .groups = "drop"
  ) %>%
  arrange(desc(prop_sign_match), desc(n_overlap))

dir_summary
```
#-----------------------------------------------------------------
#checking result similarities across age + sex top 100 and top 500
#-----------------------------------------------------------------
```{r}
sel_no  <- extract_selected(glmnet_res)   # change to your object name
sel_500 <- extract_selected(glmnet_500)  # change to your object name

dplyr::glimpse(sel_no)
dplyr::glimpse(sel_500)

# optional: map to gene symbols for readability
sel_no  <- sel_no  %>% left_join(protein2symbol, by = "feature_id")
sel_500 <- sel_500 %>% left_join(protein2symbol, by = "feature_id")
```

```{r}
overlap_per_contrast <- function(sel_a, sel_b) {
  contrasts <- intersect(unique(sel_a$contrast), unique(sel_b$contrast))

  bind_rows(lapply(contrasts, function(cc) {
    a <- sel_a %>% filter(contrast == cc) %>% pull(feature_id) %>% unique()
    b <- sel_b %>% filter(contrast == cc) %>% pull(feature_id) %>% unique()

    inter <- intersect(a, b)
    uni   <- union(a, b)

    tibble(
      contrast = cc,
      n_a = length(a),
      n_b = length(b),
      n_intersect = length(inter),
      jaccard = ifelse(length(uni) == 0, NA_real_, length(inter) / length(uni))
    )
  })) %>% arrange(desc(jaccard), desc(n_intersect))
}

overlap_tbl <- overlap_per_contrast(sel_no, sel_500)
overlap_tbl
```

```{r}
direction_overlap <- sel_no %>%
  select(contrast, feature_id, beta_no = beta, sign_no = sign) %>%
  inner_join(
    sel_500 %>% select(contrast, feature_id, beta_eth = beta, sign_500 = sign),
    by = c("contrast", "feature_id")
  ) %>%
  mutate(sign_match = (sign_no == sign_500))

# summary per contrast
dir_summary <- direction_overlap %>%
  group_by(contrast) %>%
  summarise(
    n_overlap = n(),
    n_sign_match = sum(sign_match),
    prop_sign_match = mean(sign_match),
    .groups = "drop"
  ) %>%
  arrange(desc(prop_sign_match), desc(n_overlap))

dir_summary
```