---
title: "Liver Proteomics (TODAY + TeenLABS) — Handoff README"
author: "Hailey Hampson"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

---

# Project Overview

This project investigates associations between circulating proteomics (SomaScan platform) and liver enzyme trajectories (ALT and AST) in two youth-onset obesity/diabetes cohorts:

- **TODAY** (Treatment Options for type 2 Diabetes in Adolescents and Youth) — a multi-center RCT of youth-onset T2D; provides longitudinal proteomics and ALT/AST from baseline (~M00) through ~10-year follow-up (~P120)
- **TeenLABS** — a surgical weight loss cohort; provides proteomics before and after bariatric surgery

Key scientific questions:

1. Are baseline circulating protein levels associated with ALT/AST trajectories over time?
2. Are changes in protein levels associated with changes in ALT/AST?
3. Which proteins/pathways are most predictive of liver enzyme trajectories (via elastic net feature selection)?

**PI:** Petter Bjornstad
**Biostatisticians:** Hailey Hampson; Laura Pyle (elastic net Cox model approach)
**Core:** Laura Pyle — Biostatistics Core, University of Washington

---

# Directory Structure

## Code Repository

```
Liver Proteomics (TODAY + TeenLABS)/
├── README.Rmd                                      # This file
├── Lambda_Liver_Proteomics.Rmd                     # PRIMARY analysis script (run on Lambda HPC)
├── TODAY_TeenLABS_Proteomics_Lambda.Rmd            # TODAY + TeenLABS combined analysis (Lambda)
├── Lambda_Liver_Proteomics_Functions.R             # Custom functions and gene sets
├── Lambda_Libraries_Liver_Proteomics.R             # Package imports
└── Local/
    ├── Local_Liver_Proteomics.Rmd                  # Local version (loads from OneDrive)
    ├── TODAY somalogic elastic net models baseline_heh.Rmd  # Elastic net Cox models (Laura Pyle, adapted)
    └── Local.Rproj                                 # RStudio project
```

## Results Directory (OneDrive)

```
Hailey Hampson/1_Ongoing Projects/Liver Proteomics/
├── Results/
│   ├── ALT_AST_Overtime_TODAY.png              # ALT/AST trajectory plot (visit-based x-axis)
│   ├── ALT_AST_Overtime_TODAY_Months.png       # ALT/AST trajectory plot (months x-axis)
│   └── Model_Performance_Approach1/
│       ├── n_proteins_selected.png             # Number of proteins selected by elastic net
│       ├── r2_heatmap.png                      # R² heatmap across outcomes/approaches
│       └── r2_performance.png                  # Elastic net model R² performance summary
├── Results.docx                                # Full results write-up (most recent: April 2026)
├── Methods.docx                                # Methods section draft
└── Liver Enzymes and Proteomics Project.pptx  # Presentation slides
```

---

# Data Sources

All data are stored in the **Kopah S3 object storage** (`harmonized.dataset` bucket) at UW and accessed via Python/boto3 through `reticulate` on the Lambda HPC server.

## Key Data Files (Kopah S3)

| File | Description |
|------|-------------|
| `today_proteomics_longitudinal_log2.rds` | Longitudinal TODAY proteomics (log2-transformed, baseline + follow-up) |
| `today_proteomics_log2.rds` | Cross-sectional TODAY proteomics (baseline only, log2-transformed) |
| `CBL.csv` | TODAY clinical data — includes ALT, AST, and visit information |
| `gene_prot.RDS` | Gene-protein key: maps SomaScan `AptName` (seq.XXXX) IDs to gene symbols |

## Kopah S3 Access Setup

Data loading requires Lambda HPC access with Python/boto3 credentials:

```{r}
library(reticulate)
use_python("/home/hhampson/miniconda3/bin/python", required = TRUE)
library(jsonlite)

keys <- fromJSON("/home/[your_lambda_username]/keys.json")
boto3 <- reticulate::import("boto3")
session <- boto3$session$Session(
  aws_access_key_id     = keys$MY_ACCESS_KEY,
  aws_secret_access_key = keys$MY_SECRET_KEY
)
s3 <- session$client("s3", endpoint_url = "https://s3.kopah.uw.edu")

# Example: download proteomics data
bucket    <- "harmonized.dataset"
temp_file <- tempfile(fileext = ".rds")
s3$download_file(bucket, "today_proteomics_longitudinal_log2.rds", temp_file)
prot <- readRDS(temp_file)
```

For **local use**, load directly from OneDrive (see `Local/Local_Liver_Proteomics.Rmd`):

```{r}
dir.dat <- "/Users/[your_username]/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive"
prot <- readRDS(fs::path(dir.dat, "Liver project", "Data", "today_proteomics_longitudinal_log2.rds"))
```

---

# Proteomics Platform

- **Platform:** SomaScan (SomaLogic) — ~7,288 protein aptamers
- **Identifier format:** `seq.XXXX.XX` column names in the proteomics data frames
- **Transformation:** log2-transformed prior to analysis
- **Gene mapping:** `gene_prot.RDS` maps aptamer IDs to `EntrezGeneSymbol`; used to filter to proteins with matched gene symbols from the liver scRNAseq atlas

---

# Study Populations

## TODAY

- Multi-center RCT of youth-onset T2D
- **~376 participants** with proteomics; **~265 with both proteomics and ALT/AST** (201 with ALT and/or AST at relevant timepoints)
- **Baseline visit:** M00; **~10-year follow-up:** P120
- **Outcomes:** ALT, AST, ALT/AST ratio, NAFLD status
- Proteomics filtered to genes expressed in the liver scRNAseq atlas

## TeenLABS

- Bariatric surgery cohort (adolescent/young adult)
- Proteomics before and after surgical weight loss
- Analysis in `TODAY_TeenLABS_Proteomics_Lambda.Rmd` (in progress)

---

# Analysis Workflow

## 1. Data Loading & Formatting (`§3a` in main Rmd)

- Load longitudinal proteomics and clinical (ALT/AST) data from Kopah
- Merge proteomics with ALT/AST by participant ID
- Filter proteomics to proteins with gene symbols represented in the liver scRNAseq atlas

## 2. ALT/AST Trajectory Characterization

- Spaghetti plots of individual ALT and AST trajectories over time (saved to `Results/`)
- Derive a **slope variable** per person representing their ALT or AST trajectory (linear fit)
- Construct mean ALT/AST as a summary outcome

## 3. Dataset Construction (Three Analytical Approaches)

Three protein feature datasets are constructed for the elastic net:

| Dataset | Description | Dimensions |
|---------|-------------|------------|
| **A** — Baseline proteins | Baseline protein levels | 265 × 7,288 |
| **B** — Protein changes | Change from baseline to follow-up | 265 × 7,288 |
| **C** — Combined | Baseline + change features | 265 × 14,576 |

## 4. Elastic Net Feature Selection (`§3b` in main Rmd)

Elastic net regression (α = 0.5, 10-fold cross-validation) is used to select proteins predictive of ALT/AST outcomes. Two regularization thresholds are reported:

- **`lambda.min`** — minimizes CV error (more features selected)
- **`lambda.1se`** — 1 SE above minimum (more regularized, more stable)

Results include number of selected proteins and R² performance metrics, saved to `Results/Model_Performance_Approach1/`.

## 5. Linear Regression — Proteins ~ ALT/AST

Per-protein linear regression loops across all proteins for both crude and adjusted models:

| Model | Covariates |
|-------|-----------|
| Crude | Protein only |
| Adjusted | Protein + age + sex (+ other covariates as applicable) |

Outcomes: ALT, AST, ALT/AST ratio. Results annotated with gene names via `gene_prot.RDS`.

## 6. Mixed Models — Longitudinal Associations (`TODAY_TeenLABS_Proteomics_Lambda.Rmd`)

Linear mixed effects models associating baseline proteins with mean ALT and mean AST over time, accounting for repeated measures per participant.

## 7. GSEA

Gene set enrichment analysis on ranked protein associations (crude), using custom gene sets defined in `Lambda_Liver_Proteomics_Functions.R`, including:

- Senescence-associated secretory phenotype (SASP) genes
- Metabolic pathway gene sets (TCA cycle, glycolysis)

## 8. Other Outcomes

- **NAFLD status** — protein associations with NAFLD1 (mixed model)
- Additional liver outcomes as available

---

# Custom Gene Sets (`Lambda_Liver_Proteomics_Functions.R`)

Key gene sets defined in the functions file:

| Gene Set | Description |
|----------|-------------|
| `sens_genes` | Senescence / SASP gene panel (~100 genes, including IL6, TNF, MMP family, CXCL family, VEGFA, IGFBP family) |
| TCA cycle | Tricarboxylic acid cycle genes (from external CSV) |
| Glycolysis | Glycolytic pathway genes (from external CSV) |

---

# Key R Packages

| Package | Purpose |
|---------|---------|
| `glmnet` | Elastic net regularized regression |
| `lme4` / `lmerTest` | Linear mixed effects models |
| `PGEE` | Penalized generalized estimating equations |
| `tidyverse` | Data wrangling and visualization |
| `ggplot2` / `ggrepel` | Plotting, volcano plots |
| `reticulate` | Python/boto3 interface for Kopah S3 access |
| `jsonlite` | Load Kopah credentials from JSON |
| `fs` | File path management |
| `SomaDataIO` | SomaScan data I/O (used in Laura Pyle's elastic net script) |
| `limma` | Linear models for proteomics |

---

# Results Files Summary

| File | Description |
|------|-------------|
| `Results/ALT_AST_Overtime_TODAY.png` | Individual ALT/AST trajectories by visit |
| `Results/ALT_AST_Overtime_TODAY_Months.png` | Same plot with months on x-axis |
| `Results/Model_Performance_Approach1/r2_heatmap.png` | R² across protein datasets and outcomes |
| `Results/Model_Performance_Approach1/n_proteins_selected.png` | Proteins selected by elastic net |
| `Results/Model_Performance_Approach1/r2_performance.png` | Overall elastic net model performance |
| `Results.docx` | Full results write-up (April 2026) |
| `Methods.docx` | Methods draft |
| `Liver Enzymes and Proteomics Project.pptx` | Slide deck for presentations |

---

# Notes

- **Kopah access is required** to run the Lambda scripts — data cannot be loaded locally without either Kopah credentials or downloading files manually to OneDrive first. Use `Local/Local_Liver_Proteomics.Rmd` for local runs once data files are on OneDrive.
- The `gene_prot.RDS` key is essential for translating SomaScan `seq.XXXX` aptamer IDs to gene symbols. Always load this before filtering proteins.
- Proteins are filtered to those with gene symbols present in the liver scRNAseq atlas — this substantially reduces the feature space from the full SomaScan panel.
- The TeenLABS analysis in `TODAY_TeenLABS_Proteomics_Lambda.Rmd` is in progress; TODAY analyses are more complete.
- `TODAY somalogic elastic net models baseline_heh.Rmd` (in `Local/`) was originally written by Laura Pyle and adapted for this project — it focuses on Cox proportional hazards elastic net for time-to-event ALT/AST outcomes.

---

# Contact

- **Hailey Hampson** — hehampson@uw.edu (primary biostatistician)
- **Laura Pyle** — lpyle@uw.edu (Biostatistics Core lead, elastic net Cox approach)
- **Petter Bjornstad** — petter.bjornstad@cuanschutz.edu (PI)
