---
title: "RH2 and IMPROVE Proteomics for CVD"
author: "Ye Ji Choi"
format: html
format:
  revealjs:
    embed-resources: true
    scrollable: true
    transition: slide
    controls-layout: bottom-right
    menu: true
    toc: true
    toc-depth: 3
    fontsize: 15
    number-sections: true
---

# Which SomaScan 7k plasma proteins are associated with vascular function markers, adjusting for age, sex, and kidney function (eGFR)?

```{r setup, include=FALSE}
# ── Packages ──────────────────────────────────────────────────────────────────
library(jsonlite)
library(aws.s3)
library(knitr)
library(kableExtra)
library(tidyverse)
library(lmerTest)
library(emmeans)
library(broom)
library(broom.mixed)
library(ggbeeswarm)
library(lemon)
library(quantreg)
library(readxl)
library(lubridate)
library(cowplot)
library(ggpubr)
library(Hmisc)
library(arsenal)
library(corrplot)
library(pander)
```

```{r paths, include=FALSE}
# ── User-specific paths ──────────────────────────────────────────────────────
user <- Sys.info()[["user"]]

if (user == "choiyej") {
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  # root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") {
  root_path <- ""
  git_path  <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") {
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path  <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/Users/pylell/Library/CloudStorage/OneDrive-UW/keys.json")
} else {
  stop("Unknown user: please specify root path.")
}
```

```{r s3-setup, include=FALSE}
# ── S3 credentials ────────────────────────────────────────────────────────────
Sys.setenv(
  AWS_ACCESS_KEY_ID     = keys$MY_ACCESS_KEY,
  AWS_SECRET_ACCESS_KEY = keys$MY_SECRET_KEY,
  AWS_DEFAULT_REGION    = "",
  AWS_REGION            = "",
  AWS_S3_ENDPOINT       = "s3.kopah.uw.edu"
)

harm_dat <- read.csv(
  file.path(root_path, "Data Harmonization/Data Clean/soma_harmonized_dataset.csv"),
  na = ""
)

analytes <- read.csv(
  file.path(root_path, "Data Harmonization/Combined SomaScan/analytes.csv"),
  na = ""
)
```

# Helpers

```{r}
run_limma_assoc <- function(Y, meta, outcome, covars,
                            block = NULL, study_tag = "",
                            analytes_df = analytes) {
  stopifnot(ncol(Y) == nrow(meta))
  form <- as.formula(paste("~", outcome, "+",
                           paste(covars, collapse = " + ")))

  # Drop rows missing any model variable
  use <- stats::complete.cases(meta[, c(outcome, covars), drop = FALSE])
  Y    <- Y[, use, drop = FALSE]
  meta <- meta[use, , drop = FALSE]
  if (!is.null(block)) block <- block[use]

  design    <- model.matrix(form, data = meta)
  coef_name <- colnames(design)[2]   # outcome is first non-intercept term

  if (is.null(block)) {
    fit <- lmFit(Y, design)
  } else {
    corfit <- duplicateCorrelation(Y, design, block = block)
    fit    <- lmFit(Y, design, block = block,
                    correlation = corfit$consensus)
  }
  fit <- eBayes(fit, trend = TRUE, robust = TRUE)

  tt <- topTable(fit, coef = coef_name, number = Inf, sort.by = "none") %>%
    tibble::rownames_to_column("AptName") %>%
    dplyr::mutate(FDR     = p.adjust(P.Value, method = "BH"),
                  outcome = outcome,
                  study   = study_tag,
                  n_model = ncol(Y))

  keep_cols <- intersect(
    c("AptName","EntrezGeneSymbol","Target","TargetFullName","UniProt"),
    names(analytes_df))
  tt %>% dplyr::left_join(analytes_df[, keep_cols, drop = FALSE],
                          by = "AptName")
}

cor_and_plot_soma <- function(data = df, vars) {
  # Format matrix
  d <- data[, c(vars, "acr_u", "map", "gfr_raw_plasma", "gfr_bsa_plasma")]
  n <- c(
    analytes$EntrezGeneSymbol[match(vars, analytes$AptName)],
    "acr_u", "map", "gfr_raw_plasma", "gfr_bsa_plasma"
  )
  n[duplicated(n)] <- paste0(n[duplicated(n)], "(2)")
  names(d) <- n
  # Correlations
  M <- corr.test(data.matrix(d), normal = F, method = "spearman")
  corrplot(round(M$r, 3),
    p.mat = M$p, method = "circle", type = "lower", insig = "blank",
    addCoef.col = "white", diag = FALSE
  )
  # Print a nice looking table
  r <- data.frame(round(M$r, 3))
  emphasize.strong.cells(which(M$p <= 0.05, arr.ind = TRUE))
  cat("\n")
  pander(r, caption = "Coefficients")
  cat("\n")
}

plot_volcano_proteomics <- function(data, fc, p_col, title = NULL, x_axis, y_axis, file_suffix, p_thresh = 0.05,
                                    positive_text = "Positive with PWV", 
                                    negative_text = "Negative with PWV",
                                    formula = "PWV", legend_position = c(0.8, 0.9),
                                    text_size = 15, caption_size = 8.5,
                                    top_n = 20,
                                    output_base_path = NULL,
                                    geom_text_size = 3,
                                    arrow_padding = 0.09,
                                    arrow_text_padding = 0.14,
                                    legend_text_size = 9,
                                    caption_padding = 15,
                                    x_title_padding_t = 32) {
  set.seed(1)
  top_pos <- data %>%
    dplyr::filter(!!sym(fc) > 0 & !!sym(p_col) < p_thresh) %>%
    dplyr::arrange(!!sym(p_col))
  
  n_pos <- nrow(top_pos)
  
  top_pos_n <- top_pos %>%
    slice_head(n=top_n)
  
  top_neg <- data %>%
    dplyr::filter(!!sym(fc) < 0 & !!sym(p_col) < p_thresh) %>%
    dplyr::arrange(!!sym(p_col))
  
  n_neg <- nrow(top_neg)
  
  top_neg_n <- top_neg %>%
    slice_head(n=top_n)
  
  data <- data %>%
    dplyr::mutate(top_color = case_when(data$AptName %in% top_pos$AptName ~ "#f28482",
                                        data$AptName %in% top_neg$AptName ~ "#457b9d",
                                        TRUE ~ "#ced4da"),
                  top_size = if_else(data$AptName %in% c(top_pos$AptName, top_neg$AptName), 1.3, 1),
                  EntrezGeneSymbol = if_else(data$EntrezGeneSymbol == "", data$Target, data$EntrezGeneSymbol),
                  top_lab  = if_else(data$AptName %in% c(top_pos_n$AptName, top_neg_n$AptName), EntrezGeneSymbol, ""))
  
  # Max and min for annotation arrows
  max_fc <- max(data[[fc]], na.rm = TRUE)
  min_fc <- min(data[[fc]], na.rm = TRUE)
  
  # Get y-axis max for dynamic scaling
  y_max <- max(-log10(data[[p_col]]), na.rm = TRUE) * 1.1
  
  
  p <- ggplot(data, aes(x = !!sym(fc), y = -log10(!!sym(p_col)))) +
    geom_hline(yintercept = -log10(p_thresh), linetype = "dashed", color = "darkgrey") +
    geom_point(alpha = 0.5, aes(color = top_color, size = top_size)) +
    geom_label_repel(aes(label = top_lab, color = top_color),
                    size = geom_text_size, max.overlaps = Inf,
                    force = 6, segment.alpha = 0.3, segment.size = 0.3,
                    fontface = "bold",
                    fill = fill_alpha("white", 0.7),
                    label.size = 0) +
    labs(title = paste(title),
         x = paste(x_axis),
         y = paste(y_axis),
         caption = paste0("Formula: log2(protein) ~ ", formula, 
                          "\nTotal analyzed proteins n: ", nrow(data),
                          " | Positive n = ", n_pos, " | Negative n = ", n_neg)) +
    scale_size_continuous(range = c(1, 1.3)) + 
    scale_color_manual(values = c("#457b9d"="#457b9d", "#ced4da"="#ced4da", "#f28482"="#f28482")) +
    theme_minimal() +
    guides(color = "none", size = "none")  +
    annotate("segment", 
             x=max_fc/8, 
             xend=(max_fc*7)/8, 
             y=-y_max * arrow_padding,
             col="darkgrey", arrow=arrow(length=unit(0.2, "cm"))) +
    annotate("text", 
             x=mean(c(max_fc/8, (max_fc*7)/8)), 
             y=-y_max * arrow_text_padding, 
             label=positive_text,
             size=geom_text_size, color="#343a40") +
    annotate("segment", 
             x=min_fc/8, 
             xend=(min_fc*7)/8, 
             y=-y_max * arrow_padding,
             col="darkgrey", arrow=arrow(length=unit(0.2, "cm"))) +
    annotate("text", 
             x=mean(c(min_fc/8, (min_fc*7)/8)), 
             y=-y_max * arrow_text_padding, 
             label=negative_text,
             size=geom_text_size, color="#343a40") +
    scale_y_continuous(expand=c(0,0)) +
    coord_cartesian(ylim = c(0, y_max), clip="off") +
    theme(legend.title = element_blank(),
          panel.grid = element_blank(),
          text = element_text(size = text_size, family = "Arial"),
          title = element_text(size = legend_text_size, family = "Arial"),
          legend.position = legend_position,
          legend.justification = if(is.character(legend_position) && legend_position == "bottom") c(0.5, 0) else c(1, 1),
          legend.direction = "horizontal",
          legend.spacing.x = unit(0.05, 'cm'),
          plot.margin = margin(t = 10, r = 20, b = 25, l = 20),
          axis.title.x = element_text(margin = margin(t = x_title_padding_t)),
          plot.caption = element_text(size = caption_size, hjust = 0.5, margin = margin(t = caption_padding)),
          legend.margin = margin(t = 5, b = 5))
  ggsave(paste0(output_base_path, file_suffix, ".png"), plot = p, width = 7, height = 5)
  
  return(p)
}

```

# Config and column names

```{r config}
suppressPackageStartupMessages({
  library(limma)
  library(purrr)
  library(ggplot2)
  library(tibble)
  library(tidyr)
})

cfg <- list(
  # Study / visit labels
  rh2_study_label       = "RENAL-HEIRitage",
  improve_study_label   = "IMPROVE",
  improve_visit_col     = "visit",
  improve_visit_pre     = "baseline",
  improve_visit_3mo     = "3_months_post_surgery",
  improve_visit_12mo    = "12_months_post_surgery",

  # Covariates (change if needed)
  age_col               = "age",
  sex_col               = "sex",
  egfr_col              = "eGFR_fas_cr",     # CKD-EPI / FAS creatinine eGFR
  rh2_group_col         = "group",
  bmi_col               = "bmi",
  a1c_col               = "hba1c",
  sbp_col               = "sbp",
  weight_col            = "weight",

  # CMR 4D flow markers to test
  cmr_markers = c(
    "af_pwv",
    "af_meanwss",
    "af_high5pctwss"
  ),
  
  # Output paths
  derived_dir   = file.path(root_path, "RH2_IMPROVE_proteomics/data/derived"),
  results_dir   = file.path(root_path, "RH2_IMPROVE_proteomics/results"),
  figures_dir   = file.path(root_path, "RH2_IMPROVE_proteomics/figures"),
  fdr_primary   = 0.05,
  fdr_discovery = 0.10
)

dir.create(cfg$derived_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(cfg$results_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(cfg$figures_dir, showWarnings = FALSE, recursive = TRUE)
```

```{r diagnose-columns}
expected <- c(cfg$age_col, cfg$sex_col, cfg$egfr_col, cfg$rh2_group_col,
              cfg$bmi_col, cfg$a1c_col, cfg$sbp_col,
              cfg$weight_col, cfg$improve_visit_col)
missing_cols <- setdiff(expected, colnames(harm_dat))
if (length(missing_cols))
  message("Columns not found in harm_dat — update cfg: ",
          paste(missing_cols, collapse = ", "))

cfg$cmr_markers <- intersect(cfg$cmr_markers, colnames(harm_dat))
message("CMR markers present (", length(cfg$cmr_markers), "): ",
        paste(cfg$cmr_markers, collapse = ", "))

apt_cols <- grep("^seq\\.[0-9]+\\.[0-9]+$", colnames(harm_dat), value = TRUE)
message("Aptamer columns: ", length(apt_cols))
message("Studies present: ", paste(unique(harm_dat$study), collapse = " | "))
if (cfg$improve_visit_col %in% colnames(harm_dat))
  message("Visits present: ",
          paste(unique(harm_dat[[cfg$improve_visit_col]]), collapse = " | "))
```

# Renal-HEIRitage (RH2) analysis frame

```{r build-rh2}
rh2_df <- harm_dat %>%
  filter(study == cfg$rh2_study_label) %>%
  dplyr::summarise(
      across(
        where(negate(is.numeric)),
        ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))
      ),
      across(
        where(is.numeric),
        ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))
      ),
      .by = c(record_id)
    )

rh2_df <- rh2_df %>%
  rowwise() %>%
  dplyr::mutate(pwv_avg = mean(c(pwv_val_test1, pwv_val_test2, pwv_val_test3),
                               na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::filter(
    is.finite(pwv_avg),
    rowSums(!is.na(dplyr::across(dplyr::all_of(apt_cols)))) > 0
  ) %>%
  dplyr::mutate(
    sex   = factor(.data[[cfg$sex_col]]),
    group = factor(.data[[cfg$rh2_group_col]])
  )

if ("Lean Control" %in% levels(rh2_df$group))
  rh2_df$group <- relevel(rh2_df$group, ref = "Lean Control")

saveRDS(rh2_df, file.path(cfg$derived_dir, "rh2_df.rds"))
message("RH2 analysis frame: n = ", nrow(rh2_df))
```


# IMPROVE baseline (cross-sectional) frame

```{r build-improve-baseline}
imp_base_df <- harm_dat %>%
  filter(study == cfg$improve_study_label,
         .data[[cfg$improve_visit_col]] == cfg$improve_visit_pre) %>%
  dplyr::summarise(
      across(where(negate(is.numeric)),
             ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
      across(where(is.numeric),
             ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
      .by = c(record_id)
    ) %>%
  dplyr::mutate(sex = factor(.data[[cfg$sex_col]])) %>%
  dplyr::filter(
    rowSums(!is.na(dplyr::across(dplyr::all_of(apt_cols)))) > 0,
    rowSums(!is.na(dplyr::across(dplyr::all_of(cfg$cmr_markers)))) > 0
  )

saveRDS(imp_base_df, file.path(cfg$derived_dir, "imp_base_df.rds"))
message("IMPROVE baseline frame: n = ", nrow(imp_base_df))
# 
# 
# imp_base_df_soma <- harm_dat %>%
#   filter(study == cfg$improve_study_label,
#          .data[[cfg$improve_visit_col]] == cfg$improve_visit_pre) %>%
#   dplyr::summarise(
#       across(where(negate(is.numeric)),
#              ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
#       across(where(is.numeric),
#              ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
#       .by = c(record_id)
#     ) %>%
#   dplyr::mutate(sex = factor(.data[[cfg$sex_col]])) %>%
#   dplyr::filter(
#     rowSums(!is.na(dplyr::across(dplyr::all_of(apt_cols)))) > 0
#   )
# imp_base_df_cmr <- harm_dat %>%
#   filter(study == cfg$improve_study_label,
#          .data[[cfg$improve_visit_col]] == cfg$improve_visit_pre) %>%
#   dplyr::summarise(
#       across(where(negate(is.numeric)),
#              ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
#       across(where(is.numeric),
#              ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
#       .by = c(record_id)
#     ) %>%
#   dplyr::mutate(sex = factor(.data[[cfg$sex_col]])) %>%
#   dplyr::filter(
#     rowSums(!is.na(dplyr::across(dplyr::all_of(cfg$cmr_markers)))) > 0
#   )
# 
# imp_base_df_soma$record_id
# imp_base_df_cmr$record_id
```

# IMPROVE longitudinal frames

`imp_long_df` (long) is used by the mixed-model sensitivity (C2). `imp_delta_df` (wide) is used by the Δ-on-Δ primary (C1).

```{r build-improve-longitudinal}
imp_long_df <- harm_dat %>%
  filter(study == cfg$improve_study_label,
         .data[[cfg$improve_visit_col]] %in%
           c(cfg$improve_visit_pre, cfg$improve_visit_12mo)) %>%
  dplyr::summarise(
      across(where(negate(is.numeric)),
             ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
      across(where(is.numeric),
             ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
      .by = c(record_id, visit)
    ) %>%
  dplyr::mutate(
    visit = factor(visit,
                   levels = c(cfg$improve_visit_pre, cfg$improve_visit_12mo)),
    sex   = factor(.data[[cfg$sex_col]])
  )

# Keep only participants with proteomics at BOTH visits
ids_both <- imp_long_df %>%
  dplyr::filter(rowSums(!is.na(dplyr::across(dplyr::all_of(apt_cols)))) > 0) %>%
  dplyr::count(record_id) %>%
  dplyr::filter(n == 2) %>%
  dplyr::pull(record_id)

imp_long_df <- imp_long_df %>% dplyr::filter(record_id %in% ids_both)
saveRDS(imp_long_df, file.path(cfg$derived_dir, "imp_long_df.rds"))
message("IMPROVE long frame: n_subjects = ", length(ids_both),
        ", n_rows = ", nrow(imp_long_df))

# ── Build wide Δ frame (log2 proteomics so Δ = log2 fold-change) ────────────
covars_keep <- unique(c(cfg$age_col, cfg$sex_col, cfg$egfr_col, cfg$bmi_col,
                        cfg$a1c_col, cfg$sbp_col,  cfg$weight_col))
covars_keep <- intersect(covars_keep, colnames(imp_long_df))

prot_wide <- imp_long_df %>%
  dplyr::mutate(dplyr::across(dplyr::all_of(apt_cols),
                              ~ log2(pmax(.x, 1)))) %>%
  dplyr::select(record_id, visit, dplyr::all_of(apt_cols)) %>%
  tidyr::pivot_wider(id_cols = record_id, names_from = visit,
                     values_from = dplyr::all_of(apt_cols),
                     names_sep = "__")

other_wide <- imp_long_df %>%
  dplyr::select(record_id, visit,
                dplyr::all_of(c(cfg$cmr_markers, covars_keep))) %>%
  tidyr::pivot_wider(id_cols = record_id, names_from = visit,
                     values_from = dplyr::all_of(c(cfg$cmr_markers, covars_keep)),
                     names_sep = "__")

imp_delta_df <- dplyr::left_join(prot_wide, other_wide, by = "record_id")

for (a in apt_cols) {
  n12  <- paste0(a, "__", cfg$improve_visit_12mo)
  npre <- paste0(a, "__", cfg$improve_visit_pre)
  if (all(c(n12, npre) %in% names(imp_delta_df)))
    imp_delta_df[[paste0("d_", a)]] <- imp_delta_df[[n12]] - imp_delta_df[[npre]]
}
for (m in cfg$cmr_markers) {
  n12  <- paste0(m, "__", cfg$improve_visit_12mo)
  npre <- paste0(m, "__", cfg$improve_visit_pre)
  if (all(c(n12, npre) %in% names(imp_delta_df)))
    imp_delta_df[[paste0("d_", m)]] <- imp_delta_df[[n12]] - imp_delta_df[[npre]]
}
if (cfg$weight_col %in% covars_keep) {
  n12  <- paste0(cfg$weight_col, "__", cfg$improve_visit_12mo)
  npre <- paste0(cfg$weight_col, "__", cfg$improve_visit_pre)
  if (all(c(n12, npre) %in% names(imp_delta_df)))
    imp_delta_df$d_weight <- imp_delta_df[[n12]] - imp_delta_df[[npre]]
}

# Baseline covariates for Δ models
for (cv in c(cfg$age_col, cfg$sex_col, cfg$egfr_col, cfg$bmi_col)) {
  npre <- paste0(cv, "__", cfg$improve_visit_pre)
  if (npre %in% names(imp_delta_df))
    imp_delta_df[[paste0(cv, "_pre")]] <- imp_delta_df[[npre]]
}
imp_delta_df$sex_pre <- factor(imp_delta_df[[paste0(cfg$sex_col, "_pre")]])

saveRDS(imp_delta_df, file.path(cfg$derived_dir, "imp_delta_df.rds"))
message("IMPROVE Δ frame: n = ", nrow(imp_delta_df))
```

# Proteomics QC

```{r qc-aptamers}
apt_meta <- analytes %>% dplyr::filter(AptName %in% apt_cols)
for (f in c("ColCheck","Type","Organism"))
  if (!f %in% names(apt_meta)) apt_meta[[f]] <- NA

apt_keep <- apt_meta %>%
  dplyr::filter(is.na(ColCheck) | ColCheck == "PASS",
                is.na(Type)     | Type     == "Protein",
                is.na(Organism) | Organism == "Human") %>%
  dplyr::pull(AptName)

message("Aptamers before QC: ", length(apt_cols),
        " | after QC: ", length(apt_keep))

# Y = log2 RFU matrix (rows = aptamers, cols = samples)
build_Y <- function(df, apt_ids) {
  m <- t(as.matrix(df[, apt_ids, drop = FALSE]))
  m[m <= 0 | is.infinite(m)] <- NA
  m <- log2(m)
  colnames(m) <- as.character(df$record_id)
  rownames(m) <- apt_ids
  m
}

Y_rh2     <- build_Y(rh2_df,      apt_keep)
Y_impbase <- build_Y(imp_base_df, apt_keep)
Y_implong <- build_Y(imp_long_df, apt_keep)

# Sample-level QC (median log2 RFU within 3 SD of cohort median)
sample_qc <- function(Y) {
  meds <- apply(Y, 2, median, na.rm = TRUE)
  abs(meds - median(meds, na.rm = TRUE)) < 3 * sd(meds, na.rm = TRUE)
}
keep_rh2     <- sample_qc(Y_rh2)
keep_impbase <- sample_qc(Y_impbase)
keep_implong <- sample_qc(Y_implong)

message("RH2 kept: ", sum(keep_rh2), "/", length(keep_rh2),
        " | IMPROVE base kept: ", sum(keep_impbase), "/", length(keep_impbase),
        " | IMPROVE long kept: ", sum(keep_implong), "/", length(keep_implong))

Y_rh2       <- Y_rh2[,     keep_rh2,     drop = FALSE]
Y_impbase   <- Y_impbase[, keep_impbase, drop = FALSE]
Y_implong   <- Y_implong[, keep_implong, drop = FALSE]
rh2_df      <- rh2_df[     keep_rh2,     , drop = FALSE]
imp_base_df <- imp_base_df[keep_impbase, , drop = FALSE]
imp_long_df <- imp_long_df[keep_implong, , drop = FALSE]

# Δ matrix for C1 (aptamers already log2 from step 5)
d_apt_cols <- paste0("d_", apt_keep)
d_apt_cols <- d_apt_cols[d_apt_cols %in% names(imp_delta_df)]
Y_impdelta <- t(as.matrix(imp_delta_df[, d_apt_cols, drop = FALSE]))
rownames(Y_impdelta) <- sub("^d_", "", d_apt_cols)
colnames(Y_impdelta) <- as.character(imp_delta_df$record_id)

saveRDS(Y_rh2,      file.path(cfg$derived_dir, "Y_rh2.rds"))
saveRDS(Y_impbase,  file.path(cfg$derived_dir, "Y_impbase.rds"))
saveRDS(Y_implong,  file.path(cfg$derived_dir, "Y_implong.rds"))
saveRDS(Y_impdelta, file.path(cfg$derived_dir, "Y_impdelta.rds"))
```

# Transform vascular outcomes

```{r transform-outcomes}
logz  <- function(x) { x[x <= 0] <- NA; as.numeric(scale(log(x))) }
zonly <- function(x) as.numeric(scale(x))

rh2_df$pwv_z <- logz(rh2_df$pwv_avg)

transform_cmr <- function(df, markers) {
  for (m in markers) {
    if (!m %in% names(df)) next
    # OSI is a bounded ratio (0–0.5) — z-score on native scale; everything else log-z
    if (grepl("^osi", m)) df[[paste0(m, "_z")]] <- zonly(df[[m]])
    else                   df[[paste0(m, "_z")]] <- logz(df[[m]])
  }
  df
}
imp_base_df <- transform_cmr(imp_base_df, cfg$cmr_markers)

# Δ-marker: z-score raw Δ (do not log; Δ can be negative)
for (m in cfg$cmr_markers) {
  dm <- paste0("d_", m)
  if (dm %in% names(imp_delta_df))
    imp_delta_df[[paste0(dm, "_z")]] <- zonly(imp_delta_df[[dm]])
}

# QQ plots for sanity checks
pdf(file.path(cfg$figures_dir, "qq_outcomes.pdf"), width = 8, height = 6)
op <- par(mfrow = c(2, 3))
qqnorm(rh2_df$pwv_z, main = "RH2 log-PWV z"); qqline(rh2_df$pwv_z)
for (m in head(cfg$cmr_markers, 5)) {
  zcol <- paste0(m, "_z")
  if (zcol %in% names(imp_base_df)) {
    qqnorm(imp_base_df[[zcol]], main = paste("IMPROVE", m, "z"))
    qqline(imp_base_df[[zcol]])
  }
}
par(op); dev.off()
```

# Table 1 per analytic cohort

```{r table-one, results='asis'}
t1_covars <- intersect(
  c(cfg$age_col, cfg$sex_col, cfg$egfr_col, cfg$bmi_col, cfg$a1c_col, cfg$sbp_col),
  colnames(harm_dat)
)

tab_rh2 <- tableby(
  as.formula(paste(cfg$rh2_group_col, "~",
                   paste(c(t1_covars, "pwv_avg", "dbp", "map"), collapse = " + "))),
  data = rh2_df)
summary(tab_rh2, title = "Table 1 — RH2")

tab_impb <- tableby(
  as.formula(paste("~",
                   paste(c(t1_covars, cfg$cmr_markers), collapse = " + "))),
  data = imp_base_df)
summary(tab_impb, title = "Table 1 — IMPROVE baseline")

tab_impl <- tableby(
  as.formula(paste("visit ~",
                   paste(c(t1_covars, cfg$cmr_markers, cfg$weight_col),
                         collapse = " + "))),
  data = imp_long_df)
summary(tab_impl, title = "Table 1 — IMPROVE pre vs 12 mo")
```

```{r}
rh2_df %>%
  ggplot(aes(x = group, y = pwv_avg, fill = group)) +
  geom_boxplot(size = 1) +
  geom_jitter(width = 0.1) + 
  theme_minimal() +
  theme(text = element_text(size = 15),
        panel.grid = element_blank(),
        legend.position = "none",
        axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = NULL, y = "PWV (m/s)") +
  scale_fill_manual(values = c("Lean Control" = "#c9cba3",
                               "Obese Control" = "#ffe1a8",
                               "Type 2 Diabetes" = "#e26d5c"))
ggsave(file.path(cfg$figures_dir, "/rh2_pwv_box.png"), width = 3, height = 5)
```




# RH2 PWV on all RH2

```{r analysis-A}
stopifnot(identical(as.character(rh2_df$record_id), colnames(Y_rh2)))

metaA <- rh2_df %>% dplyr::transmute(
  pwv_z,
  age  = .data[[cfg$age_col]],
  sex,
  eGFR = .data[[cfg$egfr_col]],
  group)

resA <- run_limma_assoc(
  Y         = Y_rh2,
  meta      = metaA,
  outcome   = "pwv_z",
  covars    = c("age","sex","eGFR","group"),
  study_tag = "RH2_PWV")

saveRDS(resA, file.path(cfg$results_dir, "resA.rds"))
write.csv(resA, file.path(cfg$results_dir, "resA.csv"), row.names = FALSE)

message("RH2 PWV — aptamers at FDR<0.05: ",
        sum(resA$FDR < cfg$fdr_primary, na.rm = TRUE))
head(dplyr::arrange(resA, P.Value), 10)
```

```{r}
plot_volcano_proteomics(data = resA,
                        fc = "logFC", 
                        p_col = "P.Value",
                        x_axis = "logFC", 
                        y_axis = "P-value", 
                        file_suffix = "rh2_pwv",
                        formula = "PWV + age + sex + eGFR + group",
                        output_base_path = paste0(cfg$figures_dir, "/"))
```


# RH2 PWV on T2D/OB in RH2

```{r}
keep_ids <- rh2_df %>%
  dplyr::filter(group != "Lean Control") %>%
  dplyr::pull(record_id) %>%
  as.character()

rh2_df_sub <- rh2_df %>%
  dplyr::filter(as.character(record_id) %in% keep_ids) %>%
  dplyr::mutate(group = droplevels(group),
                group = relevel(group, ref = "Obese Control"))

Y_rh2_sub <- Y_rh2[, as.character(rh2_df_sub$record_id), drop = FALSE]
stopifnot(identical(colnames(Y_rh2_sub), as.character(rh2_df_sub$record_id)))

metaB <- rh2_df_sub %>% dplyr::transmute(
  pwv_z,
  age  = .data[[cfg$age_col]],
  sex,
  eGFR = .data[[cfg$egfr_col]],
  group)

resB <- run_limma_assoc(
  Y         = Y_rh2_sub,
  meta      = metaB,
  outcome   = "pwv_z",
  covars    = c("age","sex","eGFR", "group"),
  study_tag = "RH2_PWV_T2D_OB")

saveRDS(resB, file.path(cfg$results_dir, "resB.rds"))
write.csv(resB, file.path(cfg$results_dir, "resB.csv"), row.names = FALSE)

message("RH2 PWV — aptamers at FDR<0.05: ",
        sum(resB$FDR < cfg$fdr_primary, na.rm = TRUE))
head(dplyr::arrange(resB, P.Value), 10)
```

```{r}
plot_volcano_proteomics(data = resB,
                        fc = "logFC", 
                        p_col = "P.Value",
                        x_axis = "logFC", 
                        y_axis = "P-value", 
                        file_suffix = "rh2_pwv_t2dob",
                        formula = "PWV + age + sex + eGFR + group",
                        output_base_path = paste0(cfg$figures_dir, "/")) 
```

