---
title: "Kidney Organoid scRNA-seq"
subtitle: "Pseudobulk Limma-voom: T2D vs. Lean Control"
author: "Hailey Hampson"
date: today
format:
  revealjs:
    theme: simple
    slide-number: true
    toc: true
    toc-depth: 2
    toc-title: "Outline"
    number-sections: false
    transition: slide
    background-transition: fade
    code-fold: true
    code-tools: true
    fig-align: center
    fig-width: 9
    fig-height: 6
    self-contained: true
    scrollable: true
    smaller: true
    css: |
      .reveal .slide-number { font-size: 0.7em; }
      .reveal h1 { color: #001219; }
      .reveal h2 { color: #005f73; border-bottom: 2px solid #9b2226; padding-bottom: 4px; }
      .reveal h3 { color: #9b2226; }
      .reveal .toc { font-size: 0.85em; }
      .reveal table { font-size: 0.75em; }
      .caption { font-size: 0.7em; color: #555; }
      .reveal section[data-background-color="#005f73"] h1,
      .reveal section[data-background-color="#005f73"] h2,
      .reveal section[data-background-color="#001219"] h1,
      .reveal section[data-background-color="#001219"] h2 {
        color: #e9d8a6 !important;
        border-bottom-color: #ee9b00 !important;
      }
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(patchwork)
library(kableExtra)
library(limma)
library(edgeR)
library(Seurat)
library(fgsea)
library(msigdbr)
library(stringr)
library(htmltools)

# ── Paths ──────────────────────────────────────────────────────────────────────
dir_results <- "/home/hhampson/Results"

# ── Color palette ─────────────────────────────────────────────────────────────
pal <- list(
  ink_black       = "#001219",
  dark_teal       = "#005f73",
  dark_cyan       = "#0a9396",
  pearl_aqua      = "#94d2bd",
  vanilla_custard = "#e9d8a6",
  golden_orange   = "#ee9b00",
  burnt_caramel   = "#ca6702",
  rusty_spice     = "#bb3e03",
  oxidized_iron   = "#ae2012",
  brown_red       = "#9b2226"
)
```

# Background {background-color="#005f73" color="white"}

## Background

- Diabetic nephropathy — leading cause of end-stage renal disease in United States^1^
  - Youth-onset type 2 diabetes (T2D) shows more aggressive progression to Diabetic Kidney Disease (DKD) than adult^2^
- Hyperglycemia drives kidney damage through multiple pathways including oxidative stress, inflammation, metabolic dysregulation, and podocyte injury^3^
  - Cellular mechanisms of glucose-induced kidney damage remain incompletely understood
- Kidney organoids derived from patient cells offer a controlled environment to model diabetes-related pathology in vitro^4^
- Single-cell RNA sequencing enables identification of cell type-specific responses to diabetic conditions^5^

::: {.caption}
1. Rabkin. Diabetic nephropathy 2003 · 2. Dart et al. Diabetes Care 2012 · 3. Jha et al. Antioxid Redox Signal 2016 · 4. Nakanoh et al. Life (Basel) 2025 · 5. Deng et al. Diabetes 2025
:::

## Research Aims

**Part 1: Diabetes-Driven Transcriptional Profiles**

- **Aim 1.1:** Identify diabetes-driven transcriptional signatures in kidney organoids.
  - *Research Question 1.1:* How does gene expression differ between kidney organoids derived from youth with type 2 diabetes (T2D) and lean controls (LC)?
- **Aim 1.2:** Identify diabetes-driven transcriptional signatures in kidney biopsies.
  - *Research Question 1.2:* How does gene expression differ between kidney tissue biopsies from youth with T2D and LC?
- **Aim 1.3:** Compare diabetes-driven molecular profiles between kidney organoids and kidney biopsies.
  - *Research Question 1.3:* How do diabetes-related molecular profiles compare between kidney organoids and kidney biopsies?

## Methods

:::: {.columns}
::: {.column width="50%"}
- Participants enrolled in the CROCODILE study (N = 3) & RENAL-HEIR study (N = 3)
- Collected demographics, anthropometrics, kidney tissue biopsy & blood samples
- Kidney organoids derived from blood samples (iPSC → organoid, Freedman et al. 2018)
- Single-cell RNA sequencing (scRNAseq) on matched kidney biopsies & kidney organoids
- Organoids exposed to **15 mM** (physiologic) and **30 mM** (hyperglycemic) glucose
:::
::: {.column width="50%"}
**Statistical Approach**

| Step | Method |
|------|--------|
| Aggregation | Pseudobulk (`AggregateExpression`) |
| Normalization | TMM (edgeR) |
| Filtering | `filterByExpr` |
| DE | Limma-voom |
| Correction | BH (FDR) |
| GSEA | fgsea (Reactome) |

**Significance thresholds:** adj. p < 0.05, |logFC| > 1
:::
::::

# Methods {background-color="#005f73" color="white"}

## Methods Summary

**Pseudobulk Differential Expression (Limma-voom)**

- Counts aggregated per `record_id × cell_type` using `AggregateExpression()`
- TMM normalization applied to pseudobulk count matrix
- Genes filtered with `filterByExpr()` prior to modeling
- Model: `~ treatment + group` (treatment = glucose condition, group = T2D vs. LC)
- Moderated t-statistics via empirical Bayes (`eBayes`)
- Multiple testing correction: Benjamini-Hochberg (FDR)
- **Significance:** adj.P.Val < 0.05 AND |logFC| > 1

**Gene Set Enrichment Analysis (fgsea)**

- Genes ranked by moderated t-statistic from limma-voom
- Reactome pathways via msigdbr (C2:CP:REACTOME)
- Gene set size filter: 15–500 genes
- 10,000 permutations
- Significance: adj. p < 0.05

Analysis performed using R/Bioconductor · limma · edgeR · fgsea · Seurat
---




## Table 1 — Participant Demographics {.scrollable}

```{r table1}
t1_file <- file.path(dir_results, "table1_demographics.html")

if (file.exists(t1_file)) {
  htmltools::HTML(readLines(t1_file, warn = FALSE))
} else {
  cat("Table 1 not found. Run save_table1.R first to generate:", t1_file)
}
```

# Results {background-color="#005f73" color="white"}

## Cell Types Analyzed

```{r cell-type-table}
results_file <- file.path(dir_results, "limma_T2DvsLC_allcelltypes.csv")

if (file.exists(results_file)) {
  results_all <- read.csv(results_file)

  summary_tbl <- results_all %>%
    dplyr::group_by(cell_type) %>%
    dplyr::summarise(
      Total_Genes_Tested = dplyr::n(),
      Up_in_T2D          = sum(sig == "Up in T2D",   na.rm = TRUE),
      Down_in_T2D        = sum(sig == "Down in T2D", na.rm = TRUE),
      .groups            = "drop"
    ) %>%
    dplyr::mutate(Total_DEGs = Up_in_T2D + Down_in_T2D) %>%
    dplyr::arrange(dplyr::desc(Total_DEGs))

  kable(summary_tbl,
        col.names = c("Cell Type", "Genes Tested", "Up in T2D", "Down in T2D", "Total DEGs"),
        align = "lrrrr") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE) %>%
    row_spec(0, bold = TRUE, color = "white", background = pal$dark_teal)
} else {
  cat("Results file not found at:", results_file)
}
```

# Volcano Plots {background-color="#005f73" color="white"}

## All Cell Types — Overview

```{r volcano-all, fig.width=14, fig.height=10}
if (exists("results_all")) {

  make_volcano <- function(df, ct) {
    # top 10 significant UP genes
    top_up <- df %>%
      dplyr::filter(sig == "Up in T2D") %>%
      dplyr::arrange(adj.P.Val) %>%
      dplyr::slice_head(n = 10)

    # top 10 significant DOWN genes
    top_down <- df %>%
      dplyr::filter(sig == "Down in T2D") %>%
      dplyr::arrange(adj.P.Val) %>%
      dplyr::slice_head(n = 10)

    top_genes <- dplyr::bind_rows(top_up, top_down)

    # fallback if no sig genes
    if (nrow(top_genes) == 0) {
      top_genes <- df %>%
        dplyr::arrange(P.Value) %>%
        dplyr::slice_head(n = 10)
    }

    n_genes <- format(nrow(df), big.mark = ",")

    ggplot(df, aes(x = logFC, y = -log10(adj.P.Val), color = sig)) +
      geom_point(alpha = 0.5, size = 1.5) +
      geom_label_repel(
        data         = top_genes,
        aes(label    = gene),
        size         = 3,
        max.overlaps = 25,
        show.legend  = FALSE,
        box.padding  = 0.4,
        point.padding = 0.3,
        force        = 2
      ) +
      scale_color_manual(values = c(
        "Up in T2D"   = pal$brown_red,
        "Down in T2D" = pal$dark_teal,
        "NS"          = "grey70"
      )) +
      labs(title    = ct,
           subtitle = paste0(n_genes, " genes tested"),
           x        = "log2FC (T2D vs LC)",
           y        = "-log10(adj. p)",
           color    = NULL) +
      theme_bw(base_size = 11) +
      theme(legend.position = "top",
            plot.title      = element_text(face = "bold", size = 11, color = pal$ink_black),
            plot.subtitle   = element_text(size = 8, color = "#555"))
  }

  cell_types <- unique(results_all$cell_type)
  plot_list  <- lapply(cell_types, function(ct) {
    df <- results_all %>% dplyr::filter(cell_type == ct)
    make_volcano(df, ct)
  })
  names(plot_list) <- cell_types

  n_cols <- 3
  wrap_plots(plot_list, ncol = n_cols) +
    plot_annotation(
      title    = "Limma-voom: T2D vs. Lean Control by Cell Type",
      subtitle = "adj.P.Val < 0.05, |logFC| > 1  |  Red = Up in T2D, Teal = Down in T2D",
      theme    = theme(plot.title    = element_text(face = "bold", size = 16, color = pal$ink_black),
                       plot.subtitle = element_text(size = 11, color = "#555"))
    )
}
```

# GSEA — Reactome {background-color="#005f73" color="white"}

```{r gsea-load}
# ── Load pre-computed GSEA results ────────────────────────────────────────────
gsea_file <- file.path(dir_results, "gsea_reactome_T2DvsLC_allcelltypes.csv")

gsea_results_list <- list()

if (file.exists(gsea_file)) {
  gsea_all <- read.csv(gsea_file)
  gsea_results_list <- split(gsea_all, gsea_all$cell_type)
} else {
  cat("GSEA results file not found at:", gsea_file)
}
```

## GSEA Summary

```{r gsea-summary-table}
if (exists("gsea_results_list") && length(gsea_results_list) > 0) {

  gsea_summary <- dplyr::bind_rows(gsea_results_list) %>%
    dplyr::group_by(cell_type) %>%
    dplyr::summarise(
      Pathways_Tested = dplyr::n(),
      Sig_Up          = sum(padj < 0.05 & NES > 0, na.rm = TRUE),
      Sig_Down        = sum(padj < 0.05 & NES < 0, na.rm = TRUE),
      .groups         = "drop"
    ) %>%
    dplyr::mutate(Total_Sig = Sig_Up + Sig_Down) %>%
    dplyr::arrange(dplyr::desc(Total_Sig))

  kable(gsea_summary,
        col.names = c("Cell Type", "Pathways Tested", "Up in T2D", "Down in T2D", "Total Sig"),
        align = "lrrrr") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE) %>%
    row_spec(0, bold = TRUE, color = "white", background = pal$dark_teal)
}
```

## All Cell Types — GSEA Overview

```{r gsea-all, fig.width=14, fig.height=10}
if (exists("gsea_results_list") && length(gsea_results_list) > 0) {

  make_gsea_bar <- function(gsea_df, ct, n_show = 20) {
    sig_paths <- gsea_df %>%
      dplyr::filter(padj < 0.05)

    if (nrow(sig_paths) < 3) {
      sig_paths <- gsea_df %>%
        dplyr::arrange(pval) %>%
        dplyr::slice_head(n = n_show)
    } else {
      top_up <- sig_paths %>%
        dplyr::filter(NES > 0) %>%
        dplyr::arrange(padj) %>%
        dplyr::slice_head(n = ceiling(n_show / 2))

      top_down <- sig_paths %>%
        dplyr::filter(NES < 0) %>%
        dplyr::arrange(padj) %>%
        dplyr::slice_head(n = ceiling(n_show / 2))

      sig_paths <- dplyr::bind_rows(top_up, top_down)
    }

    if (nrow(sig_paths) == 0) {
      return(
        ggplot() +
          annotate("text", x = 0.5, y = 0.5, label = paste0(ct, "\nNo pathways"),
                   size = 3, color = "grey50") +
          theme_void() +
          labs(title = ct) +
          theme(plot.title = element_text(face = "bold", size = 9))
      )
    }

    sig_paths <- sig_paths %>%
      dplyr::mutate(
        pathway_short = gsub("^REACTOME_", "", pathway),
        pathway_short = gsub("_", " ", pathway_short),
        pathway_short = stringr::str_to_title(pathway_short),
        pathway_short = stringr::str_trunc(pathway_short, 45),
        direction     = ifelse(NES > 0, "Up in T2D", "Down in T2D")
      )

    ggplot(sig_paths, aes(x = reorder(pathway_short, NES), y = NES, fill = direction)) +
      geom_col(width = 0.7) +
      coord_flip() +
      scale_fill_manual(values = c("Up in T2D" = pal$brown_red, "Down in T2D" = pal$dark_teal)) +
      labs(title = ct, x = NULL, y = "NES", fill = NULL) +
      theme_bw(base_size = 7) +
      theme(
        legend.position = "top",
        plot.title      = element_text(face = "bold", size = 9, color = pal$ink_black),
        axis.text.y     = element_text(size = 6)
      )
  }

  gsea_plot_list <- lapply(names(gsea_results_list), function(ct) {
    make_gsea_bar(gsea_results_list[[ct]], ct, n_show = 10)
  })
  names(gsea_plot_list) <- names(gsea_results_list)

  n_cols <- 3
  wrap_plots(gsea_plot_list, ncol = n_cols) +
    plot_annotation(
      title    = "Reactome GSEA: T2D vs. Lean Control by Cell Type",
      subtitle = "Top pathways by adj. p-value  |  Red = Up in T2D, Teal = Down in T2D",
      theme    = theme(plot.title    = element_text(face = "bold", size = 14, color = pal$ink_black),
                       plot.subtitle = element_text(size = 10, color = "#555"))
    )
}
```

# Per Cell Type — Volcano + GSEA {background-color="#005f73" color="white"}

```{r per-celltype-combined, fig.width=14, fig.height=6, results='asis'}
if (exists("results_all") && exists("gsea_results_list") && exists("make_volcano") && exists("make_gsea_bar")) {

  for (ct in unique(results_all$cell_type)) {

    cat("\n\n## ", ct, " {.scrollable}\n\n", sep = "")

    # volcano
    p_volc <- make_volcano(results_all %>% dplyr::filter(cell_type == ct), ct)

    # gsea bar
    if (ct %in% names(gsea_results_list)) {
      p_gsea <- make_gsea_bar(gsea_results_list[[ct]], ct, n_show = 20) +
        labs(title = paste0("Reactome GSEA — ", ct)) +
        theme(axis.text.y = element_text(size = 7))
    } else {
      p_gsea <- ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = "No GSEA results",
                 size = 4, color = "grey50") +
        theme_void()
    }

    # side by side
    combined <- p_volc + p_gsea +
      plot_layout(widths = c(1, 1.2)) +
      plot_annotation(
        title = ct,
        theme = theme(plot.title = element_text(face = "bold", size = 14, color = pal$ink_black))
      )

    print(combined)
    cat("\n\n")
  }
}
```


# Treatment Effect (30mM vs 15mM) {background-color="#005f73" color="white"}
```{r trt-load}
# ── Load pre-computed treatment effect results ────────────────────────────────
trt_results_file <- file.path(dir_results, "limma_TreatmentEffect_allcelltypes.csv")
trt_gsea_file    <- file.path(dir_results, "gsea_reactome_TreatmentEffect_allcelltypes.csv")

if (file.exists(trt_results_file)) {
  results_all_trt <- read.csv(trt_results_file)
} else {
  cat("Treatment DEG results not found at:", trt_results_file)
}

gsea_results_list_trt <- list()
if (file.exists(trt_gsea_file)) {
  gsea_all_trt <- read.csv(trt_gsea_file)
  gsea_results_list_trt <- split(gsea_all_trt, gsea_all_trt$cell_type)
} else {
  cat("Treatment GSEA results not found at:", trt_gsea_file)
}
```

## Treatment Effect — DEG Summary
```{r trt-deg-table}
if (exists("results_all_trt")) {
  trt_summary <- results_all_trt %>%
    dplyr::group_by(cell_type) %>%
    dplyr::summarise(
      Total_Genes_Tested = dplyr::n(),
      Up_in_30mM         = sum(sig == "Up in 30mM",   na.rm = TRUE),
      Down_in_30mM       = sum(sig == "Down in 30mM", na.rm = TRUE),
      .groups            = "drop"
    ) %>%
    dplyr::mutate(Total_DEGs = Up_in_30mM + Down_in_30mM) %>%
    dplyr::arrange(dplyr::desc(Total_DEGs))

  kable(trt_summary,
        col.names = c("Cell Type", "Genes Tested", "Up in 30mM", "Down in 30mM", "Total DEGs"),
        align = "lrrrr") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE) %>%
    row_spec(0, bold = TRUE, color = "white", background = pal$dark_teal)
}
```

## Treatment Effect — Volcano Overview
```{r trt-volcano-all, fig.width=18, fig.height=14}
if (exists("results_all_trt")) {

  make_volcano_trt <- function(df, ct) {
    top_up <- df %>%
      dplyr::filter(sig == "Up in 30mM") %>%
      dplyr::arrange(adj.P.Val) %>%
      dplyr::slice_head(n = 10)

    top_down <- df %>%
      dplyr::filter(sig == "Down in 30mM") %>%
      dplyr::arrange(adj.P.Val) %>%
      dplyr::slice_head(n = 10)

    top_genes <- dplyr::bind_rows(top_up, top_down)

    if (nrow(top_genes) == 0) {
      top_genes <- df %>%
        dplyr::arrange(P.Value) %>%
        dplyr::slice_head(n = 10)
    }

    n_genes <- format(nrow(df), big.mark = ",")

    ggplot(df, aes(x = logFC, y = -log10(adj.P.Val), color = sig)) +
      geom_point(alpha = 0.5, size = 1.5) +
      geom_label_repel(
        data          = top_genes,
        aes(label     = gene),
        size          = 3,
        max.overlaps  = 25,
        show.legend   = FALSE,
        box.padding   = 0.4,
        point.padding = 0.3,
        force         = 2
      ) +
      scale_color_manual(values = c(
        "Up in 30mM"   = pal$brown_red,
        "Down in 30mM" = pal$dark_teal,
        "NS"           = "grey70"
      )) +
      labs(title    = ct,
           subtitle = paste0(n_genes, " genes tested"),
           x        = "log2FC (30mM vs 15mM)",
           y        = "-log10(adj. p)",
           color    = NULL) +
      theme_bw(base_size = 11) +
      theme(legend.position = "top",
            plot.title      = element_text(face = "bold", size = 11, color = pal$ink_black),
            plot.subtitle   = element_text(size = 8, color = "#555"))
  }

  cell_types_trt <- unique(results_all_trt$cell_type)
  plot_list_trt  <- lapply(cell_types_trt, function(ct) {
    df <- results_all_trt %>% dplyr::filter(cell_type == ct)
    make_volcano_trt(df, ct)
  })
  names(plot_list_trt) <- cell_types_trt

  n_cols <- 3
  wrap_plots(plot_list_trt, ncol = n_cols) +
    plot_annotation(
      title    = "Limma-voom: Treatment Effect (30mM vs 15mM) by Cell Type",
      subtitle = "adj.P.Val < 0.05, |logFC| > 1  |  Adjusted for T2D/LC group",
      theme    = theme(plot.title    = element_text(face = "bold", size = 16, color = pal$ink_black),
                       plot.subtitle = element_text(size = 11, color = "#555"))
    )
}
```

## Treatment Effect — GSEA Summary
```{r trt-gsea-table}
if (exists("gsea_results_list_trt") && length(gsea_results_list_trt) > 0) {

  trt_gsea_summary <- dplyr::bind_rows(gsea_results_list_trt) %>%
    dplyr::group_by(cell_type) %>%
    dplyr::summarise(
      Pathways_Tested = dplyr::n(),
      Sig_Up          = sum(padj < 0.05 & NES > 0, na.rm = TRUE),
      Sig_Down        = sum(padj < 0.05 & NES < 0, na.rm = TRUE),
      .groups         = "drop"
    ) %>%
    dplyr::mutate(Total_Sig = Sig_Up + Sig_Down) %>%
    dplyr::arrange(dplyr::desc(Total_Sig))

  kable(trt_gsea_summary,
        col.names = c("Cell Type", "Pathways Tested", "Up in 30mM", "Down in 30mM", "Total Sig"),
        align = "lrrrr") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE) %>%
    row_spec(0, bold = TRUE, color = "white", background = pal$dark_teal)
}
```

## Treatment Effect — GSEA Overview
```{r trt-gsea-all, fig.width=14, fig.height=10}
if (exists("gsea_results_list_trt") && length(gsea_results_list_trt) > 0) {

  make_gsea_bar_trt <- function(gsea_df, ct, n_show = 20) {
    sig_paths <- gsea_df %>%
      dplyr::filter(padj < 0.05)

    if (nrow(sig_paths) < 3) {
      sig_paths <- gsea_df %>%
        dplyr::arrange(pval) %>%
        dplyr::slice_head(n = n_show)
    } else {
      top_up <- sig_paths %>%
        dplyr::filter(NES > 0) %>%
        dplyr::arrange(padj) %>%
        dplyr::slice_head(n = ceiling(n_show / 2))

      top_down <- sig_paths %>%
        dplyr::filter(NES < 0) %>%
        dplyr::arrange(padj) %>%
        dplyr::slice_head(n = ceiling(n_show / 2))

      sig_paths <- dplyr::bind_rows(top_up, top_down)
    }

    if (nrow(sig_paths) == 0) {
      return(
        ggplot() +
          annotate("text", x = 0.5, y = 0.5, label = paste0(ct, "\nNo pathways"),
                   size = 3, color = "grey50") +
          theme_void() +
          labs(title = ct) +
          theme(plot.title = element_text(face = "bold", size = 9))
      )
    }

    sig_paths <- sig_paths %>%
      dplyr::mutate(
        pathway_short = gsub("^REACTOME_", "", pathway),
        pathway_short = gsub("_", " ", pathway_short),
        pathway_short = stringr::str_to_title(pathway_short),
        pathway_short = stringr::str_trunc(pathway_short, 45),
        direction     = ifelse(NES > 0, "Up in 30mM", "Down in 30mM")
      )

    ggplot(sig_paths, aes(x = reorder(pathway_short, NES), y = NES, fill = direction)) +
      geom_col(width = 0.7) +
      coord_flip() +
      scale_fill_manual(values = c("Up in 30mM" = pal$brown_red, "Down in 30mM" = pal$dark_teal)) +
      labs(title = ct, x = NULL, y = "NES", fill = NULL) +
      theme_bw(base_size = 7) +
      theme(
        legend.position = "top",
        plot.title      = element_text(face = "bold", size = 9, color = pal$ink_black),
        axis.text.y     = element_text(size = 6)
      )
  }

  gsea_plot_list_trt <- lapply(names(gsea_results_list_trt), function(ct) {
    make_gsea_bar_trt(gsea_results_list_trt[[ct]], ct, n_show = 10)
  })
  names(gsea_plot_list_trt) <- names(gsea_results_list_trt)

  n_cols <- 3
  wrap_plots(gsea_plot_list_trt, ncol = n_cols) +
    plot_annotation(
      title    = "Reactome GSEA: Treatment Effect (30mM vs 15mM) by Cell Type",
      subtitle = "Top pathways by adj. p-value  |  Red = Up in 30mM, Teal = Down in 30mM",
      theme    = theme(plot.title    = element_text(face = "bold", size = 14, color = pal$ink_black),
                       plot.subtitle = element_text(size = 10, color = "#555"))
    )
}
```

## Treatment Effect — Per Cell Type {.scrollable}
```{r trt-per-celltype, fig.width=14, fig.height=6, results='asis'}
if (exists("results_all_trt") && exists("gsea_results_list_trt") &&
    exists("make_volcano_trt") && exists("make_gsea_bar_trt")) {

  for (ct in unique(results_all_trt$cell_type)) {

    cat("\n\n## ", ct, " — Treatment Effect {.scrollable}\n\n", sep = "")

    p_volc <- make_volcano_trt(results_all_trt %>% dplyr::filter(cell_type == ct), ct)

    if (ct %in% names(gsea_results_list_trt)) {
      p_gsea <- make_gsea_bar_trt(gsea_results_list_trt[[ct]], ct, n_show = 20) +
        labs(title = paste0("Reactome GSEA — ", ct)) +
        theme(axis.text.y = element_text(size = 7))
    } else {
      p_gsea <- ggplot() +
        annotate("text", x = 0.5, y = 0.5, label = "No GSEA results",
                 size = 4, color = "grey50") +
        theme_void()
    }

    combined <- p_volc + p_gsea +
      plot_layout(widths = c(1, 1.2)) +
      plot_annotation(
        title = paste0(ct, " — 30mM vs 15mM"),
        theme = theme(plot.title = element_text(face = "bold", size = 14, color = pal$ink_black))
      )

    print(combined)
    cat("\n\n")
  }
}
```

# Interaction — Group × Treatment {background-color="#005f73" color="white"}
```{r interaction-load}
# ── Load pre-computed interaction results ─────────────────────────────────────
int_results_file <- file.path(dir_results, "limma_Interaction_allcelltypes.csv")
emm_results_file <- file.path(dir_results, "limma_InteractionModel_EMM_allcelltypes.csv")
gsea_int_file    <- file.path(dir_results, "gsea_reactome_Interaction_allcelltypes.csv")

results_all_int <- if (file.exists(int_results_file)) read.csv(int_results_file) else NULL
emm_all         <- if (file.exists(emm_results_file)) read.csv(emm_results_file) else NULL

gsea_int_list <- list()
if (file.exists(gsea_int_file)) {
  gsea_int_all  <- read.csv(gsea_int_file)
  gsea_int_list <- split(gsea_int_all, gsea_int_all$cell_type)
}
```

## Interaction — Summary
```{r int-summary-table}
if (!is.null(results_all_int)) {
  int_summary <- results_all_int %>%
    dplyr::group_by(cell_type) %>%
    dplyr::summarise(
      Total_Genes = dplyr::n(),
      Sig_Pos     = sum(sig == "Sig Positive", na.rm = TRUE),
      Sig_Neg     = sum(sig == "Sig Negative", na.rm = TRUE),
      .groups     = "drop"
    ) %>%
    dplyr::mutate(Total_Sig = Sig_Pos + Sig_Neg) %>%
    dplyr::arrange(dplyr::desc(Total_Sig))

  kable(int_summary,
        col.names = c("Cell Type", "Genes Tested", "Sig Positive", "Sig Negative", "Total Sig"),
        align = "lrrrr",
        caption = "Interaction term: does the T2D vs LC difference change with glucose treatment?") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE) %>%
    row_spec(0, bold = TRUE, color = "white", background = pal$dark_teal)
}
```

## Interaction GSEA — Overview

```{r int-gsea-all, fig.width=14, fig.height=10}
if (length(gsea_int_list) > 0) {

  make_gsea_bar_int <- function(gsea_df, ct, n_show = 20) {
    sig_paths <- gsea_df %>%
      dplyr::filter(padj < 0.05)

    if (nrow(sig_paths) < 3) {
      sig_paths <- gsea_df %>%
        dplyr::arrange(pval) %>%
        dplyr::slice_head(n = n_show)
    } else {
      top_up <- sig_paths %>%
        dplyr::filter(NES > 0) %>%
        dplyr::arrange(padj) %>%
        dplyr::slice_head(n = ceiling(n_show / 2))

      top_down <- sig_paths %>%
        dplyr::filter(NES < 0) %>%
        dplyr::arrange(padj) %>%
        dplyr::slice_head(n = ceiling(n_show / 2))

      sig_paths <- dplyr::bind_rows(top_up, top_down)
    }

    if (nrow(sig_paths) == 0) {
      return(
        ggplot() +
          annotate("text", x = 0.5, y = 0.5, label = paste0(ct, "\nNo pathways"),
                   size = 3, color = "grey50") +
          theme_void() +
          labs(title = ct) +
          theme(plot.title = element_text(face = "bold", size = 9))
      )
    }

    sig_paths <- sig_paths %>%
      dplyr::mutate(
        pathway_short = gsub("^REACTOME_", "", pathway),
        pathway_short = gsub("_", " ", pathway_short),
        pathway_short = stringr::str_to_title(pathway_short),
        pathway_short = stringr::str_trunc(pathway_short, 45),
        direction     = ifelse(NES > 0, "T2D > LC response", "LC > T2D response")
      )

    ggplot(sig_paths, aes(x = reorder(pathway_short, NES), y = NES, fill = direction)) +
      geom_col(width = 0.7) +
      coord_flip() +
      scale_fill_manual(values = c("T2D > LC response" = pal$brown_red,
                                    "LC > T2D response" = pal$dark_teal)) +
      labs(title = ct, x = NULL, y = "NES", fill = NULL) +
      theme_bw(base_size = 7) +
      theme(
        legend.position = "top",
        plot.title      = element_text(face = "bold", size = 9, color = pal$ink_black),
        axis.text.y     = element_text(size = 6)
      )
  }

  gsea_int_plots <- lapply(names(gsea_int_list), function(ct) {
    make_gsea_bar_int(gsea_int_list[[ct]], ct, n_show = 10)
  })
  names(gsea_int_plots) <- names(gsea_int_list)

  n_cols <- 3
  wrap_plots(gsea_int_plots, ncol = n_cols) +
    plot_annotation(
      title    = "Reactome GSEA on Interaction Term (Group × Treatment)",
      subtitle = "Positive NES = stronger glucose response in T2D · Negative NES = stronger in LC",
      theme    = theme(plot.title    = element_text(face = "bold", size = 14, color = pal$ink_black),
                       plot.subtitle = element_text(size = 10, color = "#555"))
    )
}
```

# Per Cell Type — GSEA + EMM Heatmap {background-color="#005f73" color="white"}

```{r int-per-celltype, fig.width=14, fig.height=8, results='asis'}
library(pheatmap)

if (!is.null(results_all_int) && !is.null(emm_all)) {

  for (ct in unique(results_all_int$cell_type)) {

    # ── Slide 1: GSEA bar plot ──────────────────────────────────────────────
    if (ct %in% names(gsea_int_list)) {
      cat("\n\n## ", ct, " — Interaction GSEA {.scrollable}\n\n", sep = "")

      p_gsea <- make_gsea_bar_int(gsea_int_list[[ct]], ct, n_show = 20) +
        labs(title = paste0("Interaction GSEA — ", ct)) +
        theme(axis.text.y = element_text(size = 8))
      print(p_gsea)

      cat("\n\n")
    }

    # ── Slide 2: EMM Heatmap ────────────────────────────────────────────────
    int_ct <- results_all_int %>%
      dplyr::filter(cell_type == ct)

    sig_genes <- int_ct %>%
      dplyr::filter(sig != "NS") %>%
      dplyr::arrange(adj.P.Val) %>%
      dplyr::pull(gene)

    if (length(sig_genes) < 10) {
      top_by_p <- int_ct %>%
        dplyr::filter(!gene %in% sig_genes) %>%
        dplyr::arrange(P.Value) %>%
        dplyr::slice_head(n = 30 - length(sig_genes)) %>%
        dplyr::pull(gene)
      sig_genes <- c(sig_genes, top_by_p)
    }

    sig_genes <- head(sig_genes, 50)

    emm_ct <- emm_all %>%
      dplyr::filter(cell_type == ct, gene %in% sig_genes)

    if (nrow(emm_ct) > 0) {
      cat("\n\n## ", ct, " — EMM Heatmap {.scrollable}\n\n", sep = "")

      emm_mat <- emm_ct %>%
        dplyr::select(gene, LC_15mM_EMM, LC_30mM_EMM, T2D_15mM_EMM, T2D_30mM_EMM) %>%
        tibble::column_to_rownames("gene") %>%
        as.matrix()

      colnames(emm_mat) <- c("LC 15mM", "LC 30mM", "T2D 15mM", "T2D 30mM")
      emm_scaled <- t(scale(t(emm_mat)))

      col_anno <- data.frame(
        Group   = c("LC", "LC", "T2D", "T2D"),
        Glucose = c("15mM", "30mM", "15mM", "30mM"),
        row.names = colnames(emm_scaled)
      )

      anno_colors <- list(
        Group   = c("LC" = pal$dark_teal, "T2D" = pal$brown_red),
        Glucose = c("15mM" = pal$pearl_aqua, "30mM" = pal$golden_orange)
      )

      n_genes_hm <- nrow(emm_scaled)
      fontsize_row <- if (n_genes_hm > 40) 5 else if (n_genes_hm > 25) 6 else 7

      n_sig <- sum(int_ct$sig != "NS")
      hm_subtitle <- if (n_sig > 0) {
        paste0(n_sig, " sig interaction genes + top by p-value (", n_genes_hm, " total)")
      } else {
        paste0("Top ", n_genes_hm, " genes by interaction p-value (none adj.P < 0.05)")
      }

      p_hm <- pheatmap(
        emm_scaled,
        cluster_cols      = FALSE,
        cluster_rows      = TRUE,
        show_rownames     = TRUE,
        show_colnames     = TRUE,
        fontsize_row      = fontsize_row,
        fontsize_col      = 10,
        annotation_col    = col_anno,
        annotation_colors = anno_colors,
        color             = colorRampPalette(c(pal$dark_teal, "white", pal$brown_red))(100),
        border_color      = NA,
        main              = paste0(ct, " — Estimated Marginal Means (z-scored)\n", hm_subtitle),
        silent            = TRUE
      )

      grid::grid.newpage()
      grid::grid.draw(p_hm$gtable)

      cat("\n\n")
    }
  }
}
```
