---
title: "Kidney scRNAseq Project — Handoff README"
author: "Hailey Hampson"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

---

# Project Overview

This project investigates early kidney dysfunction in youth-onset Type 2 Diabetes (T2D) using single-cell RNA sequencing (scRNAseq) from two complementary sample types:

- **Kidney biopsies** from participants with T2D and lean controls (LC)
- **Kidney organoids** derived from participants with T2D and LC, cultured under normal (15 mM) and high (30 mM) glucose concentrations

The central scientific questions are:

1. What transcriptomic differences exist between T2D and LC kidney cells (biopsy and organoid)?
2. Does high-glucose treatment (30 mM vs. 15 mM) recapitulate T2D-associated gene expression patterns in organoids?
3. Is there a glucose-treatment × diabetes-group interaction effect in organoid gene expression?

**PI:** Petter Bjornstad
**Biostatistician:** Hailey Hampson
**Core:** Laura Pyle — Biostatistics Core, University of Washington

---

# Directory Structure

## Data Directory (OneDrive)

All data are stored on the shared OneDrive drive:

```
OneDrive: Laura Pyle - Biostatistics Core Shared Drive/
└── Kidney scRNAseq Project/
    ├── Data/                        # Seurat objects, raw/harmonized data, gene set files
    ├── IPA Results/                 # Ingenuity Pathway Analysis outputs
    ├── Organoid Results/            # Figures, DE result tables, NEBULA outputs
    │   ├── NEBULA/                  # NEBULA mixed model results (biopsy + organoid)
    │   ├── Organoid Results/        # Final organoid figures and tables
    │   ├── Organoid Results 02_12/  # Intermediate results
    │   ├── Preliminary Results/     # Early exploratory results
    │   └── Results 01_28/           # Earlier dated results
    ├── Reports/                     # Word document reports for collaborators
    └── Results/Archive/             # Archived per-cell-type DE result CSVs
```

## Code Repository (Local / GitHub)

```
Kidney scRNA/Kidney scRNA/
├── README.Rmd                          # This file
├── Kidney_Organoid_Lambda_Clean.Rmd    # PRIMARY analysis script (run on Lambda HPC)
├── Kidney_Organoid_Lambda2.Rmd         # Secondary/updated analysis version
├── Kidney_Organoid_Local2.Rmd          # Local machine version (subset of analyses)
├── Organoid_Samples_Analysis_NEBULA_06_02.Rmd  # Standalone NEBULA analysis
├── Organoid.qmd                        # Presentation slides (Quarto/Reveal.js)
├── overflow.qmd                        # Overflow slides
├── Kidney_functions_sc.R               # Custom analysis functions
├── Celltype_Annotation_Function.R      # Cell type annotation pipeline
├── Rename_Seurat_Function.R            # Seurat object renaming utilities
├── Libraries.R                         # All package imports
├── neg_binom_mixed_model.R             # NEBULA model script for HPC
├── neg_binom_mixed_model.slurm         # SLURM job submission script (Hyak/Lambda)
└── Archive/                            # Older/deprecated scripts
```

---

# Data Files

## Key Seurat Objects

Set your data directory at the top of any analysis script:

```{r}
dir.dat <- "/Users/[your_username]/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Biostatistics Core Shared Drive/Kidney scRNAseq Project"
```

| File | Description |
|------|-------------|
| `Data/cleaned_biopsy.rds` | Cleaned, QC-filtered kidney biopsy Seurat object (subset) |
| `Data/cleaned_biopsy_full.rds` | Full cleaned biopsy Seurat object (all samples) |
| `Data/cleaned_organoid.rds` | Cleaned, QC-filtered organoid Seurat object |
| `Data/annotated_cleaned_organoid.rds` | Organoid object with final cell type annotations |
| `Data/so_kidney_sc_filtered_10_mito.rds` | Early-stage filtered SC object (≤10% mitochondrial reads) |
| `Data/Test_Seurat_Object_for_Laura.rds` | Small test object for QC/debugging |

## Metadata & Reference Files

| File | Description |
|------|-------------|
| `Data/harmonized_data_kidney_sc_all_metadata2.csv` | Full harmonized metadata across all samples |
| `Data/harmonized_dataset.csv` | Harmonized gene expression + metadata |
| `Data/hailey_subset_data.csv` | Small subset for development/testing |
| `Data/Biopsies_w_mrn_Oct3.xlsx` | Biopsy sample manifest with MRNs |
| `Data/KPMP_Celltype_Markers_High_Res.csv/.xlsx` | KPMP high-resolution cell type marker gene lists |
| `Data/KPMP_Celltype_Markers_Low_Res.csv/.xlsx` | KPMP low-resolution cell type marker gene lists |
| `Data/D29 Kidney Organoid Gene Markers.pdf` | D29 organoid gene marker reference (Beno lab) |

## Gene Set Files (`Data/gmt_files/`)

These `.gmt` files are used for GSEA with `fgsea` and `enrichR`:

| File | Source |
|------|--------|
| `c2.cp.kegg_legacy.v2024.1.Hs.symbols.gmt` | KEGG Legacy pathways (MSigDB v2024.1) |
| `c2.cp.kegg_medicus.v2024.1.Hs.symbols.gmt` | KEGG Medicus pathways (MSigDB v2024.1) |
| `c2.cp.reactome.v2024.1.Hs.symbols.gmt` | Reactome pathways (MSigDB v2024.1) |
| `c5.go.v2024.1.Hs.symbols.gmt` | Gene Ontology (MSigDB v2024.1) |
| `c8.all.v2025.1.Hs.symbols.gmt` | Cell type signature gene sets (MSigDB v2025.1) |

---

# Analysis Workflow

## 1. Data Loading & Preprocessing

- Raw data originated from the **Lambda HPC cluster** (UW) and **Kopah S3 object storage** (`scrna` bucket)
- Biopsy and organoid data were processed separately, then harmonized
- QC filtering: ≤10% mitochondrial reads, doublet removal
- Cell type annotation was performed using KPMP marker gene lists (high- and low-resolution) and Beno-lab D29 organoid markers

```{r}
# Load cleaned objects (adjust path to your machine)
biopsy   <- readRDS(file.path(dir.dat, "Data/cleaned_biopsy_full.rds"))
organoid <- readRDS(file.path(dir.dat, "Data/annotated_cleaned_organoid.rds"))
```

## 2. Cell Type Annotation

Cell type annotation is handled by `Celltype_Annotation_Function.R` and validated against:

- KPMP reference markers (high- and low-resolution)
- Beno-lab D29 organoid markers
- Manual review of UMAP clustering and marker gene expression

The organoid was filtered to **6 matched organoid samples** for the primary comparisons. Sample **IT_19** was removed as an outlier prior to final analysis (see `##e.` annotation section in the main script, line ~963).

## 3. Differential Expression

Three primary DE frameworks were used:

### NEBULA (Primary Method)

Mixed-effects negative binomial model accounting for subject-level random effects. Run on Lambda HPC using `neg_binom_mixed_model.R` + `neg_binom_mixed_model.slurm`.

```{r}
# NEBULA is run per cell type; results stored as CSVs in:
# Organoid Results/NEBULA/
# Results/Archive/
```

### Pseudobulk — Limma-voom

Pseudobulk aggregation by sample × cell type, followed by `limma-voom` DE. Used for both organoid and biopsy data.

### Pseudobulk — DESeq2

Alternative pseudobulk approach using `DESeq2` for validation.

## 4. Primary Comparisons

| Comparison | Data | Model |
|-----------|------|-------|
| T2D vs. LC | Organoid (pooled glucose) | NEBULA + Pseudobulk |
| T2D vs. LC × Glucose (interaction) | Organoid | NEBULA + Pseudobulk |
| 30 mM vs. 15 mM glucose | Organoid (pooled by group) | NEBULA + Pseudobulk |
| T2D vs. LC | Biopsy | NEBULA + Pseudobulk |

## 5. Gene Set Enrichment Analysis (GSEA)

GSEA is run using both `fgsea` (rank-based) and `enrichR` (ORA) with the following databases:

- GO Biological Process 2023
- KEGG 2021 Human
- Reactome Pathways 2024
- MSigDB Hallmark 2020
- Cell type databases: CellMarker 2024, Azimuth, PanglaoDB, Descartes

Results include volcano plots, pathway bar charts, and EMM (estimated marginal means) heatmaps.

---

# Running the Analysis

## Primary Script: `Kidney_Organoid_Lambda_Clean.Rmd`

This is the main analysis script. It is designed to run on the **Lambda HPC server** but can be adapted for local use by updating file paths.

**Steps to run:**

1. Update the `dir.dat` and `dir.results` path variables in the setup chunk to match your machine
2. Source the helper scripts:

```{r}
source("Kidney_functions_sc.R")
source("Rename_Seurat_Function.R")
source("Celltype_Annotation_Function.R")
```

3. Knit the document or run chunks interactively in RStudio

**Script sections:**

| Section | Description |
|---------|-------------|
| §1 — Set Up | Libraries, paths, helper functions |
| §2 — Pre-Processing & QC | Filtering, annotation, descriptive stats |
| §3 — Organoid Analysis | DE (NEBULA + pseudobulk), GSEA, EMM heatmaps |
| §4 — Biopsy Analysis | Visualization, DE, pseudobulk, GSEA |

## HPC Job Submission (NEBULA)

To run NEBULA models on the cluster:

```bash
sbatch neg_binom_mixed_model.slurm
```

Results are written to the `Organoid Results/NEBULA/` directory.

---

# Key R Packages

| Package | Purpose |
|---------|---------|
| `Seurat` | Single-cell data structures, clustering, UMAP |
| `nebula` | Negative binomial mixed models for scRNAseq DE |
| `limma` / `edgeR` | Pseudobulk DE (voom pipeline) |
| `DESeq2` | Pseudobulk DE (count-based) |
| `fgsea` | Fast gene set enrichment analysis |
| `enrichR` | Over-representation analysis (ORA) |
| `EnhancedVolcano` | Volcano plots |
| `speckle` | Cell type proportion testing |
| `pheatmap` | Heatmaps |
| `emmeans` | Estimated marginal means (interaction models) |
| `glmmTMB` | Generalized linear mixed models |
| `msigdbr` | MSigDB gene set retrieval in R |

All packages are loaded in `Libraries.R`. Install any missing packages with:

```{r}
# Bioconductor packages
BiocManager::install(c("limma", "edgeR", "fgsea", "EnhancedVolcano",
                       "SingleCellExperiment", "scran", "speckle"))

# GitHub packages
devtools::install_github("lhe17/nebula")
devtools::install_github("immunogenomics/presto")
```

---

# Reports & Outputs

Rendered Word reports are saved in `Reports/`:

| File | Description |
|------|-------------|
| `DECODE_Kidney_scRNAseq.docx` | DECODE study kidney scRNAseq report |
| `Kidney_scRNAseq_02_09.docx` | Report dated Feb 9, 2025 |
| `Kidney_scRNAseq_02_13.docx` | Report dated Feb 13, 2025 |
| `Pathway_Genes_Results.docx` | Pathway-level gene result summary |

IPA (Ingenuity Pathway Analysis) results are in `IPA Results/`, sourced from bulk DE output files in `IPA Results/` (e.g., `Bulk_T2D_LC.txt`).

---

# Notes & Known Issues

- The organoid Seurat object is filtered to **6 samples** for the primary analysis. Sample **IT_19** is removed as an outlier (see `##e.` in main script, `sc_org <- subset(sc_org, record_id != "IT_19")`). The full organoid object retains all samples.
- File paths in the main Rmd reference Lambda HPC paths; **update `dir.dat` and `dir.results`** before running locally.
- Several packages (e.g., `MAST`, `DoubletFinder`, `ideas`) are commented out in `Libraries.R` — these were used in earlier iterations and are not required for the current analysis.
- The `python` / `reticulate` setup at the top of the script is only needed for S3 data loading on Lambda; it can be skipped if loading `.rds` files directly from OneDrive.

---

# Contact

For questions about the analysis, contact:

- **Hailey Hampson** — hehampson@uw.edu (biostatistics, code)
- **Laura Pyle** — lpyle@uw.edu (Biostatistics Core lead)
- **Petter Bjornstad** — petter.bjornstad@cuanschutz.edu (PI)
