---
title: "Christine Limonte Autophagy in ATTEMPT"
author: "Ye Ji Choi"
format: docx
---

```{r echo = F}
library(dplyr)
library(Seurat)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
library(nebula)
library(openxlsx)
library(table1)
library(lme4)
library(lmerTest)
library(purrr)
library(tibble)
library(tidyr)
library(stringr)
library(Hmisc)
library(ggtext)
library(corrplot)
library(gtools)
library(foreach)
library(emmeans)
library(knitr)
library(kableExtra)
```

```{r echo = F}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else {
  stop("Unknown user: please specify root path for this user.")
}

Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)

```

```{r echo = F}
s3read_using_region <- function(FUN, ..., object, bucket, region = NULL, opts = NULL, filename = NULL) {
  if (missing(bucket)) {
    bucket <- get_bucketname(object)
  }
  object <- get_objectkey(object)
  
  tmp <- if (is.character(filename)) {
    file.path(tempdir(TRUE), filename)
  } else {
    tempfile(fileext = paste0(".", tools::file_ext(object)))
  }
  
  on.exit(unlink(tmp))
  
  # Add region to opts if provided
  if (!is.null(region)) {
    if (is.null(opts)) {
      opts <- list(region = region)
    } else {
      opts$region <- region
    }
  }
  
  if (is.null(opts)) {
    r <- save_object(bucket = bucket, object = object, file = tmp)
  } else {
    r <- do.call("save_object", c(list(bucket = bucket, 
                                       object = object, 
                                       file = tmp), opts))
  }
  
  return(FUN(tmp, ...))
}

s3write_using_region <- function(FUN, ..., object, bucket,
                                 region = NULL, opts = NULL, filename = NULL) {
  ext  <- if (!is.null(filename)) tools::file_ext(filename) else tools::file_ext(object)
  tmp  <- tempfile(fileext = if (nchar(ext) > 0) paste0(".", ext) else "")
  on.exit(unlink(tmp))
  FUN(..., file = tmp)
  args <- list(file = tmp, object = object, bucket = bucket)
  if (!is.null(region)) args$region <- region
  if (!is.null(opts))   args        <- c(args, opts)
  do.call(aws.s3::put_object, args)
}
```

```{r echo = F}
attempt_so <- s3readRDS(bucket = "attempt", object = 'cleaned_data/attempt_clean_so.rds', region = "")
attempt_so$elevated_albuminuria <- ifelse(attempt_so$emu_urine_acr_mean > 30, "Yes", "No")
```

```{r echo = F}
celltype_groups <- list(
  PT = c("PT-S1/S2", "PT-S3", "aPT"),
  TAL = c("C-TAL-1", "C-TAL-2", "aTAL", "dTAL"),
  PC = c("CCD-PC", "CNT-PC", "dCCD-PC", "M-PC", "tPC-IC"),
  IC = c("IC-A", "IC-B", "aIC"),
  DTL_ATL = c("DTL", "aDTL", "ATL"),   # grouped thin limbs
  DCT_CNT = c("DCT", "dDCT", "CNT"),   # grouped distal tubule/connecting tubule
  EC = c("EC-AVR", "EC-GC", "EC-PTC", "EC-AEA", "EC-LYM", "EC/VSMC"), 
  Immune = c("MAC", "MON", "cDC", "pDC", "CD4+ T", "CD8+ T", "B", "NK", "cycT"),
  # Immune_Myeloid = c("MAC", "MON", "cDC", "pDC"),
  # Immune_Lymphoid = c("CD4+ T", "CD8+ T", "B", "NK", "cycT"),
  VSMC_P_FIB = c("VSMC/P", "FIB"),
  POD = "POD",
  MC = "MC",                         # mesangial cells
  PEC = "PEC",                       # parietal epithelial cells
  Schwann = "SchwannCells",
  Other = c("non-specific")          # catchall
)

map_celltype_to_general <- function(celltype, celltype_groups) {
  for (group_name in names(celltype_groups)) {
    if (celltype %in% celltype_groups[[group_name]]) {
      return(group_name)
    }
  }
  return("Other")  # For any celltype not found in groups
}

attempt_so$KPMP_celltype_general <- sapply(attempt_so$KPMP_celltype, 
                                           map_celltype_to_general, 
                                           celltype_groups = celltype_groups)
```

```{r echo = F}
gene_lists <- s3read_using_region(FUN = read.xlsx,
                                  object = "data_clean/Autophagy_gene_protein_lists_CL.xlsx", 
                                  bucket = "scrna",
                                  sheet = "Gene lists",
                                  startRow = 3,colNames = TRUE,
                                  region = "")

#create gene lists from Christine's files
overall_autophagy_genes <- c(
  gene_lists$`Overall.autophagy-related.genes.(n=589)`[!is.na(gene_lists$`Overall.autophagy-related.genes.(n=589)`)]
)
autophagy_induction_genes <- c(
  gene_lists$`Autophagy.induction.genes.(n=20)`[!is.na(gene_lists$`Autophagy.induction.genes.(n=20)`)]
)
lysosomal_biogenesis_genes <- c(
  gene_lists$`Lysosomal.biogenesis.genes.(n=26)`[!is.na(gene_lists$`Lysosomal.biogenesis.genes.(n=26)`)]
)
tfeb_genes <- c(
  gene_lists$`TFEB.genes.(n=13)`[!is.na(gene_lists$`TFEB.genes.(n=13)`)]
)

#MTORC1 pathway genes from JIC paper Supplemental Table S6 (https://www.jci.org/articles/view/164486/sd/1)
mtorc_genes<-c("STRADA",
               "STRADB",
               "RPS6",
               "TSC2",
               "TSC1",
               "PRKAB1",
               "PRKAB2",
               "MTOR",
               "PPM1A",
               "RRAGB",
               "RRAGA",
               "RRAGD",
               "RHEB",
               "RPS6KB1",
               "RRAGC",
               "AKT1S1",
               "EIF4G1",
               "LAMTOR1",
               "LAMTOR3",
               "LAMTOR2",
               "LAMTOR5",
               "LAMTOR4",
               "EIF4B",
               "PRKAA2",
               "YWHAB",
               "CAB39L",
               "CAB39",
               "EIF4EBP1",
               "PRKAG2",
               "PRKAG3",
               "PRKAG1",
               "PRKAA1",
               "STK11",
               "RPTOR",
               "EEF2K",
               "MLST8",
               "AKT1",
               "AKT2",
               "EIF4E")

signature_gs <- c("overall_autophagy", "autophagy_induction", 
                  "lysosomal_biogenesis", "tfeb", "mtorc")
```

```{r echo = F}
# Helper functions
prepare_umap_geneset_metadata <- function(
    seurat_obj,
    geneset_scores,
    celltype_col = "KPMP_celltype_general",
    umap_reduction = "umap.harmony",
    color_palette = "Set3",
    custom_colors = NULL
) {
  
  # Determine cell type levels
  celltypes <- unique(seurat_obj@meta.data[[celltype_col]])
  n_celltypes <- length(celltypes)
  
  # Build color mapping
  if (!is.null(custom_colors)) {
    celltype_colors <- custom_colors
  } else {
    umap_colors <- colorRampPalette(brewer.pal(min(brewer.pal.info[color_palette, "maxcolors"], n_celltypes), color_palette))(n_celltypes)
    celltype_colors <- setNames(umap_colors, celltypes)
  }
  
  # Get UMAP embeddings
  umap_embed_cols <- colnames(seurat_obj@reductions[[umap_reduction]]@cell.embeddings)
  
  # Build metadata with UMAP embeddings and specified geneset scores
  metadata <- seurat_obj@meta.data %>%
    cbind(seurat_obj@reductions[[umap_reduction]]@cell.embeddings) %>%
    select(all_of(c(celltype_col, geneset_scores)), all_of(umap_embed_cols))
  
  # Rename UMAP columns for consistency
  colnames(metadata)[colnames(metadata) %in% umap_embed_cols] <- c("UMAP_1", "UMAP_2")
  
  # Compute cell type centers
  centers <- metadata %>%
    group_by(.data[[celltype_col]]) %>%
    summarise(x = median(UMAP_1),
              y = median(UMAP_2), .groups = "drop") %>%
    rename(!!celltype_col := 1)  # Ensure consistent naming
  
  list(
    metadata = metadata,
    centers = centers,
    celltype_colors = celltype_colors
  )
}

plot_geneset_umap <- function(
    geneset_score,     # single column name from metadata
    metadata,          # from prepare_umap_geneset_metadata
    centers,           # centers dataframe
    celltype_colors,   # named color vector
    celltype_col = "KPMP_celltype_general",
    score_threshold = NULL,  # optional threshold to highlight high-scoring cells
    save_to_s3 = FALSE,
    s3_prefix = "",
    s3_suffix = "",
    s3_bucket = "scrna",
    plot_height = 8,
    plot_width = 8
) {
  
  # Calculate score stats per cell type
  score_stats <- metadata %>%
    group_by(.data[[celltype_col]]) %>%
    summarise(
      mean_score = round(mean(.data[[geneset_score]], na.rm = TRUE), 3),
      median_score = round(median(.data[[geneset_score]], na.rm = TRUE), 3),
      n_cells = n(),
      .groups = "drop"
    ) %>%
    arrange(desc(mean_score))
  
  # Build caption with top cell types by mean score
  caption_text <- score_stats %>%
    mutate(
      color = celltype_colors[.data[[celltype_col]]],
      label = paste0("<span style='color:", color, "'>&#9679; </span>**", 
                    .data[[celltype_col]], "**: ", mean_score),
      group = ceiling(row_number() / 6)  # Group every 6 cell types
    ) %>%
    group_by(group) %>%
    summarise(line = paste(label, collapse = " | "), .groups = "drop") %>%
    pull(line) %>%
    paste(collapse = "<br><br>") %>%
    paste0("**Mean ", gsub("_", " ", 
                           gsub("_signature1", "", geneset_score)), " by cell type:**<br><br>", .)
  
  # Create the plot with custom diverging color scale
  p <- metadata %>%
    ggplot(aes(x = UMAP_1, y = UMAP_2, color = .data[[geneset_score]])) +
    geom_point(alpha = 0.6, size = 0.3) +
    scale_color_gradient2(
      low = "#89c2d9",     # Blue for negative/low values
      mid = "white",       # White at 0
      high = "#ee7674",    # Red for positive/high values
      midpoint = 0,
      name = "Score"
    ) +
    geom_text_repel(
      data = centers,
      aes(x = x, y = y, label = .data[[celltype_col]]),
      inherit.aes = FALSE,
      size = 3.5,
      fontface = "bold",
      color = "black",
      bg.color = "white",
      bg.r = 0.1,
      max.overlaps = Inf
    ) +
    theme_void() +
    theme(
      text = element_text(size = 12),
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      plot.caption = element_markdown(size = 10, hjust = 0),
      legend.position = "right"
    ) +
    labs(title = toupper(gsub("_", " ", 
                           gsub("_signature1", "", geneset_score))), caption = caption_text)
  
  if (save_to_s3) {
    s3write_using_region(
      p,
      FUN = ggsave,
      object = paste0(s3_prefix, geneset_score, "_GeneSetScorePlot", s3_suffix, ".png"),
      bucket = s3_bucket,
      region = "",
      height = plot_height,
      width = plot_width
    )
  }
  
  return(p)
}

run_module_score <- function(seurat_obj, features, name) {
  warning_message <- NULL
  
  withCallingHandlers(
    {
      seurat_obj <- AddModuleScore(
        seurat_obj,
        features = list(features),
        name = name
      )
    },
    warning = function(w) {
      warning_message <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    }
  )
  
  list(seurat_obj = seurat_obj, warning = warning_message)
}

correlation_p_value_matrix <- function(data, relevant_vars, n_cols, cor_method = "pearson") {
  # Filter relevant variables for the correlation matrix
  dat_correlation <- data %>%
    dplyr::select(all_of(relevant_vars))
  
  # Compute p-values 
  res2 <- Hmisc::rcorr(as.matrix(dat_correlation),
                       type = cor_method)
  corr_pval <- as.data.frame(res2$P) %>%
    dplyr::select(all_of(relevant_vars[1:n_cols]))
  # corr_pval = corr_pval[(n_cols + 1):nrow(corr_pval),]
  corr_pval[is.na(corr_pval)] <- 1
  return(as.matrix(corr_pval))
}
corr_plot_modified <- function(
    data, X, Y, 
    cor_method = "pearson", 
    adj_var = NULL, 
    dict = dict,
    method = "color",
    insig = "pch", 
    coef_col = NULL,
    pch = 4, 
    pch.col = "black",
    pch.cex = 0, 
    tl.cex = .8, 
    cl.cex = 0.8, 
    number.cex = 1,
    coef_text_cex = 0.9,   # NEW: size for printed coefficients via graphics::text
    star_text_cex = 0.9,   # NEW: size for p-value stars via graphics::text
    col = colorRampPalette(c("#0096c7", "white", "#d62828"))(200)
) {
  n_cols <- length(Y)
  M <- cor(y = subset(data, select = Y),
           x = subset(data, select = X),
           use = "pairwise.complete.obs",
           method = cor_method)
  
  colnames(M) <- as.list(dict[match(colnames(M), names(dict))])
  rownames(M) <- as.list(dict[match(rownames(M), names(dict))])
  
  if (!is.null(adj_var)) {
    x_vars <- rep(X, times = length(Y))
    x_coord <- rep(seq(1, length(Y)), each = length(X))
    y_vars <- rep(Y, each = length(X))
    y_coord <- rep(seq(length(X), 1), times = length(Y))
    lm_extracted <- data.frame(yName = character(0),
                               xName = character(0),
                               x = numeric(0),
                               y = numeric(0),
                               adj_var = character(0),
                               adj_x_coef = numeric(0),
                               adj_x_pval = numeric(0))
    for (i in 1:length(x_vars)) {
      if (x_vars[i] != y_vars[i]) {
        lm_formula <- as.formula(paste0(y_vars[i], "~", x_vars[i], "+", adj_var))
        lm_model <- lm(lm_formula, data = data)
        lm_summary <- summary(lm_model)
        lm_coef <- lm_summary$coefficient[x_vars[i], "Estimate"]
        lm_p_val <- lm_summary$coefficient[x_vars[i], "Pr(>|t|)"]
        lm_extracted <- rbind(lm_extracted,
                              data.frame(
                                yName = y_vars[i],
                                xName = x_vars[i],
                                x = x_coord[i],
                                y = y_coord[i],
                                adj_var = adj_var,
                                adj_x_coef = lm_coef,
                                adj_x_pval = lm_p_val))
      }
    }
  }
  
  correlation_p_value <- correlation_p_value_matrix(
    data, 
    relevant_vars = c(Y, X), 
    n_cols = n_cols, 
    cor_method = cor_method
  )
  
  corrplot(M,
           p.mat = correlation_p_value,
           method = method,
           tl.col = "black",
           tl.srt = 45,
           tl.cex = tl.cex,         # variable label size
           insig = insig,
           addCoef.col = coef_col,  # if non-NULL, corrplot prints numbers; size via number.cex
           addgrid.col = 'white',
           number.digits = 2,
           number.cex = number.cex, # on-cell number size (if using addCoef.col)
           pch.col = pch.col,
           pch = pch,
           pch.cex = pch.cex,
           col = col,
           cl.cex = cl.cex          # legend tick label size
  )$corrPos -> p1
  
  p1_sub  <- subset(p1, p.value <= 0.05)
  p1_sub2 <- subset(p1, p.value <= 0.05 & abs(corr) >= 0.7)
  
  # Custom overlays: control size with cex =
  if (nrow(p1) > 0) {
    graphics::text(p1$x, p1$y, sprintf("%.2f", p1$corr),
                   adj = c(0.5, 0), cex = coef_text_cex)
  }
  if (nrow(p1_sub) > 0) {
    graphics::text(p1_sub$x, p1_sub$y, stars.pval(p1_sub$p.value),
                   adj = c(0.5, 2), cex = star_text_cex)
  }
  if (nrow(p1_sub2) > 0) {
    graphics::text(p1_sub2$x, p1_sub2$y, sprintf("%.2f", p1_sub2$corr),
                   col = "white", adj = c(0.5, 0), cex = coef_text_cex)
    graphics::text(p1_sub2$x, p1_sub2$y, stars.pval(p1_sub2$p.value),
                   col = "white", adj = c(0.5, 2), cex = star_text_cex)
  }
  if (!is.null(adj_var)) {
    lm_extracted <- subset(lm_extracted, adj_x_pval <= 0.05)
    if (nrow(lm_extracted) > 0){
      graphics::rect(xleft = lm_extracted$x - 0.45, 
                     ybottom = lm_extracted$y - 0.45,
                     xright = lm_extracted$x + 0.45,
                     ytop = lm_extracted$y + 0.45)
    }
  }
}

run_lmer_signatures <- function(signatures, data, outcome_suffix = "_score") {
  
  lmer_results <- map_dfr(signatures, function(sig) {
    outcome <- paste0("mean_", sig, outcome_suffix)
    mod <- lmer(as.formula(paste(outcome, "~ visit * treatment + (1|subject_id)")),
                data = data)
    
    broom.mixed::tidy(mod, effects = "fixed", conf.int = TRUE) %>%
      mutate(signature = sig, .before = 1)
  })
  
  lmer_emm <- map_dfr(signatures, function(sig) {
    outcome <- paste0("mean_", sig, outcome_suffix)
    mod <- lmer(as.formula(paste(outcome, "~ visit * treatment + (1|subject_id)")),
                data = data)
    
    emm <- emmeans(mod, ~ treatment * visit)
    contrast(emm, interaction = c("revpairwise", "revpairwise")) %>%
      summary(infer = c(TRUE, TRUE)) %>%
      as.data.frame() %>%
      mutate(signature = sig, .before = 1)
  })
  
  summary_tbl <- lmer_emm %>%
    mutate(
      estimate_ci = sprintf("%.3f (%.3f, %.3f)", estimate, lower.CL, upper.CL),
      interaction_fdr = p.adjust(p.value, method = "BH")
    ) %>%
    select(signature, estimate_ci, p.value, interaction_fdr)
  
  list(
    fixed_effects = lmer_results,
    emm_contrasts = lmer_emm,
    summary       = summary_tbl
  )
}

plot_volcano <- function(data, fc, p_col, title = NULL, x_axis, y_axis, file_suffix, p_thresh = 0.05,
                         positive_text = "Positive with Dapagliflozin", 
                         negative_text = "Negative with Dapagliflozin",
                         formula = "group", legend_position = c(0.8, 0.9),
                         full_formula = F,
                         text_size = 15, caption_size = 8.5,
                         cell_type = "",
                         bucket = "scrna",
                         output_base_path = "Projects/Christine Limonte Autophagy/Figures/",
                         geom_text_size = 3,
                         arrow_padding = 0.09,
                         arrow_text_padding = 0.14,
                         legend_text_size = 9,
                         caption_padding = 15,
                         x_title_padding_t = 32,
                         genes_to_label= NULL,
                         volcano_force = 6,
                         volcano_box_padding = 0,
                         off_chart_threshold = 0.95,  # New parameter: proportion of y_max to consider "off chart"
                         off_chart_y_position = 0.85,  # New parameter: where to place off-chart labels
                         off_chart_arrow_length = 0.02) {  # New parameter: length of off-chart arrows
  
  set.seed(1)
  
  # Add epsilon for log transformation
  epsilon <- 1e-300
  
  # Calculate -log10(p) with epsilon
  data <- data %>%
    mutate(neg_log_p = -log10(!!sym(p_col) + epsilon))
  
  # Get y-axis max for dynamic scaling
  y_max <- max(data$neg_log_p, na.rm = TRUE) * 1.1
  y_cutoff <- y_max * off_chart_threshold  # Points above this are "off chart"
  
  top_pos <- data %>%
    dplyr::filter(!!sym(fc) > 0 & !!sym(p_col) < p_thresh) %>%
    dplyr::arrange(!!sym(p_col))
  
  n_pos <- nrow(top_pos)
  
  top_pos_n <- top_pos %>%
    filter(if (!is.null(genes_to_label)) Gene %in% genes_to_label else TRUE) %>%
    slice_head(n=20)
  
  top_neg <- data %>%
    dplyr::filter(!!sym(fc) < 0 & !!sym(p_col) < p_thresh) %>%
    dplyr::arrange(!!sym(p_col))
  
  n_neg <- nrow(top_neg)
  
  top_neg_n <- top_neg %>%
    filter(if (!is.null(genes_to_label)) Gene %in% genes_to_label else TRUE) %>%
    slice_head(n=20)
  
  # Identify off-chart genes
  off_chart_genes <- data %>%
    filter(Gene %in% c(top_pos_n$Gene, top_neg_n$Gene) & neg_log_p > y_cutoff) %>%
    mutate(
      is_positive = !!sym(fc) > 0,
      # Spread out x positions for off-chart labels
      x_position = if_else(is_positive,
                           !!sym(fc) + seq(from = 0.1, by = 0.2, length.out = n()),
                           !!sym(fc) - seq(from = 0.1, by = 0.2, length.out = n())),
      y_position = y_max * off_chart_y_position
    )
  
  # Separate on-chart and off-chart genes for labeling
  on_chart_genes <- c(top_pos_n$Gene, top_neg_n$Gene)[!c(top_pos_n$Gene, top_neg_n$Gene) %in% off_chart_genes$Gene]
  
  data <- data %>%
    dplyr::mutate(
      top_color = case_when(
        Gene %in% top_pos$Gene ~ "#f28482",
        Gene %in% top_neg$Gene ~ "#457b9d",
        TRUE ~ "#ced4da"
      ),
      top_size = if_else(Gene %in% c(top_pos$Gene, top_neg$Gene), 1.3, 1),
      # Only label on-chart genes normally
      top_lab  = if_else(Gene %in% on_chart_genes, Gene, ""),
      # Cap display values at y_cutoff for plotting
      display_neg_log_p = pmin(neg_log_p, y_cutoff)
    ) %>%
    filter(abs(!!sym(fc)) < 10)
  
  # Max and min for annotation arrows
  max_fc <- max(data[[fc]], na.rm = TRUE)
  min_fc <- min(data[[fc]], na.rm = TRUE)
  
  p <- ggplot(data, aes(x = !!sym(fc), y = display_neg_log_p)) +
    geom_hline(yintercept = -log10(p_thresh), linetype = "dashed", color = "darkgrey") +
    geom_point(alpha = 0.5, aes(color = top_color), size = 3) +
    # Regular labels for on-chart genes
    geom_label_repel(seed = 1, 
      data = filter(data, top_lab != ""),
      aes(label = top_lab, color = top_color),
      fontface = "bold",
      size = geom_text_size, max.overlaps = Inf,
      force = volcano_force, segment.alpha = 0.3, segment.size = 0.3,
      box.padding = volcano_box_padding,
      fill = fill_alpha("white", 0.7),     # background color of the label box
      label.size = 0
      # bg.color = "black",
      # bg.r = .15
    ) +
    # Add arrows for off-chart genes
    {if(nrow(off_chart_genes) > 0) {
      list(
        geom_segment(
          data = off_chart_genes,
          aes(x = !!sym(fc), y = y_cutoff * 0.98,
              xend = !!sym(fc), yend = y_cutoff - (y_max * off_chart_arrow_length)),
          arrow = arrow(length = unit(0.15, "cm"), type = "closed"),
          color = "black",
          size = 0.6
        ),
        geom_text(
          data = off_chart_genes,
          aes(x = x_position, y = y_position, label = Gene),
          size = geom_text_size * 0.9,
          hjust = if_else(off_chart_genes$is_positive, 0, 1),
          color = "black",
          fontface = "italic"
        ),
        # Add p-value annotation for off-chart genes
        geom_text(
          data = off_chart_genes,
          aes(x = x_position, y = y_position - (y_max * 0.03), 
              label = paste0("p=", format(!!sym(p_col), scientific = TRUE, digits = 2))),
          size = geom_text_size * 0.7,
          hjust = if_else(off_chart_genes$is_positive, 0, 1),
          color = "darkgrey"
        )
      )
    }} +
    labs(title = paste(title),
         x = paste(x_axis),
         y = paste(y_axis),
         caption = if (!is.null(cell_type) & !full_formula & !is.null(formula)) {
           paste0("Formula: ~ ", formula, " + (1|subject)", 
                  "\n\nCell type: ", 
                  cell_type, if(cell_type != "") " | " else "", 
                  "Positive n = ", n_pos, " | Negative n = ", n_neg,
                  if(nrow(off_chart_genes) > 0) paste0("\n", nrow(off_chart_genes), " gene(s) with p-values near zero (arrows indicate off-scale values)") else "")
         } else if(!is.null(cell_type) & full_formula) {
           bquote(
             atop(
               Expression[ij] == (beta[0] + b[0*i]) +
                 beta[1]*visit[ij] +
                 beta[2]*treatment[i] +
                 beta[3]*(visit[ij]*treatment[i]) +
                 epsilon[ij],
               "Cell type:" ~ .(cell_type) ~ "|" ~ 
                 "Positive n =" ~ .(n_pos) ~ "|" ~ 
                 "Negative n =" ~ .(n_neg) ~
                 .(if(nrow(off_chart_genes) > 0) paste0(" | ", nrow(off_chart_genes), " off-scale") else "")
             )
           )
         } else {
           paste0("\nCell type: ", 
                  cell_type, if(cell_type != "") " | " else "", 
                  "Positive n = ", n_pos, " | Negative n = ", n_neg,
                  if(nrow(off_chart_genes) > 0) paste0("\n", nrow(off_chart_genes), " gene(s) with p-values near zero (arrows indicate off-scale values)") else "")
         }) +
    scale_size_continuous(range = c(1, 1.3)) + 
    scale_color_manual(values = c("#457b9d"="#457b9d", "#ced4da"="#ced4da", "#f28482"="#f28482")) +
    # theme_minimal() +
    guides(color = "none", size = "none")  +
    annotate("segment", 
             x=max_fc/8, 
             xend=(max_fc*7)/8, 
             y=-y_max * arrow_padding,
             col="darkgrey", arrow=arrow(length=unit(0.2, "cm"))) +
    annotate("text", 
             x=mean(c(max_fc/8, (max_fc*7)/8)), 
             y=-y_max * arrow_text_padding, 
             label=positive_text,
             size=geom_text_size, color="#343a40") +
    annotate("segment", 
             x=min_fc/8, 
             xend=(min_fc*7)/8, 
             y=-y_max * arrow_padding,
             col="darkgrey", arrow=arrow(length=unit(0.2, "cm"))) +
    annotate("text", 
             x=mean(c(min_fc/8, (min_fc*7)/8)), 
             y=-y_max * arrow_text_padding, 
             label=negative_text,
             size=geom_text_size, color="#343a40") +
    scale_y_continuous(expand=c(0,0)) +
    coord_cartesian(ylim = c(0, y_max), clip="off") +
    theme(legend.title = element_blank(),
          panel.grid = element_blank(),
          text = element_text(size = text_size, family = "Arial"),
          title = element_text(size = legend_text_size, family = "Arial"),
          legend.position = legend_position,
          legend.justification = if(is.character(legend_position) && legend_position == "bottom") c(0.5, 0) else c(1, 1),
          legend.direction = "horizontal",
          legend.spacing.x = unit(0.05, 'cm'),
          plot.margin = margin(t = 10, r = 20, b = 25, l = 20),
          axis.title.x = element_text(margin = margin(t = x_title_padding_t)),
          plot.caption = element_text(size = caption_size, hjust = 0.5, margin = margin(t = caption_padding)),
          legend.margin = margin(t = 5, b = 5),
          panel.background = element_blank(),
          legend.background = element_blank())
  
  if (!is.null(output_base_path)) {
    s3write_using_region(FUN = ggsave,
                         object = paste0(output_base_path, file_suffix, "_nebula_volcano.png"), bg = "transparent",
                         plot = p, width = 7, height = 5,
                         region = "",
                         bucket = bucket)
  }
  return(p)
}

```

# Methods

Single-cell RNA sequencing data from the ATTEMPT cohort were used for this analysis (young adults with T1D randomized to dapagliflozin vs. placebo, biopsies collected at baseline and 16 weeks post treatment).

Five gene signatures were scored for each cell using gene lists: overall autophagy-related genes, autophagy induction genes, lysosomal biogenesis genes, and TFEB target genes (all provided by Dr. Limonte), as well as an mTORC1 activity signature derived from Table S6 of Schaub et al. (doi: 10.1172/JCI164486). Genes signatures scores were calculated using AddModuleScore in R. Genes absent from our dataset were excluded from their respective signatures (Table 0). Gene signatures were scored separately for all cell types and for PT cells only. 

Signature scores were summarized at the individual level by averaging across all cells and across PT cells only. Group comparisons were performed using the Wilcoxon rank-sum test for pre and post treatment visits separately. UMAPs are shown to visualize the scores across all the cell types and across PT subtypes. Mixed models were used to assess the visit*treatment interaction (differences in differences; DiD). Pairwise correlations between gene signature scores were assessed using Pearson's correlation coefficient. NEBULA was performed on overall cells and PT cells only, and the DiD effects have been summarized. 


# Compute signature scores

```{r echo = F}
#add christine gene signatures to seurat object:
# have to do this separately for all cells and PT only, because AddModuleScore derives the score relative to other cells in the pool

# first do in all cells:
for (sig in signature_gs) {
  
  result <- run_module_score(attempt_so, 
                             get(paste0(sig, "_genes")), 
                             paste0(sig, "_signature"))
  
  attempt_so <- result$seurat_obj
  assign(paste0("missing_genes_", sig), result$warning)
}

# second do in PT only cells:
attempt_pt <- subset(attempt_so, KPMP_celltype %in% c("PT-S1/S2", "PT-S3", "aPT"))
for (sig in signature_gs) {
  
  result <- run_module_score(attempt_pt, 
                             get(paste0(sig, "_genes")), 
                             paste0(sig, "_signature"))
  
  attempt_pt <- result$seurat_obj
  assign(paste0("missing_genes_", sig), result$warning)
}

features <- paste0(signature_gs, "_signature1")
```

```{r echo = F}
# Create long datasets; one row per cell:
attempt_so$elevated_albuminuria <- ifelse(attempt_so$emu_urine_acr_mean > 30, "Yes", "No")
attempt_pt$elevated_albuminuria <- ifelse(attempt_pt$emu_urine_acr_mean > 30, "Yes", "No")

df_allcells <- data.frame(
  subject_id = attempt_so$subject_id,
  treatment=attempt_so$treatment,
  visit = attempt_so$visit,
  elevated_albuminuria=attempt_so$elevated_albuminuria,
  overall_score = attempt_so$overall_autophagy_signature1,
  induction_score = attempt_so$autophagy_induction_signature1,
  lysosomal_score = attempt_so$lysosomal_biogenesis_signature1,
  tfeb_score = attempt_so$tfeb_signature1,
  mtorc_score = attempt_so$mtorc_signature1
)
df_ptcells <- data.frame(
  subject_id = attempt_pt$subject_id,
  treatment=attempt_pt$treatment,
  visit = attempt_pt$visit,
  elevated_albuminuria=attempt_pt$elevated_albuminuria,
  overall_score = attempt_pt$overall_autophagy_signature1,
  induction_score = attempt_pt$autophagy_induction_signature1,
  lysosomal_score = attempt_pt$lysosomal_biogenesis_signature1,
  tfeb_score = attempt_pt$tfeb_signature1,
  mtorc_score = attempt_pt$mtorc_signature1
)

# mean scores per patient dataset:
df.1.allcells <- df_allcells %>%
  dplyr::summarise(mean_overall_autophagy_score=mean(overall_score,na.rm=T),
                   mean_autophagy_induction_score=mean(induction_score,na.rm=T),
                   mean_lysosomal_biogenesis_score=mean(lysosomal_score,na.rm=T),
                   mean_tfeb_score=mean(tfeb_score,na.rm=T),
                   mean_mtorc_score=mean(mtorc_score,na.rm=T),
                   .by = c("subject_id", "visit", "treatment"))

df.1.ptcells <- df_ptcells %>%
  summarise(mean_overall_autophagy_score_ptcells=mean(overall_score,na.rm=T),
            mean_autophagy_induction_score_ptcells=mean(induction_score,na.rm=T),
            mean_lysosomal_biogenesis_score_ptcells=mean(lysosomal_score,na.rm=T),
            mean_tfeb_score_ptcells=mean(tfeb_score,na.rm=T),
            mean_mtorc_score_ptcells=mean(mtorc_score,na.rm=T),
            .by = c("subject_id", "visit", "treatment"))

# merge together: 
df.1 <- merge(df.1.allcells,
              df.1.ptcells,
              by=c("subject_id", "visit", "treatment"),
              all=T)
```

# Tables of gene signature scores at baseline

```{r echo = F}
#table of person level comparisons across visits:
pvalue_wcx <- function(x, ...) {
  # x is a list of vectors, one per visit
  p <- wilcox.test(x[[1]], x[[2]])$p.value
  p_formatted <- ifelse(p < 0.001, "<0.001", formatC(p, digits = 3, format = "f"))
  c("", p_formatted)
}
tab_visit_pre <- table1(~ mean_overall_autophagy_score +
                          mean_autophagy_induction_score +
                          mean_lysosomal_biogenesis_score +
                          mean_tfeb_score +
                          mean_mtorc_score +
                          mean_overall_autophagy_score_ptcells +
                          mean_autophagy_induction_score_ptcells +
                          mean_lysosomal_biogenesis_score_ptcells +
                          mean_tfeb_score_ptcells +
                          mean_mtorc_score_ptcells|factor(treatment),
                        data = subset(df.1, visit == "PRE"),
                        overall=F,
                        render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)",
                                            "Mean (SD)"="MEAN (SD)"),
                        extra.col = list("P-value" = pvalue_wcx))
kable(tab_visit_pre, digits = 2)
```

*p-values from Wilcoxon rank sum test. 

# Tables of gene signature scores at follow up

```{r echo = F}
tab_visit_post <- table1(~ mean_overall_autophagy_score +
                           mean_autophagy_induction_score +
                           mean_lysosomal_biogenesis_score +
                           mean_tfeb_score +
                           mean_mtorc_score +
                           mean_overall_autophagy_score_ptcells +
                           mean_autophagy_induction_score_ptcells +
                           mean_lysosomal_biogenesis_score_ptcells +
                           mean_tfeb_score_ptcells +
                           mean_mtorc_score_ptcells|factor(treatment),
                         data = subset(df.1, visit == "POST"),
                         overall=F,
                         render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)",
                                             "Mean (SD)"="MEAN (SD)"),
                         extra.col = list("P-value" = pvalue_wcx), digits = 2)
kable(tab_visit_post)
```

*p-values from Wilcoxon rank sum test. 

# Visualize in UMAPs (Overall cells)

```{r echo = F}
plot_df <- prepare_umap_geneset_metadata(
  seurat_obj = attempt_so,
  celltype_col = "KPMP_celltype_general",
  geneset_scores = features)

umap_list <- list()
for (f in features) {
  p <- plot_geneset_umap(
    geneset_score = f,
    metadata = plot_df$metadata,
    centers = plot_df$centers,
    celltype_colors = plot_df$celltype_colors,
    celltype_col = "KPMP_celltype_general",
    s3_prefix = "Projects/Christine Limonte Autophagy/Figures/",
    s3_suffix = "",
    save_to_s3 = T
  )
  umap_list[[f]] <- p
}

```

# Visualize in UMAPs (PT cells)

```{r echo = F}
# Limited to PT
plot_df_pt <- prepare_umap_geneset_metadata(
  seurat_obj = attempt_pt,
  celltype_col = "KPMP_celltype",
  geneset_scores = features)

umap_list_pt <- list()
for (f in features) {
  p <- plot_geneset_umap(
    geneset_score = f,
    metadata = plot_df_pt$metadata,
    centers = plot_df_pt$centers,
    celltype_colors = plot_df_pt$celltype_colors,
    celltype_col = "KPMP_celltype",
    s3_prefix = "Projects/Christine Limonte Autophagy/Figures/",
    s3_suffix = "_pt",
    save_to_s3 = T
  )
  umap_list_pt[[f]] <- p
}
```

# Mixed models emmeans (DiD)

# Mixed model: Overall cells

```{r echo = F}
res_overall <- run_lmer_signatures(signature_gs, df.1, outcome_suffix = "_score")

kable(res_overall$summary,
      col.names = c("Signature", "Estimate (95% CI)", "P-value", "FDR"),
      caption = "Overall autophagy scores",
      digits = 3)
```

# Mixed model: PT cells

```{r echo = F}
# PT only
res_pt <- run_lmer_signatures(signature_gs, df.1, outcome_suffix = "_score_ptcells")

kable(res_pt$summary,
      col.names = c("Signature", "Estimate (95% CI)", "P-value", "FDR"),
      caption = "PT cell autophagy scores",
      digits = 3)
```

# Pearson correlations on delta scores: overall cells

```{r echo = F}
# pearson correlation
mean_score <- paste0("mean_", signature_gs, "_score")

deltas <- df.1 %>%
  pivot_longer(
    cols = all_of(mean_score),
    names_to = "score",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = visit,
    values_from = value
  ) %>%
  group_by(score, subject_id) %>%
  fill(POST, PRE) %>%
  ungroup() %>%
  filter(!is.na(PRE)) %>%
  dplyr::mutate(delta = POST - PRE) %>%
  dplyr::select(subject_id, treatment, score, delta) %>%
  pivot_wider(
    names_from = score,
    values_from = delta,
    names_prefix = "delta_"
  )

delta_mean_score <- paste0("delta_", mean_score)
delta_mean_score_labs <- as.data.frame(
  setNames(
    list(
      "\u0394 Overall Autophagy",
      "\u0394 Autophagy Induction",
      "\u0394 Lysosomal Biogenesis",
      "\u0394 TFEB",
      "\u0394 MTORC"),
    delta_mean_score))

temp_file <- tempfile(fileext = ".png")
png(height = 1500, width = 1500,
    file = temp_file, res = 170)
corr_plot_modified(data = deltas,
                   X = delta_mean_score,
                   Y = delta_mean_score,
                   dict = delta_mean_score_labs)
dev.off()
```

# Pearson correlations on delta scores: PT cells

```{r echo = F}
put_object(
  file = temp_file,
  object = "Projects/Christine Limonte Autophagy/Figures/delta_pearson_heatmap_overall.png",
  bucket = "scrna",
  region = ""
)

# PT cells only
mean_score_pt <- paste0("mean_", signature_gs, "_score_ptcells")

deltas_pt <- df.1 %>%
  pivot_longer(
    cols = all_of(mean_score_pt),
    names_to = "score",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = visit,
    values_from = value
  ) %>%
  group_by(score, subject_id) %>%
  fill(POST, PRE) %>%
  ungroup() %>%
  filter(!is.na(PRE)) %>%
  dplyr::mutate(delta = POST - PRE) %>%
  dplyr::select(subject_id, treatment, score, delta) %>%
  pivot_wider(
    names_from = score,
    values_from = delta,
    names_prefix = "delta_"
  )

delta_mean_score_pt <- paste0("delta_", mean_score_pt)

delta_mean_score_labs_pt <- as.data.frame(
  setNames(
    list(
      "\u0394 Overall Autophagy",
      "\u0394 Autophagy Induction",
      "\u0394 Lysosomal Biogenesis",
      "\u0394 TFEB",
      "\u0394 MTORC"),
    delta_mean_score_pt))

temp_file <- tempfile(fileext = ".png")
png(height = 1500, width = 1500, 
    file = temp_file, res = 170)
corr_plot_modified(data = deltas_pt,
                   X = delta_mean_score_pt,
                   Y = delta_mean_score_pt,
                   dict = delta_mean_score_labs_pt)
dev.off()
put_object(
  file = temp_file,
  object = "Projects/Christine Limonte Autophagy/Figures/delta_pearson_heatmap_pt.png",
  bucket = "scrna",
  region = ""
)
```

# NEBULA results (DiD)

## NEBULA results: overall cells

```{r}
# nebula
nebula_compiled <- list()
for (sig in signature_gs) {
  message("Starting ", sig)
  sig_start_time <- Sys.time()
  
  genes <- get(paste0(sig, "_genes"))
  attempt_so_subset <- subset(attempt_so, features = genes)
  
  counts_subset <- round(GetAssayData(attempt_so_subset, layer = "counts"))
  genes_list    <- rownames(counts_subset)
  meta_subset   <- attempt_so_subset@meta.data
  
  nebula_res <- lapply(genes_list, function(g) {
    tryCatch({
      count_gene <- counts_subset[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_subset)
      data_gene  <- group_cell(
        count = count_gene,
        id    = meta_subset$subject_id,
        pred  = pred_gene
      )
      
      nebula(
        count      = data_gene$count,
        id         = data_gene$id,
        pred       = data_gene$pred,
        ncore      = 1,
        output_re  = TRUE,
        covariance = TRUE,
        reml       = 1,
        model      = "NBLMM"
      )
    }, error = function(e) {
      message(sprintf("Gene %s failed: %s", g, e$message))
      NULL
    })
  })
  
  names(nebula_res) <- genes_list
  nebula_res <- Filter(Negate(is.null), nebula_res)
  
  pct_nonconv <- (length(genes_list) - length(nebula_res)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, sig))
  
  # Extract the summary table from each gene's result
  nebula_res <- lapply(nebula_res, `[[`, "summary")
  
  # save the results
  s3saveRDS(nebula_res, 
            object = paste0("Projects/Christine Limonte Autophagy/Tables/attempt_nebula_", sig, ".rds"),
            bucket = "scrna",
            region = "")
  
  nebula_compiled[[sig]] <- nebula_res
  
  message(sprintf("%s took %.1f min",
                  sig,
                  as.numeric(difftime(Sys.time(), sig_start_time, units = "mins"))))
}
```

```{r echo = F}
# Compile all nebula results into one tidy dataframe
nebula_tidy <- map_dfr(nebula_compiled, function(sig_res) {
  imap_dfr(sig_res, function(gene_summary, gene_name) {
    as_tibble(gene_summary) %>%
      mutate(Gene = gene_name, .before = 1)
  })
}, .id = "signature")


nebula_interaction <- nebula_tidy %>%
  select(
    signature,
    Gene = Gene,
    logFC = `logFC_treatmentDapagliflozin:visitPOST`,
    se    = `se_treatmentDapagliflozin:visitPOST`,
    p     = `p_treatmentDapagliflozin:visitPOST`
  ) %>%
  group_by(signature) %>%
  mutate(fdr = p.adjust(p, method = "BH")) %>%
  ungroup()

s3write_using_region(nebula_interaction, FUN = write.csv,
                     object = "Projects/Christine Limonte Autophagy/Tables/attempt_nebula_interaction_compiled.csv",
                     bucket = "scrna", region = "",
                     row.names = FALSE)

nebula_summary <- nebula_interaction %>%
  group_by(signature) %>%
  summarise(
    n_genes_tested = n(),
    n_nominal_sig  = sum(p < 0.05, na.rm = TRUE),
    n_fdr_sig      = sum(fdr < 0.05, na.rm = TRUE),
    n_up           = sum(fdr < 0.05 & logFC > 0, na.rm = TRUE),
    n_down         = sum(fdr < 0.05 & logFC < 0, na.rm = TRUE),
    .groups = "drop"
  )

kable(nebula_summary,
      col.names = c("Signature", "Genes tested", "p < 0.05", "FDR < 0.05",
                    "↑ with Dapa", "↓ with Dapa"))

volcano_plots <- map(unique(nebula_interaction$signature), function(sig) {
  dat <- nebula_interaction %>% filter(signature == sig)
  
  plot_volcano(
    data = dat,
    fc = "logFC",
    p_col = "p",
    title = toupper(sig),
    x_axis = "LogFC (interaction term)",
    y_axis = "-log10(p)",
    file_suffix = paste0("nebula_interaction_", sig),
    cell_type = "all cells",
    formula = "visit * treatment",
    positive_text = "Positive with Dapagliflozin",
    negative_text = "Negative with Dapagliflozin",
  )
})
names(volcano_plots) <- unique(nebula_interaction$signature)

volcano_plots_compiled <- plot_volcano(
  data = nebula_interaction,
  fc = "logFC",
  p_col = "p",
  x_axis = "LogFC (interaction term)",
  y_axis = "-log10(p)",
  file_suffix = "nebula_interaction_",
  cell_type = "all cells",
  formula = "visit * treatment",
  positive_text = "Positive with Dapagliflozin",
  negative_text = "Negative with Dapagliflozin",
)
```

## NEBULA results: PT only

```{r}
# nebula
nebula_compiled_pt <- list()

for (sig in signature_gs) {
  message("Starting ", sig)
  sig_start_time <- Sys.time()
  
  genes <- get(paste0(sig, "_genes"))
  attempt_pt_subset <- subset(attempt_pt, features = genes)
  
  counts_subset <- round(GetAssayData(attempt_pt_subset, layer = "counts"))
  genes_list    <- rownames(counts_subset)
  meta_subset   <- attempt_pt_subset@meta.data
  
  nebula_res_pt <- lapply(genes_list, function(g) {
    tryCatch({
      count_gene <- counts_subset[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_subset)
      data_gene  <- group_cell(
        count = count_gene,
        id    = meta_subset$subject_id,
        pred  = pred_gene
      )
      
      nebula(
        count      = data_gene$count,
        id         = data_gene$id,
        pred       = data_gene$pred,
        ncore      = 1,
        output_re  = TRUE,
        covariance = TRUE,
        reml       = 1,
        model      = "NBLMM"
      )
    }, error = function(e) {
      message(sprintf("Gene %s failed: %s", g, e$message))
      NULL
    })
  })
  
  names(nebula_res_pt) <- genes_list
  nebula_res_pt <- Filter(Negate(is.null), nebula_res_pt)
  
  pct_nonconv <- (length(genes_list) - length(nebula_res_pt)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, sig))
  
  # Extract the summary table from each gene's result
  nebula_res_pt <- lapply(nebula_res_pt, `[[`, "summary")
  
  # save the results
  s3saveRDS(nebula_res_pt, 
            object = paste0("Projects/Christine Limonte Autophagy/Tables/attempt_nebula_", sig, "_pt.rds"),
            bucket = "scrna",
            region = "")
  
  nebula_compiled_pt[[sig]] <- nebula_res_pt
  
  message(sprintf("%s took %.1f min",
                  sig, as.numeric(difftime(Sys.time(), sig_start_time, units = "mins"))))

}
# 
# attempt_nebula_overall_autophagy_pt <- s3readRDS(object = "Projects/Christine Limonte Autophagy/Tables/attempt_nebula_overall_autophagy_pt.rds",
#                                                  bucket = "scrna",
#                                                  region = "")
# 
# nebula_compiled_pt[["overall_autophagy"]] <- attempt_nebula_overall_autophagy_pt
```

```{r echo = F}
# Compile all nebula results into one tidy dataframe
nebula_tidy_pt <- map_dfr(nebula_compiled_pt, function(sig_res) {
  imap_dfr(sig_res, function(gene_summary, gene_name) {
    as_tibble(gene_summary) %>%
      mutate(Gene = gene_name, .before = 1)
  })
}, .id = "signature")


nebula_interaction_pt <- nebula_tidy_pt %>%
  select(
    signature,
    Gene = Gene,
    logFC = `logFC_treatmentDapagliflozin:visitPOST`,
    se    = `se_treatmentDapagliflozin:visitPOST`,
    p     = `p_treatmentDapagliflozin:visitPOST`
  ) %>%
  group_by(signature) %>%
  mutate(fdr = p.adjust(p, method = "BH")) %>%
  ungroup()

s3write_using_region(nebula_interaction_pt, FUN = write.csv,
                     object = "Projects/Christine Limonte Autophagy/Tables/attempt_nebula_interaction_pt_compiled_pt.csv",
                     bucket = "scrna", region = "",
                     row.names = FALSE)

nebula_summary_pt <- nebula_interaction_pt %>%
  group_by(signature) %>%
  summarise(
    n_genes_tested = n(),
    n_nominal_sig  = sum(p < 0.05, na.rm = TRUE),
    n_fdr_sig      = sum(fdr < 0.05, na.rm = TRUE),
    n_up           = sum(fdr < 0.05 & logFC > 0, na.rm = TRUE),
    n_down         = sum(fdr < 0.05 & logFC < 0, na.rm = TRUE),
    .groups = "drop"
  )

kable(nebula_summary_pt,
      col.names = c("Signature", "Genes tested", "p < 0.05", "FDR < 0.05",
                    "↑ with Dapa", "↓ with Dapa"))

volcano_plots_pt <- map(unique(nebula_interaction_pt$signature), function(sig) {
  dat <- nebula_interaction_pt %>% filter(signature == sig)
  
  plot_volcano(
    data = dat,
    fc = "logFC",
    p_col = "p",
    title = toupper(sig),
    x_axis = "LogFC (interaction term)",
    y_axis = "-log10(p)",
    file_suffix = paste0("nebula_interaction_pt_", sig),
    cell_type = "PT",
    formula = "visit * treatment",
    positive_text = "Positive with Dapagliflozin",
    negative_text = "Negative with Dapagliflozin",
  )
})
names(volcano_plots_pt) <- unique(nebula_interaction_pt$signature)

volcano_plots_pt_compiled <- 
  plot_volcano(
    data = nebula_interaction_pt,
    fc = "logFC",
    p_col = "p",
    x_axis = "LogFC (interaction term)",
    y_axis = "-log10(p)",
    file_suffix = "nebula_interaction_pt_",
    cell_type = "PT",
    formula = "visit * treatment",
    positive_text = "Positive with Dapagliflozin",
    negative_text = "Negative with Dapagliflozin",
  )
```
