---
title: "Limonte autophagy gene look up"
author: "Ye Ji Choi & Kristen Miller"
format: html
editor_options: 
  chunk_output_type: inline
---
  
```{r}
library(dplyr)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggforce)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(msigdbr)
library(ggpubr)
library(table1)
library(nlme)
library(reshape2)
library(nebula)

```

```{r}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/pylell/keys.json")
} else if (user == "kristenmiller") {
  root_path <- "/Users/kristenmiller/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive"
  git_path <- "/Users/kristenmiller/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/kristenmiller/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}

source(file.path(git_path, "Renal HEIRitage/RH_RH2_IMPROVE_functions.R"))
source(file.path(git_path, "Renal HEIRitage/RH_RH2_IMPROVE_scRNA_functions.R"))

entrez_map <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys   = keys(org.Hs.eg.db, keytype = "ENTREZID"),
  columns = c("SYMBOL", "GENENAME", "ALIAS"),
  keytype = "ENTREZID"
)


s3write_using_region <- function(FUN, ..., object, bucket, region = NULL, opts = NULL, filename = NULL) {
  if (missing(bucket)) {
    bucket <- get_bucketname(object)
  }
  object <- get_objectkey(object)
  
  tmp <- if (is.character(filename)) {
    file.path(tempdir(TRUE), filename)
  } else {
    # if object has an extension, keep it; otherwise make a generic tmp
    ext <- tools::file_ext(object)
    if (nzchar(ext)) tempfile(fileext = paste0(".", ext)) else tempfile()
  }
  
  on.exit(unlink(tmp), add = TRUE)
  
  # Add region to opts if provided
  if (!is.null(region)) {
    if (is.null(opts)) {
      opts <- list(region = region)
    } else {
      opts$region <- region
    }
  }
  
  FUN(tmp, ...)
  
  if (is.null(opts)) {
    r <- put_object(file = tmp, bucket = bucket, object = object)
  } else {
    r <- do.call("put_object", c(list(file = tmp, bucket = bucket, object = object), opts))
  }
  
  return(invisible(r))
}

# read in gene lists from Christine's excel file:
gene_lists <- openxlsx::read.xlsx(file.path(root_path, "Limonte/Autophagy_gene_protein_lists_CL.xlsx"), sheet = "Gene lists",
                                 startRow = 3,colNames = TRUE)

# read in full pb90 seurat object:
pb90_so <- s3readRDS("data_clean/pb90_processed.rds", "scrna", region = "")

# remove obese controls:
pb90_sub<- subset(pb90_so, group %in% c("Lean Control", "Type 1 Diabetes", "Type 2 Diabetes"))
# only use baseline samples: 
pb90_sub<- subset(pb90_sub, visit %in% c("baseline"))

# save two objects, one for all cells, one for PT only
pb90_allcells<-pb90_sub
# only use proximal tubular cells:
pb90_ptcells<- subset(pb90_sub, KPMP_celltype %in% c("aPT", "PT-S1/S2", "PT-S3"))

```

```{r}
#create gene lists from Christine's files
overall_autophagy_genes <- c(
  gene_lists$`Overall.autophagy-related.genes.(n=589)`[!is.na(gene_lists$`Overall.autophagy-related.genes.(n=589)`)]
)
autophagy_induction_genes <- c(
  gene_lists$`Autophagy.induction.genes.(n=20)`[!is.na(gene_lists$`Autophagy.induction.genes.(n=20)`)]
)
lysosomal_biogenesis_genes <- c(
  gene_lists$`Lysosomal.biogenesis.genes.(n=26)`[!is.na(gene_lists$`Lysosomal.biogenesis.genes.(n=26)`)]
)
tfeb_genes <- c(
  gene_lists$`TFEB.genes.(n=13)`[!is.na(gene_lists$`TFEB.genes.(n=13)`)]
)

#MTORC1 pathway genes from JIC paper Supplemental Table S6 (https://www.jci.org/articles/view/164486/sd/1)
mtorc_genes<-c("STRADA",
"STRADB",
"RPS6",
"TSC2",
"TSC1",
"PRKAB1",
"PRKAB2",
"MTOR",
"PPM1A",
"RRAGB",
"RRAGA",
"RRAGD",
"RHEB",
"RPS6KB1",
"RRAGC",
"AKT1S1",
"EIF4G1",
"LAMTOR1",
"LAMTOR3",
"LAMTOR2",
"LAMTOR5",
"LAMTOR4",
"EIF4B",
"PRKAA2",
"YWHAB",
"CAB39L",
"CAB39",
"EIF4EBP1",
"PRKAG2",
"PRKAG3",
"PRKAG1",
"PRKAA1",
"STK11",
"RPTOR",
"EEF2K",
"MLST8",
"AKT1",
"AKT2",
"EIF4E")
```

```{r}
#Calculate module scores
#create a function that returns the updated seurat object plus the warning of which genes were not found
run_module_score <- function(seurat_obj, features, name) {
  warning_message <- NULL
  
  withCallingHandlers(
    {
      seurat_obj <- AddModuleScore(
        seurat_obj,
        features = list(features),
        name = name
      )
    },
    warning = function(w) {
      warning_message <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    }
  )
  
  list(seurat_obj = seurat_obj, warning = warning_message)
}

#add christine gene signatures to seurat object:
# have to do this separately for all cells and PT only, because AddModuleScore derives the score relative to other cells in the pool

#first do in all cells:
result <- run_module_score(pb90_allcells, overall_autophagy_genes, "overall_autophagy_signature")
pb90_allcells <- result$seurat_obj
missing_genes_overall_autophagy <- result$warning

result <- run_module_score(pb90_allcells, autophagy_induction_genes, "autophagy_induction_signature")
pb90_allcells <- result$seurat_obj
missing_genes_autophagy_induction <- result$warning

result <- run_module_score(pb90_allcells, lysosomal_biogenesis_genes, "lysosomal_biogenesis_signature")
pb90_allcells <- result$seurat_obj
missing_genes_lysosomal <- result$warning

result <- run_module_score(pb90_allcells, tfeb_genes, "tfeb_signature")
pb90_allcells <- result$seurat_obj
missing_genes_tfeb <- result$warning

result <- run_module_score(pb90_allcells, mtorc_genes, "mtorc_signature")
pb90_allcells <- result$seurat_obj

#second do in PT only cells:
result <- run_module_score(pb90_ptcells, overall_autophagy_genes, "overall_autophagy_signature")
pb90_ptcells <- result$seurat_obj

result <- run_module_score(pb90_ptcells, autophagy_induction_genes, "autophagy_induction_signature")
pb90_ptcells <- result$seurat_obj

result <- run_module_score(pb90_ptcells, lysosomal_biogenesis_genes, "lysosomal_biogenesis_signature")
pb90_ptcells <- result$seurat_obj

result <- run_module_score(pb90_ptcells, tfeb_genes, "tfeb_signature")
pb90_ptcells <- result$seurat_obj

result <- run_module_score(pb90_ptcells, mtorc_genes, "mtorc_signature")
pb90_ptcells <- result$seurat_obj


features<-c("overall_autophagy_signature1", "autophagy_induction_signature1",
                         "lysosomal_biogenesis_signature1","tfeb_signature1")
```

```{r}
#Create long datasets; one row per cell:
df_allcells <- data.frame(
  record_id = pb90_allcells$record_id,
  cohort=pb90_allcells$cohort,
  group = pb90_allcells$group,
  sglt2i_ever= pb90_allcells$sglt2i_ever,
  elevated_albuminuria=pb90_allcells$elevated_albuminuria,
  overall_score = pb90_allcells$overall_autophagy_signature1,
  induction_score = pb90_allcells$autophagy_induction_signature1,
  lysosomal_score = pb90_allcells$lysosomal_biogenesis_signature1,
  tfeb_score = pb90_allcells$tfeb_signature1,
  mtorc_score = pb90_allcells$mtorc_signature1
)
df_ptcells <- data.frame(
  record_id = pb90_ptcells$record_id,
  cohort=pb90_ptcells$cohort,
  group = pb90_ptcells$group,
  sglt2i_ever= pb90_ptcells$sglt2i_ever,
  elevated_albuminuria=pb90_ptcells$elevated_albuminuria,
  overall_score = pb90_ptcells$overall_autophagy_signature1,
  induction_score = pb90_ptcells$autophagy_induction_signature1,
  lysosomal_score = pb90_ptcells$lysosomal_biogenesis_signature1,
  tfeb_score = pb90_ptcells$tfeb_signature1,
  mtorc_score = pb90_ptcells$mtorc_signature1
)

#mean scores per patient dataset:
df.1.allcells<- df_allcells %>%
    group_by(record_id) %>%
  summarise(group=group[1],
            cohort=cohort[1],
            sglt2i_ever=sglt2i_ever[1],
            mean_overall_score=mean(overall_score,na.rm=T),
            mean_induction_score=mean(induction_score,na.rm=T),
            mean_lysomal_score=mean(lysosomal_score,na.rm=T),
            mean_tfeb_score=mean(tfeb_score,na.rm=T),
            mean_mtorc_score=mean(mtorc_score,na.rm=T))

df.1.ptcells<- df_ptcells %>%
    group_by(record_id) %>%
  summarise(mean_overall_score_ptcells=mean(overall_score,na.rm=T),
            mean_induction_score_ptcells=mean(induction_score,na.rm=T),
            mean_lysomal_score_ptcells=mean(lysosomal_score,na.rm=T),
            mean_tfeb_score_ptcells=mean(tfeb_score,na.rm=T),
            mean_mtorc_score_ptcells=mean(mtorc_score,na.rm=T))

#merge together: 
df.1<-merge(df.1.allcells,df.1.ptcells,by="record_id",all=T)
#df.t2d<-subset(df,df$group=="Type 2 Diabetes")

# grouping variables
# 1. diabetes vs reference
df.1$diabetes_reference<-ifelse(df.1$group%in%c("Lean Control"),"Lean Controls","Diabetes (T1 or T2)")
# 2. SGLT2i vs. no
df.1_t2d <- subset(df.1, group %in% c("Type 2 Diabetes"))
```

```{r}
#table of person level comparisons across groups:
pvalue_wcx <- function(x, ...) {
  # x is a list of vectors, one per group
  p <- wilcox.test(x[[1]], x[[2]])$p.value
  p_formatted <- ifelse(p < 0.001, "<0.001", formatC(p, digits = 3, format = "f"))
  c("", p_formatted)
}
tab_group<-table1(~ mean_overall_score+
                    mean_induction_score+
                    mean_lysomal_score+
                    mean_tfeb_score+
                    mean_overall_score_ptcells+
                    mean_induction_score_ptcells+
                    mean_lysomal_score_ptcells+
                    mean_tfeb_score_ptcells|factor(diabetes_reference),
                  data=df.1,
                  overall=F,
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)",
                                      "Mean (SD)"="MEAN (SD)"),
                  extra.col = list("P-value" = pvalue_wcx))
tab_group

tab_sgl<-table1(~ mean_overall_score+
                    mean_induction_score+
                    mean_lysomal_score+
                    mean_tfeb_score+
                    mean_overall_score_ptcells+
                    mean_induction_score_ptcells+
                    mean_lysomal_score_ptcells+
                    mean_tfeb_score_ptcells|factor(sglt2i_ever),
                  data=df.1_t2d,
                  overall=F,
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)",
                                      "Mean (SD)"="MEAN (SD)"),
                  extra.col = list("P-value" = pvalue_wcx))
tab_sgl

```

```{r}
### mixed model, random intercept for pt
### (sensitivity analysis for comparison by group at cellular level)
## this takes a long time for the full sample models, so I'm removing for now 
# mixed <- function(data, outcome, grouping_var){
#   # data<-df
#   # grouping_var<-"group_2"
#   # outcome<-"overall_score"
#   data <- as.data.frame(data)
#   form = as.formula(paste0(outcome,"~ ", grouping_var))
#   mod <- lme(as.formula(form),random=~1|record_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|record_id))
#   anova <- anova.lme(mod, type="marginal")
#   m <- emmeans(mod,c(grouping_var))
#   prs <-  pairs(m,adjust="tukey")
#   return(list(anova,m,prs))
# }
# df_allcells$group_2<-NA
# df_allcells$group_2[df_allcells$group=="Lean Control"]<-"control"
# df_allcells$group_2[df_allcells$group=="Type 1 Diabetes"]<-"diabetes"
# df_allcells$group_2[df_allcells$group=="Type 2 Diabetes"]<-"diabetes"
# 
# mod_overall_bygroup<-mixed(df_allcells,"overall_score","group_2")
# mod_induction_bygroup<-mixed(df_allcells,"induction_score","group_2")
# mod_lysomal_bygroup<-mixed(df_allcells,"lysosomal_score","group_2")
# mod_tfeb_bygroup<-mixed(df_allcells,"tfeb_score","group_2")
# 
# mod_overall_bysgl<-mixed(df_allcells.t2d,"overall_score","sglt2i_ever")
# mod_induction_bysgl<-mixed(df_allcells.t2d,"induction_score","sglt2i_ever")
# mod_lysomal_bysgl<-mixed(df_allcells.t2d,"lysosomal_score","sglt2i_ever")
# mod_tfeb_bysgl<-mixed(df_allcells.t2d,"tfeb_score","sglt2i_ever")

```

#### Methods
Single-cell RNA sequencing data from the Bjornstad/Pyle lab cohorts were used for this analysis. Obese control and follow-up samples were excluded, and cells were subset to proximal tubular epithelial cells for this preliminary analysis.

Five gene signatures were scored for each cell using gene lists: overall autophagy-related genes, autophagy induction genes, lysosomal biogenesis genes, and TFEB target genes (all provided by Dr. Limonte), as well as an mTORC1 activity signature derived from Table S6 of Schaub et al. (doi: 10.1172/JCI164486). Genes signatures scores were calculated using AddModuleScore in R. Genes absent from our dataset were excluded from their respective signatures (Table 0).

Signature scores were summarized at the individual level by averaging across all retained proximal tubular cells. Group comparisons were performed using the Wilcoxon rank-sum test for two comparisons of interest: (1) diabetes (T1D and T2D combined) versus healthy controls, and (2) SGLT2 inhibitor (SGLT2i) use versus no SGLT2i use. As a sensitivity analysis, cell-level comparisons were conducted using a generalized linear mixed model with a random intercept for participant to account for within-subject correlation. Pairwise correlations between gene signature scores were assessed using Pearson's correlation coefficient. 

*Table 0: absent genes*
- Overall Autophagy: `r missing_genes_overall_autophagy`
- Autophagy Induction: `r missing_genes_autophagy_induction`
- Lysosomal biogenesis: `r missing_genes_lysosomal`
- TFEB: `r missing_genes_tfeb`

**Table 1: gene signature score by diabetes status**
```{r echo=FALSE}
kable(tab_group)
```
*p-values from Wilcoxon rank sum test. P-values from mixed model comparing cell-level scores were similar (all p>0.05)

**Table 2: gene signature score by SGLT2i status**
```{r echo=FALSE}
kable(tab_sgl)
```
*p-values from Wilcoxon rank sum test. P-values from mixed model comparing cell-level scores were similar (all p>0.05)

```{r}
#### visuals of scores by disease group 
par(mfrow=c(2,2))
RidgePlot(pb90_allcells, 
          features = features,
          ncol = 2, 
          group.by="diabetes_reference")


#### visuals of scores by SGLT2i
RidgePlot(pb90_allcells, 
          features = features,
          ncol = 2, 
          group.by="sglt2i_ever")


```


```{r}
#correlation heatmap between various signatures:
# full sample: 
corr.data<-df.1[,c("mean_overall_score","mean_induction_score","mean_lysomal_score","mean_tfeb_score","mean_mtorc_score")]
cormat <- round(cor(corr.data),2)
# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
upper_tri <- get_upper_tri(cormat)
melted_cormat <- reshape2::melt(upper_tri, na.rm = TRUE)

ggheatmap<-ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))

# PT cells only: 
corr.data<-df.1[,c("mean_overall_score_ptcells","mean_induction_score_ptcells","mean_lysomal_score_ptcells","mean_tfeb_score_ptcells","mean_mtorc_score_ptcells")]
cormat <- round(cor(corr.data),2)
# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
upper_tri <- get_upper_tri(cormat)
melted_cormat <- reshape2::melt(upper_tri, na.rm = TRUE)

ggheatmap<-ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))

```

```{r}
##### NEBULA FOR GENE LEVEL ANALYSIS #####

#### Diabetes vs. Healthy Controls####

### ALL CELLS ###

# Load counts data from the object and round to nearest integer
# use counts_hvg to match pipeline
seurat_object_hvg<-pb90_allcells
counts_hvg <- round(GetAssayData(seurat_object_hvg, layer = "counts"))

# Create a list of genes from the row names of the counts matrix
genes_list <- c(overall_autophagy_genes,autophagy_induction_genes,lysosomal_biogenesis_genes,tfeb_genes)

# Make sure the exposure/independent/x variable (group variable) is a factor
seurat_object_hvg$group <- as.factor(ifelse(seurat_object_hvg$group=="Lean Control",0,1))

source("nebula_pipeline_onecall.R")
write.csv(full_results,file.path(root_path, "Limonte/nebula_all_diabetes.csv"),row.names=F)
# ---- Volcano Plot ----

# Create the volcano plot using ggplot
#unadjusted:
diabetes_allcell_volcano_plot_unadjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_unadjusted, color = color_unadjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "Diabetes vs. reference (unadjusted)",  # Main title
    subtitle = "All cells",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_unadjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
diabetes_allcell_volcano_plot_unadjusted
#FDR-adjusted:
diabetes_allcell_volcano_plot_adjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_adjusted, color = color_adjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "Diabetes vs reference (FDR ADJUSTED)",  # Main title
    subtitle = "All cells",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_adjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
diabetes_allcell_volcano_plot_adjusted

### PT CELLS ###

seurat_object_hvg<-pb90_ptcells
counts_hvg <- round(GetAssayData(seurat_object_hvg, layer = "counts"))

# Create a list of genes from the row names of the counts matrix
genes_list <- c(overall_autophagy_genes,autophagy_induction_genes,lysosomal_biogenesis_genes,tfeb_genes)

# Make sure the exposure/independent/x variable (group variable) is a factor
seurat_object_hvg$group <- as.factor(ifelse(seurat_object_hvg$group=="Lean Control",0,1))

source("nebula_pipeline_onecall.R")
write.csv(full_results,file.path(root_path, "Limonte/nebula_pt_diabetes.csv"),row.names=F)
# ---- Volcano Plot ----

# Create the volcano plot using ggplot
#unadjusted:
diabetes_ptcell_volcano_plot_unadjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_unadjusted, color = color_unadjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "Diabetes vs. Healthy Controls (unadjusted)",  # Main title
    subtitle = "PT Cells Only",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_unadjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
diabetes_ptcell_volcano_plot_unadjusted
#FDR-adjusted:
diabetes_ptcell_volcano_plot_adjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_adjusted, color = color_adjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "Diabetes vs. Healthy Controls (FDR ADJUSTED)",  # Main title
    subtitle = "PT Cells Only",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_adjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
diabetes_ptcell_volcano_plot_adjusted
```

```{r}
##### NEBULA FOR GENE LEVEL ANALYSIS #####

#### SGLT2i vs. no SGLT2i####
### ALL CELLS ###
#### (2): SGLT2i vs. no; all cells ####
# Load counts data from the object and round to nearest integer
# use counts_hvg to match pipeline
pb90_allcells_t2d<-subset(pb90_allcells,pb90_allcells$group=="Type 2 Diabetes")
seurat_object_hvg<-pb90_allcells_t2d
counts_hvg <- round(GetAssayData(seurat_object_hvg, layer = "counts"))

# Create a list of genes from the row names of the counts matrix
genes_list <- c(overall_autophagy_genes,autophagy_induction_genes,lysosomal_biogenesis_genes,tfeb_genes)

# Make sure the exposure/independent/x variable (group variable) is a factor
seurat_object_hvg$group <- as.factor(ifelse(seurat_object_hvg$sglt2i_ever=="No",0,1))

source("nebula_pipeline_onecall.R")
write.csv(full_results,file.path(root_path, "Limonte/nebula_all_sgl2ti.csv"),row.names=F)

# ---- Volcano Plot ----
#unadjusted:
sglt2i_allcell_volcano_plot_unadjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_unadjusted, color = color_unadjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "sglt2i vs. no sglt2i (unadjusted)",  # Main title
    subtitle = "All cells",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_unadjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
sglt2i_allcell_volcano_plot_unadjusted
#FDR-adjusted:
sglt2i_allcell_volcano_plot_adjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_adjusted, color = color_adjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "sglt2i vs. no sglt2i (FDR ADJUSTED)",  # Main title
    subtitle = "All cells",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_adjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
sglt2i_allcell_volcano_plot_adjusted

# Load counts data from the object and round to nearest integer
# use counts_hvg to match pipeline
pb90_ptcells_t2d<-subset(pb90_ptcells,pb90_ptcells$group=="Type 2 Diabetes")
seurat_object_hvg<-pb90_ptcells_t2d
counts_hvg <- round(GetAssayData(seurat_object_hvg, layer = "counts"))

# Create a list of genes from the row names of the counts matrix
genes_list <- c(overall_autophagy_genes,autophagy_induction_genes,lysosomal_biogenesis_genes,tfeb_genes)

# Make sure the exposure/independent/x variable (group variable) is a factor
seurat_object_hvg$group <- as.factor(ifelse(seurat_object_hvg$sglt2i_ever=="No",0,1))

source("nebula_pipeline_onecall.R")
write.csv(full_results,file.path(root_path, "Limonte/nebula_pt_sgl2ti.csv"),row.names=F)
# ---- Volcano Plot ----

# Create the volcano plot using ggplot
#unadjusted:
sglt2i_ptcell_volcano_plot_unadjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_unadjusted, color = color_unadjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "sglt2i vs. no sglt2i (unadjusted)",  # Main title
    subtitle = "PT Cells Only",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_unadjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
sglt2i_ptcell_volcano_plot_unadjusted
#FDR-adjusted:
sglt2i_ptcell_volcano_plot_adjusted <- ggplot(full_results, aes(x = logFC_group1, y = p10_adjusted, color = color_adjusted)) +
  geom_point(alpha = 0.7) +  # Add points with transparency
  scale_color_identity() +  # Use color directly from 'color' column
  theme_minimal() +  # Apply minimalistic theme
  labs(
    title = "sglt2i vs. no sglt2i (FDR ADJUSTED)",  # Main title
    subtitle = "PT Cells Only",  # Subtitle
    x = "logFC",  # X-axis label
    y = "-log10(P-Value)",  # Y-axis label
    color = "LogFC Direction",
    caption = paste0(
      "Genes = ", Genes, 
      ", Nuclei = ", Nuclei, 
      ", Non-Convergence Rate: ", Nonconvergence_Rate, 
      ", Genes Filtered out for Low Expression: ", low_exp
    )  # Plot caption
  ) +
  theme(
    plot.title = element_text(hjust = 0),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) +
  xlim(min, max) +  # Set x-axis limits
  # Add gene labels for significant points
  geom_text(data = significant_df_adjusted, aes(label = gene),
            vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")

# Display the volcano plot
sglt2i_ptcell_volcano_plot_adjusted




```

