---
title: "Limonte autophagy gene look up"
author: "Ye Ji Choi & Kristen Miller"
format: html
editor_options: 
  chunk_output_type: console
---
  
```{r}
library(dplyr)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggforce)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(msigdbr)
library(ggpubr)
library(table1)
library(nlme)
library(reshape2)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/pylell/keys.json")
} else if (user == "kristenmiller") {
  root_path <- "/Users/kristenmiller/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive"
  git_path <- "/Users/kristenmiller/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/kristenmiller/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}

source(file.path(git_path, "Renal HEIRitage/RH_RH2_IMPROVE_functions.R"))
source(file.path(git_path, "Renal HEIRitage/RH_RH2_IMPROVE_scRNA_functions.R"))

entrez_map <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys   = keys(org.Hs.eg.db, keytype = "ENTREZID"),
  columns = c("SYMBOL", "GENENAME", "ALIAS"),
  keytype = "ENTREZID"
)


s3write_using_region <- function(FUN, ..., object, bucket, region = NULL, opts = NULL, filename = NULL) {
  if (missing(bucket)) {
    bucket <- get_bucketname(object)
  }
  object <- get_objectkey(object)
  
  tmp <- if (is.character(filename)) {
    file.path(tempdir(TRUE), filename)
  } else {
    # if object has an extension, keep it; otherwise make a generic tmp
    ext <- tools::file_ext(object)
    if (nzchar(ext)) tempfile(fileext = paste0(".", ext)) else tempfile()
  }
  
  on.exit(unlink(tmp), add = TRUE)
  
  # Add region to opts if provided
  if (!is.null(region)) {
    if (is.null(opts)) {
      opts <- list(region = region)
    } else {
      opts$region <- region
    }
  }
  
  FUN(tmp, ...)
  
  if (is.null(opts)) {
    r <- put_object(file = tmp, bucket = bucket, object = object)
  } else {
    r <- do.call("put_object", c(list(file = tmp, bucket = bucket, object = object), opts))
  }
  
  return(invisible(r))
}

# read in gene lists from Christine's excel file:
gene_lists <- openxlsx::read.xlsx(file.path(root_path, "./Limonte/Autophagy_gene_protein_lists_CL.xlsx"), sheet = "Gene lists",
                                 startRow = 3,colNames = TRUE)

# read in full pb90 seurat object:
pb90_so <- s3readRDS("data_clean/pb90_processed.rds", "scrna", region = "")

# remove obese controls:
pb90_sub<- subset(pb90_so, group %in% c("Lean Control", "Type 1 Diabetes", "Type 2 Diabetes"))
# only use baseline samples: 
pb90_sub<- subset(pb90_sub, visit %in% c("baseline"))

# only use proximal tubular cells:
pb90_sub<- subset(pb90_sub, KPMP_celltype %in% c("aPT", "PT-S1/S2", "PT-S3"))

# soma scan data: ignoring this for now
# urine_autophagy_soma <- openxlsx::read.xlsx(file.path(root_path, "./Limonte/Autophagy_gene_protein_lists_CL.xlsx"), sheet ="Urine_autophagy_score_PT_only",
#                                  startRow = 1,colNames = TRUE)
# blood_autophagy_soma <- openxlsx::read.xlsx(file.path(root_path, "./Limonte/Autophagy_gene_protein_lists_CL.xlsx"), sheet ="Blood_autophagy_score_kidney",
#                                  startRow = 1,colNames = TRUE)
# blood_TFEB_soma <- openxlsx::read.xlsx(file.path(root_path, "./Limonte/Autophagy_gene_protein_lists_CL.xlsx"), sheet ="Blood_TFEB_score_kidney",
#                                  startRow = 1,colNames = TRUE)
#soma<-read.csv(file.path(root_path, "./Data Harmonization/Data Clean/soma_harmonized_dataset.csv"))

```

```{r}
#create gene lists from Christine's files
overall_autophagy_genes <- c(
  gene_lists$`Overall.autophagy-related.genes.(n=589)`[!is.na(gene_lists$`Overall.autophagy-related.genes.(n=589)`)]
)
autophagy_induction_genes <- c(
  gene_lists$`Autophagy.induction.genes.(n=20)`[!is.na(gene_lists$`Autophagy.induction.genes.(n=20)`)]
)
lysosomal_biogenesis_genes <- c(
  gene_lists$`Lysosomal.biogenesis.genes.(n=26)`[!is.na(gene_lists$`Lysosomal.biogenesis.genes.(n=26)`)]
)
tfeb_genes <- c(
  gene_lists$`TFEB.genes.(n=13)`[!is.na(gene_lists$`TFEB.genes.(n=13)`)]
)

#MTORC1 pathway genes from JIC paper Supplemental Table S6 (https://www.jci.org/articles/view/164486/sd/1)
mtorc_genes<-c("STRADA",
"STRADB",
"RPS6",
"TSC2",
"TSC1",
"PRKAB1",
"PRKAB2",
"MTOR",
"PPM1A",
"RRAGB",
"RRAGA",
"RRAGD",
"RHEB",
"RPS6KB1",
"RRAGC",
"AKT1S1",
"EIF4G1",
"LAMTOR1",
"LAMTOR3",
"LAMTOR2",
"LAMTOR5",
"LAMTOR4",
"EIF4B",
"PRKAA2",
"YWHAB",
"CAB39L",
"CAB39",
"EIF4EBP1",
"PRKAG2",
"PRKAG3",
"PRKAG1",
"PRKAA1",
"STK11",
"RPTOR",
"EEF2K",
"MLST8",
"AKT1",
"AKT2",
"EIF4E")
```

```{r}
#Calculate module scores
#create a function that returns the updated seurat object plus the warning of which genes were not found
run_module_score <- function(seurat_obj, features, name) {
  warning_message <- NULL
  
  withCallingHandlers(
    {
      seurat_obj <- AddModuleScore(
        seurat_obj,
        features = list(features),
        name = name
      )
    },
    warning = function(w) {
      warning_message <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    }
  )
  
  list(seurat_obj = seurat_obj, warning = warning_message)
}

#add christine gene signatures to seurat object:
result <- run_module_score(pb90_sub, overall_autophagy_genes, "overall_autophagy_signature")
pb90_sub <- result$seurat_obj
missing_genes_overall_autophagy <- result$warning

result <- run_module_score(pb90_sub, autophagy_induction_genes, "autophagy_induction_signature")
pb90_sub <- result$seurat_obj
missing_genes_autophagy_induction <- result$warning

result <- run_module_score(pb90_sub, lysosomal_biogenesis_genes, "lysosomal_biogenesis_signature")
pb90_sub <- result$seurat_obj
missing_genes_lysosomal <- result$warning

result <- run_module_score(pb90_sub, tfeb_genes, "tfeb_signature")
pb90_sub <- result$seurat_obj
missing_genes_tfeb <- result$warning

#add MTORC signature:
result <- run_module_score(pb90_sub, mtorc_genes, "mtorc_signature")
pb90_sub <- result$seurat_obj


features<-c("overall_autophagy_signature1", "autophagy_induction_signature1",
                         "lysosomal_biogenesis_signature1","tfeb_signature1")
# grouping variables
# 1. diabetes vs reference
pb90_sub@meta.data$diabetes_reference<-ifelse(pb90_sub@meta.data$group%in%c("Lean Control"),"Lean Controls","Diabetes (T1 or T2)")
# 2. SGLT2i vs. no
pb90_t2d <- subset(pb90_sub, group %in% c("Type 2 Diabetes"))
```

```{r}
#### visuals of scores by disease group 
RidgePlot(pb90_sub, 
          features = features,
          ncol = 2, 
          group.by="diabetes_reference")

VlnPlot(pb90_sub, 
        features = features,
        ncol = 2,
        pt.size = 0,
        group.by="diabetes_reference")

FeaturePlot(pb90_sub, 
            features = features, 
            ncol = 2)

DotPlot(pb90_sub, 
        features = features, 
        group.by="diabetes_reference") + RotatedAxis()

#### visuals of scores by SGLT2i
RidgePlot(pb90_t2d, 
          features = features,
          ncol = 2, 
          group.by="sglt2i_ever")

VlnPlot(pb90_t2d, 
        features = features,
        ncol = 2,
        pt.size = 0,
        group.by="sglt2i_ever")

FeaturePlot(pb90_t2d, 
            features = features, 
            ncol = 2)

DotPlot(pb90_t2d, 
        features = features, 
        group.by="sglt2i_ever") + RotatedAxis()
```

```{r}
#Create long datasets; one row per cell:
df <- data.frame(
  record_id = pb90_sub$record_id,
  cohort=pb90_sub$cohort,
  group = pb90_sub$group,
  age=pb90_sub$age,
  sex=pb90_sub$sex,
  diabetes_reference = pb90_sub$diabetes_reference,
  sglt2i_ever= pb90_sub$sglt2i_ever,
  elevated_albuminuria=pb90_sub$elevated_albuminuria,
  overall_score = pb90_sub$overall_autophagy_signature1,
  induction_score = pb90_sub$autophagy_induction_signature1,
  lysosomal_score = pb90_sub$lysosomal_biogenesis_signature1,
  tfeb_score = pb90_sub$tfeb_signature1,
  mtorc_score = pb90_sub$mtorc_signature1
)
df.t2d<-subset(df,df$group=="Type 2 Diabetes")

```

```{r}
#Create person level dataset; mean score for each person (PT cells only):
df.1<- df %>%
    group_by(record_id) %>%
  summarise(group=group[1],
            cohort=cohort[1],
            age=age[1],
            sex=sex[1],
            diabetes_reference=diabetes_reference[1],
            sglt2i_ever=sglt2i_ever[1],
            elevated_albuminuria=elevated_albuminuria[1],
            mean_overall_score=mean(overall_score,na.rm=T),
            mean_induction_score=mean(induction_score,na.rm=T),
            mean_lysomal_score=mean(lysosomal_score,na.rm=T),
            mean_tfeb_score=mean(tfeb_score,na.rm=T),
            mean_mtorc_score=mean(mtorc_score,na.rm=T))
df.1.t2d<-subset(df.1,df.1$group=="Type 2 Diabetes")

```

```{r}
#table of person level comparisons across groups:
pvalue_kw <- function(x, ...) {
  # x is a list of vectors, one per group
  p <- kruskal.test(x)$p.value
  # Format the p-value
  p_formatted <- ifelse(p < 0.001, "<0.001", formatC(p, digits = 3, format = "f"))
  c("", p_formatted)
}
pvalue_wcx <- function(x, ...) {
  # x is a list of vectors, one per group
  p <- wilcox.test(x[[1]], x[[2]])$p.value
  p_formatted <- ifelse(p < 0.001, "<0.001", formatC(p, digits = 3, format = "f"))
  c("", p_formatted)
}
tab_group<-table1(~ cohort+age+sex+mean_overall_score+
                    mean_induction_score+
                    mean_lysomal_score+
                    mean_tfeb_score|factor(diabetes_reference),
                  data=df.1,
                  overall=F,
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)",
                                      "Mean (SD)"="MEAN (SD)"),
                  extra.col = list("P-value" = pvalue_kw))
tab_group

tab_sgl<-table1(~ mean_overall_score+
                    mean_induction_score+
                    mean_lysomal_score+
                    mean_tfeb_score|factor(sglt2i_ever),
                  data=df.1.t2d,
                  overall=F,
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)",
                                      "Mean (SD)"="MEAN (SD)"),
                  extra.col = list("P-value" = pvalue_wcx))
tab_sgl

```

#### Summary statistics by group
**Table 1: signatures by diabetes status**
```{r echo=FALSE}
kable(tab_group)
```

**Table 2: signatures by SGLT2i**
```{r echo=FALSE}
kable(tab_sgl)
```

```{r}
### mixed model, random intercept for pt

mixed <- function(data, outcome, grouping_var){
  # data<-df
  # grouping_var<-"group_2"
  # outcome<-"overall_score"
  data <- as.data.frame(data)
  form = as.formula(paste0(outcome,"~ ", grouping_var))
  mod <- lme(as.formula(form),random=~1|record_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|record_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,c(grouping_var))
  prs <-  pairs(m,adjust="tukey")
  return(list(anova,m,prs))
}
df$group_2<-NA
df$group_2[df$group=="Lean Control"]<-"control"
df$group_2[df$group=="Type 1 Diabetes"]<-"diabetes"
df$group_2[df$group=="Type 2 Diabetes"]<-"diabetes"

mod_overall_bygroup<-mixed(df,"overall_score","group_2")
mod_induction_bygroup<-mixed(df,"induction_score","group_2")
mod_lysomal_bygroup<-mixed(df,"lysosomal_score","group_2")
mod_tfeb_bygroup<-mixed(df,"tfeb_score","group_2")

mod_overall_bysgl<-mixed(df.t2d,"overall_score","sglt2i_ever")
mod_induction_bysgl<-mixed(df.t2d,"induction_score","sglt2i_ever")
mod_lysomal_bysgl<-mixed(df.t2d,"lysosomal_score","sglt2i_ever")
mod_tfeb_bysgl<-mixed(df.t2d,"tfeb_score","sglt2i_ever")

```

#### Mixed Models by group (control vs. t1s vs. t2d)
### Missing genes in our data:
- Overall Autophagy: `r missing_genes_overall_autophagy`
- Autophagy Induction: `r missing_genes_autophagy_induction`
- Lysosomal biogenesis: `r missing_genes_lysosomal`
- TFEB: `r missing_genes_tfeb`

### Overall gene score by group 

```{r echo=FALSE}
kable(mod_overall_bygroup[[1]])
```

### Induction gene score by group


```{r echo=FALSE}
kable(mod_induction_bygroup[[1]])
```

### Lysomal gene score by group

```{r echo=FALSE}
kable(mod_lysomal_bygroup[[1]])
```


### TFEB gene score by group

```{r echo=FALSE}
kable(mod_tfeb_bygroup[[1]])
```

#### Mixed Models by SGLT2i
### Overall gene score by SGLT2i 

```{r echo=FALSE}
kable(mod_overall_bysgl[[1]])
```

### Induction gene score by SGLT2i

```{r echo=FALSE}
kable(mod_induction_bysgl[[1]])
```

### Lysomal gene score by SGLT2i

```{r echo=FALSE}
kable(mod_lysomal_bysgl[[1]])
```

### TFEB gene score by SGLT2i

```{r echo=FALSE}
kable(mod_tfeb_bysgl[[1]])
```

```{r}
#signatures versus kidney outcomes:
df.1$elevated_albuminuria_yn<-ifelse(df.1$elevated_albuminuria=="Yes",1,0)
df.1.t2d$elevated_albuminuria_yn<-ifelse(df.1.t2d$elevated_albuminuria=="Yes",1,0)

# diabetes vs. reference (no interactions sig)
albu_overall<-glm(elevated_albuminuria_yn~mean_overall_score+diabetes_reference, df.1,family="binomial")
summary(albu_overall)

albu_induction<-glm(elevated_albuminuria_yn~mean_induction_score+diabetes_reference, df.1,family="binomial")
summary(albu_induction)

albu_lysosomal<-glm(elevated_albuminuria_yn~mean_lysomal_score+diabetes_reference, df.1,family="binomial")
summary(albu_lysosomal)

albu_tfeb<-glm(elevated_albuminuria_yn~mean_tfeb_score+diabetes_reference, df.1,family="binomial")
summary(albu_tfeb)

#### sglt2i: no interactions sig
albu_overall_t2d<-glm(elevated_albuminuria_yn~mean_overall_score+sglt2i_ever, df.1.t2d,family="binomial")
summary(albu_overall_t2d)

albu_induction_t2d<-glm(elevated_albuminuria_yn~mean_induction_score+sglt2i_ever, df.1.t2d,family="binomial")
summary(albu_induction_t2d)

albu_lysosomal_t2d<-glm(elevated_albuminuria_yn~mean_lysomal_score+sglt2i_ever, df.1.t2d,family="binomial")
summary(albu_lysosomal_t2d)

albu_tfeb_t2d<-glm(elevated_albuminuria_yn~mean_tfeb_score+sglt2i_ever, df.1.t2d,family="binomial")
summary(albu_tfeb_t2d)

```

```{r}
#correlation heatmap between various signatures:
corr.data<-df.1[,c("mean_overall_score","mean_induction_score","mean_lysomal_score","mean_tfeb_score","mean_mtorc_score")]
cormat <- round(cor(corr.data),2)
# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
upper_tri <- get_upper_tri(cormat)
melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggheatmap<-ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))

```

