---
title: "DECODE data request"
date: "`r lubridate::today()`"
format:
  html:
    embed-resources: true
    link-external-icon: false
    link-external-newwindow: false
    link-external-filter: '^(?:http:|https:)\/\/www\.quarto\.org\/custom'
---

```{r, include = F}
library(dplyr)
library(Hmisc)
library(purrr)
library(arsenal)
```

```{r, include = F}
load("/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive/DECODE/Data Clean/decode_harmonized_collapsed.RData")

decode_clean <- decode_collapsed %>%
  filter(study %nin% c("TEEN-LABS", "TODAY/TODAY2")) %>%
  dplyr::mutate(
    visit = case_when(
      study == "TEEN-LABS" & visit == "Month 1" ~ "baseline",
      T ~ visit
    ),
    #### Blood collection indicator
    blood_yn = case_when(
      !is.na(hba1c) | !is.na(creatinine_s) | !is.na(cholesterol) | 
        !is.na(hdl) | !is.na(ldl) | !is.na(triglycerides) |
        !is.na(cystatin_c_s) ~ "Yes", T ~ "No"),
    
    #### Urine collection indicator
    urine_yn = case_when(
      !is.na(microalbumin_u) | !is.na(acr_u) |
        !is.na(urine24h_start_time) | !is.na(urine24h_stop_time) |
        !is.na(emu_1_sample_id) | !is.na(emu_2_sample_id) | 
        !is.na(emu_3_sample_id) | !is.na(sample_id_urine24h) |
        !is.na(emu_urine_acr_mean_pooled) |
        !is.na(UAlbCreat)|
        !is.na(albuminuria) ~ "Yes", T ~ "No")
  )
```

```{r, echo = F, results = 'asis'}
summary(tableby(study ~ group + age + race + ethnicity + blood_yn + urine_yn, data = subset(decode_clean, visit == "baseline"), test = F))
```

#### ATTEMPT

Type of study: Interventional

[Adjunct-to-insulin therapy using SGLT2 inhibitors in youth with type 1 diabetes: a randomized controlled trial](https://www.nature.com/articles/s41591-025-03723-6)

#### CASPER

Type of study: Interventional

[Relative Hypoxia and Early Diabetic Kidney Disease in Type 1 Diabetes](https://diabetesjournals.org/diabetes/article/69/12/2700/16455/Relative-Hypoxia-and-Early-Diabetic-Kidney-Disease)

#### CROCODILE

Type of study: Observational

[Attenuated kidney oxidative metabolism in young adults with type 1 diabetes](https://www.jci.org/articles/view/183984)

#### IMPROVE

Type of study: Interventional

[Glycemia, Insulin Sensitivity, and Secretion Improve 3 Months Post-sleeve Gastrectomy in Youth With Type 2 Diabetes](https://academic.oup.com/jes/article/9/3/bvaf020/7985669?login=false#google_vignette)

#### PANDA

Type of study: Observational

Study active, no published paper yet.

[Study information](https://kri.washington.edu/studies/panda)

#### PENGUIN

Type of study: Observational

[Kidney Energetics and Cyst Burden in Autosomal Dominant Polycystic Kidney Disease: A Pilot Study](https://www.sciencedirect.com/science/article/abs/pii/S0272638624007169)

#### RENAL-HEIR

Type of study: Observational

All related publications listed here: [Renal Hemodynamics, Energetics and Insulin Resistance in Youth Onset Type 2 Diabetes Study](https://reporter.nih.gov/project-details/10397016)

#### RENAL-HEIRitage

Type of study: Observational

Analysis in progress. No published paper yet. 

#### Studies not listed (active): T1-DISCO, REMODEL-T1D, RPC2