---
title: "IMPROVE-T2D 12 month analysis (Dylan Changes)"
author: "Dylan W & Laura P"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(dplyr)
library(knitr)
library(ggpubr)
library(nlme)
library(emmeans)
library(gee)
library(crosstable)
library(Hmisc)
library(pROC)

knitr::opts_chunk$set(echo = FALSE,warning = FALSE,comment="")
user <- 'dylan'
options(scipen=999)
```

```{r functions, include=FALSE}
mixed <- function(data, outcome){
  data <- as.data.frame(data)
  form = as.formula(paste0(outcome,"~ visit"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  p <- pairs(m,adjust="tukey", simple="each")
  plot_dat <- data$outcome
 return(list(anova,m,p))
}

mixed_adj <- function(data, outcome){
  data <- as.data.frame(data)
  form = as.formula(paste0(outcome,"~ visit + per_wtloss"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  p <- pairs(m,adjust="tukey", simple="each")
  plot_dat <- data$outcome
  return(list(anova,m,p))
}

mixed_no3mo <- function(data, outcome){
  data <- as.data.frame(data[!(data$visit == 2),])
  data$visit <- droplevels(data$visit)
  form = as.formula(paste0(outcome,"~ visit"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  p <- pairs(m,adjust="tukey", simple="each")
  plot_dat <- data$outcome
 return(list(anova,m,p))
}

mixed_no3mo_weight <- function(data, outcome){
  data <- as.data.frame(data[!(data$visit == 2),])
  data$visit <- droplevels(data$visit)
  form = as.formula(paste0(outcome,"~ visit + mmtt_wt"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  p <- pairs(m,adjust="tukey", simple="each")
  plot_dat <- data$outcome
 return(list(anova,m,p))
}

mixed_adj_no3mo <- function(data, outcome){
  data <- as.data.frame(data[!(data$visit == 2),])
  data$visit <- droplevels(data$visit)
  form = as.formula(paste0(outcome,"~ visit + per_wtloss"))
  mod <- lme(as.formula(form),random=~1|subject_id,data = data,na.action = na.omit, correlation = corCompSymm(form = ~1|subject_id))
  anova <- anova.lme(mod, type="marginal")
  m <- emmeans(mod,"visit")
  p <- pairs(m,adjust="tukey", simple="each")
  plot_dat <- data$outcome
  return(list(anova,m,p))
}
```

```{r data, include=FALSE}
# read in screening data
if(user == 'laura'){
source('/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/IMPROVE T2D/Data_Raw/Primary analyses/IMPROVET2D-ScreeningData_R_2025-02-18_1440.r')
}else if(user == 'dylan'){
  data <- read.csv("C:/Users/netio/Documents/UofW/Projects/IMPROVE-CU/IMPROVET2D-ScreeningData_DATA_2025-09-24_1128.csv", 
                   na.strings = c(""," ","","-9999"))
  
label(data$subject_id)="Subject ID (IT_X)"
label(data$redcap_event_name)="Event Name"
label(data$redcap_repeat_instrument)="Repeat Instrument"
label(data$redcap_repeat_instance)="Repeat Instance"
label(data$co_enroll)="Co-enrolled participant? "
label(data$co_enroll_id)="Co-enrolled participant ID (RH Study ID)"
label(data$date_of_consent)="Date of Consent"
label(data$last_name)="Last Name"
label(data$first_name)="First Name"
label(data$mr_number)="Medical Record Number"
label(data$dob)="Date of Birth"
label(data$age_current)="Age at time of Consent (years)"
label(data$phone_number)="Phone number"
label(data$gender)="Gender"
label(data$race___1)="Race (choice=American Indian or Alaskan Native)"
label(data$race___2)="Race (choice=Asian)"
label(data$race___3)="Race (choice=Hawaiian or Pacific Islander)"
label(data$race___4)="Race (choice=Black or African American)"
label(data$race___5)="Race (choice=White)"
label(data$race___6)="Race (choice=Unknown)"
label(data$race___7)="Race (choice=Other)"
label(data$other_races)="Race, if Other"
label(data$ethnicity___1)="Ethnicity (choice=Hispanic)"
label(data$ethnicity___2)="Ethnicity (choice=Non-Hispanic)"
label(data$ethnicity___3)="Ethnicity (choice=Unknown/Not Reported)"
label(data$zip_code)="Zip code"
label(data$participation_status)="Participation Status"
label(data$demographics_complete)="Complete?"
label(data$study_visit_date)="Date of Medical History?"
label(data$diagnosis)="Date of Diagnosis with Diabetes"
label(data$length_of_diabetes)="Length of Diabetes Diagnosis (months)"
label(data$diabetes_age)="Age at Diagnosis of Diabetes (years)"
label(data$incl)="Does participant meet all inclusion/exclusion criteria to participate in this study?"
label(data$steriods)="Is patient currently taking steroids?"
label(data$steroid_type___1)="If yes, what kind of steroids is the patient taking? (choice=Oral steroids)"
label(data$steroid_type___2)="If yes, what kind of steroids is the patient taking? (choice=Inhaled steroids)"
label(data$steroid_type___3)="If yes, what kind of steroids is the patient taking? (choice=Steroid cream)"
label(data$steroid_type___4)="If yes, what kind of steroids is the patient taking? (choice=Steroid Injection)"
label(data$steroid_type___5)="If yes, what kind of steroids is the patient taking? (choice=Suppository)"
label(data$psychosis)="Is patient currently taking anti-psychotic meds?"
label(data$psychotic_typ)="Is patient currently taking any typical antipsychotics? (May include: Haldol, Loxitane, Thorazine, Trilafon, Stelazine, Serentil, etc.)"
label(data$psychosis_atyp)="Is patient currently taking any atypical antipsychotics? (May include: Abilify, Clozaril, Risperdal, Seroquel, Zyprexa, Solian, Goedon, etc.) "
label(data$incl_allergies)="Does patient have allergies to either seafood or iodine?"
label(data$incl_dka)="Has patient had a recent history of Diabetic Ketoacidosis (DKA)?"
label(data$incl_hh)="Has patient had a recent history of Hyperosmolar Hyperglycemia?"
label(data$incl_anemia)="Does patient have history of anemia?"
label(data$incl_meds)="Is patient currently taking any exclusionary medications? (ACE inhibitors, ARBs, diuretics, SGLT-2/1 blockers, daily NSAIDs or aspirin, sulfonamides, procaine, thiazolesulfone, probenecid)"
label(data$incl_mri)="Does patient have any MRI scanning contraindications? (claustrophobia, implantable metal devices, etc.)"
label(data$others_meds_dx)="Any other allergies, medications or conditions?"
label(data$medical_history_complete)="Complete?"
label(data$sglt2i)="Has participant ever taken an SGLT2i, including presently?"
label(data$sglt2i_history_complete)="Complete?"
label(data$phys_date)="Date of Physical Exam?"
label(data$screen_height)="Height (cm)"
label(data$screen_weight)="Weight (kg)"
label(data$screen_bmi)="BMI"
label(data$screen_bmi_percentile)="BMI Percentile (%) "
label(data$waist_circumference)="[Avg] Waist Circumference (cm)"
label(data$hip_circumference)="[Avg] Hip Circumference (cm) "
label(data$sys_bp)="Systolic Blood Pressure (SBP)"
label(data$dys_bp)="Diastolic Blood Pressure (DBP)"
label(data$map)="Mean Arterial Pressure (MAP)"
label(data$pulse)="Pulse"
label(data$activity_factor_male)="Male: Activity Factor"
label(data$activity_factor_female)="Female: Activity Factor"
label(data$schofield_male)="Male: Schofield"
label(data$schofield_female)="Female: Schofield"
label(data$phys_norm)="Physical exam normal?"
label(data$phys_no)="If no, briefly describe any clinically significant abnormal physical finding:"
label(data$breast_tanner)="Breast - Tanner Stage"
label(data$testicular_volume)="Testicular Volume (mL)?"
label(data$lmp)="Date of LMP?"
label(data$physical_exam_complete)="Complete?"
label(data$date_of_screen)="Date of Screening Labs?"
label(data$a1c_pre)="Pre-Screening A1c? (%)"
label(data$a1c_pre_date)="Date of pre-screening A1c?"
label(data$hba1c)="Screening A1c? (%)"
label(data$hemoglobin)="Hemoglobin? (g/dl)"
label(data$screen_hematocrit)="Hematocrit? (%)"
label(data$screen_serum_creatinine)="Serum Creatinine? (mg/dl)"
label(data$screen_urine_mab)="Urine Microalbumin (mcg/mL)"
label(data$screen_urine_cre)="Urine Creatinine (mg/dL)"
label(data$screen_urine_acr)="Urine ACR"
label(data$screen_pregnant)="Urine Pregnancy Test Results"
label(data$screening_labs_complete)="Complete?"
#Setting Units


#Setting Factors(will create new variable for factors)
data$redcap_event_name.factor = factor(data$redcap_event_name,levels=c("screening_arm_1","study_visits_arm_1"))
data$redcap_repeat_instrument.factor = factor(data$redcap_repeat_instrument,levels=c(""))
data$co_enroll.factor = factor(data$co_enroll,levels=c("1","0"))
data$gender.factor = factor(data$gender,levels=c("0","1","2"))
data$race___1.factor = factor(data$race___1,levels=c("0","1"))
data$race___2.factor = factor(data$race___2,levels=c("0","1"))
data$race___3.factor = factor(data$race___3,levels=c("0","1"))
data$race___4.factor = factor(data$race___4,levels=c("0","1"))
data$race___5.factor = factor(data$race___5,levels=c("0","1"))
data$race___6.factor = factor(data$race___6,levels=c("0","1"))
data$race___7.factor = factor(data$race___7,levels=c("0","1"))
data$ethnicity___1.factor = factor(data$ethnicity___1,levels=c("0","1"))
data$ethnicity___2.factor = factor(data$ethnicity___2,levels=c("0","1"))
data$ethnicity___3.factor = factor(data$ethnicity___3,levels=c("0","1"))
data$participation_status.factor = factor(data$participation_status,levels=c("1","2","3"))
data$demographics_complete.factor = factor(data$demographics_complete,levels=c("0","1","2"))
data$incl.factor = factor(data$incl,levels=c("1","0"))
data$steriods.factor = factor(data$steriods,levels=c("1","0"))
data$steroid_type___1.factor = factor(data$steroid_type___1,levels=c("0","1"))
data$steroid_type___2.factor = factor(data$steroid_type___2,levels=c("0","1"))
data$steroid_type___3.factor = factor(data$steroid_type___3,levels=c("0","1"))
data$steroid_type___4.factor = factor(data$steroid_type___4,levels=c("0","1"))
data$steroid_type___5.factor = factor(data$steroid_type___5,levels=c("0","1"))
data$psychosis.factor = factor(data$psychosis,levels=c("1","0"))
data$psychotic_typ.factor = factor(data$psychotic_typ,levels=c("1","0"))
data$psychosis_atyp.factor = factor(data$psychosis_atyp,levels=c("1","0"))
data$incl_allergies.factor = factor(data$incl_allergies,levels=c("1","0"))
data$incl_dka.factor = factor(data$incl_dka,levels=c("1","0"))
data$incl_hh.factor = factor(data$incl_hh,levels=c("1","0"))
data$incl_anemia.factor = factor(data$incl_anemia,levels=c("1","0"))
data$incl_meds.factor = factor(data$incl_meds,levels=c("1","0"))
data$incl_mri.factor = factor(data$incl_mri,levels=c("1","0"))
data$medical_history_complete.factor = factor(data$medical_history_complete,levels=c("0","1","2"))
data$sglt2i.factor = factor(data$sglt2i,levels=c("1","0"))
data$sglt2i_history_complete.factor = factor(data$sglt2i_history_complete,levels=c("0","1","2"))
data$activity_factor_male.factor = factor(data$activity_factor_male,levels=c("1.4","1.6","1.8"))
data$activity_factor_female.factor = factor(data$activity_factor_female,levels=c("1.4","1.6","1.8"))
data$phys_norm.factor = factor(data$phys_norm,levels=c("1","0"))
data$physical_exam_complete.factor = factor(data$physical_exam_complete,levels=c("0","1","2"))
data$screen_pregnant.factor = factor(data$screen_pregnant,levels=c("1","2"))
data$screening_labs_complete.factor = factor(data$screening_labs_complete,levels=c("0","1","2"))

levels(data$redcap_event_name.factor)=c("Screening","Study Visits")
levels(data$redcap_repeat_instrument.factor)=c("")
levels(data$co_enroll.factor)=c("Yes","No")
levels(data$gender.factor)=c("Female","Male","Other")
levels(data$race___1.factor)=c("Unchecked","Checked")
levels(data$race___2.factor)=c("Unchecked","Checked")
levels(data$race___3.factor)=c("Unchecked","Checked")
levels(data$race___4.factor)=c("Unchecked","Checked")
levels(data$race___5.factor)=c("Unchecked","Checked")
levels(data$race___6.factor)=c("Unchecked","Checked")
levels(data$race___7.factor)=c("Unchecked","Checked")
levels(data$ethnicity___1.factor)=c("Unchecked","Checked")
levels(data$ethnicity___2.factor)=c("Unchecked","Checked")
levels(data$ethnicity___3.factor)=c("Unchecked","Checked")
levels(data$participation_status.factor)=c("Participated","Removed (screen failed, withdrawn, dropped out, etc.)","Will participate (awaiting study visit)")
levels(data$demographics_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$incl.factor)=c("Yes","No")
levels(data$steriods.factor)=c("Yes","No")
levels(data$steroid_type___1.factor)=c("Unchecked","Checked")
levels(data$steroid_type___2.factor)=c("Unchecked","Checked")
levels(data$steroid_type___3.factor)=c("Unchecked","Checked")
levels(data$steroid_type___4.factor)=c("Unchecked","Checked")
levels(data$steroid_type___5.factor)=c("Unchecked","Checked")
levels(data$psychosis.factor)=c("Yes","No")
levels(data$psychotic_typ.factor)=c("Yes","No")
levels(data$psychosis_atyp.factor)=c("Yes","No")
levels(data$incl_allergies.factor)=c("Yes","No")
levels(data$incl_dka.factor)=c("Yes","No")
levels(data$incl_hh.factor)=c("Yes","No")
levels(data$incl_anemia.factor)=c("Yes","No")
levels(data$incl_meds.factor)=c("Yes","No")
levels(data$incl_mri.factor)=c("Yes","No")
levels(data$medical_history_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$sglt2i.factor)=c("Yes","No")
levels(data$sglt2i_history_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$activity_factor_male.factor)=c("1.4","1.6","1.8")
levels(data$activity_factor_female.factor)=c("1.4","1.6","1.8")
levels(data$phys_norm.factor)=c("Yes","No")
levels(data$physical_exam_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$screen_pregnant.factor)=c("NOT Pregnant","Pregnant")
levels(data$screening_labs_complete.factor)=c("Incomplete","Unverified","Complete")
  
   
  
  
}
screen <- data
data <- NULL
# select the participants
#screen <- screen %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_10","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20","IT_22"))
# code race ethnicity
screen$race <-  
  case_when(
  screen$race___1 == 1 ~ "American Indian or Alaskan Native",
  screen$race___2 == 1 ~ "Asian",
  screen$race___3 == 1 ~ "Hawaiian or Pacific Islander",
  screen$race___4 == 1 ~ "Black or African American",
  screen$race___5 == 1 ~ "White",
  screen$race___6 == 1 ~ "Unknown",
  screen$race___7 == 1 ~ "Other",
  .default = NA
)
screen$ethnicity <-  
  case_when(
  screen$ethnicity___1 == 1 ~ "Hispanic",
  screen$ethnicity___2 == 1 ~ "Non-Hispanic",
  screen$ethnicity___3 == 1 ~ "Unknown/Not reported",
  .default = NA
)
screen$race_eth <-  
  case_when(
  screen$race == "Black or African American" & screen$ethnicity == "Non-Hispanic" ~ "Black Non-Hispanic",
  screen$race == "Other" & screen$ethnicity == "Hispanic" ~ "Hispanic",
  screen$race == "White" & screen$ethnicity == "Hispanic" ~ "Hispanic",
  screen$race == "White" & screen$ethnicity == "Non-Hispanic" ~ "White Non-Hispanic",
  .default = NA
)


# read in repeating visit data
if(user == 'laura'){
  source('/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/IMPROVE T2D/Data_Raw/Primary analyses/IMPROVET2D-VisitDataNoMRI_R_2025-02-18_1421.r')
}else if(user == 'dylan'){
data=read.csv('C:/Users/netio/Documents/UofW/Projects/IMPROVE-CU/IMPROVET2D-VisitDataNoMRI_DATA_2025-09-24_1129.csv',na.strings = c(""," ","","-9999","-999","-99","-99999"))


label(data$subject_id)="Subject ID (IT_X)"
label(data$redcap_event_name)="Event Name"
label(data$redcap_repeat_instrument)="Repeat Instrument"
label(data$redcap_repeat_instance)="Repeat Instance"
label(data$med_date)="Date of Medication Review"
label(data$study_visit)="Associated study visit?"
label(data$current_db_dx)="Does patient currently have diabetes?"
label(data$diabetes_med___1)="What medications is patient currently taking for diabetes? (choice=Metformin)"
label(data$diabetes_med___2)="What medications is patient currently taking for diabetes? (choice=Insulin)"
label(data$diabetes_med___3)="What medications is patient currently taking for diabetes? (choice=Other)"
label(data$insulin_type___1)="If taking insulin, what type? (choice=Short acting)"
label(data$insulin_type___2)="If taking insulin, what type? (choice=Long acting)"
label(data$insulin_la_tdd)="If taking long acting insulin, what is average TDD? (units)"
label(data$diabetes_med_other___1)="If currently taking other diabetes medications, what type? (choice=Thiazolinediones (TZDs))"
label(data$diabetes_med_other___2)="If currently taking other diabetes medications, what type? (choice=GLP-1 agonists)"
label(data$diabetes_med_other___3)="If currently taking other diabetes medications, what type? (choice=SGLT-1/2 inhibitors)"
label(data$diabetes_med_other___4)="If currently taking other diabetes medications, what type? (choice=Other)"
label(data$diabetes_med_alt)="If other, what type of diabetes medications is patient taking?"
label(data$hypertension)="Does patient currently have hypertension?"
label(data$hypertenstion_controlled)="If yes, is hypertension controlled?"
label(data$htn_meds)="Is hypertension controlled by medication?"
label(data$htn_med_type___1)="What type of anti-hypertensive medication is patient taking? (choice=Angiotensin Converting Enzyme (ACE) Inhibitors)"
label(data$htn_med_type___2)="What type of anti-hypertensive medication is patient taking? (choice=Angiotensin Receptor Blockers (ARBs))"
label(data$htn_med_type___3)="What type of anti-hypertensive medication is patient taking? (choice=Beta Blockers)"
label(data$htn_med_type___4)="What type of anti-hypertensive medication is patient taking? (choice=Calcium Channel Blockers)"
label(data$htn_med_type___5)="What type of anti-hypertensive medication is patient taking? (choice=Diuretics)"
label(data$htn_med_type___6)="What type of anti-hypertensive medication is patient taking? (choice=Other)"
label(data$htn_med_other)="If other medication, what type?"
label(data$htn_manage)="If not controlled by medication, how is hypertension managed?"
label(data$addl_dx)="Does patient have additional history of cardiovascular or metabolic disease (CHF, Hyperlipidemia, Dyslipidemia, etc.)?"
label(data$addl_cardmet___1)="What history of cardiovascular or metabolic disease does patient have?  (choice=Congestive Heart Failure)"
label(data$addl_cardmet___2)="What history of cardiovascular or metabolic disease does patient have?  (choice=Hyperlipidemia/Dyslipidemia)"
label(data$addl_cardmet___3)="What history of cardiovascular or metabolic disease does patient have?  (choice=Other)"
label(data$cardmet_other)="If other, what cardiometabolic disease does patient have?"
label(data$addl_chf)="Is patient taking medications for CHF?"
label(data$addl_chf_meds___1)="If yes, what kind of medications is patient taking? (choice=Angiotensin Converting Enzyme (ACE) Inhibitors)"
label(data$addl_chf_meds___2)="If yes, what kind of medications is patient taking? (choice=Angiotensin Receptor Blockers (ARBs))"
label(data$addl_chf_meds___3)="If yes, what kind of medications is patient taking? (choice=Beta Blockers)"
label(data$addl_chf_meds___4)="If yes, what kind of medications is patient taking? (choice=Diuretics)"
label(data$addl_chf_meds___5)="If yes, what kind of medications is patient taking? (choice=Other)"
label(data$addl_chf_other)="If other medications, what type?"
label(data$addl_hld)="Is patient taking medications for hyperlipidemia/dyslipidemia?"
label(data$addl_hld_meds___1)="If yes, what kind of medications is patient taking? (choice=Statins)"
label(data$addl_hld_meds___2)="If yes, what kind of medications is patient taking? (choice=Fibrates)"
label(data$addl_hld_meds___3)="If yes, what kind of medications is patient taking? (choice=Niacin)"
label(data$addl_hld_meds___4)="If yes, what kind of medications is patient taking? (choice=Other)"
label(data$addl_hld_other)="If other medications, what type?"
label(data$meds_weight)="Is patient taking any weight loss medications?"
label(data$meds_weight_type___1)="If yes, what kind? (choice=Topiramate (Topamax))"
label(data$meds_weight_type___2)="If yes, what kind? (choice=Phentermine (Lomaira, Adipex-P))"
label(data$apnea)="Does patient have sleep apnea?"
label(data$apnea_cpap)="Does patient use CPAP?"
label(data$patient_medications_complete)="Complete?"
label(data$study_visit_accel)="Associated study visit?"
label(data$acc_wear_percent)="Accelerometer Wear Percent (%)"
label(data$accel_wear_tot)="Accelerometer Total Wear Time (min)"
label(data$accel_wear_day)="Average Wear Time/Day (min)"
label(data$accel_kcal_tot)="Total Activity Kcals"
label(data$accel_steps)="Total Steps"
label(data$accel_kcal)="Average Kcals/Day"
label(data$accel_time_sed)="Time: Sedentary (min)"
label(data$accel_per_sed)="Percent: Sedentary (%)"
label(data$accel_time_lt)="Time: Light Activity (min)"
label(data$accel_per_lt)="Percent: Light Activity (%)"
label(data$accel_time_mod)="Time: Moderate Activity (min)"
label(data$accel_per_mod)="Percent: Moderate Activity (%)"
label(data$accel_time_vig)="Time: Vigorous Activity (min)"
label(data$accel_per_vig)="Percent: Vigorous Activity (%)"
label(data$accelerometry_complete)="Complete?"
label(data$study_visit_mttt)="Associated study visit?"
label(data$mmtt_date)="Date of MMTT?"
label(data$mmtt_vitals)="Patient vitals complete?"
label(data$mmtt_ht)="Height (cm)"
label(data$mmtt_wt)="Weight (kg)"
label(data$mmtt_waist)="[Avg] Waist Circumference (cm)"
label(data$mmtt_hip)="[Avg] Hip Circumference (cm)"
label(data$mmtt_hr)="Pulse"
label(data$mmtt_sys_bp)="Systolic Blood Pressure (SBP)"
label(data$mmtt_dia_bp)="Diastolic Blood Pressure (DBP)"
label(data$mmtt_map)="Mean Arterial Pressure (MAP)"
label(data$mmtt_lmp)="Females: Date of LMP"
label(data$mmtt_pregnant)="Urine Pregnancy Test Results"
label(data$mmtt_brmr)="Resting Metabolic Rate measurements complete?"
label(data$brmr_vo2_ml)="VO2 (ml/kg/min)"
label(data$brmr_vo2_l)="VO2 (L/min)"
label(data$brmr_vco2)="VCO2 (L/min)"
label(data$brmr_rq)="Respiratory Quotient (RQ)"
label(data$brmr_ree)="Resting Energy Expenditure (REE)"
label(data$brmr_fo)="Fat Oxidation "
label(data$brmr_co)="Carbohydrate Oxidation "
label(data$mmtt_60rmr)="60 min Resting Metabolic Rate measurements complete?"
label(data$srmr_vo2_ml)="VO2 (ml/kg/min)"
label(data$srmr_vo2_l)="VO2 (L/min)"
label(data$srmr_co2)="VCO2 (L/min)"
label(data$srmr_rq)="Respiratory Quotient (RQ)"
label(data$srmr_ree)="Resting Energy Expenditure (REE)"
label(data$srmr_fo)="Fat Oxidation"
label(data$srmr_co)="Carbohydrate Oxidation"
label(data$mmtt_bg_base)="Fasting Blood Glucose (mg/dL)"
label(data$mmtt_base_labs)="CMP labs complete?"
label(data$mmtt_hba1c_base)="HbA1c (%)"
label(data$mmtt_chol_base)="Cholesterol (mg/dL)"
label(data$mmtt_hdl_base)="HDL (mg/dL)"
label(data$mmtt_ldl_base)="LDL (mg/dL)"
label(data$mmtt_tri_base)="Triglycerides (mg/dL)"
label(data$mmtt_na_base)="Na/Sodium (nmol/L)"
label(data$mmtt_k_base)="K/Potassium (nmol/L)"
label(data$mmtt_cl_base)="Cl/Chloride (nmol/L)"
label(data$mmtt_bicarb_base)="Bicarbonate/HCO3 (nmol/L)"
label(data$mmtt_gluc_base)="Glucose - CMP (mg/dL)"
label(data$mmtt_bun_base)="Blood Urea Nitrogen (mg/dL)"
label(data$mmtt_creat_base)="Creatinine (mg/dL)"
label(data$mmtt_ca_base)="Calcium (mg/dL)"
label(data$mmtt_totprot_base)="Total Protein (g/dL)"
label(data$mmtt_alb_base)="Albumin (g/dL)"
label(data$mmtt_ag_base)="A/G Ratio "
label(data$mmtt_bili_base)="Total Bilirubin (mg/dL)"
label(data$mmtt_alkphos_base)="Alkaline Phosphatase (U/L)"
label(data$mmtt_gptalt_base)="Alanine Aminotransferase (U/L)"
label(data$mmtt_gotast_base)="Aspartate Aminotransferase (U/L)"
label(data$mmtt_ffa_labs)="FFA Labs complete?"
label(data$ffa_neg_10)="FFA (T= -10)"
label(data$ffa_0)="FFA (T= 0)"
label(data$ffa_30)="FFA (T= 30)"
label(data$ffa_60)="FFA (T= 60)"
label(data$ffa_120)="FFA (T= 120)"
label(data$ffa_180)="FFA (T = 180)"
label(data$ffa_240)="FFA (T= 240)"
label(data$mmtt_insulin)="Insulin labs complete?"
label(data$mmtt_insulin_neg_10)="Insulin (T= -10)"
label(data$mmtt_insulin_0)="Insulin (T= 0)"
label(data$mmtt_insulin_10)="Insulin (T= 10)"
label(data$mmtt_insulin_20)="Insulin (T= 20)"
label(data$mmtt_insulin_30)="Insulin (T= 30)"
label(data$mmtt_insulin_45)="Insulin (T= 45)"
label(data$mmtt_insulin_60)="Insulin (T= 60)"
label(data$mmtt_insulin_90)="Insulin (T= 90)"
label(data$mmtt_insulin_120)="Insulin (T= 120)"
label(data$mmtt_insulin_150)="Insulin (T= 150)"
label(data$mmtt_insulin_180)="Insulin (T= 180)"
label(data$mmtt_insulin_210)="Insulin (T= 210)"
label(data$mmtt_insulin_240)="Insulin (T= 240)"
label(data$mmtt_glp1)="GLP-1 labs complete?"
label(data$glp1_neg_10)="GLP-1 (T= -10)"
label(data$glp1_0)="GLP-1 (T= 0)"
label(data$glp1_10)="GLP-1 (T= 10)"
label(data$glp1_30)="GLP-1 (T= 30)"
label(data$glp1_45)="GLP-1 (T= 45)"
label(data$glp1_60)="GLP-1 (T= 60)"
label(data$glp1_120)="GLP-1 (T= 120)"
label(data$glp1_240)="GLP-1 (T= 240)"
label(data$mmtt_cpep)="C-Peptide labs complete?"
label(data$mmtt_cpep_neg_10)="C-Peptide (T= -10)"
label(data$mmtt_cpep_0)="C-Peptide (T= 0)"
label(data$mmtt_cpep_10)="C-Peptide (T= 10)"
label(data$mmtt_cpep_20)="C-Peptide (T= 20)"
label(data$mmtt_cpep_30)="C-Peptide (T= 30)"
label(data$mmtt_cpep_60)="C-Peptide (T= 60)"
label(data$mmtt_cpep_90)="C-Peptide (T= 90)"
label(data$mmtt_cpep_120)="C-Peptide (T= 120)"
label(data$mmtt_cpep_180)="C-Peptide (T= 180)"
label(data$mmtt_cpep_240)="C-Peptide (T=240)"
label(data$mmtt_yy)="Peptide YY labs complete?"
label(data$mmtt_yy_neg10)="Peptide YY (T= -10)"
label(data$mmtt_yy_0)="Peptide YY (T= 0)"
label(data$mmtt_yy_10)="Peptide YY (T= 10)"
label(data$mmtt_yy_30)="Peptide YY (T= 30)"
label(data$mmtt_yy_45)="Peptide YY (T= 45)"
label(data$mmtt_yy_60)="Peptide YY (T= 60)"
label(data$mmtt_yy_120)="Peptide YY (T= 120)"
label(data$mmtt_yy_240)="Peptide YY (T= 240)"
label(data$mmtt_glucagon)="Glucagon labs complete?"
label(data$glucagon_neg_10)="Glucagon (T= -10)"
label(data$glucagon_0)="Glucagon (T= 0)"
label(data$glucagon_10)="Glucagon (T= 10)"
label(data$glucagon_30)="Glucagon (T= 30)"
label(data$glucagon_45)="Glucagon (T= 45)"
label(data$glucagon_60)="Glucagon (T= 60)"
label(data$glucagon_120)="Glucagon (T= 120)"
label(data$glucagon_240)="Glucagon (T= 240)"
label(data$mmtt_gluc)="Glucose labs complete?"
label(data$mmtt_bg_neg_10)="Blood glucose (T= -10)"
label(data$mmtt_bg_0)="Blood glucose (T= 0)"
label(data$mmtt_bg_10)="Blood glucose (T= 10)"
label(data$mmtt_bg_20)="Blood glucose (T= 20)"
label(data$mmtt_bg_30)="Blood glucose (T= 30)"
label(data$mmtt_bg_45)="Blood glucose (T= 45)"
label(data$mmtt_bg_60)="Blood glucose (T= 60)"
label(data$mmtt_bg_90)="Blood glucose (T= 90)"
label(data$mmtt_bg_120)="Blood glucose (T= 120)"
label(data$mmtt_bg_150)="Blood glucose (T= 150)"
label(data$mmtt_bg_180)="Blood glucose (T= 180)"
label(data$mmtt_bg_210)="Blood glucose (T= 210)"
label(data$mmtt_bg_240)="Blood glucose (T= 240)"
label(data$mmtt_metabolic_cart_complete)="Complete?"
label(data$study_visit_bodycomp)="Associated study visit?"
label(data$bodcomp_date)="Date of Body Composition Testing?"
label(data$dxa_complete)="DXA Values Complete?"
label(data$dxa_body_fat)="Body Fat (%)"
label(data$dxa_lean_mass)="Lean Mass (%)"
label(data$dxa_trunk_mass)="Trunk Mass (%)"
label(data$dxa_fat_kg)="Fat Mass (kg)"
label(data$dxa_lean_kg)="Lean Mass (kg)"
label(data$dxa_trunk_kg)="Trunk Mass (kg)"
label(data$dxa_bmd)="Bone Mineral Density (g/cm2)"
label(data$bodpod_complete)="BOD POD Values Complete?"
label(data$bp_body_fat)="Body Fat (%)"
label(data$bp_lean_mass)="Lean Mass (%) "
label(data$bp_fat_mass)="Fat Mass (kg)"
label(data$bp_lean_kg)="Lean Mass (kg)"
label(data$bp_est_rmr)="Estimated RMR (kcal/day)"
label(data$bia_complete)="Handheld BIA Complete?"
label(data$bia_body_fat)="Body Fat (%)"
label(data$bia_lean_mass)="Lean Mass (%)"
label(data$bia_fat_mass_kg)="Fat Mass (kg)"
label(data$bia_lean_mass_kg)="Lean Mass (kg)"
label(data$tanita_scale_complete)="Tanita Scale Complete?"
label(data$tanita_body_fat)="Body Fat (%)"
label(data$tanita_lean_mass)="Lean Mass (%)"
label(data$tanita_fat_mass_kg)="Fat Mass (kg)"
label(data$tanita_lean_mass_kg)="Lean Mass (kg)"
label(data$body_composition_dxa_bod_pod_complete)="Complete?"
label(data$study_visit_clamp)="Associated study visit?"
label(data$clamp_date)="Clamp Date"
label(data$baseline)="Baseline clamp vitals complete?"
label(data$clamp_height)="Clamp Height (cm)"
label(data$clamp_weight)="Clamp Weight (kg)"
label(data$clamp_bmi)="Clamp BMI"
label(data$clamp_sbp)="Clamp Systolic Blood Pressure (SBP)"
label(data$clamp_dbp)="Clamp Diastolic Blood Pressure (DBP)"
label(data$clamp_map)="Clamp Mean Arterial Pressure (MAP)"
label(data$clamp_pls)="Clamp Pulse"
label(data$fasting_labs)="Baseline (fasting) clamp labs complete?"
label(data$cholesterol)="Baseline Cholesterol (mg/dl)"
label(data$hdl)="Baseline HDL (mg/dl)"
label(data$ldl)="Baseline LDL (mg/dl)"
label(data$triglycerides)="Baseline Triglycerides (mg/dl)"
label(data$total_protein)="Baseline Total Protein (g/dl)"
label(data$serum_sodium)="Baseline Serum Sodium (mmol/L)"
label(data$cystatin_c)="Baseline Cystatin C (mg/L)"
label(data$serum_creatinine)="Baseline Serum Creatinine (mg/dl)"
label(data$serum_osmolality)="Baseline Serum Osmolality (mmol/kg)"
label(data$copeptin)="Baseline Copeptin (pmol/L)"
label(data$fbg)="Fasting blood glucose? (mg/dl)"
label(data$ns_bolus)="NS bolus(es) given?"
label(data$ns_number)="# of NS bolus(es) given?"
label(data$ns1_time)="Time of NS Bolus #1?"
label(data$ns1_vol)="Volume of NS Bolus #1?"
label(data$ns1_bg)="Blood glucose after NS Bolus #1?"
label(data$ns2_time)="Time of NS Bolus #2?"
label(data$ns2_vol)="Volume of NS Bolus #2?"
label(data$ns2_bg)="Blood glucose after NS Bolus #2?"
label(data$ns3_time)="Time of NS Bolus #3?"
label(data$ns3_vol)="Volume of NS Bolus #3?"
label(data$ns3_bg)="Blood glucose after NS Bolus #3?"
label(data$d20_bolus_time)="D20 Bolus - Time Started"
label(data$d20_bolus_vol)="D20 Bolus - Intial Volume"
label(data$d20_infusion)="D20 - (Steady State) Infusion Rate"
label(data$d20_infusion_rate_120)="D20 Infusion Rate (T=120)"
label(data$d20_infusion_rate_220)="D20 Infusion Rate (T=220)"
label(data$d20_infusion_rate_230)="D20 Infusion Rate (T=230)"
label(data$d20_infusion_rate_240)="D20 Infusion Rate (T=240)"
label(data$d20_volume)="Total D20 Volume?"
label(data$urine_labs)="Urine Lab values complete?"
label(data$clamp_urine_mab_baseline)="Clamp Urine Microalbumin (Baseline)"
label(data$clamp_urine_cre_baseline)="Clamp Urine Creatinine (Baseline)"
label(data$clamp_acr_baseline)="Clamp ACR (Baseline)"
label(data$clamp_urine_sodium)="Clamp Urine Sodium (Baseline)"
label(data$clamp_glucose_bl)="Clamp Urine Glucose (Baseline)"
label(data$urine_osmol)="Clamp Urine Osmolality (Baseline)"
label(data$urine_glucose)="Clamp Urine Glucose (T= 250)"
label(data$clamp_urine_mab_250)="Clamp Urine Microalbumin (T= 250)"
label(data$clamp_urine_cre_250)="Clamp Urine Creatinine (T= 250)"
label(data$clamp_acr_250)="Clamp ACR (T= 250)"
label(data$clamp_urine_vol)="Clamp Urine Volume? (mL)"
label(data$bg_labs)="Blood glucose lab values complete?"
label(data$glucose_minus_120)="Blood glucose (T= -120) "
label(data$glucose_minus_90)="Blood glucose (T= -90)"
label(data$glucose_minus_20)="Blood glucose (T= -20)"
label(data$glucose_minus_10)="Blood glucose (T= -10)"
label(data$glucose_minus_5)="Blood glucose (T= -5)"
label(data$glucose_10)="Blood glucose (T= 10)"
label(data$glucose_15)="Blood glucose (T= 15)"
label(data$glucose_20)="Blood glucose (T= 20)"
label(data$glucose_25)="Blood glucose (T= 25)"
label(data$glucose_30)="Blood glucose (T= 30)"
label(data$glucose_35)="Blood glucose (T= 35)"
label(data$glucose_40)="Blood glucose (T= 40)"
label(data$glucose_45)="Blood glucose (T= 45)"
label(data$glucose_50)="Blood glucose (T= 50)"
label(data$glucose_55)="Blood glucose (T= 55)"
label(data$glucose_60)="Blood glucose (T= 60)"
label(data$glucose_65)="Blood glucose (T= 65)"
label(data$glucose_70)="Blood glucose (T= 70)"
label(data$glucose_75)="Blood glucose (T= 75)"
label(data$glucose_80)="Blood glucose (T= 80)"
label(data$glucose_85)="Blood glucose (T= 85)"
label(data$glucose_90)="Blood glucose (T= 90)"
label(data$glucose_95)="Blood glucose (T= 95)"
label(data$glucose_100)="Blood glucose (T= 100)"
label(data$glucose_105)="Blood glucose (T= 105)"
label(data$glucose_110)="Blood glucose (T= 110)"
label(data$glucose_115)="Blood glucose (T= 115)"
label(data$glucose_120)="Blood glucose (T= 120)"
label(data$glucose_125)="Blood glucose (T= 125)"
label(data$glucose_130)="Blood glucose (T= 130)"
label(data$glucose_135)="Blood glucose (T= 135)"
label(data$glucose_140)="Blood glucose (T= 140)"
label(data$glucose_145)="Blood glucose (T= 145)"
label(data$glucose_150)="Blood glucose (T= 150)"
label(data$glucose_155)="Blood glucose (T = 155)"
label(data$glucose_160)="Blood glucose (T= 160)"
label(data$glucose_165)="Blood glucose (T= 165)"
label(data$glucose_170)="Blood glucose (T= 170)"
label(data$glucose_175)="Blood glucose (T= 175)"
label(data$glucose_180)="Blood glucose (T= 180)"
label(data$glucose_185)="Blood glucose (T= 185)"
label(data$glucose_190)="Blood glucose (T= 190)"
label(data$glucose_195)="Blood glucose (T= 195)"
label(data$glucose_200)="Blood glucose (T= 200)"
label(data$glucose_205)="Blood glucose (T= 205)"
label(data$glucose_210)="Blood glucose (T= 210)"
label(data$glucose_215)="Blood glucose (T= 215)"
label(data$glucose_220)="Blood glucose (T= 220)"
label(data$glucose_225)="Blood glucose (T= 225)"
label(data$glucose_230)="Blood glucose (T= 230)"
label(data$glucose_235)="Blood glucose (T= 235)"
label(data$glucose_240)="Blood glucose (T= 240)"
label(data$glucose_245)="Blood glucose (T= 245) [ARG - 5]"
label(data$glucose_249)="Blood glucose (T = 249) [ARG -1]"
label(data$glucose_250)="Blood glucose (T= 250)"
label(data$glucose_arg2)="Blood glucose (T= ARG +2)"
label(data$glucose_arg3)="Blood glucose (T= ARG +3)"
label(data$glucose_arg4)="Blood glucose (T= ARG +4)"
label(data$glucose_arg5)="Blood glucose (T= ARG +5)"
label(data$glucose_15post)="Blood glucose (T= +15 after clamp)"
label(data$glucose_30post)="Blood glucose (T= +30 after clamp)"
label(data$cpep_lab)="C Peptide lab values complete?"
label(data$cpeptide_minus_10)="C Peptide (T= -10)"
label(data$cpeptide_minus_5)="C Peptide (T= -5)"
label(data$cpeptide_2)="C Peptide (T= 2)"
label(data$cpeptide_4)="C Peptide (T= 4)"
label(data$cpeptide_6)="C Peptide (T= 6)"
label(data$cpeptide_8)="C Peptide (T= 8)"
label(data$cpeptide_10)="C Peptide (T= 10)"
label(data$cpeptide_120)="C Peptide (T= 120)"
label(data$cpeptide_220)="C Peptide (T= 220)"
label(data$cpeptide_230)="C Peptide (T= 230)"
label(data$cpeptide_240)="C Peptide (T= 240)"
label(data$cpeptide_245)="C Peptide (T= 245) [ARG -5]"
label(data$cpeptide_249)="C Peptide (T= 249) [ARG -1]"
label(data$cpeptide_250)="C Peptide (T= 250)"
label(data$cpeptide_252)="C Peptide (T= ARG +2)"
label(data$cpeptide_253)="C Peptide (T= ARG +3)"
label(data$cpeptide_254)="C Peptide (T= ARG +4)"
label(data$cpeptide_255)="C Peptide (T= ARG +5)"
label(data$insulin_labs)="Insulin lab values complete?"
label(data$insulin_minus_10)="Insulin (T= -10)"
label(data$insulin_minus_5)="Insulin (T= -5)"
label(data$insulin_2)="Insulin (T= 2)"
label(data$insulin_4)="Insulin (T= 4)"
label(data$insulin_6)="Insulin (T= 6)"
label(data$insulin_8)="Insulin (T= 8)"
label(data$insulin_10)="Insulin (T= 10)"
label(data$insulin_120)="Insulin (T= 120)"
label(data$insulin_220)="Insulin (T= 220)"
label(data$insulin_230)="Insulin (T= 230)"
label(data$insulin_240)="Insulin (T= 240)"
label(data$insulin_245)="Insulin (T= 245) [ARG -5]"
label(data$insulin_249)="Insulin (T= 249) [ARG -1]"
label(data$insulin_250)="Insulin (T= 250)"
label(data$insulin_252)="Insulin (T= ARG +2)"
label(data$insulin_253)="Insulin (T= ARG +3)"
label(data$insulin_254)="Insulin (T= ARG +4)"
label(data$insulin_255)="Insulin (T= ARG +5)"
label(data$iohexol_bolus)="Iohexol bolus given?"
label(data$iohexol_time)="Iohexol bolus time?"
label(data$iohexol_vol)="Iohexol bolus volume?"
label(data$iohexol_minus_10)="Iohexol (T= -10)"
label(data$iohexol_120)="Iohexol (T= 120)"
label(data$iohexol_150)="Iohexol (T= 150)"
label(data$iohexol_180)="Iohexol (T= 180)"
label(data$iohexol_210)="Iohexol (T=210)"
label(data$iohexol_240)="Iohexol (T= 240)"
label(data$pah_bolus)="PAH bolus given?"
label(data$pah_time)="PAH bolus time?"
label(data$pah_vol)="PAH bolus volume?"
label(data$pah_minus_10)="PAH (T= -10)"
label(data$pah_90)="PAH (T= 90)"
label(data$pah_120)="PAH (T= 120)"
label(data$hct_lab)="Hematocrit lab values complete?"
label(data$hematocrit_minus_10)="Hematocrit (T= -10)"
label(data$hematocrit_90)="Hematocrit (T= 90)"
label(data$hematocrit_120)="Hematocrit (T= 120)"
label(data$clamp_complete)="Complete?"
label(data$study_visit_outcomes)="Associated study visit?"
label(data$kidney_outcomes)="Kidney/Hemodynamic Outcomes Complete?"
label(data$egfr)="eGFR"
label(data$gfr)="Absolute GFR (ml/min)"
label(data$gfr_bsa)="BSA Normalized GFR (ml/min/1.72m2)"
label(data$abs_pah)="Absolute PAH Clearance (ml/min)"
label(data$pah_bsa)="BSA Normalized PAH clearance (ml/min/1.73m2)"
label(data$rpf)="Absolute Effective Renal Plamsa Flow (ERPF) (ml/min)"
label(data$erpf_bsa)="BSA Normalized Renal Plasma Flow (ml/min/1.73m2)"
label(data$ecv)="Extracellular Volume (ECV)"
label(data$gfr_ecv_percent)="GFR/ECV (%)"
label(data$gfr_ecv_std)="GFR Standardized by ECV"
label(data$ff)="Filtration Fraction (FF)"
label(data$rbf)="Absolute Renal Blood Flow (RBF)"
label(data$filtration_pressure)="Filtration Pressure "
label(data$mean_plasma_protein)="Mean Plasma Protein"
label(data$glomerular_onco_pressure)="Glomerular Oncotic Pressure"
label(data$glomerular_pressure)="Glomerular Pressure"
label(data$aff_arteriolar_resistance)="Afferent Arteriolar Resistance"
label(data$eff_arteriolar_resistance)="Efferent Arteriolar Resistance"
label(data$m_value)="Insulin Sensitivity "
label(data$metab_outcomes)="Metabolomics Outcomes complete?"
label(data$mri_date)="Date of MRI?"
label(data$o2_sats)="O2 Sat (pre-MRI?)"
label(data$asl_outcomes)="ASL MRI Outcomes complete?"
label(data$asl_right)="Right Kidney: Perfusion (cortex) (ml/min/100g)"
label(data$asl_left)="Left kidney: Perfusion (cortex) (ml/min/100g)"
label(data$bold_outcomes)="BOLD MRI Outcomes complete?"
label(data$bold_r_bl_cortex)="RIGHT: Baseline (pre-furosemide) R2* (cortex) "
label(data$bold_r_bl_medulla)="RIGHT: Baseline (pre-furosemide) R2* (medulla)"
label(data$bold_r_bl_kidney)="RIGHT: Baseline (pre-furosemide) R2* (kidney)"
label(data$bold_r_pf_cortex)="RIGHT: Post-furosemide R2* (cortex)"
label(data$bold_r_pf_medulla)="RIGHT: Post-furosemide R2* (medulla)"
label(data$bold_r_pf_kidney)="RIGHT: Post-furosemide R2* (kidney)"
label(data$bold_l_bl_cortex)="LEFT: Baseline (pre-furosemide) R2* (cortex)"
label(data$bold_l_bl_medulla)="LEFT: Baseline (pre-furosemide) R2* (medulla)"
label(data$bold_l_bl_kidney)="LEFT: Baseline (pre-furosemide) R2* (kidney)"
label(data$bold_l_pf_cortex)="LEFT: Post-furosemide R2* (cortex)"
label(data$bold_l_pf_medulla)="LEFT: Post-furosemide R2* (medulla)"
label(data$bold_l_pf_kidney)="LEFT: Post-furosemide R2* (kidney)"
label(data$adc_outcomes)="ADC MRI Outcomes complete?"
label(data$adc_right)="Right Kidney: ADC (cortex) (mm2/s) "
label(data$adc_left)="Left Kidney: ADC (cortex) (mm2/s) "
label(data$tkv_outcomes)="TKV MRI Outcomes complete?"
label(data$volume_right)="Right kidney: volume (mL)"
label(data$volume_left)="Left kidney: volume (mL)"
label(data$outcomes_complete)="Complete?"
label(data$study_id)="UM Path ID (SU Number):"
label(data$dpir_id)="DPIR ID:"
label(data$cortex_yn)="Cortex extracted for analysis?"
label(data$cortex_diagnosis___1)="Diagnosis? (choice=No significant histological abnormalities seen)"
label(data$cortex_diagnosis___2)="Diagnosis? (choice=Other)"
label(data$cortex_dx_other)="If other: "
label(data$gloms)="# Glomeruli:"
label(data$glom_enlarge___1)="Glomerular Enlargement (choice=Normal size)"
label(data$glom_enlarge___2)="Glomerular Enlargement (choice=Normal to mild)"
label(data$glom_enlarge___3)="Glomerular Enlargement (choice=Mild)"
label(data$glom_enlarge___4)="Glomerular Enlargement (choice=Mild to Moderate)"
label(data$glom_enlarge___5)="Glomerular Enlargement (choice=Moderate)"
label(data$glom_enlarge___6)="Glomerular Enlargement (choice=Moderate to Severe)"
label(data$glom_enlarge___7)="Glomerular Enlargement (choice=Severe)"
label(data$gloms_gs)="GS Gloms (%)"
label(data$ifta)="IFTA (%)"
label(data$vessels___1)="Vessels (choice=Normal)"
label(data$vessels___2)="Vessels (choice=Other)"
label(data$vessels_other)="If other: "
label(data$fia)="Fractional Interstitial Area:"
label(data$glom_tuft_area)="Glomerular Tuft Area (um^2):"
label(data$glom_volume_weibel)="Glomerular Volume - Weibel (10^6 um^3):"
label(data$glom_volume_wiggins)="Glomerular Volume - Wiggins (10^6 um^3):"
label(data$glom_volume_con)="Glomerular Volume - Consensus (10^6 um^3):"
label(data$mes_matrix_area)="Mesangial Matrix Area (um^2):"
label(data$mes_index)="Mesangial Index (%):"
label(data$mes_volume_weibel)="Mesangial Volume - Weibel (10^6 um^3):"
label(data$mes_volume_wiggins)="Mesangial Volume - Wiggins (10^6 um^3):"
label(data$mes_volume_con)="Mesangial Volume - Consensus (10^6 um^3):"
label(data$glom_nuc_count)="Glomerular Nuclear Count:"
label(data$gbm_thick_artmean)="GBM thickness nm (arithmetic mean)"
label(data$gbm_thick_harmmean)="GBM thickness nm (harmonic mean)"
label(data$mes_nuc_count)="Mesangial Nuclear Count:"
label(data$art_intima)="% Artery/arteriolar Intima:"
label(data$art_media)="% Artery/arteriolar Media:"
label(data$pod_nuc_density)="Podocyte Nuclear Density:"
label(data$pod_cell_volume)="Podocyte Cell Volume:"
label(data$biopsy_results_michigan_complete)="Complete?"
label(data$watchpat_yes_no)="WatchPAT collected? "
label(data$study_visit_accel1)="Associated study visit?"
label(data$total_ahi)="Total AHI"
label(data$rem_ahi)="REM AHI"
label(data$nonrem_ahi)="NonREM AHI"
label(data$total_sleep_time)="Total Sleep Time"
label(data$rem_sleep)="Percent REM sleep"
label(data$time_o2_90)="Total time O2 saturation less than 90%"
label(data$percent_time_o2_90)="Percent sleep time O2 less than 90"
label(data$percent_tst_o2_90)="Percent total sleep time O2 saturation less than 90"
label(data$dipper)="Overnight dipper? drop of 15bpm for at least 1 hour"
label(data$watchpat_complete)="Complete?"
label(data$ab40_avg_conc)="Ab40 Average Concentration (pg/mL)"
label(data$ab42_avg_conc)="Ab42 Average Concentration (pg/mL)"
label(data$tau_avg_conc)="Tau Average Concentration (pg/mL)"
label(data$nfl_avg_conc)="NFL Average Concentration (pg/mL)"
label(data$gfap_avg_conc)="GFAP Average Concentration (pg/mL)"
label(data$ptau_181_avg_conc)="pTau 181 Average Concentration (pg/mL)"
label(data$ptau_217_avg_conc)="pTau 217 Average Concentration (pg/mL)"
label(data$biomarkers_complete)="Complete?"
label(data$pavel_wbc)="White Blood Cells"
label(data$pavel_lymphocyte)="Lymphocyte"
label(data$pavel_neutrophil)="Neutrophils"
label(data$pavel_monocyte)="Monocyte"
label(data$pavel_rbc)="Red Blood Cells"
label(data$pavel_hgb)="Hemoglobin"
label(data$pavel_hct)="Hematocrit"
label(data$pavel_plt)="Platelets"
label(data$pavel_mpv)="Mean Platelet Volume"
label(data$pac1_baseline)="Pac1 Baseline"
label(data$pac1_adp_1_25_10_min)="Pac1 ADP 1.25 10 min"
label(data$pac1_th_0_075_5_min)="Pac1 TH 0.075 5 min"
label(data$pac1_th_0_075_2_min)="Pac1 TH 0.075 2 min"
label(data$pac1_conv_50_3_min)="Pac1 Conv 50 3 min"
label(data$psel_baseline)="Psel Baseline"
label(data$psel_adp_1_25_10_min)="Psel ADP 1.25 10 min"
label(data$psel_th_0_075_5_min)="Psel TH 0.075 5 min"
label(data$psel_th_0_075_2_min)="Psel TH 0.075 2 min"
label(data$psel_conv_50_3_min)="Psel Conv 50 3 min"
label(data$blac_baseline)="BLAC Baseline"
label(data$blac_adp_1_25_10_min)="BLAC ADP 1.25 10 min"
label(data$blac_th_0_075_5_min)="BLAC TH 0.075 5 min"
label(data$blac_th_0_075_2_min)="BLAC TH 0.075 2 min"
label(data$blac_conv_50_3_min)="BLAC Conv 50 3 min"
label(data$pavel_labs_complete)="Complete?"
#Setting Units


#Setting Factors(will create new variable for factors)
data$redcap_event_name.factor = factor(data$redcap_event_name,levels=c("screening_arm_1","study_visits_arm_1"))
data$redcap_repeat_instrument.factor = factor(data$redcap_repeat_instrument,levels=c(""))
data$study_visit.factor = factor(data$study_visit,levels=c("1","2","3"))
data$current_db_dx.factor = factor(data$current_db_dx,levels=c("1","0"))
data$diabetes_med___1.factor = factor(data$diabetes_med___1,levels=c("0","1"))
data$diabetes_med___2.factor = factor(data$diabetes_med___2,levels=c("0","1"))
data$diabetes_med___3.factor = factor(data$diabetes_med___3,levels=c("0","1"))
data$insulin_type___1.factor = factor(data$insulin_type___1,levels=c("0","1"))
data$insulin_type___2.factor = factor(data$insulin_type___2,levels=c("0","1"))
data$diabetes_med_other___1.factor = factor(data$diabetes_med_other___1,levels=c("0","1"))
data$diabetes_med_other___2.factor = factor(data$diabetes_med_other___2,levels=c("0","1"))
data$diabetes_med_other___3.factor = factor(data$diabetes_med_other___3,levels=c("0","1"))
data$diabetes_med_other___4.factor = factor(data$diabetes_med_other___4,levels=c("0","1"))
data$hypertension.factor = factor(data$hypertension,levels=c("1","2"))
data$hypertenstion_controlled.factor = factor(data$hypertenstion_controlled,levels=c("1","0"))
data$htn_meds.factor = factor(data$htn_meds,levels=c("1","0"))
data$htn_med_type___1.factor = factor(data$htn_med_type___1,levels=c("0","1"))
data$htn_med_type___2.factor = factor(data$htn_med_type___2,levels=c("0","1"))
data$htn_med_type___3.factor = factor(data$htn_med_type___3,levels=c("0","1"))
data$htn_med_type___4.factor = factor(data$htn_med_type___4,levels=c("0","1"))
data$htn_med_type___5.factor = factor(data$htn_med_type___5,levels=c("0","1"))
data$htn_med_type___6.factor = factor(data$htn_med_type___6,levels=c("0","1"))
data$addl_dx.factor = factor(data$addl_dx,levels=c("1","0"))
data$addl_cardmet___1.factor = factor(data$addl_cardmet___1,levels=c("0","1"))
data$addl_cardmet___2.factor = factor(data$addl_cardmet___2,levels=c("0","1"))
data$addl_cardmet___3.factor = factor(data$addl_cardmet___3,levels=c("0","1"))
data$addl_chf.factor = factor(data$addl_chf,levels=c("1","0"))
data$addl_chf_meds___1.factor = factor(data$addl_chf_meds___1,levels=c("0","1"))
data$addl_chf_meds___2.factor = factor(data$addl_chf_meds___2,levels=c("0","1"))
data$addl_chf_meds___3.factor = factor(data$addl_chf_meds___3,levels=c("0","1"))
data$addl_chf_meds___4.factor = factor(data$addl_chf_meds___4,levels=c("0","1"))
data$addl_chf_meds___5.factor = factor(data$addl_chf_meds___5,levels=c("0","1"))
data$addl_hld.factor = factor(data$addl_hld,levels=c("1","0"))
data$addl_hld_meds___1.factor = factor(data$addl_hld_meds___1,levels=c("0","1"))
data$addl_hld_meds___2.factor = factor(data$addl_hld_meds___2,levels=c("0","1"))
data$addl_hld_meds___3.factor = factor(data$addl_hld_meds___3,levels=c("0","1"))
data$addl_hld_meds___4.factor = factor(data$addl_hld_meds___4,levels=c("0","1"))
data$meds_weight.factor = factor(data$meds_weight,levels=c("1","0"))
data$meds_weight_type___1.factor = factor(data$meds_weight_type___1,levels=c("0","1"))
data$meds_weight_type___2.factor = factor(data$meds_weight_type___2,levels=c("0","1"))
data$apnea.factor = factor(data$apnea,levels=c("1","0"))
data$apnea_cpap.factor = factor(data$apnea_cpap,levels=c("1","0"))
data$patient_medications_complete.factor = factor(data$patient_medications_complete,levels=c("0","1","2"))
data$study_visit_accel.factor = factor(data$study_visit_accel,levels=c("1","2","3"))
data$accelerometry_complete.factor = factor(data$accelerometry_complete,levels=c("0","1","2"))
data$study_visit_mttt.factor = factor(data$study_visit_mttt,levels=c("1","2","3"))
data$mmtt_vitals.factor = factor(data$mmtt_vitals,levels=c("1","0"))
data$mmtt_pregnant.factor = factor(data$mmtt_pregnant,levels=c("1","2"))
data$mmtt_brmr.factor = factor(data$mmtt_brmr,levels=c("1","0"))
data$mmtt_60rmr.factor = factor(data$mmtt_60rmr,levels=c("1","0"))
data$mmtt_base_labs.factor = factor(data$mmtt_base_labs,levels=c("1","0"))
data$mmtt_ffa_labs.factor = factor(data$mmtt_ffa_labs,levels=c("1","0"))
data$mmtt_insulin.factor = factor(data$mmtt_insulin,levels=c("1","0"))
data$mmtt_glp1.factor = factor(data$mmtt_glp1,levels=c("1","0"))
data$mmtt_cpep.factor = factor(data$mmtt_cpep,levels=c("1","0"))
data$mmtt_yy.factor = factor(data$mmtt_yy,levels=c("1","0"))
data$mmtt_glucagon.factor = factor(data$mmtt_glucagon,levels=c("1","0"))
data$mmtt_gluc.factor = factor(data$mmtt_gluc,levels=c("1","0"))
data$mmtt_metabolic_cart_complete.factor = factor(data$mmtt_metabolic_cart_complete,levels=c("0","1","2"))
data$study_visit_bodycomp.factor = factor(data$study_visit_bodycomp,levels=c("1","2","3"))
data$dxa_complete.factor = factor(data$dxa_complete,levels=c("1","0"))
data$bodpod_complete.factor = factor(data$bodpod_complete,levels=c("1","0"))
data$bia_complete.factor = factor(data$bia_complete,levels=c("1","0"))
data$tanita_scale_complete.factor = factor(data$tanita_scale_complete,levels=c("1","0"))
data$body_composition_dxa_bod_pod_complete.factor = factor(data$body_composition_dxa_bod_pod_complete,levels=c("0","1","2"))
data$study_visit_clamp.factor = factor(data$study_visit_clamp,levels=c("1","2","3"))
data$baseline.factor = factor(data$baseline,levels=c("1","0"))
data$fasting_labs.factor = factor(data$fasting_labs,levels=c("1","0"))
data$ns_bolus.factor = factor(data$ns_bolus,levels=c("1","0"))
data$ns_number.factor = factor(data$ns_number,levels=c("1","2","3"))
data$urine_labs.factor = factor(data$urine_labs,levels=c("1","0"))
data$bg_labs.factor = factor(data$bg_labs,levels=c("1","0"))
data$cpep_lab.factor = factor(data$cpep_lab,levels=c("1","0"))
data$insulin_labs.factor = factor(data$insulin_labs,levels=c("1","0"))
data$iohexol_bolus.factor = factor(data$iohexol_bolus,levels=c("1","0"))
data$pah_bolus.factor = factor(data$pah_bolus,levels=c("1","0"))
data$hct_lab.factor = factor(data$hct_lab,levels=c("1","0"))
data$clamp_complete.factor = factor(data$clamp_complete,levels=c("0","1","2"))
data$study_visit_outcomes.factor = factor(data$study_visit_outcomes,levels=c("1","2","3"))
data$kidney_outcomes.factor = factor(data$kidney_outcomes,levels=c("1","0"))
data$metab_outcomes.factor = factor(data$metab_outcomes,levels=c("1","0"))
data$asl_outcomes.factor = factor(data$asl_outcomes,levels=c("1","0"))
data$bold_outcomes.factor = factor(data$bold_outcomes,levels=c("1","0"))
data$adc_outcomes.factor = factor(data$adc_outcomes,levels=c("1","0"))
data$tkv_outcomes.factor = factor(data$tkv_outcomes,levels=c("1","0"))
data$outcomes_complete.factor = factor(data$outcomes_complete,levels=c("0","1","2"))
data$cortex_yn.factor = factor(data$cortex_yn,levels=c("1","0"))
data$cortex_diagnosis___1.factor = factor(data$cortex_diagnosis___1,levels=c("0","1"))
data$cortex_diagnosis___2.factor = factor(data$cortex_diagnosis___2,levels=c("0","1"))
data$glom_enlarge___1.factor = factor(data$glom_enlarge___1,levels=c("0","1"))
data$glom_enlarge___2.factor = factor(data$glom_enlarge___2,levels=c("0","1"))
data$glom_enlarge___3.factor = factor(data$glom_enlarge___3,levels=c("0","1"))
data$glom_enlarge___4.factor = factor(data$glom_enlarge___4,levels=c("0","1"))
data$glom_enlarge___5.factor = factor(data$glom_enlarge___5,levels=c("0","1"))
data$glom_enlarge___6.factor = factor(data$glom_enlarge___6,levels=c("0","1"))
data$glom_enlarge___7.factor = factor(data$glom_enlarge___7,levels=c("0","1"))
data$vessels___1.factor = factor(data$vessels___1,levels=c("0","1"))
data$vessels___2.factor = factor(data$vessels___2,levels=c("0","1"))
data$biopsy_results_michigan_complete.factor = factor(data$biopsy_results_michigan_complete,levels=c("0","1","2"))
data$watchpat_yes_no.factor = factor(data$watchpat_yes_no,levels=c("1","0"))
data$study_visit_accel1.factor = factor(data$study_visit_accel1,levels=c("1","2","3"))
data$dipper.factor = factor(data$dipper,levels=c("1","2"))
data$watchpat_complete.factor = factor(data$watchpat_complete,levels=c("0","1","2"))
data$biomarkers_complete.factor = factor(data$biomarkers_complete,levels=c("0","1","2"))
data$pavel_labs_complete.factor = factor(data$pavel_labs_complete,levels=c("0","1","2"))

levels(data$redcap_event_name.factor)=c("Screening","Study Visits")
levels(data$redcap_repeat_instrument.factor)=c("")
levels(data$study_visit.factor)=c("Pre-Surgery","3 Months Post-Surgery","12 Months Post-Surgery")
levels(data$current_db_dx.factor)=c("Yes","No")
levels(data$diabetes_med___1.factor)=c("Unchecked","Checked")
levels(data$diabetes_med___2.factor)=c("Unchecked","Checked")
levels(data$diabetes_med___3.factor)=c("Unchecked","Checked")
levels(data$insulin_type___1.factor)=c("Unchecked","Checked")
levels(data$insulin_type___2.factor)=c("Unchecked","Checked")
levels(data$diabetes_med_other___1.factor)=c("Unchecked","Checked")
levels(data$diabetes_med_other___2.factor)=c("Unchecked","Checked")
levels(data$diabetes_med_other___3.factor)=c("Unchecked","Checked")
levels(data$diabetes_med_other___4.factor)=c("Unchecked","Checked")
levels(data$hypertension.factor)=c("Yes","No")
levels(data$hypertenstion_controlled.factor)=c("Yes","No")
levels(data$htn_meds.factor)=c("Yes","No")
levels(data$htn_med_type___1.factor)=c("Unchecked","Checked")
levels(data$htn_med_type___2.factor)=c("Unchecked","Checked")
levels(data$htn_med_type___3.factor)=c("Unchecked","Checked")
levels(data$htn_med_type___4.factor)=c("Unchecked","Checked")
levels(data$htn_med_type___5.factor)=c("Unchecked","Checked")
levels(data$htn_med_type___6.factor)=c("Unchecked","Checked")
levels(data$addl_dx.factor)=c("Yes","No")
levels(data$addl_cardmet___1.factor)=c("Unchecked","Checked")
levels(data$addl_cardmet___2.factor)=c("Unchecked","Checked")
levels(data$addl_cardmet___3.factor)=c("Unchecked","Checked")
levels(data$addl_chf.factor)=c("Yes","No")
levels(data$addl_chf_meds___1.factor)=c("Unchecked","Checked")
levels(data$addl_chf_meds___2.factor)=c("Unchecked","Checked")
levels(data$addl_chf_meds___3.factor)=c("Unchecked","Checked")
levels(data$addl_chf_meds___4.factor)=c("Unchecked","Checked")
levels(data$addl_chf_meds___5.factor)=c("Unchecked","Checked")
levels(data$addl_hld.factor)=c("Yes","No")
levels(data$addl_hld_meds___1.factor)=c("Unchecked","Checked")
levels(data$addl_hld_meds___2.factor)=c("Unchecked","Checked")
levels(data$addl_hld_meds___3.factor)=c("Unchecked","Checked")
levels(data$addl_hld_meds___4.factor)=c("Unchecked","Checked")
levels(data$meds_weight.factor)=c("Yes","No")
levels(data$meds_weight_type___1.factor)=c("Unchecked","Checked")
levels(data$meds_weight_type___2.factor)=c("Unchecked","Checked")
levels(data$apnea.factor)=c("Yes","No")
levels(data$apnea_cpap.factor)=c("Yes","No")
levels(data$patient_medications_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$study_visit_accel.factor)=c("Pre-Surgery","3 Months Post-Surgery","12 Months Post-Surgery")
levels(data$accelerometry_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$study_visit_mttt.factor)=c("Pre-Surgery","3 Months Post-Surgery","12 Months Post-Surgery")
levels(data$mmtt_vitals.factor)=c("Yes","No")
levels(data$mmtt_pregnant.factor)=c("NOT Pregnant","Pregnant")
levels(data$mmtt_brmr.factor)=c("Yes","No")
levels(data$mmtt_60rmr.factor)=c("Yes","No")
levels(data$mmtt_base_labs.factor)=c("Yes","No")
levels(data$mmtt_ffa_labs.factor)=c("Yes","No")
levels(data$mmtt_insulin.factor)=c("Yes","No")
levels(data$mmtt_glp1.factor)=c("Yes","No")
levels(data$mmtt_cpep.factor)=c("Yes","No")
levels(data$mmtt_yy.factor)=c("Yes","No")
levels(data$mmtt_glucagon.factor)=c("Yes","No")
levels(data$mmtt_gluc.factor)=c("Yes","No")
levels(data$mmtt_metabolic_cart_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$study_visit_bodycomp.factor)=c("Pre-Surgery","3 Months Post-Surgery","12 Months Post-Surgery")
levels(data$dxa_complete.factor)=c("Yes","No")
levels(data$bodpod_complete.factor)=c("Yes","No")
levels(data$bia_complete.factor)=c("Yes","No")
levels(data$tanita_scale_complete.factor)=c("Yes","No")
levels(data$body_composition_dxa_bod_pod_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$study_visit_clamp.factor)=c("Pre-Surgery","3 Months Post-Surgery","12 Months Post-Surgery")
levels(data$baseline.factor)=c("Yes","No")
levels(data$fasting_labs.factor)=c("Yes","No")
levels(data$ns_bolus.factor)=c("Yes","No")
levels(data$ns_number.factor)=c("1","2","3")
levels(data$urine_labs.factor)=c("Yes","No")
levels(data$bg_labs.factor)=c("Yes","No")
levels(data$cpep_lab.factor)=c("Yes","No")
levels(data$insulin_labs.factor)=c("Yes","No")
levels(data$iohexol_bolus.factor)=c("Yes","No")
levels(data$pah_bolus.factor)=c("Yes","No")
levels(data$hct_lab.factor)=c("Yes","No")
levels(data$clamp_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$study_visit_outcomes.factor)=c("Pre-Surgery","3 Months Post-Surgery","12 Months Post-Surgery")
levels(data$kidney_outcomes.factor)=c("Yes","No")
levels(data$metab_outcomes.factor)=c("Yes","No")
levels(data$asl_outcomes.factor)=c("Yes","No")
levels(data$bold_outcomes.factor)=c("Yes","No")
levels(data$adc_outcomes.factor)=c("Yes","No")
levels(data$tkv_outcomes.factor)=c("Yes","No")
levels(data$outcomes_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$cortex_yn.factor)=c("Yes","No")
levels(data$cortex_diagnosis___1.factor)=c("Unchecked","Checked")
levels(data$cortex_diagnosis___2.factor)=c("Unchecked","Checked")
levels(data$glom_enlarge___1.factor)=c("Unchecked","Checked")
levels(data$glom_enlarge___2.factor)=c("Unchecked","Checked")
levels(data$glom_enlarge___3.factor)=c("Unchecked","Checked")
levels(data$glom_enlarge___4.factor)=c("Unchecked","Checked")
levels(data$glom_enlarge___5.factor)=c("Unchecked","Checked")
levels(data$glom_enlarge___6.factor)=c("Unchecked","Checked")
levels(data$glom_enlarge___7.factor)=c("Unchecked","Checked")
levels(data$vessels___1.factor)=c("Unchecked","Checked")
levels(data$vessels___2.factor)=c("Unchecked","Checked")
levels(data$biopsy_results_michigan_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$watchpat_yes_no.factor)=c("Yes","No")
levels(data$study_visit_accel1.factor)=c("Pre-Surgery","3 Months Post-Surgery","12 Months Post-Surgery")
levels(data$dipper.factor)=c("Yes","No")
levels(data$watchpat_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$biomarkers_complete.factor)=c("Incomplete","Unverified","Complete")
levels(data$pavel_labs_complete.factor)=c("Incomplete","Unverified","Complete")










}

visit <- data
data <- NULL
# select the participants
visit <- visit %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_10","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20","IT_22"))
# FOR THIS ANALYSIS, USE BASELINE MONTH 3, MONTH 12 DATA
#visit <- visit %>% filter(redcap_repeat_instance %in% c(1,2))
visit$visit <- as.factor(visit$redcap_repeat_instance)
visit$log_clamp_acr_baseline <- log(visit$clamp_acr_baseline)

# calculate glucose AUC

if(user == 'laura'){
  source("/Users/pylell/Documents/GitHub/General-code/iauc.r")
}else if(user == 'dylan'){
  source("/Users/netio/Documents/Programs/DECODE_Demographics/Shared_Github/CHCO-Code/Kristen Nadeau/IMPROVE-T2D/iAUC.R")
}

g <- visit[!(visit$subject_id == "IT_22" & visit$redcap_repeat_instance == 2), c("subject_id","redcap_repeat_instance","mmtt_bg_0","mmtt_bg_10","mmtt_bg_20","mmtt_bg_30","mmtt_bg_45","mmtt_bg_60",
                                 "mmtt_bg_90","mmtt_bg_120","mmtt_bg_150","mmtt_bg_180","mmtt_bg_210",
                                 "mmtt_bg_240")]
g <- g %>% filter(!(subject_id == "IT_10" & redcap_repeat_instance == 2))
g$auc_glucose <- iAUC_glucose(data=g,
                                  thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                          cols=c("mmtt_bg_0","mmtt_bg_10","mmtt_bg_20","mmtt_bg_30","mmtt_bg_45","mmtt_bg_60",
                                 "mmtt_bg_90","mmtt_bg_120","mmtt_bg_150","mmtt_bg_180","mmtt_bg_210",
                                 "mmtt_bg_240"))
g <- g %>% select(subject_id, redcap_repeat_instance, auc_glucose)
visit <- left_join(visit, g, by = c("subject_id", "redcap_repeat_instance"))

# calculate insulin AUC
for (i in 1:nrow(visit)){
  visit$auc_insulin[i] <- ifelse (!is.na(visit[i, "mmtt_insulin_0"]), 
                                   iAUC_insulin(data = visit[i,], thresh=0,type="linear",times=c(0,10,20,30,45,60,90,120,150,180,210,240),
                                    cols=c("mmtt_insulin_0","mmtt_insulin_10","mmtt_insulin_20","mmtt_insulin_30","mmtt_insulin_45","mmtt_insulin_60",
                                    "mmtt_insulin_90","mmtt_insulin_120","mmtt_insulin_150","mmtt_insulin_180","mmtt_insulin_210",
                                    "mmtt_insulin_240")), 
                                   NA) 
}

# calculate HOMA-IR
visit$homa_ir <- (visit$mmtt_bg_0*visit$mmtt_insulin_0)/405

# calculate HOMA-%B
visit$homa_b <- (360*visit$mmtt_insulin_0) / (visit$mmtt_bg_0 - 63)

# calculate Matsuda Index
visit$matsuda <- 10000/(sqrt(visit$mmtt_bg_0*visit$mmtt_insulin_0*
                  (((15*visit$mmtt_insulin_0)+(30*visit$mmtt_insulin_30)+(30*visit$mmtt_insulin_60)+(30*visit$mmtt_insulin_90)+(15*visit$mmtt_insulin_120))/120)*
                  (((15*visit$mmtt_bg_0)+(30*visit$mmtt_bg_30)+(30*visit$mmtt_bg_60)+(30*visit$mmtt_bg_90)+(15*visit$mmtt_bg_120))/120)))

# calculate Disposition Index
visit$isi <- (visit$mmtt_insulin_30 - visit$mmtt_insulin_0) / (visit$mmtt_bg_30-visit$mmtt_bg_0)
visit$odi <- (1/visit$mmtt_insulin_0) * visit$isi
visit$odi <- ifelse(visit$odi < 0, NA, visit$odi)

# calculate BMI
visit$mmtt_bmi <- visit$mmtt_wt / ((visit$mmtt_ht/100)^2)

# calculate 60 - 0 min RQ
visit$rmr_rq_60_minus_0 <- visit$srmr_rq - visit$brmr_rq

# calculate mean kidney volume
visit$volume_mean <- rowMeans(visit[,c("volume_right","volume_left")])

# calculate mean ACR during clamp
visit$clamp_acr_mean <- rowMeans(visit[,c("clamp_acr_baseline","clamp_acr_250")])

# calculate weight loss
wtloss <- visit %>% select(subject_id, visit, mmtt_wt) 
wtloss_wide <- reshape(wtloss, direction = "wide", idvar = "subject_id", timevar = "visit")
wtloss_wide$wtloss <- wtloss_wide$mmtt_wt.1 - wtloss_wide$mmtt_wt.3
wtloss_wide$per_wtloss <- (wtloss_wide$wtloss / wtloss_wide$mmtt_wt.1) * 100
wtloss_wide <- wtloss_wide %>% select(subject_id, wtloss, per_wtloss)
visit <- left_join(visit, wtloss_wide, by="subject_id")

# rename variables that are also in clamp data
visit <- visit %>% rename(serum_creatinine_visit = serum_creatinine,
                          cystatin_c_visit = cystatin_c
                          )

# read in clamp data
if(user == 'laura'){
  clamp <- read.csv('/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/IMPROVE T2D/Data_Raw/Primary analyses/IMPROVET2D-ClampLP_DATA_2025-04-25_0833.csv', na.strings = c(""," ","","-9999"))
}else if(user == 'dylan'){
  clamp <- read.csv('/Users/netio/Documents/UofW/Projects/IMPROVE-CU/IMPROVET2D-ClampLP_DATA_2025-09-24_1130.csv', na.strings = c(""," ","","-9999"))
}

clamp <- clamp %>% rename(serum_creatinine_clamp = serum_creatinine, 
                          cystatin_c_clamp = cystatin_c)
visit$subject_id <- crosstable::remove_labels(visit$subject_id)
visit$redcap_repeat_instance <- crosstable::remove_labels(visit$redcap_repeat_instance)
visit <- left_join(visit, clamp, by = c("subject_id", "redcap_repeat_instance"))

# set 3 month values to missing for several outcomes
visit$fasting_glu <- ifelse(visit$study_visit==2, NA, visit$fasting_glu)

# read in MRI data
if(user == 'laura'){
  mri <- read.csv('/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/IMPROVE T2D/Data_Raw/Primary analyses/IMPROVET2D-CardiacabdominalMRIL_DATA_2025-02-18_1421.csv', na.strings = c(""," ","","-9999"))
}else if(user == 'dylan'){
  mri <- read.csv('/Users/netio/Documents/UofW/Projects/IMPROVE-CU/IMPROVET2D-CardiacabdominalMRIL_DATA_2025-09-24_1130.csv', na.strings = c(""," ","","-9999"))
}

visit <- left_join(visit, mri, by = c("subject_id", "redcap_repeat_instance"))

# calculate additional kidney variables
visit$kfg <- visit$gfr/visit$rpf

###################################################################
# next section is converted from Python harmonization
# have not updated variable names
# only a few non-missing values at visit 3 - will check with Tyler
# need to update later

# # Filtration Fraction
# out$ff <- out$gfr_raw_plasma/out$erpf_raw_plasma 
# 
# 
# # Filtration pressure across glomerular capillaries
# out$deltapf <- (out$gfr_raw_plasma/60)/out$kfg 
# 
# # Plasma protein mean concentration
# out$cm <- (out$total_protein/out$ff)*log(1/(1-out$ff)) 
# 
# # Pi G (Oncotic pressure)
# out$pg <- 5*(out$cm-2)
# 
# # Glomerular Pressure
# out$glomerular_pressure <- out$pg + out$deltapf + 10
# 
# # Renal Blood Flow
# out$rbf <- (out$erpf_raw_plasma) / (1 - out$hct/100)
# out$rbf_seconds <- (out$erpf_raw_plasma_seconds) / (1 - out$hct/100)
# 
# # Renal Vascular Resistance (mmHg*l^-1*min^-1)
# out$rvr <- out$map / out$rbf
# 
# # Efferent Arteriolar Resistance 
# out$re <- (out$gfr_raw_plasma_seconds) / (out$kfg * (out$rbf_seconds - (out$gfr_raw_plasma_seconds))) * 1328
# 
# # Afferent Arteriolar Resistance
# out$ra <- ((out$map - out$glomerular_pressure) / out$rbf_seconds) * 1328    
# out$ra[!(out$ra > 0)] <- NA    

#########################################################


# read in medication data
# metformin, insulin, SGLT2, GLP-1
if(user == 'laura'){
  meds <- read.csv('/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/IMPROVE T2D/Data_Raw/Primary analyses/IMPROVET2D-Medications_DATA_2025-02-18_1436.csv', na.strings = c(""," ","","-9999"))
}else if(user == 'dylan'){
  meds <- read.csv('/Users/netio/Documents/UofW/Projects/IMPROVE-CU/IMPROVET2D-Medications_DATA_2025-09-24_1129.csv', na.strings = c(""," ","","-9999"))
}

meds$metformin <- ifelse(meds$diabetes_med___1 == 1, "Yes","No")
meds$insulin <- ifelse(meds$diabetes_med___2 == 1, "Yes","No")
meds$sglt2 <- ifelse(meds$diabetes_med_other___3 == 1, "Yes","No")
meds$glp1 <- ifelse(meds$diabetes_med_other___2 == 1, "Yes","No")
meds$metformin_num <- ifelse(meds$diabetes_med___1 == 1, 1, 0)
meds$insulin_num <- ifelse(meds$diabetes_med___2 == 1, 1, 0)
meds$sglt2_num <- ifelse(meds$diabetes_med_other___3 == 1, 1, 0)
meds$glp1_num <- ifelse(meds$diabetes_med_other___2 == 1, 1, 0)
# select the participants
#meds <- meds %>% filter(subject_id %in% c("IT_02","IT_03","IT_05","IT_06","IT_08","IT_10","IT_11","IT_12","IT_13","IT_14","IT_15","IT_16","IT_18","IT_19","IT_20","IT_22"))
```

# Background

This analysis examines changes in outcomes from pre-surgery to the 12-month post-surgery visit in the IMPROVE study.

# Methods

Repeated measures mixed-effects models with a compound-symmetric correlations structure were used to examine changes from baseline to month 3. Unadjusted models and models adjusted for weight loss were considered.

# Results

## Descriptive statistics

```{r create_t1, include=FALSE}
t1 <- tableby( ~ age_current +  gender.factor + race_eth + hba1c + sys_bp + dys_bp + as.factor(sglt2i) + screen_height + 
                 screen_weight + screen_bmi + waist_circumference + length_of_diabetes + diabetes_age, data = screen)
```

```{r t1, results='asis'}
summary(t1)
```

## Medications

Models for medication variables would not converge because so few participants were on any medications at the 3-month and 12-month visits.

```{r create_medtable, include=FALSE}
medtable <- tableby(study_visit ~ metformin + insulin + sglt2 + glp1, data = meds, test=F)
```

```{r medtable, results='asis'}
summary(medtable)
```

```{r medsgee}
#meds_met_complete <- meds %>% select(subject_id, study_visit, metformin_num)
#meds_met_complete_ids <- meds %>% filter(study_visit == 2) %>% select(subject_id)
#meds_met_complete <- meds_met_complete %>% filter(subject_id %in% meds_met_complete_ids$subject_id)
#meds_gee_met <- gee(metformin_num ~ study_visit,
#               data = meds_met_complete, 
#               id = as.factor(subject_id), 
#               family = binomial,
#               corstr = "AR-M", Mv = 1)
```

## Unadjusted models

### Weight

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_wt")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### BMI

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_bmi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Waist circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_waist")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Hip circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hip")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Pulse

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### SBP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_sys_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### DBP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_dia_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### MAP

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_map")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### ALT

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_gptalt_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### AST

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_gotast_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### BUN

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_bun_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_creat_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Total cholesterol

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_chol_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HDL

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hdl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### LDL

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_ldl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Triglycerides

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_tri_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HbA1c

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mmtt_hba1c_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_vco2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "brmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_co2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "srmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Time 60-0 RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rmr_rq_60_minus_0")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Percent body fat

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_body_fat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Percent lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_lean_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Fat mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_fat_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "dxa_lean_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HOMA-IR

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "homa_ir")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HOMA-%B

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "homa_b")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Matsuda Index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "matsuda")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Disposition Index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "odi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp fasting glucose

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "fasting_glu")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp fasting insulin

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "fasting_insulin")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp fasting C-peptide

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "fasting_cpep")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp ACPRg

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "acpr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp AIRg

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "air")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp steady state C-peptide

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "ss_c_pep")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp steady state insulin response

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "ss_insulin_response")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp ACPR max

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "acpr_max")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp AIR max

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "air_max")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp Glucose disposal

Missing values for all except one visit.

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "glu_disposal")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### Clamp M/I

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "m_i")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp insulin clearance

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "ins_clear")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp DI

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "di")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvedv

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mri_lvedv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvedvi

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mri_lvedvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvesv

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mri_lvesv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvesvi

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mri_lvesvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lv_hr

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "lv_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lv_stroke_volume

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "lv_stroke_volume")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lvef

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "lvef")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lved_mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "lved_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lved_massi

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "lved_massi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### lves_mass

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "lves_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lvesmassi

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "lvesmassi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### lv_cardiac_output

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "lv_cardiac_output")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lv_cardiac_index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "lv_cardiac_index")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_rvedv

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mri_rvedv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rmi_rvedvi

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rmi_rvedvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rmi_rvesv

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rmi_rvesv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_rvesvi

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "mri_rvesvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_hr

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rv_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_stroke_volume

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rv_stroke_volume")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_ejection_fraction

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rv_ejection_fraction")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_cardiac_output

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rv_cardiac_output")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_cardiac_index

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "rv_cardiac_index")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### pat_sys

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "pat_sys")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### pat_dias

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "pat_dias")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### eat_sys

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "eat_sys")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### eat_dias

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "eat_dias")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_net_flow

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "aorta_net_flow")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_forward_flow

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "aorta_forward_flow")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_reverse_flow

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "aorta_reverse_flow")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_regurgitant_fraction

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "aorta_regurgitant_fraction")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_qp_qs

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "aorta_qp_qs")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_pwv

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "aa_pwv")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_pwv

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "da_pwv")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_wss_ta

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "aa_wss_ta")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_wss_ta

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "da_wss_ta")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_wss_max

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "aa_wss_max")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_wss_max

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "da_wss_max")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_rac

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "aa_rac")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_rac

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "da_rac")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_osi

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "aa_osi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_osi

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "da_osi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### abd_mean_t2

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "abd_mean_t2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### abd_elasto

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "abd_elasto")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### abd_stiffness_sd

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "abd_stiffness_sd")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### abd_fat_frac

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "abd_fat_frac")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### res_abd

```{r echo=FALSE, include=TRUE}
#mod <- mixed(data = visit, outcome = "res_abd")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### sat

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "sat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### vat

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "vat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### vat_sat_ratio

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "vat_sat_ratio")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### hepatic_ff

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "hepatic_ff")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### panc_ff

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "panc_ff")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Log(UACR)

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "log_clamp_acr_baseline")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Serum creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "serum_creatinine_visit")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Serum cystatin-c 

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "cystatin_c_visit")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### GFR

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "gfr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### GFR - BSA corrected

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "gfr_bsa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### GFR - weight adjusted

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo_weight(data = visit, outcome = "gfr_bsa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### RPF

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "rpf")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### eRPF_BSA

```{r echo=FALSE, include=TRUE}
mod <- mixed_no3mo(data = visit, outcome = "erpf_bsa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Glomerular pressure

```{r echo=FALSE, include=TRUE}
#mod <- mixed_no3mo(data = visit, outcome = "glomerular_pressure")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### Afferent Arteriolar Resistance

```{r echo=FALSE, include=TRUE}
#mod <- mixed_no3mo(data = visit, outcome = "aff_arteriolar_resistance")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### Efferent Arteriolar Resistance

```{r echo=FALSE, include=TRUE}
#mod <- mixed_no3mo(data = visit, outcome = "eff_arteriolar_resistance")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### bold_r_bl_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_r_bl_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_bl_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_r_bl_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_bl_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_r_bl_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_pf_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_r_pf_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_pf_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_r_pf_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_pf_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_r_pf_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_bl_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_l_bl_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_bl_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_l_bl_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_bl_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_l_bl_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_pf_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_l_pf_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_pf_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_l_pf_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_pf_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "bold_l_pf_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### adc_right

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "adc_right")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### adc_left

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "adc_left")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### volume_right

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "volume_right")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### volume_left

```{r echo=FALSE, include=TRUE}
mod <- mixed(data = visit, outcome = "volume_left")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

## Models adjusted for percent weight loss

### Waist circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_waist")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Hip circumference

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hip")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Pulse

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### SBP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_sys_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### DBP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_dia_bp")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### MAP

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_map")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### ALT

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_gptalt_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### AST

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_gotast_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### BUN

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_bun_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_creat_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Total cholesterol

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_chol_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HDL

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hdl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### LDL

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_ldl_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Triglycerides

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_tri_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HbA1c

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mmtt_hba1c_base")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_vco2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 0 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "brmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute VO2 (ml/kg.min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_vo2_ml")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute VO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_vo2_l")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute VCO2 (L/min)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_co2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_rq")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute REE

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_ree")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute Fat oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_fo")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### 60 minute Carb oxidation

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "srmr_co")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Time 60-0 RQ

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rmr_rq_60_minus_0")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Percent body fat

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_body_fat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Percent lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_lean_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Fat mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_fat_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Lean mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "dxa_lean_kg")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HOMA-IR

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "homa_ir")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### HOMA-%B

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "homa_b")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Matsuda Index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "matsuda")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Disposition Index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "odi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Height

### Serum sodium

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "serum_sodium")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp fasting glucose

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "fasting_glu")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp fasting insulin

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "fasting_insulin")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp fasting C-peptide

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "fasting_cpep")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp ACPRg

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "acpr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp AIRg

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "air")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp steady state C-peptide

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "ss_c_pep")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp steady state insulin response

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "ss_insulin_response")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp ACPR max

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "acpr_max")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp AIR max

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "air_max")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp Glucose disposal

Missing values for all except one visit.

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "glu_disposal")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### Clamp M/I

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "m_i")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp insulin clearance

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "ins_clear")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Clamp DI

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "di")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvedv

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mri_lvedv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvedvi

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mri_lvedvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvesv

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mri_lvesv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_lvesvi

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mri_lvesvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lv_hr

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "lv_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lv_stroke_volume

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "lv_stroke_volume")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lvef

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "lvef")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lved_mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "lved_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lved_massi

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "lved_massi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### lves_mass

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "lves_mass")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lvesmassi

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "lvesmassi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### lv_cardiac_output

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "lv_cardiac_output")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### lv_cardiac_index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "lv_cardiac_index")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_rvedv

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mri_rvedv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rmi_rvedvi

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rmi_rvedvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rmi_rvesv

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rmi_rvesv")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### mri_rvesvi

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "mri_rvesvi")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_hr

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rv_hr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_stroke_volume

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rv_stroke_volume")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_ejection_fraction

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rv_ejection_fraction")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_cardiac_output

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rv_cardiac_output")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### rv_cardiac_index

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "rv_cardiac_index")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### pat_sys

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "pat_sys")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### pat_dias

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "pat_dias")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### eat_sys

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "eat_sys")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### eat_dias

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "eat_dias")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_net_flow

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "aorta_net_flow")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_forward_flow

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "aorta_forward_flow")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_reverse_flow

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "aorta_reverse_flow")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_regurgitant_fraction

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "aorta_regurgitant_fraction")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### aorta_qp_qs

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "aorta_qp_qs")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_pwv

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "aa_pwv")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_pwv

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "da_pwv")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_wss_ta

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "aa_wss_ta")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_wss_ta

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "da_wss_ta")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_wss_max

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "aa_wss_max")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_wss_max

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "da_wss_max")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_rac

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "aa_rac")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_rac

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "da_rac")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### aa_osi

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "aa_osi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### da_osi

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "da_osi")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### abd_mean_t2

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "abd_mean_t2")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### abd_elasto

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "abd_elasto")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### abd_stiffness_sd

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "abd_stiffness_sd")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### abd_fat_frac

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "abd_fat_frac")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### res_abd

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj(data = visit, outcome = "res_abd")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### sat

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "sat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### vat

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "vat")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### vat_sat_ratio

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "vat_sat_ratio")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### hepatic_ff

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "hepatic_ff")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### panc_ff

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "panc_ff")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Log(UACR)

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "log_clamp_acr_baseline")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Serum creatinine

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "serum_creatinine_visit")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Serum cystatin-c

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "cystatin_c_visit")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### GFR

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "gfr")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### GFR - BSA adjusted

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "gfr_bsa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### RPF

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "rpf")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### eRPF_BSA

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj_no3mo(data = visit, outcome = "erpf_bsa")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### Glomerular pressure

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj_no3mo(data = visit, outcome = "glomerular_pressure")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### Afferent Arteriolar Resistance

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj_no3mo(data = visit, outcome = "aff_arteriolar_resistance")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### Efferent Arteriolar Resistance

```{r echo=FALSE, include=TRUE}
#mod <- mixed_adj_no3mo(data = visit, outcome = "eff_arteriolar_resistance")
```

```{r echo=FALSE}
#kable(mod[[1]])
#kable(mod[[2]])
#mod[[3]]
```

### bold_r_bl_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_r_bl_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_bl_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_r_bl_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_bl_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_r_bl_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_pf_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_r_pf_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_pf_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_r_pf_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_r_pf_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_r_pf_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_bl_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_l_bl_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_bl_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_l_bl_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_bl_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_l_bl_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_pf_cortex

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_l_pf_cortex")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_pf_medulla

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_l_pf_medulla")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### bold_l_pf_kidney

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "bold_l_pf_kidney")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### adc_right

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "adc_right")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### adc_left

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "adc_left")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### volume_right

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "volume_right")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```

### volume_left

```{r echo=FALSE, include=TRUE}
mod <- mixed_adj(data = visit, outcome = "volume_left")
```

```{r echo=FALSE}
kable(mod[[1]])
kable(mod[[2]])
mod[[3]]
```


