---
title: "MTAP expression"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

```{r libraries, message = F, warning=F}
#| include: false
library(scran)
library(future)
library(future.apply)
library(tidyverse)
library(colorspace)
library(patchwork)
library(ggdendro)
library(cowplot)
library(ggpubr)
library(rstatix)
library(arsenal)
library(Biobase)
library(msigdbr)
library(kableExtra)
library(knitr)
library(REDCapR)
library(data.table)
library(emmeans)
library(NMF)
library(pheatmap)
library(UpSetR)
library(enrichR)
library(WriteXLS)
library(SAVER)
library(readxl)
library(limma)
library(edgeR)
library(BiocGenerics)
library(GSEABase)
library(slingshot)
library(SingleCellExperiment)
library(MAST)
library(muscat)
library(scater)
library(Seurat)
library(jsonlite)
library(dplyr)
library(glmmTMB)
library(reshape2)
library(broom.mixed)
library(nebula)
```

```{r include = F}
reticulate::use_python("/mmfs1/gscratch/scrubbed/yejichoi/mypy/bin/python", required = T)
reticulate::repl_python()
reticulate::py_config()
# set up Kopah connection
## Load boto3 and pandas
boto3 <- reticulate::import("boto3")
pd <- reticulate::import("pandas")

## Create an S3 client
keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
session <- boto3$session$Session(
  aws_access_key_id = keys$MY_ACCESS_KEY,
  aws_secret_access_key = keys$MY_SECRET_KEY
)

## Create an S3 client with the session
s3 <- session$client("s3", endpoint_url = "https://s3.kopah.uw.edu")
```

```{r echo = F, include=F}
# read PB90 Seurat file for CROCODILE
temp_file <- tempfile(fileext = ".rds")
s3$download_file("scrna", "Kidney transcriptomics/Single cell RNA seq/PB_90_RPCAFix_Metadata_LR_RM_kpmpV1labelled.rds", temp_file)
so <- readRDS(temp_file)

so <- readRDS("C:/Users/netio/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/scRNA/data_raw/PB_90_RPCAFix_Metadata_LR_RM_kpmpV1labelled.rds")

#Check the number of Mitochondrial genes to start
sum(grepl("^MT-", rownames(so))) 

# Identify mitochondrial genes (human: start with "MT-")
mito_genes <- grep("^MT-", rownames(so), value = TRUE)
so <- subset(so, features = setdiff(rownames(so), mito_genes))

#Check the number of Mitochondrial genes after filtering to ensure filtering step was successful
sum(grepl("^MT-", rownames(so))) #Should be 0


#Ribosomal genes
# Identify ribosomal genes
ribo_genes <- c(
  "RPL22", "RPL11", "RPS8", "RPL5", "RPS27", "RPS7", "RPS27A", "RPL31", "RPL37A", "RPL32", "RPL15", "RPL14", "RPL29",
  "RPL24", "RPL22L1", "RPL35A", "RPL9", "RPL34", "RPS3A", "RPL37", "RPS23", "RPS14", "RPS18", "RPS10", "RPL10A", 
  "RPS20", "RPL7", "RPL30", "RPL8", "RPS6", "RPL35", "RPL12", "RPL7A", "RPS24", "RPLP2", "RPL27A", "RPS13", "RPS3",
  "RPS25", "RPS26", "RPL41", "RPL6", "RPLP0", "RPL21", "RPS29", "RPL4", "RPLP1", "RPS17", "RPS2", "RPS15A", "RPL13",
  "RPL26", "RPL23A", "RPL23", "RPL19", "RPL27", "RPL38", "RPL17", "RPS15", "RPL36", "RPS28", "RPL18A", "RPS16", 
  "RPS19", "RPL18", "RPL13A", "RPS11", "RPS9", "RPL28", "RPS5", "RPS21", "RPL3", "RPS4X", "RPL36A", "RPL39", 
  "RPL10", "RPS4Y1")



jpeg('C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_KPMP_celltype.jpeg', width=900, 
     height=900)
DimPlot(so, reduction = "umap.harmony", group.by = "KPMP_celltype", raster = F, alpha = 0.2, label = T)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_celltype.jpeg", width=900, 
     height=900)
DimPlot(so, reduction = "umap.harmony", group.by = "celltype_harmony", raster = F, alpha = 0.2, label = T)
dev.off()

```

# HC vs. T1D in PB90

## All cell types combined

```{r echo = F}

cgas <- data.table::fread("C:/Users/netio/Documents/UofW/cGAS_Analysis/cGAS_genes_AmiGO2.txt",
                          header=F)
cgas[2] <- 'TMEM173'



so <- subset(so, 
                      subset = nFeature_RNA > 500 & nFeature_RNA < 5000 & percent.mt < 10)

so <- subset(s0)

counts_layer <- round(GetAssayData(so, layer = 'counts'))
library_size <- Matrix::colSums(round(GetAssayData(so, layer = 'counts')))
so$library_size <- library_size
sce <- SingleCellExperiment::SingleCellExperiment(assays = list(counts = counts_layer))
sce <- computeSumFactors(sce)
# View size factors
sizeFactors(sce)
## Calculate offset → (size factors)
so$pooled_offset <- (sizeFactors(sce))



so_t1d_hc <- subset(so, group %in% c("Type_1_Diabetes","Lean_Control"))

meta.data <- so_t1d_hc@meta.data

rm(so)
save.image('C:/Users/netio/Documents/UofW/cGAS_Analysis/cGWAS_sot1d_hc_data.RData')


jpeg('C:/Users/netio/Documents/UofW/cGAS_Analysis/CellDistributions.jpeg', height= 1500, width = 1200)
ggplot(meta.data %>% filter(celltype_rpca != 'ATL'), aes(x=group, fill=celltype_harmony))+
  geom_bar(position='fill')+theme_classic()+labs(x='Condition Group', fill='Cell Type')+
  theme(text = element_text(size=25))
dev.off()



jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_Full_T1D_LC.jpeg", width=900, 
     height=900)
DimPlot(so_t1d_hc, reduction = "umap.harmony", group.by = "celltype_harmony", raster = F, alpha = 0.2, label = T)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_MTAP.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc, reduction = "umap.harmony", features = "MTAP", raster = F)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_NOS3.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc, reduction = "umap.harmony", features = "NOS3", raster = F)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_NOS2.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc, reduction = "umap.harmony", features = "NOS2", raster = F)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_NOS1.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc, reduction = "umap.harmony", features = "NOS1", raster = F)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_cGWAS_STING.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc, reduction = "umap.harmony", features = unique(c(cgas$V1, 'TMEM173')), raster = F)
dev.off()


```
### Pseudoreplication (Treating each cell independently)
```{r echo = F}
FindMarkers(so_t1d_hc, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = "MTAP")

FindMarkers(so_t1d_hc, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = "NOS3")

FindMarkers(so_t1d_hc, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = cgas$V1)





```

### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}

counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary
 
write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_MTAP_analysis_nooffset.txt', row.names=F,
            quote=F, sep='\t')


counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_nos3 <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_nos3, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_NOS3_analysis_nooffset.txt', row.names=F,
            quote=F, sep='\t')




counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_cgas <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_cgas, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary


write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_CGAS_analysis_nooffset.txt', row.names=F,
            quote=F, sep='\t')


rm(list=c(counts_t1d_hc_nos3, counts_t1d_hc_cgas, counts_t1d_hc_mtap))


##NEBULA with Offsets 


#with offset


full_analysis <- FindVariableFeatures(so_t1d_hc, selection.method = "vst", nfeatures = 2000)
hvgs <- VariableFeatures(full_analysis)
full_analysis <- subset(so_t1d_hc, features = hvgs)
full_counts <- round(GetAssayData(full_analysis, layer = "counts")) 


meta_gene <- full_analysis@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = full_analysis, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = full_counts, id = full_analysis$record_id, pred = data_g_gene$pred, 
                          offset = full_analysis$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_fullanalysis_offset.txt', row.names=F,
            quote=F, sep='\t')




counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, offset = so_t1d_hc$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary


write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_MTAP_analysis_offset.txt', row.names=F,
            quote=F, sep='\t')


counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_nos3 <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_nos3, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, offset = so_t1d_hc$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_NOS3_analysis_offset.txt', row.names=F,
            quote=F, sep='\t')

counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_cgas <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_cgas, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, offset = so_t1d_hc$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_cGAS_analysis_offset.txt', row.names=F,
            quote=F, sep='\t')

rm(list=c(counts_t1d_hc_nos3, counts_t1d_hc_cgas, counts_t1d_hc_mtap))



##########With sex analysis 



full_analysis <- FindVariableFeatures(so_t1d_hc, selection.method = "vst", nfeatures = 2000)
hvgs <- VariableFeatures(full_analysis)
full_analysis <- subset(so_t1d_hc, features = hvgs)
full_counts <- round(GetAssayData(full_analysis, layer = "counts")) 


meta_gene <- full_analysis@meta.data
pred_gene <- model.matrix(~group*sex, data = meta_gene)
data_g_gene <- list(count = full_analysis, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = full_counts, id = full_analysis$record_id, pred = data_g_gene$pred, 
                          offset = full_analysis$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_fullanalysis_offset_withsex.txt', row.names=F,
            quote=F, sep='\t')




counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group*sex, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, offset = so_t1d_hc$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary


write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_MTAP_analysis_offset_withsex.txt', row.names=F,
            quote=F, sep='\t')


counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_nos3 <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group*sex, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_nos3, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, offset = so_t1d_hc$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_NOS3_analysis_offset_withsex.txt', row.names=F,
            quote=F, sep='\t')

counts_t1d_hc <- round(GetAssayData(so_t1d_hc, layer = "counts")) # load counts and round
counts_t1d_hc_cgas <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group*sex, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_cgas, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, offset = so_t1d_hc$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_cGAS_analysis_offset_withsex.txt', row.names=F,
            quote=F, sep='\t')

rm(list=c(counts_t1d_hc_nos3, counts_t1d_hc_cgas, counts_t1d_hc_mtap))


#Plotting top 2,000 genes expressions 

plotting_function <- function(sig_markers, file_name){
  sig_markers <- sig_markers %>% dplyr::select(gene = summary.gene,
                                        logfc_sex = summary.logFC_sexMale, 
                                        logfc_group = summary.logFC_groupType_1_Diabetes, 
                                        logfc_interaction = summary.logFC_groupType_1_Diabetes.sexMale, 
                                        p_sex = summary.p_sexMale,
                                        p_group = summary.p_groupType_1_Diabetes, 
                                        p_interaction = summary.p_groupType_1_Diabetes.sexMale)
  
  variable_names <- c('Sex Effects', 'Diagnosis Group Effects', 'Interaction Effects')
  
  
  for(i in c(1:3)){
    tmp_logfc <- i + 1
    tmp_p <- i + 4
    tmp_df <- sig_markers %>% dplyr::select(gene, tmp_logfc, tmp_p)
    tmp_label <- variable_names[i]
    names(tmp_df) <- c('Gene', 'LogFC', 'Pvalue')
    tmp_df <- tmp_df %>% filter(!is.na(Pvalue))
    tmp_df$diffexp <- 'No'
    tmp_df$diffexp[tmp_df$LogFC > 0.6 & tmp_df$Pvalue < 0.05] <- 'Up'
    tmp_df$diffexp[tmp_df$LogFC < -0.6 & tmp_df$Pvalue < 0.05] <- 'Down'
    tmp_df <- tmp_df %>% arrange(Pvalue)
    tmp_df$label <- NA
    tmp_df$label[1:10] <- tmp_df$Gene[1:10]

#Making graph
    if(length(unique(tmp_df$diffexp)) > 1){
tmp_graph <- ggplot(tmp_df, aes(x= LogFC, y=-log10(Pvalue), col = diffexp, label=label))+
  geom_point()+
  geom_text(size=2, vjust = 2, color='black')+
  scale_color_manual(values = c('orange', 'grey', 'purple'),
                     labels = c('Downregulated', 'Not significant', 'Upregulated'))+
  geom_hline(yintercept = -log10(0.05), col='blue', linetype='dashed')+
  geom_vline(xintercept = c(-0.6, 0.6), col='blue', linetype ='dashed')+
  theme_classic()+labs(x='LogFC', y='-log10 pvalue', col ='Differential Expression', 
                       title = paste0(tmp_label, ' in NEBULA Analysis'))
}else{
  tmp_graph <- ggplot(tmp_df, aes(x= LogFC, y=-log10(Pvalue), col = diffexp, label=label))+
  geom_point()+
  geom_text(size=2, vjust = 2, color='black')+
  scale_color_manual(values = c('grey'),
                     labels = c('Not significant'))+
  geom_hline(yintercept = -log10(0.05), col='blue', linetype='dashed')+
  geom_vline(xintercept = c(-0.6, 0.6), col='blue', linetype ='dashed')+
  theme_classic()+labs(x='LogFC', y='-log10 P-value', col ='Differential Expression', 
                       title = paste0(tmp_label, ' in NEBULA Analysis'))
  
}
    pdf(paste0('C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_VolcanoPlot_', file_name, '_', tmp_label, '.pdf'))
    print(tmp_graph)
    dev.off()
}
}


plotting_function(nebula_full, 'Top2000')










```

## EC
```{r echo = F}
so_t1d_hc$celltype_ec <- ifelse(so_t1d_hc$celltype_harmony %in% c("EC-PTC", "EC-GC", "EC-AEA", "EC-LYM"),
                                     "EC", as.character(so_t1d_hc$celltype_harmony))
so_t1d_hc_ec <- subset(so_t1d_hc, celltype_ec == "EC")

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_EC_MTAP.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc_ec, reduction = "umap.harmony", features = "MTAP", raster = F)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_EC_NOS3.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc_ec, reduction = 'umap.harmony', features = 'NOS3', raster = F)
dev.off()

jpeg("C:/Users/netio/Documents/UofW/cGAS_Analysis/UMAP_EC_cGWAS_STING.jpeg", width=900, 
     height=900)
FeaturePlot(so_t1d_hc_ec, reduction = 'umap.harmony', features = cgas$V1)
dev.off()







```
### Pseudoreplication (Treating each cell independently)

```{r echo = F}
FindMarkers(so_t1d_hc_ec, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = "MTAP")

FindMarkers(so_t1d_hc_ec, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = "NOS3")

FindMarkers(so_t1d_hc_ec, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = cgas$V1)

```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}

counts_layer <- round(GetAssayData(so_t1d_hc_ec, layer = 'counts'))
library_size <- Matrix::colSums(counts_layer)
so_t1d_hc$library_size <- library_size


#EC Top 2,000 Analysis


full_analysis <- FindVariableFeatures(so_t1d_hc_ec, selection.method = "vst", nfeatures = 2000)
hvgs <- VariableFeatures(full_analysis)
full_analysis <- subset(so_t1d_hc_ec, features = hvgs)
full_counts <- round(GetAssayData(full_analysis, layer = "counts")) 


meta_gene <- full_analysis@meta.data
pred_gene <- model.matrix(~group*sex, data = meta_gene)
data_g_gene <- list(count = full_analysis, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = full_counts, id = full_analysis$record_id, pred = data_g_gene$pred, 
                          offset = full_analysis$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_EC_fullanalysis_offset_withsex.txt', row.names=F,
            quote=F, sep='\t')





counts_t1d_hc <- round(GetAssayData(so_t1d_hc_ec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_ec, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_ec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_ec$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_ec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_ec, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_ec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_ec$pooled_offset,
                 reml = T, model = "NBLMM")

result_ec$summary

#NOS
counts_layer <- round(GetAssayData(so_t1d_hc_ec, layer = 'counts'))
library_size <- Matrix::colSums(counts_layer)
so_t1d_hc$library_size <- library_size

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_ec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc_ec, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_ec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_ec$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_ec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc_ec, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_ec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_ec$pooled_offset,
                 reml = T, model = "NBLMM")

result_ec$summary

#cGAS-STING

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_ec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc_ec, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_ec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_ec$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_ec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc_ec, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_ec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_ec$pooled_offset,
                 reml = T, model = "NBLMM")

result_ec$summary






##Baseline vs. Follow-up Analysis 






```

## POD
```{r echo = F}
so_t1d_hc$celltype_pod <- ifelse(so_t1d_hc$celltype_harmony %in% c("POD"),
                                     "pod", as.character(so_t1d_hc$celltype_harmony))
so_t1d_hc_pod <- subset(so_t1d_hc, celltype_pod == "pod")

full_analysis <- FindVariableFeatures(so_t1d_hc_pod, selection.method = "vst", nfeatures = 2000)
hvgs <- VariableFeatures(full_analysis)
full_analysis <- subset(so_t1d_hc_pod, features = hvgs)
full_counts <- round(GetAssayData(full_analysis, layer = "counts")) 


meta_gene <- full_analysis@meta.data
pred_gene <- model.matrix(~group*sex, data = meta_gene)
data_g_gene <- list(count = full_analysis, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = full_counts, id = full_analysis$record_id, pred = data_g_gene$pred, 
                          offset = full_analysis$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_pod_fullanalysis_offset_withsex.txt', row.names=F,
            quote=F, sep='\t')


plotting_function(result_allcells, 'POD_')


counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pod, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_pod, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pod$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pod, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_pod, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_pod$pooled_offset,
                 reml = T, model = "NBLMM")

result_pod$summary

#NOS
counts_layer <- round(GetAssayData(so_t1d_hc_pod, layer = 'counts'))
library_size <- Matrix::colSums(counts_layer)
so_t1d_hc$library_size <- library_size

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pod, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc_pod, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pod$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pod, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc_pod, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_pod$pooled_offset,
                 reml = T, model = "NBLMM")

result_pod$summary

#cGAS-STING

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pod, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc_pod, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pod$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pod, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc_pod, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_pod$pooled_offset,
                 reml = T, model = "NBLMM")

result_pod$summary

























```


#PEC

so_t1d_hc$celltype_pec <- ifelse(so_t1d_hc$celltype_harmony %in% c("PEC"),
                                     "pec", as.character(so_t1d_hc$celltype_harmony))
so_t1d_hc_pec <- subset(so_t1d_hc, celltype_pec == "pec")

full_analysis <- FindVariableFeatures(so_t1d_hc_pec, selection.method = "vst", nfeatures = 2000)
hvgs <- VariableFeatures(full_analysis)
full_analysis <- subset(so_t1d_hc_pec, features = hvgs)
full_counts <- round(GetAssayData(full_analysis, layer = "counts")) 


meta_gene <- full_analysis@meta.data
pred_gene <- model.matrix(~group*sex, data = meta_gene)
data_g_gene <- list(count = full_analysis, id = meta_gene$record_id, pred = pred_gene)
result_allcells <- nebula(count = full_counts, id = full_analysis$record_id, pred = data_g_gene$pred, 
                          offset = full_analysis$pooled_offset,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary

write.table(result_allcells, 'C:/Users/netio/Documents/UofW/cGAS_Analysis/NEBULA_pec_fullanalysis_offset_withsex.txt', row.names=F,
            quote=F, sep='\t')

plotting_function(as.data.frame(result_allcells), 'PEC')



counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_pec, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pec$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_pec, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_pec$pooled_offset,
                 reml = T, model = "NBLMM")

result_pec$summary

#NOS
counts_layer <- round(GetAssayData(so_t1d_hc_pec, layer = 'counts'))
library_size <- Matrix::colSums(counts_layer)
so_t1d_hc$library_size <- library_size

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc_pec, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pec$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["NOS3",]
meta_gene <- subset(so_t1d_hc_pec, features = "NOS3")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_pec$pooled_offset,
                 reml = T, model = "NBLMM")

result_pec$summary

#cGAS-STING

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc_pec, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pec$summary

counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc[cgas$V1,]
meta_gene <- subset(so_t1d_hc_pec, features = cgas$V1)@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T, offset = so_t1d_hc_pec$pooled_offset,
                 reml = T, model = "NBLMM")

result_pec$summary


















### Pseudoreplication (Treating each cell independently)

















```{r echo = F}
FindMarkers(so_t1d_hc_pod, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = "MTAP")
```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}
counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pod, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_pod, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pod$summary
```

## PEC
```{r echo = F}
so_t1d_hc$celltype_pec <- ifelse(so_t1d_hc$celltype_harmony %in% c("PEC"),
                                     "PEC", as.character(so_t1d_hc$celltype_harmony))
so_t1d_hc_pec <- subset(so_t1d_hc, celltype_pec == "PEC")

FeaturePlot(so_t1d_hc_pec, reduction = "umap.harmony", features = "MTAP", raster = F)
```
### Pseudoreplication (Treating each cell independently)

```{r echo = F}
FindMarkers(so_t1d_hc_pec, ident.1 = "Type_1_Diabetes", ident.2 = "Lean_Control", group.by = "group", features = "MTAP")
```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}
counts_t1d_hc <- round(GetAssayData(so_t1d_hc_pec, layer = "counts")) # load counts and round
counts_t1d_hc_mtap <- counts_t1d_hc["MTAP",]
meta_gene <- subset(so_t1d_hc_pec, features = "MTAP")@meta.data
pred_gene <- model.matrix(~group, data = meta_gene)
data_g_gene <- list(count = counts_t1d_hc_mtap, id = meta_gene$record_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pec$summary
```

## PT
```{r echo = F}
so_t1d_hc$celltype_pt <- ifelse(grepl("PT-", so_t1d_hc$celltype_harmony),
                                 "PT", as.character(so_t1d_hc$celltype_harmony))
so_t1d_hc_pt <- subset(so_t1d_hc, celltype_pt == "PT" & celltype_harmony != "PT_lowQuality")
FeaturePlot(so_t1d_hc_pt, reduction = "umap.harmony", features = "MTAP", raster = F)
```

# T1D (-) vs. T1D (+) in ATTEMPT

```{r echo = F}
temp_file <- tempfile(fileext = ".RDS")
s3$download_file("scrna", "Kidney transcriptomics/Single cell RNA seq/PB_attempt_harmony_rpca_RM_kpmpV1labelled_Sept2024.RDS", temp_file)
attempt_so <- readRDS(temp_file)

DimPlot(attempt_so, reduction = "umap.harmony", group.by = "KPMP_celltype", raster = F, alpha = 0.2, label = T)
DimPlot(attempt_so, reduction = "umap.harmony", group.by = "celltype", raster = F, alpha = 0.2, label = T)

temp_file <- tempfile(fileext = ".RDS")
s3$download_file("harmonized.dataset", "attempt_clinical_data.RDS", temp_file)
attempt_dat <- readRDS(temp_file)

attempt_so_meta <- attempt_so@meta.data %>%
  mutate(subject_id = Subject.ID,
         visit = case_when(Visit == "BL" ~ "PRE", 
                           Visit == "4M" ~ "POST"))
attempt_so_meta <- left_join(attempt_so_meta, attempt_dat, by = c("subject_id", "visit"))
rownames(attempt_so_meta) <- attempt_so_meta$barcode
attempt_so <- AddMetaData(attempt_so, attempt_so_meta)

attempt_so$visit <- factor(attempt_so$visit, levels = c("PRE", "POST"))
attempt_so$treatment <- factor(attempt_so$treatment, levels = c("Placebo", "Dapagliflozin 5mg"),
                                  labels = c("Placebo", "Dapagliflozin"))

attempt_so_bl <- subset(attempt_so, visit == "PRE")
```

## All cell types combined

```{r echo = F}
FeaturePlot(attempt_so_bl, reduction = "umap.harmony", features = "MTAP", raster = F)
```
### Pseudoreplication (Treating each cell independently)

```{r echo = F}
FindMarkers(attempt_so_bl, ident.1 = "Dapagliflozin", ident.2 = "Placebo", group.by = "treatment", 
            features = "MTAP", logfc.threshold = 0)
```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}
counts_attempt_bl <- round(GetAssayData(attempt_so_bl, layer = "counts")) # load counts and round
counts_attempt_bl_mtap <- counts_attempt_bl["MTAP",]
meta_gene <- subset(attempt_so_bl, features = "MTAP")@meta.data
pred_gene <- model.matrix(~treatment, data = meta_gene)
data_g_gene <- list(count = counts_attempt_bl_mtap, id = meta_gene$subject_id, pred = pred_gene)
result_allcells <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_allcells$summary
```

## EC
```{r echo = F}
attempt_so_bl$celltype_ec <- ifelse(attempt_so_bl$celltype %in% c("EC-PTC", "EC-GC", "EC-AEA", "EC-LYM"),
                                     "EC", as.character(attempt_so_bl$celltype))
attempt_so_bl_hc_ec <- subset(attempt_so_bl, celltype_ec == "EC")

FeaturePlot(attempt_so_bl_hc_ec, reduction = "umap.harmony", features = "MTAP", raster = F)
```
### Pseudoreplication (Treating each cell independently)

```{r echo = F}
FindMarkers(attempt_so_bl_hc_ec, ident.1 = "Dapagliflozin", ident.2 = "Placebo", group.by = "treatment", features = "MTAP", logfc.threshold = 0)
```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}
counts_attempt_bl <- round(GetAssayData(attempt_so_bl_hc_ec, layer = "counts")) # load counts and round
counts_attempt_bl_mtap <- counts_attempt_bl["MTAP", , drop = F]
meta_gene <- subset(attempt_so_bl_hc_ec, features = "MTAP")@meta.data
pred_gene <- model.matrix(~treatment, data = meta_gene)
data_g_gene <- list(count = counts_attempt_bl_mtap, id = meta_gene$subject, pred = pred_gene)
result_ec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")
result_ec$summary
```

## POD
```{r echo = F}
attempt_so_bl$celltype_pod <- ifelse(attempt_so_bl$celltype %in% c("POD"),
                                     "POD", as.character(attempt_so_bl$celltype))
attempt_so_bl_hc_pod <- subset(attempt_so_bl, celltype_pod == "POD")

FeaturePlot(attempt_so_bl_hc_pod, reduction = "umap.harmony", features = "MTAP", raster = F)
```
### Pseudoreplication (Treating each cell independently)

```{r echo = F}
FindMarkers(attempt_so_bl_hc_pod, ident.1 = "Dapagliflozin", ident.2 = "Placebo", group.by = "treatment", features = "MTAP", logfc.threshold = 0)
```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}
counts_attempt_bl <- round(GetAssayData(attempt_so_bl_hc_pod, layer = "counts")) # load counts and round
counts_attempt_bl_mtap <- counts_attempt_bl["MTAP",]
meta_gene <- subset(attempt_so_bl_hc_pod, features = "MTAP")@meta.data
pred_gene <- model.matrix(~treatment, data = meta_gene)
data_g_gene <- list(count = counts_attempt_bl_mtap, id = meta_gene$subject_id, pred = pred_gene)
result_pod <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pod$summary
```

## PEC
```{r echo = F}
attempt_so_bl$celltype_pec <- ifelse(attempt_so_bl$celltype %in% c("PEC"),
                                     "PEC", as.character(attempt_so_bl$celltype))
attempt_so_bl_hc_pec <- subset(attempt_so_bl, celltype_pec == "PEC")

FeaturePlot(attempt_so_bl_hc_pec, reduction = "umap.harmony", features = "MTAP", raster = F)
```
### Pseudoreplication (Treating each cell independently)

```{r echo = F}
FindMarkers(attempt_so_bl_hc_pec, ident.1 = "Dapagliflozin", ident.2 = "Placebo", group.by = "treatment", features = "MTAP", logfc.threshold = 0)
```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}
counts_attempt_bl <- round(GetAssayData(attempt_so_bl_hc_pec, layer = "counts")) # load counts and round
counts_attempt_bl_mtap <- counts_attempt_bl["MTAP",]
meta_gene <- subset(attempt_so_bl_hc_pec, features = "MTAP")@meta.data
pred_gene <- model.matrix(~treatment, data = meta_gene)
data_g_gene <- list(count = counts_attempt_bl_mtap, id = meta_gene$subject_id, pred = pred_gene)
result_pec <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pec$summary
```

## PT
```{r echo = F}
attempt_so_bl$celltype_pt <- ifelse(grepl("PT-", attempt_so_bl$celltype),
                                 "PT", as.character(attempt_so_bl$celltype))
attempt_so_bl_hc_pt <- subset(attempt_so_bl, celltype_pt == "PT" & celltype != "PT_lowQuality")
FeaturePlot(attempt_so_bl_hc_pt, reduction = "umap.harmony", features = "MTAP", raster = F)
```
### Pseudoreplication (Treating each cell independently)

```{r echo = F}
FindMarkers(attempt_so_bl_hc_pt, ident.1 = "Dapagliflozin", ident.2 = "Placebo", group.by = "treatment", features = "MTAP", logfc.threshold = 0)
```
### Mixed model using NEBULA (Treating each subject independently)

```{r echo = F}
counts_attempt_bl <- round(GetAssayData(attempt_so_bl_hc_pt, layer = "counts")) # load counts and round
counts_attempt_bl_mtap <- counts_attempt_bl["MTAP",]
meta_gene <- subset(attempt_so_bl_hc_pt, features = "MTAP")@meta.data
pred_gene <- model.matrix(~treatment, data = meta_gene)
data_g_gene <- list(count = counts_attempt_bl_mtap, id = meta_gene$subject_id, pred = pred_gene)
result_pt <- nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred, 
                 ncore = 1, output_re = T, covariance = T,
                 reml = T, model = "NBLMM")

result_pt$summary
```