---
title: "SLC6A19 — scRNA transcript expression results (PT focus)"
author: "Ye Ji Choi"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(tidyr)
library(aws.s3)
library(jsonlite)
library(knitr)
library(kableExtra)

# ---- per-report parameters -------------------------------------------------
gene           <- "SLC6A19"
bucket         <- "scrna"
results_prefix <- "Projects/transcript search/Results/"
figures_prefix <- "Projects/transcript search/Results/Figures/"

# PT and PT-subtype cell types (sanitized filename tokens) + pooled
focus_ct <- c("PT_S1_S2", "PT_S3", "aPT", "PT", "all")
```

```{r aws-env}
# Standalone AWS/Kopah setup (mirrors the analysis doc) so this renders on its own
user <- Sys.info()[["user"]]
if (user == "yejichoi") {
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
  Sys.setenv(
    "AWS_ACCESS_KEY_ID"     = keys$MY_ACCESS_KEY,
    "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
    "AWS_DEFAULT_REGION"    = "",
    "AWS_REGION"            = "",
    "AWS_S3_ENDPOINT"       = "s3.kopah.uw.edu"
  )
}
```

```{r helpers}
s3read_using_region <- function(FUN, ..., object, bucket, region = NULL) {
  tmp <- tempfile(fileext = paste0(".", tools::file_ext(object)))
  on.exit(unlink(tmp))
  args <- list(bucket = bucket, object = object, file = tmp)
  if (!is.null(region)) args$region <- region
  do.call(aws.s3::save_object, args)
  FUN(tmp, ...)
}

# readable labels for sanitized cell-type tokens
ct_label <- c(EC_LYM = "EC-LYM", MC = "MC", aTAL = "aTAL", ATL = "ATL",
              PT_S1_S2 = "PT-S1/S2", PT_S3 = "PT-S3", aPT = "aPT",
              PT = "PT (general)", all = "all (pooled)")

# recognized cell-type tokens in filenames, longest first (avoids PT matching PT_S3)
ct_tokens <- c("PT_S1_S2", "EC_LYM", "PT_S3", "aTAL", "aPT", "ATL", "MC", "PT")

parse_neb_name <- function(key) {
  base    <- sub("_neb\\.csv$", "", basename(key))
  dataset <- dplyr::case_when(startsWith(base, "pb90_")    ~ "pb90",
                              startsWith(base, "attempt_") ~ "attempt",
                              TRUE ~ NA_character_)
  mid <- sub("^(pb90|attempt)_", "", base)
  ct  <- "all"; analysis <- mid
  for (tok in ct_tokens) {
    if (grepl(paste0("_", tok, "$"), mid)) {
      ct <- tok; analysis <- sub(paste0("_", tok, "$"), "", mid); break
    }
  }
  data.frame(dataset = dataset, analysis = analysis,
             cell_type = ct, key = key, stringsAsFactors = FALSE)
}

# read one NEBULA csv, keep this gene, reshape estimates to one row per term
read_one <- function(key, gene) {
  df <- tryCatch(
    s3read_using_region(FUN = function(f) read.csv(f, check.names = FALSE),
                        bucket = bucket, object = key, region = ""),
    error = function(e) NULL)
  if (is.null(df)) return(NULL)
  gcol <- intersect(c("gene", "gene_id", "Gene"), names(df))[1]
  if (is.na(gcol)) return(NULL)
  df <- df[df[[gcol]] == gene, , drop = FALSE]
  if (!nrow(df)) return(NULL)

  est <- df %>%
    dplyr::select(dplyr::starts_with("logFC_"),
                  dplyr::starts_with("se_"),
                  dplyr::starts_with("p_")) %>%
    dplyr::slice(1) %>%
    tidyr::pivot_longer(dplyr::everything(),
                        names_to = c(".value", "term"),
                        names_pattern = "^(logFC|se|p)_(.*)$") %>%
    dplyr::filter(term != "(Intercept)")

  est$n_cells  <- if ("n_cells"  %in% names(df)) df$n_cells[1]  else NA
  est$n_people <- if ("n_people" %in% names(df)) df$n_people[1] else NA
  est$key      <- key
  est
}

# download S3 PNGs to temp, return local paths (skip any that fail)
embed_s3_pngs <- function(keys) {
  if (!length(keys)) return(character(0))
  paths <- vapply(keys, function(k) {
    tmp <- tempfile(fileext = ".png")
    tryCatch({ aws.s3::save_object(object = k, bucket = bucket,
                                   file = tmp, region = ""); tmp },
             error = function(e) NA_character_)
  }, character(1))
  paths[!is.na(paths)]
}
```

```{r compile}
neb_keys <- aws.s3::get_bucket_df(bucket = bucket, prefix = results_prefix,
                                  region = "", max = Inf)$Key
neb_keys <- neb_keys[grepl("_neb\\.csv$", neb_keys)]

meta_tbl <- if (length(neb_keys)) do.call(rbind, lapply(neb_keys, parse_neb_name)) else NULL

est_list <- lapply(neb_keys, read_one, gene = gene)
est_tbl  <- if (any(lengths(est_list))) {
  do.call(rbind, est_list) %>% dplyr::left_join(meta_tbl, by = "key")
} else NULL

disp <- if (!is.null(est_tbl)) {
  est_tbl %>%
    dplyr::filter(cell_type %in% focus_ct) %>%
    dplyr::transmute(
      analysis,
      `cell type` = dplyr::coalesce(ct_label[cell_type], cell_type),
      term,
      logFC   = round(logFC, 3),
      se      = round(se, 3),
      p       = signif(p, 3),
      n_cells, n_people,
      dataset
    ) %>%
    dplyr::arrange(dataset, analysis, `cell type`, term)
} else NULL

show_tbl <- function(d) {
  sub <- dplyr::filter(disp, dataset == d) %>% dplyr::select(-dataset)
  if (!nrow(sub)) return(cat("_No results found for", d, "._\n"))
  sub %>% kbl(caption = paste0(d, " — ", gene, " NEBULA results")) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
    print()
}
```

# Overview

SLC6A19 (B⁰AT1) is a neutral amino-acid transporter and a canonical **proximal tubule**
identity marker, strongest in the early PT segments. This report compiles the NEBULA
results for **`r gene`** across both datasets, focused on PT and its subtypes
(PT-S1/S2, PT-S3, aPT, and the general PT grouping), alongside the pooled runs.

> **Multiplicity caution.** These tables aggregate many models (multiple contrasts × cell
> types × datasets). The `p` values are raw per-model NEBULA p-values; interpret them with
> the full size of the sweep in mind and apply a deliberate multiple-testing correction
> before drawing firm conclusions. The `n_cells` / `n_people` columns flag runs thin enough
> that estimates may be unstable.

# Compiled NEBULA results

## pb90

```{r results-pb90, results = "asis"}
show_tbl("pb90")
```

## ATTEMPT

For the DiD model (`treatment*visit`), the term to read is the **treatment × visit
interaction**; the main-effect rows are shown for completeness.

```{r results-attempt, results = "asis"}
show_tbl("attempt")
```

# Figures

```{r fig-keys}
fig_keys <- aws.s3::get_bucket_df(bucket = bucket, prefix = figures_prefix,
                                  region = "", max = Inf)$Key
fig_keys <- fig_keys[grepl(gene, fig_keys) & grepl("\\.png$", fig_keys)]

umap_keys    <- fig_keys[grepl("UMAP",  fig_keys)]
pb90_keys    <- fig_keys[grepl("pb90",  fig_keys)]
attempt_keys <- fig_keys[grepl("attempt", fig_keys)]

img_tag <- function(key, width = "70%") {
  tmp <- tempfile(fileext = ".png")
  ok  <- tryCatch({ aws.s3::save_object(object = key, bucket = bucket,
                                        file = tmp, region = ""); TRUE },
                  error = function(e) FALSE)
  if (!ok) return(invisible(NULL))
  cat(sprintf('<p><em>%s</em><br><img src="%s" style="width:%s"></p>\n\n',
              basename(key), knitr::image_uri(tmp), width))
}
```

## Feature UMAPs

```{r fig-umap}
#| label: fig-umap
#| results: asis
for (k in umap_keys) img_tag(k)
```

## pb90 — expression by group

```{r fig-pb90}
#| label: fig-pb90
#| results: asis
for (k in pb90_keys) img_tag(k)
```

## ATTEMPT — dapa vs. placebo, PRE vs. POST

```{r fig-attempt}
#| label: fig-attempt
#| results: asis
for (k in attempt_keys) img_tag(k)
```
