---
title: "scRNA transcript search (config-driven)"
author: "Ye Ji Choi"
format: html
---

<!--
================================================================================
HOW TO USE THIS SCRIPT
================================================================================
This file is split into two parts:

  1. CONFIG  — the only block you normally edit. Add transcripts, cell-type
               views, covariates/stratifications, cohorts, and palettes here.
  2. ENGINE  — stable helper functions + drivers. You should not need to touch
               this to add a new transcript or covariate.

To add a NEW TRANSCRIPT:
  - Add the gene symbol to `genes`.
  - That's it for NEBULA: every covariate model below is fit on ALL genes at
    once, so new genes automatically appear in every *_neb.csv.
  - For figures, cell types are auto-detected (top expressers per gene per
    dataset). To force specific cell types for a gene, add an entry to
    `gene_views_manual`.

To add a NEW COVARIATE / STRATIFICATION:
  - Add an entry to `pb90_covariates` and/or `attempt_covariates`.
  - type = "categorical" -> boxplot + dotplot (pb90) / handled per dataset
    type = "continuous"  -> scatter of per-subject mean expression vs covariate
    type = "did"         -> ATTEMPT difference-in-differences (treatment*visit):
                            spaghetti + delta boxplot + dotplot
  - Each covariate drives BOTH the NEBULA model (CSV) and its figure.

Outputs:
  - CSVs   -> s3://scrna/<base_dir>/<dataset>_<covariate>[_<celltype>]_neb.csv
  - Figures-> s3://scrna/<fig_dir>/...   and UMAP feature plots under <umap_dir>
================================================================================
-->

```{r echo = F}
library(dplyr)
library(Seurat)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
library(nebula)
library(openxlsx)
library(table1)
library(lme4)
library(lmerTest)
library(purrr)
library(tibble)
library(tidyr)
library(stringr)
library(Hmisc)
library(ggtext)
library(corrplot)
library(gtools)
library(foreach)
library(emmeans)
library(knitr)
library(kableExtra)
library(future)
library(BiocParallel)
library(SingleCellExperiment)
library(scran)
library(fgsea)
library(msigdbr)
library(togolab)
library(togolab.scrna)
```

```{r echo = F}
togo_paths()
```

# ============================================================================
# CONFIG  — edit this block
# ============================================================================

```{r}
## --- Transcripts -----------------------------------------------------------
## NEBULA fits ALL of these at once, so adding a gene here automatically adds
## it to every *_neb.csv. Only genes present in a dataset are used (missing
## ones are silently dropped per dataset).
genes <- c("SLC6A19", 
           "NNMT", 
           "SHISA5", 
           "STRA6", 
           "LTBP3", 
           "NPLOC4", 
           "NRP1", 
           "ARL4C")

## --- Output locations (S3) -------------------------------------------------
base_dir <- "P047_scRNA_transcript_search_yc/results/"
fig_dir  <- "P047_scRNA_transcript_search_yc/results/temp_figures/Box Plots/"
umap_dir <- "P047_scRNA_transcript_search_yc/results/temp_figures/UMAP/"

## --- Auto cell-type detection ----------------------------------------------
## For figures, cell types are chosen automatically per gene per dataset:
## any cell type expressing the gene in > pct_threshold of cells, capped at
## max_auto_celltypes (ranked by % expressed). Override per gene below.
pct_threshold     <- 10    # % of cells expressing to be "of interest"
max_auto_celltypes <- 6    # cap so facet plots stay readable

## --- Manual cell-type overrides (optional) ---------------------------------
## Map gene -> list of views. Each view: celltypes, ct_col, tag.
## Views whose cell types are absent from a dataset are skipped automatically.
## Anything NOT listed here is auto-detected.
gene_views_manual <- list(
  SLC6A19 = list(
    list(celltypes = c("PT-S1/S2", "PT-S3", "aPT"), ct_col = "KPMP_celltype",         tag = "PT_sub"),
    list(celltypes = "PT",                          ct_col = "KPMP_celltype_general", tag = "PT")
  )
  # NNMT and the 5 new genes are auto-detected (top expressers per dataset).
)

## --- Palettes --------------------------------------------------------------
group_pal  <- c("Type 2 Diabetes" = "#e07a5f", "Type 1 Diabetes" = "#f2cc8f",
                "Lean Control"    = "#81b29a", "Obese Control"   = "#3d405b")
dkd_pal    <- c("DKD" = "#f1856a", "non_DKD" = "#a3b18a")
sglt2i_pal <- c("Yes" = "#4f6d7a", "No" = "#c0d6df", "1" = "#4f6d7a", "0" = "#c0d6df")
arm_cols   <- c("Dapagliflozin 5mg" = "#a7b298", "Placebo" = "#f8ae9d")

## ===========================================================================
## PB90 covariates / stratifications
##   var        : NEBULA predictor & plot grouping column (the modelled column)
##   type        : "categorical" | "continuous"
##   cohort      : `group` levels to keep (NULL = all cohorts)
##   transform   : NULL, or a function applied to `source_var` to create `var`
##   source_var  : raw column the transform reads (continuous-with-transform)
##   palette     : named fill colors (categorical)
##   label       : x-axis label (continuous)
##   nebula/plot : TRUE/FALSE toggles (default TRUE)
## ===========================================================================
pb90_covariates <- list(
  group_all = list(var = "group",          type = "categorical",
                   cohort = NULL,           palette = group_pal),

  sglt2i_ever_t2d = list(var = "sglt2i_ever",  type = "categorical",
                   cohort = "Type 2 Diabetes", palette = sglt2i_pal),

  sglt2i_timepoint_t2d = list(var = "sglt2i_timepoint", type = "categorical",
                   cohort = "Type 2 Diabetes", palette = sglt2i_pal),

  dkd30_t2d = list(var = "dkd_group_30",    type = "categorical",
                   cohort = c("Type 2 Diabetes", "Obese Control"), palette = dkd_pal),

  dkd100_t2d = list(var = "dkd_group_100",  type = "categorical",
                   cohort = c("Type 2 Diabetes", "Obese Control"), palette = dkd_pal),

  egfr = list(var = "eGFR_CKD_epi",         type = "continuous",
                   cohort = NULL,           label = "eGFR (CKD-EPI)"),

  log_uacr = list(var = "log_acr_u",        type = "continuous",
                   cohort = NULL,           source_var = "acr_u", transform = log,
                   label = "log(uACR)")
)

## ===========================================================================
## ATTEMPT covariates (T1D only; pre/post; dapagliflozin vs placebo)
##   type "did"        : difference-in-differences, var = "treatment*visit"
##   type "continuous" : baseline-only association (visit == PRE)
## ===========================================================================
attempt_covariates <- list(
  treatment_did = list(var = "treatment*visit", type = "did", palette = arm_cols),

  egfr_bl = list(var = "eGFR_CKD_epi",      type = "continuous",
                 baseline_only = TRUE,       label = "eGFR (CKD-EPI)"),

  log_uacr_bl = list(var = "log_uacr",      type = "continuous",
                 baseline_only = TRUE,       source_var = "emu_urine_acr_mean",
                 transform = log,            label = "log(uACR)"),
  
  dkd30_bl = list(var = "dkd_group_30",    type = "categorical",
                      baseline_only = TRUE, palette = dkd_pal),
  
  dkd100_bl = list(var = "dkd_group_100",  type = "categorical",
                       baseline_only = TRUE, palette = dkd_pal)
)
```

# ============================================================================
# ENGINE  — stable helpers & drivers (no edits needed to add transcripts)
# ============================================================================

```{r}
`%||%` <- function(a, b) if (is.null(a)) b else a

theme_transparent <- theme(
  plot.background   = element_rect(fill = "transparent", color = NA),
  panel.background  = element_rect(fill = "transparent", color = NA),
  legend.background = element_rect(fill = "transparent", color = NA)
)

# genes actually present in an object (avoids subset errors for absent symbols)
genes_in <- function(obj) intersect(genes, rownames(obj))

# ---------------------------------------------------------------------------
# Cell-type view resolution: manual override else auto-detect top expressers
# ---------------------------------------------------------------------------
auto_gene_views <- function(obj, gene, ct_col = "KPMP_celltype") {
  if (!gene %in% rownames(obj)) return(list())
  expr <- Seurat::FetchData(obj, vars = gene)[[1]]
  ct   <- obj@meta.data[[ct_col]]
  pct  <- tapply(expr > 0, ct, function(x) mean(x, na.rm = TRUE) * 100)
  pct  <- pct[!is.na(pct) & pct > pct_threshold]
  if (length(pct) == 0) return(list())
  hits <- names(sort(pct, decreasing = TRUE))
  hits <- head(hits, max_auto_celltypes)
  list(list(celltypes = hits, ct_col = ct_col, tag = "auto"))
}

build_gene_views <- function(obj, gene) {
  if (!is.null(gene_views_manual[[gene]])) {
    vs <- gene_views_manual[[gene]]
    vs <- Filter(function(v) any(v$celltypes %in% obj@meta.data[[v$ct_col]]), vs)
    if (length(vs) > 0) return(vs)
  }
  auto_gene_views(obj, gene)
}

# all views for all genes on a given object
build_all_views <- function(obj) setNames(lapply(genes_in(obj), function(g) build_gene_views(obj, g)),
                                          genes_in(obj))

# flatten gene views -> named vector ct -> ct_col (for per-cell-type NEBULA)
flatten_ct_col <- function(all_views) {
  m <- list()
  for (g in names(all_views)) for (v in all_views[[g]]) for (ct in v$celltypes) m[[ct]] <- v$ct_col
  unlist(m)
}

# ---------------------------------------------------------------------------
# NEBULA core (unchanged logic)
# ---------------------------------------------------------------------------
run_nebula_var <- function(var, seurat_obj, ncore = 4, record_id = "record_id") {
  meta <- seurat_obj@meta.data
  form <- if (inherits(var, "formula")) var else as.formula(paste("~", var))
  vars_used <- all.vars(form)
  keep <- stats::complete.cases(meta[, vars_used, drop = FALSE])

  if (sum(keep) < 2 || dplyr::n_distinct(meta[[record_id]][keep]) < 2) return(NULL)

  meta_v   <- droplevels(meta[keep, , drop = FALSE])
  counts_v <- round(GetAssayData(seurat_obj, layer = "counts"))[, keep, drop = FALSE]
  offset_v <- seurat_obj$pooled_offset[keep]
  pred     <- model.matrix(form, data = meta_v)

  data_g <- group_cell(count = counts_v, id = meta_v[[record_id]], pred = pred, offset = offset_v)
  if (is.null(data_g)) data_g <- list(count = counts_v, id = meta_v[[record_id]], pred = pred, offset = offset_v)

  nebula(count = data_g$count, id = data_g$id, pred = data_g$pred,
         offset = data_g$offset, ncore = ncore, reml = 1, model = "NBLMM")
}

neb_run_and_save <- function(seurat_obj, var, out_object,
                             record_id = NULL, person_col, visit_col = NULL, ncore = 1) {
  md      <- seurat_obj@meta.data
  n_cells <- ncol(seurat_obj)
  people  <- md[[person_col]]

  args <- list(var = var, seurat_obj = seurat_obj, ncore = ncore)
  if (!is.null(record_id)) args$record_id <- record_id
  res <- do.call(run_nebula_var, args)
  if (is.null(res)) stop("NEBULA returned NULL (too few cells/subjects with complete data)")

  if (!is.null(visit_col) && visit_col %in% names(md)) {
    per_person <- as.integer(table(interaction(people, md[[visit_col]], drop = TRUE)))
  } else {
    per_person <- as.integer(table(people))
  }

  stats <- data.frame(
    n_cells = n_cells,
    n_people = length(unique(people)),
    median_cells_per_person = median(per_person)
  )

  terms      <- trimws(strsplit(var, "[*:+]")[[1]])
  terms      <- terms[terms %in% names(md)]
  group_vars <- terms[!vapply(terms, function(t) is.numeric(md[[t]]), logical(1))]
  if (length(group_vars) > 0) {
    grp       <- do.call(interaction, c(md[group_vars], list(drop = TRUE, sep = "_")))
    cells_by  <- table(grp)
    people_by <- tapply(people, grp, function(x) length(unique(x)))
    for (g in names(cells_by)) {
      stats[[paste0("cells_",  g)]] <- as.integer(cells_by[[g]])
      stats[[paste0("people_", g)]] <- as.integer(people_by[[g]])
    }
  }

  out_df <- cbind(res$summary, stats[rep(1, nrow(res$summary)), , drop = FALSE], row.names = NULL)
  s3write_using_region(out_df, FUN = write.csv, bucket = "togo.projects",
                       object = out_object, row.names = FALSE, region = "")
  message("saved: ", out_object, " | cells: ", n_cells, " | people: ", stats$n_people,
          " | converged: ", sum(res$convergence == 1), "/", length(res$convergence))
  invisible(list(summary = out_df, convergence = res$convergence, stats = stats))
}

# ---------------------------------------------------------------------------
# Prep builders: turn a covariate spec into a subsetting function
# ---------------------------------------------------------------------------
build_prep_pb90 <- function(spec) {
  function(obj) {
    keep <- if (is.null(spec$cohort)) rep(TRUE, ncol(obj)) else obj$group %in% spec$cohort
    o    <- subset(obj, cells = colnames(obj)[keep], features = genes_in(obj))
    if (!is.null(spec$transform)) {
      o[[spec$var]] <- spec$transform(o@meta.data[[spec$source_var]])
      o <- subset(o, cells = colnames(o)[is.finite(o@meta.data[[spec$var]])])
    }
    o
  }
}

build_prep_attempt <- function(spec) {
  function(obj) {
    o <- subset(obj, features = genes_in(obj))
    if (isTRUE(spec$baseline_only)) o <- subset(o, subset = visit == "PRE")
    if (!is.null(spec$transform)) {
      o[[spec$var]] <- spec$transform(o@meta.data[[spec$source_var]])
      o <- subset(o, cells = colnames(o)[is.finite(o@meta.data[[spec$var]])])
    }
    o
  }
}

# wrap a base spec so its prep additionally subsets to one cell type
add_celltype <- function(spec, ct, ct_col) {
  force(ct); force(ct_col)
  ct_safe   <- gsub("[^A-Za-z0-9]+", "_", ct)
  orig_prep <- spec$prep
  list(
    var       = spec$var,
    out       = sub("_neb\\.csv$", paste0("_", ct_safe, "_neb.csv"), spec$out),
    record_id = spec$record_id,
    prep      = function(obj) {
      o <- orig_prep(obj)
      keep <- o@meta.data[[ct_col]] %in% ct
      subset(o, cells = colnames(o)[keep])
    }
  )
}

# Build {overall + per-cell-type} specs for every NEBULA-enabled covariate,
# then run them all and save CSVs.
run_nebula_suite <- function(obj, covariates, dataset, prep_builder,
                             person_col, record_id = NULL, visit_col = NULL) {
  all_views <- build_all_views(obj)
  ct_to_col <- flatten_ct_col(all_views)

  base_specs <- list()
  for (nm in names(covariates)) {
    cv <- covariates[[nm]]
    if (isFALSE(cv$nebula)) next
    base_specs[[nm]] <- list(
      var       = cv$var,
      out       = paste0(base_dir, dataset, "_", nm, "_neb.csv"),
      record_id = record_id,
      prep      = prep_builder(cv)
    )
  }

  ct_specs <- list()
  for (ct in names(ct_to_col)) {
    ct_col <- ct_to_col[[ct]]
    for (nm in names(base_specs)) {
      key <- paste0(nm, "_", gsub("[^A-Za-z0-9]+", "_", ct))
      ct_specs[[key]] <- add_celltype(base_specs[[nm]], ct, ct_col)
    }
  }

  specs   <- c(base_specs, ct_specs)
  results <- list()
  for (nm in names(specs)) {
    message("\n=== NEBULA [", dataset, "]: ", nm, " ===")
    spec <- specs[[nm]]
    obj_s <- tryCatch(spec$prep(obj), error = function(e) NULL)
    if (is.null(obj_s) || ncol(obj_s) == 0) { message("  skipped: no cells"); next }
    results[[nm]] <- tryCatch(
      neb_run_and_save(obj_s, var = spec$var, out_object = spec$out,
                       record_id = spec$record_id, person_col = person_col, visit_col = visit_col),
      error = function(e) { message("  skipped: ", conditionMessage(e)); NULL }
    )
  }
  invisible(results)
}
```

```{r}
# ---------------------------------------------------------------------------
# UMAP feature plots (loops over all present genes)
# ---------------------------------------------------------------------------
prepare_umap_metadata <- function(seurat_obj, genes,
                                  celltype_col = "KPMP_celltype",
                                  umap_reduction = "umap.harmony",
                                  color_palette = "Set3", custom_colors = NULL) {
  ct_col_vals <- seurat_obj@meta.data[[celltype_col]]
  celltypes   <- if (is.factor(ct_col_vals)) levels(ct_col_vals) else sort(unique(as.character(ct_col_vals)))
  celltypes   <- celltypes[!is.na(celltypes)]
  n_celltypes <- length(celltypes)
  if (!is.null(custom_colors)) {
    celltype_colors <- custom_colors
  } else {
    umap_colors <- colorRampPalette(brewer.pal(min(brewer.pal.info[color_palette, "maxcolors"], n_celltypes), color_palette))(n_celltypes)
    celltype_colors <- setNames(umap_colors, celltypes)
  }
  expr_df <- Seurat::FetchData(seurat_obj, vars = genes) %>% as.data.frame()
  colnames(expr_df) <- genes
  umap_embed_cols <- colnames(seurat_obj@reductions[[umap_reduction]]@cell.embeddings)
  metadata <- seurat_obj@meta.data %>%
    cbind(seurat_obj@reductions[[umap_reduction]]@cell.embeddings) %>%
    cbind(expr_df)
  colnames(metadata)[colnames(metadata) %in% umap_embed_cols] <- c("umapharmony_1", "umapharmony_2")
  centers <- metadata %>% dplyr::group_by(.data[[celltype_col]]) %>%
    summarise(x = median(umapharmony_1), y = median(umapharmony_2), .groups = "drop")
  list(metadata = metadata, centers = centers, celltype_colors = celltype_colors)
}

plot_feature_umap <- function(genes, metadata, centers, celltype_colors,
                              pct_threshold = 10, save_to_s3 = FALSE,
                              s3_prefix = umap_dir, s3_suffix = "", s3_bucket = "togo.projects",
                              plot_height = 10, plot_width = 10) {
  plots <- list()
  for (gene in genes) {
    expr_stats <- metadata %>% dplyr::group_by(KPMP_celltype) %>%
      dplyr::summarise(pct_expressed = round(mean(.data[[gene]] > 0) * 100, 1),
                       n_total = n(), .groups = "drop") %>%
      filter(pct_expressed > pct_threshold)
    centers_filtered <- centers %>%
      mutate(pct_expressed_5 = case_when(KPMP_celltype %in% expr_stats$KPMP_celltype ~ "Y", TRUE ~ "N"))
    caption_text <- expr_stats %>% arrange(desc(pct_expressed)) %>%
      mutate(color = celltype_colors[KPMP_celltype],
             label = paste0("<span style='color:", color, "'>&#9679; </span>**", KPMP_celltype, "**: ", pct_expressed, "%"),
             group = ceiling(row_number() / 5)) %>%
      dplyr::group_by(group) %>% summarise(line = paste(label, collapse = " | "), .groups = "drop") %>%
      rbind(data.frame(group = 0, line = paste0("**% Expressed (>", pct_threshold, "%):**"))) %>%
      arrange(group) %>% pull(line) %>% paste(collapse = "<br><br>")
    feature_p <- metadata %>%
      dplyr::mutate(feature_color_logic = case_when(.data[[gene]] > 0 ~ KPMP_celltype, TRUE ~ "No")) %>%
      ggplot(aes(x = umapharmony_1, y = umapharmony_2, color = feature_color_logic)) +
      geom_point(alpha = 0.4) +
      geom_text_repel(data = centers_filtered, aes(x = x, y = y, label = KPMP_celltype),
                      inherit.aes = FALSE, size = 4,
                      fontface = ifelse(centers_filtered$pct_expressed_5 == "Y", "bold", "plain"),
                      color = ifelse(centers_filtered$pct_expressed_5 == "Y", "black", "#495057"),
                      max.overlaps = Inf) +
      scale_color_manual(values = c(celltype_colors, "No" = "#edede9")) +
      theme_minimal() +
      theme(panel.grid = element_blank(), text = element_text(size = 15), legend.position = "none",
            plot.title = element_text(hjust = 0.5, face = "bold"),
            plot.caption = element_markdown(size = 12, hjust = 0.5),
            axis.title.x = element_text(hjust = 0), axis.title.y = element_text(hjust = 0)) +
      theme_transparent +
      labs(x = "UMAP1", y = "UMAP2", title = gene, caption = caption_text)
    print(feature_p)
    plots[[gene]] <- feature_p
    if (save_to_s3) {
      s3write_using_region(feature_p, FUN = ggsave,
        object = paste0(s3_prefix, gene, "_FeaturePlot", s3_suffix, ".png"),
        bucket = s3_bucket, region = "", height = plot_height, width = plot_width)
    }
  }
  invisible(plots)
}
```

```{r}
# ---------------------------------------------------------------------------
# PB90 figure functions
# ---------------------------------------------------------------------------
# categorical: per-person boxplot + dotplot
plot_pb90_box <- function(obj, gene, celltypes, ct_col, tag, out_dir,
                          group_col = "group", person_col = "record_id",
                          cohort = NULL, cohort_col = "group", fill_values = NULL) {
  df <- FetchData(obj, vars = unique(c(gene, person_col, group_col, ct_col)))
  names(df)[names(df) == gene]       <- "expr"
  names(df)[names(df) == person_col] <- "person"
  names(df)[names(df) == group_col]  <- "strat"
  names(df)[names(df) == ct_col]     <- "celltype"
  df$cohort <- obj@meta.data[rownames(df), cohort_col]
  if (!is.null(cohort)) df <- df %>% filter(.data$cohort %in% .env$cohort)
  df <- df %>% filter(celltype %in% celltypes, !is.na(strat))

  pb <- df %>% group_by(strat, celltype, person) %>%
    summarise(mean_expr = mean(expr), .groups = "drop") %>% mutate(strat = factor(strat))

  p1 <- ggplot(pb, aes(strat, mean_expr, fill = strat)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.6) +
    geom_jitter(width = 0.15, alpha = 0.6) +
    facet_wrap(~ celltype, scales = "free_y") +
    labs(y = paste("Mean", gene, "(log-norm)"), x = NULL) +
    theme_minimal() +
    theme(text = element_text(size = 15), legend.position = "none",
          axis.text.x = element_text(angle = 60, hjust = 1), panel.grid = element_blank()) +
    theme_transparent
  if (!is.null(fill_values)) p1 <- p1 + scale_fill_manual(values = fill_values)
  print(p1)
  s3write_using_region(FUN = ggsave, bucket = "togo.projects", region = "",
    object = paste0(out_dir, "pb90_", gene, "_", tag, "_box.png"),
    width = 7, height = 5, dpi = 300, bg = "transparent")

  o <- obj
  if (!is.null(cohort)) o <- subset(o, cells = colnames(o)[o@meta.data[[cohort_col]] %in% cohort])
  o$grp <- paste(o@meta.data[[ct_col]], o@meta.data[[group_col]], sep = "__")
  p2 <- DotPlot(o, features = gene, group.by = "grp")@data %>%
    separate_wider_delim(id, delim = "__", names = c("celltype", "strat")) %>%
    filter(celltype %in% celltypes, strat != "NA") %>%
    ggplot(aes(celltype, strat, size = pct.exp, color = avg.exp.scaled)) +
    geom_point() +
    scale_color_gradient2(low = "#fff0f3", mid = "#fff0f3", high = "#ee7674") +
    theme_minimal() + theme(panel.grid = element_blank(), text = element_text(size = 15)) +
    theme_transparent + labs(y = NULL, x = NULL, size = "% expr", color = "Avg expr")
  print(p2)
  s3write_using_region(FUN = ggsave, bucket = "togo.projects", region = "",
    object = paste0(out_dir, "pb90_", gene, "_", tag, "_dotplot.png"),
    width = 7, height = 5, dpi = 300, bg = "transparent")
}

# continuous: per-person mean expression vs covariate, with lm trend
plot_pb90_scatter <- function(obj, gene, celltypes, ct_col, tag, out_dir,
                              covar_var, transform = NULL, label = NULL,
                              person_col = "record_id", cohort = NULL, cohort_col = "group") {
  df <- FetchData(obj, vars = unique(c(gene, person_col, covar_var, ct_col)))
  names(df)[names(df) == gene]       <- "expr"
  names(df)[names(df) == person_col] <- "person"
  names(df)[names(df) == covar_var]  <- "covar"
  names(df)[names(df) == ct_col]     <- "celltype"
  df$cohort <- obj@meta.data[rownames(df), cohort_col]
  if (!is.null(cohort)) df <- df %>% filter(.data$cohort %in% .env$cohort)
  if (!is.null(transform)) df$covar <- transform(df$covar)
  df <- df %>% filter(celltype %in% celltypes, is.finite(covar))
  if (nrow(df) == 0) { message("  no data for scatter: ", gene, " / ", tag); return(invisible()) }

  pb <- df %>% group_by(celltype, person) %>%
    dplyr::summarise(mean_expr = mean(expr), covar = dplyr::first(covar), .groups = "drop")

  p <- ggplot(pb, aes(covar, mean_expr)) +
    geom_point(alpha = 0.6, color = "#3d405b") +
    geom_smooth(method = "lm", se = TRUE, color = "#e07a5f", fill = "#f2cc8f") +
    facet_wrap(~ celltype, scales = "free") +
    labs(x = label %||% covar_var, y = paste("Mean", gene, "(log-norm)")) +
    theme_minimal() + theme(text = element_text(size = 15), panel.grid = element_blank()) +
    theme_transparent
  print(p)
  s3write_using_region(FUN = ggsave, bucket = "togo.projects", region = "",
    object = paste0(out_dir, "pb90_", gene, "_", tag, "_scatter.png"),
    width = 7, height = 5, dpi = 300, bg = "transparent")
}

# ---------------------------------------------------------------------------
# ATTEMPT figure functions
# ---------------------------------------------------------------------------
# DiD: spaghetti (PRE->POST by arm) + delta boxplot + dotplot
plot_attempt_did <- function(obj, gene, celltypes, ct_col, tag, out_dir, fill_values = arm_cols) {
  df <- FetchData(obj, vars = c(gene, "subject_id", "visit", "treatment", ct_col))
  names(df)[names(df) == gene]   <- "expr"
  names(df)[names(df) == ct_col] <- "celltype"
  df <- df %>% mutate(visit = factor(visit, levels = c("PRE", "POST"))) %>%
    filter(celltype %in% celltypes)
  pb <- df %>% group_by(treatment, celltype, subject_id, visit) %>%
    summarise(mean_expr = mean(expr), .groups = "drop")

  base_thm <- theme_minimal() +
    theme(text = element_text(size = 15), legend.position = "none",
          axis.text.x = element_text(angle = 60, hjust = 1), panel.grid = element_blank()) +
    theme_transparent

  p1 <- ggplot(pb, aes(visit, mean_expr, color = treatment)) +
    geom_line(aes(group = subject_id), alpha = 0.25) +
    stat_summary(aes(group = treatment), fun = mean, geom = "line", linewidth = 1.3) +
    stat_summary(aes(group = treatment), fun = mean, geom = "point", size = 2.5) +
    facet_wrap(~ celltype, scales = "free_y") +
    labs(y = paste("Mean", gene, "(log-norm)"), x = NULL) +
    base_thm + scale_color_manual(values = fill_values)
  print(p1)
  s3write_using_region(FUN = ggsave, bucket = "togo.projects", region = "",
    object = paste0(out_dir, "attempt_", gene, "_", tag, "_spaghetti.png"),
    width = 7, height = 5, dpi = 300, bg = "transparent")

  delta <- pb %>% pivot_wider(names_from = visit, values_from = mean_expr) %>% mutate(delta = POST - PRE)
  p2 <- ggplot(delta, aes(treatment, delta, fill = treatment)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.6) +
    geom_jitter(width = 0.15, alpha = 0.6) +
    geom_hline(yintercept = 0, linetype = 2) +
    facet_wrap(~ celltype, scales = "free_y") +
    labs(y = paste0("\U0394 ", gene, " (POST - PRE)"), x = NULL) +
    base_thm + scale_fill_manual(values = fill_values)
  print(p2)
  s3write_using_region(FUN = ggsave, bucket = "togo.projects", region = "",
    object = paste0(out_dir, "attempt_", gene, "_", tag, ".png"),
    width = 7, height = 5, dpi = 300, bg = "transparent")

  obj$grp <- paste(obj@meta.data[[ct_col]], obj$treatment, obj$visit, sep = "_")
  p3 <- DotPlot(obj, features = gene, group.by = "grp")@data %>%
    separate_wider_delim(id, delim = "_", names = c("celltype", "treatment", "visit")) %>%
    filter(celltype %in% celltypes) %>%
    ggplot(aes(celltype, paste0(treatment, "-", visit), size = pct.exp, color = avg.exp.scaled)) +
    geom_point() +
    scale_color_gradient2(low = "#fff0f3", mid = "#fff0f3", high = "#ee7674") +
    theme_minimal() + theme(panel.grid = element_blank(), text = element_text(size = 15)) +
    labs(y = NULL, x = NULL, size = "% expr", color = "Avg expr") + theme_transparent
  print(p3)
  s3write_using_region(FUN = ggsave, bucket = "togo.projects", region = "",
    object = paste0(out_dir, "attempt_", gene, "_", tag, "_dotplot.png"),
    width = 7, height = 5, dpi = 300, bg = "transparent")
}

# continuous baseline association
plot_attempt_scatter <- function(obj, gene, celltypes, ct_col, tag, out_dir,
                                 covar_var, transform = NULL, label = NULL) {
  o  <- subset(obj, subset = visit == "PRE")
  df <- FetchData(o, vars = unique(c(gene, "subject_id", covar_var, ct_col)))
  names(df)[names(df) == gene]   <- "expr"
  names(df)[names(df) == covar_var] <- "covar"
  names(df)[names(df) == ct_col] <- "celltype"
  if (!is.null(transform)) df$covar <- transform(df$covar)
  df <- df %>% filter(celltype %in% celltypes, is.finite(covar))
  if (nrow(df) == 0) { message("  no data for scatter: ", gene, " / ", tag); return(invisible()) }
  pb <- df %>% group_by(celltype, subject_id) %>%
    dplyr::summarise(mean_expr = mean(expr), covar = dplyr::first(covar), .groups = "drop")
  p <- ggplot(pb, aes(covar, mean_expr)) +
    geom_point(alpha = 0.6, color = "#3d405b") +
    geom_smooth(method = "lm", se = TRUE, color = "#a7b298", fill = "#f8ae9d") +
    facet_wrap(~ celltype, scales = "free") +
    labs(x = label %||% covar_var, y = paste("Mean", gene, "(log-norm)")) +
    theme_minimal() + theme(text = element_text(size = 15), panel.grid = element_blank()) +
    theme_transparent
  print(p)
  s3write_using_region(FUN = ggsave, bucket = "togo.projects", region = "",
    object = paste0(out_dir, "attempt_", gene, "_", tag, "_scatter.png"),
    width = 7, height = 5, dpi = 300, bg = "transparent")
}

# ---------------------------------------------------------------------------
# Figure drivers: loop genes x views x covariates
# ---------------------------------------------------------------------------
run_figs_pb90 <- function(obj, covariates, out_dir = fig_dir) {
  all_views <- build_all_views(obj)
  for (gene in names(all_views)) {
    for (v in all_views[[gene]]) {
      for (nm in names(covariates)) {
        cv <- covariates[[nm]]
        if (isFALSE(cv$plot)) next
        tag <- paste0(v$tag, "_", nm)
        if (identical(cv$type, "categorical")) {
          plot_pb90_box(obj, gene, v$celltypes, v$ct_col, tag, out_dir,
                        group_col = cv$var, cohort = cv$cohort,
                        fill_values = cv$palette)
        } else if (identical(cv$type, "continuous")) {
          plot_pb90_scatter(obj, gene, v$celltypes, v$ct_col, tag, out_dir,
                            covar_var = cv$source_var %||% cv$var,
                            transform = cv$transform, label = cv$label,
                            cohort = cv$cohort)
        }
      }
    }
  }
}

run_figs_attempt <- function(obj, covariates, out_dir = fig_dir) {
  all_views <- build_all_views(obj)
  for (gene in names(all_views)) {
    for (v in all_views[[gene]]) {
      for (nm in names(covariates)) {
        cv <- covariates[[nm]]
        if (isFALSE(cv$plot)) next
        tag <- paste0(v$tag, "_", nm)
        if (identical(cv$type, "did")) {
          plot_attempt_did(obj, gene, v$celltypes, v$ct_col, tag, out_dir,
                           fill_values = cv$palette %||% arm_cols)
        } else if (identical(cv$type, "continuous")) {
          plot_attempt_scatter(obj, gene, v$celltypes, v$ct_col, tag, out_dir,
                              covar_var = cv$source_var %||% cv$var,
                              transform = cv$transform, label = cv$label)
        }
      }
    }
  }
}
```

# ============================================================================
# PB90
# ============================================================================

```{r echo = F}
pb90 <- s3readRDS(bucket = "core.data", object = 'scRNA/Seurat/pb90_processed_full.rds', region = "")

# derived covariate columns used by config (DKD groups + log uACR)
pb90@meta.data <- pb90@meta.data %>%
  mutate(
    dkd_group_30 = case_when(
      eGFR_CKD_epi < 90 | acr_u >= 30 ~ "DKD",
      eGFR_CKD_epi >= 90 & acr_u < 30 ~ "non_DKD",
      TRUE ~ NA_character_),
    dkd_group_30 = factor(dkd_group_30, levels = c("non_DKD", "DKD")),
    dkd_group_100 = case_when(
      eGFR_CKD_epi < 90 | acr_u >= 100 ~ "DKD",
      eGFR_CKD_epi >= 90 & acr_u < 100 ~ "non_DKD",
      TRUE ~ NA_character_),
    dkd_group_100 = factor(dkd_group_100, levels = c("non_DKD", "DKD"))
  )
```

## UMAP feature plots (T2D)

```{r}
pb90_umap_meta <- prepare_umap_metadata(
  seurat_obj = subset(pb90, group == "Type 2 Diabetes"),
  genes = genes_in(pb90))

plot_feature_umap(
  genes = genes_in(pb90),
  metadata = pb90_umap_meta$metadata,
  centers = pb90_umap_meta$centers,
  celltype_colors = pb90_umap_meta$celltype_colors,
  pct_threshold = pct_threshold,
  save_to_s3 = TRUE, s3_prefix = umap_dir, s3_suffix = "_t2d")
```

## NEBULA models (CSV per covariate, overall + per cell type)

```{r}
pb90_neb_results <- run_nebula_suite(
  obj          = pb90,
  covariates   = pb90_covariates,
  dataset      = "pb90",
  prep_builder = build_prep_pb90,
  person_col   = "record_id",
  record_id    = NULL)
```

## Figures (boxplots for categorical, scatter for continuous)

```{r}
run_figs_pb90(pb90, pb90_covariates)
```

# ============================================================================
# ATTEMPT
# ============================================================================

```{r}
# ---- one-time cleaning (then cached to S3; skip to reload chunk below) ------
attempt_so <- s3readRDS(bucket = "core.data", 
                            object = "scRNA/Seurat/attempt_processed_full", region = "")

attempt_so_meta <- attempt_so@meta.data %>%
  mutate(dkd_group_30 = case_when(
           eGFR_CKD_epi < 90 | acr_u >= 30 ~ "DKD",
           eGFR_CKD_epi >= 90 & acr_u < 30 ~ "non_DKD",
           TRUE ~ NA_character_),
         dkd_group_30 = factor(dkd_group_30, levels = c("non_DKD", "DKD")),
         dkd_group_100 = case_when(
           eGFR_CKD_epi < 90 | acr_u >= 100 ~ "DKD",
           eGFR_CKD_epi >= 90 & acr_u < 100 ~ "non_DKD",
           TRUE ~ NA_character_),
         dkd_group_100 = factor(dkd_group_100, levels = c("non_DKD", "DKD"))
  )

assignInNamespace("FilterObjects", SeuratObject:::.FilterObjects, ns = "SeuratObject")
attempt_so <- AddMetaData(attempt_so, attempt_so_meta)
```

## UMAP feature plots (POST visit)

```{r}
attempt_umap_meta <- prepare_umap_metadata(
  seurat_obj = subset(attempt_so, visit == "POST"),
  genes = genes_in(attempt_so))

plot_feature_umap(
  genes = genes_in(attempt_so),
  metadata = attempt_umap_meta$metadata,
  centers = attempt_umap_meta$centers,
  celltype_colors = attempt_umap_meta$celltype_colors,
  pct_threshold = pct_threshold,
  save_to_s3 = TRUE, s3_prefix = umap_dir, s3_suffix = "_attempt_post")
```

## NEBULA models (DiD + baseline associations, overall + per cell type)

```{r}
attempt_neb_results <- run_nebula_suite(
  obj          = attempt_so,
  covariates   = attempt_covariates,
  dataset      = "attempt",
  prep_builder = build_prep_attempt,
  person_col   = "subject_id",
  record_id    = "subject_id",
  visit_col    = "visit")
```

## Figures (DiD spaghetti/delta + baseline scatter)

```{r}
run_figs_attempt(attempt_so, attempt_covariates)
```
