---
title: "scRNA transcript expression results"
author: "Ye Ji Choi"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(tidyr)
library(aws.s3)
library(jsonlite)
library(knitr)
library(kableExtra)

# ---- report parameters -----------------------------------------------------
bucket         <- "scrna"
results_prefix <- "Projects/transcript search/Results/"
figures_prefix <- "Projects/transcript search/Results/Figures/"

gene_notes <- list(
  SLC6A19 = "SLC6A19 (B⁰AT1): neutral amino-acid transporter and canonical proximal-tubule identity marker, strongest in early PT segments.",
  NNMT    = "NNMT (nicotinamide N-methyltransferase): metabolic enzyme tied to NAD⁺/methylation balance and, in kidney, to metabolic stress and fibrosis."
)
```

```{r aws-env}
# Standalone AWS/Kopah setup so this renders on its own
user <- Sys.info()[["user"]]
if (user == "yejichoi") {
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
}

keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
Sys.setenv(
  "AWS_ACCESS_KEY_ID"     = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION"    = "",
  "AWS_REGION"            = "",
  "AWS_S3_ENDPOINT"       = "s3.kopah.uw.edu"
)
```

```{r helpers}
s3read_using_region <- function(FUN, ..., object, bucket, region = NULL) {
  tmp <- tempfile(fileext = paste0(".", tools::file_ext(object)))
  on.exit(unlink(tmp))
  args <- list(bucket = bucket, object = object, file = tmp, url_style = "path")
  if (!is.null(region)) args$region <- region
  do.call(aws.s3::save_object, args)
  FUN(tmp, ...)
}

# ---- cell-type vocabulary (KPMP V1 fine-grained + general collapses) -------
# Used to strip the cell-type suffix from NEBULA filenames. Robust to new
# covariates because we match the cell type at the END of the name.
sanitize <- function(x) gsub("[^A-Za-z0-9]+", "_", x)
ct_vocab <- c(
  "PT-S1/S2", "PT-S3", "aPT",
  "C-TAL-1", "C-TAL-2", "aTAL", "dTAL", "ATL",
  "CCD-PC", "CNT-PC", "dCCD-PC", "M-PC", "tPC-IC",
  "IC-A", "IC-B", "aIC",
  "EC-AVR", "EC-GC", "EC-PTC", "EC-AEA", "EC-LYM", "EC/VSMC",
  "MAC", "MON", "cDC", "pDC",
  "CD4+ T", "CD8+ T", "B", "NK",
  "VSMC/P", "FIB", "POD", "MC",
  "PT", "TAL", "PC", "IC", "EC"          # general (KPMP_celltype_general)
)
ct_tokens <- sanitize(ct_vocab)
ct_label  <- setNames(ct_vocab, ct_tokens)
ct_label["all"] <- "all (pooled)"
# longest token first so e.g. PT_S1_S2 wins over PT, CNT_PC over PC
ct_tokens <- ct_tokens[order(nchar(ct_tokens), decreasing = TRUE)]

parse_neb_name <- function(key) {
  base    <- sub("_neb\\.csv$", "", basename(key))
  dataset <- dplyr::case_when(startsWith(base, "pb90_")    ~ "pb90",
                              startsWith(base, "attempt_") ~ "attempt",
                              TRUE ~ NA_character_)
  mid <- sub("^(pb90|attempt)_", "", base)
  ct  <- "all"; analysis <- mid
  for (tok in ct_tokens) {
    if (grepl(paste0("_", tok, "$"), mid)) {
      ct <- tok; analysis <- sub(paste0("_", tok, "$"), "", mid); break
    }
  }
  data.frame(dataset = dataset, analysis = analysis,
             cell_type = ct, key = key, stringsAsFactors = FALSE)
}

# read one NEBULA csv, melt ALL genes -> one row per gene x term
read_all_genes <- function(key) {
  df <- tryCatch(
    s3read_using_region(FUN = function(f) read.csv(f, check.names = FALSE),
                        bucket = bucket, object = key, region = ""),
    error = function(e) NULL)
  if (is.null(df) || !nrow(df)) return(NULL)
  gcol <- intersect(c("gene", "gene_id", "Gene"), names(df))[1]
  if (is.na(gcol)) return(NULL)
  est_cols  <- grep("^(logFC|se|p)_", names(df), value = TRUE)
  if (!length(est_cols)) return(NULL)
  meta_cols <- intersect(c("n_cells", "n_people"), names(df))

  df %>%
    dplyr::select(gene = dplyr::all_of(gcol),
                  dplyr::all_of(est_cols), dplyr::all_of(meta_cols)) %>%
    tidyr::pivot_longer(cols = dplyr::all_of(est_cols),
                        names_to = c(".value", "term"),
                        names_pattern = "^(logFC|se|p)_(.*)$") %>%
    dplyr::filter(term != "(Intercept)") %>%
    dplyr::mutate(key = key)
}

# embed an S3 PNG inline (skip silently on failure)
img_tag <- function(key, width = "70%") {
  tmp <- tempfile(fileext = ".png")
  ok  <- tryCatch({ aws.s3::save_object(object = key, bucket = bucket,
                                        file = tmp, region = "", url_style = "path"); TRUE },
                  error = function(e) FALSE)
  if (!ok) return(invisible(NULL))
  cat(sprintf('<p><em>%s</em><br><img src="%s" style="width:%s"></p>\n\n',
              basename(key), knitr::image_uri(tmp), width))
}

# match a gene token in a figure filename (delimited, avoids partial hits)
gene_match <- function(keys, g) grepl(paste0("(^|/|_)", g, "(_|\\.)"), keys)
```

```{r compile}
neb_keys <- aws.s3::get_bucket_df(bucket = bucket, prefix = results_prefix,
                                  region = "", url_style = "path", max = Inf)$Key
neb_keys <- neb_keys[grepl("_neb\\.csv$", neb_keys)]

meta_tbl <- if (length(neb_keys)) do.call(rbind, lapply(neb_keys, parse_neb_name)) else NULL

est_list <- lapply(neb_keys, read_all_genes)
est_tbl  <- if (any(lengths(est_list))) {
  do.call(rbind, est_list) %>% dplyr::left_join(meta_tbl, by = "key")
} else NULL

disp <- if (!is.null(est_tbl)) {
  est_tbl %>%
    dplyr::transmute(
      gene,
      analysis,
      `cell type` = dplyr::coalesce(ct_label[cell_type], cell_type),
      term,
      logFC   = round(logFC, 3),
      se      = round(se, 3),
      p       = signif(p, 3),
      n_cells, n_people,
      dataset
    ) %>%
    dplyr::arrange(gene, dataset, analysis, `cell type`, term)
} else NULL

# discovered transcripts (known ones first, then any others alphabetically)
genes <- if (!is.null(disp)) {
  g_all   <- sort(unique(disp$gene))
  g_known <- intersect(names(gene_notes), g_all)
  c(g_known, setdiff(g_all, g_known))
} else character(0)

# figure keys (once)
fig_keys <- aws.s3::get_bucket_df(bucket = bucket, prefix = figures_prefix,
                                  region = "", url_style = "path", max = Inf)$Key
fig_keys <- fig_keys[grepl("\\.png$", fig_keys)]

# ---- per-gene renderers ----------------------------------------------------
show_gene_tbl <- function(g, d) {
  sub <- dplyr::filter(disp, gene == g, dataset == d) %>%
    dplyr::select(-gene, -dataset)
  if (!nrow(sub)) { cat("_No results found for", d, "._\n\n"); return(invisible()) }
  sub %>%
    kbl(caption = paste0(d, " — ", g, " NEBULA results")) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
    print()
  cat("\n\n")
}

render_gene <- function(g) {
  cat("\n\n## ", g, "\n\n", sep = "")
  if (!is.null(gene_notes[[g]])) cat(gene_notes[[g]], "\n\n", sep = "")

  cat("\n\n### NEBULA results — pb90\n\n")
  show_gene_tbl(g, "pb90")
  cat("\n\n### NEBULA results — ATTEMPT\n\n")
  cat("_For the DiD model (`treatment*visit`), read the **treatment × visit interaction**; main-effect rows shown for completeness._\n\n")
  show_gene_tbl(g, "attempt")

  gk <- fig_keys[gene_match(fig_keys, g)]
  if (length(gk)) {
    cat("\n\n### Figures\n\n")
    uk <- gk[grepl("UMAP", gk)]
    pk <- gk[grepl("pb90", gk)]
    ak <- gk[grepl("attempt", gk)]
    if (length(uk)) { cat("\n\n**Feature UMAPs**\n\n");                    for (k in uk) img_tag(k) }
    if (length(pk)) { cat("\n\n**pb90 — expression by group / covariate**\n\n"); for (k in pk) img_tag(k) }
    if (length(ak)) { cat("\n\n**ATTEMPT — dapa vs. placebo / baseline**\n\n");   for (k in ak) img_tag(k) }
  }
}
```

# Overview

This report compiles **all transcripts** found in the NEBULA result files under
`r results_prefix`, across both datasets (pb90 and ATTEMPT). For each gene it shows
the per-model NEBULA estimates (every covariate × cell type, plus pooled runs) and
the associated figures.

```{r summary-tbl, results = "asis"}
if (is.null(disp)) {
  cat("_No NEBULA result files found._\n\n")
} else {
  disp %>%
    dplyr::distinct(gene, dataset, analysis, `cell type`) %>%
    dplyr::count(gene, dataset, name = "models") %>%
    tidyr::pivot_wider(names_from = dataset, values_from = models, values_fill = 0) %>%
    kbl(caption = "Models per transcript and dataset") %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
    print()
}
```

> **Multiplicity caution.** These tables aggregate many models (multiple covariates ×
> cell types × datasets × transcripts). The `p` values are raw per-model NEBULA
> p-values; interpret them with the full size of the sweep in mind and apply a
> deliberate multiple-testing correction before drawing firm conclusions. The
> `n_cells` / `n_people` columns flag runs thin enough that estimates may be unstable.

# Results by transcript

```{r per-gene, results = "asis"}
if (!length(genes)) {
  cat("_Nothing to display._\n\n")
} else {
  for (g in genes) render_gene(g)
}
```
