---
title: "CALICO Aim 1 Analyses"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(arsenal)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Linux" = "/home/tim/OneDrive/Vigers/BDC/Janet Snell-Bergeon/CALICO"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
#| include: false
# Load data
load("./Data_Clean/analysis_data.RData")
# Create additional categorical variables per Grayson and Melanie
df$cv_ft_above_uln <- df$cv_ft_perc > 100
label(df$cv_ft_above_uln) <- "Free Testosterone > 100% ULN"
df$cv_tt_above_uln <- df$cv_tt_perc > 100
label(df$cv_tt_above_uln) <- "Total Testosterone > 100% ULN"
df$cv_ft_perc_over_tt_perc_under <- df$cv_tt_perc <= 100 & df$cv_ft_perc > 100
label(df$cv_ft_perc_over_tt_perc_under) <- "TT <= 100% ULN & FT > 100% ULN"
df$cv_dheas_above_uln <- df$cv_dheas_perc > 100
label(df$cv_dheas_above_uln) <- "DHEAS > 100% ULN"
df$cv_androstendione_above_uln <- df$cv_androstendione_perc > 100
label(df$cv_androstendione_above_uln) <- "Androstendione > 100% ULN"
df$cv_amh_above_uln <- df$cv_amh_perc > 100
label(df$cv_amh_above_uln) <- "AMH > 100% ULN"
df$lh_fsh_ratio <- df$cv_lh / df$cv_fsh
label(df$lh_fsh_ratio) <- "LH:FSH ratio"
df$lh_fsh_ratio_above_2 <- df$lh_fsh_ratio > 2
label(df$lh_fsh_ratio_above_2) <- "LH:FSH ratio > 2"
df$lh_fsh_ratio_under_1 <- df$lh_fsh_ratio < 1
label(df$lh_fsh_ratio_under_1) <- "LH:FSH ratio < 1"
df$cv_shbg_under_20 <- df$cv_shbg < 20
label(df$cv_shbg_under_20) <- "SHBG < 20 "
df$cv_sbp_cat <- cut(df$cv_sbp,
  breaks = c(-Inf, 120, 130, 140, Inf),
  labels = c("<120", "120-129", "130-139", "140+"), right = F
)
label(df$cv_sbp_cat) <- "SBP category"
df$cv_a1c_cat <- cut(df$cv_a1c,
  breaks = c(-Inf, 5.6, 6.4, Inf),
  labels = c("<=5.6", "5.7-6.4", ">=6.5")
)
label(df$cv_a1c_cat) <- "HbA1c category"
df$cv_fbg_cat <- cut(df$cv_fbg,
  breaks = c(-Inf, 100, 126, Inf),
  labels = c("<100", "100-125", "126+"), right = F
)
label(df$cv_fbg_cat) <- "Fasting glucose category"
df$cv_2hrglucoseogtt_cat <- cut(df$cv_2hrglucoseogtt,
  breaks = c(-Inf, 140, 200, Inf), labels = c("<140", "140-199", "200+"),
  right = F
)
label(df$cv_2hrglucoseogtt_cat) <- "2-hr glucose category"
df$cv_tg_cat <- cut(df$cv_tg,
  breaks = c(-Inf, 90, 150, Inf),
  labels = c("<90", "90-149", "150+"),
  right = F
)
label(df$cv_tg_cat) <- "Triglycerides category"
df$cv_hdl_under_45 <- df$cv_hdl < 45
label(df$cv_hdl_under_45) <- "HDL < 45"
df$cv_alt_above_35 <- df$cv_alt > 35
label(df$cv_alt_above_35) <- "ALT > 35"
# Variable lists
demo_vars <- c(
  "site", "pcosdx_age", "combined_race", "ethnicity", "insur_type"
)
aim1_vars <- c(
  "cv_bmi", "cv_bmi_percentile", "cv_bmi_z", "cv_hirsutism_num",
  "cv_hirsutism_cat", "cv_acneface", "cv_acneother___1", "cv_acneother___2",
  "cv_acneother___0", "cv_acneother___unk", "cv_ft", "cv_ft_perc",
  "cv_ft_above_uln", "cv_tt", "cv_tt_perc", "cv_tt_above_uln",
  "cv_ft_perc_over_tt_perc_under", "cv_dheas", "cv_dheas_perc",
  "cv_dheas_above_uln", "cv_androstendione", "cv_androstendione_perc",
  "cv_androstendione_above_uln", "cv_lh", "cv_fsh", "lh_fsh_ratio",
  "lh_fsh_ratio_above_2", "lh_fsh_ratio_under_1", "cv_amh", "cv_amh_above_uln",
  "cv_a1c", "cv_a1c_cat", "cv_tg", "cv_tg_cat", "cv_hdl", "cv_hdl_under_45",
  "cv_shbg", "cv_shbg_under_20", "cv_alt", "cv_alt_above_35", "cv_fbg",
  "cv_fbg_cat", "cv_fastinsulin", "cv_2hrglucoseogtt", "cv_2hrglucoseogtt_cat",
  "cv_acanthosisneck", "cv_waist", "cv_osa_sx", "cv_medications___1",
  "cv_medications___2", "cv_medications___3", "cv_medications___4",
  "cv_medications___5", "cv_medications___6", "cv_medications___7",
  "cv_medications___8", "cv_medications___9", "cv_medications___10",
  "cv_medications___11", "cv_medications___12", "cv_medications___13",
  "cv_medications___14", "cv_medications___15", "cv_medications___16",
  "cv_medications___17", "cv_medications___18", "cv_medications___19",
  "cv_medications___20", "cv_medications___21", "cv_medications___22",
  "cv_medications___23", "cv_medications___32", "cv_medications___24",
  "cv_medications___25", "cv_medications___26", "cv_medications___27",
  "cv_medications___28", "cv_medications___29", "cv_medications___30",
  "cv_medications___31", "cv_medications___60", "cv_medications___0",
  "cv_medications___unk"
)
# Basic formula for tables
f <- as.formula(paste("~", paste(aim1_vars, collapse = "+")))
f <- update(f, . ~ . + cv_sbp + cv_sbp_cat + cv_dbp + pcosdx_age +
  pcosdx_menarche + pcosdx_famhx_parent___7 + pcosdx_famhx___7 +
  pcosdx_any_famhx___7 + pcosdx_famhx_parent___27 + pcosdx_famhx___27 +
  pcosdx_any_famhx___27 +
  combined_race + ethnicity)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct")
  )
# Diagnosis visit, remove those on hormone meds (likely data entry issue)
cv1 <- df %>%
  group_by(record_number) %>%
  filter(
    cv_monthssincepcosdx == 0, cv_medications___5 == "Unchecked",
    cv_medications___6 == "Unchecked", cv_medications___7 == "Unchecked",
    cv_medications___9 == "Unchecked", cv_medications___10 == "Unchecked",
    cv_medications___11 == "Unchecked", cv_medications___12 == "Unchecked"
  ) %>%
  ungroup()
```

# Data cleaning

Data were limited to diagnosis visit, and those without a diagnostic visit were excluded. Some participants had appeared to have incorrectly entered hormone-containing medications, so diagnostic visits were excluded if any of the following variables were checked (cyclic progesterone was not excluded):

- `cv_medications___5` (Estrogen-containing pill)
- `cv_medications___6` (Estrogen-containing patch)
- `cv_medications___7` (Estrogen-containing ring)
- `cv_medications___9` (Daily progesterone)
- `cv_medications___10` (Progesterone implant)
- `cv_medications___11` (Progesterone IUD)
- `cv_medications___12` (Progesterone injection)

# Demographics

```{r results='asis'}
#| label: tbl-table-1
#| tbl-cap: Participant demographics
f1 <-
  as.formula(paste0(
    "~",
    paste0(c(demo_vars, "combined_race"), collapse = "+")
  ))
t1 <- tableby(f1, data = cv1, control = mycontrols)
summary(t1, labelTranslations = labels(df), pfootnote = T)
```

# Aim 1 comparisons

## By age

```{r results='asis'}
#| label: tbl-table-1-age
#| tbl-cap: Stratified by age group at diagnosis
t1_age <- tableby(update(f, age_group ~ .), data = cv1, control = mycontrols)
summary(t1_age, labelTranslations = labels(df), pfootnote = T)
```

## By weight category

### Based on percentile and raw value

```{r results='asis'}
#| label: tbl-table-1-weight-perc
#| tbl-cap: Stratified by weight category (percentile and raw value)
t1_bmi <- tableby(update(f, weight_perc ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

For those with a BMI percentile included in the dataset, participants with a BMI percentile $\geq$ 85 were considered have obesity. For those with a missing value for BMI percentile, those with a raw BMI value $\geq$ 25 were considered to have obesity. 

### Based on raw value only

```{r results='asis'}
#| label: tbl-table-1-weight-raw
#| tbl-cap: Stratified by weight category (raw value only)
t1_bmi <- tableby(update(f, weight_raw ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

## By overweight status

### Based on percentile and raw value

```{r results='asis'}
#| label: tbl-table-1-obese-perc
#| tbl-cap: Stratified by obesity (percentile and raw value)
t1_bmi <- tableby(update(f, overweight_perc ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

For those with a BMI percentile included in the dataset, participants with a BMI percentile $\geq$ 85 were considered have obesity. For those with a missing value for BMI percentile, those with a raw BMI value $\geq$ 25 were considered to have obesity. 

### Based on raw value only

```{r results='asis'}
#| label: tbl-table-1-obese-raw
#| tbl-cap: Stratified by obesity (raw value only)
t1_bmi <- tableby(update(f, overweight_raw ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

## By race

```{r results='asis'}
#| label: tbl-table-1-race
#| tbl-cap: Stratified by race
t1_race <- tableby(update(f, combined_race ~ .),
  data = cv1, control = mycontrols
)
summary(t1_race, labelTranslations = labels(df), pfootnote = T)
```

## By ethnicity

```{r results='asis'}
#| label: tbl-table-1-ethnicity
#| tbl-cap: Stratified by ethnicity
t1_eth <- tableby(update(f, ethnicity ~ .), data = cv1, control = mycontrols)
summary(t1_eth, labelTranslations = labels(df), pfootnote = T)
```

## By specialty

```{r results='asis'}
#| label: tbl-table-1-specialty
#| tbl-cap: Stratified by specialty at diagnosis
t1_spec <- tableby(update(f, pcosdx_specialty ~ .),
  data = cv1, control = mycontrols
)
summary(t1_spec, labelTranslations = labels(df), pfootnote = T)
```
