---
title: "CALICO Aim 1 Analyses"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(arsenal)
library(gtsummary)
library(sjPlot)
library(pROC)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/BDC/Janet Snell-Bergeon/CALICO",
  "Linux" = "/home/tim/OneDrive/Vigers/BDC/Janet Snell-Bergeon/CALICO"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
#| include: false
# Load data
load("./Data_Clean/analysis_data.RData")
# Create additional categorical variables per Grayson and Melanie
df$cv_ft_above_uln <- df$cv_ft_perc > 100
label(df$cv_ft_above_uln) <- "Free Testosterone > 100% ULN"
df$cv_tt_above_uln <- df$cv_tt_perc > 100
label(df$cv_tt_above_uln) <- "Total Testosterone > 100% ULN"
df$cv_ft_perc_over_tt_perc_under <- df$cv_tt_perc <= 100 & df$cv_ft_perc > 100
label(df$cv_ft_perc_over_tt_perc_under) <- "TT <= 100% ULN & FT > 100% ULN"
df$cv_dheas_above_uln <- df$cv_dheas_perc > 100
label(df$cv_dheas_above_uln) <- "DHEAS > 100% ULN"
df$cv_androstendione_above_uln <- df$cv_androstendione_perc > 100
label(df$cv_androstendione_above_uln) <- "Androstendione > 100% ULN"
df$cv_amh_above_uln <- df$cv_amh_perc > 100
label(df$cv_amh_above_uln) <- "AMH > 100% ULN"
df$cv_lh_fsh <- df$cv_lh / df$cv_fsh
label(df$cv_lh_fsh) <- "LH:FSH ratio"
df$lh_fsh_ratio_above_2 <- df$cv_lh_fsh > 2
label(df$lh_fsh_ratio_above_2) <- "LH:FSH ratio > 2"
df$lh_fsh_ratio_under_1 <- df$cv_lh_fsh < 1
label(df$lh_fsh_ratio_under_1) <- "LH:FSH ratio < 1"
df$cv_shbg_under_20 <- df$cv_shbg < 20
label(df$cv_shbg_under_20) <- "SHBG < 20 "
df$cv_sbp_cat <- cut(df$cv_sbp,
  breaks = c(-Inf, 120, 130, 140, Inf),
  labels = c("<120", "120-129", "130-139", "140+"), right = F
)
label(df$cv_sbp_cat) <- "SBP category"
df$cv_a1c_cat <- cut(df$cv_a1c,
  breaks = c(-Inf, 5.6, 6.4, Inf),
  labels = c("<=5.6", "5.7-6.4", ">=6.5")
)
label(df$cv_a1c_cat) <- "HbA1c category"
df$cv_fbg_cat <- cut(df$cv_fbg,
  breaks = c(-Inf, 100, 126, Inf),
  labels = c("<100", "100-125", "126+"), right = F
)
label(df$cv_fbg_cat) <- "Fasting glucose category"
df$cv_2hrglucoseogtt_cat <- cut(df$cv_2hrglucoseogtt,
  breaks = c(-Inf, 140, 200, Inf), labels = c("<140", "140-199", "200+"),
  right = F
)
label(df$cv_2hrglucoseogtt_cat) <- "2-hr glucose category"
df$cv_tg_cat <- cut(df$cv_tg,
  breaks = c(-Inf, 90, 150, Inf),
  labels = c("<90", "90-149", "150+"),
  right = F
)
label(df$cv_tg_cat) <- "Triglycerides category"
df$cv_hdl_under_45 <- df$cv_hdl < 45
label(df$cv_hdl_under_45) <- "HDL < 45"
df$cv_alt_above_35 <- df$cv_alt > 35
label(df$cv_alt_above_35) <- "ALT > 35"
# Variable lists
demo_vars <- c(
  "site", "pcosdx_age", "combined_race", "ethnicity", "insur_type"
)
aim1_vars <- c(
  "cv_bmi", "cv_bmi_percentile", "cv_bmi_z", "cv_hirsutism_num",
  "cv_hirsutism_cat", "cv_acneface", "cv_acneother___1", "cv_acneother___2",
  "cv_acneother___0", "cv_acneother___unk", "cv_ft", "cv_ft_perc",
  "cv_ft_above_uln", "cv_tt", "cv_tt_perc", "cv_tt_above_uln",
  "cv_ft_perc_over_tt_perc_under", "cv_dheas", "cv_dheas_perc",
  "cv_dheas_above_uln", "cv_androstendione", "cv_androstendione_perc",
  "cv_androstendione_above_uln", "cv_lh", "cv_fsh", "cv_lh_fsh",
  "lh_fsh_ratio_above_2", "lh_fsh_ratio_under_1", "cv_amh", "cv_amh_above_uln",
  "cv_a1c", "cv_a1c_cat", "cv_tg", "cv_tg_cat", "cv_hdl", "cv_hdl_under_45",
  "cv_shbg", "cv_shbg_under_20", "cv_alt", "cv_alt_above_35", "cv_fbg",
  "cv_fbg_cat", "cv_fastinsulin", "cv_2hrglucoseogtt", "cv_2hrglucoseogtt_cat",
  "cv_acanthosisneck", "cv_waist", "cv_osa_sx", "cv_medications___1",
  "cv_medications___2", "cv_medications___3", "cv_medications___4",
  "cv_medications___5", "cv_medications___6", "cv_medications___7",
  "cv_medications___8", "cv_medications___9", "cv_medications___10",
  "cv_medications___11", "cv_medications___12", "cv_medications___13",
  "cv_medications___14", "cv_medications___15", "cv_medications___16",
  "cv_medications___17", "cv_medications___18", "cv_medications___19",
  "cv_medications___20", "cv_medications___21", "cv_medications___22",
  "cv_medications___23", "cv_medications___32", "cv_medications___24",
  "cv_medications___25", "cv_medications___26", "cv_medications___27",
  "cv_medications___28", "cv_medications___29", "cv_medications___30",
  "cv_medications___31", "cv_medications___60", "cv_medications___0",
  "cv_medications___unk"
)
# Basic formula for tables
f <- as.formula(paste("~", paste(aim1_vars, collapse = "+")))
f <- update(f, . ~ . + cv_sbp + cv_sbp_cat + cv_dbp + pcosdx_age +
  pcosdx_menarche + pcosdx_famhx_parent___7 + pcosdx_famhx___7 +
  pcosdx_any_famhx___7 + pcosdx_famhx_parent___27 + pcosdx_famhx___27 +
  pcosdx_any_famhx___27 +
  combined_race + ethnicity)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct")
  )
# Diagnosis visit, remove those on hormone meds (likely data entry issue)
cv1 <- df %>%
  group_by(record_number) %>%
  filter(
    cv_monthssincepcosdx == 0, cv_medications___5 == "Unchecked",
    cv_medications___6 == "Unchecked", cv_medications___7 == "Unchecked",
    cv_medications___9 == "Unchecked", cv_medications___10 == "Unchecked",
    cv_medications___11 == "Unchecked", cv_medications___12 == "Unchecked"
  ) %>%
  ungroup()
```

# Data cleaning

Data were limited to diagnosis visit, and those without a diagnostic visit were excluded. Some participants had appeared to have incorrectly entered hormone-containing medications, so diagnostic visits were excluded if any of the following variables were checked (cyclic progesterone was not excluded):

- `cv_medications___5` (Estrogen-containing pill)
- `cv_medications___6` (Estrogen-containing patch)
- `cv_medications___7` (Estrogen-containing ring)
- `cv_medications___9` (Daily progesterone)
- `cv_medications___10` (Progesterone implant)
- `cv_medications___11` (Progesterone IUD)
- `cv_medications___12` (Progesterone injection)

# Demographics

```{r results='asis'}
#| label: tbl-table-1
#| tbl-cap: Participant demographics
f1 <-
  as.formula(paste0(
    "~",
    paste0(c(demo_vars, "combined_race"), collapse = "+")
  ))
t1 <- tableby(f1, data = cv1, control = mycontrols)
summary(t1, labelTranslations = labels(df), pfootnote = T)
```

# Aim 1 comparisons

## By age

```{r results='asis'}
#| label: tbl-table-1-age
#| tbl-cap: Stratified by age group at diagnosis
t1_age <- tableby(update(f, age_group ~ .), data = cv1, control = mycontrols)
summary(t1_age, labelTranslations = labels(df), pfootnote = T)
```

## By weight category

### Based on percentile and raw value

```{r results='asis'}
#| label: tbl-table-1-weight-perc
#| tbl-cap: Stratified by weight category (percentile and raw value)
t1_bmi <- tableby(update(f, weight_perc ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

For those with a BMI percentile included in the dataset, participants with a BMI percentile $\geq$ 85 were considered have obesity. For those with a missing value for BMI percentile, those with a raw BMI value $\geq$ 25 were considered to have obesity. 

### Based on raw value only

```{r results='asis'}
#| label: tbl-table-1-weight-raw
#| tbl-cap: Stratified by weight category (raw value only)
t1_bmi <- tableby(update(f, weight_raw ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

## By overweight status

### Based on percentile and raw value

```{r results='asis'}
#| label: tbl-table-1-obese-perc
#| tbl-cap: Stratified by obesity (percentile and raw value)
t1_bmi <- tableby(update(f, overweight_perc ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

For those with a BMI percentile included in the dataset, participants with a BMI percentile $\geq$ 85 were considered have obesity. For those with a missing value for BMI percentile, those with a raw BMI value $\geq$ 25 were considered to have obesity. 

### Based on raw value only

```{r results='asis'}
#| label: tbl-table-1-obese-raw
#| tbl-cap: Stratified by obesity (raw value only)
t1_bmi <- tableby(update(f, overweight_raw ~ .), data = cv1, control = mycontrols)
summary(t1_bmi, labelTranslations = labels(df), pfootnote = T)
```

## By race

```{r results='asis'}
#| label: tbl-table-1-race
#| tbl-cap: Stratified by race
t1_race <- tableby(update(f, combined_race ~ .),
  data = cv1, control = mycontrols
)
summary(t1_race, labelTranslations = labels(df), pfootnote = T)
```

## By ethnicity

```{r results='asis'}
#| label: tbl-table-1-ethnicity
#| tbl-cap: Stratified by ethnicity
t1_eth <- tableby(update(f, ethnicity ~ .), data = cv1, control = mycontrols)
summary(t1_eth, labelTranslations = labels(df), pfootnote = T)
```

## By specialty

```{r results='asis'}
#| label: tbl-table-1-specialty
#| tbl-cap: Stratified by specialty at diagnosis
t1_spec <- tableby(update(f, pcosdx_specialty ~ .),
  data = cv1, control = mycontrols
)
summary(t1_spec, labelTranslations = labels(df), pfootnote = T)
```

# ROC Analysis

We used an ROC curve to assess total testosterone % ULN's ability to distinguish between participants with and without free testosterone > 100% ULN.

```{r results='asis'}
cv1$cv_ft_above_uln <- factor(cv1$cv_ft_above_uln,
  levels = c(F, T),
  labels = c("FT <= 100% ULN", "FT > 100% ULN")
)
t1 <- tableby(cv_ft_above_uln ~ cv_tt_perc,
  data = cv1
)
summary(t1, labelTranslations = labels(df), pfootnote = T)
```

```{r}
# Per Melanie, the outcome is free testosterone > 100% ULN vs. <= 100% ULN,
# and the predictor is total testosterone % ULN
r <- roc(cv_ft_above_uln ~ cv_tt_perc, data = cv1, quiet = T)
plot(r, print.auc = T, print.thres = T)
ci.auc(r)
# gt(data.frame(ci.coords(r, "best",
#   best.policy = "random",
#   ret = c(c("sensitivity", "specificity", "fp", "tp", "fn", "tn"))
# ))) %>% fmt_number(decimals = 3)
gt(coords(r, "best", ret = "all")) %>% fmt_number(decimals = 3)
```

# Liver Analysis

```{r}
# Load data
load("./Data_Clean/analysis_data.RData")
# LF:FSH ratio
df$cv_lh_fsh <- df$cv_lh / df$cv_fsh
label(df$cv_lh_fsh) <- "LH:FSH"
# Combine lean and overweight
df$weight_cat <- factor(df$weight_cat,
  levels = c("Normal weight", "Overweight", "Obese"),
  labels = c("Non-Obese", "Non-Obese", "Obese")
)
label(df$weight_cat) <- "Weight Category (by percentile then raw value)"
# Combine race categories
df$combined_race <- factor(df$combined_race,
  levels = c(
    "Caucasian", "African American", "American Indian or Alaska Native", "Asian",
    "More than one", "Other", "Pacific Islander"
  ),
  labels = c(
    "Caucasian", "African American", "Other/More than one", "Other/More than one",
    "Other/More than one", "Other/More than one", "Other/More than one"
  )
)
label(df$combined_race) <- "Race"
# Combine other variable levels
df$cv_acneface <- factor(df$cv_acneface,
  levels = c(
    "Mild", "Moderate (pustular)", "Severe (nodular, cystic, scarring)", "None",
    "Acne noted by clinician, severity unknown."
  ),
  labels = c(
    "None/Mild/Unknown", "Moderate/Severe", "Moderate/Severe",
    "None/Mild/Unknown", "None/Mild/Unknown"
  )
)
label(df$cv_acneface) <- "Acne severity on the face"
df$cv_acanthosisneck <- factor(df$cv_acanthosisneck,
  levels = c(
    "Back of neck, barely visible (minimal)", "Back of neck, obvious (mild)",
    "Lateral sides of neck (moderate)", "Circumferential (severe)", "None"
  ),
  labels = c("Present", "Present", "Present", "Present", "Absent")
)
label(df$cv_acanthosisneck) <- "Acanthosis at the neck"
df$cv_osa_sx <- factor(df$cv_osa_sx,
  levels = c(
    "Yes, not addressed", "Yes, prescribed CPAP but not using it",
    "Yes, referral placed for sleep study", "No, treated with CPAP",
    "No, not currently treated"
  ),
  labels = c("Present", "Present", "Present", "Absent", "Absent")
)
label(df$cv_osa_sx) <- "Symptoms of sleep apnea (snoring, frequent awakening, AM headaches, daytime fatigue, napping during day)"
# Variable lists
demo_vars <- c(
  "pcosdx_age", "combined_race", "ethnicity", "insur_type"
)
liver_vars <- c(
  "pcosdx_menarche", "cv_alt", "cv_tg", "cv_hdl", "cv_shbg",
  "cv_a1c", "cv_fbg", "cv_2hrglucoseogtt", "cv_fastinsulin", "cv_tt_perc",
  "cv_ft_perc", "cv_dheas_perc", "cv_androstendione_perc", "cv_lh", "cv_fsh",
  "cv_lh_fsh", "cv_amh_perc", "cv_sbp", "cv_dbp", "pcosdx_famhx_parent___7",
  "pcosdx_famhx___7", "cv_hirsutism_num", "cv_waist", "cv_acneface",
  "cv_acanthosisneck", "cv_osa_sx"
)
# Basic formula for tables
f <- as.formula(paste("~", paste(c(demo_vars, liver_vars), collapse = "+")))
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct")
  )
# First visit only, exclude those on hormones at Dx
cv1 <- df %>%
  group_by(record_number) %>%
  filter(
    cv_monthssincepcosdx == 0, cv_medications___5 == "Unchecked",
    cv_medications___6 == "Unchecked", cv_medications___7 == "Unchecked",
    cv_medications___9 == "Unchecked", cv_medications___10 == "Unchecked",
    cv_medications___11 == "Unchecked", cv_medications___12 == "Unchecked"
  ) %>%
  ungroup()
```

## Comparisons by weight category

```{r results='asis'}
t1 <- tableby(update(f, weight_cat ~ .), data = cv1, control = mycontrols)
summary(t1, labelTranslations = labels(df), pfootnote = T)
```

## Comparisons by race

```{r results='asis'}
t2 <- tableby(update(f, combined_race ~ weight_cat + . - combined_race),
  data = cv1, control = mycontrols
)
summary(t2, labelTranslations = labels(df), pfootnote = T)
```

## Interaction models

In the models below, the interaction term indicates whether or not the effect of the independent variable on ALT varies by either weight category or race. 

### By weight category

#### HbA1c

```{r}
m <- lm(cv_alt ~ cv_a1c * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_a1c", "weight_cat"))
```

#### Fasting glucose

```{r}
m <- lm(cv_alt ~ cv_fbg * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_fbg", "weight_cat"))
```

#### Age at time of PCOS dx

```{r}
m <- lm(cv_alt ~ pcosdx_age * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("pcosdx_age", "weight_cat"))
```

#### Race

```{r}
m <- lm(cv_alt ~ combined_race * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Ethnicity

```{r}
m <- lm(cv_alt ~ ethnicity * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Insurance type

```{r}
m <- lm(cv_alt ~ insur_type * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Menarche age

```{r}
m <- lm(cv_alt ~ pcosdx_menarche * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("pcosdx_menarche", "weight_cat"))
```

#### Triglycerides

```{r}
m <- lm(cv_alt ~ cv_tg * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_tg", "weight_cat"))
```

#### HDL

```{r}
m <- lm(cv_alt ~ cv_hdl * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_hdl", "weight_cat"))
```

#### SHBG

```{r}
m <- lm(cv_alt ~ cv_shbg * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_shbg", "weight_cat"))
```

#### Fasting Insulin

```{r}
m <- lm(cv_alt ~ cv_fastinsulin * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_fastinsulin", "weight_cat"))
```

#### Free Testosterone (% ULN)

```{r}
m <- lm(cv_alt ~ cv_ft_perc * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_ft_perc", "weight_cat"))
```

#### DHEAS (% upper)

```{r}
m <- lm(cv_alt ~ cv_dheas_perc * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_dheas_perc", "weight_cat"))
```

#### Androstenedione (% ULN)

```{r}
m <- lm(cv_alt ~ cv_androstendione_perc * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_androstendione_perc", "weight_cat"))
```

#### LH

```{r}
m <- lm(cv_alt ~ cv_lh * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_lh", "weight_cat"))
```

#### FSH

```{r}
m <- lm(cv_alt ~ cv_fsh * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_fsh", "weight_cat"))
```

#### LH:FSH

```{r}
m <- lm(cv_alt ~ cv_lh_fsh * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_lh_fsh", "weight_cat"))
```

#### Systolic BP

```{r}
m <- lm(cv_alt ~ cv_sbp * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_sbp", "weight_cat"))
```

#### Diastolic BP

```{r}
m <- lm(cv_alt ~ cv_dbp * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_dbp", "weight_cat"))
```

#### Parent Hx (DMII)

```{r}
m <- lm(cv_alt ~ pcosdx_famhx_parent___7 * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Nonparent Family Hx (DMII)

```{r}
m <- lm(cv_alt ~ pcosdx_famhx___7 * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Acne Severity on Face

```{r}
m <- lm(cv_alt ~ cv_acneface * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Acanthosis at Neck 

```{r}
m <- lm(cv_alt ~ cv_acanthosisneck * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Symptoms of Sleep Apnea

```{r}
m <- lm(cv_alt ~ cv_osa_sx * weight_cat, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

### By race

#### Weight Category

```{r}
m <- lm(cv_alt ~ weight_cat * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Ethnicity (Hispanic: Y/N)

```{r}
m <- lm(cv_alt ~ ethnicity * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Insurance type

```{r}
m <- lm(cv_alt ~ insur_type * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Triglycerides

```{r}
m <- lm(cv_alt ~ cv_tg * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_tg", "combined_race"))
```

#### HgbA1c 

```{r}
m <- lm(cv_alt ~ cv_a1c * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_a1c", "combined_race"))
```

#### Total Testosterone (% of ULN)

```{r}
m <- lm(cv_alt ~ cv_tt_perc * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_tt_perc", "combined_race"))
```

#### Free Testosterone (% of ULN)

```{r}
m <- lm(cv_alt ~ cv_ft_perc * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_ft_perc", "combined_race"))
```

#### Numerical FGS score

```{r}
m <- lm(cv_alt ~ cv_hirsutism_num * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
plot_model(m, type = "pred", terms = c("cv_hirsutism_num", "combined_race"))
```

#### Acne Severity on Face

```{r}
m <- lm(cv_alt ~ cv_acneface * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```

#### Acanthosis at the neck

```{r}
m <- lm(cv_alt ~ cv_acanthosisneck * combined_race, data = cv1)
m %>%
  tbl_regression() %>%
  add_n(location = "level")
```
