---
title: "SWEETHEART & ULTRA Imaging"
author: "Ye Ji Choi"
format:
  revealjs:
    embed-resources: true
    scrollable: true
    transition: slide
    controls-layout: bottom-right
    menu: true
    toc: true
    toc-depth: 3
    fontsize: 15
    number-sections: true
---

Left ventricular (LV), right ventricular (RV) global functions
ESV: End-systolic volume	(mL)
EDV: End-diastolic volume (mL)	
SV: Stroke volume	(mL)
CO: Cardiac output	 (L/min)
EF: Ejection fraction (%)	
HR: Heart rate (1/min)	
thickness_diast: Global myocardial thickness at diastole (mm)	
thickness_total: Global myocardial thickness assessed across the entire cardiac cycle, with values comparable to those measured at systole (mm)
MyoMass_diast: Global myocardial mass at diastole
MyoMass_syst: Global myocardial mass at systole

Left ventricular global peak strain (systolic function)
GRS: Global radial strain (%)
GCS: Global circumferential strain (%)
GLS: Global longitudinal strain (%)	

Tissue characteristics
T1 (ms)	
T2* (ms)

Intracardiac 4D flow
Mitral valve inflow (diastolic function)
E wave: Early diastolic peak inflow velocity (cm/s)
A wave: Peak inflow velocity during atrial contraction (cm/s)
E/A ratio: Ratio of E wave to A wave peak inflow velocity (–)

Aortic 4D flow (Region of interest: thoracic aorta)
PWV: Pulse wave velocity (central arterial stiffness) (m/s) 	
MeanWSS: Mean wall shear stress (Pa)	
MedianWSS: Median wall shear stress (Pa)
High5pctWSS: Median of 5% highest wall shear stress (Pa)

```{r include = F}
library(jsonlite)
library(aws.s3)
library(knitr)
library(dplyr)
library(kableExtra)
library(tidyverse)
library(lmerTest)
library(emmeans)
library(broom)
library(ggbeeswarm)
library(lemon)
library(quantreg)
library(readxl)
library(lubridate)
```

```{r include = F}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/pylell/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r include = F}
## Create an S3 client
keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")

Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```

```{r}
im_vars <- c("imaging_date",
             "imaging_hr",
             "af_high5pctwss",
             "af_medianwss",
             "af_meanwss",
             "af_pwv_xcor3",
             "mv_flow_ea",
             "mv_flow_a",
             "mv_flow_e",
             "tt2",
             "tt1",
             "long_peak",
             "circum_peak",
             "radial_peak",
             "rvef",
             "lvef",
             "ef_percent",
             "rvco",
             "lvco",
             "cardiac_output",
             "rvsv",
             "lvsv",
             "sv_ml",
             "rvedv",
             "lvedv",
             "edv_ml",
             "rvesv",
             "lvesv",
             "esv_ml",
             "myo_mass_syst",
             "myo_mass_dias",
             "myo_thickness_total",
             "myo_thickness_dias")
```

```{r}
harm_dat <- read.csv(file.path(root_path, "Data Harmonization/Data Clean/harmonized_dataset.csv"), na = "")

ultra_swht_coenrolled <- harm_dat %>%
  filter(study == "ULTRA" & !is.na(swht_id)) %>%
  pull(swht_id) %>%
  unique()

harm_dat_subset <- harm_dat %>%
  filter(study == "ULTRA" | !is.na(swht_id)) %>%
  mutate(imaging_visit = case_when(imaging_visit == 1 ~ "ultra_baseline", 
                                   imaging_visit == 2 ~ "ultra_follow_up",
                                   record_id %in% ultra_swht_coenrolled ~ "ultra_swht_follow_up",
                                   study == "SWEETHEART" ~ "swht_cs"),
         sex = case_when(study == "ULTRA" ~ "Male", T ~ sex),
         age = interval(dob, imaging_date) / years(1)) %>%
  # dplyr::select(swht_id, record_id, mrn, group, sex, age, imaging_visit, study, all_of(im_vars)) %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id, imaging_visit)) %>%
  filter(!is.na(mv_flow_ea))
```

```{r}
summary(arsenal::tableby(imaging_visit ~ group + age + sex + bmi + acr_u, data = harm_dat_subset))

```

```{r echo = F}
library(dplyr)
library(ggplot2)
library(purrr)
library(cowplot)
plots <- map(names(swht_imaging_full)[8:34], function(var) {
  ggplot(swht_imaging_full, aes(x = group, y = .data[[var]], fill = group)) +
    geom_boxplot(size = 1) +
    theme(panel.background = element_blank(),
          text = element_text(size = 18),
          axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 60, hjust = 1),
          legend.position = "none") +
    scale_fill_manual(values = c("#588157", "#f08080", "#98c1d9"))
})
  

plot_grid(plotlist = plots[1:9], ncol = 3)
plot_grid(plotlist = plots[10:18], ncol = 3)
plot_grid(plotlist = plots[19:27], ncol = 3)

table(swht_imaging_full$group)
```