---
title: "SWEETHEART & ULTRA & IMPROVE Imaging"
author: "Ye Ji Choi"
format:
  revealjs:
    embed-resources: true
    scrollable: true
    transition: slide
    controls-layout: bottom-right
    menu: true
    toc: true
    toc-depth: 3
    fontsize: 15
    number-sections: true
---

<!--
Left ventricular (LV), right ventricular (RV) global functions
  ESV: End-systolic volume (mL)
  EDV: End-diastolic volume (mL)
  SV:  Stroke volume (mL)
  CO:  Cardiac output (L/min)
  EF:  Ejection fraction (%)
  HR:  Heart rate (1/min)
  thickness_diast: Global myocardial thickness at diastole (mm)
  thickness_total: Global myocardial thickness across cardiac cycle (mm)
  MyoMass_diast: Global myocardial mass at diastole
  MyoMass_syst:  Global myocardial mass at systole

Left ventricular global peak strain (systolic function)
  GRS: Global radial strain (%)
  GCS: Global circumferential strain (%)
  GLS: Global longitudinal strain (%)

Tissue characteristics
  T1 (ms)
  T2* (ms)

Intracardiac 4D flow — Mitral valve inflow (diastolic function)
  E wave:   Early diastolic peak inflow velocity (cm/s)
  A wave:   Peak inflow velocity during atrial contraction (cm/s)
  E/A ratio

Aortic 4D flow (thoracic aorta)
  PWV:          Pulse wave velocity (m/s)
  MeanWSS:      Mean wall shear stress (Pa)
  MedianWSS:    Median wall shear stress (Pa)
  High5pctWSS:  Median of top 5% wall shear stress (Pa)
-->

```{r setup, include=FALSE}
# ── Packages ──────────────────────────────────────────────────────────────────
library(jsonlite)
library(aws.s3)
library(knitr)
library(kableExtra)
library(tidyverse)
library(lmerTest)
library(emmeans)
library(broom)
library(broom.mixed)
library(ggbeeswarm)
library(lemon)
library(quantreg)
library(readxl)
library(lubridate)
library(cowplot)
library(ggpubr)
library(Hmisc)
library(arsenal)
```

```{r paths, include=FALSE}
# ── User-specific paths ──────────────────────────────────────────────────────
user <- Sys.info()[["user"]]

if (user == "choiyej") {
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") {
  root_path <- ""
  git_path  <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") {
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path  <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/Users/pylell/Library/CloudStorage/OneDrive-UW/keys.json")
} else {
  stop("Unknown user: please specify root path.")
}
```

```{r s3-setup, include=FALSE}
# ── S3 credentials ────────────────────────────────────────────────────────────
Sys.setenv(
  AWS_ACCESS_KEY_ID     = keys$MY_ACCESS_KEY,
  AWS_SECRET_ACCESS_KEY = keys$MY_SECRET_KEY,
  AWS_DEFAULT_REGION    = "",
  AWS_REGION            = "",
  AWS_S3_ENDPOINT       = "s3.kopah.uw.edu"
)
```

```{r variable-defs, echo=FALSE}
# ── Imaging variables ─────────────────────────────────────────────────────────

# Variables requiring log-transformation
im_vars_log <- c("af_pwv", "mv_flow_a", "mv_flow_e",
                  "lv_myo_thickness_total", "lv_myo_thickness_dias")

# Full set of imaging variables
im_vars <- c(
  "lv_hr", 
  "af_high5pctwss", "af_medianwss", "af_meanwss", "af_pwv",
  "mv_flow_ea", "mv_flow_a", "mv_flow_e",
  "tt2", "tt1",
  "gls", "gcs", "grs",
  "rvef", "lvef",
  "rv_cardiac_output", "lv_cardiac_output",
  "rv_stroke_volume", "lv_stroke_volume",
  "rvedv", "lvedv", "rvesv", "lvesv",
  "lves_mass", "lved_mass",
  "lv_myo_thickness_total", "lv_myo_thickness_dias"
)

# Subset for focused analyses (raw units)
im_vars_raw <- c(
  "lvedv", "lvesv", "lv_stroke_volume", "lvef",
  "rvedv", "rvesv", "rv_stroke_volume", "rvef",
  "lved_mass", "lv_myo_thickness_dias",
  "gls", "lv_hr",
  "af_meanwss", "af_pwv",
  "tt1", "tt2"
)

# BSA-indexed versions
im_vars_bsa <- c(
  "lvedv_bsa", "lvesv_bsa", "lv_stroke_volume_bsa", "lvef",
  "rvedv_bsa", "rvesv_bsa", "rv_stroke_volume_bsa", "rvef",
  "lved_mass_bsa", "lv_myo_thickness_dias",
  "gls", "lv_hr",
  "af_meanwss", "af_pwv",
  "tt1", "tt2"
)

# ── Display labels ────────────────────────────────────────────────────────────
im_labels <- c(
  lv_hr                  = "LV Heart rate (1/min)",
  rv_hr                  = "RV Heart rate (1/min)",
  rvedv                  = "RVEDV (mL)",
  lvedv                  = "LVEDV (mL)",
  rvesv                  = "RVESV (mL)",
  lvesv                  = "LVESV (mL)",
  rv_stroke_volume       = "RV stroke\nvolume (mL)",
  lv_stroke_volume       = "LV stroke\nvolume (mL)",
  rvedv_bsa              = "RVEDV BSA (mL/m\u00b2)",
  lvedv_bsa              = "LVEDV BSA (mL/m\u00b2)",
  rvesv_bsa              = "RVESV BSA (mL/m\u00b2)",
  lvesv_bsa              = "LVESV BSA (mL/m\u00b2)",
  rv_stroke_volume_bsa   = "RV stroke\nvolume BSA (mL/m\u00b2)",
  lv_stroke_volume_bsa   = "LV stroke\nvolume BSA (mL/m\u00b2)",
  rv_cardiac_output      = "RV cardiac\noutput",
  lv_cardiac_output      = "LV cardiac\noutput",
  lves_mass              = "LV myocardial\nmass (systole)",
  lved_mass              = "LV mass (g)",
  lves_mass_bsa          = "LV myocardial\nmass BSA (systole)",
  lved_mass_bsa          = "LV mass BSA (g/m\u00b2)",
  lv_myo_thickness_total = "LV thickness\ntotal (mm)",
  lv_myo_thickness_dias  = "LV\nthickness (mm)",
  rvef                   = "RV\nejection fraction (%)",
  lvef                   = "LV\nejection fraction (%)",
  gls                    = "Global\nlongitudinal strain (%)",
  gcs                    = "Global\ncircumferential strain (%)",
  grs                    = "Global\nradial strain (%)",
  af_high5pctwss         = "Aortic wall shear\nstress (top 5%)",
  af_medianwss           = "Aortic wall shear\nstress (median)",
  af_meanwss             = "Mean wall shear\nstress (Pa)",
  af_pwv                 = "Pulse wave\nvelocity (m/s)",
  tt1                    = "CMR T1 (ms)",
  tt2                    = "CMR T2* (ms)",
  tr2                    = "R2* (1/s)",
  mv_flow_e              = "E wave (cm/s)",
  mv_flow_a              = "A wave (cm/s)",
  mv_flow_ea             = "E/A ratio"
)

# MRI variables
mri_vars <- c("avg_c_fsoc", "avg_m_fsoc", 
              "avg_c_r2", "avg_m_r2",
              "avg_c_adc", 
              "avg_pcascl", "tkv",
              "glomerular_pressure",
              "avg_c_t1")
mri_labels <- c("avg_c_fsoc" = "Cortical FSOC", 
                "avg_m_fsoc" = "Medullary FSOC", 
                "avg_c_r2" = "Cortical R2*", 
                "avg_m_r2" = "Medullary R2*",
                "avg_c_adc" = "Cortical ADC", 
                "avg_pcascl" = "ASL",
                "tkv" = "TKV",
                "glomerular_pressure" = "PGLO",
                "avg_c_t1" = "Cortical T1"
)

# Clinical variables
anchor_vars     <- c("log_acr_u", "eGFR_CKD_epi", "gfr_bsa_plasma", "bmi", "hba1c", "sbp", "dbp")
anchor_labels <- c(
  bmi = "BMI", eGFR_CKD_epi = "eGFR", gfr_bsa_plasma = "mGFR",
  hba1c = "HbA1c", log_acr_u = "log(UACR)",
  sbp = "SBP", dbp = "DBP"
)
anchor_levels <- c("bmi", "eGFR_CKD_epi", "gfr_bsa_plasma", "hba1c", "log_acr_u", "sbp", "dbp")

acr_gfr_vars <- c("ln_acr_u", "eGFR_CKD_epi", "gfr_raw_plasma", "gfr_bsa_plasma", "ff")
acr_gfr_labels <- c(
  ln_acr_u        = "ln(UACR)",
  eGFR_CKD_epi    = "eGFR\n(mL/min/1.73m²)",
  gfr_raw_plasma  = "mGFR raw\n(mL/min)",
  gfr_bsa_plasma  = "mGFR BSA\n(mL/min/1.73m²)",
  ff              = "Filtration\nfraction"
)

# Delta variables
im_vars_delta <- c("rvesv", "rvef", "rvedv", "rv_stroke_volume", "rv_cardiac_output",
                   "lvesv", "lvef", "lvedv", "lv_stroke_volume", "lv_cardiac_output",
                   "lv_hr", "lv_cardiac_output", "grs", "gls", "gcs",
                   paste0(im_vars_log, "_log10"))
mri_vars_delta <- c("avg_c_r2", "avg_m_r2", "avg_pcascl", "glomerular_pressure")

anchor_vars_delta <- anchor_vars

# ── Delta label builders ─────────────────────────────────────────────────────
# Construct named vectors that map "<var>_delta" → "Δ <label>" so that
# delta heatmaps and forest plots display readable axis text.
make_delta_labels <- function(base_labels, suffix = "_delta", prefix = "\u0394 ") {
  nms <- paste0(names(base_labels), suffix)
  labs <- paste0(prefix, gsub("\n", " ", base_labels))
  setNames(labs, nms)
}

im_labels_delta      <- make_delta_labels(im_labels)
mri_labels_delta     <- make_delta_labels(mri_labels)
anchor_labels_delta  <- make_delta_labels(anchor_labels)

# Also need labels for log-transformed delta variables
im_log_delta_labels <- setNames(
  paste0("\u0394 log10(", gsub("\n", " ", im_labels[im_vars_log]), ")"),
  paste0(im_vars_log, "_log10_delta")
)
im_labels_delta <- c(im_labels_delta, im_log_delta_labels)
```

```{r data-import, include=FALSE}
# ── Import harmonized dataset ─────────────────────────────────────────────────
harm_dat <- read.csv(
  file.path(root_path, "Data Harmonization/Data Clean/harmonized_dataset.csv"),
  na = ""
)

# ── Identify co-enrolled participants ─────────────────────────────────────────
ultra_swht_coenrolled <- harm_dat %>%
  filter(study == "ULTRA" & !is.na(swht_id)) %>% pull(swht_id) %>% unique()
ultra_swht_coenrolled_ultra <- harm_dat %>%
  filter(study == "ULTRA" & !is.na(swht_id)) %>% pull(ultra_id) %>% unique()
swht_imp_coenrolled <- harm_dat %>%
  filter(study == "SWEETHEART" & !is.na(improve_id)) %>% pull(swht_id) %>% unique()
imp_swht_coenrolled <- harm_dat %>%
  filter(study == "IMPROVE" & !is.na(swht_id)) %>% pull(improve_id) %>% unique()

primary_studies <- c("ULTRA", "IMPROVE", "SWEETHEART")
```

```{r data-subset, include=FALSE}
# ── Build analysis subset ─────────────────────────────────────────────────────
# ══════════════════════════════════════════════════════════════════════════════
# 1. Label primary study rows with imaging_visit_id
# ══════════════════════════════════════════════════════════════════════════════
harm_labeled <- harm_dat %>%
  filter(!is.na(ultra_id) | !is.na(swht_id) | !is.na(improve_id)) %>%
  filter(!is.na(group)) %>%
  dplyr::mutate(
    imaging_visit_id = case_when(
      # ULTRA
      study == "ULTRA" & visit == "baseline" &
        record_id %in% ultra_swht_coenrolled_ultra          ~ "ultra_swht_baseline",
      study == "ULTRA" & visit == "baseline"                 ~ "ultra_baseline",
      study == "ULTRA" & visit == "1_week"                   ~ "ultra_follow_up",
      # IMPROVE
      study == "IMPROVE" & visit == "baseline" &
        record_id %in% imp_swht_coenrolled                   ~ "improve_swht_baseline",
      study == "IMPROVE" & visit == "baseline"               ~ "improve_baseline",
      study == "IMPROVE" & visit == "3_months_post_surgery"  ~ "improve_3mo",
      study == "IMPROVE" & visit == "12_months_post_surgery" ~ "improve_12mo",
      # SWEETHEART
      study == "SWEETHEART" &
        record_id %in% ultra_swht_coenrolled                 ~ "ultra_swht_follow_up",
      study == "SWEETHEART" &
        record_id %in% swht_imp_coenrolled                   ~ "improve_swht_follow_up",
      study == "SWEETHEART"                                  ~ "swht_cs"
    ),
    sex       = if_else(study == "ULTRA", "Male", sex),
    cmri_date = if_else(procedure == "cardio_abdominal_mri", date, NA_character_),
    age       = coalesce(interval(dob, cmri_date) / years(1), age),
    tkv       = coalesce(total_kidney_volume_ml, total_kidney_volume_ml_manual)
  )

# ══════════════════════════════════════════════════════════════════════════════
# 2. Split into primary (labeled) vs supplementary (unlabeled)
# ══════════════════════════════════════════════════════════════════════════════
primary_rows <- harm_labeled %>% filter(!is.na(imaging_visit_id))

supplementary_rows <- harm_labeled %>%
  filter(
    is.na(imaging_visit_id),
    !(study %in% primary_studies) 
  ) %>%
  dplyr::select(-imaging_visit_id)

# ══════════════════════════════════════════════════════════════════════════════
# 3. Build a reference date per person × imaging visit from primary rows
# ══════════════════════════════════════════════════════════════════════════════
visit_dates <- primary_rows %>%
  filter(!is.na(date)) %>%
  dplyr::summarise(
    ref_date = first(na.omit(date)),
    .by = c(mrn, imaging_visit_id)
  ) %>%
  dplyr::mutate(ref_date = as.Date(ref_date))

# ══════════════════════════════════════════════════════════════════════════════
# 4. Match each supplementary row to the closest imaging visit (within window)
# ══════════════════════════════════════════════════════════════════════════════
match_supplementary <- function(supp_rows, visit_dates, max_days) {
  supp_rows %>%
    filter(!is.na(date), !is.na(mrn)) %>%
    dplyr::mutate(supp_date = as.Date(date)) %>%
    left_join(visit_dates, by = "mrn", relationship = "many-to-many") %>%
    dplyr::mutate(date_diff = abs(supp_date - ref_date)) %>%
    filter(date_diff <= max_days) %>%
    slice_min(date_diff, n = 1, by = c(mrn, record_id, date, procedure)) %>%
    select(-supp_date, -ref_date, -date_diff, mrn, record_id, date, procedure)
}

supplementary_matched     <- match_supplementary(supplementary_rows, visit_dates, 30)
supplementary_matched_500 <- match_supplementary(supplementary_rows, visit_dates, 500)

# ══════════════════════════════════════════════════════════════════════════════
# 5. Stack, collapse, and derive: one row per person × imaging visit
# ══════════════════════════════════════════════════════════════════════════════
collapse_and_derive <- function(primary, supplementary) {
  bind_rows(
    primary       %>% dplyr::mutate(study_priority = 2L),
    supplementary %>% dplyr::mutate(study_priority = 1L)
  ) %>%
    arrange(study_priority) %>%
    dplyr::summarise(
      across(
        where(negate(is.numeric)),
        ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))
      ),
      across(
        where(is.numeric),
        ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))
      ),
      .by = c(mrn, imaging_visit_id)
    ) %>%
    dplyr::mutate(
      # Kidney function groupings
      uacr_100 = if_else(acr_u >= 100, "UACR >=100 mg/g", "UACR <100 mg/g"),
      dkd_group_100 = case_when(
        eGFR_CKD_epi < 90 | acr_u >= 100 ~ "DKD",
        eGFR_CKD_epi >= 90 & acr_u < 100 ~ "non_DKD"
      ),
      dkd_group_30 = case_when(
        eGFR_CKD_epi < 90 | acr_u >= 30 ~ "DKD",
        eGFR_CKD_epi >= 90 & acr_u < 30 ~ "non_DKD"
      ),
      uacr_group = factor(
        case_when(
          eGFR_CKD_epi < 90 | acr_u >= 100  ~ "eGFR < 90 and/or UACR >=100 mg/g",
          eGFR_CKD_epi >= 90 & acr_u < 100  ~ "eGFR >= 90 and UACR <100 mg/g"
        ),
        levels = c("eGFR >= 90 and UACR <100 mg/g", "eGFR < 90 and/or UACR >=100 mg/g")
      ),
      egfr_cat = factor(
        case_when(
          eGFR_CKD_epi < 60                        ~ "<60",
          eGFR_CKD_epi >= 60 & eGFR_CKD_epi < 90  ~ "60\u2013<90",
          eGFR_CKD_epi >= 90                       ~ "\u226590"
        ),
        levels = c("<60", "60\u2013<90", "\u226590")
      ),
      acr_u_cat3 = factor(
        case_when(
          acr_u < 30                 ~ "<30",
          acr_u >= 30 & acr_u < 300  ~ "30\u2013<300",
          acr_u >= 300               ~ "\u2265300"
        ),
        levels = c("<30", "30\u2013<300", "\u2265300")
      ),
      acr_u_cat300 = if_else(acr_u >= 300, "\u2265300", "<300"),
      tr2       = (1 / tt2) * 1000,
      log_acr_u = log(acr_u),
      gls = na_if(gls, 0),
      gcs = na_if(gcs, 0),
      grs = na_if(grs, 0),
      bsa = sqrt(height * weight / 3600),
      lvesv_bsa              = lvesv / bsa,
      rvesv_bsa              = rvesv / bsa,
      lvedv_bsa              = lvedv / bsa,
      rvedv_bsa              = rvedv / bsa,
      lv_stroke_volume_bsa   = lv_stroke_volume / bsa,
      rv_stroke_volume_bsa   = rv_stroke_volume / bsa,
      lved_mass_bsa          = lved_mass / bsa,
      lves_mass_bsa          = lves_mass / bsa,
      height_m               = height / 100,
      lvesv_height            = lvesv / height_m,
      rvesv_height            = rvesv / height_m,
      lvedv_height            = lvedv / height_m,
      rvedv_height            = rvedv / height_m,
      lv_stroke_volume_height = lv_stroke_volume / height_m,
      rv_stroke_volume_height = rv_stroke_volume / height_m,
      lved_mass_height        = lved_mass / height_m,
      lves_mass_height        = lves_mass / height_m
    )
}

harm_dat_subset     <- collapse_and_derive(primary_rows, supplementary_matched)
harm_dat_subset_500 <- collapse_and_derive(primary_rows, supplementary_matched_500)
```

```{r echo = F, include = F}
# # date diff between IMPROVE BL and SWHT
# harm_dat_subset %>%
#   filter(imaging_visit_id %in% c("improve_swht_baseline", "improve_swht_follow_up")) %>%
#   summarise(
#     n_visits  = n_distinct(imaging_visit_id),
#     date_diff_yrs = as.numeric(max(as.Date(cmri_date)) - min(as.Date(cmri_date)), units = "days")/365,
#     .by = mrn
#   ) %>%
#   filter(n_visits == 2)
#     
# # date diff between ULTRA BL and SWHT
# harm_dat_subset %>%
#   filter(imaging_visit_id %in% c("ultra_swht_baseline", "ultra_swht_follow_up")) %>%
#   summarise(
#     n_visits  = n_distinct(imaging_visit_id),
#     date_diff_yrs = as.numeric(max(as.Date(cmri_date)) - min(as.Date(cmri_date)), units = "days")/365,
#     .by = mrn
#   ) %>%
#   filter(n_visits == 2) %>%
#   pull(date_diff_yrs)

```

```{r analysis-subsets, include=FALSE}
# check for CMR from Dylan's list
# marcus_proteomics_id <- c("RH-68-T", "RH-49-T", "RH-63-T", "RH-23-T", "CRC-04", "IT_11", "IT_12", 
#                           "RH-71-T", "RH-62-T", "RH-50-T", "IT_13", "RH-74-T", "RH-72-T")
# harm_dat_subset %>%
#   filter(croc_id %in% marcus_proteomics_id | rh_id %in% marcus_proteomics_id |
#            improve_id %in% marcus_proteomics_id)

# ── Analysis subsets ──────────────────────────────────────────────────────────
harm_dat_baseline <- harm_dat_subset %>%
  filter(visit == "baseline" & 
           imaging_visit_id %nin% c("improve_swht_follow_up", "ultra_swht_follow_up"))

harm_dat_obt2d_baseline <- harm_dat_baseline %>%
  filter(group != "Lean Control")

harm_dat_baseline_500 <- harm_dat_subset_500 %>%
  filter(visit == "baseline" & 
           imaging_visit_id %nin% c("improve_swht_follow_up", "ultra_swht_follow_up"))

harm_dat_obt2d_baseline_500 <- harm_dat_subset_500 %>%
  filter(group != "Lean Control") 

# IMPROVE only dataset (for VSG analysis)
improve_dat <- harm_dat_subset %>%
  filter(imaging_visit_id %in% c("improve_swht_baseline", "improve_swht_follow_up",
                                 "improve_baseline", "improve_3mo", "improve_12mo")) %>%
  group_by(mrn) %>%
  dplyr::mutate(days_since_baseline = as.numeric(as.Date(coalesce(cmri_date, date)) - 
                  min(as.Date(coalesce(cmri_date, date))))) %>% # using visit date if cmr date DNE
  ungroup() %>%
  filter(!is.na(cmri_date)) # exclude to those with CMRI data

# Longitudinal subset: ULTRA → SWEETHEART co-enrolled
harm_dat_longitudinal <- harm_dat_subset %>%
  filter(imaging_visit_id %in% c("ultra_swht_baseline", "ultra_swht_follow_up")) %>%
  arrange(mrn, imaging_visit_id) %>%
  dplyr::mutate(
    visit_time = if_else(
      imaging_visit_id == "ultra_swht_baseline", 0, 1
    )
  )

# Longitudinal subset: IMPROVE → SWEETHEART co-enrolled
harm_dat_longitudinal_imp <- harm_dat_subset %>%
  filter(imaging_visit_id %in% c("improve_swht_baseline", "improve_swht_follow_up")) %>%
  arrange(mrn, imaging_visit_id) %>%
  dplyr::mutate(
    visit_time = if_else(
      imaging_visit_id == "improve_swht_baseline", 0, 1
    )
  )

# Combined longitudinal subset: all co-enrolled → SWEETHEART
# Pools ULTRA→SWHT and IMPROVE→SWHT to maximize sample size
harm_dat_longitudinal_all <- harm_dat_subset %>%
  filter(imaging_visit_id %in% c("ultra_swht_baseline", "ultra_swht_follow_up",
                                  "improve_swht_baseline", "improve_swht_follow_up")) %>%
  arrange(mrn, imaging_visit_id) %>%
  dplyr::mutate(
    visit_time = if_else(
      imaging_visit_id %in% c("ultra_swht_baseline", "improve_swht_baseline"), 0, 1
    ),
    visit_label = factor(
      if_else(visit_time == 0, "Baseline", "SWEETHEART Follow-up"),
      levels = c("Baseline", "SWEETHEART Follow-up")
    )
  )
```

```{r delta-dfs, include=FALSE}
# ── Helper: compute baseline → 12-month deltas for IMPROVE ──────────────────
all_vars <- unique(c(im_vars, im_vars_log, mri_vars, anchor_vars))

compute_improve_deltas <- function(dat_subset, ref_cols_from = harm_dat_obt2d_baseline) {
  wide <- dat_subset %>%
    dplyr::filter(study == "IMPROVE" & visit != "3_months_post_surgery") %>%
    dplyr::select(record_id, visit, dplyr::all_of(intersect(all_vars, names(ref_cols_from)))) %>%
    dplyr::mutate(dplyr::across(
      dplyr::all_of(intersect(im_vars_log, names(.))),
      ~ dplyr::if_else(.x <= 0, NA_real_, log10(.x)),
      .names = "{.col}_log10"
    )) %>%
    dplyr::select(-dplyr::all_of(intersect(im_vars_log, names(.)))) %>%
    tidyr::pivot_wider(names_from = visit, values_from = -c(record_id, visit))

  bl_cols <- wide %>% dplyr::select(dplyr::ends_with("_baseline")) %>% names()

  deltas <- purrr::map_dfc(bl_cols, function(col) {
    follow_col <- stringr::str_replace(col, "_baseline", "_12_months_post_surgery")
    delta_name <- stringr::str_replace(col, "_baseline", "_delta")
    if (follow_col %in% names(wide)) {
      tibble::tibble(!!delta_name := wide[[follow_col]] - wide[[col]])
    } else {
      tibble::tibble(!!delta_name := NA_real_)
    }
  })

  dplyr::bind_cols(wide %>% dplyr::select(record_id), deltas)
}

harm_dat_improve_delta     <- compute_improve_deltas(harm_dat_subset)
harm_dat_improve_delta_500 <- compute_improve_deltas(harm_dat_subset_500)
```

# Descriptive tables

## All data

```{r tables, echo=FALSE, results='asis'}
cat('<div style="font-size: 0.55em; overflow-x: auto; width: 100%;">')
summary(tableby(
  imaging_visit_id ~ group + age + sex + bmi +
    kwt(acr_u, "Nmiss", "medianq1q3") + hba1c +
    dkd_group_30 + dkd_group_100 + bsa,
  data = harm_dat_subset, digits = 1, test = F
))
cat('</div>')
```

## Baseline

```{r, echo=FALSE, results='asis'}
cat('<div style="font-size: 0.55em; overflow-x: auto; width: 100%;">')
summary(tableby(
  group ~ imaging_visit_id + age + sex + bmi + sbp + dbp +
    kwt(acr_u, "Nmiss", "medianq1q3") + hba1c + eGFR_CKD_epi +
    acr_u_cat3 + dkd_group_30 + dkd_group_100,
  data = harm_dat_baseline, digits = 1
))
cat('</div>')
```

## Baseline excluding HC

```{r, echo=FALSE, results='asis'}
cat('<div style="font-size: 0.55em; overflow-x: auto; width: 100%;">')
summary(tableby(
  sex ~ group + imaging_visit_id + age + height + weight + bmi +
    kwt(acr_u, "Nmiss", "medianq1q3") + hba1c + eGFR_CKD_epi +
    dkd_group_30 + dkd_group_100,
  data = subset(harm_dat_baseline, group != "Lean Control"), digits = 1
))
cat('</div>')
```

## IMPROVE w/ CMR

```{r, echo=FALSE, results='asis'}
cat('<div style="font-size: 0.55em; overflow-x: auto; width: 100%;">')
summary(arsenal::tableby(imaging_visit_id ~ days_since_baseline, 
                         data = improve_dat), digits = 1)
cat('</div>')
```

# Box plots

```{r boxplot-function, echo=FALSE}
plot_boxgrid <- function(df,
                         vars = im_vars,
                         log_vars = NULL,
                         group_var,
                         id_var = "mrn",
                         paired = FALSE,
                         var_labels = NULL,
                         ncol = 3,
                         per_page = 9,
                         fill_colors = c("#f08080", "#98c1d9"),
                         save = TRUE,
                         save_dir = file.path(root_path, "ULTRA_SWEETHEART/Results/Figures/Box plots"),
                         file_prefix = "plot_page",
                         file_type = "png",
                         width = 12,
                         height = 12,
                         dpi = 300,
                         show_legend = FALSE,
                         fill_label = NULL,
                         text_size = 18,
                         star_size = 10,
                         p_format = "stars",
                         p_digits = 3,
                         p_size = 5,
                         test_method = "wilcox.test") {

  if (is.numeric(vars))     vars     <- names(df)[vars]
  if (!is.null(log_vars) && is.numeric(log_vars)) log_vars <- names(df)[log_vars]

  # Log-transform where requested
  if (!is.null(log_vars)) {
    for (v in log_vars) df[[paste0(v, "_log10")]] <- log10(df[[v]])
  }

  # Filter to complete pairs for paired tests

  if (paired) {
    complete_ids <- df %>%
      group_by(.data[[id_var]]) %>%
      dplyr::summarise(n_groups = n_distinct(.data[[group_var]]), .groups = "drop") %>%
      filter(n_groups == length(unique(df[[group_var]]))) %>%
      pull(.data[[id_var]])
    df <- df %>% filter(.data[[id_var]] %in% complete_ids)
    message(sprintf("Using %d complete pairs for paired analysis", length(complete_ids)))
  }

  # Build individual plots
  plots <- map(seq_along(vars), function(i) {
    v <- vars[i]
    if (!is.null(log_vars) && v %in% log_vars) {
      ycol <- paste0(v, "_log10")
      base_label <- if (!is.null(var_labels) && v %in% names(var_labels)) var_labels[[v]] else v
      ylab <- paste0("log10(", base_label, ")")
    } else {
      ycol <- v
      ylab <- if (!is.null(var_labels) && v %in% names(var_labels)) var_labels[[v]] else v
    }

    comps <- combn(unique(df[[group_var]]), 2, simplify = FALSE)

    p <- ggplot(df, aes(x = as.character(.data[[group_var]]),
                        y = .data[[ycol]],
                        fill = as.character(.data[[group_var]]))) +
      geom_boxplot(size = 1)

    # Add statistical comparisons
    stat_args <- list(
      comparisons = comps,
      method      = test_method,
      paired      = paired,
      hide.ns     = !paired && p_format == "stars"
    )
    if (!paired) stat_args$method.args <- list(var.equal = FALSE)

    if (p_format == "stars") {
      stat_args$label <- "p.signif"
      stat_args$size  <- star_size
    } else {
      stat_args$label      <- "p.format"
      stat_args$size       <- p_size
      stat_args$p.accuracy <- 10^(-p_digits)
    }
    p <- p + do.call(stat_compare_means, stat_args)
    
    p + labs(y = ylab, x = NULL, fill = fill_label) +
      scale_fill_manual(values = fill_colors) +
      scale_y_continuous(expand = expansion(mult = c(0.05, 0.35))) +
      theme(
        panel.background = element_blank(),
        text = element_text(size = text_size),
        axis.text.x = if (show_legend) element_blank() else element_text(angle = 60, hjust = 1),
        legend.position = "none",
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_rect(fill = "transparent", colour = NA)
      )
  })

  legend <- cowplot::get_legend(plots[[1]] + theme(legend.position = "right"))

  # Save pages
  if (save) dir.create(save_dir, recursive = TRUE, showWarnings = FALSE)

  n_plots <- length(plots)
  pages   <- ceiling(n_plots / per_page)

  for (page in seq_len(pages)) {
    idx <- ((page - 1) * per_page + 1):min(page * per_page, n_plots)
    page_grid <- plot_grid(plotlist = plots[idx], ncol = ncol)
    page_plot <- plot_grid(page_grid, legend, ncol = 3, rel_widths = c(1, 0.1, 0.15))
    print(page_plot)
    if (save) {
      ggsave(file.path(save_dir, sprintf("%s_%02d.%s", file_prefix, page, file_type)),
             page_plot, width = width, height = height, dpi = dpi)
    }
  }
  invisible(plots)
}
```

--- 

### Group comparison (3 groups)

```{r, echo=FALSE, fig.width=16, fig.height=14}
# Group comparison (3 groups)
plot_boxgrid(
  df = harm_dat_baseline, group_var = "group",
  var_labels = im_labels, log_vars = im_vars_log,
  fill_colors = c("Lean Control" = "#588157", "Obese Control" = "#98c1d9",
                   "Type 2 Diabetes" = "#f08080"),
  file_prefix = "baseline_group"
)
```

### Sex comparison in obesity/T2D

```{r, echo=FALSE, fig.width=12, fig.height=12}
# Sex comparison in obesity/T2D
plot_boxgrid(
  df = harm_dat_obt2d_baseline, group_var = "sex",
  vars = im_vars_raw, var_labels = im_labels,
  log_vars = im_vars_log, file_prefix = "obt2d_sex",
  ncol = 4, per_page = 16, show_legend = TRUE, fill_label = "Sex",
  width = 18, text_size = 20, star_size = 12,
  p_format = "value", test_method = "wilcox.test"
)
```

### BSA indexed sex comparison in obesity/T2D

```{r, echo=FALSE, fig.width=12, fig.height=12}
write.csv(harm_dat_obt2d_baseline, file.path(root_path, "IMPROVE T2D/Data_Cleaned/cmri_dat_obt2d_baseline_041326.csv"), row.names = F)

# BSA-indexed sex comparison
plot_boxgrid(
  df = harm_dat_obt2d_baseline, group_var = "sex",
  vars = im_vars_bsa, var_labels = im_labels,
  log_vars = im_vars_log, file_prefix = "obt2d_sex_bsa",
  ncol = 4, per_page = 16, show_legend = TRUE, fill_label = "Sex",
  width = 18, text_size = 20, star_size = 12,
  p_format = "value", test_method = "wilcox.test"
)
```

### DKD groupings (DKD defined as UACR >= 30 and/or eGFR < 90)

```{r, echo=FALSE, fig.width=12, fig.height=12}
# DKD groupings (DKD = worse → red)
dkd_colors <- c("DKD" = "#f08080", "non_DKD" = "#98c1d9")

plot_boxgrid(
  df = subset(harm_dat_obt2d_baseline, !is.na(dkd_group_30)),
  group_var = "dkd_group_30", var_labels = im_labels,
  log_vars = im_vars_log, fill_colors = dkd_colors,
  file_prefix = "obt2d_dkd30"
)
```

### DKD groupings (DKD defined as UACR >= 100 and/or eGFR < 90)

```{r, echo=FALSE, fig.width=12, fig.height=12}
plot_boxgrid(
  df = subset(harm_dat_obt2d_baseline, !is.na(dkd_group_100)),
  group_var = "dkd_group_100", var_labels = im_labels,
  log_vars = im_vars_log, fill_colors = dkd_colors,
  file_prefix = "obt2d_dkd100"
)
```

### Albuminuria categories

```{r, echo=FALSE, fig.width=12, fig.height=12}
# Albuminuria categories (≥300 = worse → red)
plot_boxgrid(
  df = subset(harm_dat_obt2d_baseline, !is.na(acr_u_cat300)),
  group_var = "acr_u_cat300", log_vars = im_vars_log,
  var_labels = im_labels, file_prefix = "obt2d_acr300",
  fill_colors = c("<300" = "#98c1d9", "\u2265300" = "#f08080"),
  fill_label = "UACR\nCategory"
)
```

### Longitudinal (paired)

```{r, echo=FALSE, fig.width=16, fig.height=14}
# Longitudinal (paired)
plot_boxgrid(
  df = harm_dat_longitudinal, group_var = "imaging_visit_id",
  log_vars = im_vars_log, var_labels = im_labels,
  file_prefix = "longitudinal", paired = TRUE
)
```

```{r, echo=FALSE}
# arteriosclerosis y/n

```

# Correlation heatmaps

## Helper function

```{r echo = F}
run_spearman_heatmap <- function(data,
                                 imaging_vars,
                                 anchor_vars,
                                 log_vars       = character(0),
                                 anchor_labels  = character(0),
                                 imaging_labels = character(0),
                                 min_obs        = 2,
                                 check_coverage = T) {
  
  # ── Log-transform columns ───────────────────────────────────────────────────
  if (length(log_vars) > 0) {
    data <- data %>%
      dplyr::mutate(dplyr::across(
        dplyr::all_of(log_vars),
        ~ dplyr::if_else(.x <= 0, NA_real_, log10(.x)),
        .names = "{.col}_log10"
      ))
  }
  
  imaging_for_cor <- if (length(log_vars) > 0) {
    c(imaging_vars, paste0(log_vars, "_log10"))
  } else {
    imaging_vars
  }
  
  # ── Spearman correlations ───────────────────────────────────────────────────
  cor_df <- expand.grid(imaging = imaging_for_cor, anchor = anchor_vars,
                        stringsAsFactors = FALSE) %>%
    dplyr::mutate(rho = NA_real_, pval = NA_real_, n_complete = NA_integer_)
  
  for (i in seq_len(nrow(cor_df))) {
    x <- data[[cor_df$imaging[i]]]
    y <- data[[cor_df$anchor[i]]]
    
    if (is.null(x) || is.null(y))           next
    if (!is.numeric(x) || !is.numeric(y))   next
    
    ok <- complete.cases(x, y)
    cor_df$n_complete[i] <- sum(ok)
    
    if (sum(ok) >= min_obs) {
      tmp <- cor.test(x[ok], y[ok], method = "spearman", exact = FALSE)
      cor_df$rho[i]  <- tmp$estimate
      cor_df$pval[i] <- tmp$p.value
    }
  }
  
  cor_df$stars <- cut(cor_df$pval,
                      breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
                      labels = c("***", "**", "*", ""))
  
  # ── Apply axis labels ───────────────────────────────────────────────────────
  cor_df$anchor_label <- if (length(anchor_labels) > 0) {
    dplyr::coalesce(anchor_labels[cor_df$anchor], cor_df$anchor)
  } else {
    cor_df$anchor
  }
  
  cor_df$imaging_label <- if (length(imaging_labels) > 0) {
    dplyr::coalesce(imaging_labels[cor_df$imaging], cor_df$imaging)
  } else {
    cor_df$imaging
  }

  # ── Enforce consistent axis ordering ────────────────────────────────────────
  # imaging axis follows the order of imaging_vars; anchor axis follows anchor_vars
  img_label_order <- unique(cor_df$imaging_label[order(match(cor_df$imaging, imaging_for_cor))])
  anc_label_order <- unique(cor_df$anchor_label[order(match(cor_df$anchor, anchor_vars))])
  cor_df$imaging_label <- factor(cor_df$imaging_label, levels = rev(img_label_order))
  cor_df$anchor_label  <- factor(cor_df$anchor_label,  levels = anc_label_order)

  # ── Coverage check ──────────────────────────────────────────────────────────
  coverage_plot <- NULL
  if (check_coverage) {
    coverage_plot <- ggplot2::ggplot(cor_df,
                                     ggplot2::aes(x = anchor_label,
                                                  y = imaging_label,
                                                  fill = n_complete)) +
      ggplot2::geom_tile(color = "white") +
      ggplot2::geom_text(ggplot2::aes(label = n_complete), size = 3) +
      ggplot2::scale_fill_gradient(low = "white", high = "#2166ac") +
      ggplot2::theme_minimal(base_size = 12) +
      ggplot2::theme(panel.grid = ggplot2::element_blank(),
                     axis.text.x = ggplot2::element_text(angle = 60, hjust = 1)) +
      ggplot2::labs(title = "Complete cases per pair",
                    x = NULL, y = NULL, fill = "n")
  }
  
  # ── Correlation heatmap ─────────────────────────────────────────────────────
  p <- ggplot2::ggplot(cor_df,
                       ggplot2::aes(x = anchor_label,
                                    y = imaging_label,
                                    fill = rho)) +
    ggplot2::geom_tile(color = "white") +
    ggplot2::geom_text(ggplot2::aes(label = stars), color = "black") +
    ggplot2::scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b",
                                  midpoint = 0, limits = c(-1, 1)) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::theme(panel.grid = ggplot2::element_blank(),
                   axis.text.x = ggplot2::element_text(angle = 60, hjust = 1)) +
    ggplot2::labs(x = NULL, y = NULL, fill = "\u03c1")
  
  list(cor_df = cor_df, plot = p, coverage_plot = coverage_plot)
}
```

## CMR vs. clinical vars in all baseline OB + T2D

### Data within 30 days

```{r, echo=FALSE, fig.width=10, fig.height=8}
cmr_clinical <- run_spearman_heatmap(
  data           = harm_dat_obt2d_baseline,
  imaging_vars   = im_vars,
  anchor_vars    = anchor_vars,
  log_vars       = im_vars_log,
  anchor_labels  = anchor_labels,
  imaging_labels = im_labels
)

# Full heatmap
# cmr_clinical
```

```{r focused-heatmap-helper, echo=FALSE}
# ── Helper: focused heatmap from a spearman result ───────────────────────────
plot_focused_heatmap <- function(cor_df,
                                 imaging_filter = NULL,
                                 img_labels     = im_labels,
                                 anc_labels     = anchor_labels,
                                 anc_levels     = anchor_levels,
                                 star_size      = 10) {
  df <- cor_df
  if (!is.null(imaging_filter)) df <- df %>% filter(imaging %in% imaging_filter)

  df %>%
    dplyr::mutate(
      imaging_label = img_labels[imaging],
      anchor        = factor(anchor, levels = anc_levels),
      anchor_label  = factor(anc_labels[anchor], levels = anc_labels[anc_levels])
    ) %>%
    ggplot(aes(x = anchor_label, y = imaging_label, fill = rho)) +
    geom_tile(color = "white") +
    geom_text(aes(label = stars), color = "black", size = star_size) +
    scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b",
                         midpoint = 0, limits = c(-1, 1)) +
    theme_minimal(base_size = 18) +
    theme(panel.grid = element_blank()) +
    labs(fill = "\u03c1", x = NULL, y = NULL)
}

cmr_focus_vars <- c("lv_myo_thickness_total", "lved_mass",
                     "lv_cardiac_output", "af_pwv", "rvedv", "rvesv", "tt2")
```

#### Heatmap 1A.1 (clin variables and CMRI)

```{r, echo=FALSE, fig.width=10, fig.height=8}
plot_focused_heatmap(cmr_clinical$cor_df, imaging_filter = cmr_focus_vars)
```

### Data within 500 days

#### Heatmap 1A.2 (clin variables and CMRI)

```{r, echo=FALSE, fig.width=10, fig.height=8}
cmr_clinical_500 <- run_spearman_heatmap(
  data           = harm_dat_obt2d_baseline_500,
  imaging_vars   = im_vars,
  anchor_vars    = anchor_vars,
  log_vars       = im_vars_log,
  anchor_labels  = anchor_labels,
  imaging_labels = im_labels
)
plot_focused_heatmap(cmr_clinical_500$cor_df, imaging_filter = cmr_focus_vars)
```

## Kidney MRI vs. clinical vars in all baseline OB + T2D

### Data within 30 days

```{r, echo=FALSE, fig.width=10, fig.height=8}
mri_clinical <- run_spearman_heatmap(
  data           = harm_dat_obt2d_baseline,
  imaging_vars   = mri_vars,
  anchor_vars    = anchor_vars,
  anchor_labels  = anchor_labels,
  imaging_labels = mri_labels
)
```

#### Heatmap 1B.1 (clin variables and kidney MRI)

```{r, echo=FALSE, fig.width=10, fig.height=8}
plot_focused_heatmap(mri_clinical$cor_df, img_labels = mri_labels)
```

### Data within 500 days

#### Heatmap 1B.2 (clin variables and kidney MRI)

```{r, echo=FALSE, fig.width=10, fig.height=8}
mri_clinical_500 <- run_spearman_heatmap(
  data           = harm_dat_obt2d_baseline_500,
  imaging_vars   = mri_vars,
  anchor_vars    = anchor_vars,
  anchor_labels  = anchor_labels,
  imaging_labels = mri_labels
)
plot_focused_heatmap(mri_clinical_500$cor_df, img_labels = mri_labels)
```

## Kidney MRI vs. CMR (heatmap)

### Data within 30 days

```{r echo=FALSE, fig.width=10, fig.height=12}
# For the baseline data, which would include not only Improve but also Sweetheart and Ultra T2D, could you associate the kidney MRI parameters with the cardiac and aortic MRI parameters? Maybe in a heat map looking at the correlations.
mri_cmr <- run_spearman_heatmap(
  data           = harm_dat_obt2d_baseline,
  imaging_vars   = im_vars,
  anchor_vars    = mri_vars,
  log_vars       = im_vars_log,
  anchor_labels  = mri_labels,
  imaging_labels = im_labels
)
mri_cmr$coverage_plot
mri_cmr$plot
```

### Data within 500 days

```{r echo=FALSE, fig.width=10, fig.height=12}
mri_cmr_500 <- run_spearman_heatmap(
  data           = harm_dat_obt2d_baseline_500,
  imaging_vars   = im_vars,
  anchor_vars    = mri_vars,
  log_vars       = im_vars_log,
  anchor_labels  = mri_labels,
  imaging_labels = im_labels
)
mri_cmr_500$coverage_plot
mri_cmr_500$plot
```

## Change in CMR vs. clinical vars (IMPROVE 12mo - BL)

### Data within 30 days

```{r echo=FALSE, fig.width=10, fig.height=12}
im_vars_delta_subset <- setdiff(
  im_vars_delta,
  c("lv_myo_thickness_total_log10", "lv_myo_thickness_dias_log10")
)
improve_delta_cmr_clin <- run_spearman_heatmap(
  data           = harm_dat_improve_delta,
  imaging_vars   = paste0(im_vars_delta_subset, "_delta"),
  anchor_vars    = paste0(anchor_vars_delta, "_delta"),
  anchor_labels  = anchor_labels_delta,
  imaging_labels = im_labels_delta
)
improve_delta_cmr_clin$plot
improve_delta_cmr_clin$coverage_plot
```

### Data within 500 days

```{r echo=FALSE, fig.width=10, fig.height=12}
improve_delta_cmr_clin_500 <- run_spearman_heatmap(
  data           = harm_dat_improve_delta_500,
  imaging_vars   = paste0(im_vars_delta_subset, "_delta"),
  anchor_vars    = paste0(anchor_vars_delta, "_delta"),
  anchor_labels  = anchor_labels_delta,
  imaging_labels = im_labels_delta
)
improve_delta_cmr_clin_500$plot
improve_delta_cmr_clin_500$coverage_plot
```

## Change in kidney MRI vs. change in clinical vars (IMPROVE 12mo - BL)

### Data within 30 days

```{r echo=FALSE, fig.width=10, fig.height=12}
improve_delta_mri_clin <- run_spearman_heatmap(
  data           = harm_dat_improve_delta,
  imaging_vars   = paste0(mri_vars_delta, "_delta"),
  anchor_vars    = paste0(anchor_vars_delta, "_delta"),
  anchor_labels  = anchor_labels_delta,
  imaging_labels = mri_labels_delta
)
improve_delta_mri_clin$plot
improve_delta_mri_clin$coverage_plot
```

### Data within 500 days

```{r echo=FALSE, fig.width=10, fig.height=12}
improve_delta_mri_clin_500 <- run_spearman_heatmap(
  data           = harm_dat_improve_delta_500,
  imaging_vars   = paste0(mri_vars_delta, "_delta"),
  anchor_vars    = paste0(anchor_vars_delta, "_delta"),
  anchor_labels  = anchor_labels_delta,
  imaging_labels = mri_labels_delta
)
improve_delta_mri_clin_500$plot
improve_delta_mri_clin_500$coverage_plot
```


## Change in kidney MRI vs. change in CMR (IMPROVE 12mo - BL)

### Data within 30 days

```{r echo=FALSE, fig.width=10, fig.height=12}
improve_delta_mri_cmr <- run_spearman_heatmap(
  data           = harm_dat_improve_delta,
  imaging_vars   = paste0(im_vars_delta_subset, "_delta"),
  anchor_vars    = paste0(mri_vars_delta, "_delta"),
  anchor_labels  = mri_labels_delta,
  imaging_labels = im_labels_delta
)
improve_delta_mri_cmr$plot
improve_delta_mri_cmr$coverage_plot
```

### Data within 500 days

```{r echo=FALSE, fig.width=10, fig.height=12}
improve_delta_mri_cmr_500 <- run_spearman_heatmap(
  data           = harm_dat_improve_delta_500,
  imaging_vars   = paste0(im_vars_delta_subset, "_delta"),
  anchor_vars    = paste0(mri_vars_delta, "_delta"),
  anchor_labels  = mri_labels_delta,
  imaging_labels = im_labels_delta
)
improve_delta_mri_cmr_500$plot
improve_delta_mri_cmr_500$coverage_plot
```

# Linear models or mixed models

## Helper function

```{r mixed-model-helpers, echo=FALSE}
# ── Helper: check whether a fitted lmer model converged properly ─────────────
# Returns TRUE if the model is usable, FALSE if it should be dropped.
# ── Helper: drop rows with non-finite estimates / CIs from results ────────────
drop_nonfinite <- function(results) {
  if (nrow(results) == 0) return(results)
  bad <- with(results, !is.finite(estimate) | !is.finite(conf.low) | !is.finite(conf.high) |
                        !is.finite(std.error) | std.error > 1e6)
  if (any(bad)) {
    dropped <- unique(results$outcome[bad])
    message("  Dropping non-finite results: ", paste(dropped, collapse = ", "))
  }
  results[!bad, ]
}

fit_lm_loop <- function(df,
                        outcomes,
                        fixed_rhs,
                        log_vars   = NULL,
                        var_labels = im_labels) {

  results <- map_dfr(outcomes, function(y) {
    if (!is.null(log_vars) && y %in% log_vars) {
      df[[y]] <- log10(df[[y]])
      y_label <- paste0("log10(", if (y %in% names(var_labels)) var_labels[y] else y, ")")
    } else {
      y_label <- if (y %in% names(var_labels)) var_labels[y] else y
    }

    df_fit <- df %>% filter(!is.na(.data[[y]]))
    if (nrow(df_fit) < 5) return(NULL)

    tryCatch({
      mod <- lm(as.formula(paste0("`", y, "` ~ ", fixed_rhs)), data = df_fit)
      broom::tidy(mod, conf.int = TRUE) %>%
        dplyr::mutate(
          outcome       = y,
          outcome_label = y_label,
          n_obs         = nrow(df_fit),
          n_subjects    = nrow(df_fit)
        )
    }, error = function(e) {
      message(sprintf("  Skipping %s: %s", y, conditionMessage(e)))
      NULL
    })
  })

  results <- drop_nonfinite(results)
  if (nrow(results) == 0) return(results)

  results %>%
    dplyr::mutate(
      stars = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ ""
      )
    )
}

# ── Helper: fit lmer across multiple imaging outcomes ─────────────────────────
fit_lmer_loop <- function(df,
                          outcomes,
                          fixed_rhs,
                          random = "(1 | mrn)",
                          log_vars = NULL,
                          var_labels = im_labels) {

  # Fits outcome ~ fixed_rhs + (1|mrn) for each imaging variable.
  # Returns a tidy data frame of fixed-effect estimates.

  results <- map_dfr(outcomes, function(y) {

    # Log-transform if requested
    if (!is.null(log_vars) && y %in% log_vars) {
      df[[y]] <- log10(df[[y]])
      y_label <- paste0("log10(", if (y %in% names(var_labels)) var_labels[y] else y, ")")
    } else {
      y_label <- if (y %in% names(var_labels)) var_labels[y] else y
    }

    formula_str <- paste0("`", y, "` ~ ", fixed_rhs, " + ", random)

    # Drop rows missing the outcome
    df_fit <- df %>% filter(!is.na(.data[[y]]))

    # Need ≥2 clusters for random effects
    n_clusters <- n_distinct(df_fit$mrn)
    if (n_clusters < 3 || nrow(df_fit) < 5) return(NULL)

    tryCatch({
      mod <- lmer(as.formula(formula_str), data = df_fit, REML = TRUE)

      broom.mixed::tidy(mod, effects = "fixed", conf.int = TRUE) %>%
        dplyr::mutate(
          outcome       = y,
          outcome_label = y_label,
          n_obs         = nrow(df_fit),
          n_subjects    = n_clusters
        )
    }, error = function(e) {
      message(sprintf("  Skipping %s: %s", y, e$message))
      NULL
    })
  })

  results <- drop_nonfinite(results)
  if (nrow(results) == 0) return(results)

  results %>%
    dplyr::mutate(
      stars = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ ""
      )
    )
}

# ── Helper: format results table for display ──────────────────────────────────
format_lm_table <- function(results, caption = "") {
  if (nrow(results) == 0) {
    message("  No converged results to display for: ", caption)
    return(invisible(NULL))
  }
  results %>%
    filter(term != "(Intercept)") %>%
    dplyr::mutate(
      estimate_ci = sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
      p_display   = ifelse(p.value < 0.001,
                           sprintf("%.1e", p.value),
                           sprintf("%.3f", p.value))
    ) %>%
    dplyr::select(outcome_label, term, estimate_ci, p_display, stars, n_obs, n_subjects) %>%
    kable(
      col.names = c("Outcome", "Predictor", "Estimate (95% CI)", "p-value", "", "N obs", "N subj"),
      caption   = caption,
      digits    = 3
    )
}

# ── Helper: fit lmer with z-scored outcomes (standardized estimates) ─────────
# Standardises each outcome to mean 0, SD 1 before fitting so that
#   • categorical predictors → Cohen's d interpretation
#   • continuous predictors  → standardised β (if predictor also z-scored)
# The raw-unit table is produced by fit_lmer_loop; this adds a column showing
# the SD used so reviewers can back-transform.
fit_lmer_loop_std <- function(df,
                              outcomes,
                              fixed_rhs,
                              random     = "(1 | mrn)",
                              log_vars   = NULL,
                              var_labels = im_labels,
                              std_predictors = FALSE) {

  results <- map_dfr(outcomes, function(y) {

    # Log-transform if requested
    if (!is.null(log_vars) && y %in% log_vars) {
      df[[y]] <- log10(df[[y]])
      y_label <- paste0("log10(", if (y %in% names(var_labels)) var_labels[y] else y, ")")
    } else {
      y_label <- if (y %in% names(var_labels)) var_labels[y] else y
    }

    df_fit <- df %>% filter(!is.na(.data[[y]]))
    n_clusters <- n_distinct(df_fit$mrn)
    if (n_clusters < 3 || nrow(df_fit) < 5) return(NULL)

    # Z-score the outcome
    y_sd   <- sd(df_fit[[y]], na.rm = TRUE)
    y_mean <- mean(df_fit[[y]], na.rm = TRUE)
    if (is.na(y_sd) || y_sd == 0) return(NULL)
    df_fit[[y]] <- (df_fit[[y]] - y_mean) / y_sd

    # Optionally z-score continuous predictors in fixed_rhs
    if (std_predictors) {
      pred_terms <- trimws(strsplit(fixed_rhs, "\\+")[[1]])
      for (p in pred_terms) {
        if (is.numeric(df_fit[[p]])) {
          p_sd <- sd(df_fit[[p]], na.rm = TRUE)
          if (!is.na(p_sd) && p_sd > 0) {
            df_fit[[p]] <- (df_fit[[p]] - mean(df_fit[[p]], na.rm = TRUE)) / p_sd
          }
        }
      }
    }

    formula_str <- paste0("`", y, "` ~ ", fixed_rhs, " + ", random)

    tryCatch({
      mod <- lmer(as.formula(formula_str), data = df_fit, REML = TRUE)

      broom.mixed::tidy(mod, effects = "fixed", conf.int = TRUE) %>%
        dplyr::mutate(
          outcome       = y,
          outcome_label = gsub("\n", " ", y_label),
          outcome_sd    = y_sd,
          n_obs         = nrow(df_fit),
          n_subjects    = n_clusters
        )
    }, error = function(e) {
      message(sprintf("  Skipping %s: %s", y, e$message))
      NULL
    })
  })

  results <- drop_nonfinite(results)
  if (nrow(results) == 0) return(results)

  results %>%
    dplyr::mutate(
      stars = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ ""
      )
    )
}

fit_lm_loop_std <- function(df,
                            outcomes,
                            fixed_rhs,
                            log_vars       = NULL,
                            var_labels     = im_labels,
                            std_predictors = FALSE) {

  results <- map_dfr(outcomes, function(y) {
    # Optional log-transform first
    if (!is.null(log_vars) && y %in% log_vars) {
      df[[y]] <- log10(df[[y]])
      y_label <- paste0("log10(", if (y %in% names(var_labels)) var_labels[y] else y, ")")
    } else {
      y_label <- if (y %in% names(var_labels)) var_labels[y] else y
    }

    df_fit <- df %>% filter(!is.na(.data[[y]]))
    if (nrow(df_fit) < 5) return(NULL)

    # z-score the outcome (Cohen's d for categorical predictors;
    # standardized β if predictors are also z-scored)
    y_sd <- sd(df_fit[[y]], na.rm = TRUE)
    if (!is.finite(y_sd) || y_sd == 0) return(NULL)
    df_fit[[y]] <- as.numeric(scale(df_fit[[y]]))

    # Optionally z-score numeric predictors so β is fully standardized
    if (std_predictors) {
      pred_terms <- all.vars(as.formula(paste("~", fixed_rhs)))
      for (p in pred_terms) {
        if (is.numeric(df_fit[[p]])) {
          df_fit[[p]] <- as.numeric(scale(df_fit[[p]]))
        }
      }
    }

    tryCatch({
      mod <- lm(as.formula(paste0("`", y, "` ~ ", fixed_rhs)), data = df_fit)
      broom::tidy(mod, conf.int = TRUE) %>%
        dplyr::mutate(
          outcome       = y,
          outcome_label = gsub("\n", " ", y_label),
          outcome_sd    = y_sd,
          n_obs         = nrow(df_fit),
          n_subjects    = nrow(df_fit)
        )
    }, error = function(e) {
      message(sprintf("  Skipping %s: %s", y, conditionMessage(e)))
      NULL
    })
  })

  results <- drop_nonfinite(results)
  if (nrow(results) == 0) return(results)

  results %>%
    dplyr::mutate(
      stars = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ ""
      )
    )
}

# ── Helper: forest plot for lmer results (raw or standardized) ────────────────
plot_forest <- function(results, term_filter, x_label = "Estimate") {
  if (nrow(results) == 0) { message("  No results for forest plot"); return(invisible(NULL)) }
  results %>%
    filter(term == term_filter) %>%
    dplyr::mutate(
      sig = if_else(p.value < 0.05, "p < 0.05", "ns"),
      outcome_label = fct_reorder(outcome_label, estimate)
    ) %>%
    ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
    geom_vline(xintercept = 0, color = "#edede9") +
    geom_point(size = 3) +
    geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
    labs(y = NULL, x = x_label, color = NULL) +
    theme_minimal() +
    theme(panel.grid = element_blank(),
          legend.position = "none",
          text = element_text(size = 15)) +
    scale_color_manual(values = c("p < 0.05" = "#284b63", "ns" = "#d9d9d9"))
}

# Convenience aliases
plot_forest_raw <- function(results, term_filter, x_label = "Estimate (raw units)")
  plot_forest(results, term_filter, x_label)
plot_forest_std <- function(results_std, term_filter, x_label = "Standardized estimate (SD units)")
  plot_forest(results_std, term_filter, x_label)

# ── Helper: % change forest plot (IMPROVE-style) ─────────────────────────────
plot_forest_pct <- function(emm_results,
                            visit_filter = NULL,
                            visit_col    = NULL,
                            ref_visit    = "Baseline",
                            x_label      = "% Change from baseline") {
  if (is.null(emm_results) || nrow(emm_results) == 0) return(invisible(NULL))

  # Auto-detect the visit column: look for "visit" or any column that isn't

  # a standard EMM output field
  if (is.null(visit_col)) {
    emm_cols <- c("estimate", "SE", "df", "conf.low", "conf.high",
                  "diff", "diff_se", "diff_low", "diff_high",
                  "pct_change", "pct_low", "pct_high", "p.value",
                  "outcome", "outcome_label", "baseline_emm",
                  "n_obs", "n_subjects", "stars")
    visit_col <- setdiff(names(emm_results), emm_cols)
    visit_col <- visit_col[1]
  }

  df <- emm_results %>%
    filter(.data[[visit_col]] != ref_visit)
  if (!is.null(visit_filter)) df <- df %>% filter(.data[[visit_col]] %in% visit_filter)

  # Rename to a standard "visit" column for ggplot aesthetics
  df <- df %>% dplyr::rename(visit = !!visit_col)

  df <- df %>%
    dplyr::mutate(outcome_label = fct_reorder(outcome_label, pct_change,
                                       .fun = function(x) mean(x, na.rm = TRUE)))

  # Build color/shape scales from the actual visit levels present
  visit_levels <- unique(df$visit)
  palette <- c("3 months" = "#e07a5f", "12 months" = "#3d405b",
               "Follow-up" = "#3d405b", "SWEETHEART Follow-up" = "#3d405b")
  # Fall back to a default color for any unlisted level
  colors <- ifelse(visit_levels %in% names(palette), palette[visit_levels], "#3d405b")
  names(colors) <- visit_levels
  shapes <- ifelse(visit_levels %in% c("3 months"), 16L, 17L)
  names(shapes) <- visit_levels

  ggplot(df, aes(x = pct_change, y = outcome_label, color = visit, shape = visit)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3, position = position_dodge(width = 0.5)) +
    geom_errorbarh(aes(xmin = pct_low, xmax = pct_high),
                   height = 0.2, position = position_dodge(width = 0.5)) +
    geom_text(aes(label = stars, x = pct_high + 0.5),
              size = 5, hjust = 0, show.legend = FALSE,
              position = position_dodge(width = 0.5)) +
    scale_color_manual(values = colors, name = "Timepoint") +
    scale_shape_manual(values = shapes, name = "Timepoint") +
    theme_minimal(base_size = 14) +
    theme(legend.position = "top") +
    labs(x = x_label, y = NULL)
}

# ── Helper: run emm for % differences ──────────────────────────────────
run_lmer_emm <- function(data,
                         outcome_vars,
                         log_vars      = character(0),
                         labels        = character(0),
                         id_col        = "mrn",
                         visit_col     = "visit",
                         ref_visit     = "Baseline",
                         min_clusters  = 3,
                         min_obs       = 2) {

  results <- map_dfr(outcome_vars, function(y) {
    df_fit <- data %>% filter(!is.na(.data[[y]]))

    # Label and optional log-transform
    y_label <- if (y %in% names(labels)) labels[y] else y
    if (y %in% log_vars) {
      df_fit[[y]] <- log10(df_fit[[y]])
      y_label     <- paste0("log10(", y_label, ")")
    }

    # Guards
    n_clusters <- n_distinct(df_fit[[id_col]])
    if (n_clusters < min_clusters || nrow(df_fit) < min_obs) return(NULL)
    if (n_distinct(df_fit[[visit_col]]) < 2)                 return(NULL)

    tryCatch({
      mod <- lmer(
        as.formula(paste0("", y, " ~ ", visit_col, " + (1 | ", id_col, ")")),
        data = df_fit, REML = TRUE
      )
      emms    <- emmeans(mod, as.formula(paste0("~", visit_col)))
      emm_df  <- as.data.frame(emms) %>%
        dplyr::rename(estimate = emmean, conf.low = lower.CL, conf.high = upper.CL)

      contrasts   <- emmeans(mod,
                             as.formula(paste0("trt.vs.ctrl ~", visit_col)),
                             ref = ref_visit)
      contr_df    <- as.data.frame(confint(contrasts$contrasts))
      contr_pvals <- as.data.frame(contrasts$contrasts)

      baseline_emm <- emm_df %>%
        filter(.data[[visit_col]] == ref_visit) %>%
        pull(estimate)
      
      # Helper to parse visit label from contrast string
      parse_visit <- function(contrast_vec) {
        non_ref <- setdiff(unique(df_fit[[visit_col]]), ref_visit)
        non_ref <- non_ref[!is.na(non_ref)]
        result  <- rep(NA_character_, length(contrast_vec))
        for (v in non_ref) {
          result <- ifelse(grepl(v, contrast_vec, fixed = TRUE), v, result)
        }
        result
      }
      
      contr_df    <- contr_df    %>% dplyr::mutate(!!visit_col := parse_visit(contrast))
      contr_pvals <- contr_pvals %>% dplyr::mutate(!!visit_col := parse_visit(contrast))

      bind_rows(
        # Baseline row
        emm_df %>%
          filter(.data[[visit_col]] == ref_visit) %>%
          dplyr::mutate(
            diff = 0, diff_se = 0, diff_low = 0, diff_high = 0,
            pct_change = 0, pct_low = 0, pct_high = 0,
            p.value = NA_real_
          ),
        # Non-baseline rows
        emm_df %>%
          filter(.data[[visit_col]] != ref_visit) %>%
          left_join(
            contr_df %>%
              dplyr::select(all_of(visit_col),
                     diff = estimate, diff_se = SE,
                     diff_low = lower.CL, diff_high = upper.CL),
            by = visit_col
          ) %>%
          left_join(
            contr_pvals %>% select(all_of(visit_col), p.value),
            by = visit_col
          ) %>%
          dplyr::mutate(
            pct_change = (diff / baseline_emm) * 100,
            pct_low    = (diff_low  / baseline_emm) * 100,
            pct_high   = (diff_high / baseline_emm) * 100
          )
      ) %>%
        dplyr::mutate(
          outcome       = y,
          outcome_label = y_label,
          baseline_emm  = baseline_emm,
          n_obs         = nrow(df_fit),
          n_subjects    = n_clusters,
          stars = case_when(
            is.na(p.value)  ~ "",
            p.value < 0.001 ~ "***",
            p.value < 0.01  ~ "**",
            p.value < 0.05  ~ "*",
            TRUE            ~ ""
          )
        )

    }, error = function(e) {
      message(sprintf("  Skipping %s: %s", y, conditionMessage(e)))
      print(e)
      NULL
    })
  })
  # Drop outcomes where any estimate / CI / pct_change is non-finite
  if (nrow(results) > 0) {
    bad <- with(results, !is.finite(estimate) | !is.finite(conf.low) | !is.finite(conf.high) |
                          !is.finite(pct_change) | !is.finite(pct_low) | !is.finite(pct_high))
    if (any(bad)) {
      dropped <- unique(results$outcome[bad])
      message("  Dropping non-finite EMM results: ", paste(dropped, collapse = ", "))
      results <- results[!bad, ]
    }
  }
  results
}
```

## Model 1: Group comparisons (Obesity vs T2D)

```{r model-group, echo=FALSE, results='asis'}
# Imaging ~ group + age + sex
# Reference group: Lean Control (alphabetically first, or relevel as needed)
harm_dat_baseline <- harm_dat_baseline %>%
  dplyr::mutate(group = factor(group, levels = c("Lean Control", "Obese Control", "Type 2 Diabetes")))

results_group <- fit_lm_loop(
  df       = subset(harm_dat_baseline, group != "Lean Control"),
  outcomes = im_vars_raw,
  fixed_rhs = "group + age + sex",
  log_vars = im_vars_log
)

format_lm_table(results_group, caption = "Imaging ~ Group + Age + Sex")
```

#### Forest plot (raw estimates)

```{r echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_group, "groupType 2 Diabetes",
                x_label = "Estimate (T2D vs Obese, raw units)")
```

#### Forest plot (standardized: Cohen's d)

```{r echo=FALSE, fig.width=10, fig.height=6}
results_group_std <- fit_lm_loop_std(
  df        = subset(harm_dat_baseline, group != "Lean Control"),
  outcomes  = im_vars_raw,
  fixed_rhs = "group + age + sex",
  log_vars  = im_vars_log
)
plot_forest_std(results_group_std, "groupType 2 Diabetes",
                x_label = "T2D vs Obese (Cohen's d)")
```

## Model 2: Kidney marker associations (Obesity/T2D only)

### log(UACR)

```{r model-kidney, echo=FALSE, results='asis'}
# ── 2a. Continuous kidney markers ─────────────────────────────────────────────
# Imaging ~ log(UACR) + age + sex + (1 | mrn)
results_uacr <- fit_lm_loop(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "log_acr_u + age + sex",
  log_vars  = im_vars_log
)
format_lm_table(results_uacr, caption = "Imaging ~ log(UACR) + Age + Sex")
```

#### Forest plot: log(UACR) (raw estimates)

```{r echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_uacr, "log_acr_u",
                x_label = "Estimate per unit log(UACR) (raw)")
```

#### Forest plot: log(UACR) (standardized β)

```{r echo=FALSE, fig.width=10, fig.height=6}
results_uacr_std <- fit_lm_loop_std(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "log_acr_u + age + sex",
  log_vars  = im_vars_log,
  std_predictors = TRUE
)
plot_forest_std(results_uacr_std, "log_acr_u",
                x_label = "Standardized \u03b2 (per SD log UACR)")
```

### eGFR

```{r echo=FALSE, fig.width=10, fig.height=6}
# Imaging ~ eGFR + age + sex + (1 | mrn)
results_egfr <- fit_lm_loop(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "eGFR_CKD_epi + age + sex",
  log_vars  = im_vars_log
)
format_lm_table(results_egfr, caption = "Imaging ~ eGFR + Age + Sex")
```

#### Forest plot: eGFR (raw estimates)

```{r echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_egfr, "eGFR_CKD_epi",
                x_label = "Estimate per unit eGFR (raw)")
```

#### Forest plot: eGFR (standardized β)

```{r echo=FALSE, fig.width=10, fig.height=6}
results_egfr_std <- fit_lm_loop_std(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "eGFR_CKD_epi + age + sex",
  log_vars  = im_vars_log,
  std_predictors = TRUE
)
plot_forest_std(results_egfr_std, "eGFR_CKD_epi",
                x_label = "Standardized \u03b2 (per SD eGFR)")
```

### DKD UACR>=30 or eGFR<90

```{r echo=FALSE}
# ── 2b. DKD group (binary) ───────────────────────────────────────────────────
# Imaging ~ DKD (UACR≥30 or eGFR<90) + age + sex
results_dkd30 <- fit_lm_loop(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(dkd_group_30)),
  outcomes  = im_vars_raw,
  fixed_rhs = "dkd_group_30 + age + sex",
  log_vars  = im_vars_log
)
format_lm_table(results_dkd30, caption = "Imaging ~ DKD group (UACR\u226530) + Age + Sex")
```

#### Forest plot: DKD (UACR≥30) (raw estimates)

```{r echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_dkd30, "dkd_group_30non_DKD",
                x_label = "Estimate (non-DKD vs DKD, raw units)")
```

#### Forest plot: DKD (UACR≥30) (standardized: Cohen's d)

```{r echo=FALSE, fig.width=10, fig.height=6}
results_dkd30_std <- fit_lm_loop_std(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(dkd_group_30)),
  outcomes  = im_vars_raw,
  fixed_rhs = "dkd_group_30 + age + sex",
  log_vars  = im_vars_log
)
plot_forest_std(results_dkd30_std, "dkd_group_30non_DKD",
                x_label = "non-DKD vs DKD (Cohen's d)")
```

### DKD UACR>=100 or eGFR<90

```{r echo = F}
results_dkd100 <- fit_lm_loop(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(dkd_group_100)),
  outcomes  = im_vars_raw,
  fixed_rhs = "dkd_group_100 + age + sex",
  log_vars  = im_vars_log
)
format_lm_table(results_dkd100, caption = "Imaging ~ DKD group (UACR\u2265100) + Age + Sex")
```

#### Forest plot: DKD (UACR≥100) (raw estimates)

```{r echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_dkd100, "dkd_group_100non_DKD",
                x_label = "Estimate (non-DKD vs DKD, raw units)")
```

#### Forest plot: DKD (UACR≥100) (standardized: Cohen's d)

```{r echo=FALSE, fig.width=10, fig.height=6}
results_dkd100_std <- fit_lm_loop_std(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(dkd_group_100)),
  outcomes  = im_vars_raw,
  fixed_rhs = "dkd_group_100 + age + sex",
  log_vars  = im_vars_log
)
plot_forest_std(results_dkd100_std, "dkd_group_100non_DKD",
                x_label = "non-DKD vs DKD (Cohen's d)")

```

### Albuminuria

```{r echo = F}
# ── 2c. Albuminuria categories ────────────────────────────────────────────────
results_acr300 <- fit_lm_loop(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(acr_u_cat300)),
  outcomes  = im_vars_raw,
  fixed_rhs = "acr_u_cat300 + age + sex",
  log_vars  = im_vars_log
)
format_lm_table(results_acr300, caption = "Imaging ~ Albuminuria (<300 vs \u2265300) + Age + Sex")
```

#### Forest plot: Albuminuria ≥300 (raw estimates)

```{r echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_acr300, "acr_u_cat300\u2265300",
                x_label = "Estimate (UACR \u2265300 vs <300, raw units)")
```

#### Forest plot: Albuminuria ≥300 (standardized: Cohen's d)

```{r echo=FALSE, fig.width=10, fig.height=6}
results_acr300_std <- fit_lm_loop_std(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(acr_u_cat300)),
  outcomes  = im_vars_raw,
  fixed_rhs = "acr_u_cat300 + age + sex",
  log_vars  = im_vars_log
)
plot_forest_std(results_acr300_std, "acr_u_cat300\u2265300",
                x_label = "UACR \u2265300 vs <300 (Cohen's d)")
```

## Model 3: Longitudinal change (co-enrolled ULTRA → SWEETHEART)

```{r model-longitudinal, echo=FALSE, results='asis', message=F}
# Imaging ~ visit_time + (1 | mrn)
results_longitudinal <- fit_lmer_loop(
  df        = harm_dat_longitudinal,
  outcomes  = im_vars_raw,
  fixed_rhs = "imaging_visit_id",
  log_vars  = im_vars_log
)
format_lm_table(results_longitudinal,
                  caption = "Longitudinal change: Imaging ~ Visit")
```

### Forest plot: raw estimates

```{r forest-longitudinal-raw, echo=FALSE, fig.width=10, fig.height=6}
# The term name is the non-reference level of imaging_visit_id
longitudinal_term <- results_longitudinal %>%
  filter(term != "(Intercept)") %>% pull(term) %>% unique()
plot_forest_raw(results_longitudinal, longitudinal_term[1],
                x_label = "Estimate (SWEETHEART \u2212 ULTRA, raw units)")
```

### Forest plot: % change from baseline (ULTRA → SWEETHEART)

```{r forest-longitudinal-pct, echo=FALSE, fig.width=12, fig.height=8}
# Prepare ULTRA→SWEETHEART longitudinal subset for EMM-based % change
harm_dat_ultra_swht_long <- harm_dat_longitudinal %>%
  dplyr::mutate(
    visit = factor(
      if_else(imaging_visit_id == "ultra_swht_baseline", "Baseline", "Follow-up"),
      levels = c("Baseline", "Follow-up")
    )
  )

ultra_swht_emm_results <- run_lmer_emm(
  data         = harm_dat_ultra_swht_long,
  outcome_vars = im_vars_raw,
  log_vars     = im_vars_log,
  labels       = im_labels,
  id_col       = "mrn",
  visit_col    = "visit",
  ref_visit    = "Baseline"
)

plot_forest_pct(ultra_swht_emm_results,
                x_label = "% Change from baseline (ULTRA \u2192 SWEETHEART)")
```

## Model 4: Sex differences in Obesity/T2D (adjusted)

```{r model-sex, echo=FALSE, results='asis'}
results_sex <- fit_lm_loop(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "sex + age",
  log_vars  = im_vars_log
)
format_lm_table(results_sex, caption = "Imaging ~ Sex + Age (Obesity/T2D only)")
```

### Forest plot: Sex (raw estimates)

```{r, echo=FALSE, fig.width=10, fig.height=6}
sex_term <- results_sex %>%
  filter(term != "(Intercept)") %>% pull(term) %>% unique()
plot_forest_raw(results_sex, sex_term[1],
                x_label = "Estimate (Male vs Female, raw units)")
```
#### Forest plot: Sex (standardized: Cohen's d)

```{r echo=FALSE, fig.width=10, fig.height=6}
results_sex_std <- fit_lm_loop_std(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "sex + age",
  log_vars  = im_vars_log
)
plot_forest_std(results_sex_std, "sexMale",
                x_label = "Male vs Female (Cohen's d)")
```

## Model 5: CMRI Longitudinal change (within IMPROVE), treating time as categorical (baseline, 3 mo, 12 mo)

```{r echo=FALSE, results='asis'}
# Imaging ~ visit_time + (1 | mrn)
improve_dat$visit <- factor(improve_dat$visit, levels = c("baseline", "3_months_post_surgery", "12_months_post_surgery"))
results_improve <- fit_lmer_loop(
  df        = improve_dat,
  outcomes  = im_vars_raw,
  fixed_rhs = "visit",
  log_vars  = im_vars_log
)
format_lm_table(results_improve,
                  caption = "Longitudinal change: Imaging ~ Visit")
```

```{r improve-data-prep, echo=FALSE}
# ── Prepare IMPROVE longitudinal subset ───────────────────────────────────────
harm_dat_improve_long <- improve_dat %>%
  dplyr::mutate(
    visit = factor(
      case_when(
        imaging_visit_id == "improve_baseline" ~ "Baseline",
        imaging_visit_id == "improve_swht_baseline" ~ "Baseline",
        imaging_visit_id == "improve_3mo"      ~ "3 months",
        imaging_visit_id == "improve_12mo"     ~ "12 months"
      ),
      levels = c("Baseline", "3 months", "12 months")
    ),
      ln_acr_u = log(acr_u)
  )
# write.csv(harm_dat_improve_long, file.path(root_path, "IMPROVE T2D/Data_Cleaned/improve_longitudinal_041326.csv"), row.names = F, na = "")
```
 
```{r improve-models, echo=FALSE, results='asis'}
# ── Fit lmer and extract EMMs with contrasts vs baseline ──────────────────────
improve_emm_results <- run_lmer_emm(
  data         = harm_dat_improve_long,
  outcome_vars = im_vars_raw,
  log_vars     = im_vars_log,
  labels       = im_labels,
  id_col       = "mrn",
  visit_col    = "visit",
  ref_visit    = "Baseline"
)
```

---

### Forest plot: IMPROVE change from baseline

```{r forest-improve, echo=FALSE}
plot_forest_pct(improve_emm_results)
```

## Model 5b: CMRI Longitudinal change (within IMPROVE), treating time as continuous months

```{r echo=FALSE, results='asis'}
# Imaging ~ visit_time + (1 | mrn)
improve_dat$months_since_baseline <- improve_dat$days_since_baseline/30
results_improve_cont <- fit_lmer_loop(
  df        = improve_dat,
  outcomes  = im_vars_raw,
  fixed_rhs = "months_since_baseline",
  log_vars  = im_vars_log
)
format_lm_table(results_improve_cont,
                  caption = "Longitudinal change (months): Imaging ~ Visit")

results_improve_cont_std <- fit_lmer_loop_std(
  df        = improve_dat,
  outcomes  = im_vars_raw,
  fixed_rhs = "months_since_baseline",
  log_vars  = im_vars_log
)

results_improve_cont_sens <- fit_lmer_loop(
  df        = subset(improve_dat, days_since_baseline < 1460),
  outcomes  = im_vars_raw,
  fixed_rhs = "months_since_baseline",
  log_vars  = im_vars_log
)
format_lm_table(results_improve_cont_sens,
                  caption = "Longitudinal change (months): Imaging ~ Visit (excluding 4.74 yr follow up)")

results_improve_cont_sens_std <- fit_lmer_loop_std(
  df        = subset(improve_dat, days_since_baseline < 1460),
  outcomes  = im_vars_raw,
  fixed_rhs = "months_since_baseline",
  log_vars  = im_vars_log
)
```

### Forest plot: raw estimates (per month)

```{r forest-improve-cont-raw, echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_improve_cont, "months_since_baseline",
                x_label = "Estimate per month (raw units)")
```

### Forest plot: standardized estimates (per month)

```{r forest-improve-cont-std, echo=FALSE, fig.width=10, fig.height=6}
plot_forest_std(results_improve_cont_std, "months_since_baseline",
                x_label = "Standardized estimate per month (SD units)")
```

### Forest plot: raw estimates (per month, excluding 4.74 yr follow up)

```{r forest-improve-cont-sens-raw, echo=FALSE, fig.width=10, fig.height=6}
plot_forest_raw(results_improve_cont_sens, "months_since_baseline",
                x_label = "Estimate per month (raw units, sensitivity)")
```

### Forest plot: standardized estimates (per month, excluding 4.74 yr follow up)

```{r forest-improve-cont-sens-std, echo=FALSE, fig.width=10, fig.height=6}
plot_forest_std(results_improve_cont_sens_std, "months_since_baseline",
                x_label = "Standardized estimate per month (SD units, sensitivity)")
```

```{r}
# spaghetti plot
colors_30 <- c(
  "#7EA8BE", "#E8A87C", "#8FB996", "#B08BBB", "#C47B7B",
  "#7DB8B0", "#D4A96A", "#9B89B4", "#88A96E", "#C4876A",
  "#6E9BAB", "#BFA76A", "#A07898", "#7DAA8A", "#C49B7B",
  "#8598BC", "#D49B8A", "#7AA89E", "#B89B6A", "#9B8AAB",
  "#8DAA7A", "#C4A07B", "#6E8FAB", "#B47A8A", "#8AAF9B",
  "#A89B7A", "#7B9BB4", "#C4907A", "#8A9B7A", "#B48A9B"
)

improve_spaghetti <- lapply(im_vars_raw, function(var) {
  y_label <- if (var %in% names(im_labels)) im_labels[[var]] else var
  improve_dat %>%
    filter(days_since_baseline < 1460) %>%
    ggplot(aes(x = months_since_baseline, y = .data[[var]], group = mrn, color = factor(mrn))) +
    geom_vline(xintercept = 3, linetype = "dashed", color = "grey") +
    geom_vline(xintercept = 12, linetype = "dashed", color = "grey") +
    geom_point() +
    geom_line(alpha = 0.8) +
    theme_minimal() +
    theme(text = element_text(size = 15),
          panel.grid = element_blank(),
          legend.position = "none") +
    labs(x = "Time since baseline (months)", y = y_label) +
    scale_color_manual(values = colors_30)
})

wrap_plots(improve_spaghetti)  # patchwork, or use cowplot::plot_grid(plotlist = plots)
```

## Model 6: Longitudinal change (within IMPROVE)

```{r echo=FALSE, results='asis'}
# log UACR and GFR
# what about eGFR and if we do BSA-adjusted GFR too, or mGFr (unadjusted) but adjusting for weight change
# ── Fit lmer and extract EMMs with contrasts vs baseline ──────────────────────
# log UACR and GFR
# what about eGFR and if we do BSA-adjusted GFR too, or mGFr (unadjusted) but adjusting for weight change
harm_dat_subset$ln_acr_u <- log(harm_dat_subset$acr_u)

acr_gfr_improve <- fit_lmer_loop(
  df         = subset(harm_dat_subset, study == "IMPROVE"),
  outcomes   = c("ln_acr_u", "eGFR_CKD_epi", "gfr_bsa_plasma"),
  fixed_rhs  = "visit",
  var_labels = acr_gfr_labels
)
format_lm_table(acr_gfr_improve,
                  caption = "Longitudinal change: ACR, GFR ~ Visit")

improve_acr_gfr_emm_results <- run_lmer_emm(
  data         = harm_dat_improve_long,
  outcome_vars = acr_gfr_vars,
  log_vars     = im_vars_log,
  labels       = acr_gfr_labels,
  id_col       = "mrn",
  visit_col    = "visit",
  ref_visit    = "Baseline"
)
```

### Forest plot: IMPROVE ACR/GFR change from baseline

```{r forest-improve2, echo=FALSE}
plot_forest_pct(improve_acr_gfr_emm_results)
```

```{r weight-change, echo=FALSE}
# Compute weight change from baseline within each IMPROVE participant
harm_dat_subset <- harm_dat_subset %>%
  group_by(mrn) %>%
  dplyr::mutate(
    weight_bl = weight[imaging_visit_id == "improve_baseline"][1],
    weight_change = weight - weight_bl
  ) %>%
  ungroup()
```

## Model 7: Longitudinal change (within IMPROVE)

```{r echo=FALSE, results='asis'}
# ── Fit lmer and extract EMMs with contrasts vs baseline ──────────────────────
improve_mri_emm_results <- run_lmer_emm(
  data         = harm_dat_improve_long,
  outcome_vars = mri_vars,
  log_vars     = im_vars_log,
  labels       = mri_labels,
  id_col       = "mrn",
  visit_col    = "visit",
  ref_visit    = "Baseline"
)
```

### Forest plot: IMPROVE MRI change from baseline (12 months)

```{r forest-improve3, echo=FALSE}
plot_forest_pct(improve_mri_emm_results, visit_filter = "12 months")
```

# Model 8: Arteriosclerosis y/n
