---
title: "SWEETHEART & ULTRA Imaging"
author: "Ye Ji Choi"
format:
  revealjs:
    embed-resources: true
    scrollable: true
    transition: slide
    controls-layout: bottom-right
    menu: true
    toc: true
    toc-depth: 3
    fontsize: 15
    number-sections: true
---

<!--
Left ventricular (LV), right ventricular (RV) global functions
  ESV: End-systolic volume (mL)
  EDV: End-diastolic volume (mL)
  SV:  Stroke volume (mL)
  CO:  Cardiac output (L/min)
  EF:  Ejection fraction (%)
  HR:  Heart rate (1/min)
  thickness_diast: Global myocardial thickness at diastole (mm)
  thickness_total: Global myocardial thickness across cardiac cycle (mm)
  MyoMass_diast: Global myocardial mass at diastole
  MyoMass_syst:  Global myocardial mass at systole

Left ventricular global peak strain (systolic function)
  GRS: Global radial strain (%)
  GCS: Global circumferential strain (%)
  GLS: Global longitudinal strain (%)

Tissue characteristics
  T1 (ms)
  T2* (ms)

Intracardiac 4D flow — Mitral valve inflow (diastolic function)
  E wave:   Early diastolic peak inflow velocity (cm/s)
  A wave:   Peak inflow velocity during atrial contraction (cm/s)
  E/A ratio

Aortic 4D flow (thoracic aorta)
  PWV:          Pulse wave velocity (m/s)
  MeanWSS:      Mean wall shear stress (Pa)
  MedianWSS:    Median wall shear stress (Pa)
  High5pctWSS:  Median of top 5% wall shear stress (Pa)
-->

# Setup

```{r setup, include=FALSE}
# ── Packages ──────────────────────────────────────────────────────────────────
library(jsonlite)
library(aws.s3)
library(knitr)
library(kableExtra)
library(tidyverse)
library(lmerTest)
library(emmeans)
library(broom)
library(broom.mixed)
library(ggbeeswarm)
library(lemon)
library(quantreg)
library(readxl)
library(lubridate)
library(cowplot)
library(ggpubr)
library(Hmisc)
library(arsenal)
```

```{r paths, include=FALSE}
# ── User-specific paths ──────────────────────────────────────────────────────
user <- Sys.info()[["user"]]

if (user == "choiyej") {
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") {
  root_path <- ""
  git_path  <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") {
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path  <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path  <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys      <- fromJSON("/Users/pylell/Library/CloudStorage/OneDrive-UW/keys.json")
} else {
  stop("Unknown user: please specify root path.")
}
```

```{r s3-setup, include=FALSE}
# ── S3 credentials ────────────────────────────────────────────────────────────
Sys.setenv(
  AWS_ACCESS_KEY_ID     = keys$MY_ACCESS_KEY,
  AWS_SECRET_ACCESS_KEY = keys$MY_SECRET_KEY,
  AWS_DEFAULT_REGION    = "",
  AWS_REGION            = "",
  AWS_S3_ENDPOINT       = "s3.kopah.uw.edu"
)
```

# Variable definitions

```{r variable-defs, echo=FALSE}
# ── Imaging variables ─────────────────────────────────────────────────────────

# Variables requiring log-transformation
im_vars_log <- c("af_pwv", "mv_flow_a", "mv_flow_e",
                  "lv_myo_thickness_total", "lv_myo_thickness_dias")

# Full set of imaging variables
im_vars <- c(
  "lv_hr", "rv_hr",
  "af_high5pctwss", "af_medianwss", "af_meanwss", "af_pwv",
  "mv_flow_ea", "mv_flow_a", "mv_flow_e",
  "tt2", "tt1",
  "gls", "gcs", "grs",
  "rvef", "lvef",
  "rv_cardiac_output", "lv_cardiac_output",
  "rv_stroke_volume", "lv_stroke_volume",
  "rvedv", "lvedv", "rvesv", "lvesv",
  "lves_mass", "lved_mass",
  "lv_myo_thickness_total", "lv_myo_thickness_dias"
)

# Subset for focused analyses (raw units)
im_vars_raw <- c(
  "lvedv", "lvesv", "lv_stroke_volume", "lvef",
  "rvedv", "rvesv", "rv_stroke_volume", "rvef",
  "lved_mass", "lv_myo_thickness_dias",
  "gls", "lv_hr",
  "af_meanwss", "af_pwv",
  "tt1", "tt2"
)

# BSA-indexed versions
im_vars_bsa <- c(
  "lvedv_bsa", "lvesv_bsa", "lv_stroke_volume_bsa", "lvef",
  "rvedv_bsa", "rvesv_bsa", "rv_stroke_volume_bsa", "rvef",
  "lved_mass_bsa", "lv_myo_thickness_dias",
  "gls", "lv_hr",
  "af_meanwss", "af_pwv",
  "tt1", "tt2"
)

# ── Display labels ────────────────────────────────────────────────────────────
im_labels <- c(
  lv_hr                  = "LV Heart rate (1/min)",
  rv_hr                  = "RV Heart rate (1/min)",
  rvedv                  = "RVEDV (mL)",
  lvedv                  = "LVEDV (mL)",
  rvesv                  = "RVESV (mL)",
  lvesv                  = "LVESV (mL)",
  rv_stroke_volume       = "RV stroke\nvolume (mL)",
  lv_stroke_volume       = "LV stroke\nvolume (mL)",
  rvedv_bsa              = "RVEDV BSA (mL/m\u00b2)",
  lvedv_bsa              = "LVEDV BSA (mL/m\u00b2)",
  rvesv_bsa              = "RVESV BSA (mL/m\u00b2)",
  lvesv_bsa              = "LVESV BSA (mL/m\u00b2)",
  rv_stroke_volume_bsa   = "RV stroke\nvolume BSA (mL/m\u00b2)",
  lv_stroke_volume_bsa   = "LV stroke\nvolume BSA (mL/m\u00b2)",
  rv_cardiac_output      = "RV cardiac\noutput",
  lv_cardiac_output      = "LV cardiac\noutput",
  lves_mass              = "LV myocardial\nmass (systole)",
  lved_mass              = "LV mass (g)",
  lves_mass_bsa          = "LV myocardial\nmass BSA (systole)",
  lved_mass_bsa          = "LV mass BSA (g/m\u00b2)",
  lv_myo_thickness_total = "LV thickness\ntotal (mm)",
  lv_myo_thickness_dias  = "LV\nthickness (mm)",
  rvef                   = "RV\nejection fraction (%)",
  lvef                   = "LV\nejection fraction (%)",
  gls                    = "Global\nlongitudinal strain (%)",
  gcs                    = "Global\ncircumferential strain (%)",
  grs                    = "Global\nradial strain (%)",
  af_high5pctwss         = "Aortic wall shear\nstress (top 5%)",
  af_medianwss           = "Aortic wall shear\nstress (median)",
  af_meanwss             = "Mean wall shear\nstress (Pa)",
  af_pwv                 = "Pulse wave\nvelocity (m/s)",
  tt1                    = "T1 (ms)",
  tt2                    = "T2* (ms)",
  tr2                    = "R2* (1/s)",
  mv_flow_e              = "E wave (cm/s)",
  mv_flow_a              = "A wave (cm/s)",
  mv_flow_ea             = "E/A ratio"
)
```

# Data import & cleaning

```{r data-import, include=FALSE}
# ── Import harmonized dataset ─────────────────────────────────────────────────
harm_dat <- read.csv(
  file.path(root_path, "Data Harmonization/Data Clean/harmonized_dataset.csv"),
  na = ""
)

# ── Identify co-enrolled participants ─────────────────────────────────────────
ultra_swht_coenrolled <- harm_dat %>%
  filter(study == "ULTRA" & !is.na(swht_id)) %>%
  pull(swht_id) %>% unique()

ultra_swht_coenrolled_ultra <- harm_dat %>%
  filter(study == "ULTRA" & !is.na(swht_id)) %>%
  pull(ultra_id) %>% unique()

swht_imp_coenrolled <- harm_dat %>%
  filter(study == "SWEETHEART" & !is.na(improve_id)) %>%
  pull(swht_id) %>% unique()

imp_swht_coenrolled <- harm_dat %>%
  filter(study == "IMPROVE" & !is.na(swht_id)) %>%
  pull(improve_id) %>% unique()

harm_dat %>%
  filter(study == "IMPROVE") %>%
  dplyr::select(record_id, visit, acr_u)
```

```{r data-subset, include=FALSE}
# ── Build analysis subset ─────────────────────────────────────────────────────
harm_dat_subset <- harm_dat %>%
  filter(study %in% c("ULTRA", "IMPROVE", "SWEETHEART")) %>%
  mutate(
    # ── Imaging visit identifiers ──
    imaging_visit_id = case_when(
      record_id %in% ultra_swht_coenrolled_ultra & study == "ULTRA" & imaging_visit == 1 ~ "ultra_swht_baseline",
      imaging_visit == 1                                                                  ~ "ultra_baseline",
      imaging_visit == 2                                                                  ~ "ultra_follow_up",
      record_id %in% ultra_swht_coenrolled & study == "SWEETHEART"                        ~ "ultra_swht_follow_up",
      record_id %in% imp_swht_coenrolled & study == "IMPROVE"                             ~ "improve_swht_baseline",
      record_id %in% swht_imp_coenrolled & study == "SWEETHEART"                          ~ "improve_swht_follow_up",
      study == "IMPROVE" & visit == "baseline"                                            ~ "improve_baseline",
      study == "IMPROVE" & visit == "3_months_post_surgery"                               ~ "improve_3mo",
      study == "IMPROVE" & visit == "12_months_post_surgery"                              ~ "improve_12mo",
      study == "SWEETHEART"                                                               ~ "swht_cs"
    ),
    # ── Co-enrollment flag ──
    coenrolled_id = case_when(
      !is.na(swht_id) & !is.na(improve_id) ~ "swht_improve_coenrolled",
      !is.na(swht_id)                       ~ "swht_coenrolled",
      !is.na(ultra_id)                      ~ "ultra_coenrolled",
      !is.na(improve_id)                    ~ "improve_coenrolled"
    ),
    sex       = if_else(study == "ULTRA", "Male", sex),
    cmri_date = if_else(procedure == "cardio_abdominal_mri", date, NA_character_),
    age       = interval(dob, cmri_date) / years(1)
  ) %>%
  summarise(
    across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
    across(where(is.numeric),         ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
    .by = c(record_id, visit)
  ) %>%  # ── Collapse duplicate rows per participant × cMRI date ──
  summarise(
    across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
    across(where(is.numeric),         ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
    .by = c(record_id, cmri_date)
  ) %>%
  filter(study != "ULTRA" | !is.na(imaging_visit_id)) %>%
  # ── Carry forward within co-enrolled participants ──
  # group_by(mrn, coenrolled_id) %>%
  # fill(-any_of(im_vars), .direction = "down") %>%
  # ungroup() %>%
  # ── Derived variables ──
  mutate(
    # Kidney function groupings
    uacr_100 = if_else(acr_u >= 100, "UACR >=100 mg/g", "UACR <100 mg/g"),
    dkd_group_100 = case_when(
      eGFR_CKD_epi < 90 | acr_u >= 100 ~ "DKD",
      eGFR_CKD_epi >= 90 & acr_u < 100 ~ "non_DKD"
    ),
    dkd_group_30 = case_when(
      eGFR_CKD_epi < 90 | acr_u >= 30 ~ "DKD",
      eGFR_CKD_epi >= 90 & acr_u < 30 ~ "non_DKD"
    ),
    uacr_group = factor(
      case_when(
        eGFR_CKD_epi < 90 | acr_u >= 100  ~ "eGFR < 90 and/or UACR >=100 mg/g",
        eGFR_CKD_epi >= 90 & acr_u < 100  ~ "eGFR >= 90 and UACR <100 mg/g"
      ),
      levels = c("eGFR >= 90 and UACR <100 mg/g", "eGFR < 90 and/or UACR >=100 mg/g")
    ),
    egfr_cat = factor(
      case_when(
        eGFR_CKD_epi < 60                        ~ "<60",
        eGFR_CKD_epi >= 60 & eGFR_CKD_epi < 90  ~ "60\u2013<90",
        eGFR_CKD_epi >= 90                       ~ "\u226590"
      ),
      levels = c("<60", "60\u2013<90", "\u226590")
    ),
    acr_u_cat3 = factor(
      case_when(
        acr_u < 30                 ~ "<30",
        acr_u >= 30 & acr_u < 300  ~ "30\u2013<300",
        acr_u >= 300               ~ "\u2265300"
      ),
      levels = c("<30", "30\u2013<300", "\u2265300")
    ),
    acr_u_cat300 = if_else(acr_u >= 300, "\u2265300", "<300"),
    # Tissue characterisation
    tr2       = (1 / tt2) * 1000,
    log_acr_u = log(acr_u),
    # Zero strain → NA
    gls = na_if(gls, 0),
    gcs = na_if(gcs, 0),
    grs = na_if(grs, 0),
    # BSA-indexed and height-indexed volumes/mass
    bsa = sqrt(height * weight / 3600),
    lvesv_bsa              = lvesv / bsa,
    rvesv_bsa              = rvesv / bsa,
    lvedv_bsa              = lvedv / bsa,
    rvedv_bsa              = rvedv / bsa,
    lv_stroke_volume_bsa   = lv_stroke_volume / bsa,
    rv_stroke_volume_bsa   = rv_stroke_volume / bsa,
    lved_mass_bsa          = lved_mass / bsa,
    lves_mass_bsa          = lves_mass / bsa,
    height_m               = height / 100,
    lvesv_height            = lvesv / height_m,
    rvesv_height            = rvesv / height_m,
    lvedv_height            = lvedv / height_m,
    rvedv_height            = rvedv / height_m,
    lv_stroke_volume_height = lv_stroke_volume / height_m,
    rv_stroke_volume_height = rv_stroke_volume / height_m,
    lved_mass_height        = lved_mass / height_m,
    lves_mass_height        = lves_mass / height_m
  )
```

```{r analysis-subsets, include=FALSE}
# ── Analysis subsets ──────────────────────────────────────────────────────────
harm_dat_baseline <- harm_dat_subset %>%
  filter(imaging_visit_id != "ultra_follow_up")

harm_dat_obt2d <- harm_dat_subset %>%
  filter(group != "Lean Control" & imaging_visit_id != "ultra_follow_up")

harm_dat_obt2d_baseline <- harm_dat_obt2d %>%
  filter(!imaging_visit_id %in% c("improve_3mo", "improve_12mo"))

harm_dat_t2d <- harm_dat_subset %>%
  filter(group == "Type 2 Diabetes" & imaging_visit_id != "ultra_follow_up")

# Longitudinal subset: co-enrolled with baseline + follow-up
harm_dat_longitudinal <- harm_dat_subset %>%
  filter(imaging_visit_id %in% c("ultra_swht_baseline", "ultra_swht_follow_up")) %>%
  arrange(mrn, imaging_visit_id) %>%
  mutate(
    visit_time = if_else(
      imaging_visit_id == "ultra_swht_baseline", 0, 1
    )
  )
```

# Descriptive tables

```{r tables, echo=FALSE, results='asis'}
summary(tableby(
  imaging_visit_id ~ group + age + sex + bmi +
    kwt(acr_u, "Nmiss", "medianq1q3") + hba1c +
    dkd_group_30 + dkd_group_100 + bsa,
  data = harm_dat_subset
))

summary(tableby(
  group ~ imaging_visit_id + age + sex + bmi + sbp + dbp +
    kwt(acr_u, "Nmiss", "medianq1q3") + hba1c + eGFR_CKD_epi +
    acr_u_cat3 + dkd_group_30 + dkd_group_100,
  data = harm_dat_baseline
))

summary(tableby(
  sex ~ group + imaging_visit_id + age + height + weight + bmi +
    kwt(acr_u, "Nmiss", "medianq1q3") + hba1c + eGFR_CKD_epi +
    dkd_group_30 + dkd_group_100,
  data = subset(harm_dat_baseline, group != "Lean Control")
))
```

# Box plots

```{r boxplot-function, echo=FALSE}
plot_boxgrid <- function(df,
                         vars = im_vars,
                         log_vars = NULL,
                         group_var,
                         id_var = "mrn",
                         paired = FALSE,
                         var_labels = NULL,
                         ncol = 3,
                         per_page = 9,
                         fill_colors = c("#f08080", "#98c1d9"),
                         save = TRUE,
                         save_dir = file.path(root_path, "ULTRA_SWEETHEART/Results/Figures/Box plots"),
                         file_prefix = "plot_page",
                         file_type = "png",
                         width = 12,
                         height = 12,
                         dpi = 300,
                         show_legend = FALSE,
                         fill_label = NULL,
                         text_size = 18,
                         star_size = 10,
                         p_format = "stars",
                         p_digits = 3,
                         p_size = 5,
                         test_method = "wilcox.test") {

  if (is.numeric(vars))     vars     <- names(df)[vars]
  if (!is.null(log_vars) && is.numeric(log_vars)) log_vars <- names(df)[log_vars]

  # Log-transform where requested
  if (!is.null(log_vars)) {
    for (v in log_vars) df[[paste0(v, "_log10")]] <- log10(df[[v]])
  }

  # Filter to complete pairs for paired tests

  if (paired) {
    complete_ids <- df %>%
      group_by(.data[[id_var]]) %>%
      summarise(n_groups = n_distinct(.data[[group_var]]), .groups = "drop") %>%
      filter(n_groups == length(unique(df[[group_var]]))) %>%
      pull(.data[[id_var]])
    df <- df %>% filter(.data[[id_var]] %in% complete_ids)
    message(sprintf("Using %d complete pairs for paired analysis", length(complete_ids)))
  }

  # Build individual plots
  plots <- map(seq_along(vars), function(i) {
    v <- vars[i]
    if (!is.null(log_vars) && v %in% log_vars) {
      ycol <- paste0(v, "_log10")
      base_label <- if (!is.null(var_labels) && v %in% names(var_labels)) var_labels[[v]] else v
      ylab <- paste0("log10(", base_label, ")")
    } else {
      ycol <- v
      ylab <- if (!is.null(var_labels) && v %in% names(var_labels)) var_labels[[v]] else v
    }

    comps <- combn(unique(df[[group_var]]), 2, simplify = FALSE)

    p <- ggplot(df, aes(x = as.character(.data[[group_var]]),
                        y = .data[[ycol]],
                        fill = as.character(.data[[group_var]]))) +
      geom_boxplot(size = 1)

    # Add statistical comparisons
    stat_args <- list(
      comparisons = comps,
      method      = test_method,
      paired      = paired,
      hide.ns     = !paired && p_format == "stars"
    )
    if (!paired) stat_args$method.args <- list(var.equal = FALSE)

    if (p_format == "stars") {
      stat_args$label <- "p.signif"
      stat_args$size  <- star_size
    } else {
      stat_args$label      <- "p.format"
      stat_args$size       <- p_size
      stat_args$p.accuracy <- 10^(-p_digits)
    }
    p <- p + do.call(stat_compare_means, stat_args)

    p + labs(y = ylab, x = NULL, fill = fill_label) +
      scale_fill_manual(values = fill_colors) +
      scale_y_continuous(expand = expansion(mult = c(0.05, 0.35))) +
      theme(
        panel.background = element_blank(),
        text = element_text(size = text_size),
        axis.text.x = if (show_legend) element_blank() else element_text(angle = 60, hjust = 1),
        legend.position = "none"
      )
  })

  legend <- cowplot::get_legend(plots[[1]] + theme(legend.position = "right"))

  # Save pages
  if (save) dir.create(save_dir, recursive = TRUE, showWarnings = FALSE)

  n_plots <- length(plots)
  pages   <- ceiling(n_plots / per_page)

  for (page in seq_len(pages)) {
    idx <- ((page - 1) * per_page + 1):min(page * per_page, n_plots)
    page_grid <- plot_grid(plotlist = plots[idx], ncol = ncol)
    page_plot <- plot_grid(page_grid, legend, ncol = 3, rel_widths = c(1, 0.1, 0.15))
    print(page_plot)
    if (save) {
      ggsave(file.path(save_dir, sprintf("%s_%02d.%s", file_prefix, page, file_type)),
             page_plot, width = width, height = height, dpi = dpi)
    }
  }
  invisible(plots)
}
```

```{r boxplots, echo=FALSE, fig.width=12, fig.height=12}
# Group comparison (3 groups)
plot_boxgrid(
  df = harm_dat_baseline, group_var = "group",
  log_vars = im_vars_log,
  fill_colors = c("#588157", "#f08080", "#98c1d9"),
  file_prefix = "baseline_group"
)

# Sex comparison in obesity/T2D
plot_boxgrid(
  df = harm_dat_obt2d_baseline, group_var = "sex",
  vars = im_vars_raw, var_labels = im_labels,
  log_vars = im_vars_log, file_prefix = "obt2d_sex",
  ncol = 4, per_page = 16, show_legend = TRUE, fill_label = "Sex",
  width = 18, text_size = 20, star_size = 12,
  p_format = "value", test_method = "wilcox.test"
)

# BSA-indexed sex comparison
plot_boxgrid(
  df = harm_dat_obt2d_baseline, group_var = "sex",
  vars = im_vars_bsa, var_labels = im_labels,
  log_vars = im_vars_log, file_prefix = "obt2d_sex_bsa",
  ncol = 4, per_page = 16, show_legend = TRUE, fill_label = "Sex",
  width = 18, text_size = 20, star_size = 12,
  p_format = "value", test_method = "wilcox.test"
)

# DKD groupings
plot_boxgrid(
  df = subset(harm_dat_obt2d_baseline, !is.na(dkd_group_30)),
  group_var = "dkd_group_30", log_vars = im_vars_log,
  file_prefix = "obt2d_dkd30"
)

plot_boxgrid(
  df = subset(harm_dat_obt2d_baseline, !is.na(dkd_group_100)),
  group_var = "dkd_group_100", log_vars = im_vars_log,
  file_prefix = "obt2d_dkd100"
)

# Albuminuria categories
plot_boxgrid(
  df = subset(harm_dat_obt2d_baseline, !is.na(acr_u_cat300)),
  group_var = "acr_u_cat300", log_vars = im_vars_log,
  var_labels = im_labels, file_prefix = "obt2d_acr300",
  fill_label = "UACR\nCategory"
)

# Longitudinal (paired)
plot_boxgrid(
  df = harm_dat_longitudinal, group_var = "imaging_visit_id",
  log_vars = im_vars_log, var_labels = im_labels,
  file_prefix = "longitudinal", paired = TRUE
)

# arteriosclerosis y/n
# plot_boxgrid(
#   df = subset(harm_dat_obt2d_baseline, !is.na(acr_u_cat300)),
#   group_var = "acr_u_cat300", log_vars = im_vars_log,
#   var_labels = im_labels, file_prefix = "obt2d_acr300",
#   fill_label = "UACR\nCategory"
# )
```

# Correlation heatmaps

```{r correlations, echo=FALSE, fig.width=10, fig.height=8}
# ── Log-transform for correlation ─────────────────────────────────────────────
harm_dat_obt2d_baseline <- harm_dat_obt2d_baseline %>%
  mutate(across(
    all_of(im_vars_log),
    ~ if_else(.x <= 0, NA_real_, log10(.x)),
    .names = "{.col}_log10"
  ))

imaging_for_cor <- c(im_vars, paste0(im_vars_log, "_log10"))
anchor_vars     <- c("log_acr_u", "eGFR_CKD_epi", "gfr_bsa_plasma", "bmi", "hba1c")

anchor_labels <- c(
  bmi = "BMI", eGFR_CKD_epi = "eGFR", gfr_bsa_plasma = "mGFR",
  hba1c = "HbA1c", log_acr_u = "log(UACR)"
)

# Compute Spearman correlations
cor_df <- expand.grid(imaging = imaging_for_cor, anchor = anchor_vars,
                      stringsAsFactors = FALSE)
cor_df$rho  <- NA_real_
cor_df$pval <- NA_real_

for (i in seq_len(nrow(cor_df))) {
  x   <- harm_dat_obt2d_baseline[[cor_df$imaging[i]]]
  y   <- harm_dat_obt2d_baseline[[cor_df$anchor[i]]]
  ok  <- complete.cases(x, y)
  if (sum(ok) >= 5) {
    tmp <- cor.test(x[ok], y[ok], method = "spearman", exact = FALSE)
    cor_df$rho[i]  <- tmp$estimate
    cor_df$pval[i] <- tmp$p.value
  }
}

cor_df$stars <- cut(cor_df$pval,
                    breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
                    labels = c("***", "**", "*", ""))

# Full heatmap
ggplot(cor_df, aes(x = anchor, y = imaging, fill = rho)) +
  geom_tile(color = "white") +
  geom_text(aes(label = stars), color = "black") +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b",
                       midpoint = 0, limits = c(-1, 1)) +
  theme_minimal(base_size = 14) +
  labs(fill = "\u03c1")

# Focused heatmap
anchor_levels <- c("bmi", "eGFR_CKD_epi", "gfr_bsa_plasma", "hba1c", "log_acr_u")

cor_df %>%
  filter(imaging %in% c("lv_myo_thickness_total", "lved_mass",
                         "lv_cardiac_output", "af_pwv", "rvedv", "tt2")) %>%
  mutate(
    imaging_label = im_labels[imaging],
    anchor        = factor(anchor, levels = anchor_levels),
    anchor_label  = factor(anchor_labels[anchor], levels = anchor_labels[anchor_levels])
  ) %>%
  ggplot(aes(x = anchor_label, y = imaging_label, fill = rho)) +
  geom_tile(color = "white") +
  geom_text(aes(label = stars), color = "black", size = 10) +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b",
                       midpoint = 0, limits = c(-1, 1)) +
  theme_minimal(base_size = 18) +
  theme(panel.grid = element_blank()) +
  labs(fill = "\u03c1", x = NULL, y = NULL)
```

# Mixed models

## Helper function

```{r mixed-model-helpers, echo=FALSE}
# ── Helper: fit lmer across multiple imaging outcomes ─────────────────────────
fit_lmer_loop <- function(df,
                          outcomes,
                          fixed_rhs,
                          random = "(1 | mrn)",
                          log_vars = NULL,
                          var_labels = im_labels) {

  # Fits outcome ~ fixed_rhs + (1|mrn) for each imaging variable.
  # Returns a tidy data frame of fixed-effect estimates.

  results <- map_dfr(outcomes, function(y) {

    # Log-transform if requested
    if (!is.null(log_vars) && y %in% log_vars) {
      df[[y]] <- log10(df[[y]])
      y_label <- paste0("log10(", if (y %in% names(var_labels)) var_labels[y] else y, ")")
    } else {
      y_label <- if (y %in% names(var_labels)) var_labels[y] else y
    }

    formula_str <- paste0("`", y, "` ~ ", fixed_rhs, " + ", random)

    # Drop rows missing the outcome
    df_fit <- df %>% filter(!is.na(.data[[y]]))

    # Need ≥2 clusters for random effects
    n_clusters <- n_distinct(df_fit$mrn)
    if (n_clusters < 3 || nrow(df_fit) < 5) return(NULL)

    tryCatch({
      mod <- lmer(as.formula(formula_str), data = df_fit, REML = TRUE)

      broom.mixed::tidy(mod, effects = "fixed", conf.int = TRUE) %>%
        mutate(
          outcome       = y,
          outcome_label = y_label,
          n_obs         = nrow(df_fit),
          n_subjects    = n_clusters
        )
    }, error = function(e) {
      message(sprintf("  Skipping %s: %s", y, e$message))
      NULL
    })
  })

  results %>%
    mutate(
      stars = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ ""
      )
    )
}

# ── Helper: format results table for display ──────────────────────────────────
format_lmer_table <- function(results, caption = "") {
  results %>%
    filter(term != "(Intercept)") %>%
    mutate(
      estimate_ci = sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
      p_display   = ifelse(p.value < 0.001,
                           sprintf("%.1e", p.value),
                           sprintf("%.3f", p.value))
    ) %>%
    select(outcome_label, term, estimate_ci, p_display, stars, n_obs, n_subjects) %>%
    kable(
      col.names = c("Outcome", "Predictor", "Estimate (95% CI)", "p-value", "", "N obs", "N subj"),
      caption   = caption,
      digits    = 3
    ) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
}
```

## Model 1: Group comparisons (Obesity vs T2D)

```{r model-group, echo=FALSE, results='asis'}
# Imaging ~ group + age + sex + (1 | mrn)
# Reference group: Lean Control (alphabetically first, or relevel as needed)
harm_dat_baseline <- harm_dat_baseline %>%
  mutate(group = factor(group, levels = c("Lean Control", "Obese Control", "Type 2 Diabetes")))

results_group <- fit_lmer_loop(
  df       = subset(harm_dat_baseline, group != "Lean Control"),
  outcomes = im_vars_raw,
  fixed_rhs = "group + age + sex",
  log_vars = im_vars_log
)

format_lmer_table(results_group, caption = "Imaging ~ Group + Age + Sex (random intercept by participant)")
```

### Forest plot

```{r echo=FALSE, fig.width=10, fig.height=6}
results_group %>%
  filter(term == "groupType 2 Diabetes") %>%
  mutate(sig = case_when(p.value < 0.05 ~ "p < 0.05", T ~ "ns")) %>%
  ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
  geom_vline(xintercept = 0, color = "#edede9") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
  labs(y = NULL, x = "Estimate") +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 15)) +
  scale_color_manual(values = c("p < 0.05" = "#284b63",
                                "ns" = "#d9d9d9"))
```

## Model 2: Kidney marker associations (Obesity/T2D only)

```{r model-kidney, echo=FALSE, results='asis'}
# ── 2a. Continuous kidney markers ─────────────────────────────────────────────
# Imaging ~ log(UACR) + age + sex + (1 | mrn)
results_uacr <- fit_lmer_loop(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "log_acr_u + age + sex",
  log_vars  = im_vars_log
)
format_lmer_table(results_uacr, caption = "Imaging ~ log(UACR) + Age + Sex")
### Forest plot
results_uacr %>%
  filter(term == "log_acr_u") %>%
  mutate(sig = case_when(p.value < 0.05 ~ "p < 0.05", T ~ "ns")) %>%
  ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
  geom_vline(xintercept = 0, color = "#edede9") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
  labs(y = NULL, x = "Estimate") +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 15)) +
  scale_color_manual(values = c("p < 0.05" = "#284b63",
                                "ns" = "#d9d9d9"))

# Imaging ~ eGFR + age + sex + (1 | mrn)
results_egfr <- fit_lmer_loop(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "eGFR_CKD_epi + age + sex",
  log_vars  = im_vars_log
)
format_lmer_table(results_egfr, caption = "Imaging ~ eGFR + Age + Sex")
### Forest plot
results_egfr %>%
  filter(term == "eGFR_CKD_epi") %>%
  mutate(sig = case_when(p.value < 0.05 ~ "p < 0.05", T ~ "ns")) %>%
  ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
  geom_vline(xintercept = 0, color = "#edede9") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
  labs(y = NULL, x = "Estimate") +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 15)) +
  scale_color_manual(values = c("p < 0.05" = "#284b63",
                                "ns" = "#d9d9d9"))

# ── 2b. DKD group (binary) ───────────────────────────────────────────────────
# Imaging ~ DKD (UACR≥30 or eGFR<90) + age + sex + (1 | mrn)
results_dkd30 <- fit_lmer_loop(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(dkd_group_30)),
  outcomes  = im_vars_raw,
  fixed_rhs = "dkd_group_30 + age + sex",
  log_vars  = im_vars_log
)
format_lmer_table(results_dkd30, caption = "Imaging ~ DKD group (UACR\u226530) + Age + Sex")
### Forest plot
results_dkd30 %>%
  filter(term == "dkd_group_30non_DKD") %>%
  mutate(sig = case_when(p.value < 0.05 ~ "p < 0.05", T ~ "ns")) %>%
  ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
  geom_vline(xintercept = 0, color = "#edede9") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
  labs(y = NULL, x = "Estimate") +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 15)) +
  scale_color_manual(values = c("p < 0.05" = "#284b63",
                                "ns" = "#d9d9d9"))


results_dkd100 <- fit_lmer_loop(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(dkd_group_100)),
  outcomes  = im_vars_raw,
  fixed_rhs = "dkd_group_100 + age + sex",
  log_vars  = im_vars_log
)
format_lmer_table(results_dkd100, caption = "Imaging ~ DKD group (UACR\u2265100) + Age + Sex")

### Forest plot
results_dkd100 %>%
  filter(term == "dkd_group_100non_DKD") %>%
  mutate(sig = case_when(p.value < 0.05 ~ "p < 0.05", T ~ "ns")) %>%
  ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
  geom_vline(xintercept = 0, color = "#edede9") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
  labs(y = NULL, x = "Estimate") +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 15)) +
  scale_color_manual(values = c("p < 0.05" = "#284b63",
                                "ns" = "#d9d9d9"))

# ── 2c. Albuminuria categories ────────────────────────────────────────────────
results_acr300 <- fit_lmer_loop(
  df        = harm_dat_obt2d_baseline %>% filter(!is.na(acr_u_cat300)),
  outcomes  = im_vars_raw,
  fixed_rhs = "acr_u_cat300 + age + sex",
  log_vars  = im_vars_log
)
format_lmer_table(results_acr300, caption = "Imaging ~ Albuminuria (<300 vs \u2265300) + Age + Sex")
### Forest plot
results_acr300 %>%
  filter(term == "acr_u_cat300≥300") %>%
  mutate(sig = case_when(p.value < 0.05 ~ "p < 0.05", T ~ "ns")) %>%
  ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
  geom_vline(xintercept = 0, color = "#edede9") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
  labs(y = NULL, x = "Estimate") +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 15)) +
  scale_color_manual(values = c("p < 0.05" = "#284b63",
                                "ns" = "#d9d9d9"))
```

## Model 3: Longitudinal change (co-enrolled ULTRA → SWEETHEART)

```{r model-longitudinal, echo=FALSE, results='asis'}
# Imaging ~ visit_time + (1 | mrn)
results_longitudinal <- fit_lmer_loop(
  df        = harm_dat_longitudinal,
  outcomes  = im_vars_raw,
  fixed_rhs = "imaging_visit_id",
  log_vars  = im_vars_log
)
format_lmer_table(results_longitudinal,
                  caption = "Longitudinal change: Imaging ~ Visit")
```

### Forest plot of longitudinal effects

```{r forest-longitudinal, echo=FALSE, fig.width=10, fig.height=6}
results_longitudinal %>%
  filter(term != "(Intercept)") %>%
  mutate(sig = case_when(p.value < 0.05 ~ "p < 0.05", T ~ "ns")) %>%
  ggplot(aes(x = estimate, y = outcome_label, color = sig)) +
  geom_vline(xintercept = 0, color = "#edede9") +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, linewidth = 0.4) +
  labs(y = NULL, x = "Estimate") +
  theme_minimal() + 
  theme(panel.grid = element_blank(),
        legend.position = "none",
        text = element_text(size = 15)) +
  scale_color_manual(values = c("p < 0.05" = "#284b63",
                                "ns" = "#d9d9d9"))
```

## Model 4: Sex differences in Obesity/T2D (adjusted)

```{r model-sex, echo=FALSE, results='asis'}
results_sex <- fit_lmer_loop(
  df        = harm_dat_obt2d_baseline,
  outcomes  = im_vars_raw,
  fixed_rhs = "sex + age",
  log_vars  = im_vars_log
)
format_lmer_table(results_sex, caption = "Imaging ~ Sex + Age (Obesity/T2D only)")
```

## Model 5: Longitudinal change (within IMPROVE)

```{r echo=FALSE, results='asis'}
# Imaging ~ visit_time + (1 | mrn)
harm_dat_subset$visit <- factor(harm_dat_subset$visit, levels = c("baseline", "3_months_post_surgery", "12_months_post_surgery"))
results_improve <- fit_lmer_loop(
  df        = subset(harm_dat_subset, study == "IMPROVE"),
  outcomes  = im_vars_raw,
  fixed_rhs = "visit",
  log_vars  = im_vars_log
)
format_lmer_table(results_improve,
                  caption = "Longitudinal change: Imaging ~ Visit")

# log UACR and GFR
# what about eGFR and if we do BSA-adjusted GFR too, or mGFr (unadjusted) but adjusting for weight change
harm_dat_subset$ln_acr_u <- log(harm_dat_subset$acr_u)

acr_gfr_improve <- fit_lmer_loop(
  df        = subset(harm_dat_subset, study == "IMPROVE"),
  outcomes  = c("ln_acr_u", "eGFR_CKD_epi", "gfr_bsa_plasma"),
  fixed_rhs = "visit"
)
```

```{r improve-data-prep, echo=FALSE}
# ── Prepare IMPROVE longitudinal subset ───────────────────────────────────────
harm_dat_improve_long <- harm_dat_subset %>%
  filter(study == "IMPROVE",
         imaging_visit_id %in% c("improve_baseline", "improve_3mo", "improve_12mo")) %>%
  mutate(
    visit = factor(
      case_when(
        imaging_visit_id == "improve_baseline" ~ "Baseline",
        imaging_visit_id == "improve_3mo"      ~ "3 months",
        imaging_visit_id == "improve_12mo"     ~ "12 months"
      ),
      levels = c("Baseline", "3 months", "12 months")
    )
  )
```
 
```{r improve-models, echo=FALSE, results='asis'}
# ── Fit lmer and extract EMMs with contrasts vs baseline ──────────────────────
improve_emm_results <- map_dfr(im_vars_raw, function(y) {
 
  df_fit <- harm_dat_improve_long %>% filter(!is.na(.data[[y]]))
 
  # Log-transform if needed
  y_label <- if (y %in% names(im_labels)) im_labels[y] else y
  if (y %in% im_vars_log) {
    df_fit[[y]] <- log10(df_fit[[y]])
    y_label <- paste0("log10(", y_label, ")")
  }
 
  n_clusters <- n_distinct(df_fit$mrn)
  if (n_clusters < 3 || nrow(df_fit) < 5) return(NULL)
 
  # Need at least 2 visit levels present
  if (n_distinct(df_fit$visit) < 2) return(NULL)
 
  tryCatch({
    mod <- lmer(as.formula(paste0("`", y, "` ~ visit + (1 | mrn)")),
                data = df_fit, REML = TRUE)
 
    # EMMs at each visit level
    emms <- emmeans(mod, ~ visit)
    emm_df <- as.data.frame(emms) %>%
      rename(estimate = emmean, conf.low = lower.CL, conf.high = upper.CL)
 
    # Contrasts: 3 months - Baseline, 12 months - Baseline
    # Use confint() to get t-based CIs consistent with the p-values
    contrasts <- emmeans(mod, trt.vs.ctrl ~ visit, ref = "Baseline")
    contr_df  <- as.data.frame(confint(contrasts$contrasts))
 
    # Get baseline EMM for percent-change calculation
    baseline_emm <- emm_df %>% filter(visit == "Baseline") %>% pull(estimate)
 
    # Combine: EMMs + raw differences + percent change from baseline
    bind_rows(
      # Baseline row (0% change by definition)
      emm_df %>%
        filter(visit == "Baseline") %>%
        mutate(
          diff = 0, diff_se = 0, diff_low = 0, diff_high = 0,
          pct_change = 0, pct_low = 0, pct_high = 0,
          p.value = NA_real_
        ),
      # 3-month and 12-month rows
      emm_df %>%
        filter(visit != "Baseline") %>%
        left_join(
          contr_df %>%
            mutate(
              visit = case_when(
                grepl("3 months", contrast)  ~ "3 months",
                grepl("12 months", contrast) ~ "12 months"
              )
            ) %>%
            select(visit, diff = estimate, diff_se = SE,
                   diff_low = lower.CL, diff_high = upper.CL),
          by = "visit"
        ) %>%
        # Recover p-values from the contrasts summary (not confint)
        left_join(
          as.data.frame(contrasts$contrasts) %>%
            mutate(
              visit = case_when(
                grepl("3 months", contrast)  ~ "3 months",
                grepl("12 months", contrast) ~ "12 months"
              )
            ) %>%
            select(visit, p.value),
          by = "visit"
        ) %>%
        mutate(
          pct_change = (diff / baseline_emm) * 100,
          pct_low    = (diff_low / baseline_emm) * 100,
          pct_high   = (diff_high / baseline_emm) * 100
        )
    ) %>%
      mutate(
        outcome       = y,
        outcome_label = y_label,
        baseline_emm  = baseline_emm,
        n_obs         = nrow(df_fit),
        n_subjects    = n_clusters,
        stars = case_when(
          is.na(p.value)  ~ "",
          p.value < 0.001 ~ "***",
          p.value < 0.01  ~ "**",
          p.value < 0.05  ~ "*",
          TRUE            ~ ""
        )
      )
  }, error = function(e) {
    message(sprintf("  Skipping %s: %s", y, e$message))
    NULL
  })
})
```
 
### Forest plot: IMPROVE change from baseline
 
```{r forest-improve, echo=FALSE, fig.width=12, fig.height=8}
if (nrow(improve_emm_results) > 0) {
  improve_forest_df <- improve_emm_results %>%
    filter(visit != "Baseline") %>%
    mutate(outcome_label = fct_reorder(outcome_label, pct_change,
                                       .fun = function(x) mean(x, na.rm = TRUE)))
 
  ggplot(improve_forest_df,
         aes(x = pct_change, y = outcome_label, color = visit, shape = visit)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3, position = position_dodge(width = 0.5)) +
    geom_errorbarh(
      aes(xmin = pct_low, xmax = pct_high),
      height = 0.2,
      position = position_dodge(width = 0.5)
    ) +
    geom_text(
      aes(label = stars,
          x = pct_high + 0.5),
      size = 5, hjust = 0, show.legend = FALSE,
      position = position_dodge(width = 0.5)
    ) +
    scale_color_manual(
      values = c("3 months" = "#e07a5f", "12 months" = "#3d405b"),
      name = "Timepoint"
    ) +
    scale_shape_manual(
      values = c("3 months" = 16, "12 months" = 17),
      name = "Timepoint"
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "top") +
    labs(
      x = "% Change from baseline",
      y = NULL
    )
}
```

## Model 6: Longitudinal change (within IMPROVE)

```{r echo=FALSE, results='asis'}
# log UACR and GFR
# what about eGFR and if we do BSA-adjusted GFR too, or mGFr (unadjusted) but adjusting for weight change
harm_dat_subset$ln_acr_u <- log(harm_dat_subset$acr_u)

acr_gfr_vars <- c("ln_acr_u", "eGFR_CKD_epi", "gfr_raw_plasma", "gfr_bsa_plasma", "ff")
```

```{r echo=FALSE, results='asis'}
# ── Fit lmer and extract EMMs with contrasts vs baseline ──────────────────────
improve_acr_gfr_emm_results <- map_dfr(acr_gfr_vars, function(y) {
 
  df_fit <- harm_dat_improve_long %>% filter(!is.na(.data[[y]]))
 
  # Log-transform if needed
  y_label <- if (y %in% names(im_labels)) im_labels[y] else y
  if (y %in% im_vars_log) {
    df_fit[[y]] <- log10(df_fit[[y]])
    y_label <- paste0("log10(", y_label, ")")
  }
 
  n_clusters <- n_distinct(df_fit$mrn)
  if (n_clusters < 3 || nrow(df_fit) < 5) return(NULL)
 
  # Need at least 2 visit levels present
  if (n_distinct(df_fit$visit) < 2) return(NULL)
 
  tryCatch({
    mod <- lmer(as.formula(paste0("`", y, "` ~ visit + (1 | mrn)")),
                data = df_fit, REML = TRUE)
 
    # EMMs at each visit level
    emms <- emmeans(mod, ~ visit)
    emm_df <- as.data.frame(emms) %>%
      rename(estimate = emmean, conf.low = lower.CL, conf.high = upper.CL)
 
    # Contrasts: 3 months - Baseline, 12 months - Baseline
    # Use confint() to get t-based CIs consistent with the p-values
    contrasts <- emmeans(mod, trt.vs.ctrl ~ visit, ref = "Baseline")
    contr_df  <- as.data.frame(confint(contrasts$contrasts))
 
    # Get baseline EMM for percent-change calculation
    baseline_emm <- emm_df %>% filter(visit == "Baseline") %>% pull(estimate)
 
    # Combine: EMMs + raw differences + percent change from baseline
    bind_rows(
      # Baseline row (0% change by definition)
      emm_df %>%
        filter(visit == "Baseline") %>%
        mutate(
          diff = 0, diff_se = 0, diff_low = 0, diff_high = 0,
          pct_change = 0, pct_low = 0, pct_high = 0,
          p.value = NA_real_
        ),
      # 3-month and 12-month rows
      emm_df %>%
        filter(visit != "Baseline") %>%
        left_join(
          contr_df %>%
            mutate(
              visit = case_when(
                grepl("3 months", contrast)  ~ "3 months",
                grepl("12 months", contrast) ~ "12 months"
              )
            ) %>%
            select(visit, diff = estimate, diff_se = SE,
                   diff_low = lower.CL, diff_high = upper.CL),
          by = "visit"
        ) %>%
        # Recover p-values from the contrasts summary (not confint)
        left_join(
          as.data.frame(contrasts$contrasts) %>%
            mutate(
              visit = case_when(
                grepl("3 months", contrast)  ~ "3 months",
                grepl("12 months", contrast) ~ "12 months"
              )
            ) %>%
            select(visit, p.value),
          by = "visit"
        ) %>%
        mutate(
          pct_change = (diff / baseline_emm) * 100,
          pct_low    = (diff_low / baseline_emm) * 100,
          pct_high   = (diff_high / baseline_emm) * 100
        )
    ) %>%
      mutate(
        outcome       = y,
        outcome_label = y_label,
        baseline_emm  = baseline_emm,
        n_obs         = nrow(df_fit),
        n_subjects    = n_clusters,
        stars = case_when(
          is.na(p.value)  ~ "",
          p.value < 0.001 ~ "***",
          p.value < 0.01  ~ "**",
          p.value < 0.05  ~ "*",
          TRUE            ~ ""
        )
      )
  }, error = function(e) {
    message(sprintf("  Skipping %s: %s", y, e$message))
    NULL
  })
})
```
 
### Forest plot: IMPROVE change from baseline
 
```{r forest-improve2, echo=FALSE, fig.width=12, fig.height=8}
if (nrow(improve_acr_gfr_emm_results) > 0) {
  improve_forest_df <- improve_acr_gfr_emm_results %>%
    filter(visit != "Baseline") %>%
    mutate(outcome_label = fct_reorder(outcome_label, pct_change,
                                       .fun = function(x) mean(x, na.rm = TRUE)))
 
  ggplot(improve_forest_df,
         aes(x = pct_change, y = outcome_label, color = visit, shape = visit)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3, position = position_dodge(width = 0.5)) +
    geom_errorbarh(
      aes(xmin = pct_low, xmax = pct_high),
      height = 0.2,
      position = position_dodge(width = 0.5)
    ) +
    geom_text(
      aes(label = stars,
          x = pct_high + 0.5),
      size = 5, hjust = 0, show.legend = FALSE,
      position = position_dodge(width = 0.5)
    ) +
    scale_color_manual(
      values = c("3 months" = "#e07a5f", "12 months" = "#3d405b"),
      name = "Timepoint"
    ) +
    scale_shape_manual(
      values = c("3 months" = 16, "12 months" = 17),
      name = "Timepoint"
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "top") +
    labs(
      x = "% Change from baseline",
      y = NULL
    )
}

# Compute weight change from baseline within each IMPROVE participant
harm_dat_subset <- harm_dat_subset %>%
  group_by(mrn) %>%
  mutate(
    weight_bl = weight[imaging_visit_id == "improve_baseline"][1],
    weight_change = weight - weight_bl
  ) %>%
  ungroup()

summary(lmer(gfr_raw_plasma ~ visit + weight_change + (1 | mrn),
             data = subset(harm_dat_subset, study == "IMPROVE")))

subset(harm_dat_subset, study == "IMPROVE")$erpf_raw_plasma
subset(harm_dat, study == "IMPROVE" & !is.na(erpf_raw_plasma))$visit
```


## Model 7: Longitudinal change (within IMPROVE)

```{r echo=FALSE, results='asis'}
# FSOC, ASL, ADC, TKV, R2* for the 12 month
mri_vars <- c("avg_c_fsoc", "avg_m_fsoc", 
              "avg_c_r2", "avg_m_r2",
              "avg_c_adc", 
              "avg_pcascl", "total_kidney_volume_ml",
              "glomerular_pressure")
```

```{r echo=FALSE, results='asis'}
# ── Fit lmer and extract EMMs with contrasts vs baseline ──────────────────────
improve_mri_emm_results <- map_dfr(mri_vars, function(y) {
 
  df_fit <- harm_dat_improve_long %>% filter(!is.na(.data[[y]]))
 
  # Log-transform if needed
  y_label <- if (y %in% names(im_labels)) im_labels[y] else y
  if (y %in% im_vars_log) {
    df_fit[[y]] <- log10(df_fit[[y]])
    y_label <- paste0("log10(", y_label, ")")
  }
 
  n_clusters <- n_distinct(df_fit$mrn)
  if (n_clusters < 3 || nrow(df_fit) < 5) return(NULL)
 
  # Need at least 2 visit levels present
  if (n_distinct(df_fit$visit) < 2) return(NULL)
 
  tryCatch({
    mod <- lmer(as.formula(paste0("`", y, "` ~ visit + (1 | mrn)")),
                data = df_fit, REML = TRUE)
 
    # EMMs at each visit level
    emms <- emmeans(mod, ~ visit)
    emm_df <- as.data.frame(emms) %>%
      rename(estimate = emmean, conf.low = lower.CL, conf.high = upper.CL)
 
    # Contrasts: 3 months - Baseline, 12 months - Baseline
    # Use confint() to get t-based CIs consistent with the p-values
    contrasts <- emmeans(mod, trt.vs.ctrl ~ visit, ref = "Baseline")
    contr_df  <- as.data.frame(confint(contrasts$contrasts))
 
    # Get baseline EMM for percent-change calculation
    baseline_emm <- emm_df %>% filter(visit == "Baseline") %>% pull(estimate)
 
    # Combine: EMMs + raw differences + percent change from baseline
    bind_rows(
      # Baseline row (0% change by definition)
      emm_df %>%
        filter(visit == "Baseline") %>%
        mutate(
          diff = 0, diff_se = 0, diff_low = 0, diff_high = 0,
          pct_change = 0, pct_low = 0, pct_high = 0,
          p.value = NA_real_
        ),
      # 3-month and 12-month rows
      emm_df %>%
        filter(visit != "Baseline") %>%
        left_join(
          contr_df %>%
            mutate(
              visit = case_when(
                grepl("3 months", contrast)  ~ "3 months",
                grepl("12 months", contrast) ~ "12 months"
              )
            ) %>%
            select(visit, diff = estimate, diff_se = SE,
                   diff_low = lower.CL, diff_high = upper.CL),
          by = "visit"
        ) %>%
        # Recover p-values from the contrasts summary (not confint)
        left_join(
          as.data.frame(contrasts$contrasts) %>%
            mutate(
              visit = case_when(
                grepl("3 months", contrast)  ~ "3 months",
                grepl("12 months", contrast) ~ "12 months"
              )
            ) %>%
            select(visit, p.value),
          by = "visit"
        ) %>%
        mutate(
          pct_change = (diff / baseline_emm) * 100,
          pct_low    = (diff_low / baseline_emm) * 100,
          pct_high   = (diff_high / baseline_emm) * 100
        )
    ) %>%
      mutate(
        outcome       = y,
        outcome_label = y_label,
        baseline_emm  = baseline_emm,
        n_obs         = nrow(df_fit),
        n_subjects    = n_clusters,
        stars = case_when(
          is.na(p.value)  ~ "",
          p.value < 0.001 ~ "***",
          p.value < 0.01  ~ "**",
          p.value < 0.05  ~ "*",
          TRUE            ~ ""
        )
      )
  }, error = function(e) {
    message(sprintf("  Skipping %s: %s", y, e$message))
    NULL
  })
})
```
 
### Forest plot: IMPROVE change from baseline
 
```{r forest-improve3, echo=FALSE, fig.width=12, fig.height=8}
if (nrow(improve_mri_emm_results) > 0) {
  improve_forest_df <- improve_mri_emm_results %>%
    filter(visit == "12 months") %>%
    mutate(outcome_label = fct_reorder(outcome_label, pct_change,
                                       .fun = function(x) mean(x, na.rm = TRUE)))
 
  ggplot(improve_forest_df,
         aes(x = pct_change, y = outcome_label, color = visit, shape = visit)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3, position = position_dodge(width = 0.5)) +
    geom_errorbarh(
      aes(xmin = pct_low, xmax = pct_high),
      height = 0.2,
      position = position_dodge(width = 0.5)
    ) +
    geom_text(
      aes(label = stars,
          x = pct_high + 0.5),
      size = 5, hjust = 0, show.legend = FALSE,
      position = position_dodge(width = 0.5)
    ) +
    scale_color_manual(
      values = c("3 months" = "#e07a5f", "12 months" = "#3d405b"),
      name = "Timepoint"
    ) +
    scale_shape_manual(
      values = c("3 months" = 16, "12 months" = 17),
      name = "Timepoint"
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "top") +
    labs(
      x = "% Change from baseline",
      y = NULL
    )
}

# Compute weight change from baseline within each IMPROVE participant
harm_dat_subset <- harm_dat_subset %>%
  group_by(mrn) %>%
  mutate(
    weight_bl = weight[imaging_visit_id == "improve_baseline"][1],
    weight_change = weight - weight_bl
  ) %>%
  ungroup()

summary(lmer(gfr_raw_plasma ~ visit + weight_change + (1 | mrn),
             data = subset(harm_dat_subset, study == "IMPROVE")))

subset(harm_dat_subset, study == "IMPROVE")$erpf_raw_plasma
subset(harm_dat, study == "IMPROVE" & !is.na(erpf_raw_plasma))$visit
```

# Model 8: Arteriosclerosis y/n