---
title: "TODAY Metabolomics Combined Analysis"
author: "Laura Pyle and Kristen Miller"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    toc: true
    toc_depth: '3'
  html_document:
    number_sections: true
    theme: readable
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(stringr)
library(table1)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

user <- Sys.info()[["user"]]
if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "kristenmiller") {
  root_path <- "/Users/kristenmiller/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive/TODAY subaward"
  git_path <- "/Users/kristenmiller/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else {
  stop("Unknown user: please specify root path for this user.")
}

knitr::opts_knit$set(root.dir = root_path)
```

```{r, include=FALSE}
# load plasma data
load("./Metabolomic data/plasma.Rdata")
# load urine data
load("./Metabolomic data/urine.Rdata")
# load ceramides
load("./Metabolomic data/plasma_ceramides.Rdata")
# load mta data'
load("./Metabolomic data/mta.Rdata")
# load comorbidity data
load("./Clinical data/comorb.Rdata")
# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# load CKM outcomes
#dat.ckm<-read.csv("./CKM/Results/CKM stage summary.csv")

# use only baseline metabolite data for this analysis: 
base_plasma <- plasma %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)
base_plasma <- base_plasma %>% arrange(Date.Drawn)
base_plasma<-data.frame(base_plasma)
# use only urine metabolites that have "nM/mM":
urine$Date.Drawn <- as.Date(urine$Date.Drawn,format = "%m/%d/%Y")
base_urine <- urine %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)
base_urine <- base_urine %>% arrange(Date.Drawn)
# create kyurenine/tryptophan ratio
base_urine <- base_urine %>% mutate(`kyu.tryp.ratio.in.nM/mM` = `Kynurenine.in.nM/mM.Creatinine`/`Tryptophan.in.nM/mM.Creatinine`)
base_urine_subset <- base_urine[, grepl("releaseid|nM/mM", names(base_urine))]

# use only baseline ceramides:
base_ceramides <- plasma_ceramides %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)
base_ceramides <- base_ceramides %>% arrange(Date.Drawn)
base_ceramides<-data.frame(base_ceramides)

mta$Date.Drawn <- as.Date(mta$Date.Drawn,format = "%m/%d/%Y")
base_mta <- mta %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)
base_mta <- base_mta %>% arrange(Date.Drawn)

# merge all metabolites:
base_metabolites<-merge(base_plasma, base_ceramides, by="releaseid",all=T)
base_metabolites<-merge(base_metabolites, base_urine_subset, by="releaseid",all=T)
base_metabolites<-merge(base_metabolites, base_mta, by="releaseid",all=T)

# merge in complication data
base <- merge(base_metabolites, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
base <- merge(base, baserisk, by="releaseid",all.x=T, all.y=F)

# merge in CKM outcomes
# dat.ckm$releaseid<-dat.ckm$RELEASEID
# base <- merge(base, dat.ckm[,c("releaseid",
#                                "CKM_syn_base","CKM_max",
#                                "progress_CKM","days_to_first_ckm_prog",
#                                "progress_CKM_stage3","days_to_first_ckm_stage3_prog",
#                                "progress_CKM_stage4","days_to_first_ckm_stage4_prog")], by="releaseid",all.x=T, all.y=F)

# identify columns corresponding to proteins
is_seq <- function(.x) grepl("uM|nM/mM|Adenine|MTA", .x)
seq <- is_seq(names(base))

# convert protein variables to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)
metabolites<-base[,seq]
# remove proteins with low variability
no_var = caret::nearZeroVar(metabolites)
if (length(no_var)>0){
  no_var_names<-names(metabolites[,no_var])
  base <- base[,-which(colnames(base)%in%no_var_names)]  
}
seq <- is_seq(names(base))

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[predictors[i]] <- base_log_scale[predictors[i]]/sd(base_log[,paste0(predictors[i])],na.rm=T)
}

# change the format of the variable names so they can be read into coxph formula
temp <- str_replace_all(colnames(base_log_scale[,seq]),"/","_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"-","_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"3_","Three_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"2_","Two_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"\\(s\\)","s")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"5","Five")
colnames(base_log_scale)[seq] <- temp
```

```{r, include=FALSE}
#### Cox PH models for comorbidity outcomes:
# Macroalbuminuria
result_MAC <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMAC, MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MAC <- rbind(result_MAC,tidymod)
}
# perform FDR adjustment separately for urine vs. plasma metabolites:
result_MAC$metabolite_type<-"plasma"
result_MAC$metabolite_type[grepl("nM_mM|Adenine|MTA", result_MAC$term)]<-"urine"
result_MAC$adj.p.value<-NA
result_MAC$adj.p.value[result_MAC$metabolite_type=="plasma"] <- p.adjust(result_MAC$p.value[result_MAC$metabolite_type=="plasma"], "fdr")
result_MAC$adj.p.value[result_MAC$metabolite_type=="urine"] <- p.adjust(result_MAC$p.value[result_MAC$metabolite_type=="urine"], "fdr")

result_MAC <- result_MAC[order(result_MAC$adj.p.value),] 

# Microalbuminuria
result_MIC <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC, MIC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC <- rbind(result_MIC,tidymod)
}
result_MIC$metabolite_type<-"plasma"
result_MIC$metabolite_type[grepl("nM_mM|Adenine|MTA", result_MIC$term)]<-"urine"
result_MIC$adj.p.value<-NA
result_MIC$adj.p.value[result_MIC$metabolite_type=="plasma"] <- p.adjust(result_MIC$p.value[result_MIC$metabolite_type=="plasma"], "fdr")
result_MIC$adj.p.value[result_MIC$metabolite_type=="urine"] <- p.adjust(result_MIC$p.value[result_MIC$metabolite_type=="urine"], "fdr")

result_MIC <- result_MIC[order(result_MIC$adj.p.value),] 

# hyperfiltration
result_HYP <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHYP, HYP)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HYP <- rbind(result_HYP,tidymod)
}
result_HYP$metabolite_type<-"plasma"
result_HYP$metabolite_type[grepl("nM_mM|Adenine|MTA", result_HYP$term)]<-"urine"
result_HYP$adj.p.value<-NA
result_HYP$adj.p.value[result_HYP$metabolite_type=="plasma"] <- p.adjust(result_HYP$p.value[result_HYP$metabolite_type=="plasma"], "fdr")
result_HYP$adj.p.value[result_HYP$metabolite_type=="urine"] <- p.adjust(result_HYP$p.value[result_HYP$metabolite_type=="urine"], "fdr")

result_HYP <- result_HYP[order(result_HYP$adj.p.value),] 

# rapid eGFR decline
result_RAPID <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORAPID, RAPID)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RAPID <- rbind(result_RAPID,tidymod)
}
result_RAPID$metabolite_type<-"plasma"
result_RAPID$metabolite_type[grepl("nM_mM|Adenine|MTA", result_RAPID$term)]<-"urine"
result_RAPID$adj.p.value<-NA
result_RAPID$adj.p.value[result_RAPID$metabolite_type=="plasma"] <- p.adjust(result_RAPID$p.value[result_RAPID$metabolite_type=="plasma"], "fdr")
result_RAPID$adj.p.value[result_RAPID$metabolite_type=="urine"] <- p.adjust(result_RAPID$p.value[result_RAPID$metabolite_type=="urine"], "fdr")

result_RAPID <- result_RAPID[order(result_RAPID$adj.p.value),] 
# hypertension
# omit SBP
result_HTN <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HTN <- rbind(result_HTN,tidymod)
}
result_HTN$metabolite_type<-"plasma"
result_HTN$metabolite_type[grepl("nM_mM|Adenine|MTA", result_HTN$term)]<-"urine"
result_HTN$adj.p.value<-NA
result_HTN$adj.p.value[result_HTN$metabolite_type=="plasma"] <- p.adjust(result_HTN$p.value[result_HTN$metabolite_type=="plasma"], "fdr")
result_HTN$adj.p.value[result_HTN$metabolite_type=="urine"] <- p.adjust(result_HTN$p.value[result_HTN$metabolite_type=="urine"], "fdr")

result_HTN <- result_HTN[order(result_HTN$adj.p.value),] 

# hypertension with SBP
results_htn_bp <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_bp <- rbind(results_htn_bp,tidymod)
}
results_htn_bp$metabolite_type<-"plasma"
results_htn_bp$metabolite_type[grepl("nM_mM|Adenine|MTA", results_htn_bp$term)]<-"urine"
results_htn_bp$adj.p.value<-NA
results_htn_bp$adj.p.value[results_htn_bp$metabolite_type=="plasma"] <- p.adjust(results_htn_bp$p.value[results_htn_bp$metabolite_type=="plasma"], "fdr")
results_htn_bp$adj.p.value[results_htn_bp$metabolite_type=="urine"] <- p.adjust(results_htn_bp$p.value[results_htn_bp$metabolite_type=="urine"], "fdr")

results_htn_bp <- results_htn_bp[order(results_htn_bp$adj.p.value),] 

# neuropathy
result_NEURO <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTONEURO, NEURO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_NEURO <- rbind(result_NEURO,tidymod)
}
result_NEURO$metabolite_type<-"plasma"
result_NEURO$metabolite_type[grepl("nM_mM|Adenine|MTA", result_NEURO$term)]<-"urine"
result_NEURO$adj.p.value<-NA
result_NEURO$adj.p.value[result_NEURO$metabolite_type=="plasma"] <- p.adjust(result_NEURO$p.value[result_NEURO$metabolite_type=="plasma"], "fdr")
result_NEURO$adj.p.value[result_NEURO$metabolite_type=="urine"] <- p.adjust(result_NEURO$p.value[result_NEURO$metabolite_type=="urine"], "fdr")

result_NEURO <- result_NEURO[order(result_NEURO$adj.p.value),] 

# retinopathy
result_RETINO <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORETINO, RETINO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RETINO <- rbind(result_RETINO,tidymod)
}
result_RETINO$metabolite_type<-"plasma"
result_RETINO$metabolite_type[grepl("nM_mM|Adenine|MTA", result_RETINO$term)]<-"urine"
result_RETINO$adj.p.value<-NA
result_RETINO$adj.p.value[result_RETINO$metabolite_type=="plasma"] <- p.adjust(result_RETINO$p.value[result_RETINO$metabolite_type=="plasma"], "fdr")
result_RETINO$adj.p.value[result_RETINO$metabolite_type=="urine"] <- p.adjust(result_RETINO$p.value[result_RETINO$metabolite_type=="urine"], "fdr")

result_RETINO <- result_RETINO[order(result_RETINO$adj.p.value),] 
# glycemic failure
result_GLYC <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_GLYC <- rbind(result_GLYC,tidymod)
}
result_GLYC$metabolite_type<-"plasma"
result_GLYC$metabolite_type[grepl("nM_mM|Adenine|MTA", result_GLYC$term)]<-"urine"
result_GLYC$adj.p.value<-NA
result_GLYC$adj.p.value[result_GLYC$metabolite_type=="plasma"] <- p.adjust(result_GLYC$p.value[result_GLYC$metabolite_type=="plasma"], "fdr")
result_GLYC$adj.p.value[result_GLYC$metabolite_type=="urine"] <- p.adjust(result_GLYC$p.value[result_GLYC$metabolite_type=="urine"], "fdr")

result_GLYC <- result_GLYC[order(result_GLYC$adj.p.value),] 

# glycemic failure with A1c
results_glyc_a1c <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+HbA1c+bmi+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_glyc_a1c <- rbind(results_glyc_a1c,tidymod)
}
results_glyc_a1c$metabolite_type<-"plasma"
results_glyc_a1c$metabolite_type[grepl("nM_mM|Adenine|MTA", results_glyc_a1c$term)]<-"urine"
results_glyc_a1c$adj.p.value<-NA
results_glyc_a1c$adj.p.value[results_glyc_a1c$metabolite_type=="plasma"] <- p.adjust(results_glyc_a1c$p.value[results_glyc_a1c$metabolite_type=="plasma"], "fdr")
results_glyc_a1c$adj.p.value[results_glyc_a1c$metabolite_type=="urine"] <- p.adjust(results_glyc_a1c$p.value[results_glyc_a1c$metabolite_type=="urine"], "fdr")

results_glyc_a1c <- results_glyc_a1c[order(results_glyc_a1c$adj.p.value),] 

#### commented out for now, pending updates from Laura on CKM dataset
# CKM progression
# results_ckm_prog <- NULL
# predictors <- colnames(base_log_scale[seq])
# for (i in 1:length(predictors)) {
#   form <- paste0("Surv(days_to_first_ckm_prog, progress_CKM)~",predictors[i],"+sex")
#   mod <- coxph(data = base_log_scale, formula = as.formula(form))
#   tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
#   tidymod <- tidymod[1,]
#   results_ckm_prog <- rbind(results_ckm_prog,tidymod)
# }
# results_ckm_prog$adj.p.value <- p.adjust(results_ckm_prog$p.value, "fdr")
# results_ckm_prog <- results_ckm_prog[order(results_ckm_prog$p.value),] 
# 
# # CKM progression stage 3
# results_ckm_prog3 <- NULL
# predictors <- colnames(base_log_scale[seq])
# for (i in 1:length(predictors)) {
#   form <- paste0("Surv(days_to_first_ckm_stage3_prog, progress_CKM_stage3)~",predictors[i],"+sex")
#   mod <- coxph(data = base_log_scale, formula = as.formula(form))
#   tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
#   tidymod <- tidymod[1,]
#   results_ckm_prog3 <- rbind(results_ckm_prog3,tidymod)
# }
# results_ckm_prog3$adj.p.value <- p.adjust(results_ckm_prog3$p.value, "fdr")
# results_ckm_prog3 <- results_ckm_prog3[order(results_ckm_prog3$p.value),] 
# 
# # CKM progression stage 4
# results_ckm_prog4 <- NULL
# predictors <- colnames(base_log_scale[seq])
# for (i in 1:length(predictors)) {
#   form <- paste0("Surv(days_to_first_ckm_stage4_prog, progress_CKM_stage4)~",predictors[i],"+sex")
#   mod <- coxph(data = base_log_scale, formula = as.formula(form))
#   tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
#   tidymod <- tidymod[1,]
#   results_ckm_prog4 <- rbind(results_ckm_prog4,tidymod)
# }
# results_ckm_prog4$adj.p.value <- p.adjust(results_ckm_prog4$p.value, "fdr")
# results_ckm_prog4 <- results_ckm_prog4[order(results_ckm_prog4$p.value),] 

# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"MAC CPH")
writeData(wb,"MAC CPH",result_MAC,rowNames = F)
addWorksheet(wb,"MIC CPH")
writeData(wb,"MIC CPH",result_MIC,rowNames = F)
addWorksheet(wb,"HYP CPH")
writeData(wb,"HYP CPH",result_HYP,rowNames = F)
addWorksheet(wb,"RAPID CPH")
writeData(wb,"RAPID CPH",result_RAPID,rowNames = F)
addWorksheet(wb,"HTN with SBP CPH")
writeData(wb,"HTN with SBP CPH",results_htn_bp,rowNames = F)
addWorksheet(wb,"HTN CPH")
writeData(wb,"HTN CPH",result_HTN,rowNames = F)
addWorksheet(wb,"NEURO CPH")
writeData(wb,"NEURO CPH",result_NEURO,rowNames = F)
addWorksheet(wb,"RETINO CPH")
writeData(wb,"RETINO CPH",result_RETINO,rowNames = F)
addWorksheet(wb,"GLYC CPH")
writeData(wb,"GLYC CPH",result_GLYC,rowNames = F)
addWorksheet(wb,"GLYC with A1c CPH")
writeData(wb,"GLYC with A1c CPH",results_glyc_a1c,rowNames = F)
#addWorksheet(wb,"CKM progression")
#writeData(wb,"CKM progression",results_ckm_prog,rowNames = F)
#addWorksheet(wb,"CKM progression stage 3")
#writeData(wb,"CKM progression stage 3",results_ckm_prog3,rowNames = F)
#addWorksheet(wb,"CKM progression stage 4")
#writeData(wb,"CKM progression stage 4",results_ckm_prog4,rowNames = F)
saveWorkbook(wb,"./Results/Metabolomics Linear and Cox models/TODAY combined metabolomics Cox models_June2026.xlsx",overwrite = TRUE)


```

# Statistical Methods

Patient demographics, clinical characteristics, and outcomes were summarized for the TODAY cohort with available metabolomics data. Summary statistics included median (interquartile range, IQR) and frequency (percentage). 

Multivariable Cox Proportional Hazard (PH) models were fit for each outcome of interest. A separate model was fit for each metabolite (plasma or urine) and was adjusted for clincial covariates. FDR adjustment was performed for the set of urine metabolites and the set of plasma metabolites separately. Volcano plots were created for the multivariable Cox PH models, using the hazard ratio for each metabolite and the FDR-adjusted p-values. 

Covariates for each model are located here: /Biostatistics Core Shared Drive/TODAY subaward/Metabolomic data/TODAY metabolomics covariates.xlsx. The resulting model results for each metabolite are located here: /Biostatistics Core Shared Drive/TODAY subaward/Results/Metabolomics linear and Cox models/TODAY combined metabolomics Cox models_June2026.xlsx.

R version 4.5.2 was used for analysis.

# Summary Statistics

```{r echo=FALSE, include=FALSE}
label(base$si_1_ins0)<-"Insulin sensitivity"
label(base$log_trig)<-"Log(triglycerides)"
tab1<-table1(~ #pt demographics:   
               AGEBASE+sex_char+racedesc+bmi+bmipct+
               dxtime+
               HbA1c+log_trig+sbp+dbp+si_1_ins0,
                  data=base,
                  overall=c(left="Total"),
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)"))
tab1

base$MAC_table<-as.factor(base$MAC)
label(base$MAC_table)<-"Macroalbuminuria"
base$MIC_table<-as.factor(base$MIC)
label(base$MIC_table)<-"Microalbuminuria"
base$HYP_table<-as.factor(base$HYP)
label(base$HYP_table)<-"Hyperfiltration"
base$RAPID_table<-as.factor(base$RAPID)
label(base$RAPID_table)<-"Rapid eGFR decline"
base$HTN_table<-as.factor(base$HTN)
label(base$HTN_table)<-"Hypertension"
base$NEURO_table<-as.factor(base$NEURO)
label(base$NEURO_table)<-"Neuropathy"
base$RETINO_table<-as.factor(base$RETINO)
label(base$RETINO_table)<-"Retinopathy"
base$GLYC_table<-as.factor(base$GLYC)
label(base$GLYC_table)<-"Loss of glycemic control"
tab_outcomes<-table1(~
             MAC_table+MIC_table+HYP_table+RAPID_table+HTN_table+NEURO_table+RETINO_table+GLYC_table,
                  data=base,
                  overall=c(left="Total"),
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)"))
tab_outcomes

```

**Table 1: Demographics of TODAY metabolomics cohort**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab1, align = c("l", "l"))
``` 

**Table 2: Outcomes of TODAY metabolomics cohort**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab_outcomes, align = c("l", "l"))
``` 

# Modeling Results
The following volcano plots display the Hazard Ratio and -log10(P-Value) for each metabolite estimate in an adjusted Cox Proportional Hazards model. The p-values are FDR-adjusted. Only metabolites with FDR-adjusted p-value<0.05 are labeled in each plot. The interpretation of the hazard ratios is as follows: for a one SD increase in the log of metabolite, the risk is multiplied by xxx.


```{r echo=FALSE, include=TRUE, fig.width = 10, fig.height = 7}
### looped code for all volcano plots (from Claude):

# Define your datasets and their labels as a named list
datasets <- list(
  "MAC"          = result_MAC,
  "MIC"          = result_MIC,
  "HYP"          = result_HYP,
  "RAPID"        = result_RAPID,
  "HTN with SBP" = results_htn_bp,
  "HTN"          = result_HTN,
  "NEURO"        = result_NEURO,
  "RETINO"       = result_RETINO,
  "GLYC"         = result_GLYC,
  "GLYC with A1c"= results_glyc_a1c
)

# Store plots if you want to access them later
volcano_plots <- list()

for (label in names(datasets)) {
  full_results <- datasets[[label]]
  
  # ---- P-value columns ----
  full_results$p_unadjusted <- full_results$p.value
  full_results$p_adjusted   <- full_results$adj.p.value
  
  # ---- -log10 transformation ----
  full_results$p10_unadjusted <- -log10(pmax(full_results$p_unadjusted, 1e-10))
  full_results$p10_adjusted   <- -log10(pmax(full_results$p_adjusted,   1e-10))
  
  # ---- Color assignment ----
  full_results$color_unadjusted <- ifelse(
    full_results$p_unadjusted < 0.05 & full_results$estimate > 1, "lightcoral",
    ifelse(full_results$p_unadjusted < 0.05 & full_results$estimate < 1, "lightblue", "gray"))
  
  full_results$color_adjusted <- ifelse(
    full_results$p_adjusted < 0.05 & full_results$estimate > 1, "lightcoral",
    ifelse(full_results$p_adjusted < 0.05 & full_results$estimate < 1, "lightblue", "gray"))
  
  # ---- Significant subsets ----
  significant_df_adjusted   <- full_results[full_results$p_adjusted   < 0.05, ]
  significant_df_unadjusted <- full_results[full_results$p_unadjusted < 0.05, ]
  
  # ---- X-axis limits ----
  x_min <- min(full_results$estimate)
  x_max <- max(full_results$estimate)
  
  # ---- Adjusted volcano plot ----
  p_adjusted <- ggplot(full_results, aes(x = estimate, y = p10_adjusted, color = color_adjusted)) +
    geom_point(alpha = 0.7) +
    scale_color_identity() +
    theme_minimal() +
    labs(
      title    = paste0(label, ": Adjusted P-values"),
      x        = "HR",
      y        = "-log10(P-Value)",
      color    = "LogFC Direction") +
    theme(
      plot.title   = element_text(hjust = 0),
      axis.text.x  = element_text(angle = 0, hjust = 1)) +
    scale_x_continuous(limits = c(x_min, x_max), expand = expansion(mult = 0.2)) +
    coord_cartesian(clip = "off") +
    geom_text_repel(
      data  = significant_df_adjusted,
      aes(label = term),
      size  = 3,
      color = "black",
      max.overlaps = Inf   # ensures all labels are drawn
)
  
  # ---- Unadjusted volcano plot ----
  p_unadjusted <- ggplot(full_results, aes(x = estimate, y = p10_unadjusted, color = color_unadjusted)) +
    geom_point(alpha = 0.7) +
    scale_color_identity() +
    theme_minimal() +
    labs(
      title    = paste0(label, ": Unadjusted P-values"),
      x        = "HR",
      y        = "-log10(P-Value)",
      color    = "LogFC Direction") +
    theme(
      plot.title   = element_text(hjust = 0),
      axis.text.x  = element_text(angle = 0, hjust = 1)) +
    scale_x_continuous(limits = c(x_min, x_max), expand = expansion(mult = 0.2)) +
    coord_cartesian(clip = "off") +
    geom_text_repel(
      data  = significant_df_adjusted,
      aes(label = term),
      size  = 3,
      color = "black",
      max.overlaps = Inf)   # ensures all labels are drawn
  
  # ---- Print both plots ----
  print(p_adjusted)
  #print(p_unadjusted)
  
  # ---- Store in list (optional) ----
  volcano_plots[[paste0(label, "_adjusted")]]   <- p_adjusted
  volcano_plots[[paste0(label, "_unadjusted")]] <- p_unadjusted
}
```

```{r echo=FALSE, include=FALSE}
# kaplan meier plot for ASN abstract: 
library(survminer)
base$yearsMAC<-base$DAYSTOMAC/365.25
base$yearsMIC<-base$DAYSTOMIC/365.25
fit_mac <- survfit(Surv(yearsMAC, MAC) ~ 1, data = base)
fit_mic <- survfit(Surv(yearsMIC, MIC) ~ 1, data = base)

# Plot the data
fit <- list(Macroalbuminuria = fit_mac, Microalbuminuria = fit_mic)
ggsurvplot(fit, data = base, combine = TRUE, # Combine curves
           risk.table = TRUE,                  # Add risk table
           conf.int = TRUE,                    # Add confidence interval
           conf.int.style = "step",            # CI style, use "step" or "ribbon"
           censor = FALSE,                     # Remove censor points
           tables.theme = theme_cleantable(),  # Clean risk table
           palette = "jco")


fit_mic_overall <- survfit(Surv(yearsMIC, MIC) ~ 1, data = base)
#by median ceramide level:
base$C24d18<-ifelse(base$C24d18_1_24_0.in.uM<=median(base$C24d18_1_24_0.in.uM,na.rm=T),"<=17.4",">17.4")
fit_mic_c24 <- survfit(Surv(yearsMIC, MIC) ~ C24d18, data = base)
jpeg("/Users/kristenmiller/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive/TODAY subaward/Results/Metabolomics Linear and Cox models/c24_micro_km.jpeg",units="in",height=6,width=8,res=300)
ggsurvplot(fit_mic_c24, data = base,
           risk.table = F,                  # Add risk table
           conf.int = F,                    # Add confidence interval
           conf.int.style = "step",            # CI style, use "step" or "ribbon"
           censor = FALSE,                     # Remove censor points
           tables.theme = theme_cleantable(),  # Clean risk table
           xlab="Years in study",
           ylab="Albuminuria-free probability",
           legend.title="C24d18",
           legend.labs= c("below median", "above median"),
           palette = "npg")
dev.off()
```
