---
title: "TODAY plasma metabolomics baseline analyses - adjusted Cox models"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    toc: true
    toc_depth: '3'
  html_document:
    number_sections: true
    theme: readable
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
library(limma)
library(dplyr)
library(caret)
library(purrr)
library(multtest)
library(openxlsx)
library(tableone)
library(EnhancedVolcano)
library(knitr)
library(survival)
library(broom)
library(stringr)
library(table1)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)

user <- Sys.info()[["user"]]
if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive/TODAY subaward"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "kristenmiller") {
  root_path <- "/Users/kristenmiller/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive/TODAY subaward"
  git_path <- "/Users/kristenmiller/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else {
  stop("Unknown user: please specify root path for this user.")
}

knitr::opts_knit$set(root.dir = root_path)
```

```{r, include=FALSE}
# load plasma data
load("./Metabolomic data/plasma_all.Rdata")

# load comorbidity data
load("./Clinical data/comorb.Rdata")

# load baseline risk factors
load("./Clinical data/TODAY/baserisk.Rdata")

# load CKM outcomes
dat.ckm<-read.csv("./CKM/Results/CKM stage summary.csv")

# use only baseline metabolite data for this analysis: 
plasma_all$Date.Drawn <- as.Date(plasma_all$Date.Drawn,format = "%m/%d/%Y")
base <- plasma_all %>% arrange(releaseid,Date.Drawn) %>% group_by(releaseid) %>% filter(row_number()==1)
base <- base %>% arrange(Date.Drawn)
base<-data.frame(base)

# merge in complication data
base <- merge(base, comorb, by="releaseid",all.x=T, all.y=F)

# merge in baseline risk factors
base <- merge(base, baserisk, by="releaseid",all.x=T, all.y=F)

# merge in CKM outcomes
dat.ckm$releaseid<-dat.ckm$RELEASEID
base <- merge(base, dat.ckm[,c("releaseid",
                               "CKM_syn_base","CKM_max",
                               "progress_CKM","days_to_first_ckm_prog",
                               "progress_CKM_stage3","days_to_first_ckm_stage3_prog",
                               "progress_CKM_stage4","days_to_first_ckm_stage4_prog")], by="releaseid",all.x=T, all.y=F)

# identify columns corresponding to proteins
is_seq <- function(.x) grepl("uM", .x)
seq <- is_seq(names(base))

# convert protein variables to numeric
base[,seq] <- apply(base[,seq],2,as.numeric)
metabolites<-base[,seq]
# remove proteins with low variability
no_var = caret::nearZeroVar(metabolites)
if (length(no_var)>0){
  no_var_names<-names(metabolites[,no_var])
  base <- base[,-which(colnames(base)%in%no_var_names)]  
}
seq <- is_seq(names(base))

# log transform
base_log <- base %>% modify_if(is_seq(names(.)), log)

# scale by SD
base_log_scale <- base_log
predictors <- colnames(base_log[seq])
for (i in 1:length(predictors)) {
  base_log_scale[predictors[i]] <- base_log_scale[predictors[i]]/sd(base_log[,paste0(predictors[i])],na.rm=T)
}

# change the format of the variable names so they can be read into coxph formula
temp <- str_replace_all(colnames(base_log_scale[,seq]),"/","_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"-","_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"3_","Three_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"2_","Two_")
colnames(base_log_scale)[seq] <- temp
temp <- str_replace_all(colnames(base_log_scale[,seq]),"\\(s\\)","s")
colnames(base_log_scale)[seq] <- temp
```

```{r, include=FALSE}
#### Cox PH models for comorbidity outcomes:
# Macroalbuminuria
result_MAC <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  print(predictors[i])
  form <- paste0("Surv(DAYSTOMAC, MAC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MAC <- rbind(result_MAC,tidymod)
}
result_MAC$adj.p.value <- p.adjust(result_MAC$p.value, "fdr")
result_MAC <- result_MAC[order(result_MAC$p.value),] 

# Microalbuminuria
result_MIC <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOMIC, MIC)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_MIC <- rbind(result_MIC,tidymod)
}
result_MIC$adj.p.value <- p.adjust(result_MIC$p.value, "fdr")
result_MIC <- result_MIC[order(result_MIC$p.value),] 

# hyperfiltration
result_HYP <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHYP, HYP)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HYP <- rbind(result_HYP,tidymod)
}
result_HYP$adj.p.value <- p.adjust(result_HYP$p.value, "fdr")
result_HYP <- result_HYP[order(result_HYP$p.value),] 

# rapid eGFR decline
result_RAPID <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORAPID, RAPID)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RAPID <- rbind(result_RAPID,tidymod)
}
result_RAPID$adj.p.value <- p.adjust(result_RAPID$p.value, "fdr")
result_RAPID <- result_RAPID[order(result_RAPID$p.value),] 

# hypertension
# omit SBP
result_HTN <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_HTN <- rbind(result_HTN,tidymod)
}
result_HTN$adj.p.value <- p.adjust(result_HTN$p.value, "fdr")
result_HTN <- result_HTN[order(result_HTN$p.value),] 

# hypertension with SBP
results_htn_bp <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOHTN, HTN)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_htn_bp <- rbind(results_htn_bp,tidymod)
}
results_htn_bp$adj.p.value <- p.adjust(results_htn_bp$p.value, "fdr")
results_htn_bp <- results_htn_bp[order(results_htn_bp$p.value),] 

# neuropathy
result_NEURO <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTONEURO, NEURO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_NEURO <- rbind(result_NEURO,tidymod)
}
result_NEURO$adj.p.value <- p.adjust(result_NEURO$p.value, "fdr")
result_NEURO <- result_NEURO[order(result_NEURO$p.value),] 

# retinopathy
result_RETINO <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTORETINO, RETINO)~",predictors[i],"+HbA1c+log_trig+sbp+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_RETINO <- rbind(result_RETINO,tidymod)
}
result_RETINO$adj.p.value <- p.adjust(result_RETINO$p.value, "fdr")
result_RETINO <- result_RETINO[order(result_RETINO$p.value),] 

# glycemic failure
result_GLYC <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+bmi+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  result_GLYC <- rbind(result_GLYC,tidymod)
}
result_GLYC$adj.p.value <- p.adjust(result_GLYC$p.value, "fdr")
result_GLYC <- result_GLYC[order(result_GLYC$p.value),] 

# glycemic failure with A1c
results_glyc_a1c <- NULL
predictors <- colnames(base_log_scale[seq])
for (i in 1:length(predictors)) {
  form <- paste0("Surv(DAYSTOGLYC, GLYC)~",predictors[i],"+HbA1c+bmi+si_1_ins0+sex")
  mod <- coxph(data = base_log_scale, formula = as.formula(form))
  tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
  tidymod <- tidymod[1,]
  results_glyc_a1c <- rbind(results_glyc_a1c,tidymod)
}
results_glyc_a1c$adj.p.value <- p.adjust(results_glyc_a1c$p.value, "fdr")
results_glyc_a1c <- results_glyc_a1c[order(results_glyc_a1c$p.value),] 

#### commented out for now, pending updates from Laura on CKM dataset
# CKM progression
# results_ckm_prog <- NULL
# predictors <- colnames(base_log_scale[seq])
# for (i in 1:length(predictors)) {
#   form <- paste0("Surv(days_to_first_ckm_prog, progress_CKM)~",predictors[i],"+sex")
#   mod <- coxph(data = base_log_scale, formula = as.formula(form))
#   tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
#   tidymod <- tidymod[1,]
#   results_ckm_prog <- rbind(results_ckm_prog,tidymod)
# }
# results_ckm_prog$adj.p.value <- p.adjust(results_ckm_prog$p.value, "fdr")
# results_ckm_prog <- results_ckm_prog[order(results_ckm_prog$p.value),] 
# 
# # CKM progression stage 3
# results_ckm_prog3 <- NULL
# predictors <- colnames(base_log_scale[seq])
# for (i in 1:length(predictors)) {
#   form <- paste0("Surv(days_to_first_ckm_stage3_prog, progress_CKM_stage3)~",predictors[i],"+sex")
#   mod <- coxph(data = base_log_scale, formula = as.formula(form))
#   tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
#   tidymod <- tidymod[1,]
#   results_ckm_prog3 <- rbind(results_ckm_prog3,tidymod)
# }
# results_ckm_prog3$adj.p.value <- p.adjust(results_ckm_prog3$p.value, "fdr")
# results_ckm_prog3 <- results_ckm_prog3[order(results_ckm_prog3$p.value),] 
# 
# # CKM progression stage 4
# results_ckm_prog4 <- NULL
# predictors <- colnames(base_log_scale[seq])
# for (i in 1:length(predictors)) {
#   form <- paste0("Surv(days_to_first_ckm_stage4_prog, progress_CKM_stage4)~",predictors[i],"+sex")
#   mod <- coxph(data = base_log_scale, formula = as.formula(form))
#   tidymod <- tidy(mod, exponentiate=TRUE, conf.int = TRUE)
#   tidymod <- tidymod[1,]
#   results_ckm_prog4 <- rbind(results_ckm_prog4,tidymod)
# }
# results_ckm_prog4$adj.p.value <- p.adjust(results_ckm_prog4$p.value, "fdr")
# results_ckm_prog4 <- results_ckm_prog4[order(results_ckm_prog4$p.value),] 

# write all results to file
wb <- createWorkbook()
addWorksheet(wb,"MAC CPH")
writeData(wb,"MAC CPH",result_MAC,rowNames = F)
addWorksheet(wb,"MIC CPH")
writeData(wb,"MIC CPH",result_MIC,rowNames = F)
addWorksheet(wb,"HYP CPH")
writeData(wb,"HYP CPH",result_HYP,rowNames = F)
addWorksheet(wb,"RAPID CPH")
writeData(wb,"RAPID CPH",result_RAPID,rowNames = F)
addWorksheet(wb,"HTN with SBP CPH")
writeData(wb,"HTN with SBP CPH",results_htn_bp,rowNames = F)
addWorksheet(wb,"HTN CPH")
writeData(wb,"HTN CPH",result_HTN,rowNames = F)
addWorksheet(wb,"NEURO CPH")
writeData(wb,"NEURO CPH",result_NEURO,rowNames = F)
addWorksheet(wb,"RETINO CPH")
writeData(wb,"RETINO CPH",result_RETINO,rowNames = F)
addWorksheet(wb,"GLYC CPH")
writeData(wb,"GLYC CPH",result_GLYC,rowNames = F)
addWorksheet(wb,"GLYC with A1c CPH")
writeData(wb,"GLYC with A1c CPH",results_glyc_a1c,rowNames = F)
#addWorksheet(wb,"CKM progression")
#writeData(wb,"CKM progression",results_ckm_prog,rowNames = F)
#addWorksheet(wb,"CKM progression stage 3")
#writeData(wb,"CKM progression stage 3",results_ckm_prog3,rowNames = F)
#addWorksheet(wb,"CKM progression stage 4")
#writeData(wb,"CKM progression stage 4",results_ckm_prog4,rowNames = F)
saveWorkbook(wb,"./Results/Metabolomics Linear and Cox models/TODAY plasma metabolomics Cox models scaled baseline adjusted_march2026.xlsx",overwrite = TRUE)


```

# Methods

Patient demographics, covariates included in modeling, and outcomes were summarized for the TODAY cohort with available metabolomics data. Summary statistics included median (interquartile range, IQR) and frequency (percentage). 

Multivariable Cox Proportional Hazard (PH) models were fit for each outcome of interest. Covariates for each model are located here: iostatistics Core Shared Drive/TODAY subaward/Metabolomic data/TODAY metabolomics covariates.xlsx. The detailed model results are here: Biostatistics Core Shared Drive/TODAY subaward/Results/Metabolomics linear and Cox models/TODAY plasma metabolomics Cox models scaled baseline adjusted_march2026.xlsx and Biostatistics Core Shared Drive/TODAY subaward/Results/Metabolomics linear and Cox models/TODAY urine metabolomics Cox models scaled baseline adjusted_march2026.xlsx. 

Volcano plots were created for the multivariable Cox PH models, using the hazard ratios for each metabolite. FDR adjusted p-values and unajdusted p-values were plotted in separate figures. 

# Summary Statistics

```{r echo=FALSE, include=FALSE}
tab1<-table1(~ #pt demographics:   
               AGEBASE+sex_char+racedesc+bmi+bmipct+
               dxtime+
               HbA1c+log_trig+sbp+dbp+si_1_ins0,
                  data=base,
                  overall=c(left="Total"),
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)"))
tab1

tab_outcomes<-table1(~
             factor(MAC)+factor(MIC)+factor(HYP)+factor(RAPID)+factor(HTN)+factor(NEURO)+factor(RETINO)+factor(GLYC),
                  data=base,
                  overall=c(left="Total"),
                  render.continuous=c("Median (IQR)"="MEDIAN (Q1, Q3)"))
tab_outcomes

```

**Table 1: Demographics of TODAY metabolomics cohort**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab1, align = c("l", "l"))
``` 

**Table 2: Outcomes of TODAY metabolomics cohort**
```{r,echo=FALSE,message=FALSE,results='asis',warning=FALSE}
kable(tab_outcomes, align = c("l", "l"))
``` 

# Volcano plots 
Need to remember that the interpretation of the hazard ratios is as follows: for a one SD increase in the log of seqXXX, the risk is multiplied by....

```{r echo=FALSE, include=TRUE, fig.width = 10, fig.height = 7}
### looped code for all volcano plots (from Claude):

# Define your datasets and their labels as a named list
datasets <- list(
  "MAC"          = result_MAC,
  "MIC"          = result_MIC,
  "HYP"          = result_HYP,
  "RAPID"        = result_RAPID,
  "HTN with SBP" = results_htn_bp,
  "HTN"          = result_HTN,
  "NEURO"        = result_NEURO,
  "RETINO"       = result_RETINO,
  "GLYC"         = result_GLYC,
  "GLYC with A1c"= results_glyc_a1c
)

# Store plots if you want to access them later
volcano_plots <- list()

for (label in names(datasets)) {
  
  full_results <- datasets[[label]]
  
  # ---- P-value columns ----
  full_results$p_unadjusted <- full_results$p.value
  full_results$p_adjusted   <- full_results$adj.p.value
  
  # ---- -log10 transformation ----
  full_results$p10_unadjusted <- -log10(pmax(full_results$p_unadjusted, 1e-10))
  full_results$p10_adjusted   <- -log10(pmax(full_results$p_adjusted,   1e-10))
  
  # ---- Color assignment ----
  full_results$color_unadjusted <- ifelse(
    full_results$p_unadjusted < 0.05 & full_results$estimate > 1, "lightcoral",
    ifelse(full_results$p_unadjusted < 0.05 & full_results$estimate < 1, "lightblue", "gray"))
  
  full_results$color_adjusted <- ifelse(
    full_results$p_adjusted < 0.05 & full_results$estimate > 1, "lightcoral",
    ifelse(full_results$p_adjusted < 0.05 & full_results$estimate < 1, "lightblue", "gray"))
  
  # ---- Significant subsets ----
  significant_df_adjusted   <- full_results[full_results$p_adjusted   < 0.05, ]
  significant_df_unadjusted <- full_results[full_results$p_unadjusted < 0.05, ]
  
  # ---- X-axis limits ----
  x_min <- min(full_results$estimate)
  x_max <- max(full_results$estimate)
  
  # ---- Adjusted volcano plot ----
  p_adjusted <- ggplot(full_results, aes(x = estimate, y = p10_adjusted, color = color_adjusted)) +
    geom_point(alpha = 0.7) +
    scale_color_identity() +
    theme_minimal() +
    labs(
      title    = paste0(label, ": Adjusted P-values"),
      x        = "HR",
      y        = "-log10(P-Value)",
      color    = "LogFC Direction") +
    theme(
      plot.title   = element_text(hjust = 0),
      axis.text.x  = element_text(angle = 0, hjust = 1)) +
    scale_x_continuous(limits = c(x_min, x_max), expand = expansion(mult = 0.2)) +
    coord_cartesian(clip = "off") +
    geom_text(data = significant_df_adjusted, aes(label = term),
              vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
  
  # ---- Unadjusted volcano plot ----
  p_unadjusted <- ggplot(full_results, aes(x = estimate, y = p10_unadjusted, color = color_unadjusted)) +
    geom_point(alpha = 0.7) +
    scale_color_identity() +
    theme_minimal() +
    labs(
      title    = paste0(label, ": Unadjusted P-values"),
      x        = "HR",
      y        = "-log10(P-Value)",
      color    = "LogFC Direction") +
    theme(
      plot.title   = element_text(hjust = 0),
      axis.text.x  = element_text(angle = 0, hjust = 1)) +
    scale_x_continuous(limits = c(x_min, x_max), expand = expansion(mult = 0.2)) +
    coord_cartesian(clip = "off") +
    geom_text(data = significant_df_unadjusted, aes(label = term),
              vjust = 1, hjust = 1, size = 3, check_overlap = TRUE, color = "black")
  
  # ---- Print both plots ----
  print(p_adjusted)
  print(p_unadjusted)
  
  # ---- Store in list (optional) ----
  volcano_plots[[paste0(label, "_adjusted")]]   <- p_adjusted
  volcano_plots[[paste0(label, "_unadjusted")]] <- p_unadjusted
}
```

```{r echo=FALSE, include=FALSE}
# kaplan meier plot for ASN abstract: 
library(survminer)
base$yearsMAC<-base$DAYSTOMAC/365.25
base$yearsMIC<-base$DAYSTOMIC/365.25
fit_mac <- survfit(Surv(yearsMAC, MAC) ~ 1, data = base)
fit_mic <- survfit(Surv(yearsMIC, MIC) ~ 1, data = base)

# Plot the data
fit <- list(Macroalbuminuria = fit_mac, Microalbuminuria = fit_mic)
ggsurvplot(fit, data = base, combine = TRUE, # Combine curves
           risk.table = TRUE,                  # Add risk table
           conf.int = TRUE,                    # Add confidence interval
           conf.int.style = "step",            # CI style, use "step" or "ribbon"
           censor = FALSE,                     # Remove censor points
           tables.theme = theme_cleantable(),  # Clean risk table
           palette = "jco")


fit_mic_overall <- survfit(Surv(yearsMIC, MIC) ~ 1, data = base)
#by median ceramide level:
base$C24d18<-ifelse(base$C24d18_1_24_0.in.uM<=median(base$C24d18_1_24_0.in.uM,na.rm=T),"<=17.4",">17.4")
fit_mic_c24 <- survfit(Surv(yearsMIC, MIC) ~ C24d18, data = base)
jpeg("/Users/kristenmiller/Library/CloudStorage/OneDrive-UW/Laura Pyle's files - Biostatistics Core Shared Drive/TODAY subaward/Results/Metabolomics Linear and Cox models/c24_micro_km.jpeg",units="in",height=6,width=8,res=300)
ggsurvplot(fit_mic_c24, data = base,
           risk.table = F,                  # Add risk table
           conf.int = F,                    # Add confidence interval
           conf.int.style = "step",            # CI style, use "step" or "ribbon"
           censor = FALSE,                     # Remove censor points
           tables.theme = theme_cleantable(),  # Clean risk table
           xlab="Years in study",
           ylab="Albuminuria-free probability",
           legend.title="C24d18",
           legend.labs= c("below median", "above median"),
           palette = "npg")
dev.off()
```
