---
title: "ATTEMPT scRNA analysis on Hyak (mixed model)"
author: "Ye Ji Choi"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
---

```{r libraries, message = F, warning=F}
#| include: false
library(future)
library(future.apply)
library(tidyverse)
library(colorspace)
library(patchwork)
library(ggdendro)
library(cowplot)
library(ggpubr)
library(rstatix)
library(arsenal)
library(Biobase)
library(msigdbr)
library(kableExtra)
# library(knitr)
library(REDCapR)
# library(data.table)
library(emmeans)
library(NMF)
library(pheatmap)
library(UpSetR)
library(enrichR)
library(WriteXLS)
library(SAVER)
library(readxl)
library(limma)
library(edgeR)
library(BiocGenerics)
library(GSEABase)
library(slingshot)
library(SingleCellExperiment)
library(MAST)
library(muscat)
library(scater)
library(Seurat)
library(jsonlite)
library(dplyr)
library(glmmTMB)
library(reshape2)
library(broom.mixed) 
library(foreach)
library(doParallel)
library(nebula)
library(Matrix)
library(scran)
library(DirichletReg)
library(BiocParallel)
library(scDC)
library(DoubletFinder)
library(scCustomize)
```

```{r include = F}
reticulate::use_python("/mmfs1/gscratch/scrubbed/yejichoi/mypy/bin/python", required = T)
reticulate::repl_python()
reticulate::py_config()
# set up Kopah connection
## Load boto3 and pandas
boto3 <- reticulate::import("boto3")
pd <- reticulate::import("pandas")

## Create an S3 client
keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
session <- boto3$session$Session(
  aws_access_key_id = keys$MY_ACCESS_KEY,
  aws_secret_access_key = keys$MY_SECRET_KEY
)

## Create an S3 client with the session
s3 <- session$client("s3", endpoint_url = "https://s3.kopah.uw.edu")
```

```{r echo = F, eval = F}
# read ATTEMPT Seurat file
temp_file <- tempfile(fileext = ".RDS")
s3$download_file("scrna", "Kidney transcriptomics/Single cell RNA seq/PB_attempt_harmony_rpca_RM_kpmpV1labelled_Sept2024.RDS", temp_file)
attempt_so <- readRDS(temp_file)

# N of cells for Petter (B-cells and POD at baseline)
# attempt_so$celltype_pt <- ifelse(grepl("PT-", attempt_so$celltype),
#                                  "PT", as.character(attempt_so$celltype))
# attempt_so_pt <- subset(attempt_so, celltype_pt == "PT" & celltype != "PT_lowQuality")
# attempt_so_b_bl <- subset(attempt_so, celltype == "B" & Visit == "BL")
# attempt_so_pod_bl <- subset(attempt_so, celltype == "POD" & Visit == "BL")


# calculate cell library size for offset

# Identify mitochondrial genes
mito_genes <- grep("^MT-", rownames(attempt_so), value = TRUE)

# Calculate the percentage of mitochondrial transcripts per cell
attempt_so[["percent.mt"]] <- PercentageFeatureSet(attempt_so, pattern = "^MT-")
hist(attempt_so$percent.mt)
# FeaturePlot(attempt_so, reduction = "umap.harmony", features = "percent.mt", cols = c("lightgray", "blue"), raster = F)
# DimPlot(attempt_so, reduction = "umap.harmony", group.by = "KPMP_celltype", raster = F, alpha = 0.2, label = T)
# DimPlot(attempt_so, reduction = "umap.harmony", group.by = "celltype", raster = F, alpha = 0.2, label = T)

# Identify ribosomal genes
ribo_genes <- c(
  "RPL22", "RPL11", "RPS8", "RPL5", "RPS27", "RPS7", "RPS27A", "RPL31", "RPL37A", "RPL32", "RPL15", "RPL14", "RPL29",
  "RPL24", "RPL22L1", "RPL35A", "RPL9", "RPL34", "RPS3A", "RPL37", "RPS23", "RPS14", "RPS18", "RPS10", "RPL10A", 
  "RPS20", "RPL7", "RPL30", "RPL8", "RPS6", "RPL35", "RPL12", "RPL7A", "RPS24", "RPLP2", "RPL27A", "RPS13", "RPS3",
  "RPS25", "RPS26", "RPL41", "RPL6", "RPLP0", "RPL21", "RPS29", "RPL4", "RPLP1", "RPS17", "RPS2", "RPS15A", "RPL13",
  "RPL26", "RPL23A", "RPL23", "RPL19", "RPL27", "RPL38", "RPL17", "RPS15", "RPL36", "RPS28", "RPL18A", "RPS16", 
  "RPS19", "RPL18", "RPL13A", "RPS11", "RPS9", "RPL28", "RPS5", "RPS21", "RPL3", "RPS4X", "RPL36A", "RPL39", 
  "RPL10", "RPS4Y1"
) # grep("^RPL|^RPS", rownames(attempt_so), value = TRUE) captures some none ribosomal genes
attempt_so[["percent.ribo"]] <- PercentageFeatureSet(attempt_so, features = ribo_genes)
hist(attempt_so$percent.ribo)

# attempt_so <- subset(attempt_so, subset = percent.mt < 50) # keep as 50%, it's how it's done with KPMP
FeaturePlot(attempt_so, reduction = "umap.harmony", features = "percent.mt", cols = c("lightgray", "blue"), raster = F)
DimPlot(attempt_so, reduction = "umap.harmony", group.by = "KPMP_celltype", raster = F, alpha = 0.2, label = T)
DimPlot(attempt_so, reduction = "umap.harmony", group.by = "celltype", raster = F, alpha = 0.2, label = T)

attempt_so <- subset(attempt_so, features = setdiff(rownames(attempt_so), mito_genes)) # remove MT- genes
attempt_so <- subset(attempt_so, features = setdiff(rownames(attempt_so), ribo_genes)) # remove ribosomal genes

expr_matrix <- as.matrix(GetAssayData(attempt_so, layer = "data"))
# Calculate the proportion of cells expressing each gene
num_cells_per_gene <- rowSums(expr_matrix > 0)  # Count nonzero cells per gene
total_cells <- ncol(expr_matrix)  # Total number of cells
gene_proportion <- num_cells_per_gene / total_cells  # Fraction of cells expressing each gene
length(gene_proportion) # 30625
# Keep genes expressed in at least 5% of cells
genes_to_keep <- names(gene_proportion[gene_proportion  >= 0.05])
length(genes_to_keep) # 8672 ( ~ 28.3% remaining)

remove(expr_matrix)
attempt_so <- subset(attempt_so, features = genes_to_keep)

# renormalize after filtering
attempt_so <- NormalizeData(attempt_so)

# # calculate cell library size for offset in NEBULA
counts_layer <- round(GetAssayData(attempt_so, layer = "counts"))
library_size <- Matrix::colSums(counts_layer)
attempt_so$library_size <- library_size
View(attempt_so@meta.data)

# TMM offset
dge <- DGEList(counts = counts_layer)
dge <- calcNormFactors(dge, method = "TMM")
tmm_offset <- (dge$samples$lib.size) * (dge$samples$norm.factors)
attempt_so$tmm_offset <- tmm_offset

# Pooled offset
bp <- MulticoreParam(workers = 63)
sce <- SingleCellExperiment(assays = list(counts = counts_layer))
# sce <- computeSumFactors(sce)
sce <- computeSumFactors(sce, BPPARAM = bp)
# # View size factors
sizeFactors(sce)
# Calculate offset → (size factors)
pooled_offset <- (sizeFactors(sce))
attempt_so$pooled_offset <- pooled_offset

## want to double check expression of some metallothionines because logFC for them in dapa is near 0
# FindMarkers(subset(attempt_so, grepl("PT-", attempt_so$celltype) & treatment == "Placebo"), features = c("MT1X", "MT2A", "MT1M", "MT1H"), group.by = "visit", 
#             min.pct = 0, logfc.threshold = 0,
#             ident.1 = "POST")
# 
# FindMarkers(subset(attempt_so, grepl("PT-", attempt_so$celltype) & treatment == "Dapagliflozin 5mg"), features = c("MT1X", "MT2A", "MT1M", "MT1H"), group.by = "visit", 
#             min.pct = 0, logfc.threshold = 0,
#             ident.1 = "POST")

# check for expression of TRHDE-AS1 (in immune), FGF14 (in immune), and PSAT1 (in fib/vsmc/p)
FeaturePlot(attempt_so, features = c("TRHDE-AS1", "FGF14", "PSAT1"), raster = F)
DimPlot(attempt_so, group.by = "celltype", raster = F, alpha = 0.2, label = T)

```

```{r echo = F, eval = F}
# read ATTEMPT clinical data file
temp_file <- tempfile(fileext = ".RDS")
s3$download_file("harmonized.dataset", "attempt_clinical_data.RDS", temp_file)
dat <- readRDS(temp_file)

attempt_so_meta <- attempt_so@meta.data %>%
  mutate(subject_id = Subject.ID,
         visit = case_when(Visit == "BL" ~ "PRE", 
                           Visit == "4M" ~ "POST"))
attempt_so_meta <- left_join(attempt_so_meta, dat, by = c("subject_id", "visit"))
rownames(attempt_so_meta) <- attempt_so_meta$barcode
attempt_so <- AddMetaData(attempt_so, attempt_so_meta)
options(future.globals.maxSize = 3000 * 1024^3)
```


```{r echo = F, eval = T}
# Read in clean seurat object saved from attempt_CROC_analysis.qmd
temp_file <- tempfile(fileext = ".rds")
s3$download_file("attempt", 'cleaned_data/attempt_clean_so.rds', temp_file)
attempt_so <- readRDS(temp_file)
```


```{r echo = F}
# remove cells with ambient RNA (indicated with high expression of PT biomarkers)
pt_markers <- c("MIOX", "GLYAT", "ALDOB", "GATM", "FABP1", "HPD", "GPX3")
other_markers <- c("MIOX", "GLYAT", "ALDOB", "GATM", "FABP1", "HPD", "GPX3")
# FeaturePlot(subset(attempt_so, celltype %in% c("IC-A", "IC-B")), features = pt_markers)
# FeaturePlot(subset(attempt_so, celltype %in% c("TAL-1", "TAL-2", "TAL-3")), features = pt_markers)

attempt_so <- AddModuleScore(attempt_so, features = list(pt_markers), name = "ambient_score")
Idents(attempt_so) <- attempt_so$celltype
VlnPlot(attempt_so, features = "ambient_score1", pt.size = 0.1, raster = F, alpha = 0.3)

attempt_so <- subset(attempt_so, subset = (ambient_score1 < 0.25 & 
                                                       celltype %nin% c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5")) 
                               | celltype %in% c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))

VlnPlot(attempt_so, features = "ambient_score1", pt.size = 0.1, raster = F, alpha = 0.3)
FeaturePlot(attempt_so, features = pt_markers)

```


```{r echo = F, eval = F}
# Doublet detection (for QC, post Michigan meeting in June '25)
# need to do this in each subject, not on integrated samples, but doublets are not a big issue if we're eliminating cells with ambient rna
# After running standard Seurat workflow (e.g., PCA)
attempt_so$SampleID <- paste0(attempt_so$subject_id, "_", attempt_so$visit)
samp_split <- SplitObject(attempt_so, split.by = "SampleID") 

# Get Doublet/Singlet IDs by DoubletFinder()
samp_split <- lapply(samp_split, run_doubletfinder_custom, multiplet_rate= 0.075) # get singlet/doublet assigned to each of the cell IDs (each element of the list is a different sample)

sglt_dblt_metadata <- data.frame(bind_rows(samp_split)) # merge to a single dataframe
rownames(sglt_dblt_metadata) <- sglt_dblt_metadata$row_names # assign cell IDs to row names to ensure match
sglt_dblt_metadata$row_names <- NULL
head(sglt_dblt_metadata)
attempt_so <- AddMetaData(attempt_so, sglt_dblt_metadata, col.name = 'doublet_finder')

# Check how doublets singlets differ in QC measures per sample.
VlnPlot(attempt_so, group.by = 'SampleID', split.by = "doublet_finder",
        features = c("nFeature_RNA", "nCount_RNA", "percent_mt", "percent_ribo", "percent_hb"), 
        ncol = 3, pt.size = 0) + theme(legend.position = 'right')

doublets_summary <- attempt_so@meta.data %>% 
  group_by(SampleID, doublet_finder) %>% 
  summarise(total_count = n(),.groups = 'drop') %>% as.data.frame() %>% ungroup() %>%
  group_by(SampleID) %>%
  mutate(countT = sum(total_count)) %>%
  group_by(doublet_finder, .add = TRUE) %>%
  mutate(percent = paste0(round(100 * total_count/countT, 2),'%')) %>%
  dplyr::select(-countT)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(doublets_summary, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'cleaned_data/attempt_doublets_summary.rds')

attempt_so <- subset(attempt_so, doublet_finder == 'Singlet') # 208817 -> 196354 (retaining 94%)


temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(attempt_so, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'cleaned_data/attempt_clean_so.rds')
```


```{r echo = F, eval = T}
# Read in clean seurat object saved from attempt_CROC_analysis.qmd
temp_file <- tempfile(fileext = ".rds")
s3$download_file("attempt", 'cleaned_data/attempt_clean_so.rds', temp_file)
attempt_so <- readRDS(temp_file)
```

# All cell types

```{r echo = F}
attempt_so$visit <- factor(attempt_so$visit, levels = c("PRE", "POST"))
attempt_so$treatment <- factor(attempt_so$treatment, levels = c("Placebo", "Dapagliflozin"),
                                  labels = c("Placebo", "Dapagliflozin"))
```

```{r echo = F}
DimPlot(attempt_so, reduction = "umap.harmony", group.by = "visit", raster = F, alpha = 0.2)
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/attempt_umap_visit.jpeg") 

DimPlot(attempt_so, reduction = "umap.harmony", group.by = "treatment", raster = F, alpha = 0.2)
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/attempt_umap_treat.jpeg") 

DimPlot(attempt_so, reduction = "umap.harmony", group.by = "celltype", raster = F, alpha = 0.2, label = T)
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/attempt_umap_celltype.jpeg") 

DimPlot(attempt_so, reduction = "umap.harmony", group.by = "KPMP_celltype", raster = F, alpha = 0.2, label = T)
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/attempt_umap_kpmpcelltype.jpeg") 

```

```{r echo = F}
# umap formatting
```

#### Proportions (Non weighted)

```{r echo = F}
attempt_so$celltype_general <- case_when(
  grepl("PT-", attempt_so$celltype) ~ "PT",
  grepl("TAL-", attempt_so$celltype) ~ "TAL",
  attempt_so$celltype %in% c("MON", "MAC") ~ "Immune_Myeloid",
  attempt_so$celltype %in% c("T", "B", "NKT/NKC") ~ "Immune_Lymphoid",
  attempt_so$celltype %in% c("PC-1", "PC-2", "tPC-IC") ~ "PC",
  attempt_so$celltype %in% c("IC-A", "IC-B") ~ "IC",
  attempt_so$celltype %in% c("EC-PTC", "EC-GC", "EC-AEA", "EC-LYM") ~ "EC",
  attempt_so$celltype %in% c("FIB", "VSMC/P") ~ "FIB/VSMC/P",
  TRUE ~ "Other - ATL, CNT, DCT, DTL, MC, PEC, POD, lowQ")

attempt_meta <- attempt_so@meta.data %>%
  mutate(visit_treatment = paste0(visit, " ", treatment))
attempt_meta$visit_treatment <- factor(attempt_meta$visit_treatment, 
                                          levels = c("PRE Placebo", "POST Placebo",
                                                     "PRE Dapagliflozin", "POST Dapagliflozin"))
attempt_meta$celltype <- factor(attempt_meta$celltype,
                                level = c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5", 
                                          "TAL-1", "TAL-2", "TAL-3", "TAL_highUMI", 
                                          "DTL-1", "DTL-2", 
                                          "CNT", "PC-1", "PC-2", 
                                          "tPC-IC", "IC-A", "IC-A_lowQuality", "IC-B", 
                                          "ATL", 
                                          "POD", "PEC", 
                                          "DCT",
                                          "MC", 
                                          "EC-PTC", "EC-GC", "EC-AEA", "EC-LYM", 
                                          "VSMC/P", "FIB", 
                                          "MAC", "MON", "NKT/NKC", "T", "B"))

attempt_meta$KPMP_celltype <- factor(attempt_meta$KPMP_celltype,
                                level = c("PT-S1/S2", "PT-S3", "aPT",
                                          "C-TAL-1", "C-TAL-2", "aTAL",
                                          "aDTL", "DTL",
                                          "CNT", "CNT-PC", "CCD-PC", "dCCD-PC", "M-PC",
                                          "tPC-IC", "IC-A", "IC-B",
                                          "ATL",
                                          "POD", "PEC",
                                          "DCT", "dDCT",
                                          "MC",
                                          "EC-PTC", "EC-GC", "EC-AEA", "EC-LYM", "EC-AVR", "EC/VSMC",
                                          "VSMC/P", "FIB",
                                          "MAC", "MON", "NK",
                                          "CD4+ T", "CD8+ T", "cycT",
                                          "B",
                                          "pDC", "cDC",
                                          "aIC", "SchwannCells", "non-specific"))


attempt_meta$celltype_general <- factor(attempt_meta$celltype_general,
                                levels = c("PT", 
                                           "TAL", 
                                           "PC", 
                                           "IC", 
                                           "EC", 
                                           "FIB/VSMC/P", 
                                           "Immune_Myeloid",
                                           "Immune_Lymphoid",
                                           "Other - ATL, CNT, DCT, DTL, MC, PEC, POD, lowQ"))

# Check counts
table(attempt_so$celltype_general)
```

```{r echo = F}
attempt_meta %>% 
  filter(!grepl("_", celltype)) %>%
  ggplot(aes(celltype, fill = visit_treatment)) + 
  geom_bar(stat = "count") +
  theme_classic() +
  labs(x = NULL,
       y = "Cell count",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("Cell type distribution") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/attempt_cellcounts_plot.jpeg") 

attempt_meta %>%
  filter(!grepl("_", celltype)) %>%
  ggplot(aes(celltype, fill = visit_treatment)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 15)
s3$upload_file(temp_file, "attempt", "Figures/attempt_cellcounts_proportions_plot.jpeg") 

attempt_meta %>% 
  filter(!grepl("_", KPMP_celltype)) %>%
  ggplot(aes(KPMP_celltype, fill = visit_treatment)) + 
  geom_bar(stat = "count") +
  theme_classic() +
  labs(x = NULL,
       y = "Cell count",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("Cell type distribution") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/attempt_kpmp_cellcounts_plot.jpeg") 

attempt_meta %>%
  filter(!grepl("_", KPMP_celltype)) %>%
  ggplot(aes(KPMP_celltype, fill = visit_treatment)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 15)
s3$upload_file(temp_file, "attempt", "Figures/attempt_kpmp_cellcounts_proportions_plot.jpeg") 


attempt_meta %>% 
  ggplot(aes(celltype_general, fill = visit_treatment)) + 
  geom_bar(stat = "count") +
  theme_classic() +
  labs(x = NULL,
       y = "Cell count",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("General cell type distribution") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file)
s3$upload_file(temp_file, "attempt", "Figures/attempt_general_cellcounts_plot.jpeg") 

attempt_meta %>%
  ggplot(aes(celltype_general, fill = visit_treatment)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/attempt_general_cellcounts_proportions_plot.jpeg") 

celltype_counts <- attempt_meta %>%
  filter(!grepl("Other", as.character(celltype_general))) %>%
  group_by(celltype_general) %>%
  summarise(n = n()) %>%
  ungroup()

attempt_meta %>% 
  filter(!grepl("Other", celltype_general)) %>%
  ggplot(aes(celltype_general, fill = visit_treatment)) + 
  geom_bar(stat = "count") +
  geom_text(data = celltype_counts, aes(x = celltype_general, y = n, label = n), 
            vjust = -0.5, size = 4, inherit.aes = FALSE) +
  theme_classic() +
  labs(x = NULL,
       y = "Cell count",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("General cell type distribution") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file)
s3$upload_file(temp_file, "attempt", "Figures/attempt_general_cellcounts_plot_clean.jpeg") 

attempt_meta %>%
  filter(!grepl("Other", celltype_general)) %>%
  ggplot(aes(celltype_general, fill = visit_treatment)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#fbc4ab", "#f4978e",
                               "#aaae8e", "#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file)
s3$upload_file(temp_file, "attempt", "Figures/attempt_general_cellcounts_proportions_plot_clean.jpeg") 

attempt_meta %>%
  filter(KPMP_celltype %in% c("PT-S1/S2", "PT-S3", "aPT")) %>%
  ggplot(aes(KPMP_celltype, fill = visit_treatment)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("PRE Placebo" = "#fbc4ab", "POST Placebo" ="#f4978e",
                               "PRE Dapagliflozin"= "#aaae8e", "POST Dapagliflozin" ="#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file)
s3$upload_file(temp_file, "attempt", "Figures/attempt_kpmp_pt_proportions_plot.jpeg") 

attempt_meta$kpmp_tal <- case_when(grepl("TAL", attempt_so$KPMP_celltype) ~ attempt_so$KPMP_celltype)
attempt_meta$kpmp_tal <- factor(attempt_meta$kpmp_tal,
                                levels = c("aTAL", "dTAL", "C-TAL-1", "C-TAL-2"))
attempt_meta %>%
  filter(!is.na(attempt_meta$kpmp_tal)) %>%
  ggplot(aes(kpmp_tal, fill = visit_treatment)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("PRE Placebo" = "#fbc4ab", "POST Placebo" ="#f4978e",
                               "PRE Dapagliflozin"= "#aaae8e", "POST Dapagliflozin" ="#828e82"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file)
s3$upload_file(temp_file, "attempt", "Figures/attempt_kpmp_tal_proportions_plot.jpeg") 
```



```{r echo = F}
attempt_so <- FindVariableFeatures(attempt_so, selection.method = "vst", nfeatures = 2000)
hvgs <- VariableFeatures(attempt_so)

#Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg <- subset(attempt_so, features = hvgs)
```

# Descriptive

```{r echo = F}
unique_attempt_meta <- attempt_meta %>%
  dplyr::select(Subject.ID, Visit, age, Sex, treatment_arm , visit_treatment, diabetes_duration) %>%
  dplyr::mutate(Subject.ID = as.character(Subject.ID, Visit)) %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(Subject.ID, Visit))
summary(arsenal::tableby(visit_treatment ~ age + Sex + diabetes_duration + treatment_arm + Visit, data = unique_attempt_meta), test = F, digits = 1)
```


# PT
```{r echo = F}
attempt_so$celltype_pt <- ifelse(grepl("PT-", attempt_so$celltype),
                                 "PT", as.character(attempt_so$celltype))
attempt_so_pt <- subset(attempt_so, celltype_pt == "PT" & celltype != "PT_lowQuality")
```

```{r echo = F}
attempt_so_pt <- FindVariableFeatures(attempt_so_pt, selection.method = "vst", nfeatures = 2000)
hvgs_pt <- VariableFeatures(attempt_so_pt)

#Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_pt <- subset(attempt_so_pt, features = hvgs_pt)
```

#### Proportions
```{r echo = F, eval = F}
attempt_pt_meta <- attempt_so_hvg_pt@meta.data %>%
  mutate(visit_treatment = paste0(visit, " ", treatment))

attempt_pt_meta$celltype <- factor(attempt_pt_meta$celltype, levels = c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))
attempt_pt_meta$visit_treatment <- factor(attempt_pt_meta$visit_treatment,
                                          levels = c("PRE Placebo", "POST Placebo",
                                                     "PRE Dapagliflozin", "POST Dapagliflozin"))
attempt_pt_meta %>%
  filter(celltype != "PT_lowQuality") %>%
  ggplot(aes(visit, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pt_visit_proportions_plot.jpeg")

attempt_pt_meta %>%
  filter(celltype != "PT_lowQuality") %>%
  ggplot(aes(treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pt_treatment_proportions_plot.jpeg")


attempt_pt_meta %>%
  filter(celltype != "PT_lowQuality") %>%
  ggplot(aes(visit_treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pt_treatment_visit_proportions_plot.jpeg")

attempt_meta %>%
  filter(KPMP_celltype %in% c("PT-S1/S2", "PT-S3", "aPT")) %>%
  ggplot(aes(visit, fill = KPMP_celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pt_visit_proportions_plot_kpmp.jpeg")

attempt_meta %>%
  filter(KPMP_celltype %in% c("PT-S1/S2", "PT-S3", "aPT")) %>%
  ggplot(aes(visit_treatment, fill = KPMP_celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pt_visit_treatment_proportions_plot_kpmp.jpeg")
```

```{r echo = F}
# test on proportions using Dirichlet
pt_props <- attempt_meta %>%
  filter(celltype_general %in% c("PT")) %>%
  group_by(subject_id, treatment, visit, celltype) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(subject_id) %>%
  mutate(prop = n / sum(n)) %>%
  dplyr::select(subject_id, treatment, celltype, visit, prop) %>%
  pivot_wider(names_from = celltype, values_from = prop, values_fill = 0)

pt_props <- pt_props %>%
  mutate(prop_sum = rowSums(across(c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))))
# Placebo
pt_props_placebo <- pt_props[pt_props$treatment == "Placebo",]
pt_props_placebo$Y <- DR_data(pt_props_placebo[,4:8])
pt_props_placebo_drichlet <- DirichReg(Y ~ visit, data = pt_props_placebo)
pt_props_placebo_drichlet_summary <- summary(pt_props_placebo_drichlet)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(pt_props_placebo_drichlet_summary, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_dirichlet_placebo.rds')

# Dapa
pt_props_dapa <- pt_props[pt_props$treatment != "Placebo",]
pt_props_dapa$Y <- DR_data(pt_props_dapa[,4:8])
pt_props_dapa_drichlet <- DirichReg(Y ~ visit, data = pt_props_dapa)
pt_props_dapa_drichlet_summary <- summary(pt_props_dapa_drichlet)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(pt_props_dapa_drichlet_summary, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_dirichlet_dapa.rds')

# POST (placebo vs. dapa)
pt_props_post <- pt_props[pt_props$visit == "POST",]
pt_props_post$Y <- DR_data(pt_props_post[,4:8])
pt_props_post_drichlet <- DirichReg(Y ~ treatment, data = pt_props_post)
pt_props_post_drichlet_summary <- summary(pt_props_post_drichlet)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(pt_props_post_drichlet_summary, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_dirichlet_post.rds')

```

```{r echo = F}
# test on proportions using Dirichlet (KPMP celltypes)
pt_props <- attempt_meta %>%
  filter(KPMP_celltype %in% c("PT-S1/S2", "PT-S3", "aPT")) %>%
  group_by(subject_id, treatment, visit, KPMP_celltype) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(subject_id) %>%
  mutate(prop = n / sum(n)) %>%
  dplyr::select(subject_id, treatment, KPMP_celltype, visit, prop) %>%
  pivot_wider(names_from = KPMP_celltype, values_from = prop, values_fill = 0)

pt_props <- pt_props %>%
  mutate(prop_sum = rowSums(across(c("PT-S1/S2", "PT-S3", "aPT"))))
# Placebo
pt_props_placebo <- pt_props[pt_props$treatment == "Placebo",]
pt_props_placebo$Y <- DR_data(pt_props_placebo[,4:6])
pt_props_placebo_drichlet <- DirichReg(Y ~ visit, data = pt_props_placebo)
pt_props_placebo_drichlet_summary <- summary(pt_props_placebo_drichlet)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(pt_props_placebo_drichlet_summary, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_dirichlet_placebo_kpmp.rds')

# Dapa
pt_props_dapa <- pt_props[pt_props$treatment != "Placebo",]
pt_props_dapa$Y <- DR_data(pt_props_dapa[,4:6])
pt_props_dapa_drichlet <- DirichReg(Y ~ visit, data = pt_props_dapa)
pt_props_dapa_drichlet_summary <- summary(pt_props_dapa_drichlet)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(pt_props_dapa_drichlet_summary, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_dirichlet_dapa_kpmp.rds')

# POST (placebo vs. dapa)
pt_props_post <- pt_props[pt_props$visit == "POST",]
pt_props_post$Y <- DR_data(pt_props_post[,4:6])
pt_props_post_drichlet <- DirichReg(Y ~ treatment, data = pt_props_post)
pt_props_post_drichlet_summary <- summary(pt_props_post_drichlet)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(pt_props_post_drichlet_summary, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_dirichlet_post_kpmp.rds')

```

```{r echo = F}
# scDC package (sc differential composition analysis)
attempt_so_pt$celltype <- factor(attempt_so_pt$celltype, levels = c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))
exprsMat <- GetAssayData(attempt_so_pt, assay = "RNA", slot = "counts")
attempt_so_pt$treatment_visit <- paste0(attempt_so_pt$treatment, "_", attempt_so_pt$visit)
attempt_so_pt$subject_visit <- paste0(attempt_so_pt$subject_id, "_", attempt_so_pt$visit)
cellTypes <- attempt_so_pt@meta.data$celltype
subject <- attempt_so_pt@meta.data$subject_visit

cond <- data.frame(subject = subject,
                   treatment = attempt_so_pt@meta.data$treatment_visit,
                   visit = attempt_so_pt@meta.data$visit) %>%
  distinct(subject, .keep_all = T)

res_scDC_noClust <- scDC_noClustering(cellTypes, subject, calCI = TRUE, 
                                     calCI_method = c("BCa"), ncores = 20)

# merge treatment and visit information to res_scDC_noClust
res_scDC_noClust$results <- left_join(res_scDC_noClust$results, cond)
```

```{r echo = F}
barplotCI_custom <- function(res, condition){
  res$results$index_original <- seq_len(nrow(res$results))
  df_toPlot <- res$results %>% 
    dplyr::mutate(cond = condition,
                  method = factor(method, levels = c("BCa", "percentile", "multinom"))) %>%
    dplyr::arrange(treatment, visit) 
  thetastar_reordered <- res$thetastar[df_toPlot$index_original, , drop = FALSE]
  
  df_toPlot$median <- apply(thetastar_reordered, 1, median)
  df_toPlot$conf_low <- apply(thetastar_reordered, 1, function(x) quantile(x, 0.025))
  df_toPlot$conf_high <- apply(thetastar_reordered, 1, function(x) quantile(x, 0.975))
  
  df_toPlot$subject <- factor(df_toPlot$subject, levels = unique(df_toPlot$subject))
  
  n_method <- length(unique(df_toPlot$method))
  n_celltype <- length(unique(df_toPlot$cellTypes))
  
  g_bar <- ggplot2::ggplot(df_toPlot, aes(x = subject, y = median, fill = cond)) +
    ggplot2::geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
    ggplot2::theme_bw() +
    ggplot2::scale_fill_brewer(palette = "Set2") +
    ggplot2::ylab("Proportion") +
    ggplot2::xlab("") +
    ggplot2::labs(fill = "") +
    ggplot2::geom_errorbar(aes(ymin = conf_low, ymax = conf_high),
                           width = 0.01, lwd = 1,
                           position = position_dodge(width = 0.5)) +
    ggplot2::theme(axis.text.x = element_text(angle = 90), text = element_text(size = 12)) +
    ggplot2::scale_color_manual(values = RColorBrewer::brewer.pal(n = 3, name = "Dark2")[1:n_method]) +
    ggplot2::facet_wrap(~cellTypes, ncol = n_celltype,
                        labeller = labeller(cellTypes = label_wrap_gen(width = 10, multi_line = TRUE))) +
    ggplot2::coord_flip() +
    ggplot2::ylim(c(0, 1))

  return(g_bar)
}

densityCI_custom <- function(res, condition){
  # Add index and reorder
  res$results$index_original <- seq_len(nrow(res$results))
  res$results <- res$results %>%
    dplyr::arrange(treatment, visit, subject)
  
  # Reorder thetastar
  thetastar_reordered <- res$thetastar[res$results$index_original, , drop = FALSE]

  # Melt the reordered thetastar
  df_toPlot <- reshape2::melt(thetastar_reordered)
  df_toPlot$cellTypes <- res$info[,"cellTypes"][df_toPlot$Var1]
  df_toPlot$subject <- res$info[,"subject"][df_toPlot$Var1]
  df_toPlot$cond <- condition

  # Match subject order to reordered results
  subject_order <- res$results$subject
  df_toPlot$subject <- factor(df_toPlot$subject, levels = unique(subject_order))
  
  # Prepare confidence line data
  conf_line <- res$results
  conf_line$cond <- condition
  conf_line$method <- factor(conf_line$method, levels = c("BCa", "percentile", "multinom"))
  
  n_method <- length(unique(conf_line$method))
  n_celltype <- length(unique(df_toPlot$cellTypes))
  
  # Plot
  if (length(unique(df_toPlot$subject)) == 1){
    g_density <- ggplot2::ggplot(df_toPlot, aes(x = value, y = subject, fill = cond)) +
      ggridges::stat_density_ridges(alpha = 0.5) +
      ggplot2::geom_vline(data = conf_line, aes(xintercept = conf_low),
                          lwd = 1, alpha = 0.8) +
      ggplot2::geom_vline(data = conf_line, aes(xintercept = conf_high),
                          lwd = 1, alpha = 0.8)
  } else {
    g_density <- ggplot2::ggplot(df_toPlot, aes(x = value, y = subject, fill = cond)) +
      ggridges::stat_density_ridges(alpha = 0.5) +
      ggplot2::geom_segment(data = conf_line, aes(x = conf_low, xend = conf_low, y = as.numeric(subject),
                                                  yend = as.numeric(subject) + .9,
                                                  color = method, linetype = method),
                            lwd = 1, alpha = 0.8) +
      ggplot2::geom_segment(data = conf_line, aes(x = conf_high, xend = conf_high, y = as.numeric(subject),
                                                  yend = as.numeric(subject) + .9,
                                                  color = method, linetype = method),
                            lwd = 1, alpha = 0.8)
  }

  g_density <- g_density +
    ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x = element_text(angle = 90), text = element_text(size = 12)) +
    ggplot2::scale_fill_brewer(palette = "Set2") +
    ggplot2::xlab("Proportion") +
    ggplot2::ylab("") +
    ggplot2::labs(fill = "") +
    ggplot2::guides(color = "none") +
    ggplot2::facet_wrap(~cellTypes, ncol = n_celltype , scales = "free_x",
                        labeller = labeller(cellTypes  = label_wrap_gen(width = 10,  multi_line = TRUE)))

  return(g_density)
}

barplotCI_custom(res_scDC_noClust, res_scDC_noClust$results$treatment)

densityCI_custom(res_scDC_noClust, res_scDC_noClust$results$treatment)

```


```{r echo = F}
# lots of non-convergence, long time to compute, and need contrasts to do comparisons between other PT cells other than the ref group (PT-1)
# # GLM in Placebo only
# attempt_so_pt_placebo <- subset(attempt_so_pt, treatment == "Placebo")
# attempt_so_pt_placebo$celltype <- factor(attempt_so_pt_placebo$celltype, levels = c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))
# exprsMat <- GetAssayData(attempt_so_pt_placebo, assay = "RNA", slot = "counts")
# attempt_so_pt_placebo$treatment_visit <- paste0(attempt_so_pt_placebo$treatment, "_", attempt_so_pt_placebo$visit)
# attempt_so_pt_placebo$subject_visit <- paste0(attempt_so_pt_placebo$subject_id, "_", attempt_so_pt_placebo$visit)
# cellTypes <- attempt_so_pt_placebo@meta.data$celltype
# subject <- attempt_so_pt_placebo@meta.data$subject_visit
# 
# cond <- data.frame(subject = subject,
#                    treatment = attempt_so_pt_placebo@meta.data$treatment_visit,
#                    visit = attempt_so_pt_placebo@meta.data$visit) %>%
#   distinct(subject, .keep_all = T)
# 
# res_scDC_placebo <- scDC_noClustering(cellTypes, subject, calCI = TRUE, 
#                                      calCI_method = c("BCa"), ncores = 20, nboot = 1000)
# 
# res_scDC_placebo$results$subject <- sub("_.*", "", res_scDC_placebo$results$subject)
# res_scDC_placebo$info$subject <- sub("_.*", "", res_scDC_placebo$info$subject)
# cond$subject <- sub("_.*", "", cond$subject)
# 
# 
# res_GLM_placebo <- fitGLM(res_scDC_placebo, cond$visit, 
#                   pairwise = T)
# 
# summary(res_GLM_placebo$pool_res_random)
```

```{r echo = F}
# # GLM in Dapa only
# attempt_so_pt_dapa <- subset(attempt_so_pt, treatment != "Placebo")
# attempt_so_pt_dapa$celltype <- factor(attempt_so_pt_dapa$celltype, levels = c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))
# exprsMat <- GetAssayData(attempt_so_pt_dapa, assay = "RNA", slot = "counts")
# attempt_so_pt_dapa$treatment_visit <- paste0(attempt_so_pt_dapa$treatment, "_", attempt_so_pt_dapa$visit)
# attempt_so_pt_dapa$subject_visit <- paste0(attempt_so_pt_dapa$subject_id, "_", attempt_so_pt_dapa$visit)
# cellTypes <- attempt_so_pt_dapa@meta.data$celltype
# subject <- attempt_so_pt_dapa@meta.data$subject_visit
# 
# cond <- data.frame(subject = subject,
#                    treatment = attempt_so_pt_dapa@meta.data$treatment_visit,
#                    visit = attempt_so_pt_dapa@meta.data$visit) %>%
#   distinct(subject, .keep_all = T)
# 
# res_scDC_dapa <- scDC_noClustering(cellTypes, subject, calCI = TRUE, 
#                                      calCI_method = c("BCa"), ncores = 20, nboot = 1000)
# 
# res_scDC_dapa$results$subject <- sub("_.*", "", res_scDC_dapa$results$subject)
# res_scDC_dapa$info$subject <- sub("_.*", "", res_scDC_dapa$info$subject)
# cond$subject <- sub("_.*", "", cond$subject)
# 
# 
# res_GLM_dapa <- fitGLM(res_scDC_dapa, cond$visit, 
#                   pairwise = T)
# 
# summary(res_GLM_dapa$pool_res_random)
# summary(res_GLM_dapa$pool_res_fixed)
```


```{r echo = F}
# Method using glmmTMB
# library(glmmTMB)
# 
# # Placebo only
# attempt_so_pt_meta_placebo <- attempt_so_pt_placebo@meta.data
# # Loop through PT subtypes
# pt_subtypes <- paste0("PT-", 1:5)
# 
# pt_prop_placebo <- lapply(pt_subtypes, function(pt) {
#   attempt_so_pt_meta_placebo$binary <- as.integer(attempt_so_pt_meta_placebo$celltype == pt)
# 
#   model <- glmmTMB(binary ~ visit + (1 | subject_id), 
#                    family = binomial(), 
#                    data = attempt_so_pt_meta_placebo)
#   
#   summary_out <- summary(model)
#   est <- summary_out$coefficients$cond["visitPOST", ]
#   data.frame(
#     subtype = pt,
#     logOR = est["Estimate"],
#     SE = est["Std. Error"],
#     p_value = est["Pr(>|z|)"]
#   )
# }) %>% bind_rows()
# 
# print(pt_prop_placebo)
```


```{r echo = F}
# attempt_so_pt_meta_dapa <- attempt_so_pt_dapa@meta.data
# # Loop through PT subtypes
# pt_subtypes <- paste0("PT-", 1:5)
# 
# pt_prop_dapa <- lapply(pt_subtypes, function(pt) {
#   attempt_so_pt_meta_dapa$binary <- as.integer(attempt_so_pt_meta_dapa$celltype == pt)
# 
#   model <- glmmTMB(binary ~ visit + (1 | subject_id), 
#                    family = binomial(), 
#                    data = attempt_so_pt_meta_dapa)
#   
#   summary_out <- summary(model)
#   est <- summary_out$coefficients$cond["visitPOST", ]
#   data.frame(
#     subtype = pt,
#     logOR = est["Estimate"],
#     SE = est["Std. Error"],
#     p_value = est["Pr(>|z|)"]
#   )
# }) %>% bind_rows()
# 
# print(pt_prop_dapa)

```

```{r echo = F}
# using propeller
library(speckle)

# Placebo

attempt_so_pt_placebo <- subset(attempt_so_pt, treatment == "Placebo")
attempt_so_pt_placebo$celltype <- factor(attempt_so_pt_placebo$celltype, levels = c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))
attempt_so_pt_meta_placebo <- attempt_so_pt_placebo@meta.data

transformed_props <- getTransformedProps(clusters = attempt_so_pt_meta_placebo$celltype,
                    sample = paste0(attempt_so_pt_meta_placebo$subject_id, "_", attempt_so_pt_meta_placebo$visit))
visit <- ifelse(grepl("_PRE$", colnames(transformed_props$Counts)), "PRE", "POST")
design <- model.matrix(~ 0 + visit)
colnames(design) <- gsub("visit", "", colnames(design)) 

contrasts <- c(POST = 1, PRE = -1)

propeller_res_placebo <- propeller.ttest(prop.list = transformed_props,
                                 design = design,
                                 contrasts = contrasts,
                                 robust=TRUE,
  trend=FALSE, sort=TRUE)
propeller_res_placebo

# Dapagliflozin
attempt_so_pt_dapa <- subset(attempt_so_pt, treatment == "Dapagliflozin")
attempt_so_pt_dapa$celltype <- factor(attempt_so_pt_dapa$celltype, levels = c("PT-1", "PT-2", "PT-3", "PT-4", "PT-5"))
attempt_so_pt_meta_dapa <- attempt_so_pt_dapa@meta.data

transformed_props <- getTransformedProps(clusters = attempt_so_pt_meta_dapa$celltype,
                    sample = paste0(attempt_so_pt_meta_dapa$subject_id, "_", attempt_so_pt_meta_dapa$visit))
visit <- ifelse(grepl("_PRE$", colnames(transformed_props$Counts)), "PRE", "POST")
design <- model.matrix(~ 0 + visit)
colnames(design) <- gsub("visit", "", colnames(design)) 

contrasts <- c(POST = 1, PRE = -1)

propeller_res_dapa <- propeller.ttest(prop.list = transformed_props,
                                 design = design,
                                 contrasts = contrasts,
                                 robust=TRUE,
  trend=FALSE, sort=TRUE)
propeller_res_dapa

propeller_res_placebo$treatment <- "Placebo"
propeller_res_placebo$celltype <- rownames(propeller_res_placebo)
propeller_res_dapa$treatment <- "Dapagliflozin"
propeller_res_dapa$celltype <- rownames(propeller_res_dapa)

combined <- rbind(propeller_res_placebo, propeller_res_dapa)

combined
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(combined, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_propeller_combined.rds')
```

```{r echo = F}
# using propeller (KPMP celltype)
# Placebo

attempt_so_pt_placebo <- subset(attempt_so_pt, treatment == "Placebo")
attempt_so_pt_placebo$KPMP_celltype <- factor(attempt_so_pt_placebo$KPMP_celltype, levels = c("PT-S1/S2", "PT-S3", "aPT"))
attempt_so_pt_meta_placebo <- attempt_so_pt_placebo@meta.data

transformed_props <- getTransformedProps(clusters = attempt_so_pt_meta_placebo$KPMP_celltype,
                    sample = paste0(attempt_so_pt_meta_placebo$subject_id, "_", attempt_so_pt_meta_placebo$visit))
visit <- ifelse(grepl("_PRE$", colnames(transformed_props$Counts)), "PRE", "POST")
design <- model.matrix(~ 0 + visit)
colnames(design) <- gsub("visit", "", colnames(design)) 

contrasts <- c(POST = 1, PRE = -1)

propeller_res_placebo <- propeller.ttest(prop.list = transformed_props,
                                 design = design,
                                 contrasts = contrasts,
                                 robust=TRUE,
  trend=FALSE, sort=TRUE)
propeller_res_placebo

# Dapagliflozin
attempt_so_pt_dapa <- subset(attempt_so_pt, treatment == "Dapagliflozin")
attempt_so_pt_dapa$KPMP_celltype <- factor(attempt_so_pt_dapa$KPMP_celltype, levels = c("PT-S1/S2", "PT-S3", "aPT"))
attempt_so_pt_meta_dapa <- attempt_so_pt_dapa@meta.data

transformed_props <- getTransformedProps(clusters = attempt_so_pt_meta_dapa$KPMP_celltype,
                    sample = paste0(attempt_so_pt_meta_dapa$subject_id, "_", attempt_so_pt_meta_dapa$visit))
visit <- ifelse(grepl("_PRE$", colnames(transformed_props$Counts)), "PRE", "POST")
design <- model.matrix(~ 0 + visit)
colnames(design) <- gsub("visit", "", colnames(design)) 

contrasts <- c(POST = 1, PRE = -1)

propeller_res_dapa <- propeller.ttest(prop.list = transformed_props,
                                 design = design,
                                 contrasts = contrasts,
                                 robust=TRUE,
  trend=FALSE, sort=TRUE)
propeller_res_dapa

propeller_res_placebo$treatment <- "Placebo"
propeller_res_placebo$KPMP_celltype <- rownames(propeller_res_placebo)
propeller_res_dapa$treatment <- "Dapagliflozin"
propeller_res_dapa$KPMP_celltype <- rownames(propeller_res_dapa)

combined_kpmp <- rbind(propeller_res_placebo, propeller_res_dapa)

temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(combined_kpmp, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_propeller_combined_kpmp.rds')
```

```{r echo = F}
# Mann whitney u testing
# Define function to run Mann-Whitney + effect size pipeline
run_mwu_effect_size <- function(meta_df, group_label, celltype = "celltype") {
  
  celltype_sym <- rlang::sym(celltype)   # create symbol from string
  
  # Count cells per sample and per PT subtype
  prop_df <- meta_df %>%
    group_by(subject_id, visit, sample_id = paste0(subject_id, "_", visit)) %>%
    dplyr::count(!!celltype_sym) %>%
    group_by(subject_id, visit, sample_id) %>%
    mutate(proportion = n / sum(n)) %>%
    ungroup()
  
  # Make wide format
  prop_wide <- prop_df %>%
    dplyr::select(sample_id, visit, !!celltype_sym, proportion) %>%
    tidyr::pivot_wider(names_from = !!celltype_sym, values_from = proportion, values_fill = 0)
  
  # Get list of PT subtypes
  celltypes <- setdiff(colnames(prop_wide), c("sample_id", "visit"))
  
  # Run Mann-Whitney + effect size per subtype
  effect_results <- lapply(celltypes, function(ct) {
    
    test_res <- wilcox.test(prop_wide[[ct]] ~ prop_wide$visit)
    
    # Effect direction → median difference
    medians <- prop_wide %>%
      group_by(visit) %>%
      summarise(median_prop = median(.data[[ct]], na.rm = TRUE)) %>%
      arrange(visit)
    
    median_diff <- medians$median_prop[medians$visit == "POST"] - 
                   medians$median_prop[medians$visit == "PRE"]
    
    # Rank-biserial correlation
    n1 <- sum(prop_wide$visit == "PRE")
    n2 <- sum(prop_wide$visit == "POST")
    U <- test_res$statistic
    rbc <- (2 * U) / (n1 * n2) - 1
    
    tibble(
      celltype = ct,
      p.value = test_res$p.value,
      W = test_res$statistic,
      median_PRE = medians$median_prop[medians$visit == "PRE"],
      median_POST = medians$median_prop[medians$visit == "POST"],
      median_diff = median_diff,
      rank_biserial_corr = rbc,
      group = group_label
    )
  })
  
  # Combine results
  bind_rows(effect_results)
}
```

```{r echo = F}
# For dapagliflozin
effect_results_dapa <- run_mwu_effect_size(attempt_so_pt_meta_dapa, "dapa")

# For placebo (replace this with your placebo meta dataframe)
effect_results_placebo <- run_mwu_effect_size(attempt_so_pt_meta_placebo, "placebo")

# Combine both
combined_effect_results <- bind_rows(effect_results_dapa, effect_results_placebo)

# Save as RDS
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(combined_effect_results, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_mwu_combined.rds')
```

```{r echo = F}
# For dapagliflozin
effect_results_dapa_kpmp <- run_mwu_effect_size(attempt_so_pt_meta_dapa, "dapa", celltype = "KPMP_celltype")

# For placebo (replace this with your placebo meta dataframe)
effect_results_placebo_kpmp <- run_mwu_effect_size(attempt_so_pt_meta_placebo, "placebo", celltype = "KPMP_celltype")

# Combine both
combined_effect_results_kpmp <- bind_rows(effect_results_dapa_kpmp, effect_results_placebo_kpmp)

# Save as RDS
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(combined_effect_results_kpmp, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/pt_mwu_combined_kpmp.rds')
```

```{r echo = F}
count_df_dapa <- attempt_so_pt_meta_dapa %>%
  group_by(subject_id, visit, sample_id = paste0(subject_id, "_", visit)) %>%
  dplyr::count(celltype) %>%
  tidyr::pivot_wider(names_from = celltype, values_from = n, values_fill = 0) %>%
  ungroup()
count_df_dapa_py <- reticulate::r_to_py(count_df_dapa)

count_df_placebo <- attempt_so_pt_meta_placebo %>%
  group_by(subject_id, visit, sample_id = paste0(subject_id, "_", visit)) %>%
  dplyr::count(celltype) %>%
  tidyr::pivot_wider(names_from = celltype, values_from = n, values_fill = 0) %>%
  ungroup()
count_df_placebo_py <- reticulate::r_to_py(count_df_placebo)
reticulate::repl_python()
```

```{python}
import pandas as pd
import numpy as np
import sccoda
from sccoda.util import cell_composition_data as dat
from sccoda.util import data_visualization as viz

# Load data
count_df_placebo = r.count_df_placebo
count_df_dapa = r.count_df_dapa

# Convert data to anndata object
print(count_df_placebo)
data_all = dat.from_pandas(count_df_placebo, covariate_columns=["sample_id"])

# Extract condition from mouse name and add it as an extra column to the covariates
data_all.obs["visit"] = data_all.obs["subject_id"].str.replace(r"_[0-9]", "", regex=True)
print(data_all)
```

#### Mixed model (hurdle, no nested random effect)

##### nebula
```{r echo = F, eval = F}
# nebula built-in testing
# data(sample_data)
# pred = model.matrix(~X1+X2+cc,data=sample_data$pred)
# re = nebula(count=sample_data$count,id=sample_data$sid,pred=pred, , output_re = T, covariance = T)
```


###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_pt <- round(GetAssayData(attempt_so_hvg_pt, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_pt)

# meta_hvg_pt <- attempt_so_hvg_pt@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_pt)
# data_g_hvg_pt = group_cell(count=counts_hvg_pt, id=meta_hvg_pt$subject, pred=pred)
# offset_hvg_pt = Matrix::colSums(data_g_hvg_pt$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()


nebula_pt_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_pt[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_pt, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_pt_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_pt_results_list_reml_pooled) <- sapply(nebula_pt_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_pt_results_list_reml_pooled <- lapply(nebula_pt_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_pt_results_list_reml_pooled <- Filter(Negate(is.null), nebula_pt_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_pt_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_pt_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PT/nebula/pt_attempt_hvg_nebula_res_reml_pooled.rds')
```

```{r echo = F}
# PT subtypes
celltype_vec <- paste0("PT-", 1:5)

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_pt, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("PT/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```

#### Average expression (pseudo-bulk within individual)

#### Sum expression (pseudo-bulk within individual)

* # of genes up/down
* upset plots of overlap in cell types
* pathways (IPA and GSEA - GO/KEGG/one that Long used from Broad)
* pseudotime (call with Kelly)
* BOLD MRI stuff in the PPT so that we show everything in one slide

# TAL

```{r echo = F}
attempt_so$celltype_tal <- ifelse(grepl("TAL-", attempt_so$celltype),
                                 "TAL", as.character(attempt_so$celltype))
attempt_so_tal <- subset(attempt_so, celltype_tal == "TAL" & celltype != "TAL_highUMI")
```

```{r echo = F}
attempt_so_tal <- FindVariableFeatures(attempt_so_tal, selection.method = "vst", nfeatures = 2000)
hvgs_tal <- VariableFeatures(attempt_so_tal)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_tal <- subset(attempt_so_tal, features = hvgs_tal)
```

#### Proportions
```{r echo = F, eval = F}
attempt_tal_meta <- attempt_so_hvg_tal@meta.data %>%
  mutate(visit_treatment = paste0(visit, " ", treatment))
attempt_tal_meta$celltype <- factor(attempt_tal_meta$celltype, levels = c("TAL-1", "TAL-2", "TAL-3"))
attempt_tal_meta$visit_treatment <- factor(attempt_tal_meta$visit_treatment,
                                          levels = c("PRE Placebo", "POST Placebo",
                                                     "PRE Dapagliflozin", "POST Dapagliflozin"))
attempt_tal_meta %>%
  filter(celltype != "TAL_highUMI") %>%
  ggplot(aes(visit, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_tal_treatment_visit_proportions_plot.jpeg")

attempt_tal_meta %>%
  filter(celltype != "tal_lowQuality") %>%
  ggplot(aes(treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_tal_treatment_proportions_plot.jpeg")

attempt_tal_meta %>%
  filter(celltype != "tal_lowQuality") %>%
  ggplot(aes(visit_treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_tal_treatment_visit_proportions_plot.jpeg")

attempt_meta %>%
  filter(grepl("TAL", KPMP_celltype)) %>%
  ggplot(aes(visit, fill = kpmp_tal)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_tal_visit_proportions_plot_kpmp.jpeg")

attempt_meta %>%
  filter(grepl("TAL", KPMP_celltype)) %>%
  ggplot(aes(treatment, fill = kpmp_tal)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_tal_treatment_proportions_plot_kpmp.jpeg")

attempt_meta %>%
  filter(grepl("TAL", KPMP_celltype)) %>%
  ggplot(aes(visit_treatment, fill = kpmp_tal)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_tal_visit_treatment_proportions_plot_kpmp.jpeg")
```

#### Mixed model (hurdle, nested random effect)
##### nebula

###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_tal <- round(GetAssayData(attempt_so_hvg_tal, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_tal)

# meta_hvg_tal <- attempt_so_hvg_tal@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_tal)
# data_g_hvg_tal = group_cell(count=counts_hvg_tal, id=meta_hvg_tal$subject, pred=pred)
# offset_hvg_tal = Matrix::colSums(data_g_hvg_tal$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_tal_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_tal[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_tal, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_tal_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_tal_results_list_reml_pooled) <- sapply(nebula_tal_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_tal_results_list_reml_pooled <- lapply(nebula_tal_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_tal_results_list_reml_pooled <- Filter(Negate(is.null), nebula_tal_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_tal_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_tal_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'TAL/nebula/tal_attempt_hvg_nebula_res_reml_pooled.rds')
```

```{r echo = F}
# TAL subtypes
celltype_vec <- paste0("TAL-", 1:3)

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_tal, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("TAL/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```


#### Average expression (pseudo-bulk within individual)

#### Sum expression (pseudo-bulk within individual)

# Immune Cells

```{r echo = F}
# immune cells: MAC, MON, T, B, NKT/NKC
attempt_so$celltype_immune <- ifelse(attempt_so$celltype %in% c("MON", "MAC", "T", "B", "NKT/NKC"),
                                     "Immune", as.character(attempt_so$celltype))
attempt_so_immune <- subset(attempt_so, celltype_immune == "Immune")
```

```{r echo = F}
attempt_so_immune <- FindVariableFeatures(attempt_so_immune, selection.method = "vst", nfeatures = 2000)
hvgs_immune <- VariableFeatures(attempt_so_immune)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_immune <- subset(attempt_so_immune, features = hvgs_immune)
```


#### Proportions
```{r echo = F, eval = F}
attempt_immune_meta <- attempt_so_hvg_immune@meta.data %>%
  mutate(visit_treatment = paste0(visit, " ", treatment))
attempt_immune_meta$celltype <- factor(attempt_immune_meta$celltype, levels = c("MON", "MAC", "T", "B", "NKT/NKC"))
attempt_immune_meta$visit_treatment <- factor(attempt_immune_meta$visit_treatment,
                                          levels = c("PRE Placebo", "POST Placebo",
                                                     "PRE Dapagliflozin", "POST Dapagliflozin"))
attempt_immune_meta %>%
  ggplot(aes(visit, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_immune_treatment_visit_proportions_plot.jpeg")

attempt_immune_meta %>%
  ggplot(aes(treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_immune_treatment_proportions_plot.jpeg")

attempt_immune_meta %>%
  ggplot(aes(visit_treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_immune_treatment_visit_proportions_plot.jpeg")
```

#### Mixed model (hurdle, nested random effect)
##### nebula

###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_immune <- round(GetAssayData(attempt_so_hvg_immune, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_immune)

# meta_hvg_immune <- attempt_so_hvg_immune@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_immune)
# data_g_hvg_immune = group_cell(count=counts_hvg_immune, id=meta_hvg_immune$subject, pred=pred)
# offset_hvg_immune = Matrix::colSums(data_g_hvg_immune$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_immune_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_immune[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_immune, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_immune_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_immune_results_list_reml_pooled) <- sapply(nebula_immune_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_immune_results_list_reml_pooled <- lapply(nebula_immune_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_immune_results_list_reml_pooled <- Filter(Negate(is.null), nebula_immune_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_immune_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_immune_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'Immune/nebula/immune_attempt_hvg_nebula_res_reml_pooled.rds')
```


#### Average expression (pseudo-bulk within individual) 

#### Sum expression (pseudo-bulk within individual) 


# Immune Cells (Myeloid)

```{r echo = F}
# immune_myeloid cells: MAC, MON
attempt_so$celltype_immune_myeloid <- ifelse(attempt_so$celltype %in% c("MON", "MAC"),
                                     "immune_myeloid", as.character(attempt_so$celltype))
attempt_so_immune_myeloid <- subset(attempt_so, celltype_immune_myeloid == "immune_myeloid")
```

```{r echo = F}
attempt_so_immune_myeloid <- FindVariableFeatures(attempt_so_immune_myeloid, selection.method = "vst", nfeatures = 2000)
hvgs_immune_myeloid <- VariableFeatures(attempt_so_immune_myeloid)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_immune_myeloid <- subset(attempt_so_immune_myeloid, features = hvgs_immune_myeloid)
```

#### Mixed model (hurdle, nested random effect)
##### nebula

###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_immune_myeloid <- round(GetAssayData(attempt_so_hvg_immune_myeloid, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_immune_myeloid)

# meta_hvg_immune_myeloid <- attempt_so_hvg_immune_myeloid@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_immune_myeloid)
# data_g_hvg_immune_myeloid = group_cell(count=counts_hvg_immune_myeloid, id=meta_hvg_immune_myeloid$subject, pred=pred)
# offset_hvg_immune_myeloid = Matrix::colSums(data_g_hvg_immune_myeloid$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_immune_myeloid_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_immune_myeloid[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_immune_myeloid, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_immune_myeloid_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_immune_myeloid_results_list_reml_pooled) <- sapply(nebula_immune_myeloid_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_immune_myeloid_results_list_reml_pooled <- lapply(nebula_immune_myeloid_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_immune_myeloid_results_list_reml_pooled <- Filter(Negate(is.null), nebula_immune_myeloid_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_immune_myeloid_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_immune_myeloid_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'Immune_Myeloid/nebula/immune_myeloid_attempt_hvg_nebula_res_reml_pooled.rds')
```


```{r echo = F}
# Immune Myeloid subtypes
celltype_vec <- c("MON", "MAC")

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_immune_myeloid, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("Immune_Myeloid/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```


# Immune Cells (Lymphoid)

```{r echo = F}
# immune_lymphoid cells: "T", "B", "NKT/NKC"
attempt_so$celltype_immune_lymphoid <- ifelse(attempt_so$celltype %in% c("T", "B", "NKT/NKC"),
                                     "immune_lymphoid", as.character(attempt_so$celltype))
attempt_so_immune_lymphoid <- subset(attempt_so, celltype_immune_lymphoid == "immune_lymphoid")
```

```{r echo = F}
attempt_so_immune_lymphoid <- FindVariableFeatures(attempt_so_immune_lymphoid, selection.method = "vst", nfeatures = 2000)
hvgs_immune_lymphoid <- VariableFeatures(attempt_so_immune_lymphoid)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_immune_lymphoid <- subset(attempt_so_immune_lymphoid, features = hvgs_immune_lymphoid)
```

#### Mixed model (hurdle, nested random effect)
##### nebula

###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_immune_lymphoid <- round(GetAssayData(attempt_so_hvg_immune_lymphoid, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_immune_lymphoid)

# meta_hvg_immune_lymphoid <- attempt_so_hvg_immune_lymphoid@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_immune_lymphoid)
# data_g_hvg_immune_lymphoid = group_cell(count=counts_hvg_immune_lymphoid, id=meta_hvg_immune_lymphoid$subject, pred=pred)
# offset_hvg_immune_lymphoid = Matrix::colSums(data_g_hvg_immune_lymphoid$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_immune_lymphoid_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_immune_lymphoid[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_immune_lymphoid, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_immune_lymphoid_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_immune_lymphoid_results_list_reml_pooled) <- sapply(nebula_immune_lymphoid_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_immune_lymphoid_results_list_reml_pooled <- lapply(nebula_immune_lymphoid_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_immune_lymphoid_results_list_reml_pooled <- Filter(Negate(is.null), nebula_immune_lymphoid_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_immune_lymphoid_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_immune_lymphoid_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'Immune_Lymphoid/nebula/immune_lymphoid_attempt_hvg_nebula_res_reml_pooled.rds')
```

```{r echo = F}
# Immune Lymphoid subtypes
celltype_vec <- c("T", "B", "NKT/NKC")

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_immune_lymphoid, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("Immune_Lymphoid/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```

# PC 

```{r echo = F} 
# PC-1 & PC-2, tPC-IC 
attempt_so$celltype_pc <- ifelse(attempt_so$celltype %in% c("PC-1", "PC-2", "tPC-IC"), 
                                     "PC", as.character(attempt_so$celltype)) 
attempt_so_pc <- subset(attempt_so, celltype_pc == "PC") 
``` 

```{r echo = F}
attempt_so_pc <- FindVariableFeatures(attempt_so_pc, selection.method = "vst", nfeatures = 2000)
hvgs_pc <- VariableFeatures(attempt_so_pc)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_pc <- subset(attempt_so_pc, features = hvgs_pc)
```

#### Proportions 
```{r echo = F, eval = F}
attempt_pc_meta <- attempt_so_hvg_pc@meta.data %>% 
  mutate(visit_treatment = paste0(visit, " ", treatment)) 
attempt_pc_meta$celltype <- factor(attempt_pc_meta$celltype, levels = c("PC-1", "PC-2", "tPC-IC")) 
attempt_pc_meta$visit_treatment <- factor(attempt_pc_meta$visit_treatment, 
                                          levels = c("PRE Placebo", "POST Placebo", 
                                                     "PRE Dapagliflozin", "POST Dapagliflozin")) 
attempt_pc_meta %>% 
  ggplot(aes(visit, fill = celltype)) + 
  geom_bar(stat = "count", position = "fill") + 
  theme_classic() + 
  labs(x = NULL, 
       y = "Proportion", 
       fill = "Cell type") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(plot.title = element_text(hjust=0.5, face="bold"), 
        text = element_text(size = 20)) + 
  ggtitle("Proportion of cells") + 
  scale_fill_manual(values = c("#264653", "#e9c46a", "#f4a261")) 
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pc_treatment_visit_proportions_plot.jpeg")

attempt_pc_meta %>% 
  ggplot(aes(treatment, fill = celltype)) + 
  geom_bar(stat = "count", position = "fill") + 
  theme_classic() + 
  labs(x = NULL, 
       y = "Proportion", 
       fill = "Cell type") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(plot.title = element_text(hjust=0.5, face="bold"), 
        text = element_text(size = 20)) + 
  ggtitle("Proportion of cells") + 
  scale_fill_manual(values = c("#264653", "#e9c46a", "#f4a261")) 
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pc_treatment_proportions_plot.jpeg")


attempt_pc_meta %>% 
  ggplot(aes(visit_treatment, fill = celltype)) + 
  geom_bar(stat = "count", position = "fill") + 
  theme_classic() + 
  labs(x = NULL, 
       y = "Proportion", 
       fill = "Cell type") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + 
  theme(plot.title = element_text(hjust=0.5, face="bold"), 
        text = element_text(size = 20)) + 
  ggtitle("Proportion of cells") + 
  scale_fill_manual(values = c("#264653", "#e9c46a", "#f4a261")) 
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_pc_treatment_visit_proportions_plot.jpeg")
``` 

#### Mixed model (hurdle, nested random effect) 
##### nebula

###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_pc <- round(GetAssayData(attempt_so_hvg_pc, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_pc)

# meta_hvg_pc <- attempt_so_hvg_pc@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_pc)
# data_g_hvg_pc = group_cell(count=counts_hvg_pc, id=meta_hvg_pc$subject, pred=pred)
# offset_hvg_pc = Matrix::colSums(data_g_hvg_pc$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_pc_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_pc[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_pc, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_pc_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_pc_results_list_reml_pooled) <- sapply(nebula_pc_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_pc_results_list_reml_pooled <- lapply(nebula_pc_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_pc_results_list_reml_pooled <- Filter(Negate(is.null), nebula_pc_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_pc_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_pc_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'PC/nebula/pc_attempt_hvg_nebula_res_reml_pooled.rds')
```


```{r echo = F}
# PC subtypes
celltype_vec <- c("tPC-IC", "PC-1", "PC-2")

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_pc, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("PC/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```


#### Average expression (pseudo-bulk within individual)

#### Sum expression (pseudo-bulk within individual)


# IC

```{r echo = F}
# IC-A, IC-B
attempt_so$celltype_ic <- ifelse(attempt_so$celltype %in% c("IC-A", "IC-B"),
                                     "IC", as.character(attempt_so$celltype))
attempt_so_ic <- subset(attempt_so, celltype_ic == "IC")
```

```{r echo = F}
attempt_so_ic <- FindVariableFeatures(attempt_so_ic, selection.method = "vst", nfeatures = 2000)
hvgs_ic <- VariableFeatures(attempt_so_ic)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_ic <- subset(attempt_so_ic, features = hvgs_ic)
```

#### Proportions
```{r echo = F, eval = F}
attempt_ic_meta <- attempt_so_hvg_ic@meta.data %>%
  mutate(visit_treatment = paste0(visit, " ", treatment))
attempt_ic_meta$celltype <- factor(attempt_ic_meta$celltype, levels = c("IC-A", "IC-B"))
attempt_ic_meta$visit_treatment <- factor(attempt_ic_meta$visit_treatment, 
                                          levels = c("PRE Placebo", "POST Placebo",
                                                     "PRE Dapagliflozin", "POST Dapagliflozin"))
attempt_ic_meta %>%
  ggplot(aes(visit, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_ic_treatment_visit_proportions_plot.jpeg")

attempt_ic_meta %>%
  ggplot(aes(treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_ic_treatment_proportions_plot.jpeg")

attempt_ic_meta %>%
  ggplot(aes(visit_treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_ic_treatment_visit_proportions_plot.jpeg")
```

#### Mixed model (hurdle, nested random effect)
##### nebula
###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_ic <- round(GetAssayData(attempt_so_hvg_ic, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_ic)

# meta_hvg_ic <- attempt_so_hvg_ic@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_ic)
# data_g_hvg_ic = group_cell(count=counts_hvg_ic, id=meta_hvg_ic$subject, pred=pred)
# offset_hvg_ic = Matrix::colSums(data_g_hvg_ic$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_ic_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_ic[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_ic, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_ic_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_ic_results_list_reml_pooled) <- sapply(nebula_ic_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_ic_results_list_reml_pooled <- lapply(nebula_ic_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_ic_results_list_reml_pooled <- Filter(Negate(is.null), nebula_ic_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_ic_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_ic_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'IC/nebula/ic_attempt_hvg_nebula_res_reml_pooled.rds')
```

```{r echo = F}
# IC subtypes
celltype_vec <- c("IC-A", "IC-B")

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_ic, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("IC/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```


#### Average expression (pseudo-bulk within individual)

#### Sum expression (pseudo-bulk within individual)

# EC

```{r echo = F}
attempt_so$celltype_ec <- ifelse(attempt_so$celltype %in% c("EC-PTC", "EC-GC", "EC-AEA", "EC-LYM"),
                                     "EC", as.character(attempt_so$celltype))
attempt_so_ec <- subset(attempt_so, celltype_ec == "EC")
```

```{r echo = F}
attempt_so_ec <- FindVariableFeatures(attempt_so_ec, selection.method = "vst", nfeatures = 2000)
hvgs_ec <- VariableFeatures(attempt_so_ec)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_ec <- subset(attempt_so_ec, features = hvgs_ec)
```

#### Proportions
```{r echo = F, eval = F}
attempt_ec_meta <- attempt_so_hvg_ec@meta.data %>%
  mutate(visit_treatment = paste0(visit, " ", treatment))
attempt_ec_meta$celltype <- factor(attempt_ec_meta$celltype, levels = c("EC-PTC", "EC-GC", "EC-AEA", "EC-LYM"))
attempt_ec_meta$visit_treatment <- factor(attempt_ec_meta$visit_treatment, 
                                          levels = c("PRE Placebo", "POST Placebo",
                                                     "PRE Dapagliflozin", "POST Dapagliflozin"))
attempt_ec_meta %>%
  ggplot(aes(visit, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_ec_treatment_visit_proportions_plot.jpeg")

attempt_ec_meta %>%
  ggplot(aes(treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_ec_treatment_proportions_plot.jpeg")

attempt_ec_meta %>%
  ggplot(aes(visit_treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_ec_treatment_visit_proportions_plot.jpeg")
```

#### Mixed model (hurdle, nested random effect)
##### nebula
###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_ec <- round(GetAssayData(attempt_so_hvg_ec, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_ec)

# meta_hvg_ec <- attempt_so_hvg_ec@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_ec)
# data_g_hvg_ec = group_cell(count=counts_hvg_ec, id=meta_hvg_ec$subject, pred=pred)
# offset_hvg_ec = Matrix::colSums(data_g_hvg_ec$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_ec_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_ec[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_ec, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_ec_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_ec_results_list_reml_pooled) <- sapply(nebula_ec_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_ec_results_list_reml_pooled <- lapply(nebula_ec_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_ec_results_list_reml_pooled <- Filter(Negate(is.null), nebula_ec_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_ec_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_ec_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'EC/nebula/ec_attempt_hvg_nebula_res_reml_pooled.rds')
```


```{r echo = F}
# EC subtypes
celltype_vec <- c("EC-PTC", "EC-GC", "EC-AEA", "EC-LYM")

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_ec, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("EC/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```


#### Average expression (pseudo-bulk within individual)

#### Sum expression (pseudo-bulk within individual)

# FIB, VSMC/P

```{r echo = F}
attempt_so$celltype_fibvsmc <- ifelse(attempt_so$celltype %in% c("FIB", "VSMC/P"),
                                     "FIBVSMC", as.character(attempt_so$celltype))
attempt_so_fibvsmc <- subset(attempt_so, celltype_fibvsmc == "FIBVSMC")
```

```{r echo = F}
attempt_so_fibvsmc <- FindVariableFeatures(attempt_so_fibvsmc, selection.method = "vst", nfeatures = 2000)
hvgs_fibvsmc <- VariableFeatures(attempt_so_fibvsmc)

# Perform remaining steps on top 2000 hvgs
# Subset Seurat object to only HVGs
attempt_so_hvg_fibvsmc <- subset(attempt_so_fibvsmc, features = hvgs_fibvsmc)
```

#### Proportions
```{r echo = F, eval = F}
attempt_fibvsmc_meta <- attempt_so_hvg_fibvsmc@meta.data %>%
  mutate(visit_treatment = paste0(visit, " ", treatment))
attempt_fibvsmc_meta$celltype <- factor(attempt_fibvsmc_meta$celltype, levels = c("FIB", "VSMC/P"))
attempt_fibvsmc_meta$visit_treatment <- factor(attempt_fibvsmc_meta$visit_treatment, 
                                          levels = c("PRE Placebo", "POST Placebo",
                                                     "PRE Dapagliflozin", "POST Dapagliflozin"))
attempt_fibvsmc_meta %>%
  ggplot(aes(visit, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_fibvsmc_treatment_visit_proportions_plot.jpeg")

attempt_fibvsmc_meta %>%
  ggplot(aes(treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_fibvsmc_treatment_proportions_plot.jpeg")

attempt_fibvsmc_meta %>%
  ggplot(aes(visit_treatment, fill = celltype)) +
  geom_bar(stat = "count", position = "fill") +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion",
       fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold"),
        text = element_text(size = 20)) +
  ggtitle("Proportion of cells") +
  scale_fill_manual(values = c("#264653", "#2a9d8f"))
temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, height = 10, width = 8)
s3$upload_file(temp_file, "attempt", "Figures/Proportions/attempt_fibvsmc_treatment_visit_proportions_plot.jpeg")
```

#### Mixed model (hurdle, nested random effect)
##### nebula
###### REML with rounded counts & pooled offset

```{r echo = F, eval = F}
counts_hvg_fibvsmc <- round(GetAssayData(attempt_so_hvg_fibvsmc, layer = "counts")) # load counts and round
genes_list <- rownames(counts_hvg_fibvsmc)

# meta_hvg_fibvsmc <- attempt_so_hvg_fibvsmc@meta.data
# pred = model.matrix(~treatment*visit, data = meta_hvg_fibvsmc)
# data_g_hvg_fibvsmc = group_cell(count=counts_hvg_fibvsmc, id=meta_hvg_fibvsmc$subject, pred=pred)
# offset_hvg_fibvsmc = Matrix::colSums(data_g_hvg_fibvsmc$count) 

cl <- makeCluster(100)
registerDoParallel(cl)

start_time <- Sys.time()

nebula_fibvsmc_results_list_reml_pooled <- foreach(g = genes_list, .packages = c("nebula", "Matrix"),
                                       .errorhandling = "pass") %dopar% {
  warn <- NULL
  err <- NULL
  res <- NULL
  
  tryCatch({
    count_gene <- counts_hvg_fibvsmc[g, , drop = FALSE]
    meta_gene <- subset(attempt_so_hvg_fibvsmc, features = g)@meta.data
    pred_gene <- model.matrix(~treatment * visit, data = meta_gene)
    data_g_gene <- group_cell(count = count_gene, id = meta_gene$subject, pred = pred_gene)

    res <- withCallingHandlers({
      nebula(count = data_g_gene$count, id = data_g_gene$id, pred = data_g_gene$pred,
             ncore = 1, output_re = TRUE, covariance = TRUE, reml = 1, model = "NBLMM", 
             offset = meta_gene$pooled_offset)
    }, warning = function(w) {
      warn <<- conditionMessage(w)
      invokeRestart("muffleWarning")
    })

  }, error = function(e) {
    err <<- conditionMessage(e)
  })

  list(gene = g, result = res, warning = warn, error = err)
}

## After the loop:
for (res in nebula_fibvsmc_results_list_reml_pooled) {
  if (!is.null(res$warning)) cat(sprintf("Warning for gene %s: %s\n", res$gene, res$warning))
  if (!is.null(res$error)) cat(sprintf("Error for gene %s: %s\n", res$gene, res$error))
}


# set the names of results based on gene names
names(nebula_fibvsmc_results_list_reml_pooled) <- sapply(nebula_fibvsmc_results_list_reml_pooled, function(x) x$gene)  # set names
nebula_fibvsmc_results_list_reml_pooled <- lapply(nebula_fibvsmc_results_list_reml_pooled, function(x) x$result)  # clean list back to just results
nebula_fibvsmc_results_list_reml_pooled <- Filter(Negate(is.null), nebula_fibvsmc_results_list_reml_pooled)  # remove NULLs first

end_time <- Sys.time()
print(end_time - start_time)
stopCluster(cl)

nebula_nonconverged_percent <- (length(genes_list)-length(nebula_fibvsmc_results_list_reml_pooled))/length(genes_list)
print(paste0(nebula_nonconverged_percent*100, "% filtered due to low expression"))
temp_file <- tempfile(fileext = "rds") # need to create a temporary file
saveRDS(nebula_fibvsmc_results_list_reml_pooled, temp_file) # save the temp file locally (will delete after R session)
s3$upload_file(temp_file, "attempt", 'FIBVSMCP/nebula/fibvsmc_attempt_hvg_nebula_res_reml_pooled.rds')
```


```{r echo = F}
# FIBVSMCP subtypes
celltype_vec <- c("FIB", "VSMC/P")

cl <- makeCluster(100)
registerDoParallel(cl)

overall_start_time <- Sys.time()
nebula_results_by_subtype <- vector("list", length(celltype_vec))
names(nebula_results_by_subtype) <- celltype_vec

for (ct in celltype_vec) {

  message("===== Running Nebula for ", ct, " =====")

  so_ct <- subset(attempt_so_hvg_fibvsmc, celltype == ct)

  counts_ct   <- round(GetAssayData(so_ct, layer = "counts"))
  genes_list  <- rownames(counts_ct)
  meta_ct     <- so_ct@meta.data

  ## Run your original foreach loop
  nebula_res_list <- foreach(
    g = genes_list, .packages = c("nebula", "Matrix"),
    .errorhandling = "pass"
  ) %dopar% {

    warn <- err <- res <- NULL
    tryCatch({
      meta_gene <- subset(so_ct, features = g)@meta.data
      count_gene <- counts_ct[g, , drop = FALSE]
      pred_gene  <- model.matrix(~ treatment * visit, data = meta_gene)
      data_gene  <- group_cell(count = count_gene,
                               id    = meta_gene$subject,
                               pred  = pred_gene)

      res <- withCallingHandlers(
        nebula(
          count  = data_gene$count,
          id     = data_gene$id,
          pred   = data_gene$pred,
          ncore  = 1,
          output_re  = TRUE,
          covariance = TRUE,
          reml   = 1,
          model  = "NBLMM",
          offset = meta_gene$pooled_offset
        ),
        warning = function(w) {
          warn <<- conditionMessage(w)
          invokeRestart("muffleWarning")
        }
      )
    }, error = function(e) {
      err <<- conditionMessage(e)
    })
    list(gene = g, result = res, warning = warn, error = err)
  }

  ## Post-process & save
  names(nebula_res_list) <- vapply(nebula_res_list, `[[`, "", "gene")
  nebula_res_list <- lapply(nebula_res_list, `[[`, "result")
  nebula_res_list <- Filter(Negate(is.null), nebula_res_list)

  nebula_results_by_subtype[[ct]] <- nebula_res_list

  pct_nonconv <- (length(genes_list) - length(nebula_res_list)) / length(genes_list)
  message(sprintf("%.2f%% models failed to converge for %s",
                  100 * pct_nonconv, ct))

  tmp <- tempfile(fileext = ".rds")
  saveRDS(nebula_res_list, tmp)
  s3_key <- sprintf("FIBVSMCP/%s/nebula/%s_nebula_res_reml_pooled.rds", ct, ct)
  s3$upload_file(tmp, "attempt", s3_key)
}

stopCluster(cl)
overall_end_time <- Sys.time()
print(overall_end_time - overall_start_time)
```


#### Average expression (pseudo-bulk within individual)

#### Sum expression (pseudo-bulk within individual)

# Expression

```{r echo = F}
prefixes <- c("pt", "tal", "pc", "ic", "ec", "fibvsmcp", "immune")
mods <- paste0("mod", 1:4)

nonconverged_genes <- setNames(lapply(prefixes, function(prefix) {
  setNames(lapply(mods, function(mod) {
    s3_key <- paste0("Nonconverged genes/", prefix, "_nonconverged_genes_", mod, ".rds")
    temp_file <- tempfile(fileext = ".rds")
    s3$download_file("attempt", s3_key, temp_file)
    readRDS(temp_file)
  }), mods)
}), prefixes)
```

### PT
```{r echo = F}
percent.nonconverged.mod1 <- PercentageFeatureSet(attempt_so_hvg_pt, features = nonconverged_genes$pt[["mod1"]])
percent.nonconverged.mod2 <- PercentageFeatureSet(attempt_so_hvg_pt, features = nonconverged_genes$pt[["mod2"]])
percent.nonconverged.mod3 <- PercentageFeatureSet(attempt_so_hvg_pt, features = nonconverged_genes$pt[["mod3"]])
percent.nonconverged.mod4 <- PercentageFeatureSet(attempt_so_hvg_pt, features = nonconverged_genes$pt[["mod4"]])

summary_list <- list(
  "Model 1" = summary(percent.nonconverged.mod1),
  "Model 2" = summary(percent.nonconverged.mod2),
  "Model 3" = summary(percent.nonconverged.mod3),
  "Model 4" = summary(percent.nonconverged.mod4)
)

# Convert each summary vector to a one-row data frame
pt_nonconverged_summary <- do.call(rbind, lapply(names(summary_list), function(model_name) {
  stats <- summary_list[[model_name]]
  # Coerce named vector to data.frame (one row)
  df <- as.data.frame(as.list(stats))
  df$Model <- model_name
  return(df)
}))

pt_nonconverged_summary <- pt_nonconverged_summary[, c(ncol(pt_nonconverged_summary), 1:(ncol(pt_nonconverged_summary)-1))]
colnames(pt_nonconverged_summary) <- gsub("\\.", "", colnames(pt_nonconverged_summary))
temp_file <- tempfile(fileext = "rds")
saveRDS(pt_nonconverged_summary, temp_file)
s3$upload_file(temp_file, "attempt", 'Nonconverged genes/pt_nonconverged_summary.rds')


pt_nonconverged_df <- rbind(data.frame(expression = percent.nonconverged.mod1, model = "Model 1"), 
                            data.frame(expression = percent.nonconverged.mod2, model = "Model 2"), 
                            data.frame(expression = percent.nonconverged.mod3, model = "Model 3"), 
                            data.frame(expression = percent.nonconverged.mod4, model = "Model 4"))

ggplot(pt_nonconverged_df, aes(x = model, y = expression, color = model)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#fca311") +
  geom_jitter(alpha = 0.1, shape = 21, stroke = 0, size = 5, aes(fill = model)) + 
  geom_boxplot(outliers = F, width = 0.6, linewidth = 1.25, aes(fill = model)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 15)) +
  labs(x = "Model", y = "% Expression",
       title = "% Expression of nonconverged genes in PT cells") +
  guides(color = "none",
         fill = "none") +
  scale_color_manual(values = c("#588157", "#55828b", 
                                "#f79d65", "#e26d5c")) +
  scale_fill_manual(values = c("#7BA179", "#76A0A9", 
                                "#FAB889", "#EB8B7E"))

temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/nonconverged/pt_nonconverged_expression_plot.jpeg") 
```

### TAL
```{r echo = F}
percent.nonconverged.mod1 <- PercentageFeatureSet(attempt_so_hvg_tal, features = nonconverged_genes$tal[["mod1"]])
percent.nonconverged.mod2 <- PercentageFeatureSet(attempt_so_hvg_tal, features = nonconverged_genes$tal[["mod2"]])
percent.nonconverged.mod3 <- PercentageFeatureSet(attempt_so_hvg_tal, features = nonconverged_genes$tal[["mod3"]])
percent.nonconverged.mod4 <- PercentageFeatureSet(attempt_so_hvg_tal, features = nonconverged_genes$tal[["mod4"]])

summary_list <- list(
  "Model 1" = summary(percent.nonconverged.mod1),
  "Model 2" = summary(percent.nonconverged.mod2),
  "Model 3" = summary(percent.nonconverged.mod3),
  "Model 4" = summary(percent.nonconverged.mod4)
)

# Convert each summary vector to a one-row data frame
tal_nonconverged_summary <- do.call(rbind, lapply(names(summary_list), function(model_name) {
  stats <- summary_list[[model_name]]
  # Coerce named vector to data.frame (one row)
  df <- as.data.frame(as.list(stats))
  df$Model <- model_name
  return(df)
}))

tal_nonconverged_summary <- tal_nonconverged_summary[, c(ncol(tal_nonconverged_summary), 1:(ncol(tal_nonconverged_summary)-1))]
colnames(tal_nonconverged_summary) <- gsub("\\.", "", colnames(tal_nonconverged_summary))
temp_file <- tempfile(fileext = "rds")
saveRDS(tal_nonconverged_summary, temp_file)
s3$upload_file(temp_file, "attempt", 'Nonconverged genes/tal_nonconverged_summary.rds')


tal_nonconverged_df <- rbind(data.frame(expression = percent.nonconverged.mod1, model = "Model 1"), 
                            data.frame(expression = percent.nonconverged.mod2, model = "Model 2"), 
                            data.frame(expression = percent.nonconverged.mod3, model = "Model 3"), 
                            data.frame(expression = percent.nonconverged.mod4, model = "Model 4"))

ggplot(tal_nonconverged_df, aes(x = model, y = expression, color = model)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#fca311") +
  geom_jitter(alpha = 0.1, shape = 21, stroke = 0, size = 5, aes(fill = model)) + 
  geom_boxplot(outliers = F, width = 0.6, linewidth = 1.25, aes(fill = model)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 15)) +
  labs(x = "Model", y = "% Expression",
       title = "% Expression of nonconverged genes in TAL cells") +
  guides(color = "none",
         fill = "none") +
  scale_color_manual(values = c("#588157", "#55828b", 
                                "#f79d65", "#e26d5c")) +
  scale_fill_manual(values = c("#7BA179", "#76A0A9", 
                                "#FAB889", "#EB8B7E"))

temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/nonconverged/tal_nonconverged_expression_plot.jpeg") 
```
### immune
```{r echo = F}
percent.nonconverged.mod1 <- PercentageFeatureSet(attempt_so_hvg_immune, features = nonconverged_genes$immune[["mod1"]])
percent.nonconverged.mod2 <- PercentageFeatureSet(attempt_so_hvg_immune, features = nonconverged_genes$immune[["mod2"]])
percent.nonconverged.mod3 <- PercentageFeatureSet(attempt_so_hvg_immune, features = nonconverged_genes$immune[["mod3"]])
percent.nonconverged.mod4 <- PercentageFeatureSet(attempt_so_hvg_immune, features = nonconverged_genes$immune[["mod4"]])

summary_list <- list(
  "Model 1" = summary(percent.nonconverged.mod1),
  "Model 2" = summary(percent.nonconverged.mod2),
  "Model 3" = summary(percent.nonconverged.mod3),
  "Model 4" = summary(percent.nonconverged.mod4)
)

# Convert each summary vector to a one-row data frame
immune_nonconverged_summary <- do.call(rbind, lapply(names(summary_list), function(model_name) {
  stats <- summary_list[[model_name]]
  # Coerce named vector to data.frame (one row)
  df <- as.data.frame(as.list(stats))
  df$Model <- model_name
  return(df)
}))

immune_nonconverged_summary <- immune_nonconverged_summary[, c(ncol(immune_nonconverged_summary), 1:(ncol(immune_nonconverged_summary)-1))]
colnames(immune_nonconverged_summary) <- gsub("\\.", "", colnames(immune_nonconverged_summary))
temp_file <- tempfile(fileext = "rds")
saveRDS(immune_nonconverged_summary, temp_file)
s3$upload_file(temp_file, "attempt", 'Nonconverged genes/immune_nonconverged_summary.rds')


immune_nonconverged_df <- rbind(data.frame(expression = percent.nonconverged.mod1, model = "Model 1"), 
                            data.frame(expression = percent.nonconverged.mod2, model = "Model 2"), 
                            data.frame(expression = percent.nonconverged.mod3, model = "Model 3"), 
                            data.frame(expression = percent.nonconverged.mod4, model = "Model 4"))

ggplot(immune_nonconverged_df, aes(x = model, y = expression, color = model)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#fca311") +
  geom_jitter(alpha = 0.1, shape = 21, stroke = 0, size = 5, aes(fill = model)) + 
  geom_boxplot(outliers = F, width = 0.6, linewidth = 1.25, aes(fill = model)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 15)) +
  labs(x = "Model", y = "% Expression",
       title = "% Expression of nonconverged genes in Immune cells") +
  guides(color = "none",
         fill = "none") +
  scale_color_manual(values = c("#588157", "#55828b", 
                                "#f79d65", "#e26d5c")) +
  scale_fill_manual(values = c("#7BA179", "#76A0A9", 
                                "#FAB889", "#EB8B7E"))

temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/nonconverged/immune_nonconverged_expression_plot.jpeg") 
```
### PC
```{r echo = F}
attempt_so_hvg_pc[["percent.nonconverged.mod1"]] <- PercentageFeatureSet(attempt_so_hvg_pc, features = nonconverged_genes$pc[[1]])

percent.nonconverged.mod1 <- PercentageFeatureSet(attempt_so_hvg_pc, features = nonconverged_genes$pc[["mod1"]])
percent.nonconverged.mod2 <- PercentageFeatureSet(attempt_so_hvg_pc, features = nonconverged_genes$pc[["mod2"]])
percent.nonconverged.mod3 <- PercentageFeatureSet(attempt_so_hvg_pc, features = nonconverged_genes$pc[["mod3"]])
percent.nonconverged.mod4 <- PercentageFeatureSet(attempt_so_hvg_pc, features = nonconverged_genes$pc[["mod4"]])

summary_list <- list(
  "Model 1" = summary(percent.nonconverged.mod1),
  "Model 2" = summary(percent.nonconverged.mod2),
  "Model 3" = summary(percent.nonconverged.mod3),
  "Model 4" = summary(percent.nonconverged.mod4)
)

# Convert each summary vector to a one-row data frame
pc_nonconverged_summary <- do.call(rbind, lapply(names(summary_list), function(model_name) {
  stats <- summary_list[[model_name]]
  # Coerce named vector to data.frame (one row)
  df <- as.data.frame(as.list(stats))
  df$Model <- model_name
  return(df)
}))

pc_nonconverged_summary <- pc_nonconverged_summary[, c(ncol(pc_nonconverged_summary), 1:(ncol(pc_nonconverged_summary)-1))]
colnames(pc_nonconverged_summary) <- gsub("\\.", "", colnames(pc_nonconverged_summary))
temp_file <- tempfile(fileext = "rds")
saveRDS(pc_nonconverged_summary, temp_file)
s3$upload_file(temp_file, "attempt", 'Nonconverged genes/pc_nonconverged_summary.rds')


pc_nonconverged_df <- rbind(data.frame(expression = percent.nonconverged.mod1, model = "Model 1"), 
                            data.frame(expression = percent.nonconverged.mod2, model = "Model 2"), 
                            data.frame(expression = percent.nonconverged.mod3, model = "Model 3"), 
                            data.frame(expression = percent.nonconverged.mod4, model = "Model 4"))

ggplot(pc_nonconverged_df, aes(x = model, y = expression, color = model)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#fca311") +
  geom_jitter(alpha = 0.1, shape = 21, stroke = 0, size = 5, aes(fill = model)) + 
  geom_boxplot(outliers = F, width = 0.6, linewidth = 1.25, aes(fill = model)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 15)) +
  labs(x = "Model", y = "% Expression",
       title = "% Expression of nonconverged genes in PC cells") +
  guides(color = "none",
         fill = "none") +
  scale_color_manual(values = c("#588157", "#55828b", 
                                "#f79d65", "#e26d5c")) +
  scale_fill_manual(values = c("#7BA179", "#76A0A9", 
                                "#FAB889", "#EB8B7E"))

temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/nonconverged/pc_nonconverged_expression_plot.jpeg") 
```

### IC
```{r echo = F}
attempt_so_hvg_ic[["percent.nonconverged.mod1"]] <- PercentageFeatureSet(attempt_so_hvg_ic, features = nonconverged_genes$ic[[1]])

percent.nonconverged.mod1 <- PercentageFeatureSet(attempt_so_hvg_ic, features = nonconverged_genes$ic[["mod1"]])
percent.nonconverged.mod2 <- PercentageFeatureSet(attempt_so_hvg_ic, features = nonconverged_genes$ic[["mod2"]])
percent.nonconverged.mod3 <- PercentageFeatureSet(attempt_so_hvg_ic, features = nonconverged_genes$ic[["mod3"]])
percent.nonconverged.mod4 <- PercentageFeatureSet(attempt_so_hvg_ic, features = nonconverged_genes$ic[["mod4"]])

summary_list <- list(
  "Model 1" = summary(percent.nonconverged.mod1),
  "Model 2" = summary(percent.nonconverged.mod2),
  "Model 3" = summary(percent.nonconverged.mod3),
  "Model 4" = summary(percent.nonconverged.mod4)
)

# Convert each summary vector to a one-row data frame
ic_nonconverged_summary <- do.call(rbind, lapply(names(summary_list), function(model_name) {
  stats <- summary_list[[model_name]]
  # Coerce named vector to data.frame (one row)
  df <- as.data.frame(as.list(stats))
  df$Model <- model_name
  return(df)
}))

ic_nonconverged_summary <- ic_nonconverged_summary[, c(ncol(ic_nonconverged_summary), 1:(ncol(ic_nonconverged_summary)-1))]
colnames(ic_nonconverged_summary) <- gsub("\\.", "", colnames(ic_nonconverged_summary))
temp_file <- tempfile(fileext = "rds")
saveRDS(ic_nonconverged_summary, temp_file)
s3$upload_file(temp_file, "attempt", 'Nonconverged genes/ic_nonconverged_summary.rds')


ic_nonconverged_df <- rbind(data.frame(expression = percent.nonconverged.mod1, model = "Model 1"), 
                            data.frame(expression = percent.nonconverged.mod2, model = "Model 2"), 
                            data.frame(expression = percent.nonconverged.mod3, model = "Model 3"), 
                            data.frame(expression = percent.nonconverged.mod4, model = "Model 4"))

ggplot(ic_nonconverged_df, aes(x = model, y = expression, color = model)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#fca311") +
  geom_jitter(alpha = 0.1, shape = 21, stroke = 0, size = 5, aes(fill = model)) + 
  geom_boxplot(outliers = F, width = 0.6, linewidth = 1.25, aes(fill = model)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 15)) +
  labs(x = "Model", y = "% Expression",
       title = "% Expression of nonconverged genes in IC cells") +
  guides(color = "none",
         fill = "none") +
  scale_color_manual(values = c("#588157", "#55828b", 
                                "#f79d65", "#e26d5c")) +
  scale_fill_manual(values = c("#7BA179", "#76A0A9", 
                                "#FAB889", "#EB8B7E"))

temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/nonconverged/ic_nonconverged_expression_plot.jpeg") 
```
### EC
```{r echo = F}
attempt_so_hvg_ec[["percent.nonconverged.mod1"]] <- PercentageFeatureSet(attempt_so_hvg_ec, features = nonconverged_genes$ec[[1]])

percent.nonconverged.mod1 <- PercentageFeatureSet(attempt_so_hvg_ec, features = nonconverged_genes$ec[["mod1"]])
percent.nonconverged.mod2 <- PercentageFeatureSet(attempt_so_hvg_ec, features = nonconverged_genes$ec[["mod2"]])
percent.nonconverged.mod3 <- PercentageFeatureSet(attempt_so_hvg_ec, features = nonconverged_genes$ec[["mod3"]])
percent.nonconverged.mod4 <- PercentageFeatureSet(attempt_so_hvg_ec, features = nonconverged_genes$ec[["mod4"]])

summary_list <- list(
  "Model 1" = summary(percent.nonconverged.mod1),
  "Model 2" = summary(percent.nonconverged.mod2),
  "Model 3" = summary(percent.nonconverged.mod3),
  "Model 4" = summary(percent.nonconverged.mod4)
)

# Convert each summary vector to a one-row data frame
ec_nonconverged_summary <- do.call(rbind, lapply(names(summary_list), function(model_name) {
  stats <- summary_list[[model_name]]
  # Coerce named vector to data.frame (one row)
  df <- as.data.frame(as.list(stats))
  df$Model <- model_name
  return(df)
}))

ec_nonconverged_summary <- ec_nonconverged_summary[, c(ncol(ec_nonconverged_summary), 1:(ncol(ec_nonconverged_summary)-1))]
colnames(ec_nonconverged_summary) <- gsub("\\.", "", colnames(ec_nonconverged_summary))
temp_file <- tempfile(fileext = "rds")
saveRDS(ec_nonconverged_summary, temp_file)
s3$upload_file(temp_file, "attempt", 'Nonconverged genes/ec_nonconverged_summary.rds')


ec_nonconverged_df <- rbind(data.frame(expression = percent.nonconverged.mod1, model = "Model 1"), 
                            data.frame(expression = percent.nonconverged.mod2, model = "Model 2"), 
                            data.frame(expression = percent.nonconverged.mod3, model = "Model 3"), 
                            data.frame(expression = percent.nonconverged.mod4, model = "Model 4"))

ggplot(ec_nonconverged_df, aes(x = model, y = expression, color = model)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#fca311") +
  geom_jitter(alpha = 0.1, shape = 21, stroke = 0, size = 5, aes(fill = model)) + 
  geom_boxplot(outliers = F, width = 0.6, linewidth = 1.25, aes(fill = model)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 15)) +
  labs(x = "Model", y = "% Expression",
       title = "% Expression of nonconverged genes in EC cells") +
  guides(color = "none",
         fill = "none") +
  scale_color_manual(values = c("#588157", "#55828b", 
                                "#f79d65", "#e26d5c")) +
  scale_fill_manual(values = c("#7BA179", "#76A0A9", 
                                "#FAB889", "#EB8B7E"))

temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/nonconverged/ec_nonconverged_expression_plot.jpeg") 
```
### FIB, VSMC/P
```{r echo = F}
attempt_so_hvg_fibvsmc[["percent.nonconverged.mod1"]] <- PercentageFeatureSet(attempt_so_hvg_fibvsmc, features = nonconverged_genes$fibvsmcp[[1]])

percent.nonconverged.mod1 <- PercentageFeatureSet(attempt_so_hvg_fibvsmc, features = nonconverged_genes$fibvsmcp[["mod1"]])
percent.nonconverged.mod2 <- PercentageFeatureSet(attempt_so_hvg_fibvsmc, features = nonconverged_genes$fibvsmcp[["mod2"]])
percent.nonconverged.mod3 <- PercentageFeatureSet(attempt_so_hvg_fibvsmc, features = nonconverged_genes$fibvsmcp[["mod3"]])
percent.nonconverged.mod4 <- PercentageFeatureSet(attempt_so_hvg_fibvsmc, features = nonconverged_genes$fibvsmcp[["mod4"]])

summary_list <- list(
  "Model 1" = summary(percent.nonconverged.mod1),
  "Model 2" = summary(percent.nonconverged.mod2),
  "Model 3" = summary(percent.nonconverged.mod3),
  "Model 4" = summary(percent.nonconverged.mod4)
)

# Convert each summary vector to a one-row data frame
fibvsmcp_nonconverged_summary <- do.call(rbind, lapply(names(summary_list), function(model_name) {
  stats <- summary_list[[model_name]]
  # Coerce named vector to data.frame (one row)
  df <- as.data.frame(as.list(stats))
  df$Model <- model_name
  return(df)
}))

fibvsmcp_nonconverged_summary <- fibvsmcp_nonconverged_summary[, c(ncol(fibvsmcp_nonconverged_summary), 1:(ncol(fibvsmcp_nonconverged_summary)-1))]
colnames(fibvsmcp_nonconverged_summary) <- gsub("\\.", "", colnames(fibvsmcp_nonconverged_summary))
temp_file <- tempfile(fileext = "rds")
saveRDS(fibvsmcp_nonconverged_summary, temp_file)
s3$upload_file(temp_file, "attempt", 'Nonconverged genes/fibvsmcp_nonconverged_summary.rds')


fibvsmcp_nonconverged_df <- rbind(data.frame(expression = percent.nonconverged.mod1, model = "Model 1"), 
                            data.frame(expression = percent.nonconverged.mod2, model = "Model 2"), 
                            data.frame(expression = percent.nonconverged.mod3, model = "Model 3"), 
                            data.frame(expression = percent.nonconverged.mod4, model = "Model 4"))

ggplot(fibvsmcp_nonconverged_df, aes(x = model, y = expression, color = model)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "#fca311") +
  geom_jitter(alpha = 0.1, shape = 21, stroke = 0, size = 5, aes(fill = model)) + 
  geom_boxplot(outliers = F, width = 0.6, linewidth = 1.25, aes(fill = model)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 15)) +
  labs(x = "Model", y = "% Expression",
       title = "% Expression of nonconverged genes in FIB, VSMC/P cells") +
  guides(color = "none",
         fill = "none") +
  scale_color_manual(values = c("#588157", "#55828b", 
                                "#f79d65", "#e26d5c")) +
  scale_fill_manual(values = c("#7BA179", "#76A0A9", 
                                "#FAB889", "#EB8B7E"))

temp_file <- tempfile(fileext = ".jpeg")
ggsave(temp_file, width = 10)
s3$upload_file(temp_file, "attempt", "Figures/nonconverged/fibvsmcp_nonconverged_expression_plot.jpeg") 
```

# HVGs (union)
```{r echo = F}
hvgs_union <- unique(c(hvgs_ec, hvgs_fibvsmc, hvgs_ic, hvgs_immune, hvgs_pc, hvgs_pt, hvgs_tal))
length(hvgs_union)
```
