---
title: "02_proteomics_analysis"
author: "Ye Ji Choi & Savanah Leidholt"
format: html
---

```{r}
library(dplyr)
library(tidyr)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggforce)
library(enrichR)
library(stringr)
library(limma)
library(EnhancedVolcano)
library(patchwork)
library(ggrepel)
library(vegan)
library(tidyverse)
library(data.table)
library(purrr)
library(haven)
library(ggVennDiagram)
library(caret)
library(kableExtra)
library(msigdbr)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/Savanah Leidholt/")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/T1D Adiposity/Results/Compiled Results/Proteomics Results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "hhampson") {
  root_path <- "/Volumes/Peds Endo"
} else {
  stop("Unknown user: please specify root path for this user.")
}
```


```{r}
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```

#no dxa on attempt
#also do waist to hip = wthr, wthr_obesity
#hba1c with age

```{r}
t1d_hc_dat_soma <- s3readRDS(object = "data_clean/t1d_hc_clinical_data_soma.csv", 
                             bucket = "t1d.adiposity",
                             region = "")
```

#-------------------------
#creating cleaned metadata
#-------------------------
```{r}
meta <- t1d_hc_dat_soma %>%
  dplyr::select(-starts_with("seq")) %>%
  dplyr::mutate(
    record_id = as.character(record_id),
    group = factor(
      group,
      levels = c("Lean Control", "Type 1 Diabetes"),
      labels = c("LC", "T1D")
    ),
    study = dplyr::recode(
      study,
      "CROCODILE"  = "CRC",
      "PANDA"      = "PAN",
      "RENAL-HEIR" = "RH",
      "CASPER"     = "CAS",
      "ATTEMPT"    = "ATT",
      .default = study
    ),
    bmi_obesity = factor(
      bmi_obesity,
      levels = c("Normal", "Overweight", "Obese")
    ),
    bmi_bin = dplyr::case_when(
      bmi_obesity == "Normal" ~ "Normal",
      bmi_obesity %in% c("Overweight", "Obese") ~ "Overweight_Obese",
      TRUE ~ NA_character_
    ),
    group_bmi = dplyr::case_when(
      group == "LC"  & bmi_obesity == "Normal" ~ "LC_Normal",
      group == "LC"  & bmi_obesity %in% c("Overweight", "Obese") ~ "LC_Overweight_Obese",
      group == "T1D" & bmi_obesity == "Normal" ~ "T1D_Normal",
      group == "T1D" & bmi_obesity == "Overweight" ~ "T1D_Overweight",
      group == "T1D" & bmi_obesity == "Obese" ~ "T1D_Obese",
      TRUE ~ NA_character_
    )
  )

meta$study <- factor(meta$study)
meta$bmi_bin <- factor(meta$bmi_bin, levels = c("Normal", "Overweight_Obese"))
meta$group_bmi <- factor(
  meta$group_bmi,
  levels = c(
    "LC_Normal",
    "LC_Overweight_Obese",
    "T1D_Normal",
    "T1D_Overweight",
    "T1D_Obese"
  )
)

# keep BMI-complete samples
hba1c_var <- "hba1c"
meta2 <- meta %>%
 dplyr::filter(
    !is.na(record_id),
    !is.na(group_bmi),
    !is.na(age),
    !is.na(.data[[hba1c_var]]),
    !is.na(sex),
    !is.na(study)
  )

table(meta2$group_bmi, useNA = "ifany")
```
#----------------------------
# Plasma protein matrix setup
#----------------------------
```{r}
dat_plasma <- t1d_hc_dat_soma %>%
  dplyr::select(record_id, starts_with("seq") &
                  !dplyr::ends_with("urine") &
                  !dplyr::ends_with("urine_cradj"))  %>% as.data.frame()

rownames(dat_plasma) <- dat_plasma$record_id
dat_plasma$record_id <- NULL

# align metadata and matrix
ids_shared <- intersect(meta2$record_id, rownames(dat_plasma))

dat_plasma2 <- dat_plasma[ids_shared, , drop = FALSE]

meta2 <- meta2 %>%
  dplyr::filter(record_id %in% ids_shared)

meta2 <- meta2[match(rownames(dat_plasma2), meta2$record_id), ]

stopifnot(identical(rownames(dat_plasma2), meta2$record_id))
```

#--------------------------------------
# Filter to SomaScan 7000 analyte panel
#--------------------------------------
```{r}
analytes <- read.csv(
  file.path(root_path, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

prot_panel <- analytes$AptName %>%
  as.character() %>%
  trimws() %>%
  unique()

prot_panel <- prot_panel[!is.na(prot_panel) & prot_panel != ""]

aptamers_keep <- intersect(prot_panel, colnames(dat_plasma2))

if (length(aptamers_keep) == 0) {
  stop("No analytes from analytes.csv matched columns in dat_plasma2.")
}

dat_plasma3 <- dat_plasma2[, aptamers_keep, drop = FALSE]

cat("Plasma features before analyte filter:", ncol(dat_plasma2), "\n")
cat("Proteins in analyte panel:", length(prot_panel), "\n")
cat("Matched proteins kept:", length(aptamers_keep), "\n")
cat("Panel proteins missing from plasma matrix:", length(setdiff(prot_panel, colnames(dat_plasma2))), "\n")

dat_plasma3 <- as.data.frame(dat_plasma3)

dat_plasma3[] <- lapply(dat_plasma3, function(x) {
  suppressWarnings(as.numeric(as.character(x)))
})

dat_plasma3 <- as.matrix(dat_plasma3)

dat_log <- log2(dat_plasma3 + 1)
dat_log[!is.finite(dat_log)] <- NA_real_

stopifnot(identical(rownames(dat_log), meta2$record_id))
```
#----------------------------------------------------
# Study presence filter
# Protein present in >=60% samples in >=2 studies
# doesn't remove anything in this case but a good check
#------------------------------------------------------
```{r}
meta2$study <- factor(meta2$study)

study_idx_all <- split(seq_len(nrow(dat_log)), meta2$study)

present_by_study_all <- sapply(study_idx_all, function(idx) {
  colMeans(!is.na(dat_log[idx, , drop = FALSE])) >= 0.6
})

present_by_study_all <- as.matrix(present_by_study_all)

n_studies_present_all <- rowSums(present_by_study_all)
keep_study_all <- n_studies_present_all >= 2

kept_proteins <- colnames(dat_log)[keep_study_all]
removed_proteins <- colnames(dat_log)[!keep_study_all]

cat("All proteomic features before study filter:", ncol(dat_log), "\n")
cat("Removed:", length(removed_proteins), "\n")
cat("Kept:", length(kept_proteins), "\n")

dat_log_filt <- dat_log[, kept_proteins, drop = FALSE]

stopifnot(identical(rownames(dat_log_filt), meta2$record_id))
```
#-------------------------
#imputing the missing data
#-------------------------
```{r}
set.seed(1835)

dat_imp <- dat_log_filt

meds <- apply(dat_imp, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(dat_imp))) {
  miss_j <- is.na(dat_imp[, j])
  if (!is.na(meds[j]) && any(miss_j)) {
    dat_imp[miss_j, j] <- meds[j]
  }
}

vars <- apply(dat_imp, 2, var)

keep_cols <- colSums(is.na(dat_imp)) == 0 &
  is.finite(vars) &
  vars > 0

dat_imp <- dat_imp[, keep_cols, drop = FALSE]

stopifnot(identical(rownames(dat_imp), meta2$record_id))
stopifnot(sum(is.na(dat_imp)) == 0)
```

#--------------------------
#checking for outliers
#--------------------------
```{r}
pca0 <- prcomp(dat_imp, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2, drop = FALSE])
scores$record_id <- rownames(scores)
scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

threshold <- mean(scores$dist, na.rm = TRUE) +
  3 * sd(scores$dist, na.rm = TRUE)

outliers <- scores$record_id[scores$dist > threshold] #2 outliers CRC-42 & 10424
outliers
dat_imp2 <- dat_imp[!rownames(dat_imp) %in% outliers, , drop = FALSE]

meta3 <- meta2 %>%
  dplyr::filter(!record_id %in% outliers)

meta3 <- meta3[match(rownames(dat_imp2), meta3$record_id), ]

stopifnot(identical(rownames(dat_imp2), meta3$record_id))
```
#----------------------------
#Data table: study x group_bmi
#-----------------------------
```{r}
study_group_bmi_table <- meta3 %>%
  dplyr::count(study, group_bmi, name = "n") %>%
  tidyr::pivot_wider(
    names_from = group_bmi,
    values_from = n,
    values_fill = 0
  ) %>%
  dplyr::mutate(
    Total = rowSums(dplyr::across(where(is.numeric)))
  )

study_group_bmi_table %>%
  kable(caption = "Participants by study and disease-adiposity group") %>%
  kable_styling(full_width = FALSE)
```
#-------------------------------------
#PCA plotting function for group_bmi
#-------------------------------------
```{r}
make_pca <- function(mat, meta_df, title_str) {
  stopifnot(identical(rownames(mat), meta_df$record_id))

  pca <- prcomp(mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1 = pca$x[, 1],
    PC2 = pca$x[, 2],
    group_bmi = meta_df$group_bmi,
    study = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = group_bmi, shape = study)) +
    scale_color_manual(
      values = c(
        LC_Normal  = "cyan4",
        LC_Overweight_Obese = "darkred",
        T1D_Normal = "gold",
        T1D_Overweight = "plum3",
        T1D_Obese = "black"
      ),
      name = "Group"
    ) +
    geom_point(size = 3, alpha = 0.85) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)"),
      color = "Disease-adiposity group"
    ) +
    theme_bw()
}
```

```{r}
stopifnot(identical(rownames(dat_imp2), meta3$record_id))
design <- model.matrix(~ group_bmi, data = meta3)

#model 0 ~ no adjustment
p0 <- make_pca(dat_imp2,meta3,"No adjustment")

#model 1 ~ age + sex adjustment
cov_age_sex <- model.matrix(~ age + sex, data = meta3)[, -1, drop = FALSE]
prot_age_sex <- t(removeBatchEffect(t(dat_imp2),
  covariates = cov_age_sex,
  design = design
))
p1 <- make_pca(prot_age_sex,meta3,"Adjusted for age + sex")

#model 2 ~ age + sex + study adjustment
cov_age_sex_study <- model.matrix(~ age + sex + study, data = meta3)[, -1, drop = FALSE]
prot_age_sex_study <- t(removeBatchEffect(t(dat_imp2),
  covariates = cov_age_sex_study,
  design = design
))
p2 <- make_pca(prot_age_sex_study, meta3, "Adjusted for age + sex + study")

#model 3 ~ age * HbA1c + sex + study adjustment
cov_hba1c_age <- model.matrix(
  as.formula(paste0("~ age * ", hba1c_var, " + sex + study")),
  data = meta3
)[, -1, drop = FALSE]

# final row checks
nrow(design)
nrow(cov_hba1c_age)
stopifnot(nrow(design) == nrow(cov_hba1c_age))
stopifnot(nrow(design) == nrow(dat_imp2))

prot_hba1c_age <- t(removeBatchEffect(t(dat_imp2), covariates = cov_hba1c_age, design = design))
p3 <- make_pca(prot_hba1c_age, meta3,"Adjusted for age * HbA1c + sex + study")

# combine plots
pca_group_bmi_all <- (p0 | p1) / (p2 | p3)
pca_group_bmi_all
```

```{r}
mat1 <- t(dat_imp2)

run_group_bmi_limma <- function(mat, meta_df, design_formula, model_label) {
  
  design <- model.matrix(design_formula, data = meta_df)
  colnames(design) <- make.names(colnames(design))
  
  message("Design columns for ", model_label, ":")
  print(colnames(design))
  
  contrast_matrix <- makeContrasts(
    
    T1DNormal_vs_LCNormal =
      group_bmiT1D_Normal - group_bmiLC_Normal,
    
    T1DOverObese_vs_LCOverObese =
      ((group_bmiT1D_Overweight + group_bmiT1D_Obese) / 2) -
      group_bmiLC_Overweight_Obese,
    
    T1DNormal_vs_T1DOverweight =
      group_bmiT1D_Normal - group_bmiT1D_Overweight,
    
    T1DNormal_vs_T1DObese =
      group_bmiT1D_Normal - group_bmiT1D_Obese,
    
    T1DObese_vs_T1DOverweight =
      group_bmiT1D_Obese - group_bmiT1D_Overweight,
    
    levels = design
  )
  
  fit <- lmFit(mat, design)
  fit2 <- contrasts.fit(fit, contrast_matrix)
  fit2 <- eBayes(fit2)
  
  results <- lapply(colnames(contrast_matrix), function(cn) {
    topTable(
      fit2,
      coef = cn,
      number = Inf,
      adjust.method = "BH"
    ) %>%
      rownames_to_column("protein_id") %>%
      mutate(
        model = model_label,
        contrast = cn,
        direction = case_when(
          adj.P.Val < 0.05 & logFC > 0 ~ "Higher in first group",
          adj.P.Val < 0.05 & logFC < 0 ~ "Higher in second group",
          TRUE ~ "Not significant"
        )
      )
  }) %>%
    bind_rows()
  
  return(results)
}
```



```{r}
# Model 0: group_bmi only / no covariate adjustment
res_m0 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi,
  model_label = "M0: group_bmi only"
)

# Model 1: group_bmi + age + sex
res_m1 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi + age + sex,
  model_label = "M1: group_bmi + age + sex"
)

#Model 2: group_bmi + age + sex + study
res_m2 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi + age + sex + study,
  model_label = "M2: group_bmi + age + sex + study"
)


# Model 3: group_bmi + age * HbA1c + sex + study
design2 <- as.formula(
  paste0("~ 0 + group_bmi + age * ", hba1c_var, " + sex + study")
)
res_m3 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = design2,
  model_label = "M3: group_bmi + age * HbA1c + sex + study"
)
```
```{r}
results_all_models <- bind_rows(
  res_m0,
  res_m1,
  res_m2,
  res_m3
)

sig_all_models <- results_all_models %>% filter(adj.P.Val < 0.05)
sig_all_models_fc <- results_all_models %>% filter(adj.P.Val < 0.05, abs(logFC) > 0.5)

sig_all_models %>% count(model, contrast, direction)
results_all_models %>% filter(model == "M3: group_bmi + age * HbA1c + sex + study")
```

```{r}
#| fig-width: 20
#| fig-height: 18

make_volcano <- function(results_df,
                         contrast_name,
                         title_prefix = NULL,
                         p_cutoff = 0.05,
                         fc_cutoff = 0.1,
                         label_n = 15) {
  
  df <- results_df %>%
    dplyr::filter(contrast == contrast_name) %>%
    dplyr::mutate(
      neg_log10_adj_p = -log10(adj.P.Val),
      sig_group = dplyr::case_when(
        adj.P.Val < p_cutoff & logFC > fc_cutoff  ~ "Up",
        adj.P.Val < p_cutoff & logFC < -fc_cutoff ~ "Down",
        TRUE ~ "Not significant"
      )
    )
  
  n_up <- sum(df$sig_group == "Up", na.rm = TRUE)
  n_down <- sum(df$sig_group == "Down", na.rm = TRUE)
  n_ns <- sum(df$sig_group == "Not significant", na.rm = TRUE)
  
  legend_labs <- c(
    "Up" = paste0("Upregulated (n = ", n_up, ")"),
    "Down" = paste0("Downregulated (n = ", n_down, ")"),
    "Not significant" = paste0("Not significant (n = ", n_ns, ")")
  )
  
  top_labels <- df %>%
    dplyr::arrange(dplyr::desc(abs(t))) %>%
    dplyr::slice_head(n = label_n)
  
  ggplot(df, aes(x = logFC, y = neg_log10_adj_p)) +
    geom_point(aes(color = sig_group), alpha = 0.75, size = 2) +
    geom_vline(xintercept = c(-fc_cutoff, fc_cutoff), linetype = "dashed") +
    geom_hline(yintercept = -log10(p_cutoff), linetype = "dashed") +
    ggrepel::geom_text_repel(
      data = top_labels,
      aes(label = protein_id),
      size = 3.5,
      max.overlaps = Inf
    ) +
    scale_color_manual(
      values = c(
        "Up" = "red3",
        "Down" = "blue3",
        "Not significant" = "grey70"
      ),
      labels = legend_labs
    ) +
    labs(
      title = paste0(title_prefix, "\n", contrast_name),
      x = "log2 Fold Change",
      y = "-log10 adjusted P-value",
      color = NULL
    ) +
    theme_bw() +
    theme(
      panel.grid = element_blank(),
      legend.position = "right"
    )
}
```


```{r}
make_volcano_list <- function(results_df, model_name, label_n = 15) {
  
  contrasts <- unique(results_df$contrast)
  
  plot_list <- lapply(contrasts, function(ct) {
    make_volcano(
      results_df = results_df,
      contrast_name = ct,
      title_prefix = model_name,
      label_n = label_n
    )
  })
  
  names(plot_list) <- contrasts
  
  plot_list
}
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m0 <- make_volcano_list(
  results_df = res_m0,
  model_name = "M0: group_bmi only",
  label_n = 5
)

top_row <- plots_m0[[1]] + plots_m0[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m0[[3]] + plots_m0[[4]] + plots_m0[[5]]

panel_m0_5 <- top_row / bottom_row

panel_m0_5
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m1 <- make_volcano_list(
  results_df = res_m1,
  model_name = "M1: group_bmi + age + sex",
  label_n = 5
)

top_row <- plots_m1[[1]] + plots_m1[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m1[[3]] + plots_m1[[4]] + plots_m1[[5]]

panel_m1_5 <- top_row / bottom_row

panel_m1_5

```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m2 <- make_volcano_list(
  results_df = res_m2,
  model_name = "M2: group_bmi + age + sex + study",
  label_n = 5
)

top_row <- plots_m2[[1]] + plots_m2[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m2[[3]] + plots_m2[[4]] + plots_m2[[5]]

panel_m2_5 <- top_row / bottom_row

panel_m2_5
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m3 <- make_volcano_list(
  results_df = res_m3,
  model_name = "M3: group_bmi + age*hba1c + sex + study",
  label_n = 5
)

top_row <- plots_m0[[1]] + plots_m0[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m0[[3]] + plots_m0[[4]] + plots_m0[[5]]

panel_m0_5 <- top_row / bottom_row

panel_m0_5
```
#-----------------
#pathway anlaysis
#-----------------
```{r}

```











#archived code
#setting up metadata and protein matrices
```{r}
meta <- t1d_hc_dat_soma %>%
  dplyr:: select (-starts_with("seq"))
#meta$group
meta$group <- factor(
  meta$group,
  levels = c("Lean Control","Type 1 Diabetes"),
  labels = c("LC","T1D")
)

meta <- meta %>%
  dplyr::mutate(
    study = dplyr::recode(study,
      "CROCODILE"   = "CRC",
      "PANDA"       = "PAN",
      "RENAL-HEIR"  = "RH",
      "CASPER" = "CAS",
      "ATTEMPT" = "ATT"
    )
  )

meta$record_id <- as.character(meta$record_id)
#meta$bmi_obesity
#meta$dxa_obesity
# keep only participants with BOTH BMI and DXA obesity
meta2 <- meta %>%
  dplyr::filter(
    !is.na(bmi_obesity) #,
    #!is.na(dxa_obesity)
  )

dat <- t1d_hc_dat_soma %>%
  dplyr::select(record_id, starts_with("seq"))


#dat_urine <- dat %>%
#  dplyr::select(ends_with("urine"), ends_with("urine_cradj"))

dat_plasma <- dat %>%
  dplyr:: select(-ends_with("urine"), -ends_with("urine_cradj"))
dat_plasma <- as.data.frame(dat_plasma)
rownames(dat_plasma) <- dat_plasma$record_id
dat_plasma$record_id <- NULL

#filtering final matrices to have the same record_ids
ids_keep <- meta2$record_id

# plasma
dat_plasma2 <- dat_plasma[ids_keep, , drop = FALSE]
meta2 <- meta2 %>%
  dplyr::filter(record_id %in% rownames(dat_plasma2))
meta2 <- meta2[match(rownames(dat_plasma2), meta2$record_id), ]

stopifnot(identical(rownames(dat_plasma2), meta2$record_id))

```


##filter to proteins that are only in the analyte panel (~7000)
##check if this is correct for the newer somascan prot pannel
##should filter ~3000 all from ATTEMPT
##then filter to keep proteins that are in at least 2 studies
```{r}
analytes <- read.csv(
  file.path(root_path, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

stopifnot("AptName" %in% colnames(analytes))

prot_panel <- analytes$AptName %>%
  as.character() %>%
  trimws() %>%
  unique()

prot_panel <- prot_panel[!is.na(prot_panel) & prot_panel != ""]

aptamers_keep <- intersect(prot_panel, colnames(dat_plasma2))

if (length(aptamers_keep) == 0) {
  stop("No analytes from analytes.csv matched columns in dat_plasma2.")
}

dat_plasma3 <- dat_plasma2[, aptamers_keep, drop = FALSE]

cat("Plasma features before analyte filter:", ncol(dat_plasma2), "\n")
cat("Proteins in analyte panel:", length(prot_panel), "\n")
cat("Matched proteins kept:", length(aptamers_keep), "\n")
cat("Panel proteins missing from plasma matrix:", length(setdiff(prot_panel, colnames(dat_plasma2))), "\n")

# force numeric safely
dat_plasma3 <- as.data.frame(dat_plasma3)

dat_plasma3[] <- lapply(dat_plasma3, function(x) {
  suppressWarnings(as.numeric(as.character(x)))
})

dat_plasma3 <- as.matrix(dat_plasma3)

# log2 transform
dat_log <- log2(dat_plasma3 + 1)
dat_log[!is.finite(dat_log)] <- NA_real_

stopifnot(identical(rownames(dat_log), meta2$record_id))
```


```{r}
#not really necessary to do as all protiens pass filter once I remove the extra 3000 from the ATTEMPT Panel
##filtering step where protein must exist in 60% of samples in a study
stopifnot(identical(rownames(dat_log), meta2$record_id))
meta2$study <- factor(meta2$study)

study_idx_all <- split(seq_len(nrow(dat_log)), meta2$study)
present_by_study_all <- sapply(study_idx_all, function(idx) {
  colMeans(!is.na(dat_log[idx, , drop = FALSE])) >= 0.6
})

present_by_study_all <- as.matrix(present_by_study_all)

n_studies_present_all <- rowSums(present_by_study_all)
keep_study_all <- n_studies_present_all >= 2

removed_proteins <- colnames(dat_log)[!keep_study_all]
kept_proteins    <- colnames(dat_log)[keep_study_all]

cat("All proteomic features before study filter:", ncol(dat_log), "\n")
cat("Removed (<2 studies meeting 60% presence threshold):", length(removed_proteins), "\n")
cat("Kept (>=2 studies meeting 60% presence threshold):", length(kept_proteins), "\n")

if (length(kept_proteins) == 0) {
  stop("No proteins passed the study presence filter.")
}

dat_log_filt <- dat_log[, kept_proteins, drop = FALSE]

stopifnot(identical(rownames(dat_log_filt), meta2$record_id))
```

#imputating the data
```{r}
set.seed(1835)
dat_imp <- dat_log_filt
meds <- apply(dat_imp, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(dat_imp))) {
  if (!is.na(meds[j])) {
    dat_imp[is.na(dat_imp[, j]), j] <- meds[j]
  }
}

keep_cols <- colSums(is.na(dat_imp)) == 0 &
  apply(dat_imp, 2, var) > 0 &
  is.finite(apply(dat_imp, 2, var))

dat_imp <- dat_imp[, keep_cols, drop = FALSE]
```

##--------------------------------
## Find and Remove Outliers in PCA
##---------------------------------
```{r}
stopifnot(identical(rownames(dat_imp), meta2$record_id))

# PCA outlier detection
pca0 <- prcomp(dat_imp, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2, drop = FALSE])
scores$record_id <- rownames(scores)

scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

threshold <- mean(scores$dist, na.rm = TRUE) + 
  3 * sd(scores$dist, na.rm = TRUE)

outliers <- scores$record_id[scores$dist > threshold]

outliers

# remove outliers from matrix
dat_imp2 <- dat_imp[
  !rownames(dat_imp) %in% outliers,, drop = FALSE]

# remove outliers from metadata
meta3 <- meta2 %>%
  dplyr::filter(!record_id %in% outliers)

# reorder metadata to match matrix
meta3 <- meta3[match(rownames(dat_imp2), meta3$record_id), ]

# checks
cat("Still present in metadata? ", any(outliers %in% meta3$record_id), "\n")
cat("Still present in matrix? ", any(outliers %in% rownames(dat_imp2)), "\n")

stopifnot(!any(outliers %in% meta3$record_id))
stopifnot(!any(outliers %in% rownames(dat_imp2)))
stopifnot(identical(rownames(dat_imp2), meta3$record_id))
```

```{r}
meta3 <- meta3 %>%
  dplyr::mutate(
    group_bmi = paste(group, bmi_obesity, sep = "_")
  )

meta3$group_bmi <- factor(meta3$group_bmi)

table(meta3$group_bmi)
```




#Data tables
```{r}
table_study_bmi <- meta3 %>%
  count(study, bmi_obesity) %>%
  pivot_wider(
    names_from = bmi_obesity,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(Total = rowSums(across(where(is.numeric))))

table_study_bmi %>%
  kable(
    caption = "Participants by study and BMI adiposity group"
  ) %>%
  kable_styling(full_width = FALSE)
```

```{r}
study_group_bmi_table <- meta3 %>%
  dplyr::count(study, group, bmi_obesity, name = "n") %>%
  tidyr::unite("group_bmi", group, bmi_obesity, sep = "_") %>%
  tidyr::pivot_wider(
    names_from = group_bmi,
    values_from = n,
    values_fill = 0
  ) %>%
  dplyr::mutate(
    Total = rowSums(dplyr::across(where(is.numeric)))
  )

study_group_bmi_table
```
#---------------
# Plotting Funciton
#---------------
```{r}
make_pca_plot <- function(mat, meta_df, title_str) {
  pca  <- prcomp(mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1    = pca$x[, 1],
    PC2    = pca$x[, 2],
    group = meta_df$group,
    study  = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = group, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        LC  = "Cyan4",
        T1D = "DarkRed"
      ),
      name = "Group"
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
##-----------------------
## PCAs by disease status
##-----------------------
```{r}
design <- model.matrix(~ group, data = meta3)
p  <- make_pca_plot(dat_imp2, meta3, "No adjustment")

cov_m1 <- model.matrix(~ age + sex, data = meta3)[, -1, drop = FALSE]
prot_m1 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m1, design = design))
p1 <- make_pca_plot(prot_m1, meta3, "Age + Sex")


cov_m2 <- model.matrix(~ age + sex + study, data = meta3)[, -1, drop = FALSE]
prot_m2 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m2, design = design))
p2 <- make_pca_plot(prot_m2, meta3, "Age + Sex + Study")

pca_disease_status <- (p | p1 | p2) 
  ggsave(pca_disease_status,
         plot = pca_disease_status,
         path = file.path(dir.results),
         width = 10,
         height = 8)
```

```{r}
make_pca_plot2 <- function(mat, meta_df, title_str) {
  pca  <- prcomp(mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1    = pca$x[, 1],
    PC2    = pca$x[, 2],
    bmi_obesity = meta_df$bmi_obesity,
    study  = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = bmi_obesity, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        Overweight  = "cyan4",
        Obese = "plum4",
        Normal = "orange3"
      ),
      name = "BMI group"
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
##------------------
## PCAs by bmi group
##------------------
```{r}
design <- model.matrix(~ bmi_obesity, data = meta3)

p  <- make_pca_plot2(dat_imp2, meta3, "No adjustment")

cov_m1 <- model.matrix(~ age + sex, data = meta3)[, -1, drop = FALSE]
prot_m1 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m1, design = design))
p1 <- make_pca_plot2(prot_m1, meta3, "Age + Sex")


cov_m2 <- model.matrix(~ age + sex + study, data = meta3)[, -1, drop = FALSE]
prot_m2 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m2, design = design))
p2 <- make_pca_plot2(prot_m2, meta3, "Age + Sex + Study")

(p | p1 | p2)
```

```{r}
#setting up limma matrices
# make sure matrix and metadata match
stopifnot(identical(rownames(dat_imp2), meta3$record_id))

#neet to transform for limma
mat1 <- t(dat_imp2)

# set "normal" as reference group
meta3$bmi_obesity <- factor(
  meta3$bmi_obesity,
  levels = c("Normal", "Overweight", "Obese")
)

meta3$study <- factor(meta3$study)
```


#---------------------------------------------
#run limma between disease status T1D vs LC
#---------------------------------------------
```{r}
design_m1 <- model.matrix(~ group + age + sex + study, data = meta3) #add/remove study

colnames(design_m1)

# fit model
disease_m1 <- lmFit(mat1, design_m1)
disease_m1 <- eBayes(disease_m1)

# extract T1D vs LC
res_disease_m1 <- topTable(
  disease_m1,
  coef = "groupT1D",
  number = Inf,
  adjust.method = "BH"
) %>%
  rownames_to_column("protein_id") %>%
  mutate(
    model = "Disease status: T1D vs LC + age + sex + study",
    contrast = "T1D vs LC",
    direction = case_when(
      adj.P.Val < 0.05 & logFC > 0 ~ "Higher in T1D",
      adj.P.Val < 0.05 & logFC < 0 ~ "Lower in T1D",
      TRUE ~ "Not significant"
    )
  )

table(res_disease_m1$direction)
```
#---------------------------------------------
#run limma for adiposity across disease status
#---------------------------------------------
```{r}
# Model 1: BMI obesity + age + sex
design_m1 <- model.matrix(~ 0 + bmi_obesity + age + sex + study, data = meta3) #add/remove study
colnames(design_m1)

contrast_matrix <- makeContrasts(
  Overweight_vs_Normal = bmi_obesityOverweight - bmi_obesityNormal,
  Obese_vs_Normal = bmi_obesityObese - bmi_obesityNormal,
  Obese_vs_Overweight = bmi_obesityObese - bmi_obesityOverweight,
  levels = design_m1
)


m <- lmFit(mat1, design_m1)
m1 <- contrasts.fit(m, contrast_matrix)
m1 <- eBayes(m1)

results_list <- lapply(colnames(contrast_matrix), function(cn) {
  topTable(m1, coef = cn, number = Inf, adjust.method = "BH") %>%
    rownames_to_column("protein_id") %>%
    mutate(
      contrast = cn,
      direction = case_when(
        adj.P.Val < 0.05 & logFC > 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][1]),
        adj.P.Val < 0.05 & logFC < 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][2]),
        TRUE ~ "Not significant"
      )
    )
})

results_bmi_all <- bind_rows(results_list)
results_sig <- results_bmi_all %>%
  dplyr::filter(
    adj.P.Val < 0.05,
    abs(logFC) > 0
  )
results_sig
```

#------------------------------------
#limma with only T1D Adiposity Groups
#------------------------------------
```{r}
#subsetting to only T1D group
meta_t1d <- meta3 %>%
  dplyr::filter(group == "T1D") %>%
  dplyr::filter(!is.na(bmi_obesity), !is.na(age), !is.na(sex), !is.na(study))

dat_t1d <- dat_imp2[meta_t1d$record_id, , drop = FALSE]

# reorder metadata to match matrix
meta_t1d <- meta_t1d[match(rownames(dat_t1d), meta_t1d$record_id), ]

stopifnot(identical(rownames(dat_t1d), meta_t1d$record_id))
```

```{r}
#resetting groups as factors
meta_t1d$bmi_obesity <- factor(
  meta_t1d$bmi_obesity,
  levels = c("Normal", "Overweight", "Obese")
)

meta_t1d$sex   <- factor(meta_t1d$sex)
meta_t1d$study <- factor(meta_t1d$study)

table(meta_t1d$bmi_obesity)
table(meta_t1d$study)
```

```{r}
mat_t1d <- t(dat_t1d)

design_t1d <- model.matrix(~ 0 + bmi_obesity + age + sex + study,data = meta_t1d
)

# optional, but protects against invalid names
colnames(design_t1d) <- make.names(colnames(design_t1d))

colnames(design_t1d)

contrast_t1d <- makeContrasts(
  Overweight_vs_Normal =
    bmi_obesityOverweight - bmi_obesityNormal,
  Obese_vs_Normal =
    bmi_obesityObese - bmi_obesityNormal,
  Obese_vs_Overweight =
    bmi_obesityObese - bmi_obesityOverweight,
  levels = design_t1d
)
```
```{r}
fit_t1d <- lmFit(mat_t1d, design_t1d)

fit_t1d2 <- contrasts.fit(fit_t1d, contrast_t1d)
fit_t1d2 <- eBayes(fit_t1d2)
```

```{r}
results_t1d_bmi_all <- lapply(colnames(contrast_t1d), function(cn) {
  topTable(
    fit_t1d2,
    coef = cn,
    number = Inf,
    adjust.method = "BH"
  ) %>%
    rownames_to_column("protein_id") %>%
    mutate(
      contrast = cn,
      direction = case_when(
        adj.P.Val < 0.05 & logFC > 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][1]),
        adj.P.Val < 0.05 & logFC < 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][2]),
        TRUE ~ "Not significant"
      )
    )
}) %>%
  bind_rows()

res_t1d_bmi_sig <- results_t1d_bmi_all %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= 0) 
res_t1d_bmi_sig
```
```{r}
sig_cutoff <- 0.05
lfc_cutoff <- 0   # change to 0.5 if you want stricter filtering

combined_sig_terms <- dplyr::bind_rows(
  
  res_disease_m1 %>%
    dplyr::mutate(
      analysis = "Disease status",
      model = "T1D vs LC + age + sex + study"
    ),
  
  results_bmi_all %>%
    dplyr::mutate(
      analysis = "BMI adiposity - all participants",
      model = "BMI group + age + sex + study"
    ),
  
  results_t1d_bmi_all %>%
    dplyr::mutate(
      analysis = "BMI adiposity - T1D only",
      model = "BMI group within T1D + age + sex + study"
    )
) %>%
  dplyr::filter(
    adj.P.Val < sig_cutoff,
    abs(logFC) > lfc_cutoff
  ) %>%
  dplyr::mutate(
    direction_simple = dplyr::case_when(
      logFC > 0 ~ "Positive logFC",
      logFC < 0 ~ "Negative logFC",
      TRUE ~ "No change"
    )
  ) %>%
  dplyr::select(
    analysis,
    model,
    contrast,
    protein_id,
    logFC,
    AveExpr,
    t,
    P.Value,
    adj.P.Val,
    B,
    direction,
    direction_simple
  ) %>%
  dplyr::arrange(analysis, contrast, adj.P.Val)

combined_sig_terms_n <- combined_sig_terms %>%
  dplyr::count(analysis, contrast, direction, name = "n_sig")
combined_sig_terms_n


write.csv(
  combined_sig_terms,
  file = file.path(dir.results, "limma_results.csv"),
  row.names = FALSE
)

write.csv(
  combined_sig_terms_n,
  file = file.path(dir.results, "sig_DEP_counts.csv"),
  row.names = FALSE
)
```


#clinical maker correlation 
#correlate with transcriptomics
T1D_normal vs LC
T1D_over_obese vs T1D normal
combined overweight/obese