---
title: "05_proteomics_analysis"
author: "Ye Ji Choi & Savanah Leidholt"
format: html
---

```{r}
library(dplyr)
library(tidyr)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggforce)
library(enrichR)
library(stringr)
library(limma)
library(EnhancedVolcano)
library(patchwork)
library(ggrepel)
library(vegan)
library(tidyverse)
library(data.table)
library(purrr)
library(haven)
library(ggVennDiagram)
library(caret)
library(kableExtra)
library(msigdbr)
library(fgsea)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/T1D Adiposity/Results/Compiled Results/Proteomics Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/T1D Adiposity/Results/Compiled Results/Proteomics Results")
    git_path <- "/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad"
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
} else if (user == "hhampson") {
  root_path <- "/Volumes/Peds Endo"
} else {
  stop("Unknown user: please specify root path for this user.")
}
```


```{r}
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```

#no dxa on attempt
#also do waist to hip = wthr, wthr_obesity
#hba1c with age
#check mrns to see if there are people are co-enrolled

```{r}
t1d_hc_dat_soma <- s3readRDS(object = "data_clean/t1d_hc_clinical_data_soma.csv", 
                             bucket = "t1d.adiposity",
                             region = "")
```

#-------------------------
#creating cleaned metadata
#-------------------------
```{r}
meta <- t1d_hc_dat_soma %>%
  dplyr::select(-starts_with("seq")) %>%
  dplyr::mutate(
    record_id = as.character(record_id),
    group = factor(
      group,
      levels = c("Lean Control", "Type 1 Diabetes"),
      labels = c("LC", "T1D")
    ),
    study = dplyr::recode(
      study,
      "CROCODILE"  = "CRC",
      "PANDA"      = "PAN",
      "RENAL-HEIR" = "RH",
      "CASPER"     = "CAS",
      "ATTEMPT"    = "ATT",
      .default = study
    ),
    bmi_obesity = factor(
      bmi_obesity,
      levels = c("Normal", "Overweight", "Obese")
    ),
    bmi_bin = dplyr::case_when(
      bmi_obesity == "Normal" ~ "Normal",
      bmi_obesity %in% c("Overweight", "Obese") ~ "Overweight_Obese",
      TRUE ~ NA_character_
    ),
    group_bmi = dplyr::case_when(
      group == "LC"  & bmi_obesity == "Normal" ~ "LC_Normal",
      group == "LC"  & bmi_obesity %in% c("Overweight", "Obese") ~ "LC_Overweight_Obese",
      group == "T1D" & bmi_obesity == "Normal" ~ "T1D_Normal",
      group == "T1D" & bmi_obesity == "Overweight" ~ "T1D_Overweight",
      group == "T1D" & bmi_obesity == "Obese" ~ "T1D_Obese",
      TRUE ~ NA_character_
    )
  )

meta$study <- factor(meta$study)
meta$bmi_bin <- factor(meta$bmi_bin, levels = c("Normal", "Overweight_Obese"))
meta$group_bmi <- factor(
  meta$group_bmi,
  levels = c(
    "LC_Normal",
    "LC_Overweight_Obese",
    "T1D_Normal",
    "T1D_Overweight",
    "T1D_Obese"
  )
)

# keep BMI-complete samples
hba1c_var <- "hba1c"
meta2 <- meta %>%
 dplyr::filter(
    !is.na(record_id),
    !is.na(group_bmi),
    !is.na(age),
    !is.na(.data[[hba1c_var]]),
    !is.na(sex),
    !is.na(study)
  )

table(meta2$group_bmi, useNA = "ifany")
```
#----------------------------
# Plasma protein matrix setup
#----------------------------
```{r}
dat_plasma <- t1d_hc_dat_soma %>%
  dplyr::select(record_id, starts_with("seq") &
                  !dplyr::ends_with("urine") &
                  !dplyr::ends_with("urine_cradj"))  %>% as.data.frame()

rownames(dat_plasma) <- dat_plasma$record_id
dat_plasma$record_id <- NULL

# align metadata and matrix
ids_shared <- intersect(meta2$record_id, rownames(dat_plasma))

dat_plasma2 <- dat_plasma[ids_shared, , drop = FALSE]

meta2 <- meta2 %>%
  dplyr::filter(record_id %in% ids_shared)

meta2 <- meta2[match(rownames(dat_plasma2), meta2$record_id), ]

stopifnot(identical(rownames(dat_plasma2), meta2$record_id))
```

#--------------------------------------
# Filter to SomaScan 7000 analyte panel
#--------------------------------------
```{r}
analytes <- read.csv(
  file.path(root_path, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

prot_panel <- analytes$AptName %>%
  as.character() %>%
  trimws() %>%
  unique()

prot_panel <- prot_panel[!is.na(prot_panel) & prot_panel != ""]

aptamers_keep <- intersect(prot_panel, colnames(dat_plasma2))

if (length(aptamers_keep) == 0) {
  stop("No analytes from analytes.csv matched columns in dat_plasma2.")
}

dat_plasma3 <- dat_plasma2[, aptamers_keep, drop = FALSE]

cat("Plasma features before analyte filter:", ncol(dat_plasma2), "\n")
cat("Proteins in analyte panel:", length(prot_panel), "\n")
cat("Matched proteins kept:", length(aptamers_keep), "\n")
cat("Panel proteins missing from plasma matrix:", length(setdiff(prot_panel, colnames(dat_plasma2))), "\n")

dat_plasma3 <- as.data.frame(dat_plasma3)

dat_plasma3[] <- lapply(dat_plasma3, function(x) {
  suppressWarnings(as.numeric(as.character(x)))
})

dat_plasma3 <- as.matrix(dat_plasma3)

dat_log <- log2(dat_plasma3 + 1)
dat_log[!is.finite(dat_log)] <- NA_real_

stopifnot(identical(rownames(dat_log), meta2$record_id))
```
#----------------------------------------------------
# Study presence filter
# Protein present in >=60% samples in >=2 studies
# doesn't remove anything in this case but a good check
#------------------------------------------------------
```{r}
meta2$study <- factor(meta2$study)

study_idx_all <- split(seq_len(nrow(dat_log)), meta2$study)

present_by_study_all <- sapply(study_idx_all, function(idx) {
  colMeans(!is.na(dat_log[idx, , drop = FALSE])) >= 0.6
})

present_by_study_all <- as.matrix(present_by_study_all)

n_studies_present_all <- rowSums(present_by_study_all)
keep_study_all <- n_studies_present_all >= 2

kept_proteins <- colnames(dat_log)[keep_study_all]
removed_proteins <- colnames(dat_log)[!keep_study_all]

cat("All proteomic features before study filter:", ncol(dat_log), "\n")
cat("Removed:", length(removed_proteins), "\n")
cat("Kept:", length(kept_proteins), "\n")

dat_log_filt <- dat_log[, kept_proteins, drop = FALSE]

stopifnot(identical(rownames(dat_log_filt), meta2$record_id))
```
#-------------------------
#imputing the missing data
#-------------------------
```{r}
set.seed(1835)

dat_imp <- dat_log_filt

meds <- apply(dat_imp, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(dat_imp))) {
  miss_j <- is.na(dat_imp[, j])
  if (!is.na(meds[j]) && any(miss_j)) {
    dat_imp[miss_j, j] <- meds[j]
  }
}

vars <- apply(dat_imp, 2, var)

keep_cols <- colSums(is.na(dat_imp)) == 0 &
  is.finite(vars) &
  vars > 0

dat_imp <- dat_imp[, keep_cols, drop = FALSE]

stopifnot(identical(rownames(dat_imp), meta2$record_id))
stopifnot(sum(is.na(dat_imp)) == 0)
```

#--------------------------
#checking for outliers
#--------------------------
```{r}
pca0 <- prcomp(dat_imp, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2, drop = FALSE])
scores$record_id <- rownames(scores)
scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

threshold <- mean(scores$dist, na.rm = TRUE) +
  3 * sd(scores$dist, na.rm = TRUE)

outliers <- scores$record_id[scores$dist > threshold] #2 outliers CRC-42 & 10424
outliers
dat_imp2 <- dat_imp[!rownames(dat_imp) %in% outliers, , drop = FALSE]

meta3 <- meta2 %>%
  dplyr::filter(!record_id %in% outliers)

meta3 <- meta3[match(rownames(dat_imp2), meta3$record_id), ]

stopifnot(identical(rownames(dat_imp2), meta3$record_id))
```
#----------------------------
#Data table: study x group_bmi
#-----------------------------
```{r}
study_group_bmi_table <- meta3 %>%
  dplyr::count(study, group_bmi, name = "n") %>%
  tidyr::pivot_wider(
    names_from = group_bmi,
    values_from = n,
    values_fill = 0
  ) %>%
  dplyr::mutate(
    Total = rowSums(dplyr::across(where(is.numeric)))
  )

study_group_bmi_table %>%
  kable(caption = "Participants by study and disease-adiposity group") %>%
  kable_styling(full_width = FALSE)
```
#-------------------------------------
#PCA plotting function for group_bmi
#-------------------------------------
```{r}
make_pca <- function(mat, meta_df, title_str) {
  stopifnot(identical(rownames(mat), meta_df$record_id))

  pca <- prcomp(mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1 = pca$x[, 1],
    PC2 = pca$x[, 2],
    group_bmi = meta_df$group_bmi,
    study = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = group_bmi, shape = study)) +
    scale_color_manual(
      values = c(
        LC_Normal  = "cyan4",
        LC_Overweight_Obese = "darkred",
        T1D_Normal = "gold",
        T1D_Overweight = "plum3",
        T1D_Obese = "black"
      ),
      name = "Group"
    ) +
    geom_point(size = 3, alpha = 0.85) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)"),
      color = "Disease-adiposity group"
    ) +
    theme_bw()
}
```

```{r}
stopifnot(identical(rownames(dat_imp2), meta3$record_id))
design <- model.matrix(~ group_bmi, data = meta3)

#model 0 ~ no adjustment
p0 <- make_pca(dat_imp2,meta3,"No adjustment")

#model 1 ~ age + sex adjustment
cov_age_sex <- model.matrix(~ age + sex, data = meta3)[, -1, drop = FALSE]
prot_age_sex <- t(removeBatchEffect(t(dat_imp2),
  covariates = cov_age_sex,
  design = design
))
p1 <- make_pca(prot_age_sex,meta3,"Adjusted for age + sex")

#model 2 ~ age + sex + study adjustment
cov_age_sex_study <- model.matrix(~ age + sex + study, data = meta3)[, -1, drop = FALSE]
prot_age_sex_study <- t(removeBatchEffect(t(dat_imp2),
  covariates = cov_age_sex_study,
  design = design
))
p2 <- make_pca(prot_age_sex_study, meta3, "Adjusted for age + sex + study")

#model 3 ~ age * HbA1c + sex + study adjustment
cov_hba1c_age <- model.matrix(
  as.formula(paste0("~ age * ", hba1c_var, " + sex + study")),
  data = meta3
)[, -1, drop = FALSE]

# final row checks
nrow(design)
nrow(cov_hba1c_age)
stopifnot(nrow(design) == nrow(cov_hba1c_age))
stopifnot(nrow(design) == nrow(dat_imp2))

prot_hba1c_age <- t(removeBatchEffect(t(dat_imp2), covariates = cov_hba1c_age, design = design))
p3 <- make_pca(prot_hba1c_age, meta3,"Adjusted for age * HbA1c + sex + study")

# combine plots
pca_group_bmi_all <- (p0 | p1) / (p2 | p3)
pca_group_bmi_all
```

```{r}
mat1 <- t(dat_imp2)

run_group_bmi_limma <- function(mat, meta_df, design_formula, model_label) {
  
  design <- model.matrix(design_formula, data = meta_df)
  colnames(design) <- make.names(colnames(design))
  
  message("Design columns for ", model_label, ":")
  print(colnames(design))
  
  contrast_matrix <- makeContrasts(
    T1DNormal_vs_LCNormal =
      group_bmiT1D_Normal - group_bmiLC_Normal,
    T1DOverObese_vs_LCOverObese =
      ((group_bmiT1D_Overweight + group_bmiT1D_Obese) / 2) -
      group_bmiLC_Overweight_Obese,
    T1DNormal_vs_T1DOverweight =
      group_bmiT1D_Normal - group_bmiT1D_Overweight,
    T1DNormal_vs_T1DObese =
      group_bmiT1D_Normal - group_bmiT1D_Obese,
    T1DObese_vs_T1DOverweight =
      group_bmiT1D_Obese - group_bmiT1D_Overweight,
    
    levels = design
  )
  
  fit <- lmFit(mat, design)
  fit2 <- contrasts.fit(fit, contrast_matrix)
  fit2 <- eBayes(fit2)
  
  
  results <- lapply(colnames(contrast_matrix), function(cn) {
    topTable(
      fit2,
      coef = cn,
      out = out,
      number = Inf,
      adjust.method = "BH"
    ) %>%
      rownames_to_column("protein_id") %>%
      mutate(
        model = model_label,
        contrast = cn,
        direction = case_when(
          adj.P.Val < 0.05 & logFC > 0 ~ "Higher in first group",
          adj.P.Val < 0.05 & logFC < 0 ~ "Higher in second group",
          TRUE ~ "Not significant"
        )
      )
  }) %>%
    bind_rows()
  
  return(results)
}
```



```{r}
# Model 0: group_bmi only / no covariate adjustment
res_m0 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi,
  model_label = "M0: group_bmi only"
)

# Model 1: group_bmi + age + sex
res_m1 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi + age + sex,
  model_label = "M1: group_bmi + age + sex"
)

#Model 2: group_bmi + age + sex + study
res_m2 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi + age + sex + study,
  model_label = "M2: group_bmi + age + sex + study"
)


# Model 3: group_bmi + age * HbA1c + sex + study
design2 <- as.formula(
  paste0("~ 0 + group_bmi + age * ", hba1c_var, " + sex + study")
)
res_m3 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = design2,
  model_label = "M3: group_bmi + age * HbA1c + sex + study"
)
```
```{r}
results_all_models <- bind_rows(
  res_m0,
  res_m1,
  res_m2,
  res_m3
)

sig_all_models <- results_all_models %>% filter(adj.P.Val < 0.05)
sig_all_models_fc <- results_all_models %>% filter(adj.P.Val < 0.05, abs(logFC) > 0.5)

sig_counts <- sig_all_models %>% 
  filter((adj.P.Val < 0.05 & abs(logFC) > 0.5)) %>%
  count(model, contrast, direction)
sig_counts
sig_counts %>% filter(model == "M3: group_bmi + age * HbA1c + sex + study")

write_csv(sig_counts, 
          file = file.path(dir.results, "sig_counts_by_model.csv"))
write_csv(sig_all_models_fc, 
          file = file.path(dir.results, "limma_results_all_models.csv"))
```

```{r}
#| fig-width: 20
#| fig-height: 18

make_volcano <- function(results_df,
                         contrast_name,
                         title_prefix = NULL,
                         p_cutoff = 0.05,
                         fc_cutoff = 0.5,
                         label_n = 15) {
  
  df <- results_df %>%
    dplyr::filter(contrast == contrast_name) %>%
    dplyr::mutate(
      neg_log10_adj_p = -log10(adj.P.Val),
      sig_group = dplyr::case_when(
        adj.P.Val < p_cutoff & logFC > fc_cutoff  ~ "Up",
        adj.P.Val < p_cutoff & logFC < -fc_cutoff ~ "Down",
        TRUE ~ "Not significant"
      )
    )
  
  n_up <- sum(df$sig_group == "Up", na.rm = TRUE)
  n_down <- sum(df$sig_group == "Down", na.rm = TRUE)
  n_ns <- sum(df$sig_group == "Not significant", na.rm = TRUE)
  
  legend_labs <- c(
    "Up" = paste0("Upregulated (n = ", n_up, ")"),
    "Down" = paste0("Downregulated (n = ", n_down, ")"),
    "Not significant" = paste0("Not significant (n = ", n_ns, ")")
  )
  
  top_labels <- df %>%
    dplyr::arrange(dplyr::desc(abs(t))) %>%
    dplyr::slice_head(n = label_n)
  
  ggplot(df, aes(x = logFC, y = neg_log10_adj_p)) +
    
    geom_point(aes(color = sig_group), alpha = 0.75, size = 2) +
    
    geom_vline(xintercept = c(-fc_cutoff, fc_cutoff), linetype = "dashed") +
    geom_vline(xintercept = 0, linetype = "solid", alpha = 0.5) +
    
    geom_hline(yintercept = -log10(p_cutoff), linetype = "dashed") +
    
    ggrepel::geom_text_repel(
      data = top_labels,
      aes(label = protein_id),
      size = 3.5,
      max.overlaps = Inf
    ) +
    
    scale_color_manual(
      values = c(
        "Up" = "darkred",
        "Down" = "cyan4",
        "Not significant" = "grey70"
      ),
      labels = legend_labs
    ) +
    
    coord_cartesian(xlim = c(-2, 3)) +   # ← THIS IS THE KEY LINE
    
    scale_x_continuous(breaks = seq(-2, 3, by = 1)) +
    
    labs(
      title = paste0(title_prefix, "\n", contrast_name),
      x = "log2 Fold Change",
      y = "-log10 adjusted P-value",
      color = NULL
    ) +
    
    theme_bw() +
    theme(
      panel.grid = element_blank(),
      legend.position = "right"
    )
}
```


```{r}
make_volcano_list <- function(results_df, model_name, label_n = 15) {
  
  contrasts <- unique(results_df$contrast)
  
  plot_list <- lapply(contrasts, function(ct) {
    make_volcano(
      results_df = results_df,
      contrast_name = ct,
      title_prefix = model_name,
      label_n = label_n
    )
  })
  
  names(plot_list) <- contrasts
  
  plot_list
}
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m0 <- make_volcano_list(
  results_df = res_m0,
  model_name = "M0: group_bmi only",
  label_n = 5
)

top_row <- plots_m0[[1]] + plots_m0[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m0[[3]] + plots_m0[[4]] + plots_m0[[5]]

panel_m0_5 <- top_row / bottom_row

panel_m0_5

ggsave("M0_pca_plots.png",
       plot = panel_m0_5,
       path = dir.results,
       width = 16,
       height =12,
       dpi = 300
       )
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m1 <- make_volcano_list(
  results_df = res_m1,
  model_name = "M1: group_bmi + age + sex",
  label_n = 5
)

top_row <- plots_m1[[1]] + plots_m1[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m1[[3]] + plots_m1[[4]] + plots_m1[[5]]

panel_m1_5 <- top_row / bottom_row

panel_m1_5

ggsave("M1_pca_plots.png",
       plot = panel_m1_5,
       path = dir.results,
       width = 16,
       height =12,
       dpi = 300
       )

```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m2 <- make_volcano_list(
  results_df = res_m2,
  model_name = "M2: group_bmi + age + sex + study",
  label_n = 5
)

top_row <- plots_m2[[1]] + plots_m2[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m2[[3]] + plots_m2[[4]] + plots_m2[[5]]

panel_m2_5 <- top_row / bottom_row

panel_m2_5
ggsave("M2_pca_plots.png",
       plot = panel_m2_5,
       path = dir.results,
       width = 16,
       height =12,
       dpi = 300
       )
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m3 <- make_volcano_list(
  results_df = res_m3,
  model_name = "M3: group_bmi + age*hba1c + sex + study",
  label_n = 5
)

top_row <- plots_m3[[1]] + plots_m3[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m3[[3]] + plots_m3[[4]] + plots_m3[[5]]

panel_m3_5 <- top_row / bottom_row

panel_m3_5
ggsave("M3_pca_plots.png",
       plot = panel_m3_5,
       path = dir.results,
       width = 16,
       height =12,
       dpi = 300
       )
```
#-----------------
#pathway anlaysis
#-----------------
```{r}

msig_kegg <- msigdbr(
  species = "Homo sapiens",
  category = "C2",
  subcategory = "CP:REACTOME"
)

kegg_list <- msig_kegg %>%
  split(x = .$gene_symbol, f = .$gs_name)

# GO Biological Process
msig_go_bp <- msigdbr(
  species = "Homo sapiens",
  category = "C5",
  subcategory = "GO:BP"
)

go_bp_list <- msig_go_bp %>%
  split(x = .$gene_symbol, f = .$gs_name)

# GO Molecular Function
msig_go_mf <- msigdbr(
  species = "Homo sapiens",
  category = "C5",
  subcategory = "GO:MF"
)

go_mf_list <- msig_go_mf %>%
  split(x = .$gene_symbol, f = .$gs_name)

# GO Cellular Component
msig_go_cc <- msigdbr(
  species = "Homo sapiens",
  category = "C5",
  subcategory = "GO:CC"
)

go_cc_list <- msig_go_cc %>%
  split(x = .$gene_symbol, f = .$gs_name)
```

```{r}
protein_gene_map <- analytes %>%
  dplyr::mutate(
    protein_id = trimws(as.character(AptName)),
    gene_symbol = trimws(as.character(EntrezGeneSymbol))  # change if needed
  ) %>%
  dplyr::select(protein_id, gene_symbol) %>%
  dplyr::filter(
    !is.na(protein_id),
    protein_id != "",
    !is.na(gene_symbol),
    gene_symbol != ""
  ) %>%
  dplyr::distinct()

results_limma_annot <- results_all_models %>%
  dplyr::left_join(protein_gene_map, by = "protein_id")

#mapping success rate
results_limma_annot %>%
  summarise(
    n_total = n(),
    n_mapped = sum(!is.na(gene_symbol)),
    pct_mapped = round(100 * mean(!is.na(gene_symbol)), 1)
  )
```
#--------------
#fgsea funciton
#--------------
```{r}
run_fgsea_model <- function(res_df,
                            model_name,
                            pathways,
                            pathway_type,
                            protein_gene_map,
                            min_size = 10,
                            max_size = 500,
                            eps = 0,
                            nPermSimple = 10000) {
  
  res_annot <- res_df %>%
    dplyr::left_join(protein_gene_map, by = "protein_id") %>%
    dplyr::filter(
      !is.na(gene_symbol),
      gene_symbol != "",
      !is.na(t)
    )
  
  fgsea_out <- res_annot %>%
    dplyr::group_by(contrast) %>%
    dplyr::group_split() %>%
    purrr::map_dfr(function(df) {
      
      contrast_name <- unique(df$contrast)
      
      ranks <- df %>%
        dplyr::group_by(gene_symbol) %>%
        dplyr::slice_max(order_by = abs(t), n = 1, with_ties = FALSE) %>%
        dplyr::ungroup() %>%
        dplyr::arrange(desc(t))
      
      ranks_vec <- ranks$t
      names(ranks_vec) <- ranks$gene_symbol
      
      fgsea::fgsea(
        pathways = pathways,
        stats = ranks_vec,
        minSize = min_size,
        maxSize = max_size
      ) %>%
        tibble::as_tibble() %>%
        dplyr::mutate(
          model = model_name,
          contrast = contrast_name,
          pathway_type = pathway_type
        )
    })
  
  fgsea_out
}
```


```{r}
#kegg results
fgsea_kegg_m0 <- run_fgsea_model(res_df = res_m0, model_name = "M0: group_bmi only", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m0 <- fgsea_kegg_m0 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_kegg_m1 <- run_fgsea_model(res_df = res_m1, model_name = "M1: group_bmi age + sex", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m1 <- fgsea_kegg_m1 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_kegg_m2 <- run_fgsea_model(res_df = res_m2, model_name = "M2: group_bmi age + sex + study", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m2 <- fgsea_kegg_m2 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_kegg_m3 <- run_fgsea_model(res_df = res_m3, model_name = "M3: group_bmi age*hba1c + sex +study", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m3 <- fgsea_kegg_m3 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

#GO:BP

#GO:MF

#GO:CC
```

#-----------------
#plotting function
#-----------------
```{r}
plot_fgsea_top20 <- function(fgsea_df,
                             model_name,
                             contrast_name,
                             pathway_type_name = NULL,
                             top_n = 20,
                             padj_cutoff = 0.05) {
  
  df <- fgsea_df %>%
    filter(
      model == model_name,
      contrast == contrast_name
    )
  
  if (!is.null(pathway_type_name)) {
    df <- df %>%
      filter(pathway_type == pathway_type_name)
  }
  
  df_top <- df %>%
    filter(!is.na(padj), !is.na(NES)) %>%
    arrange(padj) %>%
    slice_head(n = top_n) %>%
    mutate(
      pathway_clean = pathway %>%
        str_replace_all("_", " ") %>%
        str_replace("^GO ", "") %>%
        str_to_sentence(),
      
      pathway_clean = forcats::fct_reorder(pathway_clean, NES),
      
      direction = ifelse(NES > 0, "Up", "Down"),
      
      # transform FDR so smaller = bigger
      fdr_size = -log10(padj)
    )
  
  ggplot(df_top, aes(y = pathway_clean)) +
    geom_segment(
      aes(x = 0, xend = NES, yend = pathway_clean),
      color = "grey70",
      linewidth = 0.6
    ) +
    geom_point(
      aes(x = NES, size = fdr_size, color = direction)
    ) +
    geom_text(
      aes(x = 0, label = pathway_clean),
      hjust = ifelse(df_top$NES > 0, 1, 0),
      size = 3
    ) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    
    scale_color_manual(
      values = c(
        "Up" = "darkred",
        "Down" = "cyan4"
      ),
      name = "Direction"
    ) +
    
    scale_size_continuous(
      name = "-log10(FDR)"
    ) +
    
    scale_x_continuous(
      limits = c(-4, 4),
      breaks = seq(-4, 4, by = 2)
    ) +
    
    labs(
      title = paste(model_name, contrast_name, pathway_type_name, sep = "\n"),
      x = "Normalized enrichment score",
      y = NULL
    ) +
    
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      panel.border = element_blank(),
      axis.line = element_line(color = "black"),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      plot.title = element_text(face = "bold", size = 12),
      legend.position = "right"
    )
}
```

```{r}
#| fig-width: 22
#| fig-height: 16
fgsea_kegg_plots_m0 <- fgsea_kegg_m0 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m0,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

#fgsea_kegg_plots_m0$plot[[1]]
#fgsea_kegg_plots_m0$plot[[2]]
#fgsea_kegg_plots_m0$plot[[3]]
#fgsea_kegg_plots_m0$plot[[4]]
#fgsea_kegg_plots_m0$plot[[5]]

plots_m0 <- fgsea_kegg_plots_m0
top_row <- plots_m0$plot[[1]] + plots_m0$plot[[2]] + patchwork::plot_spacer()
bottom_row <- plots_m0$plot[[3]] + plots_m0$plot[[4]] + plots_m0$plot[[5]]

panel_kegg_m0 <- top_row / bottom_row

panel_kegg_m0

ggsave("GSEA_kegg_m0.png",
       plot = panel_kegg_m0,
       path = dir.results,
       width = 22,
       height =16,
       dpi = 300
       )
```

```{r}
#| fig-width: 18
#| fig-height: 16
fgsea_kegg_plots_m1 <- fgsea_kegg_m1 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m1,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

plots_m1 <- fgsea_kegg_plots_m1
top_row <- plots_m1$plot[[1]] + plots_m1$plot[[5]] + patchwork::plot_spacer()
bottom_row <- plots_m1$plot[[3]] + plots_m1$plot[[2]] + plots_m1$plot[[4]]

panel_kegg_m1 <- top_row / bottom_row

panel_kegg_m1

ggsave("GSEA_kegg_m1.png",
       plot = panel_kegg_m1,
       path = dir.results,
       width = 20,
       height =16,
       dpi = 300
       )
```

```{r}
#| fig-width: 20
#| fig-height: 16
fgsea_kegg_plots_m2 <- fgsea_kegg_m2 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m2,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

plots_m2 <- fgsea_kegg_plots_m2
top_row <- plots_m2$plot[[1]] + plots_m2$plot[[5]] + patchwork::plot_spacer()
bottom_row <- plots_m2$plot[[3]] + plots_m2$plot[[2]] + plots_m2$plot[[4]]

panel_kegg_m2 <- top_row / bottom_row

panel_kegg_m2

ggsave("GSEA_kegg_m2.png",
       plot = panel_kegg_m2,
       path = dir.results,
       width = 20,
       height =16,
       dpi = 300
       )

```

```{r}
#| fig-width: 20
#| fig-height: 16
fgsea_kegg_plots_m3 <- fgsea_kegg_m3 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m3,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

plots_m3 <- fgsea_kegg_plots_m3
plots_m3
top_row <- plots_m3$plot[[1]] + plots_m3$plot[[5]] + patchwork::plot_spacer()
bottom_row <- plots_m3$plot[[3]] + plots_m3$plot[[2]] + plots_m3$plot[[4]] 

panel_kegg_m3 <- top_row / bottom_row

panel_kegg_m3

ggsave("GSEA_kegg_m3.png",
       plot = panel_kegg_m3,
       path = dir.results,
       width = 20,
       height =16,
       dpi = 300
       )
```

#----------------------------------------
#ElasticNet/LASSO Regression
#----------------------------------------
```{r}
run_elastic_net <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  protein2symbol,
  top_n = 100,
  covars = c("age", "sex", "study"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1,
  min_per_class = 5,
  maxit = 1e6
) {
  
  cat("\n============================\n")
  cat("Running contrast:", cc, "\n")
  
  if (!cc %in% names(out_list)) {
    stop("Contrast not found in out_list: ", cc)
  }
  
  if (!grepl("_vs_", cc)) {
    stop("Contrast name must contain '_vs_'")
  }
  
  parts <- strsplit(cc, "_vs_")[[1]]
  if (length(parts) != 2) {
    stop("Contrast could not be split into two groups")
  }
  
  g_pos <- parts[1]
  g_neg <- parts[2]
  
 idx <- which(meta_df$group4 %in% c(g_pos, g_neg))
meta_sub <- meta_df[idx, , drop = FALSE]
meta_sub$record_id <- as.character(meta_sub$record_id)

meta_sub$group4 <- droplevels(meta_sub$group4)

y <- ifelse(meta_sub$group4 == g_pos, 1, 0)
y <- as.numeric(y)

tab <- table(meta_sub$group4)
if (any(tab < min_per_class)) {
  stop(
    "Too few samples in at least one class for contrast: ",
    cc,
    " | counts = ",
    paste(names(tab), tab, collapse = ", ")
  )
}
  
  t <- out_list[[cc]]
  if (is.null(t)) {
    stop("Contrast table is NULL for: ", cc)
  }
  t <- as.data.frame(t)
  
  cat("Columns in contrast table:\n")
  print(names(t))
  
  if (!"feature_id" %in% names(t)) {
    fid <- rownames(t)
    if (is.null(fid)) stop("No feature_id column or rownames found")
    t$feature_id <- fid
  }
  
  if (!all(c("adj.P.Val", "P.Value") %in% names(t))) {
    stop("Contrast table is missing adj.P.Val and/or P.Value")
  }
  
  cat("Rows in limma table:", nrow(t), "\n")
  cat("Overlap with protein matrix:", sum(t$feature_id %in% colnames(X_mat)), "\n")
  
  top_feats <- t %>%
  dplyr::filter(!is.na(adj.P.Val)) %>%
  dplyr::left_join(protein2symbol, by = "feature_id") %>%
  dplyr::mutate(
    gene_symbol = dplyr::if_else(
      is.na(gene_symbol) | gene_symbol == "",
      feature_id,
      gene_symbol
    )
  ) %>%
  dplyr::filter(feature_id %in% colnames(X_mat)) %>%
  dplyr::arrange(adj.P.Val, P.Value) %>%
  dplyr::group_by(gene_symbol) %>%
  dplyr::slice_head(n = 1) %>%   # keeps best aptamer per gene
  dplyr::ungroup() %>%
  dplyr::slice_head(n = top_n) %>%
  dplyr::pull(feature_id)
  
  cat("Top overlapping features:", length(top_feats), "\n")
  cat("First few top features:\n")
  print(head(top_feats, 10))
  
  if (length(top_feats) < 5) {
    stop("Fewer than 5 overlapping top features for contrast: ", cc)
  }
  
  meta_sub <- meta_sub %>%
    dplyr::mutate(
      sex = factor(sex),
      study6 = factor(study6)
    )
  
  covars_present <- covars[covars %in% colnames(meta_sub)]
  cat("Covariates used:", paste(covars_present, collapse = ", "), "\n")
  
  if (length(covars_present) > 0) {
    form_cov <- as.formula(paste("~", paste(covars_present, collapse = " + ")))
    Z <- model.matrix(form_cov, data = meta_sub)
  } else {
    Z <- matrix(1, nrow = nrow(meta_sub), ncol = 1)
    colnames(Z) <- "(Intercept)"
  }
  
  Xp <- X_mat[meta_sub$record_id, top_feats, drop = FALSE]
  
  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"
  
  pf <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))
  
  sds <- apply(X, 2, sd)
  keep <- which(is.finite(sds) & sds > 0)
  
  X <- X[, keep, drop = FALSE]
  pf <- pf[keep]
  
  cat("Final design matrix dim:", dim(X), "\n")
  
  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))
  cat("nfolds used:", nfolds_use, "\n")
  
  if (nfolds_use < 2) {
    stop("Not enough samples per class for CV")
  }
  
  set.seed(seed)
  
  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      cv.glmnet(
        x = X,
        y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = pf,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) {
        message("glmnet failed at alpha = ", a, ": ", conditionMessage(e))
        NULL
      }
    )
    
    if (is.null(cv)) return(NULL)
    list(alpha = a, cvfit = cv)
  })
  
  fit_list <- Filter(Negate(is.null), fit_list)
  
  if (length(fit_list) == 0) {
    stop("All glmnet fits failed")
  }
  
  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]
  
  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))
  coef_df <- tibble::tibble(
    term = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>%
    dplyr::filter(beta != 0)
  
  protein_terms <- intersect(colnames(Xp), coef_df$term)
  
  p_hat <- as.numeric(
    predict(best$cvfit, newx = X, s = "lambda.1se", type = "response")
  )
  
  auc_val <- as.numeric(
    pROC::auc(pROC::roc(y, p_hat, quiet = TRUE))
  )
  
  cat("Best alpha:", best$alpha, "\n")
  cat("AUC:", auc_val, "\n")
  cat("Nonzero coefficients:", nrow(coef_df), "\n")
  cat("Selected protein terms:", length(protein_terms), "\n")
  
  list(
    contrast         = cc,
    positive         = g_pos,
    negative         = g_neg,
    best_alpha       = best$alpha,
    lambda_1se       = best$cvfit$lambda.1se,
    lambda_min       = best$cvfit$lambda.min,
    auc_train        = auc_val,
    n_input_proteins = length(top_feats),
    coef_df          = coef_df,
    protein_terms    = protein_terms,
    cvfit            = best$cvfit,
    top_feats        = top_feats
  )
}
```

```{r}
run_elastic_net_bootstrap <- function(
  cc,
  out_list,
  meta_df,
  X_mat,
  protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  n_boot = 100,
  seed = 1835,
  min_per_class = 5, 
  verbose = F) {
  set.seed(seed)
  
  boot_res <- vector("list", n_boot)
  
  for (b in seq_len(n_boot)) {
    
    cat("Bootstrap", b, "of", n_boot, "for", cc, "\n")
    
    parts <- strsplit(cc, "_vs_")[[1]]
    g_pos <- parts[1]
    g_neg <- parts[2]
    
    idx0 <- which(meta_df$group4 %in% c(g_pos, g_neg))
    meta_sub0 <- meta_df[idx0, , drop = FALSE]
    
    # stratified bootstrap: sample within each group
    idx_boot <- unlist(lapply(split(seq_len(nrow(meta_sub0)), meta_sub0$group4), function(ii) {
      sample(ii, size = length(ii), replace = TRUE)
    }))
    
    meta_boot <- meta_sub0[idx_boot, , drop = FALSE]
    meta_boot$record_id_boot <- paste0(meta_boot$record_id, "_boot", seq_len(nrow(meta_boot)))
    
    X_boot <- X_mat[meta_boot$record_id, , drop = FALSE]
    rownames(X_boot) <- meta_boot$record_id_boot
    
    meta_boot$record_id <- meta_boot$record_id_boot
    
    fit_b <- tryCatch(
      run_elastic_net(
        cc = cc,
        out_list = out_list,
        meta_df = meta_boot,
        X_mat = X_boot,
        protein2symbol = protein2symbol,
        top_n = top_n,
        covars = covars,
        alpha_grid = alpha_grid,
        seed = seed + b,
        min_per_class = min_per_class
      ),
      error = function(e) {
        message("Bootstrap failed at b = ", b, ": ", conditionMessage(e))
        NULL
      }
    )
    
    boot_res[[b]] <- fit_b
  }
  
  boot_res <- Filter(Negate(is.null), boot_res)
  
 selected_tbl <- lapply(seq_along(boot_res), function(i) {
  tibble(
    boot = i,
    contrast = cc,
    best_alpha = boot_res[[i]]$best_alpha,
    auc_train = boot_res[[i]]$auc_train,
    feature_id = unique(boot_res[[i]]$protein_terms)
  )
}) %>%
  bind_rows() %>%
  distinct(contrast, boot, feature_id, .keep_all = TRUE)
  
  stability_tbl <- selected_tbl %>%
  distinct(contrast, boot, feature_id) %>%
  count(contrast, feature_id, name = "n_selected") %>%
  mutate(
    n_boot_success = length(boot_res),
    selection_frequency = n_selected / n_boot_success
  ) %>%
  left_join(protein2symbol, by = "feature_id") %>%
  arrange(desc(selection_frequency))
  
  stability_tbl <- stability_tbl %>%
  mutate(selection_percent = 100 * selection_frequency)
  
  model_summary <- tibble(
    contrast = cc,
    n_boot_requested = n_boot,
    n_boot_success = length(boot_res),
    mean_auc = mean(sapply(boot_res, function(x) x$auc_train), na.rm = TRUE),
    median_auc = median(sapply(boot_res, function(x) x$auc_train), na.rm = TRUE),
    mean_alpha = mean(sapply(boot_res, function(x) x$best_alpha), na.rm = TRUE),
    median_alpha = median(sapply(boot_res, function(x) x$best_alpha), na.rm = TRUE)
  )
  
  list(
    contrast = cc,
    boot_res = boot_res,
    selected_tbl = selected_tbl,
    stability_tbl = stability_tbl,
    model_summary = model_summary
  )
}
```

#----------------------
#boostrapping T1D vs LC
#----------------------
```{r}
boot_1 <- run_elastic_net_bootstrap(
  cc = "T1D_Normal_vs_LC_Normal",
  out_list = out1,
  meta_df = meta1,
  X_mat = prot_resid,
  protein2symbol = protein2symbol,
  top_n = 100,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1835,
  verbose = FALSE
)
```



#------------------------------------------
#Correlations with kidney clinical markers
#-----------------------------------------
```{r}
prot_ids <- sig_all_models_fc$protein_id
keep_prots <- intersect(colnames(prot_age_sex_study), prot_ids)
sig_prots <- prot_age_sex_study[, keep_prots, drop = FALSE]

corr_df <- bind_cols(meta3, sig_prots) %>%
  rename(HbA1c = hba1c, uACR = acr_u) %>%
  mutate(
    HbA1c = as.numeric(HbA1c), #only do this if I haven't included hba1c in the model 
    uACR  = as.numeric(uACR),
    eGFR = as.numeric(eGFR_CKiD_U25_avg),
   #ERPF = as.numeric(erpf_bsa_plasma), don't have this in attempt
    mm_si = as.numeric(mm_si),
    mm_ir = as.numeric(mm_ir),
    mm_di = as.numeric(mm_di),
    wthr = as.numeric(whtr),
    age   = as.numeric(age),
    sex   = as.factor(sex),
    BMI   = as.numeric(bmi),
   r2_avail=as.numeric(avg_c_r2),
  ) %>% filter(!is.na(HbA1c) & !is.na(uACR) & !is.na(eGFR)) 

```

```{r}
contrast_list <- c(
  "T1D_Normal_vs_LC_Normal",
  "T1D_OverObese_vs_LC_OverObese",
  "T1D_Normal_vs_T1D_Overweight",
  "T1D_Normal_vs_T1D_Obese",
  "T1D_Obese_vs_T1D_Overweight"
)

get_groups_from_contrast <- function(contrast_name) {
  
  contrast_groups <- list(
    
    T1D_Normal_vs_LC_Normal = list(
      group1 = "T1D_Normal",
      group2 = "LC_Normal"
    ),
    T1D_OverObese_vs_LC_OverObese = list(
      group1 = c("T1D_Overweight", "T1D_Obese"),
      group2 = "LC_Overweight_Obese"
    ),
    T1D_Normal_vs_T1D_Overweight = list(
      group1 = "T1D_Normal",
      group2 = "T1D_Overweight"
    ),
    T1D_Normal_vs_T1D_Obese = list(
      group1 = "T1D_Normal",
      group2 = "T1D_Obese"
    ),
    T1D_Obese_vs_T1D_Overweight = list(
      group1 = "T1D_Obese",
      group2 = "T1D_Overweight"
    )
  )
  
  if (!contrast_name %in% names(contrast_groups)) {
    stop("Contrast not found: ", contrast_name)
  }
  
  contrast_groups[[contrast_name]]
}
```

```{r}
#--------------------------------
#marginal correlations analysis
#--------------------------------
cor_test_one <- function(df, outcome, protein, method = "spearman", min_n = 10) {
  x  <- df[[outcome]]
  y  <- df[[protein]]
  ok <- is.finite(x) & is.finite(y) & !is.na(x) & !is.na(y)  
  n  <- sum(ok)
  if (n < min_n) return(tibble(n = n, rho = NA_real_, p_value = NA_real_))
  ct <- suppressWarnings(cor.test(x[ok], y[ok], method = method, exact = FALSE))
  tibble(n = n, rho = unname(ct$estimate), p_value = ct$p.value)
}

run_contrast_correlation <- function(contrast_name, corr_df, keep_prots, min_n = 10) {
  groups <- get_groups_from_contrast(contrast_name)

  df_sub <- corr_df %>%
    filter(group_bmi %in% c(groups$group1, groups$group2))
  
  prots <- intersect(keep_prots, colnames(df_sub))

  message("---- marginal: ", contrast_name, " ----")
  message("subset N samples: ", nrow(df_sub))
  message("subset N proteins: ", length(prots))

  if (nrow(df_sub) == 0 || length(prots) == 0) return(NULL)
  prots <- prots[sapply(df_sub[, prots, drop = FALSE],
                        function(v) sd(v, na.rm = TRUE) > 0)]
  if (length(prots) == 0) return(NULL)

  tidyr::crossing(outcome = c("HbA1c", "uACR", "BMI"), protein = prots) %>%
    mutate(stats = purrr::map2(outcome, protein,
                               ~cor_test_one(df_sub, .x, .y, min_n = min_n))) %>%
    tidyr::unnest(stats) %>%
    mutate(contrast = contrast_name) %>%
    group_by(contrast, outcome) %>%
    mutate(FDR = p.adjust(p_value, method = "BH")) %>%
    ungroup()
}

# ---- Run all contrasts ----
all_cor_results <- purrr::map_dfr(
  contrast_list,
  ~run_contrast_correlation(.x, corr_df, keep_prots, min_n = 10)
)

all_cor_results %>% filter(FDR <= 0.05)
```

```{r}
#-----------------------------
# Partial correlations setup
#-----------------------------

partial_outcomes <- c("HbA1c", "uACR", "BMI", "eGFR", "wthr")
partial_covars   <- c("age", "sex", "study")
stopifnot(all(partial_covars %in% names(corr_df)))

get_partial_cor <- function(df, protein, outcome, covars, min_n = 15) {
  vars_needed <- c(protein, outcome, covars)
  sub_df <- df[, vars_needed, drop = FALSE]
  sub_df <- sub_df[complete.cases(sub_df), , drop = FALSE]
  if (nrow(sub_df) < min_n) {
    return(tibble(n = nrow(sub_df), rho = NA_real_, p_value = NA_real_))
  }
  fit_p <- lm(as.formula(paste(protein, "~", paste(covars, collapse = "+"))), data = sub_df)
  fit_y <- lm(as.formula(paste(outcome, "~", paste(covars, collapse = "+"))), data = sub_df)
  res_p <- resid(fit_p)
  res_y <- resid(fit_y)
  ct <- suppressWarnings(cor.test(res_p, res_y, method = "spearman", exact = FALSE))
  tibble(n = nrow(sub_df), rho = unname(ct$estimate), p_value = ct$p.value)
}

run_contrast_partial <- function(contrast_name,         
                                 corr_df,
                                 keep_prots,             
                                 outcomes = partial_outcomes,
                                 covars   = partial_covars,
                                 min_n    = 15) {
  
  groups <- get_groups_from_contrast(contrast_name)

  df_sub <- corr_df %>%
    filter(group_bmi %in% c(groups$group1, groups$group2))
  
  prots <- intersect(keep_prots, colnames(df_sub))
  
  message("---- partial: ", contrast_name, " ----")    
  message("subset N samples: ",  nrow(df_sub))
  message("subset N proteins: ", length(prots))
  
  if (nrow(df_sub) == 0 || length(prots) == 0) return(NULL)
  prots <- prots[sapply(df_sub[, prots, drop = FALSE],
                        function(v) sd(v, na.rm = TRUE) > 0)]
  if (length(prots) == 0) return(NULL)
  
  tidyr::crossing(outcome = outcomes, protein = prots) %>%
    mutate(stats = purrr::map2(
      outcome, protein,
      ~get_partial_cor(df_sub, .y, .x, covars = covars, min_n = min_n)
    )) %>%
    tidyr::unnest(stats) %>%
    mutate(contrast = contrast_name) %>%
    group_by(contrast, outcome) %>%
    mutate(FDR = p.adjust(p_value, method = "BH")) %>%
    ungroup()
}

all_partial_results <- purrr::map_dfr(
  contrast_list,
  ~run_contrast_partial(.x, corr_df, keep_prots,
                        outcomes = partial_outcomes,
                        covars   = partial_covars,
                        min_n    = 15)
)
```

```{r}
map_tbl <- analytes %>%
  transmute(
    seq_id      = AptName,           
    gene_symbol = EntrezGeneSymbol
  ) %>%
  mutate(
    seq_id      = as.character(seq_id),
    gene_symbol = as.character(gene_symbol),
    gene_symbol = na_if(gene_symbol, ""),
    gene_symbol = if_else(is.na(gene_symbol), seq_id, gene_symbol)
  ) %>%
  distinct(seq_id, gene_symbol)
```

```{r}
cor_gene <- all_cor_results %>%
  left_join(map_tbl, by = c("protein" = "seq_id")) %>%
  mutate(gene_symbol = if_else(is.na(gene_symbol) | gene_symbol == "", protein, gene_symbol)) %>%
  group_by(contrast, outcome, gene_symbol) %>%
  summarise(
    rho = mean(rho, na.rm = TRUE),
    p_value = min(p_value, na.rm = TRUE),
    FDR = min(FDR, na.rm = TRUE),
    n = max(n, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(label = gene_symbol)

partial_gene <- all_partial_results %>%
  left_join(map_tbl, by = c("protein" = "seq_id")) %>%
  mutate(gene_symbol = if_else(is.na(gene_symbol) | gene_symbol == "", protein, gene_symbol)) %>%
  group_by(contrast, outcome, gene_symbol) %>%
  summarise(
    rho = mean(rho, na.rm = TRUE),
    p_value = min(p_value, na.rm = TRUE),
    FDR = min(FDR, na.rm = TRUE),
    n = max(n, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(label = gene_symbol)
```


#clinical associates depicted with volcanos plots
```{r}
volcano_cor_df <- partial_gene %>%
  mutate(
    neg_log10_p = -log10(FDR),
    sig_group = case_when(
      FDR < 0.05 & rho > 0  ~ "Positive",
      FDR < 0.05 & rho < 0  ~ "Negative",
      TRUE ~ "Not significant"
    )
  )
```


```{r}
make_volcano_cor<- function(df,
                              contrast_name,
                              outcome_name,
                              title_prefix = "Partial correlations",
                              p_cutoff = 0.05,
                              fdr_cutoff = 0.05,
                              rho_cutoff = 0,
                              label_n = 10) {
  
  plot_df <- df %>%
    dplyr::filter(
      contrast == contrast_name,
      outcome == outcome_name
    ) %>%
    dplyr::filter(
      is.finite(rho),
      is.finite(p_value),
      is.finite(FDR)
    ) %>%
    dplyr::mutate(
      neg_log10_p = -log10(p_value),
      p_sig_group = dplyr::case_when(
        p_value < p_cutoff & rho > rho_cutoff  ~ "Positive p < 0.05",
        p_value < p_cutoff & rho < -rho_cutoff ~ "Negative p < 0.05",
        TRUE ~ "Not significant"
      ),
      fdr_sig = FDR < fdr_cutoff,
      fdr_group = dplyr::case_when(
        fdr_sig & rho > 0 ~ "Positive FDR < 0.05",
        fdr_sig & rho < 0 ~ "Negative FDR < 0.05",
        TRUE ~ NA_character_
      )
    )
  
 
  n_fdr <- sum(plot_df$fdr_sig, na.rm = TRUE)
  
 
  top_labels <- plot_df %>%
    dplyr::filter(fdr_sig) %>%
    dplyr::arrange(FDR) %>%
    dplyr::slice_head(n = label_n)
  
  p <- ggplot(plot_df, aes(x = rho, y = neg_log10_p)) +
    
    geom_point(
      aes(color = p_sig_group),
      alpha = 0.45,
      size = 2
    ) +
    
    geom_vline(xintercept = 0, linetype = "solid", alpha = 0.5) +
    geom_hline(yintercept = -log10(p_cutoff), linetype = "dashed") +
    
    scale_color_manual(
      values = c(
        "Positive p < 0.05" = "pink3",
        "Negative p < 0.05" = "cyan2",
        "Not significant" = "grey85"
      )
    ) +
    
    labs(
      title = paste0(title_prefix, "\n", contrast_name, " - ", outcome_name),
      x = "Spearman rho",
      y = "-log10(raw p-value)",
      color = "Raw p-value"
    ) +
    
    theme_bw() +
    theme(
      panel.grid = element_blank(),
      legend.position = "right"
    )
  
  if (n_fdr > 0) {
    
    p <- p +
      geom_point(
        data = plot_df %>% dplyr::filter(fdr_sig),
        aes(fill = fdr_group),
        shape = 21,
        color = "black",
        stroke = 0.9,
        size = 2.8
      ) +
      
      ggrepel::geom_text_repel(
        data = top_labels,
        aes(label = label),
        size = 3.5,
        fontface = "bold",
        max.overlaps = Inf
      ) +
      
      scale_fill_manual(
        values = c(
          "Positive FDR < 0.05" = "darkred",
          "Negative FDR < 0.05" = "cyan4"
        ),
        name = "FDR significant"
      )
    
  } else {
    
    p <- p +
      annotate(
        "text",
        x = 0,
        y = max(plot_df$neg_log10_p, na.rm = TRUE) * 0.9,
        label = "No FDR-significant associations",
        size = 4,
        fontface = "italic",
        color = "grey40"
      )
  }
  
  return(p)
}
```

```{r}
 partial_gene %>%
  distinct(contrast) %>%
  pull(contrast)
```

```{r}
outcomes <- c("HbA1c", "uACR", "BMI", "eGFR")

plots_cor_p1 <- make_volcano_cor_list(
  df = partial_gene,
  contrast_name = "T1D_Normal_vs_LC_Normal",
  outcomes = outcomes,
  label_n = 10
)

panel_cor_p1 <- (plots_cor_p1[[1]] | plots_cor_p1[[2]]) /
             (plots_cor_p1[[3]] | plots_cor_p1[[4]])

panel_cor_p1

ggsave(
  filename = "panel_cor_p1.png",
  plot = panel_cor_p1,
  path = dir.results,
  height = 10,
  width = 12,
  dpi = 300
)
```

```{r}
plots_cor_p2 <- make_volcano_cor_list(
  df = partial_gene,
  contrast_name = "T1D_OverObese_vs_LC_OverObese",
  outcomes = outcomes,
  label_n = 10
)


panel_cor_p2 <- (plots_cor_p2[[1]] | plots_cor_p2[[2]]) /
             (plots_cor_p2[[3]] | plots_cor_p2[[4]])

panel_cor_p2

ggsave(
  filename = "panel_cor_p2.png",
  plot = panel_cor_p2,
  path = dir.results,
  height = 10,
  width = 12,
  dpi = 300
)
```

```{r}
plots_cor_p3 <- make_volcano_cor_list(
  df = partial_gene,
  contrast_name = "T1D_Normal_vs_T1D_Overweight",
  outcomes = outcomes,
  label_n = 10
)


panel_cor_p3 <- (plots_cor_p3[[1]] | plots_cor_p3[[2]]) /
             (plots_cor_p3[[3]] | plots_cor_p3[[4]])

panel_cor_p3

ggsave(
  filename = "panel_cor_p3.png",
  plot = panel_cor_p3,
  path = dir.results,
  height = 10,
  width = 12,
  dpi = 300
)
```

```{r}
plots_cor_p4 <- make_volcano_cor_list(
  df = partial_gene,
  contrast_name = "T1D_Normal_vs_T1D_Obese",
  outcomes = outcomes,
  label_n = 10
)


panel_cor_p4 <- (plots_cor_p4[[1]] | plots_cor_p4[[2]]) /
             (plots_cor_p4[[3]] | plots_cor_p4[[4]])

panel_cor_p4

ggsave(
  filename = "panel_cor_p4.png",
  plot = panel_cor_p4,
  path = dir.results,
  height = 10,
  width = 12,
  dpi = 300
)
```

```{r}
plots_cor_p5 <- make_volcano_cor_list(
  df = partial_gene,
  contrast_name = "T1D_Obese_vs_T1D_Overweight",
  outcomes = outcomes,
  label_n = 10
)


panel_cor_p5 <- (plots_cor_p5[[1]] | plots_cor_p5[[2]]) /
             (plots_cor_p5[[3]] | plots_cor_p5[[4]])

panel_cor_p5

ggsave(
  filename = "panel_cor_p5.png",
  plot = panel_cor_p5,
  path = dir.results,
  height = 10,
  width = 12,
  dpi = 300
)
```





#=========================================
#archived code
#setting up metadata and protein matrices
#=========================================
```{r}
#meta <- t1d_hc_dat_soma %>%
  dplyr:: select (-starts_with("seq"))
#meta$group
meta$group <- factor(
  meta$group,
  levels = c("Lean Control","Type 1 Diabetes"),
  labels = c("LC","T1D")
)

#meta <- meta %>%
  dplyr::mutate(
    study = dplyr::recode(study,
      "CROCODILE"   = "CRC",
      "PANDA"       = "PAN",
      "RENAL-HEIR"  = "RH",
      "CASPER" = "CAS",
      "ATTEMPT" = "ATT"
    )
  )

meta$record_id <- as.character(meta$record_id)
#meta$bmi_obesity
#meta$dxa_obesity
# keep only participants with BOTH BMI and DXA obesity
meta2 <- meta %>%
  dplyr::filter(
    !is.na(bmi_obesity) #,
    #!is.na(dxa_obesity)
  )

dat <- t1d_hc_dat_soma %>%
  dplyr::select(record_id, starts_with("seq"))


#dat_urine <- dat %>%
#  dplyr::select(ends_with("urine"), ends_with("urine_cradj"))

dat_plasma <- dat %>%
  dplyr:: select(-ends_with("urine"), -ends_with("urine_cradj"))
dat_plasma <- as.data.frame(dat_plasma)
rownames(dat_plasma) <- dat_plasma$record_id
dat_plasma$record_id <- NULL

#filtering final matrices to have the same record_ids
ids_keep <- meta2$record_id

# plasma
dat_plasma2 <- dat_plasma[ids_keep, , drop = FALSE]
meta2 <- meta2 %>%
  dplyr::filter(record_id %in% rownames(dat_plasma2))
meta2 <- meta2[match(rownames(dat_plasma2), meta2$record_id), ]

stopifnot(identical(rownames(dat_plasma2), meta2$record_id))

```


##filter to proteins that are only in the analyte panel (~7000)
##check if this is correct for the newer somascan prot pannel
##should filter ~3000 all from ATTEMPT
##then filter to keep proteins that are in at least 2 studies
```{r}
analytes <- read.csv(
  file.path(root_path, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

stopifnot("AptName" %in% colnames(analytes))

prot_panel <- analytes$AptName %>%
  as.character() %>%
  trimws() %>%
  unique()

prot_panel <- prot_panel[!is.na(prot_panel) & prot_panel != ""]

aptamers_keep <- intersect(prot_panel, colnames(dat_plasma2))

if (length(aptamers_keep) == 0) {
  stop("No analytes from analytes.csv matched columns in dat_plasma2.")
}

dat_plasma3 <- dat_plasma2[, aptamers_keep, drop = FALSE]

cat("Plasma features before analyte filter:", ncol(dat_plasma2), "\n")
cat("Proteins in analyte panel:", length(prot_panel), "\n")
cat("Matched proteins kept:", length(aptamers_keep), "\n")
cat("Panel proteins missing from plasma matrix:", length(setdiff(prot_panel, colnames(dat_plasma2))), "\n")

# force numeric safely
dat_plasma3 <- as.data.frame(dat_plasma3)

dat_plasma3[] <- lapply(dat_plasma3, function(x) {
  suppressWarnings(as.numeric(as.character(x)))
})

dat_plasma3 <- as.matrix(dat_plasma3)

# log2 transform
dat_log <- log2(dat_plasma3 + 1)
dat_log[!is.finite(dat_log)] <- NA_real_

stopifnot(identical(rownames(dat_log), meta2$record_id))
```


```{r}
#not really necessary to do as all protiens pass filter once I remove the extra 3000 from the ATTEMPT Panel
##filtering step where protein must exist in 60% of samples in a study
stopifnot(identical(rownames(dat_log), meta2$record_id))
meta2$study <- factor(meta2$study)

study_idx_all <- split(seq_len(nrow(dat_log)), meta2$study)
present_by_study_all <- sapply(study_idx_all, function(idx) {
  colMeans(!is.na(dat_log[idx, , drop = FALSE])) >= 0.6
})

present_by_study_all <- as.matrix(present_by_study_all)

n_studies_present_all <- rowSums(present_by_study_all)
keep_study_all <- n_studies_present_all >= 2

removed_proteins <- colnames(dat_log)[!keep_study_all]
kept_proteins    <- colnames(dat_log)[keep_study_all]

cat("All proteomic features before study filter:", ncol(dat_log), "\n")
cat("Removed (<2 studies meeting 60% presence threshold):", length(removed_proteins), "\n")
cat("Kept (>=2 studies meeting 60% presence threshold):", length(kept_proteins), "\n")

if (length(kept_proteins) == 0) {
  stop("No proteins passed the study presence filter.")
}

dat_log_filt <- dat_log[, kept_proteins, drop = FALSE]

stopifnot(identical(rownames(dat_log_filt), meta2$record_id))
```

#imputating the data
```{r}
set.seed(1835)
dat_imp <- dat_log_filt
meds <- apply(dat_imp, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(dat_imp))) {
  if (!is.na(meds[j])) {
    dat_imp[is.na(dat_imp[, j]), j] <- meds[j]
  }
}

keep_cols <- colSums(is.na(dat_imp)) == 0 &
  apply(dat_imp, 2, var) > 0 &
  is.finite(apply(dat_imp, 2, var))

dat_imp <- dat_imp[, keep_cols, drop = FALSE]
```

##--------------------------------
## Find and Remove Outliers in PCA
##---------------------------------
```{r}
stopifnot(identical(rownames(dat_imp), meta2$record_id))

# PCA outlier detection
pca0 <- prcomp(dat_imp, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2, drop = FALSE])
scores$record_id <- rownames(scores)

scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

threshold <- mean(scores$dist, na.rm = TRUE) + 
  3 * sd(scores$dist, na.rm = TRUE)

outliers <- scores$record_id[scores$dist > threshold]

outliers

# remove outliers from matrix
dat_imp2 <- dat_imp[
  !rownames(dat_imp) %in% outliers,, drop = FALSE]

# remove outliers from metadata
meta3 <- meta2 %>%
  dplyr::filter(!record_id %in% outliers)

# reorder metadata to match matrix
meta3 <- meta3[match(rownames(dat_imp2), meta3$record_id), ]

# checks
cat("Still present in metadata? ", any(outliers %in% meta3$record_id), "\n")
cat("Still present in matrix? ", any(outliers %in% rownames(dat_imp2)), "\n")

stopifnot(!any(outliers %in% meta3$record_id))
stopifnot(!any(outliers %in% rownames(dat_imp2)))
stopifnot(identical(rownames(dat_imp2), meta3$record_id))
```

```{r}
meta3 <- meta3 %>%
  dplyr::mutate(
    group_bmi = paste(group, bmi_obesity, sep = "_")
  )

meta3$group_bmi <- factor(meta3$group_bmi)

table(meta3$group_bmi)
```




#Data tables
```{r}
table_study_bmi <- meta3 %>%
  count(study, bmi_obesity) %>%
  pivot_wider(
    names_from = bmi_obesity,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(Total = rowSums(across(where(is.numeric))))

table_study_bmi %>%
  kable(
    caption = "Participants by study and BMI adiposity group"
  ) %>%
  kable_styling(full_width = FALSE)
```

```{r}
study_group_bmi_table <- meta3 %>%
  dplyr::count(study, group, bmi_obesity, name = "n") %>%
  tidyr::unite("group_bmi", group, bmi_obesity, sep = "_") %>%
  tidyr::pivot_wider(
    names_from = group_bmi,
    values_from = n,
    values_fill = 0
  ) %>%
  dplyr::mutate(
    Total = rowSums(dplyr::across(where(is.numeric)))
  )

study_group_bmi_table
```
#---------------
# Plotting Funciton
#---------------
```{r}
make_pca_plot <- function(mat, meta_df, title_str) {
  pca  <- prcomp(mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1    = pca$x[, 1],
    PC2    = pca$x[, 2],
    group = meta_df$group,
    study  = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = group, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        LC  = "Cyan4",
        T1D = "DarkRed"
      ),
      name = "Group"
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
##-----------------------
## PCAs by disease status
##-----------------------
```{r}
design <- model.matrix(~ group, data = meta3)
p  <- make_pca_plot(dat_imp2, meta3, "No adjustment")

cov_m1 <- model.matrix(~ age + sex, data = meta3)[, -1, drop = FALSE]
prot_m1 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m1, design = design))
p1 <- make_pca_plot(prot_m1, meta3, "Age + Sex")


cov_m2 <- model.matrix(~ age + sex + study, data = meta3)[, -1, drop = FALSE]
prot_m2 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m2, design = design))
p2 <- make_pca_plot(prot_m2, meta3, "Age + Sex + Study")

pca_disease_status <- (p | p1 | p2) 
  ggsave(pca_disease_status,
         plot = pca_disease_status,
         path = file.path(dir.results),
         width = 10,
         height = 8)
```

```{r}
make_pca_plot2 <- function(mat, meta_df, title_str) {
  pca  <- prcomp(mat, center = TRUE, scale. = TRUE)
  vexp <- round(100 * (pca$sdev^2) / sum(pca$sdev^2), 1)

  tibble(
    PC1    = pca$x[, 1],
    PC2    = pca$x[, 2],
    bmi_obesity = meta_df$bmi_obesity,
    study  = meta_df$study
  ) %>%
    ggplot(aes(PC1, PC2, color = bmi_obesity, shape = study)) +
    geom_point(size = 3, alpha = 0.85) +
    scale_color_manual(
      values = c(
        Overweight  = "cyan4",
        Obese = "plum4",
        Normal = "orange3"
      ),
      name = "BMI group"
    ) +
    labs(
      title = title_str,
      x = paste0("PC1 (", vexp[1], "%)"),
      y = paste0("PC2 (", vexp[2], "%)")
    ) +
    theme_bw()
}
```
##------------------
## PCAs by bmi group
##------------------
```{r}
design <- model.matrix(~ bmi_obesity, data = meta3)

p  <- make_pca_plot2(dat_imp2, meta3, "No adjustment")

cov_m1 <- model.matrix(~ age + sex, data = meta3)[, -1, drop = FALSE]
prot_m1 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m1, design = design))
p1 <- make_pca_plot2(prot_m1, meta3, "Age + Sex")


cov_m2 <- model.matrix(~ age + sex + study, data = meta3)[, -1, drop = FALSE]
prot_m2 <- t(removeBatchEffect(t(dat_imp2), covariates = cov_m2, design = design))
p2 <- make_pca_plot2(prot_m2, meta3, "Age + Sex + Study")

(p | p1 | p2)
```

```{r}
#setting up limma matrices
# make sure matrix and metadata match
stopifnot(identical(rownames(dat_imp2), meta3$record_id))

#neet to transform for limma
mat1 <- t(dat_imp2)

# set "normal" as reference group
meta3$bmi_obesity <- factor(
  meta3$bmi_obesity,
  levels = c("Normal", "Overweight", "Obese")
)

meta3$study <- factor(meta3$study)
```


#---------------------------------------------
#run limma between disease status T1D vs LC
#---------------------------------------------
```{r}
design_m1 <- model.matrix(~ group + age + sex + study, data = meta3) #add/remove study

colnames(design_m1)

# fit model
disease_m1 <- lmFit(mat1, design_m1)
disease_m1 <- eBayes(disease_m1)

# extract T1D vs LC
res_disease_m1 <- topTable(
  disease_m1,
  coef = "groupT1D",
  number = Inf,
  adjust.method = "BH"
) %>%
  rownames_to_column("protein_id") %>%
  mutate(
    model = "Disease status: T1D vs LC + age + sex + study",
    contrast = "T1D vs LC",
    direction = case_when(
      adj.P.Val < 0.05 & logFC > 0 ~ "Higher in T1D",
      adj.P.Val < 0.05 & logFC < 0 ~ "Lower in T1D",
      TRUE ~ "Not significant"
    )
  )

table(res_disease_m1$direction)
```
#---------------------------------------------
#run limma for adiposity across disease status
#---------------------------------------------
```{r}
# Model 1: BMI obesity + age + sex
design_m1 <- model.matrix(~ 0 + bmi_obesity + age + sex + study, data = meta3) #add/remove study
colnames(design_m1)

contrast_matrix <- makeContrasts(
  Overweight_vs_Normal = bmi_obesityOverweight - bmi_obesityNormal,
  Obese_vs_Normal = bmi_obesityObese - bmi_obesityNormal,
  Obese_vs_Overweight = bmi_obesityObese - bmi_obesityOverweight,
  levels = design_m1
)


m <- lmFit(mat1, design_m1)
m1 <- contrasts.fit(m, contrast_matrix)
m1 <- eBayes(m1)

results_list <- lapply(colnames(contrast_matrix), function(cn) {
  topTable(m1, coef = cn, number = Inf, adjust.method = "BH") %>%
    rownames_to_column("protein_id") %>%
    mutate(
      contrast = cn,
      direction = case_when(
        adj.P.Val < 0.05 & logFC > 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][1]),
        adj.P.Val < 0.05 & logFC < 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][2]),
        TRUE ~ "Not significant"
      )
    )
})

results_bmi_all <- bind_rows(results_list)
results_sig <- results_bmi_all %>%
  dplyr::filter(
    adj.P.Val < 0.05,
    abs(logFC) > 0
  )
results_sig
```

#------------------------------------
#limma with only T1D Adiposity Groups
#------------------------------------
```{r}
#subsetting to only T1D group
meta_t1d <- meta3 %>%
  dplyr::filter(group == "T1D") %>%
  dplyr::filter(!is.na(bmi_obesity), !is.na(age), !is.na(sex), !is.na(study))

dat_t1d <- dat_imp2[meta_t1d$record_id, , drop = FALSE]

# reorder metadata to match matrix
meta_t1d <- meta_t1d[match(rownames(dat_t1d), meta_t1d$record_id), ]

stopifnot(identical(rownames(dat_t1d), meta_t1d$record_id))
```

```{r}
#resetting groups as factors
meta_t1d$bmi_obesity <- factor(
  meta_t1d$bmi_obesity,
  levels = c("Normal", "Overweight", "Obese")
)

meta_t1d$sex   <- factor(meta_t1d$sex)
meta_t1d$study <- factor(meta_t1d$study)

table(meta_t1d$bmi_obesity)
table(meta_t1d$study)
```

```{r}
mat_t1d <- t(dat_t1d)

design_t1d <- model.matrix(~ 0 + bmi_obesity + age + sex + study,data = meta_t1d
)

# optional, but protects against invalid names
colnames(design_t1d) <- make.names(colnames(design_t1d))

colnames(design_t1d)

contrast_t1d <- makeContrasts(
  Overweight_vs_Normal =
    bmi_obesityOverweight - bmi_obesityNormal,
  Obese_vs_Normal =
    bmi_obesityObese - bmi_obesityNormal,
  Obese_vs_Overweight =
    bmi_obesityObese - bmi_obesityOverweight,
  levels = design_t1d
)
```
```{r}
fit_t1d <- lmFit(mat_t1d, design_t1d)

fit_t1d2 <- contrasts.fit(fit_t1d, contrast_t1d)
fit_t1d2 <- eBayes(fit_t1d2)
```

```{r}
results_t1d_bmi_all <- lapply(colnames(contrast_t1d), function(cn) {
  topTable(
    fit_t1d2,
    coef = cn,
    number = Inf,
    adjust.method = "BH"
  ) %>%
    rownames_to_column("protein_id") %>%
    mutate(
      contrast = cn,
      direction = case_when(
        adj.P.Val < 0.05 & logFC > 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][1]),
        adj.P.Val < 0.05 & logFC < 0 ~ paste("Higher in", str_split(cn, "_vs_")[[1]][2]),
        TRUE ~ "Not significant"
      )
    )
}) %>%
  bind_rows()

res_t1d_bmi_sig <- results_t1d_bmi_all %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= 0) 
res_t1d_bmi_sig
```
```{r}
sig_cutoff <- 0.05
lfc_cutoff <- 0   # change to 0.5 if you want stricter filtering

combined_sig_terms <- dplyr::bind_rows(
  
  res_disease_m1 %>%
    dplyr::mutate(
      analysis = "Disease status",
      model = "T1D vs LC + age + sex + study"
    ),
  
  results_bmi_all %>%
    dplyr::mutate(
      analysis = "BMI adiposity - all participants",
      model = "BMI group + age + sex + study"
    ),
  
  results_t1d_bmi_all %>%
    dplyr::mutate(
      analysis = "BMI adiposity - T1D only",
      model = "BMI group within T1D + age + sex + study"
    )
) %>%
  dplyr::filter(
    adj.P.Val < sig_cutoff,
    abs(logFC) > lfc_cutoff
  ) %>%
  dplyr::mutate(
    direction_simple = dplyr::case_when(
      logFC > 0 ~ "Positive logFC",
      logFC < 0 ~ "Negative logFC",
      TRUE ~ "No change"
    )
  ) %>%
  dplyr::select(
    analysis,
    model,
    contrast,
    protein_id,
    logFC,
    AveExpr,
    t,
    P.Value,
    adj.P.Val,
    B,
    direction,
    direction_simple
  ) %>%
  dplyr::arrange(analysis, contrast, adj.P.Val)

combined_sig_terms_n <- combined_sig_terms %>%
  dplyr::count(analysis, contrast, direction, name = "n_sig")
combined_sig_terms_n


write.csv(
  combined_sig_terms,
  file = file.path(dir.results, "limma_results.csv"),
  row.names = FALSE
)

write.csv(
  combined_sig_terms_n,
  file = file.path(dir.results, "sig_DEP_counts.csv"),
  row.names = FALSE
)
```


#clinical maker correlation 
#correlate with transcriptomics
T1D_normal vs LC
T1D_over_obese vs T1D normal
combined overweight/obese