---
title: "05_proteomics_analysis"
author: "Ye Ji Choi & Savanah Leidholt"
format: html
---
```{r}
library(dplyr)
library(tidyr)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggforce)
library(enrichR)
library(stringr)
library(limma)
library(EnhancedVolcano)
library(patchwork)
library(ggrepel)
library(vegan)
library(tidyverse)
library(data.table)
library(purrr)
library(haven)
library(ggVennDiagram)
library(caret)
library(kableExtra)
library(msigdbr)
library(fgsea)
library(emmeans)
library(gtsummary)
library(vegan)
library(ppcor)
library(tibble)
library(readr)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "leidholt") {
  root_path <- c("/mmfs1/gscratch/togo/leidholt/")
  git_path <- "/mmfs1/gscratch/togo/leidholt/CHCO-Code/"
  keys <- fromJSON("/mmfs1/home/leidholt/keys.json")
} else if (user == "sleidholt") {
    root_path <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/Shared/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/T1D Adiposity/Results/Compiled Results/Proteomics Results")
    git_path <- "/Users/sleidholt/Documents/GitHub/CHCO-Code/Petter Bjornstad/"
    keys <- fromJSON("/Users/Shared/OneDrive - UW/Personal_storage/keys.json")
} else if (user == "savanahleidholt") {
    root_path <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/")
    dir.results <- c("/Users/savanahleidholt/OneDrive - UW/Laura Pyle's files - Biostatistics Core Shared Drive/T1D Adiposity/Results/Compiled Results/Proteomics Results")
    git_path <- c("/Users/savanahleidholt/Desktop/CHCO-Code/Petter Bjornstad")
    keys <- fromJSON("/Users/savanahleidholt/OneDrive - UW/Personal_storage/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

#-----------------
#functions
#-----------------
```{r}
source(file = file.path(git_path,"T1D Adiposity", "functions", "proteomics_functions.R"))
```

```{r}
Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```

#no dxa on attempt
#also do waist to hip = wthr, wthr_obesity
#hba1c with age
#check mrns to see if there are people are co-enrolled
```{r}
t1d_hc_dat_soma <- s3readRDS(object = "data_clean/t1d_hc_clinical_data_soma.csv", 
                             bucket = "t1d.adiposity",
                             region = "")

harm <- readRDS(file.path(root_path,"Data Harmonization", "Data Clean", "harmonized_dataset.rds" ))
```
#----------------------
#cleaning up data & ids
#----------------------
```{r}
t1d_hc_dat_soma <- t1d_hc_dat_soma %>%
  mutate(record_id = as.character(record_id))

recordID_mrn <- harm %>%
  dplyr::select(record_id, mrn) %>%
  distinct() %>%
  mutate(record_id = as.character(record_id))

# join MRNs into the soma dataset
t1d_hc_dat_soma <- t1d_hc_dat_soma %>%
  left_join(recordID_mrn, by = "record_id")

t1d_hc_dat_soma <- t1d_hc_dat_soma %>%
  relocate(record_id, mrn)

t1d_hc_dat_soma_dedup <- t1d_hc_dat_soma %>%
  mutate(
    dedup_id = ifelse(
      !is.na(mrn) & mrn != "",
      as.character(mrn),
      paste0("record_", record_id)
    )
  ) %>%
  distinct(dedup_id, .keep_all = TRUE) %>%
  dplyr::select(-dedup_id)
```

#-------------------------
#creating cleaned metadata
#-------------------------
```{r}
meta <- t1d_hc_dat_soma_dedup %>%
  dplyr::select(-starts_with("seq")) %>%
  dplyr::mutate(
    record_id = as.character(record_id),
    group = factor(
      group,
      levels = c("Lean Control", "Type 1 Diabetes"),
      labels = c("LC", "T1D")
    ),
    study = dplyr::recode(
      study,
      "CROCODILE"  = "CRC",
      "PANDA"      = "PAN",
      "RENAL-HEIR" = "RH",
      "CASPER"     = "CAS",
      "ATTEMPT"    = "ATT",
      .default = study
    ),
    bmi_obesity = factor(
      bmi_obesity,
      levels = c("Normal", "Overweight", "Obese")
    ),
    bmi_bin = dplyr::case_when(
      bmi_obesity == "Normal" ~ "Normal",
      bmi_obesity %in% c("Overweight", "Obese") ~ "Overweight_Obese",
      TRUE ~ NA_character_
    ),
    group_bmi = dplyr::case_when(
      group == "LC"  & bmi_obesity == "Normal" ~ "LC_Normal",
      group == "LC"  & bmi_obesity %in% c("Overweight", "Obese") ~ "LC_Overweight_Obese",
      group == "T1D" & bmi_obesity == "Normal" ~ "T1D_Normal",
      group == "T1D" & bmi_obesity == "Overweight" ~ "T1D_Overweight",
      group == "T1D" & bmi_obesity == "Obese" ~ "T1D_Obese",
      TRUE ~ NA_character_
    )
  )

meta$study <- factor(meta$study)
meta$bmi_bin <- factor(meta$bmi_bin, levels = c("Normal", "Overweight_Obese"))
meta$group_bmi <- factor(
  meta$group_bmi,
  levels = c(
    "LC_Normal",
    "LC_Overweight_Obese",
    "T1D_Normal",
    "T1D_Overweight",
    "T1D_Obese"
  )
)

# keep BMI-complete samples
meta2 <- meta %>%
 dplyr::filter(
    !is.na(record_id),
    !is.na(group_bmi),
    !is.na(age),
    !is.na(sex),
    !is.na(study)
  )

table(meta2$group_bmi, useNA = "ifany")
```
#----------------------------------
#checking out clinical outcome data
#----------------------------------
```{r}
#log-transform uACR because it is usually skewed
meta_box <- meta2 %>%
  mutate(
    log_uACR = log1p(as.numeric(acr_u)),
    group_bmi = factor(group_bmi),
    sex = factor(sex),
    study = factor(study)
  )

# outcomes to plot
outcome_vars <- c(
  eGFR   = "eGFR_CKiD_U25_avg",
  avg_c_r2 = "avg_c_r2",
  avg_k_r2 = "avg_k_r2",
  HbA1c = "hba1c",
  log_uACR   = "log_uACR",
  BMI    = "bmi",
  WHtR   = "whtr"
)
```
#------------------------------
#boxplots of clinical outcomes
#add emmeans function and stats
#------------------------------
```{r}
clinical_boxplot_list <- purrr::imap(
  outcome_vars,
  ~ plot_clinical_boxplot(
    data = meta_box,
    outcome_var = .x,
    outcome_label = .y
  )
)

clinical_boxplot_panel <- (
  clinical_boxplot_list[[1]] |
  clinical_boxplot_list[[2]] |
  clinical_boxplot_list[[3]]
) / (
  clinical_boxplot_list[[4]] |
  clinical_boxplot_list[[5]] |
  clinical_boxplot_list[[6]]
)

clinical_boxplot_panel

ggsave(
  "clinical_boxplot_panels.png",
  plot = clinical_boxplot_panel,
  path = dir.results,
  width = 22,
  height = 14,
  dpi = 300
)
```
#----------------------------
# Plasma protein matrix setup
#----------------------------
```{r}
dat_plasma <- t1d_hc_dat_soma %>%
  dplyr::select(record_id, starts_with("seq") &
                  !dplyr::ends_with("urine") &
                  !dplyr::ends_with("urine_cradj"))  %>% as.data.frame()

rownames(dat_plasma) <- dat_plasma$record_id
dat_plasma$record_id <- NULL

# align metadata and matrix
ids_shared <- intersect(meta2$record_id, rownames(dat_plasma))

dat_plasma2 <- dat_plasma[ids_shared, , drop = FALSE]

meta2 <- meta2 %>%
  dplyr::filter(record_id %in% ids_shared)

meta2 <- meta2[match(rownames(dat_plasma2), meta2$record_id), ]

stopifnot(identical(rownames(dat_plasma2), meta2$record_id))
```

#--------------------------------------
# Filter to SomaScan 7000 analyte panel
#--------------------------------------
```{r}
analytes <- read.csv(
  file.path(root_path, "Data Harmonization", "Combined SomaScan", "analytes.csv")
)

prot_panel <- analytes$AptName %>%
  as.character() %>%
  trimws() %>%
  unique()

prot_panel <- prot_panel[!is.na(prot_panel) & prot_panel != ""]

aptamers_keep <- intersect(prot_panel, colnames(dat_plasma2))

if (length(aptamers_keep) == 0) {
  stop("No analytes from analytes.csv matched columns in dat_plasma2.")
}

dat_plasma3 <- dat_plasma2[, aptamers_keep, drop = FALSE]

cat("Plasma features before analyte filter:", ncol(dat_plasma2), "\n")
cat("Proteins in analyte panel:", length(prot_panel), "\n")
cat("Matched proteins kept:", length(aptamers_keep), "\n")
cat("Panel proteins missing from plasma matrix:", length(setdiff(prot_panel, colnames(dat_plasma2))), "\n")

dat_plasma3 <- as.data.frame(dat_plasma3)

dat_plasma3[] <- lapply(dat_plasma3, function(x) {
  suppressWarnings(as.numeric(as.character(x)))
})

dat_plasma3 <- as.matrix(dat_plasma3)

dat_log <- log2(dat_plasma3 + 1)
dat_log[!is.finite(dat_log)] <- NA_real_
```
#------------------------------------------------
#filtering out participants with no proteomic data
#-------------------------------------------------
```{r}
sample_nonmissing <- rowSums(!is.na(dat_log))

cat("Participants before sample filter:", nrow(dat_log), "\n")
cat("Participants with zero proteins:", sum(sample_nonmissing == 0), "\n")

keep_samples <- sample_nonmissing > 0

dat_log <- dat_log[keep_samples, , drop = FALSE]

meta2 <- meta2 %>%
  dplyr::filter(record_id %in% rownames(dat_log))

meta2 <- meta2[match(rownames(dat_log), meta2$record_id), ]

stopifnot(identical(rownames(dat_log), meta2$record_id))
stopifnot(identical(rownames(dat_log), meta2$record_id))

meta2 %>%
  dplyr::count(study, group_bmi, name = "n") %>%
  tidyr::pivot_wider(
    names_from = group_bmi,
    values_from = n,
    values_fill = 0
  )
```
#----------------------------------------------------
# Study presence filter
# Protein present in >=60% samples in >=2 studies
# doesn't remove anything in this case but a good check
#------------------------------------------------------
```{r}
meta2$study <- factor(meta2$study)

study_idx_all <- split(seq_len(nrow(dat_log)), meta2$study)

present_by_study_all <- sapply(study_idx_all, function(idx) {
  colMeans(!is.na(dat_log[idx, , drop = FALSE])) >= 0.6
})

present_by_study_all <- as.matrix(present_by_study_all)

n_studies_present_all <- rowSums(present_by_study_all)
keep_study_all <- n_studies_present_all >= 2

kept_proteins <- colnames(dat_log)[keep_study_all]
removed_proteins <- colnames(dat_log)[!keep_study_all]

cat("All proteomic features before study filter:", ncol(dat_log), "\n")
cat("Removed:", length(removed_proteins), "\n")
cat("Kept:", length(kept_proteins), "\n")

dat_log_filt <- dat_log[, kept_proteins, drop = FALSE]

stopifnot(identical(rownames(dat_log_filt), meta2$record_id))
```
#-------------------------
#imputing the missing data
#-------------------------
```{r}
set.seed(1835)

dat_imp <- dat_log_filt

meds <- apply(dat_imp, 2, function(x) {
  m <- median(x, na.rm = TRUE)
  if (is.finite(m)) m else NA_real_
})

for (j in seq_len(ncol(dat_imp))) {
  miss_j <- is.na(dat_imp[, j])
  if (!is.na(meds[j]) && any(miss_j)) {
    dat_imp[miss_j, j] <- meds[j]
  }
}

vars <- apply(dat_imp, 2, var)

keep_cols <- colSums(is.na(dat_imp)) == 0 &
  is.finite(vars) &
  vars > 0

dat_imp <- dat_imp[, keep_cols, drop = FALSE]

stopifnot(identical(rownames(dat_imp), meta2$record_id))
stopifnot(sum(is.na(dat_imp)) == 0)
```

#--------------------------
#checking for outliers
#--------------------------
```{r}
pca0 <- prcomp(dat_imp, center = TRUE, scale. = TRUE)

scores <- as.data.frame(pca0$x[, 1:2, drop = FALSE])
scores$record_id <- rownames(scores)
scores$dist <- sqrt(scores$PC1^2 + scores$PC2^2)

threshold <- mean(scores$dist, na.rm = TRUE) +
  3 * sd(scores$dist, na.rm = TRUE)

outliers <- scores$record_id[scores$dist > threshold] #2 outliers CRC-42 & 10424
outliers
dat_imp2 <- dat_imp[!rownames(dat_imp) %in% outliers, , drop = FALSE]

meta3 <- meta2 %>%
  dplyr::filter(!record_id %in% outliers)

meta3 <- meta3[match(rownames(dat_imp2), meta3$record_id), ]

stopifnot(identical(rownames(dat_imp2), meta3$record_id))
```
#----------------------------
#Data table: study x group_bmi
#-----------------------------
```{r}
study_group_bmi_table <- meta3 %>%
  dplyr::count(study, group_bmi, name = "n") %>%
  tidyr::pivot_wider(
    names_from = group_bmi,
    values_from = n,
    values_fill = 0
  ) %>%
  dplyr::mutate(
    Total = rowSums(dplyr::across(where(is.numeric)))
  )

study_group_bmi_table %>%
  kable(caption = "Participants by study and disease-adiposity group") %>%
  kable_styling(full_width = FALSE)
```
#----------------------------------------
#pretty clincial table 
#-----------------------------------------
```{r}
clinical_df <- meta3 %>%
  mutate(
    Group = as.factor(group),
    Age = as.numeric(age),
    Sex = as.factor(sex),
    HbA1c = as.numeric(hba1c),
    uACR  = as.numeric(acr_u),
    eGFR  = as.numeric(eGFR_CKiD_U25_avg),
    WHtR  = as.numeric(whtr),
    BMI   = as.numeric(bmi),
    Cortical_R = as.numeric(avg_c_r2),
    Medullary_R = as.numeric(avg_k_r2),
    Adiposity_group = factor(
      bmi_obesity,
      levels = c("Normal", "Overweight", "Obese")
    )
  ) %>%
  dplyr::select(
    Adiposity_group,
    Group, Age, Sex, HbA1c, uACR, eGFR, WHtR, BMI,
    Cortical_R, Medullary_R
  )

clinical_summary_table <- clinical_df %>%
  tbl_summary(
    by = Adiposity_group,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      uACR ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_continuous() ~ 1,
      uACR ~ 1,
      HbA1c ~ 1,
      WHtR ~ 2
    ),
    label = list(
      Group ~ "Group",
      Age ~ "Age (Years)",
      Sex ~ "Sex",
      HbA1c ~ "HbA1c (%)",
      uACR ~ "UACR (ug/g)",
      eGFR ~ "eGFR",
      WHtR ~ "WHtR",
      BMI ~ "BMI (kg/m²)",
      Cortical_R ~ "Cortical R2*",
      Medullary_R ~ "Medullary R2*"
    ),
    missing = "no"
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(
    test = list(
      all_continuous() ~ "kruskal.test",
      all_categorical() ~ "chisq.test"
    )
  ) %>%
  modify_header(
    label ~ "**Characteristic**",
    stat_0 ~ "**Total**",
    all_stat_cols() ~ "**{level}**<br>N = {n}",
    p.value ~ "**p value**"
  ) %>%
  bold_labels()

clinical_summary_table
```
#----------------------------------------
#pretty clincial table with t1d group only
#-----------------------------------------
```{r}
clinical_df_t1d <- meta3 %>%
  filter(group == "T1D") %>%
  mutate(
    Group = as.factor(group),
    Age = as.numeric(age),
    Sex = as.factor(sex),
    HbA1c = as.numeric(hba1c),
    uACR  = as.numeric(acr_u),
    eGFR  = as.numeric(eGFR_CKiD_U25_avg),
    WHtR  = as.numeric(whtr),
    BMI   = as.numeric(bmi),
    Cortical_R = as.numeric(avg_c_r2),
    Medullary_R = as.numeric(avg_k_r2),
    Adiposity_group = factor(
      bmi_obesity,
      levels = c("Normal", "Overweight", "Obese")
    )
  ) %>%
  dplyr::select(
    Adiposity_group,
    Group, Age, Sex, HbA1c, uACR, eGFR, WHtR, BMI,
    Cortical_R, Medullary_R
  )

clinical_summary_table_t1d <- clinical_df_t1d %>%
  tbl_summary(
    by = Adiposity_group,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      uACR ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_continuous() ~ 1,
      uACR ~ 1,
      HbA1c ~ 1,
      WHtR ~ 2
    ),
    label = list(
      Age ~ "Age (Years)",
      Sex ~ "Sex (Female)",
      HbA1c ~ "HbA1c (%)",
      uACR ~ "UACR (ug/g)",
      eGFR ~ "eGFR",
      WHtR ~ "WHtR",
      BMI ~ "BMI (kg/m²)",
      Cortical_R ~ "Cortical R2*",
      Medullary_R ~ "Medullary R2*"
    ),
    missing = "no"
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(
    test = list(
      all_continuous() ~ "kruskal.test",
      all_categorical() ~ "chisq.test"
    )
  ) %>%
  modify_header(
    label ~ "**Characteristic**",
    stat_0 ~ "**Total**",
    all_stat_cols() ~ "**{level}**<br>N = {n}",
    p.value ~ "**p value**"
  ) %>%
  bold_labels()

clinical_summary_table_t1d
```

```{r}
stopifnot(identical(rownames(dat_imp2), meta3$record_id))
design <- model.matrix(~ group_bmi, data = meta3)

#model 0 ~ no adjustment
p0 <- make_pca(dat_imp2,meta3,"No adjustment")

#model 1 ~ age + sex adjustment
cov_age_sex <- model.matrix(~ age + sex, data = meta3)[, -1, drop = FALSE]
prot_age_sex <- t(removeBatchEffect(t(dat_imp2),
  covariates = cov_age_sex,
  design = design
))
p1 <- make_pca(prot_age_sex,meta3,"Adjusted for age + sex")

#model 2 ~ age + sex + study adjustment
cov_age_sex_study <- model.matrix(~ age + sex + study, data = meta3)[, -1, drop = FALSE]
prot_age_sex_study <- t(removeBatchEffect(t(dat_imp2),
  covariates = cov_age_sex_study,
  design = design
))
p2 <- make_pca(prot_age_sex_study, meta3, "Adjusted for age + sex + cohort")

#model 3b ~ age * HbA1c + sex + study adjustment
hba1c_var <- "hba1c"
cov_hba1c_age_study <- model.matrix(
  as.formula(paste0("~ age * ", hba1c_var, " + sex + study")),
  data = meta3
)[, -1, drop = FALSE]

# final row checks
nrow(design)
nrow(cov_hba1c_age_study)
stopifnot(nrow(design) == nrow(cov_hba1c_age_study))
stopifnot(nrow(design) == nrow(dat_imp2))

prot_hba1c_age_study <- t(removeBatchEffect(t(dat_imp2), covariates = cov_hba1c_age_study, design = design))
p3b <- make_pca(prot_hba1c_age_study, meta3,"Adjusted for age * HbA1c + sex + study")

#model 3a
cov_hba1c_age <- model.matrix(as.formula(paste0("~ age * ", hba1c_var, " + sex")), data = meta3)[, -1, drop = FALSE]
prot_hba1c_age <- t(removeBatchEffect(t(dat_imp2), covariates = cov_hba1c_age, design = design))
p3a <- make_pca(prot_hba1c_age, meta3,"Adjusted for age * HbA1c + sex ")


# combine plots
pca_group_bmi_all <- (p0 | p1 | p3a) / (p2 | p3b | patchwork::plot_spacer())
pca_group_bmi_all

ggsave(filename = file.path(dir.results, "Plots", "pca_m2.png"),
       width = 7,
       height = 5,
       dpi = 300)
```
```{r}
#-----------------------------------------
# PERMANOVA + beta dispersion for M2 PCA
#-----------------------------------------
# distance matrix from adjusted proteomic matrix
dist_m2 <- dist(prot_age_sex_study, method = "euclidean")
set.seed(1835)

permanova_m2 <- adonis2(
  dist_m2 ~ group_bmi,
  data = meta3,
  permutations = 999
)

permanova_m2

beta_m2 <- betadisper(
  dist_m2,
  group = meta3$group_bmi
)
beta_m2
# overall dispersion test
anova(beta_m2)

# permutation test
set.seed(1835)

permutest(beta_m2, permutations = 999)
TukeyHSD(beta_m2)
```
#--------------------------------------------------
#running limma for global T1D proteomic differences
#--------------------------------------------------
```{r}
meta_t1d <- meta3 %>%
  filter(group == "T1D") %>%
  droplevels()

mat_t1d <- t(dat_imp2)
mat_t1d <- mat_t1d[, meta_t1d$record_id]

stopifnot(all(colnames(mat_t1d) == meta_t1d$record_id))
```


```{r}
res_t1d_global_m0 <- run_t1d_global_adiposity_limma(
  mat = mat_t1d,
  meta_df = meta_t1d,
  design_formula = ~ 0 + group_bmi,
  model_label = "T1D M0: no adjustment"
)

res_t1d_global_m1 <- run_t1d_global_adiposity_limma(
  mat = mat_t1d,
  meta_df = meta_t1d,
  design_formula = ~ 0 + group_bmi + age + sex,
  model_label = "T1D M1: age + sex"
)

res_t1d_global_m2 <- run_t1d_global_adiposity_limma(
  mat = mat_t1d,
  meta_df = meta_t1d,
  design_formula = ~ 0 + group_bmi + age + sex + study,
  model_label = "T1D M2: age + sex + study"
)

design_t1d_m3 <- as.formula(
  paste0("~ 0 + group_bmi + age * ", hba1c_var, " + sex + study")
)

res_t1d_global_m3 <- run_t1d_global_adiposity_limma(
  mat = mat_t1d,
  meta_df = meta_t1d,
  design_formula = design_t1d_m3,
  model_label = "T1D M3: age * HbA1c + sex + study"
)
```
#-------------------------------
#T1D all proteomic results
#-------------------------------
```{r}
res_t1d_global_all <- bind_rows(
  res_t1d_global_m0,
  res_t1d_global_m1,
  res_t1d_global_m2,
  res_t1d_global_m3
)

res_t1d_global_all

sig_t1d_global <- res_t1d_global_all %>%
  filter(adj.P.Val < 0.05)

sig_t1d_global_fc <- res_t1d_global_all %>%
  filter(adj.P.Val < 0.05, abs(adj.P.Val) >= 0.5)
```
#-------------------------------------------------------
#setting up mats for global and pairwise DEP anlaysis
#--------------------------------------------------------
```{r}
mat1 <- t(dat_imp2)
mat1<- mat1[, meta3$record_id]

stopifnot(all(colnames(mat1) == meta3$record_id))
```

#---------------------------------
#results for all models pairwise
#---------------------------------
```{r}
# Model 0: group_bmi only / no covariate adjustment
res_m0 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi,
  model_label = "M0: group_bmi only"
)

# Model 1: group_bmi + age + sex
res_m1 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi + age + sex,
  model_label = "M1: group_bmi + age + sex"
)

#Model 2: group_bmi + age + sex + study
res_m2 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = ~ 0 + group_bmi + age + sex + study,
  model_label = "M2: group_bmi + age + sex + study"
)


# Model 3: group_bmi + age * HbA1c + sex + study
design2 <- as.formula(
  paste0("~ 0 + group_bmi + age * ", hba1c_var, " + sex + study")
)

res_m3 <- run_group_bmi_limma(
  mat = mat1,
  meta_df = meta3,
  design_formula = design2,
  model_label = "M3: group_bmi + age * HbA1c + sex + study"
)
```

```{r}
clean_contrast_names <- function(df) {
  df %>%
    mutate(
      contrast = case_when(
        contrast == "Global_group_bmi" ~ "Global_group_bmi",
        contrast == "T1DNormal_vs_LCNormal" ~ "T1D_Normal_vs_LC_Normal",
        contrast == "T1DOverObese_vs_LCOverObese" ~ "T1D_OverObese_vs_LC_OverObese",
        contrast == "T1DNormal_vs_T1DOverweight" ~ "T1D_Normal_vs_T1D_Overweight",
        contrast == "T1DNormal_vs_T1DObese" ~ "T1D_Normal_vs_T1D_Obese",
        contrast == "T1DObese_vs_T1DOverweight" ~ "T1D_Obese_vs_T1D_Overweight",
        TRUE ~ contrast
      )
    )
}

res_m0 <- clean_contrast_names(res_m0)
res_m1 <- clean_contrast_names(res_m1)
res_m2 <- clean_contrast_names(res_m2)
res_m3 <- clean_contrast_names(res_m3)

results_all_models <- bind_rows(
  res_m0,
  res_m1,
  res_m2,
  res_m3
)


sig_all_models <- results_all_models %>% filter(adj.P.Val < 0.05)
sig_all_models_fc <- results_all_models %>% filter(adj.P.Val < 0.05, abs(logFC) >= 0.5)
sig_counts <- sig_all_models %>% 
  filter((adj.P.Val < 0.05 & abs(logFC) >= 0.5)) %>%
  count(model, contrast)

sig_counts %>% filter(model == "M2: group_bmi + age + sex + study")

#write_csv(sig_counts, 
 #         file = file.path(dir.results, "sig_counts_by_model.csv"))
#write_csv(sig_all_models_fc, 
 #         file = file.path(dir.results, "limma_results_all_models.csv"))
```

```{r}
map_tbl <- analytes %>%
  dplyr::transmute(
    protein_id = as.character(AptName),
    gene_symbol = as.character(EntrezGeneSymbol)
  ) %>%
  dplyr::mutate(
    gene_symbol = dplyr::if_else(
      is.na(gene_symbol) | gene_symbol == "",
      protein_id,
      gene_symbol
    )
  ) %>%
  dplyr::distinct()


```
#----------------------------------------
#heatmap of top 50 DEPs across all groups
#-----------------------------------------
```{r}
hm <- make_global_group_bmi_heatmap(
  results_df = results_all_models,
  expr_mat = mat1,
  meta_df = meta3,
  model_name = "M2: group_bmi + age + sex + study",
  sample_id_col = "record_id",
  padj_cutoff = 0.05,
  top_n = 50
)


png(
  filename = file.path(dir.results, "Plots", "Global_Heatmap.png"),
  width = 10,
  height = 8,
  units = "in",
  res = 300
)

grid::grid.newpage()
grid::grid.draw(hm$gtable)

dev.off()
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m0 <- make_volcano_list(
  results_df = res_m0,
  model_name = "M0: group_bmi only",
  label_n = 5
)

top_row <- plots_m0[[2]] + plots_m0[[3]] + patchwork::plot_spacer()
bottom_row <- plots_m0[[4]] + plots_m0[[5]] + plots_m0[[6]]

panel_m0_5 <- top_row / bottom_row

panel_m0_5

ggsave("M0_pca_plots.png",
       plot = panel_m0_5,
       path = dir.results,
       width = 16,
       height =12,
       dpi = 300
       )
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m1 <- make_volcano_list(
  results_df = res_m1,
  model_name = "M1: group_bmi + age + sex",
  label_n = 5
)

top_row <- plots_m1[[2]] + plots_m1[[3]] + patchwork::plot_spacer()
bottom_row <- plots_m1[[4]] + plots_m1[[5]] + plots_m1[[6]]

panel_m1_5 <- top_row / bottom_row

panel_m1_5

ggsave("M1_pca_plots.png",
       plot = panel_m1_5,
       path = dir.results,
       width = 16,
       height =12,
       dpi = 300
       )

```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m2 <- make_volcano_list(
  results_df = res_m2,
  model_name = "M2: group_bmi + age + sex + study",
  label_n = 10
)

top_row <- plots_m2[[2]] + plots_m2[[3]] + patchwork::plot_spacer()
bottom_row <- plots_m2[[4]] + plots_m2[[5]] + plots_m2[[6]]

panel_m2_5 <- top_row / bottom_row
panel_m2_5

ggsave("M2_volcano_plots.png",
       plot = panel_m2_5,
       path = dir.results,
       width = 20,
       height =12,
       dpi = 300
       )

#plots_m2[[1]]
#ggsave(filename = file.path(dir.results, "Plots", "T1DNormal_LCNormal_volcano.png"),
 #      width = 8,
  #     height =6,
   #    dpi = 300)

#plots_m2[[3]]
#ggsave(filename = file.path(dir.results, "Plots", "T1D_OO_LC_OO_volcano.png"),
 #      width = 8,
  #     height =6,
   #    dpi = 300)

#plots_m2[[3]]
#ggsave(filename = file.path(dir.results, "Plots", "T1DNormal_T1DOverweight_volcano.png"),
 #      width = 8,
  #     height =6,
   #    dpi = 300)
#plots_m2[[4]]
#ggsave(filename = file.path(dir.results, "Plots", "T1DNormal_T1DObese_volcano.png"),
 #      width = 8,
  #     height =6,
   #    dpi = 300)
#plots_m2[[5]]

#ggsave(filename = file.path(dir.results, "Plots", "T1DObese_T1DOverweight_volcano.png"),
     #  width = 8,
      # height =6,
      # dpi = 300)
```

```{r}
#| fig-width: 16
#| fig-height: 12

plots_m3 <- make_volcano_list(
  results_df = res_m3,
  model_name = "M3: group_bmi + age*hba1c + sex + study",
  label_n = 5
)

top_row <- plots_m3[[2]] + plots_m3[[3]] + patchwork::plot_spacer()
bottom_row <- plots_m3[[4]] + plots_m3[[5]] + plots_m3[[6]]

panel_m3_5 <- top_row / bottom_row

panel_m3_5
ggsave("M3_pca_plots.png",
       plot = panel_m3_5,
       path = dir.results,
       width = 16,
       height =12,
       dpi = 300
       )
```
#-----------------
#pathway anlaysis
#-----------------
```{r}
hallmark_msig <- msigdbr(
  species = "Homo sapiens",
  collection = "H"
)

hallmark_list <- hallmark_msig %>%
  split(x = .$gene_symbol, f = .$gs_name)

msig_kegg <- msigdbr(
  species = "Homo sapiens",
  collection = "C2",
  subcollection = "CP:REACTOME"
)

kegg_list <- msig_kegg %>%
  split(x = .$gene_symbol, f = .$gs_name)

# GO Biological Process
msig_go_bp <- msigdbr(
  species = "Homo sapiens",
  collection = "C5",
  subcollection = "GO:BP"
)

go_bp_list <- msig_go_bp %>%
  split(x = .$gene_symbol, f = .$gs_name)

# GO Molecular Function
msig_go_mf <- msigdbr(
  species = "Homo sapiens",
  collection = "C5",
  subcollection= "GO:MF"
)

go_mf_list <- msig_go_mf %>%
  split(x = .$gene_symbol, f = .$gs_name)

# GO Cellular Component
msig_go_cc <- msigdbr(
  species = "Homo sapiens",
  collection = "C5",
  subcollection = "GO:CC"
)

go_cc_list <- msig_go_cc %>%
  split(x = .$gene_symbol, f = .$gs_name)
```

```{r}
protein_gene_map <- analytes %>%
  dplyr::mutate(
    protein_id = trimws(as.character(AptName)),
    gene_symbol = trimws(as.character(EntrezGeneSymbol))  # change if needed
  ) %>%
  dplyr::select(protein_id, gene_symbol) %>%
  dplyr::filter(
    !is.na(protein_id),
    protein_id != "",
    !is.na(gene_symbol),
    gene_symbol != ""
  ) %>%
  dplyr::distinct()

results_limma_annot <- results_all_models %>%
  dplyr::left_join(protein_gene_map, by = "protein_id")

#mapping success rate
results_limma_annot %>%
  summarise(
    n_total = n(),
    n_mapped = sum(!is.na(gene_symbol)),
    pct_mapped = round(100 * mean(!is.na(gene_symbol)), 1)
  )
```

```{r}
#hallmark gene list
fgsea_hm_m2 <- run_fgsea_model(res_df = res_m2, model_name = "M2: group_bmi age + sex + study", pathways = hallmark_list, pathway_type = "H", protein_gene_map = protein_gene_map)
fgsea_hm_m2 <- fgsea_hm_m2 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::mutate(
    pathway = stringr::str_remove(pathway, "^HALLMARK_")
  ) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_hm_m2

#kegg results
fgsea_kegg_m0 <- run_fgsea_model(res_df = res_m0, model_name = "M0: group_bmi only", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m0 <- fgsea_kegg_m0 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_kegg_m1 <- run_fgsea_model(res_df = res_m1, model_name = "M1: group_bmi age + sex", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m1 <- fgsea_kegg_m1 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_kegg_m2 <- run_fgsea_model(res_df = res_m2, model_name = "M2: group_bmi age + sex + study", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m2 <- fgsea_kegg_m2 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::mutate(
    pathway = stringr::str_remove(pathway, "^REACTOME_")
  ) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_kegg_m3 <- run_fgsea_model(res_df = res_m3, model_name = "M3: group_bmi age*hba1c + sex +study", pathways = kegg_list, pathway_type = "KEGG", protein_gene_map = protein_gene_map)
fgsea_kegg_m3 <- fgsea_kegg_m3 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

#GO:BP
fgsea_BP_m2 <- run_fgsea_model(res_df = res_m2, model_name = "M2: group_bmi age + sex + study", pathways = go_bp_list, pathway_type = "GO:BP", protein_gene_map = protein_gene_map)

fgsea_BP_m2 <- fgsea_BP_m2 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

#GO:MF
fgsea_MF_m2 <- run_fgsea_model(res_df = res_m2, model_name = "M2: group_bmi age + sex + study", pathways = go_mf_list, pathway_type = "GO:MF", protein_gene_map = protein_gene_map)

fgsea_MF_m2 <- fgsea_MF_m2 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)
fgsea_MF_m2

#GO:CC
fgsea_CC_m2 <- run_fgsea_model(res_df = res_m2, model_name = "M2: group_bmi age + sex + study", pathways = go_cc_list, pathway_type = "GO:CC", protein_gene_map = protein_gene_map)

fgsea_CC_m2 <- fgsea_CC_m2 %>%
  dplyr::filter(padj < 0.05) %>%
  dplyr::arrange(model, contrast, pathway_type, padj)

fgsea_CC_m2
```

```{r}
fgsea_all_m2 <- dplyr::bind_rows(
  fgsea_kegg_m2,
  fgsea_BP_m2,
  fgsea_MF_m2,
  fgsea_CC_m2
) %>%
  dplyr::filter(!is.na(padj), !is.na(NES)) %>%
  dplyr::mutate(
    pathway_type = factor(
      pathway_type,
      levels = c("KEGG", "GO:BP", "GO:MF", "GO:CC")
    )
  )
```

```{r}
#| fig-width: 22
#| fig-height: 16
fgsea_kegg_plots_m0 <- fgsea_kegg_m0 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m0,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

#fgsea_kegg_plots_m0$plot[[1]]
#fgsea_kegg_plots_m0$plot[[2]]
#fgsea_kegg_plots_m0$plot[[3]]
#fgsea_kegg_plots_m0$plot[[4]]
#fgsea_kegg_plots_m0$plot[[5]]

plots_m0 <- fgsea_kegg_plots_m0
top_row <- plots_m0$plot[[1]] + plots_m0$plot[[5]] + patchwork::plot_spacer()
bottom_row <- plots_m0$plot[[3]] + plots_m0$plot[[2]] + plots_m0$plot[[4]]

panel_kegg_m0 <- top_row / bottom_row

panel_kegg_m0

ggsave("GSEA_kegg_m0.png",
       plot = panel_kegg_m0,
       path = dir.results,
       width = 22,
       height =16,
       dpi = 300
       )
```

```{r}
#| fig-width: 18
#| fig-height: 16
fgsea_kegg_plots_m1 <- fgsea_kegg_m1 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m1,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

plots_m1 <- fgsea_kegg_plots_m1
top_row <- plots_m1$plot[[1]] + plots_m1$plot[[5]] + patchwork::plot_spacer()
bottom_row <- plots_m1$plot[[3]] + plots_m1$plot[[2]] + plots_m1$plot[[4]]

panel_kegg_m1 <- top_row / bottom_row

panel_kegg_m1

ggsave("GSEA_kegg_m1.png",
       plot = panel_kegg_m1,
       path = dir.results,
       width = 20,
       height =16,
       dpi = 300
       )
```

```{r}
#| fig-width: 28
#| fig-height: 24
fgsea_kegg_plots_m2 <- fgsea_kegg_m2 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m2,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

plots_kegg_m2 <- fgsea_kegg_plots_m2
top_row <- plots_kegg_m2$plot[[1]] + patchwork::plot_spacer()+ patchwork::plot_spacer()
bottom_row <- plots_kegg_m2$plot[[3]] + plots_kegg_m2$plot[[2]] + plots_kegg_m2$plot[[4]]

panel_kegg_m2 <- top_row / bottom_row

panel_kegg_m2

ggsave("GSEA_kegg_m2.png",
       plot = panel_kegg_m2,
       path = dir.results,
       width = 28,
       height =24,
       dpi = 300
       )

```

```{r}
#| fig-width: 20
#| fig-height: 16
fgsea_kegg_plots_m3 <- fgsea_kegg_m3 %>%
  distinct(model, contrast, pathway_type) %>%
  mutate(
    plot = purrr::pmap(
      list(model, contrast, pathway_type),
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m3,
        model_name = ..1,
        contrast_name = ..2,
        pathway_type_name = ..3,
        top_n = 15
      )
    )
  )

plots_m3 <- fgsea_kegg_plots_m3
top_row <- plots_m3$plot[[1]] + patchwork::plot_spacer() + patchwork::plot_spacer()
bottom_row <- plots_m3$plot[[3]] + plots_m3$plot[[2]] + patchwork::plot_spacer()

panel_kegg_m3 <- top_row / bottom_row

panel_kegg_m3

ggsave("GSEA_kegg_m3.png",
       plot = panel_kegg_m3,
       path = dir.results,
       width = 20,
       height =16,
       dpi = 300
       )
```

```{r}
fgsea_kegg_plots <- fgsea_kegg_m2 %>%
  dplyr::distinct(model, contrast) %>%
  dplyr::mutate(
    plot = purrr::map2(
      model,
      contrast,
      ~ plot_fgsea_top20(
        fgsea_df = fgsea_kegg_m2,
        model_name = .x,
        contrast_name = .y,
        top_n = 20
      )
    )
  )
fgsea_kegg_plots$plot[[1]]
ggsave(filename = file.path(dir.results,"Plots", "GSEA_T1D.NW_LC.NW.png"), 
       width = 10,
       height = 12,
       dpi = 300)

fgsea_kegg_plots$plot[[2]]
ggsave(filename = file.path(dir.results,"Plots", "GSEA_T1D_NW_OB.png"), 
       width = 10,
       height =16,
       dpi = 300)


fgsea_kegg_plots$plot[[3]]
ggsave(filename = file.path(dir.results,"Plots", "GSEA_T1D_NW_OW.png"), 
       width = 10,
       height = 12,
       dpi = 300)

fgsea_kegg_plots$plot[[4]]
ggsave(filename = file.path(dir.results,"Plots", "GSEA_T1D_OB_OW.png"), 
       width = 10,
       height = 12,
       dpi = 300)

#fgsea_kegg_plots$plot[[5]]
#ggsave(filename = file.path(dir.results,"Plots", "GSEA_T1D_OB_OW.png"), 
   #    width = 10,
    #   height =12,
     #  dpi = 300)


#fgsea_kegg_plots$plot[[6]]
#ggsave(filename = file.path(dir.results,"Plots", "GSEA_T1D_NW_OB.png"), 
 #      width = 10,
  #     height = 12,
   #    dpi = 300)
```


#----------------------------------------
#ElasticNet/LASSO Regression
#----------------------------------------
```{r}
run_elastic_net <- function(
  contrast_name,
  limma_df,
  meta_df,
  X_mat,
  protein2symbol = NULL,
  group_var = "group_bmi",
  top_n = 100,
  sig_only = TRUE,
  p_cutoff = 0.05,
  fc_cutoff = 0.5,
  covars = c("age", "sex", "study"),
  alpha_grid = seq(0.1, 1, by = 0.1),
  nfolds = 10,
  seed = 1835,
  min_per_class = 5,
  maxit = 1e6
) {

  cat("\n============================\n")
  cat("Running contrast:", contrast_name, "\n")

  contrast_groups <- list(
    T1D_Normal_vs_LC_Normal = list(
      positive = "T1D_Normal",
      negative = "LC_Normal"
    ),
    T1D_OverObese_vs_LC_OverObese = list(
      positive = c("T1D_Overweight", "T1D_Obese"),
      negative = "LC_Overweight_Obese"
    ),
    T1D_Normal_vs_T1D_Overweight = list(
      positive = "T1D_Normal",
      negative = "T1D_Overweight"
    ),
    T1D_Normal_vs_T1D_Obese = list(
      positive = "T1D_Normal",
      negative = "T1D_Obese"
    ),
    T1D_Obese_vs_T1D_Overweight = list(
      positive = "T1D_Obese",
      negative = "T1D_Overweight"
    )
  )

  if (!contrast_name %in% names(contrast_groups)) {
    stop("Contrast not found in contrast_groups: ", contrast_name)
  }

  g_pos <- contrast_groups[[contrast_name]]$positive
  g_neg <- contrast_groups[[contrast_name]]$negative

  meta_sub <- meta_df %>%
    dplyr::filter(.data[[group_var]] %in% c(g_pos, g_neg)) %>%
    dplyr::mutate(
      record_id = as.character(record_id),
      y = ifelse(.data[[group_var]] %in% g_pos, 1, 0),
      sex = factor(sex),
      study = factor(study)
    )

  if (nrow(meta_sub) == 0) {
    stop("No samples found for contrast: ", contrast_name)
  }

  tab <- table(meta_sub$y)
  print(tab)

  if (any(tab < min_per_class)) {
    stop(
      "Too few samples in at least one class for contrast: ",
      contrast_name,
      " | counts = ",
      paste(names(tab), tab, collapse = ", ")
    )
  }

  ids <- intersect(meta_sub$record_id, rownames(X_mat))

  meta_sub <- meta_sub %>%
    dplyr::filter(record_id %in% ids)

  meta_sub <- meta_sub[match(ids, meta_sub$record_id), , drop = FALSE]
  X_mat_sub <- X_mat[ids, , drop = FALSE]

  stopifnot(identical(rownames(X_mat_sub), meta_sub$record_id))

  limma_sub <- limma_df %>%
    dplyr::filter(contrast == contrast_name)

  if (nrow(limma_sub) == 0) {
    stop("No limma rows found for contrast: ", contrast_name)
  }

  if (sig_only) {
    limma_sub <- limma_sub %>%
      dplyr::filter(
        adj.P.Val < p_cutoff,
        abs(logFC) > fc_cutoff
      )
  }

  if (nrow(limma_sub) == 0) {
    stop("No significant limma DEPs found for contrast: ", contrast_name)
  }

  if (!is.null(protein2symbol)) {
    limma_sub <- limma_sub %>%
      dplyr::left_join(protein2symbol, by = "protein_id") %>%
      dplyr::mutate(
        gene_symbol = dplyr::if_else(
          is.na(gene_symbol) | gene_symbol == "",
          protein_id,
          gene_symbol
        )
      ) %>%
      dplyr::filter(protein_id %in% colnames(X_mat_sub)) %>%
      dplyr::arrange(dplyr::desc(abs(t)), adj.P.Val, P.Value) %>%
      dplyr::group_by(gene_symbol) %>%
      dplyr::slice_head(n = 1) %>%
      dplyr::ungroup()
  } else {
    limma_sub <- limma_sub %>%
      dplyr::filter(protein_id %in% colnames(X_mat_sub)) %>%
      dplyr::arrange(dplyr::desc(abs(t)), adj.P.Val, P.Value) %>%
      dplyr::mutate(gene_symbol = protein_id)
  }

  top_feats_tbl <- limma_sub %>%
    dplyr::arrange(dplyr::desc(abs(t)), adj.P.Val, P.Value) %>%
    dplyr::slice_head(n = top_n)

  top_feats <- top_feats_tbl$protein_id

  cat("Input limma DEPs after redundancy filtering:", nrow(limma_sub), "\n")
  cat("Top proteins used in elastic net:", length(top_feats), "\n")

  if (length(top_feats) < 5) {
    stop("Fewer than 5 overlapping proteins for contrast: ", contrast_name)
  }

  covars_present <- covars[covars %in% colnames(meta_sub)]

  cat("Covariates used:", paste(covars_present, collapse = ", "), "\n")

  if (length(covars_present) > 0) {
    Z <- model.matrix(
      as.formula(paste("~", paste(covars_present, collapse = " + "))),
      data = meta_sub
    )
  } else {
    Z <- matrix(1, nrow = nrow(meta_sub), ncol = 1)
    colnames(Z) <- "(Intercept)"
  }

  Xp <- X_mat_sub[, top_feats, drop = FALSE]

  X <- cbind(Z, Xp)
  X <- as.matrix(X)
  storage.mode(X) <- "double"

  y <- as.numeric(meta_sub$y)

  penalty_factor <- c(rep(0, ncol(Z)), rep(1, ncol(Xp)))

  sds <- apply(X, 2, sd)
  keep <- is.finite(sds) & sds > 0

  X <- X[, keep, drop = FALSE]
  penalty_factor <- penalty_factor[keep]

  cat("Final elastic net matrix:", nrow(X), "samples x", ncol(X), "features\n")

  nfolds_use <- min(nfolds, sum(y == 0), sum(y == 1))

  cat("nfolds used:", nfolds_use, "\n")

  if (nfolds_use < 2) {
    stop("Not enough samples per class for cross-validation.")
  }

  set.seed(seed)

  fit_list <- lapply(alpha_grid, function(a) {
    cv <- tryCatch(
      glmnet::cv.glmnet(
        x = X,
        y = y,
        family = "binomial",
        alpha = a,
        nfolds = nfolds_use,
        type.measure = "deviance",
        penalty.factor = penalty_factor,
        standardize = TRUE,
        maxit = maxit
      ),
      error = function(e) {
        message("glmnet failed at alpha = ", a, ": ", conditionMessage(e))
        NULL
      }
    )

    if (is.null(cv)) return(NULL)

    list(alpha = a, cvfit = cv)
  })

  fit_list <- Filter(Negate(is.null), fit_list)

  if (length(fit_list) == 0) {
    stop("All glmnet fits failed.")
  }

  best_idx <- which.min(sapply(fit_list, function(z) min(z$cvfit$cvm)))
  best <- fit_list[[best_idx]]

  coef_mat <- as.matrix(coef(best$cvfit, s = "lambda.1se"))

  coef_df <- tibble::tibble(
    protein_id = rownames(coef_mat),
    beta = as.numeric(coef_mat)
  ) %>%
    dplyr::filter(beta != 0)

  selected_proteins <- intersect(top_feats, coef_df$protein_id)

  coef_df_annot <- coef_df %>%
    dplyr::left_join(
      top_feats_tbl %>%
        dplyr::select(protein_id, gene_symbol, logFC, t, P.Value, adj.P.Val),
      by = "protein_id"
    )

  selected_proteins_annot <- top_feats_tbl %>%
    dplyr::filter(protein_id %in% selected_proteins) %>%
    dplyr::select(protein_id, gene_symbol, logFC, t, P.Value, adj.P.Val)

  p_hat <- as.numeric(
    predict(best$cvfit, newx = X, s = "lambda.1se", type = "response")
  )

  auc_val <- as.numeric(
    pROC::auc(pROC::roc(y, p_hat, quiet = TRUE))
  )

  cat("Best alpha:", best$alpha, "\n")
  cat("Training AUC:", auc_val, "\n")
  cat("Selected proteins:", length(selected_proteins), "\n")

  list(
    contrast = contrast_name,
    positive_group = g_pos,
    negative_group = g_neg,
    best_alpha = best$alpha,
    lambda_1se = best$cvfit$lambda.1se,
    lambda_min = best$cvfit$lambda.min,
    auc_train = auc_val,
    n_input_proteins = length(top_feats),
    top_feats = top_feats,
    top_feats_tbl = top_feats_tbl,
    selected_proteins = selected_proteins,
    selected_proteins_annot = selected_proteins_annot,
    coef_df = coef_df,
    coef_df_annot = coef_df_annot,
    cvfit = best$cvfit
  )
}
```

#----------------------
#boostrapping T1D vs LC
#----------------------
```{r}
net1 <- run_elastic_net_bootstrap(
  contrast_name = "T1D_Normal_vs_LC_Normal",
  limma_df = res_m2,
  meta_df = meta3,
  X_mat = prot_age_sex_study,
  protein2symbol = map_tbl,
  top_n = 100,
  sig_only = TRUE,
  p_cutoff = 0.05,
  fc_cutoff = 0.5,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1835
)

```

```{r}
net2 <- run_elastic_net_bootstrap(
  contrast_name = "T1D_OverObese_vs_LC_OverObese",
  limma_df = res_m2,
  meta_df = meta3,
  X_mat = prot_age_sex_study,
  protein2symbol = map_tbl,
  top_n = 100,
  sig_only = TRUE,
  p_cutoff = 0.05,
  fc_cutoff = 0.5,
  covars = c("age", "sex"),
  n_boot = 100,
  seed = 1835
)

```

```{r}
#too few proteins to run elasticnet on
#net3 <- run_elastic_net_bootstrap(
#  contrast_name = "T1D_Normal_vs_T1D_Overweight",
#  limma_df = res_m2,
#  meta_df = meta3,
#  X_mat = prot_age_sex_study,
#  protein2symbol = map_tbl,
#  top_n = 100,
#  sig_only = TRUE,
#  p_cutoff = 0.05,
#  fc_cutoff = 0.5,
#  covars = c("age", "sex"),
#  n_boot = 100,
#  seed = 1835
#)
```

```{r}
net4 <- run_elastic_net_bootstrap(
  contrast_name = "T1D_Normal_vs_T1D_Obese",
  limma_df = res_m2,
  meta_df = meta3,
  X_mat = prot_age_sex_study,
  protein2symbol = map_tbl,
  top_n = 100,
  sig_only = TRUE,
  p_cutoff = 0.05,
  fc_cutoff = 0.5,
  n_boot = 100,
  covars = c("age", "sex"),
  
  seed = 1835
)

```

```{r}
net5 <- run_elastic_net_bootstrap(
 contrast_name = "T1D_Obese_vs_T1D_Overweight",
  limma_df = res_m2,
  meta_df = meta3,
  X_mat = prot_age_sex_study,
  protein2symbol = map_tbl,
  top_n = 100,
  sig_only = TRUE,
  p_cutoff = 0.05,
  fc_cutoff = 0.5,
  n_boot = 100,
  covars = c("age", "sex"),
  
  seed = 1835)
```

#checking which proteins are the most stable across bootstrapping
```{r}
net1_stability_plot <- net1$stability_tbl %>%
  filter(selection_percent >= 75) %>%
  mutate(
    feature_label = ifelse(
      !is.na(gene_symbol) & gene_symbol != "",
      gene_symbol,
      protein_id
    )
  ) %>%
  arrange(desc(selection_percent)) %>%
  mutate(
    feature_label = fct_reorder(feature_label, selection_percent)
  ) %>%
  ggplot(aes(x = selection_percent, y = feature_label)) +
  geom_col(width = 0.9, fill = "grey40") +
  scale_x_continuous(
    limits = c(0, 100),
    breaks = c(0, 25, 50, 75, 100)
  ) +
  labs(
    title = "feature stability: T1D_Normal_vs_LC_Normal",
    x = "Bootstrap selection frequency (%)",
    y = NULL
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

net1_stability_plot
 
```

```{r}
net2_stability_plot <- net2$stability_tbl %>%
  filter(selection_percent >= 75) %>%
  mutate(
    feature_label = ifelse(
      !is.na(gene_symbol) & gene_symbol != "",
      gene_symbol,
      protein_id
    )
  ) %>%
  arrange(desc(selection_percent)) %>%
  mutate(
    feature_label = fct_reorder(feature_label, selection_percent)
  ) %>%
  ggplot(aes(x = selection_percent, y = feature_label)) +
  geom_col(width = 0.9, fill = "grey40") +
  scale_x_continuous(
    limits = c(0, 100),
    breaks = c(0, 25, 50, 75, 100)
  ) +
  labs(
    title = "feature stability: T1D_OverObese_vs_LC_OverObese",
    x = "Bootstrap selection frequency (%)",
    y = NULL
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

net2_stability_plot
```
```{r fig.width=24}
net4_stability_plot <- net4$stability_tbl %>%
  filter(selection_percent >= 50) %>%
  mutate(
    feature_label = ifelse(
      !is.na(gene_symbol) & gene_symbol != "",
      gene_symbol,
      protein_id
    )
  ) %>%
  arrange(desc(selection_percent)) %>%
  mutate(
    feature_label = fct_reorder(feature_label, selection_percent)
  ) %>%
  ggplot(aes(x = selection_percent, y = feature_label)) +
  geom_col(width = 0.9, fill = "grey40") +
  scale_x_continuous(
    limits = c(0, 100),
    breaks = c(0, 25, 50, 75, 100)
  ) +
  labs(
    title = "feature stability: T1D_Normal_vs_T1D_Obese",
    x = "Bootstrap selection frequency (%)",
    y = NULL
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

net4_stability_plot
```

```{r}
net5_stability_plot <- net5$stability_tbl %>%
  filter(selection_percent >= 75) %>%
  mutate(
    feature_label = ifelse(
      !is.na(gene_symbol) & gene_symbol != "",
      gene_symbol,
      protein_id
    )
  ) %>%
  arrange(desc(selection_percent)) %>%
  mutate(
    feature_label = fct_reorder(feature_label, selection_percent)
  ) %>%
  ggplot(aes(x = selection_percent, y = feature_label)) +
  geom_col(width = 0.9, fill = "grey40") +
  scale_x_continuous(
    limits = c(0, 100),
    breaks = c(0, 25, 50, 75, 100)
  ) +
  labs(
    title = "feature stability: T1D_Normal_vs_T1D_Obese",
    x = "Bootstrap selection frequency (%)",
    y = NULL
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

net5_stability_plot 
```
#--------------------------------------------------------------------------------------------------
#table of selected proteins for correlations 
#taken from elastic net results only keeping proteins that were selected in >50% of the regressions
#fallback to all limma proteins if too few and elasticnet cant run
#--------------------------------------------------------------------------------------------------
```{r}
# elastic-net proteins by contrast
enet_prots <- list(
  T1D_Normal_vs_LC_Normal =
    net1$stability_tbl %>%
    filter(selection_percent >= 50) %>%
    pull(protein_id),

  T1D_OverObese_vs_LC_OverObese =
    net2$stability_tbl %>%
    filter(selection_percent >= 50) %>%
    pull(protein_id),

  T1D_Normal_vs_T1D_Obese =
    net4$stability_tbl %>%
    filter(selection_percent >= 50) %>%
    pull(protein_id),

  T1D_Obese_vs_T1D_Overweight =
    net5$stability_tbl %>%
    filter(selection_percent >= 50) %>%
    pull(protein_id)
)

# limma fallback for contrast where elastic net failed
limma_fallback_prots <- res_m2 %>%
  filter(
    contrast == "T1D_Normal_vs_T1D_Overweight",
    adj.P.Val < 0.05,
    abs(logFC) >= 0.5
  ) %>%
  pull(protein_id)

contrast_protein_list <- c(
  enet_prots,
  list(T1D_Normal_vs_T1D_Overweight = limma_fallback_prots)
)

# final global protein set = union of all contrast-specific final proteins
global_proteins <- contrast_protein_list %>%
  unlist(use.names = FALSE) %>%
  unique() %>%
  intersect(colnames(dat_imp2))

length(global_proteins)
global_proteins
```


#------------------------------------------
#Correlations with kidney clinical markers
#-----------------------------------------
```{r}
# outcomes for correlations
partial_outcomes <- c(
  "HbA1c",
  "uACR",
  "eGFR",
  "avg_c_r2",
  "avg_k_r2",
  "BMI"
)

partial_covars <- c("age", "sex", "study")

# clinical + protein correlation dataframe
corr_df <- bind_cols(
  meta3 %>%
    transmute(
      record_id,
      group_bmi,
      age = as.numeric(age),
      sex = factor(sex),
      study = factor(study),
      HbA1c = as.numeric(hba1c),
      uACR  = as.numeric(acr_u),
      eGFR  = as.numeric(eGFR_CKiD_U25_avg),
      avg_c_r2 = as.numeric(avg_c_r2),
      avg_k_r2 = as.numeric(avg_k_r2),
      BMI   = as.numeric(bmi)
    ),
  as.data.frame(dat_imp2)
)


```

```{r}
global_cor_results <- run_partial_spearman(
  data = corr_df,
  proteins = global_proteins,
  outcomes = partial_outcomes,
  covars = partial_covars,
  analysis_label = "Global: all adiposity groups",
  contrast_name = "All_groups",
  min_n = 15
) %>%
  left_join(map_tbl, by = "protein_id") %>%
  mutate(
    gene_symbol = if_else(
      is.na(gene_symbol) | gene_symbol == "",
      protein_id,
      gene_symbol
    )
  )

global_cor_results %>%
  filter(FDR < 0.05) %>%
  arrange(outcome, FDR)
```


```{r}
contrast_groups <- list(
  T1D_Normal_vs_LC_Normal = c("T1D_Normal", "LC_Normal"),
  T1D_OverObese_vs_LC_OverObese = c(
    "T1D_Overweight",
    "T1D_Obese",
    "LC_Overweight_Obese"
  ),
  T1D_Normal_vs_T1D_Overweight = c("T1D_Normal", "T1D_Overweight"),
  T1D_Normal_vs_T1D_Obese = c("T1D_Normal", "T1D_Obese"),
  T1D_Obese_vs_T1D_Overweight = c("T1D_Obese", "T1D_Overweight")
)

contrast_cor_results <- purrr::imap_dfr(
  contrast_protein_list,
  function(prots, contrast_name) {

    groups_use <- contrast_groups[[contrast_name]]

    data_sub <- corr_df %>%
      filter(group_bmi %in% groups_use)

    run_partial_spearman(
      data = data_sub,
      proteins = prots,
      outcomes = partial_outcomes,
      covars = partial_covars,
      analysis_label = "Contrast-specific",
      contrast_name = contrast_name,
      min_n = 10
    )
  }
) %>%
  left_join(map_tbl, by = "protein_id") %>%
  mutate(
    gene_symbol = if_else(
      is.na(gene_symbol) | gene_symbol == "",
      protein_id,
      gene_symbol
    )
  )

contrast_cor_results %>%
  filter(FDR < 0.05) %>%
  arrange(contrast, outcome, FDR)
```
```{r}
final_cor_results <- bind_rows(
  global_cor_results,
  contrast_cor_results
) %>%
  arrange(analysis, contrast, outcome, FDR)

#write_csv(
 # final_partial_results,
  #file.path(dir.results, "partial_spearman_global_and_contrast_specific.csv")
#)

#write_csv(
 # global_partial_results,
  #file.path(dir.results, "partial_spearman_global_all_adiposity_groups.csv")
#)

#write_csv(
 # contrast_partial_results,
  #file.path(dir.results, "partial_spearman_contrast_specific.csv")
#)
```


```{r fig.width=20}
contrast_volcano_panels <- purrr::map(
  names(contrast_protein_list),
  ~ make_contrast_volcano_panel(
    cor_df = final_cor_results,
    contrast_name = .x
  )
)

names(contrast_volcano_panels) <- names(contrast_protein_list)

contrast_volcano_panels[["T1D_Normal_vs_LC_Normal"]]
contrast_volcano_panels[["T1D_OverObese_vs_LC_OverObese"]]
contrast_volcano_panels[["T1D_Normal_vs_T1D_Overweight"]]
contrast_volcano_panels[["T1D_Normal_vs_T1D_Obese"]]
ggsave(filename = file.path(dir.results, "Plots", "T1D_normal_obese_correlations.png"),
       height = 8,
       width = 12,
       dpi = 300)
contrast_volcano_panels[["T1D_Obese_vs_T1D_Overweight"]]
```
#----------------------------------------------------------------------
#doing correlations for all groups (T1D & LC) and for all T1D together
#-----------------------------------------------------------------------
```{r}
# all significant proteins from any pairwise limma contrast
all_sig_pairwise_proteins <- res_m2 %>%
  filter(
    contrast != "Global_group_bmi",
    adj.P.Val < 0.05
  ) %>%
  pull(protein_id) %>%
  unique()

length(all_sig_pairwise_proteins)
```

```{r}
partial_covars2 <- c("age", "sex") #wouldn't work with study

all_groups_sig_cor_results <- run_partial_spearman(
  data = corr_df,
  proteins = all_sig_pairwise_proteins,
  outcomes = partial_outcomes,
  covars = partial_covars2,
  analysis_label = "All groups: all significant pairwise proteins",
  contrast_name = "All_groups_sig_pairwise",
  min_n = 15
) %>%
  left_join(map_tbl, by = "protein_id") %>%
  mutate(
    gene_symbol = if_else(
      is.na(gene_symbol) | gene_symbol == "",
      protein_id,
      gene_symbol
    )
  )
```

```{r}
t1d_sig_cor_results <- run_partial_spearman(
  data = corr_df %>%
    filter(group_bmi %in% c(
      "T1D_Normal",
      "T1D_Overweight",
      "T1D_Obese"
    )),
  proteins = all_sig_pairwise_proteins,
  outcomes = partial_outcomes,
  covars = partial_covars2,
  analysis_label = "T1D only: all significant pairwise proteins",
  contrast_name = "T1D_only_sig_pairwise",
  min_n = 10
) %>%
  left_join(map_tbl, by = "protein_id") %>%
  mutate(
    gene_symbol = if_else(
      is.na(gene_symbol) | gene_symbol == "",
      protein_id,
      gene_symbol
    )
  )
```
#---------------------
#remaking results table
#---------------------
```{r}
final_cor_results <- bind_rows(
  global_cor_results,
  contrast_cor_results,
  all_groups_sig_cor_results,
  t1d_sig_cor_results
) %>%
  arrange(analysis, contrast, outcome, FDR)

#sig results
final_cor_results %>%
  filter(FDR < 0.05) %>%
  arrange(analysis, outcome, FDR)
```

```{r}
all_groups_sig_volcano_panel <- purrr::map(
  partial_outcomes,
  ~ make_partial_cor_volcano(
    cor_df = final_cor_results,
    analysis_name = "All groups: all significant pairwise proteins",
    contrast_name = "All_groups_sig_pairwise",
    outcome_name = .x
  )
)

names(all_groups_sig_volcano_panel) <- partial_outcomes

all_groups_sig_panel <-
  (all_groups_sig_volcano_panel[["BMI"]] | all_groups_sig_volcano_panel[["eGFR"]]) /
  (all_groups_sig_volcano_panel[["avg_c_r2"]] | all_groups_sig_volcano_panel[["avg_k_r2"]]) /
  (all_groups_sig_volcano_panel[["HbA1c"]] | all_groups_sig_volcano_panel[["uACR"]])

all_groups_sig_panel
```


```{r}
t1d_sig_volcano_panel <- purrr::map(
  partial_outcomes,
  ~ make_partial_cor_volcano(
    cor_df = final_cor_results,
    analysis_name = "T1D only: all significant pairwise proteins",
    contrast_name = "T1D_only_sig_pairwise",
    outcome_name = .x
  )
)

names(t1d_sig_volcano_panel) <- partial_outcomes

t1d_sig_panel <-
  (t1d_sig_volcano_panel[["BMI"]] | t1d_sig_volcano_panel[["eGFR"]]) /
  (t1d_sig_volcano_panel[["avg_c_r2"]] | t1d_sig_volcano_panel[["avg_k_r2"]]) /
  (t1d_sig_volcano_panel[["HbA1c"]] | t1d_sig_volcano_panel[["uACR"]])

t1d_sig_panel

```





