---
title: "scRNA results report"
author: "Ye Ji Choi"
format:
  revealjs:
    embed-resources: true
    scrollable: true
    transition: slide
    controls-layout: bottom-right
    menu: true
    toc: true
    toc-depth: 1
    fontsize: 15
    number-sections: true
---

```{r include = F}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-SharedLibraries-UW/Laura Pyle - Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/pylell/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}
```

```{r include = F}
source(file.path(git_path, "Renal HEIRitage/RH_RH2_IMPROVE_functions.R"))
```

# Model A (non-DKD vs. HC)

Identifying early disease/obesity signatures


# Model B (DKD vs. HC)

Identifying progressed disease/obesity signatures

# Model C (DKD vs. non-DKD)

Identifying disease progression signatures within diseased cohort

# Model D (GLP- vs. HC)

Identifying non-treated disease/obesity signatures

# Model E (GLP+ vs. GLP-)

Identifying therapeutic shifts in diseased cohort

