---
title: "08_scRNA_pathway_reversal_figures"
author: "Ye Ji Choi"
format: html
---

```{r}
library(dplyr)
library(tidyr)
library(aws.s3)
library(jsonlite)
library(ggplot2)
library(ggforce)
library(enrichR)
```

```{r}
user <- Sys.info()[["user"]]

if (user == "choiyej") { # local version
  root_path <- "/Users/choiyej/Library/CloudStorage/OneDrive-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/choiyej/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/Users/choiyej/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Bjornstad Pyle Lab/keys.json")
} else if (user == "rameshsh") { # hyak version
  root_path <- ""
  git_path <- "/mmfs1/gscratch/togo/rameshsh/CHCO-Code/Petter Bjornstad"
} else if (user == "yejichoi") { # hyak version
  root_path <- "/mmfs1/gscratch/togo/yejichoi/"
  git_path <- "/mmfs1/gscratch/togo/yejichoi/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/yejichoi/keys.json")
} else if (user == "pylell") {
  root_path <- "/Users/pylell/Library/CloudStorage/OneDrive-SharedLibraries-UW/Bjornstad/Biostatistics Core Shared Drive"
  git_path <- "/Users/pylell/Documents/GitHub/CHCO-Code/Petter Bjornstad"
  keys <- fromJSON("/mmfs1/home/pylell/keys.json")
} else {
  stop("Unknown user: please specify root path for this user.")
}

source(file.path(git_path, "Renal HEIRitage/RH_RH2_IMPROVE_functions.R"))
source(file.path(git_path, "Renal HEIRitage/RH_RH2_IMPROVE_scRNA_functions.R"))
```

```{r}
# Shared analysis config (folders, celltype_groups, celltype_groups_high, celltype_groups_low)
source(file.path(git_path, "Renal HEIRitage/analysis_config.R"))
```

```{r}
# load all the necessary files (processed df and pathways)
gsea_results <- list()
for (folder in names(folders)) {
  for (cell in names(celltype_groups)) {
    
    var_name <- paste0(tolower(cell), "_", folders[folder])
    var_name <- gsub("/", "_", var_name)
    
    processed_path <- file.path(
      root_path,
      paste0("Renal HERITAGE/Results/nebula/csv/full/", var_name, "_kpmp.csv")
    )
    
    if (!file.exists(processed_path)) {
      message("Processed DF not found, skipping: ", var_name)
      next
    }
    
    processed_df <- read.csv(processed_path)
    assign(var_name, processed_df, envir = .GlobalEnv)
    
    file_path <- file.path(root_path, paste0("Renal HERITAGE/Results/GSEA/hallmark_", folders[[folder]], "_", tolower(gsub("/", "_", cell)), "_gsea.RDS"))
    
    # Check if file exists and read it
    if (file.exists(file_path)) {
      # Create a unique name for this combination
      result_name <- paste0(folders[[folder]], "_", tolower(gsub("/", "_", cell)))
      
      # Read the RDS file
      gsea_results[[result_name]] <- readRDS(file_path)
    } else {
      message(paste0("File not found: ", file_path))
    }
  }
}
```


```{r eval = F}
# check on some genes for Petter
gene_list <- c(
  "PLVAP",
  "LRG1",
  "A2M",
  "IL13RA2",
  "SOST2",
  "EHD3",
  "FAS",
  "SLC16A3",
  "FSCN1",
  "S100A6",
  "VCAM1",
  "ICAM1",
  "SELE",
  "CXCL9",
  "CXCL10",
  "CCL2",
  "FOS",
  "JUN",
  "EGR1",
  "SROHINH1",
  "ACTA2",
  "VIM",
  "TAGLN",
  "CDH5",
  "PECAM1"
)

gene_search <- data.frame()
for (folder in names(folders)) {
  for (cell in names(celltype_groups)) {
    
    var_name <- paste0(tolower(cell), "_", folders[folder])
    var_name <- gsub("/", "_", var_name)
    
    processed_path <- file.path(
      root_path,
      paste0("Renal HERITAGE/Results/nebula/csv/full/", var_name, "_kpmp.csv")
    )
    
    if (!file.exists(processed_path)) {
      message("Processed DF not found, skipping: ", var_name)
      next
    }
    
    processed_df <- read.csv(processed_path)
    
    processed_df_subset <- processed_df %>%
      dplyr::select(
        logFC = matches("^logFC_.*GLP"),
        pval  = matches("^p_.*GLP"),
        fdr   = "fdr",
        Gene  = "Gene"
      ) %>%
      dplyr::mutate(cell = cell, run = folder) %>%
      filter(Gene %in% gene_list & pval < 0.05)

    gene_search <- rbind(gene_search, processed_df_subset)
      
  }
}

gene_search <- gene_search %>%
  arrange(Gene, cell)

write.csv(gene_search, file.path(root_path, "Renal HERITAGE/Results/Summary csv/nebula_transcript_search.csv"), row.names = F, na =)
```

```{r}
# ============================================================================
# Helper functions
# ============================================================================

make_leadingedge_long <- function(
    hallmark_df,
    de_glpn_hc,
    de_glpy_glpn,
    logfc_col_glpn_hc,
    pval_col_glpn_hc,
    logfc_col_glpy_glpn,
    pval_col_glpy_glpn,
    comp1_label = "DKD GLP- vs. HC",
    comp2_label = "DKD GLP+ vs. GLP-"
) {
  lead <- hallmark_df %>%
    dplyr::transmute(pathway, leadingEdge) %>%
    tidyr::unnest(leadingEdge) %>%
    dplyr::rename(Gene = leadingEdge) %>%
    distinct(pathway, Gene)

  d1 <- de_glpn_hc %>%
    dplyr::select(Gene, logFC = all_of(logfc_col_glpn_hc), pval = all_of(pval_col_glpn_hc)) %>%
    dplyr::mutate(comparison = comp1_label)

  d2 <- de_glpy_glpn %>%
    dplyr::select(Gene, logFC = all_of(logfc_col_glpy_glpn), pval = all_of(pval_col_glpy_glpn)) %>%
    dplyr::mutate(comparison = comp2_label)

  lead %>%
    left_join(d1, by = "Gene") %>%
    bind_rows(lead %>% left_join(d2, by = "Gene")) %>%
    group_by(pathway, Gene) %>%
    filter(sum(!is.na(logFC)) == 2) %>%
    ungroup()
}

# ---------------------------------------------------------------------------
# Hallmark pathway bar plots (faceted logFC comparison)
# ---------------------------------------------------------------------------
plot_hallmark_barplots <- function(
    celltype_groups,
    gsea_results,
    glpn_hc_suffix,
    glpy_glpn_suffix,
    logfc_col_glpn_hc,
    pval_col_glpn_hc,
    logfc_col_glpy_glpn,
    pval_col_glpy_glpn,
    comp1_label,
    comp2_label,
    output_subdir,
    output_prefix
) {
  for (cell in names(celltype_groups)) {
    cellname <- tolower(gsub("/", "_", cell))

    glpn_hc_var   <- paste0(cellname, "_", glpn_hc_suffix)
    glpy_glpn_var <- paste0(cellname, "_", glpy_glpn_suffix)

    if (!exists(glpn_hc_var, envir = .GlobalEnv) || !exists(glpy_glpn_var, envir = .GlobalEnv)) {
      message("Skipping ", cellname, ": missing processed dfs")
      next
    }

    glpn_hc_df   <- get(glpn_hc_var, envir = .GlobalEnv)
    glpy_glpn_df <- get(glpy_glpn_var, envir = .GlobalEnv)

    gsea_glpn_key <- paste0(glpn_hc_suffix, "_", cellname)
    gsea_glpy_key <- paste0(glpy_glpn_suffix, "_", cellname)

    if (!gsea_glpn_key %in% names(gsea_results) || !gsea_glpy_key %in% names(gsea_results)) {
      message("Skipping ", cellname, ": missing gsea_results")
      next
    }

    glpn_obj <- gsea_results[[gsea_glpn_key]]
    glpy_obj <- gsea_results[[gsea_glpy_key]]

    if (is.null(glpn_obj$hallmark) || is.null(glpy_obj$hallmark)) {
      message("Skipping ", cellname, ": hallmark table missing")
      next
    }

    glpn_hc_hallmark   <- glpn_obj$hallmark %>% filter(pval < 0.05)
    glpy_glpn_hallmark <- glpy_obj$hallmark %>% filter(pval < 0.05)

    if (nrow(glpn_hc_hallmark) == 0 && nrow(glpy_glpn_hallmark) == 0) {
      message("Skipping ", cellname, ": no significant hallmark pathways")
      next
    }

    le_args <- list(
      de_glpn_hc = glpn_hc_df, de_glpy_glpn = glpy_glpn_df,
      logfc_col_glpn_hc = logfc_col_glpn_hc, pval_col_glpn_hc = pval_col_glpn_hc,
      logfc_col_glpy_glpn = logfc_col_glpy_glpn, pval_col_glpy_glpn = pval_col_glpy_glpn,
      comp1_label = comp1_label, comp2_label = comp2_label
    )

    plot_df <- bind_rows(
      do.call(make_leadingedge_long, c(list(hallmark_df = glpn_hc_hallmark), le_args)),
      do.call(make_leadingedge_long, c(list(hallmark_df = glpy_glpn_hallmark), le_args))
    ) %>% distinct(pathway, Gene, comparison, .keep_all = TRUE)

    if (nrow(plot_df) == 0) {
      message("Skipping ", cellname, ": plot_df empty")
      next
    }

    pathway_rev_summary <- plot_df %>%
      dplyr::select(pathway, Gene, comparison, logFC) %>%
      pivot_wider(names_from = comparison, values_from = logFC) %>%
      dplyr::mutate(reversed = sign(.data[[comp1_label]]) != sign(.data[[comp2_label]])) %>%
      group_by(pathway) %>%
      dplyr::summarise(
        n_genes = sum(!is.na(reversed)),
        n_reversed = sum(reversed, na.rm = TRUE),
        pct_reversed = 100 * n_reversed / n_genes,
        .groups = "drop"
      )

    plot_df <- plot_df %>%
      left_join(pathway_rev_summary %>% dplyr::select(pathway, pct_reversed), by = "pathway") %>%
      dplyr::mutate(pathway_lab = paste0(
        gsub("HALLMARK ", "", gsub("_", " ", pathway)),
        "\n(", sprintf("%.1f", pct_reversed), "% Reversed)"
      ))

    n_pages <- ceiling(length(unique(plot_df$pathway)) / 6)

    for (i in seq_len(n_pages)) {
      p <- ggplot(plot_df, aes(x = Gene, y = logFC, fill = comparison)) +
        geom_col(position = position_dodge(width = 0.8), width = 0.7) +
        ggforce::facet_wrap_paginate(~ pathway_lab, nrow = 3, ncol = 2, scales = "free_x", page = i) +
        labs(x = NULL, y = "logFC", title = paste0(cell, " pathways (", i, "/", n_pages, ")"), fill = NULL) +
        theme_minimal() +
        theme(panel.grid = element_blank(), axis.text.x = element_text(angle = 40, hjust = 1),
              strip.text = element_text(face = "bold"), legend.position = "bottom")

      ggsave(
        filename = file.path(root_path, paste0("Renal HERITAGE/Results/Figures/Pathway comparisons/", output_subdir, "/", output_prefix, "_", cellname, "_", i, ".png")),
        plot = p, width = 15
      )
    }
  }
}

# ---------------------------------------------------------------------------
# Transcript reversal analysis (reversed / not-reversed / enhanced / suppressed)
# ---------------------------------------------------------------------------
compute_reversal_combined <- function(
    celltype_groups,
    glpn_hc_suffix,
    glpy_glpn_suffix,
    logfc_col_glpn_hc,
    pval_col_glpn_hc,
    logfc_col_glpy_glpn,
    pval_col_glpy_glpn,
    comp1_label = "GLP (-) vs HC",
    comp2_label = "GLP (+) vs GLP(-)"
) {
  combined <- data.frame()

  for (cell in names(celltype_groups)) {
    cellname <- tolower(gsub("/", "_", cell))

    glpn_hc_var   <- paste0(cellname, "_", glpn_hc_suffix)
    glpy_glpn_var <- paste0(cellname, "_", glpy_glpn_suffix)

    if (!exists(glpn_hc_var, envir = .GlobalEnv) || !exists(glpy_glpn_var, envir = .GlobalEnv)) {
      message("Skipping ", cellname, ": missing processed dfs")
      next
    }

    glpn_hc_df   <- get(glpn_hc_var, envir = .GlobalEnv)
    glpy_glpn_df <- get(glpy_glpn_var, envir = .GlobalEnv)

    d1 <- glpn_hc_df %>%
      dplyr::select(logFC = all_of(logfc_col_glpn_hc), p.val = all_of(pval_col_glpn_hc), fdr, Gene) %>%
      dplyr::mutate(celltype = cell, comparison = comp1_label)
    d2 <- glpy_glpn_df %>%
      dplyr::select(logFC = all_of(logfc_col_glpy_glpn), p.val = all_of(pval_col_glpy_glpn), fdr, Gene) %>%
      dplyr::mutate(celltype = cell, comparison = comp2_label)

    comp <- rbind(d1, d2) %>%
      dplyr::mutate(sgn = if_else(p.val < 0.05, sign(logFC), NA_real_)) %>%
      group_by(Gene, celltype) %>%
      dplyr::mutate(
        sgn_glpn_hc   = sgn[comparison == comp1_label][1],
        sgn_glpy_glpn = sgn[comparison == comp2_label][1],
        direction = case_when(
          is.na(sgn_glpn_hc) | is.na(sgn_glpy_glpn) ~ NA_character_,
          sgn_glpn_hc == sgn_glpy_glpn               ~ "Not Reversed",
          TRUE                                        ~ "Reversed"
        ),
        rev_direction = case_when(
          direction == "Reversed" & sgn_glpy_glpn == -1 ~ "Suppressed",
          direction == "Reversed" & sgn_glpy_glpn ==  1 ~ "Enhanced"
        )
      ) %>%
      ungroup() %>%
      filter(!is.na(direction)) %>%
      dplyr::select(-sgn, -sgn_glpn_hc, -sgn_glpy_glpn) %>%
      distinct(Gene, celltype, .keep_all = TRUE)

    combined <- rbind(combined, comp)
  }

  combined
}

# ---------------------------------------------------------------------------
# Plot reversal bar charts (direction + percentage + enhanced/suppressed)
# ---------------------------------------------------------------------------
plot_reversal_barcharts <- function(rev_combined, output_dir, file_prefix) {
  # Reversed vs not reversed
  dir_summary <- rev_combined %>%
    group_by(celltype, direction) %>%
    dplyr::summarise(n = n(), .groups = "drop") %>%
    group_by(celltype) %>%
    dplyr::mutate(total_n = sum(n)) %>%
    arrange(desc(total_n)) %>% ungroup() %>%
    dplyr::mutate(celltype = factor(celltype, levels = unique(celltype)),
                  percentage = n / total_n)

  # N of transcripts
  p1 <- dir_summary %>%
    ggplot(aes(x = celltype, y = n, fill = direction)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = c("Not Reversed" = "#f2a65a", "Reversed" = "#8fb996")) +
    labs(x = NULL, y = "N of transcripts\naffected by GLP1RA", fill = NULL) +
    theme(legend.position = "top", panel.background = element_blank(), text = element_text(size = 15),
          axis.ticks.y = element_blank(), axis.text.x = element_text(angle = 50, hjust = 1))
  ggsave(file.path(output_dir, paste0(file_prefix, "_direction_barplot.png")), plot = p1, width = 7, height = 5)

  # Percentage
  p2 <- dir_summary %>%
    ggplot(aes(x = celltype, y = percentage, fill = direction)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = c("Reversed" = "#8fb996")) +
    labs(x = NULL, y = "% of transcripts\naffected by GLP1RA", fill = NULL) +
    theme(legend.position = "top", panel.background = element_blank(), text = element_text(size = 15),
          axis.ticks.y = element_blank(), axis.text.x = element_text(angle = 50, hjust = 1))
  ggsave(file.path(output_dir, paste0(file_prefix, "_direction_barplot_percentage.png")), plot = p2, width = 7, height = 5)

  # Enhanced vs suppressed (reversed only)
  rev_summary <- rev_combined %>%
    filter(direction == "Reversed") %>%
    group_by(celltype, rev_direction) %>%
    dplyr::summarise(n = n(), .groups = "drop") %>%
    group_by(celltype) %>%
    dplyr::mutate(total_n = sum(n)) %>%
    arrange(desc(total_n)) %>% ungroup() %>%
    dplyr::mutate(celltype = factor(celltype, levels = unique(celltype)))

  p3 <- rev_summary %>%
    ggplot(aes(x = celltype, y = n, fill = rev_direction)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = c("Suppressed" = "#669bbc", "Enhanced" = "#df7373")) +
    labs(x = NULL, y = "N of transcripts\nreversed by GLP1RA", fill = NULL) +
    theme(legend.position = "top", panel.background = element_blank(), text = element_text(size = 15),
          axis.ticks.y = element_blank(), axis.text.x = element_text(angle = 50, hjust = 1))
  ggsave(file.path(output_dir, paste0(file_prefix, "_reversed_barplot.png")), plot = p3, width = 7, height = 5)

  list(dir_summary = dir_summary, rev_summary = rev_summary)
}

# ---------------------------------------------------------------------------
# Reactome enrichment of reversed transcripts
# ---------------------------------------------------------------------------
compute_reactome_enrichment <- function(rev_combined, celltype_groups) {
  rev_reactome_df <- data.frame()

  for (cell in names(celltype_groups)) {
    sup_genes <- rev_combined %>% filter(celltype == cell, rev_direction == "Suppressed") %>% pull(Gene)
    enh_genes <- rev_combined %>% filter(celltype == cell, rev_direction == "Enhanced") %>% pull(Gene)

    sup_res <- enrichr(sup_genes, databases = "Reactome_Pathways_2024")
    enh_res <- enrichr(enh_genes, databases = "Reactome_Pathways_2024")

    sup_df <- sup_res$Reactome_Pathways_2024 %>%
      dplyr::mutate(bg = "Suppressed", n_genes = as.numeric(gsub("/.*", "", Overlap)),
                    celltype = cell, neg_logp = -log10(P.value))
    enh_df <- enh_res$Reactome_Pathways_2024 %>%
      dplyr::mutate(bg = "Enhanced", n_genes = as.numeric(gsub("/.*", "", Overlap)),
                    celltype = cell, neg_logp = -log10(P.value))

    rev_reactome_df <- rbind(rev_reactome_df, sup_df, enh_df)
  }

  rev_reactome_df
}

# ---------------------------------------------------------------------------
# Plot Reactome dot plots
# ---------------------------------------------------------------------------
plot_reactome_dotplots <- function(rev_reactome_df, output_dir, file_prefix, min_celltypes = 5,
                                   subset_celltypes = NULL,
                                   sup_width = 12, sup_height = 5,
                                   enh_width = 7, enh_height = 3) {
  rev_sig <- rev_reactome_df %>%
    filter(P.value < 0.05) %>%
    group_by(Term, bg) %>%
    dplyr::mutate(n = n()) %>%
    filter(n > min_celltypes)

  # Suppressed
  p1 <- rev_sig %>% filter(bg == "Suppressed") %>%
    ggplot(aes(x = celltype, y = Term, size = n_genes, color = neg_logp)) +
    geom_point() +
    theme_minimal(base_size = 15) +
    theme(panel.grid = element_blank(), axis.text.x = element_text(angle = 60, hjust = 1)) +
    labs(x = NULL, y = NULL, size = "N of genes", color = "-log10(P)") +
    scale_color_gradient2(mid = "white", high = "#669bbc", midpoint = 0)
  ggsave(file.path(output_dir, paste0(file_prefix, "_suppressed_reactome_dotplot.png")), plot = p1, 
         width = sup_width, height = sup_height)

  # Enhanced
  p2 <- rev_sig %>% filter(bg == "Enhanced") %>%
    ggplot(aes(x = celltype, y = Term, size = n_genes, color = neg_logp)) +
    geom_point() +
    theme_minimal(base_size = 15) +
    theme(panel.grid = element_blank(), axis.text.x = element_text(angle = 60, hjust = 1)) +
    labs(x = NULL, y = NULL, size = "N of genes", color = "-log10(P)") +
    scale_color_gradient2(mid = "white", high = "#df7373", midpoint = 0)
  ggsave(file.path(output_dir, paste0(file_prefix, "_enhanced_reactome_dotplot.png")), plot = p2, 
         width = enh_width, height = enh_height)

  # Cell-type-specific plots
  if (!is.null(subset_celltypes)) {
    rev_ct <- rev_reactome_df %>%
      filter(P.value < 0.05, celltype %in% subset_celltypes) %>%
      group_by(Term, bg) %>%
      dplyr::mutate(n = n())

    for (bg_type in c("Suppressed", "Enhanced")) {
      color_high <- if (bg_type == "Suppressed") "#669bbc" else "#df7373"

      for (filter_type in c("n5", "fdr")) {
        filt_data <- if (filter_type == "n5") {
          rev_ct %>% filter(bg == bg_type, n > 5)
        } else {
          rev_ct %>% filter(bg == bg_type, Adjusted.P.value < 0.1)
        }

        suffix <- if (filter_type == "fdr") "_fdr" else ""

        p <- filt_data %>%
          ggplot(aes(x = celltype, y = Term, size = n_genes, color = neg_logp)) +
          geom_point() +
          theme_minimal(base_size = 15) +
          theme(panel.grid = element_blank(), axis.text.x = element_text(angle = 60, hjust = 1)) +
          labs(x = NULL, y = NULL, size = "N of genes", color = "-log10(P)") +
          scale_color_gradient2(mid = "white", high = color_high, midpoint = 0)

        h <- if (filter_type == "fdr") 8 else if (bg_type == "Suppressed") 20 else 15
        ggsave(file.path(output_dir, paste0(file_prefix, "_", tolower(bg_type), "_reactome_dotplot_celltypes", suffix, ".png")),
               plot = p, width = 12, height = h)
      }
    }
  }

  rev_sig
}

# ---------------------------------------------------------------------------
# Reactome pathway-level logFC bar plots
# ---------------------------------------------------------------------------
plot_reactome_barplots <- function(
    celltype_groups, rev_combined, rev_reactome_sig,
    glpn_hc_suffix, glpy_glpn_suffix,
    logfc_col_glpn_hc, pval_col_glpn_hc,
    logfc_col_glpy_glpn, pval_col_glpy_glpn,
    comp1_label, comp2_label,
    output_dir, output_prefix
) {
  for (cell in names(celltype_groups)) {
    cellname <- tolower(gsub("/", "_", cell))

    glpn_hc_var   <- paste0(cellname, "_", glpn_hc_suffix)
    glpy_glpn_var <- paste0(cellname, "_", glpy_glpn_suffix)

    if (!exists(glpn_hc_var, envir = .GlobalEnv) || !exists(glpy_glpn_var, envir = .GlobalEnv)) {
      message("Skipping ", cellname, ": missing processed dfs")
      next
    }

    glpn_hc_df   <- get(glpn_hc_var, envir = .GlobalEnv)
    glpy_glpn_df <- get(glpy_glpn_var, envir = .GlobalEnv)

    rev_reactome_sig_combined <-
      rev_reactome_sig %>%
      filter(bg == "Enhanced" & celltype == cell) %>% head(5) %>%
      separate_rows(Genes, sep = ";") %>%
      rbind(rev_reactome_sig %>%
              filter(bg == "Suppressed" & celltype == cell) %>% head(5) %>%
              separate_rows(Genes, sep = ";")) %>%
      dplyr::select(pathway = Term, Gene = Genes, bg, celltype)

    if (nrow(rev_reactome_sig_combined) == 0) {
      message("Skipping ", cell, ": no significant Reactome pathways")
      next
    }

    plot_df <- bind_rows(
      glpn_hc_df %>% separate_rows(Gene, sep = ";") %>%
        dplyr::select(Gene, logFC = all_of(logfc_col_glpn_hc), p.val = all_of(pval_col_glpn_hc), fdr) %>%
        dplyr::mutate(comparison = comp1_label),
      glpy_glpn_df %>% separate_rows(Gene, sep = ";") %>%
        dplyr::select(Gene, logFC = all_of(logfc_col_glpy_glpn), p.val = all_of(pval_col_glpy_glpn), fdr) %>%
        dplyr::mutate(comparison = comp2_label)
    ) %>%
      left_join(rev_reactome_sig_combined) %>%
      filter(!is.na(pathway))

    if (nrow(plot_df) == 0) {
      message("Skipping ", cellname, ": plot_df empty")
      next
    }

    pathway_rev_summary <- plot_df %>%
      dplyr::select(pathway, Gene, comparison, logFC) %>%
      pivot_wider(names_from = comparison, values_from = logFC) %>%
      dplyr::mutate(reversal_direction = case_when(.data[[comp2_label]] > 0 ~ "Enhanced", TRUE ~ "Suppressed")) %>%
      distinct(pathway, reversal_direction)

    plot_df <- plot_df %>%
      left_join(pathway_rev_summary %>% dplyr::select(pathway, reversal_direction), by = "pathway") %>%
      dplyr::mutate(pathway_lab = paste0(pathway, "\n(", reversal_direction, ")"))

    n_pages <- ceiling(length(unique(plot_df$pathway)) / 4)

    for (i in seq_len(n_pages)) {
      p <- ggplot(plot_df, aes(x = Gene, y = logFC, fill = comparison)) +
        geom_col(position = position_dodge(width = 0.8), width = 0.7) +
        ggforce::facet_wrap_paginate(~ pathway_lab, nrow = 2, ncol = 2, scales = "free_x", page = i) +
        labs(x = NULL, y = "logFC", title = paste0(cell, " pathways (", i, "/", n_pages, ")"), fill = NULL) +
        theme_minimal() +
        theme(panel.grid = element_blank(), axis.text.x = element_text(angle = 40, hjust = 1),
              strip.text = element_text(face = "bold"), legend.position = "bottom") +
        scale_fill_manual(values = c("grey", "#e8998d"))

      ggsave(
        filename = file.path(output_dir, paste0(output_prefix, "_", cellname, "_", i, ".png")),
        plot = p, width = 15
      )
    }
  }
}
```

```{r}
# Most responsive cell type to GLP1RA
sig_df_summary_comp <- read.csv(file.path(root_path, "Renal HERITAGE/Results/Summary csv/nebula_deg_results_summary.csv"))

# DEG summary butterfly plot helper
plot_deg_butterfly <- function(sig_df_summary, analysis_type_filter, ct_names, output_file) {
  sig_df_clean <- sig_df_summary %>%
    filter(analysis_type == analysis_type_filter) %>%
    arrange(desc(Total)) %>%
    pivot_longer(cols = c(Positive, Negative), names_to = "Direction", values_to = "Count") %>%
    dplyr::mutate(
      Count = ifelse(Direction == "Negative", -Count, Count),
      celltype = factor(celltype, levels = rev(unique(celltype)))
    )

  p <- sig_df_clean %>%
    filter(celltype %in% ct_names) %>%
    ggplot(aes(x = Count, y = celltype, fill = Direction)) +
    geom_col(width = 0.7) +
    geom_vline(xintercept = 0, color = "white", linewidth = 0.6) +
    geom_text(aes(label = celltype, x = 0), size = 4) +
    scale_x_continuous(labels = abs, expand = expansion(mult = c(0.01, 0.01))) +
    scale_fill_manual(values = c("Negative" = "#c0d6df", "Positive" = "#f4978e")) +
    labs(x = "Number of DE genes", y = NULL, fill = NULL) +
    theme(legend.position = "top", panel.background = element_blank(), text = element_text(size = 15),
          axis.text.y = element_blank(), axis.ticks.y = element_blank())
  ggsave(output_file, plot = p)
  p
}

# T2D GLP Y vs GLP N (unadjusted)
plot_deg_butterfly(sig_df_summary_comp, "T2D_GLP_Y_vs_T2D_GLP_N", names(celltype_groups_high),
                   file.path(root_path, "Renal HERITAGE/Results/Figures/DEG summary/deg_summary_t2d_glpy_glpn_highres.png"))
plot_deg_butterfly(sig_df_summary_comp, "T2D_GLP_Y_vs_T2D_GLP_N", names(celltype_groups_low),
                   file.path(root_path, "Renal HERITAGE/Results/Figures/DEG summary/deg_summary_t2d_glpy_glpn_lowres.png"))

# T2D GLP Y vs GLP N (adjusted for age + sex.x)
plot_deg_butterfly(sig_df_summary_comp, "T2D_GLP_Y_vs_T2D_GLP_N_adj_age_sex", names(celltype_groups_high),
                   file.path(root_path, "Renal HERITAGE/Results/Figures/DEG summary/deg_summary_t2d_glpy_glpn_adj_highres.png"))
plot_deg_butterfly(sig_df_summary_comp, "T2D_GLP_Y_vs_T2D_GLP_N_adj_age_sex", names(celltype_groups_low),
                   file.path(root_path, "Renal HERITAGE/Results/Figures/DEG summary/deg_summary_t2d_glpy_glpn_adj_lowres.png"))
```

## (DKD30 GLP+ vs. GLP-) vs. (DKD30 GLP- vs. HC)

### Hallmark pathway bar plots

```{r}
plot_hallmark_barplots(
  celltype_groups     = celltype_groups_low,
  gsea_results        = gsea_results,
  glpn_hc_suffix      = "dkd30_glpn_hc",
  glpy_glpn_suffix    = "dkd_30_glpy_glpn",
  logfc_col_glpn_hc   = "logFC_dkd_group_30_hcDKD",
  pval_col_glpn_hc    = "p_dkd_group_30_hcDKD",
  logfc_col_glpy_glpn = "logFC_glp_t2dobGLP_Y",
  pval_col_glpy_glpn  = "p_glp_t2dobGLP_Y",
  comp1_label         = "DKD GLP- vs. HC",
  comp2_label         = "DKD GLP+ vs. GLP-",
  output_subdir       = "DKD30",
  output_prefix       = "hallmark_dkd30_glpyn_glpn_hc"
)
```



## (T2D GLP+ vs. GLP-) vs. (T2D GLP- vs. HC) — Unadjusted

### Hallmark pathway bar plots

```{r}
plot_hallmark_barplots(
  celltype_groups     = celltype_groups_low,
  gsea_results        = gsea_results,
  glpn_hc_suffix      = "t2d_glpn_hc",
  glpy_glpn_suffix    = "t2d_glpyn",
  logfc_col_glpn_hc   = "logFC_glp_t2dobGLP_N",
  pval_col_glpn_hc    = "p_glp_t2dobGLP_N",
  logfc_col_glpy_glpn = "logFC_glp_t2dobGLP_Y",
  pval_col_glpy_glpn  = "p_glp_t2dobGLP_Y",
  comp1_label         = "T2D GLP- vs. HC",
  comp2_label         = "T2D GLP+ vs. GLP-",
  output_subdir       = "T2D",
  output_prefix       = "hallmark_t2d_glpyn_glpn_hc"
)
```

### N of transcripts reversed by GLP1RA in each cell type

```{r}
t2d_glp_rev_combined <- compute_reversal_combined(
  celltype_groups     = celltype_groups_low,
  glpn_hc_suffix      = "t2d_glpn_hc",
  glpy_glpn_suffix    = "t2d_glpyn",
  logfc_col_glpn_hc   = "logFC_glp_t2dobGLP_N",
  pval_col_glpn_hc    = "p_glp_t2dobGLP_N",
  logfc_col_glpy_glpn = "logFC_glp_t2dobGLP_Y",
  pval_col_glpy_glpn  = "p_glp_t2dobGLP_Y",
  comp1_label         = "T2D GLP (-) vs HC",
  comp2_label         = "T2D GLP (+) vs T2D GLP(-)"
)

t2d_output_dir <- file.path(root_path, "Renal HERITAGE/Results/Figures/Pathway comparisons/T2D")
plot_reversal_barcharts(t2d_glp_rev_combined, t2d_output_dir, "N_transcripts")
```

### Reversal scores

```{r}
t2d_glp_rev_scores <- t2d_glp_rev_combined %>%
  dplyr::summarise(n = n(), .by = c(celltype, comparison, direction)) %>%
  group_by(celltype, comparison) %>%
  dplyr::summarise(
    n_reversed = sum(n[direction == "Reversed"]),
    n_total = sum(n),
    reversal_score = (n_reversed / n_total) * 100,
    .groups = "drop"
  ) %>%
  arrange(desc(reversal_score)) %>%
  dplyr::mutate(celltype = factor(celltype, levels = celltype))

t2d_glp_rev_scores %>%
  ggplot(aes(x = celltype, y = reversal_score)) +
  geom_col()
```

### Pathways of reversed transcripts (Reactome)

```{r}
rev_reactome_df <- compute_reactome_enrichment(t2d_glp_rev_combined, celltype_groups_low)

rev_reactome_sig <- plot_reactome_dotplots(
  rev_reactome_df, t2d_output_dir, "rev",
  min_celltypes = 5,
  subset_celltypes = c("EC", "Immune", "VSMC_P_FIB")
)
```

### Reactome pathway-level logFC bar plots

```{r}
plot_reactome_barplots(
  celltype_groups     = celltype_groups_low,
  rev_combined        = t2d_glp_rev_combined,
  rev_reactome_sig    = rev_reactome_sig,
  glpn_hc_suffix      = "t2d_glpn_hc",
  glpy_glpn_suffix    = "t2d_glpyn",
  logfc_col_glpn_hc   = "logFC_glp_t2dobGLP_N",
  pval_col_glpn_hc    = "p_glp_t2dobGLP_N",
  logfc_col_glpy_glpn = "logFC_glp_t2dobGLP_Y",
  pval_col_glpy_glpn  = "p_glp_t2dobGLP_Y",
  comp1_label         = "T2D GLP- vs. HC",
  comp2_label         = "T2D GLP+ vs. GLP-",
  output_dir          = t2d_output_dir,
  output_prefix       = "reactome_t2d_glpyn_glpn_hc"
)
```


## (T2D GLP+ vs. GLP-) vs. (T2D GLP- vs. HC) — Adjusted for age + sex.x

### Hallmark pathway bar plots (adjusted)

```{r}
plot_hallmark_barplots(
  celltype_groups     = celltype_groups_low,
  gsea_results        = gsea_results,
  glpn_hc_suffix      = "t2d_glpn_hc_adj_age_sex",
  glpy_glpn_suffix    = "t2d_glpyn_adj_age_sex",
  logfc_col_glpn_hc   = "logFC_glp_t2dobGLP_N",
  pval_col_glpn_hc    = "p_glp_t2dobGLP_N",
  logfc_col_glpy_glpn = "logFC_glp_t2dobGLP_Y",
  pval_col_glpy_glpn  = "p_glp_t2dobGLP_Y",
  comp1_label         = "T2D GLP- vs. HC (adj)",
  comp2_label         = "T2D GLP+ vs. GLP- (adj)",
  output_subdir       = "T2D_adj_age_sex",
  output_prefix       = "hallmark_t2d_glpyn_glpn_hc_adj_age_sex"
)
```

### N of transcripts reversed by GLP1RA in each cell type (adjusted)

```{r}
t2d_glp_rev_combined_adj <- compute_reversal_combined(
  celltype_groups     = celltype_groups_low,
  glpn_hc_suffix      = "t2d_glpn_hc_adj_age_sex",
  glpy_glpn_suffix    = "t2d_glpyn_adj_age_sex",
  logfc_col_glpn_hc   = "logFC_glp_t2dobGLP_N",
  pval_col_glpn_hc    = "p_glp_t2dobGLP_N",
  logfc_col_glpy_glpn = "logFC_glp_t2dobGLP_Y",
  pval_col_glpy_glpn  = "p_glp_t2dobGLP_Y",
  comp1_label         = "T2D GLP (-) vs HC (adj)",
  comp2_label         = "T2D GLP (+) vs T2D GLP(-) (adj)"
)

t2d_adj_output_dir <- file.path(root_path, "Renal HERITAGE/Results/Figures/Pathway comparisons/T2D_adj_age_sex")
dir.create(t2d_adj_output_dir, showWarnings = FALSE, recursive = TRUE)
plot_reversal_barcharts(t2d_glp_rev_combined_adj, t2d_adj_output_dir, "N_transcripts_adj")
```

### Reversal scores (adjusted)

```{r}
t2d_glp_rev_scores_adj <- t2d_glp_rev_combined_adj %>%
  dplyr::summarise(n = n(), .by = c(celltype, comparison, direction)) %>%
  group_by(celltype, comparison) %>%
  dplyr::summarise(
    n_reversed = sum(n[direction == "Reversed"]),
    n_total = sum(n),
    reversal_score = (n_reversed / n_total) * 100,
    .groups = "drop"
  ) %>%
  arrange(desc(reversal_score)) %>%
  dplyr::mutate(celltype = factor(celltype, levels = celltype))

t2d_glp_rev_scores_adj %>%
  ggplot(aes(x = celltype, y = reversal_score)) +
  geom_col()
```

### Pathways of reversed transcripts — Reactome (adjusted)

```{r}
rev_reactome_df_adj <- compute_reactome_enrichment(t2d_glp_rev_combined_adj, celltype_groups_low)

rev_reactome_sig_adj <- plot_reactome_dotplots(
  rev_reactome_df_adj, t2d_adj_output_dir, "rev_adj",
  min_celltypes = 1,
  subset_celltypes = c("EC", "Immune", "VSMC_P_FIB")
)
```

### Reactome pathway-level logFC bar plots (adjusted)

```{r}
plot_reactome_barplots(
  celltype_groups     = celltype_groups_low,
  rev_combined        = t2d_glp_rev_combined_adj,
  rev_reactome_sig    = rev_reactome_sig_adj,
  glpn_hc_suffix      = "t2d_glpn_hc_adj_age_sex",
  glpy_glpn_suffix    = "t2d_glpyn_adj_age_sex",
  logfc_col_glpn_hc   = "logFC_glp_t2dobGLP_N",
  pval_col_glpn_hc    = "p_glp_t2dobGLP_N",
  logfc_col_glpy_glpn = "logFC_glp_t2dobGLP_Y",
  pval_col_glpy_glpn  = "p_glp_t2dobGLP_Y",
  comp1_label         = "T2D GLP- vs. HC (adj)",
  comp2_label         = "T2D GLP+ vs. GLP- (adj)",
  output_dir          = t2d_adj_output_dir,
  output_prefix       = "reactome_t2d_glpyn_glpn_hc_adj_age_sex"
)
```

